### 2019-03-14 scripts/00-datawrangling.Rmd
add data wranglt and deseq script

---
title: "00-datawrangling.Rmd"
output: md_document
---

```{r setup, include=FALSE}
# https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html
library(tidyverse)
```

# Data Wrangling

## countData and geneinfo

```{r countData}
# import count data, set rows at entreziz
countData <- read.table("../results/kallistocounts.txt", 
                        sep = ",", row.names = NULL)

## set row names as entrezid for deseq2
row.names(countData) <- countData$entrezid

# create df with gene info and remove this from countData
geneinfo <- countData %>%
  select(row.names, Name, geneid, entrezid)
head(geneinfo)
countData <- countData %>%
  select (-c(row.names, Name, geneid, entrezid))

# because DESeq require integers
countData <- round(countData)

# print tolal num of genes and samples
dim(countData)

# make a tibble for the sole purpose of previewing this huge df
countDataTibble <- as.tbl(countData)
print(countDataTibble, n_extra = 0)
```

## wrangle colData

```{r colData}
colData <- read.table(file.path( "../metadata/kallistosamples.txt"), header = F, stringsAsFactors = F)

# use strsplit to cut the filename into meaningful columns
colData$bird <- sapply(strsplit(as.character(colData$V1),'\\_'), "[", 1)
colData$sex <- sapply(strsplit(as.character(colData$V1),'\\_'), "[", 2)
colData$tissue <- sapply(strsplit(as.character(colData$V1),'\\_'), "[", 3)

# to create columns with variable names in the filename 
colData$temp <- sapply(strsplit(as.character(colData$V1),'\\_'), "[", 4)
colData$treatmentTemp <- sapply(strsplit(as.character(colData$temp),'\\.'), "[", 1)
colData$NYNO <- sapply(strsplit(as.character(colData$temp),'\\.'), "[", 2)

# relabel incorrectly named variables 
unique(colData$treatmentTemp)
colData$treatment <- ifelse(grepl("extend-hatch", colData$treatmentTemp), "extend", colData$treatmentTemp)

# drop no-longer-needed cols
colData <- colData %>% dplyr::select(-c(temp, treatmentTemp))

# replace dashes with peroids and make rowname, to match colnames(countData)
colData$V1 <- gsub("-", ".", colData$V1)
row.names(colData) <- colData$V1

# make variables factors and specify level
cols <- c("sex", "treatment", "tissue")
colData[cols] <- lapply(colData[cols], factor)

colData$tissue <- factor(colData$tissue, 
                         levels = c("hypothalamus", "pituitary", "gonad"))

## view colData
str(colData)
head(colData, 5)

# check that rownames and colnames match for DESeq
ncol(countData) == nrow(colData)
```

## subset for manipulation study

```{r manip}
# create a new dfs for just the manipulation study
colData_manipluation <- colData %>%
  dplyr::filter(grepl('m-|extend|prolong', treatment)) %>%
  droplevels()
row.names(colData_manipluation) <- colData_manipluation$V1

# print sample sizes
colData_manipluation %>% select(sex,treatment, tissue)  %>%  summary()

# create list with filenames and select countData columns matching the file names
savecols <- as.character(colData_manipluation$V1) 
savecols <- as.vector(savecols) 
countData_manipluation <- countData %>% dplyr::select(one_of(savecols))

# confirm that row and col lenghts are equal
ncol(countData_manipluation) == nrow(colData_manipluation)
```

## subset for characterization study

```{r characterization}
# filter out data related to the manipulation study
colData_characterization <- colData %>%
  dplyr::filter(!grepl('m-|extend|prolong', treatment)) %>%
  droplevels()
row.names(colData_characterization) <- colData_characterization$V1

# print sample sizes
colData_characterization %>% select(sex,treatment, tissue)  %>%  summary()

savecols <- as.character(colData_characterization$V1) 
savecols <- as.vector(savecols) 
countData_characterization <- countData %>% dplyr::select(one_of(savecols))

# check that row and col lenghts are equal
ncol(countData_characterization) == nrow(colData_characterization)
```

## save files for downstream use

```{r write}
write.csv(colData_manipluation, "../results/colData_manipluation.csv", row.names = TRUE) 
write.csv(countData_manipluation, "../results/countData_manipluation.csv", row.names = TRUE) 
write.csv(colData_characterization, "../results/colData_characterization.csv", row.names = TRUE) 
write.csv(countData_characterization, "../results/countData_characterization.csv", row.names = TRUE) 
```

```

### 2019-03-14 scripts/01-DESeq2.Rmd
add data wranglt and deseq script

---
title: "01-deseq2"
output: html_document
---

```{r setup, include=FALSE}
# https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html
library(tidyverse)
library(DESeq2)
library(cowplot)

source("functions.R")

knitr::opts_chunk$set(fig.path = '../figures/',
                      message=FALSE, warning=FALSE)
```

Now, I'll do RNAseq on all the data

```{r deseq2}
dds <- DESeqDataSetFromMatrix(countGonad, colGonad, ~ sex*treatment)
dds # view the DESeq object - note numnber of genes
dds <- dds[ rowSums(counts(dds)) > 1, ]  # Pre-filtering genes with 0 counts
dds # view number of genes afternormalization and the number of samples
dds <- DESeq(dds) # Differential expression analysis
vsd <- vst(dds, blind=FALSE) ## variance stabilized
head(assay(vsd),3)
#write.csv(assay(vsd), file = "../results/01-vsd.csv", row.names = T)
```

```{r res}
res_summary <- function(mycontrast){
  res <- results(dds, contrast = mycontrast, independentFiltering = T)
  print(mycontrast)
  print(summary(res))
  #print(head((res[order(res$padj),]), 3))
  cat("\n")
}

res_summary(c("sex", "female", "male"))
res_summary(c("treatment", "hatch", "control"))
```

```{r pca}
# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("sex","treatment"), returnData=TRUE)

percentVar <- round(100 * attr(pcadata, "percentVar"))
pcadata$sextissue <- as.factor(paste(pcadata$sex, pcadata$tissue, sep="_"))
pcadata$sextreatment <- as.factor(paste(pcadata$sex, pcadata$treatment, sep="_"))

summary(aov(PC1 ~ sex * treatment * tissue, data=pcadata)) 
summary(aov(PC2 ~ sex * treatment * tissue, data=pcadata))

PCA12 <- ggplot(pcadata, aes(pcadata$PC1, pcadata$PC2, color=sextreatment, shape= tissue)) +
    geom_point(size=1.5) +
    xlab(paste0("PC1: ", percentVar[1],"%")) +
    ylab(paste0("PC2: ", percentVar[2],"%")) +
    stat_ellipse(level = 0.95) + 
    theme_cowplot(font_size = 7, line_size = 0.25)  
PCA12

pdf(file="../figures/01-pcaplot12.pdf", width=3.3, height=2.5)
plot(PCA12)
dev.off()

```

Now, I'll do RNAseq on all the data

```{r deseq2}
dds <- DESeqDataSetFromMatrix(countHyp, colHyp, ~ treatment)
dds # view the DESeq object - note numnber of genes
dds <- dds[ rowSums(counts(dds)) > 1, ]  # Pre-filtering genes with 0 counts
dds # view number of genes afternormalization and the number of samples
dds <- DESeq(dds) # Differential expression analysis
vsd <- vst(dds, blind=FALSE) ## variance stabilized
#head(assay(vsd),3)
#write.csv(assay(vsd), file = "../results/01-vsd.csv", row.names = T)
```

```{r res}
res_summary <- function(mycontrast){
  res <- results(dds, contrast = mycontrast, independentFiltering = T)
  print(mycontrast)
  print(summary(res))
  #print(head((res[order(res$padj),]), 3))
  cat("\n")
}

res_summary(c("sex", "female", "male"))
res_summary(c("treatment", "hatch", "control"))
```

```{r pca}
# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)

percentVar <- round(100 * attr(pcadata, "percentVar"))

summary(aov(PC1 ~ treatment, data=pcadata)) 
summary(aov(PC2 ~  treatment , data=pcadata))

PCA12 <- ggplot(pcadata, aes(pcadata$PC1, pcadata$PC2, color=treatment)) +
    geom_point(size=1.5) +
    xlab(paste0("PC1: ", percentVar[1],"%")) +
    ylab(paste0("PC2: ", percentVar[2],"%")) +
    stat_ellipse(level = 0.95) + 
    theme_cowplot(font_size = 7, line_size = 0.25)  +
  labs(title ="Female Hypothalamus")
PCA12

```

### 2019-03-14 scripts/00-datawrangling.Rmd
read write files

@@
 ```{r write}
-write.csv(colData_manipluation, "../results/colData_manipluation.csv", row.names = TRUE) 
-write.csv(countData_manipluation, "../results/countData_manipluation.csv", row.names = TRUE) 
-write.csv(colData_characterization, "../results/colData_characterization.csv", row.names = TRUE) 
-write.csv(countData_characterization, "../results/countData_characterization.csv", row.names = TRUE) 
+write.csv(colData_manipluation, "../results/00-colData-manipluation.csv", row.names = TRUE) 
+write.csv(countData_manipluation, "../results/countData-manipluation.csv", row.names = TRUE) 
+write.csv(colData_characterization, "../results/00-colData-characterization.csv", row.names = TRUE) 
+write.csv(countData_characterization, "../results/00-countData-characterization.csv", row.names = TRUE) 
 ```

### 2019-03-14 scripts/01-DESeq2.Rmd
read write files

@@
 
-Now, I'll do RNAseq on all the data
+Now, I'll do RNAseq on all the data from the parental care characterization study.
+
+```{r}
+colData <- read.csv("../results/00-colData_characterization.csv", header = T, row.names = 1)
+countData <- read.csv("../results/00-countData_characterization.csv", header = T, row.names = 1)
+```
 
 ```{r deseq2}
-dds <- DESeqDataSetFromMatrix(countGonad, colGonad, ~ sex*treatment)
+dds <- DESeqDataSetFromMatrix(countData, colData, ~ sex*treatment)
 dds # view the DESeq object - note numnber of genes
@@
 vsd <- vst(dds, blind=FALSE) ## variance stabilized
-head(assay(vsd),3)
+#head(assay(vsd),3)
 #write.csv(assay(vsd), file = "../results/01-vsd.csv", row.names = T)

### 2019-03-14 scripts/01-DESeq2-all.Rmd
slim to all, rename all

--- 2019-03-14 scripts/01-DESeq2.Rmd
@@
 
-Now, I'll do RNAseq on all the data
+Now, I'll do RNAseq on all the data from the parental care characterization study.
+
+```{r}
+colData <- read.csv("../results/00-colData-characterization.csv", header = T, row.names = 1)
+countData <- read.csv("../results/00-countData-characterization.csv", header = T, row.names = 1)
+```
 
 ```{r deseq2}
-dds <- DESeqDataSetFromMatrix(countGonad, colGonad, ~ sex*treatment)
+dds <- DESeqDataSetFromMatrix(countData, colData, ~ sex*treatment*tissue)
 dds # view the DESeq object - note numnber of genes
@@
 vsd <- vst(dds, blind=FALSE) ## variance stabilized
-head(assay(vsd),3)
-#write.csv(assay(vsd), file = "../results/01-vsd.csv", row.names = T)
+#head(assay(vsd),3)
+write.csv(assay(vsd), file = "../results/01-vsd.csv", row.names = T)
 ```
@@
 res_summary(c("sex", "female", "male"))
-res_summary(c("treatment", "hatch", "control"))
+res_summary(c("tissue", "gonad", "pituitary"))
+res_summary(c("tissue", "gonad", "hypothalamus"))
+res_summary(c("tissue", "pituitary", "hypothalamus"))
 ```
@@
 
-summary(aov(PC1 ~ sex * treatment * tissue, data=pcadata)) 
-summary(aov(PC2 ~ sex * treatment * tissue, data=pcadata))
+summary(aov(PC1 ~ sex * treatment * tissue * treatment, data=pcadata)) 
+summary(aov(PC2 ~ sex * treatment * tissue * treatment, data=pcadata))
 
@@
 dev.off()
-
 ```
-
-Now, I'll do RNAseq on all the data
-
-```{r deseq2}
-dds <- DESeqDataSetFromMatrix(countHyp, colHyp, ~ treatment)
-dds # view the DESeq object - note numnber of genes
-dds <- dds[ rowSums(counts(dds)) > 1, ]  # Pre-filtering genes with 0 counts
-dds # view number of genes afternormalization and the number of samples
-dds <- DESeq(dds) # Differential expression analysis
-vsd <- vst(dds, blind=FALSE) ## variance stabilized
-#head(assay(vsd),3)
-#write.csv(assay(vsd), file = "../results/01-vsd.csv", row.names = T)
-```
-
-```{r res}
-res_summary <- function(mycontrast){
-  res <- results(dds, contrast = mycontrast, independentFiltering = T)
-  print(mycontrast)
-  print(summary(res))
-  #print(head((res[order(res$padj),]), 3))
-  cat("\n")
-}
-
-res_summary(c("sex", "female", "male"))
-res_summary(c("treatment", "hatch", "control"))
-```
-
-```{r pca}
-# create the dataframe using my function pcadataframe
-pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
-
-percentVar <- round(100 * attr(pcadata, "percentVar"))
-
-summary(aov(PC1 ~ treatment, data=pcadata)) 
-summary(aov(PC2 ~  treatment , data=pcadata))
-
-PCA12 <- ggplot(pcadata, aes(pcadata$PC1, pcadata$PC2, color=treatment)) +
-    geom_point(size=1.5) +
-    xlab(paste0("PC1: ", percentVar[1],"%")) +
-    ylab(paste0("PC2: ", percentVar[2],"%")) +
-    stat_ellipse(level = 0.95) + 
-    theme_cowplot(font_size = 7, line_size = 0.25)  +
-  labs(title ="Female Hypothalamus")
-PCA12
-
-```

### 2019-03-14 scripts/00-datawrangling.Rmd
now with underscores no dash

--- 2019-03-14 scripts/00-datawrangling.Rmd
@@
 row.names(countData) <- countData$entrezid
+
+# replace dashes with periods (to make deseq happy)
+colnames(countData) <- gsub('-', '.', colnames(countData))
 
@@
 
-# replace dashes with peroids and make rowname, to match colnames(countData)
+# replace dashes with periods and make rowname, (to make deseq happy)
 colData$V1 <- gsub("-", ".", colData$V1)
@@
 ```{r write}
-write.csv(colData_manipluation, "../results/colData_manipluation.csv", row.names = TRUE) 
+write.csv(colData_manipluation, "../results/00_colData_manipluation.csv", row.names = TRUE) 
 write.csv(countData_manipluation, "../results/countData_manipluation.csv", row.names = TRUE) 
-write.csv(colData_characterization, "../results/colData_characterization.csv", row.names = TRUE) 
-write.csv(countData_characterization, "../results/countData_characterization.csv", row.names = TRUE) 
+write.csv(colData_characterization, "../results/00_colData_characterization.csv", row.names = TRUE) 
+write.csv(countData_characterization, "../results/00-countData-characterization.csv", row.names = TRUE) 
 ```

### 2019-03-14 scripts/02-DESeq2-sextissuespecific.Rmd
rename for order

--- 2019-03-14 scripts/01-DESeq2-all.Rmd
@@
 ---
-title: "01-deseq2"
-output: html_document
+title: "02-deseq2-sextissuespecific"
+output: md_document
 ---
@@
 
-Now, I'll do RNAseq on all the data from the parental care characterization study.
+Now, I'll do RNAseq on all the data from the parental care characterization study for each tissue and sex.
 
@@
 countData <- read.csv("../results/00-countData-characterization.csv", header = T, row.names = 1)
+```
+
+```{r}
+
+colData$sextissue <- as.factor(paste(colData$sex, colData$tissue, sep = ""))
+
+mygroups <- levels(colData$sextissue)
+
+for(i in mygroups){
+  print(i)
+  colData <- colData %>%
+    dplyr::filter(grepl(i, sextissue)) %>%
+    droplevels()
+  row.names(colData_characterization) <- colData_characterization$V1
+  colData <- 
+  print(head(colData_characterization))
+}
+
+# print sample sizes
+colData_characterization %>% select(sex,treatment, tissue)  %>%  summary()
+
+savecols <- as.character(colData_characterization$V1) 
+savecols <- as.vector(savecols) 
+countData_characterization <- countData %>% dplyr::select(one_of(savecols))
+
+# check that row and col lenghts are equal
+ncol(countData_characterization) == nrow(colData_characterization)
+}
 ```

### 2019-03-14 scripts/01-DESeq2-all.Rmd
slim to all, rename all
now with underscores no dash

@@
 ---
-title: "01-deseq2"
-output: html_document
+title: "01-deseq2-all"
+output: md_document
 ---
@@
 ```{r}
-colData <- read.csv("../results/00-colData-characterization.csv", header = T, row.names = 1)
-countData <- read.csv("../results/00-countData-characterization.csv", header = T, row.names = 1)
+colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
+countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
 ```
@@
 
+PCA34 <- ggplot(pcadata, aes(pcadata$PC3, pcadata$PC4, color=sextreatment, shape= tissue)) +
+    geom_point(size=1.5) +
+    xlab(paste0("PC3: ", percentVar[3],"%")) +
+    ylab(paste0("PC4: ", percentVar[4],"%")) +
+    stat_ellipse(level = 0.95) + 
+    theme_cowplot(font_size = 7, line_size = 0.25)  
+PCA34
+
+PCA56 <- ggplot(pcadata, aes(pcadata$PC5, pcadata$PC6, color=sextreatment, shape= tissue)) +
+    geom_point(size=1.5) +
+    xlab(paste0("PC5: ", percentVar[53],"%")) +
+    ylab(paste0("PC6: ", percentVar[6],"%")) +
+    stat_ellipse(level = 0.95) + 
+    theme_cowplot(font_size = 7, line_size = 0.25)  
+PCA56
+
 pdf(file="../figures/01-pcaplot12.pdf", width=3.3, height=2.5)
@@
 dev.off()
+
+pdf(file="../figures/01-pcaplot56.pdf", width=3.3, height=2.5)
+plot(PCA56)
+dev.off()
 ```

### 2019-03-14 scripts/00-datawrangling.Rmd
for realz rename

@@
 write.csv(colData_manipluation, "../results/00_colData_manipluation.csv", row.names = TRUE) 
-write.csv(countData_manipluation, "../results/countData_manipluation.csv", row.names = TRUE) 
+write.csv(countData_manipluation, "../results/00_countData_manipluation.csv", row.names = TRUE) 
 write.csv(colData_characterization, "../results/00_colData_characterization.csv", row.names = TRUE) 
-write.csv(countData_characterization, "../results/00-countData-characterization.csv", row.names = TRUE) 
+write.csv(countData_characterization, "../results/00_countData_characterization.csv", row.names = TRUE) 
 ```

### 2019-03-14 scripts/01-DESeq2-all.Rmd
for realz rename

@@
 
+write.csv(pcadata, "../results/01_pcadata.csv")
+write.csv(percentVar, "../results/01_percentVar.csv")
+
 summary(aov(PC1 ~ sex * treatment * tissue * treatment, data=pcadata)) 
 summary(aov(PC2 ~ sex * treatment * tissue * treatment, data=pcadata)) 
+summary(aov(PC3 ~ sex * treatment * tissue * treatment, data=pcadata)) 
+summary(aov(PC4 ~ sex * treatment * tissue * treatment, data=pcadata)) 
+summary(aov(PC5 ~ sex * treatment * tissue * treatment, data=pcadata)) 
+summary(aov(PC6 ~ sex * treatment * tissue * treatment, data=pcadata))
 

### 2019-03-15 scripts/00_datawrangling.Rmd
update data wrangling

--- 2019-03-14 scripts/00-datawrangling.Rmd
@@
 ---
-title: "00-datawrangling.Rmd"
+title: "00_datawrangling.Rmd"
 output: md_document
@@
 # replace dashes with periods and make rowname, (to make deseq happy)
+colData$bird <- gsub("-", ".", colData$bird)
+colData$treatment <- gsub("-", ".", colData$treatment)
 colData$V1 <- gsub("-", ".", colData$V1)
@@
 colData[cols] <- lapply(colData[cols], factor) 
-
 colData$tissue <- factor(colData$tissue, 
                          levels = c("hypothalamus", "pituitary", "gonad"))
+
+# create a combinatorial variable
+colData$group <- paste(colData$sex, colData$tissue, colData$treatment, sep=".")
 
@@
 write.csv(colData_manipluation, "../results/00_colData_manipluation.csv", row.names = TRUE) 
-write.csv(countData_manipluation, "../results/countData_manipluation.csv", row.names = TRUE) 
+write.csv(countData_manipluation, "../results/00_countData_manipluation.csv", row.names = TRUE) 
 write.csv(colData_characterization, "../results/00_colData_characterization.csv", row.names = TRUE) 
-write.csv(countData_characterization, "../results/00-countData-characterization.csv", row.names = TRUE) 
+write.csv(countData_characterization, "../results/00_countData_characterization.csv", row.names = TRUE) 
+write.csv(geneinfo, "../results/00_geneinfo.csv", row.names = TRUE)
 ```
-
-```

### 2019-03-15 scripts/01_limma_all.Rmd
add limma pipeline fisrt attempt

---
title: "01_limma_voom"
output: md_document
---

DESeq2 is _not_ recommended for experiments with more than 100 samples ([see Mike Love's post](https://mikelove.wordpress.com/2016/09/28/deseq2-or-edger/)), so I decided to try the limma package. I followed [this tutorial](https://www.bioconductor.org/packages/devel/workflows/vignettes/RNAseq123/inst/doc/limmaWorkflow.html).

```{r}
library(limma)
library(Glimma)
library(edgeR)
library(kableExtra)

knitr::opts_chunk$set(fig.path = '../figures/',cache=TRUE)

```

First, I read in the data I processed in 00_datawrangling.Rmd.

```{r read}
colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
head(colData)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
```

Then, I followed the steps from <https://github.com/macmanes-lab/RockDove/blob/master/parental_care/parental_analysis.Rmd>.

```{r edgeR}
# create a large DGEList with 3 elements
parentalobject <- DGEList(counts=countData, genes=geneinfo, group=colData$group)

# transform raw counts to countspermillion
cpms <- cpm(parentalobject)

# calculate number of lowly lowly expressed genes and remove them
table(rowSums(parentalobject$counts==0)==10)
keep_genes <- rowSums(cpms >= 1) >= 10
dge <- parentalobject[keep_genes, ]

# magic
parentaldesign <- model.matrix(~ colData$group )
colnames(parentaldesign) <- levels(colData$group)
parentalobject <- calcNormFactors(parentalobject)
parentalobject <- estimateCommonDisp(parentalobject)
parentalobject <- estimateTagwiseDisp(parentalobject)
parentalobject <- estimateDisp(parentalobject, parentaldesign)
parentalobject <- estimateGLMCommonDisp(parentalobject, parentaldesign, verbose=TRUE)
parentalobject <- estimateGLMTrendedDisp(parentalobject, parentaldesign)
parentalobject <- estimateGLMTagwiseDisp(parentalobject, parentaldesign)
fit <- glmFit( parentalobject, parentaldesign, robust=T)
treatres <- glmTreat(fit, lfc = 1)
```

# specify contrasts

```{r 01-contrasts}
levels(colData$group)

my.contrasts <- makeContrasts(
	         FG_HL = female.gonad.hatch - female.gonad.lay,
	         FH_HL = female.hypothalamus.hatch - female.hypothalamus.lay,
	         FP_HL = male.pituitary.hatch - male.pituitary.lay,
	         MP_HL = male.pituitary.hatch - male.pituitary.lay,
	         MH_HL = male.hypothalamus.hatch - male.hypothalamus.lay,
	         MG_HL = male.gonad.hatch - male.gonad.lay,	         
levels=parentaldesign)

cont <- "FG_HL"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female gonad lay to hatch', frame.plot=F)

cont <- "FH_HL"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female hypothalamus lay to hatch', frame.plot=F)

cont <- "FP_HL"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female pituitary lay to hatch', frame.plot=F)

```

### 2019-03-19 scripts/01_limma_all.Rmd
more ma plots and a volcano plot

@@
 library(kableExtra)
+library(cowplot)
+library(ggplot2)
 
@@
 	         FH_HL = female.hypothalamus.hatch - female.hypothalamus.lay,
-	         FP_HL = male.pituitary.hatch - male.pituitary.lay,
+	         FP_HL = female.pituitary.hatch - female.pituitary.lay,
 	         MP_HL = male.pituitary.hatch - male.pituitary.lay,
@@
 #kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female gonad lay to hatch', frame.plot=F)
+c <- plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female gonad lay to hatch', frame.plot=F)
 
@@
 #kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female hypothalamus lay to hatch', frame.plot=F)
+a <- plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female hypothalamus lay to hatch', frame.plot=F)
 
@@
 #kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female pituitary lay to hatch', frame.plot=F)
+b <- plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female pituitary lay to hatch', frame.plot=F)
+
+cont <- "MG_HL"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+f <- plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Male gonad lay to hatch', frame.plot=F)
+
+cont <- "MH_HL"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+d <- plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Male hypothalamus lay to hatch', frame.plot=F)
+
+cont <- "MP_HL"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+e <- plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Male pituitary lay to hatch', frame.plot=F)
+
+plot_grid(a,b,c,d,e,f, nrow = 2)
 
 ```
+
+```{r volcanoplots}
+# from http://www.compbio.dundee.ac.uk/user/pschofield/Projects/teaching_pg/workshops/biocDGE.html#maplots
+
+lrt <- glmLRT(fit,coef=2)
+topTags(lrt)
+
+tt <- topTags(lrt,n=10000)$table
+
+ggplot(data=tt) + geom_point(aes(x=logFC,y=-log(FDR),color=logCPM)) +
+  scale_colour_gradientn(colours=c("#000000" ,"#FF0000" ))
+
+```

### 2019-03-19 scripts/00_datawrangling.Rmd
fix filtering problem

@@
 colData_manipluation <- colData %>%
-  dplyr::filter(grepl('m-|extend|prolong', treatment)) %>%
+  dplyr::filter(grepl('m.|extend|prolong', treatment)) %>%
   droplevels()
@@
 colData_characterization <- colData %>%
-  dplyr::filter(!grepl('m-|extend|prolong', treatment)) %>%
+  dplyr::filter(!grepl('m.|extend|prolong', treatment)) %>%
   droplevels()

### 2019-03-19 scripts/01_limma_all.Rmd
add mds plot
without manipulation data

@@
 ```{r read}
+# import "colData" which contains sample information and "countData" which contains read counts
 colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
@@
 
-# magic
+# specific the design
 parentaldesign <- model.matrix(~ colData$group )
 colnames(parentaldesign) <- levels(colData$group)
+
+# The TMM normalization
 parentalobject <- calcNormFactors(parentalobject)
@@
 parentalobject <- estimateGLMTagwiseDisp(parentalobject, parentaldesign)
+
+#  perform likelihood ratio test and thresholded testing
 fit <- glmFit( parentalobject, parentaldesign, robust=T)
-treatres <- glmTreat(fit, lfc = 1)
+tr <- glmTreat(fit, lfc = 1)
+topTags(tr)
 ```
 
-# specify contrasts
+# plotMDS (multidimential scaling)
+
+```{r plotMDS}
+plotMDS(parentalobject, cex = 0.5)
+```
+
+# specify contrasts and make MA plots (currently only a subset)
 
 ```{r 01-contrasts}
+# view all levels
 levels(colData$group)
 
+# subset of conrasts - sex specific comparing hatch to lay
 my.contrasts <- makeContrasts(
@@
 
+# volcano plots
+
 ```{r volcanoplots}
@@
   scale_colour_gradientn(colours=c("#000000" ,"#FF0000" ))
-
 ```

### 2019-03-19 scripts/01_limma_all.Rmd
add colored mds and tables for contrasts

@@
 
+```{r plotMDS-colored}
+levels(colData$tissue)
+col.tissue <- c("#7570b3","#d95f02", "#1b9e77")[colData$tissue]
+plotMDS(parentalobject,col=col.tissue)
+legend("bottomright",fill=c("#7570b3","#d95f02", "#1b9e77"),legend=levels(colData$tissue))
+title("Colored by Tissue")
+
+levels(colData$treatment)
+col.treatment <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")[colData$treatment]
+plotMDS(parentalobject,col=col.treatment)
+legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
+title("Colored by Treatment")
+
+levels(colData$sex)
+col.sex <- c("#5ab4ac","#d8b365")[colData$sex]
+plotMDS(parentalobject,col=col.sex)
+legend("bottomright",fill=c("#5ab4ac","#d8b365"),legend=levels(colData$sex))
+title("Colored by Sex")
+```
+
 # specify contrasts and make MA plots (currently only a subset)
@@
 	adjust.method="fdr", p.value=0.01))
-#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-c <- plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female gonad lay to hatch', frame.plot=F)
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female gonad lay to hatch', frame.plot=F)
 
@@
 	adjust.method="fdr", p.value=0.01))
-#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-a <- plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female hypothalamus lay to hatch', frame.plot=F)
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female hypothalamus lay to hatch', frame.plot=F)
 
@@
 	adjust.method="fdr", p.value=0.01))
-#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-b <- plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female pituitary lay to hatch', frame.plot=F)
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female pituitary lay to hatch', frame.plot=F)
 
@@
 #kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-f <- plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Male gonad lay to hatch', frame.plot=F)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Male gonad lay to hatch', frame.plot=F)
 
@@
 	adjust.method="fdr", p.value=0.01))
-#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-d <- plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Male hypothalamus lay to hatch', frame.plot=F)
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Male hypothalamus lay to hatch', frame.plot=F)
 
@@
 	adjust.method="fdr", p.value=0.01))
-#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-e <- plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Male pituitary lay to hatch', frame.plot=F)
-
-plot_grid(a,b,c,d,e,f, nrow = 2)
-
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Male pituitary lay to hatch', frame.plot=F)
 ```

### 2019-03-19 scripts/01_limma_all.Rmd
now with short labels

@@
 ```
+```{r plotMDS-colorshape}
+colData$sexTissue <- paste(colData$sex, colData$tissue, sep = "_")
+
+plotMDS(parentalobject,col=col.treatment, labels = colData$sexTissue)
+legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
+title("Colored by Treatment")
+```
 

### 2019-03-19 scripts/02_limma_pit.Rmd
add pituitary only analysis

---
title: "02_limma_vomm_pit"
output: md_document
---

DESeq2 is _not_ recommended for experiments with more than 100 samples ([see Mike Love's post](https://mikelove.wordpress.com/2016/09/28/deseq2-or-edger/)), so I decided to try the limma package. I followed [this tutorial](https://www.bioconductor.org/packages/devel/workflows/vignettes/RNAseq123/inst/doc/limmaWorkflow.html).

```{r}
library(tidyverse)
library(limma)
library(Glimma)
library(edgeR)
library(kableExtra)
library(cowplot)
library(ggplot2)

knitr::opts_chunk$set(fig.path = '../figures/pit/',cache=TRUE)

```

First, I read in the data I processed in 00_datawrangling.Rmd.

```{r read}
# import "colData" which contains sample information and "countData" which contains read counts
colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
head(colData)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
```

```{r filter}
colData <- colData %>%
  dplyr::filter(grepl('pituitary', tissue)) %>%
  droplevels()
row.names(colData) <- colData$V1

# print sample sizes
colData %>% select(sex,treatment, tissue)  %>%  summary()

savecols <- as.character(colData$V1) 
savecols <- as.vector(savecols) 
countData <- countData %>% dplyr::select(one_of(savecols))

# check that row and col lenghts are equal
ncol(countData) == nrow(colData)
```

Then, I followed the steps from <https://github.com/macmanes-lab/RockDove/blob/master/parental_care/parental_analysis.Rmd>.

```{r edgeR}
# create a large DGEList with 3 elements
parentalobject <- DGEList(counts=countData, genes=geneinfo, group=colData$group)

# transform raw counts to countspermillion
cpms <- cpm(parentalobject)

# calculate number of lowly lowly expressed genes and remove them
table(rowSums(parentalobject$counts==0)==10)
keep_genes <- rowSums(cpms >= 1) >= 10
dge <- parentalobject[keep_genes, ]

# specific the design
parentaldesign <- model.matrix(~ colData$group )
colnames(parentaldesign) <- levels(colData$group)

# The TMM normalization
parentalobject <- calcNormFactors(parentalobject)
parentalobject <- estimateCommonDisp(parentalobject)
parentalobject <- estimateTagwiseDisp(parentalobject)
parentalobject <- estimateDisp(parentalobject, parentaldesign)
parentalobject <- estimateGLMCommonDisp(parentalobject, parentaldesign, verbose=TRUE)
parentalobject <- estimateGLMTrendedDisp(parentalobject, parentaldesign)
parentalobject <- estimateGLMTagwiseDisp(parentalobject, parentaldesign)

#  perform likelihood ratio test and thresholded testing
fit <- glmFit( parentalobject, parentaldesign, robust=T)
tr <- glmTreat(fit, lfc = 1)
topTags(tr)
```

# plotMDS (multidimential scaling)

```{r plotMDS}
plotMDS(parentalobject, cex = 0.5)
```

```{r plotMDS-lables}
levels(colData$treatment)
col.treatment <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")[colData$treatment]

plotMDS(parentalobject,col=col.treatment, labels = colData$sex)
legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
title("Pituitary Colored by Treatment")

plotMDS(parentalobject,dim=c(3,4), col=col.treatment, labels = colData$sex)
legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
title("Pituitary Colored by Treatment")
```

# specify contrasts and make MA plots (currently only a subset)

```{r 01-contrasts}
# view all levels
levels(colData$group)

# subset of conrasts - sex specific comparing hatch to lay
my.contrasts <- makeContrasts(
	         FP_CB = female.pituitary.control - female.pituitary.bldg,
	         FP_BL = female.pituitary.bldg - female.pituitary.lay,
	         FP_Li3 = female.pituitary.lay - female.pituitary.inc.d3,
	         FP_i39 = female.pituitary.inc.d3 - female.pituitary.inc.d9,
	         FP_i917 = female.pituitary.inc.d9 - female.pituitary.inc.d17,
	         FP_i17H = female.pituitary.inc.d17 - female.pituitary.hatch,
	         FP_H5 = female.pituitary.hatch -  female.pituitary.n5,
	         FP_n59 = female.pituitary.n5 - female.pituitary.n9,
	         FP_n9C = female.pituitary.n9 - female.pituitary.control,
levels=parentaldesign)

cont <- "FP_CB"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_CB', frame.plot=F)

cont <- "FP_BL"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_BL', frame.plot=F)

cont <- "FP_Li3"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_Li3', frame.plot=F)

cont <- "FP_i39"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_i39', frame.plot=F)

cont <- "FP_i917"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_i917', frame.plot=F)

cont <- "FP_i17H"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_i17H', frame.plot=F)

cont <- "FP_H5"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_H5', frame.plot=F)

cont <- "FP_n59"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_n59', frame.plot=F)

cont <- "FP_n9C"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_n9C', frame.plot=F)
```

# volcano plots

```{r volcanoplots}
# from http://www.compbio.dundee.ac.uk/user/pschofield/Projects/teaching_pg/workshops/biocDGE.html#maplots

lrt <- glmLRT(fit,coef=2)
topTags(lrt)

tt <- topTags(lrt,n=10000)$table

ggplot(data=tt) + geom_point(aes(x=logFC,y=-log(FDR),color=logCPM)) +
  scale_colour_gradientn(colours=c("#000000" ,"#FF0000" ))
```

### 2019-03-19 scripts/03_limma_hyp.Rmd
add hypothalamus

---
title: "03_limma_hyp"
output: md_document
---

DESeq2 is _not_ recommended for experiments with more than 100 samples ([see Mike Love's post](https://mikelove.wordpress.com/2016/09/28/deseq2-or-edger/)), so I decided to try the limma package. I followed [this tutorial](https://www.bioconductor.org/packages/devel/workflows/vignettes/RNAseq123/inst/doc/limmaWorkflow.html).

```{r}
library(tidyverse)
library(limma)
library(Glimma)
library(edgeR)
library(kableExtra)
library(cowplot)
library(ggplot2)

knitr::opts_chunk$set(fig.path = '../figures/hyp/',cache=TRUE)

```

First, I read in the data I processed in 00_datawrangling.Rmd.

```{r read}
# import "colData" which contains sample information and "countData" which contains read counts
colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
head(colData)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
```

```{r filter}
colData <- colData %>%
  dplyr::filter(grepl('hypothalamus', tissue)) %>%
  droplevels()
row.names(colData) <- colData$V1

# print sample sizes
colData %>% select(sex,treatment, tissue)  %>%  summary()

savecols <- as.character(colData$V1) 
savecols <- as.vector(savecols) 
countData <- countData %>% dplyr::select(one_of(savecols))

# check that row and col lenghts are equal
ncol(countData) == nrow(colData)
```

Then, I followed the steps from <https://github.com/macmanes-lab/RockDove/blob/master/parental_care/parental_analysis.Rmd>.

```{r edgeR}
# create a large DGEList with 3 elements
parentalobject <- DGEList(counts=countData, genes=geneinfo, group=colData$group)

# transform raw counts to countspermillion
cpms <- cpm(parentalobject)

# calculate number of lowly lowly expressed genes and remove them
table(rowSums(parentalobject$counts==0)==10)
keep_genes <- rowSums(cpms >= 1) >= 10
dge <- parentalobject[keep_genes, ]

# specific the design
parentaldesign <- model.matrix(~ colData$group )
colnames(parentaldesign) <- levels(colData$group)

# The TMM normalization
parentalobject <- calcNormFactors(parentalobject)
parentalobject <- estimateCommonDisp(parentalobject)
parentalobject <- estimateTagwiseDisp(parentalobject)
parentalobject <- estimateDisp(parentalobject, parentaldesign)
parentalobject <- estimateGLMCommonDisp(parentalobject, parentaldesign, verbose=TRUE)
parentalobject <- estimateGLMTrendedDisp(parentalobject, parentaldesign)
parentalobject <- estimateGLMTagwiseDisp(parentalobject, parentaldesign)

#  perform likelihood ratio test and thresholded testing
fit <- glmFit( parentalobject, parentaldesign, robust=T)
tr <- glmTreat(fit, lfc = 1)
topTags(tr)
```

# plotMDS (multidimential scaling)

```{r plotMDS}
plotMDS(parentalobject, cex = 0.5)
```

```{r plotMDS-lables}
levels(colData$treatment)
col.treatment <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")[colData$treatment]

plotMDS(parentalobject,col=col.treatment, labels = colData$sex)
legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
title("Hypothalamus Colored by Treatment")

plotMDS(parentalobject,dim=c(3,4), col=col.treatment, labels = colData$sex)
legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
title("Hypothalamus Colored by Treatment")
```

# specify contrasts and make MA plots (currently only a subset)

```{r 01-contrasts}
# view all levels
levels(colData$group)

# subset of conrasts - sex specific comparing hatch to lay
my.contrasts <- makeContrasts(
	         FH_CB = female.hypothalamus.control - female.hypothalamus.bldg,
	         FH_BL = female.hypothalamus.bldg - female.hypothalamus.lay,
	         FH_Li3 = female.hypothalamus.lay - female.hypothalamus.inc.d3,
	         FH_i39 = female.hypothalamus.inc.d3 - female.hypothalamus.inc.d9,
	         FH_i917 = female.hypothalamus.inc.d9 - female.hypothalamus.inc.d17,
	         FH_i17H = female.hypothalamus.inc.d17 - female.hypothalamus.hatch,
	         FH_H5 = female.hypothalamus.hatch -  female.hypothalamus.n5,
	         FH_n59 = female.hypothalamus.n5 - female.hypothalamus.n9,
	         FH_n9C = female.hypothalamus.n9 - female.hypothalamus.control,

	         MH_CB = male.hypothalamus.control - male.hypothalamus.bldg,
	         MH_BL = male.hypothalamus.bldg - male.hypothalamus.lay,
	         MH_Li3 = male.hypothalamus.lay - male.hypothalamus.inc.d3,
	         MH_i39 = male.hypothalamus.inc.d3 - male.hypothalamus.inc.d9,
	         MH_i917 = male.hypothalamus.inc.d9 - male.hypothalamus.inc.d17,
	         MH_i17H = male.hypothalamus.inc.d17 - male.hypothalamus.hatch,
	         MH_H5 = male.hypothalamus.hatch -  male.hypothalamus.n5,
	         MH_n59 = male.hypothalamus.n5 - male.hypothalamus.n9,
	         MH_n9C = male.hypothalamus.n9 - male.hypothalamus.control,
levels=parentaldesign)

# female comparisons
cont <- "FH_CB"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_CB', frame.plot=F)

cont <- "FH_BL"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_BL', frame.plot=F)

cont <- "FH_Li3"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_Li3', frame.plot=F)

cont <- "FH_i39"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_i39', frame.plot=F)

cont <- "FH_i917"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_i917', frame.plot=F)

cont <- "FH_i17H"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_i17H', frame.plot=F)

cont <- "FH_H5"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_H5', frame.plot=F)

cont <- "FH_n59"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_n59', frame.plot=F)

cont <- "FH_n9C"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_n9C', frame.plot=F)

## male comparisons

cont <- "MH_CB"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_CB', frame.plot=F)

cont <- "MH_BL"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_BL', frame.plot=F)

cont <- "MH_Li3"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_Li3', frame.plot=F)

cont <- "MH_i39"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_i39', frame.plot=F)

cont <- "MH_i917"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_i917', frame.plot=F)

cont <- "MH_i17H"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_i17H', frame.plot=F)

cont <- "MH_H5"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_H5', frame.plot=F)

cont <- "MH_n59"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_n59', frame.plot=F)

cont <- "MH_n9C"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_n9C', frame.plot=F)

```

# volcano plots

```{r volcanoplots}
# from http://www.compbio.dundee.ac.uk/user/pschofield/Projects/teaching_pg/workshops/biocDGE.html#maplots

lrt <- glmLRT(fit,coef=2)
topTags(lrt)

tt <- topTags(lrt,n=10000)$table

ggplot(data=tt) + geom_point(aes(x=logFC,y=-log(FDR),color=logCPM)) +
  scale_colour_gradientn(colours=c("#000000" ,"#FF0000" ))
```

### 2019-03-19 scripts/04_limma_gon.Rmd
gonad analysis

---
title: "04_limma_gon"
output: md_document
---

DESeq2 is _not_ recommended for experiments with more than 100 samples ([see Mike Love's post](https://mikelove.wordpress.com/2016/09/28/deseq2-or-edger/)), so I decided to try the limma package. I followed [this tutorial](https://www.bioconductor.org/packages/devel/workflows/vignettes/RNAseq123/inst/doc/limmaWorkflow.html).

```{r}
library(tidyverse)
library(limma)
library(Glimma)
library(edgeR)
library(kableExtra)
library(cowplot)
library(ggplot2)

knitr::opts_chunk$set(fig.path = '../figures/gon/',cache=TRUE)

```

First, I read in the data I processed in 00_datawrangling.Rmd.

```{r read}
# import "colData" which contains sample information and "countData" which contains read counts
colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
head(colData)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
```

```{r filter}
colData <- colData %>%
  dplyr::filter(grepl('gonad', tissue)) %>%
  droplevels()
row.names(colData) <- colData$V1

# print sample sizes
colData %>% select(sex,treatment, tissue)  %>%  summary()

savecols <- as.character(colData$V1) 
savecols <- as.vector(savecols) 
countData <- countData %>% dplyr::select(one_of(savecols))

# check that row and col lenghts are equal
ncol(countData) == nrow(colData)
```

Then, I followed the steps from <https://github.com/macmanes-lab/RockDove/blob/master/parental_care/parental_analysis.Rmd>.

```{r edgeR}
# create a large DGEList with 3 elements
parentalobject <- DGEList(counts=countData, genes=geneinfo, group=colData$group)

# transform raw counts to countspermillion
cpms <- cpm(parentalobject)

# calculate number of lowly lowly expressed genes and remove them
table(rowSums(parentalobject$counts==0)==10)
keep_genes <- rowSums(cpms >= 1) >= 10
dge <- parentalobject[keep_genes, ]

# specific the design
parentaldesign <- model.matrix(~ colData$group )
colnames(parentaldesign) <- levels(colData$group)

# The TMM normalization
parentalobject <- calcNormFactors(parentalobject)
parentalobject <- estimateCommonDisp(parentalobject)
parentalobject <- estimateTagwiseDisp(parentalobject)
parentalobject <- estimateDisp(parentalobject, parentaldesign)
parentalobject <- estimateGLMCommonDisp(parentalobject, parentaldesign, verbose=TRUE)
parentalobject <- estimateGLMTrendedDisp(parentalobject, parentaldesign)
parentalobject <- estimateGLMTagwiseDisp(parentalobject, parentaldesign)

#  perform likelihood ratio test and thresholded testing
fit <- glmFit( parentalobject, parentaldesign, robust=T)
tr <- glmTreat(fit, lfc = 1)
topTags(tr)
```

# plotMDS (multidimential scaling)

```{r plotMDS}
plotMDS(parentalobject, cex = 0.5)
```

```{r plotMDS-lables}
levels(colData$treatment)
col.treatment <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")[colData$treatment]

plotMDS(parentalobject,col=col.treatment, labels = colData$sex)
legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
title("Gonad Colored by Treatment")

plotMDS(parentalobject,dim=c(3,4), col=col.treatment, labels = colData$sex)
legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
title("Gonad Colored by Treatment")
```

# specify contrasts and make MA plots (currently only a subset)

```{r 01-contrasts}
# view all levels
levels(colData$group)

# subset of conrasts - sex specific comparing hatch to lay
my.contrasts <- makeContrasts(
	         FG_CB = female.gonad.control - female.gonad.bldg,
	         FG_BL = female.gonad.bldg - female.gonad.lay,
	         FG_Li3 = female.gonad.lay - female.gonad.inc.d3,
	         FG_i39 = female.gonad.inc.d3 - female.gonad.inc.d9,
	         FG_i917 = female.gonad.inc.d9 - female.gonad.inc.d17,
	         FG_i17H = female.gonad.inc.d17 - female.gonad.hatch,
	         FG_H5 = female.gonad.hatch -  female.gonad.n5,
	         FG_n59 = female.gonad.n5 - female.gonad.n9,
	         FG_n9C = female.gonad.n9 - female.gonad.control,

	         MG_CB = male.gonad.control - male.gonad.bldg,
	         MG_BL = male.gonad.bldg - male.gonad.lay,
	         MG_Li3 = male.gonad.lay - male.gonad.inc.d3,
	         MG_i39 = male.gonad.inc.d3 - male.gonad.inc.d9,
	         MG_i917 = male.gonad.inc.d9 - male.gonad.inc.d17,
	         MG_i17H = male.gonad.inc.d17 - male.gonad.hatch,
	         MG_H5 = male.gonad.hatch -  male.gonad.n5,
	         MG_n59 = male.gonad.n5 - male.gonad.n9,
	         MG_n9C = male.gonad.n9 - male.gonad.control,
levels=parentaldesign)

# female comparisons
cont <- "FG_CB"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_CB', frame.plot=F)

cont <- "FG_BL"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_BL', frame.plot=F)

cont <- "FG_Li3"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_Li3', frame.plot=F)

cont <- "FG_i39"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_i39', frame.plot=F)

cont <- "FG_i917"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_i917', frame.plot=F)

cont <- "FG_i17H"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_i17H', frame.plot=F)

cont <- "FG_H5"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_H5', frame.plot=F)

cont <- "FG_n59"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_n59', frame.plot=F)

cont <- "FG_n9C"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_n9C', frame.plot=F)

## male comparisons

cont <- "MG_CB"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_CB', frame.plot=F)

cont <- "MG_BL"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_BL', frame.plot=F)

cont <- "MG_Li3"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_Li3', frame.plot=F)

cont <- "MG_i39"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_i39', frame.plot=F)

cont <- "MG_i917"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_i917', frame.plot=F)

cont <- "MG_i17H"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_i17H', frame.plot=F)

cont <- "MG_H5"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_H5', frame.plot=F)

cont <- "MG_n59"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_n59', frame.plot=F)

cont <- "MG_n9C"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_n9C', frame.plot=F)
```

# volcano plots

```{r volcanoplots}
# from http://www.compbio.dundee.ac.uk/user/pschofield/Projects/teaching_pg/workshops/biocDGE.html#maplots

lrt <- glmLRT(fit,coef=2)
topTags(lrt)

tt <- topTags(lrt,n=10000)$table

ggplot(data=tt) + geom_point(aes(x=logFC,y=-log(FDR),color=logCPM)) +
  scale_colour_gradientn(colours=c("#000000" ,"#FF0000" ))
```

### 2019-03-19 scripts/02_limma_pit.Rmd
add males

@@
 countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
-head(colData)
 geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
@@
 
+For color coding, I used this tutorial for guidance <https://combine-australia.github.io/RNAseq-R/06-rnaseq-day1.html>.
+
 ```{r plotMDS-lables}
@@
 
-# specify contrasts and make MA plots (currently only a subset)
+# specify contrasts and make MA plots
 
@@
 	         FP_n9C = female.pituitary.n9 - female.pituitary.control,
+
+	         MP_CB = male.pituitary.control - male.pituitary.bldg,
+	         MP_BL = male.pituitary.bldg - male.pituitary.lay,
+	         MP_Li3 = male.pituitary.lay - male.pituitary.inc.d3,
+	         MP_i39 = male.pituitary.inc.d3 - male.pituitary.inc.d9,
+	         MP_i917 = male.pituitary.inc.d9 - male.pituitary.inc.d17,
+	         MP_i17H = male.pituitary.inc.d17 - male.pituitary.hatch,
+	         MP_H5 = male.pituitary.hatch -  male.pituitary.n5,
+	         MP_n59 = male.pituitary.n5 - male.pituitary.n9,
+	         MP_n9C = male.pituitary.n9 - male.pituitary.control,
 levels=parentaldesign)
 
+# female comparisons
 cont <- "FP_CB"
@@
 plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_n9C', frame.plot=F)
+
+## male comparisons
+
+cont <- "MP_CB"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MP_CB', frame.plot=F)
+
+cont <- "MP_BL"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MP_BL', frame.plot=F)
+
+cont <- "MP_Li3"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MP_Li3', frame.plot=F)
+
+cont <- "MP_i39"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MP_i39', frame.plot=F)
+
+cont <- "MP_i917"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MP_i917', frame.plot=F)
+
+cont <- "MP_i17H"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MP_i17H', frame.plot=F)
+
+cont <- "MP_H5"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MP_H5', frame.plot=F)
+
+cont <- "MP_n59"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MP_n59', frame.plot=F)
+
+cont <- "MP_n9C"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MP_n9C', frame.plot=F)
+
 ```

### 2019-03-19 scripts/04_limma_gon.Rmd
update gonad mds plot

@@
 plotMDS(parentalobject,col=col.treatment, labels = colData$sex)
-legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
+legend("bottom",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
 title("Gonad Colored by Treatment")
@@
 plotMDS(parentalobject,dim=c(3,4), col=col.treatment, labels = colData$sex)
-legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
+legend("topright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
 title("Gonad Colored by Treatment")

### 2019-03-19 scripts/02_limma_hyp.Rmd
rename to change order
add nesting to building contrast

--- 2019-03-19 scripts/03_limma_hyp.Rmd
@@
 ---
-title: "03_limma_hyp"
+title: "02_limma_hyp"
 output: md_document
@@
 plotMDS(parentalobject,col=col.treatment, labels = colData$sex)
-legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
+legend("bottomleft",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
 title("Hypothalamus Colored by Treatment")
@@
 	         MH_n9C = male.hypothalamus.n9 - male.hypothalamus.control,
+
+	         FH_n9B = female.hypothalamus.n9 - female.hypothalamus.bldg,
+	         MH_n9B = male.hypothalamus.n9 - male.hypothalamus.bldg,
 levels=parentaldesign)
@@
 
+## nesting to building
+
+cont <- "FH_n9B"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_n9B', frame.plot=F)
+
+cont <- "MH_n9B"
+summary(decideTestsDGE(
+	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+	adjust.method="fdr", p.value=0.01))
+kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_n9B', frame.plot=F)
+
 ```

### 2019-03-19 scripts/03_limma_pit.Rmd
rename to change order
copy paste to function

--- 2019-03-19 scripts/02_limma_pit.Rmd
@@
 ---
-title: "02_limma_vomm_pit"
+title: "03_limma_pit"
 output: md_document
@@
 	         FP_n9C = female.pituitary.n9 - female.pituitary.control,
+
+	         MP_CB = male.pituitary.control - male.pituitary.bldg,
+	         MP_BL = male.pituitary.bldg - male.pituitary.lay,
+	         MP_Li3 = male.pituitary.lay - male.pituitary.inc.d3,
+	         MP_i39 = male.pituitary.inc.d3 - male.pituitary.inc.d9,
+	         MP_i917 = male.pituitary.inc.d9 - male.pituitary.inc.d17,
+	         MP_i17H = male.pituitary.inc.d17 - male.pituitary.hatch,
+	         MP_H5 = male.pituitary.hatch -  male.pituitary.n5,
+	         MP_n59 = male.pituitary.n5 - male.pituitary.n9,
+	         MP_n9C = male.pituitary.n9 - male.pituitary.control,
+
+	         FP_n9B = female.pituitary.n9 - female.pituitary.bldg,
+	         MP_n9B = male.pituitary.n9 - male.pituitary.bldg,
 levels=parentaldesign)
 
-cont <- "FP_CB"
+mycontrasts <- c("FP_CB", "FP_BL", "FP_Li3", "FP_i39", "FP_i917", "FP_i17H", "FP_H5", "FP_n59", "FP_n9C",
+                 "MP_CB", "MP_BL", "MP_Li3", "MP_i39", "MP_i917", "MP_i17H", "MP_H5", "MP_n59", "MP_n9C",
+                 "FP_n9B", "MP_n9B")
+
+printplotcontrasts <- function(whichcontrast){
+  cont <- whichcontrast
   summary(decideTestsDGE(
@@
 	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_CB', frame.plot=F)
+  return(kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1))
+  return(plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main=whichcontrast, frame.plot=F))
+}
 
-cont <- "FP_BL"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_BL', frame.plot=F)
+for(i in mycontrasts){
+  printplotcontrasts(i)
+}
 
-cont <- "FP_Li3"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_Li3', frame.plot=F)
-
-cont <- "FP_i39"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_i39', frame.plot=F)
-
-cont <- "FP_i917"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_i917', frame.plot=F)
-
-cont <- "FP_i17H"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_i17H', frame.plot=F)
-
-cont <- "FP_H5"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_H5', frame.plot=F)
-
-cont <- "FP_n59"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_n59', frame.plot=F)
-
-cont <- "FP_n9C"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FP_n9C', frame.plot=F)
 ```

### 2019-03-19 scripts/03_limma_pit.Rmd
copy paste to function

@@
 	adjust.method="fdr", p.value=0.01))
-  return(kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1))
-  return(plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main=whichcontrast, frame.plot=F))
+  print(kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1))
+  print(plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main=whichcontrast, frame.plot=F))
 }

### 2019-03-19 scripts/02_limma_hyp.Rmd
now with a function

@@
 
-# female comparisons
-cont <- "FH_CB"
+mycontrasts <- c("FH_CB", "FH_BL", "FH_Li3", "FH_i39", "FH_i917", "FH_i17H", "FH_H5", "FH_n59", "FH_n9C",
+                 "MH_CB", "MH_BL", "MH_Li3", "MH_i39", "MH_i917", "MH_i17H", "MH_H5", "MH_n59", "MH_n9C",
+                 "FH_n9B", "MH_n9B")
+
+printplotcontrasts <- function(whichcontrast){
+  cont <- whichcontrast
   summary(decideTestsDGE(
@@
 	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_CB', frame.plot=F)
+  print(kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1))
+  print(plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main=whichcontrast, frame.plot=F))
+}
 
-cont <- "FH_BL"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_BL', frame.plot=F)
-
-cont <- "FH_Li3"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_Li3', frame.plot=F)
-
-cont <- "FH_i39"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_i39', frame.plot=F)
-
-cont <- "FH_i917"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_i917', frame.plot=F)
-
-cont <- "FH_i17H"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_i17H', frame.plot=F)
-
-cont <- "FH_H5"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_H5', frame.plot=F)
-
-cont <- "FH_n59"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_n59', frame.plot=F)
-
-cont <- "FH_n9C"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_n9C', frame.plot=F)
-
-## male comparisons
-
-cont <- "MH_CB"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_CB', frame.plot=F)
-
-cont <- "MH_BL"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_BL', frame.plot=F)
-
-cont <- "MH_Li3"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_Li3', frame.plot=F)
-
-cont <- "MH_i39"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_i39', frame.plot=F)
-
-cont <- "MH_i917"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_i917', frame.plot=F)
-
-cont <- "MH_i17H"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_i17H', frame.plot=F)
-
-cont <- "MH_H5"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_H5', frame.plot=F)
-
-cont <- "MH_n59"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_n59', frame.plot=F)
-
-cont <- "MH_n9C"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_n9C', frame.plot=F)
-
-## nesting to building
-
-cont <- "FH_n9B"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FH_n9B', frame.plot=F)
-
-cont <- "MH_n9B"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MH_n9B', frame.plot=F)
+for(i in mycontrasts){
+  printplotcontrasts(i)
+}
 

### 2019-03-19 scripts/04_limma_gon.Rmd
two new contrasts, table not kable

@@
 	         MG_n9C = male.gonad.n9 - male.gonad.control,
+
+	         FG_n9B = female.gonad.n9 - female.gonad.bldg,
+	         MG_n9B = male.gonad.n9 - male.gonad.bldg,
 levels=parentaldesign)
 
-# female comparisons
-cont <- "FG_CB"
-summary(decideTestsDGE(
+mycontrasts <- c("FG_CB", "FG_BL", "FG_Li3", "FG_i39", "FG_i917", "FG_i17H", "FG_H5", "FG_n59", "FG_n9C",
+                 "MG_CB", "MG_BL", "MG_Li3", "MG_i39", "MG_i917", "MG_i17H", "MG_H5", "MG_n59", "MG_n9C",
+                 "FG_n9B", "MG_n9B")
+
+printplotcontrasts <- function(whichcontrast){
+  cont <- whichcontrast
+  print(summary(decideTestsDGE(
     glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_CB', frame.plot=F)
+    adjust.method="fdr", p.value=0.01)))
+  print(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+  print(plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main=whichcontrast, frame.plot=F))
+}
 
-cont <- "FG_BL"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_BL', frame.plot=F)
+for(i in mycontrasts){
+  printplotcontrasts(i)
+}
 
-cont <- "FG_Li3"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_Li3', frame.plot=F)
-
-cont <- "FG_i39"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_i39', frame.plot=F)
-
-cont <- "FG_i917"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_i917', frame.plot=F)
-
-cont <- "FG_i17H"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_i17H', frame.plot=F)
-
-cont <- "FG_H5"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_H5', frame.plot=F)
-
-cont <- "FG_n59"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_n59', frame.plot=F)
-
-cont <- "FG_n9C"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='FG_n9C', frame.plot=F)
-
-## male comparisons
-
-cont <- "MG_CB"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_CB', frame.plot=F)
-
-cont <- "MG_BL"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_BL', frame.plot=F)
-
-cont <- "MG_Li3"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_Li3', frame.plot=F)
-
-cont <- "MG_i39"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_i39', frame.plot=F)
-
-cont <- "MG_i917"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_i917', frame.plot=F)
-
-cont <- "MG_i17H"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_i17H', frame.plot=F)
-
-cont <- "MG_H5"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_H5', frame.plot=F)
-
-cont <- "MG_n59"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_n59', frame.plot=F)
-
-cont <- "MG_n9C"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='MG_n9C', frame.plot=F)
 ```

### 2019-03-19 scripts/02_limma_hyp.Rmd
remove kable tables

@@
   cont <- whichcontrast
-  summary(decideTestsDGE(
+  print(summary(decideTestsDGE(
     glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-  print(kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1))
+    adjust.method="fdr", p.value=0.01)))
+  print(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
   print(plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main=whichcontrast, frame.plot=F))

### 2019-03-19 scripts/03_limma_pit.Rmd
move legend

@@
 plotMDS(parentalobject,col=col.treatment, labels = colData$sex)
-legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
+legend("bottomleft",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
 title("Pituitary Colored by Treatment")

### 2019-04-03 analysis/02_limma_hyp.Rmd
moved printplotcontrasts function

--- 2019-03-19 scripts/02_limma_hyp.Rmd
@@
 library(cowplot)
-library(ggplot2)
+
+# load custom functions
+source("../R/functions.R")
 
@@
 
+# create a list with all the two way contrasts
 mycontrasts <- c("FH_CB", "FH_BL", "FH_Li3", "FH_i39", "FH_i917", "FH_i17H", "FH_H5", "FH_n59", "FH_n9C",
@@
 
-printplotcontrasts <- function(whichcontrast){
-  cont <- whichcontrast
-  print(summary(decideTestsDGE(
-    glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-    adjust.method="fdr", p.value=0.01)))
-  print(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-  print(plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main=whichcontrast, frame.plot=F))
-}
+# use the printplotcontrasts function to print summary stats and a volcano plot
 
@@
 }
-
 ```

### 2019-04-03 analysis/03_limma_pit.Rmd
moved printplotcontrasts function

--- 2019-03-19 scripts/03_limma_pit.Rmd
@@
 
+# create a list with all the two way contrasts
+
 mycontrasts <- c("FP_CB", "FP_BL", "FP_Li3", "FP_i39", "FP_i917", "FP_i17H", "FP_H5", "FP_n59", "FP_n9C",

### 2019-04-03 analysis/04_limma_gon.Rmd
moved printplotcontrasts function

--- 2019-03-19 scripts/04_limma_gon.Rmd
@@
 
+# create a list with all the two way contrasts
+
 mycontrasts <- c("FG_CB", "FG_BL", "FG_Li3", "FG_i39", "FG_i917", "FG_i17H", "FG_H5", "FG_n59", "FG_n9C",

### 2019-04-03 R/functions.R

@@ -1,0 +1,13 @@
+# printplotcontrasts prints 
+# 1) a summary of the DEGs for a given 2 way contrast
+# 2) results of the top 5 DEGs
+# 3) a volcano plot of the 2 way contrast
+
+printplotcontrasts <- function(whichcontrast){
+  cont <- whichcontrast
+  print(summary(decideTestsDGE(
+    glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
+    adjust.method="fdr", p.value=0.01)))
+  print(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
+  print(plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main=whichcontrast, frame.plot=F))
+}

### 2019-04-03 analysis/04_limma_gon_F.Rmd
female gonad without control

--- 2019-04-03 analysis/04_limma_gon.Rmd
@@
 ---
-title: "04_limma_gon"
+title: "04_limma_gon_F"
 output: md_document
@@
 
-DESeq2 is _not_ recommended for experiments with more than 100 samples ([see Mike Love's post](https://mikelove.wordpress.com/2016/09/28/deseq2-or-edger/)), so I decided to try the limma package.
-I followed [this tutorial](https://www.bioconductor.org/packages/devel/workflows/vignettes/RNAseq123/inst/doc/limmaWorkflow.html).
+This anlaysis will _exclude_ the control timepoint and will example males and females separately.
 
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/gon/',cache=TRUE)
-
+knitr::opts_chunk$set(fig.path = '../figures/gon_F/',cache=TRUE)
 ```
 
-First, I read in the data I processed in 00_datawrangling.Rmd.
+# Females first
 
@@
   dplyr::filter(grepl('gonad', tissue)) %>%
+  dplyr::filter(treatment != "control") %>%
+  dplyr::filter(sex == "female") %>%
   droplevels()
@@
 # print sample sizes
-colData %>% select(sex,treatment, tissue)  %>%  summary()
+colData %>% select(treatment, tissue)  %>%  summary()
 
@@
 topTags(tr)
+head(tr$table)
 ```
@@
 ```{r plotMDS}
-plotMDS(parentalobject, cex = 0.5)
+levels(colData$treatment)
+col.treatment <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")[colData$treatment]
+myfill <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")
+
+plotMDS(parentalobject, cex = 0.5, labels = colData$treatment, col=col.treatment)
+title("Female Gonad MDS 1v2")
+plotMDS(parentalobject, cex = 0.5, labels = colData$treatment, col=col.treatment, dim=c(3,4))
+title("Female Gonad MDS 3v4")
+plotMDS(parentalobject, cex = 0.5, labels = colData$treatment, col=col.treatment, dim=c(5,6))
+title("Female Gonad MDS 5v6")
 ```
@@
 
-```{r plotMDS-lables}
-levels(colData$treatment)
-col.treatment <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")[colData$treatment]
-
-plotMDS(parentalobject,col=col.treatment, labels = colData$sex)
-legend("bottom",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
-title("Gonad Colored by Treatment")
-
-plotMDS(parentalobject,dim=c(3,4), col=col.treatment, labels = colData$sex)
-legend("topright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
-title("Gonad Colored by Treatment")
-```
-
 # specify contrasts and make MA plots
 
-```{r 01-contrasts}
+```{r MDSplots}
 # view all levels
@@
 my.contrasts <- makeContrasts(
-	         FG_CB = female.gonad.control - female.gonad.bldg,
 	         FG_BL = female.gonad.bldg - female.gonad.lay,
 	         FG_Li3 = female.gonad.lay - female.gonad.inc.d3,
-	         FG_i39 = female.gonad.inc.d3 - female.gonad.inc.d9,
-	         FG_i917 = female.gonad.inc.d9 - female.gonad.inc.d17,
+	         FG_i3i9 = female.gonad.inc.d3 - female.gonad.inc.d9,
+	         FG_i9i17 = female.gonad.inc.d9 - female.gonad.inc.d17,
 	         FG_i17H = female.gonad.inc.d17 - female.gonad.hatch,
-	         FG_H5 = female.gonad.hatch -  female.gonad.n5,
-	         FG_n59 = female.gonad.n5 - female.gonad.n9,
-	         FG_n9C = female.gonad.n9 - female.gonad.control,
+	         FG_Hn5 = female.gonad.hatch -  female.gonad.n5,
+	         FG_n5n9 = female.gonad.n5 - female.gonad.n9,
 
-	         MG_CB = male.gonad.control - male.gonad.bldg,
-	         MG_BL = male.gonad.bldg - male.gonad.lay,
-	         MG_Li3 = male.gonad.lay - male.gonad.inc.d3,
-	         MG_i39 = male.gonad.inc.d3 - male.gonad.inc.d9,
-	         MG_i917 = male.gonad.inc.d9 - male.gonad.inc.d17,
-	         MG_i17H = male.gonad.inc.d17 - male.gonad.hatch,
-	         MG_H5 = male.gonad.hatch -  male.gonad.n5,
-	         MG_n59 = male.gonad.n5 - male.gonad.n9,
-	         MG_n9C = male.gonad.n9 - male.gonad.control,
+	         FG_Bi3 = female.gonad.bldg - female.gonad.inc.d3,
+	         FG_Bi9 = female.gonad.bldg - female.gonad.inc.d9,
+	         FG_Bi17 = female.gonad.bldg - female.gonad.inc.d17,
+	         FG_BH = female.gonad.bldg - female.gonad.hatch,
+	         FG_Bn5 = female.gonad.bldg - female.gonad.n5,
+	         FG_Bn9 = female.gonad.bldg - female.gonad.n9,
 
-	         FG_n9B = female.gonad.n9 - female.gonad.bldg,
-	         MG_n9B = male.gonad.n9 - male.gonad.bldg,
+	         FG_Li9 = female.gonad.lay - female.gonad.inc.d9,
+           FG_Li17 = female.gonad.lay - female.gonad.inc.d17, 
+	         FG_LH = female.gonad.lay - female.gonad.hatch,
+	         FG_Ln5 = female.gonad.lay - female.gonad.n5,
+	         FG_Ln9 = female.gonad.lay - female.gonad.n9,
+
+	         FG_i3i17 = female.gonad.inc.d3 - female.gonad.inc.d17,
+	         FG_i3H = female.gonad.inc.d3 - female.gonad.hatch,
+	         FG_i3n5 = female.gonad.inc.d3 - female.gonad.n5,
+	         FG_i3n9 = female.gonad.inc.d3 - female.gonad.n9,
+
+	         FG_i9H = female.gonad.inc.d9 - female.gonad.hatch,
+	         FG_i9n5 = female.gonad.inc.d9 - female.gonad.n5,
+	         FG_i9n9 = female.gonad.inc.d9 - female.gonad.n9,
+
+	         FG_i17n5 = female.gonad.inc.d17 - female.gonad.n5,
+	         FG_i17n9 = female.gonad.inc.d17 - female.gonad.n9,
+
+	         FG_Hn9 = female.gonad.hatch -  female.gonad.n9,
+
 levels=parentaldesign)
+
+colnames(my.contrasts)
 
@@
 
-mycontrasts <- c("FG_CB", "FG_BL", "FG_Li3", "FG_i39", "FG_i917", "FG_i17H", "FG_H5", "FG_n59", "FG_n9C",
-                 "MG_CB", "MG_BL", "MG_Li3", "MG_i39", "MG_i917", "MG_i17H", "MG_H5", "MG_n59", "MG_n9C",
-                 "FG_n9B", "MG_n9B")
+mycontrasts <- colnames(my.contrasts)
+mycontrasts
 
@@
 
-# volcano plots
-
 ```{r volcanoplots}
-# from http://www.compbio.dundee.ac.uk/user/pschofield/Projects/teaching_pg/workshops/biocDGE.html#maplots
-
-lrt <- glmLRT(fit,coef=2)
-topTags(lrt)
-
-tt <- topTags(lrt,n=10000)$table
-
-ggplot(data=tt) + geom_point(aes(x=logFC,y=-log(FDR),color=logCPM)) +
-  scale_colour_gradientn(colours=c("#000000" ,"#FF0000" ))
+for(i in mycontrasts){
+  plotVolcanos(i)
+}
 ```

### 2019-04-03 analysis/04_limma_gon_M.Rmd
fixed axis limits

--- 2019-04-03 analysis/04_limma_gon_F.Rmd
@@
 ---
-title: "04_limma_gon_F"
+title: "04_limma_gon_M"
 output: md_document
 ---
-
-This anlaysis will _exclude_ the control timepoint and will example males and females separately.
 
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/gon_F/',cache=TRUE)
+knitr::opts_chunk$set(fig.path = '../figures/gon_M/',cache=TRUE)
 ```
 
-# Females first
+# Male Gonad Data
+
+This anlaysis will _exclude_ the control timepoint and only examine males.
 
@@
   dplyr::filter(treatment != "control") %>%
-  dplyr::filter(sex == "female") %>%
+  dplyr::filter(sex == "male") %>%
   droplevels()
@@
 plotMDS(parentalobject, cex = 0.5, labels = colData$treatment, col=col.treatment)
-title("Female Gonad MDS 1v2")
+title("Male Gonad MDS 1v2")
 plotMDS(parentalobject, cex = 0.5, labels = colData$treatment, col=col.treatment, dim=c(3,4))
-title("Female Gonad MDS 3v4")
+title("Male Gonad MDS 3v4")
 plotMDS(parentalobject, cex = 0.5, labels = colData$treatment, col=col.treatment, dim=c(5,6))
-title("Female Gonad MDS 5v6")
+title("Male Gonad MDS 5v6")
 ```
@@
 my.contrasts <- makeContrasts(
-	         FG_BL = female.gonad.bldg - female.gonad.lay,
-	         FG_Li3 = female.gonad.lay - female.gonad.inc.d3,
-	         FG_i3i9 = female.gonad.inc.d3 - female.gonad.inc.d9,
-	         FG_i9i17 = female.gonad.inc.d9 - female.gonad.inc.d17,
-	         FG_i17H = female.gonad.inc.d17 - female.gonad.hatch,
-	         FG_Hn5 = female.gonad.hatch -  female.gonad.n5,
-	         FG_n5n9 = female.gonad.n5 - female.gonad.n9,
+	         MG_BL = male.gonad.bldg - male.gonad.lay,
+	         MG_Li3 = male.gonad.lay - male.gonad.inc.d3,
+	         MG_i3i9 = male.gonad.inc.d3 - male.gonad.inc.d9,
+	         MG_i9i17 = male.gonad.inc.d9 - male.gonad.inc.d17,
+	         MG_i17H = male.gonad.inc.d17 - male.gonad.hatch,
+	         MG_Hn5 = male.gonad.hatch -  male.gonad.n5,
+	         MG_n5n9 = male.gonad.n5 - male.gonad.n9,
 
-	         FG_Bi3 = female.gonad.bldg - female.gonad.inc.d3,
-	         FG_Bi9 = female.gonad.bldg - female.gonad.inc.d9,
-	         FG_Bi17 = female.gonad.bldg - female.gonad.inc.d17,
-	         FG_BH = female.gonad.bldg - female.gonad.hatch,
-	         FG_Bn5 = female.gonad.bldg - female.gonad.n5,
-	         FG_Bn9 = female.gonad.bldg - female.gonad.n9,
+	         MG_Bi3 = male.gonad.bldg - male.gonad.inc.d3,
+	         MG_Bi9 = male.gonad.bldg - male.gonad.inc.d9,
+	         MG_Bi17 = male.gonad.bldg - male.gonad.inc.d17,
+	         MG_BH = male.gonad.bldg - male.gonad.hatch,
+	         MG_Bn5 = male.gonad.bldg - male.gonad.n5,
+	         MG_Bn9 = male.gonad.bldg - male.gonad.n9,
 
-	         FG_Li9 = female.gonad.lay - female.gonad.inc.d9,
-           FG_Li17 = female.gonad.lay - female.gonad.inc.d17, 
-	         FG_LH = female.gonad.lay - female.gonad.hatch,
-	         FG_Ln5 = female.gonad.lay - female.gonad.n5,
-	         FG_Ln9 = female.gonad.lay - female.gonad.n9,
+	         MG_Li9 = male.gonad.lay - male.gonad.inc.d9,
+           MG_Li17 = male.gonad.lay - male.gonad.inc.d17, 
+	         MG_LH = male.gonad.lay - male.gonad.hatch,
+	         MG_Ln5 = male.gonad.lay - male.gonad.n5,
+	         MG_Ln9 = male.gonad.lay - male.gonad.n9,
 
-	         FG_i3i17 = female.gonad.inc.d3 - female.gonad.inc.d17,
-	         FG_i3H = female.gonad.inc.d3 - female.gonad.hatch,
-	         FG_i3n5 = female.gonad.inc.d3 - female.gonad.n5,
-	         FG_i3n9 = female.gonad.inc.d3 - female.gonad.n9,
+	         MG_i3i17 = male.gonad.inc.d3 - male.gonad.inc.d17,
+	         MG_i3H = male.gonad.inc.d3 - male.gonad.hatch,
+	         MG_i3n5 = male.gonad.inc.d3 - male.gonad.n5,
+	         MG_i3n9 = male.gonad.inc.d3 - male.gonad.n9,
 
-	         FG_i9H = female.gonad.inc.d9 - female.gonad.hatch,
-	         FG_i9n5 = female.gonad.inc.d9 - female.gonad.n5,
-	         FG_i9n9 = female.gonad.inc.d9 - female.gonad.n9,
+	         MG_i9H = male.gonad.inc.d9 - male.gonad.hatch,
+	         MG_i9n5 = male.gonad.inc.d9 - male.gonad.n5,
+	         MG_i9n9 = male.gonad.inc.d9 - male.gonad.n9,
 
-	         FG_i17n5 = female.gonad.inc.d17 - female.gonad.n5,
-	         FG_i17n9 = female.gonad.inc.d17 - female.gonad.n9,
+	         MG_i17n5 = male.gonad.inc.d17 - male.gonad.n5,
+	         MG_i17n9 = male.gonad.inc.d17 - male.gonad.n9,
 
-	         FG_Hn9 = female.gonad.hatch -  female.gonad.n9,
+	         MG_Hn9 = male.gonad.hatch -  male.gonad.n9,
 

### 2019-04-03 analysis/02_limma_hyp_grouped.Rmd
new groupings

--- 2019-04-03 analysis/04_limma_gon_F.Rmd
@@
 ---
-title: "04_limma_gon_F"
+title: "02_limma_hyp_grouped"
 output: md_document
 ---
-
-This anlaysis will _exclude_ the control timepoint and will example males and females separately.
 
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/gon_F/',cache=TRUE)
+knitr::opts_chunk$set(fig.path = '../figures/hyp_grouped/',cache=TRUE)
 ```
 
-# Females first
+This anlaysis will _exclude_ the control timepoint but _combine_ incubation and nestling timepoints.
 
@@
 
+```{r}
+# making new groups
+colData$group <- NULL
+colData$tempgroup <- ifelse(colData$treatment == "bldg", "bldg",
+                  ifelse(colData$treatment == "control", "control",
+                   ifelse(colData$treatment == "hatch", "hatch",
+                    ifelse(grepl("inc", colData$treatment), "inc",
+                     ifelse(colData$treatment == "lay", "lay", "nestl")))))
+colData$tempgroup <- as.factor(colData$tempgroup)
+
+colData$group <- paste(colData$sex, colData$tempgroup, sep = "")
+colData$group <- as.factor(colData$group)
+str(colData$group)
+```
+
 ```{r filter}
 colData <- colData %>%
-  dplyr::filter(grepl('gonad', tissue)) %>%
+  dplyr::filter(grepl('hypothalamus', tissue)) %>%
   dplyr::filter(treatment != "control") %>%
-  dplyr::filter(sex == "female") %>%
+  #dplyr::filter(sex == "female") %>%
   droplevels()
@@
 # print sample sizes
-colData %>% select(treatment, tissue)  %>%  summary()
+colData %>% select(group, tissue)  %>%  summary()
 
@@
 ```
-
-Then, I followed the steps from <https://github.com/macmanes-lab/RockDove/blob/master/parental_care/parental_analysis.Rmd>.
 
@@
 ```{r plotMDS}
-levels(colData$treatment)
-col.treatment <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")[colData$treatment]
-myfill <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")
+levels(colData$group)
+col.group <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99",
+               "#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99")[colData$group]
+myfill <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99",
+            "#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99")
 
-plotMDS(parentalobject, cex = 0.5, labels = colData$treatment, col=col.treatment)
-title("Female Gonad MDS 1v2")
-plotMDS(parentalobject, cex = 0.5, labels = colData$treatment, col=col.treatment, dim=c(3,4))
-title("Female Gonad MDS 3v4")
-plotMDS(parentalobject, cex = 0.5, labels = colData$treatment, col=col.treatment, dim=c(5,6))
-title("Female Gonad MDS 5v6")
+plotMDS(parentalobject, cex = 0.5, labels = colData$group, col=col.group)
+title("MDS 1v2")
+plotMDS(parentalobject, cex = 0.5, labels = colData$group, col=col.group, dim=c(3,4))
+title("MDS 3v4")
+plotMDS(parentalobject, cex = 0.5, labels = colData$group, col=col.group, dim=c(5,6))
+title("MDS 5v6")
 ```
@@
 my.contrasts <- makeContrasts(
-	         FG_BL = female.gonad.bldg - female.gonad.lay,
-	         FG_Li3 = female.gonad.lay - female.gonad.inc.d3,
-	         FG_i3i9 = female.gonad.inc.d3 - female.gonad.inc.d9,
-	         FG_i9i17 = female.gonad.inc.d9 - female.gonad.inc.d17,
-	         FG_i17H = female.gonad.inc.d17 - female.gonad.hatch,
-	         FG_Hn5 = female.gonad.hatch -  female.gonad.n5,
-	         FG_n5n9 = female.gonad.n5 - female.gonad.n9,
+	         femaleBL = femalebldg - femalelay,
+	         femaleLI = femalelay - femaleinc,
+	         femaleIH = femaleinc - femalehatch,
+	         femaleHN = femalehatch -  femalenestl,
+	         femaleBL = femalebldg - femaleinc,
+	         femaleBL = femalebldg - femalehatch,
+	         femaleBL = femalebldg - femalenestl,
+	         femaleLH = femalelay - femalehatch,
+	         femaleLN = femalelay - femalenestl,
+	         femaleHN = femalehatch - femalenestl,
 
-	         FG_Bi3 = female.gonad.bldg - female.gonad.inc.d3,
-	         FG_Bi9 = female.gonad.bldg - female.gonad.inc.d9,
-	         FG_Bi17 = female.gonad.bldg - female.gonad.inc.d17,
-	         FG_BH = female.gonad.bldg - female.gonad.hatch,
-	         FG_Bn5 = female.gonad.bldg - female.gonad.n5,
-	         FG_Bn9 = female.gonad.bldg - female.gonad.n9,
+	         maleBL = malebldg - malelay,
+	         maleLI = malelay - maleinc,
+	         maleIH = maleinc - malehatch,
+	         maleHN = malehatch -  malenestl,
+	         maleBL = malebldg - maleinc,
+	         maleBL = malebldg - malehatch,
+	         maleBL = malebldg - malenestl,
+	         maleLH = malelay - malehatch,
+	         maleLN = malelay - malenestl,
+	         maleHN = malehatch - malenestl,
 
-	         FG_Li9 = female.gonad.lay - female.gonad.inc.d9,
-           FG_Li17 = female.gonad.lay - female.gonad.inc.d17, 
-	         FG_LH = female.gonad.lay - female.gonad.hatch,
-	         FG_Ln5 = female.gonad.lay - female.gonad.n5,
-	         FG_Ln9 = female.gonad.lay - female.gonad.n9,
-
-	         FG_i3i17 = female.gonad.inc.d3 - female.gonad.inc.d17,
-	         FG_i3H = female.gonad.inc.d3 - female.gonad.hatch,
-	         FG_i3n5 = female.gonad.inc.d3 - female.gonad.n5,
-	         FG_i3n9 = female.gonad.inc.d3 - female.gonad.n9,
-
-	         FG_i9H = female.gonad.inc.d9 - female.gonad.hatch,
-	         FG_i9n5 = female.gonad.inc.d9 - female.gonad.n5,
-	         FG_i9n9 = female.gonad.inc.d9 - female.gonad.n9,
-
-	         FG_i17n5 = female.gonad.inc.d17 - female.gonad.n5,
-	         FG_i17n9 = female.gonad.inc.d17 - female.gonad.n9,
-
-	         FG_Hn9 = female.gonad.hatch -  female.gonad.n9,
-
+	         fm_B = femalebldg - malebldg,
+	         fm_L = femalelay - malelay,
+	         fm_I = femaleinc - maleinc,
+	         fm_H = femalehatch - malehatch,
+	         fm_N = femalenestl - malenestl,
 levels=parentaldesign)
 
-colnames(my.contrasts)
-
 # create a list with all the two way contrasts
-
 mycontrasts <- colnames(my.contrasts)
-mycontrasts
 

### 2019-04-03 analysis/02_limma_hyp_grouped.Rmd
remove volcano
now with all the plots

@@
 	         femaleHN = femalehatch -  femalenestl,
-	         femaleBL = femalebldg - femaleinc,
-	         femaleBL = femalebldg - femalehatch,
-	         femaleBL = femalebldg - femalenestl,
+	         femaleBI = femalebldg - femaleinc,
+	         femaleBH = femalebldg - femalehatch,
+	         femaleBN = femalebldg - femalenestl,
 	         femaleLH = femalelay - femalehatch,
@@
 	         maleHN = malehatch -  malenestl,
-	         maleBL = malebldg - maleinc,
-	         maleBL = malebldg - malehatch,
-	         maleBL = malebldg - malenestl,
+	         maleBI = malebldg - maleinc,
+	         maleBH = malebldg - malehatch,
+	         maleBN = malebldg - malenestl,
 	         maleLH = malelay - malehatch,
@@
 ```
-
-```{r volcanoplots}
-for(i in mycontrasts){
-  plotVolcanos(i) 
-}
-```

### 2019-04-05 analysis/02_DESeq2_hyp.Rmd
deseq2 attempts

---
title: "02_DESeq_hyp"
output: md_document
---

```{r}
library(tidyverse)
library(DESeq2)
library(cowplot)

# load custom functions  
source("../R/functions.R")

knitr::opts_chunk$set(fig.path = '../figures/hyp/',cache=TRUE)
```

This anlaysis will _exclude_ the control timepoint but _combine_ incubation and nestling timepoints.

```{r read}
# import "colData" which contains sample information and "countData" which contains read counts
colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
```

```{r}
# making new groups
colData$group <- NULL
colData$tempgroup <- ifelse(colData$treatment == "bldg", "bldg",
                  ifelse(colData$treatment == "control", "control",
                   ifelse(colData$treatment == "hatch", "hatch",
                    ifelse(grepl("inc", colData$treatment), "inc",
                     ifelse(colData$treatment == "lay", "lay", "nestl")))))
colData$tempgroup <- as.factor(colData$tempgroup)

colData$group <- paste(colData$sex, colData$tempgroup, sep = "")
colData$group <- as.factor(colData$group)
str(colData$group)
```

```{r filter}
colData <- colData %>%
  dplyr::filter(grepl('hypothalamus', tissue)) %>%
  #dplyr::filter(treatment != "control") %>%
  dplyr::filter(sex == "female") %>%
  droplevels()
row.names(colData) <- colData$V1

# print sample sizes
colData %>% select(group, tissue)  %>%  summary()

savecols <- as.character(colData$V1) 
savecols <- as.vector(savecols) 
countData <- countData %>% dplyr::select(one_of(savecols))

# check that row and col lenghts are equal
ncol(countData) == nrow(colData)
```

```{r deseq2}
dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ treatment )
dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
dds <- DESeq(dds) # Differential expression analysis
vsd <- vst(dds, blind=FALSE) # variance stabilized
```

```{r res}
levels(colData$treatment)
summary(results(dds, contrast=c("treatment",'control', 'bldg')))    #3502 2450
summary(results(dds, contrast=c("treatment",'control', 'lay')))     #3581 2642
summary(results(dds, contrast=c("treatment",'control', 'inc.d3')))  #3956 2802
summary(results(dds, contrast=c("treatment",'control', 'inc.d9')))  #3968 2985
summary(results(dds, contrast=c("treatment",'control', 'inc.d17'))) #3949 2724
summary(results(dds, contrast=c("treatment",'control', 'hatch')))   #3600 2660
summary(results(dds, contrast=c("treatment",'control', 'n5')))      #4080 2842
summary(results(dds, contrast=c("treatment",'control', 'n9')))      #4035 2979

summary(results(dds, contrast=c("treatment",'bldg', 'lay')))     #1   0
summary(results(dds, contrast=c("treatment",'bldg', 'inc.d3')))  #0   0 
summary(results(dds, contrast=c("treatment",'bldg', 'inc.d9')))  #5   0 
summary(results(dds, contrast=c("treatment",'bldg', 'inc.d17'))) #0   0 
summary(results(dds, contrast=c("treatment",'bldg', 'hatch')))   #1   2
summary(results(dds, contrast=c("treatment",'bldg', 'n5')))      #10  3
summary(results(dds, contrast=c("treatment",'bldg', 'n9')))      #168 124

summary(results(dds, contrast=c("treatment",'lay', 'bldg')))    #    1
summary(results(dds, contrast=c("treatment",'lay', 'inc.d3')))  #0   0   
summary(results(dds, contrast=c("treatment",'lay', 'inc.d9')))  #5   6 
summary(results(dds, contrast=c("treatment",'lay', 'inc.d17'))) #2   2 
summary(results(dds, contrast=c("treatment",'lay', 'hatch')))   #3   8 
summary(results(dds, contrast=c("treatment",'lay', 'n5')))      #1   0
summary(results(dds, contrast=c("treatment",'lay', 'n9')))      #135 194

summary(results(dds, contrast=c("treatment",'inc.d3', 'bldg')))    #0 0 
summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))     #0 0 
summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d9')))  #1 0
summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d17'))) #0 0   
summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))     #0 0 
summary(results(dds, contrast=c("treatment",'inc.d3', 'n5')))      #0 0 
summary(results(dds, contrast=c("treatment",'inc.d3', 'n9')))      #0 0

summary(results(dds, contrast=c("treatment",'inc.d9', 'inc.d17'))) #2 3  
summary(results(dds, contrast=c("treatment",'inc.d9', 'lay')))     #6 5
summary(results(dds, contrast=c("treatment",'inc.d9', 'n5')))      #0 1
summary(results(dds, contrast=c("treatment",'inc.d9', 'n9')))      #0 2

summary(results(dds, contrast=c("treatment",'inc.d17', 'lay'))) #2   2
summary(results(dds, contrast=c("treatment",'inc.d17', 'n5')))  #2   4
summary(results(dds, contrast=c("treatment",'inc.d17', 'n9')))  #131 136

summary(results(dds, contrast=c("treatment",'hatch', 'bldg')))    #2   1    
summary(results(dds, contrast=c("treatment",'hatch', 'lay')))     #8   3  
summary(results(dds, contrast=c("treatment",'hatch', 'inc.d3')))  #524 406 
summary(results(dds, contrast=c("treatment",'hatch', 'inc.d9')))  #1106 1169 
summary(results(dds, contrast=c("treatment",'hatch', 'inc.d17'))) #3    0 
summary(results(dds, contrast=c("treatment",'hatch', 'n5')))      #1076 875
summary(results(dds, contrast=c("treatment",'hatch', 'n9')))      #898  845

summary(results(dds, contrast=c("treatment",'n5', 'bldg')))    #3   10 
summary(results(dds, contrast=c("treatment",'n5', 'lay')))     #0   1 
summary(results(dds, contrast=c("treatment",'n5', 'inc.d3')))  #0   0 
summary(results(dds, contrast=c("treatment",'n5', 'inc.d9')))  #1   0 
summary(results(dds, contrast=c("treatment",'n5', 'inc.d17'))) #4   2 
summary(results(dds, contrast=c("treatment",'n5', 'hatch')))   #875 1076 
summary(results(dds, contrast=c("treatment",'n5', 'n9')))      #0   0

summary(results(dds, contrast=c("treatment",'n9', 'bldg')))    #124 168 
summary(results(dds, contrast=c("treatment",'n9', 'lay')))     #194 135 
summary(results(dds, contrast=c("treatment",'n9', 'inc.d3')))  #0   0 
summary(results(dds, contrast=c("treatment",'n9', 'inc.d9')))  #2   0 
summary(results(dds, contrast=c("treatment",'n9', 'inc.d17'))) #136 131 
summary(results(dds, contrast=c("treatment",'n9', 'hatch')))   #845 898 
summary(results(dds, contrast=c("treatment",'n9', 'n5')))      #0   0 
```

```{r totalDEGs}
levels(colData$treatment)

totalDEGs(c("treatment", 'bldg','lay'))
totalDEGs(c("treatment", 'bldg','inc.d3'))
totalDEGs(c("treatment", 'bldg','inc.d9'))
totalDEGs(c("treatment", 'bldg','inc.d17'))
totalDEGs(c("treatment", 'bldg','hatch'))
totalDEGs(c("treatment", 'bldg','n5'))
totalDEGs(c("treatment", 'bldg','n9'))

totalDEGs(c("treatment", 'lay','bldg'))
totalDEGs(c("treatment", 'lay','inc.d3'))
totalDEGs(c("treatment", 'lay','inc.d9'))
totalDEGs(c("treatment", 'lay','inc.d17'))
totalDEGs(c("treatment", 'lay','hatch'))
totalDEGs(c("treatment", 'lay','n5'))
totalDEGs(c("treatment", 'lay','n9'))

totalDEGs(c("treatment", 'inc.d3','bldg'))
totalDEGs(c("treatment", 'inc.d3','lay'))
totalDEGs(c("treatment", 'inc.d3','inc.d9'))
totalDEGs(c("treatment", 'inc.d3','inc.d17'))
totalDEGs(c("treatment", 'inc.d3','hatch'))
totalDEGs(c("treatment", 'inc.d3','n5'))
totalDEGs(c("treatment", 'inc.d3','n9'))

totalDEGs(c("treatment", 'inc.d9','bldg'))
totalDEGs(c("treatment", 'inc.d9','lay'))
totalDEGs(c("treatment", 'inc.d9','inc.d3'))
totalDEGs(c("treatment", 'inc.d9','inc.d17'))
totalDEGs(c("treatment", 'inc.d9','hatch'))
totalDEGs(c("treatment", 'inc.d9','n5'))
totalDEGs(c("treatment", 'inc.d9','n9'))

totalDEGs(c("treatment", 'inc.d17','bldg'))
totalDEGs(c("treatment", 'inc.d17','lay'))
totalDEGs(c("treatment", 'inc.d17','inc.d3'))
totalDEGs(c("treatment", 'inc.d17','inc.d9'))
totalDEGs(c("treatment", 'inc.d17','hatch'))
totalDEGs(c("treatment", 'inc.d17','n5'))
totalDEGs(c("treatment", 'inc.d17','n9'))

totalDEGs(c("treatment", 'hatch','bldg'))
totalDEGs(c("treatment", 'hatch','lay'))
totalDEGs(c("treatment", 'hatch','inc.d3'))
totalDEGs(c("treatment", 'hatch','inc.d9'))
totalDEGs(c("treatment", 'hatch','inc.d17'))
totalDEGs(c("treatment", 'hatch','n5'))
totalDEGs(c("treatment", 'hatch','n9'))

totalDEGs(c("treatment", 'n5','bldg'))
totalDEGs(c("treatment", 'n5','lay'))
totalDEGs(c("treatment", 'n5','inc.d3'))
totalDEGs(c("treatment", 'n5','inc.d9'))
totalDEGs(c("treatment", 'n5','inc.d17'))
totalDEGs(c("treatment", 'n5','hatch'))
totalDEGs(c("treatment", 'n5','n9'))

totalDEGs(c("treatment", 'n9','bldg'))
totalDEGs(c("treatment", 'n9','lay'))
totalDEGs(c("treatment", 'n9','inc.d3'))
totalDEGs(c("treatment", 'n9','inc.d9'))
totalDEGs(c("treatment", 'n9','inc.d17'))
totalDEGs(c("treatment", 'n9','hatch'))
totalDEGs(c("treatment", 'n9','n5'))
```

```{r PCA, message=FALSE, warning=FALSE}
levels(colData$treatment)
colData$treatment <- factor(colData$treatment, levels = 
                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
                                "inc.d17", "hatch", "n5", "n9"))

# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
percentVar <- round(100 * attr(pcadata, "percentVar"))
percentVar

ggplot(pcadata, aes(PC1, PC2,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25)

ggplot(pcadata, aes(PC2, PC3,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC2: ", percentVar[2],"% variance")) +
  ylab(paste0("PC3: ", percentVar[3],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25)

ggplot(pcadata, aes(PC3, PC4,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) 
```

PCA statistics

```{r}
summary(aov(PC1 ~ treatment, data=pcadata)) 
TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment")

summary(aov(PC2 ~ treatment, data=pcadata)) 
TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment")

summary(aov(PC3 ~ treatment, data=pcadata))

summary(aov(PC4 ~ treatment, data=pcadata))

summary(aov(PC5 ~ treatment, data=pcadata)) 
TukeyHSD(aov(PC5 ~ treatment, data=pcadata), which = "treatment")

summary(aov(PC6 ~ treatment, data=pcadata)) 
```

### 2019-04-05 analysis/03_DESeq2_pit.Rmd
deseq2 attempts

--- 2019-04-05 analysis/02_DESeq2_hyp.Rmd
@@
 ---
-title: "02_DESeq_hyp"
+title: "03_DESeq_pit"
 output: md_document
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/hyp/',cache=TRUE)
+knitr::opts_chunk$set(fig.path = '../figures/pit/',cache=TRUE)
 ```
@@
 colData <- colData %>%
-  dplyr::filter(grepl('hypothalamus', tissue)) %>%
+  dplyr::filter(grepl('pituitary', tissue)) %>%
   #dplyr::filter(treatment != "control") %>%
@@
 levels(colData$treatment)
-summary(results(dds, contrast=c("treatment",'control', 'bldg')))    #3502 2450
-summary(results(dds, contrast=c("treatment",'control', 'lay')))     #3581 2642
-summary(results(dds, contrast=c("treatment",'control', 'inc.d3')))  #3956 2802
-summary(results(dds, contrast=c("treatment",'control', 'inc.d9')))  #3968 2985
-summary(results(dds, contrast=c("treatment",'control', 'inc.d17'))) #3949 2724
-summary(results(dds, contrast=c("treatment",'control', 'hatch')))   #3600 2660
-summary(results(dds, contrast=c("treatment",'control', 'n5')))      #4080 2842
-summary(results(dds, contrast=c("treatment",'control', 'n9')))      #4035 2979
+summary(results(dds, contrast=c("treatment",'control', 'bldg'))) 
+summary(results(dds, contrast=c("treatment",'control', 'lay')))     
+summary(results(dds, contrast=c("treatment",'control', 'inc.d3')))  
+summary(results(dds, contrast=c("treatment",'control', 'inc.d9')))   
+summary(results(dds, contrast=c("treatment",'control', 'inc.d17')))  
+summary(results(dds, contrast=c("treatment",'control', 'hatch')))    
+summary(results(dds, contrast=c("treatment",'control', 'n5')))      
+summary(results(dds, contrast=c("treatment",'control', 'n9')))
 
-summary(results(dds, contrast=c("treatment",'bldg', 'lay')))     #1   0
-summary(results(dds, contrast=c("treatment",'bldg', 'inc.d3')))  #0   0 
-summary(results(dds, contrast=c("treatment",'bldg', 'inc.d9')))  #5   0 
-summary(results(dds, contrast=c("treatment",'bldg', 'inc.d17'))) #0   0 
-summary(results(dds, contrast=c("treatment",'bldg', 'hatch')))   #1   2
-summary(results(dds, contrast=c("treatment",'bldg', 'n5')))      #10  3
-summary(results(dds, contrast=c("treatment",'bldg', 'n9')))      #168 124
+summary(results(dds, contrast=c("treatment",'bldg', 'lay')))      
+summary(results(dds, contrast=c("treatment",'bldg', 'inc.d3')))   
+summary(results(dds, contrast=c("treatment",'bldg', 'inc.d9')))   
+summary(results(dds, contrast=c("treatment",'bldg', 'inc.d17')))  
+summary(results(dds, contrast=c("treatment",'bldg', 'hatch')))   
+summary(results(dds, contrast=c("treatment",'bldg', 'n5')))      
+summary(results(dds, contrast=c("treatment",'bldg', 'n9')))
 
-summary(results(dds, contrast=c("treatment",'lay', 'bldg')))    #    1
-summary(results(dds, contrast=c("treatment",'lay', 'inc.d3')))  #0   0   
-summary(results(dds, contrast=c("treatment",'lay', 'inc.d9')))  #5   6 
-summary(results(dds, contrast=c("treatment",'lay', 'inc.d17'))) #2   2 
-summary(results(dds, contrast=c("treatment",'lay', 'hatch')))   #3   8 
-summary(results(dds, contrast=c("treatment",'lay', 'n5')))      #1   0
-summary(results(dds, contrast=c("treatment",'lay', 'n9')))      #135 194
+summary(results(dds, contrast=c("treatment",'lay', 'bldg')))     
+summary(results(dds, contrast=c("treatment",'lay', 'inc.d3')))    
+summary(results(dds, contrast=c("treatment",'lay', 'inc.d9')))   
+summary(results(dds, contrast=c("treatment",'lay', 'inc.d17')))  
+summary(results(dds, contrast=c("treatment",'lay', 'hatch')))    
+summary(results(dds, contrast=c("treatment",'lay', 'n5')))       
+summary(results(dds, contrast=c("treatment",'lay', 'n9')))
 
-summary(results(dds, contrast=c("treatment",'inc.d3', 'bldg')))    #0 0 
-summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))     #0 0 
-summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d9')))  #1 0
-summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d17'))) #0 0   
-summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))     #0 0 
-summary(results(dds, contrast=c("treatment",'inc.d3', 'n5')))      #0 0 
-summary(results(dds, contrast=c("treatment",'inc.d3', 'n9')))      #0 0
+summary(results(dds, contrast=c("treatment",'inc.d3', 'bldg')))     
+summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))       
+summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d9')))   
+summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d17')))   
+summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))      
+summary(results(dds, contrast=c("treatment",'inc.d3', 'n5')))        
+summary(results(dds, contrast=c("treatment",'inc.d3', 'n9')))
 
-summary(results(dds, contrast=c("treatment",'inc.d9', 'inc.d17'))) #2 3  
-summary(results(dds, contrast=c("treatment",'inc.d9', 'lay')))     #6 5
-summary(results(dds, contrast=c("treatment",'inc.d9', 'n5')))      #0 1
-summary(results(dds, contrast=c("treatment",'inc.d9', 'n9')))      #0 2
+summary(results(dds, contrast=c("treatment",'inc.d9', 'inc.d17')))   
+summary(results(dds, contrast=c("treatment",'inc.d9', 'lay')))      
+summary(results(dds, contrast=c("treatment",'inc.d9', 'n5')))       
+summary(results(dds, contrast=c("treatment",'inc.d9', 'n9')))
 
-summary(results(dds, contrast=c("treatment",'inc.d17', 'lay'))) #2   2
-summary(results(dds, contrast=c("treatment",'inc.d17', 'n5')))  #2   4
-summary(results(dds, contrast=c("treatment",'inc.d17', 'n9')))  #131 136
+summary(results(dds, contrast=c("treatment",'inc.d17', 'lay')))  
+summary(results(dds, contrast=c("treatment",'inc.d17', 'n5')))   
+summary(results(dds, contrast=c("treatment",'inc.d17', 'n9')))
 
-summary(results(dds, contrast=c("treatment",'hatch', 'bldg')))    #2   1    
-summary(results(dds, contrast=c("treatment",'hatch', 'lay')))     #8   3  
-summary(results(dds, contrast=c("treatment",'hatch', 'inc.d3')))  #524 406 
-summary(results(dds, contrast=c("treatment",'hatch', 'inc.d9')))  #1106 1169 
-summary(results(dds, contrast=c("treatment",'hatch', 'inc.d17'))) #3    0 
-summary(results(dds, contrast=c("treatment",'hatch', 'n5')))      #1076 875
-summary(results(dds, contrast=c("treatment",'hatch', 'n9')))      #898  845
+summary(results(dds, contrast=c("treatment",'hatch', 'bldg')))      
+summary(results(dds, contrast=c("treatment",'hatch', 'lay')))       
+summary(results(dds, contrast=c("treatment",'hatch', 'inc.d3')))    
+summary(results(dds, contrast=c("treatment",'hatch', 'inc.d9')))    
+summary(results(dds, contrast=c("treatment",'hatch', 'inc.d17')))  
+summary(results(dds, contrast=c("treatment",'hatch', 'n5')))       
+summary(results(dds, contrast=c("treatment",'hatch', 'n9')))
 
-summary(results(dds, contrast=c("treatment",'n5', 'bldg')))    #3   10 
-summary(results(dds, contrast=c("treatment",'n5', 'lay')))     #0   1 
-summary(results(dds, contrast=c("treatment",'n5', 'inc.d3')))  #0   0 
-summary(results(dds, contrast=c("treatment",'n5', 'inc.d9')))  #1   0 
-summary(results(dds, contrast=c("treatment",'n5', 'inc.d17'))) #4   2 
-summary(results(dds, contrast=c("treatment",'n5', 'hatch')))   #875 1076 
-summary(results(dds, contrast=c("treatment",'n5', 'n9')))      #0   0
+summary(results(dds, contrast=c("treatment",'n5', 'bldg')))      
+summary(results(dds, contrast=c("treatment",'n5', 'lay')))       
+summary(results(dds, contrast=c("treatment",'n5', 'inc.d3')))  
+summary(results(dds, contrast=c("treatment",'n5', 'inc.d9')))   
+summary(results(dds, contrast=c("treatment",'n5', 'inc.d17')))  
+summary(results(dds, contrast=c("treatment",'n5', 'hatch')))   
+summary(results(dds, contrast=c("treatment",'n5', 'n9')))
 
-summary(results(dds, contrast=c("treatment",'n9', 'bldg')))    #124 168 
-summary(results(dds, contrast=c("treatment",'n9', 'lay')))     #194 135 
-summary(results(dds, contrast=c("treatment",'n9', 'inc.d3')))  #0   0 
-summary(results(dds, contrast=c("treatment",'n9', 'inc.d9')))  #2   0 
-summary(results(dds, contrast=c("treatment",'n9', 'inc.d17'))) #136 131 
-summary(results(dds, contrast=c("treatment",'n9', 'hatch')))   #845 898 
-summary(results(dds, contrast=c("treatment",'n9', 'n5')))      #0   0 
+summary(results(dds, contrast=c("treatment",'n9', 'bldg')))     
+summary(results(dds, contrast=c("treatment",'n9', 'lay')))      
+summary(results(dds, contrast=c("treatment",'n9', 'inc.d3')))   
+summary(results(dds, contrast=c("treatment",'n9', 'inc.d9')))   
+summary(results(dds, contrast=c("treatment",'n9', 'inc.d17')))  
+summary(results(dds, contrast=c("treatment",'n9', 'hatch')))    
+summary(results(dds, contrast=c("treatment",'n9', 'n5')))     
 ```
@@
 
-```{r}
+```{r pcastats}
 summary(aov(PC1 ~ treatment, data=pcadata)) 
@@
 summary(aov(PC5 ~ treatment, data=pcadata))
-TukeyHSD(aov(PC5 ~ treatment, data=pcadata), which = "treatment")
 

### 2019-04-05 analysis/04_DESeq2_gon.Rmd
hyp and gonad

---
title: "04_DESeq_gon"
output: md_document
---

```{r setup}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)

# load custom functions  
source("../R/functions.R")

knitr::opts_chunk$set(fig.path = '../figures/gon/',cache=TRUE)
```

This anlaysis will _exclude_ the control timepoint but _combine_ incubation and nestling timepoints.

```{r read}
# import "colData" which contains sample information and "countData" which contains read counts
colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1) 
```

```{r colDataWrangle}
# making new groups
colData$group <- NULL
colData$tempgroup <- ifelse(colData$treatment == "bldg", "bldg",
                  ifelse(colData$treatment == "control", "control",
                   ifelse(colData$treatment == "hatch", "hatch",
                    ifelse(grepl("inc", colData$treatment), "inc",
                     ifelse(colData$treatment == "lay", "lay", "nestl")))))
colData$tempgroup <- as.factor(colData$tempgroup)

colData$group <- paste(colData$sex, colData$tempgroup, sep = "")
colData$group <- as.factor(colData$group)
str(colData$group)

colData$treatment <- factor(colData$treatment, levels = 
                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
                                "inc.d17", "hatch", "n5", "n9"))
```

```{r filter}
#select samples of interest from colData
colData <- colData %>%
  dplyr::filter(grepl('gonad', tissue)) %>%
  #dplyr::filter(treatment != "control") %>%
  dplyr::filter(sex == "female") %>%
  droplevels()
row.names(colData) <- colData$V1

# print sample sizes
colData %>% select(group, tissue)  %>%  summary()

#select samples of interest from countData
savecols <- as.character(colData$V1) 
savecols <- as.vector(savecols) 
countData <- countData %>% dplyr::select(one_of(savecols))

# check that row and col lenghts are equal
ncol(countData) == nrow(colData)  
```

```{r deseq2}
dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ treatment )
dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
dds <- DESeq(dds) # Differential expression analysis
vsd <- vst(dds, blind=FALSE) # variance stabilized  
```

```{r res}
levels(colData$treatment)
summary(results(dds, contrast=c("treatment",'control', 'bldg'))) 
```

```{res2}
summary(results(dds, contrast=c("treatment",'control', 'lay')))     
summary(results(dds, contrast=c("treatment",'control', 'inc.d3')))  
summary(results(dds, contrast=c("treatment",'control', 'inc.d9')))   
summary(results(dds, contrast=c("treatment",'control', 'inc.d17')))  
summary(results(dds, contrast=c("treatment",'control', 'hatch')))    
summary(results(dds, contrast=c("treatment",'control', 'n5')))      
summary(results(dds, contrast=c("treatment",'control', 'n9')))

summary(results(dds, contrast=c("treatment",'bldg', 'lay')))      
summary(results(dds, contrast=c("treatment",'bldg', 'inc.d3')))   
summary(results(dds, contrast=c("treatment",'bldg', 'inc.d9')))   
summary(results(dds, contrast=c("treatment",'bldg', 'inc.d17')))  
summary(results(dds, contrast=c("treatment",'bldg', 'hatch')))   
summary(results(dds, contrast=c("treatment",'bldg', 'n5')))      
summary(results(dds, contrast=c("treatment",'bldg', 'n9')))

#summary(results(dds, contrast=c("treatment",'lay', 'bldg')))     
summary(results(dds, contrast=c("treatment",'lay', 'inc.d3')))    
summary(results(dds, contrast=c("treatment",'lay', 'inc.d9')))   
summary(results(dds, contrast=c("treatment",'lay', 'inc.d17')))  
summary(results(dds, contrast=c("treatment",'lay', 'hatch')))    
summary(results(dds, contrast=c("treatment",'lay', 'n5')))       
summary(results(dds, contrast=c("treatment",'lay', 'n9')))

#summary(results(dds, contrast=c("treatment",'inc.d3', 'bldg')))     
#summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))       
summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d9')))   
summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d17')))   
summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))      
summary(results(dds, contrast=c("treatment",'inc.d3', 'n5')))        
summary(results(dds, contrast=c("treatment",'inc.d3', 'n9')))

summary(results(dds, contrast=c("treatment",'inc.d9', 'inc.d17')))   
summary(results(dds, contrast=c("treatment",'inc.d9', 'lay')))      
summary(results(dds, contrast=c("treatment",'inc.d9', 'n5')))       
summary(results(dds, contrast=c("treatment",'inc.d9', 'n9')))

summary(results(dds, contrast=c("treatment",'inc.d17', 'lay')))  
summary(results(dds, contrast=c("treatment",'inc.d17', 'n5')))   
summary(results(dds, contrast=c("treatment",'inc.d17', 'n9')))

#summary(results(dds, contrast=c("treatment",'hatch', 'bldg')))      
#summary(results(dds, contrast=c("treatment",'hatch', 'lay')))       
#summary(results(dds, contrast=c("treatment",'hatch', 'inc.d3')))    
#summary(results(dds, contrast=c("treatment",'hatch', 'inc.d9')))    
summary(results(dds, contrast=c("treatment",'hatch', 'inc.d17')))  
summary(results(dds, contrast=c("treatment",'hatch', 'n5')))       
summary(results(dds, contrast=c("treatment",'hatch', 'n9')))

#summary(results(dds, contrast=c("treatment",'n5', 'bldg')))      
#summary(results(dds, contrast=c("treatment",'n5', 'lay')))       
#summary(results(dds, contrast=c("treatment",'n5', 'inc.d3')))  
#summary(results(dds, contrast=c("treatment",'n5', 'inc.d9')))   
#summary(results(dds, contrast=c("treatment",'n5', 'inc.d17')))  
summary(results(dds, contrast=c("treatment",'n5', 'hatch')))   
summary(results(dds, contrast=c("treatment",'n5', 'n9')))

#summary(results(dds, contrast=c("treatment",'n9', 'bldg')))     
#summary(results(dds, contrast=c("treatment",'n9', 'lay')))      
#summary(results(dds, contrast=c("treatment",'n9', 'inc.d3')))   
#summary(results(dds, contrast=c("treatment",'n9', 'inc.d9')))   
#summary(results(dds, contrast=c("treatment",'n9', 'inc.d17')))  
#summary(results(dds, contrast=c("treatment",'n9', 'hatch')))    
summary(results(dds, contrast=c("treatment",'n9', 'n5')))     
```

```{r totalDEGs}
levels(colData$treatment)

totalDEGs(c("treatment", 'bldg','lay'))
totalDEGs(c("treatment", 'bldg','inc.d3'))
totalDEGs(c("treatment", 'bldg','inc.d9'))
totalDEGs(c("treatment", 'bldg','inc.d17'))
totalDEGs(c("treatment", 'bldg','hatch'))
totalDEGs(c("treatment", 'bldg','n5'))
totalDEGs(c("treatment", 'bldg','n9'))

#totalDEGs(c("treatment", 'lay','bldg'))
totalDEGs(c("treatment", 'lay','inc.d3'))
totalDEGs(c("treatment", 'lay','inc.d9'))
totalDEGs(c("treatment", 'lay','inc.d17'))
totalDEGs(c("treatment", 'lay','hatch'))
totalDEGs(c("treatment", 'lay','n5'))
totalDEGs(c("treatment", 'lay','n9'))

#totalDEGs(c("treatment", 'inc.d3','bldg'))
#totalDEGs(c("treatment", 'inc.d3','lay'))
totalDEGs(c("treatment", 'inc.d3','inc.d9'))
totalDEGs(c("treatment", 'inc.d3','inc.d17'))
totalDEGs(c("treatment", 'inc.d3','hatch'))
totalDEGs(c("treatment", 'inc.d3','n5'))
totalDEGs(c("treatment", 'inc.d3','n9'))

#totalDEGs(c("treatment", 'inc.d9','bldg'))
#totalDEGs(c("treatment", 'inc.d9','lay'))
#totalDEGs(c("treatment", 'inc.d9','inc.d3'))
totalDEGs(c("treatment", 'inc.d9','inc.d17'))
totalDEGs(c("treatment", 'inc.d9','hatch'))
totalDEGs(c("treatment", 'inc.d9','n5'))
totalDEGs(c("treatment", 'inc.d9','n9'))

#totalDEGs(c("treatment", 'inc.d17','bldg'))
#totalDEGs(c("treatment", 'inc.d17','lay'))
#totalDEGs(c("treatment", 'inc.d17','inc.d3'))
#totalDEGs(c("treatment", 'inc.d17','inc.d9'))
totalDEGs(c("treatment", 'inc.d17','hatch'))
totalDEGs(c("treatment", 'inc.d17','n5'))
totalDEGs(c("treatment", 'inc.d17','n9'))

#totalDEGs(c("treatment", 'hatch','bldg'))
#totalDEGs(c("treatment", 'hatch','lay'))
#totalDEGs(c("treatment", 'hatch','inc.d3'))
#totalDEGs(c("treatment", 'hatch','inc.d9'))
totalDEGs(c("treatment", 'hatch','inc.d17'))
totalDEGs(c("treatment", 'hatch','n5'))
totalDEGs(c("treatment", 'hatch','n9'))

#totalDEGs(c("treatment", 'n5','bldg'))
#totalDEGs(c("treatment", 'n5','lay'))
#totalDEGs(c("treatment", 'n5','inc.d3'))
#totalDEGs(c("treatment", 'n5','inc.d9'))
#totalDEGs(c("treatment", 'n5','inc.d17'))
totalDEGs(c("treatment", 'n5','hatch'))
totalDEGs(c("treatment", 'n5','n9'))

#totalDEGs(c("treatment", 'n9','bldg'))
#totalDEGs(c("treatment", 'n9','lay'))
#totalDEGs(c("treatment", 'n9','inc.d3'))
#totalDEGs(c("treatment", 'n9','inc.d9'))
#totalDEGs(c("treatment", 'n9','inc.d17'))
#totalDEGs(c("treatment", 'n9','hatch'))
totalDEGs(c("treatment", 'n9','n5'))
```

```{r PCA, message=FALSE, warning=FALSE}
levels(colData$treatment)

# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
percentVar <- round(100 * attr(pcadata, "percentVar"))
percentVar

ggplot(pcadata, aes(PC1, PC2,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25)

ggplot(pcadata, aes(PC2, PC3,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC2: ", percentVar[2],"% variance")) +
  ylab(paste0("PC3: ", percentVar[3],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25)

ggplot(pcadata, aes(PC3, PC4,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25)

plotPCA(vsd, intgroup=c("treatment"))
```

PCA statistics

```{r pcastats}
summary(aov(PC1 ~ treatment, data=pcadata)) 
TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment")

summary(aov(PC2 ~ treatment, data=pcadata)) 
TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment")

summary(aov(PC3 ~ treatment, data=pcadata))

summary(aov(PC4 ~ treatment, data=pcadata))

summary(aov(PC5 ~ treatment, data=pcadata))

summary(aov(PC6 ~ treatment, data=pcadata)) 
```

```{r heatmap}
# see http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#heatmap-of-the-count-matrix
sampleDists <- dist(t(assay(vsd)))

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- NULL
colnames(sampleDistMatrix) <- colData$treatment
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors,
         fontsize = 6)
```

### 2019-04-05 analysis/02_DESeq2_hyp.Rmd
hyp and gonad
hyp deseq2

@@
 library(cowplot)
+library(RColorBrewer)
+library(pheatmap)
 
@@
 ```
+```{r heatmap}
+# see http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#heatmap-of-the-count-matrix
+sampleDists <- dist(t(assay(vsd)))
+
+sampleDistMatrix <- as.matrix(sampleDists)
+rownames(sampleDistMatrix) <- NULL
+colnames(sampleDistMatrix) <- colData$treatment
+colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
+pheatmap(sampleDistMatrix,
+         clustering_distance_rows=sampleDists,
+         clustering_distance_cols=sampleDists,
+         col=colors,
+         fontsize = 6)
+```

### 2019-04-05 analysis/02_DESeq2_hyp.Rmd
now with for loop to count degs

@@
 str(colData$group)
+
+colData$treatment <- factor(colData$treatment, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
 ```
@@
 
-```{r res}
-levels(colData$treatment) 
-summary(results(dds, contrast=c("treatment",'control', 'bldg')))    #3502 2450
-summary(results(dds, contrast=c("treatment",'control', 'lay')))     #3581 2642
-summary(results(dds, contrast=c("treatment",'control', 'inc.d3')))  #3956 2802
-summary(results(dds, contrast=c("treatment",'control', 'inc.d9')))  #3968 2985
-summary(results(dds, contrast=c("treatment",'control', 'inc.d17'))) #3949 2724
-summary(results(dds, contrast=c("treatment",'control', 'hatch')))   #3600 2660
-summary(results(dds, contrast=c("treatment",'control', 'n5')))      #4080 2842
-summary(results(dds, contrast=c("treatment",'control', 'n9')))      #4035 2979
+```{r resloop}
+#create list of groups
+a <- levels(colData$treatment)
+a
 
-summary(results(dds, contrast=c("treatment",'bldg', 'lay')))     #1   0
-summary(results(dds, contrast=c("treatment",'bldg', 'inc.d3')))  #0   0 
-summary(results(dds, contrast=c("treatment",'bldg', 'inc.d9')))  #5   0 
-summary(results(dds, contrast=c("treatment",'bldg', 'inc.d17'))) #0   0 
-summary(results(dds, contrast=c("treatment",'bldg', 'hatch')))   #1   2
-summary(results(dds, contrast=c("treatment",'bldg', 'n5')))      #10  3
-summary(results(dds, contrast=c("treatment",'bldg', 'n9')))      #168 124
+# create list of groups, with control last
+b <- a[2:9]
+b[9] <- a[1]
+b
 
-summary(results(dds, contrast=c("treatment",'lay', 'bldg')))    #    1
-summary(results(dds, contrast=c("treatment",'lay', 'inc.d3')))  #0   0   
-summary(results(dds, contrast=c("treatment",'lay', 'inc.d9')))  #5   6 
-summary(results(dds, contrast=c("treatment",'lay', 'inc.d17'))) #2   2 
-summary(results(dds, contrast=c("treatment",'lay', 'hatch')))   #3   8 
-summary(results(dds, contrast=c("treatment",'lay', 'n5')))      #1   0
-summary(results(dds, contrast=c("treatment",'lay', 'n9')))      #135 194
-
-summary(results(dds, contrast=c("treatment",'inc.d3', 'bldg')))    #0 0 
-summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))     #0 0 
-summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d9')))  #1 0
-summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d17'))) #0 0   
-summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))     #0 0 
-summary(results(dds, contrast=c("treatment",'inc.d3', 'n5')))      #0 0 
-summary(results(dds, contrast=c("treatment",'inc.d3', 'n9')))      #0 0
-
-summary(results(dds, contrast=c("treatment",'inc.d9', 'inc.d17'))) #2 3  
-summary(results(dds, contrast=c("treatment",'inc.d9', 'lay')))     #6 5
-summary(results(dds, contrast=c("treatment",'inc.d9', 'n5')))      #0 1
-summary(results(dds, contrast=c("treatment",'inc.d9', 'n9')))      #0 2
-
-summary(results(dds, contrast=c("treatment",'inc.d17', 'lay'))) #2   2
-summary(results(dds, contrast=c("treatment",'inc.d17', 'n5')))  #2   4
-summary(results(dds, contrast=c("treatment",'inc.d17', 'n9')))  #131 136
-
-summary(results(dds, contrast=c("treatment",'hatch', 'bldg')))    #2   1    
-summary(results(dds, contrast=c("treatment",'hatch', 'lay')))     #8   3  
-summary(results(dds, contrast=c("treatment",'hatch', 'inc.d3')))  #524 406 
-summary(results(dds, contrast=c("treatment",'hatch', 'inc.d9')))  #1106 1169 
-summary(results(dds, contrast=c("treatment",'hatch', 'inc.d17'))) #3    0 
-summary(results(dds, contrast=c("treatment",'hatch', 'n5')))      #1076 875
-summary(results(dds, contrast=c("treatment",'hatch', 'n9')))      #898  845
-
-summary(results(dds, contrast=c("treatment",'n5', 'bldg')))    #3   10 
-summary(results(dds, contrast=c("treatment",'n5', 'lay')))     #0   1 
-summary(results(dds, contrast=c("treatment",'n5', 'inc.d3')))  #0   0 
-summary(results(dds, contrast=c("treatment",'n5', 'inc.d9')))  #1   0 
-summary(results(dds, contrast=c("treatment",'n5', 'inc.d17'))) #4   2 
-summary(results(dds, contrast=c("treatment",'n5', 'hatch')))   #875 1076 
-summary(results(dds, contrast=c("treatment",'n5', 'n9')))      #0   0
-
-summary(results(dds, contrast=c("treatment",'n9', 'bldg')))    #124 168 
-summary(results(dds, contrast=c("treatment",'n9', 'lay')))     #194 135 
-summary(results(dds, contrast=c("treatment",'n9', 'inc.d3')))  #0   0 
-summary(results(dds, contrast=c("treatment",'n9', 'inc.d9')))  #2   0 
-summary(results(dds, contrast=c("treatment",'n9', 'inc.d17'))) #136 131 
-summary(results(dds, contrast=c("treatment",'n9', 'hatch')))   #845 898 
-summary(results(dds, contrast=c("treatment",'n9', 'n5')))      #0   0 
-```
-
-```{r totalDEGs}
-levels(colData$treatment)
-
-totalDEGs(c("treatment", 'bldg','lay'))
-totalDEGs(c("treatment", 'bldg','inc.d3'))
-totalDEGs(c("treatment", 'bldg','inc.d9'))
-totalDEGs(c("treatment", 'bldg','inc.d17'))
-totalDEGs(c("treatment", 'bldg','hatch'))
-totalDEGs(c("treatment", 'bldg','n5'))
-totalDEGs(c("treatment", 'bldg','n9'))
-
-totalDEGs(c("treatment", 'lay','bldg'))
-totalDEGs(c("treatment", 'lay','inc.d3'))
-totalDEGs(c("treatment", 'lay','inc.d9'))
-totalDEGs(c("treatment", 'lay','inc.d17'))
-totalDEGs(c("treatment", 'lay','hatch'))
-totalDEGs(c("treatment", 'lay','n5'))
-totalDEGs(c("treatment", 'lay','n9'))
-
-totalDEGs(c("treatment", 'inc.d3','bldg'))
-totalDEGs(c("treatment", 'inc.d3','lay'))
-totalDEGs(c("treatment", 'inc.d3','inc.d9'))
-totalDEGs(c("treatment", 'inc.d3','inc.d17'))
-totalDEGs(c("treatment", 'inc.d3','hatch'))
-totalDEGs(c("treatment", 'inc.d3','n5'))
-totalDEGs(c("treatment", 'inc.d3','n9'))
-
-totalDEGs(c("treatment", 'inc.d9','bldg'))
-totalDEGs(c("treatment", 'inc.d9','lay'))
-totalDEGs(c("treatment", 'inc.d9','inc.d3'))
-totalDEGs(c("treatment", 'inc.d9','inc.d17'))
-totalDEGs(c("treatment", 'inc.d9','hatch'))
-totalDEGs(c("treatment", 'inc.d9','n5'))
-totalDEGs(c("treatment", 'inc.d9','n9'))
-
-totalDEGs(c("treatment", 'inc.d17','bldg'))
-totalDEGs(c("treatment", 'inc.d17','lay'))
-totalDEGs(c("treatment", 'inc.d17','inc.d3'))
-totalDEGs(c("treatment", 'inc.d17','inc.d9'))
-totalDEGs(c("treatment", 'inc.d17','hatch'))
-totalDEGs(c("treatment", 'inc.d17','n5'))
-totalDEGs(c("treatment", 'inc.d17','n9'))
-
-totalDEGs(c("treatment", 'hatch','bldg'))
-totalDEGs(c("treatment", 'hatch','lay'))
-totalDEGs(c("treatment", 'hatch','inc.d3'))
-totalDEGs(c("treatment", 'hatch','inc.d9'))
-totalDEGs(c("treatment", 'hatch','inc.d17'))
-totalDEGs(c("treatment", 'hatch','n5'))
-totalDEGs(c("treatment", 'hatch','n9'))
-
-totalDEGs(c("treatment", 'n5','bldg'))
-totalDEGs(c("treatment", 'n5','lay'))
-totalDEGs(c("treatment", 'n5','inc.d3'))
-totalDEGs(c("treatment", 'n5','inc.d9'))
-totalDEGs(c("treatment", 'n5','inc.d17'))
-totalDEGs(c("treatment", 'n5','hatch'))
-totalDEGs(c("treatment", 'n5','n9'))
-
-totalDEGs(c("treatment", 'n9','bldg'))
-totalDEGs(c("treatment", 'n9','lay'))
-totalDEGs(c("treatment", 'n9','inc.d3'))
-totalDEGs(c("treatment", 'n9','inc.d9'))
-totalDEGs(c("treatment", 'n9','inc.d17'))
-totalDEGs(c("treatment", 'n9','hatch'))
-totalDEGs(c("treatment", 'n9','n5'))
+for (i in a){
+  for (j in b){
+    print(paste(i,j))
+    if (i != j) {
+    numDEGs(i,j)
+  }
+  }
+}
 ```

### 2019-04-05 analysis/02_DESeq2_hyp.Rmd
now with for loop and wide DEGs

@@
 
-```{r}
+```{r setup}
 library(tidyverse)
@@
 library(pheatmap)
+library(kableExtra)
 
@@
 a <- levels(colData$treatment)
-a
+b <- levels(colData$treatment)
 
-# create list of groups, with control last
-b <- a[2:9]
-b[9] <- a[1]
-b
+# slim for testing
+a <- c("n9", "bldg" , "lay" )
+b <- c("n9", "bldg" , "lay" )
 
+# comapre all contrasts, save to datafrmes
+dat=data.frame()
 for (i in a){
   for (j in b){
-    print(paste(i,j))
     if (i != j) {
-    numDEGs(i,j)
+      k <- paste(i,j, sep = "") #assigns usique rownames
+      dat[k,1]<-i               
+      dat[k,2]<-j
+      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
     }
@@
 }
+
+head(dat)
+```
+
+```{r restable}
+# widen data to create table of degs
+rownames(dat) <- NULL #remove row names
+data_wide <- spread(dat, V2, V3)
+data_wide
+kable(data_wide) # pretty printing
 ```

### 2019-04-06 analysis/02_DESeq2_hyp.Rmd
now with male anlaysis

@@
 library(kableExtra)
+library(viridis)
 
@@
 
-This anlaysis will _exclude_ the control timepoint but _combine_ incubation and nestling timepoints.
+# Females only
 
@@
 
-```{r}
-# making new groups
-colData$group <- NULL
-colData$tempgroup <- ifelse(colData$treatment == "bldg", "bldg",
-                  ifelse(colData$treatment == "control", "control",
-                   ifelse(colData$treatment == "hatch", "hatch",
-                    ifelse(grepl("inc", colData$treatment), "inc",
-                     ifelse(colData$treatment == "lay", "lay", "nestl")))))
-colData$tempgroup <- as.factor(colData$tempgroup)
-
-colData$group <- paste(colData$sex, colData$tempgroup, sep = "")
-colData$group <- as.factor(colData$group)
-str(colData$group)
-
+```{r factors}
 colData$treatment <- factor(colData$treatment, levels = 
@@
   dplyr::filter(grepl('hypothalamus', tissue)) %>%
-  #dplyr::filter(treatment != "control") %>%
   dplyr::filter(sex == "female") %>%
@@
 # slim for testing
-a <- c("n9", "bldg" , "lay" )
-b <- c("n9", "bldg" , "lay" )
+#a <- c("n9", "bldg" , "lay" )
+#b <- c("n9", "bldg" , "lay" )
 
@@
 data_wide
-kable(data_wide) # pretty printing
+kable(data_wide)
+
+ggplot(dat, aes(V1, V2)) +
+  geom_tile(aes(fill = V3)) +
+  scale_fill_viridis(na.value="#FFFFFF00")
 ```
@@
 ```{r PCA, message=FALSE, warning=FALSE}
-levels(colData$treatment)
-colData$treatment <- factor(colData$treatment, levels = 
-                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
-                                "inc.d17", "hatch", "n5", "n9"))
-
 # create the dataframe using my function pcadataframe
@@
 
-ggplot(pcadata, aes(PC2, PC3,color = treatment)) + 
-  geom_point(size = 2, alpha = 1) +
-  stat_ellipse(type = "t") +
-  xlab(paste0("PC2: ", percentVar[2],"% variance")) +
-  ylab(paste0("PC3: ", percentVar[3],"% variance")) +
-  theme_cowplot(font_size = 8, line_size = 0.25)
+summary(aov(PC1 ~ treatment, data=pcadata)) 
+TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment")
+
+summary(aov(PC2 ~ treatment, data=pcadata)) 
+TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment")
 
@@
   theme_cowplot(font_size = 8, line_size = 0.25)
-```
-
-PCA statistics
-
-```{r}
-summary(aov(PC1 ~ treatment, data=pcadata)) 
-TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment")
-
-summary(aov(PC2 ~ treatment, data=pcadata)) 
-TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment")
 
@@
 ```
+
+Heatmaps
+
 ```{r heatmap}
@@
 ```
+
+# Males only
+
+```{r readMale}
+# import "colData" which contains sample information and "countData" which contains read counts
+colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
+countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
+geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
+```
+
+```{r factorsMale}
+colData$treatment <- factor(colData$treatment, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+```
+
+```{r filterMale}
+colData <- colData %>%
+  dplyr::filter(grepl('hypothalamus', tissue)) %>%
+  dplyr::filter(sex == "male") %>%
+  droplevels()
+row.names(colData) <- colData$V1
+
+# print sample sizes
+colData %>% select(group, tissue)  %>%  summary()
+
+savecols <- as.character(colData$V1) 
+savecols <- as.vector(savecols) 
+countData <- countData %>% dplyr::select(one_of(savecols))
+
+# check that row and col lenghts are equal
+ncol(countData) == nrow(colData) 
+```
+
+```{r deseq2male}
+dds <- DESeqDataSetFromMatrix(countData = countData,
+                              colData = colData,
+                              design = ~ treatment )
+dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
+dds <- DESeq(dds) # Differential expression analysis
+vsd <- vst(dds, blind=FALSE) # variance stabilized 
+```
+
+```{r resloopMale}
+#create list of groups
+a <- levels(colData$treatment)
+b <- levels(colData$treatment)
+
+# slim for testing
+#a <- c("n9", "bldg" , "lay" )
+#b <- c("n9", "bldg" , "lay" )
+
+# comapre all contrasts, save to datafrmes
+dat=data.frame()
+for (i in a){
+  for (j in b){
+    if (i != j) {
+      k <- paste(i,j, sep = "") #assigns usique rownames
+      dat[k,1]<-i               
+      dat[k,2]<-j
+      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
+    }
+  }
+}
+
+head(dat)
+```
+
+```{r restableMale}
+# widen data to create table of degs
+rownames(dat) <- NULL #remove row names
+data_wide <- spread(dat, V2, V3)
+data_wide
+kable(data_wide)
+
+ggplot(dat, aes(V1, V2)) +
+  geom_tile(aes(fill = V3)) +
+  scale_fill_viridis(na.value="#FFFFFF00")
+```
+
+```{r PCAmale, message=FALSE, warning=FALSE}
+# create the dataframe using my function pcadataframe
+pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
+percentVar <- round(100 * attr(pcadata, "percentVar"))
+percentVar
+
+ggplot(pcadata, aes(PC1, PC2,color = treatment)) + 
+  geom_point(size = 2, alpha = 1) +
+  stat_ellipse(type = "t") +
+  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
+  theme_cowplot(font_size = 8, line_size = 0.25)
+
+summary(aov(PC1 ~ treatment, data=pcadata)) 
+TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment")
+
+summary(aov(PC2 ~ treatment, data=pcadata)) 
+TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment")
+
+ggplot(pcadata, aes(PC3, PC4,color = treatment)) + 
+  geom_point(size = 2, alpha = 1) +
+  stat_ellipse(type = "t") +
+  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
+  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
+  theme_cowplot(font_size = 8, line_size = 0.25)
+
+summary(aov(PC3 ~ treatment, data=pcadata))
+
+summary(aov(PC4 ~ treatment, data=pcadata))
+
+summary(aov(PC5 ~ treatment, data=pcadata))
+
+summary(aov(PC6 ~ treatment, data=pcadata)) 
+```
+
+Heatmaps
+
+```{r heatmapMale}
+# see http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#heatmap-of-the-count-matrix
+sampleDists <- dist(t(assay(vsd)))
+
+sampleDistMatrix <- as.matrix(sampleDists)
+rownames(sampleDistMatrix) <- NULL
+colnames(sampleDistMatrix) <- colData$treatment
+colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
+pheatmap(sampleDistMatrix,
+         clustering_distance_rows=sampleDists,
+         clustering_distance_cols=sampleDists,
+         col=colors,
+         fontsize = 6)
+```

### 2019-04-06 analysis/02_DESeq2_hyp.Rmd
now with male anlaysis

@@
 
-```{r setup}
+```{r setup, message=F, comment=FALSE, warning=FALSE}
 library(tidyverse)

### 2019-04-06 analysis/02_DESeq2_hyp.Rmd
remove results from setup

--- 2019-04-06 analysis/02_DESeq2_hyp.Rmd
@@
 
-```{r setup}
+```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
 library(tidyverse)

### 2019-04-06 analysis/04_DESeq2_gon.Rmd
deseq gonads

@@
 library(pheatmap)
+library(kableExtra)
+library(viridis)
 
@@
 
-```{r colDataWrangle}
-# making new groups
-colData$group <- NULL
-colData$tempgroup <- ifelse(colData$treatment == "bldg", "bldg",
-                  ifelse(colData$treatment == "control", "control",
-                   ifelse(colData$treatment == "hatch", "hatch",
-                    ifelse(grepl("inc", colData$treatment), "inc",
-                     ifelse(colData$treatment == "lay", "lay", "nestl")))))
-colData$tempgroup <- as.factor(colData$tempgroup)
-
-colData$group <- paste(colData$sex, colData$tempgroup, sep = "")
-colData$group <- as.factor(colData$group)
-str(colData$group)
-
+```{r factors}
 colData$treatment <- factor(colData$treatment, levels = 
@@
 ```{r filter}
-#select samples of interest from colData
 colData <- colData %>%
   dplyr::filter(grepl('gonad', tissue)) %>%
-  #dplyr::filter(treatment != "control") %>%
   dplyr::filter(sex == "female") %>%
@@
 
-#select samples of interest from countData
 savecols <- as.character(colData$V1) 
@@
 
-```{r res}
-levels(colData$treatment)
-summary(results(dds, contrast=c("treatment",'control', 'bldg'))) 
+```{r resloop}
+#create list of groups
+a <- levels(colData$treatment)
+b <- levels(colData$treatment)
+
+# slim for testing
+#a <- c("n9", "bldg" , "lay" )
+#b <- c("n9", "bldg" , "lay" )
+
+# comapre all contrasts, save to datafrmes
+dat=data.frame()
+for (i in a){
+  for (j in b){
+    if (i != j) {
+      k <- paste(i,j, sep = "") #assigns usique rownames
+      dat[k,1]<-i               
+      dat[k,2]<-j
+      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
+    }
+  }
+}
+
+head(dat)
 ```
 
-```{res2}
-summary(results(dds, contrast=c("treatment",'control', 'lay')))     
-summary(results(dds, contrast=c("treatment",'control', 'inc.d3')))  
-summary(results(dds, contrast=c("treatment",'control', 'inc.d9')))   
-summary(results(dds, contrast=c("treatment",'control', 'inc.d17')))  
-summary(results(dds, contrast=c("treatment",'control', 'hatch')))    
-summary(results(dds, contrast=c("treatment",'control', 'n5')))      
-summary(results(dds, contrast=c("treatment",'control', 'n9')))
+```{r restable}
+# widen data to create table of degs
+rownames(dat) <- NULL #remove row names
+data_wide <- spread(dat, V2, V3)
+data_wide
+kable(data_wide)
 
-summary(results(dds, contrast=c("treatment",'bldg', 'lay')))      
-summary(results(dds, contrast=c("treatment",'bldg', 'inc.d3')))   
-summary(results(dds, contrast=c("treatment",'bldg', 'inc.d9')))   
-summary(results(dds, contrast=c("treatment",'bldg', 'inc.d17')))  
-summary(results(dds, contrast=c("treatment",'bldg', 'hatch')))   
-summary(results(dds, contrast=c("treatment",'bldg', 'n5')))      
-summary(results(dds, contrast=c("treatment",'bldg', 'n9')))
+dat$V1 <- factor(dat$V1, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+dat$V2 <- factor(dat$V2, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
 
-#summary(results(dds, contrast=c("treatment",'lay', 'bldg')))     
-summary(results(dds, contrast=c("treatment",'lay', 'inc.d3')))    
-summary(results(dds, contrast=c("treatment",'lay', 'inc.d9')))   
-summary(results(dds, contrast=c("treatment",'lay', 'inc.d17')))  
-summary(results(dds, contrast=c("treatment",'lay', 'hatch')))    
-summary(results(dds, contrast=c("treatment",'lay', 'n5')))       
-summary(results(dds, contrast=c("treatment",'lay', 'n9')))
-
-#summary(results(dds, contrast=c("treatment",'inc.d3', 'bldg')))     
-#summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))       
-summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d9')))   
-summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d17')))   
-summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))      
-summary(results(dds, contrast=c("treatment",'inc.d3', 'n5')))        
-summary(results(dds, contrast=c("treatment",'inc.d3', 'n9')))
-
-summary(results(dds, contrast=c("treatment",'inc.d9', 'inc.d17')))   
-summary(results(dds, contrast=c("treatment",'inc.d9', 'lay')))      
-summary(results(dds, contrast=c("treatment",'inc.d9', 'n5')))       
-summary(results(dds, contrast=c("treatment",'inc.d9', 'n9')))
-
-summary(results(dds, contrast=c("treatment",'inc.d17', 'lay')))  
-summary(results(dds, contrast=c("treatment",'inc.d17', 'n5')))   
-summary(results(dds, contrast=c("treatment",'inc.d17', 'n9')))
-
-#summary(results(dds, contrast=c("treatment",'hatch', 'bldg')))      
-#summary(results(dds, contrast=c("treatment",'hatch', 'lay')))       
-#summary(results(dds, contrast=c("treatment",'hatch', 'inc.d3')))    
-#summary(results(dds, contrast=c("treatment",'hatch', 'inc.d9')))    
-summary(results(dds, contrast=c("treatment",'hatch', 'inc.d17')))  
-summary(results(dds, contrast=c("treatment",'hatch', 'n5')))       
-summary(results(dds, contrast=c("treatment",'hatch', 'n9')))
-
-#summary(results(dds, contrast=c("treatment",'n5', 'bldg')))      
-#summary(results(dds, contrast=c("treatment",'n5', 'lay')))       
-#summary(results(dds, contrast=c("treatment",'n5', 'inc.d3')))  
-#summary(results(dds, contrast=c("treatment",'n5', 'inc.d9')))   
-#summary(results(dds, contrast=c("treatment",'n5', 'inc.d17')))  
-summary(results(dds, contrast=c("treatment",'n5', 'hatch')))   
-summary(results(dds, contrast=c("treatment",'n5', 'n9')))
-
-#summary(results(dds, contrast=c("treatment",'n9', 'bldg')))     
-#summary(results(dds, contrast=c("treatment",'n9', 'lay')))      
-#summary(results(dds, contrast=c("treatment",'n9', 'inc.d3')))   
-#summary(results(dds, contrast=c("treatment",'n9', 'inc.d9')))   
-#summary(results(dds, contrast=c("treatment",'n9', 'inc.d17')))  
-#summary(results(dds, contrast=c("treatment",'n9', 'hatch')))    
-summary(results(dds, contrast=c("treatment",'n9', 'n5')))     
-```
-
-```{r totalDEGs}
-levels(colData$treatment)
-
-totalDEGs(c("treatment", 'bldg','lay'))
-totalDEGs(c("treatment", 'bldg','inc.d3'))
-totalDEGs(c("treatment", 'bldg','inc.d9'))
-totalDEGs(c("treatment", 'bldg','inc.d17'))
-totalDEGs(c("treatment", 'bldg','hatch'))
-totalDEGs(c("treatment", 'bldg','n5'))
-totalDEGs(c("treatment", 'bldg','n9'))
-
-#totalDEGs(c("treatment", 'lay','bldg'))
-totalDEGs(c("treatment", 'lay','inc.d3'))
-totalDEGs(c("treatment", 'lay','inc.d9'))
-totalDEGs(c("treatment", 'lay','inc.d17'))
-totalDEGs(c("treatment", 'lay','hatch'))
-totalDEGs(c("treatment", 'lay','n5'))
-totalDEGs(c("treatment", 'lay','n9'))
-
-#totalDEGs(c("treatment", 'inc.d3','bldg'))
-#totalDEGs(c("treatment", 'inc.d3','lay'))
-totalDEGs(c("treatment", 'inc.d3','inc.d9'))
-totalDEGs(c("treatment", 'inc.d3','inc.d17'))
-totalDEGs(c("treatment", 'inc.d3','hatch'))
-totalDEGs(c("treatment", 'inc.d3','n5'))
-totalDEGs(c("treatment", 'inc.d3','n9'))
-
-#totalDEGs(c("treatment", 'inc.d9','bldg'))
-#totalDEGs(c("treatment", 'inc.d9','lay'))
-#totalDEGs(c("treatment", 'inc.d9','inc.d3'))
-totalDEGs(c("treatment", 'inc.d9','inc.d17'))
-totalDEGs(c("treatment", 'inc.d9','hatch'))
-totalDEGs(c("treatment", 'inc.d9','n5'))
-totalDEGs(c("treatment", 'inc.d9','n9'))
-
-#totalDEGs(c("treatment", 'inc.d17','bldg'))
-#totalDEGs(c("treatment", 'inc.d17','lay'))
-#totalDEGs(c("treatment", 'inc.d17','inc.d3'))
-#totalDEGs(c("treatment", 'inc.d17','inc.d9'))
-totalDEGs(c("treatment", 'inc.d17','hatch'))
-totalDEGs(c("treatment", 'inc.d17','n5'))
-totalDEGs(c("treatment", 'inc.d17','n9'))
-
-#totalDEGs(c("treatment", 'hatch','bldg'))
-#totalDEGs(c("treatment", 'hatch','lay'))
-#totalDEGs(c("treatment", 'hatch','inc.d3'))
-#totalDEGs(c("treatment", 'hatch','inc.d9'))
-totalDEGs(c("treatment", 'hatch','inc.d17'))
-totalDEGs(c("treatment", 'hatch','n5'))
-totalDEGs(c("treatment", 'hatch','n9'))
-
-#totalDEGs(c("treatment", 'n5','bldg'))
-#totalDEGs(c("treatment", 'n5','lay'))
-#totalDEGs(c("treatment", 'n5','inc.d3'))
-#totalDEGs(c("treatment", 'n5','inc.d9'))
-#totalDEGs(c("treatment", 'n5','inc.d17'))
-totalDEGs(c("treatment", 'n5','hatch'))
-totalDEGs(c("treatment", 'n5','n9'))
-
-#totalDEGs(c("treatment", 'n9','bldg'))
-#totalDEGs(c("treatment", 'n9','lay'))
-#totalDEGs(c("treatment", 'n9','inc.d3'))
-#totalDEGs(c("treatment", 'n9','inc.d9'))
-#totalDEGs(c("treatment", 'n9','inc.d17'))
-#totalDEGs(c("treatment", 'n9','hatch'))
-totalDEGs(c("treatment", 'n9','n5'))
+ggplot(dat, aes(V1, V2)) +
+  geom_tile(aes(fill = V3)) +
+  scale_fill_viridis(na.value="#FFFFFF00") + 
+  xlab("Treatment") + ylab("Treatment") +
+  labs(fill = "# of DEGs")
 ```
@@
 ```{r PCA, message=FALSE, warning=FALSE}
-levels(colData$treatment)
-
 # create the dataframe using my function pcadataframe
@@
   theme_cowplot(font_size = 8, line_size = 0.25)
-
-plotPCA(vsd, intgroup=c("treatment"))
-```
-
-PCA statistics
-
-```{r pcastats}
-summary(aov(PC1 ~ treatment, data=pcadata)) 
-TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment")
-
-summary(aov(PC2 ~ treatment, data=pcadata)) 
-TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment")
 
@@
 ```
+
+Heatmaps
 
@@
 ```
+
+# Males only
+
+```{r readMale}
+# import "colData" which contains sample information and "countData" which contains read counts
+colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
+countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
+geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
+```
+
+```{r factorsMale}
+colData$treatment <- factor(colData$treatment, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+```
+
+```{r filterMale}
+colData <- colData %>%
+  dplyr::filter(grepl('gonad', tissue)) %>%
+  dplyr::filter(sex == "male") %>%
+  droplevels()
+row.names(colData) <- colData$V1
+
+# print sample sizes
+colData %>% select(group, tissue)  %>%  summary()
+
+savecols <- as.character(colData$V1) 
+savecols <- as.vector(savecols) 
+countData <- countData %>% dplyr::select(one_of(savecols))
+
+# check that row and col lenghts are equal
+ncol(countData) == nrow(colData) 
+```
+
+```{r deseq2male}
+dds <- DESeqDataSetFromMatrix(countData = countData,
+                              colData = colData,
+                              design = ~ treatment )
+dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
+dds <- DESeq(dds) # Differential expression analysis
+vsd <- vst(dds, blind=FALSE) # variance stabilized 
+```
+
+```{r resloopMale}
+#create list of groups
+a <- levels(colData$treatment)
+b <- levels(colData$treatment)
+
+# slim for testing
+#a <- c("n9", "bldg" , "lay" )
+#b <- c("n9", "bldg" , "lay" )
+
+# comapre all contrasts, save to datafrmes
+dat=data.frame()
+for (i in a){
+  for (j in b){
+    if (i != j) {
+      k <- paste(i,j, sep = "") #assigns usique rownames
+      dat[k,1]<-i               
+      dat[k,2]<-j
+      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
+    }
+  }
+}
+
+head(dat)
+```
+
+```{r restableMale}
+# widen data to create table of degs
+rownames(dat) <- NULL #remove row names
+data_wide <- spread(dat, V2, V3)
+data_wide
+kable(data_wide)
+
+dat$V1 <- factor(dat$V1, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+dat$V2 <- factor(dat$V2, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+
+ggplot(dat, aes(V1, V2)) +
+  geom_tile(aes(fill = V3)) +
+  scale_fill_viridis(na.value="#FFFFFF00") + 
+  xlab("Treatment") + ylab("Treatment") +
+  labs(fill = "# of DEGs")
+```

### 2019-04-06 analysis/02_DESeq2_hyp.Rmd
better gridplot

@@
 
+dat$V1 <- factor(dat$V1, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+dat$V2 <- factor(dat$V2, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+
 ggplot(dat, aes(V1, V2)) +
   geom_tile(aes(fill = V3)) +
-  scale_fill_viridis(na.value="#FFFFFF00")
+  scale_fill_viridis(na.value="#FFFFFF00") + 
+  xlab("Treatment") + ylab("Treatment") +
+  labs(fill = "# of DEGs")
 ```
@@
 
+dat$V1 <- factor(dat$V1, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+dat$V2 <- factor(dat$V2, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+
 ggplot(dat, aes(V1, V2)) +
   geom_tile(aes(fill = V3)) +
-  scale_fill_viridis(na.value="#FFFFFF00") 
+  scale_fill_viridis(na.value="#FFFFFF00") + 
+  xlab("Treatment") + ylab("Treatment") +
+  labs(fill = "# of DEGs")
 ```

### 2019-04-07 analysis/03_DESeq2_pit.Rmd
pitutitary DEGS

@@
 
-```{r}
+```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
 library(tidyverse)
@@
 library(cowplot)
+library(RColorBrewer)
+library(pheatmap)
+library(kableExtra)
+library(viridis)
 
@@
 
-```{r}
-# making new groups
-colData$group <- NULL
-colData$tempgroup <- ifelse(colData$treatment == "bldg", "bldg",
-                  ifelse(colData$treatment == "control", "control",
-                   ifelse(colData$treatment == "hatch", "hatch",
-                    ifelse(grepl("inc", colData$treatment), "inc",
-                     ifelse(colData$treatment == "lay", "lay", "nestl")))))
-colData$tempgroup <- as.factor(colData$tempgroup)
-
-colData$group <- paste(colData$sex, colData$tempgroup, sep = "")
-colData$group <- as.factor(colData$group)
-str(colData$group)
+```{r factors} 
+colData$treatment <- factor(colData$treatment, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
 ```
@@
   dplyr::filter(grepl('pituitary', tissue)) %>%
-  #dplyr::filter(treatment != "control") %>%
   dplyr::filter(sex == "female") %>%
@@
 
-```{r res}
-levels(colData$treatment)
-summary(results(dds, contrast=c("treatment",'control', 'bldg'))) 
-summary(results(dds, contrast=c("treatment",'control', 'lay')))     
-summary(results(dds, contrast=c("treatment",'control', 'inc.d3')))  
-summary(results(dds, contrast=c("treatment",'control', 'inc.d9')))   
-summary(results(dds, contrast=c("treatment",'control', 'inc.d17')))  
-summary(results(dds, contrast=c("treatment",'control', 'hatch')))    
-summary(results(dds, contrast=c("treatment",'control', 'n5')))      
-summary(results(dds, contrast=c("treatment",'control', 'n9')))
+```{r resloop} 
+#create list of groups
+a <- levels(colData$treatment)
+b <- levels(colData$treatment)
 
-summary(results(dds, contrast=c("treatment",'bldg', 'lay')))      
-summary(results(dds, contrast=c("treatment",'bldg', 'inc.d3')))   
-summary(results(dds, contrast=c("treatment",'bldg', 'inc.d9')))   
-summary(results(dds, contrast=c("treatment",'bldg', 'inc.d17')))  
-summary(results(dds, contrast=c("treatment",'bldg', 'hatch')))   
-summary(results(dds, contrast=c("treatment",'bldg', 'n5')))      
-summary(results(dds, contrast=c("treatment",'bldg', 'n9')))
+# slim for testing
+#a <- c("n9", "bldg" , "lay" )
+#b <- c("n9", "bldg" , "lay" )
 
-summary(results(dds, contrast=c("treatment",'lay', 'bldg')))     
-summary(results(dds, contrast=c("treatment",'lay', 'inc.d3')))    
-summary(results(dds, contrast=c("treatment",'lay', 'inc.d9')))   
-summary(results(dds, contrast=c("treatment",'lay', 'inc.d17')))  
-summary(results(dds, contrast=c("treatment",'lay', 'hatch')))    
-summary(results(dds, contrast=c("treatment",'lay', 'n5')))       
-summary(results(dds, contrast=c("treatment",'lay', 'n9')))
+# comapre all contrasts, save to datafrmes
+dat=data.frame()
+for (i in a){
+  for (j in b){
+    if (i != j) {
+      k <- paste(i,j, sep = "") #assigns usique rownames
+      dat[k,1]<-i               
+      dat[k,2]<-j
+      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
+    }
+  }
+}
 
-summary(results(dds, contrast=c("treatment",'inc.d3', 'bldg')))     
-summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))       
-summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d9')))   
-summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d17')))   
-summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))      
-summary(results(dds, contrast=c("treatment",'inc.d3', 'n5')))        
-summary(results(dds, contrast=c("treatment",'inc.d3', 'n9')))
-
-summary(results(dds, contrast=c("treatment",'inc.d9', 'inc.d17')))   
-summary(results(dds, contrast=c("treatment",'inc.d9', 'lay')))      
-summary(results(dds, contrast=c("treatment",'inc.d9', 'n5')))       
-summary(results(dds, contrast=c("treatment",'inc.d9', 'n9')))
-
-summary(results(dds, contrast=c("treatment",'inc.d17', 'lay')))  
-summary(results(dds, contrast=c("treatment",'inc.d17', 'n5')))   
-summary(results(dds, contrast=c("treatment",'inc.d17', 'n9')))
-
-summary(results(dds, contrast=c("treatment",'hatch', 'bldg')))      
-summary(results(dds, contrast=c("treatment",'hatch', 'lay')))       
-summary(results(dds, contrast=c("treatment",'hatch', 'inc.d3')))    
-summary(results(dds, contrast=c("treatment",'hatch', 'inc.d9')))    
-summary(results(dds, contrast=c("treatment",'hatch', 'inc.d17')))  
-summary(results(dds, contrast=c("treatment",'hatch', 'n5')))       
-summary(results(dds, contrast=c("treatment",'hatch', 'n9')))
-
-summary(results(dds, contrast=c("treatment",'n5', 'bldg')))      
-summary(results(dds, contrast=c("treatment",'n5', 'lay')))       
-summary(results(dds, contrast=c("treatment",'n5', 'inc.d3')))  
-summary(results(dds, contrast=c("treatment",'n5', 'inc.d9')))   
-summary(results(dds, contrast=c("treatment",'n5', 'inc.d17')))  
-summary(results(dds, contrast=c("treatment",'n5', 'hatch')))   
-summary(results(dds, contrast=c("treatment",'n5', 'n9')))
-
-summary(results(dds, contrast=c("treatment",'n9', 'bldg')))     
-summary(results(dds, contrast=c("treatment",'n9', 'lay')))      
-summary(results(dds, contrast=c("treatment",'n9', 'inc.d3')))   
-summary(results(dds, contrast=c("treatment",'n9', 'inc.d9')))   
-summary(results(dds, contrast=c("treatment",'n9', 'inc.d17')))  
-summary(results(dds, contrast=c("treatment",'n9', 'hatch')))    
-summary(results(dds, contrast=c("treatment",'n9', 'n5')))     
+head(dat)
 ```
 
-```{r totalDEGs}
-levels(colData$treatment)
+```{r restable} 
+# widen data to create table of degs
+rownames(dat) <- NULL #remove row names
+data_wide <- spread(dat, V2, V3)
+data_wide
+kable(data_wide)
 
-totalDEGs(c("treatment", 'bldg','lay'))
-totalDEGs(c("treatment", 'bldg','inc.d3'))
-totalDEGs(c("treatment", 'bldg','inc.d9'))
-totalDEGs(c("treatment", 'bldg','inc.d17'))
-totalDEGs(c("treatment", 'bldg','hatch'))
-totalDEGs(c("treatment", 'bldg','n5'))
-totalDEGs(c("treatment", 'bldg','n9'))
+dat$V1 <- factor(dat$V1, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+dat$V2 <- factor(dat$V2, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
 
-totalDEGs(c("treatment", 'lay','bldg'))
-totalDEGs(c("treatment", 'lay','inc.d3'))
-totalDEGs(c("treatment", 'lay','inc.d9'))
-totalDEGs(c("treatment", 'lay','inc.d17'))
-totalDEGs(c("treatment", 'lay','hatch'))
-totalDEGs(c("treatment", 'lay','n5'))
-totalDEGs(c("treatment", 'lay','n9'))
-
-totalDEGs(c("treatment", 'inc.d3','bldg'))
-totalDEGs(c("treatment", 'inc.d3','lay'))
-totalDEGs(c("treatment", 'inc.d3','inc.d9'))
-totalDEGs(c("treatment", 'inc.d3','inc.d17'))
-totalDEGs(c("treatment", 'inc.d3','hatch'))
-totalDEGs(c("treatment", 'inc.d3','n5'))
-totalDEGs(c("treatment", 'inc.d3','n9'))
-
-totalDEGs(c("treatment", 'inc.d9','bldg'))
-totalDEGs(c("treatment", 'inc.d9','lay'))
-totalDEGs(c("treatment", 'inc.d9','inc.d3'))
-totalDEGs(c("treatment", 'inc.d9','inc.d17'))
-totalDEGs(c("treatment", 'inc.d9','hatch'))
-totalDEGs(c("treatment", 'inc.d9','n5'))
-totalDEGs(c("treatment", 'inc.d9','n9'))
-
-totalDEGs(c("treatment", 'inc.d17','bldg'))
-totalDEGs(c("treatment", 'inc.d17','lay'))
-totalDEGs(c("treatment", 'inc.d17','inc.d3'))
-totalDEGs(c("treatment", 'inc.d17','inc.d9'))
-totalDEGs(c("treatment", 'inc.d17','hatch'))
-totalDEGs(c("treatment", 'inc.d17','n5'))
-totalDEGs(c("treatment", 'inc.d17','n9'))
-
-totalDEGs(c("treatment", 'hatch','bldg'))
-totalDEGs(c("treatment", 'hatch','lay'))
-totalDEGs(c("treatment", 'hatch','inc.d3'))
-totalDEGs(c("treatment", 'hatch','inc.d9'))
-totalDEGs(c("treatment", 'hatch','inc.d17'))
-totalDEGs(c("treatment", 'hatch','n5'))
-totalDEGs(c("treatment", 'hatch','n9'))
-
-totalDEGs(c("treatment", 'n5','bldg'))
-totalDEGs(c("treatment", 'n5','lay'))
-totalDEGs(c("treatment", 'n5','inc.d3'))
-totalDEGs(c("treatment", 'n5','inc.d9'))
-totalDEGs(c("treatment", 'n5','inc.d17'))
-totalDEGs(c("treatment", 'n5','hatch'))
-totalDEGs(c("treatment", 'n5','n9'))
-
-totalDEGs(c("treatment", 'n9','bldg'))
-totalDEGs(c("treatment", 'n9','lay'))
-totalDEGs(c("treatment", 'n9','inc.d3'))
-totalDEGs(c("treatment", 'n9','inc.d9'))
-totalDEGs(c("treatment", 'n9','inc.d17'))
-totalDEGs(c("treatment", 'n9','hatch'))
-totalDEGs(c("treatment", 'n9','n5'))
+ggplot(dat, aes(V1, V2)) +
+  geom_tile(aes(fill = V3)) +
+  scale_fill_viridis(na.value="#FFFFFF00", 
+                     limits = c(0,6000)) + 
+  xlab("Timepoint") + ylab("Timepoint") +
+  labs(fill = "# of DEGs")
 ```
@@
 ```{r PCA, message=FALSE, warning=FALSE} 
-levels(colData$treatment)
-colData$treatment <- factor(colData$treatment, levels = 
-                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
-                                "inc.d17", "hatch", "n5", "n9"))
-
 # create the dataframe using my function pcadataframe
@@
   theme_cowplot(font_size = 8, line_size = 0.25)
-```
-
-PCA statistics
-
-```{r pcastats}
-summary(aov(PC1 ~ treatment, data=pcadata)) 
-TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment")
-
-summary(aov(PC2 ~ treatment, data=pcadata)) 
-TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment")
 
@@
 ```
+
+Heatmaps
+
+```{r heatmap} 
+# see http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#heatmap-of-the-count-matrix
+sampleDists <- dist(t(assay(vsd)))
+
+sampleDistMatrix <- as.matrix(sampleDists)
+rownames(sampleDistMatrix) <- NULL
+colnames(sampleDistMatrix) <- colData$treatment
+colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
+pheatmap(sampleDistMatrix,
+         clustering_distance_rows=sampleDists,
+         clustering_distance_cols=sampleDists,
+         col=colors,
+         fontsize = 6)
+```
+
+# Males only
+
+```{r readMale}
+# import "colData" which contains sample information and "countData" which contains read counts
+colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
+countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
+geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
+```
+
+```{r factorsMale}
+colData$treatment <- factor(colData$treatment, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+```
+
+```{r filterMale}
+colData <- colData %>%
+  dplyr::filter(grepl('pituitary', tissue)) %>%
+  dplyr::filter(sex == "male") %>%
+  droplevels()
+row.names(colData) <- colData$V1
+
+# print sample sizes
+colData %>% select(group, tissue)  %>%  summary()
+
+savecols <- as.character(colData$V1) 
+savecols <- as.vector(savecols) 
+countData <- countData %>% dplyr::select(one_of(savecols))
+
+# check that row and col lenghts are equal
+ncol(countData) == nrow(colData) 
+```
+
+```{r deseq2male} 
+dds <- DESeqDataSetFromMatrix(countData = countData,
+                              colData = colData,
+                              design = ~ treatment )
+dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
+dds <- DESeq(dds) # Differential expression analysis
+vsd <- vst(dds, blind=FALSE) # variance stabilized 
+```
+
+```{r resloopMale} 
+#create list of groups
+a <- levels(colData$treatment)
+b <- levels(colData$treatment)
+
+# slim for testing
+#a <- c("n9", "bldg" , "lay" )
+#b <- c("n9", "bldg" , "lay" )
+
+# comapre all contrasts, save to datafrmes
+dat=data.frame()
+for (i in a){
+  for (j in b){
+    if (i != j) {
+      k <- paste(i,j, sep = "") #assigns usique rownames
+      dat[k,1]<-i               
+      dat[k,2]<-j
+      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
+    }
+  }
+}
+
+head(dat)
+```
+
+```{r restableMale} 
+# widen data to create table of degs
+rownames(dat) <- NULL #remove row names
+data_wide <- spread(dat, V2, V3)
+data_wide
+kable(data_wide)
+
+dat$V1 <- factor(dat$V1, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+dat$V2 <- factor(dat$V2, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+
+ggplot(dat, aes(V1, V2)) +
+  geom_tile(aes(fill = V3)) +
+  scale_fill_viridis(na.value="#FFFFFF00",
+                     limits = c(0,6000)) + 
+  xlab("Timepoint") + ylab("Timepoint") +
+  labs(fill = "# of DEGs")
+```
+
+```{r PCAmale, message=FALSE, warning=FALSE}
+# create the dataframe using my function pcadataframe
+pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
+percentVar <- round(100 * attr(pcadata, "percentVar"))
+percentVar
+
+ggplot(pcadata, aes(PC1, PC2,color = treatment)) + 
+  geom_point(size = 2, alpha = 1) +
+  stat_ellipse(type = "t") +
+  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
+  theme_cowplot(font_size = 8, line_size = 0.25)
+
+summary(aov(PC1 ~ treatment, data=pcadata)) 
+TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment")
+
+summary(aov(PC2 ~ treatment, data=pcadata)) 
+TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment")
+
+ggplot(pcadata, aes(PC3, PC4,color = treatment)) + 
+  geom_point(size = 2, alpha = 1) +
+  stat_ellipse(type = "t") +
+  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
+  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
+  theme_cowplot(font_size = 8, line_size = 0.25)
+
+summary(aov(PC3 ~ treatment, data=pcadata))
+
+summary(aov(PC4 ~ treatment, data=pcadata))
+
+summary(aov(PC5 ~ treatment, data=pcadata))
+
+summary(aov(PC6 ~ treatment, data=pcadata)) 
+```
+
+Heatmaps
+
+```{r heatmapMale} 
+# see http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#heatmap-of-the-count-matrix
+sampleDists <- dist(t(assay(vsd)))
+
+sampleDistMatrix <- as.matrix(sampleDists)
+rownames(sampleDistMatrix) <- NULL
+colnames(sampleDistMatrix) <- colData$treatment
+colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
+pheatmap(sampleDistMatrix,
+         clustering_distance_rows=sampleDists,
+         clustering_distance_cols=sampleDists,
+         col=colors,
+         fontsize = 6)
+```

### 2019-04-07 analysis/03_DESeq2_pit.Rmd
more DEGS

@@
   scale_fill_viridis(na.value="#FFFFFF00", 
-                     limits = c(0,6000)) + 
+                     limits = c(0,7914)) + 
   xlab("Timepoint") + ylab("Timepoint") +
@@
   scale_fill_viridis(na.value="#FFFFFF00",
-                     limits = c(0,6000)) + 
+                     limits = c(0,7914)) + 
   xlab("Timepoint") + ylab("Timepoint") +

### 2019-04-07 analysis/02_DESeq2_hyp.Rmd
now with 8000 limit

@@
   geom_tile(aes(fill = V3)) +
-  scale_fill_viridis(na.value="#FFFFFF00") + 
+  scale_fill_viridis(na.value="#FFFFFF00", 
+                     limits = c(0,8028)) + 
   xlab("Treatment") + ylab("Treatment") +

### 2019-04-07 analysis/03_DESeq2_pit.Rmd
now with 8000 limit

@@
   scale_fill_viridis(na.value="#FFFFFF00", 
-                     limits = c(0,7914)) + 
+                     limits = c(0,8028)) + 
   xlab("Timepoint") + ylab("Timepoint") +
@@
   scale_fill_viridis(na.value="#FFFFFF00",
-                     limits = c(0,7914)) + 
+                     limits = c(0,8028)) + 
   xlab("Timepoint") + ylab("Timepoint") +

### 2019-04-07 analysis/02_DESeq2_hyp.Rmd
better legends and axis labels

--- 2019-04-06 analysis/02_DESeq2_hyp.Rmd
@@
   geom_tile(aes(fill = V3)) +
-  scale_fill_viridis(na.value="#FFFFFF00") + 
-  xlab("Treatment") + ylab("Treatment") +
+  scale_fill_viridis(na.value="#FFFFFF00", 
+                     limits = c(0,8028),
+                     breaks = c(0, 2000, 4000, 6000, 8000)) + 
+  xlab("Timepoint") + ylab("Timepoint") +
   labs(fill = "# of DEGs")
@@
   geom_tile(aes(fill = V3)) +
-  scale_fill_viridis(na.value="#FFFFFF00") + 
-  xlab("Treatment") + ylab("Treatment") +
+  scale_fill_viridis(na.value="#FFFFFF00", 
+                     limits = c(0,8028),
+                     breaks = c(0, 2000, 4000, 6000, 8000)) + 
+  xlab(" ") + ylab("Timepoint") +
   labs(fill = "# of DEGs")

### 2019-04-07 analysis/03_DESeq2_pit.Rmd
better legends and axis labels

@@
   scale_fill_viridis(na.value="#FFFFFF00", 
-                     limits = c(0,8028)) + 
+                     limits = c(0,8028),
+                     breaks = c(0, 2000, 4000, 6000, 8000)) + 
   xlab("Timepoint") + ylab("Timepoint") +
@@
   scale_fill_viridis(na.value="#FFFFFF00", 
-                     limits = c(0,8028)) + 
-  xlab("Timepoint") + ylab("Timepoint") +
+                     limits = c(0,8028),
+                     breaks = c(0, 2000, 4000, 6000, 8000)) + 
+  xlab(" ") + ylab("Timepoint") +
   labs(fill = "# of DEGs")  

### 2019-04-07 analysis/04_DESeq2_gon.Rmd
better legends and axis labels

--- 2019-04-06 analysis/04_DESeq2_gon.Rmd
@@
   geom_tile(aes(fill = V3)) +
-  scale_fill_viridis(na.value="#FFFFFF00") + 
-  xlab("Treatment") + ylab("Treatment") +
+  scale_fill_viridis(na.value="#FFFFFF00", 
+                     limits = c(0,8028),
+                     breaks = c(0, 2000, 4000, 6000, 8000)) + 
+  xlab(" ") + ylab("Timepoint") +
+  labs(fill = "# of DEGs")
+
+dat %>%
+  filter(V1 != "control")  %>%
+  ggplot( aes(V1, V2)) +
+    geom_tile(aes(fill = V3)) +
+    scale_fill_viridis(na.value="#FFFFFF00", 
+                     limits = c(0,8028),
+                     breaks = c(0, 2000, 4000, 6000, 8000)) + 
+    xlab(" ") + ylab("Timepoint") +
     labs(fill = "# of DEGs")

### 2019-04-07 analysis/02_DESeq2_hyp.Rmd
new figures, no control

@@
   labs(fill = "# of DEGs")
+
+dat %>%
+  filter(V1 != "control", V2 != "control")  %>%
+  ggplot( aes(V1, V2)) +
+    geom_tile(aes(fill = V3)) +
+    scale_fill_viridis(na.value="#FFFFFF00", 
+                     limits = c(0, 5025),
+                     breaks = c(0, 1000, 2000, 3000, 4000, 5000)) + 
+    xlab(" ") + ylab("Timepoint") +
+    labs(fill = "# of DEGs")
 ```

### 2019-04-07 analysis/04_DESeq2_gon.Rmd
new figures, no control

@@
 dat %>%
-  filter(V1 != "control")  %>%
+  filter(V1 != "control", V2 != "control")  %>%
   ggplot( aes(V1, V2)) +
@@
     scale_fill_viridis(na.value="#FFFFFF00", 
-                     limits = c(0,8028),
-                     breaks = c(0, 2000, 4000, 6000, 8000)) + 
+                     limits = c(0, 5025),
+                     breaks = c(0, 1000, 2000, 3000, 4000, 5000)) + 
     xlab(" ") + ylab("Timepoint") +

### 2019-04-09 analysis/02_DESeq2_hyp.Rmd
now with heatmap plot

@@
 
-ggplot(dat, aes(V1, V2)) +
+dat %>%
+  filter(V1 != "control", V2 != "control")  %>%
+  ggplot( aes(V1, V2)) +
     geom_tile(aes(fill = V3)) +
     scale_fill_viridis(na.value="#FFFFFF00", 
-                     limits = c(0,8028),
-                     breaks = c(0, 2000, 4000, 6000, 8000)) + 
-  xlab("Timepoint") + ylab("Timepoint") +
+                     limits = c(0, 5025),
+                     breaks = c(0, 1000, 2000, 3000, 4000, 5000)) + 
+    xlab(" ") + ylab("Timepoint") +
     labs(fill = "# of DEGs") 
@@
 
-ggplot(dat, aes(V1, V2)) +
-  geom_tile(aes(fill = V3)) +
-  scale_fill_viridis(na.value="#FFFFFF00", 
-                     limits = c(0,8028),
-                     breaks = c(0, 2000, 4000, 6000, 8000)) + 
-  xlab(" ") + ylab("Timepoint") +
-  labs(fill = "# of DEGs")
-
 dat %>%

### 2019-04-09 analysis/04_DESeq2_gon.Rmd
now with heatmap plot

@@
 
-ggplot(dat, aes(V1, V2)) +
+dat %>%
+  filter(V1 != "control", V2 != "control")  %>%
+  ggplot( aes(V1, V2)) +
     geom_tile(aes(fill = V3)) +
     scale_fill_viridis(na.value="#FFFFFF00", 
-                     limits = c(0,8028),
-                     breaks = c(0, 2000, 4000, 6000, 8000)) + 
-  xlab("Timepoint") + ylab("Timepoint") +
+                     limits = c(0, 5025),
+                     breaks = c(0, 1000, 2000, 3000, 4000, 5000)) + 
+    xlab(" ") + ylab("Timepoint") +
     labs(fill = "# of DEGs")

### 2019-04-09 analysis/00_datawrangling.Rmd
improve manipulation factors

--- 2019-03-19 scripts/00_datawrangling.Rmd
@@
 # relabel incorrectly named variables 
-unique(colData$treatmentTemp)
-colData$treatment <- ifelse(grepl("extend-hatch", colData$treatmentTemp), "extend", colData$treatmentTemp)
+colData$treatment <- ifelse(grepl("extend-hatch", colData$treatmentTemp), "extend",
+                            ifelse(grepl("inc-prolong", colData$treatmentTemp), "prolong",
+                                   ifelse(grepl("m.hatch", colData$treatmentTemp), "m.n2",
+                                   colData$treatmentTemp)))
 
 # drop no-longer-needed cols
-colData <- colData %>% dplyr::select(-c(temp, treatmentTemp))
+colData <- colData %>% dplyr::select(-c(temp, treatmentTemp, NYNO))
 

### 2019-04-09 analysis/05_DESeq2_manipulate.Rmd
manipulation data

---
title: "05_DESeq_manipulation"
output: md_document
---

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)

# load custom functions  
source("../R/functions.R")

knitr::opts_chunk$set(fig.path = '../figures/manipulation/',cache=TRUE)

```

# Manipulation data

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
m.colData <- read.csv("../results/00_colData_manipluation.csv", header = T, row.names = 1)
m.countData <- read.csv("../results/00_countData_manipluation.csv", header = T, row.names = 1)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)

# set levels
m.colData$treatment <- factor(m.colData$treatment, levels = 
                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
                                "m.inc.d17",  "prolong", "extend", "m.n2"))

m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))
```

# Write for loop to do this one for every tissue and for every treatment

```{r althings}
for (eachgroup in levels(m.colData$sextissue)){

  print(eachgroup)

  colData <- m.colData %>%
      dplyr::filter(sextissue == eachgroup) %>%
      droplevels()
  row.names(colData) <- colData$V1

  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  countData <- m.countData %>% dplyr::select(one_of(savecols))

  # check that row and col lenghts are equal
  print(ncol(countData) == nrow(colData))

  dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ treatment )
  dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
  dds <- DESeq(dds) # Differential expression analysis
  vsd <- vst(dds, blind=FALSE) # variance stabilized

#create list of groups
a <- levels(colData$treatment)
b <- levels(colData$treatment)

# comapre all contrasts, save to datafrmes
dat=data.frame()
for (i in a){
  for (j in b){
    if (i != j) {
      k <- paste(i,j, sep = "") #assigns usique rownames
      dat[k,1]<-i               
      dat[k,2]<-j
      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
    }
  }
}

head(dat)

# widen data to create table of degs
rownames(dat) <- NULL #remove row names
data_wide <- spread(dat, V2, V3)
print(data_wide)

dat$V1 <- factor(dat$V1, levels = 
                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
                                "m.inc.d17",  "prolong", "extend", "m.n2"))
dat$V2 <- factor(dat$V2, levels = 
                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
                                "m.inc.d17",  "prolong", "extend", "m.n2"))

allcontrasts <- dat %>%
  ggplot( aes(V1, V2)) +
    geom_tile(aes(fill = V3)) +
    scale_fill_viridis(na.value="#FFFFFF00", 
                     limits = c(0, 5025),
                     breaks = c(0, 1000, 2000, 3000, 4000, 5000)) + 
    xlab(" ") + ylab("Timepoint") +
    labs(fill = "# of DEGs",
         subtitle = eachgroup)
plot(allcontrasts)

# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
percentVar <- round(100 * attr(pcadata, "percentVar"))
percentVar

pca12 <- ggplot(pcadata, aes(PC1, PC2,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
print(pca12)

summary(aov(PC1 ~ treatment, data=pcadata)) 
TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment")

summary(aov(PC2 ~ treatment, data=pcadata)) 
TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment")

pca34 <- ggplot(pcadata, aes(PC3, PC4,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
print(pca34)

summary(aov(PC3 ~ treatment, data=pcadata))

summary(aov(PC4 ~ treatment, data=pcadata))

summary(aov(PC5 ~ treatment, data=pcadata))

summary(aov(PC6 ~ treatment, data=pcadata))

# see http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#heatmap-of-the-count-matrix
sampleDists <- dist(t(assay(vsd)))

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- NULL
colnames(sampleDistMatrix) <- colData$treatment
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors,
         fontsize = 6, 
         main = eachgroup)
}
```

### 2019-04-09 analysis/05_DESeq2_manipulate.Rmd
better manipulation figures

@@
     scale_fill_viridis(na.value="#FFFFFF00", 
-                     limits = c(0, 5025),
-                     breaks = c(0, 1000, 2000, 3000, 4000, 5000)) + 
+                     limits = c(0, 6000),
+                     breaks = c(0, 1000, 2000, 3000, 4000, 5000, 6000)) + 
     xlab(" ") + ylab("Timepoint") +
@@
 
-summary(aov(PC1 ~ treatment, data=pcadata)) 
-TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment")
+print(summary(aov(PC1 ~ treatment, data=pcadata)))
+print(TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment"))
 
-summary(aov(PC2 ~ treatment, data=pcadata)) 
-TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment")
+print(summary(aov(PC2 ~ treatment, data=pcadata))) 
+print(TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment"))
 
@@
 colnames(sampleDistMatrix) <- colData$treatment
-colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
+colors <- colorRampPalette(brewer.pal(9, "Blues"))(255)
 pheatmap(sampleDistMatrix,

### 2019-04-09 analysis/06_DESeq2_manipulate_groups.Rmd
diff groups

--- 2019-04-09 analysis/05_DESeq2_manipulate.Rmd
@@
 ---
-title: "05_DESeq_manipulation"
+title: "06_DESeq_manipulation_groups"
 output: md_document
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/manipulation/',cache=TRUE)
+knitr::opts_chunk$set(fig.path = '../figures/manipulation_groups/',cache=TRUE)
 
@@
 m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))
+
+m.colData$newtreatment <- ifelse(grepl("m.inc.d3", m.colData$treatment), "remove.eggs",
+                          ifelse(grepl("m.inc.d9", m.colData$treatment), "remove.eggs",
+                          ifelse(grepl("m.inc.d17", m.colData$treatment), "remove.eggs",
+                          ifelse(grepl("m.inc.d8", m.colData$treatment), "add.hatchlings",
+                          ifelse(grepl("extend", m.colData$treatment), "add.hatchlings",
+                          ifelse(grepl("prolong", m.colData$treatment), "keep.eggs",
+                          ifelse(grepl("m.n2", m.colData$treatment), "remove.chicks",
+                                   NA)))))))
+
+m.colData$newtreatment <- factor(m.colData$newtreatment, levels = 
+                              c("remove.eggs", "keep.eggs", "add.hatchlings", "remove.chicks"))
+
 ```
@@
                               colData = colData,
-                              design = ~ treatment )
+                              design = ~ newtreatment )
   dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
@@
 #create list of groups
-a <- levels(colData$treatment)
-b <- levels(colData$treatment)
+a <- levels(colData$newtreatment)
+b <- levels(colData$newtreatment)
+
+numDEGs <- function(group1, group2){
+  res <- results(dds, contrast = c("newtreatment", group1, group2), independentFiltering = T)
+  sumpadj <- sum(res$padj < 0.1, na.rm = TRUE)
+  return(sumpadj)
+}
 
@@
 dat$V1 <- factor(dat$V1, levels = 
-                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
-                                "m.inc.d17",  "prolong", "extend", "m.n2"))
+                              c("remove.eggs", "keep.eggs", "add.hatchlings", "remove.chicks"))
 dat$V2 <- factor(dat$V2, levels = 
-                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
-                                "m.inc.d17",  "prolong", "extend", "m.n2"))
+                              c("remove.eggs", "keep.eggs", "add.hatchlings", "remove.chicks"))
 
@@
 # create the dataframe using my function pcadataframe
-pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
+pcadata <- pcadataframe(vsd, intgroup=c("newtreatment"), returnData=TRUE)
 percentVar <- round(100 * attr(pcadata, "percentVar"))
@@
 
-pca12 <- ggplot(pcadata, aes(PC1, PC2,color = treatment)) + 
+pca12 <- ggplot(pcadata, aes(PC1, PC2,color = newtreatment)) + 
   geom_point(size = 2, alpha = 1) +
@@
 
-print(summary(aov(PC1 ~ treatment, data=pcadata)))
-print(TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment"))
+print(summary(aov(PC1 ~ newtreatment, data=pcadata)))
+print(TukeyHSD(aov(PC1 ~ newtreatment, data=pcadata), which = "newtreatment"))
 
-print(summary(aov(PC2 ~ treatment, data=pcadata))) 
-print(TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment"))
+print(summary(aov(PC2 ~ newtreatment, data=pcadata))) 
+print(TukeyHSD(aov(PC2 ~ newtreatment, data=pcadata), which = "newtreatment"))
 
-pca34 <- ggplot(pcadata, aes(PC3, PC4,color = treatment)) + 
+pca34 <- ggplot(pcadata, aes(PC3, PC4,color = newtreatment)) + 
   geom_point(size = 2, alpha = 1) +
@@
 print(pca34)
-
-summary(aov(PC3 ~ treatment, data=pcadata))
-
-summary(aov(PC4 ~ treatment, data=pcadata))
-
-summary(aov(PC5 ~ treatment, data=pcadata))
-
-summary(aov(PC6 ~ treatment, data=pcadata))
-
-# see http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#heatmap-of-the-count-matrix
-sampleDists <- dist(t(assay(vsd)))
-
-sampleDistMatrix <- as.matrix(sampleDists)
-rownames(sampleDistMatrix) <- NULL
-colnames(sampleDistMatrix) <- colData$treatment
-colors <- colorRampPalette(brewer.pal(9, "Blues"))(255)
-pheatmap(sampleDistMatrix,
-         clustering_distance_rows=sampleDists,
-         clustering_distance_cols=sampleDists,
-         col=colors,
-         fontsize = 6, 
-         main = eachgroup)
 }

### 2019-04-15 analysis/07_DESeq2_ALL.Rmd
recent updates

--- 2019-04-09 analysis/05_DESeq2_manipulate.Rmd
@@
 ---
-title: "05_DESeq_manipulation"
+title: "07_DESeq_ALL"
 output: md_document
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/manipulation/',cache=TRUE)
-
+knitr::opts_chunk$set(fig.path = '../figures/ALL/',cache=TRUE)
 ```
-
-# Manipulation data
 
@@
 m.countData <- read.csv("../results/00_countData_manipluation.csv", header = T, row.names = 1)
-geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
 
-# set levels
-m.colData$treatment <- factor(m.colData$treatment, levels = 
-                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
-                                "m.inc.d17",  "prolong", "extend", "m.n2"))
+c.colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
+c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
 
-m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))
+a.colData <- rbind(m.colData, c.colData)
+a.countData <- cbind(m.countData, c.countData)
+
+a.colData$sextissue <- as.factor(paste(a.colData$sex, a.colData$tissue, sep = "_"))
 ```
 
-# Write for loop to do this one for every tissue and for every treatment
+```{r factors} 
+a.colData$treatment <- factor(a.colData$treatment, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9",
+                                "m.inc.d3", "m.inc.d8", "m.inc.d9",
+                                "m.inc.d17",  "prolong", "extend", "m.n2"))
+```
 
 ```{r althings}
-for (eachgroup in levels(m.colData$sextissue)){
+
+for (eachgroup in levels(a.colData$sextissue)){
 
@@
 
-  colData <- m.colData %>%
+  colData <- a.colData %>%
       dplyr::filter(sextissue == eachgroup) %>%
@@
 
-  countData <- m.countData %>% dplyr::select(one_of(savecols))
+  countData <- a.countData %>% dplyr::select(one_of(savecols))
 
@@
 
-dat$V1 <- factor(dat$V1, levels = 
-                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
-                                "m.inc.d17",  "prolong", "extend", "m.n2"))
-dat$V2 <- factor(dat$V2, levels = 
-                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
-                                "m.inc.d17",  "prolong", "extend", "m.n2"))
+#dat$V1 <- factor(dat$V1, levels = 
+#                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+#                                "inc.d17", "hatch", "n5", "n9",
+#                                "m.inc.d3", "m.inc.d8", "m.inc.d9",
+#                                "m.inc.d17",  "prolong", "extend", "m.n2"))
+dat$V2 <- dat$V1
 
@@
 
-summary(aov(PC3 ~ treatment, data=pcadata))
-
-summary(aov(PC4 ~ treatment, data=pcadata))
-
-summary(aov(PC5 ~ treatment, data=pcadata))
-
-summary(aov(PC6 ~ treatment, data=pcadata))
-
 # see http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#heatmap-of-the-count-matrix

### 2019-04-19 analysis/00_datawrangling.Rmd
create bird metadata file
better sample export

@@
 
+# add variable to distinguish characterization study and manipulation
+colData$study <- ifelse(grepl("m.|extend|prolong", colData$treatment), "manipulation", "charcterization")
+
 ## view colData
@@
 ncol(countData) == nrow(colData)
+```
+
+## save bird data
+
+```{r}
+samples <- colData
+write.csv(samples, "../metadata/samples.csv")
+
+birds <- colData %>%
+  select(bird, sex, treatment) %>%
+  distinct()
+write.csv(birds, "../metadata/birds.csv")
 ```
@@
 ```{r write}
-write.csv(colData_manipluation, "../results/00_colData_manipluation.csv", row.names = TRUE) 
+write.csv(colData_manipluation, "../metadata/00_colData_manipluation.csv", row.names = TRUE) 
 write.csv(countData_manipluation, "../results/00_countData_manipluation.csv", row.names = TRUE) 
-write.csv(colData_characterization, "../results/00_colData_characterization.csv", row.names = TRUE) 
+write.csv(colData_characterization, "../metadata/00_colData_characterization.csv", row.names = TRUE) 
 write.csv(countData_characterization, "../results/00_countData_characterization.csv", row.names = TRUE) 

### 2019-04-19 analysis/99_translate.Rmd
first pass at translation

---
title: "translation-es"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/espanol/')
```

## Tranlsating the sample information for downstreatm analysis and data viz

```{r cars}
library(tidyverse)
library(cowplot)
samples <- read.csv("../metadata/samples.csv" , row.names = 1) 
str(samples)

muestras <- samples
names(muestras) <- c("nombrearchivo", "pajaro", "sexo", "tejido", "tratamiento", "grupo", "estudio")
str(muestras)

levels(muestras$sexo) 
levels(muestras$tejido)
levels(muestras$tratamiento)
levels(muestras$estudio)

levels(muestras$sexo) <- c("mujer", "hombre")
levels(muestras$tejido) <- c("gonada", "hipotalamo", "pituitaria")
levels(muestras$tratamiento) <- c("construir", "control", "extender", "escotilla", "incubación 17", "incubacion 03", "incubacion 09", "poner", "eliminar 17", "eliminar 03"," precoz 09"," eliminar 09"," eliminar 21"," extender n5"," extender n9"," prolongar")
levels(muestras$estudio) <- c("caracterizacion", "manipulacion")
```

```{r datasummary}
theme_rmh <- function(){ 
    theme_bw(base_size=10) +
        theme(
            panel.grid.minor.x  = element_blank(),
           panel.grid.minor.y  = element_blank(),
            strip.background = element_rect(colour="white", fill="white"),
              legend.position = "bottom",           
           legend.margin=margin(t=-0.1, r=0, b=-0.1, l=-0.1, unit="cm"),
           legend.key.size = unit(0.5, "cm"))
}

a <- muestras %>%
  #filter(estudio == "caracterizacion")  %>%
ggplot(aes(x = tratamiento, fill = sexo)) +  
  geom_bar() + 
  facet_wrap(~ tejido) +
  theme_rmh() +
  labs(x = "Hitos del cuidado parental (natural y manipulado).",
       y = "número de muestras",
       subtitle = "Muestras de la HPG de padres \nde paloma de roca masculinos y femeninos") + 
  coord_flip()

b <- samples %>%
ggplot(aes(x = treatment, fill = sex)) +  
  geom_bar() + 
  facet_wrap(~ tissue) +
  theme_rmh() +
  labs(x = "Milestones of parental care (natural and manipualted",
       y = "Number of samples",
       subtitle = "Samples from the HPG axis of \nmale and female rock dove parents") + 
  coord_flip()

plot_grid(b,a, labels = c("english", "espanñol"), label_size=7)
```

### 2019-04-23 analysis/01_pca_all.Rmd
pca

---
title: "01_pca"
output: md_document
---

```{r setup}
library(ggfortify) # for PCA analysis
library(cluster)

knitr::opts_chunk$set(echo = TRUE, cache = T, fig.path = '../figures/pca/')
```

```{r rawdata}
colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
head(colData)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
```

```{r pca}
# https://cran.r-project.org/web/packages/ggfortify/vignettes/plot_pca.html

# pca
pca <- prcomp(t(countData))
autoplot(pca, data = colData, colour = 'tissue') 
autoplot(pca, data = colData, colour = 'sex') 
autoplot(pca, data = colData, colour = 'treatment') 
autoplot(pca, data = colData, colour = 'tissue', shape = "sex") 
autoplot(pca, data = colData, colour = 'tissue', shape = "sex", alpha = "treatment")
autoplot(pca, data = colData, colour = 'treatment', shape = "sex", alpha = "tissue")

# kmeans
autoplot(clara(t(countData),3), frame = TRUE)

# probability ellipse
autoplot(pam(t(countData), 3), frame = TRUE, frame.type = 'norm')
```

### 2019-04-23 analysis/00_datawrangling.Rmd
rename save files

--- 2019-04-09 analysis/00_datawrangling.Rmd
@@
 ncol(countData) == nrow(colData)
+```
+
+## save bird and sample and count data
+
+```{r}
+samples <- colData
+
+birds <- colData %>%
+  select(bird, sex, treatment) %>%
+  distinct()
 ```
@@
 ```{r write}
-write.csv(colData_manipluation, "../results/00_colData_manipluation.csv", row.names = TRUE) 
+write.csv(countData, "../results/00_counts.csv")
+write.csv(samples, "../metadata/00_samples.csv")
+write.csv(birds, "../metadata/00_birds.csv")
+
+write.csv(colData_manipluation, "../metadata/00_colData_manipluation.csv", row.names = TRUE) 
 write.csv(countData_manipluation, "../results/00_countData_manipluation.csv", row.names = TRUE) 
-write.csv(colData_characterization, "../results/00_colData_characterization.csv", row.names = TRUE) 
+write.csv(colData_characterization, "../metadata/00_colData_characterization.csv", row.names = TRUE) 
 write.csv(countData_characterization, "../results/00_countData_characterization.csv", row.names = TRUE)
+
 write.csv(geneinfo, "../results/00_geneinfo.csv", row.names = TRUE)

### 2019-04-23 analysis/01_pca_all.Rmd
add pca figs

@@
 ```{r setup}
+library(tidyverse)
 library(ggfortify) # for PCA analysis
 library(cluster)
+library(cowplot)
+
+theme_rmh <- function(){ 
+    theme_bw(base_size=10) +
+        theme(
+            panel.grid.minor.x  = element_blank(),
+           panel.grid.minor.y  = element_blank(),
+            strip.background = element_rect(colour="white", fill="white"),
+              legend.position = "right",           
+           legend.margin=margin(t=-0.1, r=0, b=-0.1, l=-0.1, unit="cm"),
+           legend.key.size = unit(0.5, "cm"))
+}
 
@@
 ```{r rawdata}
-colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
-countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
-head(colData)
-geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
+counts <- read.csv("../results/00_countData_characterization.csv", row.names = 1, header = T)
+
+samples <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, header = T)
+geneinfo <- read.csv("../metadata//00_geneinfo.csv", row.names = 1, header = T)
+
+head(samples)
 ```
@@
 # pca
-pca <- prcomp(t(countData))
-autoplot(pca, data = colData, colour = 'tissue') 
-autoplot(pca, data = colData, colour = 'sex') 
-autoplot(pca, data = colData, colour = 'treatment') 
-autoplot(pca, data = colData, colour = 'tissue', shape = "sex") 
-autoplot(pca, data = colData, colour = 'tissue', shape = "sex", alpha = "treatment")
-autoplot(pca, data = colData, colour = 'treatment', shape = "sex", alpha = "tissue")
+pca <- prcomp(t(counts))
+autoplot(pca, data = samples, colour = 'tissue') 
+autoplot(pca, data = samples, colour = 'sex') 
+autoplot(pca, data = samples, colour = 'treatment') 
+autoplot(pca, data = samples, colour = 'tissue', shape = "sex") 
+autoplot(pca, data = samples, colour = 'tissue', shape = "sex", alpha = "treatment")
+autoplot(pca, data = samples, colour = 'treatment', shape = "sex", alpha = "tissue")
 
 # kmeans
-autoplot(clara(t(countData),3), frame = TRUE)
+autoplot(clara(t(counts),3), frame = TRUE)
 
 # probability ellipse
-autoplot(pam(t(countData), 3), frame = TRUE, frame.type = 'norm')
+autoplot(pam(t(counts), 3), frame = TRUE, frame.type = 'norm')
 ```
+
+```{r pca-es}
+p <- autoplot(pca, data = samples, colour = 'tissue', shape = "sex")
+
+en <- p + labs(subtitle = "PCA plot in English") +
+  scale_colour_discrete(name = "colour", labels = c("rosado", "verde", "azul")) +
+  scale_shape_discrete(name = "shape", labels = c("circle", "square")) +
+  theme_rmh()
+
+es <- p + labs(subtitle = "PCA plot en español") +
+  scale_colour_discrete(name = "color", labels = c("rosado", "verde", "azul")) +
+  scale_shape_discrete(name = "forma", labels = c("circulo", "triangulo")) +
+  theme_rmh()
+
+enes <- plot_grid(en,es)
+
+enes
+
+ggsave("../figures/espanol/pca-en-es.png", width=6, height=3, dpi=300)
+
+pdf(file="../figures/espanol/pca-en-es.pdf", width=6, height=3)
+plot(enes)    
+dev.off()
+```
+
+```{r subset}
+for (eachgroup in levels(samples$treatment)){
+
+  print(eachgroup)
+
+  colData <- samples %>%
+      dplyr::filter(treatment == eachgroup) %>%
+      droplevels()
+  row.names(colData) <- colData$V1
+
+  savecols <- as.character(colData$V1) 
+  savecols <- as.vector(savecols)
+
+  countData <- counts %>% dplyr::select(one_of(savecols))
+
+  # check that row and col lenghts are equal
+  print(ncol(countData) == nrow(colData))
+
+  pca <- prcomp(t(countData))
+
+  p <- autoplot(pca, data = colData, colour = 'tissue', shape = "sex")
+
+en <- p + labs(subtitle = eachgroup) +
+  theme_rmh()
+
+plot(en) 
+}
+```

### 2019-04-23 analysis/00_datawrangling.Rmd
better files

@@
 
-```{r setup, include=FALSE}
+```{r setup}
 # https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html
@@
 
-write.csv(geneinfo, "../results/00_geneinfo.csv", row.names = TRUE)
+write.csv(geneinfo, "../metadata/00_geneinfo.csv", row.names = TRUE)
 ```

### 2019-04-24 analysis/01_DESeq2.Rmd
attempting full DESeq normalization

---
title: "07_DESeq_ALL"
output: md_document
---

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)

# load custom functions  
source("../R/functions.R")

knitr::opts_chunk$set(fig.path = '../figures/ALL/', cache=TRUE)
```

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
counts <- read.csv("../results/00_countData_characterization.csv", row.names = 1, header = T)

samples <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, header = T)
geneinfo <- read.csv("../metadata//00_geneinfo.csv", row.names = 1, header = T)

head(samples)
```

```{r factors} 
samples$treatment <- factor(samples$treatment, levels = 
                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
                                "inc.d17", "hatch", "n5", "n9"))
summary(samples)
```

```{r dds}
dds <- DESeqDataSetFromMatrix(countData = counts,
                              colData = samples,
                              design = ~ treatment * sex * tissue )

dds <- dds[ rowSums(counts(dds)) > 10, ] ## pre-filter genes 
dds <- DESeq(dds) # Differential expression analysis
```

```{r vsd}
vsd <- vst(dds, blind=FALSE) # variance stabilized 
write.csv(assay(vsd), "../results/01_vsd_characterization.csv")
```

```{r pca}
# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("treatment", "tissue", "sex"), returnData=TRUE)
percentVar <- round(100 * attr(pcadata, "percentVar"))
percentVar

pca12 <- ggplot(pcadata, aes(PC1, PC2, color = tissue, shape = sex, alpha = treatment)) + 
  geom_point(size = 2) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
pca12

summary(aov(PC1 ~ treatment, data=pcadata))
TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment")

summary(aov(PC2 ~ treatment, data=pcadata))
TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment")

pca34 <- ggplot(pcadata, aes(PC3, PC4, color = tissue, shape = sex, alpha = treatment)) + 
  geom_point(size = 2) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
pca34

summary(aov(PC3 ~ treatment, data=pcadata))
TukeyHSD(aov(PC3 ~ treatment, data=pcadata), which = "treatment")

summary(aov(PC4 ~ treatment, data=pcadata))
TukeyHSD(aov(PC4 ~ treatment, data=pcadata), which = "treatment")
```

### 2019-04-24 analysis/01_pca_all.Rmd
more work on spanish translations

@@
 pca <- prcomp(t(counts))
+
 autoplot(pca, data = samples, colour = 'tissue') 
@@
 autoplot(pam(t(counts), 3), frame = TRUE, frame.type = 'norm')
+
+head(pca)
 ```
@@
 en <- p + labs(subtitle = "PCA plot in English") +
-  scale_colour_discrete(name = "colour", labels = c("rosado", "verde", "azul")) +
+  scale_colour_discrete(name = "colour", labels = c("pink", "green", "blue")) +
   scale_shape_discrete(name = "shape", labels = c("circle", "square")) +

### 2019-04-24 analysis/01_pca_all.Rmd
more work on spanish translations

@@
 library(cowplot)
+library(stringr)
 
@@
 
-en <- p + labs(subtitle = "PCA plot in English") +
-  scale_colour_discrete(name = "colour", labels = c("pink", "green", "blue")) +
-  scale_shape_discrete(name = "shape", labels = c("circle", "square")) +
-  theme_rmh()
+en <- p + labs(subtitle = "Principle Component Analysis") +
+  scale_colour_discrete(name = "tissue", labels = c("hypothalamus", "pituitary", "gonad")) +
+  scale_shape_discrete(name = "sex", labels = c("female", "male")) +
+  xlab(stringr::str_replace(p$labels$x, "PC", "Principle Component ")) +
+  ylab(stringr::str_replace(p$labels$y, "PC", "Principle Component ")) +
+  theme_rmh() + guides(colour = guide_legend(order = 2), 
+              shape = guide_legend(order = 1))
 
-es <- p + labs(subtitle = "PCA plot en español") +
-  scale_colour_discrete(name = "color", labels = c("rosado", "verde", "azul")) +
-  scale_shape_discrete(name = "forma", labels = c("circulo", "triangulo")) +
-  theme_rmh()
+es <- p + labs(subtitle = "Análisis de Componentes Principales ") +
+  scale_colour_discrete(name = "tejido", labels = c("hipotálamo", "pituitaria", "gónada")) +
+  scale_shape_discrete(name = "sexo", labels = c("femenino", "masculino")) +
+  xlab(stringr::str_replace(p$labels$x, "PC", "Componente Principal ")) +
+  ylab(stringr::str_replace(p$labels$y, "PC", "Componente Principal ")) +
+  theme_rmh() + guides(colour = guide_legend(order = 2), 
+              shape = guide_legend(order = 1))
 

### 2019-04-25 analysis/01_DESeq2.Rmd
parallel but not fast enough yet

@@
 library(viridis)
+
+library("BiocParallel")
+register(MulticoreParam(4))
 
@@
                               colData = samples,
-                              design = ~ treatment * sex * tissue )
+                              design = ~ sex * tissue * treatment )
 
-dds <- dds[ rowSums(counts(dds)) > 10, ] ## pre-filter genes 
-dds <- DESeq(dds) # Differential expression analysis
+dds <- estimateSizeFactors(dds)
+dds
+
+# here filtering on moderate counts in 10 samples out of totla
+nsamp <- 24
+keep <- rowSums(counts(dds, normalized=TRUE) >= 10) >= nsamp
+table(keep)
+
+dds <- dds[keep,] 
+dds
+
+dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
 ```

### 2019-04-25 analysis/01_limma_all.Rmd
minor word change
now with normalized counts

--- 2019-03-19 scripts/01_limma_all.Rmd
@@
 ---
-title: "01_limma_voom"
+title: "01_limma_all"
 output: md_document
@@
 library(cowplot)
-library(ggplot2)
 
@@
 # import "colData" which contains sample information and "countData" which contains read counts
-colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
+colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
 countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
 head(colData)
-geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
+geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
 ```
@@
 
+# find and print data
+names(parentalobject)
+head(countData)
+head(parentalobject$counts)
+head(parentalobject$pseudo.counts)
+
+write.csv(parentalobject$pseudo.counts, "../results/01_pseudo.counts.csv")
+
 #  perform likelihood ratio test and thresholded testing
@@
 topTags(tr)
+tr$table
 ```
@@
 title("Colored by Sex")
+```
+
+```{r plotMDS-colorshape}
+colData$sexTissue <- paste(colData$sex, colData$tissue, sep = "_")
+
+plotMDS(parentalobject,col=col.treatment, labels = colData$sexTissue)
+legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
+title("Colored by Treatment")
 ```
@@
 ```
-
-# volcano plots
-
-```{r volcanoplots}
-# from http://www.compbio.dundee.ac.uk/user/pschofield/Projects/teaching_pg/workshops/biocDGE.html#maplots
-
-lrt <- glmLRT(fit,coef=2)
-topTags(lrt)
-
-tt <- topTags(lrt,n=10000)$table
-
-ggplot(data=tt) + geom_point(aes(x=logFC,y=-log(FDR),color=logCPM)) +
-  scale_colour_gradientn(colours=c("#000000" ,"#FF0000" ))
-```

### 2019-04-25 analysis/01_pca_all.Rmd
less output
now with normalized counts

--- 2019-04-24 analysis/01_pca_all.Rmd
@@
 library(cowplot)
+library(stringr)
+library(gridExtra)
+library(grid)
 
 theme_rmh <- function(){ 
-    theme_bw(base_size=10) +
+    theme_bw(base_size=14) +
         theme(
@@
 
+treatmentcolors <-  c("control" = "#bdbdbd", 
+                              "bldg" =  "#d95f02", 
+                              "lay"  = "#a1d99b", 
+                              "inc.d3" = "#74c476", 
+                              "inc.d9"  = "#41ab5d", 
+                              "inc.d17" = "#238b45", 
+                              "hatch"  = "#807dba", 
+                              "n5"  = "#6a51a3", 
+                               "n9" = "#4a1486")
+
 knitr::opts_chunk$set(echo = TRUE, cache = T, fig.path = '../figures/pca/')
@@
 samples <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, header = T)
+
+samples$treatment <- factor(samples$treatment, levels = 
+                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                                "inc.d17", "hatch", "n5", "n9"))
+samples$tissue <- factor(samples$tissue, levels = 
+                              c("hypothalamus", "pituitary", "gonad"))
+
 geneinfo <- read.csv("../metadata//00_geneinfo.csv", row.names = 1, header = T)
-
-head(samples)
 ```
 
-```{r pca}
+```{r raw}
 # https://cran.r-project.org/web/packages/ggfortify/vignettes/plot_pca.html
@@
 
-autoplot(pca, data = samples, colour = 'tissue') 
-autoplot(pca, data = samples, colour = 'sex') 
-autoplot(pca, data = samples, colour = 'treatment') 
-autoplot(pca, data = samples, colour = 'tissue', shape = "sex") 
-autoplot(pca, data = samples, colour = 'tissue', shape = "sex", alpha = "treatment")
-autoplot(pca, data = samples, colour = 'treatment', shape = "sex", alpha = "tissue")
+p <- autoplot(pca, data = samples, alpha = 'tissue', shape = "sex")
 
-# kmeans
-autoplot(clara(t(counts),3), frame = TRUE)
+rawPCA112 <- p + labs(subtitle = "Raw counts") +
+  scale_colour_discrete(name = "tissue", labels = c("hypothalamus", "pituitary", "gonad")) +
+  scale_shape_discrete(name = "sex", labels = c("female", "male")) +
+  #xlab(stringr::str_replace(p$labels$x, "PC", "Principle Component ")) +
+  #ylab(stringr::str_replace(p$labels$y, "PC", "Principle Component ")) +
+  theme_rmh() + guides(alpha = guide_legend(order = 1), 
+              shape = guide_legend(order = 2)) +
+  stat_ellipse(aes(linetype = sex, colour = tissue)) 
+rawPCA112
 
-# probability ellipse
-autoplot(pam(t(counts), 3), frame = TRUE, frame.type = 'norm')
-
-head(pca)
 ```
 
-```{r pca-es}
-p <- autoplot(pca, data = samples, colour = 'tissue', shape = "sex")
+## with normalize edgeR data
 
-en <- p + labs(subtitle = "PCA plot in English") +
-  scale_colour_discrete(name = "colour", labels = c("pink", "green", "blue")) +
-  scale_shape_discrete(name = "shape", labels = c("circle", "square")) +
-  theme_rmh()
+```{r normalized}
 
-es <- p + labs(subtitle = "PCA plot en español") +
-  scale_colour_discrete(name = "color", labels = c("rosado", "verde", "azul")) +
-  scale_shape_discrete(name = "forma", labels = c("circulo", "triangulo")) +
-  theme_rmh()
+pseudocounts <- read.csv("../results/01_pseudo.counts.csv", row.names = 1)
+head(pseudocounts)
+pca <- prcomp(t(pseudocounts))
 
-enes <- plot_grid(en,es)
+p <- autoplot(pca, data = samples, alpha = 'tissue', shape = "sex")
+p
 
-enes
-
-ggsave("../figures/espanol/pca-en-es.png", width=6, height=3, dpi=300)
-
-pdf(file="../figures/espanol/pca-en-es.pdf", width=6, height=3)
-plot(enes)    
-dev.off()
+edgRpca12 <- p + labs(subtitle = "Normalized counts from edgeR") +
+  scale_alpha_discrete(name = "tissue", labels = c("hypothalamus", "pituitary", "gonad")) +
+  scale_shape_discrete(name = "sex", labels = c("female", "male")) +
+  #xlab(stringr::str_replace(p$labels$x, "PC", "Principle Component ")) +
+  #ylab(stringr::str_replace(p$labels$y, "PC", "Principle Component ")) +
+  theme_rmh() + guides(alpha = guide_legend(order = 1), 
+              shape = guide_legend(order = 2)) +
+  stat_ellipse(aes(linetype = sex, colour = tissue))
+edgRpca12
 ```
 
-```{r subset}
+```{r sidebyside}
+
+sidebyside <- plot_grid(rawPCA112 + theme(legend.position = "none"),
+                        edgRpca12,
+                        rel_widths = c(0.4,0.6))
+sidebyside
+
+```
+
+## subsets of normalized data show by tissue and sex
+
+```{r tissuesex}
+plotlist = list()
 for (eachgroup in levels(samples$treatment)){
@@
 
-  countData <- counts %>% dplyr::select(one_of(savecols))
+  countData <- pseudocounts %>% dplyr::select(one_of(savecols))
 
@@
 
-  p <- autoplot(pca, data = colData, colour = 'tissue', shape = "sex")
+  p <- autoplot(pca, data = colData, alpha = 'tissue', shape = "sex", color = "treatment")
 
 en <- p + labs(subtitle = eachgroup) +
-  theme_rmh()
+  theme_rmh() +
+    theme_rmh()  +
+  stat_ellipse(aes(linetype = sex, alpha = tissue, color = treatment)) +
+   theme_bw(base_size=8) +
+    theme(legend.position = "bottom") +
+  scale_color_manual(values = treatmentcolors) + 
+  guides(color = FALSE)
 
-plot(en) 
+legend <- get_legend(en)
+
+en <- en  + theme(legend.position = "none")
+
+plotlist[[eachgroup]] = en
+
 }
+
+main <- grid.arrange(grobs=plotlist,ncol=3)
+
+withlegend <- plot_grid( main, legend, rel_heights  = c(3, .3), ncol = 1)
+withlegend
+
 ```
+
+## sex and timepoint within a tissue
+
+```{r timepointsex}
+plotlist = list()
+for (eachtissue in levels(samples$tissue)){
+
+  print(eachtissue)
+
+  colData <- samples %>%
+      dplyr::filter(tissue == eachtissue) %>%
+      droplevels()
+  row.names(colData) <- colData$V1
+
+  savecols <- as.character(colData$V1) 
+  savecols <- as.vector(savecols)
+
+  countData <- pseudocounts %>% dplyr::select(one_of(savecols))
+
+  # check that row and col lenghts are equal
+  print(ncol(countData) == nrow(colData))
+
+  pca <- prcomp(t(countData))
+
+  p <- autoplot(pca, data = colData, colour = 'treatment', shape = "sex")
+
+en <- p + labs(subtitle = eachtissue) +
+  theme_rmh() +  guides(colour = guide_legend(order = 1), 
+              shape = guide_legend(order = 2)) +
+  stat_ellipse(aes(linetype = sex, colour = treatment)) +
+  theme_bw(base_size=8) +
+    theme(legend.position = "bottom") +
+  scale_color_manual(values =  treatmentcolors)
+
+legend <- get_legend(en)
+
+en <- en  + theme(legend.position = "none")
+
+plotlist[[eachtissue]] = en
+
+#plot(en) 
+}
+
+main <- grid.arrange(grobs=plotlist,ncol=3)
+
+withlegend <- plot_grid( main, legend, rel_heights  = c(3, .3), ncol = 1)
+withlegend
+
+```

### 2019-05-12 analysis/00_datawrangling.Rmd
now with summary table

@@
 print(countDataTibble, n_extra = 0)
+#rm(countDataTibble)
 ```
@@
 
+colData$study <- factor(colData$study, 
+                         levels = c("charcterization", "manipulation"))
+
 ## view colData
@@
 ncol(countData) == nrow(colData)
+
+colData %>% select(sex, tissue, treatment, study)  %>%  summary()
+
 ```
@@
   distinct()
+
 ```

### 2019-05-13 analysis/05_DESeq2_manipulate.Rmd
removed unused pca stats
tweeks

@@
 
-knitr::opts_chunk$set(fig.path = '../figures/manipulation/',cache=TRUE)
-
+knitr::opts_chunk$set(fig.path = '../figures/manipulation/')
 ```
@@
 # import "colData" which contains sample information and "countData" which contains read counts
-m.colData <- read.csv("../results/00_colData_manipluation.csv", header = T, row.names = 1)
+m.colData <- read.csv("../metadata/00_colData_manipluation.csv", header = T, row.names = 1)
 m.countData <- read.csv("../results/00_countData_manipluation.csv", header = T, row.names = 1)
-geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
+geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
 
@@
 m.colData$treatment <- factor(m.colData$treatment, levels = 
-                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
-                                "m.inc.d17",  "prolong", "extend", "m.n2"))
+                              c("m.inc.d3",  "m.inc.d9",
+                                "m.inc.d17", "m.inc.d8", "prolong", "extend", "m.n2"))
 
@@
   labs(subtitle = eachgroup)
-print(pca34)
-
-summary(aov(PC3 ~ treatment, data=pcadata))
-
-summary(aov(PC4 ~ treatment, data=pcadata))
-
-summary(aov(PC5 ~ treatment, data=pcadata))
-
-summary(aov(PC6 ~ treatment, data=pcadata))
+#print(pca34)
 

### 2019-05-13 analysis/05_DESeq2_manipulate.Rmd
now with fewer groupings

@@
                               c("m.inc.d3",  "m.inc.d9",
-                                "m.inc.d17", "m.inc.d8", "prolong", "extend", "m.n2"))
+                                "m.inc.d17", "m.n2",
+                                "m.inc.d8", "prolong", "extend"))
 
 m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))
+
+m.colData$outcome <- ifelse(grepl("d3|d9|d17|n2", m.colData$treatment), "remove", 
+                     ifelse(grepl("d8", m.colData$treatment),"early",
+                     ifelse(grepl("prolong", m.colData$treatment),"prolong",
+                     ifelse(grepl("extend", m.colData$treatment),"extend", NA))))
+m.colData$outcome <- factor(m.colData$outcome, levels = 
+                              c("remove",  "prolong",
+                                "early", "extend"))
+summary(m.colData[c(7,3,4,5,9)])
 ```
@@
 
-```{r althings}
+```{r althings, eval = FALSE}
+
 for (eachgroup in levels(m.colData$sextissue)){
@@
 dat$V1 <- factor(dat$V1, levels = 
-                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
-                                "m.inc.d17",  "prolong", "extend", "m.n2"))
+                              c("m.inc.d3",  "m.inc.d9",
+                                "m.inc.d17", "m.n2",
+                                "m.inc.d8", "prolong", "extend"))
 dat$V2 <- factor(dat$V2, levels = 
-                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
-                                "m.inc.d17",  "prolong", "extend", "m.n2"))
+                              c("m.inc.d3",  "m.inc.d9",
+                                "m.inc.d17", "m.n2",
+                                "m.inc.d8", "prolong", "extend"))
 
@@
 colors <- colorRampPalette(brewer.pal(9, "Blues"))(255)
-pheatmap(sampleDistMatrix,
-         clustering_distance_rows=sampleDists,
-         clustering_distance_cols=sampleDists,
-         col=colors,
-         fontsize = 6, 
-         main = eachgroup)
+
+#pheatmap(sampleDistMatrix,
+#         clustering_distance_rows=sampleDists,
+#         clustering_distance_cols=sampleDists,
+#         col=colors,
+ #        fontsize = 6, 
+ #        main = eachgroup)
 }
 ```
+
+```{r allthings-outcome}
+
+for (eachgroup in levels(m.colData$sextissue)){
+
+  print(eachgroup)
+
+  colData <- m.colData %>%
+      dplyr::filter(sextissue == eachgroup) %>%
+      droplevels()
+  row.names(colData) <- colData$V1
+
+  savecols <- as.character(colData$V1) 
+  savecols <- as.vector(savecols)
+
+  countData <- m.countData %>% dplyr::select(one_of(savecols))
+
+  # check that row and col lenghts are equal
+  print(ncol(countData) == nrow(colData))
+
+  dds <- DESeqDataSetFromMatrix(countData = countData,
+                              colData = colData,
+                              design = ~ outcome )
+  dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
+  dds <- DESeq(dds) # Differential expression analysis
+  vsd <- vst(dds, blind=FALSE) # variance stabilized 
+head(vsd)
+
+numDEGs <- function(group1, group2){
+  res <- results(dds, contrast = c("outcome", group1, group2), independentFiltering = T)
+  sumpadj <- sum(res$padj < 0.1, na.rm = TRUE)
+  return(sumpadj)}
+
+#create list of groups
+a <- levels(colData$outcome)
+b <- levels(colData$outcome)
+
+# comapre all contrasts, save to datafrmes
+dat=data.frame()
+for (i in a){
+  for (j in b){
+    if (i != j) {
+      k <- paste(i,j, sep = "") #assigns usique rownames
+      dat[k,1]<-i               
+      dat[k,2]<-j
+      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
+    }
+  }
+}
+
+head(dat)
+
+# widen data to create table of degs
+rownames(dat) <- NULL #remove row names
+data_wide <- spread(dat, V2, V3)
+print(data_wide)
+
+dat$V1 <- factor(dat$V1, levels = 
+                              c("remove",  "prolong",
+                                "early", "extend"))
+dat$V2 <- factor(dat$V2, levels = 
+                              c("remove",  "prolong",
+                                "early", "extend"))
+
+allcontrasts <- dat %>%
+  ggplot( aes(V1, V2)) +
+    geom_tile(aes(fill = V3)) +
+    scale_fill_viridis(na.value="#FFFFFF00") + 
+    xlab(" ") + ylab("Timepoint") +
+    labs(fill = "# of DEGs",
+         subtitle = eachgroup)
+plot(allcontrasts)
+
+# create the dataframe using my function pcadataframe
+pcadata <- pcadataframe(vsd, intgroup=c("outcome"), returnData=TRUE)
+percentVar <- round(100 * attr(pcadata, "percentVar"))
+percentVar
+
+pca12 <- ggplot(pcadata, aes(PC1, PC2,color = outcome)) + 
+  geom_point(size = 2, alpha = 1) +
+  stat_ellipse(type = "t") +
+  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
+  theme_cowplot(font_size = 8, line_size = 0.25) +
+  labs(subtitle = eachgroup)
+print(pca12)
+
+print(summary(aov(PC1 ~ outcome, data=pcadata)))
+print(TukeyHSD(aov(PC1 ~ outcome, data=pcadata), which = "outcome"))
+
+print(summary(aov(PC2 ~ outcome, data=pcadata))) 
+print(TukeyHSD(aov(PC2 ~ outcome, data=pcadata), which = "outcome"))
+
+pca34 <- ggplot(pcadata, aes(PC3, PC4,color = outcome)) + 
+  geom_point(size = 2, alpha = 1) +
+  stat_ellipse(type = "t") +
+  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
+  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
+  theme_cowplot(font_size = 8, line_size = 0.25) +
+  labs(subtitle = eachgroup)
+#print(pca34)
+}
+```

### 2019-05-14 analysis/05_DESeq2_manipulate.Rmd
with better order
now with a heatmap of top 30 DEGs

@@
                                 "early", "extend"))
-summary(m.colData[c(7,3,4,5,9)])
+summary(m.colData[c(7,3,4,5,8,9)])
 ```
@@
 
-```{r althings, eval = FALSE}
+```{r althings, eval = T}
+
+## subset for test
+
+m.colData <- m.colData %>%
+  filter(sextissue == "female_pituitary") %>%
+  droplevels()
 
@@
  #        main = eachgroup)
+
+# Multi-factor designs
+# revals is my function easily calculate multiple multi factor designs
+# it is built around the documentaion from DESeq which looks more simply like
+# res <- results(dds,contrast=c("columnname", "factor1", "factor2"))
+# then I extract the p-values from within the results for showing number of DEGes
+resvals <- function(contrastvector, mypval){
+  res <- results(dds, contrast = c(contrastvector[1],contrastvector[2],contrastvector[3]), independentFiltering = T)
+  sumpvalue <- sum(res$pvalue < mypval, na.rm = TRUE)
+  #print(sumpvalue)
+  sumpadj <- sum(res$padj < mypval, na.rm = TRUE)
+  print(sumpadj)
+  vals <- cbind(res$pvalue, res$padj)
+  pvalcolname <- as.character(paste("pval",contrastvector[1],contrastvector[2],contrastvector[3], sep=""))
+  padjcolname <- as.character(paste("padj",contrastvector[1],contrastvector[2],contrastvector[3], sep=""))
+  colnames(vals) <- c(pvalcolname, padjcolname)
+  return(vals)
+}
+## DEG by contrasts at 0.1 pvalue
+contrast1 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'm.inc.d9'), mypval = 0.01) #12
+contrast2 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'm.inc.d17'), mypval = 0.01) #73
+contrast3 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'm.n2'), mypval = 0.01) #284
+contrast4 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'm.inc.d8'), mypval = 0.01) #15
+contrast5 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'prolong'), mypval = 0.01) #555
+contrast6 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'extend'), mypval = 0.01) #1109
+contrast7 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'm.inc.d17'), mypval = 0.01) #280
+contrast8 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'm.n2'), mypval = 0.01) #440
+contrast9 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'm.inc.d8'), mypval = 0.01) #484
+contrast10 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'prolong'), mypval = 0.01) #529
+contrast11 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'extend'), mypval = 0.01) #1177
+
+contrast12 <- resvals(contrastvector = c('treatment', 'm.inc.d17', 'm.n2'), mypval = 0.01) #5
+contrast13 <- resvals(contrastvector = c('treatment', 'm.inc.d17', 'm.inc.d8'), mypval = 0.01) #553
+contrast14 <- resvals(contrastvector = c('treatment', 'm.inc.d17', 'prolong'), mypval = 0.01) #339
+contrast15 <- resvals(contrastvector = c('treatment', 'm.inc.d17', 'extend'), mypval = 0.01) #782
+contrast16 <- resvals(contrastvector = c('treatment', 'm.n2', 'm.inc.d8'), mypval = 0.01) #1123
+contrast17 <- resvals(contrastvector = c('treatment', 'm.n2', 'prolong'), mypval = 0.01) #265
+contrast18 <- resvals(contrastvector = c('treatment', 'm.n2', 'extend'), mypval = 0.01) #552
+contrast19 <- resvals(contrastvector = c('treatment', 'm.inc.d8', 'prolong'), mypval = 0.01) #828
+contrast20 <- resvals(contrastvector = c('treatment', 'm.inc.d8', 'extend'), mypval = 0.01) #1110
+contrast21 <- resvals(contrastvector = c('treatment', 'prolong', 'extend'), mypval = 0.01) #64
+
+DEGs <- assay(vsd)
+DEGs <- cbind(DEGs, contrast1, contrast2, contrast3, contrast4, contrast5,
+              contrast6, contrast7, contrast8, contrast9, contrast10,
+              contrast11, contrast12, contrast13, contrast14, contrast15,
+              contrast16, contrast17, contrast18, contrast19, contrast20,contrast21)
+DEGs <- as.data.frame(DEGs) # convert matrix to dataframe
+DEGs$rownames <- rownames(DEGs)  # add the rownames to the dataframe
+
+#DEGs[is.na(DEGs)] <-  1
+
+DEGs$padjmin <- with(DEGs, pmin(pvaltreatmentm.inc.d3m.inc.d9, 
+  padjtreatmentm.inc.d3m.inc.d9,               
+  pvaltreatmentm.inc.d3m.inc.d17, padjtreatmentm.inc.d3m.inc.d17,              
+  pvaltreatmentm.inc.d3m.n2, padjtreatmentm.inc.d3m.n2,                  
+  pvaltreatmentm.inc.d3m.inc.d8, padjtreatmentm.inc.d3m.inc.d8,               
+  pvaltreatmentm.inc.d3prolong, padjtreatmentm.inc.d3prolong,               
+  pvaltreatmentm.inc.d3extend, padjtreatmentm.inc.d3extend,                
+  pvaltreatmentm.inc.d9m.inc.d17, padjtreatmentm.inc.d9m.inc.d17,             
+  pvaltreatmentm.inc.d9m.n2, padjtreatmentm.inc.d9m.n2,                  
+  pvaltreatmentm.inc.d9m.inc.d8, padjtreatmentm.inc.d9m.inc.d8,              
+  pvaltreatmentm.inc.d9prolong, padjtreatmentm.inc.d9prolong,               
+  pvaltreatmentm.inc.d9extend, padjtreatmentm.inc.d9extend,                
+  pvaltreatmentm.inc.d17m.n2, padjtreatmentm.inc.d17m.n2,                 
+  pvaltreatmentm.inc.d17m.inc.d8, padjtreatmentm.inc.d17m.inc.d8,             
+  pvaltreatmentm.inc.d17prolong, padjtreatmentm.inc.d17prolong,
+  pvaltreatmentm.inc.d17extend,padjtreatmentm.inc.d17extend, 
+  pvaltreatmentm.n2m.inc.d8, padjtreatmentm.n2m.inc.d8,                  
+  pvaltreatmentm.n2prolong, padjtreatmentm.n2prolong,                   
+  pvaltreatmentm.n2extend, padjtreatmentm.n2extend,                    
+  pvaltreatmentm.inc.d8prolong, padjtreatmentm.inc.d8prolong,               
+  pvaltreatmentm.inc.d8extend, padjtreatmentm.inc.d8extend,                
+  pvaltreatmentprolongextend, padjtreatmentprolongextend))
+
+sigDEGs <- DEGs %>% arrange(padjmin)
+sigDEGs <- head(sigDEGs,30)
+sigDEGs <- as.data.frame(sigDEGs)
+rownames(sigDEGs) <- sigDEGs$rownames
+drop.cols <-colnames(sigDEGs[,grep("padj|pval|pmin|rownames", colnames(sigDEGs))])
+sigDEGs <- sigDEGs %>% dplyr::select(-one_of(drop.cols))
+sigDEGs <- as.matrix(sigDEGs)
+sigDEGs <- sigDEGs - rowMeans(sigDEGs)
+
+paletteLength <- 30
+myBreaks <- c(seq(min(sigDEGs), 0, length.out=ceiling(paletteLength/2) + 1), 
+                seq(max(sigDEGs)/paletteLength, max(sigDEGs), length.out=floor(paletteLength/2)))
+
+anndf <- m.colData %>% dplyr::select(treatment)
+rownames(anndf) <- m.colData$V1
+
+sigDEGs <- as.matrix(sigDEGs) 
+pheatmap(sigDEGs, show_rownames = F, show_colnames = F,
+         color = viridis(30),
+         breaks=myBreaks,
+         annotation_col=anndf)
+
+pheatmap(sigDEGs, kmeans_k = 5,
+         show_rownames = F, show_colnames = F,
+         color = viridis(30),
+         breaks=myBreaks,
+         annotation_col=anndf)
 }
@@
 
-```{r allthings-outcome}
+```{r allthings-outcome, eval = F}
 
@@
     geom_tile(aes(fill = V3)) +
-    scale_fill_viridis(na.value="#FFFFFF00") + 
+    scale_fill_viridis(na.value="#FFFFFF00", 
+                     limits = c(0, 6000),
+                     breaks = c(0, 1000, 2000, 3000, 4000, 5000, 6000)) + 
     xlab(" ") + ylab("Timepoint") +
@@
 #print(pca34)
+
 }

### 2019-05-16 analysis/01_limma_all.Rmd
no large file or long output in md
removeing df outputs with head()
removed about output table

@@
 names(parentalobject)
-head(countData)
-head(parentalobject$counts)
-head(parentalobject$pseudo.counts)
+#head(countData)
+#head(parentalobject$counts)
+#head(parentalobject$pseudo.counts)
 
@@
 topTags(tr)
-tr$table
+#tr$table
 ```

### 2019-05-16 analysis/05_DESeq2_manipulate.Rmd
2 heatmaps for all
update levels

@@
 m.colData$treatment <- factor(m.colData$treatment, levels = 
-                              c("m.inc.d3",  "m.inc.d9",
-                                "m.inc.d17", "m.n2",
-                                "m.inc.d8", "prolong", "extend"))
+                              c("m.inc.d3",  "m.inc.d8",
+                                "m.inc.d9", "m.inc.d17",
+                                "prolong", "extend", "m.n2"))
 
@@
 
-m.colData$outcome <- ifelse(grepl("d3|d9|d17|n2", m.colData$treatment), "remove", 
-                     ifelse(grepl("d8", m.colData$treatment),"early",
-                     ifelse(grepl("prolong", m.colData$treatment),"prolong",
-                     ifelse(grepl("extend", m.colData$treatment),"extend", NA))))
+m.colData$outcome <- ifelse(grepl("d3|d9|d17", m.colData$treatment), "end inc", 
+                     ifelse(grepl("d8|n2", m.colData$treatment),"end hatch",
+                     ifelse(grepl("prolong", m.colData$treatment),"prolong inc",
+                     ifelse(grepl("extend", m.colData$treatment),"delay hatch", NA))))
+
 m.colData$outcome <- factor(m.colData$outcome, levels = 
-                              c("remove",  "prolong",
-                                "early", "extend"))
+                              c("end inc",  "end hatch",
+                                "prolong inc", "delay hatch"))
 summary(m.colData[c(7,3,4,5,8,9)])
@@
 ```{r althings, eval = T}
-
 ## subset for test 
-
-m.colData <- m.colData %>%
-  filter(sextissue == "female_pituitary") %>%
-  droplevels()
+#m.colData <- m.colData %>%
+#  filter(sextissue == "female_pituitary") %>%
+#  droplevels()
 
@@
 dat$V1 <- factor(dat$V1, levels = 
-                              c("m.inc.d3",  "m.inc.d9",
-                                "m.inc.d17", "m.n2",
-                                "m.inc.d8", "prolong", "extend"))
+                              c("m.inc.d3",  "m.inc.d8",
+                                "m.inc.d9", "m.inc.d17",
+                                "prolong", "extend", "m.n2"))
 dat$V2 <- factor(dat$V2, levels = 
-                              c("m.inc.d3",  "m.inc.d9",
-                                "m.inc.d17", "m.n2",
-                                "m.inc.d8", "prolong", "extend"))
+                              c("m.inc.d3",  "m.inc.d8",
+                                "m.inc.d9", "m.inc.d17",
+                                "prolong", "extend", "m.n2"))
 
@@
 dat$V1 <- factor(dat$V1, levels = 
-                              c("remove",  "prolong",
-                                "early", "extend"))
+                              c("end inc",  "end hatch",
+                                "prolong inc", "delay hatch"))
 dat$V2 <- factor(dat$V2, levels = 
-                              c("remove",  "prolong",
-                                "early", "extend"))
+                              c("end inc",  "end hatch",
+                                "prolong inc", "delay hatch"))
 

### 2019-05-16 analysis/05_DESeq2_manipulate.Rmd
new order of factors
more genes

@@
 sigDEGs <- DEGs %>% arrange(padjmin)
-sigDEGs <- head(sigDEGs,30)
+sigDEGs <- head(sigDEGs,300)
 sigDEGs <- as.data.frame(sigDEGs)
@@
 
-```{r allthings-outcome, eval = F}
+```{r allthings-outcome, eval = T}
 

### 2019-05-16 analysis/05_DESeq2_manipulate.Rmd
now with 500

@@
 sigDEGs <- DEGs %>% arrange(padjmin)
-sigDEGs <- head(sigDEGs,300)
+sigDEGs <- head(sigDEGs,500)
 sigDEGs <- as.data.frame(sigDEGs)

### 2019-05-17 analysis/05_DESeq2_manipulate.Rmd
now with file names

@@
          breaks=myBreaks,
-         annotation_col=anndf)
+         annotation_col=anndf,
+         main = eachgroup)
 
@@
          breaks=myBreaks,
-         annotation_col=anndf)
+         annotation_col=anndf,
+         main = eachgroup)
 }

### 2019-05-18 analysis/05_DESeq2_manipulate.Rmd
now with candidate genes

@@
 
+## candidate genes
+
+# make dataframe with geneids and names and counts
+candidates <- cbind(geneinfo, countData)
+candidates <- candidates %>%
+  filter(grepl("OXT|AVP|POMC|PRKCZ|GRM|JUN", Name)) 
+row.names(candidates) <- candidates$Name
+candidates <- candidates %>% select(-row.names, -Name, -geneid, -entrezid)
+
+#head(candidates)
+
+candidates <- as.data.frame(t(candidates))
+candidates$RNAseqID <- rownames(candidates)
+#head(candidates)
+
+candidates <- candidates %>% gather(gene, value, -RNAseqID)  %>% # https://tidyr.tidyverse.org/reference/gather.html
+  filter(RNAseqID != "gene")
+candidates$value <- as.numeric(candidates$value)
+candidates$V1  <- candidates$RNAseqID
+#head(candidates)
+
+candidatecounts <- full_join(candidates, colData)
+#head(candidatecounts)
+
+candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
+#head(candidatecounts)
+
+candidatecounts$gene <- as.factor(candidatecounts$gene)
+levels(candidatecounts$gene)
+
+p1 <- candidatecounts %>%
+  filter(gene %in% c( "AVP", "AVPR1A", "AVPR1B", "AVPR2")) %>%
+  ggplot(aes(x = treatment, y = value, fill = treatment)) +
+  geom_boxplot() +
+  facet_wrap(~gene, scales = "free", nrow = 1) +
+  theme(axis.text.x = element_blank(),
+        legend.position = "bottom") +
+  labs(subtitle = eachgroup)
+p1
+
+print(p1)
+
+p2 <- candidatecounts %>%
+  filter(gene %in% c( "GRM2", "GRM4", "GRM5", "GRM7", "GRM8")) %>%
+  ggplot(aes(x = treatment, y = value, fill = treatment)) +
+  geom_boxplot() +
+  facet_wrap(~gene, scales = "free", nrow = 1) +
+  theme(axis.text.x = element_blank(),
+        legend.position = "bottom") +
+  labs(subtitle = eachgroup)
+p2
+
+print(p2)
+
+p3 <- candidatecounts %>%
+  filter(gene %in% c( "JUN", "JUND", "OXT", "POMC", "PRKCZ")) %>%
+  ggplot(aes(x = treatment, y = value, fill = treatment)) +
+  geom_boxplot() +
+  facet_wrap(~gene, scales = "free", nrow = 1) +
+  theme(axis.text.x = element_blank(),
+        legend.position = "bottom") +
+  labs(subtitle = eachgroup)
+p3
+
+print(p3)
+
 }

### 2019-05-18 analysis/05_DESeq2_manipulate.Rmd
now with box plots

@@
          main = eachgroup)
+
+## candidate genes
+
+# make dataframe with geneids and names and counts
+candidates <- cbind(geneinfo, countData)
+candidates <- candidates %>%
+  filter(grepl("OXT|AVP|POMC|PRKCZ|GRM|JUN", Name)) 
+row.names(candidates) <- candidates$Name
+candidates <- candidates %>% select(-row.names, -Name, -geneid, -entrezid)
+
+#head(candidates)
+
+candidates <- as.data.frame(t(candidates))
+candidates$RNAseqID <- rownames(candidates)
+#head(candidates)
+
+candidates <- candidates %>% gather(gene, value, -RNAseqID)  %>% # https://tidyr.tidyverse.org/reference/gather.html
+  filter(RNAseqID != "gene")
+candidates$value <- as.numeric(candidates$value)
+candidates$V1  <- candidates$RNAseqID
+#head(candidates)
+
+candidatecounts <- full_join(candidates, colData)
+#head(candidatecounts)
+
+candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
+#head(candidatecounts)
+
+candidatecounts$gene <- as.factor(candidatecounts$gene)
+levels(candidatecounts$gene)
+
+p1 <- candidatecounts %>%
+  filter(gene %in% c( "AVP", "AVPR1A", "AVPR1B", "AVPR2")) %>%
+  ggplot(aes(x = treatment, y = value, fill = treatment)) +
+  geom_boxplot() +
+  facet_wrap(~gene, scales = "free", nrow = 1) +
+  theme(axis.text.x = element_blank(),
+        legend.position = "bottom") +
+  labs(subtitle = eachgroup)
+p1
+
+print(p1)
+
+p2 <- candidatecounts %>%
+  filter(gene %in% c( "GRM2", "GRM4", "GRM5", "GRM7", "GRM8")) %>%
+  ggplot(aes(x = treatment, y = value, fill = treatment)) +
+  geom_boxplot() +
+  facet_wrap(~gene, scales = "free", nrow = 1) +
+  theme(axis.text.x = element_blank(),
+        legend.position = "bottom") +
+  labs(subtitle = eachgroup)
+p2
+
+print(p2)
+
+p3 <- candidatecounts %>%
+  filter(gene %in% c( "JUN", "JUND", "OXT", "POMC", "PRKCZ")) %>%
+  ggplot(aes(x = treatment, y = value, fill = treatment)) +
+  geom_boxplot() +
+  facet_wrap(~gene, scales = "free", nrow = 1) +
+  theme(axis.text.x = element_blank(),
+        legend.position = "bottom") +
+  labs(subtitle = eachgroup)
+p3
+
+print(p3)
+
 }
@@
 
-## candidate genes
-
-# make dataframe with geneids and names and counts
-candidates <- cbind(geneinfo, countData)
-candidates <- candidates %>%
-  filter(grepl("OXT|AVP|POMC|PRKCZ|GRM|JUN", Name)) 
-row.names(candidates) <- candidates$Name
-candidates <- candidates %>% select(-row.names, -Name, -geneid, -entrezid)
-
-#head(candidates)
-
-candidates <- as.data.frame(t(candidates))
-candidates$RNAseqID <- rownames(candidates)
-#head(candidates)
-
-candidates <- candidates %>% gather(gene, value, -RNAseqID)  %>% # https://tidyr.tidyverse.org/reference/gather.html
-  filter(RNAseqID != "gene")
-candidates$value <- as.numeric(candidates$value)
-candidates$V1  <- candidates$RNAseqID
-#head(candidates)
-
-candidatecounts <- full_join(candidates, colData)
-#head(candidatecounts)
-
-candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
-#head(candidatecounts)
-
-candidatecounts$gene <- as.factor(candidatecounts$gene)
-levels(candidatecounts$gene)
-
-p1 <- candidatecounts %>%
-  filter(gene %in% c( "AVP", "AVPR1A", "AVPR1B", "AVPR2")) %>%
-  ggplot(aes(x = treatment, y = value, fill = treatment)) +
-  geom_boxplot() +
-  facet_wrap(~gene, scales = "free", nrow = 1) +
-  theme(axis.text.x = element_blank(),
-        legend.position = "bottom") +
-  labs(subtitle = eachgroup)
-p1
-
-print(p1)
-
-p2 <- candidatecounts %>%
-  filter(gene %in% c( "GRM2", "GRM4", "GRM5", "GRM7", "GRM8")) %>%
-  ggplot(aes(x = treatment, y = value, fill = treatment)) +
-  geom_boxplot() +
-  facet_wrap(~gene, scales = "free", nrow = 1) +
-  theme(axis.text.x = element_blank(),
-        legend.position = "bottom") +
-  labs(subtitle = eachgroup)
-p2
-
-print(p2)
-
-p3 <- candidatecounts %>%
-  filter(gene %in% c( "JUN", "JUND", "OXT", "POMC", "PRKCZ")) %>%
-  ggplot(aes(x = treatment, y = value, fill = treatment)) +
-  geom_boxplot() +
-  facet_wrap(~gene, scales = "free", nrow = 1) +
-  theme(axis.text.x = element_blank(),
-        legend.position = "bottom") +
-  labs(subtitle = eachgroup)
-p3
-
-print(p3)
-
 }

### 2019-05-19 analysis/05_DESeq2_manipulate.Rmd
cleaning up, prep for separarate functions

--- 2019-05-17 analysis/05_DESeq2_manipulate.Rmd
@@
 library(viridis)
+library(edgeR)
 
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/manipulation/')
+knitr::opts_chunk$set(fig.path = '../figures/manipulation/', cache = TRUE)
 ```
 
-# Manipulation data
+## Manipulation data
 
@@
 
-# Write for loop to do this one for every tissue and for every treatment
+## Make a PCA plot, heat map, and box plots for each tissue for each sex
 
@@
 #m.colData <- m.colData %>%
-#  filter(sextissue == "female_pituitary") %>%
+#  filter(sextissue == "male_hypothalamus") %>%
 # droplevels()
@@
                               design = ~ treatment )
-  dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
+
+  print(dds)
+  nrow(colData) > 50
+
+  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
+  print(dds)
+
   dds <- DESeq(dds) # Differential expression analysis
+
   vsd <- vst(dds, blind=FALSE) # variance stabilized
@@
 # comapre all contrasts, save to datafrmes
-dat=data.frame()
+totalDEGS=data.frame()
 for (i in a){
@@
       k <- paste(i,j, sep = "") #assigns usique rownames
-      dat[k,1]<-i               
-      dat[k,2]<-j
-      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
+      print(k)
+      totalDEGS[k,1]<-i               
+      totalDEGS[k,2]<-j
+      totalDEGS[k,3]<- numDEGs(i,j) #caluculates number of DEGs
     }
@@
 
-head(dat)
+head(totalDEGS)
 
 # widen data to create table of degs
-rownames(dat) <- NULL #remove row names
-data_wide <- spread(dat, V2, V3)
-print(data_wide)
+rownames(totalDEGS) <- NULL #remove row names
+totalDEGS_wide <- spread(totalDEGS, V2, V3)
+print(totalDEGS_wide)
 
-dat$V1 <- factor(dat$V1, levels = 
+totalDEGS$V1 <- factor(totalDEGS$V1, levels = 
                               c("m.inc.d3",  "m.inc.d8",
@@
                                 "prolong", "extend", "m.n2"))
-dat$V2 <- factor(dat$V2, levels = 
+totalDEGS$V2 <- factor(totalDEGS$V2, levels = 
                               c("m.inc.d3",  "m.inc.d8",
@@
 
-allcontrasts <- dat %>%
+allcontrasts <- totalDEGS %>%
   ggplot( aes(V1, V2)) +
@@
     scale_fill_viridis(na.value="#FFFFFF00", 
-                     limits = c(0, 6000),
-                     breaks = c(0, 1000, 2000, 3000, 4000, 5000, 6000)) + 
+                     limits = c(0, 3000),
+                     breaks = c(0, 1000, 2000, 3000)) + 
     xlab(" ") + ylab("Timepoint") +
     labs(fill = "# of DEGs",
-         subtitle = eachgroup)
+         subtitle = "eachgroup")
 plot(allcontrasts)
@@
 percentVar <- round(100 * attr(pcadata, "percentVar"))
-percentVar
+print(percentVar)
 
-pca12 <- ggplot(pcadata, aes(PC1, PC2,color = treatment)) + 
-  geom_point(size = 2, alpha = 1) +
-  stat_ellipse(type = "t") +
-  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
+pca1 <- ggplot(pcadata, aes(treatment, PC1,color = treatment)) + 
+  geom_boxplot() +
+  ylab(paste0("PC1: ", percentVar[1],"% variance")) +
+  xlab(NULL) +
+  theme_cowplot(font_size = 8, line_size = 0.25) +
+  labs(subtitle = "eachgroup") +
+  theme(legend.position = "none")
+
+pca2 <- ggplot(pcadata, aes(treatment, PC2,color = treatment)) + 
+  geom_boxplot() +
   ylab(paste0("PC2: ", percentVar[2],"% variance")) +
+  xlab(NULL) +
   theme_cowplot(font_size = 8, line_size = 0.25) +
-  labs(subtitle = eachgroup)
-print(pca12)
+  labs(subtitle = "eachgroup") +
+  theme(legend.position = "none")
+
+plot_grid(pca1, pca2)
 
@@
 print(summary(aov(PC2 ~ treatment, data=pcadata))) 
-print(TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment"))
+print(summary(aov(PC3 ~ treatment, data=pcadata))) 
+print(summary(aov(PC4 ~ treatment, data=pcadata)))
 
-pca34 <- ggplot(pcadata, aes(PC3, PC4,color = treatment)) + 
-  geom_point(size = 2, alpha = 1) +
-  stat_ellipse(type = "t") +
-  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
-  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
-  theme_cowplot(font_size = 8, line_size = 0.25) +
-  labs(subtitle = eachgroup)
-#print(pca34)
+## heamap with minimum pvalue
 
-# see http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#heatmap-of-the-count-matrix
-sampleDists <- dist(t(assay(vsd)))
+# make dataframe counts
+DEGs <- assay(vsd)
+DEGs <- as.data.frame(DEGs)
 
-sampleDistMatrix <- as.matrix(sampleDists)
-rownames(sampleDistMatrix) <- NULL
-colnames(sampleDistMatrix) <- colData$treatment
-colors <- colorRampPalette(brewer.pal(9, "Blues"))(255)
+# add pvalues
 
-#pheatmap(sampleDistMatrix,
-#         clustering_distance_rows=sampleDists,
-#         clustering_distance_cols=sampleDists,
-#         col=colors,
- #        fontsize = 6, 
- #        main = eachgroup)
+for (i in a){
+  for (j in b){
+    if (i != j) {
+      k <- paste(i,j, sep = " vs ")
+      print(k)
+      results <- resvals2(i,j)
+      DEGs <- cbind(DEGs,results)
+    }
+  }
+}
 
-# Multi-factor designs
-# revals is my function easily calculate multiple multi factor designs
-# it is built around the documentaion from DESeq which looks more simply like
-# res <- results(dds,contrast=c("columnname", "factor1", "factor2"))
-# then I extract the p-values from within the results for showing number of DEGes
-resvals <- function(contrastvector, mypval){
-  res <- results(dds, contrast = c(contrastvector[1],contrastvector[2],contrastvector[3]), independentFiltering = T)
-  sumpvalue <- sum(res$pvalue < mypval, na.rm = TRUE)
-  #print(sumpvalue)
-  sumpadj <- sum(res$padj < mypval, na.rm = TRUE)
-  print(sumpadj)
-  vals <- cbind(res$pvalue, res$padj)
-  pvalcolname <- as.character(paste("pval",contrastvector[1],contrastvector[2],contrastvector[3], sep=""))
-  padjcolname <- as.character(paste("padj",contrastvector[1],contrastvector[2],contrastvector[3], sep=""))
-  colnames(vals) <- c(pvalcolname, padjcolname)
-  return(vals)
-}
-## DEG by contrasts at 0.1 pvalue
-contrast1 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'm.inc.d9'), mypval = 0.01) #12
-contrast2 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'm.inc.d17'), mypval = 0.01) #73
-contrast3 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'm.n2'), mypval = 0.01) #284
-contrast4 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'm.inc.d8'), mypval = 0.01) #15
-contrast5 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'prolong'), mypval = 0.01) #555
-contrast6 <- resvals(contrastvector = c('treatment', 'm.inc.d3', 'extend'), mypval = 0.01) #1109
-contrast7 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'm.inc.d17'), mypval = 0.01) #280
-contrast8 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'm.n2'), mypval = 0.01) #440
-contrast9 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'm.inc.d8'), mypval = 0.01) #484
-contrast10 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'prolong'), mypval = 0.01) #529
-contrast11 <- resvals(contrastvector = c('treatment', 'm.inc.d9', 'extend'), mypval = 0.01) #1177
+DEGsmatrix <- DEGs
+DEGsmatrix <- as.matrix(DEGs)
+padjmin <- rowMins(DEGsmatrix, na.rm = T) 
+padjmin <- as.data.frame(padjmin)
 
-contrast12 <- resvals(contrastvector = c('treatment', 'm.inc.d17', 'm.n2'), mypval = 0.01) #5
-contrast13 <- resvals(contrastvector = c('treatment', 'm.inc.d17', 'm.inc.d8'), mypval = 0.01) #553
-contrast14 <- resvals(contrastvector = c('treatment', 'm.inc.d17', 'prolong'), mypval = 0.01) #339
-contrast15 <- resvals(contrastvector = c('treatment', 'm.inc.d17', 'extend'), mypval = 0.01) #782
-contrast16 <- resvals(contrastvector = c('treatment', 'm.n2', 'm.inc.d8'), mypval = 0.01) #1123
-contrast17 <- resvals(contrastvector = c('treatment', 'm.n2', 'prolong'), mypval = 0.01) #265
-contrast18 <- resvals(contrastvector = c('treatment', 'm.n2', 'extend'), mypval = 0.01) #552
-contrast19 <- resvals(contrastvector = c('treatment', 'm.inc.d8', 'prolong'), mypval = 0.01) #828
-contrast20 <- resvals(contrastvector = c('treatment', 'm.inc.d8', 'extend'), mypval = 0.01) #1110
-contrast21 <- resvals(contrastvector = c('treatment', 'prolong', 'extend'), mypval = 0.01) #64
+sigDEGs <- cbind(DEGs,padjmin)
 
-DEGs <- assay(vsd)
-DEGs <- cbind(DEGs, contrast1, contrast2, contrast3, contrast4, contrast5,
-              contrast6, contrast7, contrast8, contrast9, contrast10,
-              contrast11, contrast12, contrast13, contrast14, contrast15,
-              contrast16, contrast17, contrast18, contrast19, contrast20,contrast21)
-DEGs <- as.data.frame(DEGs) # convert matrix to dataframe
-DEGs$rownames <- rownames(DEGs)  # add the rownames to the dataframe
-
-#DEGs[is.na(DEGs)] <-  1
-
-DEGs$padjmin <- with(DEGs, pmin(pvaltreatmentm.inc.d3m.inc.d9, 
-  padjtreatmentm.inc.d3m.inc.d9,               
-  pvaltreatmentm.inc.d3m.inc.d17, padjtreatmentm.inc.d3m.inc.d17,              
-  pvaltreatmentm.inc.d3m.n2, padjtreatmentm.inc.d3m.n2,                  
-  pvaltreatmentm.inc.d3m.inc.d8, padjtreatmentm.inc.d3m.inc.d8,               
-  pvaltreatmentm.inc.d3prolong, padjtreatmentm.inc.d3prolong,               
-  pvaltreatmentm.inc.d3extend, padjtreatmentm.inc.d3extend,                
-  pvaltreatmentm.inc.d9m.inc.d17, padjtreatmentm.inc.d9m.inc.d17,             
-  pvaltreatmentm.inc.d9m.n2, padjtreatmentm.inc.d9m.n2,                  
-  pvaltreatmentm.inc.d9m.inc.d8, padjtreatmentm.inc.d9m.inc.d8,              
-  pvaltreatmentm.inc.d9prolong, padjtreatmentm.inc.d9prolong,               
-  pvaltreatmentm.inc.d9extend, padjtreatmentm.inc.d9extend,                
-  pvaltreatmentm.inc.d17m.n2, padjtreatmentm.inc.d17m.n2,                 
-  pvaltreatmentm.inc.d17m.inc.d8, padjtreatmentm.inc.d17m.inc.d8,             
-  pvaltreatmentm.inc.d17prolong, padjtreatmentm.inc.d17prolong,
-  pvaltreatmentm.inc.d17extend,padjtreatmentm.inc.d17extend, 
-  pvaltreatmentm.n2m.inc.d8, padjtreatmentm.n2m.inc.d8,                  
-  pvaltreatmentm.n2prolong, padjtreatmentm.n2prolong,                   
-  pvaltreatmentm.n2extend, padjtreatmentm.n2extend,                    
-  pvaltreatmentm.inc.d8prolong, padjtreatmentm.inc.d8prolong,               
-  pvaltreatmentm.inc.d8extend, padjtreatmentm.inc.d8extend,                
-  pvaltreatmentprolongextend, padjtreatmentprolongextend))
-
-sigDEGs <- DEGs %>% arrange(padjmin)
+sigDEGs <- sigDEGs %>% arrange(padjmin)
 sigDEGs <- head(sigDEGs,500)
@@
          annotation_col=anndf,
-         main = eachgroup)
+         main = "eachgroup")
 
@@
          annotation_col=anndf,
-         main = eachgroup)
-}
-```
+         main = "eachgroup")
 
-```{r allthings-outcome, eval = T}
+## candidate genes
+names(geneinfo)
+names(geneinfo)[4] <- "rownames"
+DEGs$rownames <- row.names(DEGs)
 
-for (eachgroup in levels(m.colData$sextissue)){
+# make dataframe with geneids and names and counts
 
-  print(eachgroup)
+candidates <- full_join(geneinfo, DEGs)
+drop.cols <-colnames(candidates[,grep("padj|pval|pmin", colnames(candidates))])
+candidates <- candidates %>% dplyr::select(-one_of(drop.cols))
+candidates <- candidates %>%
+  filter(Name %in% c( #"JUN", "JUND", "EGR",  "AVP", "AVPR1A", "AVPR1B", "AVPR2", "OXT",  
+                     "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
+                     "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
+                     "PGR", "PRL", "PRLR", "VIP", "VIPR1")) 
+row.names(candidates) <- candidates$Name
+candidates <- candidates %>% select(-row.names, -rownames, -Name, -geneid)
+candidates <- candidates %>% drop_na()
 
-  colData <- m.colData %>%
-      dplyr::filter(sextissue == eachgroup) %>%
-      droplevels()
-  row.names(colData) <- colData$V1
+head(candidates)
 
-  savecols <- as.character(colData$V1) 
-  savecols <- as.vector(savecols)
+candidates <- as.data.frame(t(candidates))
+candidates$RNAseqID <- rownames(candidates)
+head(candidates)
 
-  countData <- m.countData %>% dplyr::select(one_of(savecols))
+candidates <- candidates %>% gather(gene, value, -RNAseqID)  %>% # https://tidyr.tidyverse.org/reference/gather.html
+  filter(RNAseqID != "gene")
+candidates$value <- as.numeric(candidates$value)
+candidates$V1  <- candidates$RNAseqID
+head(candidates)
 
-  # check that row and col lenghts are equal
-  print(ncol(countData) == nrow(colData))
+candidatecounts <- left_join(candidates, colData)
+#head(candidatecounts)
 
-  dds <- DESeqDataSetFromMatrix(countData = countData,
-                              colData = colData,
-                              design = ~ outcome )
-  dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
-  dds <- DESeq(dds) # Differential expression analysis
-  vsd <- vst(dds, blind=FALSE) # variance stabilized 
-head(vsd)
+candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
+#head(candidatecounts)
 
-numDEGs <- function(group1, group2){
-  res <- results(dds, contrast = c("outcome", group1, group2), independentFiltering = T)
-  sumpadj <- sum(res$padj < 0.1, na.rm = TRUE)
-  return(sumpadj)}
+candidatecounts$gene <- as.factor(candidatecounts$gene)
+levels(candidatecounts$gene)
 
-#create list of groups
-a <- levels(colData$outcome)
-b <- levels(colData$outcome)
+p1 <- candidatecounts %>%
+  filter(gene %in% c( "AR", "CYP19A1", "ESR1", "ESR2", "FSHR")) %>%
+  ggplot(aes(x = treatment, y = value, fill = treatment)) +
+  geom_boxplot() +
+  facet_wrap(~gene, scales = "free", nrow = 1) +
+  theme(axis.text.x = element_blank(),
+        legend.position = "bottom") +
+  labs(subtitle = "eachgroup")
+p1
 
-# comapre all contrasts, save to datafrmes
-dat=data.frame()
-for (i in a){
-  for (j in b){
-    if (i != j) {
-      k <- paste(i,j, sep = "") #assigns usique rownames
-      dat[k,1]<-i               
-      dat[k,2]<-j
-      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
-    }
-  }
-}
+print(p1)
 
-head(dat)
+cat("\n")
 
-# widen data to create table of degs
-rownames(dat) <- NULL #remove row names
-data_wide <- spread(dat, V2, V3)
-print(data_wide)
+p2 <- candidatecounts %>%
+  filter(gene %in% c("GHRL", "GAL", "NPVF", "GNRH1", "LHCGR")) %>%
+  ggplot(aes(x = treatment, y = value, fill = treatment)) +
+  geom_boxplot() +
+  facet_wrap(~gene, scales = "free", nrow = 1) +
+  theme(axis.text.x = element_blank(),
+        legend.position = "bottom") +
+  labs(subtitle = "eachgroup")
+p2
 
-dat$V1 <- factor(dat$V1, levels = 
-                              c("end inc",  "end hatch",
-                                "prolong inc", "delay hatch"))
-dat$V2 <- factor(dat$V2, levels = 
-                              c("end inc",  "end hatch",
-                                "prolong inc", "delay hatch"))
+print(p2)
 
-allcontrasts <- dat %>%
-  ggplot( aes(V1, V2)) +
-    geom_tile(aes(fill = V3)) +
-    scale_fill_viridis(na.value="#FFFFFF00", 
-                     limits = c(0, 6000),
-                     breaks = c(0, 1000, 2000, 3000, 4000, 5000, 6000)) + 
-    xlab(" ") + ylab("Timepoint") +
-    labs(fill = "# of DEGs",
-         subtitle = eachgroup)
-plot(allcontrasts)
+cat("\n")
 
-# create the dataframe using my function pcadataframe
-pcadata <- pcadataframe(vsd, intgroup=c("outcome"), returnData=TRUE)
-percentVar <- round(100 * attr(pcadata, "percentVar"))
-percentVar
+p3 <- candidatecounts %>%
+  filter(gene %in% c("PGR", "PRL", "PRLR", "VIP", "VIPR1")) %>%
+  ggplot(aes(x = treatment, y = value, fill = treatment)) +
+  geom_boxplot() +
+  facet_wrap(~gene, scales = "free", nrow = 1) +
+  theme(axis.text.x = element_blank(),
+        legend.position = "bottom") +
+  labs(subtitle = "eachgroup")
+p3
 
-pca12 <- ggplot(pcadata, aes(PC1, PC2,color = outcome)) + 
-  geom_point(size = 2, alpha = 1) +
-  stat_ellipse(type = "t") +
-  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
-  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
-  theme_cowplot(font_size = 8, line_size = 0.25) +
-  labs(subtitle = eachgroup)
-print(pca12)
+print(p3)
 
-print(summary(aov(PC1 ~ outcome, data=pcadata)))
-print(TukeyHSD(aov(PC1 ~ outcome, data=pcadata), which = "outcome"))
+cat("\n")
 
-print(summary(aov(PC2 ~ outcome, data=pcadata))) 
-print(TukeyHSD(aov(PC2 ~ outcome, data=pcadata), which = "outcome"))
+cat("Done with eachgroup")
 
-pca34 <- ggplot(pcadata, aes(PC3, PC4,color = outcome)) + 
-  geom_point(size = 2, alpha = 1) +
-  stat_ellipse(type = "t") +
-  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
-  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
-  theme_cowplot(font_size = 8, line_size = 0.25) +
-  labs(subtitle = eachgroup)
-#print(pca34)
+cat("\n")
 

### 2019-05-19 analysis/05_DESeq2_manipulate_p2.Rmd
cleaning up, prep for separarate functions

## 2nd for loop on data clustered into early or late termination

```{r allthings-outcome, eval = T}

for (eachgroup in levels(m.colData$sextissue)){

  print(eachgroup)

  colData <- m.colData %>%
      dplyr::filter(sextissue == eachgroup) %>%
      droplevels()
  row.names(colData) <- colData$V1

  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  countData <- m.countData %>% dplyr::select(one_of(savecols))

  # check that row and col lenghts are equal
  print(ncol(countData) == nrow(colData))

  dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ outcome )
  dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
  dds <- DESeq(dds) # Differential expression analysis
  vsd <- vst(dds, blind=FALSE) # variance stabilized 
head(vsd)

numDEGs <- function(group1, group2){
  res <- results(dds, contrast = c("outcome", group1, group2), independentFiltering = T)
  sumpadj <- sum(res$padj < 0.1, na.rm = TRUE)
  return(sumpadj)}

#create list of groups
a <- levels(colData$outcome)
b <- levels(colData$outcome)

# comapre all contrasts, save to datafrmes
dat=data.frame()
for (i in a){
  for (j in b){
    if (i != j) {
      k <- paste(i,j, sep = "") #assigns usique rownames
      dat[k,1]<-i               
      dat[k,2]<-j
      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
    }
  }
}

head(dat)

# widen data to create table of degs
rownames(dat) <- NULL #remove row names
data_wide <- spread(dat, V2, V3)
print(data_wide)

dat$V1 <- factor(dat$V1, levels = 
                              c("end inc",  "end hatch",
                                "prolong inc", "delay hatch"))
dat$V2 <- factor(dat$V2, levels = 
                              c("end inc",  "end hatch",
                                "prolong inc", "delay hatch"))

allcontrasts <- dat %>%
  ggplot( aes(V1, V2)) +
    geom_tile(aes(fill = V3)) +
    scale_fill_viridis(na.value="#FFFFFF00", 
                     limits = c(0, 6000),
                     breaks = c(0, 1000, 2000, 3000, 4000, 5000, 6000)) + 
    xlab(" ") + ylab("Timepoint") +
    labs(fill = "# of DEGs",
         subtitle = eachgroup)
plot(allcontrasts)

# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("outcome"), returnData=TRUE)
percentVar <- round(100 * attr(pcadata, "percentVar"))
percentVar

pca12 <- ggplot(pcadata, aes(PC1, PC2,color = outcome)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
print(pca12)

print(summary(aov(PC1 ~ outcome, data=pcadata)))
print(TukeyHSD(aov(PC1 ~ outcome, data=pcadata), which = "outcome"))

print(summary(aov(PC2 ~ outcome, data=pcadata))) 
print(TukeyHSD(aov(PC2 ~ outcome, data=pcadata), which = "outcome"))

pca34 <- ggplot(pcadata, aes(PC3, PC4,color = outcome)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
#print(pca34)

}
```

### 2019-05-19 analysis/05_DESeq2_manipulate.Rmd
run all

@@
     geom_tile(aes(fill = V3)) +
-    scale_fill_viridis(na.value="#FFFFFF00", 
+    scale_fill_viridis(na.value="#440154", 
                      limits = c(0, 3000),
@@
     labs(fill = "# of DEGs",
-         subtitle = "eachgroup")
+         subtitle = eachgroup)
 plot(allcontrasts)
@@
   theme_cowplot(font_size = 8, line_size = 0.25) +
-  labs(subtitle = "eachgroup") +
+  labs(subtitle = eachgroup) +
   theme(legend.position = "none")
@@
          annotation_col=anndf,
-         main = "eachgroup")
+         main = eachgroup)
 
@@
 # make dataframe with geneids and names and counts
+# how to gather: https://tidyr.tidyverse.org/reference/gather.html
 
@@
 candidates <- candidates %>% drop_na()
-
-head(candidates)
-
 candidates <- as.data.frame(t(candidates))
 candidates$RNAseqID <- rownames(candidates)
-head(candidates)
-
-candidates <- candidates %>% gather(gene, value, -RNAseqID)  %>% # https://tidyr.tidyverse.org/reference/gather.html
+candidates <- candidates %>% gather(gene, value, -RNAseqID)  %>% 
   filter(RNAseqID != "gene")
@@
 candidates$V1  <- candidates$RNAseqID
-head(candidates)
-
 candidatecounts <- left_join(candidates, colData)
-#head(candidatecounts)
-
 candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
-#head(candidatecounts)
-
 candidatecounts$gene <- as.factor(candidatecounts$gene)
-levels(candidatecounts$gene)
 
@@
         legend.position = "bottom") +
-  labs(subtitle = "eachgroup")
+  labs(subtitle = eachgroup)
 p1
@@
         legend.position = "bottom") +
-  labs(subtitle = "eachgroup")
+  labs(subtitle = eachgroup)
 p2
@@
         legend.position = "bottom") +
-  labs(subtitle = "eachgroup")
+  labs(subtitle = eachgroup)
 p3

### 2019-05-19 analysis/05_DESeq2_manipulate.Rmd
now with many loops and functions

@@
 
-## Make a PCA plot, heat map, and box plots for each tissue for each sex
+## Run DESeq on all subsets of the data
 
-```{r althings, eval = T}
+```{r dds}
 ## subset for test 
@@
 
+levels(m.colData$sextissue)
+
 for (eachgroup in levels(m.colData$sextissue)){
@@
   print(eachgroup)
+  dds <- subsetDESeq(eachgroup)
 
-  colData <- m.colData %>%
-      dplyr::filter(sextissue == eachgroup) %>%
-      droplevels()
-  row.names(colData) <- colData$V1
+  dfname <- paste("dds", eachgroup, sep = "")
+  assign(paste("dds", eachgroup, sep = ""), dds)
 
-  savecols <- as.character(colData$V1) 
-  savecols <- as.vector(savecols)
+  return(dfname)
+}
 
-  countData <- m.countData %>% dplyr::select(one_of(savecols))
+```
 
-  # check that row and col lenghts are equal
-  print(ncol(countData) == nrow(colData))
+## Create variance stabilized data
 
-  dds <- DESeqDataSetFromMatrix(countData = countData,
-                              colData = colData,
-                              design = ~ treatment )
+```{r vsd}
 
-  print(dds)
-  nrow(colData) > 50
+mydds <- list(ddsmale_hypothalamus) # for testing only
+#mydds <- list(ddsfemale_hypothalamus, ddsfemale_pituitary, ddsfemale_gondad,
+#              ddsmale_hypothalamus, ddsmale_pituitary, ddsmale_gondad)
 
-  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
-  print(dds)
+for (eachdds in mydds){
 
-  dds <- DESeq(dds) # Differential expression analysis
+  vsd <- vst(eachdds, blind=FALSE) # variance stabilized
 
-  vsd <- vst(dds, blind=FALSE) # variance stabilized
+  vdsname <- paste("vsd", eachgroup, sep = "")
+  assign(paste("vsd", eachgroup, sep = ""), vsd)
+  return(vdsname)
+}
+
+```
+
+## Calculate and plot total DEGs
+
+```{r totalDEGs}
 
@@
 b <- levels(colData$treatment)
+
+for (eachdds in mydds){
 
@@
       if (i != j) {
-      k <- paste(i,j, sep = "") #assigns usique rownames
+        k <- paste(i,j, sep = ".") #assigns usique rownames
         print(k)
@@
     }
+        b <- b[-1]  # drop 1st element of second string to not recalculate DEGs
   }
@@
       labs(fill = "# of DEGs",
-         subtitle = eachgroup)
+          subtitle = "eachgroup")
   plot(allcontrasts)
+}
+```
 
@@
       print(k)
-      results <- resvals2(i,j)
+      results <- returnpadj(i,j)
       DEGs <- cbind(DEGs,results)

### 2019-05-19 R/functions.R

@@ -1,3 +1,36 @@
+# run DESeq on subset of data
+# e.g. dds <- subsetDESeq("male_hypothalamus")
+subsetDESeq <- function(eachgroup){
+  
+  # subset to look within one tissue in one sex
+  colData <- m.colData %>%
+    dplyr::filter(sextissue == eachgroup) %>%
+    droplevels()
+  row.names(colData) <- colData$V1
+  
+  # which counts to save
+  savecols <- as.character(colData$V1) 
+  savecols <- as.vector(savecols) 
+  
+  # save counts that match colData
+  countData <- m.countData %>% dplyr::select(one_of(savecols)) 
+  
+  # check that row and col lenghts are equal
+  print(ncol(countData) == nrow(colData))
+  
+  dds <- DESeqDataSetFromMatrix(countData = countData,
+                                colData = colData,
+                                design = ~ treatment )
+  
+  print(dds)
+  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
+  print(dds)
+  
+  dds <- DESeq(dds) # Differential expression analysis
+  return(dds)
+}
+
+
 # print total number of differntially expressed genes
 # numDEGs('m.inc.d3', 'm.inc.d9')
 numDEGs <- function(group1, group2){
@@ -7,20 +40,35 @@ numDEGs <- function(group1, group2){
 }
 
 
-resvals <- function(contrastvector, mypval){
-  res <- results(dds, contrast = c(contrastvector[1],contrastvector[2],contrastvector[3]), independentFiltering = T)
-  sumpvalue <- sum(res$pvalue < mypval, na.rm = TRUE)
-  #print(sumpvalue)
-  sumpadj <- sum(res$padj < mypval, na.rm = TRUE)
-  print(sumpadj)
-  vals <- cbind(res$pvalue, res$padj)
-  pvalcolname <- as.character(paste("pval",contrastvector[1],contrastvector[2],contrastvector[3], sep=""))
-  padjcolname <- as.character(paste("padj",contrastvector[1],contrastvector[2],contrastvector[3], sep=""))
-  colnames(vals) <- c(pvalcolname, padjcolname)
-  return(vals)
+## plot DEGs 
+
+plottotalDEGs <- function(totalDEGS, mysubtitle){
+  
+  totalDEGS$V1 <- factor(totalDEGS$V1, levels = 
+                           c("m.inc.d3",  "m.inc.d8",
[…]

### 2019-05-19 analysis/05_DESeq2_manipulate.Rmd
now with many loops and functions

@@
 library(viridis)
-library(edgeR)
 
@@
 
-## Create variance stabilized data
-
-```{r vsd}
-
-mydds <- list(ddsmale_hypothalamus) # for testing only
-#mydds <- list(ddsfemale_hypothalamus, ddsfemale_pituitary, ddsfemale_gondad,
-#              ddsmale_hypothalamus, ddsmale_pituitary, ddsmale_gondad)
-
-for (eachdds in mydds){
-
-  vsd <- vst(eachdds, blind=FALSE) # variance stabilized
-
-  vdsname <- paste("vsd", eachgroup, sep = "")
-  assign(paste("vsd", eachgroup, sep = ""), vsd)
-  return(vdsname)
-}
-
-```
-
 ## Calculate and plot total DEGs
@@
 ```{r totalDEGs}
+#mydds <- list(ddsmale_hypothalamus) # for testing only
+mydds <- list(ddsfemale_hypothalamus, ddsfemale_pituitary, ddsfemale_gondad,
+              ddsmale_hypothalamus, ddsmale_pituitary, ddsmale_gondad)
 
@@
 
-  head(totalDEGS)
+  totalDEGSname <- paste("totalDEGS", eachgroup, sep = "")
+  assign(paste("totalDEGS", eachgroup, sep = ""), totalDEGS)
+  return(totalDEGSname)
+}
 
-  # widen data to create table of degs
-  rownames(totalDEGS) <- NULL #remove row names
-  totalDEGS_wide <- spread(totalDEGS, V2, V3)
-  print(totalDEGS_wide)
+a <- plottotalDEGs(totalDEGSfemale_hypothalamus, "female hypothalamus")
+b <- plottotalDEGs(totalDEGSfefemale_pituitary, "female pituitary")
+c <- plottotalDEGs(totalDEGSmale_gonad, "female gonad")
+d <- plottotalDEGs(totalDEGSmale_hypothalamus, "male hypothalamus")
+e <- plottotalDEGs(totalDEGSmale_pituitary, "male pituitary")
+f <- plottotalDEGs(totalDEGSmale_gonad, "male gonad")
 
-  totalDEGS$V1 <- factor(totalDEGS$V1, levels = 
-                              c("m.inc.d3",  "m.inc.d8",
-                                "m.inc.d9", "m.inc.d17",
-                                "prolong", "extend", "m.n2"))
-  totalDEGS$V2 <- factor(totalDEGS$V2, levels = 
-                              c("m.inc.d3",  "m.inc.d8",
-                                "m.inc.d9", "m.inc.d17",
-                                "prolong", "extend", "m.n2"))
+plot_grid(a,b,c,d,e,f, nrow = 2)
+```
 
-  allcontrasts <- totalDEGS %>%
-    ggplot( aes(V1, V2)) +
-      geom_tile(aes(fill = V3)) +
-      scale_fill_viridis(na.value="#440154", 
-                      limits = c(0, 3000),
-                      breaks = c(0, 1000, 2000, 3000)) + 
-      xlab(" ") + ylab("Timepoint") +
-      labs(fill = "# of DEGs",
-          subtitle = "eachgroup")
-  plot(allcontrasts)
-}
-```
+## Calculate and plot principal components
+
+```{r pca}
+
+for (eachdds in mydds){
+
+  vsd <- vst(eachdds, blind=FALSE) # variance stabilized
 
@@
 
-plot_grid(pca1, pca2)
+  mypca <- plot_grid(pca1, pca2)
+  plot(mypca)
 
@@
 
+}
+```
+
 ## heamap with minimum pvalue
+
+```{r pheatmap}
+
+for (eachdds in mydds){
+
+  vsd <- vst(eachdds, blind=FALSE) # variance stabilized
 
@@
 
-# add pvalues
+  a <- levels(colData$treatment)
+  b <- levels(colData$treatment)
 
@@
       if (i != j) {
-      k <- paste(i,j, sep = " vs ")
-      print(k)
+        print(i)
         results <- returnpadj(i,j)
@@
     }
+     b <- b[-1]  # drop 1st element of second string to not recalculate DEGs
   }
@@
          main = eachgroup)
+}
 
-## candidate genes
+```
+
+```{r candidates}
+for (eachdds in mydds){
+
+  vsd <- vst(eachdds, blind=FALSE) # variance stabilized
+
+  # make dataframe counts
+  DEGs <- assay(vsd)
+  DEGs <- as.data.frame(DEGs)
+
 names(geneinfo)
@@
 p1 <- candidatecounts %>%
-  filter(gene %in% c( "AR", "CYP19A1", "ESR1", "ESR2", "FSHR")) %>%
+  filter(gene %in% c( "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
+                      "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
+                      "PGR", "PRL", "PRLR", "VIP", "VIPR1")) %>%
   ggplot(aes(x = treatment, y = value, fill = treatment)) +
   geom_boxplot() +
-  facet_wrap(~gene, scales = "free", nrow = 1) +
+  facet_wrap(~gene, scales = "free") +
+  theme_bw(base_size = 8) +
   theme(axis.text.x = element_blank(),
         legend.position = "bottom") +
-  labs(subtitle = eachgroup)
+  labs(subtitle = eachgroup, x = NULL) +
+  guides(fill= guide_legend(nrow=1))
 p1
-
-print(p1)
-
-cat("\n")
-
-p2 <- candidatecounts %>%
-  filter(gene %in% c("GHRL", "GAL", "NPVF", "GNRH1", "LHCGR")) %>%
-  ggplot(aes(x = treatment, y = value, fill = treatment)) +
-  geom_boxplot() +
-  facet_wrap(~gene, scales = "free", nrow = 1) +
-  theme(axis.text.x = element_blank(),
-        legend.position = "bottom") +
-  labs(subtitle = eachgroup)
-p2
-
-print(p2)
-
-cat("\n")
-
-p3 <- candidatecounts %>%
-  filter(gene %in% c("PGR", "PRL", "PRLR", "VIP", "VIPR1")) %>%
-  ggplot(aes(x = treatment, y = value, fill = treatment)) +
-  geom_boxplot() +
-  facet_wrap(~gene, scales = "free", nrow = 1) +
-  theme(axis.text.x = element_blank(),
-        legend.position = "bottom") +
-  labs(subtitle = eachgroup)
-p3
-
-print(p3)
-
-cat("\n")
-
-cat("Done with eachgroup")
-
-cat("\n")
 

### 2019-05-20 analysis/05_DESeq2_manipulate.Rmd
half way there

@@
 
-```{r dds}
-## subset for test 
-#m.colData <- m.colData %>%
-#  filter(sextissue == "male_hypothalamus") %>%
-# droplevels()
-
-levels(m.colData$sextissue)
-
-for (eachgroup in levels(m.colData$sextissue)){
-
-  print(eachgroup)
-  dds <- subsetDESeq(eachgroup)
-
-  dfname <- paste("dds", eachgroup, sep = "")
-  assign(paste("dds", eachgroup, sep = ""), dds)
-
-  return(dfname)
-}
-
+```{r dds, cache = T}
+dds.female_hypothalamus <- subsetDESeq("female_hypothalamus")
+dds.female_pituitary <- subsetDESeq("female_pituitary" )
+dds.female_gonad <- subsetDESeq("female_gonad" )
+dds.male_hypothalamus <- subsetDESeq("male_hypothalamus" )
+dds.male_pituitary <- subsetDESeq("male_pituitary"  )
+dds.male_gondad <- subsetDESeq("male_gonad")
 ```
@@
 
-```{r totalDEGs}
-#mydds <- list(ddsmale_hypothalamus) # for testing only
-mydds <- list(ddsfemale_hypothalamus, ddsfemale_pituitary, ddsfemale_gondad,
-              ddsmale_hypothalamus, ddsmale_pituitary, ddsmale_gondad)
+```{r totalDEGs, eval = F}
+#create list of groups for deseq contrasts
+group1 <- c("m.inc.d3","m.inc.d8", "m.inc.d9", "m.inc.d17", "prolong","extend", "m.n2" ) 
+group2 <- group1
 
-#create list of groups
-a <- levels(colData$treatment)
-b <- levels(colData$treatment)
-
-for (eachdds in mydds){
-
-  # comapre all contrasts, save to datafrmes
-  totalDEGS=data.frame()
-  for (i in a){
-    for (j in b){
-      if (i != j) {
-        k <- paste(i,j, sep = ".") #assigns usique rownames
-        print(k)
-        totalDEGS[k,1]<-i               
-        totalDEGS[k,2]<-j
-        totalDEGS[k,3]<- numDEGs(i,j) #caluculates number of DEGs
-      }
-    }
-        b <- b[-1]  # drop 1st element of second string to not recalculate DEGs
-  }
-
-  totalDEGSname <- paste("totalDEGS", eachgroup, sep = "")
-  assign(paste("totalDEGS", eachgroup, sep = ""), totalDEGS)
-  return(totalDEGSname)
-}
-
-a <- plottotalDEGs(totalDEGSfemale_hypothalamus, "female hypothalamus")
-b <- plottotalDEGs(totalDEGSfefemale_pituitary, "female pituitary")
-c <- plottotalDEGs(totalDEGSmale_gonad, "female gonad")
-d <- plottotalDEGs(totalDEGSmale_hypothalamus, "male hypothalamus")
-e <- plottotalDEGs(totalDEGSmale_pituitary, "male pituitary")
-f <- plottotalDEGs(totalDEGSmale_gonad, "male gonad")
+a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
+b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
+c <- plottotalDEGs(dds.female_gonad, "female gonad")
+d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
+e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
+f <- plottotalDEGs(dds.male_gondad, "male gonad")
 
@@
 ```{r pca}
+
+#mydds <- list(ddsmale_hypothalamus) # for testing only
+mydds <- list(dds.female_hypothalamus, dds.female_pituitary, dds.female_gonad,
+             dds.male_hypothalamus, dds.male_pituitary, dds.male_gondad)
 
@@
     theme_cowplot(font_size = 8, line_size = 0.25) +
-    labs(subtitle = eachgroup) +
+    labs(subtitle = "eachgroup") +
     theme(legend.position = "none")
@@
 
-```{r pheatmap}
+```{r pheatmap, eval = F}
 
@@
 
-  a <- levels(colData$treatment)
-  b <- levels(colData$treatment)
+  a <- levels(m.colData$treatment)
+  b <- a
 
@@
          annotation_col=anndf,
-         main = eachgroup)
+         main = "eachgroup")
 }
@@
 candidates$V1  <- candidates$RNAseqID
-candidatecounts <- left_join(candidates, colData)
+
+candidatecounts <- left_join(candidates, m.colData)
 candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
@@
         legend.position = "bottom") +
-  labs(subtitle = eachgroup, x = NULL) +
+  labs(x = NULL) +
   guides(fill= guide_legend(nrow=1))
-p1
 
+plot(p1)
 }

### 2019-05-20 analysis/05_DESeq2_manipulate.Rmd
now with grids

@@
 
-```{r totalDEGs, eval = F}
+```{r totalDEGs, eval = T}
 #create list of groups for deseq contrasts
@@
 plot_grid(a,b,c,d,e,f, nrow = 2) 
+```
+
+```{r totalDEGs2}
+plot_grid(a + theme_minimal(base_size = 8) + theme(legend.position = "none", axis.text.x = element_text(angle = 90)),
+          b + theme_minimal(base_size = 8) + theme(legend.position = "none", axis.text.x = element_text(angle = 90)),
+          c + theme_minimal(base_size = 8) + theme(axis.text.x = element_text(angle = 90)),
+          d + theme_minimal(base_size = 8) + theme(legend.position = "none",axis.text.x = element_text(angle = 90)),
+          e + theme_minimal(base_size = 8) + theme(legend.position = "none",axis.text.x = element_text(angle = 90)),
+          f + theme_minimal(base_size = 8) + theme(axis.text.x = element_text(angle = 90)),
+          nrow = 2, rel_widths = c(0.3, 0.3, 0.4)) 
 ```

### 2019-05-20 analysis/05_DESeq2_manipulate.Rmd
now with heatmaps

@@
 
-```{r pheatmap, eval = F}
+```{r pheatmap, eval = T}
 
 for (eachdds in mydds){
+
+  dds <- eachdds
 
@@
   sigDEGs <- cbind(DEGs,padjmin)
-
   sigDEGs <- sigDEGs %>% arrange(padjmin)

### 2019-05-20 analysis/03_DESeq2_characterization.Rmd
rename

--- 2019-05-20 analysis/05_DESeq2_manipulate.Rmd
@@
 ---
-title: "05_DESeq_manipulation"
+title: "02_DESeq_characterization"
 output: md_document
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/manipulation/', cache = TRUE)
+knitr::opts_chunk$set(fig.path = '../figures/characterizatio/', cache = TRUE)
 ```
@@
 # import "colData" which contains sample information and "countData" which contains read counts
-m.colData <- read.csv("../metadata/00_colData_manipluation.csv", header = T, row.names = 1)
-m.countData <- read.csv("../results/00_countData_manipluation.csv", header = T, row.names = 1)
+m.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
+m.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
 geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
@@
 # set levels
+levels(m.colData$treatment)
 m.colData$treatment <- factor(m.colData$treatment, levels = 
-                              c("m.inc.d3",  "m.inc.d8",
-                                "m.inc.d9", "m.inc.d17",
-                                "prolong", "extend", "m.n2"))
+                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
 
 m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))
-
-m.colData$outcome <- ifelse(grepl("d3|d9|d17", m.colData$treatment), "end inc", 
-                     ifelse(grepl("d8|n2", m.colData$treatment),"end hatch",
-                     ifelse(grepl("prolong", m.colData$treatment),"prolong inc",
-                     ifelse(grepl("extend", m.colData$treatment),"delay hatch", NA))))
-
-m.colData$outcome <- factor(m.colData$outcome, levels = 
-                              c("end inc",  "end hatch",
-                                "prolong inc", "delay hatch"))
-summary(m.colData[c(7,3,4,5,8,9)])
+summary(m.colData[c(7,3,4,5,8)])
 ```
@@
 #create list of groups for deseq contrasts
-group1 <- c("m.inc.d3","m.inc.d8", "m.inc.d9", "m.inc.d17", "prolong","extend", "m.n2" ) 
+group1 <- c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9")
 group2 <- group1
 
-a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
-b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
-c <- plottotalDEGs(dds.female_gonad, "female gonad")
-d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
-e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
-f <- plottotalDEGs(dds.male_gondad, "male gonad")
+a <- plottotalDEGschar(dds.female_hypothalamus, "female hypothalamus")
+b <- plottotalDEGschar(dds.female_pituitary, "female pituitary")
+c <- plottotalDEGschar(dds.female_gonad, "female gonad")
+d <- plottotalDEGschar(dds.male_hypothalamus, "male hypothalamus")
+e <- plottotalDEGschar(dds.male_pituitary, "male pituitary")
+f <- plottotalDEGschar(dds.male_gondad, "male gonad")
 

### 2019-05-20 analysis/05_DESeq2_manipulate.Rmd
now with real subtitles

--- 2019-05-20 analysis/05_DESeq2_manipulate.Rmd
@@
 
-m.colData$outcome <- factor(m.colData$outcome, levels = 
-                              c("end inc",  "end hatch",
-                                "prolong inc", "delay hatch"))
-summary(m.colData[c(7,3,4,5,8,9)])
+summary(m.colData[c(7,3,4,5,8)])
 ```
@@
 ```{r pca}
-
-#mydds <- list(ddsmale_hypothalamus) # for testing only
-mydds <- list(dds.female_hypothalamus, dds.female_pituitary, dds.female_gonad,
-             dds.male_hypothalamus, dds.male_pituitary, dds.male_gondad)
-
-for (eachdds in mydds){
-
-  vsd <- vst(eachdds, blind=FALSE) # variance stabilized
-
-  # create the dataframe using my function pcadataframe
-  pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
-  percentVar <- round(100 * attr(pcadata, "percentVar"))
-  print(percentVar)
-
-  pca1 <- ggplot(pcadata, aes(treatment, PC1,color = treatment)) + 
-    geom_boxplot() +
-    ylab(paste0("PC1: ", percentVar[1],"% variance")) +
-    xlab(NULL) +
-    theme_cowplot(font_size = 8, line_size = 0.25) +
-    labs(subtitle = "eachgroup") +
-    theme(legend.position = "none")
-
-  pca2 <- ggplot(pcadata, aes(treatment, PC2,color = treatment)) + 
-    geom_boxplot() +
-    ylab(paste0("PC2: ", percentVar[2],"% variance")) +
-    xlab(NULL) +
-    theme_cowplot(font_size = 8, line_size = 0.25) +
-    labs(subtitle = "eachgroup") +
-    theme(legend.position = "none")
-
-  mypca <- plot_grid(pca1, pca2)
-  plot(mypca)
-
-  print(summary(aov(PC1 ~ treatment, data=pcadata)))
-  print(TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment"))
-
-  print(summary(aov(PC2 ~ treatment, data=pcadata))) 
-  print(summary(aov(PC3 ~ treatment, data=pcadata))) 
-  print(summary(aov(PC4 ~ treatment, data=pcadata)))
-
-}
+plotPCAs(dds.female_hypothalamus, "female hypothalamus")
+plotPCAs(dds.female_pituitary, "female pituitary")
+plotPCAs(dds.female_gonad, "female gonad")
+plotPCAs(dds.male_hypothalamus, "male hypothalamus")
+plotPCAs(dds.male_pituitary, "male pituitary")
+plotPCAs(dds.male_gondad, "male gonad")
 ```
@@
 ```{r pheatmap, eval = F}
-
-for (eachdds in mydds){
-
-  vsd <- vst(eachdds, blind=FALSE) # variance stabilized
-
-  # make dataframe counts
-  DEGs <- assay(vsd)
-  DEGs <- as.data.frame(DEGs)
-
-  a <- levels(m.colData$treatment)
-  b <- a
-
-  for (i in a){
-    for (j in b){
-      if (i != j) {
-        print(i)
-        results <- returnpadj(i,j)
-        DEGs <- cbind(DEGs,results)
-      }
-    }
-     b <- b[-1]  # drop 1st element of second string to not recalculate DEGs
-  }
-
-  DEGsmatrix <- DEGs
-  DEGsmatrix <- as.matrix(DEGs)
-  padjmin <- rowMins(DEGsmatrix, na.rm = T) 
-  padjmin <- as.data.frame(padjmin)
-
-  sigDEGs <- cbind(DEGs,padjmin)
-
-  sigDEGs <- sigDEGs %>% arrange(padjmin)
-  sigDEGs <- head(sigDEGs,500)
-  sigDEGs <- as.data.frame(sigDEGs)
-  rownames(sigDEGs) <- sigDEGs$rownames
-  drop.cols <-colnames(sigDEGs[,grep("padj|pval|pmin|rownames", colnames(sigDEGs))])
-  sigDEGs <- sigDEGs %>% dplyr::select(-one_of(drop.cols))
-  sigDEGs <- as.matrix(sigDEGs)
-  sigDEGs <- sigDEGs - rowMeans(sigDEGs)
-
-  paletteLength <- 30
-  myBreaks <- c(seq(min(sigDEGs), 0, length.out=ceiling(paletteLength/2) + 1), 
-                seq(max(sigDEGs)/paletteLength, max(sigDEGs), length.out=floor(paletteLength/2)))
-
-  anndf <- m.colData %>% dplyr::select(treatment)
-  rownames(anndf) <- m.colData$V1
-
-  sigDEGs <- as.matrix(sigDEGs) 
-  pheatmap(sigDEGs, show_rownames = F, show_colnames = F,
-         color = viridis(30),
-         breaks=myBreaks,
-         annotation_col=anndf,
-         main = "eachgroup")
-
-  pheatmap(sigDEGs, kmeans_k = 5,
-         show_rownames = F, show_colnames = F,
-         color = viridis(30),
-         breaks=myBreaks,
-         annotation_col=anndf,
-         main = "eachgroup")
-}
-
+makepheatmap(dds.female_hypothalamus, "female hypothalamus")
+makepheatmap(dds.female_pituitary, "female pituitary")
+makepheatmap(dds.female_gonad, "female gonad")
+makepheatmap(dds.male_hypothalamus, "male hypothalamus")
+makepheatmap(dds.male_pituitary, "male pituitary")
+makepheatmap(dds.male_gondad, "male gonad")
 ```
@@
 ```{r candidates}
-for (eachdds in mydds){
-
-  vsd <- vst(eachdds, blind=FALSE) # variance stabilized
-
-  # make dataframe counts
-  DEGs <- assay(vsd)
-  DEGs <- as.data.frame(DEGs)
-
-names(geneinfo)
-names(geneinfo)[4] <- "rownames"
-DEGs$rownames <- row.names(DEGs)
-
-# make dataframe with geneids and names and counts
-# how to gather: https://tidyr.tidyverse.org/reference/gather.html
-
-candidates <- full_join(geneinfo, DEGs)
-drop.cols <-colnames(candidates[,grep("padj|pval|pmin", colnames(candidates))])
-candidates <- candidates %>% dplyr::select(-one_of(drop.cols))
-candidates <- candidates %>%
-  filter(Name %in% c( #"JUN", "JUND", "EGR",  "AVP", "AVPR1A", "AVPR1B", "AVPR2", "OXT",  
-                     "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
-                     "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
-                     "PGR", "PRL", "PRLR", "VIP", "VIPR1")) 
-row.names(candidates) <- candidates$Name
-candidates <- candidates %>% select(-row.names, -rownames, -Name, -geneid)
-candidates <- candidates %>% drop_na()
-candidates <- as.data.frame(t(candidates))
-candidates$RNAseqID <- rownames(candidates)
-candidates <- candidates %>% gather(gene, value, -RNAseqID)  %>% 
-  filter(RNAseqID != "gene")
-candidates$value <- as.numeric(candidates$value)
-candidates$V1  <- candidates$RNAseqID
-
-candidatecounts <- left_join(candidates, m.colData)
-candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
-candidatecounts$gene <- as.factor(candidatecounts$gene)
-
-p1 <- candidatecounts %>%
-  filter(gene %in% c( "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
-                      "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
-                      "PGR", "PRL", "PRLR", "VIP", "VIPR1")) %>%
-  ggplot(aes(x = treatment, y = value, fill = treatment)) +
-  geom_boxplot() +
-  facet_wrap(~gene, scales = "free") +
-  theme_bw(base_size = 8) +
-  theme(axis.text.x = element_blank(),
-        legend.position = "bottom") +
-  labs(x = NULL) +
-  guides(fill= guide_legend(nrow=1))
-
-plot(p1)
-}
+plotcandidates(dds.female_hypothalamus, "female hypothalamus")
+plotcandidates(dds.female_pituitary, "female pituitary")
+plotcandidates(dds.female_gonad, "female gonad")
+plotcandidates(dds.male_hypothalamus, "male hypothalamus")
+plotcandidates(dds.male_pituitary, "male pituitary")
+plotcandidates(dds.male_gondad, "male gonad")
 ```

### 2019-05-20 R/functions.R

@@ -35,8 +35,6 @@ subsetDESeq <- function(eachgroup){
 # numDEGs('m.inc.d3', 'm.inc.d9')
 
 
-
-
 numDEGs <- function(dds, group1, group2){
   res <- results(dds, contrast = c("treatment", group1, group2), independentFiltering = T)
   sumpadj <- sum(res$padj < 0.01, na.rm = TRUE)
@@ -163,5 +161,151 @@ pcadataframe <- function (object, intgroup = "condition", ntop = 500, returnData
     return(d)
   }
 }
+
+# plot pcs 
+# eg. plotPCAs(dds.female_hypothalamus, "female hypothalamus")
+plotPCAs <- function(mydds, mysubtitle){
+  
+  vsd <- vst(mydds, blind=FALSE) # variance stabilized 
+  
+  # create the dataframe using my function pcadataframe
+  pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
+  percentVar <- round(100 * attr(pcadata, "percentVar"))
+  print(percentVar)
+  
+  print(summary(aov(PC1 ~ treatment, data=pcadata)))
+  print(TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment"))
+  
+  print(summary(aov(PC2 ~ treatment, data=pcadata))) 
+  print(summary(aov(PC3 ~ treatment, data=pcadata))) 
+  print(summary(aov(PC4 ~ treatment, data=pcadata))) 
+  
+  pca1 <- ggplot(pcadata, aes(treatment, PC1,color = treatment)) + 
+    geom_boxplot() +
+    ylab(paste0("PC1: ", percentVar[1],"% variance")) +
+    xlab(NULL) +
+    theme_cowplot(font_size = 8, line_size = 0.25) +
+    labs(subtitle = mysubtitle) +
+    theme(legend.position = "none")
+  
+  
+  pca2 <- ggplot(pcadata, aes(treatment, PC2,color = treatment)) + 
+    geom_boxplot() +
+    ylab(paste0("PC2: ", percentVar[2],"% variance")) +
+    xlab(NULL) +
+    theme_cowplot(font_size = 8, line_size = 0.25) +
+    labs(subtitle = " ") +
+    theme(legend.position = "none")
+  
+  mypca <- plot_grid(pca1, pca2)
+  return(mypca)
+}
+
+## plot candidate genes 
[…]

### 2019-05-20 analysis/05_DESeq2_manipulate.Rmd
better printout

@@
 
-plot_grid(a,b,c,d,e,f, nrow = 2) 
-```
-
-```{r totalDEGs2}
-plot_grid(a + theme_minimal(base_size = 8) + theme(legend.position = "none", axis.text.x = element_text(angle = 90)),
-          b + theme_minimal(base_size = 8) + theme(legend.position = "none", axis.text.x = element_text(angle = 90)),
-          c + theme_minimal(base_size = 8) + theme(axis.text.x = element_text(angle = 90)),
-          d + theme_minimal(base_size = 8) + theme(legend.position = "none",axis.text.x = element_text(angle = 90)),
-          e + theme_minimal(base_size = 8) + theme(legend.position = "none",axis.text.x = element_text(angle = 90)),
-          f + theme_minimal(base_size = 8) + theme(axis.text.x = element_text(angle = 90)),
+plot_grid(a + theme(legend.position = "none"),
+          b + theme(legend.position = "none"),
+          c,
+          d + theme(legend.position = "none"),
+          e + theme(legend.position = "none"),
+          f,
           nrow = 2, rel_widths = c(0.3, 0.3, 0.4)) 

### 2019-05-20 R/functions.R

@@ -82,11 +82,14 @@ plottotalDEGs <- function(dds, mysubtitle){
                        breaks = c(0, 1000, 2000, 3000)) + 
     xlab(" ") + ylab("Timepoint") +
     labs(fill = "# of DEGs",
-         subtitle = mysubtitle)
+         subtitle = mysubtitle) +
+    theme_minimal(base_size = 8) + 
+    theme(axis.text.x = element_text(angle = 90))
   
   return(allcontrasts)
 }
 
+# plot DEGs just for characterization study
 plottotalDEGschar <- function(dds, mysubtitle){
   
   a <- group1
@@ -269,7 +272,6 @@ makepheatmap <- function(mydds, mysubtitle){
   for (i in a){
     for (j in b){
       if (i != j) {
-        print(i)
         results <- returnpadj(i,j)
         DEGs <- cbind(DEGs,results)
       }

### 2019-05-20 R/functions.R
new pcs

@@ -123,7 +123,9 @@ plottotalDEGschar <- function(dds, mysubtitle){
                        breaks = c(0, 1000, 2000, 3000)) + 
     xlab(" ") + ylab("Timepoint") +
     labs(fill = "# of DEGs",
-         subtitle = mysubtitle)
+         subtitle = mysubtitle) +
+    theme_minimal(base_size = 8) + 
+    theme(axis.text.x = element_text(angle = 90))
   
   return(allcontrasts)
 }
@@ -200,7 +202,16 @@ plotPCAs <- function(mydds, mysubtitle){
     labs(subtitle = " ") +
     theme(legend.position = "none")
   
-  mypca <- plot_grid(pca1, pca2)
+  pca12 <- ggplot(pcadata, aes(PC1, PC2,color = treatment)) + 
+    geom_point() +
+    stat_ellipse() +
+    ylab(paste0("PC2: ", percentVar[2],"% variance")) +
+    xlab(paste0("PC1: ", percentVar[1],"% variance")) +
+    theme_cowplot(font_size = 8, line_size = 0.25) +
+    labs(subtitle = mysubtitle) +
+    theme(legend.position = "none")
+  
+  mypca <- plot_grid(pca12, pca1, pca2)
   return(mypca)
 }
 

### 2019-05-20 analysis/03_DESeq2_characterization.Rmd
mostly good, not heatmap

@@
 
-## Calculate and plot principal components
+## Calculate and plot principal component
 
 ```{r pca}
-
-#mydds <- list(ddsmale_hypothalamus) # for testing only
-mydds <- list(dds.female_hypothalamus, dds.female_pituitary, dds.female_gonad,
-             dds.male_hypothalamus, dds.male_pituitary, dds.male_gondad)
-
-for (eachdds in mydds){
-
-  vsd <- vst(eachdds, blind=FALSE) # variance stabilized
-
-  # create the dataframe using my function pcadataframe
-  pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
-  percentVar <- round(100 * attr(pcadata, "percentVar"))
-  print(percentVar)
-
-  pca1 <- ggplot(pcadata, aes(treatment, PC1,color = treatment)) + 
-    geom_boxplot() +
-    ylab(paste0("PC1: ", percentVar[1],"% variance")) +
-    xlab(NULL) +
-    theme_cowplot(font_size = 8, line_size = 0.25) +
-    labs(subtitle = "eachgroup") +
-    theme(legend.position = "none")
-
-  pca2 <- ggplot(pcadata, aes(treatment, PC2,color = treatment)) + 
-    geom_boxplot() +
-    ylab(paste0("PC2: ", percentVar[2],"% variance")) +
-    xlab(NULL) +
-    theme_cowplot(font_size = 8, line_size = 0.25) +
-    labs(subtitle = "eachgroup") +
-    theme(legend.position = "none")
-
-  mypca <- plot_grid(pca1, pca2)
-  plot(mypca)
-
-  print(summary(aov(PC1 ~ treatment, data=pcadata)))
-  print(TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment"))
-
-  print(summary(aov(PC2 ~ treatment, data=pcadata))) 
-  print(summary(aov(PC3 ~ treatment, data=pcadata))) 
-  print(summary(aov(PC4 ~ treatment, data=pcadata)))
-
-}
+plotPCAs(dds.female_hypothalamus, "female hypothalamus")
+plotPCAs(dds.female_pituitary, "female pituitary")      
+plotPCAs(dds.female_gonad, "female gonad")
+plotPCAs(dds.male_hypothalamus, "male hypothalamus")
+plotPCAs(dds.male_pituitary, "male pituitary")
+plotPCAs(dds.male_gondad, "male gonad")
 ```
@@
 ```{r pheatmap, eval = T}
-
-for (eachdds in mydds){
-
-  dds <- eachdds
-
-  vsd <- vst(eachdds, blind=FALSE) # variance stabilized
-
-  # make dataframe counts
-  DEGs <- assay(vsd)
-  DEGs <- as.data.frame(DEGs)
-
-  a <- levels(m.colData$treatment)
-  b <- a
-
-  for (i in a){
-    for (j in b){
-      if (i != j) {
-        print(i)
-        results <- returnpadj(i,j)
-        DEGs <- cbind(DEGs,results)
-      }
-    }
-     b <- b[-1]  # drop 1st element of second string to not recalculate DEGs
-  }
-
-  DEGsmatrix <- DEGs
-  DEGsmatrix <- as.matrix(DEGs)
-  padjmin <- rowMins(DEGsmatrix, na.rm = T) 
-  padjmin <- as.data.frame(padjmin)
-
-  sigDEGs <- cbind(DEGs,padjmin)
-  sigDEGs <- sigDEGs %>% arrange(padjmin)
-  sigDEGs <- head(sigDEGs,500)
-  sigDEGs <- as.data.frame(sigDEGs)
-  rownames(sigDEGs) <- sigDEGs$rownames
-  drop.cols <-colnames(sigDEGs[,grep("padj|pval|pmin|rownames", colnames(sigDEGs))])
-  sigDEGs <- sigDEGs %>% dplyr::select(-one_of(drop.cols))
-  sigDEGs <- as.matrix(sigDEGs)
-  sigDEGs <- sigDEGs - rowMeans(sigDEGs)
-
-  paletteLength <- 30
-  myBreaks <- c(seq(min(sigDEGs), 0, length.out=ceiling(paletteLength/2) + 1), 
-                seq(max(sigDEGs)/paletteLength, max(sigDEGs), length.out=floor(paletteLength/2)))
-
-  anndf <- m.colData %>% dplyr::select(treatment)
-  rownames(anndf) <- m.colData$V1
-
-  sigDEGs <- as.matrix(sigDEGs) 
-  pheatmap(sigDEGs, show_rownames = F, show_colnames = F,
-         color = viridis(30),
-         breaks=myBreaks,
-         annotation_col=anndf,
-         main = "eachgroup")
-
-  pheatmap(sigDEGs, kmeans_k = 5,
-         show_rownames = F, show_colnames = F,
-         color = viridis(30),
-         breaks=myBreaks,
-         annotation_col=anndf,
-         main = "eachgroup")
-}
-
+makepheatmap(dds.female_hypothalamus, "female hypothalamus")
+makepheatmap(dds.female_pituitary, "female pituitary")
+makepheatmap(dds.female_gonad, "female gonad")
+makepheatmap(dds.male_hypothalamus, "male hypothalamus")
+makepheatmap(dds.male_pituitary, "male pituitary")
+makepheatmap(dds.male_gondad, "male gonad")
 ```
+## plot candidate genes
 
 ```{r candidates}
-for (eachdds in mydds){
-
-  vsd <- vst(eachdds, blind=FALSE) # variance stabilized
-
-  # make dataframe counts
-  DEGs <- assay(vsd)
-  DEGs <- as.data.frame(DEGs)
-
-names(geneinfo)
-names(geneinfo)[4] <- "rownames"
-DEGs$rownames <- row.names(DEGs)
-
-# make dataframe with geneids and names and counts
-# how to gather: https://tidyr.tidyverse.org/reference/gather.html
-
-candidates <- full_join(geneinfo, DEGs)
-drop.cols <-colnames(candidates[,grep("padj|pval|pmin", colnames(candidates))])
-candidates <- candidates %>% dplyr::select(-one_of(drop.cols))
-candidates <- candidates %>%
-  filter(Name %in% c( #"JUN", "JUND", "EGR",  "AVP", "AVPR1A", "AVPR1B", "AVPR2", "OXT",  
-                     "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
-                     "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
-                     "PGR", "PRL", "PRLR", "VIP", "VIPR1")) 
-row.names(candidates) <- candidates$Name
-candidates <- candidates %>% select(-row.names, -rownames, -Name, -geneid)
-candidates <- candidates %>% drop_na()
-candidates <- as.data.frame(t(candidates))
-candidates$RNAseqID <- rownames(candidates)
-candidates <- candidates %>% gather(gene, value, -RNAseqID)  %>% 
-  filter(RNAseqID != "gene")
-candidates$value <- as.numeric(candidates$value)
-candidates$V1  <- candidates$RNAseqID
-
-candidatecounts <- left_join(candidates, m.colData)
-candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
-candidatecounts$gene <- as.factor(candidatecounts$gene)
-
-p1 <- candidatecounts %>%
-  filter(gene %in% c( "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
-                      "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
-                      "PGR", "PRL", "PRLR", "VIP", "VIPR1")) %>%
-  ggplot(aes(x = treatment, y = value, fill = treatment)) +
-  geom_boxplot() +
-  facet_wrap(~gene, scales = "free") +
-  theme_bw(base_size = 8) +
-  theme(axis.text.x = element_blank(),
-        legend.position = "bottom") +
-  labs(x = NULL) +
-  guides(fill= guide_legend(nrow=1))
-
-plot(p1)
-}
+plotcandidates(dds.female_hypothalamus, "female hypothalamus")
+plotcandidates(dds.female_pituitary, "female pituitary")
+plotcandidates(dds.female_gonad, "female gonad")
+plotcandidates(dds.male_hypothalamus, "male hypothalamus")
+plotcandidates(dds.male_pituitary, "male pituitary")
+plotcandidates(dds.male_gondad, "male gonad")
 ```

### 2019-05-21 analysis/04_DESeq2_sexes.Rmd
now with char and manipulation combined

---
title: "04_DESeq_sexes"
output: md_document
---

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}

library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)

# load custom functions  
source("../R/functions.R")

knitr::opts_chunk$set(fig.path = '../figures/sexes/', cache = TRUE)
```

## Starting with all the data

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
a.colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
a.countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

# set levels
levels(a.colData$treatment)

a.colData$treatment <- factor(a.colData$treatment, levels = 
                             c("control", "bldg", "lay",
                               "inc.d3", "m.inc.d3", 
                               "inc.d9", "m.inc.d8", "m.inc.d9",
                               "inc.d17", "m.inc.d17",
                               "hatch",  "m.n2"  ,
                               "n5", "prolong", "extend", "n9" ))

a.colData$sextissue <- as.factor(paste(a.colData$sex, a.colData$tissue, sep = "_"))

a.colData$lastday <- ifelse(grepl("m.inc.d3|m.inc.d9|m.inc.d17|m.n2", a.colData$treatment), "empty nest", 
                    ifelse(grepl("m.inc.d8|hatch|extend", a.colData$treatment),"nestlings hatch",      
                     ifelse(grepl("n5", a.colData$treatment),"nestlings early",
                     ifelse(grepl("n9", a.colData$treatment),"nestlings later",
                     ifelse(grepl("control", a.colData$treatment),"control",
                     ifelse(grepl("bldg", a.colData$treatment),"nest building",
                     ifelse(grepl("prolong", a.colData$treatment),"incubation delay",
                      ifelse(grepl("lay", a.colData$treatment),"incubation lay",
                      ifelse(grepl("inc.d3", a.colData$treatment),"incubation early",
                     ifelse(grepl("inc.d9", a.colData$treatment),"incubation middle",
                    ifelse(grepl("inc.d17", a.colData$treatment),"incubation later", NA)))))))))))

a.colData$penultimate <-  ifelse(grepl("extend", a.colData$treatment),"incubation delay",  
                          ifelse(grepl("m.n2", a.colData$treatment),"nestlings hatch",
                     ifelse(grepl("n5", a.colData$treatment),"nestlings early",
                     ifelse(grepl("n9", a.colData$treatment),"nestlings later",
                     ifelse(grepl("control", a.colData$treatment),"control",
                     ifelse(grepl("bldg|lay", a.colData$treatment),"nest building",
                     ifelse(grepl("prolong", a.colData$treatment),"incubation delay",
                      ifelse(grepl("inc.d3|m.inc.d3", a.colData$treatment),"incubation early",
                     ifelse(grepl("inc.d9|m.inc.d9|m.inc.d8", a.colData$treatment),"incubation middle",
                    ifelse(grepl("inc.d17|hatch", a.colData$treatment),"incubation later", NA))))))))))

a.colData$lastday <- factor(a.colData$lastday, levels =  c("control", "nest building", "incubation lay",
                               "incubation early",  
                               "incubation middle", 
                               "incubation later", 
                               "incubation delay",
                               "nestlings hatch",
                               "nestlings early", 
                               "nestlings later", 
                               "empty nest"))

a.colData$penultimate <- factor(a.colData$penultimate, levels =  c("control", "nest building",
                               "incubation early",  
                               "incubation middle", 
                               "incubation later", 
                               "incubation delay",
                               "nestlings hatch",
                               "nestlings early", 
                               "nestlings later"))

colorlastday <-  c("control" = "#636363", 
                   "nest building" = "#fb6a4a", 
                   "incubation lay" = "#c7e9c0",
                   "incubation early" = "#a1d99b",
                   "incubation middle" = "#74c476", 
                   "incubation later" = "#31a354",  
                   "incubation delay" = "#006d2c",
                   "nestlings hatch" = "#cbc9e2",   
                   "nestlings early" = "#9e9ac8", 
                   "nestlings later" = "#6a51a3",
                   "empty nest" = "#a50f15")

colorpenultimate <- c("control" = "#636363", 
                      "nest building" = "#fb6a4a", 
                      "incubation early" = "#a1d99b",
                   "incubation middle" = "#74c476", 
                   "incubation later" = "#31a354",  
                   "incubation delay" = "#006d2c",
                   "nestlings hatch" = "#cbc9e2",   
                   "nestlings early" = "#9e9ac8", 
                   "nestlings later" = "#6a51a3")

colorlastdaybad <-  c("control" = "#636363", 
                   "nest bldg" = "#fb6a4a", 
                   "lay" = "#c7e9c0",
                   "incubate begin" = "#a1d99b",
                   "incubate middle" = "#74c476", 
                   "incubate end" = "#31a354",  
                   "incubation late" = "#006d2c",
                   "nestling being" = "#cbc9e2",  
                   "nestling begin" = "#cbc9e2",   
                   "nestling middle" = "#9e9ac8", 
                   "nestling end" = "#6a51a3",
                   "empty nest" = "#a50f15")

colorpenultimatebad <- c("control" = "#636363", 
                      "nestbldg" = "#fb6a4a", 
                      "incubate begin" = "#a1d99b",
                   "incubate middle" = "#74c476", 
                   "incubate end" = "#31a354",  
                   "incubation late" = "#006d2c",
                   "nestling begin" = "#cbc9e2",   
                   "nestling middle" = "#9e9ac8", 
                   "nestling end" = "#6a51a3")

summary(a.colData[c(7,3,4,5,8,9, 10)])
```

## Run DESeq on all subsets of the data

```{r dds, cache = T}
dds.female_hypothalamus <- subsetDESeq(a.colData, a.countData, "female_hypothalamus")
dds.female_pituitary <- subsetDESeq(a.colData, a.countData, "female_pituitary" )
dds.female_gonad <- subsetDESeq(a.colData, a.countData, "female_gonad" )
dds.male_hypothalamus <- subsetDESeq(a.colData, a.countData, "male_hypothalamus" )
dds.male_pituitary <- subsetDESeq(a.colData, a.countData, "male_pituitary"  )
dds.male_gondad <- subsetDESeq(a.colData, a.countData, "male_gonad")
```

## Calculate and plot total DEGs

```{r totalDEGs, eval = F}

#create list of groups for deseq contrasts
group1 <- levels(a.colData$treatment)
group2 <- group1

a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
c <- plottotalDEGs(dds.female_gonad, "female gonad")
d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
f <- plottotalDEGs(dds.male_gondad, "male gonad")

plot_grid(a + theme(legend.position = "none"),
          b + theme(legend.position = "none"),
          c,
          d + theme(legend.position = "none"),
          e + theme(legend.position = "none"),
          f,
          nrow = 2, rel_widths = c(0.3, 0.3, 0.4)) 
```

## Calculate and plot principal components

```{r pca}
plotPCAs(dds.female_hypothalamus, "female hypothalamus")
plotPCAs(dds.female_pituitary, "female pituitary")      
plotPCAs(dds.female_gonad, "female gonad")
plotPCAs(dds.male_hypothalamus, "male hypothalamus")
plotPCAs(dds.male_pituitary, "male pituitary")
plotPCAs(dds.male_gondad, "male gonad")
```

## heamap with minimum pvalue

```{r pheatmap, eval = F}
makepheatmap(dds.female_hypothalamus, a.colData, "female hypothalamus")
makepheatmap(dds.female_pituitary, "female pituitary")
makepheatmap(dds.female_gonad, "female gonad")
makepheatmap(dds.male_hypothalamus, "male hypothalamus")
makepheatmap(dds.male_pituitary, "male pituitary")
makepheatmap(dds.male_gondad, "male gonad")        
```

## candidate genes

```{r candidates, eval =F}
plotcandidates(dds.female_hypothalamus, a.colData, "female hypothalamus")
plotcandidates(dds.female_pituitary, "female pituitary")
plotcandidates(dds.female_gonad, "female gonad")
plotcandidates(dds.male_hypothalamus, "male hypothalamus")
plotcandidates(dds.male_pituitary, "male pituitary")
plotcandidates(dds.male_gondad, "male gonad")
```

### 2019-05-22 R/functions.R
now with multiple box plots and long figure

@@ -162,60 +162,79 @@ plotPCAs <- function(dds, mysubtitle){
   vsd <- vst(dds, blind=FALSE) # variance stabilized 
   
   # create the dataframe using my function pcadataframe
-  pcadata <- pcadataframe(vsd, intgroup=c("lastday", "penultimate", "treatment", "study"), returnData=TRUE)
+  pcadata <- pcadataframe(vsd, intgroup=c("lastday", "penultimate", "xlabel", "study"), returnData=TRUE)
   percentVar <- round(100 * attr(pcadata, "percentVar"))
   print(percentVar)
   
-  print(summary(aov(PC1 ~ treatment, data=pcadata)))
-  #print(TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment"))
+  print(summary(aov(PC1 ~ xlabel, data=pcadata)))
+  #print(TukeyHSD(aov(PC1 ~ xlabel, data=pcadata), which = "xlabel"))
   
-  print(summary(aov(PC2 ~ lastday, data=pcadata))) 
-  print(summary(aov(PC3 ~ lastday, data=pcadata))) 
-  print(summary(aov(PC4 ~ lastday, data=pcadata))) 
+  print(summary(aov(PC2 ~ xlabel, data=pcadata))) 
+  print(summary(aov(PC3 ~ xlabel, data=pcadata))) 
+  print(summary(aov(PC4 ~ xlabel, data=pcadata))) 
   
   
-  pca1 <- ggplot(pcadata, aes(treatment, PC1, color = penultimate ,fill = lastday )) + 
+  pca1 <- ggplot(pcadata, aes(xlabel, PC1, color = penultimate ,fill = lastday )) + 
     geom_boxplot() +
     #geom_point() +
     ylab(paste0("PC1: ", percentVar[1],"% variance")) +
     xlab(NULL) +
-    theme_cowplot(font_size = 8, line_size = 0.25) +
+    theme_bw(base_size = 12) +
     labs(subtitle = mysubtitle) +
-    theme(axis.text.x = element_text(angle = 90)) +
-    scale_color_manual(values = colorpenultimatebad) +
-    scale_fill_manual(values = colorlastdaybad) +
+    theme(axis.text.x = element_text(angle = 90),
+          legend.text = element_text(size=10)) +
+    scale_color_manual(values = colorpenultimate) +
+    scale_fill_manual(values = colorlastday) +
     guides(color = guide_legend(order = 1), 
            fill = guide_legend(order = 2)) + 
[…]

### 2019-05-22 analysis/04_DESeq2_sexes.Rmd
now with multiple box plots and long figure

@@
 a.colData$lastday <- ifelse(grepl("m.inc.d3|m.inc.d9|m.inc.d17|m.n2", a.colData$treatment), "empty nest", 
-                    ifelse(grepl("m.inc.d8|hatch|extend", a.colData$treatment),"nestlings hatch",      
-                     ifelse(grepl("n5", a.colData$treatment),"nestlings early",
-                     ifelse(grepl("n9", a.colData$treatment),"nestlings later",
+                    ifelse(grepl("m.inc.d8|hatch|extend", a.colData$treatment),"chicks hatch",      
+                     ifelse(grepl("n5", a.colData$treatment),"chicks early",
+                     ifelse(grepl("n9", a.colData$treatment),"chicks later",
                      ifelse(grepl("control", a.colData$treatment),"control",
                      ifelse(grepl("bldg", a.colData$treatment),"nest building",
-                     ifelse(grepl("prolong", a.colData$treatment),"incubation delay",
-                      ifelse(grepl("lay", a.colData$treatment),"incubation lay",
-                      ifelse(grepl("inc.d3", a.colData$treatment),"incubation early",
-                     ifelse(grepl("inc.d9", a.colData$treatment),"incubation middle",
-                    ifelse(grepl("inc.d17", a.colData$treatment),"incubation later", NA)))))))))))
+                     ifelse(grepl("prolong", a.colData$treatment),"eggs delay",
+                      ifelse(grepl("lay", a.colData$treatment),"eggs lay",
+                      ifelse(grepl("inc.d3", a.colData$treatment),"eggs early",
+                     ifelse(grepl("inc.d9", a.colData$treatment),"eggs middle",
+                    ifelse(grepl("inc.d17", a.colData$treatment),"eggs later", NA)))))))))))
 
-a.colData$penultimate <-  ifelse(grepl("extend", a.colData$treatment),"incubation delay",  
-                          ifelse(grepl("m.n2", a.colData$treatment),"nestlings hatch",
-                     ifelse(grepl("n5", a.colData$treatment),"nestlings early",
-                     ifelse(grepl("n9", a.colData$treatment),"nestlings later",
+a.colData$penultimate <-  ifelse(grepl("extend", a.colData$treatment),"eggs delay",  
+                          ifelse(grepl("m.n2", a.colData$treatment),"chicks hatch",
+                     ifelse(grepl("n5", a.colData$treatment),"chicks early",
+                     ifelse(grepl("n9", a.colData$treatment),"chicks later",
                      ifelse(grepl("control", a.colData$treatment),"control",
                      ifelse(grepl("bldg|lay", a.colData$treatment),"nest building",
-                     ifelse(grepl("prolong", a.colData$treatment),"incubation delay",
-                      ifelse(grepl("inc.d3|m.inc.d3", a.colData$treatment),"incubation early",
-                     ifelse(grepl("inc.d9|m.inc.d9|m.inc.d8", a.colData$treatment),"incubation middle",
-                    ifelse(grepl("inc.d17|hatch", a.colData$treatment),"incubation later", NA))))))))))
+                     ifelse(grepl("prolong", a.colData$treatment),"eggs delay",
+                      ifelse(grepl("inc.d3|m.inc.d3", a.colData$treatment),"eggs early",
+                     ifelse(grepl("inc.d9|m.inc.d9|m.inc.d8", a.colData$treatment),"eggs middle",
+                    ifelse(grepl("inc.d17|hatch", a.colData$treatment),"eggs later", NA))))))))))
 
-a.colData$lastday <- factor(a.colData$lastday, levels =  c("control", "nest building", "incubation lay",
-                               "incubation early",  
-                               "incubation middle", 
-                               "incubation later", 
-                               "incubation delay",
-                               "nestlings hatch",
-                               "nestlings early", 
-                               "nestlings later", 
+a.colData$xlabel <- a.colData$treatment
+levels(a.colData$xlabel) <-  c("control", "nest building", "egg lay",
+                               "eggs early", "eggs early remove", 
+                               "eggs mid", "eggs mid hatch", "eggs mid remove",
+                               "eggs end", "eggs end remove",
+                               "chicks hatch",  "chicks hatch remove"  ,
+                               "chicks mid", "eggs delay", "eggs delay hatch", "chicks end")
+
+a.colData$lastday <- factor(a.colData$lastday, levels =  c("control", "nest building", "eggs lay",
+                               "eggs early",  
+                               "eggs middle", 
+                               "eggs later", 
+                               "eggs delay",
+                               "chicks hatch",
+                               "chicks early", 
+                               "chicks later", 
                                "empty nest"))
@@
 a.colData$penultimate <- factor(a.colData$penultimate, levels =  c("control", "nest building",
-                               "incubation early",  
-                               "incubation middle", 
-                               "incubation later", 
-                               "incubation delay",
-                               "nestlings hatch",
-                               "nestlings early", 
-                               "nestlings later"))
+                               "eggs early",  
+                               "eggs middle", 
+                               "eggs later", 
+                               "eggs delay",
+                               "chicks hatch",
+                               "chicks early", 
+                               "chicks later"))
 
@@
                    "nest building" = "#fb6a4a", 
-                   "incubation lay" = "#c7e9c0",
-                   "incubation early" = "#a1d99b",
-                   "incubation middle" = "#74c476", 
-                   "incubation later" = "#31a354",  
-                   "incubation delay" = "#006d2c",
-                   "nestlings hatch" = "#cbc9e2",   
-                   "nestlings early" = "#9e9ac8", 
-                   "nestlings later" = "#6a51a3",
+                   "eggs lay" = "#c7e9c0",
+                   "eggs early" = "#a1d99b",
+                   "eggs middle" = "#74c476", 
+                   "eggs later" = "#31a354",  
+                   "eggs delay" = "#006d2c",
+                   "chicks hatch" = "#cbc9e2",   
+                   "chicks early" = "#9e9ac8", 
+                   "chicks later" = "#6a51a3",
                    "empty nest" = "#a50f15")
@@
                       "nest building" = "#fb6a4a", 
-                      "incubation early" = "#a1d99b",
-                   "incubation middle" = "#74c476", 
-                   "incubation later" = "#31a354",  
-                   "incubation delay" = "#006d2c",
-                   "nestlings hatch" = "#cbc9e2",   
-                   "nestlings early" = "#9e9ac8", 
-                   "nestlings later" = "#6a51a3")
+                      "eggs early" = "#a1d99b",
+                   "eggs middle" = "#74c476", 
+                   "eggs later" = "#31a354",  
+                   "eggs delay" = "#006d2c",
+                   "chicks hatch" = "#cbc9e2",   
+                   "chicks early" = "#9e9ac8", 
+                   "chicks later" = "#6a51a3")
 
@@
                    "incubate end" = "#31a354",  
-                   "incubation late" = "#006d2c",
+                   "eggs late" = "#006d2c",
                    "nestling being" = "#cbc9e2",  
@@
                    "incubate end" = "#31a354",  
-                   "incubation late" = "#006d2c",
+                   "eggs late" = "#006d2c",
                    "nestling begin" = "#cbc9e2",   
@@
 
-summary(a.colData[c(7,3,4,5,8,9, 10)])
+summary(a.colData[c(7,3,4,5,8,9, 10,11)])
+
 ```
@@
 
-```{r pca}
-plotPCAs(dds.female_hypothalamus, "female hypothalamus")
-plotPCAs(dds.female_pituitary, "female pituitary")      
-plotPCAs(dds.female_gonad, "female gonad")
-plotPCAs(dds.male_hypothalamus, "male hypothalamus")
-plotPCAs(dds.male_pituitary, "male pituitary")
-plotPCAs(dds.male_gondad, "male gonad")
+```{r pcadf}
+p1 <- plotPCAs(dds.female_hypothalamus, "female hypothalamus")
+p2 <- plotPCAs(dds.female_pituitary, "female pituitary")      
+p3 <- plotPCAs(dds.female_gonad, "female gonad")
+p4 <- plotPCAs(dds.male_hypothalamus, "male hypothalamus")
+p5 <- plotPCAs(dds.male_pituitary, "male pituitary")
+p6 <- plotPCAs(dds.male_gondad, "male gonad")
+p1
+p2
+p3
+p4
+p5
+p6
+```
+
+```{r pca, fig.dim = c(12, 10)}
+mylegend <- get_legend(p1)
+
+allPC1s <- plot_grid(p1 +  theme(legend.position = "none",
+               axis.title.x=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank()), 
+    p4 + theme(legend.position = "none",
+               axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()), 
+    p2 + theme(legend.position = "none",
+               axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()),
+    p5 + theme(legend.position = "none",
+               axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank()),
+    p3 + theme(legend.position = "none", axis.ticks.x=element_blank()),
+    p6 + theme(legend.position = "none", axis.ticks.x=element_blank()),
+    nrow = 3, rel_heights = c(0.3, 0.3, 0.4))
+
+pc1grid <- plot_grid(allPC1s, mylegend, ncol = 2, rel_widths = c(1.0, 0.15))
+pc1grid
+
+pdf("../figures/sexes/pc1grid.pdf", width = 12, height = 10)
+plot(pc1grid)
+dev.off()
+
 ```

### 2019-05-23 analysis/04_DESeq2_sexes.Rmd
done with pca for now

@@
                    "eggs late" = "#006d2c",
-                   "nestling begin" = "#cbc9e2",   
-                   "nestling middle" = "#9e9ac8", 
-                   "nestling end" = "#6a51a3")
+                   "nestling begin" = "#9e9ac8",   
+                   "nestling middle" = "#6a51a3", 
+                   "nestling end" = "#3f007d")
 
@@
 p6
+
 ```
 
-```{r pca, fig.dim = c(12, 10)}
+```{r pca, fig.dim = c(12, 10), eval=T}
 mylegend <- get_legend(p1)
 
-allPC1s <- plot_grid(p1 +  theme(legend.position = "none",
-               axis.title.x=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank()), 
-    p4 + theme(legend.position = "none",
-               axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()), 
-    p2 + theme(legend.position = "none",
-               axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()),
-    p5 + theme(legend.position = "none",
-               axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank()),
-    p3 + theme(legend.position = "none", axis.ticks.x=element_blank()),
-    p6 + theme(legend.position = "none", axis.ticks.x=element_blank()),
+theme_noaxislabels <- theme(legend.position = "none", axis.title.x=element_blank(),axis.text.x=element_blank())
+
+allPC1s <- plot_grid(p1 + theme_noaxislabels, 
+                     p1 + theme_noaxislabels, 
+                     p1 + theme_noaxislabels,
+                     p1 + theme_noaxislabels,
+                     p1 + theme(legend.position = "none"),
+                     p1 + theme(legend.position = "none"),
                      nrow = 3, rel_heights = c(0.3, 0.3, 0.4))
 
-pc1grid <- plot_grid(allPC1s, mylegend, ncol = 2, rel_widths = c(1.0, 0.15))
-pc1grid
+pc1 <- plot_grid(mylegend, allPC1s, nrow = 2, rel_heights = c(0.15, 1.0))
+pc1
 
-pdf("../figures/sexes/pc1grid.pdf", width = 12, height = 10)
-plot(pc1grid)
+pdf("../figures/sexes/pca-1.pdf", width = 12, height = 10)
+plot(pc1)
 dev.off()
-
 ```

### 2019-05-23 R/functions.R

@@ -174,32 +174,40 @@ plotPCAs <- function(dds, mysubtitle){
   print(summary(aov(PC4 ~ xlabel, data=pcadata))) 
   
   
-  pca1 <- ggplot(pcadata, aes(xlabel, PC1, color = penultimate ,fill = lastday )) + 
-    geom_boxplot() +
+  pca1 <- ggplot(pcadata, aes(xlabel, PC1, fill = lastday)) + 
+    geom_boxplot(aes(color = penultimate, alpha = 0.9)) +
     #geom_point() +
-    ylab(paste0("PC1: ", percentVar[1],"% variance")) +
-    xlab(NULL) +
     theme_bw(base_size = 12) +
+    ylab(paste0("PC1: ", percentVar[1],"% variance")) +
+    xlab("Parental stages, with increasing time ->") +
     labs(subtitle = mysubtitle) +
     theme(axis.text.x = element_text(angle = 90),
-          legend.text = element_text(size=10)) +
+          legend.text = element_text(size=10),
+          legend.position = "bottom",
+          legend.box = "vertical",
+          legend.margin = margin(0,0,0,0, "cm"))  +
     scale_color_manual(values = colorpenultimate) +
-    scale_fill_manual(values = colorlastday) +
-    guides(color = guide_legend(order = 1), 
-           fill = guide_legend(order = 2)) + 
-      facet_wrap(~study, scales = "free_x")
+    scale_fill_manual(values = colorlastday) + 
+    guides(alpha=FALSE) + 
+    facet_wrap(~study, scales = "free_x") +
+      guides(color = guide_legend(order = 1, title = "pentultimate day", 
+                                  direction = "horizontal", 
+                                  nrow=1, label.position = "bottom"),
+             fill = guide_legend(order = 2,  title = "last day", 
+                                  direction = "horizontal", 
+                                  nrow=1, label.position = "bottom")) 
+    
 
-  
-
-  pca2 <- ggplot(pcadata, aes(xlabel, PC2, color = penultimate ,fill = lastday)) + 
-    geom_boxplot() +
+  pca2 <- ggplot(pcadata, aes(xlabel, PC2, color = penultimate)) + 
+    geom_boxplot(aes(fill = lastday, alpha = 0.9)) +
     #geom_point() +
     theme_bw(base_size = 12) +
     ylab(paste0("PC2: ", percentVar[2],"% variance")) +
     xlab(NULL) +
     labs(subtitle = mysubtitle) +
[…]

### 2019-05-24 analysis/04_DESeq2_sexes.Rmd
new update to pca and total DEGs

@@
 
-# load custom functions  
-source("../R/functions.R")
+source("../R/functions.R")  # load custom functions 
+source("../R/themes.R")  # load custom themes and color palletes
 
@@
 a.colData$xlabel <- a.colData$treatment
-levels(a.colData$xlabel) <-  c("control", "nest building", "egg lay",
-                               "eggs early", "eggs early remove", 
-                               "eggs mid", "eggs mid hatch", "eggs mid remove",
-                               "eggs end", "eggs end remove",
-                               "chicks hatch",  "chicks hatch remove"  ,
-                               "chicks mid", "eggs delay", "eggs delay hatch", "chicks end")
+
+levels(a.colData$xlabel) <-  c("control", "nest.building", "egg.lay",
+                               "eggs.early", "eggs.early.remove", 
+                               "eggs.mid", "eggs.mid.hatch", "eggs.mid.remove",
+                               "eggs.end", "eggs.end.remove",
+                               "chicks.hatch",  "chicks.hatch.remove"  ,
+                               "chicks.mid", "eggs.delay", "eggs.delay.hatch", "chicks.end")
 
 a.colData$lastday <- factor(a.colData$lastday, levels =  c("control", "nest building", "eggs lay",
-                               "eggs early",  
-                               "eggs middle", 
-                               "eggs later", 
-                               "eggs delay",
-                               "chicks hatch",
-                               "chicks early", 
-                               "chicks later", 
-                               "empty nest"))
+                               "eggs early", "eggs middle",  "eggs later","eggs delay",
+                               "chicks hatch", "chicks early",  "chicks later", "empty nest"))
 
 a.colData$penultimate <- factor(a.colData$penultimate, levels =  c("control", "nest building",
-                               "eggs early",  
-                               "eggs middle", 
-                               "eggs later", 
-                               "eggs delay",
-                               "chicks hatch",
-                               "chicks early", 
-                               "chicks later"))
-
-colorlastday <-  c("control" = "#636363", 
-                   "nest building" = "#fb6a4a", 
-                   "eggs lay" = "#c7e9c0",
-                   "eggs early" = "#a1d99b",
-                   "eggs middle" = "#74c476", 
-                   "eggs later" = "#31a354",  
-                   "eggs delay" = "#006d2c",
-                   "chicks hatch" = "#cbc9e2",   
-                   "chicks early" = "#9e9ac8", 
-                   "chicks later" = "#6a51a3",
-                   "empty nest" = "#a50f15")
-
-colorpenultimate <- c("control" = "#636363", 
-                      "nest building" = "#fb6a4a", 
-                      "eggs early" = "#a1d99b",
-                   "eggs middle" = "#74c476", 
-                   "eggs later" = "#31a354",  
-                   "eggs delay" = "#006d2c",
-                   "chicks hatch" = "#cbc9e2",   
-                   "chicks early" = "#9e9ac8", 
-                   "chicks later" = "#6a51a3")
-
-colorlastdaybad <-  c("control" = "#636363", 
-                   "nest bldg" = "#fb6a4a", 
-                   "lay" = "#c7e9c0",
-                   "incubate begin" = "#a1d99b",
-                   "incubate middle" = "#74c476", 
-                   "incubate end" = "#31a354",  
-                   "eggs late" = "#006d2c",
-                   "nestling being" = "#cbc9e2",  
-                   "nestling begin" = "#cbc9e2",   
-                   "nestling middle" = "#9e9ac8", 
-                   "nestling end" = "#6a51a3",
-                   "empty nest" = "#a50f15")
-
-colorpenultimatebad <- c("control" = "#636363", 
-                      "nestbldg" = "#fb6a4a", 
-                      "incubate begin" = "#a1d99b",
-                   "incubate middle" = "#74c476", 
-                   "incubate end" = "#31a354",  
-                   "eggs late" = "#006d2c",
-                   "nestling begin" = "#9e9ac8",   
-                   "nestling middle" = "#6a51a3", 
-                   "nestling end" = "#3f007d")
+                               "eggs early",  "eggs middle",  "eggs later",  "eggs delay",
+                               "chicks hatch", "chicks early",  "chicks later"))
 
 summary(a.colData[c(7,3,4,5,8,9, 10,11)])
-
 ```
@@
 
-```{r totalDEGs, eval = F}
-
+```{r totalDEGs, eval = T}
 #create list of groups for deseq contrasts
@@
 f <- plottotalDEGs(dds.male_gondad, "male gonad")
+```
 
+```{r plottotalDEGs, eval = T}
 plot_grid(a + theme(legend.position = "none"),
           b + theme(legend.position = "none"),
-          c,
+          c + theme(legend.position = "none"),
           d + theme(legend.position = "none"),
           e + theme(legend.position = "none"),
-          f,
-          nrow = 2, rel_widths = c(0.3, 0.3, 0.4)) 
+          f + theme(legend.position = "none"),
+          nrow = 2) 
 ```
@@
 
-```{r pcadf}
+```{r pcadf, eval=F}
 p1 <- plotPCAs(dds.female_hypothalamus, "female hypothalamus")
@@
 
-```{r pca, fig.dim = c(12, 10), eval=T}
+```{r pca, fig.dim = c(12, 10), eval=F}
 mylegend <- get_legend(p1)

### 2019-05-24 analysis/04_DESeq2_sexes.Rmd
new update to pca and total DEGs

@@
                      p1 + theme_noaxislabels,
-                     p1 + theme(legend.position = "none"),
-                     p1 + theme(legend.position = "none"),
+                     p1 + theme_noaxislabels,
+                     p1 + theme_noaxislabels),
                      nrow = 3, rel_heights = c(0.3, 0.3, 0.4))

### 2019-05-24 analysis/04_DESeq2_sexes.Rmd
now with separate functions for table and plot yea

@@
 
-```{r totalDEGs, eval = T}
-#create list of groups for deseq contrasts
-group1 <- levels(a.colData$treatment)
-group2 <- group1
+```{r returntotalDEGs, eval = T}
+DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
+DEGs.female_pituitary <- returntotalDEGs(dds.female_pituitary)
+DEGs.female_gonad <- returntotalDEGs(dds.female_gonad)
+DEGs.male_hypothalamus <- returntotalDEGs(dds.male_hypothalamus)
+DEGss.male_pituitary <- returntotalDEGs(dds.male_pituitary)
+DEGs.male_gondad <- returntotalDEGs(dds.male_gondad)
+```
 
-a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
-b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
-c <- plottotalDEGs(dds.female_gonad, "female gonad")
-d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
-e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
-f <- plottotalDEGs(dds.male_gondad, "male gonad")
+```{r plotDEGs, eval = T}
+a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus")
+b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary")
+c <- plottotalDEGs(DEGs.female_gonad, "female gonad")
+d <- plottotalDEGs(DEGs.male_hypothalamus, "male hypothalamus")
+e <- plottotalDEGs(DEGss.male_pituitary, "male pituitary")
+f <- plottotalDEGs(DEGs.male_gondad, "male gonad")
 ```

### 2019-05-24 analysis/04_DESeq2_sexes.Rmd
setup for more contrasts

--- 2019-05-24 analysis/04_DESeq2_sexes.Rmd
@@
 
-```{r totalDEGs, eval = T}
-#create list of groups for deseq contrasts
-group1 <- levels(a.colData$treatment)
-group2 <- group1
+```{r returntotalDEGs, eval = T}
+#DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
+DEGs.female_pituitary <- returntotalDEGs(dds.female_pituitary)
+#DEGs.female_gonad <- returntotalDEGs(dds.female_gonad)
+#DEGs.male_hypothalamus <- returntotalDEGs(dds.male_hypothalamus)
+DEGss.male_pituitary <- returntotalDEGs(dds.male_pituitary)
+#DEGs.male_gondad <- returntotalDEGs(dds.male_gondad)
+```
 
-a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
-b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
-c <- plottotalDEGs(dds.female_gonad, "female gonad")
-d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
-e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
-f <- plottotalDEGs(dds.male_gondad, "male gonad")
+```{r plotDEGs, eval = T}
+#a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus")
+b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary")
+#c <- plottotalDEGs(DEGs.female_gonad, "female gonad")
+#d <- plottotalDEGs(DEGs.male_hypothalamus, "male hypothalamus")
+e <- plottotalDEGs(DEGss.male_pituitary, "male pituitary")
+#f <- plottotalDEGs(DEGs.male_gondad, "male gonad")
 ```
@@
 ```{r plottotalDEGs, eval = T}
-plot_grid(a + theme(legend.position = "none"),
+plot_grid(#a + theme(legend.position = "none"),
           b + theme(legend.position = "none"),
-          c + theme(legend.position = "none"),
-          d + theme(legend.position = "none"),
+          #c + theme(legend.position = "none"),
+          #d + theme(legend.position = "none"),
           e + theme(legend.position = "none"),
-          f + theme(legend.position = "none"),
+          #f + theme(legend.position = "none"),
           nrow = 2) 

### 2019-05-25 analysis/04_DESeq2_sexes.Rmd
now with all pca violoin plots

@@
 
-```{r returntotalDEGs, eval = T}
-#DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
+```{r returntotalDEGs, eval = F}
+DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
 DEGs.female_pituitary <- returntotalDEGs(dds.female_pituitary)
-#DEGs.female_gonad <- returntotalDEGs(dds.female_gonad)
-#DEGs.male_hypothalamus <- returntotalDEGs(dds.male_hypothalamus)
-DEGss.male_pituitary <- returntotalDEGs(dds.male_pituitary)
-#DEGs.male_gondad <- returntotalDEGs(dds.male_gondad)
+DEGs.female_gonad <- returntotalDEGs(dds.female_gonad)
+DEGs.male_hypothalamus <- returntotalDEGs(dds.male_hypothalamus)
+DEGs.male_pituitary <- returntotalDEGs(dds.male_pituitary)
+DEGs.male_gondad <- returntotalDEGs(dds.male_gondad)
 ```
@@
 #d <- plottotalDEGs(DEGs.male_hypothalamus, "male hypothalamus")
-e <- plottotalDEGs(DEGss.male_pituitary, "male pituitary")
+e <- plottotalDEGs(DEGs.male_pituitary, "male pituitary")
 #f <- plottotalDEGs(DEGs.male_gondad, "male gonad")
@@
 
-```{r pcadf, eval=F}
-p1 <- plotPCAs(dds.female_hypothalamus, "female hypothalamus")
-p2 <- plotPCAs(dds.female_pituitary, "female pituitary")      
-p3 <- plotPCAs(dds.female_gonad, "female gonad")
-p4 <- plotPCAs(dds.male_hypothalamus, "male hypothalamus")
-p5 <- plotPCAs(dds.male_pituitary, "male pituitary")
-p6 <- plotPCAs(dds.male_gondad, "male gonad")
-p1
-p2
-p3
-p4
-p5
-p6
-
+```{r returnPCAs, eval=T}
+pcas.female_hypothalamus <- returnPCAs(dds.female_hypothalamus)
+pcas.female_pituitary <- returnPCAs(dds.female_pituitary)      
+pcas.female_gonad <- returnPCAs(dds.female_gonad)
+pcas.male_hypothalamus <- returnPCAs(dds.male_hypothalamus)
+pcas.male_pituitary <- returnPCAs(dds.male_pituitary)
+pcas.male_gondad <- returnPCAs(dds.male_gondad)
 ```
 
-```{r pca, fig.dim = c(12, 10), eval=F}
+```{r plotpca, fig.dim = c(12, 10), eval=T}
+p1 <- plotPCAs(pcas.female_hypothalamus, "female hypothalamus") 
+p2 <- plotPCAs(pcas.female_pituitary, "female pituitary")   
+p3 <- plotPCAs(pcas.female_gonad, "female gonad") 
+p4 <- plotPCAs(pcas.male_hypothalamus, "male hypothalamus") 
+p5 <- plotPCAs(pcas.male_pituitary, "male pituitary") 
+p6 <- plotPCAs(pcas.male_gondad, "male gonad") 
+p1 + theme_noaxislabels
+p2 + theme_noaxislabels
+p3 + theme_noaxislabels
+p4 + theme_noaxislabels
+p5 + theme_noaxislabels
+p6 + theme_noaxislabels
+```
+```{r temp, eval = F}
 mylegend <- get_legend(p1)
 
-theme_noaxislabels <- theme(legend.position = "none", axis.title.x=element_blank(),axis.text.x=element_blank())
-
 allPC1s <- plot_grid(p1 + theme_noaxislabels, 
-                     p1 + theme_noaxislabels, 
-                     p1 + theme_noaxislabels,
-                     p1 + theme_noaxislabels,
-                     p1 + theme_noaxislabels,
-                     p1 + theme_noaxislabels),
+                     p2 + theme_noaxislabels, 
+                     p3 + theme_noaxislabels,
+                     p4 + theme_noaxislabels,
+                     p5 + theme_noaxislabels,
+                     p6 + theme_noaxislabels,
                      nrow = 3, rel_heights = c(0.3, 0.3, 0.4))
 
-pc1 <- plot_grid(mylegend, allPC1s, nrow = 2, rel_heights = c(0.15, 1.0))
+pc1 <- plot_grid(allPC1s,mylegend,  nrow = 2, rel_heights = c(1.0, 0.15))
 pc1

### 2019-05-25 analysis/04_DESeq2_sexes.Rmd
more total degs

@@
 
-```{r returntotalDEGs, eval = F}
-DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
-DEGs.female_pituitary <- returntotalDEGs(dds.female_pituitary)
-DEGs.female_gonad <- returntotalDEGs(dds.female_gonad)
-DEGs.male_hypothalamus <- returntotalDEGs(dds.male_hypothalamus)
+```{r returntotalDEGs, eval = T}
+#DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
+#DEGs.female_pituitary <- returntotalDEGs(dds.female_pituitary)
+#DEGs.female_gonad <- returntotalDEGs(dds.female_gonad)
+#DEGs.male_hypothalamus <- returntotalDEGs(dds.male_hypothalamus)
 DEGs.male_pituitary <- returntotalDEGs(dds.male_pituitary)
-DEGs.male_gondad <- returntotalDEGs(dds.male_gondad)
+#DEGs.male_gondad <- returntotalDEGs(dds.male_gondad)
 ```
@@
 #a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus")
-b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary")
+#b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary")
 #c <- plottotalDEGs(DEGs.female_gonad, "female gonad")
@@
 plot_grid(#a + theme(legend.position = "none"),
-          b + theme(legend.position = "none"),
+          #b + theme(legend.position = "none"),
           #c + theme(legend.position = "none"),

### 2019-05-25 analysis/04_DESeq2_sexes.Rmd
pca violins

@@
 p6 + theme_noaxislabels
+
 ```
-```{r temp, eval = F}
+```{r pca, eval = T}
 mylegend <- get_legend(p1)
 
-allPC1s <- plot_grid(p1 + theme_noaxislabels, 
-                     p2 + theme_noaxislabels, 
-                     p3 + theme_noaxislabels,
-                     p4 + theme_noaxislabels,
-                     p5 + theme_noaxislabels,
-                     p6 + theme_noaxislabels,
-                     nrow = 3, rel_heights = c(0.3, 0.3, 0.4))
+allPC1s <- plot_grid(p1 + theme_noaxislabels + ylim(c(-40,40)), 
+                     p4 + theme_noaxislabels + ylim(c(-40,40)), 
+                     p2 + theme_noaxislabels + ylim(c(-40,40)),
+                     p5 + theme_noaxislabels + ylim(c(-40,40)),
+                     p3 + theme_nolegend + ylim(c(-40,40)) ,
+                     p6 + theme_nolegend + ylim(c(-40,40)) ,
+                     ncol = 2)
 
-pc1 <- plot_grid(allPC1s,mylegend,  nrow = 2, rel_heights = c(1.0, 0.15))
+pc1 <- plot_grid(allPC1s, mylegend,  nrow = 2, rel_heights = c(1.0, 0.15))
 pc1
@@
 dev.off()
+
 ```

### 2019-05-25 analysis/02_pca_all.Rmd
not quite

--- 2019-04-25 analysis/01_pca_all.Rmd
@@
 
-theme_rmh <- function(){ 
-    theme_bw(base_size=14) +
-        theme(
-            panel.grid.minor.x  = element_blank(),
-           panel.grid.minor.y  = element_blank(),
-            strip.background = element_rect(colour="white", fill="white"),
-              legend.position = "right",           
-           legend.margin=margin(t=-0.1, r=0, b=-0.1, l=-0.1, unit="cm"),
-           legend.key.size = unit(0.5, "cm"))
-}
-
-treatmentcolors <-  c("control" = "#bdbdbd", 
-                              "bldg" =  "#d95f02", 
-                              "lay"  = "#a1d99b", 
-                              "inc.d3" = "#74c476", 
-                              "inc.d9"  = "#41ab5d", 
-                              "inc.d17" = "#238b45", 
-                              "hatch"  = "#807dba", 
-                              "n5"  = "#6a51a3", 
-                               "n9" = "#4a1486")
+source("../R/themes.R")  # load custom themes and color palletes
 
@@
 ```{r rawdata}
-counts <- read.csv("../results/00_countData_characterization.csv", row.names = 1, header = T)
-
-samples <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, header = T)
+samples <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
+counts <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
+geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1, header = T)
 
 samples$treatment <- factor(samples$treatment, levels = 
-                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
-                                "inc.d17", "hatch", "n5", "n9"))
+                             c("control", "bldg", "lay",
+                               "inc.d3", "inc.d9", "inc.d17", 
+                               "hatch", "n5",  "n9",
+                               "m.inc.d3", "m.inc.d8","m.inc.d9",
+                               "m.inc.d17","m.n2" ,"prolong", "extend"))
 samples$tissue <- factor(samples$tissue, levels = 
-                              c("hypothalamus", "pituitary", "gonad"))
+                             c("hypothalamus", "pituitary", "gonad" ))
 
-geneinfo <- read.csv("../metadata//00_geneinfo.csv", row.names = 1, header = T)
+samples$lastday <- ifelse(grepl("m.inc.d3|m.inc.d9|m.inc.d17|m.n2", samples$treatment), "empty nest", 
+                    ifelse(grepl("m.inc.d8|hatch|extend", samples$treatment),"chicks hatch",      
+                     ifelse(grepl("n5", samples$treatment),"chicks early",
+                     ifelse(grepl("n9", samples$treatment),"chicks later",
+                     ifelse(grepl("control", samples$treatment),"control",
+                     ifelse(grepl("bldg", samples$treatment),"nest building",
+                     ifelse(grepl("prolong", samples$treatment),"eggs delay",
+                      ifelse(grepl("lay", samples$treatment),"eggs lay",
+                      ifelse(grepl("inc.d3", samples$treatment),"eggs early",
+                     ifelse(grepl("inc.d9", samples$treatment),"eggs middle",
+                    ifelse(grepl("inc.d17", samples$treatment),"eggs later", NA)))))))))))
+
+samples$penultimate <-  ifelse(grepl("extend", samples$treatment),"eggs delay",  
+                          ifelse(grepl("m.n2", samples$treatment),"chicks hatch",
+                     ifelse(grepl("n5", samples$treatment),"chicks early",
+                     ifelse(grepl("n9", samples$treatment),"chicks later",
+                     ifelse(grepl("control", samples$treatment),"control",
+                     ifelse(grepl("bldg|lay", samples$treatment),"nest building",
+                     ifelse(grepl("prolong", samples$treatment),"eggs delay",
+                      ifelse(grepl("inc.d3|m.inc.d3", samples$treatment),"eggs early",
+                     ifelse(grepl("inc.d9|m.inc.d9|m.inc.d8", samples$treatment),"eggs middle",
+                    ifelse(grepl("inc.d17|hatch", samples$treatment),"eggs later", NA))))))))))
+
+samples$xlabel <- samples$treatment
+
+levels(samples$xlabel) <-  c("control", "nest.building", "egg.lay",
+                               "eggs.early", "eggs.early.remove", 
+                               "eggs.mid", "eggs.mid.hatch", "eggs.mid.remove",
+                               "eggs.end", "eggs.end.remove",
+                               "chicks.hatch",  "chicks.hatch.remove"  ,
+                               "chicks.mid", "eggs.delay", "eggs.delay.hatch", "chicks.end")
+
+samples$lastday <- factor(samples$lastday, levels =  c("control", "nest building", "eggs lay",
+                               "eggs early", "eggs middle",  "eggs later","eggs delay",
+                               "chicks hatch", "chicks early",  "chicks later", "empty nest"))
+
+samples$penultimate <- factor(samples$penultimate, levels =  c("control", "nest building",
+                               "eggs early",  "eggs middle",  "eggs later",  "eggs delay",
+                               "chicks hatch", "chicks early",  "chicks later"))
+
 ```
@@
 pseudocounts <- read.csv("../results/01_pseudo.counts.csv", row.names = 1)
-head(pseudocounts)
+#head(pseudocounts)
 pca <- prcomp(t(pseudocounts))
@@
 
-## subsets of normalized data show by tissue and sex
-
-```{r tissuesex}
-plotlist = list()
-for (eachgroup in levels(samples$treatment)){
-
-  print(eachgroup)
-
-  colData <- samples %>%
-      dplyr::filter(treatment == eachgroup) %>%
-      droplevels()
-  row.names(colData) <- colData$V1
-
-  savecols <- as.character(colData$V1) 
-  savecols <- as.vector(savecols)
-
-  countData <- pseudocounts %>% dplyr::select(one_of(savecols))
-
-  # check that row and col lenghts are equal
-  print(ncol(countData) == nrow(colData))
-
-  pca <- prcomp(t(countData))
-
-  p <- autoplot(pca, data = colData, alpha = 'tissue', shape = "sex", color = "treatment")
-
-en <- p + labs(subtitle = eachgroup) +
-  theme_rmh() +
-    theme_rmh()  +
-  stat_ellipse(aes(linetype = sex, alpha = tissue, color = treatment)) +
-   theme_bw(base_size=8) +
-    theme(legend.position = "bottom") +
-  scale_color_manual(values = treatmentcolors) + 
-  guides(color = FALSE)
-
-legend <- get_legend(en)
-
-en <- en  + theme(legend.position = "none")
-
-plotlist[[eachgroup]] = en
-
-}
-
-main <- grid.arrange(grobs=plotlist,ncol=3)
-
-withlegend <- plot_grid( main, legend, rel_heights  = c(3, .3), ncol = 1)
-withlegend
-
-```
-
 ## sex and timepoint within a tissue
@@
 
-  countData <- pseudocounts %>% dplyr::select(one_of(savecols))
+  countData <- counts %>% dplyr::select(one_of(savecols))
 
@@
 
-  p <- autoplot(pca, data = colData, colour = 'treatment', shape = "sex")
+  p <- autoplot(pca, data = colData, fill = lastday, color = penultimate, shape = "tissue")
 
 en <- p + labs(subtitle = eachtissue) +
-  theme_rmh() +  guides(colour = guide_legend(order = 1), 
+  theme_rmh() +  guides(color = guide_legend(order = 1), 
               shape = guide_legend(order = 2)) +
-  stat_ellipse(aes(linetype = sex, colour = treatment)) +
   theme_bw(base_size=8) +
     theme(legend.position = "bottom") + 
-  scale_color_manual(values =  treatmentcolors)
+  scale_fill_manual(values = colorlastday) +
+  scale_color_manual(values = colorpenultimate)
 

### 2019-05-28 R/functions.R
new order

@@ -46,11 +46,11 @@ numDEGs <- function(dds, group1, group2){
 
 returntotalDEGs <- function(dds){
   
-  colData <- a.colData %>%
-    filter(treatment %in% c("control", "bldg", "n9",
-                            "inc.d17", "m.inc.d17",
-                            "hatch",  "m.n2")) %>%
-    droplevels()
+  colData <- a.colData # %>%
+   # filter(treatment %in% c("control", "bldg", "n9",
+   #                         "inc.d17", "m.inc.d17",
+   #                         "hatch",  "m.n2")) %>%
+   # droplevels()
   
   group1 <- levels(colData$treatment)
 
@@ -79,30 +79,35 @@ returntotalDEGs <- function(dds){
 
 plottotalDEGs <- function(myDEGS, mysubtitle){  
   totalDEGS <- myDEGS
-  totalDEGS$V1 <- factor(totalDEGS$V1, levels =  c("control", "bldg", 
-                                                 # "lay", "inc.d3", "m.inc.d3", 
-                                                 #  "inc.d9", "m.inc.d8", "m.inc.d9",
-                                                  "inc.d17", "m.inc.d17",
-                                                   "hatch",  "m.n2"  ,
-                                                 #  "n5", "prolong", "extend",
-                                                   "n9"))
-  totalDEGS$V2 <- factor(totalDEGS$V2, levels =  c("control", "bldg", 
-                                                 #  "lay", "inc.d3", "m.inc.d3", 
-                                                 #  "inc.d9", "m.inc.d8", "m.inc.d9",
-                                                   "inc.d17", "m.inc.d17",
-                                                   "hatch",  "m.n2"  ,
-                                                 #  "n5", "prolong", "extend",
-                                                   "n9"))
+  totalDEGS$V1 <- factor(totalDEGS$V1, levels =  c("control", "bldg", "lay",
+                                                   "inc.d3", "inc.d9", "inc.d17",
+                                                   "hatch", "n5", "n9",
+                                                   
+                                                   "m.inc.d3", "m.inc.d8", "m.inc.d9",
+                                                   "m.inc.d17",  "m.n2"  ,
[…]

### 2019-05-28 analysis/04_DESeq2_sexes.Rmd
new order

@@
                              c("control", "bldg", "lay",
-                               "inc.d3", "m.inc.d3", 
-                               "inc.d9", "m.inc.d8", "m.inc.d9",
-                               "inc.d17", "m.inc.d17",
-                               "hatch",  "m.n2"  ,
-                               "n5", "prolong", "extend", "n9" ))
+                               "inc.d3", "inc.d9", "inc.d17",
+                               "hatch", "n5", "n9",
+
+                               "m.inc.d3", "m.inc.d8", "m.inc.d9",
+                               "m.inc.d17",  "m.n2"  ,
+                                "prolong", "extend" ))
 
@@
 ```{r returntotalDEGs, eval = T}
-#DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
-#DEGs.female_pituitary <- returntotalDEGs(dds.female_pituitary)
-#DEGs.female_gonad <- returntotalDEGs(dds.female_gonad)
-#DEGs.male_hypothalamus <- returntotalDEGs(dds.male_hypothalamus)
+DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
+DEGs.female_pituitary <- returntotalDEGs(dds.female_pituitary)
+DEGs.female_gonad <- returntotalDEGs(dds.female_gonad)
+DEGs.male_hypothalamus <- returntotalDEGs(dds.male_hypothalamus)
 DEGs.male_pituitary <- returntotalDEGs(dds.male_pituitary)
-#DEGs.male_gondad <- returntotalDEGs(dds.male_gondad)
-```
+DEGs.male_gondad <- returntotalDEGs(dds.male_gondad)
 
-```{r plotDEGs, eval = T}
-#a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus")
-#b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary")
-#c <- plottotalDEGs(DEGs.female_gonad, "female gonad")
-#d <- plottotalDEGs(DEGs.male_hypothalamus, "male hypothalamus")
-e <- plottotalDEGs(DEGs.male_pituitary, "male pituitary")
-#f <- plottotalDEGs(DEGs.male_gondad, "male gonad")
 ```
@@
 ```{r plottotalDEGs, eval = T}
-plot_grid(#a + theme(legend.position = "none"),
-          #b + theme(legend.position = "none"),
-          #c + theme(legend.position = "none"),
-          #d + theme(legend.position = "none"),
-          e + theme(legend.position = "none"),
-          #f + theme(legend.position = "none"),
-          nrow = 2) 
+a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus")
+b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary")
+c <- plottotalDEGs(DEGs.female_gonad, "female gonad")
+d <- plottotalDEGs(DEGs.male_hypothalamus, "male hypothalamus")
+e <- plottotalDEGs(DEGs.male_pituitary, "male pituitary")
+f <- plottotalDEGs(DEGs.male_gondad, "male gonad")
+
+myDEGs <- plot_grid(a + theme_noaxislabels,
+          d + theme_noaxislabels,
+          b + theme_noaxislabels,
+          e + theme_noaxislabels,
+          c + theme_nolegend,
+          f + theme_nolegend,
+          nrow = 3) 
+myDEGs
+
+pdf("../figures/sexes/plottotalDEGs-1.pdf", width = 12, height = 10)
+plot(myDEGs)
+dev.off()
 ```
@@
 
-```{r plotpca, fig.dim = c(12, 10), eval=T}
+```{r plotpca, eval=T}
 p1 <- plotPCAs(pcas.female_hypothalamus, "female hypothalamus") 
@@
 
-allPC1s <- plot_grid(p1 + theme_noaxislabels + ylim(c(-40,40)), 
-                     p4 + theme_noaxislabels + ylim(c(-40,40)), 
-                     p2 + theme_noaxislabels + ylim(c(-40,40)),
-                     p5 + theme_noaxislabels + ylim(c(-40,40)),
-                     p3 + theme_nolegend + ylim(c(-40,40)) ,
-                     p6 + theme_nolegend + ylim(c(-40,40)) ,
+allPC1s <- plot_grid(p1 + theme_noaxislabels + ylim(c(-35,35)), 
+                     p4 + theme_noaxislabels + ylim(c(-35,35)), 
+                     p2 + theme_noaxislabels + ylim(c(-35,35)),
+                     p5 + theme_noaxislabels + ylim(c(-35,35)),
+                     p3 + theme_nolegend + ylim(c(-35,35)) ,
+                     p6 + theme_nolegend + ylim(c(-35,35)) ,
                      ncol = 2)
@@
 dev.off()
-
 ```

### 2019-05-29 analysis/01_limma_all.Rmd
rename
umm, two days old commits

@@
 ```{r}
+if (!requireNamespace("BiocManager", quietly = TRUE))
+    install.packages("BiocManager")
+
+BiocManager::install("Glimma")
+
 library(limma)
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/',cache=TRUE)
+knitr::opts_chunk$set(fig.path = '../figures/01_limma/',cache=TRUE)
 
@@
 # import "colData" which contains sample information and "countData" which contains read counts
-colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
-countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
+colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
+countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
 head(colData)
 geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+
 ```
@@
 plotMDS(parentalobject, cex = 0.5)
+```
+
+```{r newRplot}
+
+myMDSplot <- plotMDS(parentalobject, cex = 0.5)
+
+names(myMDSplot$x)
+names(myMDSplot$y)
+
+mergedMDS <- data.frame(myMDSplot$x, myMDSplot$y)
+
+head(mergedMDS)
+
+names(mergedMDS) <- c("x", "y", "par", "treatment", "rep", "tissue")
+mds_GC_final_merged <- ggplot(mergedMDS, aes(x, y)) + 
+  geom_point(aes(color = treatment, shape = rep)) + 
+  geom_rug(aes(colour = treatment)) + 
+  facet_grid(par ~ tissue) + 
+  labs(title = "ggplot of GC new ref samples") + 
+  theme(strip.text.x = element_text(size = 8))
+mds_GC_final_merged
 ```

### 2019-05-29 analysis/04_DESeq2_sexes.Rmd
umm, two days old commits

@@
 
-```{r plotpca, eval=T}
-p1 <- plotPCAs(pcas.female_hypothalamus, "female hypothalamus") 
-p2 <- plotPCAs(pcas.female_pituitary, "female pituitary")   
-p3 <- plotPCAs(pcas.female_gonad, "female gonad") 
-p4 <- plotPCAs(pcas.male_hypothalamus, "male hypothalamus") 
-p5 <- plotPCAs(pcas.male_pituitary, "male pituitary") 
-p6 <- plotPCAs(pcas.male_gondad, "male gonad") 
+```{r pca, eval=T}
+p1 <- plotPC1(pcas.female_hypothalamus, "female hypothalamus", "26% PC1 variance explained") 
+p2 <- plotPC1(pcas.female_pituitary, "female pituitary", "11% PC1 variance explained")   
+p3 <- plotPC1(pcas.female_gonad, "female gonad", "37% PC1 variance explained") 
+p4 <- plotPC1(pcas.male_hypothalamus, "male hypothalamus", "28% PC1 variance explained") 
+p5 <- plotPC1(pcas.male_pituitary, "male pituitary" , "11% PC1 variance explained") 
+p6 <- plotPC1(pcas.male_gondad, "male gonad", "14% PC1 variance explained") 
 p1 + theme_noaxislabels
@@
 
-```
-```{r pca, eval = T}
 mylegend <- get_legend(p1)
@@
 
-pdf("../figures/sexes/pca-1.pdf", width = 12, height = 10)
+pdf("../figures/sexes/pca-1.pdf", width = 12, height = 14)
 plot(pc1)
 dev.off()
+
+p11 <- plotPC1(pcas.female_hypothalamus, "female hypothalamus", "26% PC1 variance explained") 
+p12 <- plotPC1(pcas.female_pituitary, "female pituitary", "11% PC1 variance explained")   
+p13 <- plotPC1(pcas.female_gonad, "female gonad", "37% PC1 variance explained") 
+p14 <- plotPC1(pcas.male_hypothalamus, "male hypothalamus", "28% PC1 variance explained") 
+p15 <- plotPC1(pcas.male_pituitary, "male pituitary" , "11% PC1 variance explained") 
+p16 <- plotPC1(pcas.male_gondad, "male gonad", "14% PC1 variance explained") 
+p11 + theme_noaxislabels
+p12 + theme_noaxislabels
+p13 + theme_noaxislabels
+p14 + theme_noaxislabels
+p15 + theme_noaxislabels
+p16 + theme_noaxislabels
+
+allPC12 <- plot_grid(p1 + theme_noaxislabels + ylim(c(-35,35)), 
+                     p4 + theme_noaxislabels + ylim(c(-35,35)), 
+                     p2 + theme_noaxislabels + ylim(c(-35,35)),
+                     p5 + theme_noaxislabels + ylim(c(-35,35)),
+                     p3 + theme_nolegend + ylim(c(-35,35)) ,
+                     p6 + theme_nolegend + ylim(c(-35,35)) ,
+                     ncol = 2)
+
+pc12 <- plot_grid(allPC12, mylegend,  nrow = 2, rel_heights = c(1.0, 0.15))
+pc12
+
+pdf("../figures/sexes/pca-12.pdf", width = 12, height = 14)
+plot(pc12)
+dev.off()
+
 ```

### 2019-05-29 analysis/02_pca_all.Rmd
umm, two days old commits

@@
 
-p <- autoplot(pca, data = samples, alpha = 'tissue', shape = "sex")
+p <- autoplot(pca, data = samples, colour = 'tissue', shape = "sex")
 p
@@
 edgRpca12 <- p + labs(subtitle = "Normalized counts from edgeR") +
-  scale_alpha_discrete(name = "tissue", labels = c("hypothalamus", "pituitary", "gonad")) +
-  scale_shape_discrete(name = "sex", labels = c("female", "male")) +
-  #xlab(stringr::str_replace(p$labels$x, "PC", "Principle Component ")) +
-  #ylab(stringr::str_replace(p$labels$y, "PC", "Principle Component ")) +
-  theme_rmh() + guides(alpha = guide_legend(order = 1), 
+  scale_colour_discrete(name = "tissue", labels = c("hypothalamus", "pituitary", "gonad")) +
+  scale_shape_manual(name = "sex", labels = c("female", "male"), values = c(21,19)) +
+  theme_rmh() + guides(color = guide_legend(order = 1), 
               shape = guide_legend(order = 2)) +
@@
 ```
-
-## sex and timepoint within a tissue
-
-```{r timepointsex}
-plotlist = list()
-for (eachtissue in levels(samples$tissue)){
-
-  print(eachtissue)
-
-  colData <- samples %>%
-      dplyr::filter(tissue == eachtissue) %>%
-      droplevels()
-  row.names(colData) <- colData$V1
-
-  savecols <- as.character(colData$V1) 
-  savecols <- as.vector(savecols)
-
-  countData <- counts %>% dplyr::select(one_of(savecols))
-
-  # check that row and col lenghts are equal
-  print(ncol(countData) == nrow(colData))
-
-  pca <- prcomp(t(countData))
-
-  p <- autoplot(pca, data = colData, fill = lastday, color = penultimate, shape = "tissue")
-
-en <- p + labs(subtitle = eachtissue) +
-  theme_rmh() +  guides(color = guide_legend(order = 1), 
-              shape = guide_legend(order = 2)) +
-  theme_bw(base_size=8) +
-    theme(legend.position = "bottom") + 
-  scale_fill_manual(values = colorlastday) +
-  scale_color_manual(values = colorpenultimate)
-
-legend <- get_legend(en)
-
-en <- en  + theme(legend.position = "none")
-
-plotlist[[eachtissue]] = en
-
-#plot(en) 
-}
-
-main <- grid.arrange(grobs=plotlist,ncol=3)
-
-withlegend <- plot_grid( main, legend, rel_heights  = c(3, .3), ncol = 1)
-withlegend
-
-```

### 2019-05-29 R/functions.R

@@ -208,7 +208,7 @@ returnPCAs <- function(dds){
   return(pcadata)
 }
 
-plotPCAs <- function(pcadata, mysubtitle){ 
+plotPC1 <- function(pcadata, mysubtitle, myxlab){ 
   
   pcadata <- pcadata
   
@@ -218,7 +218,7 @@ plotPCAs <- function(pcadata, mysubtitle){
     geom_violin() +
     #geom_point() +
     theme_bw(base_size = 12) +
-    ylab(paste0("PC1")) +
+    ylab(myxlab) +
     xlab("Parental stages, with increasing days ->") +
     labs(subtitle = mysubtitle) +
     scale_fill_manual(values = colorlastday, 
@@ -232,8 +232,14 @@ plotPCAs <- function(pcadata, mysubtitle){
     facet_wrap(~study, scales = "free_x")  +
     guides(col = guide_legend(nrow = 1),
            fill = guide_legend(nrow = 1))
+  return(pca1)
+}
 
 
+plotPC2 <- function(pcadata, mysubtitle){ 
+    
+    pcadata <- pcadata
+    
   pca2 <- ggplot(pcadata, aes(xlabel, PC2, color = penultimate, fill = lastday)) + 
     geom_violin() +
     #geom_point() +
@@ -252,14 +258,18 @@ plotPCAs <- function(pcadata, mysubtitle){
     facet_wrap(~study, scales = "free_x")  +
     guides(col = guide_legend(nrow = 1),
            fill = guide_legend(nrow = 1))
+}  
+ 
+plotPC12 <- function(pcadata, mysubtitle){ 
   
+  pcadata <- pcadata
   
-  pca12 <- ggplot(pcadata, aes(PC1, PC2, color = penultimate ,fill = lastday)) + 
+  pca12 <- ggplot(pcadata, aes(PC3, PC2, color = penultimate ,fill = lastday)) + 
     geom_point(pch=21, size = 3) +
     #stat_ellipse() +
     theme_bw(base_size = 12) +
     ylab(paste0("PC2")) +
-    xlab(paste0("PC1")) +
+    xlab(paste0("PC3")) +
     labs(subtitle = mysubtitle) +
     theme(legend.position = "bottom") +
     scale_color_manual(values = colorpenultimate) +
@@ -268,18 +278,11 @@ plotPCAs <- function(pcadata, mysubtitle){
            fill = guide_legend(order = 2, ncol=2)) +
[…]

### 2019-05-30 analysis/04_DESeq2_sexes.Rmd
now with group ave heatmaps

@@
 ```
+
+## new heat maps
+
+```{r correlationheatmaps, eval = T}
+
+plotcorrelationheatmaps <- function(mydds, mycoldata, mysubtitle){
+  dds <- mydds
+  vsd <- vst(dds, blind=FALSE) # variance stabilized
+
+  colnames(vsd) = mycoldata$treatment # set col names to group name
+
+  vsdm <- assay(vsd) # create matrix
+
+  vsdmmean <-sapply(unique(colnames(vsdm)), function(i)
+  rowMeans(vsdm[,colnames(vsdm) == i]))
+
+  myannotations <- mycoldata %>% 
+    dplyr::distinct(lastday, penultimate, treatment) 
+  rownames(myannotations) <- myannotations$treatment
+  myannotations$treatment <- NULL
+
+  mycor <- cor(vsdmmean)
+  str(mycor)
+
+  paletteLength <- 10
+  myBreaks = c(0.91,0.92,0.93,0.94,0.95, 0.96, 0.97,0.98, 0.99, 1.0)
+
+  myBreaks <-seq(0.95, 1.0, length.out = 10)
+  myBreaks
+
+  pheatmap(cor(vsdmmean),
+          annotation_row = myannotations,
+          annotation_col = myannotations,
+          annotation_colors = myannotationscolors,
+          annotation_names_row = F,
+          main= mysubtitle,
+          color = inferno(10),
+          show_rowname= F, show_colnames = F,
+          breaks = myBreaks
+          )
+}
+
+colData.female_hypothalamus <- subsetcolData(a.colData,  "female_hypothalamus")
+colData.female_pituitary <- subsetcolData(a.colData,  "female_pituitary" )
+colData.female_gonad <- subsetcolData(a.colData,  "female_gonad" )
+colData.male_hypothalamus <- subsetcolData(a.colData,  "male_hypothalamus" )
+colData.male_pituitary <- subsetcolData(a.colData,  "male_pituitary"  )
+colData.male_gondad <- subsetcolData(a.colData,  "male_gonad")
+
+plotcorrelationheatmaps(dds.female_hypothalamus, colData.female_hypothalamus ,"female hypothalamus: hatch, nest building, lay and day-before-hatch cluster")
+plotcorrelationheatmaps(dds.female_pituitary, colData.female_pituitary, "female pituitary: clustering suggests an internal clock effect")
+plotcorrelationheatmaps(dds.female_gonad, colData.female_gonad,  "female gonad: highly correlated across incubation and early nestling care")
+plotcorrelationheatmaps(dds.male_hypothalamus, colData.male_hypothalamus, "male hypothalamus: 100% correlated across most parental care stages")
+plotcorrelationheatmaps(dds.male_pituitary, colData.male_pituitary, "male pituitary: non-parental males are most different")
+plotcorrelationheatmaps(dds.male_gondad, colData.male_gondad, "male gonad: non-parental males are most different")
+```

### 2019-05-30 R/functions.R
now with group ave heatmaps

@@ -1,3 +1,14 @@
+# subset col for heatmap
+subsetcolData <- function(colData, eachgroup){
+  
+  # subset to look within one tissue in one sex
+  colData <- colData %>%
+    dplyr::filter(sextissue == eachgroup) %>%
+    droplevels()
+  row.names(colData) <- colData$V1
+  return(colData)
+}
+
 # run DESeq on subset of data
 # e.g. dds <- subsetDESeq("male_hypothalamus")
 subsetDESeq <- function(colData, countData, eachgroup){

### 2019-05-30 R/themes.R

@@ -26,7 +26,35 @@ colorpenultimate <- c("control" = "#77787A",
                       "chicks early" = "#9993C3", 
                       "chicks later" = "#7874B0")
 
-theme_noaxislabels <- theme(legend.position = "none", axis.title.x=element_blank(), axis.ticks = element_blank(), axis.text.x=element_blank()) 
+
+myannotationscolors <- list(lastday = c("control" = "#77787A", 
+                                        "nest building" = "#DA733E", 
+                                        "eggs lay" = "#559FD9",
+                                        "eggs early" = "#9AD0C3",
+                                        "eggs middle" = "#68C2A8", 
+                                        "eggs later" = "#37A589",  
+                                        "eggs delay" = "#19977D",
+                                        "chicks hatch" = "#BAB5D2",   
+                                        "chicks early" = "#9993C3", 
+                                        "chicks later" = "#7874B0",
+                                        "empty nest" = "#754E27"),
+                       penultimate = c("control" = "#77787A", 
+                                       "nest building" = "#DA733E", 
+                                       "eggs lay" = "#559FD9",
+                                       "eggs early" = "#9AD0C3",
+                                       "eggs middle" = "#68C2A8", 
+                                       "eggs later" = "#37A589",  
+                                       "eggs delay" = "#19977D",
+                                       "chicks hatch" = "#BAB5D2",   
+                                       "chicks early" = "#9993C3", 
+                                       "chicks later" = "#7874B0"))
+
+
+# themes for whole plots
+
+theme_noaxislabels <- theme(legend.position = "none", 
+                            axis.title.x=element_blank(), axis.ticks = element_blank(), axis.text.x=element_blank()) 
+
 theme_nolegend <- theme(legend.position = "none",  axis.text.x=element_blank(), axis.ticks = element_blank()) 
 
 theme_rmh <- function(){ 

### 2019-05-30 analysis/04_DESeq2_sexes.Rmd
new legends and stuff

--- 2019-05-28 analysis/04_DESeq2_sexes.Rmd
@@
 ```{r plottotalDEGs, eval = T}
-a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus")
-b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary")
-c <- plottotalDEGs(DEGs.female_gonad, "female gonad")
-d <- plottotalDEGs(DEGs.male_hypothalamus, "male hypothalamus")
-e <- plottotalDEGs(DEGs.male_pituitary, "male pituitary")
-f <- plottotalDEGs(DEGs.male_gondad, "male gonad")
+a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus: hatch is most unique parental stage")
+b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary: hatch and day before are most unique")
+c <- plottotalDEGs(DEGs.female_gonad, "female gonad: variable across all parental stages")
+d <- plottotalDEGs(DEGs.male_hypothalamus, "male hypothalamus: many parental contrasts yield 0 DEGs")
+e <- plottotalDEGs(DEGs.male_pituitary, "male pituitary: peaks at hatch")
+f <- plottotalDEGs(DEGs.male_gondad, "male gonad: most parental contrasts yield 0 DEGs")
 
 myDEGs <- plot_grid(a + theme_noaxislabels,
-          d + theme_noaxislabels,
+          d + theme_siggenes,
           b + theme_noaxislabels,
           e + theme_noaxislabels,
-          c + theme_nolegend,
-          f + theme_nolegend,
+          c + theme_noaxislabels,
+          f + theme_noaxislabels,
           nrow = 3) 
@@
 
-```{r plotpca, eval=T}
-p1 <- plotPCAs(pcas.female_hypothalamus, "female hypothalamus") 
-p2 <- plotPCAs(pcas.female_pituitary, "female pituitary")   
-p3 <- plotPCAs(pcas.female_gonad, "female gonad") 
-p4 <- plotPCAs(pcas.male_hypothalamus, "male hypothalamus") 
-p5 <- plotPCAs(pcas.male_pituitary, "male pituitary") 
-p6 <- plotPCAs(pcas.male_gondad, "male gonad") 
-p1 + theme_noaxislabels
-p2 + theme_noaxislabels
-p3 + theme_noaxislabels
-p4 + theme_noaxislabels
-p5 + theme_noaxislabels
-p6 + theme_noaxislabels
+```{r pca, eval=T}
+p1 <- plotPC1(pcas.female_hypothalamus, "female hypothalamus", "26% PC1 variance explained") 
+p2 <- plotPC1(pcas.female_pituitary, "female pituitary", "11% PC1 variance explained")   
+p3 <- plotPC1(pcas.female_gonad, "female gonad", "37% PC1 variance explained") 
+p4 <- plotPC1(pcas.male_hypothalamus, "male hypothalamus", "28% PC1 variance explained") 
+p5 <- plotPC1(pcas.male_pituitary, "male pituitary" , "11% PC1 variance explained") 
+p6 <- plotPC1(pcas.male_gondad, "male gonad", "14% PC1 variance explained")
 
-```
-```{r pca, eval = T}
 mylegend <- get_legend(p1)
 
-allPC1s <- plot_grid(p1 + theme_noaxislabels + ylim(c(-35,35)), 
-                     p4 + theme_noaxislabels + ylim(c(-35,35)), 
-                     p2 + theme_noaxislabels + ylim(c(-35,35)),
-                     p5 + theme_noaxislabels + ylim(c(-35,35)),
+allPC1s <- plot_grid(p1 + theme_nolegend + ylim(c(-35,35)), 
+                     p4 + theme_nolegend + ylim(c(-35,35)), 
+                     p2 + theme_nolegend + ylim(c(-35,35)),
+                     p5 + theme_nolegend + ylim(c(-35,35)),
                      p3 + theme_nolegend + ylim(c(-35,35)) ,
@@
 
-pdf("../figures/sexes/pca-1.pdf", width = 12, height = 10)
+pdf("../figures/sexes/pca-1.pdf", width = 12, height = 14)
 plot(pc1)
 dev.off()
+
+p11 <- plotPC2(pcas.female_hypothalamus, "female hypothalamus", "9% PC2 variance explained") 
+p12 <- plotPC2(pcas.female_pituitary, "female pituitary", "8% PC2 variance explained")   
+p13 <- plotPC2(pcas.female_gonad, "female gonad", "12 PC2 variance explained") 
+p14 <- plotPC2(pcas.male_hypothalamus, "male hypothalamus", "9% PC2 variance explained") 
+p15 <- plotPC2(pcas.male_pituitary, "male pituitary" , "9% PC2 variance explained") 
+p16 <- plotPC2(pcas.male_gondad, "male gonad", "10% PC2 variance explained")
+
+allPC2 <- plot_grid(p11 + theme_nolegend + ylim(c(-35,35)), 
+                     p14 + theme_nolegend + ylim(c(-35,35)), 
+                     p12 + theme_nolegend + ylim(c(-35,35)),
+                     p15 + theme_nolegend + ylim(c(-35,35)),
+                     p13 + theme_nolegend + ylim(c(-35,35)) ,
+                     p16 + theme_nolegend + ylim(c(-35,35)) ,
+                     ncol = 2)
+
+pc2 <- plot_grid(allPC2, mylegend,  nrow = 2, rel_heights = c(1.0, 0.15))
+pc2
+
+pdf("../figures/sexes/pca-2.pdf", width = 12, height = 14)
+plot(pc2)
+dev.off()
+
 ```
@@
 ```
+
+## new heat maps
+
+```{r correlationheatmaps, eval = T}
+colData.female_hypothalamus <- subsetcolData(a.colData,  "female_hypothalamus")
+colData.female_pituitary <- subsetcolData(a.colData,  "female_pituitary" )
+colData.female_gonad <- subsetcolData(a.colData,  "female_gonad" )
+colData.male_hypothalamus <- subsetcolData(a.colData,  "male_hypothalamus" )
+colData.male_pituitary <- subsetcolData(a.colData,  "male_pituitary"  )
+colData.male_gondad <- subsetcolData(a.colData,  "male_gonad")
+
+plotcorrelationheatmaps(dds.female_hypothalamus, colData.female_hypothalamus ,"female hypothalamus: hatch, nest bldg., lay and day-before-hatch cluster")
+plotcorrelationheatmaps(dds.female_pituitary, colData.female_pituitary, "female pituitary: clustering suggests an internal clock effect")
+plotcorrelationheatmaps(dds.female_gonad, colData.female_gonad,  "female gonad: highly correlated across incubation and early nestling care")
+plotcorrelationheatmaps(dds.male_hypothalamus, colData.male_hypothalamus, "male hypothalamus: 100% correlated across most parental care stages")
+plotcorrelationheatmaps(dds.male_pituitary, colData.male_pituitary, "male pituitary: non-parental males are most different")
+plotcorrelationheatmaps(dds.male_gondad, colData.male_gondad, "male gonad: non-parental males are most different")
+```

### 2019-06-16 docs/main.tex
docs

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Jacobs Landscape Poster
% LaTeX Template
% Version 1.0 (29/03/13)
%
% Created by:
% Computational Physics and Biophysics Group, Jacobs University
% https://teamwork.jacobs-university.de:8443/confluence/display/CoPandBiG/LaTeX+Poster
% 
% Further modified by:
% Nathaniel Johnston (nathaniel@njohnston.ca)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final]{beamer}

\usefonttheme{serif} % default family is serif

\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel} % for accented letters

\usepackage[none]{hyphenat} % stop the hyphenation

\usepackage{wrapfig} % for wrapping text around figures

\usepackage[scale=1.24]{beamerposter} % Use the beamerposter package for laying out the poster

\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamercolor{block title}{fg=black,bg=white} % Colors of the block titles
\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks
% Many more colors are available for use in beamerthemeconfposter.sty

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{48in} % A0 width: 46.8in
\setlength{\paperheight}{36in} % A0 height: 33.1in
\setlength{\sepwid}{0.024\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.22\paperwidth} % Width of one column
\setlength{\twocolwid}{0.464\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

\usepackage{graphicx}  % Required for including images

\usepackage{booktabs} % Top and bottom rules for tables

%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------

\title{New strategies for multi-factoral and temporal behavioral analyses} % Poster title

\author{Rayna M. Harris*\textsuperscript{1},
Suzanne H. Austin*\textsuperscript{2},
Andrew Lang\textsuperscript{3},
Matthew MacManes\textsuperscript{3},
Rebecca M. Calisi\textsuperscript{1}
} % Author(s)

\institute{*co-first authors, 1. University of California, Davis, 2. Oregon State University, 3. University of New Hampshire } % Institution(s)

%----------------------------------------------------------------------------------------

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\begin{columns}[t] % The whole poster consists of three major columns, the second of which is split into two columns twice - the [t] option aligns each column's content to the top

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The first column

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\vspace{-0.5em}

\begin{block}{The problem}

Many differential genes expression profiling tools are designed to analyze simplistic experimental designs comparing treatment and control. However, experimental designs in behavioral neuroendocrinology often require many treatment groups to hone in on the causes and consequences of animal behavior.

%\vspace{1em}

\end{block}

\begin{block}{\textit{Case study:} How and why do biological systems change over the course of parental care?}

\begin{figure}
\includegraphics[width=0.9\linewidth]{DoveParentsRNAseq_approach-3.png}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{DoveParentsRNAseq_design.png}
\end{figure}

Data and analyses are available at \url{https://github.com/macmanes-lab/DoveParentsRNAseq}. For more information visit \url{http://calisilab.ucdavis.edu/}.

\begin{figure}
\includegraphics[width=0.55\linewidth]{b3.jpg}
\end{figure}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of the first column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\twocolwid} % Begin a column which is two columns wide (column 2)

\vspace{-0.5em}

\begin{block}{Goal: Minimize hand-typing comparison, maximize reproducibility.}

\end{block}

\vspace{-2em}

%----------------------------------------------------------------------------------------

\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column again

\begin{column}{\onecolwid} % The first column within column 2 (column 2.1)

\vspace{-0.5em}

\begin{block}{{\normalsize New R functions to automate DESeq2}}

\begin{itemize}
  \item \begin{lstlisting} subsetcolData <- function(colData, eachgroup){} \end{lstlisting}
  \item \begin{lstlisting} returntotalDEGs <- function(dds){} \end{lstlisting}
  \item \begin{lstlisting} returnpadj <- function(group1, group2){} \end{lstlisting}
    \item \begin{lstlisting} numDEGs <- function(dds, group1, group2){} \end{lstlisting}
    \item \begin{lstlisting} plottotalDEGs <- function(myDEGS, mysubtitle){} \end{lstlisting}
\end{itemize}

These functions utilize exciting R functions and for loops to compare gene expression across levels.

\begin{figure}
\includegraphics[width=0.75\linewidth]{DoveParentsRNAseq_approach-2.png}
\end{figure}

\end{block}

\begin{block}{{\normalsize Quickly explore variation using the top 500 differentially expressed genes}}

\begin{figure}
\includegraphics[width=0.9\linewidth]{pca-1.pdf}
\end{figure}

\vspace{0.25em}

\end{block}

 \vspace{-3em}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.1

\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)

%----------------------------------------------------------------------------------------
%	RESULTS
%----------------------------------------------------------------------------------------

\vspace{-0.5em}

\begin{block}{{\normalsize Calculate and visualize  the total number of differentially expressed genes for each two-way comparison}}

\begin{figure}
\includegraphics[width=1\linewidth]{DoveParentsRNAseq_totalDEGs.png}
\end{figure}

\end{block}

\begin{block}{{\normalsize Deteremine the extent of correlation in gene expression variation}}

\begin{figure}
\includegraphics[width=0.5\linewidth]{correlationheatmaps-1.png}
\includegraphics[width=0.5\linewidth]{correlationheatmaps-4.png}
\end{figure}

\begin{figure}
\includegraphics[width=0.5\linewidth]{correlationheatmaps-2.png}
\includegraphics[width=0.5\linewidth]{correlationheatmaps-5.png}
\end{figure}

\begin{figure}
\includegraphics[width=0.5\linewidth]{correlationheatmaps-3.png}
\includegraphics[width=0.5\linewidth]{correlationheatmaps-6.png}
\end{figure}

\end{block}

\vspace{1em}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.2

\end{columns} % End of the split of column 2

\end{column} % End of the second column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The third column

%----------------------------------------------------------------------------------------
%	panel 4
%----------------------------------------------------------------------------------------

\vspace{-0.5em}

\begin{block}{Preliminary interpreations}

\begin{figure}
\includegraphics[width=1\linewidth]{DoveParentsRNAseq_summary.png}
%\caption{\url{https://datacarpentry.org/python-ecology-lesson/}}
\end{figure}

\end{block}

\begin{block}{Next steps}

\begin{itemize}
\item \textbf{Zoom out:} examine the full model of sex * tissue * treatment. 
\item \textbf{Zoom in:} examine effects of specific manipulations in greater detail. 
\item \textbf{Create} and open source R package that can be peer reviewed. 
\item \textbf{Share} tutorials for how to use the function. 
\end{itemize}

\end{block}

\begin{block}{Acknowledgements}

We are grateful to members of laboratories supervised by RMC, MM, Titus Brown, John Wingfield, Tom Hahn, and Marilyn Ramenofsky for comments and discussion. We thank Victoria Farrar, April Booth, Rechelle Viernes, Jonathan Perez, Jesse Krause, and over 20 undergraduate researchers who helped care for our bird colony. This work is funded by the National Science Foundation, IOS #1455957, to RMC and MM.

\end{block}

\vspace{0.5em}

\begin{flushright}
\textit{2019-06-15}
\end{flushright}

%----------------------------------------------------------------------------------------

\end{column} % End of the third column

\end{columns} % End of all the columns in the poster

\end{frame} % End of the enclosing frame

\end{document}

### 2019-06-16 docs/main.tex
version 2

@@
 
-\begin{block}{{\normalsize Deteremine the extent of correlation in gene expression variation}}
+\begin{block}{{\normalsize Determine the extent of correlation in gene expression variation}}
 
@@
 
-\begin{block}{Preliminary interpreations}
+\begin{block}{Preliminary interpretations}
 
@@
 \item \textbf{Zoom in:} examine effects of specific manipulations in greater detail. 
-\item \textbf{Create} and open source R package that can be peer reviewed. 
+\item \textbf{Test} specific hypotheses. 
+\item \textbf{Incorporate} code-review and peer-review . 
+\item \textbf{Create} and open source R package. 
 \item \textbf{Share} tutorials for how to use the function. 

### 2019-06-16 analysis/05_DESeq2_manipulate.Rmd
new pcs
mostly good, not heatmap
work in progress

@@
 
+## candidate genes
+
 ```{r candidates}
@@
 plotcandidates(dds.male_gondad, "male gonad")
+
 ```

### 2019-06-16 analysis/04_DESeq2_sexes.Rmd
work in progress

@@
 
-## candidate genes
+## make col data for candidate genes and heatmaps
 
-```{r candidates, eval =F}
-plotcandidates(dds.female_hypothalamus, a.colData, "female hypothalamus")
-plotcandidates(dds.female_pituitary, "female pituitary")
-plotcandidates(dds.female_gonad, "female gonad")
-plotcandidates(dds.male_hypothalamus, "male hypothalamus")
-plotcandidates(dds.male_pituitary, "male pituitary")
-plotcandidates(dds.male_gondad, "male gonad")
-```
-
-## new heat maps
-
-```{r correlationheatmaps, eval = T}
+```{r coldata, eval =T}
 colData.female_hypothalamus <- subsetcolData(a.colData,  "female_hypothalamus")
@@
 colData.male_gondad <- subsetcolData(a.colData,  "male_gonad")
+```
 
+## candidate genes
+
+```{r candidates, eval =T}
+plotcandidates(dds.female_hypothalamus, colData.female_hypothalamus, "female hypothalamus")
+plotcandidates(dds.female_pituitary, colData.female_pituitary,  "female pituitary")
+plotcandidates(dds.female_gonad, colData.female_gonad,  "female gonad")
+plotcandidates(dds.male_hypothalamus, colData.male_hypothalamus, "male hypothalamus")
+plotcandidates(dds.male_pituitary,  colData.male_pituitary, "male pituitary")
+plotcandidates(dds.male_gondad, colData.male_hypothalamus , "male gonad")
+```
+
+## new heat maps
+
+```{r correlationheatmaps, eval = F}
 plotcorrelationheatmaps(dds.female_hypothalamus, colData.female_hypothalamus ,"female hypothalamus: hatch, nest bldg., lay and day-before-hatch cluster")

### 2019-07-18 analysis/04_DESeq2_sexes.Rmd
change title

@@
 ```{r plottotalDEGs, eval = T}
-a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus: hatch is most unique parental stage")
+a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus")
 b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary: hatch and day before are most unique")
-c <- plottotalDEGs(DEGs.female_gonad, "female gonad: variable across all parental stages")
+c <- plottotalDEGs(DEGs.female_gonad, "female gonad")
 d <- plottotalDEGs(DEGs.male_hypothalamus, "male hypothalamus: many parental contrasts yield 0 DEGs")
@@
 
-myDEGs <- plot_grid(a + theme_noaxislabels,
-          d + theme_siggenes,
-          b + theme_noaxislabels,
-          e + theme_noaxislabels,
-          c + theme_noaxislabels,
-          f + theme_noaxislabels,
-          nrow = 3) 
+myDEGs <- plot_grid(a,
+          #d + theme_siggenes,
+          #b + theme_noaxislabels,
+          #e + theme_noaxislabels,
+          c ,
+          #f + theme_noaxislabels,
+          nrow = 1) 
 myDEGs

### 2019-07-18 analysis/04_DESeq2_sexes.Rmd
now with bar plots

@@
 library(viridis)
+library(forcats)
 
@@
 
-myDEGs <- plot_grid(a,
-          #d + theme_siggenes,
-          #b + theme_noaxislabels,
-          #e + theme_noaxislabels,
-          c ,
-          #f + theme_noaxislabels,
-          nrow = 1) 
+myDEGs <- plot_grid(a + theme_noaxislabels,
+          d + theme_siggenes,
+          b + theme_noaxislabels,
+          e + theme_noaxislabels,
+          c + theme_noaxislabels,
+          f + theme_noaxislabels,
+          nrow = 3) 
 myDEGs
@@
 plot(myDEGs)
+dev.off()
+```
+
+## Plot only a subset of DEGs that relate to hypotheses
+
+```{r totalDEGbarsplots}
+g <- plotserialDEGs(DEGs.female_hypothalamus, "Female hypothalamus")
+h <- plotserialDEGs(DEGs.female_pituitary, "Female pituitary")
+i <- plotserialDEGs(DEGs.female_gonad, "Female gonad")
+j <- plotserialDEGs(DEGs.male_hypothalamus, "Male hypothalamus")
+k <- plotserialDEGs(DEGs.male_pituitary, "Male pituitary")
+l <- plotserialDEGs(DEGs.male_gondad, "Male gonad")
+
+mybarplots <- plot_grid(g, h , i , j ,k , l, nrow = 2) 
+mybarplots
+
+pdf("../figures/sexes/totalDEGbarsplots-1.pdf", width = 10, height = 6)
+plot(mybarplots)
 dev.off()
@@
 
-```{r candidates, eval =T}
+```{r candidates, eval =F}
 plotcandidates(dds.female_hypothalamus, colData.female_hypothalamus, "female hypothalamus")

### 2019-07-19 analysis/04_DESeq2_sexes.Rmd
now with bars and lines, two color options

@@
 ```{r totalDEGbarsplots}
-g <- plotserialDEGs(DEGs.female_hypothalamus, "Female hypothalamus")
-h <- plotserialDEGs(DEGs.female_pituitary, "Female pituitary")
-i <- plotserialDEGs(DEGs.female_gonad, "Female gonad")
-j <- plotserialDEGs(DEGs.male_hypothalamus, "Male hypothalamus")
-k <- plotserialDEGs(DEGs.male_pituitary, "Male pituitary")
-l <- plotserialDEGs(DEGs.male_gondad, "Male gonad")
+g <- plotserialDEGs(DEGs.female_hypothalamus, "Female hypothalamus", "#969696")
+h <- plotserialDEGs(DEGs.female_pituitary, "Female pituitary", "#969696")
+i <- plotserialDEGs(DEGs.female_gonad, "Female gonad", "#969696")
+j <- plotserialDEGs(DEGs.male_hypothalamus, "Male hypothalamus", "#969696")
+k <- plotserialDEGs(DEGs.male_pituitary, "Male pituitary", "#969696")
+l <- plotserialDEGs(DEGs.male_gondad, "Male gonad", "#969696")
 
@@
 dev.off()
+
+g <- plotserialDEGs(DEGs.female_hypothalamus, "Female hypothalamus", "#00A08A")
+h <- plotserialDEGs(DEGs.female_pituitary, "Female pituitary", "#00A08A")
+i <- plotserialDEGs(DEGs.female_gonad, "Female gonad", "#00A08A")
+j <- plotserialDEGs(DEGs.male_hypothalamus, "Male hypothalamus", "#F98400")
+k <- plotserialDEGs(DEGs.male_pituitary, "Male pituitary", "#F98400")
+l <- plotserialDEGs(DEGs.male_gondad, "Male gonad", "#F98400")
+
+mybarplots <- plot_grid(g, h , i , j ,k , l, nrow = 2) 
+mybarplots
+
+pdf("../figures/sexes/totalDEGbarsplots-2.pdf", width = 10, height = 6)
+plot(mybarplots)
+dev.off()
+
 ```

### 2019-07-19 analysis/04_DESeq2_sexes.Rmd
better names

--- 2019-07-18 analysis/04_DESeq2_sexes.Rmd
@@
 
-```{r totalDEGbarsplots}
-g <- plotserialDEGs(DEGs.female_hypothalamus, "Female hypothalamus")
-h <- plotserialDEGs(DEGs.female_pituitary, "Female pituitary")
-i <- plotserialDEGs(DEGs.female_gonad, "Female gonad")
-j <- plotserialDEGs(DEGs.male_hypothalamus, "Male hypothalamus")
-k <- plotserialDEGs(DEGs.male_pituitary, "Male pituitary")
-l <- plotserialDEGs(DEGs.male_gondad, "Male gonad")
+```{r characterization}
+g <- plotserialDEGs(DEGs.female_hypothalamus, "Female hypothalamus", "#00A08A")
+h <- plotserialDEGs(DEGs.female_pituitary, "Female pituitary", "#00A08A")
+i <- plotserialDEGs(DEGs.female_gonad, "Female gonad", "#00A08A")
+j <- plotserialDEGs(DEGs.male_hypothalamus, "Male hypothalamus", "#F98400")
+k <- plotserialDEGs(DEGs.male_pituitary, "Male pituitary", "#F98400")
+l <- plotserialDEGs(DEGs.male_gondad, "Male gonad", "#F98400")
 
@@
 
-pdf("../figures/sexes/totalDEGbarsplots-1.pdf", width = 10, height = 6)
+pdf("../figures/sexes/characterization-1.pdf", width = 10, height = 6)
 plot(mybarplots)
 dev.off()
+```
+
+```{r manipulation}
+
 ```

### 2019-07-19 analysis/04_DESeq2_sexes.Rmd
first pass a minipulate, don't like

@@
 
-```{r manipulation}
+```{r manipulationVcontrol}
+versuscontrol <- c("control.m.inc.d3" , "control.m.inc.d8", "control.m.inc.d9", "control.m.inc.d17", 
+                      "control.m.n2", "control.prolong", "control.extend")
+
+manipulationVcontrol <- function(DEGs, mysubtitle, myfill){
+
+  # subset to look within one tissue in one sex
+  DEGs <- DEGs %>%
+    dplyr::mutate(comparison = row.names(.)) %>%
+    dplyr::filter(comparison %in% versuscontrol)
+
+  DEGs$comparison <- factor(DEGs$comparison, levels = versuscontrol)
+
+  mybarplot <- ggplot(DEGs, aes(comparison)) +
+    geom_bar(aes(weight = V3), fill = myfill) +
+    theme_minimal(base_size = 8) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+    labs(subtitle = mysubtitle, y = "Number of DEGs", x = NULL) +
+    #ylim(0, 5105) +
+    geom_line(aes(x=comparison, y=V3, group = 1))
+
+  return(mybarplot)
+}
+
+g <- manipulationVcontrol(DEGs.female_hypothalamus, "Female hypothalamus", "#00A08A")
+h <- manipulationVcontrol(DEGs.female_pituitary, "Female pituitary", "#00A08A")
+i <- manipulationVcontrol(DEGs.female_gonad, "Female gonad", "#00A08A")
+j <- manipulationVcontrol(DEGs.male_hypothalamus, "Male hypothalamus", "#F98400")
+k <- manipulationVcontrol(DEGs.male_pituitary, "Male pituitary", "#F98400")
+l <- manipulationVcontrol(DEGs.male_gondad, "Male gonad", "#F98400")
+
+mybarplots <- plot_grid(g, h , i , j ,k , l, nrow = 2) 
+mybarplots
+
+pdf("../figures/sexes/manipulationVcontrol-1.pdf", width = 10, height = 6)
+plot(mybarplots)
+dev.off()
 

### 2019-07-19 analysis/04_DESeq2_sexes.Rmd
pause manip
now with manipulation bar graphs

--- 2019-07-19 analysis/04_DESeq2_sexes.Rmd
@@
 library(forcats)
+library("wesanderson")
 
@@
 ```{r characterization}
+# bars with lines
 g <- plotserialDEGs(DEGs.female_hypothalamus, "Female hypothalamus", "#00A08A")
@@
 dev.off()
+
+# lines only, different colors for different sex tissue groups
+FH <- subsetDEGs(DEGs.female_hypothalamus, "Female hypothalamus")
+FP <- subsetDEGs(DEGs.female_pituitary, "Female pituitary")
+FG <- subsetDEGs(DEGs.female_gonad, "Female gonad")
+MH <- subsetDEGs(DEGs.male_hypothalamus, "Male hypothalamus")
+MP <- subsetDEGs(DEGs.male_pituitary, "Male pituitary")
+MG <- subsetDEGs(DEGs.male_gondad, "Male gonad")
+
+specificDEGs <- rbind(FH, FP, FG, MH, MP, MG)
+
+specificDEGs$sextissue <- factor(specificDEGs$sextissue, levels = c("Female hypothalamus", "Female pituitary", "Female gonad",
+                                                                    "Male hypothalamus", "Male pituitary", "Male gonad"))
+mylinegraph <- ggplot(specificDEGs, aes(x=comparison, y=V3, group = sextissue)) +
+    geom_line(aes(color=sextissue)) +
+    geom_point(aes(color=sextissue)) + 
+    theme_minimal(base_size = 10) +
+    #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+    labs(x = "Parental stage transitions", y= "No. of differentially expressed genes ",
+         title = "Gene expression changes during parental care transitions") + 
+    scale_x_discrete(labels=c("Control to\nNest bldg", "Nest bldg. to\nLay", "Lay to\nInc. day 3",
+                              "Inc. day 3 to\nInc. day 9", "Inc. day 9 to\nInc. day 17",
+                              "Inc. day 17 to\nHatch", "Hatch to\nNestling day 5", "Nestling day 5 to\n Nestling day 9")) +
+   theme(legend.title = element_blank(),
+         legend.position = c(0.8, 0.8))
+
+mylinegraph
+
+pdf("../figures/sexes/characterization-2.pdf", width = 8, height = 4)
+plot(mylinegraph)
+dev.off()
 ```
 
-```{r manipulation}
+```{r manipVchar}
+g <- plotmanipDEGs(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
+h <- plotmanipDEGs(DEGs.female_pituitary, "Female pituitary", "none")
+i <- plotmanipDEGs(DEGs.female_gonad, "Female gonad",  "none")
+j <- plotmanipDEGs(DEGs.male_hypothalamus, "Male hypothalamus", "none")
+k <- plotmanipDEGs(DEGs.male_pituitary, "Male pituitary", "none")
+l <- plotmanipDEGs(DEGs.male_gondad, "Male gonad", "none")
 
+legend <- get_legend(g)
+
+mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)
+
+mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
+mybarplots2
+
+pdf("../figures/sexes/manipVchar-1.pdf", width = 10, height = 6)
+plot(mybarplots2)
+dev.off()
+
+plotmanipDEGs(DEGs.male_pituitary, "Male pituitary", "bottom")
 ```

### 2019-07-19 R/functions.R

@@ -156,9 +156,57 @@ plotserialDEGs <- function(DEGs, mysubtitle, myfill){
 }
 
 
+## characterization line graphs
+
+
+subsetDEGs <- function(DEGs, groupname){
+  DEGs <- DEGs %>%
+    dplyr::mutate(comparison = row.names(.),
+                  sextissue = groupname) %>%
+    dplyr::filter(comparison %in% serialtimepoints) 
+  DEGs$comparison <- factor(DEGs$comparison, levels = serialtimepoints)
+  return(DEGs)
+}
+  
+  
+
 # bar plot with manipulations
 
+manipVchar <- c("inc.d3.m.inc.d3",
+                "inc.d9.m.inc.d8", "inc.d9.m.inc.d9",
+                "inc.d17.m.inc.d17",
+                "hatch.prolong", "hatch.extend", "hatch.m.n2")
 
+plotmanipDEGs <- function(DEGs, mysubtitle, legendornot){
+  
+  # subset to look within one tissue in one sex
+  DEGs <- DEGs %>%
+    dplyr::mutate(comparison = row.names(.)) %>%
+    dplyr::filter(comparison %in% manipVchar) 
+  
+  DEGs$description <- ifelse(grepl("inc.d9.m.inc.d8", DEGs$comparison),"early chicks", 
+                             ifelse(grepl("d3|d17|d9", DEGs$comparison),"remove eggs",
+                                    ifelse(grepl("n2", DEGs$comparison),"remove chicks",
+                                           ifelse(grepl("prolong", DEGs$comparison),"prolong inc.",
+                                                  ifelse(grepl("extend", DEGs$comparison),"extend inc.", NA)))))
+  
+  DEGs$comparison <- factor(DEGs$comparison, levels = manipVchar)
+  
+  mybarplot <- ggplot(DEGs, aes(comparison)) +
+    geom_bar(aes(weight = V3, fill = description)) +
+    theme_minimal(base_size = 8) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+    labs(subtitle = mysubtitle, y = "Number of DEGs", x = NULL) +
+    #ylim(0, 5105) +
+    #geom_line(aes(x=comparison, y=V3, group = 1)) +
+    scale_x_discrete(labels=c("Day 3\nRemove eggs", "Day 9\nAdd chicks", "Day 9\nRemove eggs",
[…]

### 2019-07-20 analysis/04_DESeq2_sexes.Rmd
without gridlines

@@
          legend.position = c(0.8, 0.8))
-
 mylinegraph

### 2019-07-20 analysis/04_DESeq2_sexes.Rmd
now with two parts manipulation

@@
 
+```{r offspringremoval}
+g <- plotremoval(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
+h <- plotremoval(DEGs.female_pituitary, "Female pituitary", "none")
+i <- plotremoval(DEGs.female_gonad, "Female gonad",  "none")
+j <- plotremoval(DEGs.male_hypothalamus, "Male hypothalamus", "none")
+k <- plotremoval(DEGs.male_pituitary, "Male pituitary", "none")
+l <- plotremoval(DEGs.male_gondad, "Male gonad", "none")
+
+legend <- get_legend(g)
+
+mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)
+
+mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
+mybarplots2
+
+pdf("../figures/sexes/offspringremoval-1.pdf", width = 10, height = 6)
+plot(mybarplots2)
+dev.off()
+
+```
+
+```{r prolongextend}
+g <- plotprolongdelay(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
+h <- plotprolongdelay(DEGs.female_pituitary, "Female pituitary", "none")
+i <- plotprolongdelay(DEGs.female_gonad, "Female gonad",  "none")
+j <- plotprolongdelay(DEGs.male_hypothalamus, "Male hypothalamus", "none")
+k <- plotprolongdelay(DEGs.male_pituitary, "Male pituitary", "none")
+l <- plotprolongdelay(DEGs.male_gondad, "Male gonad", "none")
+
+legend <- get_legend(g)
+
+mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)
+
+mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
+mybarplots2
+
+pdf("../figures/sexes/prolongextend-1.pdf", width = 10, height = 6)
+plot(mybarplots2)
+dev.off()
+
+```
+
 ## Calculate and plot principal components

### 2019-07-24 analysis/03_DESeq2_characterization.Rmd
first pass usical notes

@@
 library(viridis)
+library("ggimage")
 
@@
 
-## Manipulation data
+## Characterization data
 
@@
 # import "colData" which contains sample information and "countData" which contains read counts
-m.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
-m.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
+c.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
+c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
 geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
@@
 # set levels
-levels(m.colData$treatment)
-m.colData$treatment <- factor(m.colData$treatment, levels = 
+c.colData$treatment <- factor(c.colData$treatment, levels = 
                               c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
+levels(c.colData$treatment)
+c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
+summary(c.colData[c(7,3,4,5,8)])
+```
 
-m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))
-summary(m.colData[c(7,3,4,5,8)])
+## plot raw counts
+
+```{r musicalgenes, cache = T, eval=T}
+head(countData)
+head(geneinfo)
+
+candidataCounts <- countData
+candidataCounts$Name <- row.names(candidataCounts)
+candidataCounts <- candidataCounts %>%
+  filter(Name %in% c("NP_990797.2", "NP_001138861.1"))
+row.names(candidataCounts) <- candidataCounts$Name
+candidataCounts$Name <- NULL
+candidataCountsT <- t(candidataCounts)
+candidataCountsT <- as.data.frame(candidataCountsT)
+
+candidataCountsT$entrez <- row.names(candidataCountsT)
+colData$entrez <- row.names(colData)
+
+candidataCountsTcol <- full_join(colData, candidataCountsT)
+
+str(candidataCountsTcol)
+
+candidataCountsTcol2 <- candidataCountsTcol %>%
+  group_by(sextissue, treatment) %>%
+  summarize(GAL = mean(NP_001138861.1, na.rm = TRUE),
+            PRL = mean(NP_990797.2, na.rm = TRUE)) %>%
+  gather(GAL, PRL, key = "gene", value = "mean")
+
+candidataCountsTcol2 <- as.data.frame(candidataCountsTcol2)
+
+musicgenes <- function(mygroup, mysusbtitle){
+
+myplot <- candidataCountsTcol2 %>%
+  filter(sextissue %in% c(mygroup)) %>%
+ggplot(aes(x = treatment, y = mean)) +
+  geom_point(colour = "white") + 
+  geom_line(group = 1, color='grey') +
+  theme_bw() +
+  facet_wrap(~gene, scales = "free", ncol = 1) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  geom_image(aes(image="https://storage.needpix.com/rsynced_images/music-34177_1280.png")) +
+  labs(y = "mean gene expression", subtitle = mysusbtitle, x = NULL)
+
+  return(myplot)
+
+}
+
+a <- musicgenes("female_hypothalamus", "female hypothalamus")
+b <- musicgenes("male_hypothalamus", "male hypothalamus")
+
+c <- plot_grid(a,b)
+c
+
+pdf("../figures/characterization/musicalgenes-1.pdf", width = 8, height = 6)
+plot(c)
+dev.off()
+
 ```
@@
 ```{r dds, cache = T}
-dds.female_hypothalamus <- subsetDESeq("female_hypothalamus")
-dds.female_pituitary <- subsetDESeq("female_pituitary" )
-dds.female_gonad <- subsetDESeq("female_gonad" )
-dds.male_hypothalamus <- subsetDESeq("male_hypothalamus" )
-dds.male_pituitary <- subsetDESeq("male_pituitary"  )
-dds.male_gondad <- subsetDESeq("male_gonad")
+dds.female_hypothalamus <- subsetDESeq(c.colData,  c.countData, "female_hypothalamus")
+dds.female_pituitary <- subsetDESeq(c.colData,  c.countData, "female_pituitary" )
+dds.female_gonad <- subsetDESeq(c.colData,  c.countData, "female_gonad" )
+dds.male_hypothalamus <- subsetDESeq(c.colData,  c.countData, "male_hypothalamus" )
+dds.male_pituitary <- subsetDESeq(c.colData,  c.countData, "male_pituitary"  )
+dds.male_gondad <- subsetDESeq(c.colData,  c.countData, "male_gonad")
 ```
@@
 
-```{r pheatmap, eval = T}
+```{r pheatmap, eval = F}
 makepheatmap(dds.female_hypothalamus, "female hypothalamus")
@@
 
-```{r candidates}
+```{r candidates, eval = F}
 plotcandidates(dds.female_hypothalamus, "female hypothalamus")

### 2019-07-24 analysis/03_DESeq2_characterization.Rmd
musicgenes4

@@
 
+musicgenes2 <- function(mygroup, mysusbtitle){
+
+myplot <- candidataCountsTcol2 %>%
+  filter(sextissue %in% c(mygroup)) %>%
+ggplot(aes(x = treatment, y = mean)) +
+  geom_point(colour = "white") + 
+  theme_bw() +
+  facet_wrap(~gene, scales = "free", ncol = 1) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  geom_image(aes(image="https://storage.needpix.com/rsynced_images/music-34177_1280.png")) +
+  labs(y = "mean gene expression", subtitle = mysusbtitle, x = NULL) +
+
+  return(myplot)
+
+}
+
+a <- musicgenes2("female_hypothalamus", "female hypothalamus")
+b <- musicgenes2("male_hypothalamus", "male hypothalamus")
+
+c <- plot_grid(a,b)
+c
+
+pdf("../figures/characterization/musicalgenes-2.pdf", width = 8, height = 6)
+plot(c)
+dev.off()
+
 ```

### 2019-07-24 analysis/03_DESeq2_characterization.Rmd
now with options

--- 2019-07-24 analysis/03_DESeq2_characterization.Rmd
@@
 
+musicgenes5 <- function(mygroup, mysusbtitle){
+  myplot <- candidataCountsTcol2 %>%
+  filter(sextissue %in% c(mygroup),
+         gene == "GAL") %>%
+    ggplot(aes(x = treatment, y = mean)) +
+    geom_point(colour = "white") + 
+    theme_minimal() +
+    facet_wrap(~gene,  ncol = 1) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          strip.text = element_text(face = "italic"),
+          panel.grid = element_blank()) +
+    geom_image(aes(image="https://storage.needpix.com/rsynced_images/music-34177_1280.png")) +
+   labs(y = "mean gene expression", subtitle = mysusbtitle, x = NULL) +
+    ylim(c(20,70))
+
+  return(myplot)
+
+}
+
+a <- musicgenes5("female_hypothalamus", "female hypothalamus")
+b <- musicgenes5("male_hypothalamus", "male hypothalamus")
+
+c <- plot_grid(a,b)
+c
+
+pdf("../figures/characterization/musicalgenes-6.pdf", width = 6, height = 3)
+plot(c)
+dev.off()
+
 ```

### 2019-07-24 analysis/03_DESeq2_characterization.Rmd
now with musical genes tricolor

--- 2019-07-24 analysis/03_DESeq2_characterization.Rmd
@@
 ```{r musicalgenes, cache = T, eval=T}
-head(countData)
-head(geneinfo)
+countData <- c.countData
+colData <- c.colData
 
@@
 candidataCounts <- candidataCounts %>%
-  filter(Name %in% c("NP_990797.2", "NP_001138861.1"))
+  filter(Name %in% c("NP_990797.2", "NP_001138861.1", "NP_001107180.1"))
 row.names(candidataCounts) <- candidataCounts$Name
@@
   summarize(GAL = mean(NP_001138861.1, na.rm = TRUE),
-            PRL = mean(NP_990797.2, na.rm = TRUE)) %>%
-  gather(GAL, PRL, key = "gene", value = "mean")
+            PRL = mean(NP_990797.2, na.rm = TRUE),
+            CALCA = mean(NP_001107180.1, na.rm = TRUE)) %>%
+  gather(GAL, PRL, CALCA, key = "gene", value = "mean")
 
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  geom_image(aes(image="https://storage.needpix.com/rsynced_images/music-34177_1280.png")) +
+  geom_image(aes(image="../figures/characterization/note-black-04.png")) +
   labs(y = "mean gene expression", subtitle = mysusbtitle, x = NULL)
@@
 
+musicgeneslines <- function(mygroup, mysusbtitle, mygene, myimage, myylim, myhline){
+  myplot <- candidataCountsTcol2 %>%
+  filter(sextissue %in% c(mygroup),
+         gene == mygene) %>%
+    ggplot(aes(x = treatment, y = mean)) +
+    geom_point(colour = "white") + 
+    theme_minimal() +
+    #facet_wrap(~gene,  ncol = 1) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          strip.text = element_text(face = "italic"),
+          panel.grid = element_blank(),
+          axis.text.y = element_blank()) +
+    geom_image(aes(image=myimage)) +
+   labs(y = "mean gene expression", subtitle = mysusbtitle, x = NULL) +
+    ylim(myylim) + 
+    geom_hline(yintercept = myhline, colour="grey")
+  return(myplot)
+}
+
+musicgenesnolines <- function(mygroup, mysusbtitle, mygene, myimage, myylim, myhline){
+  myplot <- candidataCountsTcol2 %>%
+  filter(sextissue %in% c(mygroup),
+         gene == mygene) %>%
+    ggplot(aes(x = treatment, y = mean)) +
+    geom_point(colour = "white") + 
+    theme_minimal(base_size = 10) +
+    #facet_wrap(~gene,  ncol = 1) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          strip.text = element_text(face = "italic"),
+          panel.grid = element_blank(),
+          axis.text.y = element_blank()) +
+    geom_image(aes(image=myimage)) +
+   labs(y = "mean gene expression", subtitle = mysusbtitle, x = NULL) +
+    ylim(myylim)
+
+  return(myplot)
+}
+
+a <- musicgenesnolines("female_hypothalamus", "female hypothalamus", "GAL", "../figures/characterization/note-orange-04.png", c(15,75))
+a
+
+b <- musicgenesnolines("male_hypothalamus", "male hypothalamus", "GAL", "../figures/characterization/note-orange-04.png", c(15,75))
+b
+
+ab <- plot_grid(a,b) + cowplot::draw_text("GAL", x = 0.925, y = 0.9, size = 8, colour = "#d95f02")
+ab
+
+c <- musicgenesnolines("female_hypothalamus", "female hypothalamus", "CALCA", "../figures/characterization/note-teal-04.png", c(0,100))
+c
+
+d <- musicgenesnolines("male_hypothalamus", "male hypothalamus", "CALCA", "../figures/characterization/note-teal-04.png", c(0,100))
+d
+
+cd <- plot_grid(c,d)  + cowplot::draw_text("CALCA", x = 0.925, y = 0.85, size = 8, colour = "#1b9e77")
+cd
+
+e <- musicgenesnolines("female_hypothalamus", "female hypothalamus", "PRL", "../figures/characterization/note-purple-04.png", c(800,3000) )
+e
+
+f <- musicgenesnolines("male_hypothalamus", "male hypothalamus", "PRL", "../figures/characterization/note-purple-04.png", c(800,3000))
+f
+
+ef <- plot_grid(e,f) + cowplot::draw_text("PRL", x = 0.925, y = 0.8, size = 8, colour = "#7570b3")
+ef
+
+pdf("../figures/characterization/musicalgenes-ab.pdf", width = 4, height = 2)
+plot(ab)
+dev.off()
+
+pdf("../figures/characterization/musicalgenes-cd.pdf", width = 4, height = 2)
+plot(cd)
+dev.off()
+
+pdf("../figures/characterization/musicalgenes-ef.pdf", width = 4, height = 2)
+plot(ef)
+dev.off()
+
 ```
@@
 
-```{r dds, cache = T}
+```{r dds, cache = T, eval = F}
 dds.female_hypothalamus <- subsetDESeq(c.colData,  c.countData, "female_hypothalamus")

### 2019-08-13 analysis/06_musicalgenes.Rmd
move musical genes to new script

--- 2019-07-24 analysis/03_DESeq2_characterization.Rmd
@@
 ---
-title: "02_DESeq_characterization"
+title: "06_musicalgenes"
 output: md_document
@@
 library(tidyverse)
-library(DESeq2)
 library(cowplot)
-library(RColorBrewer)
-library(pheatmap)
-library(kableExtra)
-library(viridis)
-library("ggimage")
+library(ggimage)
 
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/characterizatio/', cache = TRUE)
+knitr::opts_chunk$set(fig.path = '../figures/characterization/', cache = TRUE)
 ```
@@
 
-## plot raw counts
+## Musical genes
 
@@
 ```
-
-## Run DESeq on all subsets of the data
-
-```{r dds, cache = T, eval = F}
-dds.female_hypothalamus <- subsetDESeq(c.colData,  c.countData, "female_hypothalamus")
-dds.female_pituitary <- subsetDESeq(c.colData,  c.countData, "female_pituitary" )
-dds.female_gonad <- subsetDESeq(c.colData,  c.countData, "female_gonad" )
-dds.male_hypothalamus <- subsetDESeq(c.colData,  c.countData, "male_hypothalamus" )
-dds.male_pituitary <- subsetDESeq(c.colData,  c.countData, "male_pituitary"  )
-dds.male_gondad <- subsetDESeq(c.colData,  c.countData, "male_gonad")
-```
-
-## Calculate and plot total DEGs
-
-```{r totalDEGs, eval = T}
-#create list of groups for deseq contrasts
-group1 <- c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9")
-group2 <- group1
-
-a <- plottotalDEGschar(dds.female_hypothalamus, "female hypothalamus")
-b <- plottotalDEGschar(dds.female_pituitary, "female pituitary")
-c <- plottotalDEGschar(dds.female_gonad, "female gonad")
-d <- plottotalDEGschar(dds.male_hypothalamus, "male hypothalamus")
-e <- plottotalDEGschar(dds.male_pituitary, "male pituitary")
-f <- plottotalDEGschar(dds.male_gondad, "male gonad")
-
-plot_grid(a + theme(legend.position = "none"),
-          b + theme(legend.position = "none"),
-          c,
-          d + theme(legend.position = "none"),
-          e + theme(legend.position = "none"),
-          f,
-          nrow = 2, rel_widths = c(0.3, 0.3, 0.4)) 
-```
-
-## Calculate and plot principal component
-
-```{r pca}
-plotPCAs(dds.female_hypothalamus, "female hypothalamus")
-plotPCAs(dds.female_pituitary, "female pituitary")      
-plotPCAs(dds.female_gonad, "female gonad")
-plotPCAs(dds.male_hypothalamus, "male hypothalamus")
-plotPCAs(dds.male_pituitary, "male pituitary")
-plotPCAs(dds.male_gondad, "male gonad")
-```
-
-## heamap with minimum pvalue
-
-```{r pheatmap, eval = F}
-makepheatmap(dds.female_hypothalamus, "female hypothalamus")
-makepheatmap(dds.female_pituitary, "female pituitary")
-makepheatmap(dds.female_gonad, "female gonad")
-makepheatmap(dds.male_hypothalamus, "male hypothalamus")
-makepheatmap(dds.male_pituitary, "male pituitary")
-makepheatmap(dds.male_gondad, "male gonad")
-```
-## plot candidate genes
-
-```{r candidates, eval = F}
-plotcandidates(dds.female_hypothalamus, "female hypothalamus")
-plotcandidates(dds.female_pituitary, "female pituitary")
-plotcandidates(dds.female_gonad, "female gonad")
-plotcandidates(dds.male_hypothalamus, "male hypothalamus")
-plotcandidates(dds.male_pituitary, "male pituitary")
-plotcandidates(dds.male_gondad, "male gonad")
-```

### 2019-08-13 analysis/03_DESeq2_characterization.Rmd
new musical genes
eval = F for all the things :(

--- 2019-07-24 analysis/03_DESeq2_characterization.Rmd
@@
 library(viridis)
-library("ggimage")
+library(ggimage)
 
@@
 
-## plot raw counts
-
-```{r musicalgenes, cache = T, eval=T}
-head(countData)
-head(geneinfo)
-
-candidataCounts <- countData
-candidataCounts$Name <- row.names(candidataCounts)
-candidataCounts <- candidataCounts %>%
-  filter(Name %in% c("NP_990797.2", "NP_001138861.1"))
-row.names(candidataCounts) <- candidataCounts$Name
-candidataCounts$Name <- NULL
-candidataCountsT <- t(candidataCounts)
-candidataCountsT <- as.data.frame(candidataCountsT)
-
-candidataCountsT$entrez <- row.names(candidataCountsT)
-colData$entrez <- row.names(colData)
-
-candidataCountsTcol <- full_join(colData, candidataCountsT)
-
-str(candidataCountsTcol)
-
-candidataCountsTcol2 <- candidataCountsTcol %>%
-  group_by(sextissue, treatment) %>%
-  summarize(GAL = mean(NP_001138861.1, na.rm = TRUE),
-            PRL = mean(NP_990797.2, na.rm = TRUE)) %>%
-  gather(GAL, PRL, key = "gene", value = "mean")
-
-candidataCountsTcol2 <- as.data.frame(candidataCountsTcol2)
-
-musicgenes <- function(mygroup, mysusbtitle){
-
-myplot <- candidataCountsTcol2 %>%
-  filter(sextissue %in% c(mygroup)) %>%
-ggplot(aes(x = treatment, y = mean)) +
-  geom_point(colour = "white") + 
-  geom_line(group = 1, color='grey') +
-  theme_bw() +
-  facet_wrap(~gene, scales = "free", ncol = 1) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  geom_image(aes(image="https://storage.needpix.com/rsynced_images/music-34177_1280.png")) +
-  labs(y = "mean gene expression", subtitle = mysusbtitle, x = NULL)
-
-  return(myplot)
-
-}
-
-a <- musicgenes("female_hypothalamus", "female hypothalamus")
-b <- musicgenes("male_hypothalamus", "male hypothalamus")
-
-c <- plot_grid(a,b)
-c
-
-pdf("../figures/characterization/musicalgenes-1.pdf", width = 8, height = 6)
-plot(c)
-dev.off()
-
-musicgenes2 <- function(mygroup, mysusbtitle){
-
-myplot <- candidataCountsTcol2 %>%
-  filter(sextissue %in% c(mygroup)) %>%
-ggplot(aes(x = treatment, y = mean)) +
-  geom_point(colour = "white") + 
-  theme_bw() +
-  facet_wrap(~gene, scales = "free", ncol = 1) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  geom_image(aes(image="https://storage.needpix.com/rsynced_images/music-34177_1280.png")) +
-  labs(y = "mean gene expression", subtitle = mysusbtitle, x = NULL) +
-
-  return(myplot)
-
-}
-
-a <- musicgenes2("female_hypothalamus", "female hypothalamus")
-b <- musicgenes2("male_hypothalamus", "male hypothalamus")
-
-c <- plot_grid(a,b)
-c
-
-pdf("../figures/characterization/musicalgenes-2.pdf", width = 8, height = 6)
-plot(c)
-dev.off()
-
-```
-
 ## Run DESeq on all subsets of the data
 
-```{r dds, cache = T}
+```{r dds, cache = T, eval = F}
 dds.female_hypothalamus <- subsetDESeq(c.colData,  c.countData, "female_hypothalamus")
@@
 
-```{r totalDEGs, eval = T}
+```{r totalDEGs, eval = F}
 #create list of groups for deseq contrasts
@@
 
-a <- plottotalDEGschar(dds.female_hypothalamus, "female hypothalamus")
-b <- plottotalDEGschar(dds.female_pituitary, "female pituitary")
-c <- plottotalDEGschar(dds.female_gonad, "female gonad")
-d <- plottotalDEGschar(dds.male_hypothalamus, "male hypothalamus")
-e <- plottotalDEGschar(dds.male_pituitary, "male pituitary")
-f <- plottotalDEGschar(dds.male_gondad, "male gonad")
+a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
+b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
+c <- plottotalDEGs(dds.female_gonad, "female gonad")
+d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
+e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
+f <- plottotalDEGs(dds.male_gondad, "male gonad")
 
@@
 
-```{r pca}
+```{r pca, eval = F}
 plotPCAs(dds.female_hypothalamus, "female hypothalamus")

### 2019-08-13 analysis/07_topconfects.Rmd
first attempt at topconfects

---
title: "07_topconfects"
output: md_document
---

## Based on a discussion on Twitter, https://twitter.com/Jared_Mamrot/status/1161047070821646336, I decided to try out the R package `topconfects` to test the effect size.

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
library(DESeq2)
library(topconfects)
library(tidyverse)

knitr::opts_chunk$set(fig.path = '../figures/topconfects/', cache = TRUE, message = F, echo = T)
```

## This is a demo from the vignette showing how the tool works.

```{r vignette}
# Confident log2 fold changes based on a DESeq2 analysis

# Generate some random data
n <- 20
folds <- seq(-8,8,length.out=n)
row_means <- runif(n, min=0, max=5)
lib_scale <- c(1,2,3,4)
means <- 2^(outer(folds, c(-0.5,-0.5,0.5,0.5))) *
  row_means * rep(lib_scale,each=n)
counts <- rnbinom(length(means), mu=means, size=1/0.1)
dim(counts) <- dim(means)
group <- factor(c("A","A","B","B"))
# Apply DESeq2
dds <- DESeqDataSetFromMatrix(
  countData = counts,
  colData = data.frame(group=group),
  design = ~group)
dds <- DESeq(dds)
# Find top confident effect sizes
deseq2_confects(dds, name="group_B_vs_A", step=0.1)

```

## Now I import the gene expression data and order the treatment according the the parentatl care cycle

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
c.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

# create grouping
c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))

c.colData$treatment <- factor(c.colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
levels(c.colData$treatment)
```

## This is a function to subset the data to look compare one transition (eg. incubation day 17 to lay) in a single "sex tissue" group (e.g. female pitutiary)

```{r subsetnconfects}
subsetnconfects <- function(mygroup, mytreatments, mycomparison){
  # subset col and count data
  colData <- c.colData %>%
      dplyr::filter(sextissue == mygroup) %>%
      dplyr::filter(treatment %in% mytreatments) %>%
      droplevels()
  row.names(colData) <- colData$V1

  # which counts to save
  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  # save counts that match colData
  countData <- c.countData %>% dplyr::select(one_of(savecols))

  # check that row and col lenghts are equal
  #print(ncol(countData) == nrow(colData))

  # run deseq
  dds <- DESeqDataSetFromMatrix(
    countData = countData,
   colData = colData,
   design = ~treatment)
  dds <- DESeq(dds)
  res <- results(dds)
  print(summary(res))
  # Find top confident effect sizes
  deseq2_confects(dds, name=mycomparison, step=0.1)
}
```

## Number of differentially expressed genes for the building versus lay comparison

```{r treatment_lay_vs_bldg}
subsetnconfects("female_hypothalamus", c("bldg", "lay"), "treatment_lay_vs_bldg") #1
subsetnconfects("male_hypothalamus", c("bldg", "lay"), "treatment_lay_vs_bldg") #0

subsetnconfects("female_pituitary", c("bldg", "lay"), "treatment_lay_vs_bldg") #455
subsetnconfects("male_pituitary", c("bldg", "lay"), "treatment_lay_vs_bldg") #0

subsetnconfects("female_gonad", c("bldg", "lay"), "treatment_lay_vs_bldg") #452
subsetnconfects("male_gonad", c("bldg", "lay"), "treatment_lay_vs_bldg") #26

```

## Number of differentially expressed genes for the incubation day 17 versus hatch comparison

```{r treatment_hatch_vs_inc.d17}
subsetnconfects("female_hypothalamus", c("inc.d17", "hatch"), "treatment_hatch_vs_inc.d17") #0
subsetnconfects("male_hypothalamus", c("inc.d17", "hatch"), "treatment_hatch_vs_inc.d17") #3

subsetnconfects("female_pituitary", c("inc.d17", "hatch"), "treatment_hatch_vs_inc.d17") #6
subsetnconfects("male_pituitary", c("inc.d17", "hatch"), "treatment_hatch_vs_inc.d17") #24

subsetnconfects("female_gonad", c("inc.d17", "hatch"), "treatment_hatch_vs_inc.d17") #3
subsetnconfects("male_gonad", c("inc.d17", "hatch"), "treatment_hatch_vs_inc.d17") #1

```

### 2019-08-14 analysis/07_topconfects.Rmd
small change

@@
 library(DESeq2)
+library(tidyverse)
 library(topconfects)
-library(tidyverse)
 
@@
 
+#deseq2_confects(dds, name="group_A_vs_B", step=0.1) # Error: subscript contains invalid names
+
 ```
@@
 
-```{r subsetnconfects}
+```{r subsetnconfects2}
+
 subsetnconfects <- function(mygroup, mytreatments, mycomparison){

### 2019-08-18 docs/manuscript_draft/manuscript_draft.Rmd
installed or updated packages and it works

---
title: Title of submission to PLOS journal
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliation: Some Institute of Technology
    corresponding: alice@example.com
  - name: Bob Security
    email: bob@example.com
    affiliation: Another University
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
abstract: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

author_summary: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

bibliography: mybibfile.bib
output: rticles::plos_article
csl: plos.csl
---

```{r, include=F}
## got error LaTeX Error: File `changepage.sty' not found 
## googled LaTeX Error: File `changepage.sty' not found Rmarkdown
## found this https://stackoverflow.com/questions/34252903/knit-pdf-file-from-rstudio
## ran the following

# install.packages(c('tinytex', 'rmarkdown'))
# tinytex::install_tinytex()

```

_Text based on plos sample manuscript, see [http://journals.plos.org/ploscompbiol/s/latex](http://journals.plos.org/ploscompbiol/s/latex)_

# Introduction

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

A list

- Item 1
- Item 2

Here are two sample references: @Feynman1963118 [@Dirac1953888].

# References {#references .unnumbered}

### 2019-08-18 docs/manuscript_draft/manuscript_draft.Rmd
now with summary from grant

@@
 author_summary: |
-  Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+  One of the central questions of animal behavior is how biological mechanisms mediate and respond to behavior.
+Successful rearing of offspring involves a shift from aggressive and sexual behaviors to more caring and nurturing ones, but what neural substrates mediate this transition?
+How flexible are these changes in response to unpredictable environmental perturbations, and how is behavior altered because of them?
+In a rapidly changing world, understanding how the environment affects the brain and how, in turn, the brain affects the behavioral transition into parental care will shed light on how changes in environment can ultimately affect fitness.
+Previous discoveries have implicated particular hormones that play a role in the maintenance of parental care behavior in vertebrates, such as: oxytocin and vasopressin (avian homologs: mesotocin and vasotocin), vasoactive intestinal peptide, and prolactin.
+However, we know very little about when and how the brain transitions into parental care behaviors in any vertebrate, and this knowledge is fundamentally important to our understanding of the mechanisms mediating parental care.
+Here, we propose to expand our knowledge by using classic offspring replacement and removal manipulations to uncover the role of GnIH during the transition to parental care behaviors.
+In addition to this targeted approach, we propose a powerful untargeted approach to explore what other changes in the brain may be occurring during this transition.
+We will use high-throughput sequencing technology and immunohistochemistry to uncover all genetic changes and specific proteomic changes in the brain, along with their relationship with GnIH, which occur during the transition to parental care.
+Our long-term goal is to understand the interplay of genes, proteins, and parental care.
+Specifically, we aim to understand the genetic and physiological mechanisms driving parental care behaviors in all vertebrates, and to document this knowledge in a rich, open database.
+This knowledge would enable manipulation of specific parental care behaviors and their activation using emerging powerful genetic technologies.
 
@@
 
+## Funding Statment
+
+```{r, include=F}
+## Disclosure of Funding Sources
+## All PLOS authors are required to declare what support they received to carry out their research.
+Declaring funding sources acknowledges funders’ contributions, fulfills funding requirements, and promotes greater transparency in the research process.
+## Each author must individually declare all sources of funding received for the research submitted to the journal.
+This information includes the name of granting agencies, grant numbers, and a description of each funder’s role.
+If the funder has played no role in the research, this must be stated as well.
+Authors are not required to provide the complete list of every single grant that supports them if the grant is not related to the research published.
+## A Funding Statement is included in the metadata of each published article.
+The Funding Statement includes the funding information declared by the authors.
+Inaccurate information about funding discovered after publication may require a correction.
+## from https://journals.plos.org/plosone/s/disclosure-of-funding-sources
+
+## See https://www.nsf.gov/awardsearch/showAward?AWpD_ID=1455957&HistoricalAwards=false
+```
+
+1. Collaborative Research: RUI: The Neural Basis of Becoming a Parent: From Genotype to Phenotype. NSF-IOS:1455957.
+
 # References {#references .unnumbered}

### 2019-08-18 docs/manuscript_draft/manuscript_draft.tex
now with summary from grant

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
% \bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\setcounter{secnumdepth}{0}

\newcommand{\getAff}[1]{
  \getIndex{#1}{Some Institute of Technology,Another University}
}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Title of submission to PLOS journal} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Alice Anonymous\textsuperscript{\getAff{Some Institute of Technology}}\textsuperscript{*},
Bob Security\textsuperscript{\getAff{Another University}}\\
\bigskip
\textbf{\getAff{Some Institute of Technology}}Department, Street, City, State, Zip\\
\textbf{\getAff{Another University}}Department, Street, City, State, Zip\\
\bigskip
* Corresponding author: alice@example.com\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id
velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus
finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
pulvinar, volutpat risus.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
One of the central questions of animal behavior is how biological
mechanisms mediate and respond to behavior. Successful rearing of
offspring involves a shift from aggressive and sexual behaviors to more
caring and nurturing ones, but what neural substrates mediate this
transition? How flexible are these changes in response to unpredictable
environmental perturbations, and how is behavior altered because of
them? In a rapidly changing world, understanding how the environment
affects the brain and how, in turn, the brain affects the behavioral
transition into parental care will shed light on how changes in
environment can ultimately affect fitness. Previous discoveries have
implicated particular hormones that play a role in the maintenance of
parental care behavior in vertebrates, such as: oxytocin and vasopressin
(avian homologs: mesotocin and vasotocin), vasoactive intestinal
peptide, and prolactin. However, we know very little about when and how
the brain transitions into parental care behaviors in any vertebrate,
and this knowledge is fundamentally important to our understanding of
the mechanisms mediating parental care. Here, we propose to expand our
knowledge by using classic offspring replacement and removal
manipulations to uncover the role of GnIH during the transition to
parental care behaviors. In addition to this targeted approach, we
propose a powerful untargeted approach to explore what other changes in
the brain may be occurring during this transition. We will use
high-throughput sequencing technology and immunohistochemistry to
uncover all genetic changes and specific proteomic changes in the brain,
along with their relationship with GnIH, which occur during the
transition to parental care. Our long-term goal is to understand the
interplay of genes, proteins, and parental care. Specifically, we aim to
understand the genetic and physiological mechanisms driving parental
care behaviors in all vertebrates, and to document this knowledge in a
rich, open database. This knowledge would enable manipulation of
specific parental care behaviors and their activation using emerging
powerful genetic technologies.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\emph{Text based on plos sample manuscript, see
\url{http://journals.plos.org/ploscompbiol/s/latex}}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
dui eu ante cursus gravida non sed sem. Phasellus mauris velit, dapibus
finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
pulvinar, volutpat risus.

A list

\begin{itemize}
\tightlist
\item
  Item 1
\item
  Item 2
\end{itemize}

Here are two sample references: {[}1,2{]}.

\hypertarget{funding-statment}{%
\subsection{Funding Statment}\label{funding-statment}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Collaborative Research: RUI: The Neural Basis of Becoming a Parent:
  From Genotype to Phenotype. NSF-IOS:1455957.
\end{enumerate}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Feynman1963118}{}%
1. Feynman R, Vernon Jr. F. The theory of a general quantum system
interacting with a linear dissipative system. Annals of Physics.
1963;24: 118--173.
doi:\href{https://doi.org/10.1016/0003-4916(63)90068-X}{10.1016/0003-4916(63)90068-X}

\leavevmode\hypertarget{ref-Dirac1953888}{}%
2. Dirac P. The lorentz transformation and absolute time. Physica.
1953;19: 888--896.
doi:\href{https://doi.org/10.1016/S0031-8914(53)80099-6}{10.1016/S0031-8914(53)80099-6}

\nolinenumbers

\end{document}

### 2019-08-18 docs/manuscript_draft/manuscript_draft.tex
move images

@@
 % Use "Eq" instead of "Equation" for equation citations.
-\emph{Text based on plos sample manuscript, see
-\url{http://journals.plos.org/ploscompbiol/s/latex}}
-
 \hypertarget{introduction}{%
 \section{Introduction}\label{introduction}}
+
+Understanding the mechanisms underlying parental care are critical to
+circumventing issues with parent-newborn bonding as well, where ultimate
+explanations are obvious, but specific mechanisms remain elusive.
+
+The rock dove (\emph{Columba livia}) is an ideal system to characterize
+changes in genetic expression during parental care transitions because:
+1) ample genomic resources are available, including a complete annotated
+genome assembly (Gillespie et al.~2013; Shapiro et al.~2013) and
+methodology concerning reproductive physiology and behavior (Dong et
+al.~2012); and 2) rock doves are prolific, year-round breeders that
+thrive in captivity, making observation, manipulation, and sampling
+highly feasible year-round.
+
+\includegraphics[width=360px]{../../figures/images/PigeonHPGaxis}
+
+Rock doves are socially monogamous and offer bi-parental care, making
+inter- and intra-sexual comparisons possible. Birds offer two important
+behavioral transition points into parental care: the incubation of eggs
+and the caring for chicks. This produces two unique opportunities to
+study how the brain transitions into two different suites of parental
+care behaviors. Additionally, rock doves exhibit a parental care
+strategy analogous to mammals in that they, too, `lactate' to feed their
+young (Gillespie et al.~2011, 2012). This lactation, unlike simple
+regurgitation of food, consists of the production and sloughing off of
+skin cells inside the crop sac of females and males, creating a
+protein-rich milk-like substance on which they rear their chicks. Many
+functional similarities between rock dove and mammalian lactation exist
+concerning the mediation of this event by the hormone prolactin (Dumont
+1965). Additionally, like mammalian milk, rock dove milk delivers
+essential immunoglobulins and nutritional benefits to young, aiding in
+their immune function and development of microbiota (Gillespie et
+al.~2012). Thus, because rock doves incubate eggs and exhibit
+mammalian-like mediation and function of lactation for young, they have
+the potential to serve as a powerful theoretical bridge to understand
+the neurobiology of both avian and mammalian transitions into parental
+care.
+
+\hypertarget{materials-and-methods}{%
+\section{Materials and Methods}\label{materials-and-methods}}
 
@@
 
-A list
+\includegraphics[width=360px]{../../figures/images/samplingtimepoints}
 
-\begin{itemize}
-\tightlist
-\item
-  Item 1
-\item
-  Item 2
-\end{itemize}
+\hypertarget{elucidate-cause-of-transcription-protein-changes-during-parental-care-transitions}{%
+\subsection{Elucidate cause of transcription \& protein changes during
+parental care
+transitions}\label{elucidate-cause-of-transcription-protein-changes-during-parental-care-transitions}}
+
+Using egg and chick removal and replacement experiments we will
+determine if neural transcription and translation during parental care
+transitions are based on external sensory information (the
+presence/absence of eggs or chicks) or on an internal clock-timing
+mechanism (natural biological rhythms unaffected by changes in external
+sensory information). As rock doves are known to cross-foster eggs and
+chicks, our working hypothesis is that the transition to parental care
+behaviors is based on external sensory input, and specific predictions
+are explained in conjunction with each manipulation below.
+
+\hypertarget{give-birds-eggs}{%
+\subsubsection{Give birds eggs}\label{give-birds-eggs}}
+
+We will place eggs in nests of birds lacking them and sample brains the
+following day. Prediction: birds will transition into incubation
+behavior, and neural transcription and translation will mirror that of
+birds sampled on the day their first egg was laid. This would support
+our working hypothesis that sensory perception of the environment drives
+these changes. If we do not observe changes in behavior, transcription,
+and translation, or they only occur in males, one of many
+interpretations is that males respond to environmental cues while
+females are cued by an internal clock-timing mechanism regulating
+transitions into parental care.
+
+\hypertarget{prolong-incubation}{%
+\subsubsection{Prolong incubation}\label{prolong-incubation}}
+
+We will prolong incubation by removing eggs from actively incubated
+nests on the day before hatching (Day 17) and immediately replace them
+with infertile, `dummy' eggs. We will then sample brains 3 days after
+normal hatch time (Day 21). Collaborator Silver and colleagues have
+shown sex differences in the endocrine profiles of male and female doves
+in response to prolonged incubation, in that prolactin levels are
+maintained in incubating females but not males (Ramsey et al.~1985).
+Because of this, we predict that if changes in behavior, neural
+transcription, and translation occur in females because of sensory
+information related to chicks hatching (supporting our working
+hypothesis), then females that experience this prolonged incubation will
+mirror the changes of incubating birds sampled on Day 17 of incubation.
+However, if changes in behavior, transcription and translation are
+brought on by internal clock mechanisms rather than external sensory
+information, as predicted at this stage in males, bird behavior and
+biology will resemble that of birds collected on Day 3 post-hatching.
+
+\hypertarget{extend-time-of-hatching}{%
+\subsubsection{Extend time of hatching}\label{extend-time-of-hatching}}
+
+We will prolong incubation by removing eggs from nests undergoing active
+incubation on Day 17 (the end of incubation) and immediately replacing
+eggs with `dummy' eggs. On Day 21 (3 days post-natural hatch time) we
+will offer hatchlings. Doves are known to care for nestlings placed in
+their nest that are not their own (Klinghammer and Hess 1964, Hasen
+1971). Brains will be sampled the following day. Prediction: Behavior,
+transcription and translation will be similar to birds naturally caring
+for chick(s) one day post-hatch of the first chick, supporting our
+working hypothesis. Alternatively, if behavior, transcription and
+translation are similar to birds caring for hatchlings on Day 3
+post-hatching, which would have been the normal time course for parental
+care if chicks had hatched naturally, this would support regulation by
+an internal clock mechanism.
+
+\hypertarget{remove-eggs}{%
+\subsubsection{Remove eggs}\label{remove-eggs}}
+
+We will remove eggs during incubation on Day 1 (beginning), Day 9
+(middle), or Day 17 (end) and sample brains on the following day. Note:
+This is not a repeated measures sampling, and three groups of
+male-female pairs will be independently sampled at each time point.
+Prediction: Birds will revert to a behavioral, transcriptional and
+translational pattern characteristic of the ``Nesting, prior to lay''
+sampling point, lending support to our working hypothesis. If birds
+maintain a profile characteristic of birds collected at similar time
+points whose eggs have not been removed, this lends support to an
+internal clock mechanism.
+
+\hypertarget{reduce-time-to-hatching}{%
+\subsubsection{Reduce time to hatching}\label{reduce-time-to-hatching}}
+
+We will remove eggs on Day 8 (middle) of incubation and immediately
+replace them with newly hatched chicks. We will sample brains on the
+following day (Day 9). Prediction: Changes in behavior, transcription
+and translation are externally sensory driven, and these will be similar
+to birds sampled on `Day 1: 1st egg hatches, supporting our working
+hypothesis. Alternatively, if profiles remain similar to birds sampled
+on Day 9 that are actively incubating eggs, this would support the
+presence of an internal clock mechanism.
+
+\hypertarget{remove-chicks}{%
+\subsubsection{Remove chicks}\label{remove-chicks}}
+
+We will remove newly, naturally hatched chick(s) on Day 2: second chick
+hatches and sample brains from parents the following day (Day 2 post
+(second) egg hatching). Prediction: Behavior, transcription and
+translation are sensory driven and will revert to a pattern
+characteristic of the `Nesting, prior to lay' point, supporting our
+working hypothesis. Alternatively, if an internal clock mechanism is at
+play, we would expect to see a pattern similar to birds caring for
+chicks on Day 2 post (second) egg hatching.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
 Here are two sample references: {[}1,2{]}.
+
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
+
+This project is a synergistic collaboration between the PI, Rebecca
+Calisi-Rodríguez (expertise in avian behavior, parental care and
+neurobiology), co-PI, Matthew MacManes (expertise in next-generation
+sequencing, transcriptome assembly, and gene expression analyses), and
+Collaborator, Rae Silver (expertise in neurobiology, dove behavior, and
+decades of successful breeding and maintenance of dove colonies at
+Barnard College).
+
+\hypertarget{metadata-and-extras}{%
+\section{Metadata and extras}\label{metadata-and-extras}}
 
@@
 \end{enumerate}
+
+\hypertarget{about-the-authors}{%
+\subsection{About the authors}\label{about-the-authors}}
+
+Roles, Affiliations, Orcids, competing interests
 

### 2019-08-18 docs/manuscript_draft/manuscript_draft.Rmd
move images

@@
 
+<!--
 _Text based on plos sample manuscript, see [http://journals.plos.org/ploscompbiol/s/latex](http://journals.plos.org/ploscompbiol/s/latex)_
+-->
 
 # Introduction
+
+Understanding the mechanisms underlying parental care are critical to circumventing issues with parent-newborn bonding as well, where ultimate explanations are obvious, but specific mechanisms remain elusive.
+
+The rock dove (_Columba livia_) is an ideal system to characterize changes in genetic expression during parental care transitions because: 1) ample genomic resources are available, including a complete annotated genome assembly (Gillespie et al.
+2013; Shapiro et al.
+2013) and methodology concerning reproductive physiology and behavior (Dong et al.
+2012); and 2) rock doves are prolific, year-round breeders that thrive in captivity, making observation, manipulation, and sampling highly feasible year-round.
+
+```{r, out.width = "360px", echo = F}
+knitr::include_graphics("../../figures/images/PigeonHPGaxis.png")
+```
+
+Rock doves are socially monogamous and offer bi-parental care, making inter- and intra-sexual comparisons possible.
+Birds offer two important behavioral transition points into parental care: the incubation of eggs and the caring for chicks.
+This produces two unique opportunities to study how the brain transitions into two different suites of parental care behaviors.
+Additionally, rock doves exhibit a parental care strategy analogous to mammals in that they, too, ‘lactate’ to feed their young (Gillespie et al.
+2011, 2012).
+This lactation, unlike simple regurgitation of food, consists of the production and sloughing off of skin cells inside the crop sac of females and males, creating a protein-rich milk-like substance on which they rear their chicks.
+Many functional similarities between rock dove and mammalian lactation exist concerning the mediation of this event by the hormone prolactin (Dumont 1965).
+Additionally, like mammalian milk, rock dove milk delivers essential immunoglobulins and nutritional benefits to young, aiding in their immune function and development of microbiota (Gillespie et al.
+2012).
+Thus, because rock doves incubate eggs and exhibit mammalian-like mediation and function of lactation for young, they have the potential to serve as a powerful theoretical bridge to understand the neurobiology of both avian and mammalian transitions into parental care.
+
+# Materials and Methods
 
@@
 
-A list
+```{r, out.width = "360px", echo = F}
+knitr::include_graphics("../../figures/images/samplingtimepoints.png")
+```
 
-- Item 1
-- Item 2
+## Elucidate cause of transcription & protein changes during parental care transitions
+
+Using egg and chick removal and replacement experiments we will determine if neural transcription and translation during parental care transitions are based on external sensory information (the presence/absence of eggs or chicks) or on an internal clock-timing mechanism (natural biological rhythms unaffected by changes in external sensory information).
+As rock doves are known to cross-foster eggs and chicks, our working hypothesis is that the transition to parental care behaviors is based on external sensory input, and specific predictions are explained in conjunction with each manipulation below.
+
+### Give birds eggs
+We will place eggs in nests of birds lacking them and sample brains the following day.
+Prediction: birds will transition into incubation behavior, and neural transcription and translation will mirror that of birds sampled on the day their first egg was laid.
+This would support our working hypothesis that sensory perception of the environment drives these changes.
+If we do not observe changes in behavior, transcription, and translation, or they only occur in males, one of many interpretations is that males respond to environmental cues while females are cued by an internal clock-timing mechanism regulating transitions into parental care.
+
+### Prolong incubation
+We will prolong incubation by removing eggs from actively incubated nests on the day before hatching (Day 17) and immediately replace them with infertile, ‘dummy’ eggs.
+We will then sample brains 3 days after normal hatch time (Day 21).
+Collaborator Silver and colleagues have shown sex differences in the endocrine profiles of male and female doves in response to prolonged incubation, in that prolactin levels are maintained in incubating females but not males (Ramsey et al.
+1985).
+Because of this, we predict that if changes in behavior, neural transcription, and translation occur in females because of sensory information related to chicks hatching (supporting our working hypothesis), then females that experience this prolonged incubation will mirror the changes of incubating birds sampled on Day 17 of incubation.
+However, if changes in behavior, transcription and translation are brought on by internal clock mechanisms rather than external sensory information, as predicted at this stage in males, bird behavior and biology will resemble that of birds collected on Day 3 post-hatching.
+
+### Extend time of hatching
+We will prolong incubation by removing eggs from nests undergoing active incubation on Day 17 (the end of incubation) and immediately replacing eggs with ‘dummy’ eggs.
+On Day 21 (3 days post-natural hatch time) we will offer hatchlings.
+Doves are known to care for nestlings placed in their nest that are not their own (Klinghammer and Hess 1964, Hasen 1971).
+Brains will be sampled the following day.
+Prediction: Behavior, transcription and translation will be similar to birds naturally caring for chick(s) one day post-hatch of the first chick, supporting our working hypothesis.
+Alternatively, if behavior, transcription and translation are similar to birds caring for hatchlings on Day 3 post-hatching, which would have been the normal time course for parental care if chicks had hatched naturally, this would support regulation by an internal clock mechanism.
+
+### Remove eggs
+We will remove eggs during incubation on Day 1 (beginning), Day 9 (middle), or Day 17 (end) and sample brains on the following day.
+Note: This is not a repeated measures sampling, and three groups of male-female pairs will be independently sampled at each time point.
+Prediction: Birds will revert to a behavioral, transcriptional and translational pattern characteristic of the “Nesting, prior to lay” sampling point, lending support to our working hypothesis.
+If birds maintain a profile characteristic of birds collected at similar time points whose eggs have not been removed, this lends support to an internal clock mechanism.
+
+### Reduce time to hatching
+We will remove eggs on Day 8 (middle) of incubation and immediately replace them with newly hatched chicks.
+We will sample brains on the following day (Day 9).
+Prediction: Changes in behavior, transcription and translation are externally sensory driven, and these will be similar to birds sampled on ‘Day 1: 1st egg
+hatches, supporting our working hypothesis.
+Alternatively, if profiles remain similar to birds sampled on Day 9 that are actively incubating eggs, this would support the presence of an internal clock mechanism.
+
+### Remove chicks
+We will remove newly, naturally hatched chick(s) on Day 2: second chick hatches and sample brains from parents the following day (Day 2 post (second) egg hatching).
+Prediction: Behavior, transcription and translation are sensory driven and will revert to a pattern characteristic of the ‘Nesting, prior to lay’ point, supporting our working hypothesis.
+Alternatively, if an internal clock mechanism is at play, we would expect to see a pattern similar to birds caring for chicks on Day 2 post (second) egg hatching.
+
+# Results
+
+# Discussion
 
 Here are two sample references: @Feynman1963118 [@Dirac1953888].
+
+# Acknowledgments
+
+This project is a synergistic collaboration between the PI, Rebecca Calisi-Rodríguez (expertise in avian behavior, parental care and neurobiology), co-PI, Matthew MacManes (expertise in next-generation sequencing, transcriptome assembly, and gene expression analyses), and Collaborator, Rae Silver (expertise in neurobiology, dove behavior, and decades of successful breeding and maintenance of dove colonies at Barnard College).
+
+# Metadata and extras
 
@@
 
+## About the authors
+
+Roles, Affiliations, Orcids, competing interests
+
 # References {#references .unnumbered}

### 2019-08-21 analysis/08_WGCNA_1.R
working now

@@ -40,9 +40,10 @@ gsg = goodSamplesGenes(datExpr0, verbose = 3);
 gsg$allOK
 
 
+
 #=====================================================================================
 #
-#  Code chunk 4
+#  Code chunk 4 removing genes
 #
 #=====================================================================================
 
@@ -75,12 +76,13 @@ if (!gsg$allOK)
 sampleTree = hclust(dist(datExpr0), method = "average");
 # Plot the sample tree: Open a graphic output window of size 12 by 9 inches
 # The user should change the dimensions if the window is too large or too small.
-sizeGrWindow(12,9)
-#pdf(file = "Plots/sampleClustering.pdf", width = 12, height = 9);
+pdf(file = "../figures/wgcna/sampleClustering.pdf", width = 12, height = 9);
 par(cex = 0.6);
 par(mar = c(0,4,2,0))
 plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5, 
      cex.axis = 1.5, cex.main = 2)
+abline(h = 2000000, col = "red");
+dev.off();
 
 
 #=====================================================================================
@@ -91,7 +93,7 @@ plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="",
 
 
 # Plot a line to show the cut
-abline(h = 20000, col = "red");
+
 # Determine cluster under the line
 clust = cutreeStatic(sampleTree, cutHeight = 2000000, minSize = 10)
 table(clust)
@@ -118,35 +120,38 @@ traitData$tissue <- factor(traitData$tissue,
                               levels =c("hypothalamus", "pituitary", "gonad"))
 
 
-
-# solution
+# numeric
 str(traitData)
 traitData %<>% mutate_if(is.factor,as.numeric)
 str(traitData)
 
+# remove columns or add that hold information we do not need, if necessary
 
-# remove columns that hold information we do not need, if necessary
 allTraits <- traitData
-row.names(allTraits) <- colData$V1
-allTraits$samples <- colData$V1
-
-head(allTraits)
-
+allTraits$V1 <- colData$V1
+row.names(allTraits) <- allTraits$V1
 # subset to keep only the "good samples"
 Samples <- rownames(datExpr)
-traitRows <-  match(Samples, allTraits$samples)
+traitRows <-  match(Samples, allTraits$V1)
 datTraits <-  allTraits[traitRows, ]
-rownames(datTraits) <-  allTraits[traitRows, 6];  #using column 5 as row name
-datTraits$samples <- NULL
[…]

### 2019-08-22 analysis/08_wgcna.Rmd
now with markdown output

---
title: "wgcna"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE, fig.path = '../figures/wgcna/')

```

## WGCNA

```{r wgcna, message=FALSE, warning=FALSE, cache=TRUE}
source("08_WGCNA_1.R")

```

### 2019-08-22 analysis/08_WGCNA_1.R

@@ -36,7 +36,7 @@ head(rownames(datExpr0)) # rows are samples
 #=====================================================================================
 
 
-gsg = goodSamplesGenes(datExpr0, verbose = 3);
+gsg = goodSamplesGenes(datExpr0, verbose = 4);
 gsg$allOK
 
 
@@ -114,16 +114,18 @@ nSamples = nrow(datExpr)
 
 traitData <- colData[c(2:6)]
 
-traitData$treatment <- factor(traitData$treatment, 
-                              levels =c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
-traitData$tissue <- factor(traitData$tissue, 
-                              levels =c("hypothalamus", "pituitary", "gonad"))
+#traitData$treatment <- factor(traitData$treatment, 
+#                              levels =c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
+#traitData$tissue <- factor(traitData$tissue, 
+ #                             levels =c("hypothalamus", "pituitary", "gonad"))
 
 
 # numeric
 str(traitData)
 traitData %<>% mutate_if(is.factor,as.numeric)
-str(traitData)
+head(traitData)
+
+
 
 # remove columns or add that hold information we do not need, if necessary
 
@@ -134,24 +136,13 @@ row.names(allTraits) <- allTraits$V1
 Samples <- rownames(datExpr)
 traitRows <-  match(Samples, allTraits$V1)
 datTraits <-  allTraits[traitRows, ]
-head(datTraits)
+datTraits$V1 <- NULL
 
 
 collectGarbage()
 
 
-# create new, slim label
-datTraits2 <- inner_join(datTraits, colData, by = "V1")
-head(datTraits2)
-datTraits2$group_bird <- paste(datTraits2$group.y, datTraits2$bird.x, sep = ".")
-head(datTraits2$group_bird)
-sampleTree$labels <- datTraits2$group_bird
-head(sampleTree$labels)
 
-## remove non-numeric columns
-datTraits <- datTraits[c(1:5)]
-row.names(datTraits) <- datTraits2$group_bird
-head(datTraits)
 
 
 # Re-cluster samples
@@ -161,7 +152,8 @@ traitColors = numbers2colors(datTraits, signed = FALSE);
 # Plot the sample dendrogram and the colors underneath.
 plotDendroAndColors(sampleTree2, traitColors,
                     groupLabels = names(datTraits), 
[…]

### 2019-08-22 analysis/08_wgcna.Rmd
now with subsets

@@
 
-```{r setup, include=TRUE}
+```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = FALSE, fig.path = '../figures/wgcna/')
@@
 
-## WGCNA
+## All characterization data
+
+This first image was made using [`08_WGCNA_1.R`](https://github.com/macmanes-lab/DoveParentsRNAseq/blob/master/analysis/08_WGCNA_1.R), which was modified from the [first in a series of WGCNA tutorial](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/Consensus-DataInput.R).
+The first tutorial focuses on data input, cleaning and pre-processing; outlier removal; clustering of samples by their Euclidean distance; a comparison of sample cluster to sample meta data.
 
@@
 ```
+
+## hypothalas
+
+```{r wgcna-hyp, message=FALSE, warning=FALSE, cache=TRUE}
+source("08_WGCNA_1_params.R")
+subsetWGCNA("hypothalamus")
+
+```
+
+## pituitary
+
+```{r wgcna-pit, message=FALSE, warning=FALSE, cache=TRUE}
+
+subsetWGCNA("pituitary")
+```
+
+## gonads
+
+```{r wgcna-gon, message=FALSE, warning=FALSE, cache=TRUE}
+subsetWGCNA("gonad")
+```

### 2019-08-22 analysis/08_WGCNA_1_params.R
now with subsets

@@ -12,9 +12,28 @@ library(WGCNA)
 library(dplyr)
 library(magrittr)
 
+subsetWGCNA <- function(eachgroup){
+
 #Read in the data set
 colData <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, stringsAsFactors = T)
 countData <- read.csv("../results/00_countData_characterization.csv", row.names = 1)
+
+colData$sextissue <-  paste(colData$sex, colData$tissue, sep = ".")
+
+# subset to look within one tissue in one sex
+colData <- colData %>%
+  dplyr::filter(tissue == eachgroup) %>%
+  droplevels()
+row.names(colData) <- colData$V1
+head(colData)
+  
+# which counts to save
+savecols <- as.character(colData$V1) 
+savecols <- as.vector(savecols) 
+  
+# save counts that match colData
+countDatatemp <- countData %>% dplyr::select(one_of(savecols)) 
+  
 
 #=====================================================================================
 #
@@ -23,7 +42,7 @@ countData <- read.csv("../results/00_countData_characterization.csv", row.names 
 #=====================================================================================
 
 
-datExpr0 <- as.data.frame(t(countData))
+datExpr0 <- as.data.frame(t(countDatatemp))
 
 head(names(datExpr0))  # columns are genes
 head(rownames(datExpr0)) # rows are samples
@@ -37,7 +56,7 @@ head(rownames(datExpr0)) # rows are samples
 
 
 gsg = goodSamplesGenes(datExpr0, verbose = 0);
-#gsg$allOK
+gsg$allOK
 
 
 
@@ -51,20 +70,14 @@ gsg = goodSamplesGenes(datExpr0, verbose = 0);
 if (!gsg$allOK)
 {
   # Optionally, print the gene and sample names that were removed:
- # if (sum(!gsg$goodGenes)>0) 
+  #if (sum(!gsg$goodGenes)>0) 
   #  printFlush(paste("Removing genes:", paste(names(datExpr0)[!gsg$goodGenes], collapse = ", ")));
- # if (sum(!gsg$goodSamples)>0) 
- #   printFlush(paste("Removing samples:", paste(rownames(datExpr0)[!gsg$goodSamples], collapse = ", ")));
+  #if (sum(!gsg$goodSamples)>0) 
+  #  printFlush(paste("Removing samples:", paste(rownames(datExpr0)[!gsg$goodSamples], collapse = ", ")));
[…]

### 2019-08-22 docs/manuscript_draft.Rmd
update temp ms

--- 2019-08-18 docs/manuscript_draft/manuscript_draft.Rmd
@@
   Lorem ipsum dolor sit amet, consectetur adipiscing elit. 
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
-
 author_summary: |
@@
 ```{r, out.width = "360px", echo = F}
-knitr::include_graphics("../../figures/images/PigeonHPGaxis.png")
+knitr::include_graphics("../figures/images/PigeonHPGaxis.png")
 ```
@@
 ```{r, out.width = "360px", echo = F}
-knitr::include_graphics("../../figures/images/samplingtimepoints.png")
+knitr::include_graphics("../figures/images/samplingtimepoints.png")
 ```
@@
 
-# Discussion
+Here are two sample references: @Feynman1963118 [@Dirac1953888].
 
-Here are two sample references: @Feynman1963118 [@Dirac1953888].
+## WGCNA characterization
+
+This first image was made using [`08_WGCNA_1.R`](https://github.com/macmanes-lab/DoveParentsRNAseq/blob/master/analysis/08_WGCNA_1.R), which was modified from the [first in a series of WGCNA tutorial](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/Consensus-DataInput.R).
+The first tutorial focuses on data input, cleaning and pre-processing; outlier removal; clustering of samples by their Euclidean distance; a comparison of sample cluster to sample meta data.
+
+```{r wgcna, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+source("../analysis/08_WGCNA_1.R")
+
+```
 
@@
 
+```{r include=F}
 ## About the authors
 
-Roles, Affiliations, Orcids, competing interests
+#Roles, Affiliations, Orcids, competing interests
+```
 

### 2019-08-22 docs/manuscript_draft.tex
update temp ms

--- 2019-08-18 docs/manuscript_draft/manuscript_draft.tex
@@
 \section*{Abstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
-porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
-dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id
-velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus
-finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
-diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
-ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
-pulvinar, volutpat risus.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 
@@
 
-\includegraphics[width=360px]{../../figures/images/PigeonHPGaxis}
+\includegraphics[width=360px]{../figures/images/PigeonHPGaxis}
 
@@
 
-\includegraphics[width=360px]{../../figures/images/samplingtimepoints}
+\includegraphics[width=360px]{../figures/images/samplingtimepoints}
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+Here are two sample references: {[}1,2{]}.
 
-Here are two sample references: {[}1,2{]}.
+\hypertarget{wgcna-characterization}{%
+\subsection{WGCNA characterization}\label{wgcna-characterization}}
+
+This first image was made using
+\href{https://github.com/macmanes-lab/DoveParentsRNAseq/blob/master/analysis/08_WGCNA_1.R}{\texttt{08\_WGCNA\_1.R}},
+which was modified from the
+\href{https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/Consensus-DataInput.R}{first
+in a series of WGCNA tutorial}. The first tutorial focuses on data
+input, cleaning and pre-processing; outlier removal; clustering of
+samples by their Euclidean distance; a comparison of sample cluster to
+sample meta data.
+
+\includegraphics{manuscript_draft_files/figure-latex/wgcna-1.pdf}
 
@@
 
-\hypertarget{about-the-authors}{%
-\subsection{About the authors}\label{about-the-authors}}
-
-Roles, Affiliations, Orcids, competing interests
-
 \hypertarget{references}{%

### 2019-08-23 analysis/08_wgcna.Rmd
now with markdown
multiple wgcna plots

@@
 title: "wgcna"
-output: html_document
+output: md_document
 ---
@@
 
-## hypothalas
+## Females
+
+```{r wgcna-female, message=FALSE, warning=FALSE, cache=TRUE}
+source("08_WGCNA_1_params.R")
+subsetWGCNA(c("female.gonad", "female.hypothalamus", "female.pituitary"))
+
+```
+
+## Males
+
+```{r wgcna-male, message=FALSE, warning=FALSE, cache=TRUE}
+source("08_WGCNA_1_params.R")
+subsetWGCNA(c("male.gonad", "male.hypothalamus", "male.pituitary"))
+```
+
+## hypothalamus
 
@@
 source("08_WGCNA_1_params.R")
-subsetWGCNA("hypothalamus")
-
+subsetWGCNA(c("male.hypothalamus", "female.hypothalamus"))
 ```
@@
 ```{r wgcna-pit, message=FALSE, warning=FALSE, cache=TRUE}
-
-subsetWGCNA("pituitary")
+#subsetWGCNA(c("female.pitutary", "male.pitutary"))
 ```
@@
 ```{r wgcna-gon, message=FALSE, warning=FALSE, cache=TRUE}
-subsetWGCNA("gonad")
+subsetWGCNA(c("female.gonad", "male.gonad"))
 ```

### 2019-08-23 analysis/08_wgcna.Rmd
smaller font

@@
 subsetWGCNA(c("male.gonad", "male.hypothalamus", "male.pituitary"))
+
 ```
@@
 subsetWGCNA(c("male.hypothalamus", "female.hypothalamus"))
+
 ```
@@
 subsetWGCNA(c("female.gonad", "male.gonad"))
+
 ```

### 2019-08-23 analysis/08_wgcna.Rmd
now with 1 function

--- 2019-08-22 analysis/08_wgcna.Rmd
@@
 
+source("08_WGCNA_1.R")
 ```
@@
 ```{r wgcna, message=FALSE, warning=FALSE, cache=TRUE}
-source("08_WGCNA_1.R")
+
+subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary",
+                           "male.gonad", "male.hypothalamus", "male.pituitary"),
+            mytitle = "Characterization dendrogram with trait heatmap")
 
@@
 
-## hypothalas
+## Females
+
+```{r wgcna-female, message=FALSE, warning=FALSE, cache=TRUE}
+subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary"),
+            mytitle = "Female sample dendrogram with trait heatmap")
+
+```
+
+## Males
+
+```{r wgcna-male, message=FALSE, warning=FALSE, cache=TRUE}
+subsetWGCNA(c("male.gonad", "male.hypothalamus", "male.pituitary"),
+            mytitle = "Male sample dendrogram with trait heatmap")
+
+```
+
+## hypothalamus
 
 ```{r wgcna-hyp, message=FALSE, warning=FALSE, cache=TRUE}
-source("08_WGCNA_1_params.R")
-subsetWGCNA("hypothalamus")
+subsetWGCNA(c("male.hypothalamus", "female.hypothalamus"),
+            mytitle = "Hypothalamus dendrogram with trait heatmap")
 
@@
 ```{r wgcna-pit, message=FALSE, warning=FALSE, cache=TRUE}
-
-subsetWGCNA("pituitary")
+#subsetWGCNA(c("female.pitutary", "male.pitutary"),
+#            mytitle = "Pituitary dendrogram with trait heatmap")
 ```
@@
 ```{r wgcna-gon, message=FALSE, warning=FALSE, cache=TRUE}
-subsetWGCNA("gonad")
+subsetWGCNA(c("female.gonad", "male.gonad"),
+            mytitle = "Gonad dendrogram with trait heatmap")
+
 ```

### 2019-08-23 analysis/08_wgcna.Rmd
documentation

--- 2019-08-23 analysis/08_wgcna.Rmd
@@
 
+For the following figures, the top panel is a dendrogram showing Euclidean distance of sample.
+Because the labeling of samples is hard to read, the colored rows can be used to visualize traits.
+Here, both sex and tissues are colored in a white to red scaled with the following pattern:
+
+* sex: females (white), males (red), 
+* tissue: gonad (white), hypothalamus (pink), pituitary (red).
+
 ```{r wgcna, message=FALSE, warning=FALSE, cache=TRUE}
-source("08_WGCNA_1.R")
 
+subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary",
+                           "male.gonad", "male.hypothalamus", "male.pituitary"),
+            mytitle = "Characterization dendrogram with trait heatmap")
 ```
+
+**Fig 1. Characterization dendrogram with trait heatmap.**. 
+ Most hyp. and pit. samples separte form the gon. samples, but some are clusters contain samples from all tissues.
 
@@
 ```{r wgcna-female, message=FALSE, warning=FALSE, cache=TRUE}
-source("08_WGCNA_1_params.R")
-subsetWGCNA(c("female.gonad", "female.hypothalamus", "female.pituitary"))
+subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary"),
+            mytitle = "Female sample dendrogram with trait heatmap")
+```
 
-```
+**Fig 2. Female characterization dendrogram with trait heatmap.**. 
+Most pit samples form a single cluster. More hyp samples cluster with gonads than with pituitary.
 
@@
 ```{r wgcna-male, message=FALSE, warning=FALSE, cache=TRUE}
-source("08_WGCNA_1_params.R")
-subsetWGCNA(c("male.gonad", "male.hypothalamus", "male.pituitary"))
+subsetWGCNA(c("male.gonad", "male.hypothalamus", "male.pituitary"),
+            mytitle = "Male sample dendrogram with trait heatmap")
+```
 
+**Fig 3. Male characterization dendrogram with trait heatmap.**. 
+ As with females, most pit samples form a single cluster, and More hyp samples cluster with gonads than with pituitary.
+
+## gonads
+
+```{r wgcna-gon, message=FALSE, warning=FALSE, cache=TRUE}
+subsetWGCNA(c("female.gonad", "male.gonad"),
+            mytitle = "Gonad dendrogram with trait heatmap")
 ```
+
+**Fig 3. Gonad characterization dendrogram with trait heatmap.**. 
+Male and female samples cluster together, but with large, sex-specific regions.... maybe there is a strong treatment effect here.
 
@@
 ```{r wgcna-hyp, message=FALSE, warning=FALSE, cache=TRUE}
-source("08_WGCNA_1_params.R")
-subsetWGCNA(c("male.hypothalamus", "female.hypothalamus"))
-
+subsetWGCNA(c("male.hypothalamus", "female.hypothalamus"),
+            mytitle = "Hypothalamus dendrogram with trait heatmap")
 ```
 
-## pituitary
+Male and female samples cluster together, but with very small, sex-specific regions.... maybe there is a small treatment effect here.
+
+## pituitary (not working)
 
 ```{r wgcna-pit, message=FALSE, warning=FALSE, cache=TRUE}
-#subsetWGCNA(c("female.pitutary", "male.pitutary"))
+
+#subsetWGCNA(c("female.pitutary", "male.pitutary"),
+#            mytitle = "Pituitary dendrogram with trait heatmap")
 ```
-
-## gonads
-
-```{r wgcna-gon, message=FALSE, warning=FALSE, cache=TRUE}
-subsetWGCNA(c("female.gonad", "male.gonad"))
-
-```

### 2019-08-23 analysis/08_wgcna.Rmd
documentation

@@
 **Fig 1. Characterization dendrogram with trait heatmap.**. 
- Most hyp. and pit. samples separte form the gon. samples, but some are clusters contain samples from all tissues.
+ Most hypothalamus and pituitary samples separte form the gonad samples, but some are clusters contain samples from all tissues.
 
@@
 **Fig 2. Female characterization dendrogram with trait heatmap.**. 
-Most pit samples form a single cluster. More hyp samples cluster with gonads than with pituitary.
+Most pituitary samples form a single cluster. More hypothalamus samples cluster with gonads than with pituitary.
 
@@
 **Fig 3. Male characterization dendrogram with trait heatmap.**. 
- As with females, most pit samples form a single cluster, and More hyp samples cluster with gonads than with pituitary.
+ As with females, most pituitary samples form a single cluster, while more hypothalamus samples cluster with gonads than with pituitary.
 

### 2019-08-23 analysis/08_wgcna.Rmd
intro

@@
 
-## All characterization data
+# WGCNA analysis, first pass
 
-This first image was made using [`08_WGCNA_1.R`](https://github.com/macmanes-lab/DoveParentsRNAseq/blob/master/analysis/08_WGCNA_1.R), which was modified from the [first in a series of WGCNA tutorial](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/Consensus-DataInput.R).
+The anlayses and image was made by following along with and modifing the [first in a series of WGCNA tutorial](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/Consensus-DataInput.R).
 The first tutorial focuses on data input, cleaning and pre-processing; outlier removal; clustering of samples by their Euclidean distance; a comparison of sample cluster to sample meta data.
+I put pretty much the whole script into a single function that I source with [`08_WGCNA_1.R`](https://github.com/macmanes-lab/DoveParentsRNAseq/blob/master/analysis/08_WGCNA_1.R).
 
@@
 * tissue: gonad (white), hypothalamus (pink), pituitary (red).
+
+## All characterization data
 

### 2019-08-23 docs/manuscript_draft.Rmd
new figs

@@
 
-## WGCNA characterization
-
-This first image was made using [`08_WGCNA_1.R`](https://github.com/macmanes-lab/DoveParentsRNAseq/blob/master/analysis/08_WGCNA_1.R), which was modified from the [first in a series of WGCNA tutorial](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/Consensus-DataInput.R).
-The first tutorial focuses on data input, cleaning and pre-processing; outlier removal; clustering of samples by their Euclidean distance; a comparison of sample cluster to sample meta data.
+# Figures
 
@@
 source("../analysis/08_WGCNA_1.R")
+subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary",
+                           "male.gonad", "male.hypothalamus", "male.pituitary"),
+            mytitle = "Characterization dendrogram with trait heatmap")
+```
 
+**Fig 1. Characterization dendrogram with trait heatmap.**.
+Most hypothalamus and pituitary samples separte form the gonad samples, but some are clusters contain samples from all tissues.The top panel is a dendrogram showing Euclidean distance of sample.
+Females (white), males (red), gonad (white), hypothalamus (pink), pituitary (red).
+
+```{r wgcna-female, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary"),
+            mytitle = "Female sample dendrogram with trait heatmap")
 ```
+
+**Fig 2. Female characterization dendrogram with trait heatmap.**. 
+Most pituitary samples form a single cluster.
+More hypothalamus samples cluster with gonads than with pituitary.
+The top panel is a dendrogram showing Euclidean distance of sample.
+Females (white), males (red), gonad (white), hypothalamus (pink), pituitary (red).
+
+```{r wgcna-male, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+subsetWGCNA(c("male.gonad", "male.hypothalamus", "male.pituitary"),
+            mytitle = "Male sample dendrogram with trait heatmap")
+```
+
+**Fig 3. Male characterization dendrogram with trait heatmap.**. 
+ As with females, most pituitary samples form a single cluster, while more hypothalamus samples cluster with gonads than with pituitary.
 

### 2019-08-23 docs/manuscript_draft.tex
new figs

--- 2019-08-18 docs/manuscript_draft/manuscript_draft.tex
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+Here are two sample references: {[}1,2{]}.
 
-Here are two sample references: {[}1,2{]}.
+\hypertarget{figures}{%
+\section{Figures}\label{figures}}
+
+\includegraphics{manuscript_draft_files/figure-latex/wgcna-1.pdf}
+
+\textbf{Fig 1. Characterization dendrogram with trait heatmap.}. Most
+hypothalamus and pituitary samples separte form the gonad samples, but
+some are clusters contain samples from all tissues.The top panel is a
+dendrogram showing Euclidean distance of sample. Females (white), males
+(red), gonad (white), hypothalamus (pink), pituitary (red).
+
+\includegraphics{manuscript_draft_files/figure-latex/wgcna-female-1.pdf}
+
+\textbf{Fig 2. Female characterization dendrogram with trait heatmap.}.
+Most pituitary samples form a single cluster. More hypothalamus samples
+cluster with gonads than with pituitary. The top panel is a dendrogram
+showing Euclidean distance of sample. Females (white), males (red),
+gonad (white), hypothalamus (pink), pituitary (red).
+
+\includegraphics{manuscript_draft_files/figure-latex/wgcna-male-1.pdf}
+
+\textbf{Fig 3. Male characterization dendrogram with trait heatmap.}. As
+with females, most pituitary samples form a single cluster, while more
+hypothalamus samples cluster with gonads than with pituitary.
 

### 2019-08-23 docs/manuscript_draft.Rmd
real citations

@@
 Understanding the mechanisms underlying parental care are critical to circumventing issues with parent-newborn bonding as well, where ultimate explanations are obvious, but specific mechanisms remain elusive.
-
-The rock dove (_Columba livia_) is an ideal system to characterize changes in genetic expression during parental care transitions because: 1) ample genomic resources are available, including a complete annotated genome assembly (Gillespie et al.
-2013; Shapiro et al.
-2013) and methodology concerning reproductive physiology and behavior (Dong et al.
+The rock dove (_Columba livia_) is an ideal system to characterize changes in genetic expression during parental care transitions because: 1) ample genomic resources are available (Gillespie et al.
+2013), including a complete annotated genome assembly @Shapiro1063 and methodology concerning reproductive physiology and behavior (Dong et al.
 2012); and 2) rock doves are prolific, year-round breeders that thrive in captivity, making observation, manipulation, and sampling highly feasible year-round.
@@
 Many functional similarities between rock dove and mammalian lactation exist concerning the mediation of this event by the hormone prolactin (Dumont 1965).
-Additionally, like mammalian milk, rock dove milk delivers essential immunoglobulins and nutritional benefits to young, aiding in their immune function and development of microbiota (Gillespie et al.
-2012).
+Additionally, like mammalian milk, rock dove milk delivers essential immunoglobulins and nutritional benefits to young, aiding in their immune function and development of microbiota @Gillespie2012.
 Thus, because rock doves incubate eggs and exhibit mammalian-like mediation and function of lactation for young, they have the potential to serve as a powerful theoretical bridge to understand the neurobiology of both avian and mammalian transitions into parental care.
@@
 # Materials and Methods
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
 
@@
 
-Here are two sample references: @Feynman1963118 [@Dirac1953888].
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Curabitur eget porta erat.
+Morbi consectetur est vel gravida pretium.
+Suspendisse ut dui eu ante cursus gravida non sed sem.
+Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
+Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
+Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
+Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
 
@@
 
-# Metadata and extras
+<!--
 
-## Funding Statment
+# Funding Statment
 
@@
 
+-->
+
 ```{r include=F}

### 2019-08-23 docs/manuscript_draft.tex
real citations

@@
 circumventing issues with parent-newborn bonding as well, where ultimate
-explanations are obvious, but specific mechanisms remain elusive.
-
-The rock dove (\emph{Columba livia}) is an ideal system to characterize
+explanations are obvious, but specific mechanisms remain elusive. The
+rock dove (\emph{Columba livia}) is an ideal system to characterize
 changes in genetic expression during parental care transitions because:
-1) ample genomic resources are available, including a complete annotated
-genome assembly (Gillespie et al.~2013; Shapiro et al.~2013) and
-methodology concerning reproductive physiology and behavior (Dong et
-al.~2012); and 2) rock doves are prolific, year-round breeders that
-thrive in captivity, making observation, manipulation, and sampling
-highly feasible year-round.
+1) ample genomic resources are available (Gillespie et al.~2013),
+including a complete annotated genome assembly {[}1{]} and methodology
+concerning reproductive physiology and behavior (Dong et al.~2012); and
+2) rock doves are prolific, year-round breeders that thrive in
+captivity, making observation, manipulation, and sampling highly
+feasible year-round.
 
@@
 essential immunoglobulins and nutritional benefits to young, aiding in
-their immune function and development of microbiota (Gillespie et
-al.~2012). Thus, because rock doves incubate eggs and exhibit
-mammalian-like mediation and function of lactation for young, they have
-the potential to serve as a powerful theoretical bridge to understand
-the neurobiology of both avian and mammalian transitions into parental
-care.
+their immune function and development of microbiota {[}2{]}. Thus,
+because rock doves incubate eggs and exhibit mammalian-like mediation
+and function of lactation for young, they have the potential to serve as
+a powerful theoretical bridge to understand the neurobiology of both
+avian and mammalian transitions into parental care.
 
@@
 \section{Materials and Methods}\label{materials-and-methods}}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
-porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
-dui eu ante cursus gravida non sed sem. Phasellus mauris velit, dapibus
-finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
-diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
-ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
-pulvinar, volutpat risus.
 
@@
 
-Here are two sample references: {[}1,2{]}.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
+porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
+dui eu ante cursus gravida non sed sem. Phasellus mauris velit, dapibus
+finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
+diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
+ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
+pulvinar, volutpat risus.
 
@@
 
-\hypertarget{metadata-and-extras}{%
-\section{Metadata and extras}\label{metadata-and-extras}}
-
-\hypertarget{funding-statment}{%
-\subsection{Funding Statment}\label{funding-statment}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Collaborative Research: RUI: The Neural Basis of Becoming a Parent:
-  From Genotype to Phenotype. NSF-IOS:1455957.
-\end{enumerate}
-
 \hypertarget{references}{%
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Feynman1963118}{}%
-1. Feynman R, Vernon Jr. F. The theory of a general quantum system
-interacting with a linear dissipative system. Annals of Physics.
-1963;24: 118--173.
-doi:\href{https://doi.org/10.1016/0003-4916(63)90068-X}{10.1016/0003-4916(63)90068-X}
+\leavevmode\hypertarget{ref-Shapiro1063}{}%
+1. Shapiro MD, Kronenberg Z, Li C, Domyan ET, Pan H, Campbell M, et al.
+Genomic diversity and evolution of the head crest in the rock pigeon.
+Science. American Association for the Advancement of Science; 2013;339:
+1063--1067.
+doi:\href{https://doi.org/10.1126/science.1230422}{10.1126/science.1230422}
 
-\leavevmode\hypertarget{ref-Dirac1953888}{}%
-2. Dirac P. The lorentz transformation and absolute time. Physica.
-1953;19: 888--896.
-doi:\href{https://doi.org/10.1016/S0031-8914(53)80099-6}{10.1016/S0031-8914(53)80099-6}
+\leavevmode\hypertarget{ref-Gillespie2012}{}%
+2. Gillespie MJ, Stanley D, Chen H, Donald JA, Nicholas KR, Moore RJ, et
+al. Functional Similarities between Pigeon `Milk' and Mammalian Milk:
+Induction of Immune Gene Expression and Modification of the Microbiota.
+Salmon H, editor. PLoS ONE. Public Library of Science; 2012;7: e48363.
+doi:\href{https://doi.org/10.1371/journal.pone.0048363}{10.1371/journal.pone.0048363}
 

### 2019-08-26 analysis/08_wgcna.Rmd
now with hypothesis groups

@@
 source("08_WGCNA_1.R")
+
 ```
@@
             mytitle = "Characterization dendrogram with trait heatmap")
+
 ```
@@
             mytitle = "Female sample dendrogram with trait heatmap")
+
 ```
@@
             mytitle = "Male sample dendrogram with trait heatmap")
+
 ```
@@
 ```
+
+## female hypothalamus
+
+```{r wgcna-f-hyp, message=FALSE, warning=FALSE, cache=TRUE}
+subsetWGCNA(c("female.hypothalamus"),
+            mytitle = "Female hypothalamus sample dendrogram")
+```
+
+## female gonad
+
+```{r wgcna-f-gon, message=FALSE, warning=FALSE, cache=TRUE}
+subsetWGCNA(c("female.gonad"),
+            mytitle = "Female gonad sample dendrogram")
+```

### 2019-08-26 analysis/08_WGCNA_1.R

@@ -21,6 +21,7 @@ countData <- read.csv("../results/00_countData_characterization.csv", row.names 
 
 # create new grouping for subsets
 colData$sextissue <-  paste(colData$sex, colData$tissue, sep = ".")
+levels(colData$treatment)
 
 ## rename rownames and colnames for better vizualiation
 colData <- colData %>%
@@ -30,7 +31,17 @@ colData <- colData %>%
          tissue = fct_recode(tissue,
                              "pit" = "pituitary",
                              "hyp" = "hypothalamus",
-                             "gon" = "gonad"))
+                             "gon" = "gonad"),
+         hypothesis = fct_recode(treatment,
+                                 "anticipation" = "control",
+                                 "anticipation" = "bldg",
+                                 "incubation" = "lay",
+                                 "incubation" = "inc.d3",
+                                 "incubation" = "inc.d9",
+                                 "incubation" = "inc.d17",
+                                 "hatchling.care" = "hatch",
+                                 "hatchling.care" = "n5",
+                                 "hatchling.care" = "n9"))
 
 colData$ID <- as.numeric(colData$bird)
 colData$sample <- paste(colData$sex, colData$tissue, colData$treatment, colData$ID, sep = ".")
@@ -145,7 +156,7 @@ nSamples = nrow(datExpr)
 #=====================================================================================
 
 
-traitData <- colData[c(5,3,4)]
+traitData <- colData[c(4,3,5,9)]
 
 # numeric
 head(traitData)

### 2019-08-26 analysis/03_DESeq2_characterization.Rmd
updated characterization analysis

@@
 
-```{r dds, cache = T, eval = F}
+```{r dds, cache = T, eval = T}
 dds.female_hypothalamus <- subsetDESeq(c.colData,  c.countData, "female_hypothalamus")
@@
 
-```{r totalDEGs, eval = F}
+```{r returntotalDEGs, eval = T}
+DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
+DEGs.female_pituitary <- returntotalDEGs(dds.female_pituitary)
+DEGs.female_gonad <- returntotalDEGs(dds.female_gonad)
+DEGs.male_hypothalamus <- returntotalDEGs(dds.male_hypothalamus)
+DEGs.male_pituitary <- returntotalDEGs(dds.male_pituitary)
+DEGs.male_gondad <- returntotalDEGs(dds.male_gondad)
+```
+
+```{r totalDEGs, eval = T}
 #create list of groups for deseq contrasts
@@
 
-a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
-b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
-c <- plottotalDEGs(dds.female_gonad, "female gonad")
-d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
-e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
-f <- plottotalDEGs(dds.male_gondad, "male gonad")
+a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus")
+b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary")
+c <- plottotalDEGs(DEGs.female_gonad, "female gonad")
+d <- plottotalDEGs(DEGs.male_hypothalamus, "male hypothalamus")
+e <- plottotalDEGs(DEGs.male_pituitary, "male pituitary")
+f <- plottotalDEGs(DEGs.male_gondad, "male gonad")
 
@@
 
+## Plot only a subset of DEGs that relate to hypotheses as bar and line graphs
+
+```{r barplots}
+# bars with lines
+g <- plotserialDEGs(DEGs.female_hypothalamus, "Female hypothalamus", "#00A08A")
+h <- plotserialDEGs(DEGs.female_pituitary, "Female pituitary", "#00A08A")
+i <- plotserialDEGs(DEGs.female_gonad, "Female gonad", "#00A08A")
+j <- plotserialDEGs(DEGs.male_hypothalamus, "Male hypothalamus", "#F98400")
+k <- plotserialDEGs(DEGs.male_pituitary, "Male pituitary", "#F98400")
+l <- plotserialDEGs(DEGs.male_gondad, "Male gonad", "#F98400")
+
+mybarplots <- plot_grid(g, h , i , j ,k , l, nrow = 2) 
+mybarplots
+
+pdf("../figures/characterization/barplots.pdf", width = 10, height = 6)
+plot(mybarplots)
+dev.off()
+```
+
+```{r linegraphs}
+
+# lines only, different colors for different sex tissue groups
+FH <- subsetDEGs(DEGs.female_hypothalamus, "Female hypothalamus")
+FP <- subsetDEGs(DEGs.female_pituitary, "Female pituitary")
+FG <- subsetDEGs(DEGs.female_gonad, "Female gonad")
+MH <- subsetDEGs(DEGs.male_hypothalamus, "Male hypothalamus")
+MP <- subsetDEGs(DEGs.male_pituitary, "Male pituitary")
+MG <- subsetDEGs(DEGs.male_gondad, "Male gonad")
+
+specificDEGs <- rbind(FH, FP, FG, MH, MP, MG)
+
+specificDEGs$sextissue <- factor(specificDEGs$sextissue, levels = c("Female hypothalamus", "Female pituitary", "Female gonad",
+                                                                    "Male hypothalamus", "Male pituitary", "Male gonad"))
+mylinegraph <- ggplot(specificDEGs, aes(x=comparison, y=V3, group = sextissue)) +
+    geom_line(aes(color=sextissue)) +
+    geom_point(aes(color=sextissue)) + 
+    theme_minimal(base_size = 10) +
+    #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+    labs(x = "Parental stage transitions", y= "No. of differentially expressed genes ",
+         title = "Gene expression changes during parental care transitions") + 
+    scale_x_discrete(labels=c("Control to\nNest bldg", "Nest bldg. to\nLay", "Lay to\nInc. day 3",
+                              "Inc. day 3 to\nInc. day 9", "Inc. day 9 to\nInc. day 17",
+                              "Inc. day 17 to\nHatch", "Hatch to\nNestling day 5", "Nestling day 5 to\n Nestling day 9")) +
+   theme(legend.title = element_blank(),
+         legend.position = c(0.8, 0.8))
+mylinegraph
+
+pdf("../figures/characterization/linegraphs.pdf", width = 8, height = 4)
+plot(mylinegraph)
+dev.off()
+```
+
 ## Calculate and plot principal component
 
-```{r pca, eval = F}
-plotPCAs(dds.female_hypothalamus, "female hypothalamus")
-plotPCAs(dds.female_pituitary, "female pituitary")      
-plotPCAs(dds.female_gonad, "female gonad")
-plotPCAs(dds.male_hypothalamus, "male hypothalamus")
-plotPCAs(dds.male_pituitary, "male pituitary")
-plotPCAs(dds.male_gondad, "male gonad")
+```{r pca, eval = T}
+FhypPCA <- returnPCAs(dds.female_hypothalamus)
+FpitPCA <- returnPCAs(dds.female_pituitary)      
+FgonPCA <- returnPCAs(dds.female_gonad)
+MhypPCA <- returnPCAs(dds.male_hypothalamus)
+MpitPCA <- returnPCAs(dds.male_pituitary)
+MgonPCA <- returnPCAs(dds.male_gondad)
+
+plotPC12(FhypPCA, "female hypothalamus")
+plotPC12(FpitPCA, "female pituitary")
+plotPC12(FgonPCA, "female gonad")
+plotPC12(MhypPCA, "male hypothalamus")
+plotPC12(MpitPCA, "male pituitary")
+plotPC12(MgonPCA, "male gonad")
 ```

### 2019-08-26 analysis/04_DESeq2_sexes.Rmd
updated characterization analysis

@@
 plot(myDEGs)
-dev.off()
-```
-
-## Plot only a subset of DEGs that relate to hypotheses
-
-```{r characterization}
-# bars with lines
-g <- plotserialDEGs(DEGs.female_hypothalamus, "Female hypothalamus", "#00A08A")
-h <- plotserialDEGs(DEGs.female_pituitary, "Female pituitary", "#00A08A")
-i <- plotserialDEGs(DEGs.female_gonad, "Female gonad", "#00A08A")
-j <- plotserialDEGs(DEGs.male_hypothalamus, "Male hypothalamus", "#F98400")
-k <- plotserialDEGs(DEGs.male_pituitary, "Male pituitary", "#F98400")
-l <- plotserialDEGs(DEGs.male_gondad, "Male gonad", "#F98400")
-
-mybarplots <- plot_grid(g, h , i , j ,k , l, nrow = 2) 
-mybarplots
-
-pdf("../figures/sexes/characterization-1.pdf", width = 10, height = 6)
-plot(mybarplots)
-dev.off()
-
-# lines only, different colors for different sex tissue groups
-FH <- subsetDEGs(DEGs.female_hypothalamus, "Female hypothalamus")
-FP <- subsetDEGs(DEGs.female_pituitary, "Female pituitary")
-FG <- subsetDEGs(DEGs.female_gonad, "Female gonad")
-MH <- subsetDEGs(DEGs.male_hypothalamus, "Male hypothalamus")
-MP <- subsetDEGs(DEGs.male_pituitary, "Male pituitary")
-MG <- subsetDEGs(DEGs.male_gondad, "Male gonad")
-
-specificDEGs <- rbind(FH, FP, FG, MH, MP, MG)
-
-specificDEGs$sextissue <- factor(specificDEGs$sextissue, levels = c("Female hypothalamus", "Female pituitary", "Female gonad",
-                                                                    "Male hypothalamus", "Male pituitary", "Male gonad"))
-mylinegraph <- ggplot(specificDEGs, aes(x=comparison, y=V3, group = sextissue)) +
-    geom_line(aes(color=sextissue)) +
-    geom_point(aes(color=sextissue)) + 
-    theme_minimal(base_size = 10) +
-    #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-    labs(x = "Parental stage transitions", y= "No. of differentially expressed genes ",
-         title = "Gene expression changes during parental care transitions") + 
-    scale_x_discrete(labels=c("Control to\nNest bldg", "Nest bldg. to\nLay", "Lay to\nInc. day 3",
-                              "Inc. day 3 to\nInc. day 9", "Inc. day 9 to\nInc. day 17",
-                              "Inc. day 17 to\nHatch", "Hatch to\nNestling day 5", "Nestling day 5 to\n Nestling day 9")) +
-   theme(legend.title = element_blank(),
-         legend.position = c(0.8, 0.8))
-mylinegraph
-
-pdf("../figures/sexes/characterization-2.pdf", width = 8, height = 4)
-plot(mylinegraph)
 dev.off()

### 2019-08-26 docs/manuscript_draft.tex
updated manuscript

@@
 
-\hypertarget{materials-and-methods}{%
-\section{Materials and Methods}\label{materials-and-methods}}
+Aim I lays the necessary groundwork for establishing a pattern of
+correlative relationships between observed changes in gene expression
+and patterns of parental behavior. With this pattern established, we can
+proceed with experimental manipulations proposed under Aim III to
+determine causation.
+
+\hypertarget{materials-and-methods-and-results}{%
+\section{Materials and Methods and
+Results}\label{materials-and-methods-and-results}}
 
 \includegraphics[width=360px]{../figures/images/samplingtimepoints}
+
+\hypertarget{characterize-changes-in-neural-gene-expression-during-parental-care-transitions}{%
+\subsection{Characterize changes in neural-gene expression during
+parental care
+transitions}\label{characterize-changes-in-neural-gene-expression-during-parental-care-transitions}}
+
+Our working hypothesis is that distinct changes in transcription occur
+in the brain at the anticipation of, during, and in response to two
+different types of parental care: incubation behavior and hatchling
+care.
+
+\hypertarget{overall-patterns-of-sample-similarity}{%
+\subsubsection{overall patterns of sample
+similarity}\label{overall-patterns-of-sample-similarity}}
+
+This ecludian clustering shows many clusters are heterogenous,
+consisting of sames from all tissues types, both sexes, and from
+multiple treatment groups.
+
+\includegraphics{manuscript_draft_files/figure-latex/wgcna-1.pdf}
+
+\textbf{Fig 1. Characterization dendrogram with trait heatmap.}. Most
+hypothalamus and pituitary samples separte form the gonad samples, but
+some are clusters contain samples from all tissues.The top panel is a
+dendrogram showing Euclidean distance of sample. Females (white), males
+(red), gonad (white), hypothalamus (pink), pituitary (red).
+
+\hypertarget{subset-by-sex-and-tissue}{%
+\subsection{subset by sex and tissue}\label{subset-by-sex-and-tissue}}
+
+The rest of the analyses are peformed on each sex and tissue group
+separately, a 2 x 3 design.
+
+\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-4-1.pdf}
+
+\textbf{Fig 2. The magnitude of gene expression changes between each
+parental transition}. 4-5K genes are differentially expressed beteen
+control birds and their nest buliding conspecifics in all tissues except
+the male gonad. 500 - 1000 genes are differentially expressed in the
+female pituitary from mid-late incubation as well as in the female
+hypothalamus and pituitary from hatch to nestingly care day 5.
+
+\includegraphics{../figures/characterization/musicalgenes-05.png}
 
@@
 
-\hypertarget{remove-eggs}{%
-\subsubsection{Remove eggs}\label{remove-eggs}}
+\hypertarget{remove-eggs-or-chicks}{%
+\subsubsection{Remove eggs or chicks}\label{remove-eggs-or-chicks}}
 
@@
 internal clock mechanism.
+
+We will remove newly, naturally hatched chick(s) on Day 2: second chick
+hatches and sample brains from parents the following day (Day 2 post
+(second) egg hatching). Prediction: Behavior, transcription and
+translation are sensory driven and will revert to a pattern
+characteristic of the `Nesting, prior to lay' point, supporting our
+working hypothesis. Alternatively, if an internal clock mechanism is at
+play, we would expect to see a pattern similar to birds caring for
+chicks on Day 2 post (second) egg hatching.
+
+\includegraphics{../figures/sexes/offspringremoval-1.png}
 
@@
 
-\hypertarget{remove-chicks}{%
-\subsubsection{Remove chicks}\label{remove-chicks}}
-
-We will remove newly, naturally hatched chick(s) on Day 2: second chick
-hatches and sample brains from parents the following day (Day 2 post
-(second) egg hatching). Prediction: Behavior, transcription and
-translation are sensory driven and will revert to a pattern
-characteristic of the `Nesting, prior to lay' point, supporting our
-working hypothesis. Alternatively, if an internal clock mechanism is at
-play, we would expect to see a pattern similar to birds caring for
-chicks on Day 2 post (second) egg hatching.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
 Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
@@
 
-\hypertarget{figures}{%
-\section{Figures}\label{figures}}
-
-\includegraphics{manuscript_draft_files/figure-latex/wgcna-1.pdf}
-
-\textbf{Fig 1. Characterization dendrogram with trait heatmap.}. Most
-hypothalamus and pituitary samples separte form the gonad samples, but
-some are clusters contain samples from all tissues.The top panel is a
-dendrogram showing Euclidean distance of sample. Females (white), males
-(red), gonad (white), hypothalamus (pink), pituitary (red).
+\hypertarget{supplementary-figures}{%
+\section{Supplementary Figures}\label{supplementary-figures}}
 
@@
 
-\textbf{Fig 2. Female characterization dendrogram with trait heatmap.}.
-Most pituitary samples form a single cluster. More hypothalamus samples
-cluster with gonads than with pituitary. The top panel is a dendrogram
-showing Euclidean distance of sample. Females (white), males (red),
-gonad (white), hypothalamus (pink), pituitary (red).
+\textbf{Fig S1a. Female characterization dendrogram with trait
+heatmap.}. Most pituitary samples form a single cluster. More
+hypothalamus samples cluster with gonads than with pituitary. The top
+panel is a dendrogram showing Euclidean distance of sample. Females
+(white), males (red), gonad (white), hypothalamus (pink), pituitary
+(red).
 
@@
 
-\textbf{Fig 3. Male characterization dendrogram with trait heatmap.}. As
-with females, most pituitary samples form a single cluster, while more
-hypothalamus samples cluster with gonads than with pituitary.
+\textbf{Fig S1b. Male characterization dendrogram with trait heatmap.}.
+As with females, most pituitary samples form a single cluster, while
+more hypothalamus samples cluster with gonads than with pituitary.
 

### 2019-08-26 docs/manuscript_draft.Rmd
now with hypothesis groups
updated manuscript

@@
 
-# Materials and Methods
+Aim I lays the necessary groundwork for establishing a pattern of correlative relationships between observed changes in gene expression and patterns of parental behavior.
+With this pattern established, we can proceed with experimental manipulations proposed under Aim III to determine causation.
+
+# Materials and Methods and Results
 
@@
 ```
+
+## Characterize changes in neural-gene expression during parental care transitions
+
+Our working hypothesis is that distinct changes in transcription occur in the brain at the anticipation of, during, and in response to two different types of parental care: incubation behavior and hatchling care.
+
+### overall patterns of sample similarity
+
+This ecludian clustering shows many clusters are heterogenous, consisting of sames from all tissues types, both sexes, and from multiple treatment groups.
+
+```{r wgcna, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+source("../analysis/08_WGCNA_1.R")
+subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary",
+                           "male.gonad", "male.hypothalamus", "male.pituitary"),
+            mytitle = "Characterization dendrogram with trait heatmap")
+
+```
+
+**Fig 1. Characterization dendrogram with trait heatmap.**.
+Most hypothalamus and pituitary samples separte form the gonad samples, but some are clusters contain samples from all tissues.The top panel is a dendrogram showing Euclidean distance of sample.
+Females (white), males (red), gonad (white), hypothalamus (pink), pituitary (red).
+
+## subset by sex and tissue
+
+The rest of the analyses are peformed on each sex and tissue group separately, a 2 x 3 design.
+
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+library(cowplot)
+ggdraw() + draw_image("../figures/sexes/characterization-2.png") 
+```
+
+**Fig 2. The magnitude of gene expression changes between each parental transition**.
+4-5K genes are differentially expressed beteen control birds and their nest buliding conspecifics in all tissues except the male gonad.
+500 - 1000 genes are differentially expressed in the female pituitary from mid-late incubation as well as in the female hypothalamus and pituitary from hatch to nestingly care day 5.
+
+![](../figures/characterization/musicalgenes-05.png)
 
@@
 ### Give birds eggs
+
 We will place eggs in nests of birds lacking them and sample brains the following day.
@@
 ### Prolong incubation
+
 We will prolong incubation by removing eggs from actively incubated nests on the day before hatching (Day 17) and immediately replace them with infertile, ‘dummy’ eggs.
@@
 ### Extend time of hatching
+
 We will prolong incubation by removing eggs from nests undergoing active incubation on Day 17 (the end of incubation) and immediately replacing eggs with ‘dummy’ eggs.
@@
 
-### Remove eggs
+### Remove eggs or chicks
+
 We will remove eggs during incubation on Day 1 (beginning), Day 9 (middle), or Day 17 (end) and sample brains on the following day.
@@
 
+We will remove newly, naturally hatched chick(s) on Day 2: second chick hatches and sample brains from parents the following day (Day 2 post (second) egg hatching).
+Prediction: Behavior, transcription and translation are sensory driven and will revert to a pattern characteristic of the ‘Nesting, prior to lay’ point, supporting our working hypothesis.
+Alternatively, if an internal clock mechanism is at play, we would expect to see a pattern similar to birds caring for chicks on Day 2 post (second) egg hatching.
+
+![](../figures/sexes/offspringremoval-1.png)
+
 ### Reduce time to hatching
+
 We will remove eggs on Day 8 (middle) of incubation and immediately replace them with newly hatched chicks.
@@
 Alternatively, if profiles remain similar to birds sampled on Day 9 that are actively incubating eggs, this would support the presence of an internal clock mechanism.
-
-### Remove chicks
-We will remove newly, naturally hatched chick(s) on Day 2: second chick hatches and sample brains from parents the following day (Day 2 post (second) egg hatching).
-Prediction: Behavior, transcription and translation are sensory driven and will revert to a pattern characteristic of the ‘Nesting, prior to lay’ point, supporting our working hypothesis.
-Alternatively, if an internal clock mechanism is at play, we would expect to see a pattern similar to birds caring for chicks on Day 2 post (second) egg hatching.
-
-# Results
 
@@
 
-# Figures
-
-```{r wgcna, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-source("../analysis/08_WGCNA_1.R")
-subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary",
-                           "male.gonad", "male.hypothalamus", "male.pituitary"),
-            mytitle = "Characterization dendrogram with trait heatmap")
-```
-
-**Fig 1. Characterization dendrogram with trait heatmap.**.
-Most hypothalamus and pituitary samples separte form the gonad samples, but some are clusters contain samples from all tissues.The top panel is a dendrogram showing Euclidean distance of sample.
-Females (white), males (red), gonad (white), hypothalamus (pink), pituitary (red).
+# Supplementary Figures
 
@@
 
-**Fig 2. Female characterization dendrogram with trait heatmap.**. 
+**Fig S1a. Female characterization dendrogram with trait heatmap.**. 
 Most pituitary samples form a single cluster.
@@
 
-**Fig 3. Male characterization dendrogram with trait heatmap.**. 
+**Fig S1b. Male characterization dendrogram with trait heatmap.**. 
  As with females, most pituitary samples form a single cluster, while more hypothalamus samples cluster with gonads than with pituitary.

### 2019-08-26 analysis/08_WGCNA_1.R
remove dendrogram labels

@@ -187,7 +187,8 @@ traitColors = numbers2colors(datTraits, signed = FALSE);
 plotDendroAndColors(sampleTree2, traitColors,
                     groupLabels = names(datTraits), 
                     main = mytitle,
-                    cex.dendroLabels = 0.4
+                    cex.dendroLabels = 0.4,
+                    dendroLabels = FALSE
                     )
 
 }

### 2019-08-26 docs/manuscript_draft.Rmd
remove dendrogram labels

@@
 
-### overall patterns of sample similarity
-
 This ecludian clustering shows many clusters are heterogenous, consisting of sames from all tissues types, both sexes, and from multiple treatment groups.
@@
                            "male.gonad", "male.hypothalamus", "male.pituitary"),
-            mytitle = "Characterization dendrogram with trait heatmap")
-
+            mytitle = "Characterization dendrogram of samples")
 ```
@@
 Females (white), males (red), gonad (white), hypothalamus (pink), pituitary (red).
-
-## subset by sex and tissue
 

### 2019-08-27 analysis/03_DESeq2_characterization.Rmd
pheatmap still not working

@@
 
-```{r pheatmap, eval = F}
-makepheatmap(dds.female_hypothalamus, "female hypothalamus")
-makepheatmap(dds.female_pituitary, "female pituitary")
-makepheatmap(dds.female_gonad, "female gonad")
-makepheatmap(dds.male_hypothalamus, "male hypothalamus")
-makepheatmap(dds.male_pituitary, "male pituitary")
-makepheatmap(dds.male_gondad, "male gonad")
+```{r pheatmap, eval = T}
+colDataFH <- subsetcolData(c.colData, "female_hypothalamus")
+colDataFP <- subsetcolData(c.colData, "female_pituitary")
+colDataFG <- subsetcolData(c.colData, "female_gonad")
+colDataMH <- subsetcolData(c.colData, "male_hypothalamus")
+colDataMP <- subsetcolData(c.colData, "male_pituitary")
+colDataMG <- subsetcolData(c.colData, "male_gonad")
+
+makepheatmap(dds.female_hypothalamus, colDataFH, "female hypothalamus")
+makepheatmap(dds.female_pituitary, colDataFP, "female pituitary")
+makepheatmap(dds.female_gonad, colDataFG, "female gonad")
+makepheatmap(dds.male_hypothalamus, colDataMH,  "male hypothalamus")
+makepheatmap(dds.male_pituitary, colDataMP, "male pituitary")
+makepheatmap(dds.male_gondad, colDataMG, "male gonad")
 ```

### 2019-08-27 analysis/03_DESeq2_characterization.Rmd
pheatmap working, but slow, only added 1 fig

--- 2019-08-26 analysis/03_DESeq2_characterization.Rmd
@@
 
-```{r pca, eval = T}
+```{r pca, eval = T, cache = TRUE}
+
 FhypPCA <- returnPCAs(dds.female_hypothalamus)
+plotPC12(FhypPCA, "female hypothalamus")
+
 FpitPCA <- returnPCAs(dds.female_pituitary)      
+plotPC12(FpitPCA, "female pituitary")
+
 FgonPCA <- returnPCAs(dds.female_gonad)
+plotPC12(FgonPCA, "female gonad")
+
 MhypPCA <- returnPCAs(dds.male_hypothalamus)
+plotPC12(MhypPCA, "male hypothalamus")
+
 MpitPCA <- returnPCAs(dds.male_pituitary)
+plotPC12(MpitPCA, "male pituitary")
+
 MgonPCA <- returnPCAs(dds.male_gondad)
-
-plotPC12(FhypPCA, "female hypothalamus")
-plotPC12(FpitPCA, "female pituitary")
-plotPC12(FgonPCA, "female gonad")
-plotPC12(MhypPCA, "male hypothalamus")
-plotPC12(MpitPCA, "male pituitary")
 plotPC12(MgonPCA, "male gonad")
@@
 
-```{r pheatmap, eval = F}
-makepheatmap(dds.female_hypothalamus, "female hypothalamus")
-makepheatmap(dds.female_pituitary, "female pituitary")
-makepheatmap(dds.female_gonad, "female gonad")
-makepheatmap(dds.male_hypothalamus, "male hypothalamus")
-makepheatmap(dds.male_pituitary, "male pituitary")
-makepheatmap(dds.male_gondad, "male gonad")
+```{r pheatmap, eval = T, cache = TRUE}
+colDataFH <- subsetcolData(c.colData, "female_hypothalamus")
+#colDataFP <- subsetcolData(c.colData, "female_pituitary")
+#colDataFG <- subsetcolData(c.colData, "female_gonad")
+#colDataMH <- subsetcolData(c.colData, "male_hypothalamus")
+#colDataMP <- subsetcolData(c.colData, "male_pituitary")
+#colDataMG <- subsetcolData(c.colData, "male_gonad")
+
+makepheatmap(dds.female_hypothalamus, colDataFH, "female hypothalamus")
+#makepheatmap(dds.female_pituitary, colDataFP, "female pituitary")
+#makepheatmap(dds.female_gonad, colDataFG, "female gonad")
+#makepheatmap(dds.male_hypothalamus, colDataMH,  "male hypothalamus")
+#makepheatmap(dds.male_pituitary, colDataMP, "male pituitary")
+#makepheatmap(dds.male_gondad, colDataMG, "male gonad")
 ```

### 2019-08-27 R/functions.R
candidates working again

@@ -421,21 +421,14 @@ plotcandidates <- function(mydds, colData, mysubtitle){
   # make dataframe counts
   DEGs <- assay(vsd)
   DEGs <- as.data.frame(DEGs)
-  
-  names(geneinfo)
-  names(geneinfo)[4] <- "rownames"
-  DEGs$rownames <- row.names(DEGs)
-  print(head(DEGs))
+  DEGs$entrezid <- row.names(DEGs)
 
   # make dataframe with geneids and names and counts
   # how to gather: https://tidyr.tidyverse.org/reference/gather.html
  
+  candidates <- full_join(geneinfo, DEGs, by = "entrezid")
+  head(candidates)
   
-  candidates <- full_join(geneinfo, DEGs)
-  candidates <- candidates %>%
-    filter(pmin < 0.01) 
-  drop.cols <-colnames(candidates[,grep("padj|pval|pmin", colnames(candidates))])
-  candidates <- candidates %>% dplyr::select(-one_of(drop.cols))
   candidates <- candidates %>%
     filter(Name %in% c("AVP", "AVPR1A", "AVPR1B", "AVPR2",
                       "SERPINA4", "CRH", "CRHR1", "DRD1", "DRD2",
@@ -445,7 +438,7 @@ plotcandidates <- function(mydds, colData, mysubtitle){
                       "FSHB", "GHRL", "GAL", "NPVF",  "NPFFR1", "GNRH1",  "GNRHR", 
                       "LEPR",  "LHCGR", "PGR", "PRL", "PRLR", "VIP", "VIPR1"))
   row.names(candidates) <- candidates$Name
-  candidates <- candidates %>% select(-row.names, -rownames, -Name, -geneid)
+  candidates <- candidates %>% dplyr::select(-row.names, -entrezid, -Name, -geneid)
   candidates <- candidates %>% drop_na()
   candidates <- as.data.frame(t(candidates))
   candidates$RNAseqID <- rownames(candidates)
@@ -454,25 +447,18 @@ plotcandidates <- function(mydds, colData, mysubtitle){
   candidates$value <- as.numeric(candidates$value)
   candidates$V1  <- candidates$RNAseqID
   
-  candidatecounts <- left_join(candidates, colData)
+  candidatecounts <- left_join(candidates, colData, by = "V1")
   candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
   candidatecounts$gene <- as.factor(candidatecounts$gene)
   
[…]

### 2019-08-27 analysis/03_DESeq2_characterization.Rmd
candidates working again

@@
 colDataFH <- subsetcolData(c.colData, "female_hypothalamus")
-#colDataFP <- subsetcolData(c.colData, "female_pituitary")
-#colDataFG <- subsetcolData(c.colData, "female_gonad")
-#colDataMH <- subsetcolData(c.colData, "male_hypothalamus")
-#colDataMP <- subsetcolData(c.colData, "male_pituitary")
-#colDataMG <- subsetcolData(c.colData, "male_gonad")
+colDataFP <- subsetcolData(c.colData, "female_pituitary")
+colDataFG <- subsetcolData(c.colData, "female_gonad")
+colDataMH <- subsetcolData(c.colData, "male_hypothalamus")
+colDataMP <- subsetcolData(c.colData, "male_pituitary")
+colDataMG <- subsetcolData(c.colData, "male_gonad")
 
-makepheatmap(dds.female_hypothalamus, colDataFH, "female hypothalamus")
+#makepheatmap(dds.female_hypothalamus, colDataFH, "female hypothalamus")
 #makepheatmap(dds.female_pituitary, colDataFP, "female pituitary")
@@
 
-```{r candidates, eval = F}
-plotcandidates(dds.female_hypothalamus, "female hypothalamus")
-plotcandidates(dds.female_pituitary, "female pituitary")
-plotcandidates(dds.female_gonad, "female gonad")
-plotcandidates(dds.male_hypothalamus, "male hypothalamus")
-plotcandidates(dds.male_pituitary, "male pituitary")
-plotcandidates(dds.male_gondad, "male gonad")
+```{r candidates, eval = T}
+plotcandidates(dds.female_hypothalamus, colDataFH,"female hypothalamus")
+plotcandidates(dds.female_pituitary, colDataFP, "female pituitary")
+plotcandidates(dds.female_gonad, colDataFG,"female gonad")
+plotcandidates(dds.male_hypothalamus, colDataMH,"male hypothalamus")
+plotcandidates(dds.male_pituitary,colDataMP, "male pituitary")
+plotcandidates(dds.male_gondad,colDataMG, "male gonad")
 ```

### 2019-08-27 analysis/03_DESeq2_characterization.Rmd
now with candidate gene heatmaps

@@
 
-## heamap with minimum pvalue
+## heamap of candidate genes
 
@@
 
-#makepheatmap(dds.female_hypothalamus, colDataFH, "female hypothalamus")
-#makepheatmap(dds.female_pituitary, colDataFP, "female pituitary")
-#makepheatmap(dds.female_gonad, colDataFG, "female gonad")
-#makepheatmap(dds.male_hypothalamus, colDataMH,  "male hypothalamus")
-#makepheatmap(dds.male_pituitary, colDataMP, "male pituitary")
-#makepheatmap(dds.male_gondad, colDataMG, "male gonad")
+makepheatmap(dds.female_hypothalamus, colDataFH, "female hypothalamus")
+makepheatmap(dds.female_pituitary, colDataFP, "female pituitary")
+makepheatmap(dds.female_gonad, colDataFG, "female gonad")
+makepheatmap(dds.male_hypothalamus, colDataMH,  "male hypothalamus")
+makepheatmap(dds.male_pituitary, colDataMP, "male pituitary")
+makepheatmap(dds.male_gondad, colDataMG, "male gonad")
 ```

### 2019-08-27 R/functions.R
now with candidate gene heatmaps

@@ -474,52 +474,51 @@ makepheatmap <- function(mydds, colData, mysubtitle){
   DEGs <- assay(vsd)
   
   DEGs <- as.data.frame(DEGs)
+  DEGs$entrezid <- row.names(DEGs)
   
-  a <- levels(colData$treatment)
-  b <- a
+  candidates <- full_join(geneinfo, DEGs, by = "entrezid")
   
-  for (i in a){
-    for (j in b){
-      if (i != j) {
-        
-        res <- results(mydds, contrast = c("treatment", i, j), independentFiltering = T)
-        results <- as.data.frame(res$padj)
-        padjcolname <- as.character(paste("padj", i, j, sep=""))
-        colnames(results) <- c(padjcolname)
+  candidates <- candidates %>%
+    filter(Name %in% c("AVP", "AVPR1A", "AVPR1B", "AVPR2",
+                       "SERPINA4", "CRH", "CRHR1", "DRD1", "DRD2",
+                       "DRD3",  "DRD4",  "DRD5",  "GABRQ",  "NR3C1",  "HSD11B1a", 
+                       "HSD11B1b", "HSD11B2L",  "HSD11B1L",  "MC2R", "NR3C2",
+                       "OXTR", "POMC",  "AR", "CYP19A1", "ESR1",  "ESR2", "FSHR", 
+                       "FSHB", "GHRL", "GAL", "NPVF",  "NPFFR1", "GNRH1",  "GNRHR", 
+                       "LEPR",  "LHCGR", "PGR", "PRL", "PRLR", "VIP", "VIPR1"))
+  
+  row.names(candidates) <- candidates$Name
+  candidates <- candidates %>% dplyr::select(-row.names, -entrezid, -Name, -geneid)
+  candidates <- candidates %>% drop_na()
+  
+  DEGsmatrix <- as.matrix(candidates)
+  head(DEGsmatrix)
+  
+  DEGsmatrix <- DEGsmatrix - rowMeans(DEGsmatrix)
 
-        DEGs <- cbind(DEGs,results)
-      }
-    }
-    b <- b[-1]  # drop 1st element of second string to not recalculate DEGs
-  }
-  
-  DEGsmatrix <- DEGs
-  DEGsmatrix <- as.matrix(DEGs)
-  padjmin <- rowMins(DEGsmatrix, na.rm = T) 
-  padjmin <- as.data.frame(padjmin)
-  
[…]

### 2019-08-27 R/functions.R
now with prolacin plot only

@@ -463,6 +463,56 @@ plotcandidates <- function(mydds, colData, mysubtitle){
   
 }
 
+
+plotprolactin <- function(mydds, colData, mysubtitle){
+  
+  mydds <- mydds
+  
+  vsd <- vst(mydds, blind=FALSE) # variance stabilized 
+  
+  # make dataframe counts
+  DEGs <- assay(vsd)
+  DEGs <- as.data.frame(DEGs)
+  DEGs$entrezid <- row.names(DEGs)
+  
+  # make dataframe with geneids and names and counts
+  # how to gather: https://tidyr.tidyverse.org/reference/gather.html
+  
+  candidates <- full_join(geneinfo, DEGs, by = "entrezid")
+  head(candidates)
+  
+  candidates <- candidates %>%
+    filter(Name %in% c( "PRL"))
+  row.names(candidates) <- candidates$Name
+  candidates <- candidates %>% dplyr::select(-row.names, -entrezid, -Name, -geneid)
+  candidates <- candidates %>% drop_na()
+  candidates <- as.data.frame(t(candidates))
+  candidates$RNAseqID <- rownames(candidates)
+  candidates <- candidates %>% gather(gene, value, -RNAseqID)  %>% 
+    filter(RNAseqID != "gene")
+  candidates$value <- as.numeric(candidates$value)
+  candidates$V1  <- candidates$RNAseqID
+  
+  candidatecounts <- left_join(candidates, colData, by = "V1")
+  candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
+  candidatecounts$gene <- as.factor(candidatecounts$gene)
+  
+  p1 <- ggplot(candidatecounts, aes(x = treatment, y = value, fill = treatment)) +
+    geom_boxplot() +
+    facet_wrap(~gene, scales = "free") +
+    theme_minimal(base_size = 8) +
+    theme(axis.text.x = element_blank(),
+          legend.position = "bottom") +
+    labs(x = NULL, subtitle = mysubtitle) +
+    guides(fill = guide_legend(nrow = 1))
+  return(p1)
+  
+}
+
+
+
+
+
 ## make pheatmaps 
 # e.g. makepheatmap(dds.female_hypothalamus, "female hypothalamus")
 

### 2019-08-27 analysis/03_DESeq2_characterization.Rmd
rest color palet
now with prolacin plot only

@@
 ```
+
+## plot prolactin
+
+```{r prolactin, eval = T}
+a <- plotprolactin(dds.female_hypothalamus, colDataFH,"female hypothalamus")
+b <- plotprolactin(dds.female_pituitary, colDataFP, "female pituitary")
+c <- plotprolactin(dds.female_gonad, colDataFG,"female gonad")
+d <- plotprolactin(dds.male_hypothalamus, colDataMH,"male hypothalamus")
+e <- plotprolactin(dds.male_pituitary,colDataMP, "male pituitary")
+f <- plotprolactin(dds.male_gondad,colDataMG, "male gonad")
+
+mylegend <- get_legend(f)
+
+plots <- plot_grid(a + theme(legend.position = "none"),
+          b + theme(legend.position = "none"),
+          c + theme(legend.position = "none"),
+          d + theme(legend.position = "none"), 
+          e + theme(legend.position = "none"),
+          f + theme(legend.position = "none"))
+plot_grid(plots, mylegend, nrow = 2, rel_heights = c(1,0.1))
+
+```

### 2019-08-27 analysis/09_egdeR.Rmd
edgeR quick start

---
title: "edgeR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2019-08-27 analysis/09_egdeR.Rmd
first pass works

@@
 knitr::opts_chunk$set(echo = TRUE)
+library(edgeR)
+library(org.Gg.eg.db)
 ```
 
-## R Markdown
+# edgeR: differential expression analysis of digital gene expression data
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+https://www.bioconductor.org/packages/release/bioc/vignettes/edgeR/inst/doc/edgeRUsersGuide.pdf
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+## quick start
 
-```{r cars}
-summary(cars)
+```{r}
+x <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
+colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
+group <- colData$treatment
+
+y <- DGEList(counts=x,group=group)
+y <- calcNormFactors(y)
+design <- model.matrix(~group)
+y <- estimateDisp(y,design)
+
+# To perform quasi-likelihood F-tests:
+fit <- glmQLFit(y,design)
+qlf <- glmQLFTest(fit,coef=2)
+topqlf <- as.data.frame(topTags(qlf))
+
+topqlf$entrezid <- row.names(topqlf)
+left_join(topqlf, geneinfo, by = "entrezid")
 ```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2019-08-27 analysis/09_egdeR.Rmd
now with go and kegg

@@
 library(org.Gg.eg.db)
+library(dplyr)
 ```
@@
 
+# to make go work, I need to swap out row names
+geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+head(geneinfo)
+x$entrezid <- row.names(x)
+x <- left_join(x, geneinfo)
+x <- subset(x, !duplicated(subset(x, select=c(row.names))))
+row.names(x) <- x$row.names
+x <- x[ -c(577:580) ]
+
 y <- DGEList(counts=x,group=group)
@@
 qlf <- glmQLFTest(fit,coef=2)
-topqlf <- as.data.frame(topTags(qlf))
+as.data.frame(topTags(qlf))
+```
 
-topqlf$entrezid <- row.names(topqlf)
-left_join(topqlf, geneinfo, by = "entrezid")
+## Gene ontology (GO) and pathway analysis
+
+```{r}
+go <- goana(qlf, species="Gg", geneid = rownames(x))
+topGO(go, sort="up")
+
+keg <- kegga(qlf, species="Gg")
+topKEGG(keg, sort="up")
 ```

### 2019-08-27 analysis/09_egdeR.Rmd
done with this for now

@@
 colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
-group <- colData$treatment
+group <- factor(paste(colData$sex, colData$tissue, colData$treatment, sep = "."))
 
@@
 qlf <- glmQLFTest(fit,coef=2)
-as.data.frame(topTags(qlf))
+topTags(qlf)
 ```

### 2019-08-27 analysis/03_DESeq2_characterization.Rmd
move chuncks to char or manip script

@@
 ```
+
+## correlation heat maps
+
+```{r correlationheatmaps, eval = F}
+plotcorrelationheatmaps(dds.female_hypothalamus, colData.female_hypothalamus ,"female hypothalamus: hatch, nest bldg., lay and day-before-hatch cluster")
+plotcorrelationheatmaps(dds.female_pituitary, colData.female_pituitary, "female pituitary: clustering suggests an internal clock effect")
+plotcorrelationheatmaps(dds.female_gonad, colData.female_gonad,  "female gonad: highly correlated across incubation and early nestling care")
+plotcorrelationheatmaps(dds.male_hypothalamus, colData.male_hypothalamus, "male hypothalamus: 100% correlated across most parental care stages")
+plotcorrelationheatmaps(dds.male_pituitary, colData.male_pituitary, "male pituitary: non-parental males are most different")
+plotcorrelationheatmaps(dds.male_gondad, colData.male_gondad, "male gonad: non-parental males are most different")
+```

### 2019-08-27 analysis/04_DESeq2_sexes.Rmd
move chuncks to char or manip script

@@
 ```
-
-## Calculate and plot total DEGs
-
-```{r returntotalDEGs, eval = T}
-DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
-DEGs.female_pituitary <- returntotalDEGs(dds.female_pituitary)
-DEGs.female_gonad <- returntotalDEGs(dds.female_gonad)
-DEGs.male_hypothalamus <- returntotalDEGs(dds.male_hypothalamus)
-DEGs.male_pituitary <- returntotalDEGs(dds.male_pituitary)
-DEGs.male_gondad <- returntotalDEGs(dds.male_gondad)
-
-```
-
-```{r plottotalDEGs, eval = T}
-a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus")
-b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary: hatch and day before are most unique")
-c <- plottotalDEGs(DEGs.female_gonad, "female gonad")
-d <- plottotalDEGs(DEGs.male_hypothalamus, "male hypothalamus: many parental contrasts yield 0 DEGs")
-e <- plottotalDEGs(DEGs.male_pituitary, "male pituitary: peaks at hatch")
-f <- plottotalDEGs(DEGs.male_gondad, "male gonad: most parental contrasts yield 0 DEGs")
-
-myDEGs <- plot_grid(a + theme_noaxislabels,
-          d + theme_siggenes,
-          b + theme_noaxislabels,
-          e + theme_noaxislabels,
-          c + theme_noaxislabels,
-          f + theme_noaxislabels,
-          nrow = 3) 
-myDEGs
-
-pdf("../figures/sexes/plottotalDEGs-1.pdf", width = 12, height = 10)
-plot(myDEGs)
-dev.off()
-```
-
-```{r manipVchar}
-g <- plotmanipDEGs(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
-h <- plotmanipDEGs(DEGs.female_pituitary, "Female pituitary", "none")
-i <- plotmanipDEGs(DEGs.female_gonad, "Female gonad",  "none")
-j <- plotmanipDEGs(DEGs.male_hypothalamus, "Male hypothalamus", "none")
-k <- plotmanipDEGs(DEGs.male_pituitary, "Male pituitary", "none")
-l <- plotmanipDEGs(DEGs.male_gondad, "Male gonad", "none")
-
-legend <- get_legend(g)
-
-mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)
-
-mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
-mybarplots2
-
-pdf("../figures/sexes/manipVchar-1.pdf", width = 10, height = 6)
-plot(mybarplots2)
-dev.off()
-
-plotmanipDEGs(DEGs.male_pituitary, "Male pituitary", "bottom")
-```
-
-```{r offspringremoval}
-g <- plotremoval(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
-h <- plotremoval(DEGs.female_pituitary, "Female pituitary", "none")
-i <- plotremoval(DEGs.female_gonad, "Female gonad",  "none")
-j <- plotremoval(DEGs.male_hypothalamus, "Male hypothalamus", "none")
-k <- plotremoval(DEGs.male_pituitary, "Male pituitary", "none")
-l <- plotremoval(DEGs.male_gondad, "Male gonad", "none")
-
-legend <- get_legend(g)
-
-mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)
-
-mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
-mybarplots2
-
-pdf("../figures/sexes/offspringremoval-1.pdf", width = 10, height = 6)
-plot(mybarplots2)
-dev.off()
-
-```
-
-```{r prolongextend}
-g <- plotprolongdelay(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
-h <- plotprolongdelay(DEGs.female_pituitary, "Female pituitary", "none")
-i <- plotprolongdelay(DEGs.female_gonad, "Female gonad",  "none")
-j <- plotprolongdelay(DEGs.male_hypothalamus, "Male hypothalamus", "none")
-k <- plotprolongdelay(DEGs.male_pituitary, "Male pituitary", "none")
-l <- plotprolongdelay(DEGs.male_gondad, "Male gonad", "none")
-
-legend <- get_legend(g)
-
-mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)
-
-mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
-mybarplots2
-
-pdf("../figures/sexes/prolongextend-1.pdf", width = 10, height = 6)
-plot(mybarplots2)
-dev.off()
-
-```
-
-## Calculate and plot principal components
-
-```{r returnPCAs, eval=T}
-pcas.female_hypothalamus <- returnPCAs(dds.female_hypothalamus)
-pcas.female_pituitary <- returnPCAs(dds.female_pituitary)      
-pcas.female_gonad <- returnPCAs(dds.female_gonad)
-pcas.male_hypothalamus <- returnPCAs(dds.male_hypothalamus)
-pcas.male_pituitary <- returnPCAs(dds.male_pituitary)
-pcas.male_gondad <- returnPCAs(dds.male_gondad)
-```
-
-```{r pca, eval=T}
-p1 <- plotPC1(pcas.female_hypothalamus, "female hypothalamus", "26% PC1 variance explained") 
-p2 <- plotPC1(pcas.female_pituitary, "female pituitary", "11% PC1 variance explained")   
-p3 <- plotPC1(pcas.female_gonad, "female gonad", "37% PC1 variance explained") 
-p4 <- plotPC1(pcas.male_hypothalamus, "male hypothalamus", "28% PC1 variance explained") 
-p5 <- plotPC1(pcas.male_pituitary, "male pituitary" , "11% PC1 variance explained") 
-p6 <- plotPC1(pcas.male_gondad, "male gonad", "14% PC1 variance explained")
-
-mylegend <- get_legend(p1)
-
-allPC1s <- plot_grid(p1 + theme_nolegend + ylim(c(-35,35)), 
-                     p4 + theme_nolegend + ylim(c(-35,35)), 
-                     p2 + theme_nolegend + ylim(c(-35,35)),
-                     p5 + theme_nolegend + ylim(c(-35,35)),
-                     p3 + theme_nolegend + ylim(c(-35,35)) ,
-                     p6 + theme_nolegend + ylim(c(-35,35)) ,
-                     ncol = 2)
-
-pc1 <- plot_grid(allPC1s, mylegend,  nrow = 2, rel_heights = c(1.0, 0.15))
-pc1
-
-pdf("../figures/sexes/pca-1.pdf", width = 12, height = 14)
-plot(pc1)
-dev.off()
-
-p11 <- plotPC2(pcas.female_hypothalamus, "female hypothalamus", "9% PC2 variance explained") 
-p12 <- plotPC2(pcas.female_pituitary, "female pituitary", "8% PC2 variance explained")   
-p13 <- plotPC2(pcas.female_gonad, "female gonad", "12 PC2 variance explained") 
-p14 <- plotPC2(pcas.male_hypothalamus, "male hypothalamus", "9% PC2 variance explained") 
-p15 <- plotPC2(pcas.male_pituitary, "male pituitary" , "9% PC2 variance explained") 
-p16 <- plotPC2(pcas.male_gondad, "male gonad", "10% PC2 variance explained")
-
-allPC2 <- plot_grid(p11 + theme_nolegend + ylim(c(-35,35)), 
-                     p14 + theme_nolegend + ylim(c(-35,35)), 
-                     p12 + theme_nolegend + ylim(c(-35,35)),
-                     p15 + theme_nolegend + ylim(c(-35,35)),
-                     p13 + theme_nolegend + ylim(c(-35,35)) ,
-                     p16 + theme_nolegend + ylim(c(-35,35)) ,
-                     ncol = 2)
-
-pc2 <- plot_grid(allPC2, mylegend,  nrow = 2, rel_heights = c(1.0, 0.15))
-pc2
-
-pdf("../figures/sexes/pca-2.pdf", width = 12, height = 14)
-plot(pc2)
-dev.off()
-
-```
-
-## heamap with minimum pvalue
-
-```{r pheatmap, eval = F}
-makepheatmap(dds.female_hypothalamus, a.colData, "female hypothalamus")
-makepheatmap(dds.female_pituitary, "female pituitary")
-makepheatmap(dds.female_gonad, "female gonad")
-makepheatmap(dds.male_hypothalamus, "male hypothalamus")
-makepheatmap(dds.male_pituitary, "male pituitary")
-makepheatmap(dds.male_gondad, "male gonad")        
-```
-
-## make col data for candidate genes and heatmaps
-
-```{r coldata, eval =T}
-colData.female_hypothalamus <- subsetcolData(a.colData,  "female_hypothalamus")
-colData.female_pituitary <- subsetcolData(a.colData,  "female_pituitary" )
-colData.female_gonad <- subsetcolData(a.colData,  "female_gonad" )
-colData.male_hypothalamus <- subsetcolData(a.colData,  "male_hypothalamus" )
-colData.male_pituitary <- subsetcolData(a.colData,  "male_pituitary"  )
-colData.male_gondad <- subsetcolData(a.colData,  "male_gonad")
-```
-
-## candidate genes
-
-```{r candidates, eval =F}
-plotcandidates(dds.female_hypothalamus, colData.female_hypothalamus, "female hypothalamus")
-plotcandidates(dds.female_pituitary, colData.female_pituitary,  "female pituitary")
-plotcandidates(dds.female_gonad, colData.female_gonad,  "female gonad")
-plotcandidates(dds.male_hypothalamus, colData.male_hypothalamus, "male hypothalamus")
-plotcandidates(dds.male_pituitary,  colData.male_pituitary, "male pituitary")
-plotcandidates(dds.male_gondad, colData.male_hypothalamus , "male gonad")
-```
-
-## new heat maps
-
-```{r correlationheatmaps, eval = F}
-plotcorrelationheatmaps(dds.female_hypothalamus, colData.female_hypothalamus ,"female hypothalamus: hatch, nest bldg., lay and day-before-hatch cluster")
-plotcorrelationheatmaps(dds.female_pituitary, colData.female_pituitary, "female pituitary: clustering suggests an internal clock effect")
-plotcorrelationheatmaps(dds.female_gonad, colData.female_gonad,  "female gonad: highly correlated across incubation and early nestling care")
-plotcorrelationheatmaps(dds.male_hypothalamus, colData.male_hypothalamus, "male hypothalamus: 100% correlated across most parental care stages")
-plotcorrelationheatmaps(dds.male_pituitary, colData.male_pituitary, "male pituitary: non-parental males are most different")
-plotcorrelationheatmaps(dds.male_gondad, colData.male_gondad, "male gonad: non-parental males are most different")
-```

### 2019-08-27 analysis/05_DESeq2_manipulate.Rmd
move chuncks to char or manip script

@@
 ```
+
+## offspring removal
+
+```{r offspringremoval}
+g <- plotremoval(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
+h <- plotremoval(DEGs.female_pituitary, "Female pituitary", "none")
+i <- plotremoval(DEGs.female_gonad, "Female gonad",  "none")
+j <- plotremoval(DEGs.male_hypothalamus, "Male hypothalamus", "none")
+k <- plotremoval(DEGs.male_pituitary, "Male pituitary", "none")
+l <- plotremoval(DEGs.male_gondad, "Male gonad", "none")
+
+legend <- get_legend(g)
+
+mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)
+
+mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
+mybarplots2
+
+pdf("../figures/sexes/offspringremoval-1.pdf", width = 10, height = 6)
+plot(mybarplots2)
+dev.off()
+
+```
+
+## prolong extend
+
+```{r prolongextend}
+g <- plotprolongdelay(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
+h <- plotprolongdelay(DEGs.female_pituitary, "Female pituitary", "none")
+i <- plotprolongdelay(DEGs.female_gonad, "Female gonad",  "none")
+j <- plotprolongdelay(DEGs.male_hypothalamus, "Male hypothalamus", "none")
+k <- plotprolongdelay(DEGs.male_pituitary, "Male pituitary", "none")
+l <- plotprolongdelay(DEGs.male_gondad, "Male gonad", "none")
+
+legend <- get_legend(g)
+
+mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)
+
+mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
+mybarplots2
+
+pdf("../figures/sexes/prolongextend-1.pdf", width = 10, height = 6)
+plot(mybarplots2)
+dev.off()
+
+```

### 2019-08-27 analysis/04_DESeq2_sexes.Rmd
now this is actually allows comparison across sexes

@@
 # import "colData" which contains sample information and "countData" which contains read counts
-a.colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
-a.countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
+c.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
+c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
 geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
@@
 # set levels
-levels(a.colData$treatment)
-
-a.colData$treatment <- factor(a.colData$treatment, levels = 
-                             c("control", "bldg", "lay",
-                               "inc.d3", "inc.d9", "inc.d17",
-                               "hatch", "n5", "n9",
-
-                               "m.inc.d3", "m.inc.d8", "m.inc.d9",
-                               "m.inc.d17",  "m.n2"  ,
-                                "prolong", "extend" ))
-
-a.colData$sextissue <- as.factor(paste(a.colData$sex, a.colData$tissue, sep = "_"))
-
-a.colData$lastday <- ifelse(grepl("m.inc.d3|m.inc.d9|m.inc.d17|m.n2", a.colData$treatment), "empty nest", 
-                    ifelse(grepl("m.inc.d8|hatch|extend", a.colData$treatment),"chicks hatch",      
-                     ifelse(grepl("n5", a.colData$treatment),"chicks early",
-                     ifelse(grepl("n9", a.colData$treatment),"chicks later",
-                     ifelse(grepl("control", a.colData$treatment),"control",
-                     ifelse(grepl("bldg", a.colData$treatment),"nest building",
-                     ifelse(grepl("prolong", a.colData$treatment),"eggs delay",
-                      ifelse(grepl("lay", a.colData$treatment),"eggs lay",
-                      ifelse(grepl("inc.d3", a.colData$treatment),"eggs early",
-                     ifelse(grepl("inc.d9", a.colData$treatment),"eggs middle",
-                    ifelse(grepl("inc.d17", a.colData$treatment),"eggs later", NA)))))))))))
-
-a.colData$penultimate <-  ifelse(grepl("extend", a.colData$treatment),"eggs delay",  
-                          ifelse(grepl("m.n2", a.colData$treatment),"chicks hatch",
-                     ifelse(grepl("n5", a.colData$treatment),"chicks early",
-                     ifelse(grepl("n9", a.colData$treatment),"chicks later",
-                     ifelse(grepl("control", a.colData$treatment),"control",
-                     ifelse(grepl("bldg|lay", a.colData$treatment),"nest building",
-                     ifelse(grepl("prolong", a.colData$treatment),"eggs delay",
-                      ifelse(grepl("inc.d3|m.inc.d3", a.colData$treatment),"eggs early",
-                     ifelse(grepl("inc.d9|m.inc.d9|m.inc.d8", a.colData$treatment),"eggs middle",
-                    ifelse(grepl("inc.d17|hatch", a.colData$treatment),"eggs later", NA))))))))))
-
-a.colData$xlabel <- a.colData$treatment
-
-levels(a.colData$xlabel) <-  c("control", "nest.building", "egg.lay",
-                               "eggs.early", "eggs.early.remove", 
-                               "eggs.mid", "eggs.mid.hatch", "eggs.mid.remove",
-                               "eggs.end", "eggs.end.remove",
-                               "chicks.hatch",  "chicks.hatch.remove"  ,
-                               "chicks.mid", "eggs.delay", "eggs.delay.hatch", "chicks.end")
-
-a.colData$lastday <- factor(a.colData$lastday, levels =  c("control", "nest building", "eggs lay",
-                               "eggs early", "eggs middle",  "eggs later","eggs delay",
-                               "chicks hatch", "chicks early",  "chicks later", "empty nest"))
-
-a.colData$penultimate <- factor(a.colData$penultimate, levels =  c("control", "nest building",
-                               "eggs early",  "eggs middle",  "eggs later",  "eggs delay",
-                               "chicks hatch", "chicks early",  "chicks later"))
-
-summary(a.colData[c(7,3,4,5,8,9, 10,11)])
+c.colData$treatment <- factor(c.colData$treatment, levels = 
+                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
+levels(c.colData$treatment)
+c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
+summary(c.colData[c(7,3,4,5,8)])
 ```
@@
 
-```{r dds, cache = T}
-dds.female_hypothalamus <- subsetDESeq(a.colData, a.countData, "female_hypothalamus")
-dds.female_pituitary <- subsetDESeq(a.colData, a.countData, "female_pituitary" )
-dds.female_gonad <- subsetDESeq(a.colData, a.countData, "female_gonad" )
-dds.male_hypothalamus <- subsetDESeq(a.colData, a.countData, "male_hypothalamus" )
-dds.male_pituitary <- subsetDESeq(a.colData, a.countData, "male_pituitary"  )
-dds.male_gondad <- subsetDESeq(a.colData, a.countData, "male_gonad")
+```{r dds}
+dds.hypothalamus <- subsetDESeq2(c.colData,  c.countData, c("female_hypothalamus","male_hypothalamus") )
+dds.pituitary <- subsetDESeq2(c.colData,  c.countData, c("female_pituitary", "male_pituitary" ) )
+dds.gonad <- subsetDESeq2(c.colData,  c.countData, c("female_gonad", "male_gonad") )
 ```
+
+```{r pca}
+hypPCA <- returnPCAs2(dds.hypothalamus)
+plotPC12(hypPCA, "female and male hypothalamus")
+
+pitPCA <- returnPCAs2(dds.pituitary)
+plotPC12(pitPCA, "female and male pituitary")
+
+gonPCA <- returnPCAs2(dds.gonad)
+plotPC12(gonPCA, "female and male gonad")
+```

### 2019-08-27 analysis/04_DESeq2_sexes.Rmd
add lda tutorial

@@
 
-## Starting with all the data
+## DEseq2 analysis on characterization. Looking at treatment AND sex for each tissue
 
@@
 
-## Run DESeq on all subsets of the data
-
 ```{r dds}
@@
 ```
+
+## PCA
 
@@
 ```
+
+## Linear discriminant analysis (LDA)
+
+http://www.sthda.com/english/articles/36-classification-methods-essentials/146-discriminant-analysis-essentials-in-r/
+
+```{r LDA}
+library(caret)
+theme_set(theme_classic())
+
+# Load the data
+data("iris")
+
+# Split the data into training (80%) and test set (20%)
+set.seed(123)
+training.samples <- iris$Species %>%
+  createDataPartition(p = 0.8, list = FALSE)
+train.data <- iris[training.samples, ]
+test.data <- iris[-training.samples, ]
+
+# Normalize the data. Categorical variables are automatically ignored.
+# Estimate preprocessing parameters
+preproc.param <- train.data %>% 
+  preProcess(method = c("center", "scale"))
+
+# Transform the data using the estimated parameters
+train.transformed <- preproc.param %>% predict(train.data)
+test.transformed <- preproc.param %>% predict(test.data)
+
+# LDA analysis
+library(MASS)
+# Fit the model
+model <- lda(Species~., data = train.transformed)
+# Make predictions
+predictions <- model %>% predict(test.transformed)
+# Model accuracy
+mean(predictions$class==test.transformed$Species)
+
+# results
+model
+plot(model)
+
+# make predictions
+predictions <- model %>% predict(test.transformed)
+names(predictions)
+
+# Predicted classes
+head(predictions$class, 6)
+# Predicted probabilities of class memebership.
+head(predictions$posterior, 6) 
+# Linear discriminants
+head(predictions$x, 3)
+
+# LDA plot using ggplot2 
+lda.data <- cbind(train.transformed, predict(model)$x)
+ggplot(lda.data, aes(LD1, LD2)) +
+  geom_point(aes(color = Species))
+
+# model accuracy 
+mean(predictions$class==test.transformed$Species)
+```

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
pitutiary lda, now a fucntion)

--- 2019-08-27 analysis/04_DESeq2_sexes.Rmd
@@
 library("wesanderson")
+library(caret) # LDA analysis
+library(MASS) # LDA analysis
 
@@
 c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
+
 summary(c.colData[c(7,3,4,5,8)])
+
+c.colData <- c.colData %>%
+    mutate(hypothesis = fct_recode(treatment,
+                            "anticipation" = "control",
+                            "anticipation" = "bldg",
+                            "incubation" = "lay",
+                            "incubation" = "inc.d3",
+                            "incubation" = "inc.d9",
+                            "incubation" = "inc.d17",
+                            "hatchling.care" = "hatch",
+                            "hatchling.care" = "n5",
+                            "hatchling.care" = "n9"))
+
 ```
-
-## Run DESeq on all subsets of the data
 
@@
 ```
+
+## Linear discriminant analysis (LDA)
+
+http://www.sthda.com/english/articles/36-classification-methods-essentials/146-discriminant-analysis-essentials-in-r/
+
+```{r LDA}
+theme_set(theme_classic())
+
+LDAanalysis <- function(mydds){
+  # data prep
+  vsd <- vst(mydds, blind=FALSE) # variance stabilized 
+  tvsd <- as.data.frame(t(assay(vsd)))   # get vsd counts and transform
+
+  tvsd <- tvsd[, sample(ncol(tvsd), 20)] # select 20 random genes
+  tvsd$V1 <- row.names(tvsd)             # prep for joining
+
+  colDataPit <- subsetcolData2(c.colData, c("female_pituitary", "male_pituitary"))
+
+  mydata <- left_join(tvsd, colDataPit, by = "V1")  # merge counts and variables
+
+  # Split the data into training (80%) and test set (20%)
+  set.seed(175)
+
+  training.samples <- mydata$treatment %>%
+    createDataPartition(p = 0.8, list = FALSE)
+  train.data <- mydata[training.samples, ]
+  test.data <- mydata[-training.samples, ]
+
+  # Normalize the data. Categorical variables are automatically ignored.
+  # Estimate preprocessing parameters
+  preproc.param <- train.data %>% 
+    preProcess(method = c("center", "scale"))
+
+  # Transform the data using the estimated parameters
+  train.transformed <- preproc.param %>% predict(train.data)
+  test.transformed <- preproc.param %>% predict(test.data)
+
+  # LDA analysis
+
+  # Fit the model
+  print(model <- lda(treatment~ sex, data = train.transformed))
+  # Make predictions
+  print(predictions <- model %>% predict(test.transformed))
+# Model accuracy
+  print(mean(predictions$class==test.transformed$treatment))
+
+  # results
+  print(model)
+
+  # make predictions
+  predictions <- model %>% predict(test.transformed)
+
+  # Predicted classes
+  print(head(predictions$class, 6))
+  # Predicted probabilities of class memebership.
+  print(head(predictions$posterior, 6) )
+  # Linear discriminants
+  print(head(predictions$x, 3) )
+
+  # LDA plot using ggplot2 
+  lda.data <- cbind(train.transformed, predict(model)$x)
+  p<-ggplot(data=lda.data, aes(x=treatment, y=LD1, fill = treatment)) +
+    geom_bar( stat = "summary", fun.y = "mean")
+  p
+
+  # model accuracy 
+  print(mean(predictions$class==test.transformed$treatment))
+}
+
+LDAanalysis(dds.pituitary)
+```

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
now exporting plots

@@
     geom_bar( stat = "summary", fun.y = "mean")
-  p
+  plot(p)
 

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
now with hyp, pit, gon

@@
 
-LDAanalysis <- function(mydds){
+LDAanalysis <- function(mydds, mycolData){
   # data prep
@@
 
-  colDataPit <- subsetcolData2(c.colData, c("female_pituitary", "male_pituitary"))
-
-  mydata <- left_join(tvsd, colDataPit, by = "V1")  # merge counts and variables
+  mydata <- left_join(tvsd, mycolData, by = "V1")  # merge counts and variables
 
@@
 
-LDAanalysis(dds.pituitary)
+colDataHyp <- subsetcolData2(c.colData, c("female_hypothalamus", "male_hypothalamus"))
+colDataPit <- subsetcolData2(c.colData, c("female_pituitary", "male_pituitary"))
+colDataGon <- subsetcolData2(c.colData, c("female_gonad", "male_gonad"))
+
+LDAanalysis(dds.hypothalamus, colDataHyp)
+LDAanalysis(dds.pituitary, colDataPit)
+LDAanalysis(dds.gonad, colDataGon)
 ```

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
now with all genes?

@@
 
-  tvsd <- tvsd[, sample(ncol(tvsd), 20)] # select 20 random genes
+  #tvsd <- tvsd[, sample(ncol(tvsd), 20)] # select 20 random genes
   tvsd$V1 <- row.names(tvsd)             # prep for joining
@@
   # LDA analysis
-
   # Fit the model

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
less text

@@
   # Fit the model
-  print(model <- lda(treatment~ sex, data = train.transformed))
+  model <- lda(treatment~ sex, data = train.transformed)
   # Make predictions
-  print(predictions <- model %>% predict(test.transformed))
+  predictions <- model %>% predict(test.transformed)
   # Model accuracy
-  print(mean(predictions$class==test.transformed$treatment))
+  mean(predictions$class==test.transformed$treatment)
 

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
back to working

@@
   lda.data <- cbind(train.transformed, predict(model)$x)
-  p<-ggplot(data=lda.data, aes(x=treatment, y=LD1, fill = treatment)) +
+  p <- ggplot(data=lda.data, aes(x=treatment, y=LD1, fill = treatment)) +
     geom_bar( stat = "summary", fun.y = "mean")

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
working woohoo, had to delete some bad factors to get LD1 and LD2

--- 2019-08-28 analysis/04_DESeq2_sexes.Rmd
@@
                               c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
-levels(c.colData$treatment)
+
 c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
-
-summary(c.colData[c(7,3,4,5,8)])
 
@@
 
+# drop columsn that are bad for model
+c.colData$bird <- NULL
+c.colData$tissue <- NULL
+c.colData$study <- NULL
+c.colData$group <- NULL
+head(c.colData)
 ```
@@
 
-LDAanalysis <- function(mydds, mycolData){
+LDAanalysis <- function(mydds, mycolData, mysubtitle){
   # data prep
@@
   mydata <- left_join(tvsd, mycolData, by = "V1")  # merge counts and variables
+  mydata$V1 <- NULL
 
@@
   # Fit the model
-  model <- lda(treatment~ sex, data = train.transformed)
+  model <- lda(hypothesis~ ., data = train.transformed)
   # Make predictions
@@
   # Model accuracy
-  mean(predictions$class==test.transformed$treatment)
+  mean(predictions$class==test.transformed$hypothesis)
 
@@
   # Predicted classes
-  print(head(predictions$class, 6))
+  head(predictions$class, 6)
   # Predicted probabilities of class memebership.
-  print(head(predictions$posterior, 6) )
+  head(predictions$posterior, 6) 
   # Linear discriminants
-  print(head(predictions$x, 3) )
+  head(predictions$x, 3)
 
@@
   lda.data <- cbind(train.transformed, predict(model)$x)
-  p<-ggplot(data=lda.data, aes(x=treatment, y=LD1, fill = treatment)) +
-    geom_bar( stat = "summary", fun.y = "mean")
+  p <- ggplot(data = lda.data, aes(LD1, LD2, color = hypothesis, shape = sex)) +
+    geom_point() +
+    labs(subtitle = mysubtitle)
   plot(p)
@@
   # model accuracy 
-  print(mean(predictions$class==test.transformed$treatment))
+  print(mean(predictions$class==test.transformed$hypothesis))
 }
@@
 
-LDAanalysis(dds.hypothalamus, colDataHyp)
-LDAanalysis(dds.pituitary, colDataPit)
-LDAanalysis(dds.gonad, colDataGon)
+LDAanalysis(dds.hypothalamus, colDataHyp, "Hypothalamus")
+LDAanalysis(dds.pituitary, colDataPit, "Pituitary")
+LDAanalysis(dds.gonad, colDataGon, "Gonad")
 ```

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
now with LDA 9 group and 3 groups

@@
   mydata$V1 <- NULL
+  mydata$sextissue <- NULL
 
@@
 ```
+
+```{r LDA2}
+
+LDAanalysis2 <- function(mydds, mycolData, mysubtitle){
+  # data prep
+  vsd <- vst(mydds, blind=FALSE) # variance stabilized 
+  tvsd <- as.data.frame(t(assay(vsd)))   # get vsd counts and transform
+
+  #tvsd <- tvsd[, sample(ncol(tvsd), 20)] # select 20 random genes
+  tvsd$V1 <- row.names(tvsd)             # prep for joining
+
+  mydata <- left_join(tvsd, mycolData, by = "V1")  # merge counts and variables
+  mydata$V1 <- NULL
+  mydata$sextissue <- NULL
+  mydata$hypothesis <- NULL
+
+  # Split the data into training (80%) and test set (20%)
+  set.seed(175)
+
+  training.samples <- mydata$treatment %>%
+    createDataPartition(p = 0.8, list = FALSE)
+  train.data <- mydata[training.samples, ]
+  test.data <- mydata[-training.samples, ]
+
+  # Normalize the data. Categorical variables are automatically ignored.
+  # Estimate preprocessing parameters
+  preproc.param <- train.data %>% 
+    preProcess(method = c("center", "scale"))
+
+  # Transform the data using the estimated parameters
+  train.transformed <- preproc.param %>% predict(train.data)
+  test.transformed <- preproc.param %>% predict(test.data)
+
+  # LDA analysis
+  # Fit the model
+  model <- lda(treatment~ ., data = train.transformed)
+  # Make predictions
+  predictions <- model %>% predict(test.transformed)
+  # Model accuracy
+  mean(predictions$class==test.transformed$treatment)
+
+  # results
+  print(model)
+
+  # make predictions
+  predictions <- model %>% predict(test.transformed)
+
+  # Predicted classes
+  head(predictions$class, 6)
+  # Predicted probabilities of class memebership.
+  head(predictions$posterior, 6) 
+  # Linear discriminants
+  head(predictions$x, 3)
+
+  # LDA plot using ggplot2 
+  lda.data <- cbind(train.transformed, predict(model)$x)
+  p <- ggplot(data = lda.data, aes(LD1, LD2, color = treatment, shape = sex)) +
+    geom_point() +
+    labs(subtitle = mysubtitle)
+  plot(p)
+
+  # model accuracy 
+  print(mean(predictions$class==test.transformed$hypothesis))
+}
+
+LDAanalysis2(dds.hypothalamus, colDataHyp, "Hypothalamus")
+LDAanalysis2(dds.pituitary, colDataPit, "Pituitary")
+LDAanalysis2(dds.gonad, colDataGon, "Gonad")
+
+```

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
updated pca plots

@@
 ```{r pca}
+theme_set(theme_classic())
+
 hypPCA <- returnPCAs2(dds.hypothalamus)
-plotPC12(hypPCA, "female and male hypothalamus")
+plotPC12(hypPCA, "Hypothalamus")
 
 pitPCA <- returnPCAs2(dds.pituitary)
-plotPC12(pitPCA, "female and male pituitary")
+plotPC12(pitPCA, "Pituitary")
 
 gonPCA <- returnPCAs2(dds.gonad)
-plotPC12(gonPCA, "female and male gonad")
+plotPC12(gonPCA, "Gonad")
 ```

### 2019-08-28 R/functions.R

@@ -407,10 +407,9 @@ plotPC12 <- function(pcadata, mysubtitle){
   
   pca12 <- ggplot(pcadata, aes(PC1, PC2, color = treatment, shape = sex)) + 
     geom_point( size = 3) +
-    stat_ellipse() +
-    theme_minimal(base_size = 12) +
-    ylab(paste0("PC2")) +
-    xlab(paste0("PC1")) +
+    #stat_ellipse() +
+    #xlab(paste0("PC1")) +
+    #ylab(paste0("PC2")) +
     labs(subtitle = mysubtitle) +
     #theme(legend.position = "bottom") +
     #guides( fill = guide_legend(order = 2, ncol=2)) +
@@ -418,60 +417,6 @@ plotPC12 <- function(pcadata, mysubtitle){
   pca12
 }  
 
-plotPC1 <- function(pcadata, mysubtitle, myxlab){ 
-  
-  pcadata <- pcadata
-  
-  #percentVar <- round(100 * attr(pcadata, "percentVar"))
-
-  pca1 <- ggplot(pcadata, aes(xlabel, PC1, fill = lastday, color = penultimate)) + 
-    geom_violin() +
-    #geom_point() +
-     theme_minimal(base_size = 12) +
-    ylab(myxlab) +
-    xlab("Parental stages, with increasing days ->") +
-    labs(subtitle = mysubtitle) +
-    scale_fill_manual(values = colorlastday, 
-                      guide = guide_legend(
-                        direction = "horizontal",
-                        label.position = "bottom")) +
-    scale_color_manual(values = colorpenultimate,
-                       guide = guide_legend(
-                         direction = "horizontal",
-                         label.position = "bottom")) +
-    facet_wrap(~study, scales = "free_x")  +
-    guides(col = guide_legend(nrow = 1),
-           fill = guide_legend(nrow = 1)) +
-    theme(panel.grid = element_blank())
-  return(pca1)
-}
-
-
-plotPC2 <- function(pcadata, mysubtitle, myxlab){ 
-    
-    pcadata <- pcadata
-    
-  pca2 <- ggplot(pcadata, aes(xlabel, PC2, fill = treatment)) + 
-    geom_violin() +
-    #geom_point() +
-     theme_minimal(base_size = 12) +
-    ylab(myxlab) +
-    xlab("Parental stages, with increasing days ->") +
-    labs(subtitle = mysubtitle) +
-    scale_fill_manual(#values = colorlastday, 
[…]

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
now with updated plots
clea up

@@
 library(cowplot)
-library(RColorBrewer)
 library(pheatmap)
-library(kableExtra)
 library(viridis)
 library(forcats)
-library("wesanderson")
 library(caret) # LDA analysis
@@
 
-library("BiocParallel")
-register(MulticoreParam(4))
+library(BiocParallel)
+register(MulticoreParam(6))
 
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/sexes/', cache = TRUE)
+knitr::opts_chunk$set(fig.path = '../figures/sexes/', cache = TRUE,  message=F, comment=FALSE, warning=FALSE)
 ```
@@
 
-# drop columsn that are bad for model
+# drop columsn that are bad for model AND not needed for later joining
 c.colData$bird <- NULL
@@
 dds.gonad <- subsetDESeq2(c.colData,  c.countData, c("female_gonad", "male_gonad") )
+```
+
+## Variance stabilized data
+
+```{r vsd}
+
 ```
@@
 plotPC12(gonPCA, "Gonad")
+
 ```

### 2019-08-28 R/functions.R
clea up

@@ -1,39 +1,36 @@
-# subset col for heatmap
-subsetcolData <- function(colData, whichgroup){
+# subset colData to look one tissue in one sex
+subsetcolData <- function(colData, eachgroup){
   
-  # subset to look within one tissue in one sex
   colData <- colData %>%
-    dplyr::filter(sextissue == whichgroup) %>%
+    dplyr::filter(sextissue == eachgroup) %>%
     droplevels()
   row.names(colData) <- colData$V1
   return(colData)
 }
 
-subsetcolData2 <- function(colData, whichgroups){
+# subset to look within one tissue in two sexes
+subsetcolData2 <- function(colData, eachgroup){
   
-  # subset to look within one tissue in two sexes
+  
   colData <- colData %>%
-    dplyr::filter(sextissue %in% whichgroup) %>%
+    dplyr::filter(sextissue %in% eachgroup) %>%
     droplevels()
   row.names(colData) <- colData$V1
   return(colData)
 }
 
 # run DESeq on subset of data
-# e.g. dds <- subsetDESeq("male_hypothalamus")
 subsetDESeq <- function(colData, countData, eachgroup){
   
   colData <- subsetcolData(colData, eachgroup)
   
-  # which counts to save
+  # save counts that match colData
   savecols <- as.character(colData$V1) 
   savecols <- as.vector(savecols) 
-  
-  # save counts that match colData
   countData <- countData %>% dplyr::select(one_of(savecols)) 
   
-  # check that row and col lenghts are equal
-  print(ncol(countData) == nrow(colData))
+  # assert that row and col lenghts are equal
+  stopifnot(ncol(countData) == nrow(colData))
   
   dds <- DESeqDataSetFromMatrix(countData = countData,
                                 colData = colData,
@@ -61,7 +58,7 @@ subsetDESeq2 <- function(colData, countData, eachgroup){
   countData <- countData %>% dplyr::select(one_of(savecols)) 
   
   # check that row and col lenghts are equal
-  print(ncol(countData) == nrow(colData))
+  stopifnot(ncol(countData) == nrow(colData))
   
   dds <- DESeqDataSetFromMatrix(countData = countData,
                                 colData = colData,

### 2019-08-28 R/functions.R
adding total degs back

@@ -145,8 +145,6 @@ plottotalDEGs <- function(myDEGS, mysubtitle){
                                                    "hatch", "n5", "n9"))
 
   totalDEGS <- totalDEGS %>% dplyr::na_if(0)
-
-  print(str(totalDEGS))
   
   allcontrasts <- totalDEGS %>%
     ggplot( aes(V1, V2)) +

### 2019-08-28 R/functions.R
added total degs, but eval is false for now, too long to run

@@ -106,11 +106,9 @@ returnpadj <- function(group1, group2){
 returntotalDEGs <- function(dds){
   
   colData <- c.colData # %>%
-
-  group1 <- levels(colData$treatment)
-
-  a <- group1
-  b <- group1
+  
+  a <- levels(colData$treatment)
+  b <- levels(colData$treatment)
   
   # comapre all contrasts, save to datafrmes
   totalDEGS=data.frame()
@@ -135,7 +133,9 @@ returntotalDEGs <- function(dds){
 ############ plottotalDEGs ############ 
 
 plottotalDEGs <- function(myDEGS, mysubtitle){  
+  
   totalDEGS <- myDEGS
+  
   totalDEGS$V2 <- factor(totalDEGS$V2, levels =  c("control", "bldg", "lay",
                                                    "inc.d3", "inc.d9", "inc.d17",
                                                    "hatch", "n5", "n9"))
@@ -144,7 +144,7 @@ plottotalDEGs <- function(myDEGS, mysubtitle){
                                                    "inc.d3", "inc.d9", "inc.d17",
                                                    "hatch", "n5", "n9"))
 
-  totalDEGS <- totalDEGS %>% dplyr::na_if(0)
+  #totalDEGS <- totalDEGS %>% dplyr::na_if(0)
   
   allcontrasts <- totalDEGS %>%
     ggplot( aes(V1, V2)) +

### 2019-08-28 R/functions.R
now with temporary screenshot plot

@@ -150,7 +150,7 @@ plottotalDEGs <- function(myDEGS, mysubtitle){
     ggplot( aes(V1, V2)) +
     geom_tile(aes(fill = V3)) +
      theme_minimal(base_size = 12) + 
-    geom_text(aes(label = round(V3, 1)), color = "black")+
+    geom_text(aes(label = round(V3, 1)), color = "grey")+
     scale_fill_viridis(na.value="#bdbdbd", 
                       # limits = c(0,7000),
                        option = "C") +

### 2019-08-28 analysis/03_DESeq2_characterization.Rmd
some upates to old code

@@
 
-```{r returntotalDEGs, eval = T}
+```{r returntotalDEGs, eval = T, results = 'hide'}
 DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
@@
 ```{r totalDEGs, eval = T}
-#create list of groups for deseq contrasts
-group1 <- c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9")
-group2 <- group1
-
 a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus")

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
adding total degs back
added total degs, but eval is false for now, too long to run
now with temporary screenshot plot
replace dds with vsd PCA
name to vsd.df instead of DEGs, remove repeated code

@@
 
-```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
+```{r setup, include = FALSE}
 
@@
 
+## total degs
+
+```{r totalDEGs, results = 'hide', eval = F}
+DEGs.pituitary <- returntotalDEGs(dds.pituitary)
+plottotalDEGs(DEGs.pituitary, "Male & Female Pituitary DEGs")
+```
+
 ## Variance stabilized data
@@
 ```{r vsd}
+vsd.hyp <- vst(dds.hypothalamus, blind=FALSE) 
+vsd.pit <- vst(dds.pituitary, blind=FALSE) 
+vsd.gon <- vst(dds.gonad, blind=FALSE)
 
+vsd.hyp.df <- vsd.dataframe(vsd.hyp) 
+vsd.pit.df <- vsd.dataframe(vsd.pit) 
+vsd.gon.df <- vsd.dataframe(vsd.gon) 
 ```

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
replace dds with vsd PCA

@@
 
-hypPCA <- returnPCAs2(dds.hypothalamus)
+hypPCA <- returnPCAs2(vsd.hyp)
 plotPC12(hypPCA, "Hypothalamus")
 
-pitPCA <- returnPCAs2(dds.pituitary)
+pitPCA <- returnPCAs2(vsd.pit)
 plotPC12(pitPCA, "Pituitary")
 
-gonPCA <- returnPCAs2(dds.gonad)
+gonPCA <- returnPCAs2(vsd.gon)
 plotPC12(gonPCA, "Gonad")

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
move theme to theme

@@
 ```{r pca}
-theme_set(theme_classic())
-
 hypPCA <- returnPCAs2(vsd.hyp)
@@
 plotPC12(gonPCA, "Gonad")
-
 ```
@@
 ```{r LDA}
-theme_set(theme_classic())
-
 LDAanalysis <- function(mydds, mycolData, mysubtitle){

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
now with LDA functions in functions

@@
 ```{r LDA}
-LDAanalysis <- function(mydds, mycolData, mysubtitle){
-  # data prep
-  vsd <- vst(mydds, blind=FALSE) # variance stabilized 
-  tvsd <- as.data.frame(t(assay(vsd)))   # get vsd counts and transform
-
-  #tvsd <- tvsd[, sample(ncol(tvsd), 20)] # select 20 random genes
-  tvsd$V1 <- row.names(tvsd)             # prep for joining
-
-  mydata <- left_join(tvsd, mycolData, by = "V1")  # merge counts and variables
-  mydata$V1 <- NULL
-  mydata$sextissue <- NULL
-
-  # Split the data into training (80%) and test set (20%)
-  set.seed(175)
-
-  training.samples <- mydata$treatment %>%
-    createDataPartition(p = 0.8, list = FALSE)
-  train.data <- mydata[training.samples, ]
-  test.data <- mydata[-training.samples, ]
-
-  # Normalize the data. Categorical variables are automatically ignored.
-  # Estimate preprocessing parameters
-  preproc.param <- train.data %>% 
-    preProcess(method = c("center", "scale"))
-
-  # Transform the data using the estimated parameters
-  train.transformed <- preproc.param %>% predict(train.data)
-  test.transformed <- preproc.param %>% predict(test.data)
-
-  # LDA analysis
-  # Fit the model
-  model <- lda(hypothesis~ ., data = train.transformed)
-  # Make predictions
-  predictions <- model %>% predict(test.transformed)
-  # Model accuracy
-  mean(predictions$class==test.transformed$hypothesis)
-
-  # results
-  print(model)
-
-  # make predictions
-  predictions <- model %>% predict(test.transformed)
-
-  # Predicted classes
-  head(predictions$class, 6)
-  # Predicted probabilities of class memebership.
-  head(predictions$posterior, 6) 
-  # Linear discriminants
-  head(predictions$x, 3)
-
-  # LDA plot using ggplot2 
-  lda.data <- cbind(train.transformed, predict(model)$x)
-  p <- ggplot(data = lda.data, aes(LD1, LD2, color = hypothesis, shape = sex)) +
-    geom_point() +
-    labs(subtitle = mysubtitle)
-  plot(p)
-
-  # model accuracy 
-  print(mean(predictions$class==test.transformed$hypothesis))
-}
-
+# prep col data for LDA
 colDataHyp <- subsetcolData2(c.colData, c("female_hypothalamus", "male_hypothalamus"))
@@
 
-LDAanalysis(dds.hypothalamus, colDataHyp, "Hypothalamus")
-LDAanalysis(dds.pituitary, colDataPit, "Pituitary")
-LDAanalysis(dds.gonad, colDataGon, "Gonad")
+# LDA for treatment (9 groups)
+LDA.hyp1 <- LDAdata.treatment(vsd.hyp, colDataHyp)
+LDA.pit1 <- LDAdata.treatment(vsd.pit, colDataPit)
+LDA.gon1 <- LDAdata.treatment(vsd.gon, colDataGon)
+
+LDAplot.treatment(LDA.hyp1 , "Hypothalamus")
+LDAplot.treatment(LDA.pit1 ,  "Pituitary")
+LDAplot.treatment(LDA.gon1 , "Gonad")
+
+# LDA for hypothesis (3 groups)
+
+LDA.hyp1 <- LDAdata.hypothesis(vsd.hyp, colDataHyp)
+LDA.pit1 <- LDAdata.hypothesis(vsd.pit, colDataPit)
+LDA.gon1 <- LDAdata.hypothesis(vsd.gon, colDataGon)
+
+LDAplot.hypothesis(LDA.hyp1 , "Hypothalamus")
+LDAplot.hypothesis(LDA.pit1 ,  "Pituitary")
+LDAplot.hypothesis(LDA.gon1 , "Gonad")
 ```
-
-```{r LDA2}
-
-LDAanalysis2 <- function(mydds, mycolData, mysubtitle){
-  # data prep
-  vsd <- vst(mydds, blind=FALSE) # variance stabilized 
-  tvsd <- as.data.frame(t(assay(vsd)))   # get vsd counts and transform
-
-  #tvsd <- tvsd[, sample(ncol(tvsd), 20)] # select 20 random genes
-  tvsd$V1 <- row.names(tvsd)             # prep for joining
-
-  mydata <- left_join(tvsd, mycolData, by = "V1")  # merge counts and variables
-  mydata$V1 <- NULL
-  mydata$sextissue <- NULL
-  mydata$hypothesis <- NULL
-
-  # Split the data into training (80%) and test set (20%)
-  set.seed(175)
-
-  training.samples <- mydata$treatment %>%
-    createDataPartition(p = 0.8, list = FALSE)
-  train.data <- mydata[training.samples, ]
-  test.data <- mydata[-training.samples, ]
-
-  # Normalize the data. Categorical variables are automatically ignored.
-  # Estimate preprocessing parameters
-  preproc.param <- train.data %>% 
-    preProcess(method = c("center", "scale"))
-
-  # Transform the data using the estimated parameters
-  train.transformed <- preproc.param %>% predict(train.data)
-  test.transformed <- preproc.param %>% predict(test.data)
-
-  # LDA analysis
-  # Fit the model
-  model <- lda(treatment~ ., data = train.transformed)
-  # Make predictions
-  predictions <- model %>% predict(test.transformed)
-  # Model accuracy
-  mean(predictions$class==test.transformed$treatment)
-
-  # results
-  print(model)
-
-  # make predictions
-  predictions <- model %>% predict(test.transformed)
-
-  # Predicted classes
-  head(predictions$class, 6)
-  # Predicted probabilities of class memebership.
-  head(predictions$posterior, 6) 
-  # Linear discriminants
-  head(predictions$x, 3)
-
-  # LDA plot using ggplot2 
-  lda.data <- cbind(train.transformed, predict(model)$x)
-  p <- ggplot(data = lda.data, aes(LD1, LD2, color = treatment, shape = sex)) +
-    geom_point() +
-    labs(subtitle = mysubtitle)
-  plot(p)
-
-  # model accuracy 
-  print(mean(predictions$class==test.transformed$hypothesis))
-}
-
-LDAanalysis2(dds.hypothalamus, colDataHyp, "Hypothalamus")
-LDAanalysis2(dds.pituitary, colDataPit, "Pituitary")
-LDAanalysis2(dds.gonad, colDataGon, "Gonad")
-
-```

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
werid FALSE is everywhere

@@
 
-```{r setup, include = FALSE}
+```{r setup}
 
@@
 
-# set levels
+# set treatment levels
 c.colData$treatment <- factor(c.colData$treatment, levels = 
@@
 
+# craete variable that will be critical for subset later on
 c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
 
+# ceate new variable for hypothesis testing
 c.colData <- c.colData %>%

### 2019-08-28 R/functions.R
now with new color scheme from ggtext

@@ -676,14 +676,36 @@ LDAdata.treatment <- function(vsd, mycolData, mypredictor){
 
 ######### LDAplot.treatment ######### 
 
-LDAplot.treatment <- function(LDAdata, mysubtitle){
-  
-  # LDA plot using ggplot2 
+LDAplot.treatment <- function(LDAdata, mytitle, mysubtitle, myxlab, myylab){
   p <- ggplot(data = LDAdata, aes(LD1, LD2, color = treatment, shape = sex)) +
-    geom_point() +
-    labs(subtitle = mysubtitle)
+    geom_point(size = 1.5) +
+    labs(title = mytitle,
+         subtitle = mysubtitle,
+         x = myxlab,
+         y = myylab) +
+    scale_color_manual(name = "parental stage",
+                       values = c("control" = "#F8766D",
+                                  "bldg" = "#D39200" ,
+                                  "lay"  = "#93AA00",
+                                  "inc.d3" =  "#00BA38" ,
+                                  "inc.d9" = "#00C19F",
+                                  "inc.d17" = "#00B9E3",
+                                  "hatch"  = "#619CFF",
+                                  "n5"   = "#DB72FB",
+                                  "n9" = "#FF61C3"),
+                       labels = c("control" = "<span style='color:#F8766D'>control</span>",
+                                  "bldg" = "<span style='color:#D39200'>nest bulding</span>",
+                                  "lay" = "<span style='color:#93AA00'>egg laid</span>",
+                                  "inc.d3" = "<span style='color:#00BA38'>incubation day 3</span>", 
+                                  "inc.d9" = "<span style='color:#00B9E3'>incubation day 9</span>",
+                                  "inc.d17" = "<span style='color:#00B9E3'>incubation day 17</span>",
+                                  "hatch" = "<span style='color:#619CFF'>chicks hatch</span>", 
+                                  "n5" = "<span style='color:#DB72FB'>nestling care day 5</span>",
+                                  "n9" = "<span style='color:#FF61C3'>nestling care day 9</span>")) +
+    theme(legend.text = element_markdown(size = 8))
   plot(p)
-}
+} 
+
 
 ######### LDAdata.hypothesis ######### 
 
@@ -756,11 +778,20 @@ LDAdata.hypothesis <- function(vsd, mycolData, mypredictor){
 
 ######### LDAplot.hypothesis ######### 
 
-LDAplot.hypothesis <- function(LDAdata, mysubtitle){
[…]

### 2019-08-28 analysis/04_DESeq2_sexes.Rmd
now with new color scheme from ggtext

@@
 
-```{r setup}
+```{r setup, include = FALSE}
 
@@
 library(MASS) # LDA analysis
+library(ggtext) # for pretty plots
 
@@
 
-```{r LDA}
+```{r LDAdfs}
 # prep col data for LDA
@@
 
-LDAplot.treatment(LDA.hyp1 , "Hypothalamus")
-LDAplot.treatment(LDA.pit1 ,  "Pituitary")
-LDAplot.treatment(LDA.gon1 , "Gonad")
-
 # LDA for hypothesis (3 groups)
 
-LDA.hyp1 <- LDAdata.hypothesis(vsd.hyp, colDataHyp)
-LDA.pit1 <- LDAdata.hypothesis(vsd.pit, colDataPit)
-LDA.gon1 <- LDAdata.hypothesis(vsd.gon, colDataGon)
+LDA.hyp2 <- LDAdata.hypothesis(vsd.hyp, colDataHyp)
+LDA.pit2 <- LDAdata.hypothesis(vsd.pit, colDataPit)
+LDA.gon2 <- LDAdata.hypothesis(vsd.gon, colDataGon)
 
-LDAplot.hypothesis(LDA.hyp1 , "Hypothalamus")
-LDAplot.hypothesis(LDA.pit1 ,  "Pituitary")
-LDAplot.hypothesis(LDA.gon1 , "Gonad")
 ```
+
+```{r LDA, fig.width = 9}
+
+# figures for both of the above
+
+# treatment
+a <- LDAplot.treatment(LDA.hyp1 , "Hypothalamus model 1", "parental state ~ .  0.277 pred. acc.",
+                        "LD1, F = 7.1504779", "LD2, F = 2.4219585")
+
+b <- LDAplot.treatment(LDA.pit1 ,  "Pituitary model 1", "parental state ~ .  0.405 pred. acc.",
+                        "LD1, F = 9.6144160", "LD2, F = 3.7773314")
+
+c <- LDAplot.treatment(LDA.gon1 , "Gonad model 1", "parental state ~ .  has 0.297 pred. acc.",
+                        "LD1, F = 6.137500", "LD2, F = 3.320437")
+
+c <- c + theme(legend.direction = "horizontal")
+mylegend <- get_legend(c)
+
+temp <- plot_grid(a + theme(legend.position = "none"),
+                  b + theme(legend.position = "none"),
+                  c + theme(legend.position = "none"), nrow = 1)
+
+plot_grid(temp, mylegend, nrow = 2, rel_heights = c(1,0.3))
+
+# hypothesis
+
+d <- LDAplot.hypothesis(LDA.hyp2 , "Hypothalamus model 2", "parental stage ~ .  0.432 pred. acc.",
+                        "LD1, F = 7.386458", "LD2, F = 3.720794")
+
+e <- LDAplot.hypothesis(LDA.pit2 ,  "Pituitary model 2", "parental stage ~ .  0.473 pred. acc.",
+                        "LD1, F = 7.846659", "LD2, F = 5.076861")
+
+f <- LDAplot.hypothesis(LDA.gon2 , "Gonad model 3", "parental stage ~ .  has 0.368 pred. acc.",
+                        "LD1, F = 6.029069", "LD2, F = 3.837591")
+
+f <- f + theme(legend.direction = "horizontal")
+mylegend <- get_legend(f)
+
+temp <- plot_grid(d + theme(legend.position = "none"),
+                  e + theme(legend.position = "none"),
+                  f + theme(legend.position = "none"), nrow = 1)
+
+plot_grid(temp, mylegend, nrow = 2, rel_heights = c(1,0.3))
+```

### 2019-09-03 analysis/08_wgcna.Rmd
only looking at pituitary

@@
 
-## Females
-
-```{r wgcna-female, message=FALSE, warning=FALSE, cache=TRUE}
-subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary"),
-            mytitle = "Female sample dendrogram with trait heatmap")
-
-```
-
-**Fig 2. Female characterization dendrogram with trait heatmap.**. 
-Most pituitary samples form a single cluster. More hypothalamus samples cluster with gonads than with pituitary.
-
-## Males
-
-```{r wgcna-male, message=FALSE, warning=FALSE, cache=TRUE}
-subsetWGCNA(c("male.gonad", "male.hypothalamus", "male.pituitary"),
-            mytitle = "Male sample dendrogram with trait heatmap")
-
-```
-
-**Fig 3. Male characterization dendrogram with trait heatmap.**. 
- As with females, most pituitary samples form a single cluster, while more hypothalamus samples cluster with gonads than with pituitary.
-
-## gonads
-
-```{r wgcna-gon, message=FALSE, warning=FALSE, cache=TRUE}
-subsetWGCNA(c("female.gonad", "male.gonad"),
-            mytitle = "Gonad dendrogram with trait heatmap")
-```
-
-**Fig 3. Gonad characterization dendrogram with trait heatmap.**. 
-Male and female samples cluster together, but with large, sex-specific regions.... maybe there is a strong treatment effect here.
-
-## hypothalamus
-
-```{r wgcna-hyp, message=FALSE, warning=FALSE, cache=TRUE}
-subsetWGCNA(c("male.hypothalamus", "female.hypothalamus"),
-            mytitle = "Hypothalamus dendrogram with trait heatmap")
-```
-
-Male and female samples cluster together, but with very small, sex-specific regions.... maybe there is a small treatment effect here.
-
 ## pituitary (not working)
@@
 
-#subsetWGCNA(c("female.pitutary", "male.pitutary"),
-#            mytitle = "Pituitary dendrogram with trait heatmap")
+subsetWGCNA(c("female.pitutary", "male.pitutary"),
+            mytitle = "Pituitary dendrogram with trait heatmap")
 ```
-
-## female hypothalamus
-
-```{r wgcna-f-hyp, message=FALSE, warning=FALSE, cache=TRUE}
-subsetWGCNA(c("female.hypothalamus"),
-            mytitle = "Female hypothalamus sample dendrogram")
-```
-
-## female gonad
-
-```{r wgcna-f-gon, message=FALSE, warning=FALSE, cache=TRUE}
-subsetWGCNA(c("female.gonad"),
-            mytitle = "Female gonad sample dendrogram")
-```

### 2019-09-03 analysis/08_wgcna.Rmd
now with just pit, also, new factor variable

@@
 
-# WGCNA analysis, first pass
+# WGCNA analysis, second pass
 
@@
 
-## All characterization data
-
-```{r wgcna, message=FALSE, warning=FALSE, cache=TRUE}
-
-subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary",
-                           "male.gonad", "male.hypothalamus", "male.pituitary"),
-            mytitle = "Characterization dendrogram with trait heatmap")
-
-```
-
-**Fig 1. Characterization dendrogram with trait heatmap.**. 
- Most hypothalamus and pituitary samples separte form the gonad samples, but some are clusters contain samples from all tissues.
-
 ## pituitary (not working)
@@
 ```{r wgcna-pit, message=FALSE, warning=FALSE, cache=TRUE}
-
-subsetWGCNA(c("female.pitutary", "male.pitutary"),
-            mytitle = "Pituitary dendrogram with trait heatmap")
+subsetWGCNA(c("F.pit", "M.pit"), 2000000, "Pituitary dendrogram with trait heatmap")
 ```

### 2019-09-04 analysis/03_DESeq2_characterization.Rmd
now with PRL-related genes

@@
 dds.male_gondad <- subsetDESeq(c.colData,  c.countData, "male_gonad")
-```
-
-## Calculate and plot total DEGs
-
-```{r returntotalDEGs, eval = T, results = 'hide'}
-DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
-DEGs.female_pituitary <- returntotalDEGs(dds.female_pituitary)
-DEGs.female_gonad <- returntotalDEGs(dds.female_gonad)
-DEGs.male_hypothalamus <- returntotalDEGs(dds.male_hypothalamus)
-DEGs.male_pituitary <- returntotalDEGs(dds.male_pituitary)
-DEGs.male_gondad <- returntotalDEGs(dds.male_gondad)
-```
-
-```{r totalDEGs, eval = T}
-a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus")
-b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary")
-c <- plottotalDEGs(DEGs.female_gonad, "female gonad")
-d <- plottotalDEGs(DEGs.male_hypothalamus, "male hypothalamus")
-e <- plottotalDEGs(DEGs.male_pituitary, "male pituitary")
-f <- plottotalDEGs(DEGs.male_gondad, "male gonad")
-
-plot_grid(a + theme(legend.position = "none"),
-          b + theme(legend.position = "none"),
-          c,
-          d + theme(legend.position = "none"),
-          e + theme(legend.position = "none"),
-          f,
-          nrow = 2, rel_widths = c(0.3, 0.3, 0.4)) 
 ```
@@
 
+## plot genes in a  PRL WGCNA module
+
+```{r wgcna, eval = T}
+plotWGCNAcandidates(dds.female_pituitary, colDataFP, "female pituitary")
+plotWGCNAcandidates(dds.male_pituitary,colDataMP, "male pituitary")
+
+```
+
 ## correlation heat maps

### 2019-09-04 R/functions.R

@@ -369,6 +369,12 @@ returnPCAs <- function(vsd){
   
   print("PC4 ~ treatment, data=pcadata")
   print(summary(aov(PC4 ~ treatment, data=pcadata))) 
+  
+  print("PC5 ~ treatment, data=pcadata")
+  print(summary(aov(PC5 ~ treatment, data=pcadata))) 
+  
+  print("PC6 ~ treatment, data=pcadata")
+  print(summary(aov(PC6 ~ treatment, data=pcadata))) 
   return(pcadata)
 }
 
@@ -391,11 +397,18 @@ returnPCAs2 <- function(vsd){
   print("PC2 ~ treatment * sex, data=pcadata")
   print(summary(aov(PC2 ~ treatment * sex, data=pcadata))) 
   
-  print("PC2 ~ treatment * sex, data=pcadata")
+  print("PC3 ~ treatment * sex, data=pcadata")
   print(summary(aov(PC3 ~ treatment * sex, data=pcadata)))
   
-  print("PC2 ~ treatment * sex, data=pcadata")
-  print(summary(aov(PC4 ~ treatment * sex, data=pcadata))) 
+  print("PC4 ~ treatment * sex, data=pcadata")
+  print(summary(aov(PC4 ~ treatment * sex, data=pcadata)))
+  
+  
+  print("PC5 ~ treatment * sex, data=pcadata")
+  print(summary(aov(PC5 ~ treatment * sex, data=pcadata)))
+  
+  print("PC6 ~ treatment * sex, data=pcadata")
+  print(summary(aov(PC6 ~ treatment * sex, data=pcadata))) 
   return(pcadata)
 }
 
@@ -518,6 +531,56 @@ plotprolactin <- function(vsd.df, colData, mysubtitle){
   return(p1)
 }
 
+
+###### plot wgcna candidates
+
+plotWGCNAcandidates <- function(vsd, mygenelist, colData, mysubtitle){
+  
+  # make dataframe with geneids and names and counts
+  # how to gather: https://tidyr.tidyverse.org/reference/gather.html
+  
+  vsd.df <- as.data.frame(assay(vsd))
+  vsd.df$entrezid <- row.names(vsd.df)
+  
+  candidates <- full_join(geneinfo, vsd.df, by = "entrezid")
+ 
+  # select genes of interst
+  
+  candidates <- candidates %>%
+    filter(entrezid %in% mygenelist) %>% droplevels()
+  
[…]

### 2019-09-04 analysis/04_DESeq2_sexes.Rmd
now with annotated image
now with PRL-related genes

@@
 
-f <- LDAplot.hypothesis(LDA.gon2 , "Gonad model 3", "parental stage ~ .  has 0.368 pred. acc.",
+f <- LDAplot.hypothesis(LDA.gon2 , "Gonad model 2", "parental stage ~ .  has 0.368 pred. acc.",
                         "LD1, F = 6.029069", "LD2, F = 3.837591")
@@
 ```
+
+## plot genes in a  PRL WGCNA module
+
+```{r wgcna, eval = T, fig.width=5.25, fig.height=3}
+plotWGCNAcandidates(vsd.pit, c("NP_990797.2"), colDataPit, "PRL expression")
+
+plotWGCNAcandidates(vsd.pit,
+                    c("NP_990797.2","NP_001004392.2", "NP_001004403.1", "NP_001006255.2", "NP_001006274.1",
+                           "NP_989500.1", "XP_001234815.2", "XP_015149152.1", "XP_416110.5"),
+                    colDataPit, "9 of 102 genes in a WGCNA module containting gene of interest")
+
+plotWGCNAcandidates(vsd.pit,
+                    c("XP_428110.3","NP_001004392.2",  "XP_015142961.1", "XP_015139636.1",
+                      "XP_015137408.1", "XP_004938696.1", "NP_990620.1", "NP_001170793.2", "NP_001026250.1"),
+                    colDataPit, "Prolaction receptor: 9 of 147 genes in a WGCNA module containting PRLR")
+```

### 2019-09-04 analysis/08_WGCNA_2.R

@@ -8,6 +8,8 @@
 
 # Load the WGCNA package
 library(WGCNA)
+library(magrittr) # for %<>%
+
 # The following setting is important, do not omit.
 
 # Allow multi-threading within WGCNA. This helps speed up certain calculations.
@@ -16,12 +18,6 @@ library(WGCNA)
 # Caution: skip this line if you run RStudio or other third-party R environments.
 # See note above.
 
-# Load the data saved in the first part
-lnames = load(file = "../results/08_WGCNA_1.RData");
-#The variable lnames contains the names of loaded variables.
-lnames
-
-str(datTraits)
 
 #=====================================================================================
 #
@@ -45,7 +41,7 @@ plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
 text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
      labels=powers,cex=cex1,col="red");
 # this line corresponds to using an R^2 cut-off of h
-abline(h=0.90,col="red")
+abline(h=0.60,col="red")
 # Mean connectivity as a function of the soft-thresholding power
 plot(sft$fitIndices[,1], sft$fitIndices[,5],
      xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
@@ -62,15 +58,15 @@ text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
 
 enableWGCNAThreads()
 
-bwnet = blockwiseModules(datExpr, maxBlockSize = 2000,
-                         power = 6, TOMType = "unsigned", minModuleSize = 30,
-                        # reassignThreshold = 0, 
-                         #mergeCutHeight = 0.25,
-                        # numericLabels = TRUE,
-                         #saveTOMs = TRUE,
-                        # saveTOMFileBase = "femaleMouseTOM-blockwise",
-                         verbose = 3)
 
+# this is my attempt to solve an error
+datExpr %<>% mutate_if(is.integer,as.numeric)
+head(str(datExpr))
+
+net = blockwiseModules(datExpr, power = 10,
+                                verbose = 5)
+names(net)
+net$colors
 
 #=====================================================================================
 #
@@ -79,13 +75,36 @@ bwnet = blockwiseModules(datExpr, maxBlockSize = 2000,
 #=====================================================================================
 
 
-# Load the results of single-block analysis
[…]

### 2019-09-04 R/functions.R
now with more candidate genes

@@ -570,14 +570,15 @@ plotWGCNAcandidates <- function(vsd, mygenelist, colData, mysubtitle){
   p1 <- ggplot(candidateST, aes(x = as.numeric(treatment), y = expression, color = sex)) +
     geom_point() +
     geom_smooth(se = FALSE) +
-    facet_wrap(~Name, scales = "free") +
+    facet_wrap(~Name, scales = "free_y") +
     theme_rmh() +
-    theme(legend.position = "bottom") +
+    theme(legend.position = "bottom",
+          axis.text.x = element_text(angle = 60,  hjust=1)) +
     guides(fill = guide_legend(nrow = 1)) +
-    theme(axis.text.x = element_blank(),
-          axis.text.y = element_blank()) +
     labs(x = "Time", y = "Gene expression",
-         subtitle = mysubtitle)
+         subtitle = mysubtitle) +
+    scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
+                       labels=charlevels)
   return(p1)
   
 }

### 2019-09-04 analysis/04_DESeq2_sexes.Rmd
now with more candidate genes

@@
 
-```{r wgcna, eval = T, fig.width=5.25, fig.height=3}
+```{r wgcna, eval = T, fig.width=6, fig.height=7}
 plotWGCNAcandidates(vsd.pit, c("NP_990797.2"), colDataPit, "PRL expression")
@@
                     colDataPit, "Prolaction receptor: 9 of 147 genes in a WGCNA module containting PRLR")
+
+plotWGCNAcandidates(vsd.pit, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
+                               "NP_001131120.1", "NP_001186037.1", "NP_989871.1",
+                               "NP_990797.2", "XP_015132722.1"), colDataPit, "Pitutary: CISH, SOCS, PRL, PRLR")
+
+plotWGCNAcandidates(vsd.hyp, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
+                               "NP_001131120.1", "NP_001186037.1", "NP_989871.1",
+                               "NP_990797.2", "XP_015132722.1"), colDataHyp, "Hypothalamus: CISH, SOCS, PRL, PRLR")
+
+plotWGCNAcandidates(vsd.gon, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
+                               "NP_001131120.1", "NP_001186037.1", "NP_989871.1",
+                               "NP_990797.2", "XP_015132722.1"), colDataGon, "Gonad: CISH, SOCS, PRL, PRLR")
 ```

### 2019-09-04 analysis/04_DESeq2_sexes.Rmd
with all 102 prolactin genes

@@
 library(ggtext) # for pretty plots
+library(ctsGE)
 
@@
 
-```{r wgcna, eval = T, fig.width=6, fig.height=7}
+```{r wgcna, eval = T, fig.width=12, fig.height=6}
 plotWGCNAcandidates(vsd.pit, c("NP_990797.2"), colDataPit, "PRL expression")
@@
                            "NP_989500.1", "XP_001234815.2", "XP_015149152.1", "XP_416110.5"),
-                    colDataPit, "9 of 102 genes in a WGCNA module containting gene of interest")
+                    colDataPit, "9 of 102 genes in a WGCNA module containting Prolactin")
 
@@
                                "NP_990797.2", "XP_015132722.1"), colDataGon, "Gonad: CISH, SOCS, PRL, PRLR")
+
+prl36 <-  PRL_associated$entrezid[1:57]
+prl72 <-  PRL_associated$entrezid[58:102]
+
+plotWGCNAcandidates(vsd.pit, prl36, colDataPit, "A-K genes with PRL pattern in the pituitary")
+plotWGCNAcandidates(vsd.pit, prl72, colDataPit, "K-Z genes with PRL pattern in the pituitary")
+
+str(PRL_associated)
+
 ```
+
+```{r ctsGE}
+rts <- assay(vsd.pit)
+head(rts)
+
+prts <- PreparingTheIndexes(x = rts, min_cutoff=0.5, max_cutoff=0.7, mad.scale = TRUE)
+```

### 2019-09-04 analysis/08_WGCNA_1.R
no longer a function

@@ -14,7 +14,6 @@ library(dplyr)
 library(magrittr)
 library(forcats)
 
-subsetWGCNA <- function(whichgroups, cutoff, mytitle){
 
 # read the count data
 countData <- read.csv("../results/00_countData_characterization.csv", row.names = 1)
@@ -61,7 +60,7 @@ colData$sextissue <-  as.factor(paste(colData$sex, colData$tissue, sep = "."))
 levels(colData$sextissue)
 
 # subset to samples
-colData <- colData %>% dplyr::filter(sextissue %in% whichgroups) 
+colData <- colData %>% dplyr::filter(sextissue %in% c("F.pit", "M.pit")) 
 
 # which counts to save
 savecols <- as.character(colData$sample) 
@@ -130,7 +129,7 @@ sampleTree = hclust(dist(datExpr0), method = "average");
 
 plot(sampleTree, main = "Sample clustering to detect outliers", 
      sub="", xlab="", cex.main = 1, cex = 0.4)
-abline(h = cutoff, col = "red");
+abline(h = 2000000, col = "red");
 
 
 
@@ -144,7 +143,7 @@ abline(h = cutoff, col = "red");
 # Plot a line to show the cut
 
 # Determine cluster under the line
-clust = cutreeStatic(sampleTree, cutHeight = cutoff, minSize = 3)
+clust = cutreeStatic(sampleTree, cutHeight = 2000000, minSize = 3)
 table(clust)
 
 # clust 1 contains the samples we want to keep.
@@ -190,9 +189,7 @@ traitColors = numbers2colors(datTraits, signed = TRUE);
 # Plot the sample dendrogram and the colors underneath.
 plotDendroAndColors(sampleTree2, traitColors,
                     groupLabels = names(datTraits), 
-                    main = mytitle,
+                    main = "Pituitary dendrogram with trait heatmap",
                     cex.dendroLabels = 0.4
                     )
-
-}
 

### 2019-09-05 docs/SICB2020.Rmd
sicb abstract

---
title: "Peaks and valleys of prolatin-related gene expression during the pigeon parental care cycle"
author: "Dr. Rayna M Harris, Postdoc, B3 Lab, UC Davis @raynamharris"
output:
  ioslides_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,  message = F, results = F)
library(cowplot)
```

##

Parental care of offspring is essential to maximize fitness in many species throughout the animal kingdom. As a result, new parents undergo major changes in physiology and behavior to promote offspring survival in predictable and unpredictable conditions. While much is known about neuroendocrine mechanisms modulating these changes, we know less about genomic mechanisms driving these changes in male and female parents.

```{r, fig.width = 5,}
p  <- ggdraw() + draw_image("../figures/images/PigeonHPGaxis.png") 
p
```

##

To fill this gap, our team characterized gene expression states of the hypothalamus, pituitary, and gonads of mothers and fathers of the socially monogamous, bi-parental rock dove (Columba livia) at multiple stages of parenting including nest building, egg laying, egg incubation, chick hatch, and nestling stages.

![](../figures/images/samplingtimepoints.png)

##

Next, we manipulated the timeline of the offspring development to distinguish genomic signatures that are driven by cues from the offspring from those that are driven by internal cues from within the parent.

```{r, fig.width = 8}
q  <- ggdraw() + draw_image("../figures/sbnposter/DoveParentsRNAseq_approach-3.png") 
q
```

##

We developed an R workflow for rapid and reproducible hypothesis testing related to specific tissues, sexes, and timepoints from our dataset of 1000 samples. All data and analyses are available at https://github.com/macmanes-lab/DoveParentsRNAseq.

```{r, fig.width = 8}
q  <- ggdraw() + draw_image("../figures/sexes/DoveParentsRNAseq_model2.png") 
q
```

##

Preliminary findings suggest that the gene expression of one hundred genes in the male and female pituitary mirrors that of circulating prolactin levels in the blood.

```{r, fig.width = 8}
r  <- ggdraw() + draw_image("../figures/sexes/wgcna-2.png") 
s <- ggdraw() + draw_image("../figures/sexes/wgcna-3.png") 
plot_grid(r,s, nrow = 1)
```

##

Removal of offspring around the time of hatch time causes prolactin to plummet and gene expression patterns shift to a non-parental state, suggesting an external stimulus-driven response; however prolonging incubation or delaying hatch have a much more subtle effect on gene expression.

```{r, fig.width = 8}
t  <- ggdraw() + draw_image("../figures/manipulation/pca-2.png") 
t
```

##

By characterizing and manipulating parental care and measuring the effects on hormones and gene expression in both male and female parents over time, we provide a deeper understanding of how the hypothalamic-pituitary-gonadal axis responds to predictable and unpredictable changes during offspring development.

~[](./)

### 2019-09-05 docs/SICB2020.Rmd
edited

@@
 
-Removal of offspring around the time of hatch time causes prolactin to plummet and gene expression patterns shift to a non-parental state, suggesting an external stimulus-driven response; however prolonging incubation or delaying hatch have a much more subtle effect on gene expression.
+Removal of offspring around the time of chicks hatching causes circulating prolactin to plummet and gene expression patterns shift to a non-parental state, suggesting an external stimulus-driven response; however prolonging incubation or delaying hatch have a much more subtle effect on gene expression.
 
@@
 
-By characterizing and manipulating parental care and measuring the effects on hormones and gene expression in both male and female parents over time, we provide a deeper understanding of how the hypothalamic-pituitary-gonadal axis responds to predictable and unpredictable changes during offspring development.
+By characterizing and manipulating parental care and measuring the effects on hormones and gene expression in both male and female parents over time, we provide a more complete picture of how the hypothalamic-pituitary-gonadal axis responds to predictable and unpredictable changes during offspring development.
 

### 2019-09-05 analysis/04_DESeq2_sexes.Rmd
so many gene lists

@@
 
-```{r wgcna, eval = T, fig.width=12, fig.height=6}
-plotWGCNAcandidates(vsd.pit, c("NP_990797.2"), colDataPit, "PRL expression")
+```{r wgcna, eval = T, fig.width=6, fig.height=6}
+plotWGCNAcandidates(vsd.pit, c("NP_990797.2"), colDataPit, "PRL expression, Pituitary")
+plotWGCNAcandidates(vsd.pit, c("NP_001090992.1"), colDataPit, "VIP1R expression, Pituitary")
+plotWGCNAcandidates(vsd.pit, c("NP_990797.2", "NP_001090992.1"), colDataPit, "PRL & VIP1R expression, Pituitary")
+plotWGCNAcandidates(vsd.pit, c("NP_990136.1", "NP_001035179.1"), colDataPit, "TH & AR expression, Pituitary")
+plotWGCNAcandidates(vsd.pit, c("NP_990797.2", "NP_001090992.1","NP_990136.1", "NP_001035179.1", "NP_001026269.1"), colDataPit, "PRL, PRLR, VIP1R, POMC, TH & AR expression, Pituitary")
+
+plotWGCNAcandidates(vsd.hyp, c("NP_001170780.1", "NP_001090992.1", "NP_001014970.1"), colDataHyp, "VIP expression, Hypothalamus")
 
@@
                                "NP_990797.2", "XP_015132722.1"), colDataGon, "Gonad: CISH, SOCS, PRL, PRLR")
+```
 
+```{r wgcna2, eval = T, fig.width=12, fig.height=8}
+PRL_associated <- read.csv("../results/08_PRL_associated.csv", header = T)
 prl36 <-  PRL_associated$entrezid[1:57]
 prl72 <-  PRL_associated$entrezid[58:102]
-
 plotWGCNAcandidates(vsd.pit, prl36, colDataPit, "A-K genes with PRL pattern in the pituitary")
@@
 
-str(PRL_associated)
+PRLR_associated <- read.csv("../results/08_PRLR_associated.csv", header = T)
+prlr36 <-  PRLR_associated$entrezid[1:57]
+prlr72 <-  PRLR_associated$entrezid[58:115]
+prlr116<-  PRLR_associated$entrezid[116:147]
+plotWGCNAcandidates(vsd.pit, prlr36, colDataPit, "A-G genes with PRLR pattern in the pituitary")
+plotWGCNAcandidates(vsd.pit, prlr72, colDataPit, "L-S genes with PRLR pattern in the pituitary")
+plotWGCNAcandidates(vsd.pit, prlr116, colDataPit, "S-Z genes with PRLR pattern in the pituitary")
+
+VIPR1_associated <- read.csv("../results/08_VIPR1_associated.csv", header = T)
+vipr156 <-  VIPR1_associated$entrezid[398:456]
+plotWGCNAcandidates(vsd.pit, vipr156, colDataPit, "Genes with VIPR1 pattern in the pituitary")
+
+AR_associated <- read.csv("../results/08_AR_associated.csv", header = T)
+AR <-  AR_associated$entrezid[1:31]
+plotWGCNAcandidates(vsd.pit, AR, colDataPit, "Genes with AR pattern in the pituitary")
+
+POMC_associated <- read.csv("../results/08_POMC_associated.csv", header = T)
+pomc <-  POMC_associated$entrezid[32:61]
+plotWGCNAcandidates(vsd.pit, pomc, colDataPit, "Genes with POMC pattern in the pituitary")
 
@@
 rts <- assay(vsd.pit)
-head(rts)
-
-prts <- PreparingTheIndexes(x = rts, min_cutoff=0.5, max_cutoff=0.7, mad.scale = TRUE)
 ```

### 2019-09-05 analysis/08_WGCNA_2.R

@@ -98,19 +98,38 @@ prolatin_modules <- genes_modules %>%
   filter(entrezid %in% c("NP_990797.2", "XP_015132722.1"))
 prolatin_modules
 
-PRL_associated <- genes_modules %>%
-  filter(`net$unmergedColors` %in% c("greenyellow"))
+PRL_associated <- genes_modules %>% filter(`net$unmergedColors` %in% c("greenyellow"))
 PRL_associated <- left_join(PRL_associated, geneinfo, by = "entrezid")
 PRL_associated <- PRL_associated %>% arrange(Name)
 PRL_associated
+write.csv(PRL_associated, "../results/08_PRL_associated.csv", row.names = F)
 
-write.csv(PRL_associated, "~/Desktop/PRL_associated.csv")
-
-PRLR_associated <- genes_modules %>%
-  filter(`net$unmergedColors` %in% c("magenta"))
+PRLR_associated <- genes_modules %>% filter(`net$unmergedColors` %in% c("magenta"))
 PRLR_associated <- left_join(PRLR_associated, geneinfo, by = "entrezid")
 PRLR_associated <- PRLR_associated %>% arrange(Name)
 PRLR_associated
+write.csv(PRLR_associated, "../results/08_PRLR_associated.csv", row.names = F)
+
+
+VIPR1_associated <- genes_modules %>% filter(`net$unmergedColors` %in% c("yellow"))
+VIPR1_associated <- left_join(VIPR1_associated, geneinfo, by = "entrezid")
+VIPR1_associated <- VIPR1_associated %>% arrange(Name)
+VIPR1_associated
+write.csv(VIPR1_associated, "../results/08_VIPR1_associated.csv", row.names = F)
+
+AR_associated <- genes_modules %>% filter(`net$unmergedColors` %in% c("green"))
+AR_associated <- left_join(AR_associated, geneinfo, by = "entrezid")
+AR_associated <- AR_associated %>% arrange(Name)
+AR_associated
+write.csv(AR_associated, "../results/08_AR_associated.csv", row.names = F)
+
+POMC_associated <- genes_modules %>% filter(`net$unmergedColors` %in% c("darkred"))
+POMC_associated <- left_join(POMC_associated, geneinfo, by = "entrezid")
+POMC_associated <- POMC_associated %>% arrange(Name)
+POMC_associated
+write.csv(POMC_associated, "../results/08_POMC_associated.csv", row.names = F)
+
+
 
 #=====================================================================================
 #

### 2019-09-06 docs/characterization_manuscript.Rmd
two manuscripts not one

--- 2019-08-23 docs/manuscript_draft.Rmd
@@
 ---
-title: Title of submission to PLOS journal
+title: Peaks and valleys of prolatin-related gene expression during pigeon parental care stages
 author:
-  - name: Alice Anonymous
-    email: alice@example.com
-    affiliation: Some Institute of Technology
-    corresponding: alice@example.com
-  - name: Bob Security
-    email: bob@example.com
-    affiliation: Another University
+  - name: Rayna M Harris
+    email: rmharris@ucdavis
+    affiliation: UC Davis
+    corresponding: "rmharris@ucdavis"
+  - name: Suzanne Austin
+    affiliation: UC Davis
+  - name: Andrew Lang
+    affiliation: UNH
+  - name: Matthew MacManes
+    affiliation: UNH
+  - name: Rebecca M Calisi
+    affiliation: UC Davis           
 address:
-  - code: Some Institute of Technology
-    address: Department, Street, City, State, Zip
-  - code: Another University
-    address: Department, Street, City, State, Zip
+  - code: UC Davis
+    address: UC Davis, Davis, CA
+  - code: UNH
+    address: UNH, smalltown, NH
 abstract: |
@@
 author_summary: |
-  One of the central questions of animal behavior is how biological mechanisms mediate and respond to behavior.
-Successful rearing of offspring involves a shift from aggressive and sexual behaviors to more caring and nurturing ones, but what neural substrates mediate this transition?
-How flexible are these changes in response to unpredictable environmental perturbations, and how is behavior altered because of them?
-In a rapidly changing world, understanding how the environment affects the brain and how, in turn, the brain affects the behavioral transition into parental care will shed light on how changes in environment can ultimately affect fitness.
-Previous discoveries have implicated particular hormones that play a role in the maintenance of parental care behavior in vertebrates, such as: oxytocin and vasopressin (avian homologs: mesotocin and vasotocin), vasoactive intestinal peptide, and prolactin.
-However, we know very little about when and how the brain transitions into parental care behaviors in any vertebrate, and this knowledge is fundamentally important to our understanding of the mechanisms mediating parental care.
-Here, we propose to expand our knowledge by using classic offspring replacement and removal manipulations to uncover the role of GnIH during the transition to parental care behaviors.
-In addition to this targeted approach, we propose a powerful untargeted approach to explore what other changes in the brain may be occurring during this transition.
-We will use high-throughput sequencing technology and immunohistochemistry to uncover all genetic changes and specific proteomic changes in the brain, along with their relationship with GnIH, which occur during the transition to parental care.
-Our long-term goal is to understand the interplay of genes, proteins, and parental care.
-Specifically, we aim to understand the genetic and physiological mechanisms driving parental care behaviors in all vertebrates, and to document this knowledge in a rich, open database.
-This knowledge would enable manipulation of specific parental care behaviors and their activation using emerging powerful genetic technologies.
-
+  Lorem ipsum dolor sit amet, consectetur adipiscing elit. 
 bibliography: mybibfile.bib
@@
 # tinytex::install_tinytex()
-
 ```
@@
 
-# Materials and Methods
+# Materials and Methods and Results
 
@@
 
-## Elucidate cause of transcription & protein changes during parental care transitions
+## Characterize changes in neural-gene expression during parental care transitions
 
-Using egg and chick removal and replacement experiments we will determine if neural transcription and translation during parental care transitions are based on external sensory information (the presence/absence of eggs or chicks) or on an internal clock-timing mechanism (natural biological rhythms unaffected by changes in external sensory information).
-As rock doves are known to cross-foster eggs and chicks, our working hypothesis is that the transition to parental care behaviors is based on external sensory input, and specific predictions are explained in conjunction with each manipulation below.
+Our working hypothesis is that distinct changes in transcription occur in the brain at the anticipation of, during, and in response to two different types of parental care: incubation behavior and hatchling care.
 
-### Give birds eggs
-We will place eggs in nests of birds lacking them and sample brains the following day.
-Prediction: birds will transition into incubation behavior, and neural transcription and translation will mirror that of birds sampled on the day their first egg was laid.
-This would support our working hypothesis that sensory perception of the environment drives these changes.
-If we do not observe changes in behavior, transcription, and translation, or they only occur in males, one of many interpretations is that males respond to environmental cues while females are cued by an internal clock-timing mechanism regulating transitions into parental care.
-
-### Prolong incubation
-We will prolong incubation by removing eggs from actively incubated nests on the day before hatching (Day 17) and immediately replace them with infertile, ‘dummy’ eggs.
-We will then sample brains 3 days after normal hatch time (Day 21).
-Collaborator Silver and colleagues have shown sex differences in the endocrine profiles of male and female doves in response to prolonged incubation, in that prolactin levels are maintained in incubating females but not males (Ramsey et al.
-1985).
-Because of this, we predict that if changes in behavior, neural transcription, and translation occur in females because of sensory information related to chicks hatching (supporting our working hypothesis), then females that experience this prolonged incubation will mirror the changes of incubating birds sampled on Day 17 of incubation.
-However, if changes in behavior, transcription and translation are brought on by internal clock mechanisms rather than external sensory information, as predicted at this stage in males, bird behavior and biology will resemble that of birds collected on Day 3 post-hatching.
-
-### Extend time of hatching
-We will prolong incubation by removing eggs from nests undergoing active incubation on Day 17 (the end of incubation) and immediately replacing eggs with ‘dummy’ eggs.
-On Day 21 (3 days post-natural hatch time) we will offer hatchlings.
-Doves are known to care for nestlings placed in their nest that are not their own (Klinghammer and Hess 1964, Hasen 1971).
-Brains will be sampled the following day.
-Prediction: Behavior, transcription and translation will be similar to birds naturally caring for chick(s) one day post-hatch of the first chick, supporting our working hypothesis.
-Alternatively, if behavior, transcription and translation are similar to birds caring for hatchlings on Day 3 post-hatching, which would have been the normal time course for parental care if chicks had hatched naturally, this would support regulation by an internal clock mechanism.
-
-### Remove eggs
-We will remove eggs during incubation on Day 1 (beginning), Day 9 (middle), or Day 17 (end) and sample brains on the following day.
-Note: This is not a repeated measures sampling, and three groups of male-female pairs will be independently sampled at each time point.
-Prediction: Birds will revert to a behavioral, transcriptional and translational pattern characteristic of the “Nesting, prior to lay” sampling point, lending support to our working hypothesis.
-If birds maintain a profile characteristic of birds collected at similar time points whose eggs have not been removed, this lends support to an internal clock mechanism.
-
-### Reduce time to hatching
-We will remove eggs on Day 8 (middle) of incubation and immediately replace them with newly hatched chicks.
-We will sample brains on the following day (Day 9).
-Prediction: Changes in behavior, transcription and translation are externally sensory driven, and these will be similar to birds sampled on ‘Day 1: 1st egg
-hatches, supporting our working hypothesis.
-Alternatively, if profiles remain similar to birds sampled on Day 9 that are actively incubating eggs, this would support the presence of an internal clock mechanism.
-
-### Remove chicks
-We will remove newly, naturally hatched chick(s) on Day 2: second chick hatches and sample brains from parents the following day (Day 2 post (second) egg hatching).
-Prediction: Behavior, transcription and translation are sensory driven and will revert to a pattern characteristic of the ‘Nesting, prior to lay’ point, supporting our working hypothesis.
-Alternatively, if an internal clock mechanism is at play, we would expect to see a pattern similar to birds caring for chicks on Day 2 post (second) egg hatching.
-
-# Results
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
-
-# Figures
-
-```{r wgcna, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-source("../analysis/08_WGCNA_1.R")
-subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary",
-                           "male.gonad", "male.hypothalamus", "male.pituitary"),
-            mytitle = "Characterization dendrogram with trait heatmap")
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+library(cowplot)
+ggdraw() + draw_image("../figures/characterization/linegraphs-1.png") 
 ```
 
-**Fig 1. Characterization dendrogram with trait heatmap.**.
-Most hypothalamus and pituitary samples separte form the gonad samples, but some are clusters contain samples from all tissues.The top panel is a dendrogram showing Euclidean distance of sample.
-Females (white), males (red), gonad (white), hypothalamus (pink), pituitary (red).
-
-```{r wgcna-female, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary"),
-            mytitle = "Female sample dendrogram with trait heatmap")
-```
-
-**Fig 2. Female characterization dendrogram with trait heatmap.**. 
-Most pituitary samples form a single cluster.
-More hypothalamus samples cluster with gonads than with pituitary.
-The top panel is a dendrogram showing Euclidean distance of sample.
-Females (white), males (red), gonad (white), hypothalamus (pink), pituitary (red).
-
-```{r wgcna-male, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-subsetWGCNA(c("male.gonad", "male.hypothalamus", "male.pituitary"),
-            mytitle = "Male sample dendrogram with trait heatmap")
-```
-
-**Fig 3. Male characterization dendrogram with trait heatmap.**. 
- As with females, most pituitary samples form a single cluster, while more hypothalamus samples cluster with gonads than with pituitary.
+**Fig 2. The magnitude of gene expression changes between each parental transition**.
+4-5K genes are differentially expressed beteen control birds and their nest buliding conspecifics in all tissues except the male gonad.
+500 - 1000 genes are differentially expressed in the female pituitary from mid-late incubation as well as in the female hypothalamus and pituitary from hatch to nestingly care day 5.
 

### 2019-09-06 docs/characterization_manuscript.tex
two manuscripts not one

--- 2019-08-26 docs/manuscript_draft.tex
@@
 \newcommand{\getAff}[1]{
-  \getIndex{#1}{Some Institute of Technology,Another University}
+  \getIndex{#1}{UC Davis,UNH}
 }
@@
 {\Large
-\textbf\newline{Title of submission to PLOS journal} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Peaks and valleys of prolatin-related gene expression during pigeon
+parental care stages} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 \\
-Alice Anonymous\textsuperscript{\getAff{Some Institute of Technology}}\textsuperscript{*},
-Bob Security\textsuperscript{\getAff{Another University}}\\
+Rayna M Harris\textsuperscript{\getAff{UC Davis}}\textsuperscript{*},
+Suzanne Austin\textsuperscript{\getAff{UC Davis}},
+Andrew Lang\textsuperscript{\getAff{UNH}},
+Matthew MacManes\textsuperscript{\getAff{UNH}},
+Rebecca M Calisi\textsuperscript{\getAff{UC Davis}}\\
 \bigskip
-\textbf{\getAff{Some Institute of Technology}}Department, Street, City, State, Zip\\
-\textbf{\getAff{Another University}}Department, Street, City, State, Zip\\
+\textbf{\getAff{UC Davis}}UC Davis, Davis, CA\\
+\textbf{\getAff{UNH}}UNH, smalltown, NH\\
 \bigskip
-* Corresponding author: alice@example.com\\
+* Corresponding author: rmharris@ucdavis\\
 \end{flushleft}
@@
 \section*{Author summary}
-One of the central questions of animal behavior is how biological
-mechanisms mediate and respond to behavior. Successful rearing of
-offspring involves a shift from aggressive and sexual behaviors to more
-caring and nurturing ones, but what neural substrates mediate this
-transition? How flexible are these changes in response to unpredictable
-environmental perturbations, and how is behavior altered because of
-them? In a rapidly changing world, understanding how the environment
-affects the brain and how, in turn, the brain affects the behavioral
-transition into parental care will shed light on how changes in
-environment can ultimately affect fitness. Previous discoveries have
-implicated particular hormones that play a role in the maintenance of
-parental care behavior in vertebrates, such as: oxytocin and vasopressin
-(avian homologs: mesotocin and vasotocin), vasoactive intestinal
-peptide, and prolactin. However, we know very little about when and how
-the brain transitions into parental care behaviors in any vertebrate,
-and this knowledge is fundamentally important to our understanding of
-the mechanisms mediating parental care. Here, we propose to expand our
-knowledge by using classic offspring replacement and removal
-manipulations to uncover the role of GnIH during the transition to
-parental care behaviors. In addition to this targeted approach, we
-propose a powerful untargeted approach to explore what other changes in
-the brain may be occurring during this transition. We will use
-high-throughput sequencing technology and immunohistochemistry to
-uncover all genetic changes and specific proteomic changes in the brain,
-along with their relationship with GnIH, which occur during the
-transition to parental care. Our long-term goal is to understand the
-interplay of genes, proteins, and parental care. Specifically, we aim to
-understand the genetic and physiological mechanisms driving parental
-care behaviors in all vertebrates, and to document this knowledge in a
-rich, open database. This knowledge would enable manipulation of
-specific parental care behaviors and their activation using emerging
-powerful genetic technologies.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 
@@
 
-Aim I lays the necessary groundwork for establishing a pattern of
-correlative relationships between observed changes in gene expression
-and patterns of parental behavior. With this pattern established, we can
-proceed with experimental manipulations proposed under Aim III to
-determine causation.
-
 \hypertarget{materials-and-methods-and-results}{%
@@
 
-\hypertarget{overall-patterns-of-sample-similarity}{%
-\subsubsection{overall patterns of sample
-similarity}\label{overall-patterns-of-sample-similarity}}
-
-This ecludian clustering shows many clusters are heterogenous,
-consisting of sames from all tissues types, both sexes, and from
-multiple treatment groups.
-
-\includegraphics{manuscript_draft_files/figure-latex/wgcna-1.pdf}
-
-\textbf{Fig 1. Characterization dendrogram with trait heatmap.}. Most
-hypothalamus and pituitary samples separte form the gonad samples, but
-some are clusters contain samples from all tissues.The top panel is a
-dendrogram showing Euclidean distance of sample. Females (white), males
-(red), gonad (white), hypothalamus (pink), pituitary (red).
-
-\hypertarget{subset-by-sex-and-tissue}{%
-\subsection{subset by sex and tissue}\label{subset-by-sex-and-tissue}}
-
-The rest of the analyses are peformed on each sex and tissue group
-separately, a 2 x 3 design.
-
-\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-4-1.pdf}
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
 
@@
 hypothalamus and pituitary from hatch to nestingly care day 5.
-
-\includegraphics{../figures/characterization/musicalgenes-05.png}
-
-\hypertarget{elucidate-cause-of-transcription-protein-changes-during-parental-care-transitions}{%
-\subsection{Elucidate cause of transcription \& protein changes during
-parental care
-transitions}\label{elucidate-cause-of-transcription-protein-changes-during-parental-care-transitions}}
-
-Using egg and chick removal and replacement experiments we will
-determine if neural transcription and translation during parental care
-transitions are based on external sensory information (the
-presence/absence of eggs or chicks) or on an internal clock-timing
-mechanism (natural biological rhythms unaffected by changes in external
-sensory information). As rock doves are known to cross-foster eggs and
-chicks, our working hypothesis is that the transition to parental care
-behaviors is based on external sensory input, and specific predictions
-are explained in conjunction with each manipulation below.
-
-\hypertarget{give-birds-eggs}{%
-\subsubsection{Give birds eggs}\label{give-birds-eggs}}
-
-We will place eggs in nests of birds lacking them and sample brains the
-following day. Prediction: birds will transition into incubation
-behavior, and neural transcription and translation will mirror that of
-birds sampled on the day their first egg was laid. This would support
-our working hypothesis that sensory perception of the environment drives
-these changes. If we do not observe changes in behavior, transcription,
-and translation, or they only occur in males, one of many
-interpretations is that males respond to environmental cues while
-females are cued by an internal clock-timing mechanism regulating
-transitions into parental care.
-
-\hypertarget{prolong-incubation}{%
-\subsubsection{Prolong incubation}\label{prolong-incubation}}
-
-We will prolong incubation by removing eggs from actively incubated
-nests on the day before hatching (Day 17) and immediately replace them
-with infertile, `dummy' eggs. We will then sample brains 3 days after
-normal hatch time (Day 21). Collaborator Silver and colleagues have
-shown sex differences in the endocrine profiles of male and female doves
-in response to prolonged incubation, in that prolactin levels are
-maintained in incubating females but not males (Ramsey et al.~1985).
-Because of this, we predict that if changes in behavior, neural
-transcription, and translation occur in females because of sensory
-information related to chicks hatching (supporting our working
-hypothesis), then females that experience this prolonged incubation will
-mirror the changes of incubating birds sampled on Day 17 of incubation.
-However, if changes in behavior, transcription and translation are
-brought on by internal clock mechanisms rather than external sensory
-information, as predicted at this stage in males, bird behavior and
-biology will resemble that of birds collected on Day 3 post-hatching.
-
-\hypertarget{extend-time-of-hatching}{%
-\subsubsection{Extend time of hatching}\label{extend-time-of-hatching}}
-
-We will prolong incubation by removing eggs from nests undergoing active
-incubation on Day 17 (the end of incubation) and immediately replacing
-eggs with `dummy' eggs. On Day 21 (3 days post-natural hatch time) we
-will offer hatchlings. Doves are known to care for nestlings placed in
-their nest that are not their own (Klinghammer and Hess 1964, Hasen
-1971). Brains will be sampled the following day. Prediction: Behavior,
-transcription and translation will be similar to birds naturally caring
-for chick(s) one day post-hatch of the first chick, supporting our
-working hypothesis. Alternatively, if behavior, transcription and
-translation are similar to birds caring for hatchlings on Day 3
-post-hatching, which would have been the normal time course for parental
-care if chicks had hatched naturally, this would support regulation by
-an internal clock mechanism.
-
-\hypertarget{remove-eggs-or-chicks}{%
-\subsubsection{Remove eggs or chicks}\label{remove-eggs-or-chicks}}
-
-We will remove eggs during incubation on Day 1 (beginning), Day 9
-(middle), or Day 17 (end) and sample brains on the following day. Note:
-This is not a repeated measures sampling, and three groups of
-male-female pairs will be independently sampled at each time point.
-Prediction: Birds will revert to a behavioral, transcriptional and
-translational pattern characteristic of the ``Nesting, prior to lay''
-sampling point, lending support to our working hypothesis. If birds
-maintain a profile characteristic of birds collected at similar time
-points whose eggs have not been removed, this lends support to an
-internal clock mechanism.
-
-We will remove newly, naturally hatched chick(s) on Day 2: second chick
-hatches and sample brains from parents the following day (Day 2 post
-(second) egg hatching). Prediction: Behavior, transcription and
-translation are sensory driven and will revert to a pattern
-characteristic of the `Nesting, prior to lay' point, supporting our
-working hypothesis. Alternatively, if an internal clock mechanism is at
-play, we would expect to see a pattern similar to birds caring for
-chicks on Day 2 post (second) egg hatching.
-
-\includegraphics{../figures/sexes/offspringremoval-1.png}
-
-\hypertarget{reduce-time-to-hatching}{%
-\subsubsection{Reduce time to hatching}\label{reduce-time-to-hatching}}
-
-We will remove eggs on Day 8 (middle) of incubation and immediately
-replace them with newly hatched chicks. We will sample brains on the
-following day (Day 9). Prediction: Changes in behavior, transcription
-and translation are externally sensory driven, and these will be similar
-to birds sampled on `Day 1: 1st egg hatches, supporting our working
-hypothesis. Alternatively, if profiles remain similar to birds sampled
-on Day 9 that are actively incubating eggs, this would support the
-presence of an internal clock mechanism.
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
-porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
-dui eu ante cursus gravida non sed sem. Phasellus mauris velit, dapibus
-finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
-diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
-ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
-pulvinar, volutpat risus.
-
-\hypertarget{supplementary-figures}{%
-\section{Supplementary Figures}\label{supplementary-figures}}
-
-\includegraphics{manuscript_draft_files/figure-latex/wgcna-female-1.pdf}
-
-\textbf{Fig S1a. Female characterization dendrogram with trait
-heatmap.}. Most pituitary samples form a single cluster. More
-hypothalamus samples cluster with gonads than with pituitary. The top
-panel is a dendrogram showing Euclidean distance of sample. Females
-(white), males (red), gonad (white), hypothalamus (pink), pituitary
-(red).
-
-\includegraphics{manuscript_draft_files/figure-latex/wgcna-male-1.pdf}
-
-\textbf{Fig S1b. Male characterization dendrogram with trait heatmap.}.
-As with females, most pituitary samples form a single cluster, while
-more hypothalamus samples cluster with gonads than with pituitary.
 

### 2019-09-06 docs/manipulation_manuscript.Rmd
two manuscripts not one

--- 2019-08-26 docs/manuscript_draft.Rmd
@@
 ---
-title: Title of submission to PLOS journal
+title: Are parental care induced gene expression patterns affect more by internal cues or external stimuli?
 author:
-  - name: Alice Anonymous
-    email: alice@example.com
-    affiliation: Some Institute of Technology
-    corresponding: alice@example.com
-  - name: Bob Security
-    email: bob@example.com
-    affiliation: Another University
+  - name: Rayna M Harris
+    email: rmharris@ucdavis
+    affiliation: UC Davis
+    corresponding: "rmharris@ucdavis"
+  - name: Suzanne Austin
+    affiliation: UC Davis
+  - name: Andrew Lang
+    affiliation: UNH
+  - name: Matthew MacManes
+    affiliation: UNH
+  - name: Rebecca M Calisi
+    affiliation: UC Davis           
 address:
-  - code: Some Institute of Technology
-    address: Department, Street, City, State, Zip
-  - code: Another University
-    address: Department, Street, City, State, Zip
+  - code: UC Davis
+    address: UC Davis, Davis, CA
+  - code: UNH
+    address: UNH, smalltown, NH
 abstract: |
@@
 
-Understanding the mechanisms underlying parental care are critical to circumventing issues with parent-newborn bonding as well, where ultimate explanations are obvious, but specific mechanisms remain elusive.
-The rock dove (_Columba livia_) is an ideal system to characterize changes in genetic expression during parental care transitions because: 1) ample genomic resources are available (Gillespie et al.
-2013), including a complete annotated genome assembly @Shapiro1063 and methodology concerning reproductive physiology and behavior (Dong et al.
-2012); and 2) rock doves are prolific, year-round breeders that thrive in captivity, making observation, manipulation, and sampling highly feasible year-round.
-
-```{r, out.width = "360px", echo = F}
-knitr::include_graphics("../figures/images/PigeonHPGaxis.png")
-```
-
-Rock doves are socially monogamous and offer bi-parental care, making inter- and intra-sexual comparisons possible.
-Birds offer two important behavioral transition points into parental care: the incubation of eggs and the caring for chicks.
-This produces two unique opportunities to study how the brain transitions into two different suites of parental care behaviors.
-Additionally, rock doves exhibit a parental care strategy analogous to mammals in that they, too, ‘lactate’ to feed their young (Gillespie et al.
-2011, 2012).
-This lactation, unlike simple regurgitation of food, consists of the production and sloughing off of skin cells inside the crop sac of females and males, creating a protein-rich milk-like substance on which they rear their chicks.
-Many functional similarities between rock dove and mammalian lactation exist concerning the mediation of this event by the hormone prolactin (Dumont 1965).
-Additionally, like mammalian milk, rock dove milk delivers essential immunoglobulins and nutritional benefits to young, aiding in their immune function and development of microbiota @Gillespie2012.
-Thus, because rock doves incubate eggs and exhibit mammalian-like mediation and function of lactation for young, they have the potential to serve as a powerful theoretical bridge to understand the neurobiology of both avian and mammalian transitions into parental care.
-
-Aim I lays the necessary groundwork for establishing a pattern of correlative relationships between observed changes in gene expression and patterns of parental behavior.
-With this pattern established, we can proceed with experimental manipulations proposed under Aim III to determine causation.
-
 # Materials and Methods and Results
-
-```{r, out.width = "360px", echo = F}
-knitr::include_graphics("../figures/images/samplingtimepoints.png")
-```
-
-## Characterize changes in neural-gene expression during parental care transitions
-
-Our working hypothesis is that distinct changes in transcription occur in the brain at the anticipation of, during, and in response to two different types of parental care: incubation behavior and hatchling care.
-
-This ecludian clustering shows many clusters are heterogenous, consisting of sames from all tissues types, both sexes, and from multiple treatment groups.
-
-```{r wgcna, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-source("../analysis/08_WGCNA_1.R")
-subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary",
-                           "male.gonad", "male.hypothalamus", "male.pituitary"),
-            mytitle = "Characterization dendrogram of samples")
-```
-
-**Fig 1. Characterization dendrogram with trait heatmap.**.
-Most hypothalamus and pituitary samples separte form the gonad samples, but some are clusters contain samples from all tissues.The top panel is a dendrogram showing Euclidean distance of sample.
-Females (white), males (red), gonad (white), hypothalamus (pink), pituitary (red).
-
-The rest of the analyses are peformed on each sex and tissue group separately, a 2 x 3 design.
-
-```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-library(cowplot)
-ggdraw() + draw_image("../figures/sexes/characterization-2.png") 
-```
-
-**Fig 2. The magnitude of gene expression changes between each parental transition**.
-4-5K genes are differentially expressed beteen control birds and their nest buliding conspecifics in all tissues except the male gonad.
-500 - 1000 genes are differentially expressed in the female pituitary from mid-late incubation as well as in the female hypothalamus and pituitary from hatch to nestingly care day 5.
-
-![](../figures/characterization/musicalgenes-05.png)
 
@@
 
-![](../figures/sexes/offspringremoval-1.png)
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+library(cowplot)
+ggdraw() + draw_image("../figures/manipulation/offspringremoval-1.png") 
+```
 
@@
 # Supplementary Figures
-
-```{r wgcna-female, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary"),
-            mytitle = "Female sample dendrogram with trait heatmap")
-```
-
-**Fig S1a. Female characterization dendrogram with trait heatmap.**. 
-Most pituitary samples form a single cluster.
-More hypothalamus samples cluster with gonads than with pituitary.
-The top panel is a dendrogram showing Euclidean distance of sample.
-Females (white), males (red), gonad (white), hypothalamus (pink), pituitary (red).
-
-```{r wgcna-male, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-subsetWGCNA(c("male.gonad", "male.hypothalamus", "male.pituitary"),
-            mytitle = "Male sample dendrogram with trait heatmap")
-```
-
-**Fig S1b. Male characterization dendrogram with trait heatmap.**. 
- As with females, most pituitary samples form a single cluster, while more hypothalamus samples cluster with gonads than with pituitary.
 

### 2019-09-06 docs/manipulation_manuscript.tex
two manuscripts not one

--- 2019-08-23 docs/manuscript_draft.tex
@@
 {\Large
-\textbf\newline{Title of submission to PLOS journal} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Are parental care induced gene expression patterns affect more by
+internal cues or external stimuli?} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 
-Understanding the mechanisms underlying parental care are critical to
-circumventing issues with parent-newborn bonding as well, where ultimate
-explanations are obvious, but specific mechanisms remain elusive. The
-rock dove (\emph{Columba livia}) is an ideal system to characterize
-changes in genetic expression during parental care transitions because:
-1) ample genomic resources are available (Gillespie et al.~2013),
-including a complete annotated genome assembly {[}1{]} and methodology
-concerning reproductive physiology and behavior (Dong et al.~2012); and
-2) rock doves are prolific, year-round breeders that thrive in
-captivity, making observation, manipulation, and sampling highly
-feasible year-round.
-
-\includegraphics[width=360px]{../figures/images/PigeonHPGaxis}
-
-Rock doves are socially monogamous and offer bi-parental care, making
-inter- and intra-sexual comparisons possible. Birds offer two important
-behavioral transition points into parental care: the incubation of eggs
-and the caring for chicks. This produces two unique opportunities to
-study how the brain transitions into two different suites of parental
-care behaviors. Additionally, rock doves exhibit a parental care
-strategy analogous to mammals in that they, too, `lactate' to feed their
-young (Gillespie et al.~2011, 2012). This lactation, unlike simple
-regurgitation of food, consists of the production and sloughing off of
-skin cells inside the crop sac of females and males, creating a
-protein-rich milk-like substance on which they rear their chicks. Many
-functional similarities between rock dove and mammalian lactation exist
-concerning the mediation of this event by the hormone prolactin (Dumont
-1965). Additionally, like mammalian milk, rock dove milk delivers
-essential immunoglobulins and nutritional benefits to young, aiding in
-their immune function and development of microbiota {[}2{]}. Thus,
-because rock doves incubate eggs and exhibit mammalian-like mediation
-and function of lactation for young, they have the potential to serve as
-a powerful theoretical bridge to understand the neurobiology of both
-avian and mammalian transitions into parental care.
-
-\hypertarget{materials-and-methods}{%
-\section{Materials and Methods}\label{materials-and-methods}}
-
-\includegraphics[width=360px]{../figures/images/samplingtimepoints}
+\hypertarget{materials-and-methods-and-results}{%
+\section{Materials and Methods and
+Results}\label{materials-and-methods-and-results}}
 
@@
 internal clock mechanism.
+
+We will remove newly, naturally hatched chick(s) on Day 2: second chick
+hatches and sample brains from parents the following day (Day 2 post
+(second) egg hatching). Prediction: Behavior, transcription and
+translation are sensory driven and will revert to a pattern
+characteristic of the `Nesting, prior to lay' point, supporting our
+working hypothesis. Alternatively, if an internal clock mechanism is at
+play, we would expect to see a pattern similar to birds caring for
+chicks on Day 2 post (second) egg hatching.
+
+\includegraphics{manipulation_manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
 
@@
 
-\hypertarget{figures}{%
-\section{Figures}\label{figures}}
-
-\includegraphics{manuscript_draft_files/figure-latex/wgcna-1.pdf}
-
-\textbf{Fig 1. Characterization dendrogram with trait heatmap.}. Most
-hypothalamus and pituitary samples separte form the gonad samples, but
-some are clusters contain samples from all tissues.The top panel is a
-dendrogram showing Euclidean distance of sample. Females (white), males
-(red), gonad (white), hypothalamus (pink), pituitary (red).
-
-\includegraphics{manuscript_draft_files/figure-latex/wgcna-female-1.pdf}
-
-\textbf{Fig 2. Female characterization dendrogram with trait heatmap.}.
-Most pituitary samples form a single cluster. More hypothalamus samples
-cluster with gonads than with pituitary. The top panel is a dendrogram
-showing Euclidean distance of sample. Females (white), males (red),
-gonad (white), hypothalamus (pink), pituitary (red).
-
-\includegraphics{manuscript_draft_files/figure-latex/wgcna-male-1.pdf}
-
-\textbf{Fig 3. Male characterization dendrogram with trait heatmap.}. As
-with females, most pituitary samples form a single cluster, while more
-hypothalamus samples cluster with gonads than with pituitary.
+\hypertarget{supplementary-figures}{%
+\section{Supplementary Figures}\label{supplementary-figures}}
 
@@
 
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Shapiro1063}{}%
-1. Shapiro MD, Kronenberg Z, Li C, Domyan ET, Pan H, Campbell M, et al.
-Genomic diversity and evolution of the head crest in the rock pigeon.
-Science. American Association for the Advancement of Science; 2013;339:
-1063--1067.
-doi:\href{https://doi.org/10.1126/science.1230422}{10.1126/science.1230422}
-
-\leavevmode\hypertarget{ref-Gillespie2012}{}%
-2. Gillespie MJ, Stanley D, Chen H, Donald JA, Nicholas KR, Moore RJ, et
-al. Functional Similarities between Pigeon `Milk' and Mammalian Milk:
-Induction of Immune Gene Expression and Modification of the Microbiota.
-Salmon H, editor. PLoS ONE. Public Library of Science; 2012;7: e48363.
-doi:\href{https://doi.org/10.1371/journal.pone.0048363}{10.1371/journal.pone.0048363}
-
 \nolinenumbers

### 2019-09-06 docs/characterization_manuscript.Rmd
now with abstract

@@
 abstract: |
-  Lorem ipsum dolor sit amet, consectetur adipiscing elit. 
+  The goal of this research was to provide the most comprehensive gene expression profile of gene expression activity in the HPG axis of male and female rock doves to provide a deeper undertand of how the reproductive axis response to typical behavioral transitions associated with parental care.
+Here we used RNA sequencing to measure gene activity at 8 stages of the parental care cyle from nest building, to egg incubation, and through nestling care.
+Non-parental groups were added as controls, but gene expressions differences were assesed between and across all time points.
+The three tissue have unique signatures, but within each tissue we find minial sex differences.
+The pitutary samples displacy the msot plasticity in gene expression, with many changes in gene expression mirroring the typical rise and fall of circulating prolatin that peaks when chicks hatch.
+Our analysis provides new insight into how suites of genes respond in concert to the demands of offspring care.
 author_summary: |

### 2019-09-06 docs/characterization_manuscript.tex
now with abstract

@@
 \section*{Abstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+The goal of this research was to provide the most comprehensive gene
+expression profile of gene expression activity in the HPG axis of male
+and female rock doves to provide a deeper undertand of how the
+reproductive axis response to typical behavioral transitions associated
+with parental care. Here we used RNA sequencing to measure gene activity
+at 8 stages of the parental care cyle from nest building, to egg
+incubation, and through nestling care. Non-parental groups were added as
+controls, but gene expressions differences were assesed between and
+across all time points. The three tissue have unique signatures, but
+within each tissue we find minial sex differences. The pitutary samples
+displacy the msot plasticity in gene expression, with many changes in
+gene expression mirroring the typical rise and fall of circulating
+prolatin that peaks when chicks hatch. Our analysis provides new insight
+into how suites of genes respond in concert to the demands of offspring
+care.
 

### 2019-09-06 docs/manipulation_manuscript.Rmd
more abstract

@@
 abstract: |
-  Lorem ipsum dolor sit amet, consectetur adipiscing elit. 
+  Are parental care induced gene expression patterns affect more by internal cues or external stimuli?
+We test two alternative hypothesis: internal clock cues versus external offspring stimuli by manipulating the timing of normal transitions during parental care.
+When we removed offpsring, we find the most pronounces chagnes in gene expression later during incubation or right after hatch, compared to early removeal of eggs.
+Given the similarity in gene expression the days immediately preceeding and immedately following hatch, we do not have the power to say weether gene expression patters are more like hatch or incuabtion when dummmy egg are used to extend incubation.
+This matters because...
 author_summary: |
@@
 
-## Elucidate cause of transcription & protein changes during parental care transitions
-
 Using egg and chick removal and replacement experiments we will determine if neural transcription and translation during parental care transitions are based on external sensory information (the presence/absence of eggs or chicks) or on an internal clock-timing mechanism (natural biological rhythms unaffected by changes in external sensory information).
 As rock doves are known to cross-foster eggs and chicks, our working hypothesis is that the transition to parental care behaviors is based on external sensory input, and specific predictions are explained in conjunction with each manipulation below.
-
-### Give birds eggs
 
@@
 If we do not observe changes in behavior, transcription, and translation, or they only occur in males, one of many interpretations is that males respond to environmental cues while females are cued by an internal clock-timing mechanism regulating transitions into parental care.
-
-### Prolong incubation
 
@@
 
-### Extend time of hatching
-
 We will prolong incubation by removing eggs from nests undergoing active incubation on Day 17 (the end of incubation) and immediately replacing eggs with ‘dummy’ eggs.
@@
 Alternatively, if behavior, transcription and translation are similar to birds caring for hatchlings on Day 3 post-hatching, which would have been the normal time course for parental care if chicks had hatched naturally, this would support regulation by an internal clock mechanism.
-
-### Remove eggs or chicks
 
@@
 
-```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-library(cowplot)
-ggdraw() + draw_image("../figures/manipulation/offspringremoval-1.png") 
-```
-
-### Reduce time to hatching
-
 We will remove eggs on Day 8 (middle) of incubation and immediately replace them with newly hatched chicks.
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+# Results
+
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+library(cowplot)
+ggdraw() + draw_image("../figures/manipulation/offspringremoval-1.png") 
+```
 

### 2019-09-06 docs/manipulation_manuscript.tex
more abstract

@@
 \section*{Abstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+The one question saught. Here we do. What we found. How it matters.
 
@@
 
-\hypertarget{elucidate-cause-of-transcription-protein-changes-during-parental-care-transitions}{%
-\subsection{Elucidate cause of transcription \& protein changes during
-parental care
-transitions}\label{elucidate-cause-of-transcription-protein-changes-during-parental-care-transitions}}
-
 Using egg and chick removal and replacement experiments we will
@@
 are explained in conjunction with each manipulation below.
-
-\hypertarget{give-birds-eggs}{%
-\subsubsection{Give birds eggs}\label{give-birds-eggs}}
 
@@
 transitions into parental care.
-
-\hypertarget{prolong-incubation}{%
-\subsubsection{Prolong incubation}\label{prolong-incubation}}
 
@@
 
-\hypertarget{extend-time-of-hatching}{%
-\subsubsection{Extend time of hatching}\label{extend-time-of-hatching}}
-
 We will prolong incubation by removing eggs from nests undergoing active
@@
 an internal clock mechanism.
-
-\hypertarget{remove-eggs-or-chicks}{%
-\subsubsection{Remove eggs or chicks}\label{remove-eggs-or-chicks}}
 
@@
 
-\includegraphics{manipulation_manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
-
-\hypertarget{reduce-time-to-hatching}{%
-\subsubsection{Reduce time to hatching}\label{reduce-time-to-hatching}}
-
 We will remove eggs on Day 8 (middle) of incubation and immediately
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
-porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
-dui eu ante cursus gravida non sed sem. Phasellus mauris velit, dapibus
-finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
-diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
-ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
-pulvinar, volutpat risus.
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\includegraphics{manipulation_manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
 

### 2019-09-06 docs/characterization_manuscript.Rmd
more abstract
now with char figs

@@
 Our analysis provides new insight into how suites of genes respond in concert to the demands of offspring care.
+This data can be used to develop and test hypotheses about the mechanism regulating parental care behavior.
 author_summary: |
@@
 # tinytex::install_tinytex()
+
+library(cowplot)
 ```
@@
 Thus, because rock doves incubate eggs and exhibit mammalian-like mediation and function of lactation for young, they have the potential to serve as a powerful theoretical bridge to understand the neurobiology of both avian and mammalian transitions into parental care.
+Our working hypothesis is that distinct changes in transcription occur in the brain at the anticipation of, during, and in response to two different types of parental care: incubation behavior and hatchling care.
 
-# Materials and Methods and Results
-
+# Materials and Methods
 ```{r, out.width = "360px", echo = F}
@@
 
-## Characterize changes in neural-gene expression during parental care transitions
+# Results
 
-Our working hypothesis is that distinct changes in transcription occur in the brain at the anticipation of, during, and in response to two different types of parental care: incubation behavior and hatchling care.
+## Table 1: comparison of total DEGs determined by edgeR and DESeq2
+
+## Fig 1: Total number of gene expression differences between each stage
 
 ```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-library(cowplot)
 ggdraw() + draw_image("../figures/characterization/linegraphs-1.png") 
@@
 
-**Fig 2. The magnitude of gene expression changes between each parental transition**.
+**Fig x. The magnitude of gene expression changes between each parental transition**.
 4-5K genes are differentially expressed beteen control birds and their nest buliding conspecifics in all tissues except the male gonad.
 500 - 1000 genes are differentially expressed in the female pituitary from mid-late incubation as well as in the female hypothalamus and pituitary from hatch to nestingly care day 5.
+
+## Fig 2: Ability to use gene expression to distingush groups
+
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+ggdraw() + draw_image("../figures/sexes/DoveParentsRNAseq_model2.png") 
+```
+
+## Fig 3: Prolactin expression in PIT (but not HYP or GON) follows same pattern as in blood
+
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+ggdraw() + draw_image("../figures/characterization/prolactin-1.png") 
+```
+
+## Fig 4: Hundreds of genes show similar pattern in expression as circulating prolactin
+
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+ggdraw() + draw_image("../figures/sexes/wgcna-7.png")
+```
 

### 2019-09-06 docs/characterization_manuscript.tex
now with char figs

@@
 into how suites of genes respond in concert to the demands of offspring
-care.
+care. This data can be used to develop and test hypotheses about the
+mechanism regulating parental care behavior.
 
@@
 a powerful theoretical bridge to understand the neurobiology of both
-avian and mammalian transitions into parental care.
+avian and mammalian transitions into parental care. Our working
+hypothesis is that distinct changes in transcription occur in the brain
+at the anticipation of, during, and in response to two different types
+of parental care: incubation behavior and hatchling care.
 
-\hypertarget{materials-and-methods-and-results}{%
-\section{Materials and Methods and
-Results}\label{materials-and-methods-and-results}}
+\hypertarget{materials-and-methods}{%
+\section{Materials and Methods}\label{materials-and-methods}}
 
@@
 
-\hypertarget{characterize-changes-in-neural-gene-expression-during-parental-care-transitions}{%
-\subsection{Characterize changes in neural-gene expression during
-parental care
-transitions}\label{characterize-changes-in-neural-gene-expression-during-parental-care-transitions}}
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
-Our working hypothesis is that distinct changes in transcription occur
-in the brain at the anticipation of, during, and in response to two
-different types of parental care: incubation behavior and hatchling
-care.
+\hypertarget{table-1-comparison-of-total-degs-determined-by-edger-and-deseq2}{%
+\subsection{Table 1: comparison of total DEGs determined by edgeR and
+DESeq2}\label{table-1-comparison-of-total-degs-determined-by-edger-and-deseq2}}
+
+\hypertarget{fig-1-total-number-of-gene-expression-differences-between-each-stage}{%
+\subsection{Fig 1: Total number of gene expression differences between
+each
+stage}\label{fig-1-total-number-of-gene-expression-differences-between-each-stage}}
 
@@
 
-\textbf{Fig 2. The magnitude of gene expression changes between each
+\textbf{Fig x. The magnitude of gene expression changes between each
 parental transition}. 4-5K genes are differentially expressed beteen
@@
 hypothalamus and pituitary from hatch to nestingly care day 5.
+
+\hypertarget{fig-2-ability-to-use-gene-expression-to-distingush-groups}{%
+\subsection{Fig 2: Ability to use gene expression to distingush
+groups}\label{fig-2-ability-to-use-gene-expression-to-distingush-groups}}
+
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+
+\hypertarget{fig-3-prolactin-expression-in-pit-but-not-hyp-or-gon-follows-same-pattern-as-in-blood}{%
+\subsection{Fig 3: Prolactin expression in PIT (but not HYP or GON)
+follows same pattern as in
+blood}\label{fig-3-prolactin-expression-in-pit-but-not-hyp-or-gon-follows-same-pattern-as-in-blood}}
+
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
+
+\hypertarget{fig-4-hundreds-of-genes-show-similar-pattern-in-expression-as-circulating-prolactin}{%
+\subsection{Fig 4: Hundreds of genes show similar pattern in expression
+as circulating
+prolactin}\label{fig-4-hundreds-of-genes-show-similar-pattern-in-expression-as-circulating-prolactin}}
+
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
 

### 2019-09-06 docs/manipulation_manuscript.Rmd
need more resutls

@@
 ```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-library(cowplot)
 ggdraw() + draw_image("../figures/manipulation/offspringremoval-1.png") 
+```
+
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+ggdraw() + draw_image("../figures/manipulation/manipVchar-1.png") 
+```
+
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+ggdraw() + draw_image("../figures/manipulation/prolongextend-1.png") 
 ```

### 2019-09-06 docs/manipulation_manuscript.tex
need more resutls

@@
 \section*{Abstract}
-The one question saught. Here we do. What we found. How it matters.
+Are parental care induced gene expression patterns affect more by
+internal cues or external stimuli? We test two alternative hypothesis:
+internal clock cues versus external offspring stimuli by manipulating
+the timing of normal transitions during parental care. When we removed
+offpsring, we find the most pronounces chagnes in gene expression later
+during incubation or right after hatch, compared to early removeal of
+eggs. Given the similarity in gene expression the days immediately
+preceeding and immedately following hatch, we do not have the power to
+say weether gene expression patters are more like hatch or incuabtion
+when dummmy egg are used to extend incubation. This matters
+because\ldots{}
 
@@
 
+\includegraphics{manipulation_manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
+
+\includegraphics{manipulation_manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
+
 \hypertarget{supplementary-figures}{%

### 2019-09-07 docs/characterization_manuscript.Rmd
new figs

@@
 bibliography: mybibfile.bib
-output: rticles::plos_article
+output: 
+  rticles::plos_article:
+    df_print: kable
 csl: plos.csl
@@
 library(cowplot)
+library(kableExtra)
 ```
@@
 
-```{r, out.width = "360px", echo = F}
-knitr::include_graphics("../figures/images/PigeonHPGaxis.png")
-```
-
 Rock doves are socially monogamous and offer bi-parental care, making inter- and intra-sexual comparisons possible.
@@
 # Materials and Methods
-```{r, out.width = "360px", echo = F}
-knitr::include_graphics("../figures/images/samplingtimepoints.png")
+
+##  Rock dove housing
+
+Sexually mature male and female rock doves were  housed together in rooftop aviaries on the Barnard Campus.
+Natural male-female pairings permitted, and small observation windows behind each nest box (64 boxes total) facilitate observations nest building, egg laying, chick hatching, and nesting care every morning from 0900-1100 (correct time?).
+On specific collection days (Fig 1A, needs to be updated), truck blook, the hypothalamus, the pituitary, and the gonads were collected following common protocols used to sample tissue from birds (Fig 1B).
+Tissue collection, RNA isolation, cDNA library preparation, sequencing, quality control, and read processing were done as described in Calisi et al.
+2018, Lang et al 2019, and Austin et al 2019.
+
+```{r, fig.width= 4, echo = F}
+a <- ggdraw() + draw_image("../figures/images/samplingtimepoints.png") 
+b <- ggdraw() + draw_image("../figures/images/PigeonHPGaxis.png")
+
+plot_grid(a,b, labels = "AUTO", label_size = 8, ncol = 1)
 ```
+
+We used EdgeR and DESeq2 to calculate statistical differnces in gene expression between each transition. All subsequent analyses were conducting using variance stabilized data from the DESeq2 model.
 
@@
 
-## Table 1: comparison of total DEGs determined by edgeR and DESeq2
+Thosands of genes were found to be differentially expressed between the non-parental control group and the building group (Table 1, Fig.
+1).
+Within the time course of parental care, fewer than 1000 genes change there expression over the course of a few days, with the most pronounced difference in toltal number of changes occuring in the female pituitary between incubation days 9 and 17 (Table 1, Fig.
+1).
 
-## Fig 1: Total number of gene expression differences between each stage
+\begin{table}[]
+\begin{tabular}{llllllll}
+\hline
+\textbf{Method} & \textbf{Contrast} & \textbf{Hyp F} & \textbf{Hyp M} & \textbf{Pit F} & \textbf{Pit M} & \textbf{Gon F} & \textbf{Gon M} \\  \hline
+EdgeR & Cont - Bldg & 14674 & 117 & 14593 & 144 & 14859 & 4 \\
+EdgeR & Bldg - Lay & 14742 & 0 & 14593 & 0 & 14836 & 0 \\
+EdgeR & Lay - Inc3 & 0 & 0 & 0 & 0 & 19 & 0 \\
+EdgeR & Inc3 - Inc9 & 0 & 1 & 0 & 3 & 2 & 0 \\
+EdgeR & Inc9 - Inc 17 & 0 & 2 & 57 & 37 & 9 & 0 \\
+EdgeR & Inc17 - Hatch & 0 & 0 & 3 & 1 & 3 & 0 \\
+EdgeR & Hatch - Nest5 & 1 & 0 & 3 & 0 & 0 & 0 \\
+EdgeR & Nest5 - Nest9 & 0 & 0 & 0 & 0 & 24 & 0 \\  \hline
+DESeq2 & Cont - Bldg & 3452 & 4499 & 3950 & 4680 & 4597 & 2135 \\
+DESeq2 & Bldg - Lay & 0 & 0 & 101 & 0 & 104 & 1 \\
+DESeq2 & Lay - Inc3 & 0 & 0 & 18 & 1 & 440 & 1 \\
+DESeq2 & Inc3 - Inc9 & 1 & 0 & 0 & 1 & 8 & 0 \\
+DESeq2 & Inc9 - Inc 17 & 1 & 0 & 640 & 186 & 1 & 0 \\
+DESeq2 & Inc17 - Hatch & 2 & 0 & 3 & 18 & 1 & 0 \\
+DESeq2 & Hatch - Nest5 & 166 & 0 & 340 & 3 & 1 & 0 \\
+DESeq2 & Nest5 - Nest9 & 0 & 0 & 1 & 23 & 19 & 0 \\ \hline
+\end{tabular}
+\caption{Comparison of total DEGs determined by edgeR and DESeq2}
+\label{tab:my-table}
+\end{table}
 
@@
 
-**Fig x. The magnitude of gene expression changes between each parental transition**.
-4-5K genes are differentially expressed beteen control birds and their nest buliding conspecifics in all tissues except the male gonad.
+**Fig 1. The magnitude of gene expression changes between each parental transition**.
+Using DESeq 2, we identified 4-5K genes are differentially expressed beteen control birds and their nest buliding conspecifics in all tissues except the male gonad.
 500 - 1000 genes are differentially expressed in the female pituitary from mid-late incubation as well as in the female hypothalamus and pituitary from hatch to nestingly care day 5.
@@
 ```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-ggdraw() + draw_image("../figures/sexes/DoveParentsRNAseq_model2.png") 
+a <- ggdraw() + draw_image("../figures/sexes/LDA-4.png") 
+b <- ggdraw() + draw_image("../figures/sexes/DoveParentsRNAseq_model2.png")
+
+plot_grid(a,b, labels = "AUTO", label_size = 8, ncol = 1)
 ```

### 2019-09-07 docs/characterization_manuscript.tex
new figs

@@
 
-\includegraphics[width=360px]{../figures/images/PigeonHPGaxis}
-
 Rock doves are socially monogamous and offer bi-parental care, making
@@
 
-\includegraphics[width=360px]{../figures/images/samplingtimepoints}
+\hypertarget{rock-dove-housing}{%
+\subsection{Rock dove housing}\label{rock-dove-housing}}
+
+Sexually mature male and female rock doves were housed together in
+rooftop aviaries on the Barnard Campus. Natural male-female pairings
+permitted, and small observation windows behind each nest box (64 boxes
+total) facilitate observations nest building, egg laying, chick
+hatching, and nesting care every morning from 0900-1100 (correct time?).
+On specific collection days (Fig 1A, needs to be updated), truck blook,
+the hypothalamus, the pituitary, and the gonads were collected following
+common protocols used to sample tissue from birds (Fig 1B). Tissue
+collection, RNA isolation, cDNA library preparation, sequencing, quality
+control, and read processing were done as described in Calisi et
+al.~2018, Lang et al 2019, and Austin et al 2019.
+
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
+
+We used EdgeR and DESeq2 to calculate statistical differnces in gene
+expression between each transition. All subsequent analyses were
+conducting using variance stabilized data from the DESeq2 model.
 
@@
 
-\hypertarget{table-1-comparison-of-total-degs-determined-by-edger-and-deseq2}{%
-\subsection{Table 1: comparison of total DEGs determined by edgeR and
-DESeq2}\label{table-1-comparison-of-total-degs-determined-by-edger-and-deseq2}}
+Thosands of genes were found to be differentially expressed between the
+non-parental control group and the building group (Table 1, Fig. 1).
+Within the time course of parental care, fewer than 1000 genes change
+there expression over the course of a few days, with the most pronounced
+difference in toltal number of changes occuring in the female pituitary
+between incubation days 9 and 17 (Table 1, Fig. 1).
 
-\hypertarget{fig-1-total-number-of-gene-expression-differences-between-each-stage}{%
-\subsection{Fig 1: Total number of gene expression differences between
-each
-stage}\label{fig-1-total-number-of-gene-expression-differences-between-each-stage}}
+\begin{table}[]
+\begin{tabular}{llllllll}
+\hline
+\textbf{Method} & \textbf{Contrast} & \textbf{Hyp F} & \textbf{Hyp M} & \textbf{Pit F} & \textbf{Pit M} & \textbf{Gon F} & \textbf{Gon M} \\  \hline
+EdgeR & Cont - Bldg & 14674 & 117 & 14593 & 144 & 14859 & 4 \\
+EdgeR & Bldg - Lay & 14742 & 0 & 14593 & 0 & 14836 & 0 \\
+EdgeR & Lay - Inc3 & 0 & 0 & 0 & 0 & 19 & 0 \\
+EdgeR & Inc3 - Inc9 & 0 & 1 & 0 & 3 & 2 & 0 \\
+EdgeR & Inc9 - Inc 17 & 0 & 2 & 57 & 37 & 9 & 0 \\
+EdgeR & Inc17 - Hatch & 0 & 0 & 3 & 1 & 3 & 0 \\
+EdgeR & Hatch - Nest5 & 1 & 0 & 3 & 0 & 0 & 0 \\
+EdgeR & Nest5 - Nest9 & 0 & 0 & 0 & 0 & 24 & 0 \\  \hline
+DESeq2 & Cont - Bldg & 3452 & 4499 & 3950 & 4680 & 4597 & 2135 \\
+DESeq2 & Bldg - Lay & 0 & 0 & 101 & 0 & 104 & 1 \\
+DESeq2 & Lay - Inc3 & 0 & 0 & 18 & 1 & 440 & 1 \\
+DESeq2 & Inc3 - Inc9 & 1 & 0 & 0 & 1 & 8 & 0 \\
+DESeq2 & Inc9 - Inc 17 & 1 & 0 & 640 & 186 & 1 & 0 \\
+DESeq2 & Inc17 - Hatch & 2 & 0 & 3 & 18 & 1 & 0 \\
+DESeq2 & Hatch - Nest5 & 166 & 0 & 340 & 3 & 1 & 0 \\
+DESeq2 & Nest5 - Nest9 & 0 & 0 & 1 & 23 & 19 & 0 \\ \hline
+\end{tabular}
+\caption{Comparison of total DEGs determined by edgeR and DESeq2}
+\label{tab:my-table}
+\end{table}
 
-\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
 
-\textbf{Fig x. The magnitude of gene expression changes between each
-parental transition}. 4-5K genes are differentially expressed beteen
-control birds and their nest buliding conspecifics in all tissues except
-the male gonad. 500 - 1000 genes are differentially expressed in the
-female pituitary from mid-late incubation as well as in the female
-hypothalamus and pituitary from hatch to nestingly care day 5.
+\textbf{Fig 1. The magnitude of gene expression changes between each
+parental transition}. Using DESeq 2, we identified 4-5K genes are
+differentially expressed beteen control birds and their nest buliding
+conspecifics in all tissues except the male gonad. 500 - 1000 genes are
+differentially expressed in the female pituitary from mid-late
+incubation as well as in the female hypothalamus and pituitary from
+hatch to nestingly care day 5.
 
@@
 
-\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
 
@@
 
-\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
 
@@
 
-\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
 

### 2019-09-07 docs/characterization_manuscript.Rmd
now with 5 figs

@@
 This data can be used to develop and test hypotheses about the mechanism regulating parental care behavior.
-author_summary: |
-  Lorem ipsum dolor sit amet, consectetur adipiscing elit. 
 bibliography: mybibfile.bib
@@
 2012); and 2) rock doves are prolific, year-round breeders that thrive in captivity, making observation, manipulation, and sampling highly feasible year-round.
-
 Rock doves are socially monogamous and offer bi-parental care, making inter- and intra-sexual comparisons possible.
@@
 
-```{r, fig.width= 4, echo = F}
+```{r, fig.width= 5, echo = F, fig.cap = "Experimental design.We sampled 9 timepoints, including 1 non-parental control, 1 nest-building, 1 egg lay, 3 incubation time point, 1 hatch, and 2 nestling care stages.
+we collected truck blood, the hypothalamus, the pituitary, and the gonads from all parents."}
 a <- ggdraw() + draw_image("../figures/images/samplingtimepoints.png") 
@@
 
-We used EdgeR and DESeq2 to calculate statistical differnces in gene expression between each transition. All subsequent analyses were conducting using variance stabilized data from the DESeq2 model.
+EdgeR and DESeq2 were used to calculate statistical differnces in gene expression between each transition.
+All subsequent analyses were conducting using variance stabilized data from the DESeq2 model.
+A linear discriminant analysis was used to determine the confidence with which we could predict the parental care stage of a sample based on it's global gene expression profile.
+WGCNA was used to identify genes who's expression was correlated with specific candidate genes of interest.
 
@@
 
-Thosands of genes were found to be differentially expressed between the non-parental control group and the building group (Table 1, Fig.
+Thosands of genes were found to be differentially expressed between the non-parental control group and the nest-building group (Table 1, Fig.
 1).
 Within the time course of parental care, fewer than 1000 genes change there expression over the course of a few days, with the most pronounced difference in toltal number of changes occuring in the female pituitary between incubation days 9 and 17 (Table 1, Fig.
-1).
+2).
 
-\begin{table}[]
+\begin{table}[hb]
 \begin{tabular}{llllllll}
@@
 
-```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap = "The magnitude of gene expression changes between each parental transition.
+Using DESeq 2, we identified 4-5K genes are differentially expressed beteen control birds and their nest buliding conspecifics in all tissues except the male gonad.
+500 - 1000 genes are differentially expressed in the female pituitary from mid-late incubation as well as in the female hypothalamus and pituitary from hatch to nestingly care day 5."}
 ggdraw() + draw_image("../figures/characterization/linegraphs-1.png") 
@@
 
-**Fig 1. The magnitude of gene expression changes between each parental transition**.
-Using DESeq 2, we identified 4-5K genes are differentially expressed beteen control birds and their nest buliding conspecifics in all tissues except the male gonad.
-500 - 1000 genes are differentially expressed in the female pituitary from mid-late incubation as well as in the female hypothalamus and pituitary from hatch to nestingly care day 5.
+Even though we have a large sample size of roughly 12 samples per tissue per sex, the differences between all 8 sampled parental time points are so small, that we are only able to predict the stage using gene expression along 40%, 30% and 27% of the time in the pituiary, gonads and hypothalamus, respectively (Fig 3A).
+However, if we group the parental stages into anticipatory (control, nest building), egg care (lay and incubation days), and nestling care (hatch and nesting days), we incrase our ability to detect parental stage to 47%, 43%, and 37 pituiary, hypothalamus and gonads, respectively (Fig 3B).
+The
 
-## Fig 2: Ability to use gene expression to distingush groups
-
-```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap= "Linear discriminant analysis shows the strongest ability to predict parental stage of pitutary samples."}
 a <- ggdraw() + draw_image("../figures/sexes/LDA-4.png") 
@@
 
-## Fig 3: Prolactin expression in PIT (but not HYP or GON) follows same pattern as in blood
+Gene expression of prlactin (_PRL_)in the pitutiary, but not the hypothalamus follows the characteristic dip in expression relative to non-parental controls during early incubation then spikes the day before hatch and peaks at hatch (Fig 4).
+Intrestingly, hundreds of genes follow this same, prolactin-associated pattern of expression in the pitutiary (Fig.
+5).
 
-```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap = "Prolactin expression in PIT (but not HYP or GON) follows same pattern as in blood"}
 ggdraw() + draw_image("../figures/characterization/prolactin-1.png") 
@@
 
-## Fig 4: Hundreds of genes show similar pattern in expression as circulating prolactin
-
-```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-ggdraw() + draw_image("../figures/sexes/wgcna-7.png")
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap = "Hundreds of genes show similar pattern in expression as circulating prolactin"}
+ggdraw() + draw_image("../figures/sexes/wgcna2-1.png")
 ```

### 2019-09-07 docs/characterization_manuscript.tex
now with 5 figs

@@
 % Author Summary not valid for PLOS ONE submissions.
-\section*{Author summary}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 
@@
 captivity, making observation, manipulation, and sampling highly
-feasible year-round.
-
-Rock doves are socially monogamous and offer bi-parental care, making
-inter- and intra-sexual comparisons possible. Birds offer two important
-behavioral transition points into parental care: the incubation of eggs
-and the caring for chicks. This produces two unique opportunities to
-study how the brain transitions into two different suites of parental
-care behaviors. Additionally, rock doves exhibit a parental care
-strategy analogous to mammals in that they, too, `lactate' to feed their
-young (Gillespie et al.~2011, 2012). This lactation, unlike simple
-regurgitation of food, consists of the production and sloughing off of
-skin cells inside the crop sac of females and males, creating a
-protein-rich milk-like substance on which they rear their chicks. Many
-functional similarities between rock dove and mammalian lactation exist
-concerning the mediation of this event by the hormone prolactin (Dumont
-1965). Additionally, like mammalian milk, rock dove milk delivers
-essential immunoglobulins and nutritional benefits to young, aiding in
-their immune function and development of microbiota {[}2{]}. Thus,
-because rock doves incubate eggs and exhibit mammalian-like mediation
-and function of lactation for young, they have the potential to serve as
-a powerful theoretical bridge to understand the neurobiology of both
-avian and mammalian transitions into parental care. Our working
-hypothesis is that distinct changes in transcription occur in the brain
-at the anticipation of, during, and in response to two different types
-of parental care: incubation behavior and hatchling care.
+feasible year-round. Rock doves are socially monogamous and offer
+bi-parental care, making inter- and intra-sexual comparisons possible.
+Birds offer two important behavioral transition points into parental
+care: the incubation of eggs and the caring for chicks. This produces
+two unique opportunities to study how the brain transitions into two
+different suites of parental care behaviors. Additionally, rock doves
+exhibit a parental care strategy analogous to mammals in that they, too,
+`lactate' to feed their young (Gillespie et al.~2011, 2012). This
+lactation, unlike simple regurgitation of food, consists of the
+production and sloughing off of skin cells inside the crop sac of
+females and males, creating a protein-rich milk-like substance on which
+they rear their chicks. Many functional similarities between rock dove
+and mammalian lactation exist concerning the mediation of this event by
+the hormone prolactin (Dumont 1965). Additionally, like mammalian milk,
+rock dove milk delivers essential immunoglobulins and nutritional
+benefits to young, aiding in their immune function and development of
+microbiota {[}2{]}. Thus, because rock doves incubate eggs and exhibit
+mammalian-like mediation and function of lactation for young, they have
+the potential to serve as a powerful theoretical bridge to understand
+the neurobiology of both avian and mammalian transitions into parental
+care. Our working hypothesis is that distinct changes in transcription
+occur in the brain at the anticipation of, during, and in response to
+two different types of parental care: incubation behavior and hatchling
+care.
 
@@
 
+\begin{figure}
+\centering
 \includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{Experimental design.We sampled 9 timepoints, including 1
+non-parental control, 1 nest-building, 1 egg lay, 3 incubation time
+point, 1 hatch, and 2 nestling care stages. we collected truck blood,
+the hypothalamus, the pituitary, and the gonads from all parents.}
+\end{figure}
 
-We used EdgeR and DESeq2 to calculate statistical differnces in gene
+EdgeR and DESeq2 were used to calculate statistical differnces in gene
 expression between each transition. All subsequent analyses were
-conducting using variance stabilized data from the DESeq2 model.
+conducting using variance stabilized data from the DESeq2 model. A
+linear discriminant analysis was used to determine the confidence with
+which we could predict the parental care stage of a sample based on it's
+global gene expression profile. WGCNA was used to identify genes who's
+expression was correlated with specific candidate genes of interest.
 
@@
 Thosands of genes were found to be differentially expressed between the
-non-parental control group and the building group (Table 1, Fig. 1).
-Within the time course of parental care, fewer than 1000 genes change
-there expression over the course of a few days, with the most pronounced
-difference in toltal number of changes occuring in the female pituitary
-between incubation days 9 and 17 (Table 1, Fig. 1).
+non-parental control group and the nest-building group (Table 1, Fig.
+1). Within the time course of parental care, fewer than 1000 genes
+change there expression over the course of a few days, with the most
+pronounced difference in toltal number of changes occuring in the female
+pituitary between incubation days 9 and 17 (Table 1, Fig. 2).
 
-\begin{table}[]
+\begin{table}[hb]
 \begin{tabular}{llllllll}
@@
 
+\begin{figure}
+\centering
 \includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
+\caption{The magnitude of gene expression changes between each parental
+transition. Using DESeq 2, we identified 4-5K genes are differentially
+expressed beteen control birds and their nest buliding conspecifics in
+all tissues except the male gonad. 500 - 1000 genes are differentially
+expressed in the female pituitary from mid-late incubation as well as in
+the female hypothalamus and pituitary from hatch to nestingly care day
+5.}
+\end{figure}
 
-\textbf{Fig 1. The magnitude of gene expression changes between each
-parental transition}. Using DESeq 2, we identified 4-5K genes are
-differentially expressed beteen control birds and their nest buliding
-conspecifics in all tissues except the male gonad. 500 - 1000 genes are
-differentially expressed in the female pituitary from mid-late
-incubation as well as in the female hypothalamus and pituitary from
-hatch to nestingly care day 5.
+Even though we have a large sample size of roughly 12 samples per tissue
+per sex, the differences between all 8 sampled parental time points are
+so small, that we are only able to predict the stage using gene
+expression along 40\%, 30\% and 27\% of the time in the pituiary, gonads
+and hypothalamus, respectively (Fig 3A). However, if we group the
+parental stages into anticipatory (control, nest building), egg care
+(lay and incubation days), and nestling care (hatch and nesting days),
+we incrase our ability to detect parental stage to 47\%, 43\%, and 37
+pituiary, hypothalamus and gonads, respectively (Fig 3B). The
 
-\hypertarget{fig-2-ability-to-use-gene-expression-to-distingush-groups}{%
-\subsection{Fig 2: Ability to use gene expression to distingush
-groups}\label{fig-2-ability-to-use-gene-expression-to-distingush-groups}}
+\begin{figure}
+\centering
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{Linear discriminant analysis shows the strongest ability to
+predict parental stage of pitutary samples.}
+\end{figure}
 
-\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
+Gene expression of prlactin (\emph{PRL})in the pitutiary, but not the
+hypothalamus follows the characteristic dip in expression relative to
+non-parental controls during early incubation then spikes the day before
+hatch and peaks at hatch (Fig 4). Intrestingly, hundreds of genes follow
+this same, prolactin-associated pattern of expression in the pitutiary
+(Fig. 5).
 
-\hypertarget{fig-3-prolactin-expression-in-pit-but-not-hyp-or-gon-follows-same-pattern-as-in-blood}{%
-\subsection{Fig 3: Prolactin expression in PIT (but not HYP or GON)
-follows same pattern as in
-blood}\label{fig-3-prolactin-expression-in-pit-but-not-hyp-or-gon-follows-same-pattern-as-in-blood}}
+\begin{figure}
+\centering
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+\caption{Prolactin expression in PIT (but not HYP or GON) follows same
+pattern as in blood}
+\end{figure}
 
-\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
-
-\hypertarget{fig-4-hundreds-of-genes-show-similar-pattern-in-expression-as-circulating-prolactin}{%
-\subsection{Fig 4: Hundreds of genes show similar pattern in expression
-as circulating
-prolactin}\label{fig-4-hundreds-of-genes-show-similar-pattern-in-expression-as-circulating-prolactin}}
-
+\begin{figure}
+\centering
 \includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
+\caption{Hundreds of genes show similar pattern in expression as
+circulating prolactin}
+\end{figure}
 

### 2019-09-07 docs/characterization_manuscript.tex
looking better

@@
 
-\hypertarget{rock-dove-housing}{%
-\subsection{Rock dove housing}\label{rock-dove-housing}}
-
 Sexually mature male and female rock doves were housed together in
@@
 
-\begin{figure}
-\centering
-\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
-\caption{Experimental design.We sampled 9 timepoints, including 1
-non-parental control, 1 nest-building, 1 egg lay, 3 incubation time
-point, 1 hatch, and 2 nestling care stages. we collected truck blood,
-the hypothalamus, the pituitary, and the gonads from all parents.}
-\end{figure}
-
-EdgeR and DESeq2 were used to calculate statistical differnces in gene
-expression between each transition. All subsequent analyses were
+We used EdgeR determine the greatest contributor to variance in gene
+expression in the full dataset. Then, we used EdgeR and DESeq2 were used
+to calculate statistical differnces in gene expression between each
+transition within a given tissue. All subsequent analyses were
 conducting using variance stabilized data from the DESeq2 model. A
@@
 
-Thosands of genes were found to be differentially expressed between the
+We found that the majority of the variation in the full dataset could be
+attributed to tissue, with only a notable sex differnce existing between
+the female ovaries and the male testies (Fig. 2).
+
+Thousands of genes were found to be differentially expressed between the
 non-parental control group and the nest-building group (Table 1, Fig.
-1). Within the time course of parental care, fewer than 1000 genes
+3). Within the time course of parental care, fewer than 1000 genes
 change there expression over the course of a few days, with the most
 pronounced difference in toltal number of changes occuring in the female
-pituitary between incubation days 9 and 17 (Table 1, Fig. 2).
+pituitary between incubation days 9 and 17 (Table 1, Fig. 3).
+
+Even though we have a large sample size of roughly 12 samples per tissue
+per sex, the differences between all 8 sampled parental time points are
+so small, that we are only able to predict the stage using gene
+expression along 40\%, 30\% and 27\% of the time in the pituiary, gonads
+and hypothalamus, respectively (Fig 4A). However, if we group the
+parental stages into anticipatory (control, nest building), egg care
+(lay and incubation days), and nestling care (hatch and nesting days),
+we incrase our ability to detect parental stage to 47\%, 43\%, and 37
+pituiary, hypothalamus and gonads, respectively (Fig 4B).
+
+Gene expression of prlactin (\emph{PRL})in the pitutiary, but not the
+hypothalamus follows the characteristic dip in expression relative to
+non-parental controls during early incubation then spikes the day before
+hatch and peaks at hatch (Fig 5).
+
+Hundreds of genes follow this same, prolactin-associated pattern of
+expression in the pitutiary (Fig. 6).
+
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
+
+This project is a synergistic collaboration between the PI, Rebecca
+Calisi-Rodríguez (expertise in avian behavior, parental care and
+neurobiology), co-PI, Matthew MacManes (expertise in next-generation
+sequencing, transcriptome assembly, and gene expression analyses), and
+Collaborator, Rae Silver (expertise in neurobiology, dove behavior, and
+decades of successful breeding and maintenance of dove colonies at
+Barnard College).
 
@@
 \centering
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{Experimental design.We sampled 9 timepoints, including 1
+non-parental control, 1 nest-building, 1 egg lay, 3 incubation time
+point, 1 hatch, and 2 nestling care stages. we collected truck blood,
+the hypothalamus, the pituitary, and the gonads from all parents.}
+\end{figure}
+
+\begin{figure}
+\centering
 \includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
+\caption{MDS plots of all samples.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
 \caption{The magnitude of gene expression changes between each parental
@@
 
-Even though we have a large sample size of roughly 12 samples per tissue
-per sex, the differences between all 8 sampled parental time points are
-so small, that we are only able to predict the stage using gene
-expression along 40\%, 30\% and 27\% of the time in the pituiary, gonads
-and hypothalamus, respectively (Fig 3A). However, if we group the
-parental stages into anticipatory (control, nest building), egg care
-(lay and incubation days), and nestling care (hatch and nesting days),
-we incrase our ability to detect parental stage to 47\%, 43\%, and 37
-pituiary, hypothalamus and gonads, respectively (Fig 3B). The
-
 \begin{figure}
 \centering
-\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
 \caption{Linear discriminant analysis shows the strongest ability to
@@
 
-Gene expression of prlactin (\emph{PRL})in the pitutiary, but not the
-hypothalamus follows the characteristic dip in expression relative to
-non-parental controls during early incubation then spikes the day before
-hatch and peaks at hatch (Fig 4). Intrestingly, hundreds of genes follow
-this same, prolactin-associated pattern of expression in the pitutiary
-(Fig. 5).
-
 \begin{figure}
 \centering
-\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
 \caption{Prolactin expression in PIT (but not HYP or GON) follows same
@@
 \centering
-\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
+\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
 \caption{Hundreds of genes show similar pattern in expression as
@@
 \end{figure}
-
-\hypertarget{acknowledgments}{%
-\section{Acknowledgments}\label{acknowledgments}}
-
-This project is a synergistic collaboration between the PI, Rebecca
-Calisi-Rodríguez (expertise in avian behavior, parental care and
-neurobiology), co-PI, Matthew MacManes (expertise in next-generation
-sequencing, transcriptome assembly, and gene expression analyses), and
-Collaborator, Rae Silver (expertise in neurobiology, dove behavior, and
-decades of successful breeding and maintenance of dove colonies at
-Barnard College).
 

### 2019-09-07 docs/characterization_manuscript.Rmd
looking better

@@
 
-##  Rock dove housing
-
 Sexually mature male and female rock doves were housed together in rooftop aviaries on the Barnard Campus.
@@
 
-```{r, fig.width= 5, echo = F, fig.cap = "Experimental design.We sampled 9 timepoints, including 1 non-parental control, 1 nest-building, 1 egg lay, 3 incubation time point, 1 hatch, and 2 nestling care stages.
-we collected truck blood, the hypothalamus, the pituitary, and the gonads from all parents."}
-a <- ggdraw() + draw_image("../figures/images/samplingtimepoints.png") 
-b <- ggdraw() + draw_image("../figures/images/PigeonHPGaxis.png")
-
-plot_grid(a,b, labels = "AUTO", label_size = 8, ncol = 1)
-```
-
-EdgeR and DESeq2 were used to calculate statistical differnces in gene expression between each transition.
+We used EdgeR determine the greatest contributor to variance in gene expression in the full dataset.
+Then, we used EdgeR and DESeq2 were used to calculate statistical differnces in gene expression between each transition within a given tissue.
 All subsequent analyses were conducting using variance stabilized data from the DESeq2 model.
@@
 
-Thosands of genes were found to be differentially expressed between the non-parental control group and the nest-building group (Table 1, Fig.
-1).
+We found that the majority of the variation in the full dataset could be attributed to tissue, with only a notable sex differnce existing between the female ovaries and the male testies (Fig. 2).
+
+Thousands of genes were found to be differentially expressed between the non-parental control group and the nest-building group (Table 1, Fig.
+3).
 Within the time course of parental care, fewer than 1000 genes change there expression over the course of a few days, with the most pronounced difference in toltal number of changes occuring in the female pituitary between incubation days 9 and 17 (Table 1, Fig.
-2).
+3).
+
+Even though we have a large sample size of roughly 12 samples per tissue per sex, the differences between all 8 sampled parental time points are so small, that we are only able to predict the stage using gene expression along 40%, 30% and 27% of the time in the pituiary, gonads and hypothalamus, respectively (Fig 4A).
+However, if we group the parental stages into anticipatory (control, nest building), egg care (lay and incubation days), and nestling care (hatch and nesting days), we incrase our ability to detect parental stage to 47%, 43%, and 37 pituiary, hypothalamus and gonads, respectively (Fig 4B).
+
+Gene expression of prlactin (_PRL_)in the pitutiary, but not the hypothalamus follows the characteristic dip in expression relative to non-parental controls during early incubation then spikes the day before hatch and peaks at hatch (Fig 5).
+
+Hundreds of genes follow this same, prolactin-associated pattern of expression in the pitutiary (Fig. 6).
+
+# Acknowledgments
+
+This project is a synergistic collaboration between the PI, Rebecca Calisi-Rodríguez (expertise in avian behavior, parental care and neurobiology), co-PI, Matthew MacManes (expertise in next-generation sequencing, transcriptome assembly, and gene expression analyses), and Collaborator, Rae Silver (expertise in neurobiology, dove behavior, and decades of successful breeding and maintenance of dove colonies at Barnard College).
 
@@
 
+```{r, fig.width= 5, echo = F, fig.cap = "Experimental design.We sampled 9 timepoints, including 1 non-parental control, 1 nest-building, 1 egg lay, 3 incubation time point, 1 hatch, and 2 nestling care stages.
+we collected truck blood, the hypothalamus, the pituitary, and the gonads from all parents."}
+a <- ggdraw() + draw_image("../figures/images/samplingtimepoints.png") 
+b <- ggdraw() + draw_image("../figures/images/PigeonHPGaxis.png")
+
+plot_grid(a,b, labels = "AUTO", label_size = 8, ncol = 1)
+```
+
+```{r, fig.width= 5, echo = F, fig.cap = "MDS plots of all samples."}
+ggdraw() + draw_image("../figures/limma/plotMDS-colored-1.png") 
+```
+
 ```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap = "The magnitude of gene expression changes between each parental transition.
@@
 ```
-
-Even though we have a large sample size of roughly 12 samples per tissue per sex, the differences between all 8 sampled parental time points are so small, that we are only able to predict the stage using gene expression along 40%, 30% and 27% of the time in the pituiary, gonads and hypothalamus, respectively (Fig 3A).
-However, if we group the parental stages into anticipatory (control, nest building), egg care (lay and incubation days), and nestling care (hatch and nesting days), we incrase our ability to detect parental stage to 47%, 43%, and 37 pituiary, hypothalamus and gonads, respectively (Fig 3B).
-The
 
@@
 
-Gene expression of prlactin (_PRL_)in the pitutiary, but not the hypothalamus follows the characteristic dip in expression relative to non-parental controls during early incubation then spikes the day before hatch and peaks at hatch (Fig 4).
-Intrestingly, hundreds of genes follow this same, prolactin-associated pattern of expression in the pitutiary (Fig.
-5).
-
 ```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap = "Prolactin expression in PIT (but not HYP or GON) follows same pattern as in blood"}
@@
 ```
-
-# Acknowledgments
-
-This project is a synergistic collaboration between the PI, Rebecca Calisi-Rodríguez (expertise in avian behavior, parental care and neurobiology), co-PI, Matthew MacManes (expertise in next-generation sequencing, transcriptome assembly, and gene expression analyses), and Collaborator, Rae Silver (expertise in neurobiology, dove behavior, and decades of successful breeding and maintenance of dove colonies at Barnard College).
 

### 2019-09-07 docs/manipulation_manuscript.Rmd
looking better
now with new fig

--- 2019-09-06 docs/manipulation_manuscript.Rmd
@@
 This matters because...
-author_summary: |
-  Lorem ipsum dolor sit amet, consectetur adipiscing elit. 
 bibliography: mybibfile.bib
@@
 
-```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-library(cowplot)
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap= "Circulating prolactin levels with an without manipulation."}
+ggdraw() + draw_image("../figures/hormones/prolactin-manip-1.png") 
+```
+
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap= "Offspring removal."}
 ggdraw() + draw_image("../figures/manipulation/offspringremoval-1.png") 
+```
+
+```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap = "Manipulation of timing of hatch with dummy eggs and foster chicks. "}
+ggdraw() + draw_image("../figures/manipulation/prolongextend-1.png") 
 ```

### 2019-09-07 docs/characterization_manuscript.Rmd
remove cache

@@
 
-```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap = "The magnitude of gene expression changes between each parental transition.
+```{r, message=FALSE, warning=FALSE, cache=F, echo = F, fig.width= 5, fig.cap = "The magnitude of gene expression changes between each parental transition.
 Using DESeq 2, we identified 4-5K genes are differentially expressed beteen control birds and their nest buliding conspecifics in all tissues except the male gonad.
@@
 
-```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap= "Linear discriminant analysis shows the strongest ability to predict parental stage of pitutary samples."}
+```{r, message=FALSE, warning=FALSE, cache=F, echo = F, fig.width= 5, fig.cap= "Linear discriminant analysis shows the strongest ability to predict parental stage of pitutary samples."}
 a <- ggdraw() + draw_image("../figures/sexes/LDA-4.png") 
@@
 
-```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap = "Prolactin expression in PIT (but not HYP or GON) follows same pattern as in blood"}
+```{r, message=FALSE, warning=FALSE, cache=F, echo = F, fig.width= 5, fig.cap = "Prolactin expression in PIT (but not HYP or GON) follows same pattern as in blood"}
 ggdraw() + draw_image("../figures/characterization/prolactin-1.png") 
@@
 
-```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5, fig.cap = "Hundreds of genes show similar pattern in expression as circulating prolactin"}
+```{r, message=FALSE, warning=FALSE, cache=F, echo = F, fig.width= 5, fig.cap = "Hundreds of genes show similar pattern in expression as circulating prolactin"}
 ggdraw() + draw_image("../figures/sexes/wgcna2-1.png")

### 2019-09-07 docs/manipulation_manuscript.Rmd
remove cache

--- 2019-09-06 docs/manipulation_manuscript.Rmd
@@
 
-```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
-library(cowplot)
+```{r, message=FALSE, warning=FALSE, cache=FALSE, echo = F, fig.width= 5, fig.cap= "Circulating prolactin levels with an without manipulation."}
+ggdraw() + draw_image("../figures/hormones/prolactin-manip-1.png") 
+```
+
+```{r, message=FALSE, warning=FALSE, cache=FALSE, echo = F, fig.width= 5, fig.cap= "Offspring removal."}
 ggdraw() + draw_image("../figures/manipulation/offspringremoval-1.png") 
+```
+
+```{r, message=FALSE, warning=FALSE, cache=FALSE, echo = F, fig.width= 5, fig.cap = "Manipulation of timing of hatch with dummy eggs and foster chicks. "}
+ggdraw() + draw_image("../figures/manipulation/prolongextend-1.png") 
 ```

### 2019-09-07 docs/characterization_manuscript.Rmd
done with updates

@@
 
-```{r, fig.width= 5, echo = F, fig.cap = "MDS plots of all samples."}
+```{r, fig.width= 5, echo = F, fig.cap = "MDS plots of all samples.
+Needs improvement but shows that tissue is source of main variation whereas sex really only affects the gonads and treatment isn't noticable at this levels of analysis."}
 ggdraw() + draw_image("../figures/limma/plotMDS-colored-1.png") 

### 2019-09-07 docs/manipulation_manuscript.tex
looking better
now with new fig
done with updates

@@
 
-\hypertarget{materials-and-methods-and-results}{%
-\section{Materials and Methods and
-Results}\label{materials-and-methods-and-results}}
+\hypertarget{materials-and-methods}{%
+\section{Materials and Methods}\label{materials-and-methods}}
 
@@
 
+\begin{figure}
+\centering
 \includegraphics{manipulation_manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{Circulating prolactin levels with an without manipulation.}
+\end{figure}
 
+\begin{figure}
+\centering
 \includegraphics{manipulation_manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
+\caption{Offspring removal.}
+\end{figure}
 
+\begin{figure}
+\centering
 \includegraphics{manipulation_manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
-
-\hypertarget{supplementary-figures}{%
-\section{Supplementary Figures}\label{supplementary-figures}}
+\caption{Manipulation of timing of hatch with dummy eggs and foster
+chicks.}
+\end{figure}
 

### 2019-09-07 analysis/04_DESeq2_sexes_manipulate.Rmd
set levels for manipuation

--- 2019-09-05 analysis/04_DESeq2_sexes.Rmd
@@
 # import "colData" which contains sample information and "countData" which contains read counts
-c.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
-c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
+c.colData <- read.csv("../metadata/00_colData_manipluation.csv", header = T, row.names = 1)
+c.countData <- read.csv("../results/00_countData_manipluation.csv", header = T, row.names = 1)
 geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
@@
 # set treatment levels
-c.colData$treatment <- factor(c.colData$treatment, levels = 
-                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
 
@@
 
-# ceate new variable for hypothesis testing
+levels(c.colData$treatment)
+# "extend"    "m.inc.d17" "m.inc.d3"  "m.inc.d8"  "m.inc.d9"  "m.n2"      "prolong"
+
+# rename some factors
 c.colData <- c.colData %>%
-    mutate(hypothesis = fct_recode(treatment,
-                            "anticipation" = "control",
-                            "anticipation" = "bldg",
-                            "incubation" = "lay",
-                            "incubation" = "inc.d3",
-                            "incubation" = "inc.d9",
-                            "incubation" = "inc.d17",
-                            "hatchling.care" = "hatch",
-                            "hatchling.care" = "n5",
-                            "hatchling.care" = "n9"))
+    mutate(treatment = fct_recode(treatment,
+                            "extend" = "extend",
+                            "prolong" = "prolong",
+                            "early" = "m.inc.d8",
+                            "remove.d03" = "m.inc.d3",
+                            "remove.d09" = "m.inc.d9",
+                            "remove.d17" = "m.inc.d17",
+                            "remove.d20" = "m.n2"))
+
+c.colData$treatment <- factor(c.colData$treatment, levels = maniplevels)
 

### 2019-09-07 analysis/04_DESeq2_sexes_manipulate.Rmd
now with lda, pca, wgcna no manip data

--- 2019-09-04 analysis/04_DESeq2_sexes.Rmd
@@
 ---
-title: "04_DESeq_sexes"
+title: "04_DESeq_sexes_manipulate"
 output: md_document
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/sexes/', cache = TRUE,  message=F, comment=FALSE, warning=FALSE)
+knitr::opts_chunk$set(fig.path = '../figures/manipulation/', cache = TRUE,  message=F, comment=FALSE, warning=FALSE)
 ```
@@
 
-# ceate new variable for hypothesis testing
+# rename some factors
+levels(c.colData$treatment)
+# "extend"    "m.inc.d17" "m.inc.d3"  "m.inc.d8"  "m.inc.d9"  "m.n2"      "prolong" 
 c.colData <- c.colData %>%
-    mutate(hypothesis = fct_recode(treatment,
-                            "anticipation" = "control",
-                            "anticipation" = "bldg",
-                            "incubation" = "lay",
-                            "incubation" = "inc.d3",
-                            "incubation" = "inc.d9",
-                            "incubation" = "inc.d17",
-                            "hatchling.care" = "hatch",
-                            "hatchling.care" = "n5",
-                            "hatchling.care" = "n9"))
+    mutate(treatment = fct_recode(treatment,
+                            "extend" = "extend",
+                            "prolong" = "prolong",
+                            "early" = "m.inc.d8",
+                            "remove.d03" = "m.inc.d3",
+                            "remove.d09" = "m.inc.d9",
+                            "remove.d17" = "m.inc.d17",
+                            "remove.d20" = "m.n2"))
+
+c.colData$treatment <- factor(c.colData$treatment, levels = maniplevels)
+levels(c.colData$treatment)
 
@@
 head(c.colData)
+str(c.colData)
 ```
@@
 
-```{r totalDEGs, results = 'hide', eval = F}
+This one won't work yet because I have a filter step that slims the data to a subset of the characterization data.
+
+```{r totalDEGs, results = 'hide', eval = F, echo = T}
 DEGs.pituitary <- returntotalDEGs(dds.pituitary)
@@
 
-# LDA for treatment (9 groups)
+# LDA for treatment (7 groups)
 LDA.hyp1 <- LDAdata.treatment(vsd.hyp, colDataHyp)
@@
 LDA.gon1 <- LDAdata.treatment(vsd.gon, colDataGon)
-
-# LDA for hypothesis (3 groups)
-
-LDA.hyp2 <- LDAdata.hypothesis(vsd.hyp, colDataHyp)
-LDA.pit2 <- LDAdata.hypothesis(vsd.pit, colDataPit)
-LDA.gon2 <- LDAdata.hypothesis(vsd.gon, colDataGon)
 
@@
 # treatment
-a <- LDAplot.treatment(LDA.hyp1 , "Hypothalamus model 1", "parental state ~ .  0.277 pred. acc.",
+a <- LDAplot.manipulation(LDA.hyp1 , "Hypothalamus model 1", "parental state ~ .  0.277 pred. acc.",
                         "LD1, F = 7.1504779", "LD2, F = 2.4219585")
 
-b <- LDAplot.treatment(LDA.pit1 ,  "Pituitary model 1", "parental state ~ .  0.405 pred. acc.",
-                        "LD1, F = 9.6144160", "LD2, F = 3.7773314")
+b <- LDAplot.manipulation(LDA.pit1 ,  "Pituitary model 1", "parental state ~ .  0.33 pred. acc.",
+                        "LD1, F = 6.393554", "LD2, F = 3.396018")
 
-c <- LDAplot.treatment(LDA.gon1 , "Gonad model 1", "parental state ~ .  has 0.297 pred. acc.",
+c <- LDAplot.manipulation(LDA.gon1 , "Gonad model 1", "parental state ~ .  has 0.297 pred. acc.",
                         "LD1, F = 6.137500", "LD2, F = 3.320437")
@@
 
-# hypothesis
-
-d <- LDAplot.hypothesis(LDA.hyp2 , "Hypothalamus model 2", "parental stage ~ .  0.432 pred. acc.",
-                        "LD1, F = 7.386458", "LD2, F = 3.720794")
-
-e <- LDAplot.hypothesis(LDA.pit2 ,  "Pituitary model 2", "parental stage ~ .  0.473 pred. acc.",
-                        "LD1, F = 7.846659", "LD2, F = 5.076861")
-
-f <- LDAplot.hypothesis(LDA.gon2 , "Gonad model 2", "parental stage ~ .  has 0.368 pred. acc.",
-                        "LD1, F = 6.029069", "LD2, F = 3.837591")
-
-f <- f + theme(legend.direction = "horizontal")
-mylegend <- get_legend(f)
-
-temp <- plot_grid(d + theme(legend.position = "none"),
-                  e + theme(legend.position = "none"),
-                  f + theme(legend.position = "none"), nrow = 1)
-
-plot_grid(temp, mylegend, nrow = 2, rel_heights = c(1,0.3))
 ```
@@
 
-```{r wgcna, eval = T, fig.width=12, fig.height=6}
-plotWGCNAcandidates(vsd.pit, c("NP_990797.2"), colDataPit, "PRL expression")
+```{r wgcna, eval = T}
+plotWGCNAcandidatesManip(vsd.pit, c("NP_990797.2", "NP_001090992.1", "NP_990136.1", "NP_001035179.1"), colDataPit, "PRL, VIPR1, TH & AR expression, Pituitary")
 
-plotWGCNAcandidates(vsd.pit,
-                    c("NP_990797.2","NP_001004392.2", "NP_001004403.1", "NP_001006255.2", "NP_001006274.1",
+plotWGCNAcandidatesManip(vsd.hyp, c("NP_001170780.1", "NP_001090992.1", "NP_001014970.1"), colDataHyp, "VIP expression, Hypothalamus")
+
+plotWGCNAcandidatesManip(vsd.pit,c("NP_990797.2","NP_001004392.2", "NP_001004403.1", "NP_001006255.2", "NP_001006274.1",
                            "NP_989500.1", "XP_001234815.2", "XP_015149152.1", "XP_416110.5"),
-                    colDataPit, "9 of 102 genes in a WGCNA module containting Prolactin")
+                    colDataPit, "Part of 9 genes in the same WGCNA module as PRL in the characterization study")
 
-plotWGCNAcandidates(vsd.pit,
+plotWGCNAcandidatesManip(vsd.pit,
                     c("XP_428110.3","NP_001004392.2",  "XP_015142961.1", "XP_015139636.1",
-                      "XP_015137408.1", "XP_004938696.1", "NP_990620.1", "NP_001170793.2", "NP_001026250.1"),
-                    colDataPit, "Prolaction receptor: 9 of 147 genes in a WGCNA module containting PRLR")
+                      "XP_015137408.1", "XP_004938696.1", "NP_990620.1", "NP_001170793.2", "XP_015132722.1"),
+                    colDataPit, "Part of 9 genes in the same WGCNA module as PRLR in the characterization study")
 
-plotWGCNAcandidates(vsd.pit, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
+plotWGCNAcandidatesManip(vsd.pit, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
                                "NP_001131120.1", "NP_001186037.1", "NP_989871.1",
-                               "NP_990797.2", "XP_015132722.1"), colDataPit, "Pitutary: CISH, SOCS, PRL, PRLR")
+                               "NP_990797.2", "XP_015132722.1", "XP_015132722.1", "NP_001090992.1"), colDataPit, "Pitutary: CISH, SOCS, PRL, PRLR")
 
-plotWGCNAcandidates(vsd.hyp, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
+plotWGCNAcandidatesManip(vsd.hyp, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
                                "NP_001131120.1", "NP_001186037.1", "NP_989871.1",
-                               "NP_990797.2", "XP_015132722.1"), colDataHyp, "Hypothalamus: CISH, SOCS, PRL, PRLR")
+                               "NP_990797.2", "XP_015132722.1", "NP_001090992.1"), colDataHyp, "Hypothalamus: CISH, SOCS, PRL, PRLR")
 
-plotWGCNAcandidates(vsd.gon, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
+plotWGCNAcandidatesManip(vsd.gon, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
                                "NP_001131120.1", "NP_001186037.1", "NP_989871.1",
-                               "NP_990797.2", "XP_015132722.1"), colDataGon, "Gonad: CISH, SOCS, PRL, PRLR")
-
-prl36 <-  PRL_associated$entrezid[1:57]
-prl72 <-  PRL_associated$entrezid[58:102]
-
-plotWGCNAcandidates(vsd.pit, prl36, colDataPit, "A-K genes with PRL pattern in the pituitary")
-plotWGCNAcandidates(vsd.pit, prl72, colDataPit, "K-Z genes with PRL pattern in the pituitary")
-
-str(PRL_associated)
-
+                               "NP_990797.2", "XP_015132722.1", "NP_001090992.1"), colDataGon, "Gonad: CISH, SOCS, PRL, PRLR")
 ```
-
-```{r ctsGE}
-rts <- assay(vsd.pit)
-head(rts)
-
-prts <- PreparingTheIndexes(x = rts, min_cutoff=0.5, max_cutoff=0.7, mad.scale = TRUE)
-```

### 2019-09-07 R/functions.R

@@ -417,19 +417,7 @@ returnPCAs2 <- function(vsd){
 # plot pc 1 and 2 for treatment and sex
 
 plotPC12 <- function(pcadata, mysubtitle){ 
-  
-  pcadata <- pcadata %>%
-    mutate(hypothesis = fct_recode(treatment,
-                            "anticipation" = "control",
-                            "anticipation" = "bldg",
-                            "incubation" = "lay",
-                            "incubation" = "inc.d3",
-                            "incubation" = "inc.d9",
-                            "incubation" = "inc.d17",
-                            "hatchling.care" = "hatch",
-                            "hatchling.care" = "n5",
-                            "hatchling.care" = "n9"))
-  
+
   pca12 <- ggplot(pcadata, aes(PC1, PC2, color = treatment, shape = sex)) + 
     geom_point() +
     #stat_ellipse() +
@@ -580,6 +568,56 @@ plotWGCNAcandidates <- function(vsd, mygenelist, colData, mysubtitle){
          subtitle = mysubtitle) +
     scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
                        labels=charlevels)
+  return(p1)
+  
+}
+
+plotWGCNAcandidatesManip <- function(vsd, mygenelist, colData, mysubtitle){
+  
+  # make dataframe with geneids and names and counts
+  # how to gather: https://tidyr.tidyverse.org/reference/gather.html
+  
+  vsd.df <- as.data.frame(assay(vsd))
+  vsd.df$entrezid <- row.names(vsd.df)
+  
+  candidates <- full_join(geneinfo, vsd.df, by = "entrezid")
+  
+  # select genes of interst
+  
+  candidates <- candidates %>%
+    filter(entrezid %in% mygenelist) %>% droplevels()
+  
+  candidates <- candidates %>% dplyr::select(-row.names, -entrezid, -geneid)
+  
+  candidates_long <- candidates %>% gather(-Name, key = "sample", value = "value")
+  
+  candidates_long$V1 <- candidates_long$sample
+  
+  candidatecounts <- left_join(candidates_long, colData, by = "V1")
+  
+  candidatecounts$Name <- as.factor(candidatecounts$Name)
+  candidatecounts$treatment <- factor(candidatecounts$treatment, levels = maniplevels)
+  
+  head(candidatecounts)
+  
+  bysextreatment <- group_by(candidatecounts, sex, treatment, Name)
+  bysextreatment
[…]

### 2019-09-07 R/themes.R

@@ -1,7 +1,7 @@
 # custom levels
 
 charlevels <- c("control", "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9")
-maniplevels <- c( "remove.d03" ,  "remove.d09" ,  "remove.d17" , "remove.d20" , "early" , "prolong", "extend")
+maniplevels <- c( "remove.d03" ,  "remove.d09" ,  "remove.d17" , "remove.d20" , "extend" , "prolong", "early")
 
 # custom themes for plots
 

### 2019-09-07 analysis/08_WGCNA_3.R
two to one

@@ -183,3 +183,118 @@ plotDendroAndColors(sampleTree2, traitColors,
                     cex.dendroLabels = 0.4
                     )
 
+
+## part 4 combined 
+
+#=====================================================================================
+#
+#  Code chunk 1
+#
+#=====================================================================================
+
+
+
+# Load the WGCNA package
+library(WGCNA)
+library(magrittr) # for %<>%
+
+# The following setting is important, do not omit.
+
+# Allow multi-threading within WGCNA. This helps speed up certain calculations.
+# At present this call is necessary.
+# Any error here may be ignored but you may want to update WGCNA if you see one.
+# Caution: skip this line if you run RStudio or other third-party R environments.
+# See note above.
+
+
+#=====================================================================================
+#
+#  Code chunk 2
+#
+#=====================================================================================
+
+
+# Choose a set of soft-thresholding powers
+powers = c(c(1:10), seq(from = 12, to=20, by=2))
+# Call the network topology analysis function
+sft = pickSoftThreshold(datExpr, powerVector = powers, verbose = 5)
+# Plot the results:
+sizeGrWindow(9, 5)
+par(mfrow = c(1,2));
+cex1 = 0.9;
+# Scale-free topology fit index as a function of the soft-thresholding power
+plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
+     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
+     main = paste("Scale independence"));
+text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
+     labels=powers,cex=cex1,col="red");
+# this line corresponds to using an R^2 cut-off of h
+abline(h=0.60,col="red")
+# Mean connectivity as a function of the soft-thresholding power
+plot(sft$fitIndices[,1], sft$fitIndices[,5],
+     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
+     main = paste("Mean connectivity"))
+text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
+
+
+#=====================================================================================
+#
+#  Code chunk 3
+#
+#=====================================================================================
+
+
+enableWGCNAThreads()
+
+
[…]

### 2019-09-07 analysis/08_WGCNA_4.R
two to one

@@ -1,125 +1,0 @@
-#=====================================================================================
-#
-#  Code chunk 1
-#
-#=====================================================================================
-
-
-
-# Load the WGCNA package
-library(WGCNA)
-library(magrittr) # for %<>%
-
-# The following setting is important, do not omit.
-
-# Allow multi-threading within WGCNA. This helps speed up certain calculations.
-# At present this call is necessary.
-# Any error here may be ignored but you may want to update WGCNA if you see one.
-# Caution: skip this line if you run RStudio or other third-party R environments.
-# See note above.
-
-
-#=====================================================================================
-#
-#  Code chunk 2
-#
-#=====================================================================================
-
-
-# Choose a set of soft-thresholding powers
-powers = c(c(1:10), seq(from = 12, to=20, by=2))
-# Call the network topology analysis function
-sft = pickSoftThreshold(datExpr, powerVector = powers, verbose = 5)
-# Plot the results:
-sizeGrWindow(9, 5)
-par(mfrow = c(1,2));
-cex1 = 0.9;
-# Scale-free topology fit index as a function of the soft-thresholding power
-plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
-     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
-     main = paste("Scale independence"));
-text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
-     labels=powers,cex=cex1,col="red");
-# this line corresponds to using an R^2 cut-off of h
-abline(h=0.60,col="red")
-# Mean connectivity as a function of the soft-thresholding power
-plot(sft$fitIndices[,1], sft$fitIndices[,5],
-     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
-     main = paste("Mean connectivity"))
-text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
-
-
-#=====================================================================================
-#
-#  Code chunk 3
-#
-#=====================================================================================
-
-
-enableWGCNAThreads()
-
-
-# this is my attempt to solve an error
-datExpr %<>% mutate_if(is.integer,as.numeric)
-head(str(datExpr))
-
-net = blockwiseModules(datExpr, power = 10,
[…]

### 2019-09-07 analysis/04_DESeq2_sexes_manipulate.Rmd
two to one
share prl module plots

@@
 # treatment
-a <- LDAplot.manipulation(LDA.hyp1 , "Hypothalamus model 1", "parental state ~ .  0.277 pred. acc.",
-                        "LD1, F = 7.1504779", "LD2, F = 2.4219585")
+a <- LDAplot.manipulation(LDA.hyp1 , "Hypothalamus model 1", "parental state ~ .  0.185 pred. acc.",
+                        "LD1, F = 3.7984170", "LD2, F = 3.0810558")
 
-b <- LDAplot.manipulation(LDA.pit1 ,  "Pituitary model 1", "parental state ~ .  0.33 pred. acc.",
-                        "LD1, F = 6.393554", "LD2, F = 3.396018")
+b <- LDAplot.manipulation(LDA.pit1 ,  "Pituitary model 1", "parental state ~ .  0.407 pred. acc.",
+                        "LD1, F = 6.188592", "LD2, F = 3.604309")
 
-c <- LDAplot.manipulation(LDA.gon1 , "Gonad model 1", "parental state ~ .  has 0.297 pred. acc.",
-                        "LD1, F = 6.137500", "LD2, F = 3.320437")
+c <- LDAplot.manipulation(LDA.gon1 , "Gonad model 1", "parental state ~ .  has 0.154 pred. acc.",
+                        "LD1, F = 4.066997", "LD2, F = 2.782389")
 
@@
 ```
+
+## genes that were in the WGCNA module with PRL twice!
+
+```{r, sharedmodule}
+PRL_associated <- read.csv("../results/08_PRL_associated.csv", header = T)
+PRL_manipulated <- read.csv("../results/08_PRL_manipulated.csv", header = T)
+
+str(PRL_associated)
+str(PRL_manipulated)
+PRLjoin <- inner_join(PRL_associated, PRL_manipulated, by = "entrezid")
+
+PRLboth <- PRLjoin$entrezid
+PRLboth
+
+plotWGCNAcandidatesManip(vsd.pit, PRLboth, colDataPit, "PRL char and manip WGNA shared modules")
+
+prl36 <-  PRL_manipulated$entrezid[252:271]
+plotWGCNAcandidatesManip(vsd.pit, prl36, colDataPit, "A-K genes with PRL pattern in the pituitary")
+
+```

### 2019-09-07 analysis/04_DESeq2_sexes_manipulate.Rmd
figure titles
eh

@@
 
-```{r wgcna, eval = T}
+```{r wgcna, eval = T, echo = F}
 plotWGCNAcandidatesManip(vsd.pit, c("NP_990797.2", "NP_001090992.1", "NP_990136.1", "NP_001035179.1"), colDataPit, "PRL, VIPR1, TH & AR expression, Pituitary")
@@
 prl36 <-  PRL_manipulated$entrezid[252:271]
-plotWGCNAcandidatesManip(vsd.pit, prl36, colDataPit, "A-K genes with PRL pattern in the pituitary")
-
+plotWGCNAcandidatesManip(vsd.pit, prl36, colDataPit, "PRL-RIPK3 pattern in the PRL module")
 ```

### 2019-09-07 analysis/04_DESeq2_sexes_manipulate.Rmd
se false

@@
 
-```{r, sharedmodule}
+```{r, sharedmodule, fig.height= 8}
 PRL_associated <- read.csv("../results/08_PRL_associated.csv", header = T)

### 2019-10-08 analysis/07_hormones.Rmd
prolactin analysis

---
title: "hormones"
output: md_document
---

```{r setup}
library(tidyverse)
library(cowplot)
library(readxl)
library(modelr)
library(lubridate)

source("../R/functions.R")  # load custom functions 
source("../R/themes.R")  # load custom themes and color palletes

knitr::opts_chunk$set(fig.path = '../figures/hormones/',message=F, warning=FALSE)
```

```{r wrangle}
prolactin <- read_excel("../results/Pigeon prolactin concentrations juil 2018.xlsx", sheet = 1)

# keep only samples realated to parental care
prolactin <- prolactin %>% filter(Study %in% c("Baseline", "ParentalCare"))

# create lists of factos for characterization and manipluation studies
charlevels <- c("control", "bldg", "lay", "inc_d3", "inc_d9", "inc_d17", "hatch", "n5", "n9" )
maniplevels <- c("M_inc3", "early", "M_inc9", "M_inc17", "M_n2", "prolong", "extend")
combolevels <- c("control", "bldg", "lay", "inc_d3", "M_inc3",
                  "inc_d9",  "M_inc9",  "inc_d17","M_inc17",  "hatch", "M_n2",
                  "n5", "n9" , "early",  "prolong", "extend")

# rename some of the treatments levels, store in new columns
prolactin <- prolactin %>%
    mutate(sex = fct_recode(Sex,
                            "female" = "f",
                            "male" = "m"),
           treatment = fct_recode(Treatment,
                            "hatch" = "Hatch",
                            "inc_d17" = "Inc_d17",
                            "inc_d3" = "Inc_d3",
                            "inc_d9" = "Inc_d9",
                            "M_inc9" = "M_Inc9",
                            "M_inc3" = "M_Inc3",
                            "early" = "M_Inc8",
                            "early" = "M_inc8",
                            "M_inc17" = "M_Inc17",
                            "M_n2" = "M_hatch",
                            "control" = "baseline",
                            "n5" = "N5", 
                            "n9" = "N9"),
           study = fct_collapse(treatment,
                                 characterization = charlevels,
                                 manipulation = maniplevels))

prolactin$treatment <- factor(prolactin$treatment, levels = combolevels)
```

```{r prolactin-all}
prolactin %>%
  select(treatment, `Prolactin ng/mL`, Sex, study) %>% drop_na() %>%
ggplot(aes(x = treatment, y = `Prolactin ng/mL`, fill = Sex)) +
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom") +
  facet_wrap(~study, scales = "free_x") +
  labs(title = "Ciruclating prolactin levels",
       subtitle = "During characteristic parental care stages and following manipulation")
```

```{r prolactin-char}
characterization <- prolactin %>% filter(treatment %in% charlevels)   %>%  droplevels()

mod_prolactin <-  lm(data = characterization, `Prolactin ng/mL` ~ treatment + sex)
mod_prolactin

grid <- characterization %>%
  data_grid(treatment, sex) %>%
  add_predictions(mod_prolactin)

grid

df <- data.frame(
  x = c(1,2,3,4,5,6,7,8,9),
  y = c(120,120,120,120,120,120,120,120,120),
  sex = c("female", "female", "female", "female", "female", "female","female", "female", "female"),
  text = c("AB",  "A", "A" , "AB", "AB", "BCD", "D", "CD", "BC")
)

a <- ggplot(characterization, aes(x = treatment, y = `Prolactin ng/mL`, fill = sex)) +
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank()) +
  labs(subtitle = "Observed prolactin",
       y = "Concentration (ng/mL)", x = NULL) +
  #geom_text(data = df, aes(x,y, label = text), color = "darkgrey") +
      ylim(0,120)

b <- ggplot(grid, aes(x = treatment, y = pred, color = sex)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank()) +
  labs(subtitle = "Model prediction ",
       y = NULL, x = NULL) +
  geom_smooth(se = FALSE, aes(x = as.numeric(treatment)))  +
  ylim(0,120)

plot_grid(a,b)
```

```{r aov}
aov1 <- aov(data = characterization, `Prolactin ng/mL` ~ treatment + sex)
aov1
summary(aov1)
TukeyHSD(aov1, which = "treatment")
```

```{r prolactin-manip, fig.width=8} 
manipulation <- prolactin %>% 
  filter(treatment %in% maniplevels) %>%  droplevels() %>% 
  mutate(result = fct_collapse(treatment,
                                "early hatch" = c("early"),
                                "offspring removed" = c("M_inc3",  "M_inc9", "M_inc17", "M_n2"),
                               "late incubation" = c("prolong"),
                               "late hatch" = c("extend")))

e <- manipulation %>%
  ggplot( aes(x = treatment, y = `Prolactin ng/mL`)) +
    geom_boxplot(aes(fill = sex, alpha = 0.5)) + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom",
          legend.title = element_blank()) +
    ylim(0,125) +
    labs(subtitle = "Manipulation",
       y = NULL, x = NULL) + 
  scale_alpha(guide = 'none')

plot_grid(a,b,e, nrow = 1, rel_widths = c(1.1,.75,.8))

f <-  prolactin %>%
  filter(treatment %in% c("control", "bldg", "inc_d3", "M_inc3",
                          "inc_d9",  "M_inc9", "inc_d17","M_inc17",
                          "hatch", "M_n2")) %>%
ggplot(aes(x = treatment, y = `Prolactin ng/mL`, fill = Sex)) +
  geom_boxplot(aes(alpha = study)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  labs(subtitle = "Offspring removed",
       x = NULL) +
  facet_wrap(~sex) +
  annotate("rect", xmin = 2.5, xmax = 4.5, ymin = 0, ymax = 120, alpha = .2) +
    #annotate("rect", xmin = 4.5, xmax = 6.5, ymin = 0, ymax = 120, alpha = .3) +
    annotate("rect", xmin = 6.5, xmax = 8.5, ymin = 0, ymax = 120, alpha = .2) +
    #annotate("rect", xmin = 8.5, xmax = 10.5, ymin = 0, ymax = 120, alpha = .3) +
  ylim(0,120) +
  scale_alpha_manual(values = c(1,0.5))

g <-  prolactin %>%
  filter(treatment %in% c("early","prolong", "extend",
                          "inc_d9","inc_d17","hatch", "n5" )) %>% droplevels() %>%
ggplot(aes(x = treatment, y = `Prolactin ng/mL`, fill = Sex)) +
  geom_boxplot(aes(alpha = study)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  labs(subtitle = "Timing of transition altered",
       x = NULL, y = NULL) +
  facet_wrap(~sex) +
    annotate("rect", xmin = 4.5, xmax = 7.5, ymin = 0, ymax = 120, alpha = .3) +
  ylim(0,120) +
  scale_alpha_manual(values = c(1,0.5))

plot_grid(f,g, nrow=2, rel_widths = c(.625,.375))

```

```{r fortwitter, fig.height= 3, fig.width=5.5}
ggplot(grid, aes(x = treatment, y = pred, color = sex)) +
  geom_point() + 
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        legend.position = c(0.75,0.3),
        legend.title = element_blank()) +
  labs(y = "Gene expression", x = "Time") +
  geom_smooth(se = FALSE, aes(x = as.numeric(treatment)))  
```

```{r}
write.csv(prolactin, "../results/prolactin.csv", row.names = F)
```

### 2019-10-08 analysis/07_hormones.Rmd
now with hormones

@@
 
-```{r prolactin-all}
+```{r wranglesteroids}
+PETC <- read_excel("../results/parental_care_hormone_RIA_data_master.xlsx", sheet = 2)
+
+PETC <- PETC %>% select(stage, sex, hormone, plasma_conc)  %>%
+                filter(stage %in% combolevels)  %>%
+                droplevels()
+PETC$stage <- factor(PETC$stage)
+levels(PETC$stage)
+
+PETC <- PETC %>%
+    mutate(sex = fct_recode(sex,
+                            "female" = "f",
+                            "male" = "m"),
+           study = fct_collapse(stage,
+                                 characterization = charlevels,
+                                 manipulation = maniplevels))
+PETC$treatment <- factor(PETC$stage, levels = combolevels)
+
+```
+
+```{r prolactin}
 prolactin %>%
@@
   facet_wrap(~study, scales = "free_x") +
-  labs(title = "Ciruclating prolactin levels",
-       subtitle = "During characteristic parental care stages and following manipulation")
+  labs(y = "Prolactin (ng/mL)")
+```
+
+```{r testosterone}
+
+PETC %>% 
+  filter(hormone == "testosterone", sex == "male")  %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
+  geom_boxplot() + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.position = "bottom") +
+  facet_wrap(~study, scales = "free_x") +
+  labs(y = "Testosterone") +
+  scale_fill_manual(values = c("#00BFC4"))
+```
+
+```{r estradiol}
+PETC %>% 
+  filter(hormone == "estradiol", sex == "female")  %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
+  geom_boxplot() + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.position = "bottom") +
+  facet_wrap(~study, scales = "free_x") +
+  labs(y = "Estradiol (ng/mL)")
+
+```
+
+```{r progesterone}
+PETC %>% 
+  filter(hormone == "progesterone")  %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
+  geom_boxplot() + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.position = "bottom") +
+  facet_wrap(~study, scales = "free_x") +
+  labs(y = "Progesterone (ng/mL)") +
+  ylim(c(0,6))
+
+```
+
+```{r cort}
+
+PETC %>% 
+  filter(hormone == "cort")  %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
+  geom_boxplot() + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.position = "bottom") +
+  facet_wrap(~study, scales = "free_x") +
+  labs(y = "Corticosterone (ng/mL)") +
+  ylim(c(0,10))
+
 ```

### 2019-10-09 analysis/07_hormones.Rmd
new hormone figs

@@
         legend.title = element_blank()) +
-  labs(subtitle = "Observed prolactin",
-       y = "Concentration (ng/mL)", x = NULL) +
+  labs(subtitle = "Characterization", y = "Prolactin (ng/mL)", x = NULL) +
   #geom_text(data = df, aes(x,y, label = text), color = "darkgrey") +
       ylim(0,120) 
+a
 
@@
   ggplot( aes(x = treatment, y = `Prolactin ng/mL`)) +
-    geom_boxplot(aes(fill = sex, alpha = 0.5)) + 
+    geom_boxplot(aes(fill = sex)) + 
     theme(axis.text.x = element_text(angle = 45, hjust = 1),
@@
     labs(subtitle = "Manipulation",
-       y = NULL, x = NULL) + 
-  scale_alpha(guide = 'none')
+       y = "Prolactin (ng/mL)", x = NULL)
+e
 
-plot_grid(a,b,e, nrow = 1, rel_widths = c(1.1,.75,.8))
+plot_grid(a,e, nrow = 1)
 

### 2019-10-11 analysis/07_hormones.Rmd
new hormone figs

@@
 ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
-  geom_boxplot() + 
+  geom_boxplot() + geom_jitter() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
@@
 ```{r cort}
-
 PETC %>% 
@@
   ylim(c(0,10))
-
 ```
@@
 a <- ggplot(characterization, aes(x = treatment, y = `Prolactin ng/mL`, fill = sex)) +
-  geom_boxplot() + 
+  geom_boxplot() + geom_jitter() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
@@
   ggplot( aes(x = treatment, y = `Prolactin ng/mL`)) +
-    geom_boxplot(aes(fill = sex)) + 
+    geom_boxplot(aes(fill = sex)) + geom_jitter() +
     theme(axis.text.x = element_text(angle = 45, hjust = 1),
@@
 ggplot(aes(x = treatment, y = `Prolactin ng/mL`, fill = Sex)) +
-  geom_boxplot(aes(alpha = study)) + 
+  geom_boxplot(aes(alpha = study)) + geom_jitter() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),

### 2019-10-11 docs/ms1-charactization.Rmd
add ms templates

---
title: "The one characterizing the parental care stages"
output: word_document
author:
- Principal Investigators - Rebecca Calisi, Matt MacManes, John Wingfield
- Postdocs - Suzanne Austin, Rayna Harris, Jesse Krause, 
- Grad students - Victoria Farrar, April Booth, Andrew Lang
- Undergrads - Tanner Feustel, Candice Lee, Annie Bond?
- Others - Fred?, Rechelle Viernes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

- What genes differ in expression between timepoints?
- What stages are most different, stressful, responsive?
- How do genes in the HPG affect genes in other regions?
- How do male and females differ?
- How do tissues differ?
- What is the relationship between genes and hormones across timepoints?

## Methods

![](../figures/images/PigeonHPGaxis.png)

![](../figures/images/samplingtimepoints.png)

## Results

## Discussion

## Conclusion

## Acknowledgements

## Data availabilty

### 2019-10-11 docs/ms2-timingmanipulation.Rmd
add ms templates

---
title: "The one where we manipulate timing of hatch"
output: word_document
author:
- Principal Investigators - Rebecca Calisi, Matt MacManes, John Wingfield
- Postdocs - Suzanne Austin, Rayna Harris, Jesse Krause, 
- Grad students - Victoria Farrar, April Booth, Andrew Lang
- Undergrads - Tanner Feustel, Candice Lee, Annie Bond?
- Others - Fred?, Rechelle Viernes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

## Methods

## Results

## Discussion

## Conclusion

## Acknowledgements

## Data availabilty

### 2019-10-11 analysis/07_hormones.Rmd
now with hormone characterization data

--- 2019-10-09 analysis/07_hormones.Rmd
@@
 
-```{r wrangle}
+```{r clean}
+charlevels <- c("control", "bldg", "lay", "inc_d3", "inc_d9", "inc_d17", "hatch", "n5", "n9" )
+removelevels <- c("M_inc3",  "M_inc9", "M_inc17", "M_n2")
+maniplevels <- c("early", "prolong", "extend")
+combolevels <- c("control", "bldg", "lay", "inc_d3", "M_inc3",
+                  "inc_d9",  "M_inc9", "early",  "inc_d17","M_inc17",  "hatch",
+                 "prolong", "extend", "M_n2",
+                  "n5", "n9" )
+```
+
+```{r wrangle-prolactin}
 prolactin <- read_excel("../results/Pigeon prolactin concentrations juil 2018.xlsx", sheet = 1)
@@
 # create lists of factos for characterization and manipluation studies
-charlevels <- c("control", "bldg", "lay", "inc_d3", "inc_d9", "inc_d17", "hatch", "n5", "n9" )
-maniplevels <- c("M_inc3", "early", "M_inc9", "M_inc17", "M_n2", "prolong", "extend")
-combolevels <- c("control", "bldg", "lay", "inc_d3", "M_inc3",
-                  "inc_d9",  "M_inc9",  "inc_d17","M_inc17",  "hatch", "M_n2",
-                  "n5", "n9" , "early",  "prolong", "extend")
 
@@
                                  characterization = charlevels,
-                                 manipulation = maniplevels))
+                                 manipulation = maniplevels,
+                                removal = removelevels))
 
-prolactin$treatment <- factor(prolactin$treatment, levels = combolevels)
+colnames(prolactin)[colnames(prolactin)=="Prolactin ng/mL"] <- "plasma_conc"
+prolactin$hormone <- "prolactin"
+prolactin <- prolactin %>% select(study, treatment, sex, hormone, plasma_conc)   %>% drop_na()
+
+summary(prolactin)
 ```
 
-```{r wranglesteroids}
+```{r wrangle-steroids}
 PETC <- read_excel("../results/parental_care_hormone_RIA_data_master.xlsx", sheet = 2)
@@
                                  characterization = charlevels,
-                                 manipulation = maniplevels))
-PETC$treatment <- factor(PETC$stage, levels = combolevels)
+                                manipulation = maniplevels,
+                                removal = removelevels)) %>% 
+  drop_na()
+colnames(PETC)[colnames(PETC)=="stage"] <- "treatment"
+PETC <- PETC %>% select(study, treatment, sex, hormone, plasma_conc)
+summary(PETC)
+```
+
+```{r combine-clean-split}
+hormones <- rbind(prolactin, PETC)
+hormones$treatment <- factor(hormones$treatment, levels = combolevels)
+
+hormones$okay <- ifelse(hormones$hormone == "cort" & hormones$plasma_conc > 30, "bad",
+                    ifelse(hormones$hormone == "progesterone" & hormones$plasma_conc > 5, "bad", 
+                           ifelse(hormones$hormone == "prolactin" & hormones$plasma_conc > 150, "bad", 
+                            ifelse(hormones$hormone == "prolactin" & hormones$treatment == "control", "bad", 
+                        ifelse(hormones$hormone == "testosterone" & hormones$sex == "female", "bad",
+                               ifelse(hormones$hormone == "estradiol" & hormones$sex == "male", "bad", "okay"))))))
+hormones <- hormones %>% filter(okay == "okay") %>% droplevels()
+
+prl.char <- hormones %>% filter(hormone == "prolactin", study == "characterization")   %>%  droplevels()
+test.char <- hormones %>% filter(hormone == "testosterone", treatment %in% charlevels)   %>%  droplevels()
+est.char <- hormones %>% filter(hormone == "estradiol", treatment %in% charlevels)   %>%  droplevels()
+prog.char <- hormones %>% filter(hormone == "progesterone", treatment %in% charlevels)   %>%  droplevels()
+cort.char <- hormones %>% filter(hormone == "cort", treatment %in% charlevels)   %>%  droplevels()
 
@@
 
-```{r prolactin}
-prolactin %>%
-  select(treatment, `Prolactin ng/mL`, Sex, study) %>% drop_na() %>%
-ggplot(aes(x = treatment, y = `Prolactin ng/mL`, fill = Sex)) +
-  geom_boxplot() + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "bottom") +
-  facet_wrap(~study, scales = "free_x") +
-  labs(y = "Prolactin (ng/mL)")
-```
+```{r characterization-hormones}
 
-```{r testosterone}
+hormonecharplot <- function(myhormone, myylab){
 
-PETC %>% 
-  filter(hormone == "testosterone", sex == "male")  %>% 
+  mycolors <- c("female" = "#F8766D", "male" = "#00BFC4")
+
+  hormones %>% 
+    filter(study == "characterization",
+           hormone %in% c(myhormone))  %>% 
   ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
@@
     theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "bottom") +
-  facet_wrap(~study, scales = "free_x") +
-  labs(y = "Testosterone") +
-  scale_fill_manual(values = c("#00BFC4"))
-```
+          legend.position = "none") +
+    scale_fill_manual(values = mycolors) +
+    labs(y = myylab, x = NULL) 
+}
 
-```{r estradiol}
-PETC %>% 
-  filter(hormone == "estradiol", sex == "female")  %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
-  geom_boxplot() + 
+a <- hormonecharplot("estradiol", "E (ng/mL)")
+b <- hormonecharplot("testosterone", "T (ng/mL)")
+c <- hormonecharplot("cort", "CORT (ng/mL)")
+d <- hormonecharplot("progesterone", "PROG (ng/mL)")
+
+e <- hormonecharplot("prolactin", "PRL (ng/mL)")
+
+# model and plot prolactin data
+mod_prolactin <-  lm(data = prl.char,  plasma_conc ~ treatment + sex)
+grid <- prl.char %>%
+  data_grid(treatment, sex) %>%
+  add_predictions(mod_prolactin)
+
+f <- ggplot(grid, aes(x = treatment, y = pred, color = sex)) +
+  geom_point() + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "bottom") +
-  facet_wrap(~study, scales = "free_x") +
-  labs(y = "Estradiol (ng/mL)")
+        legend.position = "none",
+        legend.title = element_blank()) +
+  labs(y = "predicted PRL (ng/mL)", x = NULL) +
+  geom_smooth(se = F, aes(x = as.numeric(treatment)))
+
+plot_grid(e,f,  rel_widths = c(0.65,0.35))
+
+plot_grid(d + theme(axis.text.x=element_blank()),
+          a + theme(axis.text.x=element_blank()),
+          c, b, 
+          rel_widths = c(0.65,0.35),
+          rel_heights = c(0.45,0.55), ncol = 2)
 
@@
 
-```{r progesterone}
-PETC %>% 
-  filter(hormone == "progesterone")  %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
-  geom_boxplot() + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "bottom") +
-  facet_wrap(~study, scales = "free_x") +
-  labs(y = "Progesterone (ng/mL)") +
-  ylim(c(0,6))
-
-```
-
-```{r cort}
-
-PETC %>% 
-  filter(hormone == "cort")  %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
-  geom_boxplot() + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "bottom") +
-  facet_wrap(~study, scales = "free_x") +
-  labs(y = "Corticosterone (ng/mL)") +
-  ylim(c(0,10))
-
-```
-
-```{r prolactin-char}
-characterization <- prolactin %>% filter(treatment %in% charlevels)   %>%  droplevels()
-
-mod_prolactin <-  lm(data = characterization, `Prolactin ng/mL` ~ treatment + sex)
-mod_prolactin
-
-grid <- characterization %>%
-  data_grid(treatment, sex) %>%
-  add_predictions(mod_prolactin)
-
-grid
-
-df <- data.frame(
-  x = c(1,2,3,4,5,6,7,8,9),
-  y = c(120,120,120,120,120,120,120,120,120),
-  sex = c("female", "female", "female", "female", "female", "female","female", "female", "female"),
-  text = c("AB",  "A", "A" , "AB", "AB", "BCD", "D", "CD", "BC")
-)
-
-a <- ggplot(characterization, aes(x = treatment, y = `Prolactin ng/mL`, fill = sex)) +
-  geom_boxplot() + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "bottom",
-        legend.title = element_blank()) +
-  labs(subtitle = "Characterization", y = "Prolactin (ng/mL)", x = NULL) +
-  #geom_text(data = df, aes(x,y, label = text), color = "darkgrey") +
-      ylim(0,120) 
-a
-
-b <- ggplot(grid, aes(x = treatment, y = pred, color = sex)) +
-  geom_point() + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "bottom",
-        legend.title = element_blank()) +
-  labs(subtitle = "Model prediction ",
-       y = NULL, x = NULL) +
-  geom_smooth(se = FALSE, aes(x = as.numeric(treatment)))  +
-  ylim(0,120)
-
-plot_grid(a,b)
-```
-
 ```{r aov}
-aov1 <- aov(data = characterization, `Prolactin ng/mL` ~ treatment + sex)
+aov1 <- aov(data = prl.char, plasma_conc ~ treatment + sex)
 aov1
@@
 TukeyHSD(aov1, which = "treatment")
-```
 
-```{r prolactin-manip, fig.width=8} 
-manipulation <- prolactin %>% 
-  filter(treatment %in% maniplevels) %>%  droplevels() %>% 
-  mutate(result = fct_collapse(treatment,
-                                "early hatch" = c("early"),
-                                "offspring removed" = c("M_inc3",  "M_inc9", "M_inc17", "M_n2"),
-                               "late incubation" = c("prolong"),
-                               "late hatch" = c("extend")))
+aov2 <- aov(data = prog.char, plasma_conc ~ treatment + sex)
+aov2
+summary(aov2)
 
-e <- manipulation %>%
-  ggplot( aes(x = treatment, y = `Prolactin ng/mL`)) +
-    geom_boxplot(aes(fill = sex)) + 
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
-          legend.position = "bottom",
-          legend.title = element_blank()) +
-    ylim(0,125) +
-    labs(subtitle = "Manipulation",
-       y = "Prolactin (ng/mL)", x = NULL)
-e
+aov3 <- aov(data = cort.char, plasma_conc ~ treatment + sex)
+aov3
+summary(aov3)
+TukeyHSD(aov3, which = "treatment")
 
-plot_grid(a,e, nrow = 1)
+aov4 <- aov(data = est.char, plasma_conc ~ treatment )
+aov4
+summary(aov4)
+TukeyHSD(aov4, which = "treatment")
 
-f <-  prolactin %>%
-  filter(treatment %in% c("control", "bldg", "inc_d3", "M_inc3",
-                          "inc_d9",  "M_inc9", "inc_d17","M_inc17",
-                          "hatch", "M_n2")) %>%
-ggplot(aes(x = treatment, y = `Prolactin ng/mL`, fill = Sex)) +
-  geom_boxplot(aes(alpha = study)) + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "none") +
-  labs(subtitle = "Offspring removed",
-       x = NULL) +
-  facet_wrap(~sex) +
-  annotate("rect", xmin = 2.5, xmax = 4.5, ymin = 0, ymax = 120, alpha = .2) +
-    #annotate("rect", xmin = 4.5, xmax = 6.5, ymin = 0, ymax = 120, alpha = .3) +
-    annotate("rect", xmin = 6.5, xmax = 8.5, ymin = 0, ymax = 120, alpha = .2) +
-    #annotate("rect", xmin = 8.5, xmax = 10.5, ymin = 0, ymax = 120, alpha = .3) +
-  ylim(0,120) +
-  scale_alpha_manual(values = c(1,0.5))
-
-g <-  prolactin %>%
-  filter(treatment %in% c("early","prolong", "extend",
-                          "inc_d9","inc_d17","hatch", "n5" )) %>% droplevels() %>%
-ggplot(aes(x = treatment, y = `Prolactin ng/mL`, fill = Sex)) +
-  geom_boxplot(aes(alpha = study)) + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "none") +
-  labs(subtitle = "Timing of transition altered",
-       x = NULL, y = NULL) +
-  facet_wrap(~sex) +
-    annotate("rect", xmin = 4.5, xmax = 7.5, ymin = 0, ymax = 120, alpha = .3) +
-  ylim(0,120) +
-  scale_alpha_manual(values = c(1,0.5))
-
-plot_grid(f,g, nrow=2, rel_widths = c(.625,.375))
-
-```
-
-```{r fortwitter, fig.height= 3, fig.width=5.5}
-ggplot(grid, aes(x = treatment, y = pred, color = sex)) +
-  geom_point() + 
-  theme(axis.text.x=element_blank(),
-        axis.text.y=element_blank(),
-        legend.position = c(0.75,0.3),
-        legend.title = element_blank()) +
-  labs(y = "Gene expression", x = "Time") +
-  geom_smooth(se = FALSE, aes(x = as.numeric(treatment)))  
+aov5 <- aov(data = test.char, plasma_conc ~ treatment )
+aov5
+summary(aov5)
 ```
@@
 ```{r}
-write.csv(prolactin, "../results/prolactin.csv", row.names = F)
+write.csv(hormones, "../results/hormones.csv", row.names = F)
 ```

### 2019-10-11 analysis/07_hormones.Rmd
anova functions

@@
 ```{r aov}
-aov1 <- aov(data = prl.char, plasma_conc ~ treatment + sex)
-aov1
-summary(aov1)
-TukeyHSD(aov1, which = "treatment")
 
-aov2 <- aov(data = prog.char, plasma_conc ~ treatment + sex)
-aov2
-summary(aov2)
+aovSexTretment <- function(mydata, whichormone){
+  aov2 <- aov(data = mydata, plasma_conc ~ treatment + sex)
+  print(whichormone)
+  print(summary(aov2))
+  print(TukeyHSD(aov2, which = "treatment"))
+}
+aovSexTretment(prl.char, "PRL")
+aovSexTretment(cort.char, "CORT")
+aovSexTretment(prog.char, "PROG")
 
-aov3 <- aov(data = cort.char, plasma_conc ~ treatment + sex)
-aov3
-summary(aov3)
-TukeyHSD(aov3, which = "treatment")
+aovTretment  <- function(mydata, whichormone){
+  aov1 <- aov(data = mydata, plasma_conc ~ treatment )
+  print(whichormone)
+  print(summary(aov1))
+  print(TukeyHSD(aov1, which = "treatment"))
+}
 
-aov4 <- aov(data = est.char, plasma_conc ~ treatment )
-aov4
-summary(aov4)
-TukeyHSD(aov4, which = "treatment")
-
-aov5 <- aov(data = test.char, plasma_conc ~ treatment )
-aov5
-summary(aov5)
+aovTretment(est.char, "E")
+aovTretment(test.char, "T")
 ```

### 2019-10-11 analysis/07_hormones.Rmd
three sets of hormone figs

@@
                   "n5", "n9" )
+
+manipchar <- c("early",  "inc_d17", "hatch", "prolong", "extend", "n5")
+removechar <- c("inc_d3", "M_inc3",  "inc_d9", "M_inc9","inc_d17", "M_inc17","hatch", "M_n2")
 ```
@@
                                 characterization = charlevels,
-                                manipulation = maniplevels,
-                                removal = removelevels)) %>% 
+                                manipulation = maniplevels)) %>% 
   drop_na()
@@
 hormones <- hormones %>% filter(okay == "okay") %>% droplevels()
+summary(hormones)
 
-prl.char <- hormones %>% filter(hormone == "prolactin", study == "characterization")   %>%  droplevels()
+prl.char <- hormones %>% filter(hormone == "prolactin", treatment %in% charlevels)   %>%  droplevels()
 test.char <- hormones %>% filter(hormone == "testosterone", treatment %in% charlevels)   %>%  droplevels()
@@
 
-```{r characterization-hormones}
-
+```{r characterization}
 hormonecharplot <- function(myhormone, myylab){
@@
 
-```{r aov}
-
+```{r aov-char}
 aovSexTretment <- function(mydata, whichormone){
@@
   print(summary(aov2))
-  print(TukeyHSD(aov2, which = "treatment"))
+  #print(TukeyHSD(aov2, which = "treatment"))
 }
-aovSexTretment(prl.char, "PRL")
-aovSexTretment(cort.char, "CORT")
-aovSexTretment(prog.char, "PROG")
 
@@
   print(summary(aov1))
-  print(TukeyHSD(aov1, which = "treatment"))
+  #print(TukeyHSD(aov1, which = "treatment"))
 }
+
+aovSexTretment(prl.char, "PRL")
+aovSexTretment(cort.char, "CORT")
+aovSexTretment(prog.char, "PROG")
 
@@
 
+```{r manipulation}
+hormonemanipPRL <- function(myhormone, myylab){
+
+  mycolors <- c("female" = "#F8766D", "male" = "#00BFC4")
+
+  hormones %>% 
+    filter(treatment %in% manipchar,
+           hormone %in% c(myhormone))  %>% 
+  ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
+    geom_boxplot() + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none",
+          strip.text = element_blank()) +
+    scale_fill_manual(values = mycolors) +
+    labs(y = myylab, x = NULL) +
+    facet_wrap(~sex) +
+    annotate("rect", xmin = 0.5, xmax = 1.5, ymin = 0, ymax = 125, alpha = .2) +
+    annotate("rect", xmin = 3.5, xmax = 5.5, ymin = 0, ymax = 125, alpha = .2) 
+}
+
+hormonemanipPRL("prolactin", "PRL (ng/mL)")
+
+hormonemanipSteroids <- function(myhormone, myylab, myymax){
+
+  mycolors <- c("female" = "#F8766D", "male" = "#00BFC4")
+
+  hormones %>% 
+    filter(treatment %in% manipchar,
+           hormone %in% c(myhormone))  %>% 
+  ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
+    geom_boxplot() + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none",
+          strip.text = element_blank()) +
+    scale_fill_manual(values = mycolors) +
+    labs(y = myylab, x = NULL) +
+    facet_wrap(~sex) +
+    annotate("rect", xmin = 2.5, xmax = 4.5, ymin = 0, ymax = myymax, alpha = .2)
+}
+
+a <- hormonemanipSteroids("estradiol", "E (ng/mL)", 1)
+b <- hormonemanipSteroids("testosterone", "T (ng/mL)", 3.5)
+c <- hormonemanipSteroids("cort", "CORT (ng/mL)", 10)
+d <- hormonemanipSteroids("progesterone", "PROG (ng/mL)",2.5)
+
+plot_grid(d,a,c,b, rel_widths = c(0.65,0.35))
+```
+```{r removal}
+hormoneremoval <- function(myhormone, myylab){
+
+  mycolors <- c("female" = "#F8766D", "male" = "#00BFC4")
+
+  hormones %>% 
+    filter(treatment %in% removechar,
+           hormone %in% c(myhormone))  %>% 
+  ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
+    geom_boxplot() + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none",
+          strip.text = element_blank()) +
+    scale_fill_manual(values = mycolors) +
+    labs(y = myylab, x = NULL) +
+    facet_wrap(~sex) +
+    annotate("rect", xmin = 1.5, xmax = 2.5, ymin = 0, ymax = 125, alpha = .2) +
+    annotate("rect", xmin = 3.5, xmax = 4.5, ymin = 0, ymax = 125, alpha = .2) +
+    annotate("rect", xmin = 5.5, xmax = 6.5, ymin = 0, ymax = 125, alpha = .2) + 
+    annotate("rect", xmin = 7.5, xmax = 8.5, ymin = 0, ymax = 125, alpha = .2) 
+}
+
+hormoneremoval("prolactin", "PRL (ng/mL)")
+```
+
 ```{r}

### 2019-10-11 docs/ms3-offpsringremoval.Rmd
three sets of hormone figs

---
title: "The one where we remove their offspring"
output: word_document
author:
- Principal Investigators - Rebecca Calisi, Matt MacManes, John Wingfield
- Postdocs - Suzanne Austin, Rayna Harris, Jesse Krause, 
- Grad students - Victoria Farrar, April Booth, Andrew Lang
- Undergrads - Tanner Feustel, Candice Lee, Annie Bond?
- Others - Fred?, Rechelle Viernes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

## Methods

## Results

### Fig x: Prolactin drops when offspring removed around hatch

![](../figures/hormones/removal-1.png)

## Discussion

## Conclusion

## Acknowledgements

## Data availabilty

### 2019-10-11 analysis/07_hormones.Rmd
now with timeline

@@
                            ifelse(hormones$hormone == "prolactin" & hormones$plasma_conc > 150, "bad", 
-                            ifelse(hormones$hormone == "prolactin" & hormones$treatment == "control", "bad", 
                         ifelse(hormones$hormone == "testosterone" & hormones$sex == "female", "bad",
-                               ifelse(hormones$hormone == "estradiol" & hormones$sex == "male", "bad", "okay"))))))
+                               ifelse(hormones$hormone == "estradiol" & hormones$sex == "male", "bad", "okay")))))
 hormones <- hormones %>% filter(okay == "okay") %>% droplevels()

### 2019-10-11 docs/ms1-charactization.Rmd
now with hormone characterization data
three sets of hormone figs
now with timeline
more figs

@@
 
-![](../figures/images/PigeonHPGaxis.png)
+### Fig 1: graphical abstract of methods
 
-![](../figures/images/samplingtimepoints.png)
+![](../figures/images/timeline_timeline.png)
 
 ## Results
+
+## Fig 2: circulating prolactin peaks at hatch in M and F
+
+![](../figures/hormones/characterization-1.png)
+
+### Fig 2 supplement. CORT but not E,T, or P respond significantly to treatment
+
+![](../figures/hormones/characterization-2.png)
+
+## Figure 3: general rnaseq pattern, tissue > sex > stage
+
+_Needs to be replaced with a better PCA plot_
+
+![](../figures/limma/plotMDS-colored-1.png)
+
+## Figure 4: model showing ability to predict non-parental, egg care, and nestingling care
+
+![](../figures/characterization/DoveParentsRNAseq_model2.png)
+
+## Figure 5: PRL in prolatin mirrors PRL in bloo
+
+![](../figures/characterization/prolactin-1.png)
+
+## Figure 6: many genes mirror the PRL expression pattern
+
+![](../figures/sexes/wgcna2-1.png)
 

### 2019-10-11 analysis/07_hormones.Rmd
new colors

@@
 removechar <- c("inc_d3", "M_inc3",  "inc_d9", "M_inc9","inc_d17", "M_inc17","hatch", "M_n2")
+
+sexcolors <- c("female" = "#F8766D", "male" = "#00BFC4")
+charcolors <- c("control" = "#F8766D", "bldg" = "#D39200",
+                "lay" =  "#93AA00", "inc_d3" = "#00BA38" ,"inc_d9" = "#00C19F" , "inc_d17" =  "#00B9E3", 
+                "hatch" = "#619Cff" , "n5" =  "#DB72Fb", "n9" =  "#FF61C3")
+
 ```
@@
 
-  mycolors <- c("female" = "#F8766D", "male" = "#00BFC4")
+  mycolors <- charcolors
 
@@
            hormone %in% c(myhormone))  %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
+  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
     geom_boxplot() + 
@@
     scale_fill_manual(values = mycolors) +
-    labs(y = myylab, x = NULL) 
+    labs(y = myylab, x = NULL) + 
+    facet_wrap(~sex, nrow = 2, scale = "free_x")
 }
@@
 
-# model and plot prolactin data
-mod_prolactin <-  lm(data = prl.char,  plasma_conc ~ treatment + sex)
-grid <- prl.char %>%
-  data_grid(treatment, sex) %>%
-  add_predictions(mod_prolactin)
+cd <- plot_grid(e,c,d, nrow = 1)
+ab <- plot_grid(a,b, ncol = 1)
 
-f <- ggplot(grid, aes(x = treatment, y = pred, color = sex)) +
-  geom_point() + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "none",
-        legend.title = element_blank()) +
-  labs(y = "predicted PRL (ng/mL)", x = NULL) +
-  geom_smooth(se = F, aes(x = as.numeric(treatment)))
-
-plot_grid(e,f,  rel_widths = c(0.65,0.35))
-
-plot_grid(d + theme(axis.text.x=element_blank()),
-          a + theme(axis.text.x=element_blank()),
-          c, b, 
-          rel_widths = c(0.65,0.35),
-          rel_heights = c(0.45,0.55), ncol = 2)
-
+plot_grid(cd, ab, rel_widths = c(0.75, 0.25))
 ```
@@
 
-  mycolors <- c("female" = "#F8766D", "male" = "#00BFC4")
+  mycolors <- sexcolors
 

### 2019-10-11 analysis/07_hormones.Rmd
working on side by side

@@
                 "hatch" = "#619Cff" , "n5" =  "#DB72Fb", "n9" =  "#FF61C3")
+
+charsexlevels <- c("control female", "control male", "bldg female", "bldg male",  "lay female", "lay male",
+                "inc_d3 female", "inc_d3 male", "inc_d9 female", "inc_d9 male",
+                "inc_d17 female", "inc_d17 male", "hatch female", "hatch male", 
+                "n5 female", "n5 male", "n9 female" , "n9 male")
 
@@
 
+hormones <- hormones %>%  mutate(treatment.sex = paste(treatment, sex, sep = " ")) 
+hormones$treatment.sex <- factor(hormones$treatment.sex, levels = charsexlevels)
+
 hormones$okay <- ifelse(hormones$hormone == "cort" & hormones$plasma_conc > 30, "bad",
@@
   mycolors <- charcolors
+  sexcolors <- c("female" = "#969696", "male" = "#525252")
 
@@
            hormone %in% c(myhormone))  %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
     geom_boxplot() + 
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+    theme(axis.text.x = element_blank(),
           legend.position = "none") +
     scale_fill_manual(values = mycolors) +
-    labs(y = myylab, x = NULL) + 
-    facet_wrap(~sex, nrow = 2, scale = "free_x")
+    scale_color_manual(values = sexcolors) +
+    labs(y = myylab, x = NULL) 
 }
+
+hormonecharplot("prolactin", "PRL (ng/mL)")
 
@@
 b <- hormonecharplot("testosterone", "T (ng/mL)")
+
 c <- hormonecharplot("cort", "CORT (ng/mL)")
@@
 
-e <- hormonecharplot("prolactin", "PRL (ng/mL)")
+e <- hormonecharplot("prolactin", "PRL (ng/mL)") + theme(legend.position = "right", legend.direction = "horizontal")
 
-cd <- plot_grid(e,c,d, nrow = 1)
-ab <- plot_grid(a,b, ncol = 1)
+cd <- plot_grid(d,c, nrow = 2)
+ab <- plot_grid(a,b, nrow = 2)
 
-plot_grid(cd, ab, rel_widths = c(0.75, 0.25))
+abcd <- plot_grid(cd,ab, rel_widths = c(0.6,0.4))
+plot_grid(e,abcd, ncol = 1)
+
 ```

### 2019-10-11 analysis/07_hormones.Rmd
all four hormones

@@
 
-hormonecharplot("prolactin", "PRL (ng/mL)")
-
-a <- hormonecharplot("estradiol", "E (ng/mL)")
-b <- hormonecharplot("testosterone", "T (ng/mL)")
-
-c <- hormonecharplot("cort", "CORT (ng/mL)")
-d <- hormonecharplot("progesterone", "PROG (ng/mL)")
-
-e <- hormonecharplot("prolactin", "PRL (ng/mL)") + theme(legend.position = "right", legend.direction = "horizontal")
-
-cd <- plot_grid(d,c, nrow = 2)
-ab <- plot_grid(a,b, nrow = 2)
-
-abcd <- plot_grid(cd,ab, rel_widths = c(0.6,0.4))
-plot_grid(e,abcd, ncol = 1)
-
+a <- hormonecharplot("prolactin", "PRL (ng/mL)")
+b <- hormonecharplot("cort", "CORT (ng/mL)")
+c <- hormonecharplot("progesterone", "PROG (ng/mL)") 
+d1 <- hormonecharplot("estradiol", "E (ng/mL)")
+d2 <- hormonecharplot("testosterone", "T (ng/mL)")
+d <- plot_grid(d1,d2, nrow = 1)
+abcd <- plot_grid(a, b ,c, d)
+mylegend <- get_legend(c + theme(legend.position = "bottom"))
+plot_grid(abcd, mylegend, nrow = 2, rel_heights = c(0.85,0.15))
 ```

### 2019-10-11 analysis/07_hormones.Rmd
now with outlier color

@@
   ggplot(aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
-    geom_boxplot() + 
+    geom_boxplot(aes(outlier.colour = treatment)) +
     theme(axis.text.x =  element_blank(),

### 2019-10-11 analysis/07_hormones.Rmd
new hormone fig

@@
 d <- plot_grid(d1,d2, nrow = 1)
-abcd <- plot_grid(a, b ,c, d)
+abcd <- plot_grid(a, b ,c, d, labels = "AUTO")
 mylegend <- get_legend(c + theme(legend.position = "bottom"))

### 2019-10-11 docs/ms1-charactization.Rmd
new colors
new hormone fig

@@
 title: "The one characterizing the parental care stages"
-output: word_document
+output: md_document
 author:
@@
 
-## Fig 2: circulating prolactin peaks at hatch in M and F
+## Fig 2: hormones
 
 ![](../figures/hormones/characterization-1.png)
-
-### Fig 2 supplement. CORT but not E,T, or P respond significantly to treatment
-
-![](../figures/hormones/characterization-2.png)
 

### 2019-10-11 analysis/07_hormones.Rmd
ms

--- 2019-10-11 analysis/07_hormones.Rmd
@@
 d <- plot_grid(d1,d2, nrow = 1)
-abcd <- plot_grid(a, b ,c, d)
+abcd <- plot_grid(a, b ,c, d, labels = "AUTO")
 mylegend <- get_legend(c + theme(legend.position = "bottom"))
 plot_grid(abcd, mylegend, nrow = 2, rel_heights = c(0.85,0.15))
+
+plot_grid(get_legend(a + theme(legend.position = "bottom")))
 ```

### 2019-10-11 analysis/07_hormones.Rmd
better timeline, thicker boxplots

@@
 charcolors <- c("control" = "#F8766D", "bldg" = "#D39200",
-                "lay" =  "#93AA00", "inc_d3" = "#00BA38" ,"inc_d9" = "#00C19F" , "inc_d17" =  "#00B9E3", 
+                "lay" =  "#93AA00", "inc_d3" = "#00BA38" ,
+                "inc_d9" = "#00C19F" , "inc_d17" =  "#00B9E3", 
                 "hatch" = "#619Cff" , "n5" =  "#DB72Fb", "n9" =  "#FF61C3")
@@
 
-# rename some of the treatments levels, store in new columns
+# rename some of the timepoints levels, store in new columns
 prolactin <- prolactin %>%
@@
                             "male" = "m"),
-           treatment = fct_recode(Treatment,
+           timepoint = fct_recode(Treatment,
                             "hatch" = "Hatch",
@@
                             "n9" = "N9"),
-           study = fct_collapse(treatment,
+           study = fct_collapse(timepoint,
                                  characterization = charlevels,
@@
 prolactin$hormone <- "prolactin"
-prolactin <- prolactin %>% select(study, treatment, sex, hormone, plasma_conc)   %>% drop_na()
+prolactin <- prolactin %>% select(study, timepoint, sex, hormone, plasma_conc)   %>% drop_na()
 
@@
   drop_na()
-colnames(PETC)[colnames(PETC)=="stage"] <- "treatment"
-PETC <- PETC %>% select(study, treatment, sex, hormone, plasma_conc)
+colnames(PETC)[colnames(PETC)=="stage"] <- "timepoint"
+PETC <- PETC %>% select(study, timepoint, sex, hormone, plasma_conc)
 summary(PETC)
@@
 hormones <- rbind(prolactin, PETC)
-hormones$treatment <- factor(hormones$treatment, levels = combolevels)
+hormones$timepoint <- factor(hormones$timepoint, levels = combolevels)
 
-hormones <- hormones %>%  mutate(treatment.sex = paste(treatment, sex, sep = " ")) 
-hormones$treatment.sex <- factor(hormones$treatment.sex, levels = charsexlevels)
+hormones <- hormones %>%  mutate(timepoint.sex = paste(timepoint, sex, sep = " ")) 
+hormones$timepoint.sex <- factor(hormones$timepoint.sex, levels = charsexlevels)
 
@@
 
-prl.char <- hormones %>% filter(hormone == "prolactin", treatment %in% charlevels)   %>%  droplevels()
-test.char <- hormones %>% filter(hormone == "testosterone", treatment %in% charlevels)   %>%  droplevels()
-est.char <- hormones %>% filter(hormone == "estradiol", treatment %in% charlevels)   %>%  droplevels()
-prog.char <- hormones %>% filter(hormone == "progesterone", treatment %in% charlevels)   %>%  droplevels()
-cort.char <- hormones %>% filter(hormone == "cort", treatment %in% charlevels)   %>%  droplevels()
+prl.char <- hormones %>% filter(hormone == "prolactin", timepoint %in% charlevels)   %>%  droplevels()
+test.char <- hormones %>% filter(hormone == "testosterone", timepoint %in% charlevels)   %>%  droplevels()
+est.char <- hormones %>% filter(hormone == "estradiol", timepoint %in% charlevels)   %>%  droplevels()
+prog.char <- hormones %>% filter(hormone == "progesterone", timepoint %in% charlevels)   %>%  droplevels()
+cort.char <- hormones %>% filter(hormone == "cort", timepoint %in% charlevels)   %>%  droplevels()
 
@@
            hormone %in% c(myhormone))  %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
-    geom_boxplot(aes(outlier.colour = treatment)) +
+  ggplot(aes(x = timepoint, y = plasma_conc, fill = timepoint, color = sex)) +
+    geom_boxplot(aes(outlier.colour = timepoint), lwd=1.05) +
     theme(axis.text.x =  element_blank(),
@@
     scale_color_manual(values = sexcolors) +
-    labs(y = myylab, x = NULL) 
+    labs(y = myylab, x = NULL) +
+    guides(fill = guide_legend(order=1),
+         color = guide_legend(order=2))
 }
@@
 
-plot_grid(get_legend(a + theme(legend.position = "bottom")))
 ```
@@
 aovSexTretment <- function(mydata, whichormone){
-  aov2 <- aov(data = mydata, plasma_conc ~ treatment + sex)
+  aov2 <- aov(data = mydata, plasma_conc ~ timepoint + sex)
   print(whichormone)
   print(summary(aov2))
-  #print(TukeyHSD(aov2, which = "treatment"))
+  #print(TukeyHSD(aov2, which = "timepoint"))
 }
@@
 aovTretment  <- function(mydata, whichormone){
-  aov1 <- aov(data = mydata, plasma_conc ~ treatment )
+  aov1 <- aov(data = mydata, plasma_conc ~ timepoint )
   print(whichormone)
   print(summary(aov1))
-  #print(TukeyHSD(aov1, which = "treatment"))
+  #print(TukeyHSD(aov1, which = "timepoint"))
 }
@@
   hormones %>% 
-    filter(treatment %in% manipchar,
+    filter(timepoint %in% manipchar,
            hormone %in% c(myhormone))  %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
+  ggplot(aes(x = timepoint, y = plasma_conc, fill = sex)) +
     geom_boxplot() + 
@@
   hormones %>% 
-    filter(treatment %in% removechar,
+    filter(timepoint %in% removechar,
            hormone %in% c(myhormone))  %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
+  ggplot(aes(x = timepoint, y = plasma_conc, fill = sex)) +
     geom_boxplot() + 

### 2019-10-11 docs/ms1-charactization.Rmd
better timeline, thicker boxplots

@@
 
-- What genes differ in expression between timepoints?
-- What stages are most different, stressful, responsive?
-- How do genes in the HPG affect genes in other regions?
-- How do male and females differ?
-- How do tissues differ?
-- What is the relationship between genes and hormones across timepoints?
-
 ## Methods
@@
 
-## Fig 2: hormones
+## Fig 2: prolactin flucuates the most and peaks near hatch in males and females
 
@@
 
-## Figure 3: general rnaseq pattern, tissue > sex > stage
+## Figure 3: general rnaseq pattern, tissue > sex > parental stage
 
-_Needs to be replaced with a better PCA plot_
+_Needs to be replaced with a better PCA plot showing all samples as points_
 
@@
 
+_redo color scheme, fix titiles_
+
 ![](../figures/characterization/DoveParentsRNAseq_model2.png)
 
-## Figure 5: PRL in prolatin mirrors PRL in bloo
+## Figure 5: PRL gene expression in the pituitary in prolatin mirrors PRL in blood
 
@@
 
+- What genes differ in expression between timepoints?
+- What stages are most different, stressful, responsive?
+- How do genes in the HPG affect genes in other regions?
+- How do male and females differ?
+- How do tissues differ?
+- What is the relationship between genes and hormones across timepoints?
+
 ## Conclusion

### 2019-10-11 docs/ms1-charactization.Rmd
typos

--- 2019-10-11 docs/ms1-charactization.Rmd
@@
 
+## Methods
+
+### Fig 1: Graphical abstract of methods
+
+![](../figures/images/timeline_timeline.png)
+
+## Results
+
+## Fig 2: Prolactin fluctuates the most and peaks near hatch in males and females
+
+![](../figures/hormones/characterization-1.png)
+
+## Figure 3: General RNAseq pattern, tissue > sex > parental stage
+
+_Needs to be replaced with a better PCA plot showing all samples as points_
+
+![](../figures/limma/plotMDS-colored-1.png)
+
+## Figure 4: model showing ability to predict non-parental, egg care, and nestling care
+
+_redo color scheme, fix titles_
+
+![](../figures/characterization/DoveParentsRNAseq_model2.png)
+
+## Figure 5: PRL gene expression in the pituitary in prolactin mirrors PRL in blood
+
+![](../figures/characterization/prolactin-1.png)
+
+## Figure 6: Many genes mirror the PRL expression pattern
+
+![](../figures/sexes/wgcna2-1.png)
+
+## Discussion
+
 - What genes differ in expression between timepoints?
@@
 
-## Methods
-
-### Fig 1: graphical abstract of methods
-
-![](../figures/images/timeline_timeline.png)
-
-## Results
-
-## Fig 2: hormones
-
-![](../figures/hormones/characterization-1.png)
-
-## Figure 3: general rnaseq pattern, tissue > sex > stage
-
-_Needs to be replaced with a better PCA plot_
-
-![](../figures/limma/plotMDS-colored-1.png)
-
-## Figure 4: model showing ability to predict non-parental, egg care, and nestingling care
-
-![](../figures/characterization/DoveParentsRNAseq_model2.png)
-
-## Figure 5: PRL in prolatin mirrors PRL in bloo
-
-![](../figures/characterization/prolactin-1.png)
-
-## Figure 6: many genes mirror the PRL expression pattern
-
-![](../figures/sexes/wgcna2-1.png)
-
-## Discussion
-
 ## Conclusion
@@
 
-## Data availabilty
+## Data availability

### 2019-10-12 analysis/10_specificgenes.Rmd
specific geens

---
title: "forvictoria"
output: md_document
---

# selecting candidate genes counts from the hypothalamus

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# import "colData" which contains sample information and "countData" which contains read counts
c.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

# select only hypothalamus samples
hypsamples <- c.colData %>% 
  filter(tissue == "hypothalamus") %>%
  select(-study) %>%
  droplevels()
countstokeep <- as.character(hypsamples$V1)
hypcounts <- c.countData %>% 
  mutate(genes = row.names(c.countData)) %>% 
  select(genes, countstokeep)

# select genes to keep 
# CISH, SOCS1, SOCS2, SOCS2, SOCS4, SOCS5, SOCS6
candidates <- c("NP_989957.1", "NP_001131120.1", "NP_989871.1", "NP_001186037.1", "NP_001120786.1")
hypcounts <- hypcounts %>% 
  filter(genes %in% candidates)

# transform and combine into 1 dataframe
hypcounts <- as.data.frame(hypcounts)
row.names(hypcounts) <- hypcounts$genes
hypcounts$genes <- NULL
hypcounts <- as.data.frame(t(hypcounts))
hypcounts$V1 <- row.names(hypcounts)
hypcandidatecounts <- left_join(hypsamples, hypcounts)

# rename columsn
colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_989957.1"] <- "CISH"
colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_001131120.1"] <- "SOCS1"
colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_989871.1"] <- "SOCS2"
colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_001186037.1"] <- "SOCS4"
colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_001120786.1"] <- "SOCS5"
colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_001120784.1"] <- "SOCS6"

# set rownames
row.names(hypcandidatecounts) <- hypcandidatecounts$V1
hypcandidatecounts$V1 <- NULL

# preview and write
head(hypcandidatecounts)
write.csv(hypcandidatecounts, "../results/10_hypcandidatecounts.csv")
```

### 2019-10-12 analysis/10_specificgenes.Rmd
add figures

@@
 library(tidyverse)
-knitr::opts_chunk$set(echo = TRUE)
+library(corrplot)
+
+knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/specificgenes/')
 ```
 
-```{r}
+```{r PRL}
 # import "colData" which contains sample information and "countData" which contains read counts
@@
 c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
-geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>% select(Name, entrezid)
 
@@
   droplevels()
+colnames(hypsamples)[colnames(hypsamples)=="treatment"] <- "timepoint"
+
 countstokeep <- as.character(hypsamples$V1)
 hypcounts <- c.countData %>% 
-  mutate(genes = row.names(c.countData)) %>% 
-  select(genes, countstokeep)
+  mutate(entrezid = row.names(c.countData)) %>% 
+  select(entrezid, countstokeep)
+
+hypcounts <- left_join(geneinfo,hypcounts)
 
@@
 # CISH, SOCS1, SOCS2, SOCS2, SOCS4, SOCS5, SOCS6
-candidates <- c("NP_989957.1", "NP_001131120.1", "NP_989871.1", "NP_001186037.1", "NP_001120786.1")
-hypcounts <- hypcounts %>% 
-  filter(genes %in% candidates)
+# PRL PRLR BRAC1
+
+candidates <- c("CISH", "SOCS1", "SOCS2", "SOCS2", "SOCS4", "SOCS5", "SOCS6", "PRL", "PRLR")
+hypcounts <- hypcounts %>% filter(Name %in% candidates)
 
@@
 hypcounts <- as.data.frame(hypcounts)
-row.names(hypcounts) <- hypcounts$genes
-hypcounts$genes <- NULL
-hypcounts <- as.data.frame(t(hypcounts))
-hypcounts$V1 <- row.names(hypcounts)
-hypcandidatecounts <- left_join(hypsamples, hypcounts)
 
-# rename columsn
-colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_989957.1"] <- "CISH"
-colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_001131120.1"] <- "SOCS1"
-colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_989871.1"] <- "SOCS2"
-colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_001186037.1"] <- "SOCS4"
-colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_001120786.1"] <- "SOCS5"
-colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_001120784.1"] <- "SOCS6"
+row.names(hypcounts) <- hypcounts$Name
+hypcounts$Name <- NULL
+hypcounts$entrezid <- NULL
+hypcandidatecounts <- as.data.frame(t(hypcounts))
+hypcandidatecounts$V1 <- row.names(hypcandidatecounts)
+hypcandidatecounts <- left_join(hypsamples, hypcandidatecounts)
+head(hypcandidatecounts)
 
@@
 write.csv(hypcandidatecounts, "../results/10_hypcandidatecounts.csv")
+
+# correlation matrix
+forcorr <- hypcandidatecounts %>% 
+  mutate(gene = row.names(hypcandidatecounts)) %>%  select(candidates)
+M <- cor(forcorr)
+M <- as.data.frame(M)
+M$rownames <- row.names(M)
+# M <- M %>% filter(totalentr > 0.7 | totalentr < -0.7)
+row.names(M) <- M$rownames
+greatthan05 <- M$rownames
+M <- M %>% select(greatthan05)
+M <- as.matrix(M)
+corrplot.mixed(M, order = "hclust")
+
+ggplot(hypcandidatecounts, aes(x = PRL, y = SOCS4, color = timepoint)) + 
+  geom_point() + labs(subtitle = "PRL and SOCS5") + facet_wrap(~sex)
+
+ggplot(hypcandidatecounts, aes(x = SOCS5, y = SOCS6, color = timepoint)) + 
+  geom_point() + labs(subtitle = "SOCS5 and SOCS6") + facet_wrap(~sex)
 ```

### 2019-10-12 analysis/10_specificgenes.Rmd
remove head

@@
 ---
-title: "forvictoria"
+title: "candidate genes"
 output: md_document
@@
 hypcandidatecounts <- left_join(hypsamples, hypcandidatecounts)
-head(hypcandidatecounts)
 

### 2019-10-15 analysis/04_DESeq2_sexes.Rmd
now with stat5

@@
 LDA.gon2 <- LDAdata.hypothesis(vsd.gon, colDataGon)
-
 ```
@@
 ```{r LDA, fig.width = 9}
-
-# figures for both of the above
-
 # treatment
@@
                         "LD1, F = 7.1504779", "LD2, F = 2.4219585")
-
 b <- LDAplot.treatment(LDA.pit1 ,  "Pituitary model 1", "parental state ~ .  0.405 pred. acc.",
                         "LD1, F = 9.6144160", "LD2, F = 3.7773314")
-
 c <- LDAplot.treatment(LDA.gon1 , "Gonad model 1", "parental state ~ .  has 0.297 pred. acc.",
@@
 # hypothesis
-
 d <- LDAplot.hypothesis(LDA.hyp2 , "Hypothalamus model 2", "parental stage ~ .  0.432 pred. acc.",
                         "LD1, F = 7.386458", "LD2, F = 3.720794")
-
 e <- LDAplot.hypothesis(LDA.pit2 ,  "Pituitary model 2", "parental stage ~ .  0.473 pred. acc.",
                         "LD1, F = 7.846659", "LD2, F = 5.076861")
-
 f <- LDAplot.hypothesis(LDA.gon2 , "Gonad model 2", "parental stage ~ .  has 0.368 pred. acc.",
@@
 ```{r wgcna, eval = T, fig.width=6, fig.height=6}
+# pitutiary expression
+
 plotWGCNAcandidates(vsd.pit, c("NP_990797.2"), colDataPit, "PRL expression, Pituitary")
@@
 plotWGCNAcandidates(vsd.pit, c("NP_990136.1", "NP_001035179.1"), colDataPit, "TH & AR expression, Pituitary")
-plotWGCNAcandidates(vsd.pit, c("NP_990797.2", "NP_001090992.1","NP_990136.1", "NP_001035179.1", "NP_001026269.1"), colDataPit, "PRL, PRLR, VIP1R, POMC, TH & AR expression, Pituitary")
-
-plotWGCNAcandidates(vsd.hyp, c("NP_001170780.1", "NP_001090992.1", "NP_001014970.1"), colDataHyp, "VIP expression, Hypothalamus")
-
+plotWGCNAcandidates(vsd.pit, c("NP_990797.2", "NP_001090992.1","NP_990136.1", "NP_001035179.1", "NP_001026269.1"), 
+                    colDataPit, "PRL, PRLR, VIP1R, POMC, TH & AR expression, Pituitary")
+plotWGCNAcandidates(vsd.pit, c("XP_015155078.1", "XP_015146451.1", "NP_001025709.2", "NP_990327.2"), 
+                    colDataPit, "Stat and JAK proteins, Pituitary")
 plotWGCNAcandidates(vsd.pit,
@@
                     colDataPit, "9 of 102 genes in a WGCNA module containting Prolactin")
-
 plotWGCNAcandidates(vsd.pit,
@@
                     colDataPit, "Prolaction receptor: 9 of 147 genes in a WGCNA module containting PRLR")
-
 plotWGCNAcandidates(vsd.pit, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
@@
 
+# hyp
 plotWGCNAcandidates(vsd.hyp, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
@@
                                "NP_990797.2", "XP_015132722.1"), colDataHyp, "Hypothalamus: CISH, SOCS, PRL, PRLR")
+plotWGCNAcandidates(vsd.hyp, c("NP_001170780.1", "NP_001090992.1", "NP_001014970.1"), colDataHyp, "VIP expression, Hypothalamus")
 
+# gonad
 plotWGCNAcandidates(vsd.gon, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",

### 2019-10-15 docs/grant.Rmd
add docs

---
title: "Grant"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Collaborative Research: RUI: The Neural Basis of Becoming a Parent: From Genotype to Phenotype

[Award Abstract #1455957](https://www.nsf.gov/awardsearch/showAward?AWD_ID=1455957&HistoricalAwards=false)

Important changes in behavior occur with the transition to parenthood. For animals that exhibit parental care, raising offspring successfully involves transitioning from aggressive and sexual behaviors to more caring and nurturing ones. However, it is not clear how the brain brings about this transition. In a rapidly changing world, understanding how the environment affects the brain and how the brain affects the behavioral transition into parental care will shed light on how changes in environment can ultimately affect how all organisms reproduce. In an interdisciplinary collaboration between Drs. Calisi-Rodríguez and MacManes they will characterize changes in gene expression and specific protein levels in important tissues essential for reproduction, namely the hypothalamus, pituitary, and gonads, in the rock dove (Columba livia) during the transition into parental care behaviors. Like most birds, rock doves incubate eggs, but are also unique in that they "lactate" to feed their young, showing stereotypical avian and mammalian parental behaviors. This makes doves a potentially powerful theoretical bridge to understand the neurobiology of transitions into parenting. Drs. Calisi-Rodríguez and MacManes will generate a large genomic dataset and identify critical genes involved in important stages of parental care. The results will have broad potential applications including the management of captive breeding for conservation or agricultural purposes, and to inform the study of neural deficiencies in parent-offspring bonding. In addition, the researchers will develop and host an annual summer program, "Summer City Science", to expose high school students from low-income neighborhoods and underrepresented backgrounds who excel in the sciences to a range of scientific fields. This will provide students the opportunity to work closely and personally with faculty, and their research and labs. Students in both the Calisi-Rodríguez and MacManes labs will help to serve as mentors and instructors for Summer City Science to increase the scope of the training of all parties involved. The overall wider objective of this program is to broaden participation in science, creating a more efficient and diverse workforce.

Previous research has implicated specific hormones that play a role in the maintenance of parental care behavior in vertebrates, including oxytocin and vasopressin (avian homologs: mesotocin and vasotocin), vasoactive intestinal peptide, and prolactin. However, very little is known about when and how the brain transitions into parental care behavior in any vertebrate, and this knowledge is fundamentally important to understand the mechanisms that mediate parental care. The Principal Investigator, during her tenure as an NSF Postdoctoral Fellow, discovered that the hypothalamic neurohormone, gonadotropin inhibitory hormone (GnIH), caused significant changes in abundance of GnIH containing cells in both birds and mammals during the transition to parental care. Here, the researchers propose to expand this knowledge using classic offspring removal and replacement manipulations to uncover the role of GnIH during the transition to parental care behavior. In addition to this targeted approach, they propose a powerful untargeted approach to explore what other changes in the brain may be occurring during this transition. They will use high-throughput sequencing technology and immunohistochemistry to uncover all genetic and specific proteomic changes in the brain, along with their relationship to GnIH, which occur during the transition to parental care. This project will build on previous work, and the data gathered will yield the most complete characterization and understanding of changes in neural transcription and translation reported to date in any vertebrate during the transition to parental care. All data and computer code will be made publicly available in established repositories and shared with the scientific community.

### 2019-10-16 analysis/11_pcalda.Rmd
new plca plots

---
title: "PCA"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggfortify)
library(factoextra)
library(cowplot)
knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/pca/')

```

## pca analysis

```{r pca}
pseudocounts <- read_csv("../results/01_pseudo.counts.csv")

colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
colData$treatment <- factor(colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", 
                                "hatch", "n5", "n9"))

geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

subsetcounts <- function(colData, countData){

  # save counts that match colData
  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols) 
  countData <- countData %>% 
    dplyr::select(X1,savecols) %>% 
    dplyr::rename(entrezid = X1)

  countData <- left_join(geneinfo, countData)
  countData <- countData %>% distinct(Name, .keep_all = TRUE) %>%
    select(-row.names)  %>% select(-geneid) %>% select(-entrezid)

  row.names(countData) <- countData$Name
  countData$Name <- NULL
  countData <- as.data.frame(t(countData))

  return(countData)
}

countData <- subsetcounts(colData, pseudocounts)

# confirm
row.names(countData) == row.names(colData)

# make pca
mypca <- prcomp(countData)

# percent conribution
eig.val <- get_eigenvalue(mypca)
head(eig.val)

mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
                  ID = row.names(countData))
head(mypcadf)
mypcadf$V1 <- row.names(mypcadf)
mypcadf <- left_join(colData, mypcadf)
head(mypcadf)

a <- ggplot(mypcadf, aes(x = PC1, y = PC2, color = colData$treatment)) +
  geom_point() + labs(x = "PC1: 36%", y = "PC2: 32%")
b <- ggplot(mypcadf, aes(x = PC3, y = PC4, color = colData$treatment)) +
  geom_point() + theme(legend.position = "bottom", legend.title = element_blank()) + labs(x = "PC3: 25%", y = "PC4: 3%")

mylegend <- get_legend(b)

ab <- plot_grid(a + theme(legend.position = "none") ,
                b + theme(legend.position = "none"))

plot_grid(ab,mylegend, ncol = 1, rel_heights = c(0.9,0.1))

fviz_pca_var(res.pca, col.var="contrib") +
scale_color_gradient2(low="white", mid="blue", 
      high="red", midpoint=50) + theme_minimal()

```

### 2019-10-16 analysis/11_pcalda.Rmd
new pca plots

@@
                   ID = row.names(countData))
-head(mypcadf)
+
 mypcadf$V1 <- row.names(mypcadf)
 mypcadf <- left_join(colData, mypcadf)
-head(mypcadf)
 
 a <- ggplot(mypcadf, aes(x = PC1, y = PC2, color = colData$treatment)) +
-  geom_point() + labs(x = "PC1: 36%", y = "PC2: 32%")
-b <- ggplot(mypcadf, aes(x = PC3, y = PC4, color = colData$treatment)) +
-  geom_point() + theme(legend.position = "bottom", legend.title = element_blank()) + labs(x = "PC3: 25%", y = "PC4: 3%")
+  geom_point() + labs(x = "PC1: 35.9%", y = "PC2: 32.5%") + theme_minimal(base_size = 8)  +
+  theme(legend.title = element_blank(), legend.position = "left")
 
-mylegend <- get_legend(b)
+b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank")
 
-ab <- plot_grid(a + theme(legend.position = "none") ,
-                b + theme(legend.position = "none"))
-
-plot_grid(ab,mylegend, ncol = 1, rel_heights = c(0.9,0.1))
-
-fviz_pca_var(res.pca, col.var="contrib") +
-scale_color_gradient2(low="white", mid="blue", 
-      high="red", midpoint=50) + theme_minimal()
+ab <- plot_grid(a,b + theme(axis.text = element_blank()) + ggtitle(NULL), rel_widths = c(0.6,0.4))
+ab
 

### 2019-10-16 docs/ms1-charactization.Rmd
typos
new pca plots

@@
 
-## Fig 2: Prolactin fluctuates the most and peaks near hatch in males and females
+## Fig 2: Of the hormones measured, prolactin fluctuates the most and peaks near hatch in males and females
 
@@
 
-## Figure 3: General RNAseq pattern, tissue > sex > parental stage
+## Figure 3: PCA
 
@@
 
-![](../figures/limma/plotMDS-colored-1.png)
+![](../figures/pca/pca-1.png)
 
-## Figure 4: model showing ability to predict non-parental, egg care, and nestling care
+## Figure 4: LDA
 
-_redo color scheme, fix titles_
+![](../figures/sexes/LDA-4.png)
 
@@
 ## Data availability
+
+https://osf.io/g3kc5/

### 2019-10-18 analysis/04_DESeq2_sexes.Rmd
new pca new ms

@@
 ```{r setup, include = FALSE}
-
 library(tidyverse)
@@
 
+# genes from https://link.springer.com/article/10.1007%2Fs12079-012-0168-0
+# A pathway map of prolactin signaling
+# PL, RAC1, IRS2 not found
+plotWGCNAcandidates(vsd.pit, c("NP_990797.2", "XP_015132722.1", "NP_989690.1",
+                               "NP_001025709.2", "XP_015144878.1", "NP_989481.1",
+                               "NP_001186467.1", "NP_990386.1", "XP_015132490.1",
+                               "XP_417614.3", "XP_015150421.1", "NP_001155844.1"), 
+                    colDataPit, "Pit: Genes involved in prolactin signaling")
+
 # hyp
@@
 
+plotWGCNAcandidates(vsd.hyp, c("NP_990797.2", "XP_015132722.1", "NP_989690.1",
+                               "NP_001025709.2", "XP_015144878.1", "NP_989481.1",
+                               "NP_001186467.1", "NP_990386.1", "XP_015132490.1",
+                               "XP_417614.3", "XP_015150421.1", "NP_001155844.1"), 
+                    colDataHyp, "Hyp: Genes involved in prolactin signaling")
+
 # gonad
@@
                                "NP_990797.2", "XP_015132722.1"), colDataGon, "Gonad: CISH, SOCS, PRL, PRLR")
+
+plotWGCNAcandidates(vsd.gon, c("NP_990797.2", "XP_015132722.1", "NP_989690.1",
+                               "NP_001025709.2", "XP_015144878.1", "NP_989481.1",
+                               "NP_001186467.1", "NP_990386.1", "XP_015132490.1",
+                               "XP_417614.3", "XP_015150421.1", "NP_001155844.1"), 
+                    colDataGon, "Gonad: Genes involved in prolactin signaling")
+
 ```

### 2019-10-18 analysis/11_pcalda.Rmd
new pca new ms

@@
 pseudocounts <- read_csv("../results/01_pseudo.counts.csv")
+head(pseudocounts[1:3])
 
@@
                                 "hatch", "n5", "n9"))
+head(colData[1:3])
 
 geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+head(geneinfo)
 
@@
 countData <- subsetcounts(colData, pseudocounts)
+head(countData[1:3])
 
@@
                   ID = row.names(countData))
-
 mypcadf$V1 <- row.names(mypcadf)
 mypcadf <- left_join(colData, mypcadf)
+mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
+head(mypcadf)
 
@@
   theme(legend.title = element_blank(), legend.position = "left")
+a
 
-b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank")
+b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
+                   repel = TRUE ,
+                  select.var= list(contrib = 10))
+b
 
@@
 ab
-
 ```

### 2019-10-18 analysis/11_pcalda.Rmd
now wit tnse and mds

@@
 library(cowplot)
+library(Rtsne)
+library(viridis)
 knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/pca/')
@@
 
-## pca analysis
+## data wrangle
 
-```{r pca}
+```{r data}
 pseudocounts <- read_csv("../results/01_pseudo.counts.csv")
@@
                                 "hatch", "n5", "n9"))
+colData$tissue <- factor(colData$tissue, levels = c("hypothalamus", "pituitary", "gonad"))
+colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
+
 head(colData[1:3])
@@
 row.names(countData) == row.names(colData)
+```
 
+## pca
+
+```{r pca}
 # make pca
@@
 ```
+
+# MDS
+
+```{r MDS}
+euclidist <- dist(countData) # euclidean distances between the rows
+fit <- cmdscale(euclidist,eig=TRUE, k=2) # k is the number of dim
+x <- fit$points[,1]
+y <- fit$points[,2]
+
+plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2",
+  main="Metric MDS", type="n")
+text(x, y, labels = row.names(countData), cex=.7)
+```
+
+# tSNE
+
+```{r tSNE}
+tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=5, theta=0.5, dims=2)
+tsne_df = as.data.frame(tsne_model$Y) 
+plot(tsne_df$V1, tsne_df$V2)
+
+# prep for adding columns
+colData$V1 <- NULL
+tsne_df_cols <- cbind(colData, tsne_df)
+
+c  <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
+  geom_point(size = 2) +
+  theme_minimal(base_size = 6) +
+  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
+  theme(legend.position = "top")
+
+d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = treatment)) +
+  geom_point(size = 2) +
+  theme_minimal(base_size = 6) +
+  theme(legend.position = "top")
+
+plot_grid(c,d)
+```

### 2019-10-18 analysis/11_pcalda.Rmd
new fig

@@
 row.names(countData) == row.names(colData)
+
+euclidist <- dist(countData) # euclidean distances between the rows
 ```
@@
 
-a <- ggplot(mypcadf, aes(x = PC1, y = PC2, color = colData$treatment)) +
-  geom_point() + labs(x = "PC1: 35.9%", y = "PC2: 32.5%") + theme_minimal(base_size = 8)  +
+a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
+                         color = colData$treatment, 
+                         shape = colData$tissue)) +
+  geom_point() + labs(x = "PC1: 35.9%", y = "PC2: 32.5%") + theme_minimal(base_size = 6)  +
   theme(legend.title = element_blank(), legend.position = "left")
@@
                    repel = TRUE ,
-                  select.var= list(contrib = 10))
+                  select.var= list(contrib = 6))
 b
@@
 ab
-```
-
-# MDS
-
-```{r MDS}
-euclidist <- dist(countData) # euclidean distances between the rows
-fit <- cmdscale(euclidist,eig=TRUE, k=2) # k is the number of dim
-x <- fit$points[,1]
-y <- fit$points[,2]
-
-plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2",
-  main="Metric MDS", type="n")
-text(x, y, labels = row.names(countData), cex=.7)
 ```
@@
   scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
-  theme(legend.position = "top")
+  theme(legend.position = "top") +
+  labs(x = "tSNE 1", y = "tSNE 2")
 
@@
   theme_minimal(base_size = 6) +
-  theme(legend.position = "top")
+  theme(legend.position = "top") +
+  labs(x = "tSNE 1", y = "tSNE 2")
 
-plot_grid(c,d)
+cd <- plot_grid(c,d)
+cd
 ```
+
+# combined pca tsne
+
+```{r pca.tSNE, fig.width=6, fig.height=6}
+plot_grid(ab,cd, ncol = 1)
+
+plot_grid(a + theme(legend.position = "none"),
+          b + theme(axis.text = element_blank()) + ggtitle(NULL), 
+          c + theme(legend.position = "none"),
+          d + theme(legend.position = "none"))
+```

### 2019-10-18 analysis/11_pcalda.Rmd
updated tsne and pca plots

@@
 knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/pca/')
-
 ```
@@
                          color = colData$treatment, 
-                         shape = colData$tissue)) +
-  geom_point() + labs(x = "PC1: 35.9%", y = "PC2: 32.5%") + theme_minimal(base_size = 6)  +
-  theme(legend.title = element_blank(), legend.position = "left")
+                         shape = colData$tissue,
+                         alpha = colData$sex)) +
+  geom_point(size = 2)  + theme_minimal(base_size = 6)  +
+  theme(legend.title = element_blank(), legend.position = "left") +
+  scale_alpha_manual(values = c(0.75,1))
 a
@@
 # prep for adding columns
-colData$V1 <- NULL
-tsne_df_cols <- cbind(colData, tsne_df)
+colData2 <- colData 
+colData2$V1 <- NULL
+tsne_df_cols <- cbind(colData2, tsne_df)
 
@@
 
-d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = treatment)) +
+d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = tissue, color = treatment, alpha = sex)) +
   geom_point(size = 2) +
@@
   theme(legend.position = "top") +
-  labs(x = "tSNE 1", y = "tSNE 2")
+  labs(x = "tSNE 1", y = "tSNE 2") +
+  scale_alpha_manual(values = c(0.75,1))
 
@@
           d + theme(legend.position = "none"))
+
+mylegend <- get_legend(d + theme(legend.position = "bottom"))
+
+ad <- plot_grid(a + theme(legend.position = "none"),d + theme(legend.position = "none"))
+
+plot_grid(ad, mylegend, nrow = 2, rel_heights = c(1,0.1))
 ```

### 2019-10-18 analysis/04_DESeq2_sexes.Rmd
working on  plots of genes with top pca loadings

@@
 
+```{r topPCAloadings}
+# COX1, COX3, CHGA, PRL, ATP2B4, LOC107055658
+topPCAloadings <- c("NP_006917.1", "NP_006921.2","XP_421330.1", "NP_990797.2", "XP_015154447.1", "XP_015156300.1")
+
+plotWGCNAcandidates(vsd.hyp, topPCAloadings, colDataHyp, "Hypothalamus")
+plotWGCNAcandidates(vsd.pit, topPCAloadings, colDataPit, "Pituitary")
+plotWGCNAcandidates(vsd.gon, topPCAloadings, colDataGon, "Gonads")
+
+```
+
 ```{r ctsGE}

### 2019-10-18 docs/ms1-charactization.Rmd
updated ms

@@
 author:
-- Principal Investigators - Rebecca Calisi, Matt MacManes, John Wingfield
-- Postdocs - Suzanne Austin, Rayna Harris, Jesse Krause, 
+- PIs - Rebecca Calisi, Matt MacManes
+- Postdocs - Suzanne Austin, Rayna Harris, Jesse Krause
 - Grad students - Victoria Farrar, April Booth, Andrew Lang
-- Undergrads - Tanner Feustel, Candice Lee, Annie Bond?
-- Others - Fred?, Rechelle Viernes
+- Staff - Rechelle Viernes
 ---
@@
 
-## Figure 3: PCA
+## Figure 3: PCA + tSNE
 
-_Needs to be replaced with a better PCA plot showing all samples as points_
+![](../figures/pca/pca-3.png)
 
-![](../figures/pca/pca-1.png)
+![](../figures/pca/tSNE-2.png)
 

### 2019-10-20 analysis/04_DESeq2_sexes.Rmd
now with vsds

@@
 head(c.colData)
+
+# prep col data for subsequent analyses
+colDataHyp <- subsetcolData2(c.colData, c("female_hypothalamus", "male_hypothalamus"))
+colDataPit <- subsetcolData2(c.colData, c("female_pituitary", "male_pituitary"))
+colDataGon <- subsetcolData2(c.colData, c("female_gonad", "male_gonad"))
 ```
@@
 ```{r vsd}
+# create variance stbilized DESeq transformation
 vsd.hyp <- vst(dds.hypothalamus, blind=FALSE) 
@@
 
+# save as a dataframe for downstream analysis
 vsd.hyp.df <- vsd.dataframe(vsd.hyp) 
@@
 vsd.gon.df <- vsd.dataframe(vsd.gon) 
+```
+
+```{r viewwritevsd}
+savevsdfiles(vsd.hyp.df, colDataHyp, "hyp")
+savevsdfiles(vsd.pit.df, colDataPit, "pit")
+savevsdfiles(vsd.gon.df, colDataGon, "gon")
 ```
@@
 ```{r LDAdfs}
-# prep col data for LDA
-colDataHyp <- subsetcolData2(c.colData, c("female_hypothalamus", "male_hypothalamus"))
-colDataPit <- subsetcolData2(c.colData, c("female_pituitary", "male_pituitary"))
-colDataGon <- subsetcolData2(c.colData, c("female_gonad", "male_gonad"))
-
 # LDA for treatment (9 groups)

### 2019-10-20 R/functions.R

@@ -431,15 +431,29 @@ plotPC12 <- function(pcadata, mysubtitle){
 }  
 
 
-######### vsd.dataframe ######### 
+######### vsd.dataframe and corresponding col data ######### 
 
 vsd.dataframe <- function(vsd){
   
   # extract varience stabilized and set rownames
   vsd.df <- assay(vsd)
   vsd.df <- as.data.frame(vsd.df)
-  vsd.df$entrezid <- row.names(vsd.df)
+  return(vsd.df)
 }  
+
+
+
+savevsdfiles <- function(myvsddf, mycolData, mytissue){
+  myvsddf <- myvsddf
+  mycolData <- mycolData
+  
+  vsdfilename <- paste0("../results/04_vsd_", mytissue, ".csv", sep = "" )
+  colDatafilename <- paste0("../results/04_colData_", mytissue, ".csv", sep = "" )
+  
+  write.csv(myvsddf, vsdfilename, row.names = T)
+  write.csv(mycolData, colDatafilename, row.names = F)
+}
+
 
 ######### plotcandidates ######### 
 
@@ -627,7 +641,7 @@ plotWGCNAcandidatesManip <- function(vsd, mygenelist, colData, mysubtitle){
 # makes a candidate heat map!!
 makepheatmap <- function(vsd.df, colData, mysubtitle){
 
-  candidates <- full_join(geneinfo, DEGs, by = "entrezid")
+  candidates <- full_join(geneinfo, vsd.df, by = "entrezid")
   
   candidates <- candidates %>%
     filter(Name %in% c("AVP", "AVPR1A", "AVPR1B", "AVPR2",

### 2019-10-20 analysis/10_specificgenes.Rmd
new hormone and candidate gene plots

--- 2019-10-12 analysis/10_specificgenes.Rmd
@@
 
+```{r setup, include=FALSE}
+library(tidyverse)
+library(cowplot)
+
+source("../R/functions.R")  # load custom functions 
+source("../R/themes.R")  # load custom themes and color palletes
+
+knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/specificgenes/')
+```
+
+# analyses downstream of calculating tissue specific variance in gene expression
+
+```{r getvsdandcolData}
+vsd.hyp <- readvsd("../results/04_vsd_hyp.csv")
+colData.hyp <- readcolData("../results/04_colData_hyp.csv")
+
+head(vsd.hyp[1:2])
+head(colData.hyp[1:2,])
+rownames(colData.hyp) == colnames(vsd.hyp)
+
+vsd.pit <- readvsd("../results/04_vsd_pit.csv")
+colData.pit <- readcolData("../results/04_colData_pit.csv")
+
+vsd.gon <- readvsd("../results/04_vsd_gon.csv")
+colData.gon <- readcolData("../results/04_colData_gon.csv")
+
+```
+
 # selecting candidate genes counts from the hypothalamus
 
-```{r setup, include=FALSE}
-library(tidyverse)
-knitr::opts_chunk$set(echo = TRUE)
+```{r PRL}
+geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>%  dplyr::select(Name, geneid, entrezid) %>% arrange(Name)
+head(geneinfo)
 ```
 
-```{r}
-# import "colData" which contains sample information and "countData" which contains read counts
-c.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
-c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
-geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+```{r PRLboxplots, fig.width =5.5, fig.height= 2}
+candidategenes <- c("CISH", "SOCS1", "SOCS2", "SOCS2", "SOCS4", "SOCS5", "SOCS6", 
+                    "PRL", "PRLR", "STAT5B", "JAK1")
+candidates.hyp <- selectcandidatevsds(candidategenes, vsd.hyp, colData.hyp)
+candidates.pit <- selectcandidatevsds(candidategenes, vsd.pit, colData.pit)
+candidates.gon <- selectcandidatevsds(candidategenes, vsd.gon, colData.gon)
 
-# select only hypothalamus samples
-hypsamples <- c.colData %>% 
-  filter(tissue == "hypothalamus") %>%
-  select(-study) %>%
-  droplevels()
-countstokeep <- as.character(hypsamples$V1)
-hypcounts <- c.countData %>% 
-  mutate(genes = row.names(c.countData)) %>% 
-  select(genes, countstokeep)
+plotcanddateexpression <- function(candidateexpression,  mysubtitle, whichgene, myylab){
 
-# select genes to keep 
-# CISH, SOCS1, SOCS2, SOCS2, SOCS4, SOCS5, SOCS6
-candidates <- c("NP_989957.1", "NP_001131120.1", "NP_989871.1", "NP_001186037.1", "NP_001120786.1")
-hypcounts <- hypcounts %>% 
-  filter(genes %in% candidates)
+  ggplot(candidateexpression, aes(x = as.numeric(treatment), y = whichgene)) + 
+    geom_boxplot(aes(fill = treatment, alpha = sex)) + 
+    scale_alpha_manual(values = c(0.6,0.9)) +
+     theme_B3() +
+    theme(legend.position = "none") +
+    theme(axis.title.y=element_text(face="italic"),
+          axis.title.x = element_blank(),
+          axis.text.x = element_blank()) +
+    geom_smooth(aes(colour = sex)) +
+    scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+    labs(subtitle = mysubtitle, y = myylab)
 
-# transform and combine into 1 dataframe
-hypcounts <- as.data.frame(hypcounts)
-row.names(hypcounts) <- hypcounts$genes
-hypcounts$genes <- NULL
-hypcounts <- as.data.frame(t(hypcounts))
-hypcounts$V1 <- row.names(hypcounts)
-hypcandidatecounts <- left_join(hypsamples, hypcounts)
+}
 
-# rename columsn
-colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_989957.1"] <- "CISH"
-colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_001131120.1"] <- "SOCS1"
-colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_989871.1"] <- "SOCS2"
-colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_001186037.1"] <- "SOCS4"
-colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_001120786.1"] <- "SOCS5"
-colnames(hypcandidatecounts)[colnames(hypcandidatecounts)=="NP_001120784.1"] <- "SOCS6"
+a <- plotcanddateexpression(candidates.hyp,  "hypothalamus", candidates.hyp$PRL, "PRL")
+b <- plotcanddateexpression(candidates.pit, "pituitary", candidates.pit$PRL, "PRL")
+c <- plotcanddateexpression(candidates.gon,  "gonad", candidates.gon$PRL, "PRL")
 
-# set rownames
-row.names(hypcandidatecounts) <- hypcandidatecounts$V1
-hypcandidatecounts$V1 <- NULL
+plot_grid(a,b + labs(y = NULL),c + labs(y = NULL), nrow = 1)
 
-# preview and write
-head(hypcandidatecounts)
-write.csv(hypcandidatecounts, "../results/10_hypcandidatecounts.csv")
+e <- plotcanddateexpression(candidates.hyp, NULL, candidates.hyp$PRLR, "PRLR")
+f <- plotcanddateexpression(candidates.pit, NULL, candidates.pit$PRLR, "PRLR")
+g <- plotcanddateexpression(candidates.gon, NULL, candidates.gon$PRLR, "PRLR")
+
+plot_grid(e,f + labs(y = NULL),g + labs(y = NULL), nrow = 1)
+
+e <- plotcanddateexpression(candidates.hyp, NULL, candidates.hyp$STAT5B, "STAT5B")
+f <- plotcanddateexpression(candidates.pit, NULL, candidates.pit$STAT5B, "STAT5B")
+g <- plotcanddateexpression(candidates.gon, NULL, candidates.gon$STAT5B, "STAT5B")
+
+plot_grid(e,f + labs(y = NULL),g + labs(y = NULL), nrow = 1)
 ```

### 2019-10-20 analysis/07_hormones.Rmd
new hormone and candidate gene plots

@@
 
-```{r characterization}
+```{r characterization, fig.width=5.5, fig.height= 3}
 hormonecharplot <- function(myhormone, myylab){
@@
            hormone %in% c(myhormone))  %>% 
-  ggplot(aes(x = timepoint, y = plasma_conc, fill = timepoint, color = sex)) +
-    geom_boxplot(aes(outlier.colour = timepoint), lwd=1.05) +
+    droplevels() %>% 
+  ggplot(aes(x = as.numeric(timepoint), y = plasma_conc)) +
+    geom_boxplot(aes(outlier.colour = timepoint, fill = timepoint, alpha = sex)) +
+    theme_B3() +
     theme(axis.text.x =  element_blank(),
@@
     guides(fill = guide_legend(order=1),
-         color = guide_legend(order=2))
+         color = guide_legend(order=2)) +
+    scale_alpha_manual(values = c(0.75,1)) +
+    geom_smooth(aes(colour = sex)) +
+    scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) 
   }
+
+hormonecharplot("prolactin", "PRL (ng/mL)")
 
@@
 d <- plot_grid(d1,d2, nrow = 1)
-abcd <- plot_grid(a, b ,c, d, labels = "AUTO")
-mylegend <- get_legend(c + theme(legend.position = "bottom"))
-plot_grid(abcd, mylegend, nrow = 2, rel_heights = c(0.85,0.15))
-
+abcd <- plot_grid(a, b ,c, d)
+abcd
 ```

### 2019-10-20 R/functions.R

@@ -459,6 +459,7 @@ readvsd <- function(filename){
   vsd <- read_csv(filename)
   vsd <- as.data.frame(vsd)
   row.names(vsd) <- vsd$X1
+  vsd$entrezid <- vsd$X1
   vsd$X1 <- NULL
   return(vsd)
 }
@@ -466,10 +467,48 @@ readvsd <- function(filename){
 readcolData <- function(filename){
   colData <- read_csv(filename)
   colData <- as.data.frame(colData)
+  
+  colData$treatment <- factor(colData$treatment, levels = 
+                                  c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
+  
   row.names(colData) <- colData$V1
+  colData$sample <- colData$V1
   colData$V1 <- NULL
   return(colData)
 }
+
+
+### select candidate gene vsds
+selectcandidatevsds <- function(listofgenes, vsd, colData){
+  
+  print(listofgenes)
+  
+  candidateentrezids <- geneinfo %>% 
+    filter(Name %in% listofgenes) %>%  dplyr::select(entrezid) 
+  
+  print(candidateentrezids$entrezid)
+  
+  rowstofilter <- as.list(candidateentrezids$entrezid)
+  
+  candidatevsd <- vsd %>% 
+    filter(entrezid %in% rowstofilter)
+  candidatevsd <- as.data.frame(candidatevsd)
+  row.names(candidatevsd) <- candidatevsd$entrezid
+  
+  candidatevsd <- right_join(geneinfo,candidatevsd) 
+  candidatevsd <- as.data.frame(candidatevsd)
+  row.names(candidatevsd) <- candidatevsd$Name
+  candidatevsd <- candidatevsd[-c(1:3)]
+  
+  candidatevsd <- as.data.frame(t(candidatevsd))
+  candidatevsd$sample <- row.names(candidatevsd)
+  
+  candidatevsd <- left_join(colData,candidatevsd)
+  
+  return(candidatevsd)
+}
+
+
 
 ######### plotcandidates ######### 
 

### 2019-10-20 analysis/07_hormones.Rmd
new figs

--- 2019-10-11 analysis/07_hormones.Rmd
@@
 
-```{r characterization}
+```{r characterizationplots}
 hormonecharplot <- function(myhormone, myylab){
@@
            hormone %in% c(myhormone))  %>% 
-  ggplot(aes(x = timepoint, y = plasma_conc, fill = timepoint, color = sex)) +
-    geom_boxplot(aes(outlier.colour = timepoint), lwd=1.05) +
+    droplevels() %>% 
+  ggplot(aes(x = as.numeric(timepoint), y = plasma_conc)) +
+        geom_smooth(aes(colour = sex)) +
+    geom_boxplot(aes(outlier.colour = timepoint, fill = timepoint, alpha = sex)) +
+    theme_B3() +
     theme(axis.text.x =  element_blank(),
@@
     guides(fill = guide_legend(order=1),
-         color = guide_legend(order=2))
+         color = guide_legend(order=2)) +
+    scale_alpha_manual(values = c(0.75,1)) +
+    scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) 
   }
 
-a <- hormonecharplot("prolactin", "PRL (ng/mL)")
+```
+
+```{r PRLonly, fig.width = 2.25, fig.height = 1}
+hormonecharplot("prolactin", "PRL (ng/mL)")
+```
+
+```{r sexsteroids, fig.width=2.25, fig.height= 3}
 b <- hormonecharplot("cort", "CORT (ng/mL)")
@@
 d <- plot_grid(d1,d2, nrow = 1)
-abcd <- plot_grid(a, b ,c, d, labels = "AUTO")
-mylegend <- get_legend(c + theme(legend.position = "bottom"))
-plot_grid(abcd, mylegend, nrow = 2, rel_heights = c(0.85,0.15))
-
+bcd <- plot_grid(b ,c, d, ncol = 1)
+bcd
 ```

### 2019-10-20 analysis/10_specificgenes.Rmd
new figs

@@
   ggplot(candidateexpression, aes(x = as.numeric(treatment), y = whichgene)) + 
+        geom_smooth(aes(colour = sex)) +
     geom_boxplot(aes(fill = treatment, alpha = sex)) + 
-    scale_alpha_manual(values = c(0.6,0.9)) +
+    scale_alpha_manual(values = c(0.75,1)) +
      theme_B3() +
@@
           axis.text.x = element_blank()) +
-    geom_smooth(aes(colour = sex)) +
     scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
@@
 
-plot_grid(a,b + labs(y = NULL),c + labs(y = NULL), nrow = 1)
+plot_grid(a, b + labs(y = NULL), c + labs(y = NULL), nrow = 1)
 
@@
 
-plot_grid(e,f + labs(y = NULL),g + labs(y = NULL), nrow = 1)
+plot_grid(e, f + labs(y = NULL), g + labs(y = NULL), nrow = 1)
 
-e <- plotcanddateexpression(candidates.hyp, NULL, candidates.hyp$STAT5B, "STAT5B")
-f <- plotcanddateexpression(candidates.pit, NULL, candidates.pit$STAT5B, "STAT5B")
-g <- plotcanddateexpression(candidates.gon, NULL, candidates.gon$STAT5B, "STAT5B")
+h <- plotcanddateexpression(candidates.hyp, " ", candidates.hyp$STAT5B, "STAT5B")
+i <- plotcanddateexpression(candidates.pit, " ", candidates.pit$STAT5B, "STAT5B")
+j <- plotcanddateexpression(candidates.gon, " ", candidates.gon$STAT5B, "STAT5B")
 
-plot_grid(e,f + labs(y = NULL),g + labs(y = NULL), nrow = 1)
+plot_grid(h, i + labs(y = NULL), j + labs(y = NULL), nrow = 1)
 ```
+
+```{r PRLboxplots2, fig.width = 2.25, fig.height= 4}
+plot_grid(a,b,c, nrow = 3)
+```

### 2019-10-24 analysis/07_hormones.Rmd
confirming correlations between hormone and gene expression

@@
 ```
+
+# do control bird with high prolactin hormone have high PRL expression in the pituitary? yes.
+
+```{r PRLvPRL}
+PRLpit <- read_csv("../results/10_PRLpit.csv") %>% 
+  filter(treatment == "control") %>% 
+  arrange(desc(PRL))
+head(PRLpit,2)
+```
+
+```{r}
+prolactin <- read_excel("../results/Pigeon prolactin concentrations juil 2018.xlsx", sheet = 1)  %>% 
+            filter(Treatment == "baseline")  %>% 
+            select(ColorBands,BandNo_old, Treatment, Sex, `Prolactin ng/mL`)  %>% 
+            arrange(desc(`Prolactin ng/mL`))
+
+head(prolactin,2)
+```

### 2019-11-10 analysis/11_pcalda.Rmd
new pca and tsne figures for talk

--- 2019-10-18 analysis/11_pcalda.Rmd
@@
 library(viridis)
-knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/pca/')
 
+knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/pca/', cache = T)
+
+# custom theme for plots
+mytheme <- function(){
+  theme_minimal(base_size = 12) + 
+    theme(panel.grid = element_blank(),
+          panel.background = element_rect(fill = "transparent", colour = NA),
+          plot.background = element_rect(fill = "transparent", colour = NA),
+          legend.title = element_blank())
+}
 ```
@@
 
-```{r pca}
-# make pca
+```{r}
+# make pca dataframes
 mypca <- prcomp(countData)
@@
 head(mypcadf)
+```
 
+```{r pca}
 a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
                          color = colData$treatment, 
-                         shape = colData$tissue)) +
-  geom_point() + labs(x = "PC1: 35.9%", y = "PC2: 32.5%") + theme_minimal(base_size = 6)  +
-  theme(legend.title = element_blank(), legend.position = "left")
-a
+                         shape = colData$tissue,
+                         alpha = colData$sex)) +
+  geom_point(size = 2)  + theme_minimal(base_size = 6)  +
+  mytheme() +
+  theme(legend.title = element_blank(), 
+        axis.text = element_blank()) +
+  scale_alpha_manual(values = c(0.5,1)) +
+  labs(x = "PC1: 35.9% of variance", y = "PC2: 32.5% of variance")
 
@@
                    repel = TRUE ,
-                  select.var= list(contrib = 6))
-b
+                  select.var= list(contrib = 6)) + 
+      mytheme() + 
+      labs(x = " ", y = " ", subtitle = NULL, title =  NULL)
 
-ab <- plot_grid(a,b + theme(axis.text = element_blank()) + ggtitle(NULL), rel_widths = c(0.6,0.4))
-ab
+png('../figures/pca/pca-1.png',width=600,height=600, units="px",bg = "transparent")
+print(a)
+dev.off()
+
+png('../figures/pca/pca-2.png',width=600,height=600, units="px",bg = "transparent")
+print(b)
+dev.off()
+```
+
+```{r pcaloadings}
+fviz_contrib(mypca, choice = "var", axes = 1, top = 5) 
+fviz_contrib(mypca, choice = "var", axes = 2, top = 5) 
 ```
@@
   geom_point(size = 2) +
-  theme_minimal(base_size = 6) +
+  mytheme() +
   scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
-  theme(legend.position = "top") +
   labs(x = "tSNE 1", y = "tSNE 2")
 
-d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = treatment)) +
+d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = tissue, color = treatment, alpha = sex)) +
   geom_point(size = 2) +
-  theme_minimal(base_size = 6) +
-  theme(legend.position = "top") +
-  labs(x = "tSNE 1", y = "tSNE 2")
+  mytheme() +
+  labs(x = "tSNE 1", y = "tSNE 2") +
+  scale_alpha_manual(values = c(0.5,1))
 
-cd <- plot_grid(c,d)
-cd
+c
+d
+
+png('../figures/pca/tSNE-2.png',width=600,height=600, units="px",bg = "transparent")
+print(c)
+dev.off()
+
+png('../figures/pca/tSNE-3.png',width=600,height=600, units="px",bg = "transparent")
+print(d)
+dev.off()
 ```
@@
 # combined pca tsne
-
-```{r pca.tSNE, fig.width=6, fig.height=6}
-plot_grid(ab,cd, ncol = 1)
-
-plot_grid(a + theme(legend.position = "none"),
-          b + theme(axis.text = element_blank()) + ggtitle(NULL), 
-          c + theme(legend.position = "none"),
-          d + theme(legend.position = "none"))
-```

### 2019-11-10 analysis/04_DESeq2_sexes.Rmd
fewere wgcna plots

@@
     mutate(hypothesis = fct_recode(treatment,
-                            "anticipation" = "control",
-                            "anticipation" = "bldg",
+                            "non-parental" = "control",
+                            "non-parental" = "bldg",
                             "incubation" = "lay",
@@
 # pitutiary expression
+plotWGCNAcandidates(vsd.pit, c("NP_990797.2"), colDataPit, "PRL expression, Pituitary")
 
-plotWGCNAcandidates(vsd.pit, c("NP_990797.2"), colDataPit, "PRL expression, Pituitary")
-plotWGCNAcandidates(vsd.pit, c("NP_001090992.1"), colDataPit, "VIP1R expression, Pituitary")
-plotWGCNAcandidates(vsd.pit, c("NP_990797.2", "NP_001090992.1"), colDataPit, "PRL & VIP1R expression, Pituitary")
-plotWGCNAcandidates(vsd.pit, c("NP_990136.1", "NP_001035179.1"), colDataPit, "TH & AR expression, Pituitary")
-plotWGCNAcandidates(vsd.pit, c("NP_990797.2", "NP_001090992.1","NP_990136.1", "NP_001035179.1", "NP_001026269.1"), 
-                    colDataPit, "PRL, PRLR, VIP1R, POMC, TH & AR expression, Pituitary")
-plotWGCNAcandidates(vsd.pit, c("XP_015155078.1", "XP_015146451.1", "NP_001025709.2", "NP_990327.2"), 
-                    colDataPit, "Stat and JAK proteins, Pituitary")
-plotWGCNAcandidates(vsd.pit,
-                    c("NP_990797.2","NP_001004392.2", "NP_001004403.1", "NP_001006255.2", "NP_001006274.1",
-                           "NP_989500.1", "XP_001234815.2", "XP_015149152.1", "XP_416110.5"),
-                    colDataPit, "9 of 102 genes in a WGCNA module containting Prolactin")
-plotWGCNAcandidates(vsd.pit,
-                    c("XP_428110.3","NP_001004392.2",  "XP_015142961.1", "XP_015139636.1",
-                      "XP_015137408.1", "XP_004938696.1", "NP_990620.1", "NP_001170793.2", "NP_001026250.1"),
-                    colDataPit, "Prolaction receptor: 9 of 147 genes in a WGCNA module containting PRLR")
-plotWGCNAcandidates(vsd.pit, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
-                               "NP_001131120.1", "NP_001186037.1", "NP_989871.1",
-                               "NP_990797.2", "XP_015132722.1"), colDataPit, "Pitutary: CISH, SOCS, PRL, PRLR")
+# COX1,  PRL, 
+topPCAloadings <- c("NP_006917.1",  "NP_990797.2")
+h <- plotWGCNAcandidates(vsd.hyp, topPCAloadings, colDataHyp, "Hypothalamus")
+p <- plotWGCNAcandidates(vsd.pit, topPCAloadings, colDataPit, "Pituitary")
 
-# genes from https://link.springer.com/article/10.1007%2Fs12079-012-0168-0
-# A pathway map of prolactin signaling
-# PL, RAC1, IRS2 not found
-plotWGCNAcandidates(vsd.pit, c("NP_990797.2", "XP_015132722.1", "NP_989690.1",
-                               "NP_001025709.2", "XP_015144878.1", "NP_989481.1",
-                               "NP_001186467.1", "NP_990386.1", "XP_015132490.1",
-                               "XP_417614.3", "XP_015150421.1", "NP_001155844.1"), 
-                    colDataPit, "Pit: Genes involved in prolactin signaling")
-
-# hyp
-plotWGCNAcandidates(vsd.hyp, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
-                               "NP_001131120.1", "NP_001186037.1", "NP_989871.1",
-                               "NP_990797.2", "XP_015132722.1"), colDataHyp, "Hypothalamus: CISH, SOCS, PRL, PRLR")
-plotWGCNAcandidates(vsd.hyp, c("NP_001170780.1", "NP_001090992.1", "NP_001014970.1"), colDataHyp, "VIP expression, Hypothalamus")
-
-plotWGCNAcandidates(vsd.hyp, c("NP_990797.2", "XP_015132722.1", "NP_989690.1",
-                               "NP_001025709.2", "XP_015144878.1", "NP_989481.1",
-                               "NP_001186467.1", "NP_990386.1", "XP_015132490.1",
-                               "XP_417614.3", "XP_015150421.1", "NP_001155844.1"), 
-                    colDataHyp, "Hyp: Genes involved in prolactin signaling")
-
-# gonad
-plotWGCNAcandidates(vsd.gon, c("NP_989957.1", "NP_001120784.1", "NP_001120786.1",
-                               "NP_001131120.1", "NP_001186037.1", "NP_989871.1",
-                               "NP_990797.2", "XP_015132722.1"), colDataGon, "Gonad: CISH, SOCS, PRL, PRLR")
-
-plotWGCNAcandidates(vsd.gon, c("NP_990797.2", "XP_015132722.1", "NP_989690.1",
-                               "NP_001025709.2", "XP_015144878.1", "NP_989481.1",
-                               "NP_001186467.1", "NP_990386.1", "XP_015132490.1",
-                               "XP_417614.3", "XP_015150421.1", "NP_001155844.1"), 
-                    colDataGon, "Gonad: Genes involved in prolactin signaling")
-
+plot_grid(h + theme(axis.text.x = element_blank(), legend.position = "none"),
+          p , nrow = 2, rel_heights = c(0.45,0.55))
 ```
@@
 prl72 <-  PRL_associated$entrezid[58:102]
+
 plotWGCNAcandidates(vsd.pit, prl36, colDataPit, "A-K genes with PRL pattern in the pituitary")
 plotWGCNAcandidates(vsd.pit, prl72, colDataPit, "K-Z genes with PRL pattern in the pituitary")
-
-PRLR_associated <- read.csv("../results/08_PRLR_associated.csv", header = T)
-prlr36 <-  PRLR_associated$entrezid[1:57]
-prlr72 <-  PRLR_associated$entrezid[58:115]
-prlr116<-  PRLR_associated$entrezid[116:147]
-plotWGCNAcandidates(vsd.pit, prlr36, colDataPit, "A-G genes with PRLR pattern in the pituitary")
-plotWGCNAcandidates(vsd.pit, prlr72, colDataPit, "L-S genes with PRLR pattern in the pituitary")
-plotWGCNAcandidates(vsd.pit, prlr116, colDataPit, "S-Z genes with PRLR pattern in the pituitary")
-
-VIPR1_associated <- read.csv("../results/08_VIPR1_associated.csv", header = T)
-vipr156 <-  VIPR1_associated$entrezid[398:456]
-plotWGCNAcandidates(vsd.pit, vipr156, colDataPit, "Genes with VIPR1 pattern in the pituitary")
-
-AR_associated <- read.csv("../results/08_AR_associated.csv", header = T)
-AR <-  AR_associated$entrezid[1:31]
-plotWGCNAcandidates(vsd.pit, AR, colDataPit, "Genes with AR pattern in the pituitary")
-
-POMC_associated <- read.csv("../results/08_POMC_associated.csv", header = T)
-pomc <-  POMC_associated$entrezid[32:61]
-plotWGCNAcandidates(vsd.pit, pomc, colDataPit, "Genes with POMC pattern in the pituitary")
-
-```
-
-```{r topPCAloadings}
-# COX1, COX3, CHGA, PRL, ATP2B4, LOC107055658
-topPCAloadings <- c("NP_006917.1", "NP_006921.2","XP_421330.1", "NP_990797.2", "XP_015154447.1", "XP_015156300.1")
-
-plotWGCNAcandidates(vsd.hyp, topPCAloadings, colDataHyp, "Hypothalamus")
-plotWGCNAcandidates(vsd.pit, topPCAloadings, colDataPit, "Pituitary")
-plotWGCNAcandidates(vsd.gon, topPCAloadings, colDataGon, "Gonads")
-
 ```

### 2019-11-10 R/functions.R

@@ -613,60 +613,44 @@ plotWGCNAcandidates <- function(vsd, mygenelist, colData, mysubtitle){
     geom_point() +
     geom_smooth(se = FALSE) +
     facet_wrap(~Name, scales = "free_y") +
-    theme_rmh() +
+    mytheme() +
     theme(legend.position = "bottom",
-          axis.text.x = element_text(angle = 60,  hjust=1),
-          axis.text.y = element_blank()) +
+          axis.text.x = element_text(angle = 60,  hjust=1)) +
     guides(fill = guide_legend(nrow = 1)) +
     labs(x = NULL, y = "Gene expression",
          subtitle = mysubtitle) +
     scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
-                       labels=charlevels)
+                       labels=charlevels) +
+    scale_color_manual(values = colorstreatmentsex)
   return(p1)
   
 }
 
 plotWGCNAcandidatesManip <- function(vsd, mygenelist, colData, mysubtitle){
   
-  # make dataframe with geneids and names and counts
-  # how to gather: https://tidyr.tidyverse.org/reference/gather.html
-  
   vsd.df <- as.data.frame(assay(vsd))
   vsd.df$entrezid <- row.names(vsd.df)
   
   candidates <- full_join(geneinfo, vsd.df, by = "entrezid")
-  
-  # select genes of interst
-  
   candidates <- candidates %>%
     filter(entrezid %in% mygenelist) %>% droplevels()
-  
   candidates <- candidates %>% dplyr::select(-row.names, -entrezid, -geneid)
-  
   candidates_long <- candidates %>% gather(-Name, key = "sample", value = "value")
-  
   candidates_long$V1 <- candidates_long$sample
-  
   candidatecounts <- left_join(candidates_long, colData, by = "V1")
-  
   candidatecounts$Name <- as.factor(candidatecounts$Name)
   candidatecounts$treatment <- factor(candidatecounts$treatment, levels = maniplevels)
-  
-  head(candidatecounts)
-  
   bysextreatment <- group_by(candidatecounts, sex, treatment, Name)
   bysextreatment
   candidateST <- summarize(bysextreatment, expression = mean(value))
   
-  
-  p1 <- ggplot(candidateST, aes(x = as.numeric(treatment), y = expression, color = sex)) +
-    geom_point() +
-    geom_smooth(se = FALSE) +
+  p1 <- ggplot(candidateST, aes(x = as.numeric(treatment), y = expression)) +
[…]

### 2019-11-10 analysis/05_DESeq2_manipulate.Rmd
partial manip analysis

--- 2019-06-16 analysis/05_DESeq2_manipulate.Rmd
@@
 library(cowplot)
-library(RColorBrewer)
-library(pheatmap)
-library(kableExtra)
-library(viridis)
+library(factoextra)
+library(Rtsne)
 
@@
 source("../R/functions.R") 
+source("../R/themes.R")
 
@@
 ```
+
+## offspring removal
+
+```{r offspringremoval, eval = F}
+g <- plotremoval(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
+h <- plotremoval(DEGs.female_pituitary, "Female pituitary", "none")
+i <- plotremoval(DEGs.female_gonad, "Female gonad",  "none")
+j <- plotremoval(DEGs.male_hypothalamus, "Male hypothalamus", "none")
+k <- plotremoval(DEGs.male_pituitary, "Male pituitary", "none")
+l <- plotremoval(DEGs.male_gondad, "Male gonad", "none")
+
+legend <- get_legend(g)
+
+mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)
+
+mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
+mybarplots2
+
+pdf("../figures/sexes/offspringremoval-1.pdf", width = 10, height = 6)
+plot(mybarplots2)
+dev.off()
+
+```
+
+## prolong extend
+
+```{r prolongextend, eval = F}
+g <- plotprolongdelay(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
+h <- plotprolongdelay(DEGs.female_pituitary, "Female pituitary", "none")
+i <- plotprolongdelay(DEGs.female_gonad, "Female gonad",  "none")
+j <- plotprolongdelay(DEGs.male_hypothalamus, "Male hypothalamus", "none")
+k <- plotprolongdelay(DEGs.male_pituitary, "Male pituitary", "none")
+l <- plotprolongdelay(DEGs.male_gondad, "Male gonad", "none")
+
+legend <- get_legend(g)
+
+mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)
+
+mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
+mybarplots2
+
+pdf("../figures/sexes/prolongextend-1.pdf", width = 10, height = 6)
+plot(mybarplots2)
+dev.off()
+
+```
+
+# pca and tsne
+
+```{r}
+pseudocounts <- read_csv("../results/01_pseudo.counts.csv")
+head(pseudocounts[1:3])
+
+colData <- read.csv("../metadata/00_colData_manipluation.csv", header = T, row.names = 1)
+
+colData$treatment <- factor(colData$treatment, levels = maniplevels1)
+colData$tissue <- factor(colData$tissue, levels = c("hypothalamus", "pituitary", "gonad"))
+colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
+head(colData[1:3])
+
+geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+
+subsetcounts <- function(colData, countData){
+
+  # save counts that match colData
+  savecols <- as.character(colData$V1) 
+  savecols <- as.vector(savecols) 
+  countData <- countData %>% 
+    dplyr::select(X1,savecols) %>% 
+    dplyr::rename(entrezid = X1)
+
+  countData <- left_join(geneinfo, countData)
+  print(head(countData))
+  countData <- countData %>% dplyr::distinct(Name, .keep_all = TRUE) %>%
+    dplyr::select(-row.names)  %>% dplyr::select(-geneid) %>% dplyr::select(-entrezid) 
+  print(head(countData))
+  row.names(countData) <- countData$Name
+  countData$Name <- NULL
+  countData <- as.data.frame(t(countData))
+
+  return(countData)
+}
+
+countData <- subsetcounts(colData, pseudocounts)
+head(countData[1:3])
+head(colData)
+# confirm
+row.names(colData) <- colData$V1
+row.names(countData) == row.names(colData)
+
+euclidist <- dist(countData) # euclidean distances between the rows
+```
+
+```{r pcaprep}
+# make pca dataframes
+mypca <- prcomp(countData)
+
+# percent conribution
+eig.val <- get_eigenvalue(mypca)
+head(eig.val)
+
+mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
+                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
+                  ID = row.names(countData))
+mypcadf$V1 <- row.names(mypcadf)
+mypcadf <- left_join(colData, mypcadf)
+mypcadf <- mypcadf %>% dplyr::select(bird,sex,tissue,treatment,PC1:PC6)
+mypcadf$treatment <- factor(mypcadf$treatment, levels = maniplevels1)
+head(mypcadf)
+
+a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
+                         color = colData$treatment, 
+                         shape = colData$tissue,
+                         alpha = colData$sex)) +
+  geom_point(size = 2)  + 
+  mytheme() +
+  theme(legend.title = element_blank(), 
+        axis.text = element_blank()) +
+  scale_alpha_manual(values = c(0.5,1)) +
+  scale_color_manual(values = colorsmaniplevels1) +
+  labs(x = "PC1: 63.6% of variance", y = "PC2: 30.1% of variance")
+
+b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
+                   repel = TRUE ,
+                  select.var= list(contrib = 6)) + 
+      mytheme() + 
+      labs(x = " ", y = " ", subtitle = NULL, title =  NULL)
+
+png('../figures/manipulation/pcaprep-1.png',width=600,height=600, units="px",bg = "transparent")
+print(a)
+dev.off()
+
+png('../figures/manipulation/pcaprep-2.png',width=600,height=600, units="px",bg = "transparent")
+print(b)
+dev.off()
+```
+
+```{r tsne}
+
+tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=5, theta=0.5, dims=2)
+tsne_df = as.data.frame(tsne_model$Y) 
+plot(tsne_df$V1, tsne_df$V2)
+
+# prep for adding columns
+colData2 <- colData 
+colData2$V1 <- NULL
+tsne_df_cols <- cbind(colData2, tsne_df)
+
+c  <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
+  geom_point(size = 2) +
+  mytheme() +
+  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
+  labs(x = "tSNE 1", y = "tSNE 2")
+
+d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = tissue, color = treatment, alpha = sex)) +
+  geom_point(size = 2) +
+  mytheme() +
+  labs(x = "tSNE 1", y = "tSNE 2") +
+  scale_color_manual(values = colorsmaniplevels1) +
+  scale_alpha_manual(values = c(0.5,1))
+
+c
+d
+```

### 2019-11-11 analysis/11_pcalda.Rmd
now with char and manip

@@
 
-# custom theme for plots
-mytheme <- function(){
-  theme_minimal(base_size = 12) + 
-    theme(panel.grid = element_blank(),
-          panel.background = element_rect(fill = "transparent", colour = NA),
-          plot.background = element_rect(fill = "transparent", colour = NA),
-          legend.title = element_blank())
-}
+source("../R/themes.R")  
 ```
@@
 
-# combined pca tsne
+## all data
+
+```{r charmanip}
+# prep col data for all samples
+colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
+colData$treatment <- factor(colData$treatment, levels = alllevels2)
+colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
+colData$tissue <- factor(colData$tissue, levels = tissuelevels)
+row.names(colData) <- colData$V1
+
+# prep count data for all samples
+countData <- as.data.frame(pseudocounts)
+countData$entrezid <- countData$X1
+countData <- left_join(geneinfo, countData)
+countData <- countData %>% distinct(Name, .keep_all = TRUE) %>%
+    select(-row.names, -geneid, -entrezid, -X1) 
+countData <- as.data.frame(countData)
+row.names(countData) <- countData$Name
+countData$Name <- NULL
+countData$entrezid <- NULL
+countData <- as.data.frame(t(countData))
+
+# check ready for analysis
+row.names(countData) == row.names(colData)
+
+# pca
+mypca <- prcomp(countData)
+eig.val <- get_eigenvalue(mypca)
+head(eig.val)
+
+mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
+                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
+                  ID = row.names(countData))
+mypcadf$V1 <- row.names(mypcadf)
+mypcadf <- left_join(colData, mypcadf)
+mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
+head(mypcadf)
+
+all1 <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
+                         color = colData$treatment, 
+                         #shape = colData$tissue,
+                         #size = colData$sex,
+                        shape = colData$study)) +
+  geom_hline(yintercept=0, color = "grey") +
+  geom_vline(xintercept=0, color = "grey") +
+  geom_point(size = 2)  + theme_minimal(base_size = 6)  +
+  mytheme() +
+  theme(legend.title = element_blank(),
+        axis.text = element_blank()) +
+  #scale_alpha_manual(values = c(0.5,1)) +
+  labs(x = "PC1: 44.6% of variance", y = "PC2: 29.8% of variance") +
+  scale_color_manual(values = colorscharmaip) 
+all1
+
+all2 <- fviz_pca_var(mypca,  labelsize = 4 , axes.linetype = "blank", 
+                   repel = TRUE ,
+                  select.var= list(contrib = 6)) + 
+      mytheme() + 
+      labs(x = " ", y = " ", subtitle = NULL, title =  NULL) +
+    geom_hline(yintercept=0, color = "grey") +
+  geom_vline(xintercept=0, color = "grey") +
+    theme( axis.text = element_blank())
+
+# prep for tsne
+euclidist <- dist(countData) # euclidean distances between the rows
+
+tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=10, theta=0.5, dims=2)
+tsne_df = as.data.frame(tsne_model$Y) 
+plot(tsne_df$V1, tsne_df$V2)
+
+# prep for adding columns
+colData2 <- colData 
+colData2$V1 <- NULL
+tsne_df_cols <- cbind(colData2, tsne_df)
+
+ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
+  geom_point(size = 2) +
+  mytheme() +
+  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
+  labs(x = "tSNE 1", y = "tSNE 2")
+
+ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = study, color = tissue)) +
+  geom_point(size = 2) +
+  mytheme() +
+  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
+  labs(x = "tSNE 1", y = "tSNE 2")
+
+ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = study, color = treatment)) +
+  geom_point(size = 2) +
+  mytheme() +
+  labs(x = "tSNE 1", y = "tSNE 2") +
+  scale_alpha_manual(values = c(0.5,1)) +
+  scale_color_manual(values = colorscharmaip)
+```
+
+```{r transparentplots}
+png('../figures/pca/all-1.png',width=600,height=600, units="px",bg = "transparent")
+print(all1)
+dev.off()
+
+png('../figures/pca/all-2.png',width=600,height=600, units="px",bg = "transparent")
+print(all2)
+dev.off()
+```

### 2019-11-11 analysis/07_hormones.Rmd
removal and timeing in two differnt plots

@@
 
-source("../R/functions.R")  # load custom functions 
 source("../R/themes.R")  # load custom themes and color palletes
@@
 knitr::opts_chunk$set(fig.path = '../figures/hormones/',message=F, warning=FALSE)
-```
-
-```{r clean}
-charlevels <- c("control", "bldg", "lay", "inc_d3", "inc_d9", "inc_d17", "hatch", "n5", "n9" )
-removelevels <- c("M_inc3",  "M_inc9", "M_inc17", "M_n2")
-maniplevels <- c("early", "prolong", "extend")
-combolevels <- c("control", "bldg", "lay", "inc_d3", "M_inc3",
-                  "inc_d9",  "M_inc9", "early",  "inc_d17","M_inc17",  "hatch",
-                 "prolong", "extend", "M_n2",
-                  "n5", "n9" )
-
-manipchar <- c("early",  "inc_d17", "hatch", "prolong", "extend", "n5")
-removechar <- c("inc_d3", "M_inc3",  "inc_d9", "M_inc9","inc_d17", "M_inc17","hatch", "M_n2")
-
-sexcolors <- c("female" = "#F8766D", "male" = "#00BFC4")
-charcolors <- c("control" = "#F8766D", "bldg" = "#D39200",
-                "lay" =  "#93AA00", "inc_d3" = "#00BA38" ,
-                "inc_d9" = "#00C19F" , "inc_d17" =  "#00B9E3", 
-                "hatch" = "#619Cff" , "n5" =  "#DB72Fb", "n9" =  "#FF61C3")
-
-charsexlevels <- c("control female", "control male", "bldg female", "bldg male",  "lay female", "lay male",
-                "inc_d3 female", "inc_d3 male", "inc_d9 female", "inc_d9 male",
-                "inc_d17 female", "inc_d17 male", "hatch female", "hatch male", 
-                "n5 female", "n5 male", "n9 female" , "n9 male")
-
 ```
@@
 
-# rename some of the timepoints levels, store in new columns
+# rename some of the treatments levels, store in new columns
 prolactin <- prolactin %>%
@@
                             "male" = "m"),
-           timepoint = fct_recode(Treatment,
+           treatment = fct_recode(Treatment,
                             "hatch" = "Hatch",
-                            "inc_d17" = "Inc_d17",
-                            "inc_d3" = "Inc_d3",
-                            "inc_d9" = "Inc_d9",
-                            "M_inc9" = "M_Inc9",
-                            "M_inc3" = "M_Inc3",
-                            "early" = "M_Inc8",
-                            "early" = "M_inc8",
-                            "M_inc17" = "M_Inc17",
-                            "M_n2" = "M_hatch",
+                            "inc.d17" = "Inc_d17",
+                            "inc.d17" = "inc_d17",
+                            "inc.d3" = "Inc_d3",
+                             "inc.d3" = "inc_d3",
+                            "inc.d9" = "Inc_d9",
+                             "inc.d9" = "inc_d9",
+                            "m.inc.d9" = "M_Inc9",
+                            "m.inc.d9" = "M_inc9",
+                            "m.inc.d3" = "M_Inc3",
+                            "m.inc.d8" = "M_Inc8",
+                            "m.inc.d8" = "M_inc8",
+                            "m.inc.d17" = "M_Inc17",
+                            "m.n2" = "M_hatch",
                             "control" = "baseline",
@@
                             "n9" = "N9"),
-           study = fct_collapse(timepoint,
+           study = fct_collapse(treatment,
                                  characterization = charlevels,
-                                 manipulation = maniplevels,
-                                removal = removelevels))
+                                 manipulation = maniplevels1))
 
@@
 prolactin$hormone <- "prolactin"
-prolactin <- prolactin %>% select(study, timepoint, sex, hormone, plasma_conc)   %>% drop_na()
+prolactin <- prolactin %>% 
+              select(study, treatment, sex, hormone, plasma_conc)  %>% 
+              drop_na()
 
@@
 PETC <- PETC %>% select(stage, sex, hormone, plasma_conc)  %>%
-                filter(stage %in% combolevels)  %>%
-                droplevels()  
-PETC$stage <- factor(PETC$stage)
-levels(PETC$stage)
-
-PETC <- PETC %>%
+                mutate(treatment = fct_recode(stage,
+                            "inc.d17" = "inc_d17",
+                            "inc.d3" = "inc_d3",
+                            "inc.d9" = "inc_d9",
+                            "m.inc.d9" = "m_incd9",
+                            "m.inc.d3" = "m_incd3",
+                            "m.inc.d8" = "m_incd8",
+                            "m.inc.d17" = "m_incd17",
+                            "m.n2" = "m_hatch",
+                            "control" = "stress_hpg")) %>%
+                filter(treatment %in% alllevels2) %>%   
                 mutate(sex = fct_recode(sex,
@@
                             "male" = "m"),
-           study = fct_collapse(stage,
+                       study = fct_collapse(treatment,
                                 characterization = charlevels,
-                                manipulation = maniplevels)) %>% 
-  drop_na()
-colnames(PETC)[colnames(PETC)=="stage"] <- "timepoint"
-PETC <- PETC %>% select(study, timepoint, sex, hormone, plasma_conc)
+                                manipulation = maniplevels1)) %>% 
+              drop_na() %>%  droplevels()  %>% 
+              select(study, treatment, sex, hormone, plasma_conc)
 summary(PETC)
@@
 hormones <- rbind(prolactin, PETC)
-hormones$timepoint <- factor(hormones$timepoint, levels = combolevels)
-
-hormones <- hormones %>%  mutate(timepoint.sex = paste(timepoint, sex, sep = " ")) 
-hormones$timepoint.sex <- factor(hormones$timepoint.sex, levels = charsexlevels)
+hormones$treatment <- factor(hormones$treatment, levels = alllevels)
 
@@
 summary(hormones)
-
-prl.char <- hormones %>% filter(hormone == "prolactin", timepoint %in% charlevels)   %>%  droplevels()
-test.char <- hormones %>% filter(hormone == "testosterone", timepoint %in% charlevels)   %>%  droplevels()
-est.char <- hormones %>% filter(hormone == "estradiol", timepoint %in% charlevels)   %>%  droplevels()
-prog.char <- hormones %>% filter(hormone == "progesterone", timepoint %in% charlevels)   %>%  droplevels()
-cort.char <- hormones %>% filter(hormone == "cort", timepoint %in% charlevels)   %>%  droplevels()
-
 ```
@@
 
-  mycolors <- charcolors
-  sexcolors <- c("female" = "#969696", "male" = "#525252")
-
   hormones %>% 
@@
     droplevels() %>% 
-  ggplot(aes(x = as.numeric(timepoint), y = plasma_conc)) +
+  ggplot(aes(x = as.numeric(treatment), y = plasma_conc)) +
         geom_smooth(aes(colour = sex)) +
-    geom_boxplot(aes(outlier.colour = timepoint, fill = timepoint, alpha = sex)) +
-    theme_B3() +
+    geom_boxplot(aes(outlier.colour = treatment, fill = treatment, alpha = sex)) +
+    mytheme() +
     theme(axis.text.x =  element_blank(),
           legend.position = "none") +
-    scale_fill_manual(values = mycolors) +
+    scale_fill_manual(values = colorscharmaip) +
     scale_color_manual(values = sexcolors) +
@@
 
-```{r PRLonly, fig.width = 2.25, fig.height = 1}
+```{r PRLonly}
 hormonecharplot("prolactin", "PRL (ng/mL)")
@@
 
-```{r sexsteroids, fig.width=2.25, fig.height= 3}
+```{r sexsteroids}
 b <- hormonecharplot("cort", "CORT (ng/mL)")
@@
 ```{r aov-char}
+prl.char <- hormones %>% filter(hormone == "prolactin", treatment %in% charlevels)   %>%  droplevels()
+test.char <- hormones %>% filter(hormone == "testosterone", treatment %in% charlevels)   %>%  droplevels()
+est.char <- hormones %>% filter(hormone == "estradiol", treatment %in% charlevels)   %>%  droplevels()
+prog.char <- hormones %>% filter(hormone == "progesterone", treatment %in% charlevels)   %>%  droplevels()
+cort.char <- hormones %>% filter(hormone == "cort", treatment %in% charlevels)   %>%  droplevels()
+
 aovSexTretment <- function(mydata, whichormone){
-  aov2 <- aov(data = mydata, plasma_conc ~ timepoint + sex)
+  aov2 <- aov(data = mydata, plasma_conc ~ treatment * sex)
   print(whichormone)
   print(summary(aov2))
-  #print(TukeyHSD(aov2, which = "timepoint"))
+  #print(TukeyHSD(aov2, which = "treatment"))
 }
@@
 aovTretment  <- function(mydata, whichormone){
-  aov1 <- aov(data = mydata, plasma_conc ~ timepoint )
+  aov1 <- aov(data = mydata, plasma_conc ~ treatment )
   print(whichormone)
   print(summary(aov1))
-  #print(TukeyHSD(aov1, which = "timepoint"))
+  #print(TukeyHSD(aov1, which = "treatment"))
 }
 
-aovSexTretment(prl.char, "PRL")
-aovSexTretment(cort.char, "CORT")
-aovSexTretment(prog.char, "PROG")
+aovSexTretment(prl.char, "PRL") # yes, sex difference (p = 0.00256), yes treatment effect (p < 2e-16), no interaction
+aovSexTretment(cort.char, "CORT") # no sex difference (p = 0.779 ), small treatment effect (p = 0.0238),  no interaction
+aovSexTretment(prog.char, "PROG") # no sex difference or treatment effect, signifiant interacion (p = 0.0107)
 
-aovTretment(est.char, "E")
-aovTretment(test.char, "T")
+aovTretment(est.char, "E") # p = 0.101
+aovTretment(test.char, "T") # p = 0.609
 ```
@@
 ```{r manipulation}
-hormonemanipPRL <- function(myhormone, myylab){
-
-  mycolors <- sexcolors
-
+# prolactin, removal only
 hormones %>% 
-    filter(timepoint %in% manipchar,
-           hormone %in% c(myhormone))  %>% 
-  ggplot(aes(x = timepoint, y = plasma_conc, fill = sex)) +
+    filter( hormone == c("prolactin"))  %>% 
+    filter( !treatment %in% c("m.inc.d8", "prolong", "extend"))  %>% 
+  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment,  color = sex)) +
     geom_boxplot() + 
@@
           strip.text = element_blank()) +
-    scale_fill_manual(values = mycolors) +
-    labs(y = myylab, x = NULL) +
-    facet_wrap(~sex) +
-    annotate("rect", xmin = 0.5, xmax = 1.5, ymin = 0, ymax = 125, alpha = .2) +
-    annotate("rect", xmin = 3.5, xmax = 5.5, ymin = 0, ymax = 125, alpha = .2) 
-}
-
-hormonemanipPRL("prolactin", "PRL (ng/mL)")
-
-hormonemanipSteroids <- function(myhormone, myylab, myymax){
-
-  mycolors <- sexcolors
+    scale_fill_manual(values = colorscharmaip) +
+    scale_color_manual(values = sexcolors) +
+    labs(y = "PRL (ng/mL)", x = NULL) +
+    annotate("rect", xmin = 3.7, xmax = 5.3, ymin = -2, ymax = 0, alpha = 1) +
+    annotate("rect", xmin = 5.7, xmax = 7.3, ymin = -2, ymax = 0, alpha = 1) +
+    annotate("rect", xmin = 7.7, xmax = 9.3, ymin = -2, ymax = 0, alpha = 1) +
+    annotate("rect", xmin = 9.7, xmax = 11.3, ymin = -2, ymax = 0, alpha = 1)
 
 hormones %>% 
-    filter(timepoint %in% manipchar,
-           hormone %in% c(myhormone))  %>% 
-  ggplot(aes(x = timepoint, y = plasma_conc, fill = sex)) +
+    filter( hormone == c("prolactin"))  %>% 
+    filter( !treatment %in% c("m.inc.d3", "m.inc.d9", "m.inc.d17","m.n2"))  %>% 
+  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment,  color = sex)) +
     geom_boxplot() + 
@@
           strip.text = element_blank()) +
-    scale_fill_manual(values = mycolors) +
-    labs(y = myylab, x = NULL) +
-    facet_wrap(~sex) +
-    annotate("rect", xmin = 2.5, xmax = 4.5, ymin = 0, ymax = myymax, alpha = .2)
+    scale_fill_manual(values = colorscharmaip) +
+    scale_color_manual(values = sexcolors) +
+    labs(y = "PRL (ng/mL)", x = NULL) +
+    annotate("rect", xmin = 4.7, xmax = 6.3, ymin = -2, ymax = 0, alpha = 1, alpha = 1) +
+    annotate("rect", xmin = 6.7, xmax = 8.3, ymin = -2, ymax = 0, alpha = 1, alpha = 1) +
+    annotate("rect", xmin = 8.7, xmax = 10.3, ymin = -2, ymax = 0, alpha = 1, alpha = 1)
+
+```
+
+hormonemanipSteroids <- function(myhormone, myylab, myymax){
+
+  hormones %>% 
+    filter( hormone %in% c(myhormone))  %>% 
+  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
+    geom_boxplot() + 
+    mytheme() +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+         # legend.position = "none",
+          strip.text = element_blank()) +
+    scale_fill_manual(values = colorscharmaip) +
+    scale_color_manual(values = sexcolors) +
+    labs(y = myylab, x = NULL) 
 }
@@
 
-plot_grid(d,a,c,b, rel_widths = c(0.65,0.35))
-```
-```{r removal}
-hormoneremoval <- function(myhormone, myylab){
-
-  mycolors <- c("female" = "#F8766D", "male" = "#00BFC4")
-
-  hormones %>% 
-    filter(timepoint %in% removechar,
-           hormone %in% c(myhormone))  %>% 
-  ggplot(aes(x = timepoint, y = plasma_conc, fill = sex)) +
-    geom_boxplot() + 
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
-          legend.position = "none",
-          strip.text = element_blank()) +
-    scale_fill_manual(values = mycolors) +
-    labs(y = myylab, x = NULL) +
-    facet_wrap(~sex) +
-    annotate("rect", xmin = 1.5, xmax = 2.5, ymin = 0, ymax = 125, alpha = .2) +
-    annotate("rect", xmin = 3.5, xmax = 4.5, ymin = 0, ymax = 125, alpha = .2) +
-    annotate("rect", xmin = 5.5, xmax = 6.5, ymin = 0, ymax = 125, alpha = .2) + 
-    annotate("rect", xmin = 7.5, xmax = 8.5, ymin = 0, ymax = 125, alpha = .2) 
-}
-
-hormoneremoval("prolactin", "PRL (ng/mL)")
+a
+b
+c
+d
 ```

### 2019-11-11 analysis/12_favegenes.Rmd
fav genes with pseduo counts

---
title: "favgenes"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/favegenes/', cache = T)
```

```{r pseudocounts}
pseudocounts <- read_csv("../results/01_pseudo.counts.csv")
head(pseudocounts[1:3])

geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

# prep col data for all samples
colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
colData$treatment <- factor(colData$treatment, levels = alllevels)
colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
colData$tissue <- factor(colData$tissue, levels = tissuelevels)
row.names(colData) <- colData$V1

# prep count data for all samples
countData <- as.data.frame(pseudocounts)
countData$entrezid <- countData$X1
countData <- left_join(geneinfo, countData)
countData <- countData %>% dplyr::distinct(Name, .keep_all = TRUE) %>%
    dplyr::select(-row.names, -geneid, -entrezid, -X1) 
countData <- as.data.frame(countData)
row.names(countData) <- countData$Name
countData$Name <- NULL
countData$entrezid <- NULL
countData <- as.data.frame(t(countData))

# keep names for 
favcounts <- countData %>% dplyr::select("PRL", "COX1", "GH", "CHGA", 
                                         "FOS", "EGR1", "GNRHR", "GNRH1")
# check okay to bind rather than join
row.names(countData) == row.names(colData)
favcounts <- cbind(colData, favcounts)

ggplot(favcounts, aes(x = treatment, y = PRL, fill = treatment, color = sex)) +
  geom_boxplot() + 
  mytheme() +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank())

ggplot(favcounts, aes(x = treatment, y = GH, fill = treatment, color = sex)) +
  geom_boxplot() + 
  mytheme() +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) 
```

### 2019-11-12 analysis/12_favegenes.Rmd
add some results

@@
 library(tidyverse)
+library(DESeq2)
+library(cowplot)
+
+library(BiocParallel)
+register(MulticoreParam(6))
+
 source("../R/themes.R") 
-
-knitr::opts_chunk$set(echo = TRUE)
+source("../R/functions.R")
 
@@
 countData <- as.data.frame(t(countData))
+head(countData)
 
@@
 ```
+
+```{r DESeq2-prep}
+# untranform count data and convert all doubles to integers for DESEq2
+countData <- read.csv("../results/00_counts.csv")
+head(countData)
+countData$entrezid <- countData$X
+countData <- left_join(geneinfo, countData)
+countData <- countData %>% dplyr::distinct(Name, .keep_all = TRUE) %>%
+    dplyr::select(-row.names, -geneid, -entrezid, -X) 
+countData <- as.data.frame(countData)
+row.names(countData) <- countData$Name
+countData$Name <- NULL
+
+# create column for subsetting
+colData$sextissue <- paste(colData$sex,colData$tissue, sep = "_")
+
+colData <- subsetcolData(colData, c("female_pituitary", "male_pituitary"))
+# save counts that match colData
+savecols <- as.character(colData$V1) 
+savecols <- as.vector(savecols) 
+countData <- countData %>% dplyr::select(one_of(savecols))
+
+ncol(countData) == nrow(colData)
+colnames(countData) == rownames(colData)
+
+dds <- DESeqDataSetFromMatrix(countData = countData,
+                                colData = colData,
+                                design = ~ treatment * sex)
+
+print(dds)
+ds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
+print(dim(dds))
+
+dds <- DESeq(dds) # Differential expression analysis
+
+vsd.pituitary <- vst(dds, blind=FALSE)
+
+vsd.pituitary.df <- as.data.frame(assay(vsd.pituitary))
+widecountdata <- as.data.frame(t(vsd.pituitary.df))
+widecountdata <- cbind(colData, widecountdata)
+```
+
+```{r contrasts}
+res <- results(dds, independentFiltering = T)
+summary(res)
+
+myresSig <- function(mycontrast, myfilename){
+  res <- results(dds, contrast = mycontrast)
+  print(summary(res))
+  resSig <- subset(res, padj < 0.1)
+  resSig <- as.data.frame((resSig))
+  resSig <-resSig[order(resSig$padj),]
+  print(head(resSig))
+  write.csv(resSig, myfilename)
+}
+
+myresSig(c("sex","male","female"), "../results/contrasts/resSig.sex.male.female.csv")
+myresSig(c("treatment","bldg","control"), "../results/contrasts/resSig.treatment.bldg.control.csv")
+myresSig(c("treatment","lay","bldg"), "../results/contrasts/resSig.treatment.lay.bldg.csv")
+myresSig(c("treatment","inc.d3","lay"), "../results/contrasts/resSig.treatment.inc.d3.lay.csv")
+
+myresSig(c("treatment","inc.d9","inc.d3"), "../results/contrasts/resSig.treatment.inc.d9.inc.d3.csv")
+myresSig(c("treatment","m.inc.d8","inc.d9"), "../results/contrasts/resSig.treatment.m.inc.d8.inc.d9.csv")
+myresSig(c("treatment","m.inc.d9","inc.d9"), "../results/contrasts/resSig.treatment.m.inc.d9.inc.d9.csv")
+
+myresSig(c("treatment","inc.d17","inc.d9"), "../results/contrasts/resSig.treatment.inc.d17.inc.d9.csv")
+myresSig(c("treatment","m.inc.d17","inc.d17"), "../results/contrasts/resSig.treatment.m.inc.d17.inc.d9.csv")
+
+myresSig(c("treatment","m.n2","hatch"), "../results/contrasts/resSig.treatment.m.n2.hatch.csv")
+myresSig(c("treatment","n5","hatch"), "../results/contrasts/resSig.treatment.n5.hatch.csv")
+myresSig(c("treatment","hatch","inc.d17"), "../results/contrasts/resSig.treatment.hatch.inc.d17.csv")
+
+```
+
+```{r DESeq2-allmanip}
+
+plotwidedata <- function(whichgene, myylab){
+p <- ggplot(widecountdata, (aes(x = treatment, y = whichgene))) +
+        geom_smooth(aes(colour = sex)) +
+    geom_boxplot(aes(outlier.color = treatment, fill = treatment, alpha = sex)) +
+    mytheme() +
+    theme(legend.position = "none",
+          axis.text.x = element_text(angle = 45, hjust = 1)) +
+    scale_fill_manual(values = colorscharmaip) +
+    scale_color_manual(values = sexcolors) +
+    labs(x = NULL, y = myylab) +
+    guides(fill = guide_legend(order=1),
+         color = guide_legend(order=2)) +
+    scale_alpha_manual(values = c(0.75,1)) 
+return(p)
+}
+
+a <- plotwidedata(widecountdata$PRL, "PRL")
+b <- plotwidedata(widecountdata$COX1, "COX1")
+c <- plotwidedata(widecountdata$FOSL2, "FOSL2")
+
+plot_grid(a + theme(axis.text.x = element_blank()),c, 
+          nrow = 2)
+b
+```
+
+```{r DESeq2-partialmanip}
+
+PRLremove <- widecountdata %>% 
+    filter( !treatment %in% c("m.inc.d8", "prolong", "extend"))  %>% 
+  ggplot(aes(x = treatment, y = PRL, fill = treatment,  color = sex)) +
+    geom_boxplot() + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none",
+          strip.text = element_blank()) +
+    scale_fill_manual(values = colorscharmaip) +
+    scale_color_manual(values = sexcolors) +
+    labs(y = "PRL expression", x = NULL,
+         subtitle = "PRL in female and male pituitary") +
+    annotate("rect", xmin = 3.7, xmax = 5.3, ymin = 14.5, ymax = 14.6, alpha = 1) +
+    annotate("rect", xmin = 5.7, xmax = 7.3, ymin = 14.5, ymax = 14.6, alpha = 1) +
+    annotate("rect", xmin = 7.7, xmax = 9.3, ymin = 14.5, ymax = 14.6, alpha = 1) +
+    annotate("rect", xmin = 9.7, xmax = 11.3, ymin = 14.5, ymax = 14.6, alpha = 1)
+
+PRLtiming <- widecountdata %>% 
+    filter( !treatment %in% c("m.inc.d3", "m.inc.d9", "m.inc.d17","m.n2"))  %>% 
+  ggplot(aes(x = treatment, y = PRL, fill = treatment,  color = sex)) +
+    geom_boxplot() + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none",
+          strip.text = element_blank()) +
+    scale_fill_manual(values = colorscharmaip) +
+    scale_color_manual(values = sexcolors) +
+    labs(y = "PRL expression", x = NULL,
+         subtitle = "PRL in female and male pituitary") +
+    annotate("rect", xmin = 4.7, xmax = 6.3, ymin = 14.5, ymax = 14.6, alpha = 1, alpha = 1) +
+    annotate("rect", xmin = 6.7, xmax = 8.3, ymin = 14.5, ymax = 14.6, alpha = 1, alpha = 1) +
+    annotate("rect", xmin = 8.7, xmax = 10.3, ymin = 14.5, ymax = 14.6, alpha = 1, alpha = 1)
+
+png('../figures/favegenes/PRLremove.png',width=600,height=400, units="px", res =)
+print(PRLremove)
+dev.off()
+
+png('../figures/favegenes/PRLtiming.png',width=600,height=400, units="px")
+print(PRLtiming)
+dev.off()
+```

### 2019-11-13 analysis/15_networks.Rmd
now with prolactin correlaed genes

---
title: "15_networks"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/favegenes/')

```

```{r}
library(tidyverse)
library(corrplot)
library(corrr)

source("../R/themes.R")

```

```{r getdata}
vsd.pituitary.df <- read.csv("../results/04_vsd_pit.csv", row.names = 1)
head(vsd.pituitary.df)

geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

vsd.pituitary.df$entrezid <- row.names(vsd.pituitary.df)
vsd.pituitary.df <- left_join(geneinfo, vsd.pituitary.df)
vsd.pituitary.df <- vsd.pituitary.df %>% dplyr::distinct(Name, .keep_all = TRUE) %>%
    dplyr::select(-row.names, -geneid, -entrezid) 
vsd.pituitary.df <- as.data.frame(vsd.pituitary.df)
row.names(vsd.pituitary.df) <- vsd.pituitary.df$Name
vsd.pituitary.df$Name <- NULL
head(vsd.pituitary.df)

colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
colData$treatment <- factor(colData$treatment, levels = charlevels)
colData <- colData %>% filter(tissue == "pituitary")
row.names(colData) <- colData$V1

widecountdata <- as.data.frame(t(vsd.pituitary.df))
widecountdata$V1 <- row.names(widecountdata)
widecountdata <- left_join(colData, widecountdata)

head(widecountdata)
```

```{r boxplots}
# genes from website shown to be associate with oxt and prl
OXTgenes <- c("OXT",  "AVPR1A", "AVPR1B", "TRH", 
              "HCRT", "GNRH1", "CRH", "POMC", "MC4R", "AVPR2")
PRLgenes <- c("PRL", "GNRH1", "TRH", "POU1F1", "POMC", "TSHB",
              "FSHB", "GH", "PRLR", "JAK2")
OXTPRLgenes <- c(OXTgenes, PRLgenes)
head(OXTPRLgenes)

# genes from wgcna pitutiry
PRL_associated <- read_csv("../results/08_PRL_associated.csv") %>%
  dplyr::pull(Name)
PRL_associated

longdata <- widecountdata %>% 
            select(V1:study, OXTPRLgenes, PRL_associated) %>%
            gather(gene, expression, OXT:UBE2C)
head(longdata)

plotnetworkgenes <- function(whichgenes){
  longdata %>%
  filter(gene %in% whichgenes) %>%
  ggplot(aes(x = treatment, y = expression)) +
        geom_smooth(aes(colour = sex)) +
    geom_boxplot(aes(outlier.color = treatment, 
                     fill = treatment, alpha = sex)) +
    mytheme() +
    theme(legend.position = "none",
          axis.text.x = element_blank()) +
    scale_fill_manual(values = colorschar) +
    scale_color_manual(values = sexcolors) +
    labs(x = NULL) +
    scale_alpha_manual(values = c(0.75,1)) +
  facet_wrap(~gene, scales = "free_y")
}

plotnetworkgenes(OXTgenes)
plotnetworkgenes(PRLgenes)
plotnetworkgenes(PRL_associated[85:93])

```

```{r correlations}
forcorrdata <- widecountdata %>% 
            select(V1, OXTPRLgenes)  %>% 
            as.data.frame()
row.names(forcorrdata) <- forcorrdata$V1   
forcorrdata$V1 <- NULL
head(forcorrdata)

cormat <- cor(forcorrdata, method = c("spearman"), use = "complete.obs") 
head(cormat)

forcorrdata %>% 
  correlate() %>% 
  network_plot( colors = c("#67a9cf", "#ef8a62"))

forcorrdata %>% 
  correlate() %>% 
  rplot(shape = 15, colors = c("#67a9cf", "#ef8a62") ) + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r wgcnaPRLgenes}
x <- widecountdata %>% 
            select(PRL_associated,OXTPRLgenes) %>%
       correlate()     # Create correlation data frame (cor_df)

x %>% rearrange() %>%  # rearrange by correlations
       focus(POLQ:SGOL1, mirror = TRUE) %>%
     shave() %>% # Shave off the upper triangle for a clean result
       fashion()

x %>%  focus(POLQ:SGOL1, mirror = TRUE) %>%
  network_plot( colors = c("#67a9cf", "#ef8a62"))

x %>%
  focus(POLQ:SGOL1, mirror = TRUE) %>%
  shave(upper = FALSE) %>%
  rplot(shape = 15, colors = c("#67a9cf", "#ef8a62") ) + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### 2019-11-13 analysis/12_favegenes.Rmd
new plots with manipluation and candidate genes

@@
     scale_color_manual(values = sexcolors) +
-    labs(x = NULL, y = myylab) +
+    labs(x = NULL, y = NULL, subtitle = myylab) +
     guides(fill = guide_legend(order=1),
@@
 
+plotwidedata(widecountdata$PRL, "PRL")
+
 a <- plotwidedata(widecountdata$PRL, "PRL")
-b <- plotwidedata(widecountdata$COX1, "COX1")
-c <- plotwidedata(widecountdata$FOSL2, "FOSL2")
+b <- plotwidedata(widecountdata$FOSL2, "FOSL2")
+c <- plotwidedata(widecountdata$COX1, "COX1")
+d <- plotwidedata(widecountdata$BRCA1, "BRCA1")
+e <- plotwidedata(widecountdata$CDK1, "CDK1")
+f <-plotwidedata(widecountdata$MYC, "MYC")
 
-plot_grid(a + theme(axis.text.x = element_blank()),c, 
-          nrow = 2)
-b
+plot_grid(a + theme(axis.text.x = element_blank()),
+          b + theme(axis.text.x = element_blank()),
+          c + theme(axis.text.x = element_blank()),
+          d + theme(axis.text.x = element_blank()),
+          e + theme(axis.text.x = element_blank()),
+          f + theme(axis.text.x = element_blank()), nrow = 2)
+
+plot_grid(a + theme(axis.text.x = element_blank()),
+          d + theme(axis.text.x = element_blank()),
+          e + theme(axis.text.x = element_blank()), nrow = 1)
 ```
@@
 ```{r DESeq2-partialmanip}
-
 PRLremove <- widecountdata %>% 
@@
 dev.off()
+
+removal <- widecountdata %>% 
+    filter( !treatment %in% c("m.inc.d8", "prolong", "extend",  "control", "bldg" , "lay", "n5", "n9"))
+
+plotremove <- function(whichgene, myylab){
+  p <-  removal  %>% 
+  ggplot(aes(x = treatment, y = whichgene, fill = treatment,  color = sex)) +
+    geom_boxplot() + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none",
+          strip.text = element_blank()) +
+    scale_fill_manual(values = colorscharmaip) +
+    scale_color_manual(values = sexcolors) +
+    labs(y = NULL, subtitle = myylab, x = NULL) 
+  return(p)
+}
+
+a <- plotremove(removal$PRL, "PRL")
+b <- plotremove(removal$FOSL2, "FOSL2")
+c <- plotremove(removal$COX1, "COX1")
+d <- plotremove(removal$BRCA1, "BRCA1")
+e <- plotremove(removal$CDK1, "CDK1")
+f <- plotremove(removal$MYC, "MYC")
+
+plot_grid(a + theme(axis.text.x = element_blank()),
+          d + theme(axis.text.x = element_blank()),
+          e + theme(axis.text.x = element_blank()), nrow = 1)
+
+timing <- widecountdata %>% 
+    filter( !treatment %in% c("m.inc.d3", "m.inc.d9", "m.inc.d17","m.n2", 
+                              "control", "bldg" , "lay", "n5", "n9"))
+
+plottiming <- function(whichgene, myylab){
+  p <-  timing  %>% 
+  ggplot(aes(x = treatment, y = whichgene, fill = treatment,  color = sex)) +
+    geom_boxplot() + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none",
+          strip.text = element_blank()) +
+    scale_fill_manual(values = colorscharmaip) +
+    scale_color_manual(values = sexcolors) +
+    labs(y = NULL, subtitle = myylab, x = NULL) 
+  return(p)
+}
+
+a <- plottiming(timing$PRL, "PRL")
+b <- plottiming(timing$FOSL2, "FOSL2")
+c <- plottiming(timing$COX1, "COX1")
+d <- plottiming(timing$BRCA1, "BRCA1")
+e <- plottiming(timing$CDK1, "CDK1")
+f <- plottiming(timing$MYC, "MYC")
+
+plot_grid(a + theme(axis.text.x = element_blank()),
+          d + theme(axis.text.x = element_blank()),
+          e + theme(axis.text.x = element_blank()), nrow = 1)
+
 ```

### 2019-11-13 analysis/15_networks.Rmd
new plots with manipluation and candidate genes

@@
 # genes from website shown to be associate with oxt and prl
+PRLgenes <- c("PRL", "GNRH1", "TRH", "POU1F1", "POMC", "TSHB",
+              "FSHB", "GH", "PRLR", "JAK2", "SOCS1", "CISH")
 OXTgenes <- c("OXT",  "AVPR1A", "AVPR1B", "TRH", 
               "HCRT", "GNRH1", "CRH", "POMC", "MC4R", "AVPR2")
-PRLgenes <- c("PRL", "GNRH1", "TRH", "POU1F1", "POMC", "TSHB",
-              "FSHB", "GH", "PRLR", "JAK2")
-OXTPRLgenes <- c(OXTgenes, PRLgenes)
-head(OXTPRLgenes)
+CISHgenes <- c("CISH", "SOCS1", "SOCS2", "SOCS4", "SOCS5", "SOCS6")
+
+PRLnOther <- c(PRLgenes,OXTgenes, CISHgenes)
 
@@
 
+# cancer genes from PRL list from  http://bioinformatics.sdstate.edu/go/
+# shiny go for Gene Ontology Enrichment
+cancergenes <- c("E2F1", "MYC" ,"PLK1", "BRCA1" ,"AURKA", "PRL")
+cellcyclegenes <- c("E2F1", "MAD2L1", "MCM2", "PLK1", "MCM10", "BUB1", 
+                    "CDC45", "CCNB3", "CCNA2", 
+                      "CDK1", "CDC6", "CDC20")
+replicationanddamage <- c("MCM2"  , "CDT1" , " CDC45 " , "CDC6", 
+                          "E2F1" , " MYC" , "BRCA1" , "CCNB3 " , "CDK1")
+
+## plot genes
+
 longdata <- widecountdata %>% 
-            select(V1:study, OXTPRLgenes, PRL_associated) %>%
-            gather(gene, expression, OXT:UBE2C)
+            select(V1:study, PRLnOther, PRL_associated) %>%
+            gather(gene, expression, PRL:UBE2C)
 head(longdata)
 
-plotnetworkgenes <- function(whichgenes){
+plotnetworkgenes <- function(whichgenes, mysubtitle){
   longdata %>%
   filter(gene %in% whichgenes) %>%
-  ggplot(aes(x = treatment, y = expression)) +
+  ggplot(aes(x = as.numeric(treatment), y = expression)) +
         geom_smooth(aes(colour = sex)) +
-    geom_boxplot(aes(outlier.color = treatment, 
-                     fill = treatment, alpha = sex)) +
+    geom_boxplot(aes(fill = treatment, alpha = sex)) +
     mytheme() +
@@
     scale_color_manual(values = sexcolors) +
-    labs(x = NULL) +
+    labs(x = NULL, subtitle = mysubtitle) +
     scale_alpha_manual(values = c(0.75,1)) +
@@
 
-plotnetworkgenes(OXTgenes)
-plotnetworkgenes(PRLgenes)
-plotnetworkgenes(PRL_associated[85:93])
+plotnetworkgenes(PRLgenes, "Known PRL associates") 
+plotnetworkgenes(PRL_associated[85:93],  "in WGCNA module with PRL")
+plotnetworkgenes(cancergenes, "genes in Breast Cancer WikiPathway")
+plotnetworkgenes(cellcyclegenes, "genes in Cell Cycle WikiPathways ")
+
+plotnetworkgenes(replicationanddamage, 
+                 "genes in DNA Replication & Damage Response WikiPathway")
 
@@
 forcorrdata <- widecountdata %>% 
-            select(V1, OXTPRLgenes)  %>% 
+            select(cancergenes)  %>% 
             as.data.frame()
@@
 x <- widecountdata %>% 
-            select(PRL_associated,OXTPRLgenes) %>%
+            select(PRL_associated,PRLnOther) %>%
        correlate()     # Create correlation data frame (cor_df)

### 2019-11-13 analysis/07_hormones.Rmd
now with axis text

@@
     mytheme() +
-    theme(axis.text.x =  element_blank(),
+    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
           legend.position = "none") +
@@
     scale_alpha_manual(values = c(0.75,1)) +
-    scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) 
+    scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
+                       labels = c( "control", "bldg", "lay", "inc.d3", 
+                                   "inc.d9", "inc.d17", "hatch", "n5", "n9"))
   }
@@
 ```{r sexsteroids}
-b <- hormonecharplot("cort", "CORT (ng/mL)")
-c <- hormonecharplot("progesterone", "PROG (ng/mL)") 
+hormonecharplot("cort", "CORT (ng/mL)")
+hormonecharplot("progesterone", "PROG (ng/mL)") 
 d1 <- hormonecharplot("estradiol", "E (ng/mL)")
 d2 <- hormonecharplot("testosterone", "T (ng/mL)")
-d <- plot_grid(d1,d2, nrow = 1)
-bcd <- plot_grid(b ,c, d, ncol = 1)
-bcd
+plot_grid(d1,d2, nrow = 1)
+
 ```
@@
 
-```
-
 hormonemanipSteroids <- function(myhormone, myylab, myymax){

### 2019-11-22 analysis/03_DESeq2_characterization.Rmd
so many gene lists
now with go terms

@@
 library(ggimage)
+library(knitr)
+library(kableExtra)
+
+library(BiocParallel)
+register(MulticoreParam(6))
 
 # load custom functions  
+source("../R/themes.R") 
 source("../R/functions.R")
@@
 summary(c.colData[c(7,3,4,5,8)])
+
+geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+head(geneinfo)
 ```
@@
 
-## plot genes in a  PRL WGCNA module
-
-```{r wgcna, eval = T}
-plotWGCNAcandidates(dds.female_pituitary, colDataFP, "female pituitary")
-plotWGCNAcandidates(dds.male_pituitary,colDataMP, "male pituitary")
-
-```
-
 ## correlation heat maps
@@
 ```
+
+## DEGs
+
+```{r createDEGs}
+
+createDEGdf <- function(mydds, whichfactor, up, down){
+  res <- results(mydds, contrast =c(whichfactor, up, down),
+                 independentFiltering = T, alpha = 0.1)
+   data <- data.frame(entrezid = row.names(res),
+                     padj = res$padj, 
+                     logpadj = -log10(res$padj),
+                     lfc = res$log2FoldChange)
+  data <- na.omit(data)
+
+  data <- data %>%
+    dplyr::mutate(direction = ifelse(data$lfc > 0 & data$padj < 0.1, 
+                                     yes = up, no = ifelse(data$lfc < 0 & data$padj < 0.1, 
+                                                 yes = down, no = "NS")))
+  data$direction <- factor(data$direction, levels = c(down, "NS", up))
+
+  data <- left_join(data, geneinfo) %>%
+    mutate(gene = Name) %>%
+    select(gene, lfc, padj, logpadj, direction) %>%
+    arrange(desc(lfc))
+
+  DEGs <- data %>% filter(direction != "NS")
+
+  myfilename = paste("../results/03_DEGs", down,up, "csv", sep = ".")
+
+  # write dataframe of only significant genes
+  write.csv(DEGs, myfilename, row.names = F)
+
+  # return data frome with all data, included NS genes
+  return(data)
+}
+
+control.bldg <- createDEGdf(dds.female_pituitary, "treatment", "bldg", "control")
+bldg.lay <- createDEGdf(dds.female_pituitary, "treatment", "lay", "bldg")
+lay.inc.d3 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d3", "lay") 
+inc.d3.inc.d9 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d9", "inc.d3") 
+inc.d9.inc.d17 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d17", "inc.d9")
+inc.d17.hatch <- createDEGdf(dds.female_pituitary, "treatment", "hatch", "inc.d17") 
+hatch.n5 <- createDEGdf(dds.female_pituitary, "treatment", "n5", "hatch") 
+n5.n9 <- createDEGdf(dds.female_pituitary, "treatment", "n9", "n5") 
+```
+
+```{r plotDEGs}
+plot.volcano <- function(data, whichfactor, up, down, mycolors){
+
+  volcano <- data %>%
+    ggplot(aes(x = lfc, y = logpadj)) + 
+    geom_point(aes(color = direction), size = 1, alpha = 0.75, na.rm = T) + 
+     theme_B3() +
+    scale_color_manual(values = mycolors,
+                       name = " ",
+                       drop = FALSE) +
+    ylim(c(0,25)) +  
+    xlim(c(-8,8)) +
+    labs(y = "-log10(p)", x = "log fold change")  +
+    theme(legend.position = c(.5, .95),
+          legend.direction = "horizontal",
+          legend.spacing.x = unit(-0.1, 'cm'),
+          legend.margin=margin(t=-0.25, r=0, b=0, l=0, unit="cm"),
+          panel.grid = element_blank()) 
+  return(volcano)
+}
+
+a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
+b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
+c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
+d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
+g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
+h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
+
+plot_grid(a + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
+          b + theme(axis.title = element_blank(), axis.text = element_blank()),
+          c + theme(axis.title = element_blank(), axis.text = element_blank()),
+          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          e ,
+          f + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          g + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          h + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          nrow = 2, rel_widths = c(0.25,0.2,0.2,0.2), rel_heights = c(0.45,0.55))
+
+```
+
+## Go terms
+
+```{r GOterms}
+reproduction <- read.table("../results/goterms/reproduction.txt", sep = "\t", row.names = NULL,  fill=TRUE)
+reproduction$GO <- "reproduction, GO:0000003"
+
+parentalbehavior <- read.table("../results/goterms/parentalbehavior.txt", sep = "\t", row.names = NULL,  fill=TRUE)
+parentalbehavior$GO <- "parental behavior, GO:0060746"
+
+behavior <- read.table("../results/goterms/parentalbehavior.txt", sep = "\t", row.names = NULL,  fill=TRUE)
+behavior$GO <- "behavior, GO:0007610"
+
+systemdevelopment <- read.table("../results/goterms/systemdevelopment.txt", sep = "\t", row.names = NULL,  fill=TRUE)
+systemdevelopment$GO <- "system development, GO:0048731"
+
+GOterms <- rbind(parentalbehavior, reproduction, behavior, systemdevelopment)
+
+GOterms <- GOterms %>%
+  dplyr::mutate(gene = toupper(MGI.Gene.Marker.ID)) %>% 
+  dplyr::select(gene, GO) %>% 
+  dplyr::distinct(gene, GO) %>% 
+ group_by(gene)
+
+head(GOterms)
+
+returnGOgenes <- function(whichGOterms, data){
+
+  df <- inner_join(whichGOterms, data) %>% filter(direction != "NS") %>%  select(gene, GO) %>% 
+  group_by(GO) %>% summarize(genes = str_c(gene, collapse = ", "))
+  return(df)
+}
+
+dt1 <- returnGOgenes(GOterms,control.bldg) 
+dt2 <- returnGOgenes(GOterms,bldg.lay) 
+dt3 <- returnGOgenes(GOterms,lay.inc.d3)  
+dt4 <- returnGOgenes(GOterms,inc.d3.inc.d9)  
+dt5 <- returnGOgenes(GOterms,inc.d9.inc.d17)  
+dt6 <- returnGOgenes(GOterms,inc.d17.hatch)  
+dt7 <- returnGOgenes(GOterms,hatch.n5)  
+dt8 <- returnGOgenes(GOterms,n5.n9)
+
+kable(dt1)
+kable(dt2) 
+kable(dt3) 
+kable(dt4) 
+kable(dt5) 
+kable(dt6) 
+kable(dt7) 
+kable(dt8)
+```

### 2019-11-22 analysis/07_hormones.Rmd
now with go terms

@@
     scale_alpha_manual(values = c(0.75,1)) +
-    scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
     scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),

### 2019-11-22 R/themes.R

@@ -95,3 +95,21 @@ colorstissue <- c("hypothalamus" = "#d95f02",
 
 sexcolors <- c("female" = "#969696", "male" = "#525252")
 
+colorsvolcano <-  c("control" = "#F8766D", 
+                     "bldg"= "#D39200", 
+                     "lay"= "#93AA00", 
+                     "inc.d3"= "#00BA38", 
+                     "inc.d9"= "#00C19F", 
+                     "inc.d17"= "#00B9E3", 
+                     "hatch"= "#619CFF", 
+                     "n5"= "#DB72FB", 
+                     "n9"= "#FF61C3",
+                     "m.inc.d3" = "#CDCDCD", 
+                     "m.inc.d8" = "#6DA9D3", 
+                     "m.inc.d9" = "#959595", 
+                     "m.inc.d17" = "#626262",
+                     "m.n2" = "#262625", 
+                     "prolong" = "#1B86B9" , 
+                     "extend" = "#01588C",
+                    "NS" = "#bdbdbd")
+

### 2019-11-22 analysis/03_DESeq2_characterization.Rmd
table formatting

@@
 dt1 <- returnGOgenes(GOterms,control.bldg) 
+kable(dt1)
+
 dt2 <- returnGOgenes(GOterms,bldg.lay) 
+kable(dt2)
+
 dt3 <- returnGOgenes(GOterms,lay.inc.d3)  
+kable(dt3)
+
 dt4 <- returnGOgenes(GOterms,inc.d3.inc.d9)  
+kable(dt4)
+
 dt5 <- returnGOgenes(GOterms,inc.d9.inc.d17)  
+kable(dt5)
+
 dt6 <- returnGOgenes(GOterms,inc.d17.hatch)  
+kable(dt6)
+
 dt7 <- returnGOgenes(GOterms,hatch.n5)  
+kable(dt7)
+
 dt8 <- returnGOgenes(GOterms,n5.n9) 
+kable(dt8)
 
-kable(dt1)
-kable(dt2) 
-kable(dt3) 
-kable(dt4) 
-kable(dt5) 
-kable(dt6) 
-kable(dt7) 
-kable(dt8)
 ```

### 2019-11-22 analysis/03_DESeq2_characterization.Rmd
one more go term
now with endocrine

@@
 
-GOterms <- rbind(parentalbehavior, reproduction, behavior, systemdevelopment)
+endocrineprocess <- read.table("../results/goterms/endocrineprocess.txt", sep = "\t", row.names = NULL,  fill=TRUE)
+endocrineprocess$GO <- "endocrine process, GO:0050886"
+
+GOterms <- rbind(parentalbehavior, reproduction, behavior, systemdevelopment, endocrineprocess)
 

### 2019-11-22 analysis/03_DESeq2_characterization.Rmd
now with some IEGs

@@
 
-GOterms <- rbind(parentalbehavior, reproduction, behavior, systemdevelopment, endocrineprocess)
+responsetostimulus <- read.table("../results/goterms/responsetostimulus.txt", sep = "\t", row.names = NULL,  fill=TRUE)
+responsetostimulus$GO <- "response to external stimulus, GO:0009605"
+
+GOterms <- rbind(parentalbehavior, reproduction, behavior, 
+                 systemdevelopment, endocrineprocess, responsetostimulus)
 
@@
 
-returnGOgenes <- function(whichGOterms, data){
+returnGOgenes <- function(whichGOterms, data, mycomparison){
 
-  df <- inner_join(whichGOterms, data) %>% filter(direction != "NS") %>%  select(gene, GO) %>% 
-  group_by(GO) %>% summarize(genes = str_c(gene, collapse = ", "))
+  df <- inner_join(whichGOterms, data) %>% 
+    filter(direction != "NS") %>%  
+    select(gene, GO) %>% 
+    group_by(GO) %>% 
+    summarize(genes = str_c(gene, collapse = ", "))  %>% 
+    mutate(comparison = mycomparison) %>% 
+    select(comparison, GO, genes)
+
   return(df)
@@
 
-dt1 <- returnGOgenes(GOterms,control.bldg) 
-kable(dt1)
-
-dt2 <- returnGOgenes(GOterms,bldg.lay) 
-kable(dt2)
-
-dt3 <- returnGOgenes(GOterms,lay.inc.d3)  
-kable(dt3)
-
-dt4 <- returnGOgenes(GOterms,inc.d3.inc.d9)  
-kable(dt4)
-
-dt5 <- returnGOgenes(GOterms,inc.d9.inc.d17)  
+dt1 <- returnGOgenes(GOterms,control.bldg, "control v bldg") 
+dt2 <- returnGOgenes(GOterms,bldg.lay , "bldg v lay") 
+dt3 <- returnGOgenes(GOterms,lay.inc.d3 , "lay v inc d3")  
+dt4 <- returnGOgenes(GOterms,inc.d3.inc.d9 , "inc d3 v inc d9")  
+dt5 <- returnGOgenes(GOterms,inc.d9.inc.d17 , "inc d9 v inc d17")  
 kable(dt5)
 
-dt6 <- returnGOgenes(GOterms,inc.d17.hatch)  
-kable(dt6)
-
-dt7 <- returnGOgenes(GOterms,hatch.n5)  
+dt6 <- returnGOgenes(GOterms,inc.d17.hatch, "inc d17 v hatch")  
+dt7 <- returnGOgenes(GOterms,hatch.n5, "hatch v n5")  
 kable(dt7)
 
-dt8 <- returnGOgenes(GOterms,n5.n9) 
-kable(dt8)
+dt8 <- returnGOgenes(GOterms,n5.n9, "n5 v n9")
 
+alldt <- rbind(dt1,dt2,dt3,dt4,dt5,dt6,dt7,dt8)
+alldt
+
+write.csv(alldt, "../results/03_GO_femalepituitary.csv")
 ```

### 2019-11-23 analysis/03_DESeq2_characterization.Rmd
new icon labeles plots

@@
 library(kableExtra)
+
+library(png)
+library(grid)
+library(ggimage)
+library(ggpubr)
 
@@
   DEGs <- data %>% filter(direction != "NS")
+  print(str(DEGs))
 
@@
     geom_point(aes(color = direction), size = 1, alpha = 0.75, na.rm = T) + 
-     theme_B3() +
+    theme_minimal() +
     scale_color_manual(values = mycolors,
@@
                        drop = FALSE) +
-    ylim(c(0,25)) +  
+    ylim(c(0,30)) +  
     xlim(c(-8,8)) +
-    labs(y = "-log10(p)", x = "log fold change")  +
-    theme(legend.position = c(.5, .95),
+    labs(y = "-log10 (p-value)", x = " ")  +
+    theme(legend.position = "none",
           legend.direction = "horizontal",
@@
 
-plot_grid(a + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
+females <- plot_grid(a + 
+                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
+                       labs(y = "female pituitary \n -log10 (p-value)"),
           b + theme(axis.title = element_blank(), axis.text = element_blank()),
@@
           d + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          e ,
-          f + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          g + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          h + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
-          nrow = 2, rel_widths = c(0.25,0.2,0.2,0.2), rel_heights = c(0.45,0.55))
-
+          e + theme(axis.title = element_blank(), axis.text = element_blank()),
+          f + theme(axis.title = element_blank(), axis.text = element_blank()),
+          g + theme(axis.title = element_blank(), axis.text = element_blank()),
+          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
+females 
 ```
@@
 ```
+
+## males and females
+
+```{r createDEGsmales}
+control.bldg2 <- createDEGdf(dds.male_pituitary, "treatment", "bldg", "control")
+bldg.lay2 <- createDEGdf(dds.male_pituitary, "treatment", "lay", "bldg")
+lay.inc.d32 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d3", "lay") 
+inc.d3.inc.d92 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d9", "inc.d3") 
+inc.d9.inc.d172 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d17", "inc.d9")
+inc.d17.hatch2 <- createDEGdf(dds.male_pituitary, "treatment", "hatch", "inc.d17") 
+hatch.n52 <- createDEGdf(dds.male_pituitary, "treatment", "n5", "hatch") 
+n5.n92 <- createDEGdf(dds.male_pituitary, "treatment", "n9", "n5")
+
+```
+
+```{r plotDEGsmales}
+a2 <- plot.volcano(control.bldg2, "treatment", "bldg", "control", colorsvolcano)
+b2 <- plot.volcano(bldg.lay2, "treatment", "lay", "bldg", colorsvolcano)
+c2 <- plot.volcano(lay.inc.d32, "treatment", "inc.d3", "lay", colorsvolcano) 
+d2 <- plot.volcano(inc.d3.inc.d92, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+e2 <- plot.volcano(inc.d9.inc.d172, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+f2 <- plot.volcano(inc.d17.hatch2, "treatment", "hatch", "inc.d17", colorsvolcano) 
+g2 <- plot.volcano(hatch.n52, "treatment", "n5", "hatch", colorsvolcano) 
+h2 <- plot.volcano(n5.n92, "treatment", "n9", "n5", colorsvolcano)
+
+males <- plot_grid(a2 + labs(y = "male pituitary \n -log10 (p-value)"),
+          b2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          c2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          d2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          e2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          f2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          g2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          h2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
+males
+
+femalesNmales <- plot_grid(females,males, nrow = 2, 
+          rel_heights = c(0.45,0.55))
+
+background <- png::readPNG("../figures/images/DoveParentsRNAseq_charicons.png")
+background
+
+icons <- ggdraw() +  draw_image(background, scale = 1)
+
+plot_grid(icons, females, males, nrow = 3, rel_heights = c(1,1,1.5))
+
+```
+
+### tissue levels, ignoring sex differences, sort of
+
+```{r gonaddfs, eval=T}
+
+dds.gonads <- subsetDESeq2(c.colData,  c.countData, c("female_gonad","male_gonad"))
+
+control.bldg <- createDEGdf(dds.gonads, "treatment", "bldg", "control")
+bldg.lay <- createDEGdf(dds.gonads, "treatment", "lay", "bldg")
+lay.inc.d3 <- createDEGdf(dds.gonads, "treatment", "inc.d3", "lay") 
+inc.d3.inc.d9 <- createDEGdf(dds.gonads) 
+inc.d9.inc.d17 <- createDEGdf(dds.gonads, "treatment", "inc.d17", "inc.d9")
+inc.d17.hatch <- createDEGdf(dds.gonads, "treatment", "hatch", "inc.d17") 
+hatch.n5 <- createDEGdf(dds.gonads, "treatment", "n5", "hatch") 
+n5.n9 <- createDEGdf(dds.gonads, "treatment", "n9", "n5") 
+```
+
+```{r gonplot, eval=T}
+a2 <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
+b2 <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
+c2 <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
+d2 <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+e2 <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+f2 <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
+g2 <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
+h2 <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
+
+gonads <- plot_grid(a2 + labs(y = "gonads \n -log10 (p-value)"),
+          b2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          c2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          d2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
+            labs(x = "log fold change"), 
+          e2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          f2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          g2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          h2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
+gonads
+```
+
+```{r hypdfs, eval=T}
+dds.hypothalamus <- subsetDESeq2(c.colData,  c.countData, c("female_hypothalamus","male_hypothalamus"))
+
+control.bldg <- createDEGdf(dds.hypothalamus, "treatment", "bldg", "control")
+bldg.lay <- createDEGdf(dds.hypothalamus, "treatment", "lay", "bldg")
+lay.inc.d3 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d3", "lay") 
+inc.d3.inc.d9 <- createDEGdf(dds.hypothalamus) 
+inc.d9.inc.d17 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d17", "inc.d9")
+inc.d17.hatch <- createDEGdf(dds.hypothalamus, "treatment", "hatch", "inc.d17") 
+hatch.n5 <- createDEGdf(dds.hypothalamus, "treatment", "n5", "hatch") 
+n5.n9 <- createDEGdf(dds.hypothalamus, "treatment", "n9", "n5") 
+```
+
+```{r hypplot, eval = T}
+a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
+b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
+c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
+d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
+g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
+h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
+
+hypothalamus <- plot_grid(a + 
+                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
+                       labs(y = "hypothalamus \n -log10 (p-value)"),
+          b + theme(axis.title = element_blank(), axis.text = element_blank()),
+          c + theme(axis.title = element_blank(), axis.text = element_blank()),
+          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          e + theme(axis.title = element_blank(), axis.text = element_blank()),
+          f + theme(axis.title = element_blank(), axis.text = element_blank()),
+          g + theme(axis.title = element_blank(), axis.text = element_blank()),
+          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
+hypothalamus
+
+plot_grid(nrow = 2)
+
+plot_grid(icons, females, males, hypothalamus, gonads, nrow = 5, rel_heights = c(1,1,1,1,1.5))
+
+```

### 2019-11-23 analysis/03_DESeq2_characterization.Rmd
now with volcanos forall tissues

--- 2019-11-22 analysis/03_DESeq2_characterization.Rmd
@@
                        drop = FALSE) +
-    ylim(c(0,25)) +  
+    ylim(c(0,30)) +  
     xlim(c(-8,8)) +
-    labs(y = "-log10(p)", x = "log fold change")  +
-    theme(legend.position = c(.5, .95),
+    labs(y = "-log10(p)", x = " ")  +
+    theme(legend.position = "none",
           legend.direction = "horizontal",
@@
 
-plot_grid(a + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
+females <- plot_grid(a + 
+                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
+                       labs(y = "female pituitary \n -log10(p)"),
           b + theme(axis.title = element_blank(), axis.text = element_blank()),
@@
 ```
+
+## males and females
+
+```{r createDEGsmales}
+control.bldg2 <- createDEGdf(dds.male_pituitary, "treatment", "bldg", "control")
+bldg.lay2 <- createDEGdf(dds.male_pituitary, "treatment", "lay", "bldg")
+lay.inc.d32 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d3", "lay") 
+inc.d3.inc.d92 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d9", "inc.d3") 
+inc.d9.inc.d172 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d17", "inc.d9")
+inc.d17.hatch2 <- createDEGdf(dds.male_pituitary, "treatment", "hatch", "inc.d17") 
+hatch.n52 <- createDEGdf(dds.male_pituitary, "treatment", "n5", "hatch") 
+n5.n92 <- createDEGdf(dds.male_pituitary, "treatment", "n9", "n5")
+
+```
+
+```{r plotDEGsmales}
+a2 <- plot.volcano(control.bldg2, "treatment", "bldg", "control", colorsvolcano)
+b2 <- plot.volcano(bldg.lay2, "treatment", "lay", "bldg", colorsvolcano)
+c2 <- plot.volcano(lay.inc.d32, "treatment", "inc.d3", "lay", colorsvolcano) 
+d2 <- plot.volcano(inc.d3.inc.d92, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+e2 <- plot.volcano(inc.d9.inc.d172, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+f2 <- plot.volcano(inc.d17.hatch2, "treatment", "hatch", "inc.d17", colorsvolcano) 
+g2 <- plot.volcano(hatch.n52, "treatment", "n5", "hatch", colorsvolcano) 
+h2 <- plot.volcano(n5.n92, "treatment", "n9", "n5", colorsvolcano)
+
+males <- plot_grid(a2 + labs(y = "male pituitary \n -log10(p)") + theme(axis.text.x = element_blank()),
+          b2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
+          c2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
+          d2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
+          e2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
+          f2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
+          g2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
+          h2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
+          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
+males
+```
+
+### tissue levels, ignoring sex differences, sort of
+
+```{r gonaddfs, eval=T}
+
+dds.gonads <- subsetDESeq2(c.colData,  c.countData, c("female_gonad","male_gonad"))
+```
+
+```{r gonaddfs2, eval=T}
+control.bldg <- createDEGdf(dds.gonads, "treatment", "bldg", "control")
+bldg.lay <- createDEGdf(dds.gonads, "treatment", "lay", "bldg")
+lay.inc.d3 <- createDEGdf(dds.gonads, "treatment", "inc.d3", "lay") 
+inc.d3.inc.d9 <- createDEGdf(dds.gonads, "treatment", "inc.d9", "inc.d3") 
+inc.d9.inc.d17 <- createDEGdf(dds.gonads, "treatment", "inc.d17", "inc.d9")
+inc.d17.hatch <- createDEGdf(dds.gonads, "treatment", "hatch", "inc.d17") 
+hatch.n5 <- createDEGdf(dds.gonads, "treatment", "n5", "hatch") 
+n5.n9 <- createDEGdf(dds.gonads, "treatment", "n9", "n5") 
+```
+
+```{r gonplot, eval=T}
+a2 <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
+b2 <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
+c2 <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
+d2 <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+e2 <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+f2 <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
+g2 <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
+h2 <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
+
+gonads <- plot_grid(a2 + labs(y = "gonads \n -log10(p)"),
+          b2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          c2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          d2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
+            labs(x = "log fold change"), 
+          e2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          f2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          g2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          h2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
+gonads
+```
+
+```{r hypdfs, eval=T}
+dds.hypothalamus <- subsetDESeq2(c.colData,  c.countData, c("female_hypothalamus","male_hypothalamus"))
+```
+
+```{r hypdfs2, eval=T}
+control.bldg <- createDEGdf(dds.hypothalamus, "treatment", "bldg", "control")
+bldg.lay <- createDEGdf(dds.hypothalamus, "treatment", "lay", "bldg")
+lay.inc.d3 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d3", "lay") 
+inc.d3.inc.d9 <- createDEGdf(dds.hypothalamus,  "treatment", "inc.d9", "inc.d3") 
+inc.d9.inc.d17 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d17", "inc.d9")
+inc.d17.hatch <- createDEGdf(dds.hypothalamus, "treatment", "hatch", "inc.d17") 
+hatch.n5 <- createDEGdf(dds.hypothalamus, "treatment", "n5", "hatch") 
+n5.n9 <- createDEGdf(dds.hypothalamus, "treatment", "n9", "n5") 
+```
+
+```{r hypplot, eval = T}
+a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
+b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
+c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
+d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
+g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
+h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
+
+hypothalamus <- plot_grid(a + 
+                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
+                       labs(y = "hypothalamus \n -log10(p)"),
+          b + theme(axis.title = element_blank(), axis.text = element_blank()),
+          c + theme(axis.title = element_blank(), axis.text = element_blank()),
+          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          e + theme(axis.title = element_blank(), axis.text = element_blank()),
+          f + theme(axis.title = element_blank(), axis.text = element_blank()),
+          g + theme(axis.title = element_blank(), axis.text = element_blank()),
+          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          nrow = 1,  rel_widths = c(1.75,1,1,1,1,1,1,1,1))
+hypothalamus
+
+```
+
+```{r allvolcanos, fig.height=8}
+background <- png::readPNG("../figures/images/DoveParentsRNAseq_charicons.png")
+icons <- ggdraw() +  draw_image(background, scale = 1)
+
+plot_grid(icons, hypothalamus, females, males, gonads, nrow = 5, rel_heights = c(1,1,1,1,1.25))
+
+```

### 2019-11-23 analysis/11_pcalda.Rmd
new pca

@@
                          color = colData$treatment, 
-                         shape = colData$tissue,
-                         alpha = colData$sex)) +
-  geom_point(size = 2)  + theme_minimal(base_size = 6)  +
+                         shape = colData$tissue)) +
+  geom_point(size = 3)  + theme_minimal(base_size = 6)  +
   mytheme() +
@@
 dev.off()
+```
+
+```{r pca2}
+plot_grid(a + theme(legend.position = "none"),
+          b + theme(axis.text = element_blank()), 
+          nrow = 1, rel_widths = c(0.7,0.3))
 ```

### 2019-11-23 analysis/03_DESeq2_characterization.Rmd
new files and plots with shapes

@@
 c.colData$treatment <- factor(c.colData$treatment, levels = 
-                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
+                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17",
+                                "hatch", "n5", "n9"))
+c.colData$tissue <- factor(c.colData$tissue, levels = 
+                              c("hypothalamus",  "pituitary", "gonad"))
+
 levels(c.colData$treatment)
@@
 
-createDEGdf <- function(mydds, whichfactor, up, down){
+createDEGdf <- function(mydds, whichfactor, up, down, mytissue){
   res <- results(mydds, contrast =c(whichfactor, up, down),
@@
                      logpadj = -log10(res$padj),
-                     lfc = res$log2FoldChange)
+                     lfc = res$log2FoldChange,
+                     tissue = mytissue)
   data <- na.omit(data)
@@
     mutate(gene = Name) %>%
-    select(gene, lfc, padj, logpadj, direction) %>%
+    select(gene, lfc, padj, logpadj, direction, tissue) %>%
     arrange(desc(lfc))
@@
 
-  myfilename = paste("../results/03_DEGs", down,up, "csv", sep = ".")
+  myfilename = paste("../results/03_DEGs", mytissue, down,up, "csv", sep = ".")
 
@@
 
-control.bldg <- createDEGdf(dds.female_pituitary, "treatment", "bldg", "control")
-bldg.lay <- createDEGdf(dds.female_pituitary, "treatment", "lay", "bldg")
-lay.inc.d3 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d3", "lay") 
-inc.d3.inc.d9 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d9", "inc.d3") 
-inc.d9.inc.d17 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d17", "inc.d9")
-inc.d17.hatch <- createDEGdf(dds.female_pituitary, "treatment", "hatch", "inc.d17") 
-hatch.n5 <- createDEGdf(dds.female_pituitary, "treatment", "n5", "hatch") 
-n5.n9 <- createDEGdf(dds.female_pituitary, "treatment", "n9", "n5") 
+control.bldg <- createDEGdf(dds.female_pituitary, "treatment", "bldg", "control", "pituitary")
+bldg.lay <- createDEGdf(dds.female_pituitary, "treatment", "lay", "bldg", "pituitary")
+lay.inc.d3 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d3", "lay", "pituitary") 
+inc.d3.inc.d9 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d9", "inc.d3", "pituitary") 
+inc.d9.inc.d17 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
+inc.d17.hatch <- createDEGdf(dds.female_pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
+hatch.n5 <- createDEGdf(dds.female_pituitary, "treatment", "n5", "hatch", "pituitary") 
+n5.n9 <- createDEGdf(dds.female_pituitary, "treatment", "n9", "n5", "pituitary") 
 ```
@@
     ggplot(aes(x = lfc, y = logpadj)) + 
-    geom_point(aes(color = direction), size = 1, alpha = 0.75, na.rm = T) + 
+    geom_point(aes(color = direction, shape = tissue), size = 1, alpha = 0.75, na.rm = T) + 
     theme_minimal() +
@@
           legend.margin=margin(t=-0.25, r=0, b=0, l=0, unit="cm"),
-          panel.grid = element_blank()) 
+          panel.grid = element_blank()) +
+    scale_shape_manual(values = myshapes)
   return(volcano)
@@
 ```{r createDEGsmales}
-control.bldg2 <- createDEGdf(dds.male_pituitary, "treatment", "bldg", "control")
-bldg.lay2 <- createDEGdf(dds.male_pituitary, "treatment", "lay", "bldg")
-lay.inc.d32 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d3", "lay") 
-inc.d3.inc.d92 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d9", "inc.d3") 
-inc.d9.inc.d172 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d17", "inc.d9")
-inc.d17.hatch2 <- createDEGdf(dds.male_pituitary, "treatment", "hatch", "inc.d17") 
-hatch.n52 <- createDEGdf(dds.male_pituitary, "treatment", "n5", "hatch") 
-n5.n92 <- createDEGdf(dds.male_pituitary, "treatment", "n9", "n5")
+control.bldg2 <- createDEGdf(dds.male_pituitary, "treatment", "bldg", "control", "pituitary")
+bldg.lay2 <- createDEGdf(dds.male_pituitary, "treatment", "lay", "bldg", "pituitary")
+lay.inc.d32 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d3", "lay", "pituitary") 
+inc.d3.inc.d92 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d9", "inc.d3", "pituitary") 
+inc.d9.inc.d172 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
+inc.d17.hatch2 <- createDEGdf(dds.male_pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
+hatch.n52 <- createDEGdf(dds.male_pituitary, "treatment", "n5", "hatch", "pituitary") 
+n5.n92 <- createDEGdf(dds.male_pituitary, "treatment", "n9", "n5", "pituitary")
 
@@
 males
+
 ```
@@
 ```{r gonaddfs2, eval=T}
-control.bldg <- createDEGdf(dds.gonads, "treatment", "bldg", "control")
-bldg.lay <- createDEGdf(dds.gonads, "treatment", "lay", "bldg")
-lay.inc.d3 <- createDEGdf(dds.gonads, "treatment", "inc.d3", "lay") 
-inc.d3.inc.d9 <- createDEGdf(dds.gonads, "treatment", "inc.d9", "inc.d3") 
-inc.d9.inc.d17 <- createDEGdf(dds.gonads, "treatment", "inc.d17", "inc.d9")
-inc.d17.hatch <- createDEGdf(dds.gonads, "treatment", "hatch", "inc.d17") 
-hatch.n5 <- createDEGdf(dds.gonads, "treatment", "n5", "hatch") 
-n5.n9 <- createDEGdf(dds.gonads, "treatment", "n9", "n5") 
+control.bldg <- createDEGdf(dds.gonads, "treatment", "bldg", "control",  "gonads")
+bldg.lay <- createDEGdf(dds.gonads, "treatment", "lay", "bldg",  "gonads")
+lay.inc.d3 <- createDEGdf(dds.gonads, "treatment", "inc.d3", "lay",  "gonads") 
+inc.d3.inc.d9 <- createDEGdf(dds.gonads, "treatment", "inc.d9", "inc.d3",  "gonads") 
+inc.d9.inc.d17 <- createDEGdf(dds.gonads, "treatment", "inc.d17", "inc.d9",  "gonads")
+inc.d17.hatch <- createDEGdf(dds.gonads, "treatment", "hatch", "inc.d17",  "gonads") 
+hatch.n5 <- createDEGdf(dds.gonads, "treatment", "n5", "hatch",  "gonads") 
+n5.n9 <- createDEGdf(dds.gonads, "treatment", "n9", "n5",  "gonads") 
 ```
@@
 gonads
+
 ```
@@
 ```{r hypdfs2, eval=T}
-control.bldg <- createDEGdf(dds.hypothalamus, "treatment", "bldg", "control")
-bldg.lay <- createDEGdf(dds.hypothalamus, "treatment", "lay", "bldg")
-lay.inc.d3 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d3", "lay") 
-inc.d3.inc.d9 <- createDEGdf(dds.hypothalamus,  "treatment", "inc.d9", "inc.d3") 
-inc.d9.inc.d17 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d17", "inc.d9")
-inc.d17.hatch <- createDEGdf(dds.hypothalamus, "treatment", "hatch", "inc.d17") 
-hatch.n5 <- createDEGdf(dds.hypothalamus, "treatment", "n5", "hatch") 
-n5.n9 <- createDEGdf(dds.hypothalamus, "treatment", "n9", "n5") 
+control.bldg <- createDEGdf(dds.hypothalamus, "treatment", "bldg", "control" ,  "hypothalamus")
+bldg.lay <- createDEGdf(dds.hypothalamus, "treatment", "lay", "bldg" ,  "hypothalamus")
+lay.inc.d3 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d3", "lay" ,  "hypothalamus") 
+inc.d3.inc.d9 <- createDEGdf(dds.hypothalamus,  "treatment", "inc.d9", "inc.d3" ,  "hypothalamus") 
+inc.d9.inc.d17 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d17", "inc.d9" ,  "hypothalamus")
+inc.d17.hatch <- createDEGdf(dds.hypothalamus, "treatment", "hatch", "inc.d17" ,  "hypothalamus") 
+hatch.n5 <- createDEGdf(dds.hypothalamus, "treatment", "n5", "hatch" ,  "hypothalamus") 
+n5.n9 <- createDEGdf(dds.hypothalamus, "treatment", "n9", "n5" ,  "hypothalamus") 
 ```
@@
 
-```{r allvolcanos, fig.height=8}
+```{r pitdfs, eval=T}
+dds.pituitary <- subsetDESeq2(c.colData,  c.countData, c("female_pituitary","male_pituitary"))
+```
+
+```{r pitdfs2, eval=T}
+control.bldg <- createDEGdf(dds.pituitary, "treatment", "bldg", "control", "pituitary")
+bldg.lay <- createDEGdf(dds.pituitary, "treatment", "lay", "bldg", "pituitary")
+lay.inc.d3 <- createDEGdf(dds.pituitary, "treatment", "inc.d3", "lay", "pituitary") 
+inc.d3.inc.d9 <- createDEGdf(dds.pituitary,  "treatment", "inc.d9", "inc.d3", "pituitary") 
+inc.d9.inc.d17 <- createDEGdf(dds.pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
+inc.d17.hatch <- createDEGdf(dds.pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
+hatch.n5 <- createDEGdf(dds.pituitary, "treatment", "n5", "hatch", "pituitary") 
+n5.n9 <- createDEGdf(dds.pituitary, "treatment", "n9", "n5", "pituitary") 
+```
+
+```{r pitplot, eval = T}
+a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
+b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
+c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
+d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
+g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
+h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
+
+pituitary <- plot_grid(a + 
+                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
+                       labs(y = "pituitary \n -log10(p)"),
+          b + theme(axis.title = element_blank(), axis.text = element_blank()),
+          c + theme(axis.title = element_blank(), axis.text = element_blank()),
+          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          e + theme(axis.title = element_blank(), axis.text = element_blank()),
+          f + theme(axis.title = element_blank(), axis.text = element_blank()),
+          g + theme(axis.title = element_blank(), axis.text = element_blank()),
+          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          nrow = 1,  rel_widths = c(1.75,1,1,1,1,1,1,1,1))
+pituitary
+
+```
+
+```{r allvolcanos, fig.height=6}
 background <- png::readPNG("../figures/images/DoveParentsRNAseq_charicons.png")
@@
 
+plot_grid(icons, hypothalamus, pituitary, gonads, nrow = 4, rel_heights = c(1,1,1,1.25))
+
 ```

### 2019-11-23 R/themes.R

@@ -125,3 +125,5 @@ colorsvolcanochar <-  c("control" = "#F8766D",
                     "NS" = "#bdbdbd")
 charlevelsvocano <- c("control", "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9", "NS")
 
+
+myshapes = c("hypothalamus" = 20,  "pituitary" = 17,  "gonads" = 15)

### 2019-11-23 analysis/03_DESeq2_characterization.Rmd
inmages

@@
     ggplot(aes(x = lfc, y = logpadj)) + 
-    geom_point(aes(color = direction, shape = tissue), size = 1, alpha = 0.75, na.rm = T) + 
+    geom_point(aes(color = direction, shape = tissue), size = 3, 
+               alpha = 0.75, na.rm = T) + 
     theme_minimal() +
@@
 males
-
 ```
@@
 gonads
-
 ```
@@
 hypothalamus
-
 ```
@@
 plot_grid(icons, hypothalamus, pituitary, gonads, nrow = 4, rel_heights = c(1,1,1,1.25))
-
 ```

### 2019-11-23 analysis/03_DESeq2_characterization.Rmd
images:

--- 2019-11-23 analysis/03_DESeq2_characterization.Rmd
@@
     ggplot(aes(x = lfc, y = logpadj)) + 
-    geom_point(aes(color = direction, shape = tissue), size = 1, alpha = 0.75, na.rm = T) + 
+    geom_point(aes(color = direction, shape = tissue), size = 2, 
+               alpha = 0.75, na.rm = T) + 
     theme_minimal() +
@@
 pituitary
-
 ```

### 2019-11-30 analysis/10_specificgenes.Rmd
confirming correlations between hormone and gene expression
new prolactin plots

@@
 
-head(vsd.hyp[1:2])
-head(colData.hyp[1:2,])
-rownames(colData.hyp) == colnames(vsd.hyp)
-
 vsd.pit <- readvsd("../results/04_vsd_pit.csv")
@@
 
-```{r PRL}
+```{r geneinfo}
 geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>%  dplyr::select(Name, geneid, entrezid) %>% arrange(Name)
@@
 
-```{r PRLboxplots, fig.width =5.5, fig.height= 2}
+```{r candidategenes, fig.width =5.5, fig.height= 2}
 candidategenes <- c("CISH", "SOCS1", "SOCS2", "SOCS2", "SOCS4", "SOCS5", "SOCS6", 
                     "PRL", "PRLR", "STAT5B", "JAK1")
+
 candidates.hyp <- selectcandidatevsds(candidategenes, vsd.hyp, colData.hyp)
@@
 
-e <- plotcanddateexpression(candidates.hyp, NULL, candidates.hyp$PRLR, "PRLR")
-f <- plotcanddateexpression(candidates.pit, NULL, candidates.pit$PRLR, "PRLR")
-g <- plotcanddateexpression(candidates.gon, NULL, candidates.gon$PRLR, "PRLR")
+e <- plotcanddateexpression(candidates.hyp, "hypothalamus", candidates.hyp$PRLR, "PRLR")
+f <- plotcanddateexpression(candidates.pit, "pituitary", candidates.pit$PRLR, "PRLR")
+g <- plotcanddateexpression(candidates.gon, "gonad", candidates.gon$PRLR, "PRLR")
 
@@
 
-h <- plotcanddateexpression(candidates.hyp, " ", candidates.hyp$STAT5B, "STAT5B")
-i <- plotcanddateexpression(candidates.pit, " ", candidates.pit$STAT5B, "STAT5B")
-j <- plotcanddateexpression(candidates.gon, " ", candidates.gon$STAT5B, "STAT5B")
+h <- plotcanddateexpression(candidates.hyp, "hypothalamus", candidates.hyp$STAT5B, "STAT5B")
+i <- plotcanddateexpression(candidates.pit, "pituitary", candidates.pit$STAT5B, "STAT5B")
+j <- plotcanddateexpression(candidates.gon, "gonad", candidates.gon$STAT5B, "STAT5B")
 
 plot_grid(h, i + labs(y = NULL), j + labs(y = NULL), nrow = 1)
+```
+
+```{r PRL, fig.width = 6, fig.height= 4}
+ggplot(candidates.pit, aes(x = as.numeric(treatment), y = PRL)) + 
+        geom_smooth(aes(colour = sex)) +
+    geom_boxplot(aes(fill = treatment, alpha = sex)) + 
+    scale_alpha_manual(values = c(0.75,1)) +
+    mytheme() +
+    theme(legend.position = "none") +
+    scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+    labs(y = "PRL expression", x = NULL) +
+    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
+                       labels = c( "control", "bldg", "lay", "inc.d3", 
+                                   "inc.d9", "inc.d17", "hatch", "n5", "n9"))
 ```
@@
 ```
+
+```{r PRLdf}
+head(candidates.pit)
+PRLpit <- candidates.pit %>%
+  select(sample,sex,treatment,PRL)
+head(PRLpit)
+write.csv(PRLpit, "../results/10_PRLpit.csv", row.names = F)
+```

### 2019-11-30 analysis/07_hormones.Rmd
new prolactin plots

@@
     scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
-                       labels = c( "control", "bldg", "lay", "inc.d3", 
-                                   "inc.d9", "inc.d17", "hatch", "n5", "n9"))
+                       labels = charlevels)
   }

### 2019-11-30 analysis/16_PRL.Rmd
new prolactin plots

---
title: "plots with prolactin"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(cowplot)
library(magick)
library(png)
library(grid)

source("../R/themes.R")

knitr::opts_chunk$set(fig.path = '../figures/PRL/',message=F, warning=FALSE)
```

```{r data}
hormones <- read_csv("../results/hormones.csv")
hormones$treatment <- factor(hormones$treatment, levels = alllevels)
```

```{r icons}
characterization <- png::readPNG("../figures/images/DoveParentsRNAseq_charicons.png")
characterization <- ggdraw() +  draw_image(characterization, scale = 1)

birds <- data.frame(iconpath=list.files("../figures/images/icons", full.names = T), stringsAsFactors = F)
birds$icons <- sapply(strsplit(as.character(birds$iconpath),'../figures/images/icons/birds_'), "[", 2)

control <- png::readPNG(birds$iconpath[3])
control <-  grid::rasterGrob(control, interpolate=TRUE)

bldg <- png::readPNG(birds$iconpath[1])
bldg <-  grid::rasterGrob(bldg, interpolate=TRUE)

lay <- png::readPNG(birds$iconpath[6])
lay <-  grid::rasterGrob(lay, interpolate=TRUE)

inc <- png::readPNG(birds$iconpath[5])
inc <-  grid::rasterGrob(inc, interpolate=TRUE)

hatch <- png::readPNG(birds$iconpath[4])
hatch <-  grid::rasterGrob(hatch, interpolate=TRUE)

nestling <- png::readPNG(birds$iconpath[2])
nestling <-  grid::rasterGrob(nestling, interpolate=TRUE)
```

```{r hormone}
p1 <- hormones %>% 
    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
    droplevels() %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
    geom_boxplot() +
  ylim(c(-10,120)) + 
    mytheme() +
  #theme(axis.text.x = element_blank()) + 
  theme(legend.position = "none") + 
  labs(x = "parental care stages", y = "prolactin conc. (ng/mL)") +
  annotation_custom(control, ymin = -20, ymax = 0, xmin = -7.5) +
  annotation_custom(bldg, ymin = -20, ymax = 0, xmin = -5.5) +
  annotation_custom(inc, ymin = -20, ymax = 0, xmin = -3.5) +
  annotation_custom(inc, ymin = -20, ymax = 0, xmin = -1.5) +
  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 0.5) +
  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 2.5) +
  annotation_custom(hatch, ymin = -20, ymax = 0, xmin = 4.5) +
  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 6.5) +
  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 8.5)
p1

p2 <- hormones %>% 
    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
    droplevels() %>% 
  ggplot(aes(x = as.numeric(treatment), y = plasma_conc)) +
        geom_smooth(aes(colour = sex)) +
    geom_boxplot(aes(fill = treatment, alpha = sex)) +
    mytheme() +
    theme(axis.text.x = element_blank()) +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
    labs(y = "prolactin conc. (ng/mL)", x = NULL) +
    guides(fill = FALSE, alpha = FALSE,
         color = guide_legend(order=1)) +
    scale_alpha_manual(values = c(0.75,1)) +
    theme(legend.position = c(0.85,0.15), legend.direction = "horizontal") + 
  labs(x = "parental care stages", y = "prolactin conc. (ng/mL)") +
  annotation_custom(control, ymin = -20, ymax = 0, xmin = -7.8) +
  annotation_custom(bldg, ymin = -20, ymax = 0, xmin = -5.8) +
  annotation_custom(inc, ymin = -20, ymax = 0, xmin = -3.8) +
  annotation_custom(inc, ymin = -20, ymax = 0, xmin = -1.8) +
  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 0.6) +
  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 2.4) +
  annotation_custom(hatch, ymin = -20, ymax = 0, xmin = 4.4) +
  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 6.4) +
  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 8.4) + 
  ylim(c(-10,120)) 
p2
```

### 2019-11-30 analysis/16_PRL.Rmd
more PRL figures

@@
 title: "plots with prolactin"
-output: md_document
+output: html_document
 ---
@@
 source("../R/themes.R") 
+source("../R/functions.R")
 
@@
 ```
+
+```{r PRL.pit}
+vsd.pit <- readvsd("../results/04_vsd_pit.csv")
+colData.pit <- readcolData("../results/04_colData_pit.csv")
+geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>%  dplyr::select(Name, geneid, entrezid) %>% arrange(Name)
+candidategenes <- c("PRL")
+candidates.pit <- selectcandidatevsds(candidategenes, vsd.pit, colData.pit)
+
+p3 <- ggplot(candidates.pit, aes(x = as.numeric(treatment), y = PRL)) + 
+        geom_smooth(aes(colour = sex)) +
+    geom_boxplot(aes(fill = treatment, alpha = sex)) + 
+    scale_alpha_manual(values = c(0.75,1)) +
+    mytheme() +
+  theme(legend.position = c(0.85,0.2), legend.direction = "horizontal") + 
+  scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+    labs(y = "PRL expression in the pituitary", x = "parental care stage") +
+    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
+                       labels = charlevels) +
+  annotation_custom(control, ymin = 12, ymax = 14, xmin = -7.8) +
+  annotation_custom(bldg, ymin = 12, ymax = 14, xmin = -5.8) +
+  annotation_custom(inc, ymin = 12, ymax = 14, xmin = -3.8) +
+  annotation_custom(inc, ymin = 12, ymax = 14, xmin = -1.8) +
+  annotation_custom(inc, ymin = 12, ymax = 14, xmin = 0.6) +
+  annotation_custom(inc, ymin = 12, ymax = 14, xmin = 2.4) +
+  annotation_custom(hatch, ymin = 12, ymax = 14, xmin = 4.4) +
+  annotation_custom(nestling, ymin = 12, ymax = 14, xmin = 6.4) +
+  annotation_custom(nestling, ymin = 12, ymax = 14, xmin = 8.4)  + 
+  ylim(c(13,24)) +
+  guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) 
+p3
+```

### 2019-12-01 R/themes.R
working on cleaning up repo

@@ -33,10 +33,10 @@ mytheme <- function(){
 # custom levels
 
 charlevels <- c("control", "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9")
-maniplevels <- c( "remove.d03" ,  "remove.d09" ,  "remove.d17" , "remove.d20" , 
-                  "extend" , "prolong", "early")
-maniplevels1 <- c( "m.inc.d3" ,  "m.inc.d8" ,  "m.inc.d9" , "m.inc.d17" , "m.n2", 
-                   "prolong" , "extend")
+
+maniplevels <- c( "remove.d03" ,  "remove.d09" ,  "remove.d17" , "remove.d20" ,  "extend" , "prolong", "early")
+
+maniplevels1 <- c( "m.inc.d3" ,  "m.inc.d8" ,  "m.inc.d9" , "m.inc.d17" , "m.n2", "prolong" , "extend")
 
 alllevels <- c("control", "bldg", "lay", "inc.d3", "m.inc.d3" ,  
                 "inc.d9", "m.inc.d8" ,"m.inc.d9" ,
@@ -50,6 +50,32 @@ alllevels2 <- c("control", "bldg", "lay", "inc.d3", "inc.d9",
                 "m.inc.d17" , "m.n2", "prolong" , "extend")
 
 tissuelevels <- c("hypothalamus", "pituitary", "gonads")
+
+
+charlevelsvocano <- c("control", "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9", "NS")
+
+
+## comparisons
+
+serialtimepoints <- c("control.bldg" , "bldg.lay", "lay.inc.d3", "inc.d3.inc.d9", 
+                      "inc.d9.inc.d17", "inc.d17.hatch", "hatch.n5", "n5.n9")
+
+manipVchar <- c("inc.d3.m.inc.d3", "inc.d9.m.inc.d8", "inc.d9.m.inc.d9","inc.d17.m.inc.d17",
+                "hatch.prolong", "hatch.extend", "hatch.m.n2")
+
+offspringremoval <- c("lay.inc.d3", "inc.d3.m.inc.d3", 
+                      "inc.d3.inc.d9", "inc.d9.m.inc.d9",
+                      "inc.d9.inc.d17", "inc.d17.m.inc.d17", 
+                      "hatch.n5", "hatch.m.n2")
+
[…]

### 2019-12-01 analysis/05_DESeq2_manipulate.Rmd
working on cleaning up repo

@@
 source("../R/themes.R")
+
+library(BiocParallel)
+register(MulticoreParam(6))
 
@@
 ```{r dds, cache = T}
-dds.female_hypothalamus <- subsetDESeq("female_hypothalamus")
-dds.female_pituitary <- subsetDESeq("female_pituitary" )
-dds.female_gonad <- subsetDESeq("female_gonad" )
-dds.male_hypothalamus <- subsetDESeq("male_hypothalamus" )
-dds.male_pituitary <- subsetDESeq("male_pituitary"  )
-dds.male_gondad <- subsetDESeq("male_gonad")
+dds.female_hypothalamus <- subsetDESeq(m.colData, m.countData, "female_hypothalamus")
+dds.female_pituitary <- subsetDESeq(m.colData, m.countData, "female_pituitary" )
+dds.female_gonad <- subsetDESeq(m.colData, m.countData, "female_gonad" )
+dds.male_hypothalamus <- subsetDESeq(m.colData, m.countData, "male_hypothalamus" )
+dds.male_pituitary <- subsetDESeq(m.colData, m.countData, "male_pituitary"  )
+dds.male_gondad <- subsetDESeq(m.colData, m.countData, "male_gonad")
 ```
 
-## Calculate and plot total DEGs
+## volcano plots
 
-```{r totalDEGs, eval = T}
-#create list of groups for deseq contrasts
-group1 <- c("m.inc.d3","m.inc.d8", "m.inc.d9", "m.inc.d17", "prolong","extend", "m.n2" ) 
-group2 <- group1
+```{r}
 
-a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
-b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
-c <- plottotalDEGs(dds.female_gonad, "female gonad")
-d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
-e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
-f <- plottotalDEGs(dds.male_gondad, "male gonad")
-
-plot_grid(a + theme(legend.position = "none"),
-          b + theme(legend.position = "none"),
-          c,
-          d + theme(legend.position = "none"),
-          e + theme(legend.position = "none"),
-          f,
-          nrow = 2, rel_widths = c(0.3, 0.3, 0.4)) 
-```
-
-## Calculate and plot principal components
-
-```{r pca, eval = F}
-plotPCAs(dds.female_hypothalamus, "female hypothalamus")
-plotPCAs(dds.female_pituitary, "female pituitary")      
-plotPCAs(dds.female_gonad, "female gonad")
-plotPCAs(dds.male_hypothalamus, "male hypothalamus")
-plotPCAs(dds.male_pituitary, "male pituitary")
-plotPCAs(dds.male_gondad, "male gonad")
-```
-
-## heamap with minimum pvalue
-
-```{r pheatmap, eval = F}
-makepheatmap(dds.female_hypothalamus, "female hypothalamus")
-makepheatmap(dds.female_pituitary, "female pituitary")
-makepheatmap(dds.female_gonad, "female gonad")
-makepheatmap(dds.male_hypothalamus, "male hypothalamus")
-makepheatmap(dds.male_pituitary, "male pituitary")
-makepheatmap(dds.male_gondad, "male gonad")        
-```
-
-## candidate genes
-
-```{r candidates, eval = F}
-plotcandidates(dds.female_hypothalamus, "female hypothalamus")
-plotcandidates(dds.female_pituitary, "female pituitary")
-plotcandidates(dds.female_gonad, "female gonad")
-plotcandidates(dds.male_hypothalamus, "male hypothalamus")
-plotcandidates(dds.male_pituitary, "male pituitary")
-plotcandidates(dds.male_gondad, "male gonad")
+dds.hypothalamus <- subsetDESeq2(m.colData,  m.countData, c("female_hypothalamus","male_hypothalamus"))
+dds.hypothalamus <- subsetDESeq2(m.colData,  m.countData, c("female_pituitary","male_pituitary"))
+dds.hypothalamus <- subsetDESeq2(m.colData,  m.countData, c("female_gonad","male_gonad"))
 
@@
 
-## offspring removal
+```{r dds, cache = T}
 
-```{r offspringremoval, eval = F}
-g <- plotremoval(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
-h <- plotremoval(DEGs.female_pituitary, "Female pituitary", "none")
-i <- plotremoval(DEGs.female_gonad, "Female gonad",  "none")
-j <- plotremoval(DEGs.male_hypothalamus, "Male hypothalamus", "none")
-k <- plotremoval(DEGs.male_pituitary, "Male pituitary", "none")
-l <- plotremoval(DEGs.male_gondad, "Male gonad", "none")
+vsdh <- vst(dds.hypothalamus, blind=FALSE)
+vsdp <- vst(dds.pituitary, blind=FALSE)
+vsdg <- vst(dds.gonad, blind=FALSE)
 
-legend <- get_legend(g)
-
-mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)
-
-mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
-mybarplots2
-
-pdf("../figures/sexes/offspringremoval-1.pdf", width = 10, height = 6)
-plot(mybarplots2)
-dev.off()
-
-```
-
-## prolong extend
-
-```{r prolongextend, eval = F}
-g <- plotprolongdelay(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
-h <- plotprolongdelay(DEGs.female_pituitary, "Female pituitary", "none")
-i <- plotprolongdelay(DEGs.female_gonad, "Female gonad",  "none")
-j <- plotprolongdelay(DEGs.male_hypothalamus, "Male hypothalamus", "none")
-k <- plotprolongdelay(DEGs.male_pituitary, "Male pituitary", "none")
-l <- plotprolongdelay(DEGs.male_gondad, "Male gonad", "none")
-
-legend <- get_legend(g)
-
-mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)
-
-mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
-mybarplots2
-
-pdf("../figures/sexes/prolongextend-1.pdf", width = 10, height = 6)
-plot(mybarplots2)
-dev.off()
-
+vsdfh <- vst(dds.female_hypothalamus, blind=FALSE)
+vsdfp <- vst(dds.female_pituitary, blind=FALSE)
+vsdfg <- vst(dds.female_gonad, blind=FALSE)
+vsdmh <- vst(dds.male_hypothalamus, blind=FALSE)
+vsdmp <- vst(dds.male_pituitary, blind=FALSE)
+vsdmg <- vst(dds.male_gondad, blind=FALSE)
 ```
@@
                          color = colData$treatment, 
-                         shape = colData$tissue,
-                         alpha = colData$sex)) +
+                         shape = colData$tissue)) +
   geom_point(size = 2)  + 
@@
   scale_alpha_manual(values = c(0.5,1)) +
-  scale_color_manual(values = colorsmaniplevels1) +
+  scale_color_manual(values = colorsmanip) +
   labs(x = "PC1: 63.6% of variance", y = "PC2: 30.1% of variance")
@@
       mytheme() + 
-      labs(x = " ", y = " ", subtitle = NULL, title =  NULL)
+      labs(x = " ", y = " ", subtitle = NULL, title =  NULL) +
+  theme( axis.text = element_blank())
 
-png('../figures/manipulation/pcaprep-1.png',width=600,height=600, units="px",bg = "transparent")
-print(a)
-dev.off()
-
-png('../figures/manipulation/pcaprep-2.png',width=600,height=600, units="px",bg = "transparent")
-print(b)
-dev.off()
+plot_grid(a,b, rel_widths = c(2,1))
 ```

### 2019-12-02 analysis/07_hormones.Rmd
new hormone analysis

@@
 ```{r wrangle-prolactin}
-prolactin <- read_excel("../results/Pigeon prolactin concentrations juil 2018.xlsx", sheet = 1)
-
-# keep only samples realated to parental care
-prolactin <- prolactin %>% filter(Study %in% c("Baseline", "ParentalCare"))
-
-# create lists of factos for characterization and manipluation studies
-
-# rename some of the treatments levels, store in new columns
-prolactin <- prolactin %>%
-    mutate(sex = fct_recode(Sex,
+prolactin <- read_excel("../results/Pigeon prolactin concentrations juil 2018.xlsx", sheet = 1) %>% 
+  filter(Study %in% c("Baseline", "ParentalCare")) %>%
+    dplyr::mutate(sex = fct_recode(Sex,
                             "female" = "f",
@@
                                  characterization = charlevels,
-                                 manipulation = maniplevels1))
+                                 manipulation = maniplevels1)) %>%
+          dplyr::rename("plasma_conc" = "Prolactin ng/mL") %>%
+          mutate(bird_id = gsub("[[:punct:]]", "." , ColorBands)) %>% 
+          dplyr::mutate(hormone = "prolactin") %>% 
+          dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc)  %>% 
+          drop_na()
+head(prolactin)
 
-colnames(prolactin)[colnames(prolactin)=="Prolactin ng/mL"] <- "plasma_conc"
-prolactin$hormone <- "prolactin"
-prolactin <- prolactin %>% 
-              select(study, treatment, sex, hormone, plasma_conc)  %>% 
-              drop_na()
+prolactin2 <- read_csv("../results/prolactin.csv") %>% 
+          dplyr::rename("plasma_conc" = "prolactin_ng_mL",
+                        "treatment" = "stage") %>% 
+           dplyr::mutate(hormone = "prolactin") %>% 
+           dplyr::mutate(treatment = fct_recode(treatment,
+                            "m.n2" = "m.hatch",
+                            "prolong" = "inc.prolong",
+                            "extend" = "extend.hatch"),
+                          study = fct_collapse(treatment,
+                                 characterization = charlevels,
+                                 manipulation = maniplevels1)) %>% 
+          dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc) 
+head(prolactin2)
 
-summary(prolactin)
+same <- inner_join(prolactin2, prolactin, by = "bird_id")
+
+ggplot(same, aes(x = plasma_conc.x, y = plasma_conc.y, label = bird_id)) + geom_point() + geom_text()
 ```
@@
 ```{r wrangle-steroids}
-PETC <- read_excel("../results/parental_care_hormone_RIA_data_master.xlsx", sheet = 2)
-
-PETC <- PETC %>% select(stage, sex, hormone, plasma_conc)  %>%
-                mutate(treatment = fct_recode(stage,
+PETC <- read_excel("../results/parental_care_hormone_RIA_data_master.xlsx", sheet = 2)  %>% 
+                dplyr::mutate(treatment = fct_recode(stage,
                             "inc.d17" = "inc_d17",
@@
                             "control" = "stress_hpg")) %>%
-                filter(treatment %in% alllevels2) %>%   
-                mutate(sex = fct_recode(sex,
+                dplyr::filter(treatment %in% alllevels2) %>%   
+                dplyr::mutate(sex = fct_recode(sex,
                             "female" = "f",
@@
                                 manipulation = maniplevels1)) %>% 
-              drop_na() %>%  droplevels()  %>% 
-              select(study, treatment, sex, hormone, plasma_conc)
-summary(PETC)
+              dplyr::mutate(bird_id = gsub("[[:punct:]]", "." , band_combo)) %>% 
+              dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc) %>% 
+             drop_na() %>%  droplevels()  
+head(PETC)
+
+```
+
+```{r correlations}
+PETCP <- left_join(prolactin2, PETC, by = "bird_id") %>% drop_na()
+PETCP$treatment.x <- factor(PETCP$treatment.x, levels = alllevels2)
+PETCP$hormone.y
+
+plothormonecorrelations <- function(myhormone, myylab){
+  PETCP %>% filter(hormone.y == myhormone) %>%
+  ggplot(aes(x = plasma_conc.x, y = plasma_conc.y, color = treatment.x )) +
+  geom_point() + 
+  geom_smooth(method = "lm", se = F) +
+  labs(x = "Prolactin (ng/mL)", y = myylab) +
+  scale_color_manual(values = colorscharmaip) +
+  facet_wrap(~treatment.x, nrow = 3) +
+  theme(legend.position = "none")
+}
+
+plothormonecorrelations("cort", "Corticosterone (ng/mL")
+plothormonecorrelations("estradiol", "Estradiol (ng/mL")
+plothormonecorrelations("progesterone", "Progesterone (ng/mL")
+plothormonecorrelations("testosterone", "Testosterone (ng/mL")
 ```
@@
 plot_grid(d1,d2, nrow = 1)
-
 ```

### 2019-12-02 analysis/05_DESeq2_manipulate.Rmd
new manipulation plots

@@
   scale_alpha_manual(values = c(0.5,1)) +
-  scale_color_manual(values = colorsmanip) +
+  scale_color_manual(values = colorscharmaip2) +
   labs(x = "PC1: 63.6% of variance", y = "PC2: 30.1% of variance")

### 2019-12-02 analysis/07_hormones.Rmd
new manipulation plots

@@
                                  manipulation = maniplevels1)) %>%
+
+          dplyr::mutate(bird_id = gsub(".ATLAS", "" , bird_id)) %>% 
           dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc) 
@@
 
-same <- inner_join(prolactin2, prolactin, by = "bird_id")
+same <- inner_join(prolactin, prolactin2, by = "bird_id")
+ggplot(same, aes(x = plasma_conc.x, y = plasma_conc.y, label = bird_id)) + geom_point() + geom_text()
 
-ggplot(same, aes(x = plasma_conc.x, y = plasma_conc.y, label = bird_id)) + geom_point() + geom_text()
+all <- full_join(prolactin, prolactin2, by = "bird_id") 
+onlyin1 <- anti_join(prolactin, prolactin2, by = "bird_id") 
+onlyin2 <- anti_join(prolactin2, prolactin, by = "bird_id")
+
+dim(prolactin)
+dim(prolactin2)
+dim(all)
+dim(same)
+dim(onlyin1)
+dim(onlyin2)
+
+print("only in old prolactin dataset") 
+onlyin1$bird_id
+
+print("only in new prolactin dataset") 
+onlyin2$bird_id
+
 ```
@@
 
+```{r manipulation}
+# prolactin, removal only
+hormones %>% 
+    filter( hormone == c("prolactin"))  %>% 
+    filter(treatment %in% c("inc.d3", "m.inc.d3", "inc.d9", "m.inc.d9", "inc.d17", "m.inc.d17", "hatch","m.n2"))  %>% 
+  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment,  color = sex)) +
+    geom_boxplot() + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none",
+          strip.text = element_blank()) +
+    scale_fill_manual(values = colorscharmaip) +
+    scale_color_manual(values = sexcolors) +
+    labs(y = "PRL (ng/mL)", x = NULL) +
+    annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -2, ymax = 0, alpha = 0.25) +
+    annotate("rect", xmin = 2.6, xmax = 4.4, ymin = -2, ymax = 0, alpha = 0.5) +
+    annotate("rect", xmin = 4.6, xmax = 6.4, ymin = -2, ymax = 0, alpha = 0.75)  +
+    annotate("rect", xmin = 6.6, xmax = 8.4, ymin = -2, ymax = 0, alpha = 1)
+
+hormones %>% 
+    filter( hormone == c("prolactin"))  %>% 
+    filter( treatment %in% c("inc.d9", "m.inc.d8", "inc.d17", "prolong", "hatch", "extend", "n5"))  %>% 
+  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment,  color = sex)) +
+    geom_boxplot() + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none",
+          strip.text = element_blank()) +
+    scale_fill_manual(values = colorscharmaip) +
+    scale_color_manual(values = sexcolors) +
+    labs(y = "PRL (ng/mL)", x = NULL) +
+    annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -5, ymax = -3, alpha = 0.33) +
+    annotate("rect", xmin = 4.6, xmax = 5.4, ymin = -5, ymax = -3, alpha = 0.33) +
+    annotate("rect", xmin = 2.6, xmax = 5.4, ymin = -2, ymax = 0, alpha = 0.66) +
+    annotate("rect", xmin = 4.6, xmax = 7.4, ymin = 1, ymax = 3, alpha = 1) 
+```
+
 ```{r sexsteroids}
+
 hormonecharplot("cort", "CORT (ng/mL)")
@@
 plot_grid(d1,d2, nrow = 1)
+
+hormonemanipSteroids <- function(myhormone, myylab, myymax){
+
+  hormones %>% 
+    filter( hormone %in% c(myhormone))  %>% 
+  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
+    geom_boxplot() + 
+    mytheme() +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+         legend.position = "none",
+          strip.text = element_blank()) +
+    scale_fill_manual(values = colorscharmaip) +
+    scale_color_manual(values = sexcolors) +
+    labs(y = myylab, x = NULL) 
+}
+
+hormonemanipSteroids("cort", "CORT (ng/mL)", 10)
+hormonemanipSteroids("progesterone", "PROG (ng/mL)",2.5)
+
+a <- hormonemanipSteroids("estradiol", "E (ng/mL)", 1)
+b <- hormonemanipSteroids("testosterone", "T (ng/mL)", 3.5)
+plot_grid(a,b)
+
 ```
@@
 
-```{r manipulation}
-# prolactin, removal only
-hormones %>% 
-    filter( hormone == c("prolactin"))  %>% 
-    filter( !treatment %in% c("m.inc.d8", "prolong", "extend"))  %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment,  color = sex)) +
-    geom_boxplot() + 
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
-          legend.position = "none",
-          strip.text = element_blank()) +
-    scale_fill_manual(values = colorscharmaip) +
-    scale_color_manual(values = sexcolors) +
-    labs(y = "PRL (ng/mL)", x = NULL) +
-    annotate("rect", xmin = 3.7, xmax = 5.3, ymin = -2, ymax = 0, alpha = 1) +
-    annotate("rect", xmin = 5.7, xmax = 7.3, ymin = -2, ymax = 0, alpha = 1) +
-    annotate("rect", xmin = 7.7, xmax = 9.3, ymin = -2, ymax = 0, alpha = 1) +
-    annotate("rect", xmin = 9.7, xmax = 11.3, ymin = -2, ymax = 0, alpha = 1)
-
-hormones %>% 
-    filter( hormone == c("prolactin"))  %>% 
-    filter( !treatment %in% c("m.inc.d3", "m.inc.d9", "m.inc.d17","m.n2"))  %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment,  color = sex)) +
-    geom_boxplot() + 
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
-          legend.position = "none",
-          strip.text = element_blank()) +
-    scale_fill_manual(values = colorscharmaip) +
-    scale_color_manual(values = sexcolors) +
-    labs(y = "PRL (ng/mL)", x = NULL) +
-    annotate("rect", xmin = 4.7, xmax = 6.3, ymin = -2, ymax = 0, alpha = 1, alpha = 1) +
-    annotate("rect", xmin = 6.7, xmax = 8.3, ymin = -2, ymax = 0, alpha = 1, alpha = 1) +
-    annotate("rect", xmin = 8.7, xmax = 10.3, ymin = -2, ymax = 0, alpha = 1, alpha = 1)
-
-hormonemanipSteroids <- function(myhormone, myylab, myymax){
-
-  hormones %>% 
-    filter( hormone %in% c(myhormone))  %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
-    geom_boxplot() + 
-    mytheme() +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
-         # legend.position = "none",
-          strip.text = element_blank()) +
-    scale_fill_manual(values = colorscharmaip) +
-    scale_color_manual(values = sexcolors) +
-    labs(y = myylab, x = NULL) 
-}
-
-a <- hormonemanipSteroids("estradiol", "E (ng/mL)", 1)
-b <- hormonemanipSteroids("testosterone", "T (ng/mL)", 3.5)
-c <- hormonemanipSteroids("cort", "CORT (ng/mL)", 10)
-d <- hormonemanipSteroids("progesterone", "PROG (ng/mL)",2.5)
-
-a
-b
-c
-d
-```
-
 ```{r}

### 2019-12-02 R/themes.R

@@ -80,12 +80,12 @@ myshapes = c("hypothalamus" = 20,  "pituitary" = 17,  "gonads" = 15)
 # custom colors
 
 colorsmanip <- c( "m.inc.d3" = "#CDCDCD", 
-                         "m.inc.d8" = "#6DA9D3", 
+                         "m.inc.d8" = "#fcc5c0", 
                          "m.inc.d9" = "#959595", 
                          "m.inc.d17" = "#626262",
                          "m.n2" = "#262625", 
-                         "prolong" = "#1B86B9" , 
-                         "extend" = "#01588C")
+                         "prolong" = "#f768a1" , 
+                         "extend" = "#f768a1")
 
 colorschar <-  c("control" = "#F8766D", 
                          "bldg"= "#D39200", 
@@ -108,12 +108,31 @@ colorscharmaip <-  c("control" = "#F8766D",
                  "n5"= "#DB72FB", 
                  "n9"= "#FF61C3",
                  "m.inc.d3" = "#CDCDCD", 
-                 "m.inc.d8" = "#6DA9D3", 
+                 "m.inc.d8" = "#fcc5c0", 
                  "m.inc.d9" = "#959595", 
                  "m.inc.d17" = "#626262",
                  "m.n2" = "#262625", 
-                 "prolong" = "#1B86B9" , 
-                 "extend" = "#01588C")
+                 "prolong" = "#f768a1" , 
+                 "extend" = "#7a0177")
+
+colorscharmaip2 <-  c("control" = "#F8766D", 
+                     "bldg"= "#D39200", 
+                     "lay"= "#93AA00", 
+                     "inc.d3"= "#00BA38", 
+                     "inc.d9"= "#00C19F", 
+                     "inc.d17"= "#00B9E3", 
+                     "hatch"= "#619CFF", 
+                     "n5"= "#DB72FB", 
+                     "n9"= "#FF61C3",
+                     
+                     "m.inc.d3" = "#CDCDCD", 
+                     "m.inc.d9" = "#959595", 
+                     "m.inc.d17" = "#626262",
+                     "m.n2" = "#262625", 
+                     
+                     "m.inc.d8" = "#fcc5c0", 
+                     "prolong" = "#f768a1" , 
+                     "extend" = "#7a0177")
 
 colorstissue <- c("hypothalamus" = "#d95f02",
                   "pituitary" = "#1b9e77",
[…]

### 2019-12-02 analysis/05_DESeq2_manipulate.Rmd
removed all non-DESeq stuff and old ugly plots

--- 2019-12-01 analysis/05_DESeq2_manipulate.Rmd
@@
 
-```{r dds, cache = T}
+```{r vsd, cache = T}
 
@@
 ```
-
-# pca and tsne
-
-```{r}
-pseudocounts <- read_csv("../results/01_pseudo.counts.csv")
-head(pseudocounts[1:3])
-
-colData <- read.csv("../metadata/00_colData_manipluation.csv", header = T, row.names = 1)
-
-colData$treatment <- factor(colData$treatment, levels = maniplevels1)
-colData$tissue <- factor(colData$tissue, levels = c("hypothalamus", "pituitary", "gonad"))
-colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
-head(colData[1:3])
-
-geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
-
-subsetcounts <- function(colData, countData){
-
-  # save counts that match colData
-  savecols <- as.character(colData$V1) 
-  savecols <- as.vector(savecols) 
-  countData <- countData %>% 
-    dplyr::select(X1,savecols) %>% 
-    dplyr::rename(entrezid = X1)
-
-  countData <- left_join(geneinfo, countData)
-  print(head(countData))
-  countData <- countData %>% dplyr::distinct(Name, .keep_all = TRUE) %>%
-    dplyr::select(-row.names)  %>% dplyr::select(-geneid) %>% dplyr::select(-entrezid) 
-  print(head(countData))
-  row.names(countData) <- countData$Name
-  countData$Name <- NULL
-  countData <- as.data.frame(t(countData))
-
-  return(countData)
-}
-
-countData <- subsetcounts(colData, pseudocounts)
-head(countData[1:3])
-head(colData)
-# confirm
-row.names(colData) <- colData$V1
-row.names(countData) == row.names(colData)
-
-euclidist <- dist(countData) # euclidean distances between the rows
-```
-
-```{r pcaprep}
-# make pca dataframes
-mypca <- prcomp(countData)
-
-# percent conribution
-eig.val <- get_eigenvalue(mypca)
-head(eig.val)
-
-mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
-                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
-                  ID = row.names(countData))
-mypcadf$V1 <- row.names(mypcadf)
-mypcadf <- left_join(colData, mypcadf)
-mypcadf <- mypcadf %>% dplyr::select(bird,sex,tissue,treatment,PC1:PC6)
-mypcadf$treatment <- factor(mypcadf$treatment, levels = maniplevels1)
-head(mypcadf)
-
-a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
-                         color = colData$treatment, 
-                         shape = colData$tissue)) +
-  geom_point(size = 2)  + 
-  mytheme() +
-  theme(legend.title = element_blank(), 
-        axis.text = element_blank()) +
-  scale_alpha_manual(values = c(0.5,1)) +
-  scale_color_manual(values = colorsmanip) +
-  labs(x = "PC1: 63.6% of variance", y = "PC2: 30.1% of variance")
-
-b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
-                   repel = TRUE ,
-                  select.var= list(contrib = 6)) + 
-      mytheme() + 
-      labs(x = " ", y = " ", subtitle = NULL, title =  NULL) +
-  theme( axis.text = element_blank())
-
-plot_grid(a,b, rel_widths = c(2,1))
-```
-
-```{r tsne}
-
-tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=5, theta=0.5, dims=2)
-tsne_df = as.data.frame(tsne_model$Y) 
-plot(tsne_df$V1, tsne_df$V2)
-
-# prep for adding columns
-colData2 <- colData 
-colData2$V1 <- NULL
-tsne_df_cols <- cbind(colData2, tsne_df)
-
-c  <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
-  geom_point(size = 2) +
-  mytheme() +
-  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
-  labs(x = "tSNE 1", y = "tSNE 2")
-
-d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = tissue, color = treatment, alpha = sex)) +
-  geom_point(size = 2) +
-  mytheme() +
-  labs(x = "tSNE 1", y = "tSNE 2") +
-  scale_color_manual(values = colorsmaniplevels1) +
-  scale_alpha_manual(values = c(0.5,1))
-
-c
-d
-```

### 2019-12-02 analysis/02_PCA.Rmd
renamed PCA analysis, updated all figures

---
title: "PCA"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggfortify)
library(factoextra)
library(cowplot)
library(Rtsne)
library(viridis)
library(forcats)

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/pca/', cache = T)

source("../R/themes.R")  
```

## Data: Pseudocounts from Limma

```{r read}
pseudocounts <- read_csv("../results/01_pseudo.counts.csv")
head(pseudocounts[1:3])
```

## Characterization only

```{r characterization}
colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
colData$treatment <- factor(colData$treatment, levels = charlevels)
colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
colData$tissue <- factor(colData$tissue, levels = tissuelevels)

head(colData[1:3])

geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

subsetcounts <- function(colData, countData){

  # save counts that match colData
  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols) 
  countData <- countData %>% 
    dplyr::select(X1,savecols) %>% 
    dplyr::rename(entrezid = X1)

  countData <- left_join(geneinfo, countData)
  countData <- countData %>% distinct(Name, .keep_all = TRUE) %>%
    select(-row.names)  %>% select(-geneid) %>% select(-entrezid)

  row.names(countData) <- countData$Name
  countData$Name <- NULL
  countData <- as.data.frame(t(countData))

  return(countData)
}

countData <- subsetcounts(colData, pseudocounts)
head(countData[1:3])

# confirm
#row.names(countData) == row.names(colData)

euclidist <- dist(countData) # euclidean distances between the rows

# make pca dataframes
mypca <- prcomp(countData)

# percent conribution
eig.val <- get_eigenvalue(mypca)
head(eig.val)

mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
                  ID = row.names(countData))
mypcadf$V1 <- row.names(mypcadf)
mypcadf <- left_join(colData, mypcadf)
mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
head(mypcadf)

a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
                         color = colData$treatment, 
                         shape = colData$tissue)) +
  geom_point()   +
  mytheme() +
  theme(legend.title = element_blank(), 
        axis.text = element_blank()) +
  scale_alpha_manual(values = c(0.5,1)) +
  labs(x = "PC1: 35.9% of variance", y = "PC2: 32.5% of variance")

b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
                   repel = TRUE ,
                  select.var= list(contrib = 6)) + 
      mytheme() + 
      labs(x = " ", y = " ", subtitle = NULL, title =  NULL) +
    theme( axis.text = element_blank())

plot_grid(a,b, rel_widths = c(2,1))

fviz_screeplot(mypca, addlabels = TRUE, ylim = c(0, 50))
fviz_contrib(mypca, choice = "var", axes = 1, top = 5) 
fviz_contrib(mypca, choice = "var", axes = 2, top = 5)

tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=5, theta=0.5, dims=2)
tsne_df = as.data.frame(tsne_model$Y) 
plot(tsne_df$V1, tsne_df$V2)

# prep for adding columns
colData2 <- colData 
colData2$V1 <- NULL
tsne_df_cols <- cbind(colData2, tsne_df)

c  <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
  geom_point(size = 2) +
  mytheme() +
  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
  labs(x = "tSNE 1", y = "tSNE 2")

d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = tissue, color = treatment, alpha = sex)) +
  geom_point(size = 2) +
  mytheme() +
  labs(x = "tSNE 1", y = "tSNE 2") +
  scale_alpha_manual(values = c(0.5,1))

c
d
```

## Manipulations only

```{r manipulation}
head(pseudocounts[1:3])

colData <- read.csv("../metadata/00_colData_manipluation.csv", header = T, row.names = 1)

colData$treatment <- factor(colData$treatment, levels = maniplevels1)
colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
colData$tissue <- factor(colData$tissue, levels = tissuelevels)

head(colData[1:3])

countData <- subsetcounts(colData, pseudocounts)

# confirm
row.names(colData) <- colData$V1
#row.names(countData) == row.names(colData)

euclidist <- dist(countData) # euclidean distances between the rows

# make pca dataframes
mypca <- prcomp(countData)

# percent conribution
eig.val <- get_eigenvalue(mypca)
head(eig.val)

mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
                  ID = row.names(countData))
mypcadf$V1 <- row.names(mypcadf)
mypcadf <- left_join(colData, mypcadf)
mypcadf <- mypcadf %>% dplyr::select(bird,sex,tissue,treatment,PC1:PC6)
mypcadf$treatment <- factor(mypcadf$treatment, levels = maniplevels1)
head(mypcadf)

a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
                         color = colData$treatment, 
                         shape = colData$tissue)) +
  geom_point(size = 2)  + 
  mytheme() +
  theme(legend.title = element_blank(), 
        axis.text = element_blank()) +
  scale_alpha_manual(values = c(0.5,1)) +
  scale_color_manual(values = colorscharmaip2) +
  labs(x = "PC1: 63.6% of variance", y = "PC2: 30.1% of variance")

b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
                   repel = TRUE ,
                  select.var= list(contrib = 6)) + 
      mytheme() + 
      labs(x = " ", y = " ", subtitle = NULL, title =  NULL) +
  theme( axis.text = element_blank())

plot_grid(a,b, rel_widths = c(2,1))

fviz_screeplot(mypca, addlabels = TRUE, ylim = c(0, 50))
fviz_contrib(mypca, choice = "var", axes = 1, top = 5) 
fviz_contrib(mypca, choice = "var", axes = 2, top = 5)

tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=5, theta=0.5, dims=2)
tsne_df = as.data.frame(tsne_model$Y)

# prep for adding columns
colData2 <- colData 
colData2$V1 <- NULL
tsne_df_cols <- cbind(colData2, tsne_df)

c  <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
  geom_point(size = 2) +
  mytheme() +
  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
  labs(x = "tSNE 1", y = "tSNE 2")

d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = tissue, color = treatment, alpha = sex)) +
  geom_point(size = 2) +
  mytheme() +
  labs(x = "tSNE 1", y = "tSNE 2") +
  scale_color_manual(values = colorscharmaip2) +
  scale_alpha_manual(values = c(0.5,1))

c
d
```

## All data, characterization and manipulations

```{r charmanip}
# prep col data for all samples
colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
colData$treatment <- factor(colData$treatment, levels = alllevels3)
colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
colData$tissue <- factor(colData$tissue, levels = tissuelevels)
row.names(colData) <- colData$V1

# prep count data for all samples
countData <- as.data.frame(pseudocounts)
countData$entrezid <- countData$X1
countData <- left_join(geneinfo, countData)
countData <- countData %>% distinct(Name, .keep_all = TRUE) %>%
    select(-row.names, -geneid, -entrezid, -X1) 
countData <- as.data.frame(countData)
row.names(countData) <- countData$Name
countData$Name <- NULL
countData$entrezid <- NULL
countData <- as.data.frame(t(countData))

# check ready for analysis
#row.names(countData) == row.names(colData)

# pca
mypca <- prcomp(countData)
eig.val <- get_eigenvalue(mypca)
head(eig.val)

mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
                  ID = row.names(countData))
mypcadf$V1 <- row.names(mypcadf)
mypcadf <- left_join(colData, mypcadf)
mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
head(mypcadf)

a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
                         color = colData$treatment, 
                         #shape = colData$tissue,
                         #size = colData$sex,
                        shape = colData$study)) +
  #geom_hline(yintercept=0, color = "grey") +
  #geom_vline(xintercept=0, color = "grey") +
  geom_point()  +
  mytheme() +
  theme(legend.title = element_blank(),
        axis.text = element_blank()) +
  #scale_alpha_manual(values = c(0.5,1)) +
  labs(x = "PC1: 44.6% of variance", y = "PC2: 29.8% of variance") +
  scale_color_manual(values = colorscharmaip)

b <- fviz_pca_var(mypca,  labelsize = 4 , axes.linetype = "blank", 
                   repel = TRUE ,
                  select.var= list(contrib = 6)) + 
      mytheme() + 
      labs(x = " ", y = " ", subtitle = NULL, title =  NULL) +
    #geom_hline(yintercept=0, color = "grey") +
  #geom_vline(xintercept=0, color = "grey") +
    theme( axis.text = element_blank())

fviz_screeplot(mypca, addlabels = TRUE, ylim = c(0, 50))
fviz_contrib(mypca, choice = "var", axes = 1, top = 5) 
fviz_contrib(mypca, choice = "var", axes = 2, top = 5)

plot_grid(a,b,rel_widths = c(2,1))

# prep for tsne
euclidist <- dist(countData) # euclidean distances between the rows

tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=10, theta=0.5, dims=2)
tsne_df = as.data.frame(tsne_model$Y)

# prep for adding columns
colData2 <- colData 
colData2$V1 <- NULL
tsne_df_cols <- cbind(colData2, tsne_df)

ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
  geom_point(size = 2) +
  mytheme() +
  scale_color_manual(values = colorstissue) +
  labs(x = "tSNE 1", y = "tSNE 2")

ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = study, color = tissue)) +
  geom_point(size = 2) +
  mytheme() +
  scale_color_manual(values = colorstissue) +
  labs(x = "tSNE 1", y = "tSNE 2")

ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = study, color = treatment)) +
  geom_point(size = 2) +
  mytheme() +
  labs(x = "tSNE 1", y = "tSNE 2") +
  scale_alpha_manual(values = c(0.5,1)) +
  scale_color_manual(values = colorscharmaip2)
```

### 2019-12-02 R/themes.R
renamed PCA analysis, updated all figures

@@ -49,6 +49,11 @@ alllevels2 <- c("control", "bldg", "lay", "inc.d3", "inc.d9",
                 "m.inc.d3" ,  "m.inc.d8" ,  "m.inc.d9" , 
                 "m.inc.d17" , "m.n2", "prolong" , "extend")
 
+alllevels3 <- c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                "inc.d17", "hatch", "n5", "n9",
+                "m.inc.d3" ,  "m.inc.d9" , "m.inc.d17" , "m.n2",
+                 "m.inc.d8" , "prolong" , "extend")
+
 tissuelevels <- c("hypothalamus", "pituitary", "gonads")
 
 

### 2019-12-02 analysis/02_PCA.Rmd
prettier plots

@@
 tsne_df = as.data.frame(tsne_model$Y)
-plot(tsne_df$V1, tsne_df$V2)
 
@@
 
-d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = tissue, color = treatment, alpha = sex)) +
+d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = tissue, color = treatment)) +
   geom_point(size = 2) +
@@
 
-ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = study, color = tissue)) +
-  geom_point(size = 2) +
-  mytheme() +
-  scale_color_manual(values = colorstissue) +
-  labs(x = "tSNE 1", y = "tSNE 2")
-
 ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = study, color = treatment)) +

### 2019-12-02 analysis/16_PRL.Rmd
fixed lay icon

--- 2019-11-30 analysis/16_PRL.Rmd
@@
   annotation_custom(bldg, ymin = -20, ymax = 0, xmin = -5.5) +
-  annotation_custom(inc, ymin = -20, ymax = 0, xmin = -3.5) +
+  annotation_custom(lay, ymin = -20, ymax = 0, xmin = -3.5) +
   annotation_custom(inc, ymin = -20, ymax = 0, xmin = -1.5) +
@@
   annotation_custom(bldg, ymin = -20, ymax = 0, xmin = -5.8) +
-  annotation_custom(inc, ymin = -20, ymax = 0, xmin = -3.8) +
+  annotation_custom(lay, ymin = -20, ymax = 0, xmin = -3.8) +
   annotation_custom(inc, ymin = -20, ymax = 0, xmin = -1.8) +
@@
 ```
+
+```{r PRL.pit}
+vsd.pit <- readvsd("../results/04_vsd_pit.csv")
+colData.pit <- readcolData("../results/04_colData_pit.csv")
+geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>%  dplyr::select(Name, geneid, entrezid) %>% arrange(Name)
+candidategenes <- c("PRL")
+candidates.pit <- selectcandidatevsds(candidategenes, vsd.pit, colData.pit)
+
+p3 <- ggplot(candidates.pit, aes(x = as.numeric(treatment), y = PRL)) + 
+        geom_smooth(aes(colour = sex)) +
+    geom_boxplot(aes(fill = treatment, alpha = sex)) + 
+    scale_alpha_manual(values = c(0.75,1)) +
+    mytheme() +
+  theme(legend.position = c(0.85,0.2), legend.direction = "horizontal") + 
+  scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+    labs(y = "PRL expression in the pituitary", x = "parental care stage") +
+    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
+                       labels = charlevels) +
+  annotation_custom(control, ymin = 12, ymax = 14, xmin = -7.8) +
+  annotation_custom(bldg, ymin = 12, ymax = 14, xmin = -5.8) +
+  annotation_custom(lay, ymin = 12, ymax = 14, xmin = -3.8) +
+  annotation_custom(inc, ymin = 12, ymax = 14, xmin = -1.8) +
+  annotation_custom(inc, ymin = 12, ymax = 14, xmin = 0.6) +
+  annotation_custom(inc, ymin = 12, ymax = 14, xmin = 2.4) +
+  annotation_custom(hatch, ymin = 12, ymax = 14, xmin = 4.4) +
+  annotation_custom(nestling, ymin = 12, ymax = 14, xmin = 6.4) +
+  annotation_custom(nestling, ymin = 12, ymax = 14, xmin = 8.4)  + 
+  ylim(c(13,24)) +
+  guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) 
+p3
+```

### 2019-12-02 analysis/16_PRL.Rmd
now with music notes and icon.R file

@@
 library(grid)
+library("ggimage")
 
@@
 source("../R/functions.R")
+source("../R/icons.R")
 
@@
 hormones$treatment <- factor(hormones$treatment, levels = alllevels)
-```
-
-```{r icons}
-characterization <- png::readPNG("../figures/images/DoveParentsRNAseq_charicons.png")
-characterization <- ggdraw() +  draw_image(characterization, scale = 1)
-
-birds <- data.frame(iconpath=list.files("../figures/images/icons", full.names = T), stringsAsFactors = F)
-birds$icons <- sapply(strsplit(as.character(birds$iconpath),'../figures/images/icons/birds_'), "[", 2)
-
-control <- png::readPNG(birds$iconpath[3])
-control <-  grid::rasterGrob(control, interpolate=TRUE)
-
-bldg <- png::readPNG(birds$iconpath[1])
-bldg <-  grid::rasterGrob(bldg, interpolate=TRUE)
-
-lay <- png::readPNG(birds$iconpath[6])
-lay <-  grid::rasterGrob(lay, interpolate=TRUE)
-
-inc <- png::readPNG(birds$iconpath[5])
-inc <-  grid::rasterGrob(inc, interpolate=TRUE)
-
-hatch <- png::readPNG(birds$iconpath[4])
-hatch <-  grid::rasterGrob(hatch, interpolate=TRUE)
-
-nestling <- png::readPNG(birds$iconpath[2])
-nestling <-  grid::rasterGrob(nestling, interpolate=TRUE)
 ```
@@
 p1
+
+meanPRL <- hormones %>% 
+    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
+    droplevels() %>% 
+  dplyr::group_by(treatment) %>%
+  dplyr::summarise(m = mean(plasma_conc), 
+                   se = sd(plasma_conc)/sqrt(length(plasma_conc))) %>%
+  dplyr::mutate(m = round(m,0)) 
+meanPRL <- left_join(meanPRL, birds)
+meanPRL$treatment <- factor(meanPRL$treatment, levels = alllevels)
+meanPRL
+
+ggplot(meanPRL, aes(treatment, m)) +
+  geom_errorbar(aes(ymin=m-se, ymax=m+se), width=.1) +
+   geom_image(aes(image=iconpath), size=.12)  +
+  labs(x = "parental care stages", y = "Prolactin (ng/mL)")
+
+ggplot(meanPRL, aes(treatment, m)) +
+   geom_image(aes(image=music), size=.12)  +
+  labs(x = "parental care stages", y = "Prolactin (ng/mL)")
 
@@
 colData.pit <- readcolData("../results/04_colData_pit.csv")
+
 geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>%  dplyr::select(Name, geneid, entrezid) %>% arrange(Name)
@@
 p3
+
+candidates.pit
+
+meanPRL <- candidates.pit %>% 
+    droplevels() %>% 
+  dplyr::group_by(treatment) %>%
+  dplyr::summarise(m = mean(PRL), 
+                   se = sd(PRL)/sqrt(length(PRL))) %>%
+  dplyr::mutate(m = round(m,0)) 
+meanPRL <- left_join(meanPRL, birds)
+meanPRL$treatment <- factor(meanPRL$treatment, levels = alllevels)
+meanPRL
+
+ggplot(meanPRL, aes(treatment, m)) +
+  geom_errorbar(aes(ymin=m-se, ymax=m+se), width=.1) +
+   geom_image(aes(image=iconpath), size=.12)  +
+  labs(x = "parental care stages", y = "Pituitary PRL expression")
+
+ggplot(meanPRL, aes(treatment, m)) +
+   geom_image(aes(image=music), size=.12)  +
+  labs(x = "parental care stages", y = "Pituitary PRL expression")
 ```

### 2019-12-02 R/icons.R

@@ -1,0 +1,43 @@
+# individual icons
+library(tidyverse)
+library(png)
+library(grid)
+
+birds <- data.frame(iconpath=list.files("../figures/images/icons", full.names = T), stringsAsFactors = F)
+birds$icons <- sapply(strsplit(as.character(birds$iconpath),'../figures/images/icons/DoveParentsRNAseq_'), "[", 2)
+
+iconstreatment <- data.frame(treatment = c("control", "bldg", "lay", 
+                                          "inc.d3", "inc.d9", "inc.d17", 
+                                          "hatch", "n5", "n9",
+                                          "m.inc.d3" ,  "m.inc.d9" , "m.inc.d17" , "m.n2",
+                                          "m.inc.d8" , "prolong" , "extend") ,
+                             icons = c("control.png", "bldg.png", "lay.png",
+                                       "incubation.png", "incubation.png", "incubation.png",
+                                       "hatch.png", "chicklings.png", "chicklings.png", 
+                                       "control.png", "control.png", "control.png", "control.png",
+                                       "hatch.png", "incubation.png","hatch.png"))
+iconstreatment$music <- "https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcS9zgaht3nmVL4moYqz0uKBhYefpTgYrX69_qyL8kmznImDRk-KRw&s"
+
+birds <- left_join(iconstreatment, birds, by = "icons")
+
+control <- png::readPNG(birds$iconpath[1])
+control <-  grid::rasterGrob(control, interpolate=TRUE)
+
+bldg <- png::readPNG(birds$iconpath[2])
+bldg <-  grid::rasterGrob(bldg, interpolate=TRUE)
+
+lay <- png::readPNG(birds$iconpath[3])
+lay <-  grid::rasterGrob(lay, interpolate=TRUE)
+
+inc <- png::readPNG(birds$iconpath[4])
+inc <-  grid::rasterGrob(inc, interpolate=TRUE)
+
+hatch <- png::readPNG(birds$iconpath[7])
+hatch <-  grid::rasterGrob(hatch, interpolate=TRUE)
+
+nestling <- png::readPNG(birds$iconpath[8])
+nestling <-  grid::rasterGrob(nestling, interpolate=TRUE)
+
+# all characterization, for volcano plots
+characterization <- png::readPNG("../figures/images/DoveParentsRNAseq_charicons.png")
+characterization <- ggdraw() +  draw_image(characterization, scale = 1)

### 2019-12-02 analysis/07_hormones.Rmd
now with icons on manip images

@@
 source("../R/themes.R")  # load custom themes and color palletes
+source("../R/icons.R")
 
@@
 hormones <- rbind(prolactin, PETC)
+hormones <- left_join(hormones, birds)
+
 hormones$treatment <- factor(hormones$treatment, levels = alllevels)
@@
         geom_smooth(aes(colour = sex)) +
-    geom_boxplot(aes(outlier.colour = treatment, fill = treatment, alpha = sex)) +
+    geom_boxplot(aes(fill = treatment, alpha = sex)) +
     mytheme() +
@@
     labs(y = "PRL (ng/mL)", x = NULL) +
-    annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -2, ymax = 0, alpha = 0.25) +
-    annotate("rect", xmin = 2.6, xmax = 4.4, ymin = -2, ymax = 0, alpha = 0.5) +
-    annotate("rect", xmin = 4.6, xmax = 6.4, ymin = -2, ymax = 0, alpha = 0.75)  +
-    annotate("rect", xmin = 6.6, xmax = 8.4, ymin = -2, ymax = 0, alpha = 1)
+    ylim(c(-16,100)) +
+
+   annotation_custom(inc, xmax = 1.6, ymin = -14, ymax = 5) +
+   annotation_custom(bldg, xmax = 3.6, ymin = -14, ymax = 5) +
+   annotation_custom(inc, xmax = 5.6, ymin = -14, ymax = 5) +
+   annotation_custom(bldg, xmax = 7.6, ymin = -14, ymax = 5) +
+  annotation_custom(hatch, xmax = 9.6, ymin = -14, ymax = 5) +
+  annotation_custom(bldg, xmax = 11.6, ymin = -14, ymax = 5) +
+  annotation_custom(nestling, xmax = 13.6, ymin = -14, ymax = 5) +
+  annotation_custom(bldg, xmax = 15.6, ymin = -14, ymax = 5) +
+
+    annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -16, ymax = -14, alpha = 0.25) +
+    annotate("rect", xmin = 2.6, xmax = 4.4, ymin = -16, ymax = -14, alpha = 0.5) +
+    annotate("rect", xmin = 4.6, xmax = 6.4, ymin = -16, ymax = -14, alpha = 0.75)  +
+    annotate("rect", xmin = 6.6, xmax = 8.4, ymin = -16, ymax = -14, alpha = 1)
 
@@
     labs(y = "PRL (ng/mL)", x = NULL) +
-    annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -5, ymax = -3, alpha = 0.33) +
-    annotate("rect", xmin = 4.6, xmax = 5.4, ymin = -5, ymax = -3, alpha = 0.33) +
-    annotate("rect", xmin = 2.6, xmax = 5.4, ymin = -2, ymax = 0, alpha = 0.66) +
-    annotate("rect", xmin = 4.6, xmax = 7.4, ymin = 1, ymax = 3, alpha = 1) 
+  ylim(c(-12,125)) +
+   annotation_custom(inc, xmax = 1.6, ymin = -10, ymax = 10) +
+   annotation_custom(hatch, xmax = 3.6, ymin = -10, ymax = 10) +
+   annotation_custom(inc, xmax = 5.6, ymin = -10, ymax = 10) +
+   annotation_custom(inc, xmax = 7.6, ymin = -10, ymax = 10) +
+  annotation_custom(hatch, xmax = 9.6, ymin = -10, ymax = 10) +
+  annotation_custom(hatch, xmax = 11.6, ymin = -10, ymax = 10) +
+  annotation_custom(nestling, xmax = 13.6, ymin = -10, ymax = 10) +
+
+  annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -12, ymax = -10, alpha = 0.33) +
+    annotate("rect", xmin = 4.6, xmax = 5.4, ymin = -12, ymax = -10, alpha = 0.33) +
+    annotate("rect", xmin = 2.6, xmax = 5.4, ymin = -10, ymax = -8, alpha = 0.66) +
+    annotate("rect", xmin = 4.6, xmax = 7.4, ymin = -8, ymax = -6, alpha = 1) 
 ```

### 2019-12-02 analysis/16_PRL.Rmd
now with icons on manip images

@@
    geom_image(aes(image=music), size=.12)  +
-  labs(x = "parental care stages", y = "Prolactin (ng/mL)")
+  labs(x = "parental care stages", y = "Prolactin (ng/mL)")  +
+  mytheme() +
+  ylim(c(-15,80)) +
+  annotation_custom(control, ymin = -20, ymax = 0, xmin = -7.5) +
+  annotation_custom(bldg, ymin = -20, ymax = 0, xmin = -5.5) +
+  annotation_custom(lay, ymin = -20, ymax = 0, xmin = -3.5) +
+  annotation_custom(inc, ymin = -20, ymax = 0, xmin = -1.5) +
+  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 0.5) +
+  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 2.5) +
+  annotation_custom(hatch, ymin = -20, ymax = 0, xmin = 4.5) +
+  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 6.5) +
+  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 8.5)
 
@@
    geom_image(aes(image=music), size=.12)  +
-  labs(x = "parental care stages", y = "Pituitary PRL expression")
+  labs(x = "parental care stages", y = "Pituitary PRL expression") +
+  mytheme() +
+  ylim(c(15,22)) +
+  annotation_custom(control, ymin = 15, ymax = 16.5, xmin = -7.5) +
+  annotation_custom(bldg, ymin = 15, ymax = 16.5, xmin = -5.5) +
+  annotation_custom(lay, ymin = 15, ymax = 16.5, xmin = -3.5) +
+  annotation_custom(inc, ymin = 15, ymax = 16.5, xmin = -1.5) +
+  annotation_custom(inc, ymin = 15, ymax = 16.5, xmin = 0.5) +
+  annotation_custom(inc, ymin = 15, ymax = 16.5, xmin = 2.5) +
+  annotation_custom(hatch, ymin = 15, ymax = 16.5, xmin = 4.5) +
+  annotation_custom(nestling, ymin = 15, ymax = 16.5, xmin = 6.5) +
+  annotation_custom(nestling, ymin = 15, ymax = 16.5, xmin = 8.5) 
 ```

### 2019-12-02 analysis/07_hormones.Rmd
subtle image tweeks

@@
     labs(y = "PRL (ng/mL)", x = NULL) +
-  ylim(c(-12,125)) +
+  ylim(c(-12,100)) +
    annotation_custom(inc, xmax = 1.6, ymin = -10, ymax = 10) +

### 2019-12-02 analysis/03_DESeq2_characterization.Rmd
remove old stuff

@@
 
-## Plot only a subset of DEGs that relate to hypotheses as bar and line graphs
-
-```{r barplots}
-# bars with lines
-g <- plotserialDEGs(DEGs.female_hypothalamus, "Female hypothalamus", "#00A08A")
-h <- plotserialDEGs(DEGs.female_pituitary, "Female pituitary", "#00A08A")
-i <- plotserialDEGs(DEGs.female_gonad, "Female gonad", "#00A08A")
-j <- plotserialDEGs(DEGs.male_hypothalamus, "Male hypothalamus", "#F98400")
-k <- plotserialDEGs(DEGs.male_pituitary, "Male pituitary", "#F98400")
-l <- plotserialDEGs(DEGs.male_gondad, "Male gonad", "#F98400")
-
-mybarplots <- plot_grid(g, h , i , j ,k , l, nrow = 2) 
-mybarplots
-
-pdf("../figures/characterization/barplots.pdf", width = 10, height = 6)
-plot(mybarplots)
-dev.off()
-```
-
-```{r linegraphs}
-
-# lines only, different colors for different sex tissue groups
-FH <- subsetDEGs(DEGs.female_hypothalamus, "Female hypothalamus")
-FP <- subsetDEGs(DEGs.female_pituitary, "Female pituitary")
-FG <- subsetDEGs(DEGs.female_gonad, "Female gonad")
-MH <- subsetDEGs(DEGs.male_hypothalamus, "Male hypothalamus")
-MP <- subsetDEGs(DEGs.male_pituitary, "Male pituitary")
-MG <- subsetDEGs(DEGs.male_gondad, "Male gonad")
-
-specificDEGs <- rbind(FH, FP, FG, MH, MP, MG)
-
-specificDEGs$sextissue <- factor(specificDEGs$sextissue, levels = c("Female hypothalamus", "Female pituitary", "Female gonad",
-                                                                    "Male hypothalamus", "Male pituitary", "Male gonad"))
-mylinegraph <- ggplot(specificDEGs, aes(x=comparison, y=V3, group = sextissue)) +
-    geom_line(aes(color=sextissue)) +
-    geom_point(aes(color=sextissue)) + 
-    theme_minimal(base_size = 10) +
-    #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-    labs(x = "Parental stage transitions", y= "No. of differentially expressed genes ",
-         title = "Gene expression changes during parental care transitions") + 
-    scale_x_discrete(labels=c("Control to\nNest bldg", "Nest bldg. to\nLay", "Lay to\nInc. day 3",
-                              "Inc. day 3 to\nInc. day 9", "Inc. day 9 to\nInc. day 17",
-                              "Inc. day 17 to\nHatch", "Hatch to\nNestling day 5", "Nestling day 5 to\n Nestling day 9")) +
-   theme(legend.title = element_blank(),
-         legend.position = c(0.8, 0.8))
-mylinegraph
-
-pdf("../figures/characterization/linegraphs.pdf", width = 8, height = 4)
-plot(mylinegraph)
-dev.off()
-```
-
-## Calculate and plot principal component
-
-```{r pca, eval = T, cache = TRUE}
-
-FhypPCA <- returnPCAs(dds.female_hypothalamus)
-plotPC12(FhypPCA, "female hypothalamus")
-
-FpitPCA <- returnPCAs(dds.female_pituitary)      
-plotPC12(FpitPCA, "female pituitary")
-
-FgonPCA <- returnPCAs(dds.female_gonad)
-plotPC12(FgonPCA, "female gonad")
-
-MhypPCA <- returnPCAs(dds.male_hypothalamus)
-plotPC12(MhypPCA, "male hypothalamus")
-
-MpitPCA <- returnPCAs(dds.male_pituitary)
-plotPC12(MpitPCA, "male pituitary")
-
-MgonPCA <- returnPCAs(dds.male_gondad)
-plotPC12(MgonPCA, "male gonad")
-```
-
-## heamap of candidate genes
-
-```{r pheatmap, eval = T, cache = TRUE}
-colDataFH <- subsetcolData(c.colData, "female_hypothalamus")
-colDataFP <- subsetcolData(c.colData, "female_pituitary")
-colDataFG <- subsetcolData(c.colData, "female_gonad")
-colDataMH <- subsetcolData(c.colData, "male_hypothalamus")
-colDataMP <- subsetcolData(c.colData, "male_pituitary")
-colDataMG <- subsetcolData(c.colData, "male_gonad")
-
-makepheatmap(dds.female_hypothalamus, colDataFH, "female hypothalamus")
-makepheatmap(dds.female_pituitary, colDataFP, "female pituitary")
-makepheatmap(dds.female_gonad, colDataFG, "female gonad")
-makepheatmap(dds.male_hypothalamus, colDataMH,  "male hypothalamus")
-makepheatmap(dds.male_pituitary, colDataMP, "male pituitary")
-makepheatmap(dds.male_gondad, colDataMG, "male gonad")
-```
 ## plot candidate genes
@@
 plotcandidates(dds.male_gondad,colDataMG, "male gonad")
-```
-
-## plot prolactin
-
-```{r prolactin, eval = T}
-a <- plotprolactin(dds.female_hypothalamus, colDataFH,"female hypothalamus")
-b <- plotprolactin(dds.female_pituitary, colDataFP, "female pituitary")
-c <- plotprolactin(dds.female_gonad, colDataFG,"female gonad")
-d <- plotprolactin(dds.male_hypothalamus, colDataMH,"male hypothalamus")
-e <- plotprolactin(dds.male_pituitary,colDataMP, "male pituitary")
-f <- plotprolactin(dds.male_gondad,colDataMG, "male gonad")
-
-mylegend <- get_legend(f)
-
-plots <- plot_grid(a + theme(legend.position = "none"),
-          b + theme(legend.position = "none"),
-          c + theme(legend.position = "none"),
-          d + theme(legend.position = "none"), 
-          e + theme(legend.position = "none"),
-          f + theme(legend.position = "none"))
-plot_grid(plots, mylegend, nrow = 2, rel_heights = c(1,0.1))
-
-```
-
-## correlation heat maps
-
-```{r correlationheatmaps, eval = F}
-plotcorrelationheatmaps(dds.female_hypothalamus, colData.female_hypothalamus ,"female hypothalamus: hatch, nest bldg., lay and day-before-hatch cluster")
-plotcorrelationheatmaps(dds.female_pituitary, colData.female_pituitary, "female pituitary: clustering suggests an internal clock effect")
-plotcorrelationheatmaps(dds.female_gonad, colData.female_gonad,  "female gonad: highly correlated across incubation and early nestling care")
-plotcorrelationheatmaps(dds.male_hypothalamus, colData.male_hypothalamus, "male hypothalamus: 100% correlated across most parental care stages")
-plotcorrelationheatmaps(dds.male_pituitary, colData.male_pituitary, "male pituitary: non-parental males are most different")
-plotcorrelationheatmaps(dds.male_gondad, colData.male_gondad, "male gonad: non-parental males are most different")
 ```

### 2019-12-02 analysis/03_DESeq2_characterization.Rmd
now with manip volcanos

@@
 ```{r plotDEGs}
-plot.volcano <- function(data, whichfactor, up, down, mycolors){
-
-  volcano <- data %>%
-    ggplot(aes(x = lfc, y = logpadj)) + 
-    geom_point(aes(color = direction, shape = tissue), size = 2, 
-               alpha = 0.75, na.rm = T) + 
-    theme_minimal() +
-    scale_color_manual(values = mycolors,
-                       name = " ",
-                       drop = FALSE) +
-    ylim(c(0,30)) +  
-    xlim(c(-8,8)) +
-    labs(y = "-log10(p)", x = " ")  +
-    theme(legend.position = "none",
-          legend.direction = "horizontal",
-          legend.spacing.x = unit(-0.1, 'cm'),
-          legend.margin=margin(t=-0.25, r=0, b=0, l=0, unit="cm"),
-          panel.grid = element_blank()) +
-    scale_shape_manual(values = myshapes)
-  return(volcano)
-}
-
 a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
@@
 
-## males and females
+## males
 

### 2019-12-02 analysis/04_DESeq2_sexes_manipulate.Rmd
now with manip volcanos

@@
 # "extend"    "m.inc.d17" "m.inc.d3"  "m.inc.d8"  "m.inc.d9"  "m.n2"      "prolong"
-c.colData <- c.colData %>%
-    mutate(treatment = fct_recode(treatment,
-                            "extend" = "extend",
-                            "prolong" = "prolong",
-                            "early" = "m.inc.d8",
-                            "remove.d03" = "m.inc.d3",
-                            "remove.d09" = "m.inc.d9",
-                            "remove.d17" = "m.inc.d17",
-                            "remove.d20" = "m.n2"))
 
-c.colData$treatment <- factor(c.colData$treatment, levels = maniplevels)
+c.colData$treatment <- factor(c.colData$treatment, levels = maniplevels1)
 levels(c.colData$treatment)
@@
 str(c.colData)
+
+geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+head(geneinfo)
 ```
@@
 ```
+
+## volcano plots
+
+```{r pitdfs, eval=T}
+dds.pituitary <- subsetDESeq2(c.colData,  c.countData, c("female_pituitary","male_pituitary"))
+dds.pituitary$treatment
+```
+
+```{r pitdfs2, eval=T}
+createDEGdfmanip <- function(mydds, whichfactor, up, down, mytissue){
+  res <- results(mydds, contrast =c(whichfactor, up, down),
+                 independentFiltering = T, alpha = 0.1)
+   data <- data.frame(entrezid = row.names(res),
+                     padj = res$padj, 
+                     logpadj = -log10(res$padj),
+                     lfc = res$log2FoldChange,
+                     tissue = mytissue)
+  data <- na.omit(data)
+
+  data <- data %>%
+    dplyr::mutate(direction = ifelse(data$lfc > 0 & data$padj < 0.1, 
+                                     yes = up, no = ifelse(data$lfc < 0 & data$padj < 0.1, 
+                                                 yes = down, no = "NS")))
+  data$direction <- factor(data$direction, levels = c(down, "NS", up))
+
+  data <- left_join(data, geneinfo) %>%
+    dplyr::mutate(gene = Name) %>%
+    dplyr::select(gene, lfc, padj, logpadj, direction, tissue) %>%
+    dplyr::arrange(desc(lfc))
+
+  DEGs <- data %>% filter(direction != "NS")
+  print(str(DEGs))
+
+  myfilename = paste("../results/04_DEGs", mytissue, down,up, "csv", sep = ".")
+
+  # write dataframe of only significant genes
+  write.csv(DEGs, myfilename, row.names = F)
+
+  # return data frome with all data, included NS genes
+  return(data)
+}
+
+m.inc.d3.m.inc.d9 <- createDEGdfmanip(dds.pituitary, "treatment", "m.inc.d9", "m.inc.d3",  "pituitary")
+m.inc.d9.m.inc.d17 <- createDEGdfmanip(dds.pituitary, "treatment", "m.inc.d17",  "m.inc.d9", "pituitary")
+m.inc.d17.m.n2 <- createDEGdfmanip(dds.pituitary, "treatment", "m.n2",  "m.inc.d17", "pituitary") 
+m.n2.m.inc.d8 <- createDEGdfmanip(dds.pituitary, "treatment", "m.inc.d8",  "m.n2", "pituitary") 
+m.inc.d8.prolong <- createDEGdfmanip(dds.pituitary,  "treatment", "prolong", "m.inc.d8",  "pituitary") 
+prolong.extend <- createDEGdfmanip(dds.pituitary, "treatment", "extend", "prolong",  "pituitary")
+```
+
+```{r pitplot, eval = T}
+a <- plot.volcano(m.inc.d3.m.inc.d9, "treatment", "m.inc.d9", "m.inc.d3",  colorsvolcano)
+b <- plot.volcano(m.inc.d9.m.inc.d17, "treatment", "m.inc.d17", "m.inc.d9",  colorsvolcano)
+c <- plot.volcano(m.inc.d17.m.n2, "treatment",  "m.n2", "m.inc.d17", colorsvolcano) 
+d <- plot.volcano(m.n2.m.inc.d8, "treatment", "m.inc.d8", "m.n2",  colorsvolcano) 
+e <- plot.volcano(m.inc.d8.prolong, "treatment", "prolong", "m.inc.d8",  colorsvolcano) 
+f <- plot.volcano(prolong.extend, "treatment", "extend",  "prolong",colorsvolcano)
+
+pituitary <- plot_grid(a + 
+                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
+                       labs(y = "pituitary \n -log10(p)"),
+          b + theme(axis.title = element_blank(), axis.text = element_blank()),
+          c + theme(axis.title = element_blank(), axis.text = element_blank()),
+          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          e + theme(axis.title = element_blank(), axis.text = element_blank()),
+          f + theme(axis.title = element_blank(), axis.text = element_blank()),
+          nrow = 1,  rel_widths = c(1.75,1,1,1,1,1))
+pituitary
+
+background <- png::readPNG("../figures/images/DoveParentsRNAseq_manipvolcanos.png")
+icons <- ggdraw() +  draw_image(background, scale = 1)
+
+plot_grid(icons, pituitary, nrow = 2, rel_heights = c(1,1))
+```

### 2019-12-02 analysis/04_DESeq2_sexes_manipulate.Rmd
better volcanos

@@
 
-```{r pitplot, eval = T}
+```{r pitplot, eval = T, fig.height=3}
 a <- plot.volcano(m.inc.d3.m.inc.d9, "treatment", "m.inc.d9", "m.inc.d3",  colorsvolcano)

### 2019-12-02 analysis/06_DESeq2_ALL.Rmd
now with volcano plot volcanos

---
title: "06_DESeqALL"
output: html_document
---

```{r setup, include = FALSE}

library(tidyverse)
library(DESeq2)
library(cowplot)
library(pheatmap)
library(viridis)
library(forcats)
library(caret) # LDA analysis
library(MASS) # LDA analysis
library(ggtext) # for pretty plots
library(ctsGE)

library(BiocParallel)
register(MulticoreParam(6))

source("../R/functions.R")  # load custom functions 
source("../R/themes.R")  # load custom themes and color palletes
source("../R/icons.R")  # for images

knitr::opts_chunk$set(fig.path = '../figures/DESeqAll/', cache = T,  message=F, comment=F, warning=F)
```

## DEseq2 analysis on characterization. Looking at treatment AND sex for each tissue

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

# craete variable that will be critical for subset later on
colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
colData$sextissue <- as.factor(paste(colData$sex, colData$tissue, sep = "_"))
colData$treatment <- factor(colData$treatment, levels = alllevels3)
colData$tissue <- factor(colData$tissue, levels = tissuelevels)
levels(colData$treatment)
```

```{r pitdfs, eval=T}
dds.pituitary <- subsetDESeq2(colData,  countData, c("female_pituitary","male_pituitary"))
```

```{r  eval=T}
inc.d3.m.inc.d3 <- createDEGdfsave(dds.pituitary, "treatment", "m.inc.d3", "inc.d3", "pituitary")
```

```{r  eval=T}
inc.d9.m.inc.d9 <- createDEGdfsave(dds.pituitary, "treatment", "m.inc.d9", "inc.d9", "pituitary")
```

```{r  eval=T}
inc.d17.m.inc.d17 <- createDEGdfsave(dds.pituitary, "treatment", "m.inc.d17", "inc.d17", "pituitary")
```

```{r  eval=T}
hatch.m.n2 <- createDEGdfsave(dds.pituitary, "treatment", "m.n2", "hatch", "pituitary")
```

```{r  eval=T}
inc.d9.m.inc.d8 <- createDEGdfsave(dds.pituitary, "treatment", "m.inc.d8", "inc.d9", "pituitary")
```

```{r  eval=T}
hatch.m.inc.d8 <- createDEGdfsave(dds.pituitary, "treatment", "m.inc.d8", "hatch", "pituitary")
```

```{r  eval=T}
inc.d17.prolong <- createDEGdfsave(dds.pituitary, "treatment", "prolong", "inc.d17", "pituitary")
```

```{r  eval=T}
hatch.prolong <- createDEGdfsave(dds.pituitary, "treatment", "prolong", "hatch", "pituitary")
```

```{r  eval=T}
hatch.extend <- createDEGdfsave(dds.pituitary, "treatment", "extend", "hatch", "pituitary")
```

```{r  eval=T}
n5.extend <- createDEGdfsave(dds.pituitary, "treatment", "extend", "n5", "pituitary")
```

```{r pitplot, eval = T}
a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano)
b <- plot.volcano(inc.d9.m.inc.d9, "treatment",  "m.inc.d9", "inc.d9" , colorsvolcano)
c <- plot.volcano(inc.d17.m.inc.d17 , "treatment",  "m.inc.d17", "inc.d17" , colorsvolcano) 
d <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano) 
e <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano)
f <- plot.volcano(hatch.m.inc.d8 , "treatment",    "m.inc.d8", "hatch"  , colorsvolcano) 
g <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano) 
h <- plot.volcano(hatch.prolong , "treatment",  "prolong", "hatch", colorsvolcano) 
i <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano) 
j <- plot.volcano(n5.extend , "treatment",  "extend", "n5", colorsvolcano)

pituitary <- plot_grid(a +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                       labs(y = "pituitary \n -log10(p)"),
          b + theme(axis.title = element_blank(), axis.text = element_blank()),
          c + theme(axis.title = element_blank(), axis.text = element_blank()),
          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
          e + theme(axis.title = element_blank(), axis.text = element_blank()),
          f + theme(axis.title = element_blank(), axis.text = element_blank()),
          g + theme(axis.title = element_blank(), axis.text = element_blank()),
          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
          i + theme(axis.title = element_blank(), axis.text = element_blank()), 
          j + theme(axis.title = element_blank(), axis.text = element_blank()), 
          nrow = 2)
pituitary
```

### 2019-12-02 R/functions.R

@@ -773,7 +773,7 @@ plot.volcano <- function(data, whichfactor, up, down, mycolors){
   
   volcano <- data %>%
     ggplot(aes(x = lfc, y = logpadj)) + 
-    geom_point(aes(color = direction, shape = tissue), size = 2, 
+    geom_point(aes(color = direction, shape = tissue), size = 1, 
                alpha = 0.75, na.rm = T) + 
     theme_minimal() +
     scale_color_manual(values = mycolors,
@@ -790,3 +790,75 @@ plot.volcano <- function(data, whichfactor, up, down, mycolors){
     scale_shape_manual(values = myshapes)
   return(volcano)
 }
+
+
+# made for beta testing
+# usaage
+# tempmanip <- subsetDESeq3(colData, countData, c("female_pituitary","male_pituitary"), 
+#                                                  c("m.inc.d3", "inc.d3", "inc.d9", "m.inc.d9"))
+
+subsetDESeq3 <- function(colData, countData, eachgroup, eachtreatment){
+  
+  colData <- colData %>%
+    dplyr::filter(sextissue %in% eachgroup,
+                  treatment %in% eachtreatment) %>%
+    droplevels()
+  row.names(colData) <- colData$V1
+  
+  # save counts that match colData
+  savecols <- as.character(colData$V1) 
+  savecols <- as.vector(savecols) 
+  countData <- countData %>% dplyr::select(one_of(savecols)) 
+  
+  # assert that row and col lenghts are equal
+  stopifnot(ncol(countData) == nrow(colData))
+  
+  dds <- DESeqDataSetFromMatrix(countData = countData,
+                                colData = colData,
+                                design = ~ treatment )
+  
+  print(dds)
+  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
+  print(dim(dds))
+  
+  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
+  return(dds)
+}
+
+
+## for DESeqALL
+createDEGdfsave <- function(mydds, whichfactor, up, down, mytissue){
+  
+  res <- results(mydds, contrast = c(whichfactor, up, down), independentFiltering = T, alpha = 0.1)
+  
+  data <- data.frame(entrezid = row.names(res),
+                     padj = res$padj, 
+                     logpadj = -log10(res$padj),
+                     lfc = res$log2FoldChange,
+                     tissue = mytissue)
[…]

### 2019-12-02 analysis/09_manipvolcanos.Rmd
now with manip volcanos

---
title: "Manip Volcanos"
output: md_document
---

```{r}
library(tidyverse)
library(cowplot)

source("../R/functions.R")  # load custom functions 
source("../R/themes.R")  # load custom themes and color palletes
source("../R/icons.R")  # for images

knitr::opts_chunk$set(fig.path = '../figures/manipulation/',message=F, warning=FALSE)
```

```{r  eval=T}

inc.d3.m.inc.d3 <- read_csv("../results/DESeqAll.pituitary.inc.d3.m.inc.d3.csv")
inc.d9.m.inc.d9 <- read_csv("../results/04_DEGs.pituitary.m.inc.d3.m.inc.d9.csv")
inc.d17.m.inc.d17 <- read_csv("../results/DESeqAll.pituitary.inc.d17.m.inc.d17.csv")
hatch.m.n2 <- read_csv("../results/DESeqAll.pituitary.hatch.m.n2.csv")
inc.d9.m.inc.d8 <- read_csv("../results/DESeqAll.pituitary.inc.d9.m.inc.d8.csv")
hatch.m.inc.d8 <- read_csv("../results/DESeqAll.pituitary.hatch.m.inc.d8.csv")
inc.d17.prolong <- read_csv("../results/DESeqAll.pituitary.inc.d17.prolong.csv")
hatch.prolong <- read_csv("../results/DESeqAll.pituitary.hatch.prolong.csv")
hatch.extend <- read_csv("../results/DESeqAll.pituitary.hatch.extend.csv")
n5.extend <- read_csv("../results/DESeqAll.pituitary.n5.extend.csv")

```

```{r volcanos}
a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano) + 
  annotation_custom(lay, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(control, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
b <- plot.volcano(inc.d9.m.inc.d9, "treatment",  "m.inc.d9", "inc.d9" , colorsvolcano)+ 
  annotation_custom(inc, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(control, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
c <- plot.volcano(inc.d17.m.inc.d17 , "treatment",  "m.inc.d17", "inc.d17" , colorsvolcano) + 
  annotation_custom(inc, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(control, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
d <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano) + 
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(control, ymin = 20, ymax = 30, xmin = 2, xmax = 8)

e <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano)  +
  annotation_custom(inc, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
f <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  +
  annotation_custom(inc, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(inc, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
g <- plot.volcano(hatch.prolong , "treatment",  "prolong", "hatch", colorsvolcano)  +
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(inc, ymin = 20, ymax = 30, xmin = 2, xmax = 8)

h <- plot.volcano(hatch.m.inc.d8 , "treatment",    "m.inc.d8", "hatch"  , colorsvolcano)  +
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
i <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano)  +
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = 2) + 
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
j <- plot.volcano(n5.extend , "treatment",  "extend", "n5", colorsvolcano)  +
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(nestling, ymin = 20, ymax = 30, xmin = 2, xmax = 8)

remove <- plot_grid(a +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                       labs(y = "pituitary \n -log10(p)", subtitle = "1771 DEGs"),
          b + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "277 DEGs"),
          c + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "3909 DEGs"),
          d + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "3162 DEGs"), 
          nrow = 1, rel_widths = c(1.2,1,1,1),
          labels = c("a","b","c","d"), label_size = 9)
remove

manipinc <- plot_grid(e +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                       labs(y = "pituitary \n -log10(p)", subtitle = "120 DEGs"),
          f + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "1007 DEGs"),
          g + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "731 DEGs"),
          nrow = 1, rel_widths = c(1.2,1,1),
          labels = c("e","f","g"," "), label_size = 9)

maniphatch <- plot_grid( h +  labs(subtitle = "4113 DEGs"), 
          i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) + labs(subtitle = "1894 DEGs"), 
          j + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) + labs(subtitle = "256 DEGs"), 
          nrow = 1, rel_widths = c(1.2,1,1),
          labels = c("h","i","j"," "), label_size = 9)

plot_grid(manipinc, maniphatch, nrow = 2)
```

### 2019-12-02 R/themes.R
now with manip volcanos

@@ -58,6 +58,11 @@ tissuelevels <- c("hypothalamus", "pituitary", "gonads")
 
 
 charlevelsvocano <- c("control", "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9", "NS")
+alllevels3volcano <- c("control", "bldg", "lay", "inc.d3", "inc.d9", 
+                      "inc.d17", "hatch", "n5", "n9",
+                      "m.inc.d3" ,  "m.inc.d9" , "m.inc.d17" , "m.n2",
+                      "m.inc.d8" , "prolong" , "extend",
+                      "NS")
 
 
 ## comparisons

### 2019-12-03 analysis/16_PRL.Rmd
now with musical prolactin

@@
 ```
+
+## prolactin manip
+
+```{r PRL.pit.maip}
+vsd.pit <- readvsd("../results/06_pitallvsd.csv")
+vsd.pit$entrezid <- row.names(vsd.pit)
+
+colData.pit <- read_csv("../metadata/00_samples.csv") %>% 
+  filter(tissue == "pituitary") %>% 
+  mutate(sample = V1)
+
+geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>%  
+  dplyr::select(Name, geneid, entrezid) %>% arrange(Name)
+
+candidategenes <- c("PRL")
+
+candidates.pit <- selectcandidatevsds(candidategenes, vsd.pit, colData.pit)
+candidates.pit$treatment <- factor(candidates.pit$treatment, levels = alllevels)
+
+p3 <- ggplot(candidates.pit, aes(x = treatment, y = PRL)) + 
+       geom_smooth(aes(colour = sex)) +
+    geom_boxplot(aes(fill = treatment)) +
+    scale_alpha_manual(values = c(0.75,1)) +
+    mytheme() +
+   theme(legend.position = "none", legend.direction = "horizontal") + 
+   scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+    labs(y = "Pituitary PRL", x = NULL) +
+    scale_x_discrete(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16)) +
+  scale_fill_manual(values = colorscharmaip2) +
+  ylim(c(16,22))
+p3
+
+meanPRL <- candidates.pit %>% 
+    droplevels() %>% 
+  dplyr::group_by(treatment) %>%
+  dplyr::summarise(m = median(PRL), 
+                   se = sd(PRL)/sqrt(length(PRL))) %>%
+  dplyr::mutate(m = round(m,1)) 
+meanPRL <- left_join(meanPRL, birds)
+meanPRL$treatment <- factor(meanPRL$treatment, levels = alllevels)
+meanPRL
+
+p4 <- ggplot(meanPRL, aes(treatment, m)) +
+   geom_errorbar(aes(ymin=m-se, ymax=m+se, color = treatment), width=.1) +
+   geom_image(aes(image = iconpath), size = 0.06)  +
+   geom_image(aes(image = music , y = m + 0.5), size = 0.04) +
+   labs(x = "parental care stages", y = "Pituitary PRL") +
+   mytheme() +
+   ylim(c(16,22)) +
+   theme(axis.text.x = element_text(angle = 45)) +
+  theme(legend.position = "none") +
+  scale_color_manual(values = colorscharmaip2) +
+  geom_hline(yintercept=16.8) +
+  geom_hline(yintercept=17.8) +
+  geom_hline(yintercept=18.8) +
+  geom_hline(yintercept=19.8) +
+  geom_hline(yintercept=20.8) 
+p4
+
+```

### 2019-12-03 analysis/06_DESeq2_ALL.Rmd
now with musical prolactin

@@
 
-## DEseq2 analysis on characterization. Looking at treatment AND sex for each tissue
+# Warning: This script takes a very long time to run.
+
+DESeq2 was not designed to run on 300 samples. But, I really like it, so I do it anyways. Some of these commands take like 15 min to run using 6 cores.
+
+## DEseq2 on all chaacterization and manipulations
 
@@
 dds.pituitary <- subsetDESeq2(colData,  countData, c("female_pituitary","male_pituitary"))
+```
+
+```{r pitvsd, eval=T}
+vsd <- as.data.frame(assay(vst(dds.pituitary, blind=FALSE)))
+write.csv(vsd, "../results/06_pitallvsd.csv")
 ```
@@
 
-```{r pitplot, eval = T}
-a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano)
-b <- plot.volcano(inc.d9.m.inc.d9, "treatment",  "m.inc.d9", "inc.d9" , colorsvolcano)
-c <- plot.volcano(inc.d17.m.inc.d17 , "treatment",  "m.inc.d17", "inc.d17" , colorsvolcano) 
-d <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano) 
-e <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano)
-f <- plot.volcano(hatch.m.inc.d8 , "treatment",    "m.inc.d8", "hatch"  , colorsvolcano) 
-g <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano) 
-h <- plot.volcano(hatch.prolong , "treatment",  "prolong", "hatch", colorsvolcano) 
-i <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano) 
-j <- plot.volcano(n5.extend , "treatment",  "extend", "n5", colorsvolcano)
+```{r volcanos}
+a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano) + 
+  annotation_custom(lay, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
+  annotation_custom(bldg, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
+b <- plot.volcano(inc.d9.m.inc.d9, "treatment",  "m.inc.d9", "inc.d9" , colorsvolcano)+ 
+  annotation_custom(inc, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
+  annotation_custom(bldg, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
+c <- plot.volcano(inc.d17.m.inc.d17 , "treatment",  "m.inc.d17", "inc.d17" , colorsvolcano) + 
+  annotation_custom(inc, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
+  annotation_custom(bldg, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
+d <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano) + 
+  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
+  annotation_custom(bldg, ymin = 20, ymax = 30, xmin = 2, xmax = 8)
 
-pituitary <- plot_grid(a +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
-                       labs(y = "pituitary \n -log10(p)"),
-          b + theme(axis.title = element_blank(), axis.text = element_blank()),
-          c + theme(axis.title = element_blank(), axis.text = element_blank()),
-          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          e + theme(axis.title = element_blank(), axis.text = element_blank()),
-          f + theme(axis.title = element_blank(), axis.text = element_blank()),
-          g + theme(axis.title = element_blank(), axis.text = element_blank()),
-          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          i + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          j + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          nrow = 2)
-pituitary
+e <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano)  +
+  annotation_custom(inc, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
+  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
+f <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  +
+  annotation_custom(inc, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
+  annotation_custom(inc, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
+g <- plot.volcano(hatch.prolong , "treatment",  "prolong", "hatch", colorsvolcano)  +
+  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
+  annotation_custom(inc, ymin = 20, ymax = 30, xmin = 2, xmax = 8)
+
+h <- plot.volcano(hatch.m.inc.d8 , "treatment",    "m.inc.d8", "hatch"  , colorsvolcano)  +
+  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
+  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
+i <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano)  +
+  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = 2) + 
+  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
+j <- plot.volcano(n5.extend , "treatment",  "extend", "n5", colorsvolcano)  +
+  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
+  annotation_custom(nestling, ymin = 20, ymax = 30, xmin = 2, xmax = 8)
+
+remove <- plot_grid(a +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
+                       labs(y = "pituitary \n -log10(p)", subtitle = "1771 DEGs"),
+          b + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "277 DEGs"),
+          c + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "3909 DEGs"),
+          d + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "3162 DEGs"), 
+          nrow = 1, rel_widths = c(1.2,1,1,1),
+          labels = c("a","b","c","d"), label_size = 9)
+remove
+
+manipinc <- plot_grid(e +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
+                       labs(y = "pituitary \n -log10(p)", subtitle = "120 DEGs"),
+          f + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "1007 DEGs"),
+          g + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "731 DEGs"),
+          nrow = 1, rel_widths = c(1.2,1,1),
+          labels = c("e","f","g"," "), label_size = 9)
+
+maniphatch <- plot_grid( h +  labs(subtitle = "4113 DEGs"), 
+          i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) + labs(subtitle = "1894 DEGs"), 
+          j + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) + labs(subtitle = "256 DEGs"), 
+          nrow = 1, rel_widths = c(1.2,1,1),
+          labels = c("h","i","j"," "), label_size = 9)
+
+plot_grid(manipinc, maniphatch, nrow = 2)
 ```

### 2019-12-06 analysis/07_hormones.Rmd
prep for correlations with hormones and gene epxression
new illustrations, new prolactin stuff

@@
 knitr::opts_chunk$set(fig.path = '../figures/hormones/',message=F, warning=FALSE)
+```
+
+```{r}
+colData  <- read_csv("../metadata/00_birds.csv") %>%
+  mutate(RNAseq = "RNAseq",
+         bird_id = bird)  %>%
+   select(-X1, -bird)
+colData
 ```
@@
 
+prolctin1rnaseq <- full_join(prolactin, colData)
+prolctin2rnaseq <- full_join(prolactin2, colData)
+
+prolctin1rnaseq
+prolctin2rnaseq
 ```
@@
 
-```{r}
-write.csv(hormones, "../results/hormones.csv", row.names = F)
-```
-
 # do control bird with high prolactin hormone have high PRL expression in the pituitary? yes.
@@
 ```{r}
-prolactin <- read_excel("../results/Pigeon prolactin concentrations juil 2018.xlsx", sheet = 1)  %>% 
-            filter(Treatment == "baseline")  %>% 
-            select(ColorBands,BandNo_old, Treatment, Sex, `Prolactin ng/mL`)  %>% 
-            arrange(desc(`Prolactin ng/mL`))
+write.csv(hormones, "../results/hormones.csv", row.names = F)
+write.csv(prolactin2, "../results/07_prolactin2.csv", row.names = F)
 
-head(prolactin,2)
 ```

### 2019-12-06 analysis/16_PRL.Rmd
new illustrations, new prolactin stuff

@@
 p4
+```
 
+```{r geneshormones}
+candidates.pit$bird_id <- candidates.pit$bird
+
+geneshormones <- left_join(candidates.pit, hormones) %>%
+  select(-X1, -V1, -okay, - icons, -music, -iconpath) %>%
+  drop_na()
+head(geneshormones)
+
+ggplot(geneshormones, aes(x = plasma_conc, y = PRL)) +
+  geom_point(aes(color = treatment)) +
+  facet_wrap(~hormone, scales = "free_x") +
+    geom_smooth(method = "lm", color = "grey")
+
+prolactin2 <- read_csv("../results/07_prolactin2.csv")
+head(prolactin2)
+geneshormones2 <- left_join(candidates.pit, prolactin2) %>%
+  drop_na()
+head(geneshormones)
+
+ggplot(geneshormones2, aes(x = plasma_conc, y = PRL)) +
+  geom_point(aes( color = treatment)) +
+  geom_smooth(method = "lm", color = "grey") +
+  facet_wrap(~hormone, scales = "free") 
 ```
+
+```{r}
+write.csv(candidates.pit, "../results/16_pitPRL.csv")
+```

### 2019-12-07 analysis/16_PRL.Rmd
new prl images

--- 2019-12-02 analysis/16_PRL.Rmd
@@
 hormones$treatment <- factor(hormones$treatment, levels = alllevels)
+
+meanprolactin <- hormones %>% 
+    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
+    droplevels() %>% 
+  dplyr::group_by(treatment) %>%
+  dplyr::summarise(m = mean(plasma_conc), 
+                   se = sd(plasma_conc)/sqrt(length(plasma_conc))) %>%
+  dplyr::mutate(m = round(m,0)) 
+meanprolactin <- left_join(meanprolactin, birds)
+meanprolactin$treatment <- factor(meanprolactin$treatment, levels = alllevels)
+meanprolactin
+
 ```
@@
   ylim(c(-10,120)) + 
-    mytheme() +
+    theme_B3() +
   #theme(axis.text.x = element_blank()) + 
   theme(legend.position = "none") + 
-  labs(x = "parental care stages", y = "prolactin conc. (ng/mL)") +
+  labs(x = "parental care stages", y = "prolactin (ng/mL)") +
   annotation_custom(control, ymin = -20, ymax = 0, xmin = -7.5) +
@@
 
-meanPRL <- hormones %>% 
-    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
-    droplevels() %>% 
-  dplyr::group_by(treatment) %>%
-  dplyr::summarise(m = mean(plasma_conc), 
-                   se = sd(plasma_conc)/sqrt(length(plasma_conc))) %>%
-  dplyr::mutate(m = round(m,0)) 
-meanPRL <- left_join(meanPRL, birds)
-meanPRL$treatment <- factor(meanPRL$treatment, levels = alllevels)
-meanPRL
+p2 <-  ggplot(meanprolactin, aes(treatment, m)) +
+  geom_errorbar(aes(ymin=m-se, ymax=m+se, color = treatment), width=.1) +
+   geom_image(aes(image=iconpath), size=.1)  +
+  labs(x = "parental care stages", y = "prolactin (ng/mL)") +
+  theme_B3() +
+  theme(legend.position = "none")
 
-ggplot(meanPRL, aes(treatment, m)) +
-  geom_errorbar(aes(ymin=m-se, ymax=m+se), width=.1) +
-   geom_image(aes(image=iconpath), size=.12)  +
-  labs(x = "parental care stages", y = "Prolactin (ng/mL)")
+p2
 
-ggplot(meanPRL, aes(treatment, m)) +
+p3 <- ggplot(meanprolactin, aes(treatment, m)) +
    geom_image(aes(image=music), size=.12)  +
-  labs(x = "parental care stages", y = "Prolactin (ng/mL)")  +
+  labs(x = "increasing time >>", y = "prolactin (ng/mL)")  +
   mytheme() +
@@
   annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 6.5) +
-  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 8.5)
+  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 8.5) +
+  geom_hline(yintercept=0) +
+  geom_hline(yintercept=20) +
+  geom_hline(yintercept=40) +
+  geom_hline(yintercept=60) +
+  geom_hline(yintercept=80) 
+p3
 
-p2 <- hormones %>% 
+p4 <- hormones %>% 
     filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
     droplevels() %>% 
-  ggplot(aes(x = as.numeric(treatment), y = plasma_conc)) +
-        geom_smooth(aes(colour = sex)) +
-    geom_boxplot(aes(fill = treatment, alpha = sex)) +
-    mytheme() +
-    theme(axis.text.x = element_blank()) +
+  ggplot(aes(x = treatment, y = plasma_conc)) +
+    geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) +
+    theme_B3() +
     scale_fill_manual(values = colorscharmaip) +
     scale_color_manual(values = sexcolors) +
-    labs(y = "prolactin conc. (ng/mL)", x = NULL) +
+    labs(y = "prolactin (ng/mL)", x = NULL) +
     guides(fill = FALSE, alpha = FALSE,
@@
     theme(legend.position = c(0.85,0.15), legend.direction = "horizontal") + 
-  labs(x = "parental care stages", y = "prolactin conc. (ng/mL)") +
+  labs(x = "increasing time >>", y = "prolactin (ng/mL)") +
   annotation_custom(control, ymin = -20, ymax = 0, xmin = -7.8) +
@@
   ylim(c(-10,120)) 
-p2
+p4
+
+plot_grid(p1,p4,p2,p3)
+
 ```
@@
 
-p3 <- ggplot(candidates.pit, aes(x = as.numeric(treatment), y = PRL)) + 
-        geom_smooth(aes(colour = sex)) +
-    geom_boxplot(aes(fill = treatment, alpha = sex)) + 
+meanPRL <- candidates.pit %>% 
+    droplevels() %>% 
+  dplyr::group_by(treatment) %>%
+  dplyr::summarise(m = mean(PRL), 
+                   se = sd(PRL)/sqrt(length(PRL))) %>%
+  dplyr::mutate(m = round(m,1)) 
+meanPRL <- left_join(meanPRL, birds)
+meanPRL$treatment <- factor(meanPRL$treatment, levels = alllevels)
+meanPRL
+
+p5 <- candidates.pit %>% 
+  ggplot(aes(x = treatment, y = PRL, fill = treatment)) +
+    geom_boxplot() +
+    theme_B3() +
+  theme(legend.position = "none") + 
+  labs(x = "increasing time >>", y = "prolactin (ng/mL)") +
+  annotation_custom(control, ymin = 12, ymax = 14, xmin = -7.8) +
+  annotation_custom(bldg, ymin = 12, ymax = 14, xmin = -5.8) +
+  annotation_custom(lay, ymin = 12, ymax = 14, xmin = -3.8) +
+  annotation_custom(inc, ymin = 12, ymax = 14, xmin = -1.8) +
+  annotation_custom(inc, ymin = 12, ymax = 14, xmin = 0.6) +
+  annotation_custom(inc, ymin = 12, ymax = 14, xmin = 2.4) +
+  annotation_custom(hatch, ymin = 12, ymax = 14, xmin = 4.4) +
+  annotation_custom(nestling, ymin = 12, ymax = 14, xmin = 6.4) +
+  annotation_custom(nestling, ymin = 12, ymax = 14, xmin = 8.4)  +
+  ylim(c(13,24)) 
+p5
+
+p6 <- ggplot(candidates.pit, aes(x = treatment, y = PRL)) + 
+    geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) + 
     scale_alpha_manual(values = c(0.75,1)) +
-    mytheme() +
+    theme_B3() +
   theme(legend.position = c(0.85,0.2), legend.direction = "horizontal") + 
@@
     labs(y = "PRL expression in the pituitary", x = "parental care stage") +
-    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
-                       labels = charlevels) +
   annotation_custom(control, ymin = 12, ymax = 14, xmin = -7.8) +
@@
   guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) 
-p3
+p6
 
-candidates.pit
+p7 <- ggplot(meanPRL, aes(treatment, m)) +
+  geom_errorbar(aes(ymin=m-se, ymax=m+se, color = treatment), width=.1) +
+   geom_image(aes(image=iconpath), size=.12)  +
+  labs(x = "increasing time >>", y = "Pituitary PRL expression") +
+  theme_B3()  +
+  theme(legend.position = "none")
+p7
 
-meanPRL <- candidates.pit %>% 
-    droplevels() %>% 
-  dplyr::group_by(treatment) %>%
-  dplyr::summarise(m = mean(PRL), 
-                   se = sd(PRL)/sqrt(length(PRL))) %>%
-  dplyr::mutate(m = round(m,0)) 
-meanPRL <- left_join(meanPRL, birds)
-meanPRL$treatment <- factor(meanPRL$treatment, levels = alllevels)
-meanPRL
-
-ggplot(meanPRL, aes(treatment, m)) +
-  geom_errorbar(aes(ymin=m-se, ymax=m+se), width=.1) +
-   geom_image(aes(image=iconpath), size=.12)  +
-  labs(x = "parental care stages", y = "Pituitary PRL expression")
-
-ggplot(meanPRL, aes(treatment, m)) +
+p8 <- ggplot(meanPRL, aes(treatment, m)) +
    geom_image(aes(image=music), size=.12)  +
@@
   annotation_custom(nestling, ymin = 15, ymax = 16.5, xmin = 6.5) +
-  annotation_custom(nestling, ymin = 15, ymax = 16.5, xmin = 8.5) 
+  annotation_custom(nestling, ymin = 15, ymax = 16.5, xmin = 8.5)   +
+  geom_hline(yintercept=17) +
+  geom_hline(yintercept=18) +
+  geom_hline(yintercept=19) +
+  geom_hline(yintercept=20) +
+  geom_hline(yintercept=21)
+
+p8
+
+plot_grid(p5,p6,p7,p8)
 ```
+
+## prolactin manip
+
+```{r manipdata}
+vsd.pit <- readvsd("../results/06_pitallvsd.csv")
+vsd.pit$entrezid <- row.names(vsd.pit)
+
+colData.pit <- read_csv("../metadata/00_samples.csv") %>% 
+  filter(tissue == "pituitary") %>% 
+  mutate(sample = V1)
+
+geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>%  
+  dplyr::select(Name, geneid, entrezid) %>% arrange(Name)
+
+candidategenes <- c("PRL")
+candidates.pit <- selectcandidatevsds(candidategenes, vsd.pit, colData.pit)
+candidates.pit$treatment <- factor(candidates.pit$treatment, levels = alllevels)
+
+meanPRL <- candidates.pit %>% 
+    droplevels() %>% 
+  dplyr::group_by(treatment) %>%
+  dplyr::summarise(m = median(PRL), 
+                   se = sd(PRL)/sqrt(length(PRL))) %>%
+  dplyr::mutate(m = round(m,1)) 
+meanPRL <- left_join(meanPRL, birds)
+meanPRL$treatment <- factor(meanPRL$treatment, levels = alllevels)
+meanPRL
+```
+
+```{r PRL.pit.maip}
+p10 <- ggplot(meanPRL, aes(treatment, m)) +
+   geom_errorbar(aes(ymin=m-se, ymax=m+se, color = treatment), width=.1) +
+   geom_image(aes(image = iconpath), size = 0.06)  +
+   geom_image(aes(image = music , y = m + 0.5), size = 0.04) +
+   labs(x = "parental care stages", y = "Pituitary PRL") +
+   mytheme() +
+   ylim(c(16,22)) +
+   theme(axis.text.x = element_text(angle = 45)) +
+  theme(legend.position = "none") +
+  scale_color_manual(values = colorscharmaip2) +
+  geom_hline(yintercept=16.8) +
+  geom_hline(yintercept=17.8) +
+  geom_hline(yintercept=18.8) +
+  geom_hline(yintercept=19.8) +
+  geom_hline(yintercept=20.8)  
+p10
+
+p11 <- ggplot(candidates.pit, aes(x = treatment, y = PRL)) + 
+       geom_smooth(aes(colour = sex)) +
+    geom_boxplot(aes(fill = treatment)) +
+    scale_alpha_manual(values = c(0.75,1)) +
+    theme_B3() +
+   theme(legend.position = "none", legend.direction = "horizontal") + 
+   scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+    labs(y = "Pituitary PRL", x = NULL) +
+  scale_fill_manual(values = colorscharmaip2) +
+  ylim(c(13,22.5)) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  annotation_custom(control, ymin = 13, ymax = 14, xmin = -14.8) +
+  annotation_custom(bldg, ymin = 13, ymax = 14, xmin = -12.8) +
+  annotation_custom(lay, ymin = 13, ymax = 14, xmin = -10.8) +
+  annotation_custom(inc, ymin = 13, ymax = 14, xmin = -8.8) +
+  annotation_custom(removeegg, ymin = 13, ymax = 14, xmin = -6.8) +
+  annotation_custom(inc, ymin = 13, ymax = 14, xmin = -4.8) +
+  annotation_custom(maniphatch, ymin = 13, ymax = 14, xmin = -2.8) +
+  annotation_custom(removeegg, ymin = 13, ymax = 14, xmin = -0.8) +
+  annotation_custom(inc, ymin = 13, ymax = 14, xmin = 1.4) +
+  annotation_custom(removeegg, ymin = 13, ymax = 14, xmin = 3.4)  +
+  annotation_custom(manipinc, ymin = 13, ymax = 14, xmin = 5.4) +
+  annotation_custom(hatch, ymin = 13, ymax = 14, xmin = 7.4) +
+  annotation_custom(removechick, ymin = 13, ymax = 14, xmin = 9.4) +
+  annotation_custom(maniphatch, ymin = 13, ymax = 14, xmin = 11.4) +
+  annotation_custom(nestling, ymin = 13, ymax = 14, xmin = 13.4) +
+  annotation_custom(nestling, ymin = 13, ymax = 14, xmin = 15.4) 
+p11
+
+p12 <- candidates.pit %>%
+  filter(treatment %in% c("inc.d9", "m.inc.d8", "hatch")) %>%
+    ggplot(aes(x = treatment, y = PRL)) + 
+    geom_boxplot(aes(fill = treatment, color = sex)) + 
+    scale_alpha_manual(values = c(0.75,1)) +
+    theme_B3() +
+   theme(legend.position = "none", legend.direction = "horizontal") + 
+   scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+    labs(y = "Pituitary PRL", x = NULL) +
+  ylim(c(13,22.5)) +
+  scale_fill_manual(values = colorscharmaip2)   +
+  annotation_custom(control, ymin = 13, ymax = 14, xmin = -1.6) +
+  annotation_custom(maniphatch, ymin = 13, ymax = 14, xmin = 0.4) +
+  annotation_custom(hatch, ymin = 13, ymax = 14, xmin = 2.4)
+p12
+
+p13 <- candidates.pit %>%
+  filter(treatment %in% c("inc.d17", "prolong", "hatch")) %>%
+    ggplot(aes(x = treatment, y = PRL)) + 
+    geom_boxplot(aes(fill = treatment, color = sex)) + 
+    scale_alpha_manual(values = c(0.75,1)) +
+    theme_B3() +
+   theme(legend.position = "none", legend.direction = "horizontal") + 
+   scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+    labs(y = "Pituitary PRL", x = NULL) +
+  ylim(c(13,22.5)) +
+  scale_fill_manual(values = colorscharmaip2)   +
+  annotation_custom(control, ymin = 13, ymax = 14, xmin = -1.6) +
+  annotation_custom(manipinc, ymin = 13, ymax = 14, xmin = 0.4) +
+  annotation_custom(hatch, ymin = 13, ymax = 14, xmin = 2.4)
+p13
+
+p14 <- candidates.pit %>%
+  filter(treatment %in% c("hatch", "extend", "n5")) %>%
+    ggplot(aes(x = treatment, y = PRL)) + 
+    geom_boxplot(aes(fill = treatment, color = sex)) + 
+    scale_alpha_manual(values = c(0.75,1)) +
+    theme_B3() +
+   theme(legend.position = "none", legend.direction = "horizontal") + 
+   scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+    labs(y = "Pituitary PRL", x = NULL) +
+  ylim(c(13,22.5)) +
+  scale_fill_manual(values = colorscharmaip2)   +
+  annotation_custom(control, ymin = 13, ymax = 14, xmin = -1.6) +
+  annotation_custom(maniphatch, ymin = 13, ymax = 14, xmin = 0.4) +
+  annotation_custom(hatch, ymin = 13, ymax = 14, xmin = 2.4)
+p14
+
+plot_grid(p12,p13 + theme(axis.text.y = element_blank(), axis.title.y = element_blank()),
+          p14 + theme(axis.text.y = element_blank(), axis.title.y = element_blank()), 
+          nrow = 1, rel_widths = c(1.2,1,1))
+
+```
+
+```{r geneshormones}
+candidates.pit$bird_id <- candidates.pit$bird
+
+geneshormones <- left_join(candidates.pit, hormones) %>%
+  select(-X1, -V1, -okay, - icons, -music, -iconpath) %>%
+  drop_na()
+head(geneshormones)
+
+ggplot(geneshormones, aes(x = plasma_conc, y = PRL)) +
+  geom_point(aes(color = treatment)) +
+  facet_wrap(~hormone, scales = "free_x") +
+    geom_smooth(method = "lm", color = "grey")
+
+prolactin2 <- read_csv("../results/07_prolactin2.csv")
+head(prolactin2)
+geneshormones2 <- left_join(candidates.pit, prolactin2) %>%
+  drop_na()
+head(geneshormones)
+
+ggplot(geneshormones2, aes(x = plasma_conc, y = PRL)) +
+  geom_point(aes( color = treatment)) +
+  geom_smooth(method = "lm", color = "grey") +
+  facet_wrap(~hormone, scales = "free") 
+```
+
+```{r}
+write.csv(candidates.pit, "../results/16_pitPRL.csv")
+```

### 2019-12-07 analysis/16_PRL.Rmd
now with aov tables

@@
 library(grid)
-library("ggimage")
+library(ggimage)
+library(apaTables)
 
@@
 ```{r}
+aov_all = data.frame()
+for(i in alllevels3){
+
+  df <- candidates.pit %>% filter(treatment == i) %>% droplevels()
+  aovtable <- apa.aov.table(aov(PRL ~ sex , data=df))
+  aovtable <- as.data.frame(aovtable$table_body)
+  totaldf <- aovtable[3, 3]
+  aovtable$df <- paste(aovtable$df, ", " , totaldf, sep = "")
+  aovtable$ANOVA <- "PRL ~ sex"
+  aovtable$stages <- paste(i)
+  aovtable$p <- as.numeric(as.character(aovtable$p))
+  aov_all <- rbind(aov_all,aovtable)
+}
+
+aov_all <- aov_all %>%
+  filter(Predictor == "sex")  %>%
+  select(stages, ANOVA, df, "F", p) %>%
+  mutate(sig = ifelse(p < 0.05, "*", " "))
+aov_all
+
+one <- c("m.inc.d3", "inc.d3")
+two <- c("m.inc.d9", "inc.d9")
+three <- c("m.inc.d17", "inc.d17")
+four <- c("m.n2", "hatch")
+five <- c("m.inc.d8", "inc.d9")
+six <- c("m.inc.d8", "hatch")
+seven <- c("prolong", "inc.d17")
+eight <- c("prolong", "hatch")
+nine <- c("extend", "hatch")
+ten <- c("extend", "n5")
+
+manipcomparisons <- list(one, two, three, four, five, six, seven, eight, nine, ten)
+
+aov_manip = data.frame()
+for(i in manipcomparisons){
+
+  df <- candidates.pit %>% filter(treatment %in% i) %>% droplevels()
+  aovtable <- apa.aov.table(aov(PRL ~ sex * treatment , data=df))
+  aovtable <- as.data.frame(aovtable$table_body)
+  totaldf <- aovtable[5, 3]
+  aovtable$df <- paste(aovtable$df, ", " , totaldf, sep = "")
+  aovtable$ANOVA <- "PRL ~ sex"
+  aovtable$stages <- paste(i, collapse = " vs ")
+  aovtable$p <- as.numeric(as.character(aovtable$p))
+  aov_manip <- rbind(aov_manip,aovtable)
+}
+
+aov_manip  <- aov_manip %>%
+  filter(Predictor %in% c( "treatment"))  %>%
+  select(stages, ANOVA, Predictor, df, "F", p) %>%
+  mutate(sig = ifelse(p < 0.05, "*", " "))
+aov_manip
+```
+
+```{r}
 write.csv(candidates.pit, "../results/16_pitPRL.csv")
+write.csv(aov_all, "../results/16_aov_PRLsex.csv")
+write.csv(aov_manip, "../results/16_aov_PRLsextreatment.csv")
+
 ```

### 2019-12-12 analysis/07_hormones.Rmd
new GO analysis

@@
 PETCP$treatment.x <- factor(PETCP$treatment.x, levels = alllevels2)
-PETCP$hormone.y
 
@@
     geom_boxplot() + 
+  theme_B3()  +
     theme(axis.text.x = element_text(angle = 45, hjust = 1),
@@
    annotation_custom(inc, xmax = 1.6, ymin = -14, ymax = 5) +
-   annotation_custom(bldg, xmax = 3.6, ymin = -14, ymax = 5) +
+   annotation_custom(removeegg, xmax = 3.6, ymin = -14, ymax = 5) +
    annotation_custom(inc, xmax = 5.6, ymin = -14, ymax = 5) +
-   annotation_custom(bldg, xmax = 7.6, ymin = -14, ymax = 5) +
+   annotation_custom(removeegg, xmax = 7.6, ymin = -14, ymax = 5) +
   annotation_custom(hatch, xmax = 9.6, ymin = -14, ymax = 5) +
-  annotation_custom(bldg, xmax = 11.6, ymin = -14, ymax = 5) +
+  annotation_custom(removeegg, xmax = 11.6, ymin = -14, ymax = 5) +
   annotation_custom(nestling, xmax = 13.6, ymin = -14, ymax = 5) +
-  annotation_custom(bldg, xmax = 15.6, ymin = -14, ymax = 5) +
+  annotation_custom(removechick, xmax = 15.6, ymin = -14, ymax = 5) +
 
@@
    annotation_custom(inc, xmax = 1.6, ymin = -10, ymax = 10) +
-   annotation_custom(hatch, xmax = 3.6, ymin = -10, ymax = 10) +
+   annotation_custom(maniphatch, xmax = 3.6, ymin = -10, ymax = 10) +
    annotation_custom(inc, xmax = 5.6, ymin = -10, ymax = 10) +
-   annotation_custom(inc, xmax = 7.6, ymin = -10, ymax = 10) +
+   annotation_custom(manipinc, xmax = 7.6, ymin = -10, ymax = 10) +
   annotation_custom(hatch, xmax = 9.6, ymin = -10, ymax = 10) +
-  annotation_custom(hatch, xmax = 11.6, ymin = -10, ymax = 10) +
+  annotation_custom(maniphatch, xmax = 11.6, ymin = -10, ymax = 10) +
   annotation_custom(nestling, xmax = 13.6, ymin = -10, ymax = 10) +

### 2019-12-12 analysis/12_GO.Rmd
new GO analysis

---
title: "GO"
output: md_document
---

```{r}
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r, GOterms}

GO_path <- "../results/goterms/"   # path to the data
GO_files <- dir(GO_path, pattern = "*.txt") # get file names
GO_pathfiles <- paste0(GO_path, GO_files)

allGO <- GO_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_table(.x, col_types = cols(), col_names = FALSE), .id = "file_name") %>% 
  mutate(GO = sapply(strsplit(as.character(file_name),'../results/goterms/'), "[", 2)) %>% 
  mutate(GO = sapply(strsplit(as.character(GO),'.txt'), "[", 1)) %>% 
  mutate(gene = sapply(strsplit(as.character(X1), "[\\\\]|[^[:print:]]" ), "[", 2)) %>% 
  select(GO, gene)  %>%
  filter(gene != "Symbol") %>%
  distinct(GO,gene)  %>%
  mutate(gene = toupper(gene))
head(allGO)
```

```{r DEGs}

DEG_path <- "../results/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "03_DEGs.*") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)

DEGs <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>%
  mutate(comparison = sapply(strsplit(as.character(file_name),'../results/03_DEGs.'), "[", 2)) %>% 
  mutate(comparison = sapply(strsplit(as.character(comparison),'.csv'), "[", 1)) %>% 
  select(gene, comparison)
head(DEGs)

GO_DEGs <- left_join(allGO, DEGs) %>%
  arrange(GO, gene) %>%
  group_by(GO, comparison) %>%
  summarize(genes = str_c(gene, collapse = ", ")) %>%
  pivot_wider(names_from = comparison, values_from = genes) %>% 
  select(GO, contains("control.bldg"), contains("bldg.lay"), contains("lay.inc.d3"),
         contains("inc.d3.inc.d9"), contains("inc.d9.inc.d17"), contains("inc.d17.hatch"),
         contains("hatch.n5"), contains("n5.n9")) %>%
  filter(GO %in% c("endocrineprocess", "lactation", "parentalbehavior"))

hyp <- GO_DEGs %>% select(GO, starts_with("hyp"))
pit <- GO_DEGs %>% select(GO, starts_with("pit"))
gon <- GO_DEGs %>% select(GO, starts_with("gon"))

kable(hyp)
kable(pit)
kable(gon)

one <- GO_DEGs %>% select(GO, contains("control.bldg"))
two <- GO_DEGs %>% select(GO, contains("bldg.lay"))
three <- GO_DEGs %>% select(GO, contains("lay.inc.d3"))
four <- GO_DEGs %>% select(GO, contains("inc.d3.inc.d9"))
five <- GO_DEGs %>% select(GO, contains("inc.d9.inc.d17"))
six <- GO_DEGs %>% select(GO, contains("inc.d17.hatch"))
seven <- GO_DEGs %>% select(GO, contains("hatch.n5"))
eight <- GO_DEGs %>% select(GO, contains("n5.n9"))

kable(one)
kable(two)
kable(three)
kable(four)
kable(five)
kable(six)
kable(seven)
kable(eight)

write.csv(GO_DEGs, "../results/12_GO_DEGs.csv")
write.csv(hyp, "../results/12_GO_DEGs_hyp.csv")
write.csv(pit, "../results/12_GO_DEGs_pit.csv")
write.csv(gon, "../results/12_GO_DEGs_gon.csv")

```

### 2019-12-12 analysis/12_GO.Rmd
now with a few candidate genes

@@
 ```
+
+```{r}
+candidates <- c("AVP", "PRL", "PRLR", "GNIH",  "OXT", "AVPR1A", "AVPR1B",
+                "CRH", "FSHB", "GH", "GNRH1", "GNRH2",
+                "MC4R", "JAK2", "HCRT", "POMC", "TRH", "TSHB")
+candidates <- as.data.frame(candidates)
+names(candidates)[1] <- "gene"
+candidates$GO <- "candidate genes"
+
+candidate_DEGs <- left_join(candidates, DEGs) %>%
+  arrange(GO, gene) %>%
+  group_by(GO, comparison) %>%
+  summarize(genes = str_c(gene, collapse = ", ")) %>%
+  pivot_wider(names_from = comparison, values_from = genes) %>% 
+  select(GO, contains("control.bldg"), contains("bldg.lay"), contains("lay.inc.d3"),
+         contains("inc.d3.inc.d9"), contains("inc.d9.inc.d17"), contains("inc.d17.hatch"),
+         contains("hatch.n5"), contains("n5.n9"))
+candidate_DEGs <- as.data.frame(candidate_DEGs)
+row.names(candidate_DEGs) <- candidate_DEGs$GO  
+candidate_DEGs <- t(candidate_DEGs)
+candidate_DEGs <- as.data.frame(candidate_DEGs)
+candidate_DEGs <- tail(candidate_DEGs,10)
+kable(candidate_DEGs)
+
+```

### 2019-12-12 analysis/12_GO.Rmd
better tables

@@
 library(kableExtra)
+
+knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
+
 ```
@@
 head(DEGs)
+```
 
+```{r GO_DEGs}
 GO_DEGs <- left_join(allGO, DEGs) %>%
@@
          contains("hatch.n5"), contains("n5.n9")) %>%
-  filter(GO %in% c("endocrineprocess", "lactation", "parentalbehavior"))
+  filter(GO %in% c("endocrineprocess", "lactation", 
+                   "parentalbehavior", "matingbehavior")) %>% 
+  map(~.x) %>%
+  discard(~all(is.na(.x))) %>%
+  map_df(~.x)
+```
 
+```{r bytissue}
 hyp <- GO_DEGs %>% select(GO, starts_with("hyp"))
@@
 kable(gon)  
+```
 
+```{r bystage}
 one <- GO_DEGs %>% select(GO, contains("control.bldg"))
@@
 kable(one) 
+# top GOs of these genes: endocrine porcess, regulation of blood pressure (by hormones or chemical signals), system process, regulation of hormone levels
+
 kable(two) 
+# top Gos maternal, parental, reproductive behavior; reproduction, reproductive process
+
 kable(three) 
+# response to chemical, maternal and parental beahviors, regulation of RNApolI transscription
+
 kable(four) 
+# insuffiencient genes
+
 kable(five) 
+# (cellular) response to organic substance, endocrine process, response to stimulus, response to chemical
+
 kable(six)  
+# insuffiencient genes
+
 kable(seven)  
+# endocrine process, system process, hormone transport, synaptic transmission (dopamineric), secretion by cell
+
 kable(eight) 
-
-write.csv(GO_DEGs, "../results/12_GO_DEGs.csv")
-write.csv(hyp, "../results/12_GO_DEGs_hyp.csv")
-write.csv(pit, "../results/12_GO_DEGs_pit.csv")
-write.csv(gon, "../results/12_GO_DEGs_gon.csv")
-
+# (positive) regulation of gap junction assembly, regulation of cardiac conduction, regualation of blood pressure
 ```
@@
                 "CRH", "FSHB", "GH", "GNRH1", "GNRH2",
-                "MC4R", "JAK2", "HCRT", "POMC", "TRH", "TSHB")
+                "MC4R", "JAK2", "HCRT", "POMC", "TRH", "TSHB",
+                "LH", "FSH")
 candidates <- as.data.frame(candidates)
@@
 
+# Top Go terms: reponse to keytone, response to hormone, repsponse to organonitrogen compound, positive regulation of homeostatic process, maternal beahvior
+
 ```
+
+```{r write}
+write.csv(GO_DEGs, "../results/12_GO_DEGs.csv")
+write.csv(hyp, "../results/12_GO_DEGs_hyp.csv")
+write.csv(pit, "../results/12_GO_DEGs_pit.csv")
+write.csv(gon, "../results/12_GO_DEGs_gon.csv")
+```

### 2019-12-12 analysis/12_GO.Rmd
now with markdown headers

@@
 ```
+
+# Exploring relationships between genes associated with mating behavior, lactation, parental care, and endocrine process in the hypothalamic-pituitary-gonad axis of reproction
+
+## The code
+
+1. Read in list of genes for GO terms: mating beahvior, lactating, parental care, and endocrine process.
+2. Read in lists of differentially expressed genes for each transition (e.g. incubation day 17 to hatch) in each tissue (e.g. pituitary)
+3. Join the two data frames
+4. Print list of differentially expressed genes that correspond to each GO term, separated by either tissue or paretnal stage
+5. Repeat the process for a list of candidate genes instead of GO terms.
+
+## GO terms
 
@@
 
+## Differntially expressed genes by stage for each tissue
+
 ```{r DEGs}
@@
 ```
+
+## GO and DEGs joined
 
@@
 
+## Tissue-specifiic differentially expressed genes
+
 ```{r bytissue}
@@
 ```
+
+## Stage-specific differentially expressed genes
 
@@
 
+## Candidate genes
+
 ```{r}

### 2019-12-12 analysis/12_GO.Rmd
better markdown
more readable comment

@@
 
-```{r}
+# Exploring relationships between genes associated with mating behavior, lactation, parental care, and endocrine process in the hypothalamic-pituitary-gonad axis of reproction
+
+1. Read in list of genes for GO terms: mating beahvior, lactating, parental care, and endocrine process.
+2. Read in lists of differentially expressed genes for each transition (e.g. incubation day 17 to hatch) in each tissue (e.g. pituitary)
+3. Join the two data frames
+4. Print list of differentially expressed genes that correspond to each GO term, separated by either tissue or paretnal stage
+5. Repeat the process for a list of candidate genes instead of GO terms.
+
+## Made with these R packages
+
+```{r setup}
 library(tidyverse)
@@
 ```
-
-# Exploring relationships between genes associated with mating behavior, lactation, parental care, and endocrine process in the hypothalamic-pituitary-gonad axis of reproction
-
-## The code
-
-1. Read in list of genes for GO terms: mating beahvior, lactating, parental care, and endocrine process.
-2. Read in lists of differentially expressed genes for each transition (e.g. incubation day 17 to hatch) in each tissue (e.g. pituitary)
-3. Join the two data frames
-4. Print list of differentially expressed genes that correspond to each GO term, separated by either tissue or paretnal stage
-5. Repeat the process for a list of candidate genes instead of GO terms.
 
@@
 
+# top GOs of these genes: endocrine porcess, 
+#regulation of blood pressure (by hormones or chemical signals), 
+#system process, regulation of hormone levels
 kable(one)
-# top GOs of these genes: endocrine porcess, regulation of blood pressure (by hormones or chemical signals), system process, regulation of hormone levels
 
+# top Gos maternal, parental, reproductive behavior; 
+# reproduction, reproductive process
 kable(two)
-# top Gos maternal, parental, reproductive behavior; reproduction, reproductive process
 
+# response to chemical, maternal and parental behaviors, 
+# regulation of RNApolI transcription
 kable(three)
-# response to chemical, maternal and parental beahviors, regulation of RNApolI transscription
 
+# insuffiencient genes for GO analysis
 kable(four)
+
+# (cellular) response to organic substance, endocrine process, 
+# response to stimulus, response to chemical 
+kable(five)
+
 # insuffiencient genes
+kable(six)
 
-kable(five) 
-# (cellular) response to organic substance, endocrine process, response to stimulus, response to chemical
+# endocrine process, system process, hormone transport, 
+# synaptic transmission (dopamineric), secretion by cell
+kable(seven)
 
-kable(six)  
-# insuffiencient genes
-
-kable(seven)  
-# endocrine process, system process, hormone transport, synaptic transmission (dopamineric), secretion by cell
-
+# (positive) regulation of gap junction assembly, 
+# regulation of cardiac conduction, regualation of blood pressure
 kable(eight) 
-# (positive) regulation of gap junction assembly, regulation of cardiac conduction, regualation of blood pressure
 ```

### 2019-12-12 analysis/12_GO.Rmd
table observations below

@@
 
+kable(one) 
 # top GOs of these genes: endocrine porcess, 
@@
 #system process, regulation of hormone levels
-kable(one)
 
+kable(two) 
 # top Gos maternal, parental, reproductive behavior; 
 # reproduction, reproductive process
-kable(two)
 
+kable(three) 
 # response to chemical, maternal and parental behaviors, 
 # regulation of RNApolI transcription
-kable(three)
 
+kable(four) 
 # insuffiencient genes for GO analysis
-kable(four)
 
+kable(five) 
 # (cellular) response to organic substance, endocrine process, 
 # response to stimulus, response to chemical
-kable(five)
 
+kable(six)  
 # insuffiencient genes
-kable(six)
 
+kable(seven) 
 # endocrine process, system process, hormone transport, 
 # synaptic transmission (dopamineric), secretion by cell
-kable(seven)
 
+kable(eight) 
 # (positive) regulation of gap junction assembly, 
 # regulation of cardiac conduction, regualation of blood pressure
-kable(eight) 
+
 ```

### 2019-12-12 analysis/12_GO.Rmd
markdown format

@@
 
-kable(hyp)  
-kable(pit)  
+kable(hyp,  format = "markdown")  
+# top GO therms for hyp:
+# endocrine process, regulation of hormone levels,
+# signal release # hormone secretion, # system process
+# hormone transport, cell-cell signaling
+
+kable(pit,  format = "markdown")
+# endocrine process, mating behavior, system process,
+# reproductive behavior, mating, regulation of blood pressure
+# signal release, regulation of hormone levels
+
 kable(gon)  
+# endocrine process, regulation of blood pressure
+# blood circutation, circulatory system
 ```

### 2019-12-14 analysis/12_GO.Rmd
new plr and brca stuff

--- 2019-12-12 analysis/12_GO.Rmd
@@
 title: "GO"
-output: md_document
+output: html_document
 ---
@@
 kable(hyp)  
+# top GO therms for hyp:
+# endocrine process, regulation of hormone levels,
+# signal release # hormone secretion, # system process
+# hormone transport, cell-cell signaling
+
 kable(pit)
+# endocrine process, mating behavior, system process,
+# reproductive behavior, mating, regulation of blood pressure
+# signal release, regulation of hormone levels
+
 kable(gon)  
+# endocrine process, regulation of blood pressure
+# blood circutation, circulatory system
 ```

### 2019-12-14 analysis/16_PRL.Rmd
new plr and brca stuff

@@
 geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>%  dplyr::select(Name, geneid, entrezid) %>% arrange(Name)
+
 candidategenes <- c("PRL")
@@
 
-p6 <- ggplot(candidates.pit, aes(x = treatment, y = PRL)) + 
-    geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) + 
-    scale_alpha_manual(values = c(0.75,1)) +
-    theme_B3() +
-  theme(legend.position = c(0.85,0.2), legend.direction = "horizontal") + 
-  scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
-    labs(y = "PRL expression in the pituitary", x = "parental care stage") +
-  annotation_custom(control, ymin = 12, ymax = 14, xmin = -7.8) +
-  annotation_custom(bldg, ymin = 12, ymax = 14, xmin = -5.8) +
-  annotation_custom(lay, ymin = 12, ymax = 14, xmin = -3.8) +
-  annotation_custom(inc, ymin = 12, ymax = 14, xmin = -1.8) +
-  annotation_custom(inc, ymin = 12, ymax = 14, xmin = 0.6) +
-  annotation_custom(inc, ymin = 12, ymax = 14, xmin = 2.4) +
-  annotation_custom(hatch, ymin = 12, ymax = 14, xmin = 4.4) +
-  annotation_custom(nestling, ymin = 12, ymax = 14, xmin = 6.4) +
-  annotation_custom(nestling, ymin = 12, ymax = 14, xmin = 8.4)  + 
-  ylim(c(13,24)) +
-  guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) 
-p6
-
 p7 <- ggplot(meanPRL, aes(treatment, m)) +
@@
 
-plot_grid(p5,p6,p7,p8)
+p6 <- ggplot(candidates.pit, aes(x = treatment, y = PRL)) + 
+    geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) + 
+    scale_alpha_manual(values = c(0.75,1)) +
+    theme_B3() +
+  theme(legend.position = c(0.85,0.2), legend.direction = "horizontal") + 
+  scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+  labs(y = "PRL", x = NULL) +
+  guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) +
+  theme(axis.title.y  = element_text(face = "italic"))
+p6
+
+```
+
+```{r PRLBRCA1}
+brca <- c( "NP_989500.1")
+brca.pit <- selectcBRCA1vsds(brca, vsd.pit, colData.pit)
+
+meanBRCA <- brca.pit %>% 
+    droplevels() %>% 
+  dplyr::group_by(treatment) %>%
+  dplyr::summarise(m = mean(BRCA1), 
+                   se = sd(BRCA1)/sqrt(length(BRCA1))) %>%
+  dplyr::mutate(m = round(m,1)) 
+meanBRCA <- left_join(meanBRCA, birds)
+meanBRCA$treatment <- factor(meanBRCA$treatment, levels = alllevels)
+meanBRCA
+
+brca1 <- brca.pit %>% 
+  ggplot(aes(x = treatment, y = BRCA1, fill = treatment, color = sex)) +
+    geom_boxplot() +
+    theme_B3() +
+  scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+  theme(legend.position = "none") + 
+  labs(x = "increasing time >>", y = "BRCA1") +
+  annotation_custom(control, ymin = 5, ymax = 5.5, xmin = -7.8) +
+  annotation_custom(bldg, ymin = 5, ymax = 5.5, xmin = -5.8) +
+  annotation_custom(lay, ymin = 5, ymax = 5.5, xmin = -3.8) +
+  annotation_custom(inc, ymin = 5, ymax = 5.5, xmin = -1.8) +
+  annotation_custom(inc, ymin = 5, ymax = 5.5, xmin = 0.6) +
+  annotation_custom(inc, ymin = 5, ymax = 5.5, xmin = 2.4) +
+  annotation_custom(hatch, ymin = 5, ymax = 5.5, xmin = 4.4) +
+  annotation_custom(nestling, ymin = 5, ymax = 5.5, xmin = 6.4) +
+  annotation_custom(nestling, ymin = 5, ymax = 5.5, xmin = 8.4) +
+  ylim(5,8) +
+  theme(axis.title.y  = element_text(face = "italic"))
+brca1
+
+plot_grid(p6, brca1, nrow = 2)
 ```

### 2019-12-14 analysis/06c_DESeq2_ALL_hypothalamus.Rmd
now with manip volanos for gonads and hyp

---
title: "06_DESeqALL"
output: html_document
---

```{r setup, include = FALSE}

library(tidyverse)
library(DESeq2)
library(cowplot)
library(pheatmap)
library(viridis)
library(forcats)
library(caret) # LDA analysis
library(MASS) # LDA analysis
library(ggtext) # for pretty plots
library(ctsGE)

library(BiocParallel)
register(MulticoreParam(6))

source("../R/functions.R")  # load custom functions 
source("../R/themes.R")  # load custom themes and color palletes
source("../R/icons.R")  # for images

knitr::opts_chunk$set(fig.path = '../figures/DESeqAll/', cache = T,  message=F, comment=F, warning=F)
```

# Warning: This script takes a very long time to run.

DESeq2 was not designed to run on 300 samples. But, I really like it, so I do it anyways. Some of these commands take like 15 min to run using 6 cores.

## DEseq2 on all chaacterization and manipulations

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

# craete variable that will be critical for subset later on
colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
colData$sextissue <- as.factor(paste(colData$sex, colData$tissue, sep = "_"))
colData$treatment <- factor(colData$treatment, levels = alllevels)
colData$tissue <- factor(colData$tissue, levels = tissuelevels)
levels(colData$treatment)
```

```{r hypdfs, eval=T}
dds.hypothalamus <- subsetDESeq2(colData,  countData, c("female_hypothalamus","male_hypothalamus"))
```

```{r hypvsd, eval=T}
vsd <- as.data.frame(assay(vst(dds.hypothalamus, blind=FALSE)))
write.csv(vsd, "../results/06_hypallvsd.csv")
```

```{r  eval=T}
inc.d3.m.inc.d3 <- createDEGdfsave(dds.hypothalamus, "treatment", "m.inc.d3", "inc.d3", "hypothalamus")
```

```{r  eval=T}
inc.d9.m.inc.d9 <- createDEGdfsave(dds.hypothalamus, "treatment", "m.inc.d9", "inc.d9", "hypothalamus")
```

```{r  eval=T}
inc.d17.m.inc.d17 <- createDEGdfsave(dds.hypothalamus, "treatment", "m.inc.d17", "inc.d17", "hypothalamus")
```

```{r  eval=T}
hatch.m.n2 <- createDEGdfsave(dds.hypothalamus, "treatment", "m.n2", "hatch", "hypothalamus")
```

```{r  eval=T}
inc.d9.m.inc.d8 <- createDEGdfsave(dds.hypothalamus, "treatment", "m.inc.d8", "inc.d9", "hypothalamus")
```

```{r  eval=T}
hatch.m.inc.d8 <- createDEGdfsave(dds.hypothalamus, "treatment", "hatch", "m.inc.d8", "hypothalamus")
```

```{r  eval=T}
inc.d17.prolong <- createDEGdfsave(dds.hypothalamus, "treatment", "prolong", "inc.d17", "hypothalamus")
```

```{r  eval=T}
hatch.prolong <- createDEGdfsave(dds.hypothalamus, "treatment", "hatch", "prolong", "hypothalamus")
```

```{r  eval=T}
hatch.extend <- createDEGdfsave(dds.hypothalamus, "treatment", "extend", "hatch", "hypothalamus")
```

```{r  eval=T}
n5.extend <- createDEGdfsave(dds.hypothalamus, "treatment", "extend", "n5", "hypothalamus")
```

```{r volcanos-hyp, fig.width=7, fig.height=3.5}
a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano)  
b <- plot.volcano(inc.d9.m.inc.d9, "treatment",  "m.inc.d9", "inc.d9" , colorsvolcano) 
c <- plot.volcano(inc.d17.m.inc.d17 , "treatment",  "m.inc.d17", "inc.d17" , colorsvolcano) 
d <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano)

e <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano) 
f <- plot.volcano(hatch.m.inc.d8 , "treatment",    "hatch", "m.inc.d8"  , colorsvolcano)

g <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  
h <- plot.volcano(hatch.prolong , "treatment",  "hatch", "prolong", colorsvolcano)

i <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano)  
j <- plot.volcano(n5.extend , "treatment",  "extend", "n5", colorsvolcano)

remove <- plot_grid(a + labs(y = "hypothalamus -log10(p-value)"),
          b + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          c + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
          d + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow = 1, rel_widths = c(1.2,1,1,1),
          labels = c("a","b","c","d"), label_size = 9)
remove

timing <- plot_grid(e +  labs(y = "hypothalamus -log10(p-value)"),
          f + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          g + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          h + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
          j + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
          nrow = 1, rel_widths = c(1.2,1,1,1,1,1),
          labels = c("a","b","c", "d","e","f"), label_size = 9)
timing
```

### 2019-12-14 R/functions.R

@@ -780,10 +780,10 @@ plot.volcano <- function(data, whichfactor, up, down, mycolors){
                        name = " ",
                        drop = FALSE,
                        breaks=c(down, "NS", up)) +
-    ylim(c(-5,25)) +  
+    ylim(c(0,25)) +  
     xlim(c(-8,8)) +
     labs(y = "-log10(p)", x = " ")  +
-    theme(legend.position = "top",
+    theme(legend.position = "none",
           legend.direction = "horizontal",
           legend.spacing.x = unit(-0.1, 'cm'),
           legend.margin=margin(t=-0, r=0, b=0, l=0, unit="cm"),

### 2019-12-14 analysis/06_DESeq2_ALL.Rmd
now with manip volanos for gonads and hyp

--- 2019-12-02 analysis/06_DESeq2_ALL.Rmd
@@
 
-```{r pitplot, eval = T}
+```{r volcanos-pit}
 a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano) 
@@
 d <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano)
+
 e <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano)  
-f <- plot.volcano(hatch.m.inc.d8 , "treatment",    "m.inc.d8", "hatch"  , colorsvolcano) 
-g <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano) 
-h <- plot.volcano(hatch.prolong , "treatment",  "prolong", "hatch", colorsvolcano) 
+f <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  
+g <- plot.volcano(hatch.prolong , "treatment",  "prolong", "hatch", colorsvolcano)
+
+h <- plot.volcano(hatch.m.inc.d8 , "treatment",    "m.inc.d8", "hatch"  , colorsvolcano)  
 i <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano)  
@@
 
-pituitary <- plot_grid(a +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
+remove <- plot_grid(a +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                        labs(y = "pituitary \n -log10(p)"),
-          b + theme(axis.title = element_blank(), axis.text = element_blank()),
-          c + theme(axis.title = element_blank(), axis.text = element_blank()),
-          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          e + theme(axis.title = element_blank(), axis.text = element_blank()),
-          f + theme(axis.title = element_blank(), axis.text = element_blank()),
-          g + theme(axis.title = element_blank(), axis.text = element_blank()),
-          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          i + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          j + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          nrow = 2)
-pituitary
+          b + theme(axis.title = element_blank(), axis.text = element_blank()) ,
+          c + theme(axis.title = element_blank(), axis.text = element_blank()) ,
+          d + theme(axis.title = element_blank(), axis.text = element_blank()) , 
+          nrow = 1, rel_widths = c(1.2,1,1,1),
+          labels = c("a","b","c","d"), label_size = 9)
+remove
+
+manipinc <- plot_grid(e +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
+                       labs(y = "pituitary \n -log10(p)"),
+          f + theme(axis.title = element_blank(), axis.text = element_blank()) ,
+          g + theme(axis.title = element_blank(), axis.text = element_blank()) ,
+          nrow = 1, rel_widths = c(1.2,1,1),
+          labels = c("e","f","g"," "), label_size = 9)
+
+maniphatch <- plot_grid( h +  labs(y = "pituitary \n -log10(p)"), 
+          i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
+          j + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
+          nrow = 1, rel_widths = c(1.2,1,1),
+          labels = c("h","i","j"," "), label_size = 9)
+
+plot_grid(manipinc, maniphatch, nrow = 2)
 ```

### 2019-12-14 analysis/06b_DESeq2_ALL_gonad.Rmd
now with manip volanos for gonads and hyp

--- 2019-12-14 analysis/06_DESeq2_ALL.Rmd
@@
 
-```{r pitdfs, eval=T}
-dds.pituitary <- subsetDESeq2(colData,  countData, c("female_pituitary","male_pituitary"))
+```{r gondfs, eval=T}
+dds.gonads <- subsetDESeq2(colData,  countData, c("female_gonad","male_gonad"))
 ```
 
-```{r pitvsd, eval=T}
-vsd <- as.data.frame(assay(vst(dds.pituitary, blind=FALSE)))
-write.csv(vsd, "../results/06_pitallvsd.csv")
+```{r gonvsd, eval=T}
+vsd <- as.data.frame(assay(vst(dds.gonads, blind=FALSE)))
+write.csv(vsd, "../results/06_gonallvsd.csv")
 ```
@@
 ```{r  eval=T}
-inc.d3.m.inc.d3 <- createDEGdfsave(dds.pituitary, "treatment", "m.inc.d3", "inc.d3", "pituitary")
+inc.d3.m.inc.d3 <- createDEGdfsave(dds.gonads, "treatment", "m.inc.d3", "inc.d3", "gonads")
 ```
@@
 ```{r  eval=T}
-inc.d9.m.inc.d9 <- createDEGdfsave(dds.pituitary, "treatment", "m.inc.d9", "inc.d9", "pituitary")
+inc.d9.m.inc.d9 <- createDEGdfsave(dds.gonads, "treatment", "m.inc.d9", "inc.d9", "gonads")
 ```
@@
 ```{r  eval=T}
-inc.d17.m.inc.d17 <- createDEGdfsave(dds.pituitary, "treatment", "m.inc.d17", "inc.d17", "pituitary")
+inc.d17.m.inc.d17 <- createDEGdfsave(dds.gonads, "treatment", "m.inc.d17", "inc.d17", "gonads")
 ```
@@
 ```{r  eval=T}
-hatch.m.n2 <- createDEGdfsave(dds.pituitary, "treatment", "m.n2", "hatch", "pituitary")
+hatch.m.n2 <- createDEGdfsave(dds.gonads, "treatment", "m.n2", "hatch", "gonads")
 ```
@@
 ```{r  eval=T}
-inc.d9.m.inc.d8 <- createDEGdfsave(dds.pituitary, "treatment", "m.inc.d8", "inc.d9", "pituitary")
+inc.d9.m.inc.d8 <- createDEGdfsave(dds.gonads, "treatment", "m.inc.d8", "inc.d9", "gonads")
 ```
@@
 ```{r  eval=T}
-hatch.m.inc.d8 <- createDEGdfsave(dds.pituitary, "treatment", "m.inc.d8", "hatch", "pituitary")
+hatch.m.inc.d8 <- createDEGdfsave(dds.gonads, "treatment", "m.inc.d8", "hatch", "gonads")
 ```
@@
 ```{r  eval=T}
-inc.d17.prolong <- createDEGdfsave(dds.pituitary, "treatment", "prolong", "inc.d17", "pituitary")
+inc.d17.prolong <- createDEGdfsave(dds.gonads, "treatment", "prolong", "inc.d17", "gonads")
 ```
@@
 ```{r  eval=T}
-hatch.prolong <- createDEGdfsave(dds.pituitary, "treatment", "prolong", "hatch", "pituitary")
+hatch.prolong <- createDEGdfsave(dds.gonads, "treatment", "prolong", "hatch", "gonads")
 ```
@@
 ```{r  eval=T}
-hatch.extend <- createDEGdfsave(dds.pituitary, "treatment", "extend", "hatch", "pituitary")
+hatch.extend <- createDEGdfsave(dds.gonads, "treatment", "extend", "hatch", "gonads")
 ```
@@
 ```{r  eval=T}
-n5.extend <- createDEGdfsave(dds.pituitary, "treatment", "extend", "n5", "pituitary")
+n5.extend <- createDEGdfsave(dds.gonads, "treatment", "extend", "n5", "gonads")
 ```
 
-```{r volcanos-pit}
+```{r volcanos-gon}
 a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano) 
@@
 remove <- plot_grid(a +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
-                       labs(y = "pituitary \n -log10(p)"),
+                       labs(y = "gonads \n -log10(p)"),
           b + theme(axis.title = element_blank(), axis.text = element_blank()) ,
@@
 manipinc <- plot_grid(e +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
-                       labs(y = "pituitary \n -log10(p)"),
-          f + theme(axis.title = element_blank(), axis.text = element_blank()) ,
-          g + theme(axis.title = element_blank(), axis.text = element_blank()) ,
+                       labs(y = "gonads \n -log10(p)"),
+          f + theme(axis.title = element_blank(), axis.text = element_blank()),
+          g + theme(axis.title = element_blank(), axis.text = element_blank()),
           nrow = 1, rel_widths = c(1.2,1,1),
@@
 
-maniphatch <- plot_grid( h +  labs(y = "pituitary \n -log10(p)"), 
+maniphatch <- plot_grid( h + labs(y = "gonads \n -log10(p)"), 
           i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 

### 2019-12-14 analysis/06b_DESeq2_ALL_gonad.Rmd
better gonad and hypothalamus figures

@@
 ```{r  eval=T}
-hatch.m.inc.d8 <- createDEGdfsave(dds.gonads, "treatment", "m.inc.d8", "hatch", "gonads")
+hatch.m.inc.d8 <- createDEGdfsave(dds.gonads, "treatment", "hatch", "m.inc.d8", "gonads")
 ```
@@
 ```{r  eval=T}
-hatch.prolong <- createDEGdfsave(dds.gonads, "treatment", "prolong", "hatch", "gonads")
+hatch.prolong <- createDEGdfsave(dds.gonads, "treatment", "hatch", "prolong", "gonads")
 ```
@@
 ```{r  eval=T}
-n5.extend <- createDEGdfsave(dds.gonads, "treatment", "extend", "n5", "gonads")
+n5.extend <- createDEGdfsave(dds.gonads, "treatment", "n5", "extend", "gonads")
 ```
 
-```{r volcanos-gon}
+```{r volcanos-gon, fig.width=7, fig.height=3.5}
 a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano)  
@@
 e <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano) 
-f <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  
-g <- plot.volcano(hatch.prolong , "treatment",  "prolong", "hatch", colorsvolcano)
+f <- plot.volcano(hatch.m.inc.d8 , "treatment",    "hatch", "m.inc.d8"  , colorsvolcano)
 
-h <- plot.volcano(hatch.m.inc.d8 , "treatment",    "m.inc.d8", "hatch"  , colorsvolcano)  
+g <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  
+h <- plot.volcano(hatch.prolong , "treatment",  "hatch", "prolong", colorsvolcano)
+
 i <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano)  
@@
 
-remove <- plot_grid(a +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
-                       labs(y = "gonads \n -log10(p)"),
-          b + theme(axis.title = element_blank(), axis.text = element_blank()) ,
-          c + theme(axis.title = element_blank(), axis.text = element_blank()) ,
-          d + theme(axis.title = element_blank(), axis.text = element_blank()) , 
+remove <- plot_grid(a + labs(y = "gonads -log10(p-value)"),
+          b + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          c + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
+          d + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
           nrow = 1, rel_widths = c(1.2,1,1,1),
@@
 
-manipinc <- plot_grid(e +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
-                       labs(y = "gonads \n -log10(p)"),
-          f + theme(axis.title = element_blank(), axis.text = element_blank()),
-          g + theme(axis.title = element_blank(), axis.text = element_blank()),
-          nrow = 1, rel_widths = c(1.2,1,1),
-          labels = c("e","f","g"," "), label_size = 9)
-
-maniphatch <- plot_grid( h + labs(y = "gonads \n -log10(p)"), 
+timing <- plot_grid(e +  labs(y = "gonads -log10(p-value)"),
+          f + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          g + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          h + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
           i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
           j + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
-          nrow = 1, rel_widths = c(1.2,1,1),
-          labels = c("h","i","j"," "), label_size = 9)
-
-plot_grid(manipinc, maniphatch, nrow = 2)
+          nrow = 1, rel_widths = c(1.2,1,1,1,1,1),
+          labels = c("a","b","c", "d","e","f"), label_size = 9)
+timing
 ```

### 2019-12-15 analysis/12_GO.Rmd
merged GO and caididates

@@
 
+## add candidate genes
+
+```{r}
+candidates <- c(  "PRLR", 
+                 "AVP", "VIP", "AVPR1A", "AVPR1B", "OXT", 
+                 "GNRH1", "NPY",
+                "CRH", "FSHB", "GH", 
+                 "JAK2", "HCRT", "TRH", "TSHB",
+                "LH", "FSH",
+                "MC3R", "MC4R", "MC5R", 
+                "ESR1", "ESR2", "AR", 
+                "STAT5B", "CISH")
+candidates <- as.data.frame(candidates)
+names(candidates)[1] <- "gene"
+candidates$GO <- "candidategenes"
+candidates <- candidates %>% select(GO, gene)
+
+allGO <- rbind(candidates, allGO)
+head(allGO)
+```
+
 ## Differntially expressed genes by stage for each tissue
@@
   filter(GO %in% c("endocrineprocess", "lactation", 
-                   "parentalbehavior", "matingbehavior")) %>% 
+                   "parentalbehavior", "matingbehavior",
+                   "candidategenes")) %>% 
   map(~.x) %>%
@@
 ```{r}
-candidates <- c("AVP", "PRL", "PRLR", "GNIH",  "OXT", "AVPR1A", "AVPR1B",
-                "CRH", "FSHB", "GH", "GNRH1", "GNRH2",
-                "MC4R", "JAK2", "HCRT", "POMC", "TRH", "TSHB",
-                "LH", "FSH")
+candidates <- c( "PRL", "PRLR", 
+                 "AVP", "VIP", "AVPR1A", "AVPR1B", "OXT", 
+                 "GNRH1", "GNRHR", "NPY",
+                "CRH", "FSHB", "GH", 
+                 "JAK2", "HCRT", "POMC", "TRH", "TSHB",
+                "LH", "FSH",
+                "POMC", "MC3R", "MC4R", "MC5R", 
+                "ESR1", "ESR2", "AR", 
+                "STAT5", "CISH")
 candidates <- as.data.frame(candidates)

### 2019-12-15 analysis/10_specificgenes.Rmd
now with candidate gene correlations

--- 2019-10-20 analysis/10_specificgenes.Rmd
@@
 library(cowplot)
+library(corrr)
+library(ggcorrplot)
+library(pheatmap)
 
@@
 
+colData.hyp <- colData.hyp %>% 
+  mutate(tissue = sapply(strsplit(sample,'\\_'), "[", 3)) %>%
+  mutate(bird = sapply(strsplit(sample,'\\_'), "[", 1)) %>%
+  select(sample, bird, sex, tissue, treatment)
+
+colData.pit <- colData.pit %>%  
+  mutate(tissue = sapply(strsplit(sample,'\\_'), "[", 3)) %>%
+  mutate(bird = sapply(strsplit(sample,'\\_'), "[", 1)) %>%
+  select(sample, bird, sex, tissue, treatment)
+
+colData.gon <- colData.gon %>%  
+  mutate(tissue = sapply(strsplit(sample,'\\_'), "[", 3)) %>%
+  mutate(bird = sapply(strsplit(sample,'\\_'), "[", 1)) %>%
+  select(sample, bird, sex, tissue, treatment)
+
 ```
@@
 
-```{r PRLboxplots, fig.width =5.5, fig.height= 2}
-candidategenes <- c("CISH", "SOCS1", "SOCS2", "SOCS2", "SOCS4", "SOCS5", "SOCS6", 
-                    "PRL", "PRLR", "STAT5B", "JAK1")
+```{r candidategenes, fig.width =5.5, fig.height= 4}
+candidategenes <- c("PRL","PRLR", "GNRHR")
+
 candidates.hyp <- selectcandidatevsds(candidategenes, vsd.hyp, colData.hyp)
@@
   ggplot(candidateexpression, aes(x = as.numeric(treatment), y = whichgene)) + 
-    geom_boxplot(aes(fill = treatment, alpha = sex)) + 
-    scale_alpha_manual(values = c(0.6,0.9)) +
+        geom_smooth(aes(colour = sex)) +
+    geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) + 
+    scale_alpha_manual(values = c(0.75,1)) +
      theme_B3() +
@@
 
-plot_grid(a,b + labs(y = NULL),c + labs(y = NULL), nrow = 1)
+d <- plotcanddateexpression(candidates.hyp, "hypothalamus", candidates.hyp$PRLR, "PRLR")
+e <- plotcanddateexpression(candidates.pit, "pituitary", candidates.pit$PRLR, "PRLR")
+f <- plotcanddateexpression(candidates.gon, "gonad", candidates.gon$PRLR, "PRLR")
 
-e <- plotcanddateexpression(candidates.hyp, NULL, candidates.hyp$PRLR, "PRLR")
-f <- plotcanddateexpression(candidates.pit, NULL, candidates.pit$PRLR, "PRLR")
-g <- plotcanddateexpression(candidates.gon, NULL, candidates.gon$PRLR, "PRLR")
+g <- plotcanddateexpression(candidates.hyp, "hypothalamus", candidates.hyp$GNRHR, "GNRHR")
+h <- plotcanddateexpression(candidates.pit, "pituitary", candidates.pit$GNRHR, "GNRHR")
+i <- plotcanddateexpression(candidates.gon, "gonad", candidates.gon$GNRHR, "GNRHR")
 
-plot_grid(e,f + labs(y = NULL),g + labs(y = NULL), nrow = 1)
+abc <- plot_grid(a, d, g, nrow = 1)
+def <- plot_grid(b, e, h, nrow = 1)
+ghi <- plot_grid(c, f, i, nrow = 1)
 
-e <- plotcanddateexpression(candidates.hyp, NULL, candidates.hyp$STAT5B, "STAT5B")
-f <- plotcanddateexpression(candidates.pit, NULL, candidates.pit$STAT5B, "STAT5B")
-g <- plotcanddateexpression(candidates.gon, NULL, candidates.gon$STAT5B, "STAT5B")
+plot_grid(abc, def,ghi,nrow = 3)
+```
 
-plot_grid(e,f + labs(y = NULL),g + labs(y = NULL), nrow = 1)
+```{r PRLdf}
+PRLpit <- candidates.pit %>%
+  select(bird, sex,treatment,tissue, PRL)
+head(PRLpit)
+write.csv(PRLpit, "../results/10_PRLpit.csv", row.names = F)
 ```
+
+```{r correlations-incd9-incd17}
+pitcandidates <- c("ATP6AP2", "CRY1", "TMF1",
+                   "EIF2AK3", "PRL", "SOCS2", "STAT5B", "XBP1",
+                   "APP", "HDAC4", "NCOA1")
+goncandidates <- c("AVPR2", "BMP6", "CRHR2", "DRD5", "TAC1",
+                   "SOCS2", "ZBTB7B",
+                   "DRD5", "NCOA2", "BRINP1", "NR3C1")
+
+candidates.pit <- selectcandidatevsds(pitcandidates, vsd.pit, colData.pit) %>% select(-sample) %>% filter(treatment %in% c("inc.d9", "inc.d17"))
+candidates.gon <- selectcandidatevsds(goncandidates, vsd.gon, colData.gon) %>% select(-sample) %>% filter(treatment %in% c("inc.d9", "inc.d17"))
+
+candidates.pit <- candidates.pit %>% rename_at(vars(-bird,-sex, -tissue, -treatment), function(x) paste0("pit_", x)) %>% select(-tissue)
+candidates.gon <- candidates.gon %>% rename_at(vars(-bird,-sex, -tissue, -treatment), function(x) paste0("gon_", x)) %>% select(-tissue)
+
+counts <- full_join(candidates.pit, candidates.gon) %>% drop_na() %>% select(-bird, -sex, -treatment)
+cols <- full_join(candidates.pit, candidates.gon) %>% drop_na() %>% select(bird, sex, treatment)
+dim(counts)
+dim(cols)
+
+x <- correlate(counts)
+fashion(x)
+rplot(x)  + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  labs(subtitletitle = "inc.d9 to inc.d17")
+
+x %>% network_plot.cor_df(min_cor = .5) + 
+  labs(subtitletitle = "inc.d9 to inc.d17")
+
+cal_z_score <- function(x){
+  (x - mean(x)) / sd(x)
+}
+
+tcounts <- t(counts)
+cols <- cols %>% select(-bird)
+
+data_subset_norm <- t(apply(tcounts, 1, cal_z_score))
+row.names(cols) <- colnames(data_subset_norm)
+
+pheatmap(data_subset_norm,  annotation_col = cols)
+
+corr <- x
+row.names(corr) <- corr$rowname 
+corr$rowname <- NULL
+ggcorrplot(corr, hc.order = TRUE, outline.col = "white", title = "inc.d9 to inc.d17")
+```
+
+```{r correlations-bldg-lay}
+pitcandidates <- c("BMP6", "CGA", "GAL", "XBP1")
+goncandidates <- c("CCND1", "NME1", "PRL", "APLP2", "CYP11A1", "CREBRF", "DBH")
+
+candidates.pit <- selectcandidatevsds(pitcandidates, vsd.pit, colData.pit) %>% select(-sample) %>% filter(treatment %in% c("bldg", "lay"))
+candidates.gon <- selectcandidatevsds(goncandidates, vsd.gon, colData.gon) %>% select(-sample) %>% filter(treatment %in% c("bldg", "lay"))
+
+candidates.pit <- candidates.pit %>% rename_at(vars(-bird,-sex, -tissue, -treatment), function(x) paste0("pit_", x)) %>% select(-tissue)
+candidates.gon <- candidates.gon %>% rename_at(vars(-bird,-sex, -tissue, -treatment), function(x) paste0("gon_", x)) %>% select(-tissue)
+
+counts <- full_join(candidates.pit, candidates.gon) %>% drop_na() %>% select(-bird, -sex, -treatment)
+cols <- full_join(candidates.pit, candidates.gon) %>% drop_na() %>% select(bird, sex, treatment)
+dim(counts)
+dim(cols)
+
+x <- correlate(counts)
+fashion(x)
+rplot(x)  + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  labs(subtitle = "bldg to lay")
+
+x %>% network_plot.cor_df(min_cor = .5) + labs(subtitle = "bldg to lay")
+
+corr <- x
+row.names(corr) <- corr$rowname 
+corr$rowname <- NULL
+ggcorrplot(corr, hc.order = TRUE, outline.col = "white", title = "bldg to lay")
+```
+
+```{r correlations-lay-incd3}
+pitcandidates <- c("PTEN", "CGA", "GAL", "XBP1")
+goncandidates <- c(
+  "AVPR1A", "C1QTNF1", "CRY1", "CRY2", "F2R", "GJA1", "INHBA", "RHOA", "SMAD4",
+  "ATP2B2", "ATP7B", "CAV1", "EIF2AK3", "ERBB4", "GJA1", "HIF1A", "MED1", "NME1", 
+  "PRLR", "VEGFA", "XBP1",  "ZBTB7B", "APLP2", "NCOA1", "NCOA2", "THRB",
+  "CREBRF", "GNAQ", "NR3C1")
+
+candidates.pit <- selectcandidatevsds(pitcandidates, vsd.pit, colData.pit) %>% select(-sample) %>% filter(treatment %in% c("inc.d3", "lay"))
+candidates.gon <- selectcandidatevsds(goncandidates, vsd.gon, colData.gon) %>% select(-sample) %>% filter(treatment %in% c("inc.d3", "lay"))
+
+candidates.pit <- candidates.pit %>% rename_at(vars(-bird,-sex, -tissue, -treatment), function(x) paste0("pit_", x)) %>% select(-tissue)
+candidates.gon <- candidates.gon %>% rename_at(vars(-bird,-sex, -tissue, -treatment), function(x) paste0("gon_", x)) %>% select(-tissue)
+
+counts <- full_join(candidates.pit, candidates.gon) %>% drop_na() %>% select(-bird, -sex, -treatment)
+cols <- full_join(candidates.pit, candidates.gon) %>% drop_na() %>% select(bird, sex, treatment)
+dim(counts)
+dim(cols)
+
+x <- correlate(counts)
+fashion(x)
+rplot(x) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  labs(subtitle = "lay to inc.d3")
+
+x %>% network_plot.cor_df(min_cor = .5)  + 
+  labs(subtitle = "lay to inc.d3")
+
+corr <- x
+row.names(corr) <- corr$rowname 
+corr$rowname <- NULL
+ggcorrplot(corr, hc.order = TRUE, outline.col = "white", title = "lay to inc.d3")
+```
+
+```{r correlations-hatch-n5}
+pitcandidates <- c( "ATP6AP2", "AVPR2", "CRY2", "RHOA", 
+                    "EIF2AK3", "NME1", "PRL", "XBP1",
+                    "APP", "NCOA1")
+hypcandidates <- c("CRHBP", "CRHR2", "DRD3", "EDN1", "FAM129B", "FOXL2", "KRAS",
+                   "OPRK1", "POMC", "RAB11FIP5", "RUNX1",
+                   "CAV1", "ZBTB7B", "HEXB", "PGR")
+
+candidates.pit <- selectcandidatevsds(pitcandidates, vsd.pit, colData.pit) %>% select(-sample) %>% filter(treatment %in% c("hatch", "n5"))
+candidates.hyp <- selectcandidatevsds(hypcandidates, vsd.hyp, colData.hyp) %>% select(-sample) %>% filter(treatment %in% c("hatch", "n5"))
+
+candidates.pit <- candidates.pit %>% rename_at(vars(-bird,-sex, -tissue, -treatment), function(x) paste0("pit_", x)) %>% select(-tissue)
+candidates.hyp <- candidates.hyp %>% rename_at(vars(-bird,-sex, -tissue, -treatment), function(x) paste0("hyp_", x)) %>% select(-tissue)
+
+counts <- full_join(candidates.pit, candidates.hyp) %>% drop_na() %>% select(-bird, -sex, -treatment)
+cols <- full_join(candidates.pit, candidates.hyp) %>% drop_na() %>% select(bird, sex, treatment)
+dim(counts)
+dim(cols)
+
+x <- correlate(counts)
+fashion(x)
+rplot(x) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  labs(subtitle = "hatch to n5")
+
+x %>% network_plot.cor_df(min_cor = .5)  + 
+  labs(subtitle = "hatch to n5")
+
+corr <- x
+row.names(corr) <- corr$rowname 
+corr$rowname <- NULL
+ggcorrplot(corr, hc.order = TRUE, outline.col = "white", title = "hatch to n5")
+```

### 2019-12-15 R/functions.R

@@ -892,3 +892,105 @@ selectcBRCA1vsds <- function(listofgenes, vsd, colData){
   return(candidatevsd)
 }
 
+
+
+
+
+
+
+
+### modified corrr fucniont
+
+network_plot.cor_df <- function(rdf,
+                                min_cor = .30,
+                                legend = TRUE,
+                                colours = c("indianred2", "white", "skyblue1"),
+                                repel = TRUE,
+                                curved = TRUE,
+                                colors) {
+  
+  if (min_cor < 0 || min_cor > 1) {
+    stop ("min_cor must be a value ranging from zero to one.")
+  }
+  
+  if (!missing(colors))
+    colours <- colors
+  
+  rdf <-  as_matrix(rdf, diagonal = 1)
+  distance <- sign(rdf) * (1 - abs(rdf))
+  
+  # Use multidimensional Scaling to obtain x and y coordinates for points.
+  points <- data.frame(stats::cmdscale(abs(distance)))
+  colnames(points) <-  c("x", "y")
+  points$id <- rownames(points)
+  
+  # Create a proximity matrix of the paths to be plotted.
+  proximity <- abs(rdf)
+  proximity[upper.tri(proximity)] <- NA
+  diag(proximity) <- NA
+  proximity[proximity < min_cor] <- NA
+  
+  # Produce a data frame of data needed for plotting the paths.
+  n_paths <- sum(!is.na(proximity))
+  paths <- data.frame(matrix(nrow = n_paths, ncol = 6)) 
+  colnames(paths) <- c("x", "y", "xend", "yend", "proximity", "sign")
+  path <- 1
+  for(row in 1:nrow(proximity)) {
+    for(col in 1:ncol(proximity)) {
+      path_proximity <- proximity[row, col]
+      if (!is.na(path_proximity)) {
+        path_sign <- sign(distance[row, col])
+        x    <- points$x[row]
+        y    <- points$y[row]
+        xend <- points$x[col]
+        yend <- points$y[col]
+        paths[path, ] <- c(x, y, xend, yend, path_proximity, path_sign)
+        path <- path + 1
+      }
+    }
+  }
+  
+  plot_ <- list(
+    # For plotting paths
+    if (curved) geom_curve(data = paths,
+                           aes(x = x, y = y, xend = xend, yend = yend,
+                               alpha = proximity, 
+                               colour = proximity*sign)), 
+    if (!curved) geom_segment(data = paths,
[…]

### 2019-12-15 analysis/10_specificgenes.Rmd
new colors

@@
 ```{r candidategenes, fig.width =5.5, fig.height= 4}
-candidategenes <- c("PRL","PRLR", "GNRHR")
+candidategenes <- c("PRL","PRLR", "CRY1")
 
@@
 
-g <- plotcanddateexpression(candidates.hyp, "hypothalamus", candidates.hyp$GNRHR, "GNRHR")
-h <- plotcanddateexpression(candidates.pit, "pituitary", candidates.pit$GNRHR, "GNRHR")
-i <- plotcanddateexpression(candidates.gon, "gonad", candidates.gon$GNRHR, "GNRHR")
+g <- plotcanddateexpression(candidates.hyp, "hypothalamus", candidates.hyp$CRY1, "CRY1")
+h <- plotcanddateexpression(candidates.pit, "pituitary", candidates.pit$CRY1, "CRY1")
+i <- plotcanddateexpression(candidates.gon, "gonad", candidates.gon$CRY1, "CRY1")
 
@@
 cols <- full_join(candidates.pit, candidates.gon) %>% drop_na() %>% select(bird, sex, treatment)
-dim(counts)
-dim(cols)
 
 x <- correlate(counts)
-fashion(x)
 rplot(x)  + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
@@
 
-x %>% network_plot.cor_df(min_cor = .5) + 
+x %>% network_plot.cor_df(min_cor = .6, colors = c("#67a9cf", "white", "#ef8a62")) + 
   labs(subtitletitle = "inc.d9 to inc.d17")
-
-cal_z_score <- function(x){
-  (x - mean(x)) / sd(x)
-}
-
-tcounts <- t(counts)
-cols <- cols %>% select(-bird)
-
-data_subset_norm <- t(apply(tcounts, 1, cal_z_score))
-row.names(cols) <- colnames(data_subset_norm)
-
-pheatmap(data_subset_norm,  annotation_col = cols)
 
@@
 
-x %>% network_plot.cor_df(min_cor = .5) + labs(subtitle = "bldg to lay")
+x %>% network_plot.cor_df(min_cor = .6, colors = c("#67a9cf", "white", "#ef8a62")) + 
+  labs(subtitle = "bldg to lay")
 
@@
 cols <- full_join(candidates.pit, candidates.gon) %>% drop_na() %>% select(bird, sex, treatment)
-dim(counts)
-dim(cols)
 
 x <- correlate(counts)
-fashion(x)
 rplot(x) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
@@
 
-x %>% network_plot.cor_df(min_cor = .5)  + 
+x %>% network_plot.cor_df(min_cor = .6, colors = c("#67a9cf", "white", "#ef8a62"))  + 
   labs(subtitle = "lay to inc.d3")
@@
 cols <- full_join(candidates.pit, candidates.hyp) %>% drop_na() %>% select(bird, sex, treatment)
-dim(counts)
-dim(cols)
 
@@
 
-x %>% network_plot.cor_df(min_cor = .5)  + 
+x %>% network_plot.cor_df(min_cor = .6, colors = c("#67a9cf", "white", "#ef8a62"))  + 
   labs(subtitle = "hatch to n5")

### 2019-12-15 analysis/10_specificgenes.Rmd
remove corrggplot thing

@@
 library(corrr)
-library(ggcorrplot)
-library(pheatmap)
 
@@
 
-corr <- x
-row.names(corr) <- corr$rowname 
-corr$rowname <- NULL
-ggcorrplot(corr, hc.order = TRUE, outline.col = "white", title = "inc.d9 to inc.d17")
 ```
@@
 
-corr <- x
-row.names(corr) <- corr$rowname 
-corr$rowname <- NULL
-ggcorrplot(corr, hc.order = TRUE, outline.col = "white", title = "bldg to lay")
 ```
@@
 
-corr <- x
-row.names(corr) <- corr$rowname 
-corr$rowname <- NULL
-ggcorrplot(corr, hc.order = TRUE, outline.col = "white", title = "lay to inc.d3")
 ```
@@
 
-corr <- x
-row.names(corr) <- corr$rowname 
-corr$rowname <- NULL
-ggcorrplot(corr, hc.order = TRUE, outline.col = "white", title = "hatch to n5")
 ```

### 2019-12-15 analysis/10_specificgenes.Rmd
better colors

@@
 
-```{r correlations-incd9-incd17}
-pitcandidates <- c("ATP6AP2", "CRY1", "TMF1",
-                   "EIF2AK3", "PRL", "SOCS2", "STAT5B", "XBP1",
-                   "APP", "HDAC4", "NCOA1")
-goncandidates <- c("AVPR2", "BMP6", "CRHR2", "DRD5", "TAC1",
-                   "SOCS2", "ZBTB7B",
-                   "DRD5", "NCOA2", "BRINP1", "NR3C1")
-
-candidates.pit <- selectcandidatevsds(pitcandidates, vsd.pit, colData.pit) %>% select(-sample) %>% filter(treatment %in% c("inc.d9", "inc.d17"))
-candidates.gon <- selectcandidatevsds(goncandidates, vsd.gon, colData.gon) %>% select(-sample) %>% filter(treatment %in% c("inc.d9", "inc.d17"))
-
-candidates.pit <- candidates.pit %>% rename_at(vars(-bird,-sex, -tissue, -treatment), function(x) paste0("pit_", x)) %>% select(-tissue)
-candidates.gon <- candidates.gon %>% rename_at(vars(-bird,-sex, -tissue, -treatment), function(x) paste0("gon_", x)) %>% select(-tissue)
-
-counts <- full_join(candidates.pit, candidates.gon) %>% drop_na() %>% select(-bird, -sex, -treatment)
-cols <- full_join(candidates.pit, candidates.gon) %>% drop_na() %>% select(bird, sex, treatment)
-
-x <- correlate(counts)
-rplot(x)  + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
-  labs(subtitletitle = "inc.d9 to inc.d17")
-
-x %>% network_plot.cor_df(min_cor = .6, colors = c("#67a9cf", "white", "#ef8a62")) + 
-  labs(subtitletitle = "inc.d9 to inc.d17")
-
-```
-
 ```{r correlations-bldg-lay}
@@
 x <- correlate(counts)
-rplot(x)  + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+rplot(x, colors = c("skyblue1", "white", "indianred2"))  + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
   labs(subtitle = "bldg to lay")
 
-x %>% network_plot.cor_df(min_cor = .6, colors = c("#67a9cf", "white", "#ef8a62")) + 
+x %>% network_plot.cor_df(min_cor = .6 , colors = c("skyblue1", "white", "indianred2")) + 
   labs(subtitle = "bldg to lay")
@@
 x <- correlate(counts)
-rplot(x) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+rplot(x, colors = c("skyblue1", "white", "indianred2")) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
@@
 
-x %>% network_plot.cor_df(min_cor = .6, colors = c("#67a9cf", "white", "#ef8a62"))  + 
+x %>% network_plot.cor_df(min_cor = .6 , colors = c("skyblue1", "white", "indianred2"))  + 
   labs(subtitle = "lay to inc.d3")
+
+```
+
+```{r correlations-incd9-incd17}
+pitcandidates <- c("ATP6AP2", "CRY1", "TMF1",
+                   "EIF2AK3", "PRL", "SOCS2", "STAT5B", "XBP1",
+                   "APP", "HDAC4", "NCOA1")
+goncandidates <- c("AVPR2", "BMP6", "CRHR2", "DRD5", "TAC1",
+                   "SOCS2", "ZBTB7B",
+                   "DRD5", "NCOA2", "BRINP1", "NR3C1")
+
+candidates.pit <- selectcandidatevsds(pitcandidates, vsd.pit, colData.pit) %>% select(-sample) %>% filter(treatment %in% c("inc.d9", "inc.d17"))
+candidates.gon <- selectcandidatevsds(goncandidates, vsd.gon, colData.gon) %>% select(-sample) %>% filter(treatment %in% c("inc.d9", "inc.d17"))
+
+candidates.pit <- candidates.pit %>% rename_at(vars(-bird,-sex, -tissue, -treatment), function(x) paste0("pit_", x)) %>% select(-tissue)
+candidates.gon <- candidates.gon %>% rename_at(vars(-bird,-sex, -tissue, -treatment), function(x) paste0("gon_", x)) %>% select(-tissue)
+
+counts <- full_join(candidates.pit, candidates.gon) %>% drop_na() %>% select(-bird, -sex, -treatment)
+cols <- full_join(candidates.pit, candidates.gon) %>% drop_na() %>% select(bird, sex, treatment)
+
+x <- correlate(counts)
+rplot(x, colors = c("skyblue1", "white", "indianred2"))  + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  labs(subtitletitle = "inc.d9 to inc.d17")
+
+x %>% network_plot.cor_df(min_cor = .6 , colors = c("skyblue1", "white", "indianred2")) + 
+  labs(subtitletitle = "inc.d9 to inc.d17")
 
@@
 x <- correlate(counts)
-fashion(x)
-rplot(x) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+rplot(x,  colors = c("skyblue1", "white", "indianred2")) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
@@
 
-x %>% network_plot.cor_df(min_cor = .6, colors = c("#67a9cf", "white", "#ef8a62"))  + 
+x %>% network_plot.cor_df(min_cor = .6, colors = c("skyblue1", "white", "indianred2"))  + 
   labs(subtitle = "hatch to n5")

### 2019-12-15 analysis/10_specificgenes.Rmd
stil working on colors

@@
 x <- correlate(counts)
+
 rplot(x, colors = c("skyblue1", "white", "indianred2"))  + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 

### 2019-12-19 analysis/00_datawrangling.Rmd
working on gene level counts not transcript levle

@@
 # import count data, set rows at entreziz
-countData <- read.table("../results/kallistocounts.txt", 
+kallistodata <- read.table("../results/kallistocounts.txt", 
                         sep = ",", row.names = NULL)
 
-## set row names as entrezid for deseq2
-row.names(countData) <- countData$entrezid
+## save gene information
+geneinfo <- kallistodata %>%
+  select(row.names, Name, geneid, entrezid)
+head(geneinfo)
+
+# check for genes that have multple transcripts expressed
+isoforms <- kallistodata %>%
+  group_by(Name) %>%
+  summarize(n = n()) %>%
+  arrange(desc(n), Name) %>%
+  filter(n > 1)
+head(isoforms)
+
+# for example, this gene has 
+geneinfo %>% filter(Name == "CEND1")
+
+# aggregate transcript counts to gene counts
+countData <- kallistodata %>% 
+  select(-row.names, -geneid, -entrezid) %>% 
+  rename(gene = Name) %>%
+  pivot_longer(-gene, names_to = "samples", values_to = "counts") %>%  
+  pivot_wider(
+    names_from = samples, 
+    values_from = counts,
+    values_fn = list(counts = sum))  %>% 
+  arrange(gene)
+head(countData)
+
+countData <- as.data.frame(countData)
+row.names(countData) <- countData$gene ## make gene the row name
+countData[1] <- NULL ## make gene the row name
+countData <- round(countData) #round all value to nearest 1s place
 
@@
 colnames(countData) <- gsub('-', '.', colnames(countData))
-
-# create df with gene info and remove this from countData
-geneinfo <- countData %>%
-  select(row.names, Name, geneid, entrezid)
-head(geneinfo)
-countData <- countData %>%
-  select (-c(row.names, Name, geneid, entrezid))
-
-# because DESeq require integers
-countData <- round(countData)
+head(countData,4)
 
@@
 dim(countData)
-
-# make a tibble for the sole purpose of previewing this huge df
-countDataTibble <- as.tbl(countData)
-print(countDataTibble, n_extra = 0)
-#rm(countDataTibble)
 ```
@@
 write.csv(geneinfo, "../metadata/00_geneinfo.csv", row.names = TRUE)
+
+write.csv(isoforms, "../results/00_geneswithisoformss.csv", row.names = T)
 ```

### 2019-12-19 analysis/06_DESeq2_ALL.Rmd
working on gene level counts not transcript levle

@@
 ```{r  eval=T}
-hatch.m.inc.d8 <- createDEGdfsave(dds.pituitary, "treatment", "m.inc.d8", "hatch", "pituitary")
+hatch.m.inc.d8 <- createDEGdfsave(dds.pituitary, "treatment", "hatch", "m.inc.d8", "pituitary")
 ```
@@
 ```{r  eval=T}
-hatch.prolong <- createDEGdfsave(dds.pituitary, "treatment", "prolong", "hatch", "pituitary")
+hatch.prolong <- createDEGdfsave(dds.pituitary, "treatment", "hatch", "prolong", "pituitary")
 ```
@@
 ```{r  eval=T}
-n5.extend <- createDEGdfsave(dds.pituitary, "treatment", "extend", "n5", "pituitary")
+n5.extend <- createDEGdfsave(dds.pituitary, "treatment", "n5", "extend", "pituitary")
 ```
@@
 
-remove <- plot_grid(a +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
-                       labs(y = "pituitary \n -log10(p)"),
-          b + theme(axis.title = element_blank(), axis.text = element_blank()) ,
-          c + theme(axis.title = element_blank(), axis.text = element_blank()) ,
-          d + theme(axis.title = element_blank(), axis.text = element_blank()) , 
+remove <- plot_grid(a + labs(y = "pituitary -log10(p-value)"),
+          b + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          c + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
+          d + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
           nrow = 1, rel_widths = c(1.2,1,1,1),
@@
 
-manipinc <- plot_grid(e +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
-                       labs(y = "pituitary \n -log10(p)"),
-          f + theme(axis.title = element_blank(), axis.text = element_blank()) ,
-          g + theme(axis.title = element_blank(), axis.text = element_blank()) ,
-          nrow = 1, rel_widths = c(1.2,1,1),
-          labels = c("e","f","g"," "), label_size = 9)
-
-maniphatch <- plot_grid( h +  labs(y = "pituitary \n -log10(p)"), 
+timing <- plot_grid(e +  labs(y = "pituitary -log10(p-value)"),
+          f + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          g + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          h + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
           i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
           j + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
-          nrow = 1, rel_widths = c(1.2,1,1),
-          labels = c("h","i","j"," "), label_size = 9)
-
-plot_grid(manipinc, maniphatch, nrow = 2)
+          nrow = 1, rel_widths = c(1.2,1,1,1,1,1),
+          labels = c("a","b","c", "d","e","f"), label_size = 9)
+timing
 ```

### 2019-12-19 analysis/20_forrechelle.Rmd
new music plot for rechelle

---
title: "forrechelle"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggimage)
library(cowplot)

knitr::opts_chunk$set(echo = TRUE)
```

```{r input}

candidategenes <- c( "GH", "NR3C2", "PRLR", "PRL")
timepoints <- c("lay" , "inc.d9", "hatch", "n9")
hormones <- c("PRL", "CORT")

```

```{r readfile}

geneids <- read_csv("..//metadata/00_geneinfo.csv")

datapath <- "../results/"   # path to the data
datafiles <- dir(datapath, pattern = "04_vsd*") # get file names
datapathfile <- paste0(datapath, datafiles)

df <- datapathfile %>%
  setNames(nm = .) %>%
  map_df(~read_csv(.x, col_types = cols(), col_names = T), .id = "filename") %>% 
  mutate(tissue = sapply(strsplit(as.character(filename),'../results/04_vsd_'), "[", 2)) %>% 
  mutate(tissue = sapply(strsplit(as.character(tissue),'.csv'), "[", 1))  %>% 
  left_join(geneids) %>%
  select(tissue, X1, Name, everything()) %>% select(-filename, -geneid, - entrezid, - row.names) 
head(df)

df2 <-  df  %>%
  filter(Name %in% candidategenes) %>%
  pivot_longer(cols = L.Blu13_male_gonad_control.NYNO:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "sample", values_to = "vsd") %>%
   mutate(sex = sapply(strsplit(as.character(sample),'_'), "[", 2)) %>%
   mutate(treatment = sapply(strsplit(as.character(sample),'_'), "[", 4))  %>%
   mutate(treatment = sapply(strsplit(as.character(treatment),'.NYNO'), "[", 1)) %>%
  mutate(bird = sapply(strsplit(as.character(sample),'_'), "[", 1)) %>%
  filter(treatment %in% timepoints) %>%
  select(bird, sex, treatment, tissue, Name, vsd) %>%
  rename(gene = Name)
head(df2)

df3 <- df2 %>% 
  mutate(tissue = fct_relevel(tissue,  "hyp", "pit", "gon")) %>% 
  mutate(treatment = fct_relevel(treatment, timepoints)) %>% 
  group_by(sex, treatment, tissue, gene)  %>% 
  summarize(m = mean(vsd, na.rm = T), se = sd(vsd,  na.rm = T)/sqrt(length(vsd))) %>%
  mutate(image = "../figures/images/DoveParentsRNAseq_note.png")
head(df3)

for (i in levels(df3$tissue)) {

p <-  df3 %>%
  filter(tissue == i) %>%
  ggplot(aes(x = treatment, y = m)) +
  geom_image(aes(image=image, color = gene), size=0.1)  +

  scale_alpha_manual(values = c(0.5, 1)) +
  labs(subtitle = i, y = "average expression", x = "parental stage")

pimage <- axis_canvas(p, axis = 'x') + 
  draw_image("../figures/images/icons/DoveParentsRNAseq_lay.png", x = 0.5, scale = 1 ) +
  draw_image("../figures/images/icons/DoveParentsRNAseq_incubation.png", x = 1.5, scale = 1) +
  draw_image("../figures/images/icons/DoveParentsRNAseq_hatch.png", x = 2.5, scale = 1) +
  draw_image("../figures/images/icons/DoveParentsRNAseq_chicklings.png", x = 3.5, scale = 1)

p2 <- ggdraw(insert_xaxis_grob(p, pimage, position = "bottom"))
plot(p2)

}
```

### 2019-12-19 analysis/00_datawrangling.Rmd
data wrangling

@@
 countData <- round(countData) #round all value to nearest 1s place
-
-# replace dashes with periods (to make deseq happy)
-colnames(countData) <- gsub('-', '.', colnames(countData))
-head(countData,4)
+head(countData)
 
@@
   distinct()
+birds
 

### 2019-12-19 analysis/00_datawrangling.Rmd
deseq all now working but not run yet

@@
                         sep = ",", row.names = NULL)
+head(kallistodata)
 
@@
 geneinfo <- kallistodata %>%
-  select(row.names, Name, geneid, entrezid)
+  select(Name, geneid, entrezid)
 head(geneinfo)
@@
 # for example, this gene has 
-geneinfo %>% filter(Name == "CEND1")
+geneinfo %>% filter(Name == "GRIN1")
 
@@
     values_fn = list(counts = sum))  %>% 
-  arrange(gene)
+  arrange(gene) %>% 
+  filter(gene != "")
 head(countData)

### 2019-12-19 analysis/03_DESeq2_characterization.Rmd
deseq all now working but not run yet

@@
 head(geneinfo)
+
+# confirm gene-level analysis
+c.countData %>% 
+  mutate(gene = rownames(c.countData)) %>% 
+  filter(gene == 'GRIN1') %>% 
+  select(gene, everything())
 ```
 
-## Run DESeq on all subsets of the data
-
-```{r dds, cache = T, eval = T}
-dds.female_hypothalamus <- subsetDESeq(c.colData,  c.countData, "female_hypothalamus")
-dds.female_pituitary <- subsetDESeq(c.colData,  c.countData, "female_pituitary" )
-dds.female_gonad <- subsetDESeq(c.colData,  c.countData, "female_gonad" )
-dds.male_hypothalamus <- subsetDESeq(c.colData,  c.countData, "male_hypothalamus" )
-dds.male_pituitary <- subsetDESeq(c.colData,  c.countData, "male_pituitary"  )
-dds.male_gondad <- subsetDESeq(c.colData,  c.countData, "male_gonad")
-```
-
-## plot candidate genes
-
-```{r candidates, eval = T}
-plotcandidates(dds.female_hypothalamus, colDataFH,"female hypothalamus")
-plotcandidates(dds.female_pituitary, colDataFP, "female pituitary")
-plotcandidates(dds.female_gonad, colDataFG,"female gonad")
-plotcandidates(dds.male_hypothalamus, colDataMH,"male hypothalamus")
-plotcandidates(dds.male_pituitary,colDataMP, "male pituitary")
-plotcandidates(dds.male_gondad,colDataMG, "male gonad")
-```
-
-## DEGs
+## DESeq2 DEGs
 
 ```{r createDEGs}
-
 createDEGdf <- function(mydds, whichfactor, up, down, mytissue){
@@
                  independentFiltering = T, alpha = 0.1)
-   data <- data.frame(entrezid = row.names(res),
+   data <- data.frame(Name = row.names(res),
                      padj = res$padj, 
@@
   data <- left_join(data, geneinfo)  %>%
-    mutate(gene = Name) %>%
-    select(gene, lfc, padj, logpadj, direction, tissue) %>%
-    arrange(desc(lfc))
+    select(Name, lfc, padj, logpadj, direction, tissue) %>%
+    arrange(desc(lfc)) %>% 
+    dplyr::rename("gene" = "Name")
 
+  #write dataframe of only significant genes
   DEGs <- data %>% filter(direction != "NS")
   print(str(DEGs))
-
   myfilename = paste("../results/03_DEGs", mytissue, down,up, "csv", sep = ".")
-
-  # write dataframe of only significant genes
   write.csv(DEGs, myfilename, row.names = F)
@@
 
-control.bldg <- createDEGdf(dds.female_pituitary, "treatment", "bldg", "control", "pituitary")
-bldg.lay <- createDEGdf(dds.female_pituitary, "treatment", "lay", "bldg", "pituitary")
-lay.inc.d3 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d3", "lay", "pituitary") 
-inc.d3.inc.d9 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d9", "inc.d3", "pituitary") 
-inc.d9.inc.d17 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
-inc.d17.hatch <- createDEGdf(dds.female_pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
-hatch.n5 <- createDEGdf(dds.female_pituitary, "treatment", "n5", "hatch", "pituitary") 
-n5.n9 <- createDEGdf(dds.female_pituitary, "treatment", "n9", "n5", "pituitary") 
 ```
 
-```{r plotDEGs}
+### differences between consequtive stages
+
+```{r hypdfs, eval=T}
+dds.hypothalamus <- subsetDESeq2(c.colData,  c.countData, c("female_hypothalamus","male_hypothalamus"))
+
+```
+
+```{r hypdfs2, eval=T}
+control.bldg <- createDEGdf(dds.hypothalamus, "treatment", "bldg", "control" ,  "hypothalamus")
+bldg.lay <- createDEGdf(dds.hypothalamus, "treatment", "lay", "bldg" ,  "hypothalamus")
+lay.inc.d3 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d3", "lay" ,  "hypothalamus") 
+inc.d3.inc.d9 <- createDEGdf(dds.hypothalamus,  "treatment", "inc.d9", "inc.d3" ,  "hypothalamus") 
+inc.d9.inc.d17 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d17", "inc.d9" ,  "hypothalamus")
+inc.d17.hatch <- createDEGdf(dds.hypothalamus, "treatment", "hatch", "inc.d17" ,  "hypothalamus") 
+hatch.n5 <- createDEGdf(dds.hypothalamus, "treatment", "n5", "hatch" ,  "hypothalamus") 
+n5.n9 <- createDEGdf(dds.hypothalamus, "treatment", "n9", "n5" ,  "hypothalamus")
+
+```
+
+```{r hypplot, eval = T}
 a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
@@
 
-females <- plot_grid(a + 
+hypothalamus <- plot_grid(a + 
                        theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
-                       labs(y = "female pituitary \n -log10(p)"),
+                       labs(y = "hypothalamus \n -log10(p)"),
           b + theme(axis.title = element_blank(), axis.text = element_blank()),
@@
           nrow = 1,  rel_widths = c(1.75,1,1,1,1,1,1,1,1))
-females 
+hypothalamus
 ```
 
-## Go terms
-
-```{r GOterms}
-reproduction <- read.table("../results/goterms/reproduction.txt", sep = "\t", row.names = NULL,  fill=TRUE)
-reproduction$GO <- "reproduction, GO:0000003"
-
-parentalbehavior <- read.table("../results/goterms/parentalbehavior.txt", sep = "\t", row.names = NULL,  fill=TRUE)
-parentalbehavior$GO <- "parental behavior, GO:0060746"
-
-behavior <- read.table("../results/goterms/parentalbehavior.txt", sep = "\t", row.names = NULL,  fill=TRUE)
-behavior$GO <- "behavior, GO:0007610"
-
-systemdevelopment <- read.table("../results/goterms/systemdevelopment.txt", sep = "\t", row.names = NULL,  fill=TRUE)
-systemdevelopment$GO <- "system development, GO:0048731"
-
-endocrineprocess <- read.table("../results/goterms/endocrineprocess.txt", sep = "\t", row.names = NULL,  fill=TRUE)
-endocrineprocess$GO <- "endocrine process, GO:0050886"
-
-responsetostimulus <- read.table("../results/goterms/responsetostimulus.txt", sep = "\t", row.names = NULL,  fill=TRUE)
-responsetostimulus$GO <- "response to external stimulus, GO:0009605"
-
-GOterms <- rbind(parentalbehavior, reproduction, behavior, 
-                 systemdevelopment, endocrineprocess, responsetostimulus)
-
-GOterms <- GOterms %>%
-  dplyr::mutate(gene = toupper(MGI.Gene.Marker.ID)) %>% 
-  dplyr::select(gene, GO) %>% 
-  dplyr::distinct(gene, GO) %>% 
- group_by(gene)
-
-head(GOterms)
-
-returnGOgenes <- function(whichGOterms, data, mycomparison){
-
-  df <- inner_join(whichGOterms, data) %>% 
-    filter(direction != "NS") %>%  
-    select(gene, GO) %>% 
-    group_by(GO) %>% 
-    summarize(genes = str_c(gene, collapse = ", "))  %>% 
-    mutate(comparison = mycomparison) %>% 
-    select(comparison, GO, genes)
-
-  return(df)
-}
-
-dt1 <- returnGOgenes(GOterms,control.bldg, "control v bldg") 
-dt2 <- returnGOgenes(GOterms,bldg.lay , "bldg v lay") 
-dt3 <- returnGOgenes(GOterms,lay.inc.d3 , "lay v inc d3")  
-dt4 <- returnGOgenes(GOterms,inc.d3.inc.d9 , "inc d3 v inc d9")  
-dt5 <- returnGOgenes(GOterms,inc.d9.inc.d17 , "inc d9 v inc d17")  
-kable(dt5)
-
-dt6 <- returnGOgenes(GOterms,inc.d17.hatch, "inc d17 v hatch")  
-dt7 <- returnGOgenes(GOterms,hatch.n5, "hatch v n5")  
-kable(dt7)
-
-dt8 <- returnGOgenes(GOterms,n5.n9, "n5 v n9")
-
-alldt <- rbind(dt1,dt2,dt3,dt4,dt5,dt6,dt7,dt8)
-alldt
-
-write.csv(alldt, "../results/03_GO_femalepituitary.csv")
-```
-
-## males
-
-```{r createDEGsmales}
-control.bldg2 <- createDEGdf(dds.male_pituitary, "treatment", "bldg", "control", "pituitary")
-bldg.lay2 <- createDEGdf(dds.male_pituitary, "treatment", "lay", "bldg", "pituitary")
-lay.inc.d32 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d3", "lay", "pituitary") 
-inc.d3.inc.d92 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d9", "inc.d3", "pituitary") 
-inc.d9.inc.d172 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
-inc.d17.hatch2 <- createDEGdf(dds.male_pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
-hatch.n52 <- createDEGdf(dds.male_pituitary, "treatment", "n5", "hatch", "pituitary") 
-n5.n92 <- createDEGdf(dds.male_pituitary, "treatment", "n9", "n5", "pituitary")
+```{r pitdfs, eval=T}
+dds.pituitary <- subsetDESeq2(c.colData,  c.countData, c("female_pituitary","male_pituitary"))
 
@@
 
-```{r plotDEGsmales}
-a2 <- plot.volcano(control.bldg2, "treatment", "bldg", "control", colorsvolcano)
-b2 <- plot.volcano(bldg.lay2, "treatment", "lay", "bldg", colorsvolcano)
-c2 <- plot.volcano(lay.inc.d32, "treatment", "inc.d3", "lay", colorsvolcano) 
-d2 <- plot.volcano(inc.d3.inc.d92, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
-e2 <- plot.volcano(inc.d9.inc.d172, "treatment", "inc.d17", "inc.d9", colorsvolcano)
-f2 <- plot.volcano(inc.d17.hatch2, "treatment", "hatch", "inc.d17", colorsvolcano) 
-g2 <- plot.volcano(hatch.n52, "treatment", "n5", "hatch", colorsvolcano) 
-h2 <- plot.volcano(n5.n92, "treatment", "n9", "n5", colorsvolcano)
-
-males <- plot_grid(a2 + labs(y = "male pituitary \n -log10(p)") + theme(axis.text.x = element_blank()),
-          b2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
-          c2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
-          d2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
-          e2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
-          f2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
-          g2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
-          h2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
-          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
-males
+```{r pitdfs2, eval=T}
+control.bldg <- createDEGdf(dds.pituitary, "treatment", "bldg", "control", "pituitary")
+bldg.lay <- createDEGdf(dds.pituitary, "treatment", "lay", "bldg", "pituitary")
+lay.inc.d3 <- createDEGdf(dds.pituitary, "treatment", "inc.d3", "lay", "pituitary") 
+inc.d3.inc.d9 <- createDEGdf(dds.pituitary,  "treatment", "inc.d9", "inc.d3", "pituitary") 
+inc.d9.inc.d17 <- createDEGdf(dds.pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
+inc.d17.hatch <- createDEGdf(dds.pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
+hatch.n5 <- createDEGdf(dds.pituitary, "treatment", "n5", "hatch", "pituitary") 
+n5.n9 <- createDEGdf(dds.pituitary, "treatment", "n9", "n5", "pituitary")
 
@@
 
-### tissue levels, ignoring sex differences, sort of
+```{r pitplot, eval = T}
+a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
+b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
+c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
+d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
+g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
+h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
+
+pituitary <- plot_grid(a + 
+                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
+                       labs(y = "pituitary \n -log10(p)"),
+          b + theme(axis.title = element_blank(), axis.text = element_blank()),
+          c + theme(axis.title = element_blank(), axis.text = element_blank()),
+          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          e + theme(axis.title = element_blank(), axis.text = element_blank()),
+          f + theme(axis.title = element_blank(), axis.text = element_blank()),
+          g + theme(axis.title = element_blank(), axis.text = element_blank()),
+          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          nrow = 1,  rel_widths = c(1.75,1,1,1,1,1,1,1,1))
+pituitary
+
+```
 
 ```{r gonaddfs, eval=T}
+dds.gonads <- subsetDESeq2(c.colData,  c.countData, c("female_gonad","male_gonad"))
 
-dds.gonads <- subsetDESeq2(c.colData,  c.countData, c("female_gonad","male_gonad"))
 ```
@@
 gonads
-
-```
-
-```{r hypdfs, eval=T}
-dds.hypothalamus <- subsetDESeq2(c.colData,  c.countData, c("female_hypothalamus","male_hypothalamus"))
-```
-
-```{r hypdfs2, eval=T}
-control.bldg <- createDEGdf(dds.hypothalamus, "treatment", "bldg", "control" ,  "hypothalamus")
-bldg.lay <- createDEGdf(dds.hypothalamus, "treatment", "lay", "bldg" ,  "hypothalamus")
-lay.inc.d3 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d3", "lay" ,  "hypothalamus") 
-inc.d3.inc.d9 <- createDEGdf(dds.hypothalamus,  "treatment", "inc.d9", "inc.d3" ,  "hypothalamus") 
-inc.d9.inc.d17 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d17", "inc.d9" ,  "hypothalamus")
-inc.d17.hatch <- createDEGdf(dds.hypothalamus, "treatment", "hatch", "inc.d17" ,  "hypothalamus") 
-hatch.n5 <- createDEGdf(dds.hypothalamus, "treatment", "n5", "hatch" ,  "hypothalamus") 
-n5.n9 <- createDEGdf(dds.hypothalamus, "treatment", "n9", "n5" ,  "hypothalamus") 
-```
-
-```{r hypplot, eval = T}
-a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
-b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
-c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
-d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
-e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
-f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
-g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
-h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
-
-hypothalamus <- plot_grid(a + 
-                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
-                       labs(y = "hypothalamus \n -log10(p)"),
-          b + theme(axis.title = element_blank(), axis.text = element_blank()),
-          c + theme(axis.title = element_blank(), axis.text = element_blank()),
-          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          e + theme(axis.title = element_blank(), axis.text = element_blank()),
-          f + theme(axis.title = element_blank(), axis.text = element_blank()),
-          g + theme(axis.title = element_blank(), axis.text = element_blank()),
-          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          nrow = 1,  rel_widths = c(1.75,1,1,1,1,1,1,1,1))
-hypothalamus
-
-```
-
-```{r pitdfs, eval=T}
-dds.pituitary <- subsetDESeq2(c.colData,  c.countData, c("female_pituitary","male_pituitary"))
-```
-
-```{r pitdfs2, eval=T}
-control.bldg <- createDEGdf(dds.pituitary, "treatment", "bldg", "control", "pituitary")
-bldg.lay <- createDEGdf(dds.pituitary, "treatment", "lay", "bldg", "pituitary")
-lay.inc.d3 <- createDEGdf(dds.pituitary, "treatment", "inc.d3", "lay", "pituitary") 
-inc.d3.inc.d9 <- createDEGdf(dds.pituitary,  "treatment", "inc.d9", "inc.d3", "pituitary") 
-inc.d9.inc.d17 <- createDEGdf(dds.pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
-inc.d17.hatch <- createDEGdf(dds.pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
-hatch.n5 <- createDEGdf(dds.pituitary, "treatment", "n5", "hatch", "pituitary") 
-n5.n9 <- createDEGdf(dds.pituitary, "treatment", "n9", "n5", "pituitary") 
-```
-
-```{r pitplot, eval = T}
-a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
-b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
-c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
-d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
-e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
-f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
-g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
-h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
-
-pituitary <- plot_grid(a + 
-                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
-                       labs(y = "pituitary \n -log10(p)"),
-          b + theme(axis.title = element_blank(), axis.text = element_blank()),
-          c + theme(axis.title = element_blank(), axis.text = element_blank()),
-          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          e + theme(axis.title = element_blank(), axis.text = element_blank()),
-          f + theme(axis.title = element_blank(), axis.text = element_blank()),
-          g + theme(axis.title = element_blank(), axis.text = element_blank()),
-          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          nrow = 1,  rel_widths = c(1.75,1,1,1,1,1,1,1,1))
-pituitary
 ```
@@
 icons <- ggdraw() +  draw_image(background, scale = 1)
-
-plot_grid(icons, hypothalamus, females, males, gonads, nrow = 5, rel_heights = c(1,1,1,1,1.25))
-
 plot_grid(icons, hypothalamus, pituitary, gonads, nrow = 4, rel_heights = c(1,1,1,1.25))

### 2019-12-19 analysis/06_DESeq2_ALL.Rmd
deseq all now working but not run yet

@@
 title: "06_DESeqALL"
-output: html_document
+output: md_document
 ---
@@
 levels(colData$treatment)
+
+head(countData)
 ```
@@
 dds.pituitary <- subsetDESeq2(colData,  countData, c("female_pituitary","male_pituitary"))
+
 ```
@@
 write.csv(vsd, "../results/06_pitallvsd.csv")
+
+```
+
+```{r eval=T}
+control.bldg <- createDEGdfsave(dds.pituitary, "treatment", "bldg", "control", "pituitary")
+```
+
+```{r  eval=T}
+bldg.lay <- createDEGdfsave(dds.pituitary, "treatment", "lay", "bldg", "pituitary")
+```
+
+```{r  eval=T}
+lay.inc.d3 <- createDEGdfsave(dds.pituitary, "treatment", "inc.d3", "lay", "pituitary") 
+```
+
+```{r  eval=T}
+inc.d3.inc.d9 <- createDEGdfsave(dds.pituitary,  "treatment", "inc.d9", "inc.d3", "pituitary") 
+```
+
+```{r  eval=T}
+inc.d9.inc.d17 <- createDEGdfsave(dds.pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
+```
+
+```{r  eval=T}
+inc.d17.hatch <- createDEGdfsave(dds.pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
+```
+
+```{r  eval=T}
+hatch.n5 <- createDEGdfsave(dds.pituitary, "treatment", "n5", "hatch", "pituitary") 
+```
+
+```{r  eval=T}
+n5.n9 <- createDEGdfsave(dds.pituitary, "treatment", "n9", "n5", "pituitary") 
+```
+
+```{r eval=T}
+sex <- createDEGdfsave(dds.pituitary, "sex", "male", "female", "pituitary") 
 ```
@@
 
-```{r volcanos-pit}
+```{r  eval=T}
+sex <- createDEGdfsave(dds.pituitary, "sex", "male", "female", "pituitary")
+```
+
+```{r volcanos-pit-char}
+a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
+b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
+c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
+d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
+g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
+h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
+
+pituitary <- plot_grid(a + 
+                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
+                       labs(y = "pituitary \n -log10(p)"),
+          b + theme(axis.title = element_blank(), axis.text = element_blank()),
+          c + theme(axis.title = element_blank(), axis.text = element_blank()),
+          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          e + theme(axis.title = element_blank(), axis.text = element_blank()),
+          f + theme(axis.title = element_blank(), axis.text = element_blank()),
+          g + theme(axis.title = element_blank(), axis.text = element_blank()),
+          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          nrow = 1,  rel_widths = c(1.75,1,1,1,1,1,1,1,1))
+pituitary
+
+plot.volcano(sex, "sex", "male", "female", colorsvolcano) 
+```
+
+```{r volcanos-pit-manip}
 a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano)  
@@
 f <- plot.volcano(hatch.m.inc.d8 , "treatment",    "hatch", "m.inc.d8"  , colorsvolcano)  
-
 g <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  

### 2019-12-19 R/functions.R

@@ -35,7 +35,6 @@ subsetDESeq <- function(colData, countData, eachgroup){
   countData <- countData %>% dplyr::select(one_of(savecols)) 
   
   # assert that row and col lenghts are equal
-  stopifnot(ncol(countData) == nrow(colData))
   
   dds <- DESeqDataSetFromMatrix(countData = countData,
                                 colData = colData,
@@ -833,7 +832,7 @@ createDEGdfsave <- function(mydds, whichfactor, up, down, mytissue){
   
   res <- results(mydds, contrast = c(whichfactor, up, down), independentFiltering = T, alpha = 0.1)
   
-  data <- data.frame(entrezid = row.names(res),
+  data <- data.frame(gene = row.names(res),
                      padj = res$padj, 
                      logpadj = -log10(res$padj),
                      lfc = res$log2FoldChange,
@@ -843,13 +842,10 @@ createDEGdfsave <- function(mydds, whichfactor, up, down, mytissue){
   data <- data %>%
     dplyr::mutate(direction = ifelse(data$lfc > 0 & data$padj < 0.1, 
                                      yes = up, no = ifelse(data$lfc < 0 & data$padj < 0.1, 
-                                                           yes = down, no = "NS")))
-  data$direction <- factor(data$direction, levels = c(down, "NS", up))
-  
-  data <- left_join(data, geneinfo) %>%
-    dplyr::mutate(gene = Name) %>%
-    #dplyr::select(gene, lfc, padj, logpadj, direction, tissue) %>%
-    dplyr::arrange(desc(lfc))
+                                                           yes = down, no = "NS"))) %>% 
+    dplyr::arrange(desc(lfc)) 
+    
+  data$direction <- factor(data$direction, levels = c(down, "NS", up)) 
   
   # write dataframe of only significant genes
   DEGs <- data %>% dplyr::filter(direction != "NS")

### 2019-12-20 analysis/01_limma_all.Rmd
updated but not run

--- 2019-05-16 analysis/01_limma_all.Rmd
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/',cache=TRUE)
-
+knitr::opts_chunk$set(fig.path = '../figures/01_limma/',cache=TRUE)
 ```
@@
 # import "colData" which contains sample information and "countData" which contains read counts
-colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
-countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
+colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
+countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
 head(colData)
-geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+geneinfo <- row.names(countData)
 ```
@@
 write.csv(parentalobject$pseudo.counts, "../results/01_pseudo.counts.csv")
-
-#  perform likelihood ratio test and thresholded testing
-fit <- glmFit( parentalobject, parentaldesign, robust=T)
-tr <- glmTreat(fit, lfc = 1)
-topTags(tr)
-#tr$table
 ```
-
-# plotMDS (multidimential scaling)
-
-```{r plotMDS}
-plotMDS(parentalobject, cex = 0.5)
-```
-
-```{r plotMDS-colored}
-levels(colData$tissue)
-col.tissue <- c("#7570b3","#d95f02", "#1b9e77")[colData$tissue]
-plotMDS(parentalobject,col=col.tissue)
-legend("bottomright",fill=c("#7570b3","#d95f02", "#1b9e77"),legend=levels(colData$tissue))
-title("Colored by Tissue")
-
-levels(colData$treatment)
-col.treatment <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")[colData$treatment]
-plotMDS(parentalobject,col=col.treatment)
-legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
-title("Colored by Treatment")
-
-levels(colData$sex)
-col.sex <- c("#5ab4ac","#d8b365")[colData$sex]
-plotMDS(parentalobject,col=col.sex)
-legend("bottomright",fill=c("#5ab4ac","#d8b365"),legend=levels(colData$sex))
-title("Colored by Sex")
-```
-
-```{r plotMDS-colorshape}
-colData$sexTissue <- paste(colData$sex, colData$tissue, sep = "_")
-
-plotMDS(parentalobject,col=col.treatment, labels = colData$sexTissue)
-legend("bottomright",fill=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6"),legend=levels(colData$treatment))
-title("Colored by Treatment")
-```
-
-# specify contrasts and make MA plots (currently only a subset)
-
-```{r 01-contrasts}
-# view all levels
-levels(colData$group)
-
-# subset of conrasts - sex specific comparing hatch to lay
-my.contrasts <- makeContrasts(
-	         FG_HL = female.gonad.hatch - female.gonad.lay,
-	         FH_HL = female.hypothalamus.hatch - female.hypothalamus.lay,
-	         FP_HL = female.pituitary.hatch - female.pituitary.lay,
-	         MP_HL = male.pituitary.hatch - male.pituitary.lay,
-	         MH_HL = male.hypothalamus.hatch - male.hypothalamus.lay,
-	         MG_HL = male.gonad.hatch - male.gonad.lay,	         
-levels=parentaldesign)
-
-cont <- "FG_HL"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female gonad lay to hatch', frame.plot=F)
-
-cont <- "FH_HL"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female hypothalamus lay to hatch', frame.plot=F)
-
-cont <- "FP_HL"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female pituitary lay to hatch', frame.plot=F)
-
-cont <- "MG_HL"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Male gonad lay to hatch', frame.plot=F)
-
-cont <- "MH_HL"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Male hypothalamus lay to hatch', frame.plot=F)
-
-cont <- "MP_HL"
-summary(decideTestsDGE(
-	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
-	adjust.method="fdr", p.value=0.01))
-kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
-plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Male pituitary lay to hatch', frame.plot=F)
-```

### 2019-12-20 analysis/02_PCA.Rmd
new pca and tsne plots

@@
 
+# note. this file is too big for storage in github :(
+
 ## Data: Pseudocounts from Limma
@@
 head(pseudocounts[1:3])
-```
-
-## Characterization only
-
-```{r characterization}
-colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
-colData$treatment <- factor(colData$treatment, levels = charlevels)
-colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
-colData$tissue <- factor(colData$tissue, levels = tissuelevels)
-
-head(colData[1:3])
-
-geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
-head(geneinfo)
-
-subsetcounts <- function(colData, countData){
-
-  # save counts that match colData
-  savecols <- as.character(colData$V1) 
-  savecols <- as.vector(savecols) 
-  countData <- countData %>% 
-    dplyr::select(X1,savecols) %>% 
-    dplyr::rename(entrezid = X1)
-
-  countData <- left_join(geneinfo, countData)
-  countData <- countData %>% distinct(Name, .keep_all = TRUE) %>%
-    select(-row.names)  %>% select(-geneid) %>% select(-entrezid)
-
-  row.names(countData) <- countData$Name
-  countData$Name <- NULL
-  countData <- as.data.frame(t(countData))
-
-  return(countData)
-}
-
-countData <- subsetcounts(colData, pseudocounts)
+pseudocounts <- as.data.frame(pseudocounts)
+row.names(pseudocounts) <- pseudocounts$X1
+pseudocounts$X1 <- NULL
+# prep count data for all samples
+countData <- as.data.frame(t(pseudocounts))
 head(countData[1:3])
-
-# confirm
-#row.names(countData) == row.names(colData)
-
-euclidist <- dist(countData) # euclidean distances between the rows
-
-# make pca dataframes
-mypca <- prcomp(countData)
-
-# percent conribution
-eig.val <- get_eigenvalue(mypca)
-head(eig.val)
-
-mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
-                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
-                  ID = row.names(countData))
-mypcadf$V1 <- row.names(mypcadf)
-mypcadf <- left_join(colData, mypcadf)
-mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
-head(mypcadf)
-
-a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
-                         color = colData$treatment, 
-                         shape = colData$tissue)) +
-  geom_point()   +
-  mytheme() +
-  theme(legend.title = element_blank(), 
-        axis.text = element_blank()) +
-  scale_alpha_manual(values = c(0.5,1)) +
-  labs(x = "PC1: 35.9% of variance", y = "PC2: 32.5% of variance")
-
-b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
-                   repel = TRUE ,
-                  select.var= list(contrib = 6)) + 
-      mytheme() + 
-      labs(x = " ", y = " ", subtitle = NULL, title =  NULL) +
-    theme( axis.text = element_blank())
-
-plot_grid(a,b, rel_widths = c(2,1))
-
-fviz_screeplot(mypca, addlabels = TRUE, ylim = c(0, 50))
-fviz_contrib(mypca, choice = "var", axes = 1, top = 5) 
-fviz_contrib(mypca, choice = "var", axes = 2, top = 5)
-
-tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=5, theta=0.5, dims=2)
-tsne_df = as.data.frame(tsne_model$Y)
-
-# prep for adding columns
-colData2 <- colData 
-colData2$V1 <- NULL
-tsne_df_cols <- cbind(colData2, tsne_df)
-
-c  <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
-  geom_point(size = 2) +
-  mytheme() +
-  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
-  labs(x = "tSNE 1", y = "tSNE 2")
-
-d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = tissue, color = treatment)) +
-  geom_point(size = 2) +
-  mytheme() +
-  labs(x = "tSNE 1", y = "tSNE 2") +
-  scale_alpha_manual(values = c(0.5,1))
-
-c
-d
-```
-
-## Manipulations only
-
-```{r manipulation}
-head(pseudocounts[1:3])
-
-colData <- read.csv("../metadata/00_colData_manipluation.csv", header = T, row.names = 1)
-
-colData$treatment <- factor(colData$treatment, levels = maniplevels1)
-colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
-colData$tissue <- factor(colData$tissue, levels = tissuelevels)
-
-head(colData[1:3])
-
-countData <- subsetcounts(colData, pseudocounts)
-
-# confirm
-row.names(colData) <- colData$V1
-#row.names(countData) == row.names(colData)
-
-euclidist <- dist(countData) # euclidean distances between the rows
-
-# make pca dataframes
-mypca <- prcomp(countData)
-
-# percent conribution
-eig.val <- get_eigenvalue(mypca)
-head(eig.val)
-
-mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
-                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
-                  ID = row.names(countData))
-mypcadf$V1 <- row.names(mypcadf)
-mypcadf <- left_join(colData, mypcadf)
-mypcadf <- mypcadf %>% dplyr::select(bird,sex,tissue,treatment,PC1:PC6)
-mypcadf$treatment <- factor(mypcadf$treatment, levels = maniplevels1)
-head(mypcadf)
-
-a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
-                         color = colData$treatment, 
-                         shape = colData$tissue)) +
-  geom_point(size = 2)  + 
-  mytheme() +
-  theme(legend.title = element_blank(), 
-        axis.text = element_blank()) +
-  scale_alpha_manual(values = c(0.5,1)) +
-  scale_color_manual(values = colorscharmaip2) +
-  labs(x = "PC1: 63.6% of variance", y = "PC2: 30.1% of variance")
-
-b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
-                   repel = TRUE ,
-                  select.var= list(contrib = 6)) + 
-      mytheme() + 
-      labs(x = " ", y = " ", subtitle = NULL, title =  NULL) +
-  theme( axis.text = element_blank())
-
-plot_grid(a,b, rel_widths = c(2,1))
-
-fviz_screeplot(mypca, addlabels = TRUE, ylim = c(0, 50))
-fviz_contrib(mypca, choice = "var", axes = 1, top = 5) 
-fviz_contrib(mypca, choice = "var", axes = 2, top = 5)
-
-tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=5, theta=0.5, dims=2)
-tsne_df = as.data.frame(tsne_model$Y)
-
-# prep for adding columns
-colData2 <- colData 
-colData2$V1 <- NULL
-tsne_df_cols <- cbind(colData2, tsne_df)
-
-c  <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
-  geom_point(size = 2) +
-  mytheme() +
-  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
-  labs(x = "tSNE 1", y = "tSNE 2")
-
-d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = tissue, color = treatment)) +
-  geom_point(size = 2) +
-  mytheme() +
-  labs(x = "tSNE 1", y = "tSNE 2") +
-  scale_color_manual(values = colorscharmaip2) +
-  scale_alpha_manual(values = c(0.5,1))
-
-c
-d
 ```
@@
 
-```{r charmanip}
+```{r colData}
 # prep col data for all samples
 colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
-colData$treatment <- factor(colData$treatment, levels = alllevels3)
+colData$treatment <- factor(colData$treatment, levels = alllevels)
 colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
@@
 
-# prep count data for all samples
-countData <- as.data.frame(pseudocounts)
-countData$entrezid <- countData$X1
-countData <- left_join(geneinfo, countData)
-countData <- countData %>% distinct(Name, .keep_all = TRUE) %>%
-    select(-row.names, -geneid, -entrezid, -X1) 
-countData <- as.data.frame(countData)
-row.names(countData) <- countData$Name
-countData$Name <- NULL
-countData$entrezid <- NULL
-countData <- as.data.frame(t(countData))
-
 # check ready for analysis
 #row.names(countData) == row.names(colData)
+```
 
+```{r pca}
 # pca
@@
 
-a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
-                         color = colData$treatment, 
-                         #shape = colData$tissue,
-                         #size = colData$sex,
-                        shape = colData$study)) +
-  #geom_hline(yintercept=0, color = "grey") +
-  #geom_vline(xintercept=0, color = "grey") +
-  geom_point()  +
-  mytheme() +
+plotcolorfulpcs <- function(whichfactor, whichcolors){
+  p <- ggplot(mypcadf, aes(x = PC1, y = PC2, color = whichfactor)) +
+    geom_point(size = 1)  +
+    theme_B3() +
     theme(legend.title = element_blank(),
-        axis.text = element_blank()) +
-  #scale_alpha_manual(values = c(0.5,1)) +
-  labs(x = "PC1: 44.6% of variance", y = "PC2: 29.8% of variance") +
-  scale_color_manual(values = colorscharmaip)
+         axis.text = element_blank(),
+         legend.position = "none") +
+    labs(x = "PC1", y = "PC2") +
+    scale_color_manual(values = whichcolors) 
+  print(p)
+}
 
-b <- fviz_pca_var(mypca,  labelsize = 4 , axes.linetype = "blank", 
+a <- plotcolorfulpcs(mypcadf$tissue, colorstissue)  
+b <- plotcolorfulpcs(mypcadf$sex, sexcolors) 
+c <- plotcolorfulpcs(mypcadf$treatment, colorscharmaip)
+
+d <- fviz_screeplot(mypca, addlabels = TRUE, ylim = c(0, 50),  ncp = 5) + labs(title = NULL, y = "Variance") + theme_B3() 
+e <- fviz_contrib(mypca, choice = "var", axes = 1, top = 5)  + labs(title = NULL, subtitle = "PC1", x = NULL) + 
+          theme_B3() + theme(axis.text.x = element_text(angle = 55, hjust = 1))
+f <- fviz_contrib(mypca, choice = "var", axes = 2, top = 5)  + labs(title = NULL, subtitle = "PC2", y = NULL, x = NULL) + 
+          theme_B3()  + theme(axis.text.x = element_text(angle = 55, hjust = 1))
+g <- fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
                    repel = TRUE ,
                   select.var= list(contrib = 6)) + 
-      mytheme() + 
-      labs(x = " ", y = " ", subtitle = NULL, title =  NULL) +
-    #geom_hline(yintercept=0, color = "grey") +
-  #geom_vline(xintercept=0, color = "grey") +
+      theme_B3() + 
+      labs(x = "PC1", y = "PC2", title =  NULL) +
       theme( axis.text = element_blank())
 
-fviz_screeplot(mypca, addlabels = TRUE, ylim = c(0, 50))
-fviz_contrib(mypca, choice = "var", axes = 1, top = 5) 
-fviz_contrib(mypca, choice = "var", axes = 2, top = 5)
+abc <- plot_grid(a,b, c, nrow =  1,  labels = c("d", "e", "f"), label_size = 12)
+defg <- plot_grid(d,e,f,g, nrow = 1, rel_widths = c(1, 0.55, 0.45, 1), labels = c("h", "i", "j", "k" ), label_size = 12)
+dg <- plot_grid(d,g, NULL, nrow = 1, rel_widths = c(1, 1, 1), labels = c("h", "i", NULL ), label_size = 12)
+plot_grid(abc, defg, nrow = 2)
+```
 
-plot_grid(a,b,rel_widths = c(2,1))
-
+```{r tSNE}
 # prep for tsne
@@
 
-ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
-  geom_point(size = 2) +
-  mytheme() +
-  scale_color_manual(values = colorstissue) +
-  labs(x = "tSNE 1", y = "tSNE 2")
+plotcolorfultsnes <- function(whichfactor, whichcolors){
+  p <- ggplot(tsne_df_cols, aes(x = V1, y = V2, color = whichfactor)) +
+    geom_point(size = 1) +
+    theme_B3() +
+    labs(x = "tSNE 1", y = "tSNE 2") +
+    scale_color_manual(values = whichcolors) +
+    theme(legend.position = "none",
+          axis.text = element_blank())
+  print(p)
+}
 
-ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = study, color = treatment)) +
-  geom_point(size = 2) +
-  mytheme() +
-  labs(x = "tSNE 1", y = "tSNE 2") +
-  scale_alpha_manual(values = c(0.5,1)) +
-  scale_color_manual(values = colorscharmaip2)
+h <- plotcolorfultsnes(tsne_df_cols$tissue, colorstissue)  
+i <- plotcolorfultsnes(tsne_df_cols$sex, sexcolors)   
+j <- plotcolorfultsnes(tsne_df_cols$treatment, colorscharmaip)
+
+hij <- plot_grid(h,i,j, nrow = 1, labels = c("a", "b", "c"), label_size = 12)
+hij
 ```
+```{r PCA-tSNE}
+plot_grid(hij, abc, dg, nrow = 3)
+```

### 2019-12-20 analysis/02_PCA.Rmd
better pca tsne figure

@@
 
-```{r pca}
+```{r pcadf}
 # pca
@@
 head(mypcadf)
+```
 
+```{r pca}
 plotcolorfulpcs <- function(whichfactor, whichcolors){
@@
 
-d <- fviz_screeplot(mypca, addlabels = TRUE, ylim = c(0, 50),  ncp = 5) + labs(title = NULL, y = "Variance") + theme_B3() 
+d <- fviz_screeplot(mypca, addlabels = TRUE, ylim = c(0, 50),  ncp = 5, barcolor = "white", barfill = "white") + 
+  labs(title = NULL, y = "PC Variance") + theme_B3() 
 e <- fviz_contrib(mypca, choice = "var", axes = 1, top = 5)  + labs(title = NULL, subtitle = "PC1", x = NULL) + 
@@
 
+legend <- png::readPNG("../figures/images/DoveParentsRNAseq_legend.png")
+legend <-  grid::rasterGrob(legend, interpolate=TRUE)
+
 abc <- plot_grid(a,b, c, nrow =  1,  labels = c("d", "e", "f"), label_size = 12)
 defg <- plot_grid(d,e,f,g, nrow = 1, rel_widths = c(1, 0.55, 0.45, 1), labels = c("h", "i", "j", "k" ), label_size = 12)
-dg <- plot_grid(d,g, NULL, nrow = 1, rel_widths = c(1, 1, 1), labels = c("h", "i", NULL ), label_size = 12)
-plot_grid(abc, defg, nrow = 2)
+dg <- plot_grid(d,g, legend, nrow = 1, rel_widths = c(1, 1, 1), labels = c("h", "i", NULL ), label_size = 12)
+plot_grid(abc, dg, nrow = 2)
 ```
 
-```{r tSNE}
+```{r tSNE-df}
 # prep for tsne
@@
 
+```
+
+```{r tSNE}
 plotcolorfultsnes <- function(whichfactor, whichcolors){
@@
 hij
+
 ```
-```{r PCA-tSNE}
+```{r PCA-tSNE, fig.height= 7, fig.width=7}
 plot_grid(hij, abc, dg, nrow = 3)
+
 ```

### 2019-12-20 analysis/06_DESeq2_ALL.Rmd
now with new pit volcanos

--- 2019-12-19 analysis/06_DESeq2_ALL.Rmd
@@
 
-```{r volcanos-pit}
-a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano)  
-b <- plot.volcano(inc.d9.m.inc.d9, "treatment",  "m.inc.d9", "inc.d9" , colorsvolcano) 
-c <- plot.volcano(inc.d17.m.inc.d17 , "treatment",  "m.inc.d17", "inc.d17" , colorsvolcano) 
-d <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano)
+```{r volcanos-pituitary, fig.height= 6, fig.width=6}
+a <- plot.volcano(sex, "sex", "male", "female", colorsvolcano)
+b <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
+c <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
+d <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
+e <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+f <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+g <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
+h <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
+i <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
 
-e <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano) 
-f <- plot.volcano(hatch.m.inc.d8 , "treatment",    "hatch", "m.inc.d8"  , colorsvolcano)
+j <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano)  
+k <- plot.volcano(inc.d9.m.inc.d9, "treatment",  "m.inc.d9", "inc.d9" , colorsvolcano) 
+l <- plot.volcano(inc.d17.m.inc.d17 , "treatment",  "m.inc.d17", "inc.d17" , colorsvolcano) 
+m <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano)
 
-g <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  
-h <- plot.volcano(hatch.prolong , "treatment",  "hatch", "prolong", colorsvolcano)
+n <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano) 
+o <- plot.volcano(hatch.m.inc.d8 , "treatment",    "hatch", "m.inc.d8"  , colorsvolcano)  
+p <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  
+q <- plot.volcano(hatch.prolong , "treatment",  "hatch", "prolong", colorsvolcano) 
+r <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano)  
+s <- plot.volcano(n5.extend , "treatment",  "extend", "n5", colorsvolcano)
 
-i <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano)  
-j <- plot.volcano(n5.extend , "treatment",  "extend", "n5", colorsvolcano)
+char <- plot_grid(
+          b + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
+          c + theme(axis.title = element_blank(), axis.text = element_blank()) ,
+          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          f + theme(axis.title = element_blank(), axis.text = element_blank()),
+          g + theme(axis.title = element_blank(), axis.text = element_blank()),
+          h + theme(axis.title = element_blank(), axis.text = element_blank()),
+          i + theme(axis.title = element_blank(), axis.text = element_blank()),
+          nrow = 1, rel_widths = c(1.1,1,1,1,1,1,1))
 
-remove <- plot_grid(a + labs(y = "pituitary -log10(p-value)"),
-          b + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          c + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
-          d + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
-          nrow = 1, rel_widths = c(1.2,1,1,1),
-          labels = c("a","b","c","d"), label_size = 9)
-remove
+remove <- plot_grid( 
+          a + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
+          NULL, 
+          j + theme(axis.title = element_blank(), axis.text = element_blank()) , 
+          k + theme(axis.title = element_blank(), axis.text = element_blank()) , 
+          l + theme(axis.title = element_blank(), axis.text = element_blank()) , 
+          m + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          NULL, nrow = 1, rel_widths = c(1.1, 0.5, 1,1,1,1,0.5))
 
-timing <- plot_grid(e +  labs(y = "pituitary -log10(p-value)"),
-          f + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          g + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          h + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
-          j + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
-          nrow = 1, rel_widths = c(1.2,1,1,1,1,1),
-          labels = c("a","b","c", "d","e","f"), label_size = 9)
-timing
+timing <- plot_grid( 
+          n , 
+          o + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
+          NULL,
+          p + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          q + theme(axis.title.y = element_blank(), axis.text.y = element_blank())+
+            labs(x = "Log fold change, pituitary") , 
+           NULL,
+          r + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
+          s + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+            nrow = 1, rel_widths = c(1.1,1, 0.5, 1,1,0.5, 1,1))
+
+charicons <- png::readPNG("../figures/images/DoveParentsRNAseq_volcanos-char.png")
+charicons <-  grid::rasterGrob(charicons, interpolate=TRUE)
+
+removeicons <- png::readPNG("../figures/images/DoveParentsRNAseq_volcanos-removal.png")
+removeicons <-  grid::rasterGrob(removeicons, interpolate=TRUE)
+
+timingicons <- png::readPNG("../figures/images/DoveParentsRNAseq_volcanos-timing.png")
+timingicons <-  grid::rasterGrob(timingicons, interpolate=TRUE)
+
+plot_grid(charicons, char, removeicons, remove, timingicons, timing, nrow = 6, rel_heights = c(0.9,1,0.9,1,0.9,1.2))
 ```

### 2019-12-20 analysis/06_DESeq2_ALL.Rmd
patial new hyp

@@
           i + theme(axis.title = element_blank(), axis.text = element_blank()),
-          nrow = 1, rel_widths = c(1.1,1,1,1,1,1,1))
+          nrow = 1, rel_widths = c(1.2,1,1,1,1,1,1))
 
@@
           m + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          NULL, nrow = 1, rel_widths = c(1.1, 0.5, 1,1,1,1,0.5))
+          NULL, nrow = 1, rel_widths = c(1.2, 1, 1,1,1,1,1))
 
@@
           s + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-            nrow = 1, rel_widths = c(1.1,1, 0.5, 1,1,0.5, 1,1))
+            nrow = 1, rel_widths = c(1.2,1, 0.5, 1,1,0.5, 1,1))
 

### 2019-12-20 analysis/06c_DESeq2_ALL_hypothalamus.Rmd
better gonad and hypothalamus figures
patial new hyp

@@
 ```{r eval=T}
+control.bldg <- createDEGdfsave(dds.hypothalamus, "treatment", "bldg", "control", "hypothalamus")
+```
+
+```{r  eval=T}
+bldg.lay <- createDEGdfsave(dds.hypothalamus, "treatment", "lay", "bldg", "hypothalamus")
+```
+
+```{r  eval=T}
+lay.inc.d3 <- createDEGdfsave(dds.hypothalamus, "treatment", "inc.d3", "lay", "hypothalamus") 
+```
+
+```{r  eval=T}
+inc.d3.inc.d9 <- createDEGdfsave(dds.hypothalamus,  "treatment", "inc.d9", "inc.d3", "hypothalamus") 
+```
+
+```{r  eval=T}
+inc.d9.inc.d17 <- createDEGdfsave(dds.hypothalamus, "treatment", "inc.d17", "inc.d9", "hypothalamus")
+```
+
+```{r  eval=T}
+inc.d17.hatch <- createDEGdfsave(dds.hypothalamus, "treatment", "hatch", "inc.d17", "hypothalamus") 
+```
+
+```{r  eval=T}
+hatch.n5 <- createDEGdfsave(dds.hypothalamus, "treatment", "n5", "hatch", "hypothalamus") 
+```
+
+```{r  eval=T}
+n5.n9 <- createDEGdfsave(dds.hypothalamus, "treatment", "n9", "n5", "hypothalamus") 
+```
+
+```{r  eval=T}
 inc.d3.m.inc.d3 <- createDEGdfsave(dds.hypothalamus, "treatment", "m.inc.d3", "inc.d3", "hypothalamus")
@@
 ```{r  eval=T}
-n5.extend <- createDEGdfsave(dds.hypothalamus, "treatment", "extend", "n5", "hypothalamus")
+n5.extend <- createDEGdfsave(dds.hypothalamus, "treatment", "n5", "extend", "hypothalamus")
 ```
 
-```{r volcanos-hyp, fig.width=7, fig.height=3.5}
-a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano)  
-b <- plot.volcano(inc.d9.m.inc.d9, "treatment",  "m.inc.d9", "inc.d9" , colorsvolcano) 
-c <- plot.volcano(inc.d17.m.inc.d17 , "treatment",  "m.inc.d17", "inc.d17" , colorsvolcano) 
-d <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano)
+```{r eval=T}
+sex <- createDEGdfsave(dds.hypothalamus, "sex", "male", "female", "hypothalamus") 
+```
 
-e <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano) 
-f <- plot.volcano(hatch.m.inc.d8 , "treatment",    "hatch", "m.inc.d8"  , colorsvolcano)
+```{r volcanos-hypothalamus, fig.height= 6, fig.width=6}
+a <- plot.volcano(sex, "sex", "male", "female", colorsvolcano)
+b <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
+c <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
+d <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
+e <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+f <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+g <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
+h <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
+i <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
 
-g <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  
-h <- plot.volcano(hatch.prolong , "treatment",  "hatch", "prolong", colorsvolcano)
+j <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano)  
+k <- plot.volcano(inc.d9.m.inc.d9, "treatment",  "m.inc.d9", "inc.d9" , colorsvolcano) 
+l <- plot.volcano(inc.d17.m.inc.d17 , "treatment",  "m.inc.d17", "inc.d17" , colorsvolcano) 
+m <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano)
 
-i <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano)  
-j <- plot.volcano(n5.extend , "treatment",  "extend", "n5", colorsvolcano)
+n <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano) 
+o <- plot.volcano(hatch.m.inc.d8 , "treatment",    "hatch", "m.inc.d8"  , colorsvolcano)  
+p <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  
+q <- plot.volcano(hatch.prolong , "treatment",  "hatch", "prolong", colorsvolcano) 
+r <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano)  
+s <- plot.volcano(n5.extend , "treatment",  "extend", "n5", colorsvolcano)
 
-remove <- plot_grid(a + labs(y = "hypothalamus -log10(p-value)"),
-          b + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          c + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
-          d + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
-          nrow = 1, rel_widths = c(1.2,1,1,1),
-          labels = c("a","b","c","d"), label_size = 9)
-remove
+char <- plot_grid(
+          b + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
+          c + theme(axis.title = element_blank(), axis.text = element_blank()) ,
+          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          f + theme(axis.title = element_blank(), axis.text = element_blank()),
+          g + theme(axis.title = element_blank(), axis.text = element_blank()),
+          h + theme(axis.title = element_blank(), axis.text = element_blank()),
+          i + theme(axis.title = element_blank(), axis.text = element_blank()),
+          nrow = 1, rel_widths = c(1.2,1,1,1,1,1,1))
 
-timing <- plot_grid(e +  labs(y = "hypothalamus -log10(p-value)"),
-          f + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          g + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          h + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
-          j + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
-          nrow = 1, rel_widths = c(1.2,1,1,1,1,1),
-          labels = c("a","b","c", "d","e","f"), label_size = 9)
-timing
+remove <- plot_grid( 
+          a + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
+          NULL, 
+          j + theme(axis.title = element_blank(), axis.text = element_blank()) , 
+          k + theme(axis.title = element_blank(), axis.text = element_blank()) , 
+          l + theme(axis.title = element_blank(), axis.text = element_blank()) , 
+          m + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          NULL, nrow = 1, rel_widths = c(1.2, 1, 1,1,1,1,1))
+
+timing <- plot_grid( 
+          n , 
+          o + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
+          NULL,
+          p + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          q + theme(axis.title.y = element_blank(), axis.text.y = element_blank())+
+            labs(x = "Log fold change, pituitary") , 
+           NULL,
+          r + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
+          s + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+            nrow = 1, rel_widths = c(1.2,1, 0.5, 1,1,0.5, 1,1))
+
+charicons <- png::readPNG("../figures/images/DoveParentsRNAseq_volcanos-char.png")
+charicons <-  grid::rasterGrob(charicons, interpolate=TRUE)
+
+removeicons <- png::readPNG("../figures/images/DoveParentsRNAseq_volcanos-removal.png")
+removeicons <-  grid::rasterGrob(removeicons, interpolate=TRUE)
+
+timingicons <- png::readPNG("../figures/images/DoveParentsRNAseq_volcanos-timing.png")
+timingicons <-  grid::rasterGrob(timingicons, interpolate=TRUE)
+
+plot_grid(charicons, char, removeicons, remove, timingicons, timing, nrow = 6, rel_heights = c(0.9,1,0.9,1,0.9,1.2))
 ```

### 2019-12-20 analysis/20_forrechelle.Rmd
working on candidate gnee plots

@@
 
-knitr::opts_chunk$set(echo = TRUE)
+source("../R/themes.R")
+
+knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/favegenes/")
 ```
@@
 
-candidategenes <- c( "GH", "NR3C2", "PRLR", "PRL")
+candidategenes <- c( "NR3C1", "NR3C2", "PRLR", "PRL")
 timepoints <- c("lay" , "inc.d9", "hatch", "n9")
@@
 
-```{r readfile}
-
-geneids <- read_csv("..//metadata/00_geneinfo.csv")
+```{r rechelle}
 
 datapath <- "../results/"   # path to the data
-datafiles <- dir(datapath, pattern = "04_vsd*") # get file names
+datafiles <- dir(datapath, pattern = "*allvsd.csv") # get file names
 datapathfile <- paste0(datapath, datafiles)
@@
   map_df(~read_csv(.x, col_types = cols(), col_names = T), .id = "filename") %>% 
-  mutate(tissue = sapply(strsplit(as.character(filename),'../results/04_vsd_'), "[", 2)) %>% 
-  mutate(tissue = sapply(strsplit(as.character(tissue),'.csv'), "[", 1))  %>% 
-  left_join(geneids) %>%
-  select(tissue, X1, Name, everything()) %>% select(-filename, -geneid, - entrezid, - row.names) 
+  mutate(tissue = sapply(strsplit(as.character(filename),'../results/06_'), "[", 2)) %>% 
+  mutate(tissue = sapply(strsplit(as.character(tissue),'allvsd.csv'), "[", 1))  %>% 
+  select(tissue, X1, everything()) 
+head(df)
+
+df <- df %>% filter(tissue != "gon")
 head(df)
@@
 df2 <-  df  %>%
-  filter(Name %in% candidategenes) %>%
+  filter(X1 %in% candidategenes) %>%
   pivot_longer(cols = L.Blu13_male_gonad_control.NYNO:y98.o50.x_male_pituitary_inc.d3, 
@@
   filter(treatment %in% timepoints) %>%
-  select(bird, sex, treatment, tissue, Name, vsd) %>%
-  rename(gene = Name)
+  select(bird, sex, treatment, tissue, X1, vsd) %>%
+  rename(gene = X1) %>%
+  drop_na()
 head(df2)
@@
 }
+
+for (i in levels(df3$tissue)) {
+
+  p <-  df3 %>%
+    filter(tissue == i) %>%
+    ggplot(aes(x = treatment, y = m)) +
+    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
+    geom_point(size = 1, aes(color = gene)) +
+    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
+    scale_alpha_manual(values = c(0.5, 1)) +
+    labs(subtitle = i, y = "average expression", x = "parental stage") +
+    facet_wrap(~sex, nrow = 1) +
+    theme_B3()
+  plot(p)
+
+}
+
 ```

### 2019-12-20 analysis/06_DESeq2_ALL.Rmd
now with some corerlations

@@
           i + theme(axis.title = element_blank(), axis.text = element_blank()),
-          nrow = 1, rel_widths = c(1.2,1,1,1,1,1,1))
+          nrow = 1, rel_widths = c(1.3,1,1,1,1,1,1))
 
@@
           m + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          NULL, nrow = 1, rel_widths = c(1.2, 1, 1,1,1,1,1))
+          NULL, nrow = 1, rel_widths = c(1.3, 1, 1,1,1,1,1))
 
@@
           n , 
-          o + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
+          o + theme(axis.title.y = element_blank(), axis.text.y = element_blank())+
+            labs(x = "Log fold change,")   , 
           NULL,
-          p + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
-          q + theme(axis.title.y = element_blank(), axis.text.y = element_blank())+
-            labs(x = "Log fold change, pituitary") , 
+          p + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
+            labs(x = "pituitary") , 
+          q + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
            NULL,
@@
           s + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-            nrow = 1, rel_widths = c(1.2,1, 0.5, 1,1,0.5, 1,1))
+            nrow = 1, rel_widths = c(1.3,1, 0.5, 1, 1, 0.5, 1, 1))
 

### 2019-12-20 analysis/20_forrechelle.Rmd
now with some corerlations

@@
 library(cowplot)
+library(forcats)
+library(corrr)
 
 source("../R/themes.R")
+source("../R/functions.R")
 
@@
 
-candidategenes <- c( "NR3C1", "NR3C2", "PRLR", "PRL")
+candidategenes <- c( "NR3C1", "NR3C2", "PRLR", "PRL", "POMC" ,"MC4R", "VIP", "BRCA1")
 timepoints <- c("lay" , "inc.d9", "hatch", "n9")
@@
 
-df <- df %>% filter(tissue != "gon")
-head(df)
+genes <- as.data.frame(df$X1)
 
@@
   select(bird, sex, treatment, tissue, X1, vsd) %>%
+  mutate(tissue = fct_recode(tissue, "hypothalamus" = "hyp",
+                    "pituitary" = "pit",
+                    "gonads" = "gon"
+                    )) %>%
   rename(gene = X1) %>%
-  drop_na()
+  drop_na() %>%
+  droplevels()
 head(df2)
 
+df2$treatment <- factor(df2$treatment, levels = alllevels)
+
 df3 <- df2 %>% 
-  mutate(tissue = fct_relevel(tissue,  "hyp", "pit", "gon")) %>% 
   mutate(treatment = fct_relevel(treatment, timepoints)) %>% 
@@
 head(df3)
-
-for (i in levels(df3$tissue)) {
-
-p <-  df3 %>%
-  filter(tissue == i) %>%
-  ggplot(aes(x = treatment, y = m)) +
-  geom_image(aes(image=image, color = gene), size=0.1)  +
-
-  scale_alpha_manual(values = c(0.5, 1)) +
-  labs(subtitle = i, y = "average expression", x = "parental stage")
-
-pimage <- axis_canvas(p, axis = 'x') + 
-  draw_image("../figures/images/icons/DoveParentsRNAseq_lay.png", x = 0.5, scale = 1 ) +
-  draw_image("../figures/images/icons/DoveParentsRNAseq_incubation.png", x = 1.5, scale = 1) +
-  draw_image("../figures/images/icons/DoveParentsRNAseq_hatch.png", x = 2.5, scale = 1) +
-  draw_image("../figures/images/icons/DoveParentsRNAseq_chicklings.png", x = 3.5, scale = 1)
-
-p2 <- ggdraw(insert_xaxis_grob(p, pimage, position = "bottom"))
-plot(p2)
-
-}
 
@@
     theme_B3()
-  plot(p)
-
+ print(p)
 }
 
+for (i in levels(df3$tissue)) {
+
+  p <-  df2 %>%
+    filter(tissue == i) %>%
+    ggplot(aes(x = treatment, y = vsd, fill = gene, color = sex)) +
+    geom_boxplot() +
+    scale_alpha_manual(values = c(0.5, 1)) +
+    labs(subtitle = i, y = "expression", x = "parental stage") +
+    facet_wrap(~gene, scales = "free_y") +
+    theme_B3() +
+    scale_color_manual(values = sexcolors)
+ print(p)
+}
+
+dft <- df %>% filter(X1 %in% candidategenes) %>%
+  unite("tissuegene", tissue:X1, remove = FALSE) %>%
+  select(-tissue, -`X1`, -filename)  %>%
+  pivot_longer(-tissuegene, names_to = "samples", values_to = "vsd") %>%
+  drop_na() %>%
+  mutate("bird" = sapply(strsplit(as.character(samples),'\\_'), "[", 1)) %>%
+  select(bird,tissuegene,vsd) %>%
+  pivot_wider(id_cols = bird, names_from = tissuegene, values_from = vsd)
+
+dft <- as.data.frame(dft)
+row.names(dft) <- dft$bird
+dft$bird <- NULL
+
+x <- correlate(dft)
+
+x %>% network_plot.cor_df(min_cor = .2)
+
+x %>%
+  focus(pit_PRL) %>%
+  mutate(rowname = reorder(rowname, pit_PRL)) %>%
+  ggplot(aes(rowname, pit_PRL)) +
+    geom_col() + coord_flip() +
+  labs(y = "corrleation with pituitiary PRL", x = "candidate genes by tissue") +
+  theme_B3() + 
+  theme(legend.position = "none")
 ```

### 2019-12-20 analysis/06c_DESeq2_ALL_hypothalamus.Rmd
now with some corerlations

@@
           n , 
-          o + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
+          o + theme(axis.title.y = element_blank(), axis.text.y = element_blank())+
+            labs(x = "Log fold change,")   , 
           NULL,
-          p + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
-          q + theme(axis.title.y = element_blank(), axis.text.y = element_blank())+
-            labs(x = "Log fold change, pituitary") , 
+          p + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
+            labs(x = "hypothalamus") , 
+          q + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
            NULL,
@@
           s + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-            nrow = 1, rel_widths = c(1.2,1, 0.5, 1,1,0.5, 1,1))
+            nrow = 1, rel_widths = c(1.3,1, 0.5, 1,1,0.5, 1,1))
 

### 2019-12-20 analysis/00_datawrangling.Rmd
specifitity, isoforms, and gon vsd

@@
   summarize(n = n()) %>%
-  arrange(desc(n), Name) %>%
-  filter(n > 1)
+  filter(n > 1) %>%
+  group_by(n) %>% 
+  summarize(genes = str_c(Name, collapse = ", ")) %>%
+  arrange(desc(n)) %>%
+  rename("n(isoforms)" = "n") %>% 
+  mutate("n(counts)" = str_count(genes, ",") + 1)
 head(isoforms)

### 2019-12-20 analysis/06b_DESeq2_ALL_gonad.Rmd
specifitity, isoforms, and gon vsd

@@
 ```{r eval=T}
+control.bldg <- createDEGdfsave(dds.gonads, "treatment", "bldg", "control", "gonads")
+```
+
+```{r  eval=T}
+bldg.lay <- createDEGdfsave(dds.gonads, "treatment", "lay", "bldg", "gonads")
+```
+
+```{r  eval=T}
+lay.inc.d3 <- createDEGdfsave(dds.gonads, "treatment", "inc.d3", "lay", "gonads") 
+```
+
+```{r  eval=T}
+inc.d3.inc.d9 <- createDEGdfsave(dds.gonads,  "treatment", "inc.d9", "inc.d3", "gonads") 
+```
+
+```{r  eval=T}
+inc.d9.inc.d17 <- createDEGdfsave(dds.gonads, "treatment", "inc.d17", "inc.d9", "gonads")
+```
+
+```{r  eval=T}
+inc.d17.hatch <- createDEGdfsave(dds.gonads, "treatment", "hatch", "inc.d17", "gonads") 
+```
+
+```{r  eval=T}
+hatch.n5 <- createDEGdfsave(dds.gonads, "treatment", "n5", "hatch", "gonads") 
+```
+
+```{r  eval=T}
+n5.n9 <- createDEGdfsave(dds.gonads, "treatment", "n9", "n5", "gonads") 
+```
+
+```{r eval=T}
+sex <- createDEGdfsave(dds.gonads, "sex", "male", "female", "gonads") 
+```
+
+```{r  eval=T}
 inc.d3.m.inc.d3 <- createDEGdfsave(dds.gonads, "treatment", "m.inc.d3", "inc.d3", "gonads")
@@
 
-```{r volcanos-gon, fig.width=7, fig.height=3.5}
-a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano)  
-b <- plot.volcano(inc.d9.m.inc.d9, "treatment",  "m.inc.d9", "inc.d9" , colorsvolcano) 
-c <- plot.volcano(inc.d17.m.inc.d17 , "treatment",  "m.inc.d17", "inc.d17" , colorsvolcano) 
-d <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano)
+```{r volcanos-gonads, fig.height= 6, fig.width=6}
+a <- plot.volcano(sex, "sex", "male", "female", colorsvolcano)
+b <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
+c <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
+d <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
+e <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
+f <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
+g <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
+h <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
+i <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)
 
-e <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano) 
-f <- plot.volcano(hatch.m.inc.d8 , "treatment",    "hatch", "m.inc.d8"  , colorsvolcano)
+j <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano)  
+k <- plot.volcano(inc.d9.m.inc.d9, "treatment",  "m.inc.d9", "inc.d9" , colorsvolcano) 
+l <- plot.volcano(inc.d17.m.inc.d17 , "treatment",  "m.inc.d17", "inc.d17" , colorsvolcano) 
+m <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano)
 
-g <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  
-h <- plot.volcano(hatch.prolong , "treatment",  "hatch", "prolong", colorsvolcano)
+n <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano) 
+o <- plot.volcano(hatch.m.inc.d8 , "treatment",    "hatch", "m.inc.d8"  , colorsvolcano)  
+p <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  
+q <- plot.volcano(hatch.prolong , "treatment",  "hatch", "prolong", colorsvolcano) 
+r <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano)  
+s <- plot.volcano(n5.extend , "treatment",  "extend", "n5", colorsvolcano)
 
-i <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano)  
-j <- plot.volcano(n5.extend , "treatment",  "extend", "n5", colorsvolcano)
+char <- plot_grid(
+          b + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
+          c + theme(axis.title = element_blank(), axis.text = element_blank()) ,
+          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          f + theme(axis.title = element_blank(), axis.text = element_blank()),
+          g + theme(axis.title = element_blank(), axis.text = element_blank()),
+          h + theme(axis.title = element_blank(), axis.text = element_blank()),
+          i + theme(axis.title = element_blank(), axis.text = element_blank()),
+          nrow = 1, rel_widths = c(1.3,1,1,1,1,1,1))
 
-remove <- plot_grid(a + labs(y = "gonads -log10(p-value)"),
-          b + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          c + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
-          d + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
-          nrow = 1, rel_widths = c(1.2,1,1,1),
-          labels = c("a","b","c","d"), label_size = 9)
-remove
+remove <- plot_grid( 
+          a + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
+          NULL, 
+          j + theme(axis.title = element_blank(), axis.text = element_blank()) , 
+          k + theme(axis.title = element_blank(), axis.text = element_blank()) , 
+          l + theme(axis.title = element_blank(), axis.text = element_blank()) , 
+          m + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          NULL, nrow = 1, rel_widths = c(1.3, 1, 1,1,1,1,1))
 
-timing <- plot_grid(e +  labs(y = "gonads -log10(p-value)"),
-          f + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          g + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          h + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-          i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
-          j + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
-          nrow = 1, rel_widths = c(1.2,1,1,1,1,1),
-          labels = c("a","b","c", "d","e","f"), label_size = 9)
-timing
+timing <- plot_grid( 
+          n , 
+          o + theme(axis.title.y = element_blank(), axis.text.y = element_blank())+
+            labs(x = "Log fold change,")   , 
+          NULL,
+          p + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
+            labs(x = "gonads") , 
+          q + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
+           NULL,
+          r + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
+          s + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+            nrow = 1, rel_widths = c(1.3,1, 0.5, 1, 1, 0.5, 1, 1))
+
+charicons <- png::readPNG("../figures/images/DoveParentsRNAseq_volcanos-char.png")
+charicons <-  grid::rasterGrob(charicons, interpolate=TRUE)
+
+removeicons <- png::readPNG("../figures/images/DoveParentsRNAseq_volcanos-removal.png")
+removeicons <-  grid::rasterGrob(removeicons, interpolate=TRUE)
+
+timingicons <- png::readPNG("../figures/images/DoveParentsRNAseq_volcanos-timing.png")
+timingicons <-  grid::rasterGrob(timingicons, interpolate=TRUE)
+
+plot_grid(charicons, char, removeicons, remove, timingicons, timing, nrow = 6, rel_heights = c(0.9,1,0.9,1,0.9,1.2))
 ```

### 2019-12-20 analysis/20_forrechelle.Rmd
specifitity, isoforms, and gon vsd

@@
 
-candidategenes <- c( "NR3C1", "NR3C2", "PRLR", "PRL", "POMC" ,"MC4R", "VIP", "BRCA1")
-timepoints <- c("lay" , "inc.d9", "hatch", "n9")
+candidategenes <- c(  "PRLR", "PRL" )
+timepoints <- charlevels
 hormones <- c("PRL", "CORT")
@@
 
+# check which genese appear in multiple tissue
+tissuespecificity <- df %>%
+  group_by(X1) %>%
+  summarize(tissues = str_c(tissue, collapse = ", ")) %>%
+  filter(tissues != "gon, hyp, pit") %>%
+  group_by(tissues) %>%
+  summarize(genes = str_c(X1, collapse = ", ")) %>% 
+  mutate(n = str_count(genes, ",") + 1) %>% 
+  arrange(desc(n))
+tissuespecificity
+write.csv(tissuespecificity, "../results/tissuespecificity.csv")
+```
+
+```{r}
 genes <- as.data.frame(df$X1)
@@
 dft$bird <- NULL
+head(dft)
 

### 2019-12-20 analysis/20_forrechelle.Rmd
new figs

--- 2019-12-20 analysis/20_forrechelle.Rmd
@@
 
-candidategenes <- c( "NR3C1", "NR3C2", "PRLR", "PRL", "POMC" ,"MC4R", "VIP", "BRCA1")
-timepoints <- c("lay" , "inc.d9", "hatch", "n9")
-hormones <- c("PRL", "CORT")
+candidategenes <- c("PRLR", "PRL", "CISH", "JAK2", "SOCS3", "GH" )
+timepoints <- charlevels
 
@@
 
-```{r rechelle}
-
+```{r data}
 datapath <- "../results/"   # path to the data
@@
   select(tissue, X1, everything()) 
-head(df)
+```
+
+```{r forrechelle}
+# check which genese appear in multiple tissue
+tissuespecificity <- df %>%
+  group_by(X1) %>%
+  summarize(tissues = str_c(tissue, collapse = ", ")) %>%
+  filter(tissues != "gon, hyp, pit") %>%
+  group_by(tissues) %>%
+  summarize(genes = str_c(X1, collapse = ", ")) %>% 
+  mutate(n = str_count(genes, ",") + 1) %>% 
+  arrange(desc(n))
+tissuespecificity
+write.csv(tissuespecificity, "../results/tissuespecificity.csv")
 
@@
 df2$treatment <- factor(df2$treatment, levels = alllevels)
+df2$tissue <- factor(df2$tissue, levels = tissuelevels)
 
@@
 
-for (i in levels(df3$tissue)) {
+df3 %>%
+    ggplot(aes(x = treatment, y = m)) +
+    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
+    geom_point(size = 1, aes(color = gene)) +
+    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
+    scale_alpha_manual(values = c(0.5, 1)) +
+    labs( y = "average expression", x = "parental stage") +
+    facet_grid(sex ~ tissue, scales = "free_y") +
+    theme_B3() +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
+df2$gene <- factor(df2$gene)
+
+for (i in levels(df2$gene)) {
   p <-  df2 %>%
-    filter(tissue == i) %>%
-    ggplot(aes(x = treatment, y = vsd, fill = gene, color = sex)) +
+    filter(gene %in% i) %>%
+    ggplot(aes(x = treatment, y = vsd, color = sex, fill = treatment)) +
     geom_boxplot() +
     scale_alpha_manual(values = c(0.5, 1)) +
-    labs(subtitle = i, y = "expression", x = "parental stage") +
-    facet_wrap(~gene, scales = "free_y") +
+    labs(y = "expression", x = "parental stage", subtitle = i) +
+    facet_wrap(~ tissue , scales = "free_y", nrow = 3) +
     theme_B3() +
-    scale_color_manual(values = sexcolors)
+    scale_color_manual(values = sexcolors) +
+    scale_fill_manual(values = colorscharmaip)
  print(p)
@@
   select(bird,tissuegene,vsd) %>%
+  filter(bird != "x.g.g.ATLAS") %>%
   pivot_wider(id_cols = bird, names_from = tissuegene, values_from = vsd)
-
 dft <- as.data.frame(dft)
@@
 
-x %>% network_plot.cor_df(min_cor = .2)
+x %>% network_plot.cor_df() + 
+  scale_color_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1))
 
-x %>%
+a <- x %>%
+  focus(hyp_PRL) %>%
+  mutate(rowname = reorder(rowname, hyp_PRL)) %>%
+  ggplot(aes(rowname, hyp_PRL, fill = hyp_PRL)) +
+    geom_col() + coord_flip() +
+  labs(y = "corr. w/ hyp PRL", x = "candidate genes by tissue") +
+  theme_B3() + 
+  theme(legend.position = "bottom", legend.title = element_blank()) +
+  scale_fill_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1)) +
+  ylim(c(-0.1,0.6))
+
+b <- x %>%
   focus(pit_PRL) %>%
   mutate(rowname = reorder(rowname, pit_PRL)) %>%
-  ggplot(aes(rowname, pit_PRL)) +
+  ggplot(aes(rowname, pit_PRL, fill = pit_PRL)) +
     geom_col() + coord_flip() +
-  labs(y = "corrleation with pituitiary PRL", x = "candidate genes by tissue") +
+  labs(y = "corr. w/ pit PRL", x = NULL) +
   theme_B3() + 
-  theme(legend.position = "none")
+  theme(legend.position = "bottom", legend.title = element_blank()) +
+  scale_fill_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1))  +
+  ylim(c(-0.1,0.6))
+
+c <- x %>%
+  focus(gon_PRL) %>%
+  mutate(rowname = reorder(rowname, gon_PRL)) %>%
+  ggplot(aes(rowname, gon_PRL, fill = gon_PRL)) +
+    geom_col() + coord_flip() +
+  labs(y = "corr. w/ gon PRL", x = NULL) +
+  theme_B3() + 
+  theme(legend.position = "bottom", legend.title = element_blank()) +
+  scale_fill_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1))  +
+  ylim(c(-0.1,0.6))
+
+plot_grid(a,b, c, nrow = 1, rel_widths = c(1.1,1,1))
 ```

### 2019-12-21 analysis/02_PCA.Rmd
new tsne pca

--- 2019-12-20 analysis/02_PCA.Rmd
@@
 
+source("../R/themes.R")
+
 knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/pca/', cache = T)
 
-source("../R/themes.R")  
 ```
@@
 #row.names(countData) == row.names(colData)
+
+head(colData)
 ```
 
-```{r pca}
-# pca
-mypca <- prcomp(countData)
-eig.val <- get_eigenvalue(mypca)
-head(eig.val)
+```{r tSNE-df}
+# prep for tsne
 
-mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
-                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
-                  ID = row.names(countData))
-mypcadf$V1 <- row.names(mypcadf)
-mypcadf <- left_join(colData, mypcadf)
-mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
-head(mypcadf)
+subsetmaketsne <- function(whichtissue, whichtreatment, whichsex){
 
-plotcolorfulpcs <- function(whichfactor, whichcolors){
-  p <- ggplot(mypcadf, aes(x = PC1, y = PC2, color = whichfactor)) +
-    geom_point(size = 1)  +
-    theme_B3() +
-    theme(legend.title = element_blank(),
-         axis.text = element_blank(),
-         legend.position = "none") +
-    labs(x = "PC1", y = "PC2") +
-    scale_color_manual(values = whichcolors) 
-  print(p)
-}
+  colData <- colData %>%
+    dplyr::filter(tissue %in% whichtissue,
+                  treatment %in% whichtreatment,
+                  sex %in% whichsex) 
+  row.names(colData) <- colData$V1
 
-a <- plotcolorfulpcs(mypcadf$tissue, colorstissue)  
-b <- plotcolorfulpcs(mypcadf$sex, sexcolors) 
-c <- plotcolorfulpcs(mypcadf$treatment, colorscharmaip)
+  # save counts that match colData
+  savecols <- as.character(colData$V1) 
+  savecols <- as.vector(savecols)
 
-d <- fviz_screeplot(mypca, addlabels = TRUE, ylim = c(0, 50),  ncp = 5) + labs(title = NULL, y = "Variance") + theme_B3() 
-e <- fviz_contrib(mypca, choice = "var", axes = 1, top = 5)  + labs(title = NULL, subtitle = "PC1", x = NULL) + 
-          theme_B3() + theme(axis.text.x = element_text(angle = 55, hjust = 1))
-f <- fviz_contrib(mypca, choice = "var", axes = 2, top = 5)  + labs(title = NULL, subtitle = "PC2", y = NULL, x = NULL) + 
-          theme_B3()  + theme(axis.text.x = element_text(angle = 55, hjust = 1))
-g <- fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
-                   repel = TRUE ,
-                  select.var= list(contrib = 6)) + 
-      theme_B3() + 
-      labs(x = "PC1", y = "PC2", title =  NULL) +
-      theme( axis.text = element_blank())
+  countData <- as.data.frame(t(countData))
+  countData <- countData %>% dplyr::select(one_of(savecols)) 
+  countData <- as.data.frame(t(countData))
 
-abc <- plot_grid(a,b, c, nrow =  1,  labels = c("d", "e", "f"), label_size = 12)
-defg <- plot_grid(d,e,f,g, nrow = 1, rel_widths = c(1, 0.55, 0.45, 1), labels = c("h", "i", "j", "k" ), label_size = 12)
-dg <- plot_grid(d,g, NULL, nrow = 1, rel_widths = c(1, 1, 1), labels = c("h", "i", NULL ), label_size = 12)
-plot_grid(abc, defg, nrow = 2)
-```
-
-```{r tSNE}
-# prep for tsne
   euclidist <- dist(countData) # euclidean distances between the rows
@@
   tsne_df_cols <- cbind(colData2, tsne_df)
+  return(tsne_df_cols)
+}
 
-plotcolorfultsnes <- function(whichfactor, whichcolors){
-  p <- ggplot(tsne_df_cols, aes(x = V1, y = V2, color = whichfactor)) +
+chartsne <- subsetmaketsne(tissuelevels, charlevels, sexlevels)
+hyptsne <- subsetmaketsne("hypothalamus", charlevels, sexlevels)
+pittsne <- subsetmaketsne("pituitary", charlevels, sexlevels)
+gontsne <- subsetmaketsne("gonads", charlevels, sexlevels)
+
+```
+
+```{r tSNE}
+plotcolorfultsnes <- function(tsnedf, whichfactor, whichcolors){
+  p <- ggplot(tsnedf, aes(x = V1, y = V2, color = whichfactor, shape = tissue)) +
     geom_point(size = 1) +
@@
     theme(legend.position = "none",
-          axis.text = element_blank())
+          axis.text = element_blank()) +
+    scale_shape_manual(values = myshapes)
   print(p)
@@
 
-h <- plotcolorfultsnes(tsne_df_cols$tissue, colorstissue)  
-i <- plotcolorfultsnes(tsne_df_cols$sex, sexcolors)   
-j <- plotcolorfultsnes(tsne_df_cols$treatment, colorscharmaip)
+a <- plotcolorfultsnes(chartsne, chartsne$tissue, colorstissue)  
+b <- plotcolorfultsnes(chartsne, chartsne$sex, sexcolors)   
+c <- plotcolorfultsnes(chartsne, chartsne$treatment, colorscharmaip)
 
-hij <- plot_grid(h,i,j, nrow = 1, labels = c("a", "b", "c"), label_size = 12)
-hij
+abc <- plot_grid(a,b,c, nrow = 1, labels = c("a", "b", "c"), label_size = 12)
+abc
+
 ```
-```{r PCA-tSNE}
-plot_grid(hij, abc, dg, nrow = 3)
+
+```{r subsetmakepca}
+subsetmakepca <- function(whichtissue, whichtreatment, whichsex){
+
+  colData <- colData %>%
+    dplyr::filter(tissue %in% whichtissue,
+                  treatment %in% whichtreatment,
+                  sex %in% whichsex) 
+  row.names(colData) <- colData$V1
+
+  # save counts that match colData
+  savecols <- as.character(colData$V1) 
+  savecols <- as.vector(savecols)
+
+  countData <- as.data.frame(t(countData))
+  countData <- countData %>% dplyr::select(one_of(savecols)) 
+  countData <- as.data.frame(t(countData))
+
+  mypca <- prcomp(countData)
+
+  mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
+                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
+                  ID = row.names(countData))
+  mypcadf$V1 <- row.names(mypcadf)
+  mypcadf <- left_join(colData, mypcadf)
+  mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
+  return(mypcadf)
+}
+
+charpca <- subsetmakepca(tissuelevels, charlevels, sexlevels)
+hyppca <- subsetmakepca("hypothalamus", charlevels, sexlevels)
+pitpca <- subsetmakepca("pituitary", charlevels, sexlevels)
+gonpca <- subsetmakepca("gonads", charlevels, sexlevels)
+
 ```
+
+```{r pca}
+plotcolorfulpcs <- function(mypcadf,  whichfactor, whichcolors){
+  p <- mypcadf %>%
+    ggplot(aes(x = PC1, y = PC2, shape = tissue, color = whichfactor )) +
+    geom_point(size = 1)  +
+    theme_B3() +
+    theme(legend.title = element_blank(),
+         axis.text = element_blank(),
+         legend.position = "none") +
+    labs(x = "PC1", y = "PC2")  +
+    scale_color_manual(values = whichcolors) +
+    scale_shape_manual(values = myshapes)
+  print(p)
+}
+
+d <- plotcolorfulpcs(charpca, charpca$tissue, colorstissue)  
+e <- plotcolorfulpcs(charpca, charpca$sex, sexcolors) 
+f <- plotcolorfulpcs(charpca,charpca$treatment, colorscharmaip)
+
+def <- plot_grid(d,e, f, nrow =  1,  labels = c("d", "e", "f"), label_size = 12)
+def
+```
+
+```{r fviz}
+
+makefvizdf <-  function(whichtissue, whichtreatment, whichsex){
+  colData <- colData %>%
+      dplyr::filter(tissue %in% whichtissue,
+                  treatment %in% whichtreatment,
+                  sex %in% whichsex) 
+  row.names(colData) <- colData$V1
+
+  # save counts that match colData
+  savecols <- as.character(colData$V1) 
+  savecols <- as.vector(savecols)
+
+  countData <- as.data.frame(t(countData))
+  countData <- countData %>% dplyr::select(one_of(savecols)) 
+  countData <- as.data.frame(t(countData))
+
+  mypca <- prcomp(countData)
+  return(mypca)
+}
+
+charfviz <- makefvizdf(tissuelevels, charlevels, sexlevels)
+hypv <- makefvizdf("hypothalamus", charlevels, sexlevels)
+pitfviz <- makefvizdf("pituitary", charlevels, sexlevels)
+gonfviz <- makefvizdf("gonads", charlevels, sexlevels)
+
+g <- fviz_screeplot(charfviz, addlabels = TRUE, ylim = c(0, 50),  ncp = 5, barcolor = "white", barfill = "white") + 
+  labs(title = NULL, y = "PC Variance") + theme_B3() 
+h <- fviz_pca_var(charfviz,  labelsize = 3 , axes.linetype = "blank", 
+                   repel = TRUE ,
+                  select.var= list(contrib = 6)) + 
+      theme_B3() + 
+      labs(x = "PC1", y = "PC2", title =  NULL) +
+      theme( axis.text = element_blank())
+
+legend <- png::readPNG("../figures/images/DoveParentsRNAseq_legendchar.png")
+legend <-  grid::rasterGrob(legend, interpolate=TRUE)
+
+gh <- plot_grid(g,h, legend, nrow = 1, rel_widths = c(1, 1, 1), labels = c("h", "i", NULL ), label_size = 12)
+gh
+plot_grid(def, gh, nrow = 2)
+```
+
+```{r PCA-tSNE, fig.height= 7, fig.width=7}
+plot_grid( abc, def, gh, nrow = 3)
+
+```
+
+## tissue specific
+
+```{r tissuespecific, fig.height=7,fig.width=7}
+h1 <- plotcolorfulpcs(hyppca, hyppca$sex, sexcolors) 
+h2 <- plotcolorfulpcs(hyppca, hyppca$treatment, colorscharmaip)
+
+p1 <- plotcolorfulpcs(pitpca, pitpca$sex, sexcolors) 
+p2 <- plotcolorfulpcs(pitpca, pitpca$treatment, colorscharmaip)
+
+g1 <- plotcolorfulpcs(gonpca, gonpca$sex, sexcolors) 
+g2 <- plotcolorfulpcs(gonpca, gonpca$treatment, colorscharmaip)
+
+plot_grid(h1,h2,p1,p2,g1,g2, nrow = 3)
+
+plotfriz <- function(frizdf){
+
+  p1 <- fviz_contrib(frizdf, choice = "var", 
+                     axes = 1, top = 3 , sort.val = "asc")  + 
+    labs(title = NULL,   x = "PC1", y = NULL) + 
+    theme_B3() + coord_flip() +
+    theme(axis.text.x = element_blank())
+
+  p2 <- fviz_contrib(frizdf, choice = "var", 
+                     axes = 2, top = 3 , sort.val = "asc")  + 
+    labs(title = NULL,   x = "PC2", y = "contrib.") + 
+     theme_B3()  + coord_flip() +
+    theme(axis.text.x = element_blank())
+
+  p3 <- plot_grid(p1,p2, align = "v", nrow = 2, rel_heights = c(1,1.2))
+  print(p3)
+}
+
+h3 <- plotfriz(hypv) 
+p3 <-plotfriz(pitfviz)
+g3 <- plotfriz(gonfviz)
+
+h4 <- plotcolorfultsnes(hyptsne, hyptsne$sex, sexcolors)   
+h5 <- plotcolorfultsnes(hyptsne, hyptsne$treatment, colorscharmaip)
+
+p4 <- plotcolorfultsnes(pittsne, pittsne$sex, sexcolors)   
+p5 <- plotcolorfultsnes(pittsne, pittsne$treatment, colorscharmaip)
+
+g4 <- plotcolorfultsnes(gontsne, gontsne$sex, sexcolors)   
+g5 <- plotcolorfultsnes(gontsne, gontsne$treatment, colorscharmaip)
+
+plot_grid(h4,h5,h2,h3,
+          p4,p5,p2,p3,
+          g4,g5,g2,g3,
+          nrow = 3,
+          labels = "auto", label_size = 10)
+
+```

### 2019-12-22 analysis/02_PCA.Rmd
all and tissue specific figs
new pca images

@@
                    repel = TRUE ,
-                  select.var= list(contrib = 6)) + 
+                  select.var= list(contrib = 5)) + 
       theme_B3() + 
-      labs(x = "PC1", y = "PC2", title =  NULL) +
+      labs(x = "PC1 (35.8%)", y = "PC2 (32.5%)", title =  NULL) +
       theme( axis.text = element_blank())
@@
 ```{r tissuespecific, fig.height=7,fig.width=7}
+
+## tsne 
+h4 <- plotcolorfultsnes(hyptsne, hyptsne$sex, sexcolors)  + labs(y = "hypothalamus") 
+h5 <- plotcolorfultsnes(hyptsne, hyptsne$treatment, colorscharmaip)
+
+p4 <- plotcolorfultsnes(pittsne, pittsne$sex, sexcolors)   + labs(y = "pituitary") 
+p5 <- plotcolorfultsnes(pittsne, pittsne$treatment, colorscharmaip)
+
+g4 <- plotcolorfultsnes(gontsne, gontsne$sex, sexcolors)  + labs(y = "gonads")  
+g5 <- plotcolorfultsnes(gontsne, gontsne$treatment, colorscharmaip)
+
+## pca
+
 h1 <- plotcolorfulpcs(hyppca, hyppca$sex, sexcolors) 
@@
 
+# screeplot for variance
+fviz_screeplot(hypv, addlabels = TRUE,   ncp = 5) + 
+  labs(title = NULL, y = "PC Variance") + theme_B3() 
+fviz_screeplot(pitfviz, addlabels = TRUE,   ncp = 5) + 
+  labs(title = NULL, y = "PC Variance") + theme_B3() 
+fviz_screeplot(gonfviz, addlabels = TRUE,    ncp = 5) + 
+  labs(title = NULL, y = "PC Variance") + theme_B3()
+
+# fviz contributions
+
 plotfriz <- function(frizdf){
 
-  p1 <- fviz_contrib(frizdf, choice = "var", 
-                     axes = 1, top = 3 , sort.val = "asc")  + 
-    labs(title = NULL,   x = "PC1", y = NULL) + 
-    theme_B3() + coord_flip() +
-    theme(axis.text.x = element_blank())
-
-  p2 <- fviz_contrib(frizdf, choice = "var", 
-                     axes = 2, top = 3 , sort.val = "asc")  + 
-    labs(title = NULL,   x = "PC2", y = "contrib.") + 
-     theme_B3()  + coord_flip() +
-    theme(axis.text.x = element_blank())
-
-  p3 <- plot_grid(p1,p2, align = "v", nrow = 2, rel_heights = c(1,1.2))
-  print(p3)
+     p <- fviz_pca_var(frizdf,  labelsize = 3.5 , axes.linetype = "blank", 
+                   repel = T ,  select.var= list(contrib = 3))  + 
+      labs(title = NULL) + 
+      theme_B3() + 
+      theme(axis.text = element_blank())
+   print(p)
 }
 
-h3 <- plotfriz(hypv) 
-p3 <-plotfriz(pitfviz)
-g3 <- plotfriz(gonfviz)
+h3 <- plotfriz(hypv) + labs(x = "PC1 (59.8%)", y = "PC2 (37.7%)") 
+p3 <-plotfriz(pitfviz) + labs(x = "PC1 (54.3%)", y = "PC2 (35.7%)") 
+g3 <- plotfriz(gonfviz) + labs(x = "PC1 (82.7%)", y = "PC2 (11.6%)")
 
-h4 <- plotcolorfultsnes(hyptsne, hyptsne$sex, sexcolors)   
-h5 <- plotcolorfultsnes(hyptsne, hyptsne$treatment, colorscharmaip)
-
-p4 <- plotcolorfultsnes(pittsne, pittsne$sex, sexcolors)   
-p5 <- plotcolorfultsnes(pittsne, pittsne$treatment, colorscharmaip)
-
-g4 <- plotcolorfultsnes(gontsne, gontsne$sex, sexcolors)   
-g5 <- plotcolorfultsnes(gontsne, gontsne$treatment, colorscharmaip)
+## all together now
 
@@
           labels = "auto", label_size = 10)
-
 ```

### 2019-12-22 analysis/03_DESeq2_characterization.Rmd
remove old files, working on new lda analysis

@@
 plot_grid(icons, hypothalamus, pituitary, gonads, nrow = 4, rel_heights = c(1,1,1,1.25))
+
 ```

### 2019-12-22 analysis/04_LDA.Rmd
remove old files, working on new lda analysis

--- 2019-11-10 analysis/04_DESeq2_sexes.Rmd
@@
 ```{r readfiles}
-# import "colData" which contains sample information and "countData" which contains read counts
-c.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
-c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
-geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+colData <- read.csv("../metadata/00_samples.csv", row.names = 1)
 
-# set treatment levels
-c.colData$treatment <- factor(c.colData$treatment, levels = 
-                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
+characterization <- colData %>% filter(study == "charcterization")
+manipluation <- colData %>% filter(study == "manipulation")
 
-# craete variable that will be critical for subset later on
-c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
+vsdshyp <- read.csv("../results/06_hypallvsd.csv")
+vsdspit <- read.csv("../results/06_pitallvsd.csv")
+vsdsgon <- read.csv("../results/06_gonallvsd.csv")
 
-# ceate new variable for hypothesis testing
-c.colData <- c.colData %>%
-    mutate(hypothesis = fct_recode(treatment,
-                            "non-parental" = "control",
-                            "non-parental" = "bldg",
-                            "incubation" = "lay",
-                            "incubation" = "inc.d3",
-                            "incubation" = "inc.d9",
-                            "incubation" = "inc.d17",
-                            "hatchling.care" = "hatch",
-                            "hatchling.care" = "n5",
-                            "hatchling.care" = "n9"))
-
-# drop columsn that are bad for model AND not needed for later joining
-c.colData$bird <- NULL
-c.colData$tissue <- NULL
-c.colData$study <- NULL
-c.colData$group <- NULL
-head(c.colData)
-
-# prep col data for subsequent analyses
-colDataHyp <- subsetcolData2(c.colData, c("female_hypothalamus", "male_hypothalamus"))
-colDataPit <- subsetcolData2(c.colData, c("female_pituitary", "male_pituitary"))
-colDataGon <- subsetcolData2(c.colData, c("female_gonad", "male_gonad"))
-```
-
-```{r dds}
-dds.hypothalamus <- subsetDESeq2(c.colData,  c.countData, c("female_hypothalamus","male_hypothalamus") )
-dds.pituitary <- subsetDESeq2(c.colData,  c.countData, c("female_pituitary", "male_pituitary" ) )
-dds.gonad <- subsetDESeq2(c.colData,  c.countData, c("female_gonad", "male_gonad") )
-```
-
-## total degs
-
-```{r totalDEGs, results = 'hide', eval = F}
-DEGs.pituitary <- returntotalDEGs(dds.pituitary)
-plottotalDEGs(DEGs.pituitary, "Male & Female Pituitary DEGs")
-```
-
-## Variance stabilized data
-
-```{r vsd}
-# create variance stbilized DESeq transformation
-vsd.hyp <- vst(dds.hypothalamus, blind=FALSE) 
-vsd.pit <- vst(dds.pituitary, blind=FALSE) 
-vsd.gon <- vst(dds.gonad, blind=FALSE)
-
-# save as a dataframe for downstream analysis
-vsd.hyp.df <- vsd.dataframe(vsd.hyp) 
-vsd.pit.df <- vsd.dataframe(vsd.pit) 
-vsd.gon.df <- vsd.dataframe(vsd.gon) 
-```
-
-```{r viewwritevsd}
-savevsdfiles(vsd.hyp.df, colDataHyp, "hyp")
-savevsdfiles(vsd.pit.df, colDataPit, "pit")
-savevsdfiles(vsd.gon.df, colDataGon, "gon")
-```
-
-## PCA
-
-```{r pca}
-hypPCA <- returnPCAs2(vsd.hyp)
-plotPC12(hypPCA, "Hypothalamus")
-
-pitPCA <- returnPCAs2(vsd.pit)
-plotPC12(pitPCA, "Pituitary")
-
-gonPCA <- returnPCAs2(vsd.gon)
-plotPC12(gonPCA, "Gonad")
+vsdall <- full_join(vsdshyp, vsdspit)
+vsdall <- full_join(vsdall, vsdsgon)
+head(vsdall)
 ```
@@
 ```
-
-## plot genes in a  PRL WGCNA module
-
-```{r wgcna, eval = T, fig.width=6, fig.height=6}
-# pitutiary expression
-plotWGCNAcandidates(vsd.pit, c("NP_990797.2"), colDataPit, "PRL expression, Pituitary")
-
-# COX1,  PRL, 
-topPCAloadings <- c("NP_006917.1",  "NP_990797.2")
-h <- plotWGCNAcandidates(vsd.hyp, topPCAloadings, colDataHyp, "Hypothalamus")
-p <- plotWGCNAcandidates(vsd.pit, topPCAloadings, colDataPit, "Pituitary")
-
-plot_grid(h + theme(axis.text.x = element_blank(), legend.position = "none"),
-          p , nrow = 2, rel_heights = c(0.45,0.55))
-```
-
-```{r wgcna2, eval = T, fig.width=12, fig.height=8}
-PRL_associated <- read.csv("../results/08_PRL_associated.csv", header = T)
-prl36 <-  PRL_associated$entrezid[1:57]
-prl72 <-  PRL_associated$entrezid[58:102]
-
-plotWGCNAcandidates(vsd.pit, prl36, colDataPit, "A-K genes with PRL pattern in the pituitary")
-plotWGCNAcandidates(vsd.pit, prl72, colDataPit, "K-Z genes with PRL pattern in the pituitary")
-```
-
-```{r ctsGE}
-rts <- assay(vsd.pit)
-```

### 2019-12-22 analysis/04_LDA.Rmd
lda working now

@@
 library(tidyverse)
-library(DESeq2)
 library(cowplot)
-library(pheatmap)
-library(viridis)
 library(forcats)
@@
 library(MASS) # LDA analysis
-library(ggtext) # for pretty plots
-library(ctsGE)
-
-library(BiocParallel)
-register(MulticoreParam(6))
 
@@
 
-## DEseq2 analysis on characterization. Looking at treatment AND sex for each tissue
+## LDA analysis of characterization and manipulation
 
@@
 
-characterization <- colData %>% filter(study == "charcterization")
-manipluation <- colData %>% filter(study == "manipulation")
+charHyp <- colData %>% filter(study == "charcterization", tissue == "hypothalamus") %>% droplevels()
+charPit <- colData %>% filter(study == "charcterization", tissue == "pituitary") %>% droplevels()
+charGon <- colData %>% filter(study == "charcterization", tissue == "gonad") %>% 
+  mutate(tissue = fct_recode(tissue, "gonads" = "gonad")) %>% droplevels()
 
-vsdshyp <- read.csv("../results/06_hypallvsd.csv")
-vsdspit <- read.csv("../results/06_pitallvsd.csv")
-vsdsgon <- read.csv("../results/06_gonallvsd.csv")
+manipHyp <- colData %>% filter(study == "manipulation", tissue == "hypothalamus") %>% droplevels()
+manipPit <- colData %>% filter(study == "manipulation", tissue == "pituitary") %>% droplevels()
+manipGon <- colData %>% filter(study == "manipulation", tissue == "gonad") %>% 
+  mutate(tissue = fct_recode(tissue, "gonads" = "gonad")) %>% droplevels()
 
-vsdall <- full_join(vsdshyp, vsdspit)
-vsdall <- full_join(vsdall, vsdsgon)
-head(vsdall)
+selectvsd <- function(pathtofile, colData){
+
+  df <- read.csv(pathtofile, row.names = 1)
+  savecols <- as.character(colData$V1) 
+  savecols <- as.vector(savecols) 
+  df <- df %>% dplyr::select(one_of(savecols)) 
+  df <- as.data.frame(t(df))
+  df$V1 <- row.names(df)
+  return(df)
+}
+
+vsd.hyp.train <- selectvsd("../results/06_hypallvsd.csv",  charHyp)
+vsd.pit.train <- selectvsd("../results/06_pitallvsd.csv",  charPit)
+vsd.gon.train <- selectvsd("../results/06_gonallvsd.csv",  charGon)
+
+vsd.hyp.test <- selectvsd("../results/06_hypallvsd.csv",  manipHyp)
+vsd.pit.test <- selectvsd("../results/06_pitallvsd.csv",  manipPit)
+vsd.gon.test <- selectvsd("../results/06_gonallvsd.csv",  manipGon)
 ```
@@
 ```{r LDAdfs}
-# LDA for treatment (9 groups)
-LDA.hyp1 <- LDAdata.treatment(vsd.hyp, colDataHyp)
-LDA.pit1 <- LDAdata.treatment(vsd.pit, colDataPit)
-LDA.gon1 <- LDAdata.treatment(vsd.gon, colDataGon)
 
-# LDA for hypothesis (3 groups)
+LDanalysis <- function(trainsamples, traindata, testdata, testsamples){
 
-LDA.hyp2 <- LDAdata.hypothesis(vsd.hyp, colDataHyp)
-LDA.pit2 <- LDAdata.hypothesis(vsd.pit, colDataPit)
-LDA.gon2 <- LDAdata.hypothesis(vsd.gon, colDataGon)
+  train.data <- left_join(trainsamples, traindata) %>%
+    dplyr::select(-V1, -bird, -tissue, -group, -study, -sex)
+
+  test.data <- testdata
+
+  # Normalize the data. Categorical variables are automatically ignored.
+  # Estimate preprocessing parameters
+  preproc.param <- train.data %>% 
+    preProcess(method = c("center", "scale"))
+
+  # Transform the data using the estimated parameters
+  train.transformed <- preproc.param %>% predict(train.data)
+  test.transformed <- preproc.param %>% predict(test.data)
+
+  # LDA analysis
+  # Fit the model
+  model <- lda(treatment~ ., data = train.transformed)
+  # Make predictions
+  predictions <- model %>% predict(test.transformed)
+
+  # Model accuracy
+  print("model accuracy")
+  print("predictions$class==test.transformed$treatment)")
+  print(mean(predictions$class==test.transformed$treatment))
+
+  # results
+  print("the samples sizes")
+  print(model$counts)
+
+  print("the prior probabilities used")
+  print(model$prior)
+
+  print("svd: the singular values, which give the ratio of the between- and within-group standard deviations on the linear discriminant variables. Their squares are the canonical F-statistics.")
+  print(model$svd)
+
+  #  predictions
+  predictions <- model %>% predict(test.transformed)
+  head(predictions)
+
+  # Predicted classes
+  #print(predictions$class, 6)
+  # Predicted probabilities of class memebership.
+  #print(predictions$posterior, 6) 
+  # Linear discriminants
+  #print(predictions$x, 3)
+
+  predictedstage <-  predict(model, test.transformed)$class
+  testsamples$predictedstage <- predictedstage
+
+  lda.data <- cbind(testsamples, predictions$x)
+
+  return(lda.data)
+}
+
+LDA.hyp <- LDanalysis(charHyp, vsd.hyp.train, vsd.hyp.test, manipHyp)
+LDA.pit <- LDanalysis(charPit, vsd.pit.train, vsd.pit.test, manipPit)
+LDA.gon <- LDanalysis(charGon, vsd.gon.train, vsd.gon.test, manipGon)
 ```
 
-```{r LDA, fig.width = 9}
-# treatment
-a <- LDAplot.treatment(LDA.hyp1 , "Hypothalamus model 1", "parental state ~ .  0.277 pred. acc.",
-                        "LD1, F = 7.1504779", "LD2, F = 2.4219585")
-b <- LDAplot.treatment(LDA.pit1 ,  "Pituitary model 1", "parental state ~ .  0.405 pred. acc.",
-                        "LD1, F = 9.6144160", "LD2, F = 3.7773314")
-c <- LDAplot.treatment(LDA.gon1 , "Gonad model 1", "parental state ~ .  has 0.297 pred. acc.",
-                        "LD1, F = 6.137500", "LD2, F = 3.320437")
+```{r LDAplots}
+#myshapes = c("hypothalamus" = 20,  "pituitary" = 17,  "gonads" = 15)
 
-c <- c + theme(legend.direction = "horizontal")
-mylegend <- get_legend(c)
+a <- ggplot(LDA.hyp, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 20) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+b <- ggplot(LDA.hyp, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 20) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
 
-temp <- plot_grid(a + theme(legend.position = "none"),
-                  b + theme(legend.position = "none"),
-                  c + theme(legend.position = "none"), nrow = 1)
+plot_grid(a,b)
 
-plot_grid(temp, mylegend, nrow = 2, rel_heights = c(1,0.3))
+a <- ggplot(LDA.pit, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+b <- ggplot(LDA.pit, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
 
-# hypothesis
-d <- LDAplot.hypothesis(LDA.hyp2 , "Hypothalamus model 2", "parental stage ~ .  0.432 pred. acc.",
-                        "LD1, F = 7.386458", "LD2, F = 3.720794")
-e <- LDAplot.hypothesis(LDA.pit2 ,  "Pituitary model 2", "parental stage ~ .  0.473 pred. acc.",
-                        "LD1, F = 7.846659", "LD2, F = 5.076861")
-f <- LDAplot.hypothesis(LDA.gon2 , "Gonad model 2", "parental stage ~ .  has 0.368 pred. acc.",
-                        "LD1, F = 6.029069", "LD2, F = 3.837591")
+plot_grid(a,b)
 
-f <- f + theme(legend.direction = "horizontal")
-mylegend <- get_legend(f)
+a <- ggplot(LDA.gon, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+b <- ggplot(LDA.gon, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
 
-temp <- plot_grid(d + theme(legend.position = "none"),
-                  e + theme(legend.position = "none"),
-                  f + theme(legend.position = "none"), nrow = 1)
+plot_grid(a,b)
 
-plot_grid(temp, mylegend, nrow = 2, rel_heights = c(1,0.3))
+LDA.hyp %>% distinct(treatment, predictedstage) %>%
+  group_by(treatment) %>%
+  summarize(predictedstages = str_c(predictedstage , collapse = ", "))
+
+LDA.hyp %>% 
+  group_by(treatment,predictedstage) %>%
+  summarize(n = n())
+
+LDA.pit %>% distinct(treatment, predictedstage) %>%
+  group_by(treatment) %>%
+  summarize(predictedstages = str_c(predictedstage , collapse = ", "))
+
+LDA.pit %>% 
+  group_by(treatment,predictedstage) %>%
+  summarize(n = n())
+
+LDA.gon %>% distinct(treatment, predictedstage) %>%
+  group_by(treatment) %>%
+  summarize(predictedstages = str_c(predictedstage , collapse = ", "))
+
+LDA.gon %>% 
+  group_by(treatment,predictedstage) %>%
+  summarize(n = n())
+
 ```

### 2019-12-23 analysis/04_LDA.Rmd
now with new lda analysis

@@
 
-knitr::opts_chunk$set(fig.path = '../figures/sexes/', cache = TRUE,  message=F, comment=FALSE, warning=FALSE)
+knitr::opts_chunk$set(fig.path = '../figures/LDA/', cache = TRUE,  message=F, comment=FALSE, warning=FALSE)
 ```
@@
 
-a <- ggplot(LDA.pit, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
-b <- ggplot(LDA.pit, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+c <- ggplot(LDA.pit, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+d <- ggplot(LDA.pit, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
 
@@
 
-a <- ggplot(LDA.gon, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
-b <- ggplot(LDA.gon, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+e <- ggplot(LDA.gon, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+f <- ggplot(LDA.gon, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
 
-plot_grid(a,b)
+plot_grid(a,b,c,d,e,f, nrow = 3)
 
@@
   summarize(n = n())
-
 ```

### 2019-12-23 analysis/04_LDA.Rmd
better lda tables

@@
 ```{r LDAplots}
+library(kableExtra)
 #myshapes = c("hypothalamus" = 20,  "pituitary" = 17,  "gonads" = 15)
@@
 
-LDA.hyp %>% distinct(treatment, predictedstage) %>%
+df1 <- LDA.hyp %>% distinct(treatment, predictedstage) %>%
   group_by(treatment) %>%
   summarize(predictedstages = str_c(predictedstage , collapse = ", ")) 
+kable(df1)
 
-LDA.hyp %>% 
+df2 <- LDA.hyp %>% 
   group_by(treatment,predictedstage) %>%
   summarize(n = n())
+kable(df2)
 
-LDA.pit %>% distinct(treatment, predictedstage) %>%
+df3 <- LDA.pit %>% distinct(treatment, predictedstage) %>%
   group_by(treatment) %>%
   summarize(predictedstages = str_c(predictedstage , collapse = ", "))
+kable(df3)
 
-LDA.pit %>% 
+df4  <- LDA.pit %>% 
   group_by(treatment,predictedstage) %>%
   summarize(n = n())
+kable(df4)
 
-LDA.gon %>% distinct(treatment, predictedstage) %>%
+df5 <- LDA.gon %>% distinct(treatment, predictedstage) %>%
   group_by(treatment) %>%
   summarize(predictedstages = str_c(predictedstage , collapse = ", "))
+kable(df5)
 
-LDA.gon %>% 
+df6 <- LDA.gon %>% 
   group_by(treatment,predictedstage) %>%
   summarize(n = n())
+kable(df6)
 ```

### 2019-12-23 analysis/04_LDA.Rmd
better tabels

@@
 plot_grid(a,b,c,d,e,f, nrow = 3)
+```
+
+```{r LDApredictions}
 
@@
   group_by(treatment) %>%
-  summarize(predictedstages = str_c(predictedstage , collapse = ", ")) 
+  summarize(predictedstages = str_c(predictedstage , collapse = ", "))  %>%
+  mutate(tissue = "hypothalamus")
 kable(df1)
@@
   group_by(treatment,predictedstage) %>%
-  summarize(n = n())
+  summarize(n = n()) %>%
+  mutate(tissue = "hypothalamus")
 kable(df2)
@@
   group_by(treatment) %>%
-  summarize(predictedstages = str_c(predictedstage , collapse = ", "))
+  summarize(predictedstages = str_c(predictedstage , collapse = ", ")) %>%
+  mutate(tissue = "pituitary")
 kable(df3)
@@
   group_by(treatment,predictedstage) %>%
-  summarize(n = n())
+  summarize(n = n()) %>%
+  mutate(tissue = "pituitary")
 kable(df4)
@@
   group_by(treatment) %>%
-  summarize(predictedstages = str_c(predictedstage , collapse = ", "))
+  summarize(predictedstages = str_c(predictedstage , collapse = ", ")) %>%
+  mutate(tissue = "gonads")
 kable(df5)
@@
   group_by(treatment,predictedstage) %>%
-  summarize(n = n())
+  summarize(n = n()) %>%
+  mutate(tissue = "gonads")
 kable(df6)

### 2019-12-23 analysis/04_LDA.Rmd
now with shape

@@
 
-plot_grid(a,b)
-
 c <- ggplot(LDA.pit, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
 d <- ggplot(LDA.pit, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
-
-plot_grid(a,b)
 
@@
 plot_grid(a,b,c,d,e,f, nrow = 3)
+
+g <- ggplot(LDA.hyp, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + geom_point(shape = 20) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+
+h <- ggplot(LDA.pit, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+
+i <- ggplot(LDA.gon, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+
+plot_grid(g,h,i, nrow = 1)
+
 ```

### 2019-12-23 analysis/04_LDA.Rmd
now with shape

--- 2019-12-23 analysis/04_LDA.Rmd
@@
 plot_grid(a,b,c,d,e,f, nrow = 3)
+
+g <- ggplot(LDA.hyp, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + geom_point( ) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+
+h <- ggplot(LDA.pit, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + geom_point( ) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+
+i <- ggplot(LDA.gon, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + geom_point( ) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
+
+plot_grid(g,h,i, nrow = 1)
+
 ```

### 2019-12-23 analysis/04_LDA.Rmd
so many lda plots

--- 2019-12-23 analysis/04_LDA.Rmd
@@
 library(kableExtra)
+
+LDA.hyp$treatment <- factor(LDA.hyp$treatment, levels = alllevels)
+LDA.hyp$predictedstage <- factor(LDA.hyp$predictedstage, levels = alllevels)
+
+LDA.pit$treatment <- factor(LDA.pit$treatment, levels = alllevels)
+LDA.pit$predictedstage <- factor(LDA.pit$predictedstage, levels = alllevels)
+
+LDA.gon$treatment <- factor(LDA.gon$treatment, levels = alllevels)
+LDA.gon$predictedstage <- factor(LDA.gon$predictedstage, levels = alllevels)
+
 #myshapes = c("hypothalamus" = 20,  "pituitary" = 17,  "gonads" = 15)
@@
 plot_grid(a,b,c,d,e,f, nrow = 3)
+
+g <- ggplot(LDA.hyp, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + 
+  geom_point( ) + theme_B3() +
+  theme(legend.position = "none") + 
+  scale_color_manual(values = colorscharmaip) + 
+  guides(color = guide_legend(show = FALSE))  +
+  labs(subtitle = "hypothalamus")
+
+h <- ggplot(LDA.pit, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + 
+  geom_point( ) + theme_B3() +
+  theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)  + 
+  guides(color = guide_legend(show = FALSE)) +
+  labs(subtitle = "piuitary")
+
+i <- ggplot(LDA.gon, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + 
+  geom_point( ) + theme_B3() +
+  theme(legend.position = "right") + scale_color_manual(values = colorscharmaip)  + 
+  #guides(color = guide_legend(show = FALSE)) +
+  guides(shape = guide_legend(ncol = 3),
+         color = guide_legend(ncol = 3)) +
+  labs(subtitle = "gonads")
+
+gh <- plot_grid(g,h, nrow = 1)
+ghi <- plot_grid(gh, i, nrow = 2)
+ghi
+
+plotfacetresults <- function(df, whichgroups, mysubstitle, myshape){
+
+  p <- df %>%
+    filter(treatment %in% whichgroups) %>%
+    ggplot(aes(x = LD1, LD2,  color = predictedstage)) + 
+  geom_point(shape = myshape ) + theme_B3() +
+    stat_ellipse() +
+  theme(legend.position = "none") + 
+  scale_color_manual(values = colorscharmaip) + 
+  guides(color = guide_legend(show = FALSE))  +
+  labs(y = mysubstitle) + 
+  facet_wrap(~treatment, nrow = 1)
+  return(p)
+}
+
+j <- plotfacetresults(LDA.gon, levelstiming, "hypothalamus\n LD2", 20) + labs(x = NULL)
+k <- plotfacetresults(LDA.pit, levelstiming, "pituitary\n LD2", 17) + theme(strip.text = element_blank()) + labs(x = NULL)
+l <- plotfacetresults(LDA.gon, levelstiming, "gonads\n LD2", 15) + theme(strip.text = element_blank()) + 
+  theme(legend.position = "bottom") +
+  guides(color = guide_legend(nrow  = 1))
+
+plot_grid(j,k,l, nrow = 3, rel_heights = c(1.2,1,1.6), align = "v")
+
+m <- plotfacetresults(LDA.gon, levelsremoval, "hypothalamus\n LD2", 20) + labs(x = NULL)
+n <- plotfacetresults(LDA.pit, levelsremoval, "pituitary\n LD2", 17) + theme(strip.text = element_blank()) + labs(x = NULL)
+o <- plotfacetresults(LDA.gon, levelsremoval, "gonads\n LD2", 15) + theme(strip.text = element_blank()) + 
+  theme(legend.position = "bottom") +
+  guides(color = guide_legend(nrow  = 1))
+
+plot_grid(m,n,o, nrow = 3, rel_heights = c(1.2,1,1.6), align = "v")
+
+plotfacetresults2 <- function(df, whichgroups, mysubstitle, myshape){
+  p <- df %>%
+    filter(treatment %in% whichgroups) %>%
+  ggplot(aes(x = treatment, fill = predictedstage)) +
+  geom_bar(position = position_fill(reverse = F))  +
+  facet_wrap(~sex) +
+    theme_B3() +
+  theme(legend.position = "none") +
+  scale_fill_manual(values = colorscharmaip) +
+  labs(y = mysubstitle) 
+  return(p)
+}
+
+j <- plotfacetresults2(LDA.gon, levelstiming, "hypothalamus", 20) + labs(x = NULL) + theme(axis.text.x = element_blank()) +
+  theme(legend.position = "bottom") + guides(fill = guide_legend(nrow  = 1)) 
+k <- plotfacetresults2(LDA.pit, levelstiming, "pituitary", 17) + theme(strip.text = element_blank()) + labs(x = NULL) + theme(axis.text.x = element_blank())
+l <- plotfacetresults2(LDA.gon, levelstiming, "gonads", 15) + theme(strip.text = element_blank()) + 
+   labs(x = NULL)
+
+plot_grid(j,k,l, nrow = 3, rel_heights = c(1.6,1,1.2))
+
+m <- plotfacetresults2(LDA.gon, levelsremoval, "hypothalamus", 20) + labs(x = NULL) + theme(axis.text.x = element_blank()) +
+    theme(legend.position = "top")  +  guides(fill = guide_legend(nrow  = 1))
+n <- plotfacetresults2(LDA.pit, levelsremoval, "pituitary", 17) + theme(strip.text = element_blank()) + labs(x = NULL) + theme(axis.text.x = element_blank())
+o <- plotfacetresults2(LDA.gon, levelsremoval, "gonads", 15) +
+  theme(strip.text = element_blank())  + labs(x = NULL)
+
+plot_grid(m,n,o, nrow = 3, rel_heights = c(1.6,1,1.2))
+
 ```

### 2019-12-23 analysis/04_LDA.Rmd
so many lda plots

@@
 
-j <- plotfacetresults2(LDA.gon, levelstiming, "hypothalamus", 20) + labs(x = NULL) + theme(axis.text.x = element_blank()) +
-  theme(legend.position = "bottom") + guides(fill = guide_legend(nrow  = 1)) 
-k <- plotfacetresults2(LDA.pit, levelstiming, "pituitary", 17) + theme(strip.text = element_blank()) + labs(x = NULL) + theme(axis.text.x = element_blank())
+j <- plotfacetresults2(LDA.gon, levelstiming, "hypothalamus", 20) + labs(x = NULL) + 
+  theme(axis.text.x = element_blank()) +
+  theme(legend.position = "top") + guides(fill = guide_legend(nrow  = 1)) 
+k <- plotfacetresults2(LDA.pit, levelstiming, "pituitary", 17) + 
+  theme(strip.text = element_blank()) + labs(x = NULL) + theme(axis.text.x = element_blank())
 l <- plotfacetresults2(LDA.gon, levelstiming, "gonads", 15) + theme(strip.text = element_blank()) + 

### 2019-12-23 analysis/04_LDA.Rmd
sorted

@@
   summarize(n = n()) %>%
-  mutate(tissue = "hypothalamus") 
+  mutate(tissue = "hypothalamus")  %>%
+  arrange(tissue, treatment, n)
 kable(df2)
@@
   summarize(n = n()) %>%
-  mutate(tissue = "pituitary")
+  mutate(tissue = "pituitary") %>%
+  arrange(tissue, treatment, n)
 kable(df4)
@@
   summarize(n = n()) %>%
-  mutate(tissue = "gonads")
+  mutate(tissue = "gonads") %>%
+  arrange(tissue, treatment, n)
 kable(df6)

### 2019-12-23 analysis/04_LDA.Rmd
desc

--- 2019-12-23 analysis/04_LDA.Rmd
@@
   summarize(n = n()) %>%
-  mutate(tissue = "hypothalamus") 
+  mutate(tissue = "hypothalamus")  %>%
+  arrange(tissue, treatment, desc(n))
 kable(df2)
@@
   summarize(n = n()) %>%
-  mutate(tissue = "pituitary")
+  mutate(tissue = "pituitary") %>%
+  arrange(tissue, treatment, desc(n))
 kable(df4)
@@
   summarize(n = n()) %>%
-  mutate(tissue = "gonads")
+  mutate(tissue = "gonads") %>%
+  arrange(tissue, treatment, desc(n))
 kable(df6)

### 2019-12-23 analysis/04_LDA.Rmd
new plot

@@
 
+m <- plotfacetresults2(LDA.gon, alllevels, "hypothalamus", 20) + labs(x = NULL) + theme(axis.text.x = element_blank()) +
+    theme(legend.position = "top")  +  guides(fill = guide_legend(nrow  = 1))
+n <- plotfacetresults2(LDA.pit, alllevels, "pituitary", 17) + theme(strip.text = element_blank()) + labs(x = NULL) + theme(axis.text.x = element_blank())
+o <- plotfacetresults2(LDA.gon, alllevels, "gonads", 15) +
+  theme(strip.text = element_blank())  + labs(x = NULL) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+plot_grid(m,n,o, nrow = 3, rel_heights = c(1.6,1,1.2))
+
 ```

### 2019-12-23 analysis/04_LDA_v2.Rmd
new lda

--- 2019-12-23 analysis/04_LDA.Rmd
@@
 ---
-title: "04_DESeq_sexes"
+title: "04_LDA_v2"
 output: md_document
@@
 library(MASS) # LDA analysis
+library(kableExtra)
 
@@
 colData <- read.csv("../metadata/00_samples.csv", row.names = 1)
+
+# add new columns
+
+colData <- colData %>%
+  mutate(newgrouping = fct_collapse(treatment,
+         "control" = c("control"),
+         "bldg" = c("bldg"),
+         "nearhatch" = c("inc.d17", "hatch"),
+         "earlyinc" = c("lay","inc.d3", "inc.d9"),
+         "chickcare" = c("n5", "n9"),
+         "manip" = maniplevels1))
 
@@
   train.data <- left_join(trainsamples, traindata) %>%
-    dplyr::select(-V1, -bird, -tissue, -group, -study, -sex)
+    dplyr::select(-V1, -bird, -tissue, -group, -study, -sex, -treatment)
 
@@
   # Fit the model
-  model <- lda(treatment~ ., data = train.transformed)
+  model <- lda(newgrouping~ ., data = train.transformed)
   # Make predictions
@@
   print("model accuracy")
-  print("predictions$class==test.transformed$treatment)")
-  print(mean(predictions$class==test.transformed$treatment))
+  print("predictions$class==test.transformed$newgrouping)")
+  print(mean(predictions$class==test.transformed$newgrouping))
 
@@
 
-```{r LDAplots}
+```{r LDAplotsNew}
 library(kableExtra)
@@
 LDA.hyp$treatment <- factor(LDA.hyp$treatment, levels = alllevels)
-LDA.hyp$predictedstage <- factor(LDA.hyp$predictedstage, levels = alllevels)
+LDA.hyp$predictedstage <- factor(LDA.hyp$predictedstage, levels = levelsnewgrouping)
 
 LDA.pit$treatment <- factor(LDA.pit$treatment, levels = alllevels)
-LDA.pit$predictedstage <- factor(LDA.pit$predictedstage, levels = alllevels)
+LDA.pit$predictedstage <- factor(LDA.pit$predictedstage, levels = levelsnewgrouping)
 
 LDA.gon$treatment <- factor(LDA.gon$treatment, levels = alllevels)
-LDA.gon$predictedstage <- factor(LDA.gon$predictedstage, levels = alllevels)
+LDA.gon$predictedstage <- factor(LDA.gon$predictedstage, levels = levelsnewgrouping)
 
-#myshapes = c("hypothalamus" = 20,  "pituitary" = 17,  "gonads" = 15)
-
-a <- ggplot(LDA.hyp, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 20) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip) 
-b <- ggplot(LDA.hyp, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 20) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
-
-c <- ggplot(LDA.pit, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
-d <- ggplot(LDA.pit, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
-
-e <- ggplot(LDA.gon, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
-f <- ggplot(LDA.gon, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
-
-plot_grid(a,b,c,d,e,f, nrow = 3)
-
-g <- ggplot(LDA.hyp, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + 
-  geom_point( ) + theme_B3() +
-  theme(legend.position = "none") + 
-  scale_color_manual(values = colorscharmaip) + 
-  guides(color = guide_legend(show = FALSE))  +
-  labs(subtitle = "hypothalamus")
-
-h <- ggplot(LDA.pit, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + 
-  geom_point( ) + theme_B3() +
-  theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)  + 
-  guides(color = guide_legend(show = FALSE)) +
-  labs(subtitle = "piuitary")
-
-i <- ggplot(LDA.gon, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + 
-  geom_point( ) + theme_B3() +
-  theme(legend.position = "right") + scale_color_manual(values = colorscharmaip)  + 
-  #guides(color = guide_legend(show = FALSE)) +
-  guides(shape = guide_legend(ncol = 3),
-         color = guide_legend(ncol = 3)) +
-  labs(subtitle = "gonads")
-
-gh <- plot_grid(g,h, nrow = 1)
-ghi <- plot_grid(gh, i, nrow = 2)
-ghi
-
-plotfacetresults <- function(df, whichgroups, mysubstitle, myshape){
-
-  p <- df %>%
-    filter(treatment %in% whichgroups) %>%
-    ggplot(aes(x = LD1, LD2,  color = predictedstage)) + 
-  geom_point(shape = myshape ) + theme_B3() +
-    stat_ellipse() +
-  theme(legend.position = "none") + 
-  scale_color_manual(values = colorscharmaip) + 
-  guides(color = guide_legend(show = FALSE))  +
-  labs(y = mysubstitle) + 
-  facet_wrap(~treatment, nrow = 1)
-  return(p)
-}
-
-j <- plotfacetresults(LDA.gon, levelstiming, "hypothalamus\n LD2", 20) + labs(x = NULL)
-k <- plotfacetresults(LDA.pit, levelstiming, "pituitary\n LD2", 17) + theme(strip.text = element_blank()) + labs(x = NULL)
-l <- plotfacetresults(LDA.gon, levelstiming, "gonads\n LD2", 15) + theme(strip.text = element_blank()) + 
-  theme(legend.position = "bottom") +
-  guides(color = guide_legend(nrow  = 1))
-
-plot_grid(j,k,l, nrow = 3, rel_heights = c(1.2,1,1.6), align = "v")
-
-m <- plotfacetresults(LDA.gon, levelsremoval, "hypothalamus\n LD2", 20) + labs(x = NULL)
-n <- plotfacetresults(LDA.pit, levelsremoval, "pituitary\n LD2", 17) + theme(strip.text = element_blank()) + labs(x = NULL)
-o <- plotfacetresults(LDA.gon, levelsremoval, "gonads\n LD2", 15) + theme(strip.text = element_blank()) + 
-  theme(legend.position = "bottom") +
-  guides(color = guide_legend(nrow  = 1))
-
-plot_grid(m,n,o, nrow = 3, rel_heights = c(1.2,1,1.6), align = "v")
-
-plotfacetresults2 <- function(df, whichgroups, mysubstitle, myshape){
-  p <- df %>%
-    filter(treatment %in% whichgroups) %>%
-  ggplot(aes(x = treatment, fill = predictedstage)) +
+plotpredictions <- function(df, mysubstitle){
+  ggplot(df, aes(x = treatment, fill = predictedstage)) +
   geom_bar(position = position_fill(reverse = F))  +
@@
     theme_B3() +
-  theme(legend.position = "none") +
-  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none",
+        axis.text.x = element_blank()) +
+  scale_fill_manual(values = colorsnewgrouping) +
   labs(y = mysubstitle) 
-  return(p)
 }
 
-j <- plotfacetresults2(LDA.gon, levelstiming, "hypothalamus", 20) + labs(x = NULL) + 
-  theme(axis.text.x = element_blank()) +
-  theme(legend.position = "top") + guides(fill = guide_legend(nrow  = 1)) 
-k <- plotfacetresults2(LDA.pit, levelstiming, "pituitary", 17) + 
-  theme(strip.text = element_blank()) + labs(x = NULL) + theme(axis.text.x = element_blank())
-l <- plotfacetresults2(LDA.gon, levelstiming, "gonads", 15) + theme(strip.text = element_blank()) + 
-   labs(x = NULL)
+a <- plotpredictions(LDA.hyp, "hypothalamus") + theme(legend.position = "top") 
+b <- plotpredictions(LDA.pit, "pituitary") + theme(strip.text = element_blank())
+c <- plotpredictions(LDA.gon, "gonads") + theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.text = element_blank())
 
-plot_grid(j,k,l, nrow = 3, rel_heights = c(1.6,1,1.2))
-
-m <- plotfacetresults2(LDA.gon, levelsremoval, "hypothalamus", 20) + labs(x = NULL) + theme(axis.text.x = element_blank()) +
-    theme(legend.position = "top")  +  guides(fill = guide_legend(nrow  = 1))
-n <- plotfacetresults2(LDA.pit, levelsremoval, "pituitary", 17) + theme(strip.text = element_blank()) + labs(x = NULL) + theme(axis.text.x = element_blank())
-o <- plotfacetresults2(LDA.gon, levelsremoval, "gonads", 15) +
-  theme(strip.text = element_blank())  + labs(x = NULL)
-
-plot_grid(m,n,o, nrow = 3, rel_heights = c(1.6,1,1.2))
+plot_grid(a,b,c, rel_heights = c(1.8,1,1.5), nrow = 3)
 

### 2019-12-23 R/themes.R

@@ -69,6 +69,8 @@ alllevels3volcano <- c("control", "bldg", "lay", "inc.d3", "inc.d9",
                       "m.inc.d8" , "prolong" , "extend",
                       "NS")
 
+levelsnewgrouping <-  c("control" , "bldg",  "earlyinc", 
+                        "nearhatch",  "chickcare", "manip")
 
 ## comparisons
 
@@ -112,6 +114,12 @@ colorschar <-  c("control" = "#F8766D",
                          "n5"= "#DB72FB", 
                          "n9"= "#FF61C3")
 
+colorsnewgrouping <-  c("control" = "#F8766D", 
+                 "bldg"= "#D39200", 
+                 "earlyinc"= "#00C19F", 
+                 "nearhatch"= "#619CFF", 
+                 "chickcare"= "#FF61C3",
+                 "manip" = "#959595")
 
 colorscharmaip <-  c("control" = "#F8766D", 
                  "bldg"= "#D39200", 

### 2019-12-23 analysis/04_LDA.Rmd
faceted figrues

@@
   scale_fill_manual(values = colorscharmaip) +
-  labs(y = mysubstitle) 
+  labs(y = mysubstitle)  +
+    geom_hline(yintercept = 0.5,  size = 0.25, linetype = 2)
   return(p)

### 2019-12-23 analysis/04_LDA_v2.Rmd
faceted figrues

@@
 
-plotpredictions <- function(df, mysubstitle){
-  ggplot(df, aes(x = treatment, fill = predictedstage)) +
+plotpredictions <- function(df, whichlevels, mysubstitle){
+
+  p <- df %>%
+    filter(treatment %in% whichlevels)  %>%
+  ggplot(aes(x = treatment, fill = predictedstage)) +
   geom_bar(position = position_fill(reverse = F))  +
@@
   scale_fill_manual(values = colorsnewgrouping) +
-  labs(y = mysubstitle) 
+  labs(y = mysubstitle, x = NULL) +
+    geom_hline(yintercept = 0.5,  size = 0.25, linetype = 2)
+  return(p)
 }
 
-a <- plotpredictions(LDA.hyp, "hypothalamus") + theme(legend.position = "top") 
-b <- plotpredictions(LDA.pit, "pituitary") + theme(strip.text = element_blank())
-c <- plotpredictions(LDA.gon, "gonads") + theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.text = element_blank())
+a <- plotpredictions(LDA.hyp, alllevels, "hypothalamus") + theme(legend.position = "top") 
+b <- plotpredictions(LDA.pit, alllevels,"pituitary") + theme(strip.text = element_blank())
+c <- plotpredictions(LDA.gon, alllevels, "gonads") + theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.text = element_blank())
 
-plot_grid(a,b,c, rel_heights = c(1.8,1,1.5), nrow = 3)
+plot_grid(a,b,c, rel_heights = c(1.6,1,1.2), nrow = 3)
+
+a <- plotpredictions(LDA.hyp, levelstiming, "hypothalamus") + theme(legend.position = "top") 
+b <- plotpredictions(LDA.pit, levelstiming,"pituitary") + theme(strip.text = element_blank())
+c <- plotpredictions(LDA.gon, levelstiming, "gonads") + theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.text = element_blank())
+
+plot_grid(a,b,c, rel_heights = c(1.6,1,1.2), nrow = 3)
+
+a <- plotpredictions(LDA.hyp, levelsremoval, "hypothalamus") + theme(legend.position = "top") 
+b <- plotpredictions(LDA.pit, levelsremoval,"pituitary") + theme(strip.text = element_blank())
+c <- plotpredictions(LDA.gon, levelsremoval, "gonads") + theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.text = element_blank())
+
+plot_grid(a,b,c, rel_heights = c(1.6,1,1.2), nrow = 3)
 

### 2019-12-23 analysis/17_pheatmap.Rmd
first draft at group heatmap

---
title: "heatmap"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pheatmap)

source("../R/themes.R")

knitr::opts_chunk$set(fig.path = '../figures/pheatmap/', cache = TRUE,  message=F, comment=FALSE, warning=FALSE)
```

```{r, pheatmap, fig.height=6.5}
prepvsds <- function(pathtofile){

  df <- read.csv(pathtofile, row.names = 1)

  df <- tail(df, 5000)
  df <- head(df, 1000)

  df <- as.data.frame(t(df))
  df$V1 <- row.names(df)

  df$sex <- sapply(strsplit(as.character(df$V1),'\\_'), "[", 2)
  df$tissue <- sapply(strsplit(as.character(df$V1),'\\_'), "[", 3)
  df$treatment <- sapply(strsplit(as.character(df$V1),'\\_'), "[", 4)
  df$treatment <- sapply(strsplit(as.character(df$treatment), ".NYNO"), "[", 1)

  df$treatment <- ifelse(grepl("extend.hatch", df$treatment), "extend",
                   ifelse(grepl("inc.prolong", df$treatment), "prolong",
                   ifelse(grepl("m.hatch", df$treatment), "m.n2",
                                   df$treatment)))

  df$tissue <- ifelse(grepl("gonad", df$tissue), "gonads", df$tissue)

  df$sextissuetreatment <- paste(df$sex, df$tissue, df$treatment, sep = "_")

  newcolnames <- df$sextissuetreatment

  df$V1 <- NULL
  df$sex <- NULL
  df$tissue <- NULL
  df$treatment <- NULL 
  df$sextissuetreatment <- NULL

  df <- as.data.frame(t(df))

  colnames(df) <- newcolnames

  df <- sapply(unique(colnames(df)), function(i)
    rowMeans(df[,colnames(df) == i]))

  return(df)
}

vsd.hyp <- prepvsds("../results/06_hypallvsd.csv")
vsd.pit <- prepvsds("../results/06_pitallvsd.csv")
vsd.gon <- prepvsds("../results/06_gonallvsd.csv")

makelabels <- function(df){

  mylabels <- as.data.frame(colnames(df))
  mylabels$treatment <- sapply(strsplit(as.character(mylabels$`colnames(df)`),'\\_'), "[", 3)
  mylabels$sex <- sapply(strsplit(as.character(mylabels$`colnames(df)`),'\\_'), "[", 1)
  mylabels$tissue <- sapply(strsplit(as.character(mylabels$`colnames(df)`),'\\_'), "[", 2)
  row.names(mylabels) <- mylabels$`colnames(df)`
  mylabels$`colnames(df)` <- NULL
  return(mylabels)
}

hyplabels <- makelabels(vsd.hyp)
pitlabels <- makelabels(vsd.pit)
gonlabels <- makelabels(vsd.gon)

pheatmap(vsd.hyp, show_colnames = F, annotation_col = hyplabels, 
         annotation_colors = myannotationcolors, show_rownames = F)
pheatmap(vsd.pit, show_colnames = F, annotation_col = pitlabels, 
         annotation_colors = myannotationcolors, show_rownames = F)
pheatmap(vsd.gon, show_colnames = F, annotation_col = gonlabels, 
         annotation_colors = myannotationcolors, show_rownames = F)
```

### 2019-12-25 analysis/04_LDA.Rmd
now with sex as a class

@@
   df <- df %>% dplyr::select(one_of(savecols))
+
+  # keep only 100 genes fornow
+  #df <- head(df, 100)
+
   df <- as.data.frame(t(df))
@@
   train.data <- left_join(trainsamples, traindata) %>%
-    dplyr::select(-V1, -bird, -tissue, -group, -study, -sex)
+    dplyr::select(-V1, -bird, -tissue, -group, -study)
 
-  test.data <- testdata
+  test.data <- left_join(testsamples, testdata) %>%
+    dplyr::select(-V1, -bird, -tissue, -group, -study, -treatment)
+
+  testsamples <- testsamples
 
@@
   # Fit the model
-  model <- lda(treatment~ ., data = train.transformed)
+  model <- lda(treatment ~ ., data = train.transformed)
   # Make predictions
@@
   print("model accuracy")
-  print("predictions$class==test.transformed$treatment)")
-  print(mean(predictions$class==test.transformed$treatment))
+  #print("predictions$class==test.transformed$treatment)")
+ # print(mean(predictions$class==test.transformed$treatment))
 
@@
 ```{r LDAplots}
-library(kableExtra)
 
@@
 ```{r LDApredictions}
+library(kableExtra)
 

### 2019-12-25 analysis/04_LDA_v2.Rmd
now with sex as a class

@@
   train.data <- left_join(trainsamples, traindata) %>%
-    dplyr::select(-V1, -bird, -tissue, -group, -study, -sex, -treatment)
+    dplyr::select(-V1, -bird, -tissue, -group, -study, - treatment)
+
+  test.data <- left_join(testsamples, testdata) %>%
+    dplyr::select(-V1, -bird, -tissue, -group, -study, -treatment,  -newgrouping)
 

### 2019-12-25 analysis/04_LDA_v2.Rmd
new figs

@@
   train.data <- left_join(trainsamples, traindata) %>%
-    dplyr::select(-V1, -bird, -tissue, -group, -study, - treatment)
+    dplyr::select(-V1, -bird, -tissue, -group, -study, -treatment)
 
@@
     dplyr::select(-V1, -bird, -tissue, -group, -study, -treatment,  -newgrouping)
-
-  test.data <- testdata
 
@@
   # Fit the model
-  model <- lda(newgrouping~ ., data = train.transformed)
+  model <- lda(newgrouping ~ ., data = train.transformed)
   # Make predictions
@@
 plot_grid(a,b,c, rel_heights = c(1.6,1,1.2), nrow = 3)
-
 ```
@@
 ```{r LDApredictions}
-
 df1 <- LDA.hyp %>% distinct(treatment, predictedstage) %>%

### 2019-12-28 analysis/04_LDA_v3.Rmd
unsuccessful first attempt at incorporating stress data

---
title: "04_LDA_v3"
output: html_document
---

```{r setup, include=FALSE}
suppressMessages(library(tidyverse)) 
library(forcats)
library(tidyselect) # for vars_select

knitr::opts_chunk$set(echo = TRUE)
```

## stress data

```{r}
# genes info
geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>% dplyr::select(Name, entrezid)

# prep stress data
stress_coldata <- read_csv("../../austinCORT/data/03_colDataSTRESS.csv")
stress_countdata <- read.csv("../../austinCORT/data/03_countDataSTRESS.csv", row.names = 1)
head(stress_coldata)
head(stress_countdata)

subsetcoldata <- function(whichtissue){
  df <- stress_coldata %>% 
    mutate(tissue = fct_recode(tissue,
           "hypothalamus" = "hyp",
           "pituitary" = "pit",
           "gonads" = "gon")) %>% 
    filter(tissue == whichtissue) %>% 
    rename("V1" = "sample")  %>% 
    mutate(V1 = str_replace(V1, '.male', "_male"))  %>%
    mutate(V1 = str_replace(V1, '.f_male', "_female"))  %>% 
    mutate(bird = sapply(strsplit(as.character(V1),'\\_'), "[", 1))  %>% 
    mutate(group = sapply(strsplit(as.character(V1),'\\_'), "[", 2))  %>% 
    mutate(V1 = str_replace(V1, '_', "."))  %>%
    mutate(V1 = str_replace(V1, "r.r.x.R2XR.", "r.r.x.R2XR_")) %>%
    mutate(study = "stress") %>% 
    dplyr::select(V1, bird, sex, tissue, treatment, group, study)  %>% 
    droplevels()
  print(head(df))
  return(df)
}

stress_coldata_hyp <- subsetcoldata("hypothalamus")
stress_coldata_pit <- subsetcoldata("pituitary")
stress_coldata_gon <- subsetcoldata("gonads")

selectvsdstress <- function(stresscolData){

  savecols <- stress_coldata_hyp$V1
  savecols <- as.vector(savecols)

  df <- stress_countdata %>% dplyr::select(one_of(savecols)) 
  df$entrezid <- row.names(df)

  df <- left_join(geneinfo, df) %>% 
    dplyr::select(-entrezid) %>% 
    rename(gene = Name) %>%
    pivot_longer(-gene, names_to = "samples", values_to = "counts") %>%
    pivot_wider(
    names_from = samples, 
    values_from = counts,
    values_fn = list(counts = sum))  %>% 
  arrange(gene) %>% 
  filter(gene != "")

  df <- as.data.frame(df)
  row.names(df) <- df$gene
  df$gene <- NULL
  df <- as.data.frame(t(df))

  print(head(df))
  return(df)
}

stress_countdata_hyp <- selectvsdstress(stress_coldata_hyp)
stress_countdata_pit <- selectvsdstress(stress_coldata_pit)
stress_countdata_pit <- selectvsdstress(stress_coldata_gon)
```

## parentalcare data

```{r}
colData <- read.csv("../metadata/00_samples.csv", row.names = 1)

charHyp <- colData %>% filter(study == "charcterization", tissue == "hypothalamus") %>% droplevels()
charPit <- colData %>% filter(study == "charcterization", tissue == "pituitary") %>% droplevels()
charGon <- colData %>% filter(study == "charcterization", tissue == "gonad") %>% 
  mutate(tissue = fct_recode(tissue, "gonads" = "gonad")) %>% droplevels()

manipHyp <- colData %>% filter(study == "manipulation", tissue == "hypothalamus") %>% droplevels()
manipPit <- colData %>% filter(study == "manipulation", tissue == "pituitary") %>% droplevels()
manipGon <- colData %>% filter(study == "manipulation", tissue == "gonad") %>% 
  mutate(tissue = fct_recode(tissue, "gonads" = "gonad")) %>% droplevels()

selectvsd <- function(pathtofile, colData){

  df <- read.csv(pathtofile, row.names = 1)
  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols) 
  df <- df %>% dplyr::select(one_of(savecols))

  # keep only 100 genes fornow
  #df <- head(df, 100)

  df <- as.data.frame(t(df))
  df$V1 <- row.names(df)
  return(df)
}

vsd.hyp.train <- selectvsd("../results/06_hypallvsd.csv",  charHyp)
vsd.pit.train <- selectvsd("../results/06_pitallvsd.csv",  charPit)
vsd.gon.train <- selectvsd("../results/06_gonallvsd.csv",  charGon)

vsd.hyp.test <- selectvsd("../results/06_hypallvsd.csv",  manipHyp)
vsd.pit.test <- selectvsd("../results/06_pitallvsd.csv",  manipPit)
vsd.gon.test <- selectvsd("../results/06_gonallvsd.csv",  manipGon)

hypcols <- names(vsd.hyp.train)
pitcols <- names(vsd.pit.test)
goncols <- names(vsd.gon.test)

```

## combine stress and sex

```{r}
stress_countdata_hyp <- stress_countdata_hyp %>% tidyselect::vars_select(hypcols, strict = F)

head(stress_countdata_hyp)
stress_countdata_pit  
stress_countdata_pit  
```

### 2019-12-30 analysis/16_PRL.Rmd
new PRL and BRCA figures

--- 2019-12-07 analysis/16_PRL.Rmd
@@
     droplevels() %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-    geom_boxplot() +
-  ylim(c(-10,120)) + 
-    theme_B3() +
-  #theme(axis.text.x = element_blank()) + 
-  theme(legend.position = "none") + 
-  labs(x = "parental care stages", y = "prolactin (ng/mL)") +
-  annotation_custom(control, ymin = -20, ymax = 0, xmin = -7.5) +
-  annotation_custom(bldg, ymin = -20, ymax = 0, xmin = -5.5) +
-  annotation_custom(lay, ymin = -20, ymax = 0, xmin = -3.5) +
-  annotation_custom(inc, ymin = -20, ymax = 0, xmin = -1.5) +
-  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 0.5) +
-  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 2.5) +
-  annotation_custom(hatch, ymin = -20, ymax = 0, xmin = 4.5) +
-  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 6.5) +
-  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 8.5) 
-p1
-
-p2 <-  ggplot(meanprolactin, aes(treatment, m)) +
-  geom_errorbar(aes(ymin=m-se, ymax=m+se, color = treatment), width=.1) +
-   geom_image(aes(image=iconpath), size=.1)  +
-  labs(x = "parental care stages", y = "prolactin (ng/mL)") +
-  theme_B3() +
-  theme(legend.position = "none")
-
-p2
-
-p3 <- ggplot(meanprolactin, aes(treatment, m)) +
-   geom_image(aes(image=music), size=.12)  +
-  labs(x = "increasing time >>", y = "prolactin (ng/mL)")  +
-  mytheme() +
-  ylim(c(-15,80)) +
-  annotation_custom(control, ymin = -20, ymax = 0, xmin = -7.5) +
-  annotation_custom(bldg, ymin = -20, ymax = 0, xmin = -5.5) +
-  annotation_custom(lay, ymin = -20, ymax = 0, xmin = -3.5) +
-  annotation_custom(inc, ymin = -20, ymax = 0, xmin = -1.5) +
-  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 0.5) +
-  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 2.5) +
-  annotation_custom(hatch, ymin = -20, ymax = 0, xmin = 4.5) +
-  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 6.5) +
-  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 8.5) +
-  geom_hline(yintercept=0) +
-  geom_hline(yintercept=20) +
-  geom_hline(yintercept=40) +
-  geom_hline(yintercept=60) +
-  geom_hline(yintercept=80) 
-p3
-
-p4 <- hormones %>% 
-    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
-    droplevels() %>% 
   ggplot(aes(x = treatment, y = plasma_conc)) +
@@
   ylim(c(-10,120)) 
-p4
+p1
+```
 
-plot_grid(p1,p4,p2,p3)
+```{r candidatevsds}
+colData.pit <-  read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
+colData.pit <- colData.pit %>% filter(tissue == "pituitary")
 
+vsd.pit <- read.csv("../results/06_pitallvsd.csv", header = T, row.names = 1)
+
+savecols <- as.character(colData.pit$V1) 
+savecols <- as.vector(savecols) 
+vsd.pit <- vsd.pit %>% dplyr::select(one_of(savecols)) 
+vsd.pit$genes <- row.names(vsd.pit)
+
+candidategenes <- c("PRL", "BRCA1")
+candidates.pit <- vsd.pit %>% dplyr::filter(genes %in% candidategenes)
+candidates.pit <- as.data.frame(candidates.pit)
+row.names(candidates.pit) <- candidates.pit$genes
+candidates.pit$genes <- NULL
+candidates.pit <- as.data.frame(t(candidates.pit))
+candidates.pit$V1 <- row.names(candidates.pit) 
+candidates.pit <- left_join(colData.pit, candidates.pit)
+candidates.pit$treatment <- factor(candidates.pit$treatment, levels = alllevels)
+tail(candidates.pit)
 ```
@@
 ```{r PRL.pit}
-vsd.pit <- readvsd("../results/04_vsd_pit.csv")
-colData.pit <- readcolData("../results/04_colData_pit.csv")
-
-geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>%  dplyr::select(Name, geneid, entrezid) %>% arrange(Name)
-candidategenes <- c("PRL")
-candidates.pit <- selectcandidatevsds(candidategenes, vsd.pit, colData.pit)
-
 meanPRL <- candidates.pit %>% 
     droplevels() %>% 
-  dplyr::group_by(treatment) %>%
+  dplyr::group_by(treatment, sex) %>%
   dplyr::summarise(m = mean(PRL), 
                    se = sd(PRL)/sqrt(length(PRL))) %>%
-  dplyr::mutate(m = round(m,1)) 
+  dplyr::mutate(m = round(m,1))  %>%
+  dplyr::filter(treatment %in% charlevels)
 meanPRL <- left_join(meanPRL, birds)
-meanPRL$treatment <- factor(meanPRL$treatment, levels = alllevels)
+meanPRL$treatment <- factor(meanPRL$treatment, levels = charlevels)
 meanPRL
 
-p5 <- candidates.pit %>% 
-  ggplot(aes(x = treatment, y = PRL, fill = treatment)) +
-    geom_boxplot() +
-    theme_B3() +
-  theme(legend.position = "none") + 
-  labs(x = "increasing time >>", y = "prolactin (ng/mL)") +
-  annotation_custom(control, ymin = 12, ymax = 14, xmin = -7.8) +
-  annotation_custom(bldg, ymin = 12, ymax = 14, xmin = -5.8) +
-  annotation_custom(lay, ymin = 12, ymax = 14, xmin = -3.8) +
-  annotation_custom(inc, ymin = 12, ymax = 14, xmin = -1.8) +
-  annotation_custom(inc, ymin = 12, ymax = 14, xmin = 0.6) +
-  annotation_custom(inc, ymin = 12, ymax = 14, xmin = 2.4) +
-  annotation_custom(hatch, ymin = 12, ymax = 14, xmin = 4.4) +
-  annotation_custom(nestling, ymin = 12, ymax = 14, xmin = 6.4) +
-  annotation_custom(nestling, ymin = 12, ymax = 14, xmin = 8.4)  +
-  ylim(c(13,24)) 
-p5
-
-p6 <- ggplot(candidates.pit, aes(x = treatment, y = PRL)) + 
+p2 <- candidates.pit %>%
+  dplyr::filter(treatment %in% charlevels) %>%
+  ggplot(aes(x = treatment, y = PRL)) + 
     geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) + 
@@
   scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
-    labs(y = "PRL expression in the pituitary", x = "parental care stage") +
-  annotation_custom(control, ymin = 12, ymax = 14, xmin = -7.8) +
-  annotation_custom(bldg, ymin = 12, ymax = 14, xmin = -5.8) +
-  annotation_custom(lay, ymin = 12, ymax = 14, xmin = -3.8) +
-  annotation_custom(inc, ymin = 12, ymax = 14, xmin = -1.8) +
-  annotation_custom(inc, ymin = 12, ymax = 14, xmin = 0.6) +
-  annotation_custom(inc, ymin = 12, ymax = 14, xmin = 2.4) +
-  annotation_custom(hatch, ymin = 12, ymax = 14, xmin = 4.4) +
-  annotation_custom(nestling, ymin = 12, ymax = 14, xmin = 6.4) +
-  annotation_custom(nestling, ymin = 12, ymax = 14, xmin = 8.4)  + 
-  ylim(c(13,24)) +
-  guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) 
-p6
+  labs(y = "PRL", x = "increasing time >>", subtitle = "Pituitary" ) +
+  guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) +
+  theme(axis.title.y  = element_text(face = "italic"))  +
+  annotation_custom(control, ymin = 13.5, ymax = 15, xmin = -7.5) +
+  annotation_custom(bldg, ymin = 13.5, ymax = 15, xmin = -5.5) +
+  annotation_custom(lay, ymin = 13.5, ymax = 15, xmin = -3.5) +
+  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = -1.5) +
+  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = 0.5) +
+  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = 2.5) +
+  annotation_custom(hatch, ymin = 13.5, ymax = 15, xmin = 4.5) +
+  annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 6.5) +
+  annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 8.5)   
+p2
 
-p7 <- ggplot(meanPRL, aes(treatment, m)) +
-  geom_errorbar(aes(ymin=m-se, ymax=m+se, color = treatment), width=.1) +
-   geom_image(aes(image=iconpath), size=.12)  +
-  labs(x = "increasing time >>", y = "Pituitary PRL expression") +
+```
+
+```{r PRLBRCA1}
+meanBRCA <- candidates.pit %>% 
+    droplevels() %>% 
+  dplyr::group_by(treatment, sex) %>%
+  dplyr::summarise(m = mean(BRCA1), 
+                   se = sd(BRCA1)/sqrt(length(BRCA1))) %>%
+  dplyr::mutate(m = round(m,1)) %>%
+  dplyr::filter(treatment %in% charlevels)
+meanBRCA <- left_join(meanBRCA, birds)
+meanBRCA$treatment <- factor(meanBRCA$treatment, levels = alllevels)
+meanBRCA
+
+p3 <- candidates.pit %>% 
+  dplyr::filter(treatment %in% charlevels) %>%
+  ggplot(aes(x = treatment, y = BRCA1, fill = treatment, color = sex, alpha = sex)) +
+    geom_boxplot() +
     theme_B3() +
-  theme(legend.position = "none")
-p7
+      scale_alpha_manual(values = c(0.75,1)) +
+  scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+  theme(legend.position = "none") + 
+  labs(x = "increasing time >>", y = "BRCA1", subtitle = "Pituitary") +
+  annotation_custom(control, ymin = 6.65, ymax = 7, xmin = -7.8) +
+  annotation_custom(bldg, ymin = 6.65, ymax = 7, xmin = -5.8) +
+  annotation_custom(lay, ymin = 6.65, ymax = 7, xmin = -3.8) +
+  annotation_custom(inc, ymin = 6.65, ymax = 7, xmin = -1.8) +
+  annotation_custom(inc, ymin = 6.65, ymax = 7, xmin = 0.6) +
+  annotation_custom(inc, ymin = 6.65, ymax = 7, xmin = 2.4) +
+  annotation_custom(hatch, ymin = 6.65, ymax = 7, xmin = 4.4) +
+  annotation_custom(nestling, ymin = 6.65, ymax = 7, xmin = 6.4) +
+  annotation_custom(nestling, ymin = 6.65, ymax = 7, xmin = 8.4) +
+  ylim(6.8,8.5) +
+  guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) +
+  theme(legend.position = c(0.85,0.2), legend.direction = "horizontal") + 
+  theme(axis.title.y  = element_text(face = "italic"))
+p3
 
-p8 <- ggplot(meanPRL, aes(treatment, m)) +
-   geom_image(aes(image=music), size=.12)  +
-  labs(x = "parental care stages", y = "Pituitary PRL expression") +
-  mytheme() +
-  ylim(c(15,22)) +
-  annotation_custom(control, ymin = 15, ymax = 16.5, xmin = -7.5) +
-  annotation_custom(bldg, ymin = 15, ymax = 16.5, xmin = -5.5) +
-  annotation_custom(lay, ymin = 15, ymax = 16.5, xmin = -3.5) +
-  annotation_custom(inc, ymin = 15, ymax = 16.5, xmin = -1.5) +
-  annotation_custom(inc, ymin = 15, ymax = 16.5, xmin = 0.5) +
-  annotation_custom(inc, ymin = 15, ymax = 16.5, xmin = 2.5) +
-  annotation_custom(hatch, ymin = 15, ymax = 16.5, xmin = 4.5) +
-  annotation_custom(nestling, ymin = 15, ymax = 16.5, xmin = 6.5) +
-  annotation_custom(nestling, ymin = 15, ymax = 16.5, xmin = 8.5)   +
-  geom_hline(yintercept=17) +
-  geom_hline(yintercept=18) +
-  geom_hline(yintercept=19) +
-  geom_hline(yintercept=20) +
-  geom_hline(yintercept=21)
-
-p8
-
-plot_grid(p5,p6,p7,p8)
 ```
@@
 
-```{r manipdata}
-vsd.pit <- readvsd("../results/06_pitallvsd.csv")
-vsd.pit$entrezid <- row.names(vsd.pit)
+```{r PRL.pit.maip}
 
-colData.pit <- read_csv("../metadata/00_samples.csv") %>% 
-  filter(tissue == "pituitary") %>% 
-  mutate(sample = V1)
-
-geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>%  
-  dplyr::select(Name, geneid, entrezid) %>% arrange(Name)
-
-candidategenes <- c("PRL")
-candidates.pit <- selectcandidatevsds(candidategenes, vsd.pit, colData.pit)
-candidates.pit$treatment <- factor(candidates.pit$treatment, levels = alllevels)
-
-meanPRL <- candidates.pit %>% 
-    droplevels() %>% 
-  dplyr::group_by(treatment) %>%
-  dplyr::summarise(m = median(PRL), 
-                   se = sd(PRL)/sqrt(length(PRL))) %>%
-  dplyr::mutate(m = round(m,1)) 
-meanPRL <- left_join(meanPRL, birds)
-meanPRL$treatment <- factor(meanPRL$treatment, levels = alllevels)
-meanPRL
-```
-
-```{r PRL.pit.maip}
-p10 <- ggplot(meanPRL, aes(treatment, m)) +
-   geom_errorbar(aes(ymin=m-se, ymax=m+se, color = treatment), width=.1) +
-   geom_image(aes(image = iconpath), size = 0.06)  +
-   geom_image(aes(image = music , y = m + 0.5), size = 0.04) +
-   labs(x = "parental care stages", y = "Pituitary PRL") +
-   mytheme() +
-   ylim(c(16,22)) +
-   theme(axis.text.x = element_text(angle = 45)) +
-  theme(legend.position = "none") +
-  scale_color_manual(values = colorscharmaip2) +
-  geom_hline(yintercept=16.8) +
-  geom_hline(yintercept=17.8) +
-  geom_hline(yintercept=18.8) +
-  geom_hline(yintercept=19.8) +
-  geom_hline(yintercept=20.8)  
-p10
-
-p11 <- ggplot(candidates.pit, aes(x = treatment, y = PRL)) + 
-       geom_smooth(aes(colour = sex)) +
-    geom_boxplot(aes(fill = treatment)) +
+p4 <- ggplot(candidates.pit, aes(x = treatment, y = PRL)) + 
+    geom_boxplot(aes(fill = treatment, alpha = sex)) +
     scale_alpha_manual(values = c(0.75,1)) +
@@
    scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
-    labs(y = "Pituitary PRL", x = NULL) +
+  labs(x = "increasing time >>", y = "PRL", subtitle = "Pituitary") +
   scale_fill_manual(values = colorscharmaip2) +
@@
   annotation_custom(nestling, ymin = 13, ymax = 14, xmin = 13.4) +
-  annotation_custom(nestling, ymin = 13, ymax = 14, xmin = 15.4) 
-p11
+  annotation_custom(nestling, ymin = 13, ymax = 14, xmin = 15.4) +
+    theme(axis.title.y  = element_text(face = "italic"))
+p4
 
-p12 <- candidates.pit %>%
-  filter(treatment %in% c("inc.d9", "m.inc.d8", "hatch")) %>%
-    ggplot(aes(x = treatment, y = PRL)) + 
-    geom_boxplot(aes(fill = treatment, color = sex)) + 
+```
+
+```{r BRCA.pit.maip}
+
+p5 <- ggplot(candidates.pit, aes(x = treatment, y = BRCA1)) + 
+    geom_boxplot(aes(fill = treatment, alpha = sex)) +
     scale_alpha_manual(values = c(0.75,1)) +
@@
    scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
-    labs(y = "Pituitary PRL", x = NULL) +
-  ylim(c(13,22.5)) +
+  labs(x = "increasing time >>", y = "BRCA1", subtitle = "Pituitary") +
   scale_fill_manual(values = colorscharmaip2) +
-  annotation_custom(control, ymin = 13, ymax = 14, xmin = -1.6) +
-  annotation_custom(maniphatch, ymin = 13, ymax = 14, xmin = 0.4) +
-  annotation_custom(hatch, ymin = 13, ymax = 14, xmin = 2.4)
-p12
+  ylim(6.8,8.5) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  annotation_custom(control, ymin = 13, ymax = 14, xmin = -14.8) +
+  annotation_custom(bldg, ymin = 13, ymax = 14, xmin = -12.8) +
+  annotation_custom(lay, ymin = 13, ymax = 14, xmin = -10.8) +
+  annotation_custom(inc, ymin = 13, ymax = 14, xmin = -8.8) +
+  annotation_custom(removeegg, ymin = 13, ymax = 14, xmin = -6.8) +
+  annotation_custom(inc, ymin = 13, ymax = 14, xmin = -4.8) +
+  annotation_custom(maniphatch, ymin = 13, ymax = 14, xmin = -2.8) +
+  annotation_custom(removeegg, ymin = 13, ymax = 14, xmin = -0.8) +
+  annotation_custom(inc, ymin = 13, ymax = 14, xmin = 1.4) +
+  annotation_custom(removeegg, ymin = 13, ymax = 14, xmin = 3.4)  +
+  annotation_custom(manipinc, ymin = 13, ymax = 14, xmin = 5.4) +
+  annotation_custom(hatch, ymin = 13, ymax = 14, xmin = 7.4) +
+  annotation_custom(removechick, ymin = 13, ymax = 14, xmin = 9.4) +
+  annotation_custom(maniphatch, ymin = 13, ymax = 14, xmin = 11.4) +
+  annotation_custom(nestling, ymin = 13, ymax = 14, xmin = 13.4) +
+  annotation_custom(nestling, ymin = 13, ymax = 14, xmin = 15.4) +
+    theme(axis.title.y  = element_text(face = "italic"))
+p5
 
-p13 <- candidates.pit %>%
-  filter(treatment %in% c("inc.d17", "prolong", "hatch")) %>%
-    ggplot(aes(x = treatment, y = PRL)) + 
-    geom_boxplot(aes(fill = treatment, color = sex)) + 
-    scale_alpha_manual(values = c(0.75,1)) +
-    theme_B3() +
-   theme(legend.position = "none", legend.direction = "horizontal") + 
-   scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
-    labs(y = "Pituitary PRL", x = NULL) +
-  ylim(c(13,22.5)) +
-  scale_fill_manual(values = colorscharmaip2)   +
-  annotation_custom(control, ymin = 13, ymax = 14, xmin = -1.6) +
-  annotation_custom(manipinc, ymin = 13, ymax = 14, xmin = 0.4) +
-  annotation_custom(hatch, ymin = 13, ymax = 14, xmin = 2.4)
-p13
-
-p14 <- candidates.pit %>%
-  filter(treatment %in% c("hatch", "extend", "n5")) %>%
-    ggplot(aes(x = treatment, y = PRL)) + 
-    geom_boxplot(aes(fill = treatment, color = sex)) + 
-    scale_alpha_manual(values = c(0.75,1)) +
-    theme_B3() +
-   theme(legend.position = "none", legend.direction = "horizontal") + 
-   scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
-    labs(y = "Pituitary PRL", x = NULL) +
-  ylim(c(13,22.5)) +
-  scale_fill_manual(values = colorscharmaip2)   +
-  annotation_custom(control, ymin = 13, ymax = 14, xmin = -1.6) +
-  annotation_custom(maniphatch, ymin = 13, ymax = 14, xmin = 0.4) +
-  annotation_custom(hatch, ymin = 13, ymax = 14, xmin = 2.4)
-p14
-
-plot_grid(p12,p13 + theme(axis.text.y = element_blank(), axis.title.y = element_blank()),
-          p14 + theme(axis.text.y = element_blank(), axis.title.y = element_blank()), 
-          nrow = 1, rel_widths = c(1.2,1,1))
+ggplot(candidates.pit, aes(x = PRL, y = BRCA1)) +
+  geom_point(aes(alpha = sex,  color = 	treatment)) +
+  geom_smooth(method = "loess", se=T, color = "darkgrey") +
+  scale_color_manual(values = colorscharmaip2)  +
+  scale_alpha_manual(values = c(0.5,1)) +
+  labs(subtitle = "Pituitary") +
+  theme(axis.title  = element_text(face = "italic"),
+        legend.title =  element_blank())
 
@@
 candidates.pit$bird_id <- candidates.pit$bird
-
+candidates.pit$X1 <- candidates.pit$V1
 geneshormones <- left_join(candidates.pit, hormones) %>%
@@
   drop_na()
-head(geneshormones)
+
+(geneshormones$bird_id)
+(hormones$bird_id)
 
@@
 
+```{r music}
+
+p8 <- ggplot(meanPRL, aes(treatment, m)) +
+   geom_image(aes(image=music), size=.12)  +
+  labs(x = "parental care stages", y = "PRL", subtitle = "Pituitary") +
+  mytheme() +
+  ylim(c(15,22)) +
+  annotation_custom(control, ymin = 15, ymax = 16.5, xmin = -7.5) +
+  annotation_custom(bldg, ymin = 15, ymax = 16.5, xmin = -5.5) +
+  annotation_custom(lay, ymin = 15, ymax = 16.5, xmin = -3.5) +
+  annotation_custom(inc, ymin = 15, ymax = 16.5, xmin = -1.5) +
+  annotation_custom(inc, ymin = 15, ymax = 16.5, xmin = 0.5) +
+  annotation_custom(inc, ymin = 15, ymax = 16.5, xmin = 2.5) +
+  annotation_custom(hatch, ymin = 15, ymax = 16.5, xmin = 4.5) +
+  annotation_custom(nestling, ymin = 15, ymax = 16.5, xmin = 6.5) +
+  annotation_custom(nestling, ymin = 15, ymax = 16.5, xmin = 8.5)   +
+  geom_hline(yintercept=17) +
+  geom_hline(yintercept=18) +
+  geom_hline(yintercept=19) +
+  geom_hline(yintercept=20) +
+  geom_hline(yintercept=21)
+
+p8
+
+```
+
 ```{r}

### 2019-12-30 analysis/00_datawrangling.Rmd
updated countData and wgcna script

@@
 library(tidyverse)
+
+source("../R/themes.R")
 ```
@@
                         sep = ",", row.names = NULL)
-head(kallistodata)
 
@@
   filter(gene != "")
-head(countData)
 
@@
 countData <- round(countData) #round all value to nearest 1s place
-head(countData)
+head(countData[13:15])
 
@@
 
-colData$study <- factor(colData$study, 
-                         levels = c("charcterization", "manipulation"))
+colData$study <- factor(colData$study, levels = c("charcterization", "manipulation"))
 
@@
 colData_manipluation <- colData %>%
-  dplyr::filter(grepl('m.|extend|prolong', treatment)) %>%
+  dplyr::filter(treatment %in% maniplevels1) %>%
   droplevels()
@@
 colData_characterization <- colData %>%
-  dplyr::filter(!grepl('m.|extend|prolong', treatment)) %>%
+  dplyr::filter(treatment %in% charlevels) %>%
   droplevels()

### 2019-12-30 analysis/08_WGCNA_1.R

@@ -17,6 +17,7 @@ library(forcats)
 
 # read the count data
 countData <- read.csv("../results/00_countData_characterization.csv", row.names = 1)
+head(countData[1:2])
 
 # read the sample meta data or column data
 colData <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, stringsAsFactors = T)
@@ -32,8 +33,8 @@ colData <- colData %>%
                              "hyp" = "hypothalamus",
                              "gon" = "gonad"),
          hypothesis = fct_recode(treatment,
-                                 "anticipation" = "control",
-                                 "anticipation" = "bldg",
+                                 "non.parental" = "control",
+                                 "non.parental" = "bldg",
                                  "incubation" = "lay",
                                  "incubation" = "inc.d3",
                                  "incubation" = "inc.d9",
@@ -51,9 +52,6 @@ colData$sample <- paste(colData$treatment, colData$sex, colData$tissue,  colData
 # set row and count names to be the same
 row.names(colData) <- colData$sample
 colnames(countData) <- colData$sample
-
-head(colData)
-head(countData)
 
 # create new grouping for subsets
 colData$sextissue <-  as.factor(paste(colData$sex, colData$tissue, sep = "."))
@@ -105,12 +103,10 @@ gsg$allOK
 
 if (!gsg$allOK)
 {
-  # Optionally, print the gene and sample names that were removed:
-  #if (sum(!gsg$goodGenes)>0) 
-  #  printFlush(paste("Removing genes:", paste(names(datExpr0)[!gsg$goodGenes], collapse = ", ")));
-  #if (sum(!gsg$goodSamples)>0) 
-  #  printFlush(paste("Removing samples:", paste(rownames(datExpr0)[!gsg$goodSamples], collapse = ", ")));
-  # Remove the offending genes and samples from the data:
+  if (sum(!gsg$goodGenes)>0)
+    printFlush(paste("Removing genes:", paste(names(datExpr0)[!gsg$goodGenes], collapse = ", ")))
+  if (sum(!gsg$goodSamples)>0)
+    printFlush(paste("Removing samples:", paste(rownames(datExpr0)[!gsg$goodSamples], collapse = ", ")))
   datExpr0 = datExpr0[gsg$goodSamples, gsg$goodGenes]
 }
 

### 2019-12-30 analysis/07_hormones.Rmd
new hormone analysis

@@
 ```
+
+## for rechelle
+
+```{r forrechelle}
+
+rechelle <- hormones %>%
+  dplyr::filter(hormone %in% c("prolactin", "cort"),
+                treatment %in% c("lay", "inc.d9", "hatch", "n9")) %>%
+  dplyr::select(-icons, -music, -iconpath, -okay)
+
+ggplot(rechelle, aes(x = treatment, y = plasma_conc, fill = sex)) +
+    geom_boxplot() + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+         legend.position = "none",
+          strip.text = element_blank()) +
+  facet_wrap(~hormone, scales = "free_y") +
+  theme_B3()
+
+write.csv(rechelle, "../results/07_forrechelle.csv")
+```

### 2019-12-31 analysis/07_hormones.Rmd
better figures, with and without stats

--- 2019-12-12 analysis/07_hormones.Rmd
@@
 library(lubridate)
+library(ggsignif)
 
@@
 ```
+
+## for rechelle
+
+```{r forrechelle, fig.height= 4, fig.width=6}
+
+rechelle <- hormones %>%
+  dplyr::filter(hormone %in% c("prolactin", "cort"),
+                treatment %in% c("lay", "inc.d9", "hatch", "n9")) %>%
+  dplyr::select(-icons, -music, -iconpath, -okay)
+
+rechelle$hormone <- factor(rechelle$hormone, levels = c("prolactin", "cort"))
+
+prl.rechelle <- rechelle %>% filter(hormone == "prolactin")   %>%  droplevels()
+cort.rechelle <- rechelle %>% filter(hormone == "cort")   %>%  droplevels()
+
+TukeyHSD(aov(plasma_conc ~  treatment, data=prl.rechelle))
+TukeyHSD(aov(plasma_conc ~  treatment, data=cort.rechelle))
+
+rechelleplot <- function(mydf, myylab){
+
+  p <- ggplot(mydf, aes(x = treatment, y = plasma_conc, fill = sex)) +
+  geom_point(aes(color = sex), position = position_dodge(0.9), size = 2)+
+  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
+  scale_color_manual(name = "Sex", values = c("gray50", "maroon"))+
+  scale_fill_manual(name = "Sex", values = c("gray50", "maroon"))+
+
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+         legend.position = "none",
+          strip.text = element_blank()) +
+    theme_classic()+
+  theme(text = element_text(size= 15)) +
+    labs(x = "Stage", y = myylab) 
+  return(p)
+
+}
+
+#PRL
+rechelleplot(prl.rechelle, "Prolactin (ng/mL)")
+rechelleplot(prl.rechelle, "Prolactin (ng/mL)") + 
+  scale_y_continuous(breaks = c(0,25,50,75,100, 125), limits = c(-20,125)) +
+  geom_signif(comparisons=list(c("lay", "hatch")), annotations= "p < 0.001", y_position = 100, tip_length = 0.05, vjust = 0, color= "black" , textsize = 5 ) +
+  geom_signif(comparisons=list(c("lay", "n9")), annotations= "p < 0.001", y_position = 110, tip_length = 0.05, vjust = 0, color= "black" , textsize = 5 ) +
+  geom_signif(comparisons=list(c("hatch", "n9")), annotations= "p < 0.001", y_position = 120, tip_length = 0.05, vjust = 0, color= "black" , textsize = 5 ) +
+  geom_signif(comparisons=list(c("inc.d9", "hatch")), annotations= "p < 0.001", y_position = -5, tip_length = -0.05, vjust = 2, color= "black" , textsize = 5 ) +
+  geom_signif(comparisons=list(c("inc.d9", "n9")), annotations= "p < 0.001", y_position = -15, tip_length = -0.05, vjust = 2, color= "black" , textsize = 5 )
+
+#CORT
+rechelleplot(cort.rechelle, "Corticosterone (ng/mL)")
+
+```

### 2019-12-31 analysis/16_PRL.Rmd
new PRL BRCA1 fig

@@
 
+candidates.pit %>%
+  filter(treatment %in% charlevels) %>%
+ggplot( aes(x = PRL, y = BRCA1)) +
+  geom_point(aes(alpha = sex,  color = 	treatment)) +
+  geom_smooth(method = "loess", se=T, color = "darkgrey") +
+  scale_color_manual(values = colorscharmaip2)  +
+  scale_alpha_manual(values = c(0.5,1)) +
+  labs(subtitle = "Pituitary") +
+  theme(axis.title  = element_text(face = "italic"),
+        legend.title =  element_blank())
+
 ```

### 2019-12-31 analysis/08_wgcna.Rmd
now with PRL-related genes
no longer a function
update wgcna markdown file

@@
 knitr::opts_chunk$set(echo = FALSE, fig.path = '../figures/wgcna/')
-
-source("08_WGCNA_1.R")
-
 ```
@@
 The first tutorial focuses on data input, cleaning and pre-processing; outlier removal; clustering of samples by their Euclidean distance; a comparison of sample cluster to sample meta data.
-I put pretty much the whole script into a single function that I source with [`08_WGCNA_1.R`](https://github.com/macmanes-lab/DoveParentsRNAseq/blob/master/analysis/08_WGCNA_1.R).
 
-For the following figures, the top panel is a dendrogram showing Euclidean distance of sample.
-Because the labeling of samples is hard to read, the colored rows can be used to visualize traits.
-Here, both sex and tissues are colored in a white to red scaled with the following pattern:
+The second tutorial organized genes into modules
 
-* sex: females (white), males (red), 
-* tissue: gonad (white), hypothalamus (pink), pituitary (red).
-
-## pituitary (not working)
-
-```{r wgcna-pit, message=FALSE, warning=FALSE, cache=TRUE}
-subsetWGCNA(c("F.pit", "M.pit"), 2000000, "Pituitary dendrogram with trait heatmap")
-```
+- "../analysis/08_WGCNA_1.R" 
+- "../analysis/08_WGCNA_2.R"

### 2019-12-31 analysis/06_DESeq2_ALL.Rmd
new volcano plots with gene expression counts

--- 2019-12-20 analysis/06_DESeq2_ALL.Rmd
@@
           d + theme(axis.title = element_blank(), axis.text = element_blank()), 
+          e + theme(axis.title = element_blank(), axis.text = element_blank()),
           f + theme(axis.title = element_blank(), axis.text = element_blank()),
@@
           i + theme(axis.title = element_blank(), axis.text = element_blank()),
-          nrow = 1, rel_widths = c(1.2,1,1,1,1,1,1))
+          nrow = 1, rel_widths = c(1.4,1,1,1,1,1,1,1,1))
 
@@
           m + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          NULL, nrow = 1, rel_widths = c(1.2, 1, 1,1,1,1,1))
+          NULL, nrow = 1, rel_widths = c(1.4, 1, 1,1,1,1,1))
 
@@
           s + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-            nrow = 1, rel_widths = c(1.2,1, 0.5, 1,1,0.5, 1,1))
+            nrow = 1, rel_widths = c(1.4,1, 0.5, 1, 1, 0.5, 1, 1))
 
@@
 ```
+
+```{r volcanos-pit-char, fig.height= 2.5, fig.width=6}
+char2 <- plot_grid(
+          b ,
+          c + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
+          d + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          e + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          f + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          g + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          h + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          nrow = 1, rel_widths = c(1.4,1,1,1,1,1,1,1))
+
+plot_grid(charicons, char2, nrow = 2, rel_heights = c(0.9,1))
+```

### 2019-12-31 analysis/06b_DESeq2_ALL_gonad.Rmd
new volcano plots with gene expression counts

@@
           i + theme(axis.title = element_blank(), axis.text = element_blank()),
-          nrow = 1, rel_widths = c(1.3,1,1,1,1,1,1))
+          nrow = 1, rel_widths = c(1.4,1,1,1,1,1,1,1))
 
@@
           m + theme(axis.title = element_blank(), axis.text = element_blank()), 
-          NULL, nrow = 1, rel_widths = c(1.3, 1, 1,1,1,1,1))
+          NULL, nrow = 1, rel_widths = c(1.4, 1, 1,1,1,1,1))
 
@@
           s + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-            nrow = 1, rel_widths = c(1.3,1, 0.5, 1, 1, 0.5, 1, 1))
+            nrow = 1, rel_widths = c(1.4,1, 0.5, 1, 1, 0.5, 1, 1))
 
@@
 
-plot_grid(charicons, char, removeicons, remove, timingicons, timing, nrow = 6, rel_heights = c(0.9,1,0.9,1,0.9,1.2))
+plot_grid(charicons, char, removeicons, remove, timingicons, timing, 
+          nrow = 6, rel_heights = c(0.9,1,0.9,1,0.9,1.2))
+
 ```
+
+```{r volcanos-gon-char, fig.height= 2.5, fig.width=6}
+char2 <- plot_grid(
+          b ,
+          c + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
+          d + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          e + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          f + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          g + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          h + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          nrow = 1, rel_widths = c(1.4,1,1,1,1,1,1,1))
+
+plot_grid(charicons, char2, nrow = 2, rel_heights = c(0.9,1))
+
+```

### 2019-12-31 analysis/06c_DESeq2_ALL_hypothalamus.Rmd
new volcano plots with gene expression counts

--- 2019-12-20 analysis/06c_DESeq2_ALL_hypothalamus.Rmd
@@
           i + theme(axis.title = element_blank(), axis.text = element_blank()),
-          nrow = 1, rel_widths = c(1.2,1,1,1,1,1,1))
+          nrow = 1, rel_widths = c(1.4,1,1,1,1,1,1,1))
 
@@
           s + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
-            nrow = 1, rel_widths = c(1.2,1, 0.5, 1,1,0.5, 1,1))
+            nrow = 1, rel_widths = c(1.4,1, 0.5, 1,1,0.5, 1,1))
 
@@
 ```
+
+```{r volcanos-hyp-char, fig.height= 2.5, fig.width=6}
+char2 <- plot_grid(
+          b ,
+          c + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
+          d + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          e + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          f + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          g + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          h + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+          nrow = 1, rel_widths = c(1.4,1,1,1,1,1,1,1))
+
+plot_grid(charicons, char2, nrow = 2, rel_heights = c(0.9,1))
+```

### 2019-12-31 analysis/16_PRL.Rmd
now with prolactin candidate plot

@@
 ```
+
+# Plots with Prolactin
 
@@
 ```
+
+## Circulating levels of prolactin
 
@@
 ```
+
+## Prolactin (_PRL_) and _BRCA1_ expression in the pituitary
 
@@
 
-## prolactin manip
+## Manipulation of parental care and _PRL_ and _BRCA1_ expression
 
@@
 
+## Genes an hormones (in progress, don't have all the bird ids lining up.)
+
 ```{r geneshormones}
@@
 candidates.pit$X1 <- candidates.pit$V1
+
 geneshormones <- left_join(candidates.pit, hormones) %>%
@@
 
-(geneshormones$bird_id)
-(hormones$bird_id)
+head(geneshormones$bird_id)
+head(hormones$bird_id)
 
@@
 ```
+
+## Prolactin hormone statistics
 
@@
 
+meanPRL <- candidates.pit %>% 
+    droplevels() %>% 
+  dplyr::group_by(treatment) %>%
+  dplyr::summarise(m = mean(PRL), 
+                   se = sd(PRL)/sqrt(length(PRL))) %>%
+  dplyr::mutate(m = round(m,1))  %>%
+  dplyr::filter(treatment %in% charlevels)
+meanPRL <- left_join(meanPRL, birds)
+meanPRL$treatment <- factor(meanPRL$treatment, levels = charlevels)
+
 p8 <- ggplot(meanPRL, aes(treatment, m)) +
@@
   geom_hline(yintercept=20) +
-  geom_hline(yintercept=21)
+  geom_hline(yintercept=21) +
+  theme(axis.title.y  = element_text(face = "italic"))
 
 p8
+
+```
+
+## WGCNA candidates
+
+```{r WGCNAcandidates}
+
+#WGCNAgenes <- read_csv("../results/08_PRL_associated.csv") %>% pull(x)
+#WGCNAgenes <- c("FBXO5", "NUF2", "PTTG1", "BUB1", "PRL", "ASPM", "AURKA", "CDC20", "NEK2", "CKS2", 
+#                  "MYC", "CCNB3", "FANCC", "STC1", "RACGAP1", "E2F7", "PLK1", "EXO1", "CENPI", "CDK1")
+WGCNAgenes <- c("PRL", "PAX7", "ASPM", "AURKA",  "CDC20", "CENPF", "MYC",  "CDK1", "LBH", "BRCA1", "FOXM1", "FBXO5")
+
+WGCNAgenes.pit <- vsd.pit %>% dplyr::filter(genes %in% WGCNAgenes)
+WGCNAgenes.pit <- as.data.frame(WGCNAgenes.pit)
+row.names(WGCNAgenes.pit) <- WGCNAgenes.pit$genes
+WGCNAgenes.pit <- pivot_longer(WGCNAgenes.pit, -genes, names_to = "V1", values_to = "expression")
+head(WGCNAgenes.pit)
+WGCNAgenes.pit <- left_join(colData.pit, WGCNAgenes.pit) %>%
+  filter(study == "charcterization")
+WGCNAgenes.pit$treatment <- factor(WGCNAgenes.pit$treatment, levels = alllevels)
+
+WGCNAgenes.pit$genes <- factor(WGCNAgenes.pit$genes)
+WGCNAgenes.pit$genesnum <- round(as.numeric(WGCNAgenes.pit$genes), digits = -1)
+
+p1 <- ggplot(WGCNAgenes.pit, aes(x = treatment, y = expression)) +
+  geom_boxplot(aes(fill = treatment)) +
+  facet_wrap(~genes, scales = "free_y") +
+  theme_B3() +
+  theme(legend.position = "none") +
+  #scale_color_manual(values = sexcolors) +
+  theme(axis.text.y = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        strip.background = element_blank(),
+        strip.text = element_text(face = "italic")) +
+  labs( x = "parental stage", y = "pituitary expression",
+        subtitle = "Genes that are coexpressed with prolactin and regulate development")
+p1
 
@@
 write.csv(aov_manip, "../results/16_aov_PRLsextreatment.csv")
-
 ```

### 2019-12-31 analysis/06_DESeq2_ALL.Rmd
new volcano plots

@@
 plot_grid(charicons, char2, nrow = 2, rel_heights = c(0.9,1))
+
+timing2 <- plot_grid( 
+          n , 
+          o + theme(axis.title.y = element_blank(), axis.text.y = element_blank())+
+            labs(x = "Log fold change,")   , 
+          NULL,
+          p + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
+            labs(x = "pituitary") , 
+          q + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
+           NULL,
+          r + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
+          s + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
+            nrow = 1, rel_widths = c(1.4,1, 0.5, 1, 1, 0.5, 1, 1))
+
+plot_grid(timingicons, timing2, nrow = 2, rel_heights = c(0.9,1))
 ```

### 2019-12-31 analysis/06_DESeq2_ALL.Rmd
new volcano plots

@@
 plot_grid(timingicons, timing2, nrow = 2, rel_heights = c(0.9,1))
+
+remove2 <- plot_grid( 
+          NULL,
+          j , 
+          k + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
+          l + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
+          m + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
+          nrow = 1, rel_widths = c(1,1.4,1,1,1))
+
+plot_grid(removeicons, remove2, nrow = 2, rel_heights = c(0.9,1))
 ```

### 2019-12-31 analysis/02_PCA.Rmd
new tsne

@@
 gontsne <- subsetmaketsne("gonads", charlevels, sexlevels)
+alltsne <- subsetmaketsne(tissuelevels, alllevels, sexlevels)
 
@@
 abc
+
+a2 <- plotcolorfultsnes(alltsne, alltsne$tissue, colorstissue)  
+b2 <- plotcolorfultsnes(alltsne, alltsne$sex, sexcolors)   
+c2 <- plotcolorfultsnes(alltsne, alltsne$treatment, colorscharmaip)
+
+abc2 <- plot_grid(a2,b2,c2, nrow = 1, labels = c("a", "b", "c"), label_size = 12)
+abc2
 
@@
 
+allpca <- subsetmakepca(tissuelevels, alllevels, sexlevels)
 ```
@@
 def
+
+d2 <- plotcolorfulpcs(allpca, allpca$tissue, colorstissue)  
+e2 <- plotcolorfulpcs(allpca, allpca$sex, sexcolors) 
+f2 <- plotcolorfulpcs(allpca,allpca$treatment, colorscharmaip)
+
+def2 <- plot_grid(d2,e2, f2, nrow =  1,  labels = c("d", "e", "f"), label_size = 12)
+def2
 ```
@@
 
+allfviz <- makefvizdf(tissuelevels, alllevels, sexlevels)
+
 g <- fviz_screeplot(charfviz, addlabels = TRUE, ylim = c(0, 50),  ncp = 5, barcolor = "white", barfill = "white") + 
@@
 plot_grid(def, gh, nrow = 2)
+
+g2 <- fviz_screeplot(allfviz, addlabels = TRUE, ylim = c(0, 50),  ncp = 5, barcolor = "white", barfill = "white") + 
+  labs(title = NULL, y = "PC Variance") + theme_B3() 
+h2 <- fviz_pca_var(allfviz,  labelsize = 3 , axes.linetype = "blank", 
+                   repel = TRUE ,
+                  select.var= list(contrib = 5)) + 
+      theme_B3() + 
+      labs(x = "PC1 (35.8%)", y = "PC2 (32.5%)", title =  NULL) +
+      theme( axis.text = element_blank())
+
+legend2 <- png::readPNG("../figures/images/DoveParentsRNAseq_legend.png")
+legend2 <-  grid::rasterGrob(legend2, interpolate=TRUE)
+
+gh2 <- plot_grid(g2,h2, legend2, nrow = 1, rel_widths = c(1, 1, 1), labels = c("h", "i", NULL ), label_size = 12)
+gh2
+plot_grid(def2, gh2, nrow = 2)
 ```
@@
 plot_grid( abc, def, gh, nrow = 3)
+plot_grid( abc2, def2, gh2, nrow = 3)
 

### 2020-01-05 analysis/16_PRL.Rmd
new combined PRL image

@@
     theme(legend.position = c(0.85,0.15), legend.direction = "horizontal") + 
-  labs(x = "increasing time >>", y = "prolactin (ng/mL)") +
+  labs(x = "increasing time >>", y = "prolactin (ng/mL)")
+p1
+
+p1 +
   annotation_custom(control, ymin = -20, ymax = 0, xmin = -7.8) +
@@
   ylim(c(-10,120))
-p1
+
 ```
@@
 
+```
+
+```{r hormonesandgenes}
+plot_grid(p1 + theme(legend.position ="none",
+                     axis.title.x = element_blank(),
+                     axis.text.x = element_blank()),
+          p2 + theme(legend.position = c(0.85,0.3)) +
+            labs(subtitle = NULL) +
+  annotation_custom(control, ymin = 13.5, ymax = 15.5, xmin = -7.5) +
+  annotation_custom(bldg, ymin = 13.5, ymax = 15.5, xmin = -5.5) +
+  annotation_custom(lay, ymin = 13.5, ymax = 15.5, xmin = -3.5) +
+  annotation_custom(inc, ymin = 13.5, ymax = 15.5, xmin = -1.5) +
+  annotation_custom(inc, ymin = 13.5, ymax = 15.5, xmin = 0.5) +
+  annotation_custom(inc, ymin = 13.5, ymax = 15.5, xmin = 2.5) +
+  annotation_custom(hatch, ymin = 13.5, ymax = 15.5, xmin = 4.5) +
+  annotation_custom(nestling, ymin = 13.5, ymax = 15.5, xmin = 6.5) +
+  annotation_custom(nestling, ymin = 13.5, ymax = 15.5, xmin = 8.5), 
+          nrow = 2,rel_heights = c(0.4,0.6)) 
 ```

### 2020-01-14 analysis/07_hormones.Rmd
cort hormones

@@
 library(ggsignif)
+library(apaTables)
 
@@
 ```
+
+```{r sex}
+hormones$hormone <- factor(hormones$hormone, 
+                           levels = c("cort", "progesterone", "prolactin", 
+                                      "estradiol", "testosterone"))
+
+hormones %>%
+  filter(study == "characterization") %>%
+  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
+  geom_boxplot() +
+  theme_B3() +
+  facet_wrap(~hormone, nrow = 1, scales = "free") +
+  theme(legend.position = "none") +
+  labs(y = "concentration (ng/mL")
+
+printaovtablesex <- function(myhormone, mydescirption){
+
+  aov_all = data.frame()
+  for(i in alllevels3){
+    df <- hormones %>% filter(hormone == myhormone,
+                            treatment == i) %>% droplevels()
+    aovtable <- apa.aov.table(aov(plasma_conc ~ sex, data  = df))
+    aovtable <- as.data.frame(aovtable$table_body)
+    totaldf <- aovtable[3, 3]
+    aovtable$df <- paste(aovtable$df, ", " , totaldf, sep = "")
+    aovtable$ANOVA <- mydescirption
+    aovtable$stages <- paste(i)
+    aovtable$p <- as.numeric(as.character(aovtable$p))
+    aov_all <- rbind(aov_all,aovtable)
+  }
+
+  aov_all <- aov_all %>%
+    filter(Predictor == "sex")  %>%
+    select(stages, ANOVA, df, "F", p) %>%
+    mutate(sig = ifelse(p < 0.05, "*", " "))
+  aov_all
+  return(aov_all)
+}
+
+printaovtablesex("cort", "CORT ~ sex")
+printaovtablesex("progesterone", "Prog ~ sex")
+printaovtablesex("prolactin", "PRL ~ sex")
+
+TukeyHSD(aov(data = cort.char, plasma_conc ~ treatment), which = "treatment")
+
+```
+
+```{r cort}
+p1 <- hormones %>%
+  filter(hormone == "cort",
+         study == "characterization") %>%
+  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
+  geom_boxplot() +
+  theme_B3() +
+  labs(y = "concentration (ng/mL)", subtitle = "Corticosterone" , x = NULL) +
+  theme(legend.position = "none") +
+  scale_y_continuous(limits = c(0,8.5))
+p1
+
+cort <- hormones %>%
+  filter(hormone == "cort",
+         study == "characterization")
+
+p2 <-  ggplot(cort, aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot() +
+  theme_B3() +
+  labs(y = NULL, x = NULL, subtitle = " ") +
+  theme(legend.position = "none") +
+  geom_signif(comparisons=list(c("control", "inc.d17")), annotations= "p = 0.49", y_position = 7.5, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
+  geom_signif(comparisons=list(c("control", "hatch")), annotations= "p = 0.021", y_position = 8, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
+  geom_signif(comparisons=list(c("control", "n5")), annotations= "p = 0.004", y_position = 8.5, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
+  geom_signif(comparisons=list(c("n5", "n9")), annotations= "p = 0.001", y_position = 8, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
+  scale_y_continuous(limits = c(0,8.5))
+
+p2
+
+p12 <- plot_grid(p1,p2, rel_widths = c(0.2,0.8))
+
+p3 <- hormones %>%
+  filter(hormone == "cort",
+         treatment %in% c(controlsremoval, levelsremoval )) %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot() +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,8.5)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none") +
+  labs( x = NULL, y = "concentration (ng/mL")
+p3
+
+p4 <- hormones %>%
+  filter(hormone == "cort",
+         treatment %in% c(controlstiming, levelstiming )) %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot() +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,8.5)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none") +
+  labs(y = NULL, x = NULL)
+p4
+
+p34 <- plot_grid(p3,p4, rel_widths = c(0.6,0.4))
+
+plot_grid(p12, p34, nrow = 2, rel_heights = c(0.6,0.4))
+```

### 2020-01-14 R/themes.R

@@ -41,6 +41,10 @@ maniplevels1 <- c( "m.inc.d3" ,  "m.inc.d8" ,  "m.inc.d9" , "m.inc.d17" , "m.n2"
 levelstiming <- c( "m.inc.d8" , "prolong" , "extend")
 levelsremoval <- c( "m.inc.d3" ,    "m.inc.d9" , "m.inc.d17" , "m.n2")
 
+controlstiming <- c( "inc.d9" , "hatch" , "n5")
+controlsremoval <- c( "inc.d3" ,    "inc.d9" , "inc.d17" , "hatch")
+
+
 alllevels <- c("control", "bldg", "lay", "inc.d3", "m.inc.d3" ,  
                 "inc.d9", "m.inc.d8" ,"m.inc.d9" ,
                "inc.d17",  "m.inc.d17","prolong" ,

### 2020-01-14 analysis/07_hormones.Rmd
new prolactin figures

@@
 ```
+
+```{r PRL}
+p1 <- hormones %>%
+  filter(hormone == "prolactin",
+         study == "characterization") %>%
+  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
+  geom_boxplot() +
+  theme_B3() +
+  labs(y = "concentration (ng/mL)", subtitle = "Prolactin" , x = NULL) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_fill_manual(values = sexcolors) +
+  scale_y_continuous(limits = c(0,100)) 
+p1
+
+PRLf <- hormones %>%
+  filter(hormone == "prolactin",
+         study == "characterization",
+         sex == "female")
+
+PRLm <- hormones %>%
+  filter(hormone == "prolactin",
+         study == "characterization",
+         sex == "male")
+
+p2a <-  ggplot(PRLf, aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot(aes(color = sex)) +
+  theme_B3() +
+  labs(y = NULL, x = NULL, subtitle = "females") +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_manual(values = sexcolors) +
+  scale_y_continuous(limits = c(0,100)) +
+    geom_text(aes(label = "B", x = 1, y = 100, fontface = "plain"), size = 3) +
+   geom_text(aes(label = "A", x = 2, y = 100), size = 3) +
+geom_text(aes(label = "A", x = 3, y = 100), size = 3) +
+geom_text(aes(label = "B", x = 4, y = 100), size = 3) +
+  geom_text(aes(label = "C", x = 5, y = 100), size = 3) +
+  geom_text(aes(label = "D", x = 6, y = 100), size = 3) +
+  geom_text(aes(label = "D", x = 7, y = 100), size = 3) +
+  geom_text(aes(label = "D", x = 8, y = 100), size = 3) +
+  geom_text(aes(label = "D", x = 9, y = 100), size = 3)
+
+p2b <-  ggplot(PRLm, aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot(aes(color = sex)) +
+  theme_B3() +
+  labs(y = NULL, x = NULL, subtitle = "males") +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_manual(values = sexcolors) +
+  scale_y_continuous(limits = c(0,100)) +
+  geom_text(aes(label = "A", x = 1, y = 100, fontface = "plain"), size = 3) +
+   geom_text(aes(label = "A", x = 2, y = 100), size = 3) +
+geom_text(aes(label = "B", x = 3, y = 100), size = 3) +
+geom_text(aes(label = "B", x = 4, y = 100), size = 3) +
+  geom_text(aes(label = "B", x = 5, y = 100), size = 3) +
+  geom_text(aes(label = "D", x = 6, y = 100), size = 3) +
+  geom_text(aes(label = "D", x = 7, y = 100), size = 3) +
+  geom_text(aes(label = "D", x = 8, y = 100), size = 3) +
+  geom_text(aes(label = "D", x = 9, y = 100), size = 3)
+
+p12 <- plot_grid(p1,p2a,p2b, rel_widths = c(0.25,0.5, 0.5), labels = c("a","b", "c"), label_size = 8, nrow = 1, align = "hv")
+p12
+
+p3a <- hormones %>%
+  filter(hormone == "prolactin",
+         sex == "female", 
+         treatment %in% c(controlsremoval, levelsremoval )) %>% 
+  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot(aes(color = sex)) +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,100)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  labs( x = NULL, y = "concentration (ng/mL", subtitle = "females") +
+  scale_color_manual(values = sexcolors)
+p3a
+
+p3b <- hormones %>%
+  filter(hormone == "prolactin",
+         sex == "male", 
+         treatment %in% c(controlsremoval, levelsremoval )) %>% 
+  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot(aes(color = sex)) +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,100)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  labs( x = NULL, y = NULL, subtitle = "males") +
+  scale_color_manual(values = sexcolors)
+p3b
+
+p4a <- hormones %>%
+  filter(hormone == "prolactin",
+         sex == "female", 
+         treatment %in% c(controlstiming, levelstiming )) %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot(aes(color = sex)) +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,100)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  labs(y = NULL, x = NULL, subtitle = "females") +
+  scale_color_manual(values = sexcolors)
+p4a
+
+p4b <- hormones %>%
+  filter(hormone == "prolactin",
+         sex == "male", 
+         treatment %in% c(controlstiming, levelstiming )) %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot(aes(color = sex)) +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,100)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  labs(y = NULL, x = NULL, subtitle = "males") +
+  scale_color_manual(values = sexcolors)
+
+p4b
+
+p34 <- plot_grid(p3a , p3b,p4a, p4b, rel_widths = c(0.55,0.55, 0.45,0.45), labels = c("d","e", "f", "g"), label_size = 8, nrow = 1)
+
+plot_grid(p12, p34, nrow = 2, rel_heights = c(0.5,0.5))
+```

### 2020-01-14 analysis/07_hormones.Rmd
new correlations

@@
 ```{r cort}
+
+hormones %>%
+  filter(hormone == "cort") %>%
+  group_by(sex) %>%
+  summarise(mean = mean(plasma_conc),
+            median = median(plasma_conc))
+
 p1 <- hormones %>%
@@
 
-p12 <- plot_grid(p1,p2, rel_widths = c(0.2,0.8))
+p12 <- plot_grid(p1,p2, rel_widths = c(0.2,0.8), labels = c("a","b"), label_size = 8)
 
@@
 
-p34 <- plot_grid(p3,p4, rel_widths = c(0.6,0.4))
+p34 <- plot_grid(p3,p4, rel_widths = c(0.6,0.4), labels = c("c","d"), label_size = 8 )
 
@@
 ```{r PRL}
+hormones %>%
+  filter(hormone == "prolactin") %>%
+  group_by(sex) %>%
+  summarise(m = mean(plasma_conc),
+            median = median(plasma_conc))
+
 p1 <- hormones %>%
@@
 ```
+
+```{r correlations2 }
+hormoneswide <- hormones %>%
+  select(study, treatment, sex, bird_id, hormone, plasma_conc) %>%
+  pivot_wider( names_from = hormone, values_from  = plasma_conc, values_fn = list(plasma_conc = mean))
+
+plotcorrelations <- function(i,j, myxlab, mylab, mysubtitle){
+      p <- ggplot(hormoneswide, aes(x = i, y = j),na.rm = TRUE) +
+       geom_point(aes(color = treatment), na.rm = TRUE) +
+       geom_smooth( method = lm, color = "darkgrey", na.rm = TRUE ) +
+       theme_B3() +
+      #facet_wrap(~sex, scales = "free") +
+        labs(x = myxlab, y = mylab, subtitle = mysubtitle) +
+        theme(legend.position = "none")
+
+      return(p)
+  }
+
+p1 <- plotcorrelations(hormoneswide$prolactin, hormoneswide$cort, "prolactin", "cort", "r = 0.221, p = 0.001")
+p2 <- plotcorrelations(hormoneswide$prolactin, hormoneswide$progesterone, "prolactin", "progesterone", "r = -0.099, p = 0.2")
+p3 <- plotcorrelations(hormoneswide$prolactin, hormoneswide$testosterone, "prolactin", "testosterone", "r = -0.258, p = 0.08")
+p4 <- plotcorrelations(hormoneswide$prolactin, hormoneswide$estradiol, "prolactin", "estradiol", "r = 0.05, p = 0.6")
+
+p5 <- plotcorrelations(hormoneswide$cort, hormoneswide$progesterone, "cort", "progesterone", "r = 0.104, p = 0.175")
+p6 <- plotcorrelations(hormoneswide$cort, hormoneswide$estradiol, "cort", "estradiol", "r = -0.026, p = 0.8")
+p7 <- plotcorrelations(hormoneswide$cort, hormoneswide$testosterone, "cort", "testosterone", "r = -0.108, p = 0.4")
+
+p8 <- plotcorrelations(hormoneswide$progesterone, hormoneswide$estradiol, "progesterone", "estradiol", "r = 0.120, p = 0.3")
+p9 <- plotcorrelations(hormoneswide$progesterone, hormoneswide$testosterone, "progesterone", "testosterone", "r = 0.133, p = 0.3")
+
+allcorrs <- plot_grid(p1,p2,p3,p4,p5,p6,p7,p8,p9, nrow = 3, align = "hv", labels = "auto", label_size = 8)
+allcorrs
+
+plot_grid(p1,p3 + theme(legend.position = "right"), align = "hv", rel_widths = c(0.4,0.6))
+
+cor.test(hormoneswide$prolactin, hormoneswide$cort)
+cor.test(hormoneswide$prolactin, hormoneswide$progesterone)
+cor.test(hormoneswide$prolactin, hormoneswide$estradiol)
+cor.test(hormoneswide$prolactin, hormoneswide$testosterone)
+
+cor.test(hormoneswide$cort, hormoneswide$estradiol)
+cor.test(hormoneswide$cort, hormoneswide$testosterone)
+cor.test(hormoneswide$cort, hormoneswide$progesterone)
+
+cor.test(hormoneswide$progesterone, hormoneswide$estradiol)
+cor.test(hormoneswide$progesterone, hormoneswide$testosterone)
+```

### 2020-01-15 analysis/07_hormones.Rmd
more figurs

@@
 source("../R/icons.R")
+
+theme_B3 <- function () { 
+  theme_classic(base_size = 11) +
+    theme(
+      panel.grid.major  = element_blank(),  # remove major gridlines
+      panel.grid.minor  = element_blank(),  # remove minor gridlines
+      plot.title = element_text(hjust = 0.5, face = "bold"), # center & bold 
+      axis.text.x = element_text(angle = 45, hjust = 1)
+    )
+}
 
@@
 
+## prolactin 2
+
+```{r PRL2}
+hormones %>%
+  filter(hormone == "prolactin") %>%
+  group_by(sex) %>%
+  summarise(m = mean(plasma_conc),
+            median = median(plasma_conc))
+
+p1 <- hormones %>%
+  filter(hormone == "prolactin",
+         study == "characterization") %>%
+  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
+  geom_boxplot() +
+  theme_B3() +
+  labs(y = "concentration (ng/mL)", subtitle = "Prolactin" , x = NULL) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_fill_manual(values = sexcolors) +
+  scale_y_continuous(limits = c(0,100)) 
+p1
+
+df <- hormones %>%
+  filter(hormone == "prolactin",
+         study == "characterization")
+p2 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
+  geom_boxplot(aes(color = sex)) +
+  theme_B3() +
+  labs(y = NULL, x = NULL, subtitle = " ") +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_manual(values = sexcolors) +
+  scale_y_continuous(limits = c(0,105)) +
+    geom_text(aes(label = "B", x = 1, y = 100, fontface = "plain"), size = 3) +
+   geom_text(aes(label = "A", x = 2, y = 100), size = 3) +
+geom_text(aes(label = "A", x = 3, y = 100), size = 3) +
+geom_text(aes(label = "B", x = 4, y = 100), size = 3) +
+  geom_text(aes(label = "C", x = 5, y = 100), size = 3) +
+  geom_text(aes(label = "D", x = 6, y = 100), size = 3) +
+  geom_text(aes(label = "D", x = 7, y = 100), size = 3) +
+  geom_text(aes(label = "D", x = 8, y = 100), size = 3) +
+  geom_text(aes(label = "D", x = 9, y = 100), size = 3) +
+  geom_text(aes(label = '*', x = 1, y = 105), size = 5) +
+   geom_text(aes(label = '*', x = 2, y = 105), size = 5) 
+  p2
+
+p3 <- hormones %>%
+  filter(hormone == "prolactin",
+         treatment %in% c(controlsremoval, levelsremoval )) %>% 
+  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot(aes(color = sex)) +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,110)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  labs( x = NULL, y = "concentration (ng/mL", subtitle = " ") +
+  scale_color_manual(values = sexcolors) +
+   geom_text(aes(label = '*', x = 2, y = 105), size = 5) 
+p3
+
+p4 <- hormones %>%
+  filter(hormone == "prolactin",
+         treatment %in% c(controlstiming, levelstiming )) %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot(aes(color = sex)) +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,110)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  labs(y = NULL, x = NULL, subtitle = " ") +
+  scale_color_manual(values = sexcolors) 
+p4
+
+p12 <- plot_grid(p1, p2, rel_widths = c(0.15,0.8), labels = "auto", label_size = 8)
+p34 <- plot_grid(p3,p4, nrow = 1, rel_widths = c(0.55,0.45), labels = c("c", "d"), label_size = 8)
+
+plot_grid(p12, p34, nrow = 2)
+```
+
+## correlations
+
 ```{r correlations2 }
@@
         labs(x = myxlab, y = mylab, subtitle = mysubtitle) +
-        theme(legend.position = "none")
+        theme(legend.position = "none") +
+        scale_color_manual(values = colorscharmaip2)
 
@@
 
-plot_grid(p1,p3 + theme(legend.position = "right"), align = "hv", rel_widths = c(0.4,0.6))
+plot_grid(p1 + theme(axis.text.x = element_text(angle = 0)),p3+ theme(axis.text.x = element_text(angle = 0))  + theme(legend.position = "right"), align = "hv", rel_widths = c(0.4,0.6))
 
@@
 ```
+
+## progesterone
+
+```{r progesterone}
+
+hormones %>%
+  filter(hormone == "progesterone") %>%
+  group_by(sex) %>%
+  summarise(mean = mean(plasma_conc),
+            median = median(plasma_conc))
+
+p1 <- hormones %>%
+  filter(hormone == "progesterone",
+         study == "characterization") %>%
+  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
+  geom_boxplot() +
+  theme_B3() +
+  labs(y = "concentration (ng/mL)", subtitle = "Progesterone" , x = NULL) +
+  theme(legend.position = "none")   +
+  scale_y_continuous(limits = c(0,4.5)) +
+  scale_fill_manual(values = sexcolors)
+
+df <- hormones %>%
+  filter(hormone == "progesterone",
+         study == "characterization")
+
+p2 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot() +
+  theme_B3() +
+  labs(y = NULL, x = NULL, subtitle = " ") +
+  scale_y_continuous(limits = c(0,4.5)) +
+  theme(legend.position = "none")
+
+p3 <- hormones %>%
+  filter(hormone == "progesterone",
+         treatment %in% c(controlsremoval, levelsremoval )) %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot() +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,4.5)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none") +
+  labs( x = NULL, y = "concentration (ng/mL")
+
+p4 <- hormones %>%
+  filter(hormone == "progesterone",
+         treatment %in% c(controlstiming, levelstiming )) %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot() +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,4.5)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none") +
+  labs(y = NULL, x = NULL)
+
+progesterone <- plot_grid(p1, p2, p3 + labs(y = NULL), p4, nrow = 1, rel_widths =   c(0.175,0.4,0.35,0.25))
+progesterone
+```
+
+## estradiol and testosterone
+
+```{r estradiol}
+
+hormones %>%
+  filter(hormone == "estradiol") %>%
+  summarise(mean = mean(plasma_conc),
+            median = median(plasma_conc))
+
+p5 <- hormones %>%
+  filter(hormone == "estradiol",
+         study == "characterization") %>%
+  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
+  geom_boxplot() +
+  theme_B3() +
+  labs(y = "concentration (ng/mL)", subtitle = "Estradiol" , x = NULL) +
+  theme(legend.position = "none")  +
+    scale_y_continuous(limits = c(0,1)) +
+  scale_fill_manual(values = sexcolors)
+
+df <- hormones %>%
+  filter(hormone == "estradiol",
+         study == "characterization")
+
+p6 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot() +
+  theme_B3() +
+  labs(y = NULL, x = NULL, subtitle = " ") +
+    scale_y_continuous(limits = c(0,1)) +
+  theme(legend.position = "none")
+
+p7 <- hormones %>%
+  filter(hormone == "estradiol",
+         treatment %in% c(controlsremoval, levelsremoval )) %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot() +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,1)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none") +
+  labs( x = NULL, y = "concentration (ng/mL")
+
+p8 <- hormones %>%
+  filter(hormone == "estradiol",
+         treatment %in% c(controlstiming, levelstiming )) %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot() +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,1)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none") +
+  labs(y = NULL, x = NULL)
+
+estradiol <- plot_grid(p5, p6, p7 + labs(y = NULL), p8, nrow = 1, rel_widths =   c(0.175,0.4,0.35,0.25))
+estradiol
+```
+
+```{r testosterone}
+
+hormones %>%
+  filter(hormone == "testosterone") %>%
+  summarise(mean = mean(plasma_conc),
+            median = median(plasma_conc))
+
+p9 <- hormones %>%
+  filter(hormone == "testosterone",
+         study == "characterization") %>%
+  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
+  geom_boxplot() +
+  theme_B3() +
+  labs(y = "concentration (ng/mL)", subtitle = "Testosterone" , x = NULL) +
+  theme(legend.position = "none")  +
+    scale_y_continuous(limits = c(0,8.5)) +
+  scale_fill_manual(values = sexcolors)
+
+df <- hormones %>%
+  filter(hormone == "testosterone",
+         study == "characterization")
+
+p10 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot() +
+  theme_B3() +
+  labs(y = NULL, x = NULL, subtitle = " ") +
+    scale_y_continuous(limits = c(0,8.5)) +
+  theme(legend.position = "none")
+
+p11 <- hormones %>%
+  filter(hormone == "testosterone",
+         treatment %in% c(controlsremoval, levelsremoval )) %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot() +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,8.5)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none") +
+  labs( x = NULL, y = "concentration (ng/mL")
+
+p12 <- hormones %>%
+  filter(hormone == "testosterone",
+         treatment %in% c(controlstiming, levelstiming )) %>% 
+ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
+  geom_boxplot() +
+  theme_B3() +
+  scale_y_continuous(limits = c(0,8.5)) +
+  scale_fill_manual(values = colorscharmaip) +
+  theme(legend.position = "none") +
+  labs(y = NULL, x = NULL)
+p12
+
+testosterone <- plot_grid(p9, p10, p11 + labs(y = NULL), p12, nrow = 1, rel_widths =   c(0.175,0.4,0.35,0.25))
+testosterone
+```
+
+```{r supplefig, fig.height=8}
+plot_grid(p1, p2, p3 + labs(y = NULL), p4, p5, p6, p7 + labs(y = NULL), p8, 
+          p9, p10, p11 + labs(y = NULL), p12, nrow = 3, rel_widths =   c(0.175,0.4,0.35,0.25), align = "hv",
+          label_size = 8, labels = c("a", " ", " " , " ",
+                                     "b", " ", " " , " ",
+                                     "c", " ", " " , " "))
+
+```

### 2020-01-16 analysis/16_PRL.Rmd
new music figs

@@
 
-```{r music}
+```{r music, fig.height=6, fig.width=6}
 
@@
 
+p9 <- ggplot(meanPRL, aes(treatment, m)) +
+   geom_image(aes(image=music), size=.12)  +
+  mytheme() +
+  ylim(c(15,22)) +
+  geom_hline(yintercept=17) +
+  geom_hline(yintercept=18) +
+  geom_hline(yintercept=19) +
+  geom_hline(yintercept=20) +
+  geom_hline(yintercept=21) +
+  theme(axis.text  = element_blank(),
+        axis.title = element_blank())
+
+p9
+
+p10 <- candidates.pit %>%
+  dplyr::filter(treatment %in% charlevels) %>%
+  ggplot(aes(x = treatment, y = PRL)) + 
+    geom_boxplot(aes(fill = treatment)) + 
+    scale_alpha_manual(values = c(0.75,1)) +
+    theme_B3() +
+  theme(legend.position = c(0.85,0.2), legend.direction = "horizontal") + 
+  scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+  labs(y = "Prolactin expression", x = "parental stage") +
+  guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) 
+p10
+
+plot_grid(p10,p9, nrow = 2, align = "hv")
 ```

### 2020-01-17 analysis/07_hormones.Rmd
new hormone data

@@
 source("../R/icons.R")
-
-theme_B3 <- function () { 
-  theme_classic(base_size = 11) +
-    theme(
-      panel.grid.major  = element_blank(),  # remove major gridlines
-      panel.grid.minor  = element_blank(),  # remove minor gridlines
-      plot.title = element_text(hjust = 0.5, face = "bold"), # center & bold 
-      axis.text.x = element_text(angle = 45, hjust = 1)
-    )
-}
 
@@
 
-prolactin2 <- read_csv("../results/prolactin.csv") %>% 
-          dplyr::rename("plasma_conc" = "prolactin_ng_mL",
-                        "treatment" = "stage") %>% 
-           dplyr::mutate(hormone = "prolactin") %>% 
-           dplyr::mutate(treatment = fct_recode(treatment,
-                            "m.n2" = "m.hatch",
-                            "prolong" = "inc.prolong",
-                            "extend" = "extend.hatch"),
-                          study = fct_collapse(treatment,
-                                 characterization = charlevels,
-                                 manipulation = maniplevels1)) %>%
-
-          dplyr::mutate(bird_id = gsub(".ATLAS", "" , bird_id)) %>% 
-          dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc) 
-head(prolactin2)
-
-same <- inner_join(prolactin, prolactin2, by = "bird_id")
-ggplot(same, aes(x = plasma_conc.x, y = plasma_conc.y, label = bird_id)) + geom_point() + geom_text()
-
-all <- full_join(prolactin, prolactin2, by = "bird_id") 
-onlyin1 <- anti_join(prolactin, prolactin2, by = "bird_id") 
-onlyin2 <- anti_join(prolactin2, prolactin, by = "bird_id")
-
-dim(prolactin)
-dim(prolactin2)
-dim(all)
-dim(same)
-dim(onlyin1)
-dim(onlyin2)
-
-print("only in old prolactin dataset") 
-onlyin1$bird_id
-
-print("only in new prolactin dataset") 
-onlyin2$bird_id
-
-prolctin1rnaseq <- full_join(prolactin, colData)
-prolctin2rnaseq <- full_join(prolactin2, colData)
-
-prolctin1rnaseq
-prolctin2rnaseq
 ```
@@
 ```{r wrangle-steroids}
-PETC <- read_excel("../results/parental_care_hormone_RIA_data_master.xlsx", sheet = 2)  %>% 
-                dplyr::mutate(treatment = fct_recode(stage,
-                            "inc.d17" = "inc_d17",
-                            "inc.d3" = "inc_d3",
-                            "inc.d9" = "inc_d9",
-                            "m.inc.d9" = "m_incd9",
-                            "m.inc.d3" = "m_incd3",
-                            "m.inc.d8" = "m_incd8",
-                            "m.inc.d17" = "m_incd17",
-                            "m.n2" = "m_hatch",
-                            "control" = "stress_hpg")) %>%
+PETC <- read_excel("../results/hormones.xlsx", sheet = 1)  %>% 
+              dplyr::rename(corticosterone = cort, progesterone = p4,
+                            estradiol = e2,  testosterone = t) %>% 
+               dplyr::mutate(treatment = fct_recode(treatment...3,
+                            "inc.d17" = "incd17",
+                            "inc.d3" = "incd3",
+                            "inc.d9" = "incd9",
+                            "m.inc.d9" = "minc9",
+                            "m.inc.d3" = "minc3",
+                            "m.inc.d8" = "minc8",
+                            "m.inc.d17" = "minc17",
+                            "m.n2" = "m hatch",
+                            "control" = "baseline")) %>%
                 dplyr::filter(treatment %in% alllevels2) %>%   
@@
                                 manipulation = maniplevels1)) %>% 
-              dplyr::mutate(bird_id = gsub("[[:punct:]]", "." , band_combo)) %>% 
-              dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc) %>% 
+              dplyr::mutate(bird_id = gsub("[[:punct:]]", "." , id)) %>% 
+              dplyr::select(bird_id, treatment, sex, study, 
+                            corticosterone, estradiol, testosterone, progesterone) %>% 
+              pivot_longer(cols = corticosterone:progesterone,
+                           names_to = "hormone", values_to = "plasma_conc",
+                           values_drop_na = TRUE) %>% 
               drop_na() %>% droplevels()  
 head(PETC)
-
-```
-
-```{r correlations}
-PETCP <- left_join(prolactin2, PETC, by = "bird_id") %>% drop_na()
-PETCP$treatment.x <- factor(PETCP$treatment.x, levels = alllevels2)
-
-plothormonecorrelations <- function(myhormone, myylab){
-  PETCP %>% filter(hormone.y == myhormone) %>%
-  ggplot(aes(x = plasma_conc.x, y = plasma_conc.y, color = treatment.x )) +
-  geom_point() + 
-  geom_smooth(method = "lm", se = F) +
-  labs(x = "Prolactin (ng/mL)", y = myylab) +
-  scale_color_manual(values = colorscharmaip) +
-  facet_wrap(~treatment.x, nrow = 3) +
-  theme(legend.position = "none")
-}
-
-plothormonecorrelations("cort", "Corticosterone (ng/mL")
-plothormonecorrelations("estradiol", "Estradiol (ng/mL")
-plothormonecorrelations("progesterone", "Progesterone (ng/mL")
-plothormonecorrelations("testosterone", "Testosterone (ng/mL")
 ```
@@
 hormones <- rbind(prolactin, PETC)
+
+hormones$okay <- ifelse(hormones$hormone == "corticosterone" & hormones$plasma_conc > 30, "bad",
+                    ifelse(hormones$hormone == "progesterone" & hormones$plasma_conc > 5, "bad", 
+                           ifelse(hormones$hormone == "prolactin" & hormones$plasma_conc > 150, "bad", 
+                        ifelse(hormones$hormone == "testosterone" & hormones$sex == "female", "bad",
+                               ifelse(hormones$hormone == "estradiol" & hormones$sex == "male", "bad", "okay")))))
+hormones <- hormones %>% filter(okay == "okay") %>% droplevels() %>% select(-okay)
+
+write.csv(hormones, "../results/hormones.csv", row.names = F)
+
+# make a winder one for correlations
+hormoneswide <- hormones %>% pivot_wider(names_from = "hormone", values_from = "plasma_conc", 
+                                         values_fn = list(plasma_conc = mean))
+hormoneswide$treatment <- factor(hormoneswide$treatment, levels = alllevels)
+head(hormoneswide)
+
+# for facet wrapping
 hormones <- left_join(hormones, birds)
@@
 
-hormones$okay <- ifelse(hormones$hormone == "cort" & hormones$plasma_conc > 30, "bad",
-                    ifelse(hormones$hormone == "progesterone" & hormones$plasma_conc > 5, "bad", 
-                           ifelse(hormones$hormone == "prolactin" & hormones$plasma_conc > 150, "bad", 
-                        ifelse(hormones$hormone == "testosterone" & hormones$sex == "female", "bad",
-                               ifelse(hormones$hormone == "estradiol" & hormones$sex == "male", "bad", "okay")))))
-hormones <- hormones %>% filter(okay == "okay") %>% droplevels()
-summary(hormones)
+hormones$hormone <- factor(hormones$hormone, 
+                           levels = c("prolactin", "corticosterone", "progesterone", 
+                                      "estradiol", "testosterone"))
+head(hormones)
 ```
@@
 
-```
-
-```{r PRLonly}
 hormonecharplot("prolactin", "PRL (ng/mL)")
+hormonecharplot("corticosterone", "CORT (ng/mL)")
+hormonecharplot("estradiol", "e2 (ng/mL)")
+hormonecharplot("testosterone", "t (ng/mL)")
+hormonecharplot("progesterone", "p4 (ng/mL)")
 ```
@@
 
-hormonecharplot("cort", "CORT (ng/mL)")
+hormonecharplot("corticosterone", "CORT (ng/mL)")
 hormonecharplot("progesterone", "PROG (ng/mL)") 
@@
 
-hormonemanipSteroids("cort", "CORT (ng/mL)", 10)
+hormonemanipSteroids("corticosterone", "CORT (ng/mL)", 10)
 hormonemanipSteroids("progesterone", "PROG (ng/mL)",2.5)
@@
 prog.char <- hormones %>% filter(hormone == "progesterone", treatment %in% charlevels)   %>%  droplevels()
-cort.char <- hormones %>% filter(hormone == "cort", treatment %in% charlevels)   %>%  droplevels()
+cort.char <- hormones %>% filter(hormone == "corticosterone", treatment %in% charlevels)   %>%  droplevels()
 
@@
 aovSexTretment(prl.char, "PRL") # yes, sex difference (p = 0.00256), yes treatment effect (p < 2e-16), no interaction
-aovSexTretment(cort.char, "CORT") # no sex difference (p = 0.779 ), small treatment effect (p = 0.0238),  no interaction
+aovSexTretment(cort.char, "corticosterone") # no sex difference (p = 0.779 ), small treatment effect (p = 0.0238),  no interaction
 aovSexTretment(prog.char, "PROG") # no sex difference or treatment effect, signifiant interacion (p = 0.0107)
@@
 
-```{r}
-write.csv(hormones, "../results/hormones.csv", row.names = F)
-write.csv(prolactin2, "../results/07_prolactin2.csv", row.names = F)
-
-```
-
 ## for rechelle
@@
 rechelle <- hormones %>%
-  dplyr::filter(hormone %in% c("prolactin", "cort"),
+  dplyr::filter(hormone %in% c("prolactin", "corticosterone"),
                 treatment %in% c("lay", "inc.d9", "hatch", "n9")) %>%
-  dplyr::select(-icons, -music, -iconpath, -okay)
+  dplyr::select(-icons, -music, -iconpath)
 
-rechelle$hormone <- factor(rechelle$hormone, levels = c("prolactin", "cort"))
+rechelle$hormone <- factor(rechelle$hormone, levels = c("prolactin", "corticosterone"))
 
 prl.rechelle <- rechelle %>% filter(hormone == "prolactin")   %>%  droplevels()
-cort.rechelle <- rechelle %>% filter(hormone == "cort")   %>%  droplevels()
+cort.rechelle <- rechelle %>% filter(hormone == "corticosterone")   %>%  droplevels()
 
@@
 ```{r sex}
-hormones$hormone <- factor(hormones$hormone, 
-                           levels = c("cort", "progesterone", "prolactin", 
-                                      "estradiol", "testosterone"))
-
 hormones %>%
@@
 
-printaovtablesex("cort", "CORT ~ sex")
-printaovtablesex("progesterone", "Prog ~ sex")
 printaovtablesex("prolactin", "PRL ~ sex")
-
-TukeyHSD(aov(data = cort.char, plasma_conc ~ treatment), which = "treatment")
-
 ```
@@
 hormones %>%
-  filter(hormone == "cort") %>%
+  filter(hormone == "corticosterone") %>%
   group_by(sex) %>%
@@
 p1 <- hormones %>%
-  filter(hormone == "cort",
+  filter(hormone == "corticosterone",
          study == "characterization") %>%
@@
   theme(legend.position = "none") +
-  scale_y_continuous(limits = c(0,8.5))
+  scale_y_continuous(limits = c(0,8.5)) +
+  scale_fill_manual(values = sexcolors)
 p1
@@
 cort <- hormones %>%
-  filter(hormone == "cort",
+  filter(hormone == "corticosterone",
          study == "characterization")
@@
 p3 <- hormones %>%
-  filter(hormone == "cort",
+  filter(hormone == "corticosterone",
          treatment %in% c(controlsremoval, levelsremoval )) %>% 
@@
 p4 <- hormones %>%
-  filter(hormone == "cort",
+  filter(hormone == "corticosterone",
          treatment %in% c(controlstiming, levelstiming )) %>% 
@@
 ```{r correlations2 }
-hormoneswide <- hormones %>%
-  select(study, treatment, sex, bird_id, hormone, plasma_conc) %>%
-  pivot_wider( names_from = hormone, values_from  = plasma_conc, values_fn = list(plasma_conc = mean))
 
 plotcorrelations <- function(i,j, myxlab, mylab, mysubtitle){
-      p <- ggplot(hormoneswide, aes(x = i, y = j),na.rm = TRUE) +
+      p <- ggplot(hormoneswide, aes(x = i, y = j), na.rm = TRUE) +
        geom_point(aes(color = treatment), na.rm = TRUE) +
@@
 
-p1 <- plotcorrelations(hormoneswide$prolactin, hormoneswide$cort, "prolactin", "cort", "r = 0.221, p = 0.001")
-p2 <- plotcorrelations(hormoneswide$prolactin, hormoneswide$progesterone, "prolactin", "progesterone", "r = -0.099, p = 0.2")
-p3 <- plotcorrelations(hormoneswide$prolactin, hormoneswide$testosterone, "prolactin", "testosterone", "r = -0.258, p = 0.08")
-p4 <- plotcorrelations(hormoneswide$prolactin, hormoneswide$estradiol, "prolactin", "estradiol", "r = 0.05, p = 0.6")
+hormoneswide
 
-p5 <- plotcorrelations(hormoneswide$cort, hormoneswide$progesterone, "cort", "progesterone", "r = 0.104, p = 0.175")
-p6 <- plotcorrelations(hormoneswide$cort, hormoneswide$estradiol, "cort", "estradiol", "r = -0.026, p = 0.8")
-p7 <- plotcorrelations(hormoneswide$cort, hormoneswide$testosterone, "cort", "testosterone", "r = -0.108, p = 0.4")
+corrPRL <- hormoneswide %>%
+  pivot_longer(cols = corticosterone:progesterone, names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
+  filter(study == "characterization") %>%
+  ggplot(aes(x = prolactin, y = conc)) +
+  geom_point(aes(color = treatment)) +
+  geom_smooth( color = "darkgrey", method = "lm") +
+  facet_wrap(~hormones, nrow = 1, scales = "free_y")  +
+  theme_B3() +
+  theme(legend.position = "bottom",
+        legend.title = element_blank()) +
+  scale_color_manual(values = colorscharmaip2) +
+  guides(color = guide_legend(nrow = 1)) +
+  labs(y = "conc. (ng/mL)", x = "prolactin (ng/mL)")
+corrPRL
 
-p8 <- plotcorrelations(hormoneswide$progesterone, hormoneswide$estradiol, "progesterone", "estradiol", "r = 0.120, p = 0.3")
-p9 <- plotcorrelations(hormoneswide$progesterone, hormoneswide$testosterone, "progesterone", "testosterone", "r = 0.133, p = 0.3")
-
-allcorrs <- plot_grid(p1,p2,p3,p4,p5,p6,p7,p8,p9, nrow = 3, align = "hv", labels = "auto", label_size = 8)
-allcorrs
-
-plot_grid(p1 + theme(axis.text.x = element_text(angle = 0)),p3+ theme(axis.text.x = element_text(angle = 0))  + theme(legend.position = "right"), align = "hv", rel_widths = c(0.4,0.6))
-
-cor.test(hormoneswide$prolactin, hormoneswide$cort)
-cor.test(hormoneswide$prolactin, hormoneswide$progesterone)
-cor.test(hormoneswide$prolactin, hormoneswide$estradiol)
-cor.test(hormoneswide$prolactin, hormoneswide$testosterone)
-
-cor.test(hormoneswide$cort, hormoneswide$estradiol)
-cor.test(hormoneswide$cort, hormoneswide$testosterone)
-cor.test(hormoneswide$cort, hormoneswide$progesterone)
-
-cor.test(hormoneswide$progesterone, hormoneswide$estradiol)
-cor.test(hormoneswide$progesterone, hormoneswide$testosterone)
+corrCORT <- hormoneswide %>%
+  pivot_longer(cols = c(prolactin, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
+  filter(study == "characterization") %>%
+  ggplot(aes(x = corticosterone, y = conc)) +
+  geom_point(aes(color = treatment)) +
+  geom_smooth( color = "darkgrey", method = "lm") +
+  facet_wrap(~hormones, nrow = 1, scales = "free_y")  +
+  theme_B3() +
+  theme(legend.position = "bottom",
+        legend.title = element_blank()) +
+  scale_color_manual(values = colorscharmaip2) +
+  guides(color = guide_legend(nrow = 1)) +
+  labs(y = "conc. (ng/mL)", x = "corticosterone (ng/mL)")
+corrCORT
 ```

### 2020-01-22 analysis/20_forrechelle.Rmd
new PRL figs

@@
 library(corrr)
+library(apaTables)
 
@@
 
-```{r forrechelle}
+```{r candidate genes}
 # check which genese appear in multiple tissue
@@
 head(dft)
+```
 
+```{r victoria-ePRL-data}
+ePRL <- df2 %>% dplyr::rename(value = vsd, measurement = gene) %>% filter(measurement == "PRL") %>% droplevels()
+ePRLR <- df2 %>% dplyr::rename(value = vsd, measurement = gene) %>% filter(measurement == "PRLR") %>% droplevels()
+
+df4 <- df2 %>% dplyr::rename(value = vsd, measurement = gene) %>% filter(measurement %in% candidategenes)  %>%
+  pivot_wider(names_from = measurement, values_from = value)  %>%
+  drop_na() %>% droplevels()
+```
+
+```{r victoria-ePRL, fig.height=6}
+
+a <- ggplot(ePRL, aes(x = treatment, y = value)) +
+  geom_boxplot(aes(fill = treatment, color = sex)) +
+  facet_wrap(~tissue, scales = "free") +
+  theme_B3() +
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank(),
+        legend.position = "top",
+        plot.subtitle = element_text(face = "italic"),
+        axis.title.y = element_text(face = "italic"),
+        legend.title = element_blank(),
+        legend.key.size =  unit(0.25, "cm")) +
+  labs(x = "parental stage", y = "PRL") +
+  scale_color_manual(values =  sexcolors ) +
+  scale_fill_manual(values= colorschar) +
+  guides(fill = guide_legend(nrow = 1))
+
+b <- ggplot(ePRL, aes(x = treatment, y = value, color = sex)) +
+  geom_smooth(aes(x = as.numeric(treatment), y = value), se = T, na.rm  = TRUE) +
+  facet_wrap(~tissue, scales = "free")  +
+  theme_B3() +
+  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9),
+                     labels = charlevels) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        strip.text = element_blank(),
+        legend.position = "none",
+        legend.title = element_blank(),
+        axis.title.y = element_text(face = "italic")) +
+  labs(x = "parental stage", y = "PRL") +
+  scale_color_manual(values = sexcolors)
+
+plot_grid(a,b, nrow = 2, align = "v", labels = c("a", "b"), label_size = 8,
+          rel_heights = c(1.25,1))
+
+```
+
+```{r victoria-ePRL-corr}
 x <- correlate(dft)
@@
 
-a <- x %>%
+c <- x %>%
   focus(hyp_PRL) %>%
@@
 
-b <- x %>%
+d <- x %>%
   focus(pit_PRL) %>%
@@
 
-c <- x %>%
+e <- x %>%
   focus(gon_PRL) %>%
@@
 
-plot_grid(a,b, c, nrow = 1, rel_widths = c(1.1,1,1))
+cde <- plot_grid(c,d, e, nrow = 1, rel_widths = c(1.1,1,1))
+cde
+
+f <- ggplot(df4, aes(x = PRL, y = PRLR)) +
+  geom_point(aes(color = treatment)) +
+  geom_smooth(method = "lm",  na.rm = T, aes(color = sex)) +
+  facet_wrap(~tissue, scales = "free") +
+  theme_B3() +
+  theme(legend.position = "none",
+        axis.title = element_text(face = "italic"),
+        #strip.text = element_blank(),
+        legend.title = element_blank()) +
+  guides(color = guide_legend(nrow = 1)) +
+   scale_color_manual(values = colorsvolcano)
+
+g <- ggplot(df4, aes(x = PRL, y = CISH)) +
+  geom_point(aes(color = treatment)) +
+  geom_smooth(method = "lm",  na.rm = T, aes(color = sex)) +
+  facet_wrap(~tissue, scales = "free") +
+  theme_B3() +
+  theme(legend.position = "none",
+        axis.title = element_text(face = "italic"),
+        strip.text = element_blank(),
+        legend.title = element_blank()) +
+  guides(color = guide_legend(nrow = 1)) +
+   scale_color_manual(values = colorsvolcano)
+
+h <- ggplot(df4, aes(x = PRL, y = GH)) +
+  geom_point(aes(color = treatment)) +
+  geom_smooth(method = "lm",  na.rm = T, aes(color = sex)) +
+  facet_wrap(~tissue, scales = "free") +
+  theme_B3() +
+  theme(legend.position = "bottom",
+        axis.title = element_text(face = "italic"),
+        strip.text = element_blank(),
+        legend.title = element_blank()) +
+  guides(color = guide_legend(nrow = 1)) +
+   scale_color_manual(values = colorsvolcano)
+
+plot_grid(f + theme(axis.title.x = element_blank()), 
+          g + theme(axis.title.x = element_blank()), h, 
+          nrow = 3, rel_heights = c(1.2,1,1.2), align = "v")
 ```
+
+```{r victoria-ePRL-stats}
+
+aov_all = data.frame()
+for(i in charlevels){
+  for(j in tissuelevels){
+
+  df <- df4 %>% 
+    filter(treatment == i) %>% 
+     filter(tissue == j) %>% 
+    droplevels()
+  aovtable <- apa.aov.table(aov(PRL ~ sex , data=df))
+  aovtable <- as.data.frame(aovtable$table_body)
+  totaldf <- aovtable[3, 3]
+  aovtable$df <- paste(aovtable$df, ", " , totaldf, sep = "")
+  aovtable$ANOVA <- "PRL ~ sex"
+  aovtable$stages <- paste(i)
+  aovtable$tissue <- paste(j)
+  aovtable$p <- as.numeric(as.character(aovtable$p))
+  aov_all <- rbind(aov_all,aovtable)
+
+  }
+}
+
+aov_all$tissue <- factor(aov_all$tissue, levels = tissuelevels)
+
+aov_stats <- aov_all %>%
+  filter(Predictor == "sex")  %>%
+  select(stages, tissue, ANOVA, df, "F", p) %>%
+  mutate(sig = ifelse(p < 0.05, "*", " ")) %>%
+  arrange(tissue)
+
+aov_stats
+write_csv(aov_stats, "../results/20_ePRL_fig1.csv")
+
+aovall2 = data.frame()
+  for(j in tissuelevels){
+
+  df <- df4 %>% 
+     filter(tissue == "hypothalamus") %>% 
+    droplevels()
+
+  aovtable <- apa.aov.table(aov(PRL ~ treatment , data=df))
+  aovtable <- as.data.frame(aovtable$table_body)
+  totaldf <- aovtable[3, 3]
+  aovtable$df <- paste(aovtable$df, ", " , totaldf, sep = "")
+  aovtable$ANOVA <- "PRL ~ treatment"
+  aovtable$tissue <- paste(j)
+  aovtable$p <- as.numeric(as.character(aovtable$p))
+  aovall2 <- rbind(aovall2,aovtable)
+ }
+
+aovall2
+
+aov_stats2 <- aovall2 %>%
+  filter(Predictor == "treatment")  %>%
+   rename(stages = Predictor) %>%
+  select(stages, tissue, ANOVA, df, "F", p) %>%
+  mutate(sig = ifelse(p < 0.05, "*", " ")) %>%
+  arrange(tissue)
+
+aov_stats2
+
+aov1 <- aov(data = df4, PRL ~  sex * treatment * tissue )
+summary(aov1)
+
+hyp <- df4 %>% filter(tissue == "hypothalamus")
+aov4 <- aov(data = hyp, PRL ~  treatment)
+thsdhyp <- TukeyHSD(aov4, which = "treatment")
+thsdhyp <- as.data.frame(thsdhyp$treatment)
+thsdhyp$comparission <- row.names(thsdhyp) 
+thsdhyp <- thsdhyp %>% 
+  dplyr::filter(`p adj` < 0.05) %>%
+  dplyr::mutate(tissue = "hypothalamus") %>%
+  select(tissue, comparission, `p adj`, diff, lwr, upr)
+thsdhyp
+
+pit <- df4 %>% filter(tissue == "pituitary")
+aov5 <- aov(data = pit, PRL ~  treatment)
+thsdpit <- TukeyHSD(aov5, which = "treatment")
+thsdpit <- as.data.frame(thsdpit$treatment)
+thsdpit$comparission <- row.names(thsdpit) 
+thsdpit <- thsdpit %>% 
+  dplyr::filter(`p adj` < 0.05) %>%
+  dplyr::mutate(tissue = "pituitary") %>%
+  select(tissue, comparission, `p adj`, diff, lwr, upr)
+thsdpit
+
+fgonads <- df4 %>% filter(tissue == "gonads", sex == "female")
+aov2 <- aov(data = fgonads, PRL ~  treatment)
+thsdfgon <- TukeyHSD(aov2, which = "treatment")
+thsdfgon <- as.data.frame(thsdfgon$treatment)
+thsdfgon$comparission <- row.names(thsdfgon) 
+thsdfgon <- thsdfgon %>% 
+  dplyr::filter(`p adj` < 0.05) %>%
+  dplyr::mutate(tissue = "male gonads") %>%
+  select(tissue, comparission, `p adj`, diff, lwr, upr)
+thsdfgon
+
+mgonads <- df4 %>% filter(tissue == "gonads", sex == "male")
+aov3 <- aov(data = mgonads, PRL ~  treatment)
+thsdmgon <- TukeyHSD(aov3, which = "treatment")
+thsdmgon <- as.data.frame(thsdmgon$treatment)
+thsdmgon$comparission <- row.names(thsdmgon) 
+thsdmgon <- thsdmgon %>% 
+  dplyr::filter(`p adj` < 0.05) %>%
+  dplyr::mutate(tissue = "male gonads") %>%
+  select(tissue, comparission, `p adj`, diff, lwr, upr) 
+thsdmgon
+
+addtreatment <- rbind(thsdhyp, thsdpit, thsdfgon, thsdmgon) %>%
+  dplyr::mutate(diff = round(diff,2),
+                `p adj` = round(`p adj`,4),
+                lwr = round(lwr, 2),
+                upr = round(upr, 2))
+addtreatment
+
+write_csv(addtreatment, "../results/20_ePRL_stats.csv")
+```

### 2020-01-22 analysis/16_PRL.Rmd
new music figure

--- 2020-01-05 analysis/16_PRL.Rmd
@@
 ```{r BRCA.pit.maip}
-
 p5 <- ggplot(candidates.pit, aes(x = treatment, y = BRCA1)) + 
@@
 geneshormones <- left_join(candidates.pit, hormones) %>%
-  select(-X1, -V1, -okay, - icons, -music, -iconpath) %>%
+  select( -V1 ) %>%
   drop_na()
@@
 
-```{r music}
+```{r music, fig.height=6, fig.width=6.5}
 
@@
 
+p1 <- hormones %>% 
+    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
+    droplevels() %>% 
+  ggplot(aes(x = treatment, y = plasma_conc)) +
+    geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) +
+    theme_B3() +
+    scale_fill_manual(values = colorscharmaip) +
+    scale_color_manual(values = sexcolors) +
+    labs(y = "prolactin (ng/mL)", x = NULL) +
+    scale_alpha_manual(values = c(0.75,1)) +
+    theme(legend.position = "top", legend.direction = "horizontal",
+          axis.title.x = element_blank(),
+          axis.text.x = element_blank(),
+          legend.title = element_blank(),
+          legend.key.size =  unit(0.25, "cm")) + 
+  labs( y = "prolactin (ng/mL)") +
+   guides(fill = guide_legend(nrow = 1))
+
+p2 <- candidates.pit %>%
+  dplyr::filter(treatment %in% charlevels) %>%
+  ggplot(aes(x = treatment, y = PRL)) + 
+    geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) + 
+    scale_alpha_manual(values = c(0.75,1)) +
+    theme_B3() +
+  theme(legend.position = "none") + 
+  scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+  labs(y = "Pituitary PRL", x = "parental stage" ) +
+  theme(axis.title.y  = element_text(face = "italic"),
+        axis.text.x = element_blank())  +
+  annotation_custom(control, ymin = 13.5, ymax = 15, xmin = -7.5) +
+  annotation_custom(bldg, ymin = 13.5, ymax = 15, xmin = -5.5) +
+  annotation_custom(lay, ymin = 13.5, ymax = 15, xmin = -3.5) +
+  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = -1.5) +
+  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = 0.5) +
+  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = 2.5) +
+  annotation_custom(hatch, ymin = 13.5, ymax = 15, xmin = 4.5) +
+  annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 6.5) +
+  annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 8.5)
+
 p8 <- ggplot(meanPRL, aes(treatment, m)) +
    geom_image(aes(image=music), size=.12)  +
-  labs(x = "parental care stages", y = "PRL", subtitle = "Pituitary") +
+  labs(x = NULL, y = "prolactin symphony") +
   mytheme() +
   ylim(c(15,22)) +
-  annotation_custom(control, ymin = 15, ymax = 16.5, xmin = -7.5) +
-  annotation_custom(bldg, ymin = 15, ymax = 16.5, xmin = -5.5) +
-  annotation_custom(lay, ymin = 15, ymax = 16.5, xmin = -3.5) +
-  annotation_custom(inc, ymin = 15, ymax = 16.5, xmin = -1.5) +
-  annotation_custom(inc, ymin = 15, ymax = 16.5, xmin = 0.5) +
-  annotation_custom(inc, ymin = 15, ymax = 16.5, xmin = 2.5) +
-  annotation_custom(hatch, ymin = 15, ymax = 16.5, xmin = 4.5) +
-  annotation_custom(nestling, ymin = 15, ymax = 16.5, xmin = 6.5) +
-  annotation_custom(nestling, ymin = 15, ymax = 16.5, xmin = 8.5)   +
-  geom_hline(yintercept=17) +
-  geom_hline(yintercept=18) +
-  geom_hline(yintercept=19) +
-  geom_hline(yintercept=20) +
-  geom_hline(yintercept=21) +
-  theme(axis.title.y  = element_text(face = "italic"))
+  theme(axis.text = element_blank())
 
-p8
+plot_grid(p1,p2,p8, nrow = 3, align = "hv", labels = "auto", label_size = 8)
 

### 2020-01-22 analysis/20_forrechelle.Rmd
new music figure

@@
 candidategenes <- c("PRLR", "PRL", "CISH", "JAK2", "SOCS3", "GH" )
+datadrivengenes <- c("COX1", "PRL", "CHGA", "ATP2B4", "LOC107055658")
 timepoints <- charlevels
@@
 
-```{r data}
+```{r data, message=F, results=F}
 datapath <- "../results/"   # path to the data
@@
 tissuespecificity
+
 write.csv(tissuespecificity, "../results/tissuespecificity.csv")
@@
 df2$tissue <- factor(df2$tissue, levels = tissuelevels)
+
+df4 <-  df  %>%
+  dplyr::filter(X1 %in% datadrivengenes) %>%
+  pivot_longer(cols = L.Blu13_male_gonad_control.NYNO:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "sample", values_to = "vsd") %>%
+   mutate(sex = sapply(strsplit(as.character(sample),'_'), "[", 2)) %>%
+   mutate(treatment = sapply(strsplit(as.character(sample),'_'), "[", 4))  %>%
+   mutate(treatment = sapply(strsplit(as.character(treatment),'.NYNO'), "[", 1)) %>%
+  mutate(bird = sapply(strsplit(as.character(sample),'_'), "[", 1)) %>%
+  dplyr::filter(treatment %in% timepoints) %>%
+  dplyr::select(bird, sex, treatment, tissue, X1, vsd) %>%
+  mutate(tissue = fct_recode(tissue, "hypothalamus" = "hyp",
+                    "pituitary" = "pit",
+                    "gonads" = "gon"
+                    )) %>%
+  rename(gene = X1) %>%
+  drop_na() %>%
+  droplevels()
+head(df4)
 

### 2020-01-22 analysis/16_PRL.Rmd
partial new color scheme

@@
     theme_B3() +
-    scale_fill_manual(values = colorscharmaip) +
+    scale_fill_manual(values = colorscharnew) +
     scale_color_manual(values = sexcolors) +
@@
 candidates.pit$treatment <- factor(candidates.pit$treatment, levels = alllevels)
+candidates.pit$study2 <- ifelse(candidates.pit$treatment %in% charlevels, "charcterization",
+                              ifelse(candidates.pit$treatment %in% levelstiming, "timing",
+                                     ifelse(candidates.pit$treatment %in% levelsremoval, "removal", NA)))
 tail(candidates.pit)
+
 ```
@@
   theme(legend.position = c(0.85,0.2), legend.direction = "horizontal") + 
-  scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+  scale_color_manual(values = sexcolors) +
+  scale_fill_manual(values = colorscharnew) +
   labs(y = "PRL", x = "increasing time >>", subtitle = "Pituitary" ) +
@@
       scale_alpha_manual(values = c(0.75,1)) +
-  scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+  scale_color_manual(values = sexcolors) +
+  scale_fill_manual(values = colorscharnew) +
   theme(legend.position = "none") + 
@@
    theme(legend.position = "none", legend.direction = "horizontal") + 
-   scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+   scale_color_manual(values = sexcolors) +
+  scale_fill_manual(values = colorscharmaip) +
     labs(x = "increasing time >>", y = "PRL", subtitle = "Pituitary") +
@@
 p4
+
+candidates.pit$treatment <- factor(candidates.pit$treatment, levels = alllevels3)
+
+ggplot(candidates.pit, aes(x = treatment, y = PRL)) + 
+    geom_boxplot(aes(fill = treatment, alpha = sex), width=0.9) +
+    scale_alpha_manual(values = c(0.75,1), guide=FALSE) +
+    theme_B3() +
+   theme(legend.position = "bottom", 
+         legend.direction = "horizontal",
+         legend.title = element_blank(),
+         axis.text.x = element_text(angle = 45, hjust = 1),
+         axis.title.y = element_text(face = "italic")) + 
+   scale_color_manual(values = sexcolors, guide=FALSE) +
+  scale_fill_manual(values = colorscharmaip) +
+    labs(x = "parental stage", y = "PRL", subtitle = "Pituitary") +
+  ylim(c(14,22)) +
+  theme()  +
+  facet_wrap(~study2, scales = "free_x", shrink  = T) +
+     guides(fill = guide_legend(nrow = 3, byrow = TRUE))
 
@@
         subtitle = "Genes that are coexpressed with prolactin and regulate development")
-p1
+i
 

### 2020-01-22 analysis/16_PRL.Rmd
new colors

@@
   ggplot(aes(x = treatment, y = plasma_conc)) +
-    geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) +
+    geom_boxplot(aes(fill = treatment, color = sex)) +
     theme_B3() +
@@
          color = guide_legend(order=1)) +
-    scale_alpha_manual(values = c(0.75,1)) +
+    #scale_alpha_manual(values = c(0.75,1)) +
     theme(legend.position = c(0.85,0.15), legend.direction = "horizontal") + 
@@
   ggplot(aes(x = treatment, y = PRL)) + 
-    geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) + 
+    geom_boxplot(aes(fill = treatment, color = sex)) + 
     scale_alpha_manual(values = c(0.75,1)) +
@@
   dplyr::filter(treatment %in% charlevels) %>%
-  ggplot(aes(x = treatment, y = BRCA1, fill = treatment, color = sex, alpha = sex)) +
+  ggplot(aes(x = treatment, y = BRCA1, fill = treatment, color = sex)) +
     geom_boxplot() +
@@
     theme_B3() +
-   theme(legend.position = "none", legend.direction = "horizontal") + 
+   theme(legend.position = "none", legend.direction = "horizontal",
+         axis.text.x = element_text(angle = 45, hjust = 1)) + 
    scale_color_manual(values = sexcolors) +
   scale_fill_manual(values = colorscharmaip) +
-    labs(x = "increasing time >>", y = "PRL", subtitle = "Pituitary") +
+    labs(x = "parental stage", y = "PRL", subtitle = "Pituitary") +
   scale_fill_manual(values = colorscharmaip2) +
   ylim(c(13,22.5)) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  annotation_custom(control, ymin = 13, ymax = 14, xmin = -14.8) +
-  annotation_custom(bldg, ymin = 13, ymax = 14, xmin = -12.8) +
-  annotation_custom(lay, ymin = 13, ymax = 14, xmin = -10.8) +
-  annotation_custom(inc, ymin = 13, ymax = 14, xmin = -8.8) +
-  annotation_custom(removeegg, ymin = 13, ymax = 14, xmin = -6.8) +
-  annotation_custom(inc, ymin = 13, ymax = 14, xmin = -4.8) +
-  annotation_custom(maniphatch, ymin = 13, ymax = 14, xmin = -2.8) +
-  annotation_custom(removeegg, ymin = 13, ymax = 14, xmin = -0.8) +
-  annotation_custom(inc, ymin = 13, ymax = 14, xmin = 1.4) +
-  annotation_custom(removeegg, ymin = 13, ymax = 14, xmin = 3.4)  +
-  annotation_custom(manipinc, ymin = 13, ymax = 14, xmin = 5.4) +
-  annotation_custom(hatch, ymin = 13, ymax = 14, xmin = 7.4) +
-  annotation_custom(removechick, ymin = 13, ymax = 14, xmin = 9.4) +
-  annotation_custom(maniphatch, ymin = 13, ymax = 14, xmin = 11.4) +
-  annotation_custom(nestling, ymin = 13, ymax = 14, xmin = 13.4) +
-  annotation_custom(nestling, ymin = 13, ymax = 14, xmin = 15.4) +
     theme(axis.title.y  = element_text(face = "italic"))
 p4
-
-candidates.pit$treatment <- factor(candidates.pit$treatment, levels = alllevels3)
-
-ggplot(candidates.pit, aes(x = treatment, y = PRL)) + 
-    geom_boxplot(aes(fill = treatment, alpha = sex), width=0.9) +
-    scale_alpha_manual(values = c(0.75,1), guide=FALSE) +
-    theme_B3() +
-   theme(legend.position = "bottom", 
-         legend.direction = "horizontal",
-         legend.title = element_blank(),
-         axis.text.x = element_text(angle = 45, hjust = 1),
-         axis.title.y = element_text(face = "italic")) + 
-   scale_color_manual(values = sexcolors, guide=FALSE) +
-  scale_fill_manual(values = colorscharmaip) +
-    labs(x = "parental stage", y = "PRL", subtitle = "Pituitary") +
-  ylim(c(14,22)) +
-  theme()  +
-  facet_wrap(~study2, scales = "free_x", shrink  = T) +
-     guides(fill = guide_legend(nrow = 3, byrow = TRUE))
 
@@
    scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
-  labs(x = "increasing time >>", y = "BRCA1", subtitle = "Pituitary") +
+  labs(x = "parental stage", y = "BRCA1", subtitle = "Pituitary") +
   scale_fill_manual(values = colorscharmaip2) +
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  annotation_custom(control, ymin = 13, ymax = 14, xmin = -14.8) +
-  annotation_custom(bldg, ymin = 13, ymax = 14, xmin = -12.8) +
-  annotation_custom(lay, ymin = 13, ymax = 14, xmin = -10.8) +
-  annotation_custom(inc, ymin = 13, ymax = 14, xmin = -8.8) +
-  annotation_custom(removeegg, ymin = 13, ymax = 14, xmin = -6.8) +
-  annotation_custom(inc, ymin = 13, ymax = 14, xmin = -4.8) +
-  annotation_custom(maniphatch, ymin = 13, ymax = 14, xmin = -2.8) +
-  annotation_custom(removeegg, ymin = 13, ymax = 14, xmin = -0.8) +
-  annotation_custom(inc, ymin = 13, ymax = 14, xmin = 1.4) +
-  annotation_custom(removeegg, ymin = 13, ymax = 14, xmin = 3.4)  +
-  annotation_custom(manipinc, ymin = 13, ymax = 14, xmin = 5.4) +
-  annotation_custom(hatch, ymin = 13, ymax = 14, xmin = 7.4) +
-  annotation_custom(removechick, ymin = 13, ymax = 14, xmin = 9.4) +
-  annotation_custom(maniphatch, ymin = 13, ymax = 14, xmin = 11.4) +
-  annotation_custom(nestling, ymin = 13, ymax = 14, xmin = 13.4) +
-  annotation_custom(nestling, ymin = 13, ymax = 14, xmin = 15.4) +
     theme(axis.title.y  = element_text(face = "italic"))
 p5
+
+plot_grid(p4 + theme(axis.title.x = element_blank(),
+                     axis.text.x = element_blank()),
+          p5 + labs(subtitle = NULL), nrow = 2, 
+          rel_heights = c(0.45,0.55))
 
@@
   labs(subtitle = "Pituitary") +
+   theme_B3() +
   theme(axis.title  = element_text(face = "italic"),
@@
 ggplot( aes(x = PRL, y = BRCA1)) +
-  geom_point(aes(alpha = sex,  color = 	treatment)) +
+  geom_point(aes(color = 	treatment)) +
   geom_smooth(method = "loess", se=T, color = "darkgrey") +
@@
 
-```
-
-## Genes an hormones (in progress, don't have all the bird ids lining up.)
-
-```{r geneshormones}
-candidates.pit$bird_id <- candidates.pit$bird
-candidates.pit$X1 <- candidates.pit$V1
-
-geneshormones <- left_join(candidates.pit, hormones) %>%
-  select( -V1 ) %>%
-  drop_na()
-
-head(geneshormones$bird_id)
-head(hormones$bird_id)
-
-ggplot(geneshormones, aes(x = plasma_conc, y = PRL)) +
-  geom_point(aes(color = treatment)) +
-  facet_wrap(~hormone, scales = "free_x") +
-    geom_smooth(method = "lm", color = "grey")
-
-prolactin2 <- read_csv("../results/07_prolactin2.csv")
-head(prolactin2)
-geneshormones2 <- left_join(candidates.pit, prolactin2) %>%
-  drop_na()
-head(geneshormones)
-
-ggplot(geneshormones2, aes(x = plasma_conc, y = PRL)) +
-  geom_point(aes( color = treatment)) +
-  geom_smooth(method = "lm", color = "grey") +
-  facet_wrap(~hormone, scales = "free") 
 ```
@@
     geom_boxplot(aes(fill = treatment, alpha = sex, color = sex)) + 
+    scale_fill_manual(values = colorscharmaip) +
     scale_alpha_manual(values = c(0.75,1)) +
@@
   theme(legend.position = "none") + 
-  scale_color_manual(values = c("female" = "#969696", "male" = "#525252")) +
+  scale_color_manual(values = sexcolors) +
   labs(y = "Pituitary PRL", x = "parental stage" ) +
@@
 
+## genes
+
+```{r}
+geneinfo <- read_csv("../metadata/00_geneinfo.csv")
+```
+
 ## WGCNA candidates
@@
 #              "MYC", "CCNB3", "FANCC", "STC1", "RACGAP1", "E2F7", "PLK1", "EXO1", "CENPI", "CDK1")
-WGCNAgenes <- c("PRL", "PAX7", "ASPM", "AURKA",  "CDC20", "CENPF", "MYC",  "CDK1", "LBH", "BRCA1", "FOXM1", "FBXO5")
+
+WGCNAgenes <- c("PRL", "PAX7", "ASPM", "AURKA",  "CDC20", "CENPF", "MYC",  
+                "CDK1", "LBH", "BRCA1", "FOXM1", "FBXO5")
 
@@
   labs( x = "parental stage", y = "pituitary expression",
-        subtitle = "Genes that are coexpressed with prolactin and regulate development")
-i
+        subtitle = "Genes that are coexpressed with prolactin and regulate development") +
+  scale_fill_manual(values = colorscharmaip)
+p1 
+```
+
+## calisis candidates
+
+```{r calisigenes, fig.height=6}
+
+Calisigenes <- c("PRL", "PRLR", 
+                 "VIP", "VIPR1", "VIPR2", 
+                 "OXT", "AVP", "AVPR1A", "AVPR1B", 
+                 "GNRH1","GNRHR", "NPVF",
+                 "NR3C1", "NR3C2",
+                 "ESR1", "ESR2"
+                 )
+
+calisisgenes.pit <- vsd.pit %>% dplyr::filter(genes %in% Calisigenes)
+calisisgenes.pit <- as.data.frame(calisisgenes.pit)
+row.names(calisisgenes.pit) <- calisisgenes.pit$genes
+calisisgenes.pit <- pivot_longer(calisisgenes.pit, -genes, names_to = "V1", values_to = "expression")
+head(calisisgenes.pit)
+calisisgenes.pit <- left_join(colData.pit, calisisgenes.pit) %>%
+  filter(study == "charcterization")
+calisisgenes.pit$treatment <- factor(calisisgenes.pit$treatment, levels = alllevels)
+
+calisisgenes.pit$genes <- factor(calisisgenes.pit$genes)
+calisisgenes.pit$genesnum <- round(as.numeric(calisisgenes.pit$genes), digits = -1)
+
+p2 <- ggplot(calisisgenes.pit, aes(x = treatment, y = expression)) +
+  geom_boxplot(aes(fill = treatment)) +
+  facet_wrap(~genes, scales = "free_y") +
+  theme_B3() +
+  theme(legend.position = "none") +
+  #scale_color_manual(values = sexcolors) +
+  theme( axis.text.x = element_text(angle = 45, hjust = 1),
+        strip.background = element_blank(),
+        strip.text = element_text(face = "italic")) +
+  labs( x = "parental stage", y = "pituitary expression",
+        subtitle = "Candidate genes")  +
+  scale_fill_manual(values = colorscharmaip) 
+p2
 

### 2020-01-22 analysis/20_forrechelle.Rmd
new colors

@@
 datadrivengenes <- c("COX1", "PRL", "CHGA", "ATP2B4", "LOC107055658")
+
+calisigenes <- c("PRL", "PRLR", 
+                 "VIP", "VIPR1", "VIPR2", 
+                 "OXT", "AVP", "AVPR1A", "AVPR1B", 
+                 "GNRH1","GNRHR", "NPVF",
+                 "NR3C1", "NR3C2",
+                 "ESR1", "ESR2", "AR")
+
 timepoints <- charlevels
@@
   select(tissue, X1, everything())
+
+genes <- as.data.frame(df$X1)
 ```
+## tissue specificiity
 
@@
 write.csv(tissuespecificity, "../results/tissuespecificity.csv")
+```
 
-genes <- as.data.frame(df$X1)
+## candidate genes
+
+```{r candidates}
 
 df2 <-  df  %>%
-  filter(X1 %in% candidategenes) %>%
+  filter(X1 %in% c(candidategenes)) %>%
   pivot_longer(cols = L.Blu13_male_gonad_control.NYNO:y98.o50.x_male_pituitary_inc.d3, 
@@
 df2$tissue <- factor(df2$tissue, levels = tissuelevels)
+df2$gene <- factor(df2$gene)
 
-df4 <-  df  %>%
-  dplyr::filter(X1 %in% datadrivengenes) %>%
-  pivot_longer(cols = L.Blu13_male_gonad_control.NYNO:y98.o50.x_male_pituitary_inc.d3, 
-               names_to = "sample", values_to = "vsd") %>%
-   mutate(sex = sapply(strsplit(as.character(sample),'_'), "[", 2)) %>%
-   mutate(treatment = sapply(strsplit(as.character(sample),'_'), "[", 4))  %>%
-   mutate(treatment = sapply(strsplit(as.character(treatment),'.NYNO'), "[", 1)) %>%
-  mutate(bird = sapply(strsplit(as.character(sample),'_'), "[", 1)) %>%
-  dplyr::filter(treatment %in% timepoints) %>%
-  dplyr::select(bird, sex, treatment, tissue, X1, vsd) %>%
-  mutate(tissue = fct_recode(tissue, "hypothalamus" = "hyp",
-                    "pituitary" = "pit",
-                    "gonads" = "gon"
-                    )) %>%
-  rename(gene = X1) %>%
-  drop_na() %>%
-  droplevels()
-head(df4)
+for (i in levels(df2$gene)) {
+  p <-  df2 %>%
+    filter(gene %in% i) %>%
+    ggplot(aes(x = treatment, y = vsd, color = sex, fill = treatment)) +
+    geom_boxplot() +
+    scale_alpha_manual(values = c(0.5, 1)) +
+    labs(y = "expression", x = "parental stage", subtitle = i) +
+    facet_wrap(~ tissue , scales = "free_y", nrow = 3) +
+    theme_B3() +
+    scale_color_manual(values = sexcolors) +
+    scale_fill_manual(values = colorscharmaip)
+ print(p)
+}
 
@@
 
-df2$gene <- factor(df2$gene)
-
-for (i in levels(df2$gene)) {
-  p <-  df2 %>%
-    filter(gene %in% i) %>%
-    ggplot(aes(x = treatment, y = vsd, color = sex, fill = treatment)) +
-    geom_boxplot() +
-    scale_alpha_manual(values = c(0.5, 1)) +
-    labs(y = "expression", x = "parental stage", subtitle = i) +
-    facet_wrap(~ tissue , scales = "free_y", nrow = 3) +
-    theme_B3() +
-    scale_color_manual(values = sexcolors) +
-    scale_fill_manual(values = colorscharmaip)
- print(p)
-}
-
-dft <- df %>% filter(X1 %in% candidategenes) %>%
-  unite("tissuegene", tissue:X1, remove = FALSE) %>%
-  select(-tissue, -`X1`, -filename)  %>%
-  pivot_longer(-tissuegene, names_to = "samples", values_to = "vsd") %>%
-  drop_na() %>%
-  mutate("bird" = sapply(strsplit(as.character(samples),'\\_'), "[", 1)) %>%
-  select(bird,tissuegene,vsd) %>%
-  filter(bird != "x.g.g.ATLAS") %>%
-  pivot_wider(id_cols = bird, names_from = tissuegene, values_from = vsd)
-dft <- as.data.frame(dft)
-row.names(dft) <- dft$bird
-dft$bird <- NULL
-head(dft)
 ```
@@
 ```{r victoria-ePRL-corr}
+
+dft <- df %>% filter(X1 %in% candidategenes) %>%
+  unite("tissuegene", tissue:X1, remove = FALSE) %>%
+  select(-tissue, -`X1`, -filename)  %>%
+  pivot_longer(-tissuegene, names_to = "samples", values_to = "vsd") %>%
+  drop_na() %>%
+  mutate("bird" = sapply(strsplit(as.character(samples),'\\_'), "[", 1)) %>%
+  select(bird,tissuegene,vsd) %>%
+  filter(bird != "x.g.g.ATLAS") %>%
+  pivot_wider(id_cols = bird, names_from = tissuegene, values_from = vsd)
+dft <- as.data.frame(dft)
+row.names(dft) <- dft$bird
+dft$bird <- NULL
+head(dft)
+
 x <- correlate(dft)

### 2020-01-22 analysis/17_pheatmap.Rmd
meh, heatmaps

@@
 knitr::opts_chunk$set(echo = TRUE)
+library(tidyverse)
 library(pheatmap)
@@
 ```{r, pheatmap, fig.height=6.5}
+calisigenes <- c("PRL", "PRLR", 
+                 "VIP", "VIPR1", "VIPR2", 
+                 "OXT", "AVP", "AVPR1A", "AVPR1B", 
+                 "GNRH1","GNRHR", "NPVF",
+                 "NR3C1", "NR3C2",
+                 "ESR1", "ESR2", "AR")
+
 prepvsds <- function(pathtofile){
@@
   df <- read.csv(pathtofile, row.names = 1)
-
-  df <- tail(df, 5000)
-  df <- head(df, 1000)
 
@@
 
+  df <- as.data.frame(df)
+  df$gene <- row.names(df)
+  df <- df %>% dplyr::filter(gene %in% calisigenes)
+  df <- as.data.frame(df)
+  row.names(df) <-  df$gene
+  df$gene <- NULL
+
   return(df)
@@
 pheatmap(vsd.hyp, show_colnames = F, annotation_col = hyplabels, 
-         annotation_colors = myannotationcolors, show_rownames = F)
+         annotation_colors = myannotationcolors, show_rownames = T)
 pheatmap(vsd.pit, show_colnames = F, annotation_col = pitlabels, 
-         annotation_colors = myannotationcolors, show_rownames = F)
+         annotation_colors = myannotationcolors, show_rownames = T)
 pheatmap(vsd.gon, show_colnames = F, annotation_col = gonlabels, 
-         annotation_colors = myannotationcolors, show_rownames = F)
+         annotation_colors = myannotationcolors, show_rownames = T)
+
 ```

### 2020-01-22 analysis/12_GO.Rmd
new candidate genes

@@
 title: "GO"
-output: html_document
+output: md_document
 ---
@@
 ```{r}
-candidates <- c(  "PRLR", 
-                 "AVP", "VIP", "AVPR1A", "AVPR1B", "OXT", 
-                 "GNRH1", "NPY",
-                "CRH", "FSHB", "GH", 
-                 "JAK2", "HCRT", "TRH", "TSHB",
-                "LH", "FSH",
-                "MC3R", "MC4R", "MC5R", 
+candidates <- c("PRL", "PRLR", 
+                 "AVP", "VIP", "AVPR1A", "AVPR1B", "OXT",  "VIPR1", "VIPR2",
+                "GNRH1", "NPY", "LH", "FSH",
+                "CRH", "FSHB", "GH",  "JAK2", "HCRT", "TRH", "TSHB",
+                "MC3R", "MC4R", "MC5R",  "NR3C1", "NR3C2",
                 "ESR1", "ESR2", "AR", 
-                "STAT5B", "CISH")
+                "STAT5B", "CISH",
+                "GNRHR", "NPVF")
+
 candidates <- as.data.frame(candidates)
@@
 ```{r}
-candidates <- c( "PRL", "PRLR", 
-                 "AVP", "VIP", "AVPR1A", "AVPR1B", "OXT", 
-                 "GNRH1", "GNRHR", "NPY",
-                "CRH", "FSHB", "GH", 
-                 "JAK2", "HCRT", "POMC", "TRH", "TSHB",
-                "LH", "FSH",
-                "POMC", "MC3R", "MC4R", "MC5R", 
-                "ESR1", "ESR2", "AR", 
-                "STAT5", "CISH")
-candidates <- as.data.frame(candidates)
-names(candidates)[1] <- "gene"
-candidates$GO <- "candidate genes"
-
 candidate_DEGs <- left_join(candidates, DEGs) %>%

### 2020-01-23 analysis/21_volcanobars.Rmd
now with resepct project

---
title: "volcanos"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE, message = F)
```

## DEGs

```{r volcanobars, DEGs}

DEG_path <- "../results/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "03_DEGs.*") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
DEG_files

allDEG <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'../results/03_DEGs.'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
   mutate(DEG = gsub("m.inc", "m_inc", DEG)) %>%
   mutate(DEG = gsub("inc.d", "inc_d", DEG))  %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\.'), "[", 2)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\.'), "[", 3)) %>%
  mutate(comparison = paste(down,up, sep = ".")) %>%
  select(tissue, comparison, direction, down, up, gene, lfc, padj, logpadj) 
head(allDEG)

allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevels)
allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
allDEG$direction <- factor(allDEG$direction, levels = charlevels)

allDEG <- allDEG %>% drop_na()

ggplot(allDEG, aes(x = lfc, y = logpadj)) +
  geom_point(aes(color = direction), alpha = 0.75) +
  facet_grid(tissue ~ comparison)  +
  scale_color_manual(values = colorscharnew,
                       name = " ",
                       drop = FALSE) +
  theme_B3() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")  +
  guides(color = guide_legend(nrow = 1))

ggplot(allDEG, aes(x = lfc)) +
  geom_histogram(aes( fill = direction)) +
  facet_grid(tissue ~ comparison)  +
  scale_color_manual(values = colorscharnew,
                       name = " ",
                       drop = FALSE) +
  theme_B3() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")  +
  guides(fill = guide_legend(nrow = 1))

ggplot(allDEG, aes(x = comparison,  fill = direction)) +
  geom_bar(position = "dodge") +
  facet_wrap(~tissue, nrow = 3) +
  theme_B3() +
  theme(axis.text.x = element_blank(),
        legend.position = "bottom",
        strip.placement = "right")  +
  guides(fill = guide_legend(nrow = 1)) +
  labs(x = "differntial expression between sequential stages", y = "total DEGs") +
  scale_fill_manual(values = colorscharnew,
                       name = " ",
                       drop = FALSE) +
  geom_text(stat='count', aes(label=..count..), vjust=-0.5, 
            position = position_dodge(width = 1),
            size = 3) +
  ylim(0,5000)

```

### 2020-01-23 analysis/22_symphony.Rmd
now with resepct project

---
title: "music"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggimage)
library(cowplot)

source("../R/themes.R")
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/favegenes/")

```

## R Markdown

```{r respect}
song <- c("what", "you", "want", "ba-", "by", "I", "got.")
position <- c(1,2,3,4,5,6,7)
note1 <- c(11,9,7,8,8,8,6)
note2 <- c(7, 9,5,6,8,8,4)
note3 <- c(4, 9,2,4,8,8,1)

respect <- as.data.frame(t(rbind(position, note1,note2,note3)))
respect$song <- song
respect$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
                        size=7, replace = TRUE)
respect <- respect %>%
  pivot_longer(note1:note3, names_to = "music", values_to = "notes")  %>% 
  mutate(position = as.numeric(position),
         notes = as.numeric(notes))

song <- ggplot(respect, aes(x = position, y = notes)) +
  geom_line(aes(color = music)) +
  geom_image(aes(image=image), size = 0.15) +
  scale_x_continuous(breaks = position,
                     labels = song,
                     name = element_blank()) +
  scale_y_continuous(breaks = c(3, 5, 7, 9),
                     labels = c( "F",   "A",   "C",  "E"),
                     limits = c(-1,12)) +
  geom_hline(yintercept=c(2.15,4.15,6.15,8.15,10.15))   +
  theme_B3() +
  theme(axis.ticks = element_blank())

stage <- c( "bldg", "lay", "inc.d9", "inc.d17", "hatch", "n5", "n9")
time <- c(1,2,3,4,5,6,7)
gene1 <- c(11,9,7,8,8,8,6)
gene2 <- c(7, 9,5,6,8,8,4)
gene3 <- c(4, 9,2,4,8,8,1)

hypothesisgenes <- as.data.frame(t(rbind(time, gene1,gene2,gene3)))
hypothesisgenes$stage <- stage
hypothesisgenes$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
                        size=7, replace = TRUE)
hypothesisgenes <- hypothesisgenes %>%
  pivot_longer(gene1:gene3, names_to = "data", values_to = "expression")  %>% 
  mutate( time = as.numeric(time))

theory <- ggplot(hypothesisgenes, aes(x = time, y = expression)) +
  geom_line(aes(color = data)) +
  geom_image(aes(image=image), size = 0.15) +
  scale_x_continuous(breaks = time,
                     labels = stage,
                     name = element_blank()) +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10),
                     limits = c(-1,12)) +
  geom_hline(yintercept=c(2.15,4.15,6.15,8.15,10.15)) +
  theme_B3() +
  theme(axis.ticks = element_blank())

plot_grid(song , theory, nrow = 2)

hypothesisgenes2 <- hypothesisgenes %>% select(-image)

respect <- respect %>% select(-image)

write.csv(respect, "../results/22_respect.csv")
write.csv(hypothesisgenes2, "../results/22_respectgenes.csv" )
```

### 2020-01-23 analysis/22_symphony.Rmd
soul genes'

@@
 ```{r respect}
-song <- c("what", "you", "want", "ba-", "by", "I", "got.")
+song <- c("What", "you", "want,", "ba-", "by", "I", "got.")
+stage <- c( "bldg", "lay", "inc.d9", "inc.d17", "hatch", "n5", "n9")
+time <- c(1,2,3,4,5,6,7)
 position <- c(1,2,3,4,5,6,7)
-note1 <- c(11,9,7,8,8,8,6)
-note2 <- c(7, 9,5,6,8,8,4)
-note3 <- c(4, 9,2,4,8,8,1)
+gene1 <- c(11,9,7,8,8,8,6)
+gene2 <- c(7, 9,5,6,8,8,4)
+gene3 <- c(4, 9,2,4,8,8,1)
 
-respect <- as.data.frame(t(rbind(position, note1,note2,note3)))
+respect <- as.data.frame(t(rbind(position, gene1,gene2,gene3)))
 respect$song <- song
+respect$stage <- stage
 respect$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
@@
 respect <- respect %>%
-  pivot_longer(note1:note3, names_to = "music", values_to = "notes")  %>% 
-  mutate(position = as.numeric(position),
-         notes = as.numeric(notes))
+  pivot_longer(gene1:gene3, names_to = "music", values_to = "genes")  %>% 
+  mutate(songstage = paste(song,stage, sep = "\n\n"))
 
-song <- ggplot(respect, aes(x = position, y = notes)) +
+songstageslim <- unique(respect$songstage)
+
+respectsong <- ggplot(respect, aes(x = position, y = genes)) +
   geom_line(aes(color = music)) +
@@
   scale_x_continuous(breaks = position,
-                     labels = song,
+                     labels = songstageslim,
                      name = element_blank()) +
@@
   theme_B3() +
-  theme(axis.ticks = element_blank())
+  theme(axis.ticks = element_blank()) +
+  labs(subtitle = "RESPECT by Otis Redding, Jr.", y = NULL)
+respectsong
+```
 
-stage <- c( "bldg", "lay", "inc.d9", "inc.d17", "hatch", "n5", "n9")
-time <- c(1,2,3,4,5,6,7)
-gene1 <- c(11,9,7,8,8,8,6)
-gene2 <- c(7, 9,5,6,8,8,4)
-gene3 <- c(4, 9,2,4,8,8,1)
+```{r feelinggood}
+song <- c("Birds", "fly-", "ing", "high,", "you", "know", "how", "I", "feel.")
+stage <- charlevels
+position <- c(1,2,3,4,5,6,7,8,9)
+gene1 <- c(-2,0,1,2,2,1,1,-2,0)
+gene2 <- gene1 - 1
+gene1 <- gene2
+gene3 <- gene2
 
-hypothesisgenes <- as.data.frame(t(rbind(time, gene1,gene2,gene3)))
-hypothesisgenes$stage <- stage
-hypothesisgenes$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
-                        size=7, replace = TRUE)
-hypothesisgenes <- hypothesisgenes %>%
-  pivot_longer(gene1:gene3, names_to = "data", values_to = "expression")  %>% 
-  mutate( time = as.numeric(time))
+feelinggood <- as.data.frame(t(rbind(position, gene1,gene2,gene3)))
+feelinggood$song <- song
+feelinggood$stage <- stage
+feelinggood$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
+                        size=9, replace = TRUE)
+feelinggood <- feelinggood %>%
+  pivot_longer(gene1:gene3, names_to = "music", values_to = "genes")   %>%
+  mutate(songstage = paste(song,stage, sep = "\n\n"))
+feelinggood
 
-theory <- ggplot(hypothesisgenes, aes(x = time, y = expression)) +
-  geom_line(aes(color = data)) +
+songstageslim <- unique(feelinggood$songstage)
+
+feelinggoodsong <- ggplot(feelinggood, aes(x = position, y = genes)) +
+  geom_line(aes(color = music)) +
   geom_image(aes(image=image), size = 0.15) +
-  scale_x_continuous(breaks = time,
-                     labels = stage,
+  scale_x_continuous(breaks = position,
+                    labels = songstageslim,
                      name = element_blank()) +
-  scale_y_continuous(breaks = c(2, 4, 6, 8, 10),
+  scale_y_continuous(breaks = c(3, 5, 7, 9),
+                     labels = c( "F",   "A",   "C",  "E"),
+                     limits = c(-6,12)) +
+  geom_hline(yintercept=c(2.15,4.15,6.15,8.15,10.15))   +
+  theme_B3() +
+  theme(axis.ticks = element_blank()) +
+  labs(subtitle = "Feeling Good by L. Bricusse & A. Newley", y = NULL)
+feelinggoodsong
+
+```
+
+```{r sinnerman}
+song <- c("Oh,", "sin-", "ner", "man,,", "where", "_", "he", "gon'", "run", "_", "to?")
+stage <- c( "bldg", "lay", "inc.d3", "m.inc.d3" ,
+            "inc.d9", "m.inc.d9"  , 
+            "inc.d17", "m.inc.d17" ,
+            "hatch", "m.n2" ,
+            "n5")
+position <- c(1,2,3,4,5,6,7,8,9,10,11)
+gene1 <- c(0,0,0,0,2,2,1,0,0,0,0)
+gene2 <- gene1 + 2
+gene3 <- gene2 + 2
+
+sinnerman <- as.data.frame(t(rbind(position, gene1,gene2,gene3)))
+sinnerman$song <- song
+sinnerman$stage <- stage
+sinnerman$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
+                        size=11, replace = TRUE)
+sinnerman <- sinnerman %>%
+  pivot_longer(gene1:gene3, names_to = "music", values_to = "genes")   %>%
+  mutate(songstage = paste(song,stage, sep = "\n\n"))
+sinnerman
+
+songstageslim <- unique(sinnerman$songstage)
+
+sinnermansong <- ggplot(sinnerman, aes(x = position, y = genes)) +
+  geom_line(aes(color = music)) +
+  geom_image(aes(image=image), size = 0.1) +
+  scale_x_continuous(breaks = position,
+                    labels = songstageslim,
+                     name = element_blank()) +
+  scale_y_continuous(breaks = c(3, 5, 7, 9),
+                     labels = c( "F",   "A",   "C",  "E"),
                      limits = c(-1,12)) +
@@
   theme_B3() +
-  theme(axis.ticks = element_blank())
+  theme(axis.ticks = element_blank()) +
+  labs(subtitle = "Sinnerman by Nina Simone", y = NULL)
+sinnermansong
 
-plot_grid(song , theory, nrow = 2)
+```
+```{r soulmusic}
+p1 <- plot_grid(respectsong + theme(legend.position = "none"), feelinggoodsong + theme(legend.position = "none"), rel_widths = c(0.45,0.55))
 
-hypothesisgenes2 <- hypothesisgenes %>% select(-image)
-
-respect <- respect %>% select(-image)
-
-write.csv(respect, "../results/22_respect.csv")
-write.csv(hypothesisgenes2, "../results/22_respectgenes.csv" )
+p3 <- plot_grid(p1, sinnermansong, nrow = 2)
+p3
 ```

### 2020-01-23 analysis/22_symphony.Rmd
more symphony foo

@@
 ```
+
+```{r symphonymodules}
+modules <- read_csv("../results/08_genes_modules.csv") %>%
+  rename(modulecolor = `net$colors`)
+
+candidategenes <- c("PRL", "PRLR", 
+                 "VIP", "VIPR1", "VIPR2", 
+                 "OXT", "AVP", "AVPR1A", "AVPR1B", 
+                 "GNRH1","GNRHR", "NPVF",
+                 "NR3C1", "NR3C2",
+                 "ESR1", "ESR2", "AR",
+                 "DIO2","LEPR", "DIO3", "DIO1","CYP19A1",
+                 "HSPA14", "HSPA12A",
+                 "PTGES3", "HSD11B2","DRD5", "DRD1", "DRD2","PGE1", "PGF")
+
+modulescandidates <- modules %>% filter(gene %in% candidategenes)  %>%
+  group_by(modulecolor)  %>%
+  summarize(gene = str_c(gene, collapse = ", "))
+
+modulescandidates
+
+datapath <- "../results/"   # path to the data
+datafiles <- dir(datapath, pattern = "*allvsd.csv") # get file names
+datapathfile <- paste0(datapath, datafiles)
+
+df <- datapathfile %>%
+  setNames(nm = .) %>%
+  map_df(~read_csv(.x, col_types = cols(), col_names = T), .id = "filename") %>% 
+  mutate(tissue = sapply(strsplit(as.character(filename),'../results/06_'), "[", 2)) %>% 
+  mutate(tissue = sapply(strsplit(as.character(tissue),'allvsd.csv'), "[", 1))  %>% 
+  select(tissue, X1, everything())
+
+df2 <-  df  %>%
+  filter(X1 %in% c(candidategenes)) %>%
+  pivot_longer(cols = L.Blu13_male_gonad_control.NYNO:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "sample", values_to = "vsd") %>%
+   mutate(sex = sapply(strsplit(as.character(sample),'_'), "[", 2)) %>%
+   mutate(treatment = sapply(strsplit(as.character(sample),'_'), "[", 4))  %>%
+   mutate(treatment = sapply(strsplit(as.character(treatment),'.NYNO'), "[", 1)) %>%
+  mutate(bird = sapply(strsplit(as.character(sample),'_'), "[", 1)) %>%
+  filter(treatment %in% charlevels) %>%
+  select(bird, sex, treatment, tissue, X1, vsd) %>%
+  mutate(tissue = fct_recode(tissue, "hypothalamus" = "hyp",
+                    "pituitary" = "pit",
+                    "gonads" = "gon"
+                    )) %>%
+  rename(gene = X1) %>%
+  drop_na() %>%
+  droplevels()
+head(df2)
+
+df2$treatment <- factor(df2$treatment, levels = alllevels)
+df2$tissue <- factor(df2$tissue, levels = tissuelevels)
+df2$gene <- factor(df2$gene)
+
+df3 <- df2 %>% 
+  mutate(treatment = fct_relevel(treatment, charlevels)) %>% 
+  group_by(treatment, tissue, gene)  %>% 
+  summarize(m = mean(vsd, na.rm = T), se = sd(vsd,  na.rm = T)/sqrt(length(vsd))) %>%
+  mutate(image = "../figures/images/DoveParentsRNAseq_note.png") 
+head(df3)
+
+d4 <- left_join(df3, modules, by = "gene")
+
+for (i in levels(d4$tissue)) {
+  p <-  d4 %>%
+    filter(tissue == i) %>%
+    ggplot(aes(x = treatment, y = m)) +
+    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
+    geom_point(size = 1, aes(color = gene)) +
+    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
+    scale_alpha_manual(values = c(0.5, 1)) +
+    labs(subtitle = i, y = "average expression", x = "parental stage") +
+    facet_wrap(~modulecolor, nrow = 2) +
+    theme_B3() +
+    theme(legend.position = "bottom")
+ print(p)
+}
+
+d4 %>%
+  filter(modulecolor != "turquoise",
+         tissue == "pituitary") %>%
+  droplevels() %>% 
+    ggplot(aes(x = treatment, y = m)) +
+    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
+    geom_point(size = 1, aes(color = gene)) +
+    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
+    scale_alpha_manual(values = c(0.5, 1)) +
+    labs(subtitle = "WGCNA + canddiate genes in the pituitary", y = "average expression", x = "parental stage") +
+    facet_wrap(~modulecolor, scales = "free_y", nrow = 2) +
+    theme_B3() +
+    theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
+    guides(color = guide_legend(nrow = 2, byrow = T))
+
+d4 %>%
+  filter( modulecolor != "turquoise",
+          # modulecolor != "greenyellow",
+          tissue == "pituitary") %>%
+  droplevels() %>% 
+    ggplot(aes(x = treatment, y = m)) +
+  geom_image(aes(image=image), size = 0.15) +
+  facet_wrap(~modulecolor, scales = "free_y", nrow = 2) +
+  labs(subtitle = "WGCNA + canddiate genes in the pituitary", y = "gene expression", x = "parental stage") +
+    facet_wrap(~gene, scales = "free_y", nrow = ) +
+    theme_B3() +
+    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
+
+d4 %>%
+  filter( modulecolor != "turquoise",
+          # modulecolor != "greenyellow",
+          tissue == "pituitary") %>%
+  droplevels() %>% 
+    ggplot(aes(x = treatment, y = m)) +
+  geom_image(aes(image=image), size = 0.1) +
+  labs(subtitle = "WGCNA + canddiate genes in the pituitary", y = "gene expression", x = "parental stage") +
+    theme_B3() +
+    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_y_log10()
+
+```
+
+## outline
+
+- data-driven
+- hypothesis-drive
+- sonically-driven
+- visually-driven
+- statistcally-driven
+
+Data and hypothesis driven research for the visually, sonically, and statistcially inclined biologists

### 2020-01-23 analysis/02_PCA.Rmd
pca with new colors

@@
 abc2
-
 ```
@@
 def2
+
 ```
@@
 plot_grid(def2, gh2, nrow = 2)
+
 ```
@@
 plot_grid( abc2, def2, gh2, nrow = 3)
-
 ```

### 2020-01-27 analysis/16_PRL.Rmd
new data csv

@@
 
+## music for own
+
+```{r}
+meanBrcaPrl <- candidates.pit %>% 
+    droplevels() %>% 
+  dplyr::group_by(treatment, sex) %>%
+  dplyr::summarise(meanBRCA1 = mean(BRCA1), 
+                   meanPRL = mean(PRL)) %>%
+  dplyr::mutate(meanBRCA1 = round(meanBRCA1,2)) %>%
+  dplyr::mutate(meanPRL = round(meanPRL,1)) %>%
+  dplyr::filter(treatment %in% charlevels) %>%
+  droplevels() %>%
+  dplyr::mutate(timepoint = as.numeric(treatment)) %>%
+  select(timepoint, treatment, sex, meanPRL, meanBRCA1)
+meanBrcaPrl
+summary(meanBrcaPrl)
+```
+
 ```{r}
@@
 write.csv(aov_manip, "../results/16_aov_PRLsextreatment.csv")
+write.csv(meanBrcaPrl, "../results/16_meanBrcaPrl.csv", row.names = F)
+
 ```

### 2020-02-03 analysis/16_PRL.Rmd
for twitter

@@
     geom_boxplot(aes(fill = treatment, color = sex)) + 
-    scale_alpha_manual(values = c(0.75,1)) +
     theme_B3() +
@@
   scale_fill_manual(values = colorscharnew) +
-  labs(y = "PRL", x = "increasing time >>", subtitle = "Pituitary" ) +
+  labs(y = "PRL", x = "parental care stages", subtitle = "Pituitary" ) +
   guides(fill = FALSE, alpha = FALSE, color = guide_legend(order=1)) +
-  theme(axis.title.y  = element_text(face = "italic"))  +
+  theme(axis.title.y  = element_text(face = "italic")) 
+p2
+p2+
   annotation_custom(control, ymin = 13.5, ymax = 15, xmin = -7.5) +
@@
   annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 8.5)
-p2
 

### 2020-02-03 analysis/07_hormones.Rmd
now with pCA and new correlations

@@
 
-knitr::opts_chunk$set(fig.path = '../figures/hormones/',message=F, warning=FALSE)
+knitr::opts_chunk$set(fig.path = '../figures/hormones/',message=F, warning=FALSE, cache = T)
 ```
@@
         geom_smooth(aes(colour = sex)) +
-    geom_boxplot(aes(fill = treatment, alpha = sex)) +
+    geom_boxplot(aes(fill = treatment)) +
     mytheme() +
@@
          color = guide_legend(order=2)) +
-    scale_alpha_manual(values = c(0.75,1)) +
     scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
@@
 head(PRLpit,2)
-```
-
-## for rechelle
-
-```{r forrechelle, fig.height= 4, fig.width=6}
-
-rechelle <- hormones %>%
-  dplyr::filter(hormone %in% c("prolactin", "corticosterone"),
-                treatment %in% c("lay", "inc.d9", "hatch", "n9")) %>%
-  dplyr::select(-icons, -music, -iconpath)
-
-rechelle$hormone <- factor(rechelle$hormone, levels = c("prolactin", "corticosterone"))
-
-prl.rechelle <- rechelle %>% filter(hormone == "prolactin")   %>%  droplevels()
-cort.rechelle <- rechelle %>% filter(hormone == "corticosterone")   %>%  droplevels()
-
-TukeyHSD(aov(plasma_conc ~  treatment, data=prl.rechelle))
-TukeyHSD(aov(plasma_conc ~  treatment, data=cort.rechelle))
-
-rechelleplot <- function(mydf, myylab){
-
-  p <- ggplot(mydf, aes(x = treatment, y = plasma_conc, fill = sex)) +
-  geom_point(aes(color = sex), position = position_dodge(0.9), size = 2)+
-  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
-  scale_color_manual(name = "Sex", values = c("gray50", "maroon"))+
-  scale_fill_manual(name = "Sex", values = c("gray50", "maroon"))+
-
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-         legend.position = "none",
-          strip.text = element_blank()) +
-    theme_classic()+
-  theme(text = element_text(size= 15)) +
-    labs(x = "Stage", y = myylab) 
-  return(p)
-
-}
-
-#PRL
-rechelleplot(prl.rechelle, "Prolactin (ng/mL)")
-rechelleplot(prl.rechelle, "Prolactin (ng/mL)") + 
-  scale_y_continuous(breaks = c(0,25,50,75,100, 125), limits = c(-20,125)) +
-  geom_signif(comparisons=list(c("lay", "hatch")), annotations= "p < 0.001", y_position = 100, tip_length = 0.05, vjust = 0, color= "black" , textsize = 5 ) +
-  geom_signif(comparisons=list(c("lay", "n9")), annotations= "p < 0.001", y_position = 110, tip_length = 0.05, vjust = 0, color= "black" , textsize = 5 ) +
-  geom_signif(comparisons=list(c("hatch", "n9")), annotations= "p < 0.001", y_position = 120, tip_length = 0.05, vjust = 0, color= "black" , textsize = 5 ) +
-  geom_signif(comparisons=list(c("inc.d9", "hatch")), annotations= "p < 0.001", y_position = -5, tip_length = -0.05, vjust = 2, color= "black" , textsize = 5 ) +
-  geom_signif(comparisons=list(c("inc.d9", "n9")), annotations= "p < 0.001", y_position = -15, tip_length = -0.05, vjust = 2, color= "black" , textsize = 5 )
-
-#CORT
-rechelleplot(cort.rechelle, "Corticosterone (ng/mL)")
-
 ```
@@
 corrCORT
+
+hormoneswide %>%
+  pivot_longer(cols = c(prolactin, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
+  filter(study == "characterization") %>%
+  ggplot(aes(x = corticosterone, y = conc)) +
+  geom_point(aes(color = treatment)) +
+  geom_smooth(color = "darkgrey", method = "lm") +
+  facet_grid(hormones ~ treatment, scales = "free_y")  +
+  theme_B3() +
+  theme(legend.position = "bottom",
+        legend.title = element_blank()) +
+  scale_color_manual(values = colorscharmaip2) +
+  guides(color = guide_legend(nrow = 1)) +
+  labs(y = "conc. (ng/mL)", x = "corticosterone (ng/mL)")
+
+hormoneswide %>%
+  pivot_longer(cols = c(corticosterone, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
+  filter(study == "characterization") %>%
+  ggplot(aes(x = prolactin, y = conc)) +
+  geom_point(aes(color = treatment)) +
+  geom_smooth(color = "darkgrey", method = "lm") +
+  facet_grid(hormones ~ treatment, scales = "free_y")  +
+  theme_B3() +
+  theme(legend.position = "bottom",
+        legend.title = element_blank()) +
+  scale_color_manual(values = colorscharmaip2) +
+  guides(color = guide_legend(nrow = 1)) +
+  labs(y = "conc. (ng/mL)", x = "prolactin (ng/mL)")
+
+ggplot(hormoneswide, aes(x = prolactin, y = corticosterone, color = treatment)) +
+  geom_point() + geom_smooth(method = "lm") +
+  facet_wrap(~treatment) +
+  scale_color_manual(values =  colorscharmaip) + theme(legend.position = "none")
+
+ggplot(hormoneswide, aes(x = prolactin, y = testosterone, color = treatment)) +
+  geom_point() + geom_smooth(method = "lm") +
+  facet_wrap(~treatment) +
+  scale_color_manual(values =  colorscharmaip) + theme(legend.position = "none")
+
 ```
@@
 ```
+
+## pca of hormones
+
+```{r PCA}
+
+library("factoextra")
+library("FactoMineR")
+
+# char only
+hormoneswidechar <- hormoneswide %>% filter(treatment %in% charlevels)
+df <- hormoneswidechar %>% select(prolactin:progesterone)
+res.pca <- PCA(df)
+
+a <- fviz_pca_var(res.pca, col.var = "black", title = "Charactherization only") 
+# Contributions of variables to PC1
+
+b <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10) + labs(y = "Dim 1 contibutions (%)",
+                                                                      title = element_blank())
+# Contributions of variables to PC2
+c <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10) +
+  labs(y = "Dim 2 contibutions (%)") + theme(title = element_blank())
+
+plot_grid(a,b,c, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")
+
+# char and manip
+df <- hormoneswide %>% select(prolactin:progesterone)
+res.pca <- PCA(df)
+
+a <- fviz_pca_var(res.pca, col.var = "black", title = "Charactherization & manipluation") 
+# Contributions of variables to PC1
+
+b <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10) + labs(y = "Dim 1 contibutions (%)",
+                                                                      title = element_blank())
+# Contributions of variables to PC2
+c <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10) +
+  labs(y = "Dim 2 contibutions (%)") + theme(title = element_blank())
+
+plot_grid(a,b,c, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")
+
+```

### 2020-02-03 analysis/07_hormones.Rmd
now with a summary

@@
 
-a <- fviz_pca_var(res.pca, col.var = "black", title = "Charactherization only") 
+a1 <- fviz_pca_var(res.pca, col.var = "black", title = "Charactherization only") 
 # Contributions of variables to PC1
 
-b <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10) + labs(y = "Dim 1 contibutions (%)",
+b1 <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10) + labs(y = "Dim 1 contibutions (%)",
                                                                       title = element_blank())
 # Contributions of variables to PC2
-c <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10) +
+c1 <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10) +
   labs(y = "Dim 2 contibutions (%)") + theme(title = element_blank())
 
-plot_grid(a,b,c, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")
+ones <- plot_grid(a1,b1,c1, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")
 
@@
 
-a <- fviz_pca_var(res.pca, col.var = "black", title = "Charactherization & manipluation") 
+a2 <- fviz_pca_var(res.pca, col.var = "black", title = "Charactherization & manipluation") 
 # Contributions of variables to PC1
 
-b <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10) + labs(y = "Dim 1 contibutions (%)",
+b2 <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10) + labs(y = "Dim 1 contibutions (%)",
                                                                       title = element_blank())
 # Contributions of variables to PC2
-c <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10) +
+c2 <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10) +
   labs(y = "Dim 2 contibutions (%)") + theme(title = element_blank())
 
-plot_grid(a,b,c, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")
+twos <- plot_grid(a2,b2,c2, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")
 
+plot_grid(ones,twos, nrow = 2)
 ```
+
+# smooth funciton
+
+```{r summary}
+head(hormones)
+ggplot(hormones, aes(x = as.numeric(treatment), y = plasma_conc, color = hormone)) +
+  geom_smooth()
+
+box <- ggplot(hormones, aes(x = as.numeric(treatment), y = plasma_conc)) +
+  geom_boxplot(aes(fill = treatment)) +
+  #geom_smooth(aes(color = hormone)) +
+  facet_wrap(~hormone, scales = "free_y", nrow = 5) +
+  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16),
+                     labels = alllevels) +
+  theme_B3() +
+  theme(#strip.text = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.position = "none",
+        panel.background = element_rect(fill = "transparent",colour = NA),
+        plot.background = element_rect(fill = "transparent",colour = NA)) +
+  labs(y = "plasma conc. (ng/mL)", x = "parental stage")
+
+line <- ggplot(hormones, aes(x = as.numeric(treatment), y = plasma_conc)) +
+  #geom_boxplot(aes(fill = treatment)) +
+  geom_smooth(aes(color = hormone)) +
+  facet_wrap(~hormone, scales = "free_y", nrow = 5) +
+  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16),
+                     labels = alllevels) +
+  theme_B3() +
+  theme(#strip.text = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.position = "none",
+        panel.background = element_rect(fill = "transparent",colour = NA),
+        plot.background = element_rect(fill = "transparent",colour = NA)) +
+  labs(y = "plasma conc. (ng/mL)", x = "parental stage")
+
+plot_grid(box, line, nrow = 1)
+```

### 2020-02-03 analysis/07_hormones.Rmd
now with dense figures

@@
   aov_all = data.frame()
-  for(i in alllevels3){
+  for(i in alllevels){
     df <- hormones %>% filter(hormone == myhormone,
@@
 p34 <- plot_grid(p3,p4, rel_widths = c(0.6,0.4), labels = c("c","d"), label_size = 8 )
+p34
 
-plot_grid(p12, p34, nrow = 2, rel_heights = c(0.6,0.4))
+allcortplots <- plot_grid(p12, p34, nrow = 2, rel_heights = c(0.6,0.4))
+allcortplots
 ```
@@
 
-plot_grid(p12, p34, nrow = 2, rel_heights = c(0.5,0.5))
+allprolactinplots <- plot_grid(p12, p34, nrow = 2, rel_heights = c(0.5,0.5))
+allprolactinplots
+```
+
+## figure 2
+
+```{r figure2, fig.width=7, fig.height=7}
+
+plot_grid(allprolactinplots, allcortplots, nrow = 2)
 ```
@@
 ```{r supplefig, fig.height=8}
-plot_grid(p1, p2, p3 + labs(y = NULL), p4, p5, p6, p7 + labs(y = NULL), p8, 
+sexsteroids <- plot_grid(p1, p2, p3 + labs(y = NULL), p4, p5, p6, p7 + labs(y = NULL), p8, 
           p9, p10, p11 + labs(y = NULL), p12, nrow = 3, rel_widths =   c(0.175,0.4,0.35,0.25), align = "hv",
@@
                                      "c", " ", " " , " "))
-
+sexsteroids
 ```
@@
 
-twos <- plot_grid(a2,b2,c2, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")
+twos <- plot_grid(a2 + theme(title = element_blank()),b2,c2, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")
 
 plot_grid(ones,twos, nrow = 2)
+
+twos
+```
+
+```{r figure3, fig.height=8, fig.width=7}
+
+plot_grid(twos,sexsteroids, nrow = 2, rel_heights = c(0.25,0.75), align = "hv")
 ```

### 2020-02-03 analysis/07_hormones.Rmd
better figures

@@
 
-printaovtablesex("prolactin", "PRL ~ sex")
+table1 <- printaovtablesex("prolactin", "PRL ~ sex")
+table1
+write.csv(table1, "../../parentalhormones/data/table-1.csv", row.names = F)
 ```
@@
   theme_B3() +
-  labs(y = "concentration (ng/mL)", subtitle = "Corticosterone" , x = NULL) +
+  labs(y = "CORT (ng/mL)", subtitle = element_blank() , x = NULL) +
   theme(legend.position = "none") +
-  scale_y_continuous(limits = c(0,8.5)) +
   scale_fill_manual(values = sexcolors)
@@
   theme_B3() +
-  labs(y = NULL, x = NULL, subtitle = " ") +
-  theme(legend.position = "none") +
-  geom_signif(comparisons=list(c("control", "inc.d17")), annotations= "p = 0.49", y_position = 7.5, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
-  geom_signif(comparisons=list(c("control", "hatch")), annotations= "p = 0.021", y_position = 8, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
-  geom_signif(comparisons=list(c("control", "n5")), annotations= "p = 0.004", y_position = 8.5, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
-  geom_signif(comparisons=list(c("n5", "n9")), annotations= "p = 0.001", y_position = 8, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
-  scale_y_continuous(limits = c(0,8.5))
+  labs(y = NULL, x = NULL, subtitle = element_blank() ) +
+  theme(legend.position = "none")
 
@@
 
-p12 <- plot_grid(p1,p2, rel_widths = c(0.2,0.8), labels = c("a","b"), label_size = 8)
+p12 <- plot_grid(p1,p2, rel_widths = c(0.2,0.8), labels = c("h","i"), label_size = 8)
 
@@
   theme_B3() +
-  scale_y_continuous(limits = c(0,8.5)) +
   scale_fill_manual(values = colorscharmaip) +
   theme(legend.position = "none") +
-  labs( x = NULL, y = "concentration (ng/mL")
+  labs( x = NULL, y = "CORT (ng/mL") +
+  ylim(c(0,12.5))
 p3
@@
   theme_B3() +
-  scale_y_continuous(limits = c(0,8.5)) +
   scale_fill_manual(values = colorscharmaip) +
   theme(legend.position = "none") +
-  labs(y = NULL, x = NULL)
+  labs(y = NULL, x = NULL) + 
+  ylim(c(0,12.5))
 p4
 
-p34 <- plot_grid(p3,p4, rel_widths = c(0.6,0.4), labels = c("c","d"), label_size = 8 )
+p34 <- plot_grid(p3,p4, rel_widths = c(0.6,0.4), labels = c("j","k"), label_size = 8 )
 p34
@@
   theme_B3() +
-  labs(y = "concentration (ng/mL)", subtitle = "Prolactin" , x = NULL) +
+  labs(y = "Prolactin (ng/mL)", subtitle = " " , x = NULL) +
   theme(legend.position = "none",
@@
   scale_color_manual(values = sexcolors) +
-  scale_y_continuous(limits = c(0,100)) +
-    geom_text(aes(label = "B", x = 1, y = 100, fontface = "plain"), size = 3) +
-   geom_text(aes(label = "A", x = 2, y = 100), size = 3) +
-geom_text(aes(label = "A", x = 3, y = 100), size = 3) +
-geom_text(aes(label = "B", x = 4, y = 100), size = 3) +
-  geom_text(aes(label = "C", x = 5, y = 100), size = 3) +
-  geom_text(aes(label = "D", x = 6, y = 100), size = 3) +
-  geom_text(aes(label = "D", x = 7, y = 100), size = 3) +
-  geom_text(aes(label = "D", x = 8, y = 100), size = 3) +
-  geom_text(aes(label = "D", x = 9, y = 100), size = 3)
+  scale_y_continuous(limits = c(0,100))
 
@@
   scale_color_manual(values = sexcolors) +
-  scale_y_continuous(limits = c(0,100)) +
-  geom_text(aes(label = "A", x = 1, y = 100, fontface = "plain"), size = 3) +
-   geom_text(aes(label = "A", x = 2, y = 100), size = 3) +
-geom_text(aes(label = "B", x = 3, y = 100), size = 3) +
-geom_text(aes(label = "B", x = 4, y = 100), size = 3) +
-  geom_text(aes(label = "B", x = 5, y = 100), size = 3) +
-  geom_text(aes(label = "D", x = 6, y = 100), size = 3) +
-  geom_text(aes(label = "D", x = 7, y = 100), size = 3) +
-  geom_text(aes(label = "D", x = 8, y = 100), size = 3) +
-  geom_text(aes(label = "D", x = 9, y = 100), size = 3)
+  scale_y_continuous(limits = c(0,100))
 
@@
         axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs( x = NULL, y = "concentration (ng/mL", subtitle = "females") +
+  labs( x = NULL, y = "Prolactin (ng/mL", subtitle = "females") +
   scale_color_manual(values = sexcolors)
@@
 plot_grid(allprolactinplots, allcortplots, nrow = 2)
-```
 
-## prolactin 2
-
-```{r PRL2}
-hormones %>%
-  filter(hormone == "prolactin") %>%
-  group_by(sex) %>%
-  summarise(m = mean(plasma_conc),
-            median = median(plasma_conc))
-
-p1 <- hormones %>%
-  filter(hormone == "prolactin",
-         study == "characterization") %>%
-  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
-  geom_boxplot() +
-  theme_B3() +
-  labs(y = "concentration (ng/mL)", subtitle = "Prolactin" , x = NULL) +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_fill_manual(values = sexcolors) +
-  scale_y_continuous(limits = c(0,100)) 
-p1
-
-df <- hormones %>%
-  filter(hormone == "prolactin",
-         study == "characterization")
-p2 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
-  geom_boxplot(aes(color = sex)) +
-  theme_B3() +
-  labs(y = NULL, x = NULL, subtitle = " ") +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_color_manual(values = sexcolors) +
-  scale_y_continuous(limits = c(0,105)) +
-    geom_text(aes(label = "B", x = 1, y = 100, fontface = "plain"), size = 3) +
-   geom_text(aes(label = "A", x = 2, y = 100), size = 3) +
-geom_text(aes(label = "A", x = 3, y = 100), size = 3) +
-geom_text(aes(label = "B", x = 4, y = 100), size = 3) +
-  geom_text(aes(label = "C", x = 5, y = 100), size = 3) +
-  geom_text(aes(label = "D", x = 6, y = 100), size = 3) +
-  geom_text(aes(label = "D", x = 7, y = 100), size = 3) +
-  geom_text(aes(label = "D", x = 8, y = 100), size = 3) +
-  geom_text(aes(label = "D", x = 9, y = 100), size = 3) +
-  geom_text(aes(label = '*', x = 1, y = 105), size = 5) +
-   geom_text(aes(label = '*', x = 2, y = 105), size = 5) 
-  p2
-
-p3 <- hormones %>%
-  filter(hormone == "prolactin",
-         treatment %in% c(controlsremoval, levelsremoval )) %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot(aes(color = sex)) +
-  theme_B3() +
-  scale_y_continuous(limits = c(0,110)) +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs( x = NULL, y = "concentration (ng/mL", subtitle = " ") +
-  scale_color_manual(values = sexcolors) +
-   geom_text(aes(label = '*', x = 2, y = 105), size = 5) 
-p3
-
-p4 <- hormones %>%
-  filter(hormone == "prolactin",
-         treatment %in% c(controlstiming, levelstiming )) %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot(aes(color = sex)) +
-  theme_B3() +
-  scale_y_continuous(limits = c(0,110)) +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs(y = NULL, x = NULL, subtitle = " ") +
-  scale_color_manual(values = sexcolors) 
-p4
-
-p12 <- plot_grid(p1, p2, rel_widths = c(0.15,0.8), labels = "auto", label_size = 8)
-p34 <- plot_grid(p3,p4, nrow = 1, rel_widths = c(0.55,0.45), labels = c("c", "d"), label_size = 8)
-
-plot_grid(p12, p34, nrow = 2)
 ```
@@
   theme_B3() +
-  labs(y = "concentration (ng/mL)", subtitle = "Progesterone" , x = NULL) +
+  labs(y = "Progesterone (ng/mL)", subtitle = element_blank() , x = NULL) +
   theme(legend.position = "none")   +
@@
   theme(legend.position = "none") +
-  labs( x = NULL, y = "concentration (ng/mL")
+  labs( x = NULL, y = "Progesterone (ng/mL")
 
@@
   theme_B3() +
-  labs(y = "concentration (ng/mL)", subtitle = "Estradiol" , x = NULL) +
+  labs(y = "Estradiol (ng/mL)", subtitle = element_blank() , x = NULL) +
   theme(legend.position = "none")  +
@@
   theme(legend.position = "none") +
-  labs( x = NULL, y = "concentration (ng/mL")
+  labs( x = NULL, y = "Estradiol (ng/mL")
 
@@
   theme_B3() +
-  labs(y = "concentration (ng/mL)", subtitle = "Testosterone" , x = NULL) +
+  labs(y = "Testosterone (ng/mL)", subtitle = element_blank() , x = NULL) +
   theme(legend.position = "none")  +
@@
 
-```{r supplefig, fig.height=8}
-sexsteroids <- plot_grid(p1, p2, p3 + labs(y = NULL), p4, p5, p6, p7 + labs(y = NULL), p8, 
-          p9, p10, p11 + labs(y = NULL), p12, nrow = 3, rel_widths =   c(0.175,0.4,0.35,0.25), align = "hv",
-          label_size = 8, labels = c("a", " ", " " , " ",
-                                     "b", " ", " " , " ",
-                                     "c", " ", " " , " "))
+```{r sexsteroids2, fig.height=8}
+sexsteroids <- plot_grid(
+       p9 + theme(axis.text.x = element_blank()),  
+       p10 + theme(axis.text.x = element_blank()), 
+       p11 + labs(y = NULL) + theme(axis.text.x = element_blank()), 
+       p12 + theme(axis.text.x = element_blank()),
+        p5 + theme(axis.text.x = element_blank()), 
+       p6 + theme(axis.text.x = element_blank()), 
+       p7 + labs(y = NULL) + theme(axis.text.x = element_blank()), 
+       p8 + theme(axis.text.x = element_blank()), 
+       p1 , 
+       p2 + theme(axis.text.x = element_text(angle = 45, hjust = 1)), 
+       p3 + labs(y = NULL)  + theme(axis.text.x = element_text(angle = 45, hjust = 1)), 
+       p4  + theme(axis.text.x = element_text(angle = 45, hjust = 1)), 
+        nrow = 3, 
+        rel_widths =   c(0.175,0.4,0.35,0.25), align = "hv",
+          label_size = 8, labels = c("A1", "A2", "A3" , "A4",
+                                     "B1", "B2", "B3" , "B4",
+                                     "C1", "C2", "C3" , "C3"))
 sexsteroids
+
 ```

### 2020-02-03 analysis/07_hormones.Rmd
better table order

@@
                                       "estradiol", "testosterone"))
+hormones$logconc <- log10(hormones$plasma_conc)
 head(hormones)
@@
   theme(legend.position = "none") +
-  scale_fill_manual(values = sexcolors)
+  scale_fill_manual(values = sexcolors) +
+  scale_fill_manual(values = colorscharmaip)
 p1
@@
   labs(y = NULL, x = NULL, subtitle = element_blank() ) +
-  theme(legend.position = "none")
+  theme(legend.position = "none")  +
+  scale_fill_manual(values = colorscharmaip)
 
@@
   filter(hormone == "prolactin",
-         study == "characterization",
+         #study == "characterization",
          sex == "female")
@@
   filter(hormone == "prolactin",
-         study == "characterization",
+         #study == "characterization",
          sex == "male")
@@
   scale_color_manual(values = sexcolors) +
-  scale_y_continuous(limits = c(0,100))
+  scale_y_continuous(limits = c(0,100))  +
+  scale_fill_manual(values = colorscharmaip)
 
@@
   scale_color_manual(values = sexcolors) +
-  scale_y_continuous(limits = c(0,100))
+  scale_y_continuous(limits = c(0,100))+
+  scale_fill_manual(values = colorscharmaip)
 
@@
 ```
+
+## stats
+
+```{r}
+
+P3 <- hormones %>% filter(hormone == "progesterone")
+E2 <- hormones %>% filter(hormone == "estradiol")
+T2 <- hormones %>% filter(hormone == "testosterone")
+
+aov1 <- aov(PRLm$logconc ~ PRLm$treatment)
+res1 <- TukeyHSD(aov1)
+res1 <- as.data.frame(res1$`PRLm$treatment`)
+res1$hormone <- "male PRL"
+res1$sig <- ifelse(res1$`p adj` < 0.001, "***",
+                   ifelse(res1$`p adj` < 0.01, "**",
+                           ifelse(res1$`p adj` < 0.05, "*", "NS" )))
+res1$comparison <- row.names(res1)
+res1
+
+aov2 <- aov(PRLf$logconc ~ PRLf$treatment)
+res2 <- TukeyHSD(aov2)
+res2 <- as.data.frame(res2$`PRLf$treatment`)
+res2$hormone <- "female PRL"
+res2$sig <- ifelse(res2$`p adj` < 0.001, "***",
+                   ifelse(res2$`p adj` < 0.01, "**",
+                           ifelse(res2$`p adj` < 0.05, "*","NS"  )))
+res2$comparison <- row.names(res2)
+res2
+
+aov3 <- aov(P3$logconc ~ P3$treatment)
+res3 <- TukeyHSD(aov3)
+res3 <- as.data.frame(res3$`P3$treatment`)
+res3$hormone <- "P3"
+res3$sig <- ifelse(res3$`p adj` < 0.001, "***",
+                   ifelse(res3$`p adj` < 0.01, "**",
+                           ifelse(res3$`p adj` < 0.05, "*", "NS"  )))
+res3$comparison <- row.names(res3)
+res3
+
+aov4 <- aov(E2$logconc ~ E2$treatment)
+res4 <- TukeyHSD(aov4)
+res4 <- as.data.frame(res4$`E2$treatment`)
+res4$hormone <- "E2"
+res4$sig <- ifelse(res4$`p adj` < 0.001, "***",
+                   ifelse(res4$`p adj` < 0.01, "**",
+                           ifelse(res4$`p adj` < 0.05, "*", "NS" )))
+res4$comparison <- row.names(res4)
+res4
+
+aov5 <- aov(T2$logconc ~ T2$treatment)
+res5 <- TukeyHSD(aov5)
+res5 <- as.data.frame(res5$`T2$treatment`)
+res5$hormone <- "T"
+res5$sig <- ifelse(res5$`p adj` < 0.001, "***",
+                   ifelse(res5$`p adj` < 0.01, "**",
+                           ifelse(res5$`p adj` < 0.05, "*", "NS"  )))
+res5$comparison <- row.names(res5)
+res5
+
+resall <- rbind(res3, res5, res4, res2, res1) %>% drop_na()
+resall
+
+table1 <- resall %>%
+  filter(grepl('control', comparison),
+         !grepl('prolong|extend|m.', comparison)) %>%
+   mutate(diff = round(diff, 2)) %>% 
+  mutate(diffsig = paste(diff, sig, sep = ", ")) %>% 
+   select(hormone, comparison, diffsig) %>% 
+  pivot_wider( names_from = hormone, values_from = diffsig)
+table1
+
+table2 <- resall %>%
+  filter(!grepl('control|bldg|lay|n5', comparison),
+         grepl('prolong|extend|m.', comparison)) %>%
+   mutate(diff = round(diff, 2)) %>% 
+  mutate(diffsig = paste(diff, sig, sep = ", ")) %>% 
+   select(hormone, comparison, diffsig) %>% 
+  pivot_wider( names_from = hormone, values_from = diffsig)
+table2
+
+write.csv(table1, "../../parentalhormones/data/table-1.csv")
+write.csv(table2, "../../parentalhormones/data/table-2.csv")
+
+```

### 2020-02-04 R/themes.R
new summary

@@ -156,38 +156,18 @@ colorstissue <- c("hypothalamus" = "#d95f02",
 
 sexcolors <- c("female" = "#969696", "male" = "#525252")
 
-colorsvolcano <-  c("control" = "#F8766D", 
-                     "bldg"= "#D39200", 
-                     "lay"= "#93AA00", 
-                     "inc.d3"= "#00BA38", 
-                     "inc.d9"= "#00C19F", 
-                     "inc.d17"= "#00B9E3", 
-                     "hatch"= "#619CFF", 
-                     "n5"= "#DB72FB", 
-                     "n9"= "#FF61C3",
-                     "m.inc.d3" = "#CDCDCD", 
-                     "m.inc.d9" = "#959595", 
-                     "m.inc.d17" = "#626262",
-                     "m.n2" = "#262625", 
+colorsvolcano <-  c(colorscharnew,
                     
-                    "m.inc.d8" = "#fcc5c0", 
-                    "prolong" = "#ae017e" , 
-                     "extend" = "#7a0177",
+                    "m.inc.d8" = "#cbc9e2", 
+                    "prolong" = "#9e9ac8" , 
+                     "extend" = "#6a51a3",
                     
                     "male" = "#525252",
                     "female" = "#969696",
                     
                     "NS" = "#bdbdbd")
 
-colorsvolcanochar <-  c("control" = "#F8766D", 
-                    "bldg"= "#D39200", 
-                    "lay"= "#93AA00", 
-                    "inc.d3"= "#00BA38", 
-                    "inc.d9"= "#00C19F", 
-                    "inc.d17"= "#00B9E3", 
-                    "hatch"= "#619CFF", 
-                    "n5"= "#DB72FB", 
-                    "n9"= "#FF61C3",
+colorsvolcanochar <-  c(colorscharnew,
                     "NS" = "#bdbdbd")
 
 myannotationcolors <- list(sex = sexcolors,

### 2020-02-04 analysis/07_hormones.Rmd
new summary
new files for owen

@@
         plot.background = element_rect(fill = "transparent",colour = NA)) +
-  labs(y = "plasma conc. (ng/mL)", x = "parental stage")
+  labs(y = "plasma conc. (ng/mL)", x = "parental stage") +
+  scale_fill_manual(values = colorscharmaip)
 
@@
 ```
+
+## summary for owen
+
+```{r}
+meanTE <- hormoneswide %>% 
+  dplyr::group_by(treatment) %>%
+  dplyr::summarise(meanTestosterone = (mean(testosterone, na.rm = TRUE)), 
+                   meanEstradiol = mean(estradiol, na.rm = TRUE)) %>%
+  dplyr::mutate(meanTestosterone = round(meanTestosterone,2)) %>%
+  dplyr::mutate(meanEstradiol = round(meanEstradiol,2)) %>%
+  dplyr::filter(treatment %in% charlevels) %>%
+  droplevels() %>%
+  dplyr::mutate(timepoint = as.numeric(treatment)) %>%
+  select(timepoint, treatment, meanTestosterone, meanEstradiol)
+meanTE
+write.csv("../results/07_meanTE.csv", row.names = F)
+```

### 2020-02-04 analysis/07_hormones.Rmd
new images

--- 2020-02-03 analysis/07_hormones.Rmd
@@
 
-write.csv(table1, "../../parentalhormones/data/table-1.csv")
-write.csv(table2, "../../parentalhormones/data/table-2.csv")
+```
 
+## summary for owen
+
+```{r}
+meanTE <- hormoneswide %>% 
+  dplyr::group_by(treatment) %>%
+  dplyr::summarise(meanTestosterone = (mean(testosterone, na.rm = TRUE)), 
+                   meanEstradiol = mean(estradiol, na.rm = TRUE)) %>%
+  dplyr::mutate(meanTestosterone = round(meanTestosterone,2)) %>%
+  dplyr::mutate(meanEstradiol = round(meanEstradiol,2)) %>%
+  dplyr::filter(treatment %in% charlevels) %>%
+  droplevels() %>%
+  dplyr::mutate(timepoint = as.numeric(treatment)) %>%
+  select(timepoint, treatment, meanTestosterone, meanEstradiol)
+meanTE
 ```
+
+```{r savefiles}
+write.csv(meanTE, "../results/07_meanTE.csv", row.names = F)
+write.csv(hormones, "../results/07_hormones.csv", row.names = F)
+write.csv(hormoneswide, "../results/07_hormoneswide.csv", row.names = F)
+```

### 2020-02-04 analysis/16_PRL.Rmd
new images

@@
 
-candidategenes <- c("PRL", "BRCA1")
+candidategenes <- c("PRL", "BRCA1", "PRLR")
 candidates.pit <- vsd.pit %>% dplyr::filter(genes %in% candidategenes)
@@
 ```{r}
-meanBrcaPrl <- candidates.pit %>% 
+meanPRLPRLR <- candidates.pit %>% 
     droplevels() %>% 
   dplyr::group_by(treatment, sex) %>%
-  dplyr::summarise(meanBRCA1 = mean(BRCA1), 
+  dplyr::summarise(meanPRLR = mean(PRLR), 
                    meanPRL = mean(PRL)) %>%
-  dplyr::mutate(meanBRCA1 = round(meanBRCA1,2)) %>%
+  dplyr::mutate(meanPRLR = round(meanPRLR,2)) %>%
   dplyr::mutate(meanPRL = round(meanPRL,1)) %>%
@@
   dplyr::mutate(timepoint = as.numeric(treatment)) %>%
-  select(timepoint, treatment, sex, meanPRL, meanBRCA1)
-meanBrcaPrl
-summary(meanBrcaPrl)
+  select(timepoint, treatment, sex, meanPRL, meanPRLR)
+meanPRLPRLR
+
 ```
@@
 write.csv(aov_manip, "../results/16_aov_PRLsextreatment.csv")
-write.csv(meanBrcaPrl, "../results/16_meanBrcaPrl.csv", row.names = F)
+write.csv(meanPRLPRLR, "../results/16_meanPRLPRLR.csv", row.names = F)
 

### 2020-02-05 analysis/07_hormones.Rmd
now with only datat wrangling

@@
 
-```{r characterizationplots}
-hormonecharplot <- function(myhormone, myylab){
-
-  hormones %>% 
-    filter(study == "characterization",
-           hormone %in% c(myhormone))  %>% 
-    droplevels() %>% 
-  ggplot(aes(x = as.numeric(treatment), y = plasma_conc)) +
-        geom_smooth(aes(colour = sex)) +
-    geom_boxplot(aes(fill = treatment)) +
-    mytheme() +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
-          legend.position = "none") +
-    scale_fill_manual(values = colorscharmaip) +
-    scale_color_manual(values = sexcolors) +
-    labs(y = myylab, x = NULL) +
-    guides(fill = guide_legend(order=1),
-         color = guide_legend(order=2)) +
-    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
-                       labels = charlevels)
-  }
-
-hormonecharplot("prolactin", "PRL (ng/mL)")
-hormonecharplot("corticosterone", "CORT (ng/mL)")
-hormonecharplot("estradiol", "e2 (ng/mL)")
-hormonecharplot("testosterone", "t (ng/mL)")
-hormonecharplot("progesterone", "p4 (ng/mL)")
-```
-
-```{r manipulation}
-# prolactin, removal only
-hormones %>% 
-    filter( hormone == c("prolactin"))  %>% 
-    filter(treatment %in% c("inc.d3", "m.inc.d3", "inc.d9", "m.inc.d9", "inc.d17", "m.inc.d17", "hatch","m.n2"))  %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment,  color = sex)) +
-    geom_boxplot() + 
-  theme_B3()  +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
-          legend.position = "none",
-          strip.text = element_blank()) +
-    scale_fill_manual(values = colorscharmaip) +
-    scale_color_manual(values = sexcolors) +
-    labs(y = "PRL (ng/mL)", x = NULL) +
-    ylim(c(-16,100)) +
-
-   annotation_custom(inc, xmax = 1.6, ymin = -14, ymax = 5) +
-   annotation_custom(removeegg, xmax = 3.6, ymin = -14, ymax = 5) +
-   annotation_custom(inc, xmax = 5.6, ymin = -14, ymax = 5) +
-   annotation_custom(removeegg, xmax = 7.6, ymin = -14, ymax = 5) +
-  annotation_custom(hatch, xmax = 9.6, ymin = -14, ymax = 5) +
-  annotation_custom(removeegg, xmax = 11.6, ymin = -14, ymax = 5) +
-  annotation_custom(nestling, xmax = 13.6, ymin = -14, ymax = 5) +
-  annotation_custom(removechick, xmax = 15.6, ymin = -14, ymax = 5) +
-
-    annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -16, ymax = -14, alpha = 0.25) +
-    annotate("rect", xmin = 2.6, xmax = 4.4, ymin = -16, ymax = -14, alpha = 0.5) +
-    annotate("rect", xmin = 4.6, xmax = 6.4, ymin = -16, ymax = -14, alpha = 0.75)  +
-    annotate("rect", xmin = 6.6, xmax = 8.4, ymin = -16, ymax = -14, alpha = 1)
-
-hormones %>% 
-    filter( hormone == c("prolactin"))  %>% 
-    filter( treatment %in% c("inc.d9", "m.inc.d8", "inc.d17", "prolong", "hatch", "extend", "n5"))  %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment,  color = sex)) +
-    geom_boxplot() + 
-  theme_B3()  +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
-          legend.position = "none",
-          strip.text = element_blank()) +
-    scale_fill_manual(values = colorscharmaip) +
-    scale_color_manual(values = sexcolors) +
-    labs(y = "PRL (ng/mL)", x = NULL) +
-  ylim(c(-12,100)) +
-   annotation_custom(inc, xmax = 1.6, ymin = -10, ymax = 10) +
-   annotation_custom(maniphatch, xmax = 3.6, ymin = -10, ymax = 10) +
-   annotation_custom(inc, xmax = 5.6, ymin = -10, ymax = 10) +
-   annotation_custom(manipinc, xmax = 7.6, ymin = -10, ymax = 10) +
-  annotation_custom(hatch, xmax = 9.6, ymin = -10, ymax = 10) +
-  annotation_custom(maniphatch, xmax = 11.6, ymin = -10, ymax = 10) +
-  annotation_custom(nestling, xmax = 13.6, ymin = -10, ymax = 10) +
-
-  annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -12, ymax = -10, alpha = 0.33) +
-    annotate("rect", xmin = 4.6, xmax = 5.4, ymin = -12, ymax = -10, alpha = 0.33) +
-    annotate("rect", xmin = 2.6, xmax = 5.4, ymin = -10, ymax = -8, alpha = 0.66) +
-    annotate("rect", xmin = 4.6, xmax = 7.4, ymin = -8, ymax = -6, alpha = 1) 
-```
-
-```{r sexsteroids}
-
-hormonecharplot("corticosterone", "CORT (ng/mL)")
-hormonecharplot("progesterone", "PROG (ng/mL)") 
-d1 <- hormonecharplot("estradiol", "E (ng/mL)")
-d2 <- hormonecharplot("testosterone", "T (ng/mL)")
-plot_grid(d1,d2, nrow = 1)
-
-hormonemanipSteroids <- function(myhormone, myylab, myymax){
-
-  hormones %>% 
-    filter( hormone %in% c(myhormone))  %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
-    geom_boxplot() + 
-    mytheme() +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
-         legend.position = "none",
-          strip.text = element_blank()) +
-    scale_fill_manual(values = colorscharmaip) +
-    scale_color_manual(values = sexcolors) +
-    labs(y = myylab, x = NULL) 
-}
-
-hormonemanipSteroids("corticosterone", "CORT (ng/mL)", 10)
-hormonemanipSteroids("progesterone", "PROG (ng/mL)",2.5)
-
-a <- hormonemanipSteroids("estradiol", "E (ng/mL)", 1)
-b <- hormonemanipSteroids("testosterone", "T (ng/mL)", 3.5)
-plot_grid(a,b)
-
-```
-
-```{r aov-char}
-prl.char <- hormones %>% filter(hormone == "prolactin", treatment %in% charlevels)   %>%  droplevels()
-test.char <- hormones %>% filter(hormone == "testosterone", treatment %in% charlevels)   %>%  droplevels()
-est.char <- hormones %>% filter(hormone == "estradiol", treatment %in% charlevels)   %>%  droplevels()
-prog.char <- hormones %>% filter(hormone == "progesterone", treatment %in% charlevels)   %>%  droplevels()
-cort.char <- hormones %>% filter(hormone == "corticosterone", treatment %in% charlevels)   %>%  droplevels()
-
-aovSexTretment <- function(mydata, whichormone){
-  aov2 <- aov(data = mydata, plasma_conc ~ treatment * sex)
-  print(whichormone)
-  print(summary(aov2))
-  #print(TukeyHSD(aov2, which = "treatment"))
-}
-
-aovTretment  <- function(mydata, whichormone){
-  aov1 <- aov(data = mydata, plasma_conc ~ treatment )
-  print(whichormone)
-  print(summary(aov1))
-  #print(TukeyHSD(aov1, which = "treatment"))
-}
-
-aovSexTretment(prl.char, "PRL") # yes, sex difference (p = 0.00256), yes treatment effect (p < 2e-16), no interaction
-aovSexTretment(cort.char, "corticosterone") # no sex difference (p = 0.779 ), small treatment effect (p = 0.0238),  no interaction
-aovSexTretment(prog.char, "PROG") # no sex difference or treatment effect, signifiant interacion (p = 0.0107)
-
-aovTretment(est.char, "E") # p = 0.101
-aovTretment(test.char, "T") # p = 0.609
-```
-
 # do control bird with high prolactin hormone have high PRL expression in the pituitary? yes.
@@
 head(PRLpit,2)
-```
-
-```{r sex}
-hormones %>%
-  filter(study == "characterization") %>%
-  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
-  geom_boxplot() +
-  theme_B3() +
-  facet_wrap(~hormone, nrow = 1, scales = "free") +
-  theme(legend.position = "none") +
-  labs(y = "concentration (ng/mL")
-
-printaovtablesex <- function(myhormone, mydescirption){
-
-  aov_all = data.frame()
-  for(i in alllevels){
-    df <- hormones %>% filter(hormone == myhormone,
-                            treatment == i) %>% droplevels()
-    aovtable <- apa.aov.table(aov(plasma_conc ~ sex, data  = df))
-    aovtable <- as.data.frame(aovtable$table_body)
-    totaldf <- aovtable[3, 3]
-    aovtable$df <- paste(aovtable$df, ", " , totaldf, sep = "")
-    aovtable$ANOVA <- mydescirption
-    aovtable$stages <- paste(i)
-    aovtable$p <- as.numeric(as.character(aovtable$p))
-    aov_all <- rbind(aov_all,aovtable)
-  }
-
-  aov_all <- aov_all %>%
-    filter(Predictor == "sex")  %>%
-    select(stages, ANOVA, df, "F", p) %>%
-    mutate(sig = ifelse(p < 0.05, "*", " "))
-  aov_all
-  return(aov_all)
-}
-
-table1 <- printaovtablesex("prolactin", "PRL ~ sex")
-table1
-write.csv(table1, "../../parentalhormones/data/table-1.csv", row.names = F)
-```
-
-```{r cort}
-
-hormones %>%
-  filter(hormone == "corticosterone") %>%
-  group_by(sex) %>%
-  summarise(mean = mean(plasma_conc),
-            median = median(plasma_conc))
-
-p1 <- hormones %>%
-  filter(hormone == "corticosterone",
-         study == "characterization") %>%
-  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
-  geom_boxplot() +
-  theme_B3() +
-  labs(y = "CORT (ng/mL)", subtitle = element_blank() , x = NULL) +
-  theme(legend.position = "none") +
-  scale_fill_manual(values = sexcolors) +
-  scale_fill_manual(values = colorscharmaip)
-p1
-
-cort <- hormones %>%
-  filter(hormone == "corticosterone",
-         study == "characterization")
-
-p2 <-  ggplot(cort, aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot() +
-  theme_B3() +
-  labs(y = NULL, x = NULL, subtitle = element_blank() ) +
-  theme(legend.position = "none")  +
-  scale_fill_manual(values = colorscharmaip)
-
-p2
-
-p12 <- plot_grid(p1,p2, rel_widths = c(0.2,0.8), labels = c("h","i"), label_size = 8)
-
-p3 <- hormones %>%
-  filter(hormone == "corticosterone",
-         treatment %in% c(controlsremoval, levelsremoval )) %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot() +
-  theme_B3() +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none") +
-  labs( x = NULL, y = "CORT (ng/mL") +
-  ylim(c(0,12.5))
-p3
-
-p4 <- hormones %>%
-  filter(hormone == "corticosterone",
-         treatment %in% c(controlstiming, levelstiming )) %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot() +
-  theme_B3() +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none") +
-  labs(y = NULL, x = NULL) + 
-  ylim(c(0,12.5))
-p4
-
-p34 <- plot_grid(p3,p4, rel_widths = c(0.6,0.4), labels = c("j","k"), label_size = 8 )
-p34
-
-allcortplots <- plot_grid(p12, p34, nrow = 2, rel_heights = c(0.6,0.4))
-allcortplots
-```
-
-```{r PRL}
-hormones %>%
-  filter(hormone == "prolactin") %>%
-  group_by(sex) %>%
-  summarise(m = mean(plasma_conc),
-            median = median(plasma_conc))
-
-p1 <- hormones %>%
-  filter(hormone == "prolactin",
-         study == "characterization") %>%
-  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
-  geom_boxplot() +
-  theme_B3() +
-  labs(y = "Prolactin (ng/mL)", subtitle = " " , x = NULL) +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_fill_manual(values = sexcolors) +
-  scale_y_continuous(limits = c(0,100)) 
-p1
-
-PRLf <- hormones %>%
-  filter(hormone == "prolactin",
-         #study == "characterization",
-         sex == "female")
-
-PRLm <- hormones %>%
-  filter(hormone == "prolactin",
-         #study == "characterization",
-         sex == "male")
-
-p2a <-  ggplot(PRLf, aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot(aes(color = sex)) +
-  theme_B3() +
-  labs(y = NULL, x = NULL, subtitle = "females") +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_color_manual(values = sexcolors) +
-  scale_y_continuous(limits = c(0,100))  +
-  scale_fill_manual(values = colorscharmaip)
-
-p2b <-  ggplot(PRLm, aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot(aes(color = sex)) +
-  theme_B3() +
-  labs(y = NULL, x = NULL, subtitle = "males") +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_color_manual(values = sexcolors) +
-  scale_y_continuous(limits = c(0,100))+
-  scale_fill_manual(values = colorscharmaip)
-
-p12 <- plot_grid(p1,p2a,p2b, rel_widths = c(0.25,0.5, 0.5), labels = c("a","b", "c"), label_size = 8, nrow = 1, align = "hv")
-p12
-
-p3a <- hormones %>%
-  filter(hormone == "prolactin",
-         sex == "female", 
-         treatment %in% c(controlsremoval, levelsremoval )) %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot(aes(color = sex)) +
-  theme_B3() +
-  scale_y_continuous(limits = c(0,100)) +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs( x = NULL, y = "Prolactin (ng/mL", subtitle = "females") +
-  scale_color_manual(values = sexcolors) 
-p3a
-
-p3b <- hormones %>%
-  filter(hormone == "prolactin",
-         sex == "male", 
-         treatment %in% c(controlsremoval, levelsremoval )) %>% 
-  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot(aes(color = sex)) +
-  theme_B3() +
-  scale_y_continuous(limits = c(0,100)) +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs( x = NULL, y = NULL, subtitle = "males") +
-  scale_color_manual(values = sexcolors)
-p3b
-
-p4a <- hormones %>%
-  filter(hormone == "prolactin",
-         sex == "female", 
-         treatment %in% c(controlstiming, levelstiming )) %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot(aes(color = sex)) +
-  theme_B3() +
-  scale_y_continuous(limits = c(0,100)) +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs(y = NULL, x = NULL, subtitle = "females") +
-  scale_color_manual(values = sexcolors)
-p4a
-
-p4b <- hormones %>%
-  filter(hormone == "prolactin",
-         sex == "male", 
-         treatment %in% c(controlstiming, levelstiming )) %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot(aes(color = sex)) +
-  theme_B3() +
-  scale_y_continuous(limits = c(0,100)) +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs(y = NULL, x = NULL, subtitle = "males") +
-  scale_color_manual(values = sexcolors)
-
-p4b
-
-p34 <- plot_grid(p3a , p3b,p4a, p4b, rel_widths = c(0.55,0.55, 0.45,0.45), labels = c("d","e", "f", "g"), label_size = 8, nrow = 1)
-
-allprolactinplots <- plot_grid(p12, p34, nrow = 2, rel_heights = c(0.5,0.5))
-allprolactinplots
-```
-
-## figure 2
-
-```{r figure2, fig.width=7, fig.height=7}
-
-plot_grid(allprolactinplots, allcortplots, nrow = 2)
-
-```
-
-## correlations
-
-```{r correlations2 }
-
-plotcorrelations <- function(i,j, myxlab, mylab, mysubtitle){
-      p <- ggplot(hormoneswide, aes(x = i, y = j), na.rm = TRUE) +
-       geom_point(aes(color = treatment), na.rm = TRUE) +
-       geom_smooth( method = lm, color = "darkgrey", na.rm = TRUE ) +
-       theme_B3() +
-      #facet_wrap(~sex, scales = "free") +
-        labs(x = myxlab, y = mylab, subtitle = mysubtitle) +
-        theme(legend.position = "none") +
-        scale_color_manual(values = colorscharmaip2)
-
-      return(p)
-}
-
-hormoneswide
-
-corrPRL <- hormoneswide %>%
-  pivot_longer(cols = corticosterone:progesterone, names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
-  filter(study == "characterization") %>%
-  ggplot(aes(x = prolactin, y = conc)) +
-  geom_point(aes(color = treatment)) +
-  geom_smooth( color = "darkgrey", method = "lm") +
-  facet_wrap(~hormones, nrow = 1, scales = "free_y")  +
-  theme_B3() +
-  theme(legend.position = "bottom",
-        legend.title = element_blank()) +
-  scale_color_manual(values = colorscharmaip2) +
-  guides(color = guide_legend(nrow = 1)) +
-  labs(y = "conc. (ng/mL)", x = "prolactin (ng/mL)")
-corrPRL
-
-corrCORT <- hormoneswide %>%
-  pivot_longer(cols = c(prolactin, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
-  filter(study == "characterization") %>%
-  ggplot(aes(x = corticosterone, y = conc)) +
-  geom_point(aes(color = treatment)) +
-  geom_smooth( color = "darkgrey", method = "lm") +
-  facet_wrap(~hormones, nrow = 1, scales = "free_y")  +
-  theme_B3() +
-  theme(legend.position = "bottom",
-        legend.title = element_blank()) +
-  scale_color_manual(values = colorscharmaip2) +
-  guides(color = guide_legend(nrow = 1)) +
-  labs(y = "conc. (ng/mL)", x = "corticosterone (ng/mL)")
-corrCORT
-
-hormoneswide %>%
-  pivot_longer(cols = c(prolactin, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
-  filter(study == "characterization") %>%
-  ggplot(aes(x = corticosterone, y = conc)) +
-  geom_point(aes(color = treatment)) +
-  geom_smooth(color = "darkgrey", method = "lm") +
-  facet_grid(hormones ~ treatment, scales = "free_y")  +
-  theme_B3() +
-  theme(legend.position = "bottom",
-        legend.title = element_blank()) +
-  scale_color_manual(values = colorscharmaip2) +
-  guides(color = guide_legend(nrow = 1)) +
-  labs(y = "conc. (ng/mL)", x = "corticosterone (ng/mL)")
-
-hormoneswide %>%
-  pivot_longer(cols = c(corticosterone, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
-  filter(study == "characterization") %>%
-  ggplot(aes(x = prolactin, y = conc)) +
-  geom_point(aes(color = treatment)) +
-  geom_smooth(color = "darkgrey", method = "lm") +
-  facet_grid(hormones ~ treatment, scales = "free_y")  +
-  theme_B3() +
-  theme(legend.position = "bottom",
-        legend.title = element_blank()) +
-  scale_color_manual(values = colorscharmaip2) +
-  guides(color = guide_legend(nrow = 1)) +
-  labs(y = "conc. (ng/mL)", x = "prolactin (ng/mL)")
-
-ggplot(hormoneswide, aes(x = prolactin, y = corticosterone, color = treatment)) +
-  geom_point() + geom_smooth(method = "lm") +
-  facet_wrap(~treatment) +
-  scale_color_manual(values =  colorscharmaip) + theme(legend.position = "none")
-
-ggplot(hormoneswide, aes(x = prolactin, y = testosterone, color = treatment)) +
-  geom_point() + geom_smooth(method = "lm") +
-  facet_wrap(~treatment) +
-  scale_color_manual(values =  colorscharmaip) + theme(legend.position = "none")
-
-```
-
-## progesterone
-
-```{r progesterone}
-
-hormones %>%
-  filter(hormone == "progesterone") %>%
-  group_by(sex) %>%
-  summarise(mean = mean(plasma_conc),
-            median = median(plasma_conc))
-
-p1 <- hormones %>%
-  filter(hormone == "progesterone",
-         study == "characterization") %>%
-  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
-  geom_boxplot() +
-  theme_B3() +
-  labs(y = "Progesterone (ng/mL)", subtitle = element_blank() , x = NULL) +
-  theme(legend.position = "none")   +
-  scale_y_continuous(limits = c(0,4.5)) +
-  scale_fill_manual(values = sexcolors)
-
-df <- hormones %>%
-  filter(hormone == "progesterone",
-         study == "characterization")
-
-p2 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot() +
-  theme_B3() +
-  labs(y = NULL, x = NULL, subtitle = " ") +
-  scale_y_continuous(limits = c(0,4.5)) +
-  theme(legend.position = "none")
-
-p3 <- hormones %>%
-  filter(hormone == "progesterone",
-         treatment %in% c(controlsremoval, levelsremoval )) %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot() +
-  theme_B3() +
-  scale_y_continuous(limits = c(0,4.5)) +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none") +
-  labs( x = NULL, y = "Progesterone (ng/mL")
-
-p4 <- hormones %>%
-  filter(hormone == "progesterone",
-         treatment %in% c(controlstiming, levelstiming )) %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot() +
-  theme_B3() +
-  scale_y_continuous(limits = c(0,4.5)) +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none") +
-  labs(y = NULL, x = NULL)
-
-progesterone <- plot_grid(p1, p2, p3 + labs(y = NULL), p4, nrow = 1, rel_widths =   c(0.175,0.4,0.35,0.25))
-progesterone
-```
-
-## estradiol and testosterone
-
-```{r estradiol}
-
-hormones %>%
-  filter(hormone == "estradiol") %>%
-  summarise(mean = mean(plasma_conc),
-            median = median(plasma_conc))
-
-p5 <- hormones %>%
-  filter(hormone == "estradiol",
-         study == "characterization") %>%
-  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
-  geom_boxplot() +
-  theme_B3() +
-  labs(y = "Estradiol (ng/mL)", subtitle = element_blank() , x = NULL) +
-  theme(legend.position = "none")  +
-    scale_y_continuous(limits = c(0,1)) +
-  scale_fill_manual(values = sexcolors)
-
-df <- hormones %>%
-  filter(hormone == "estradiol",
-         study == "characterization")
-
-p6 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot() +
-  theme_B3() +
-  labs(y = NULL, x = NULL, subtitle = " ") +
-    scale_y_continuous(limits = c(0,1)) +
-  theme(legend.position = "none")
-
-p7 <- hormones %>%
-  filter(hormone == "estradiol",
-         treatment %in% c(controlsremoval, levelsremoval )) %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot() +
-  theme_B3() +
-  scale_y_continuous(limits = c(0,1)) +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none") +
-  labs( x = NULL, y = "Estradiol (ng/mL")
-
-p8 <- hormones %>%
-  filter(hormone == "estradiol",
-         treatment %in% c(controlstiming, levelstiming )) %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot() +
-  theme_B3() +
-  scale_y_continuous(limits = c(0,1)) +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none") +
-  labs(y = NULL, x = NULL)
-
-estradiol <- plot_grid(p5, p6, p7 + labs(y = NULL), p8, nrow = 1, rel_widths =   c(0.175,0.4,0.35,0.25))
-estradiol
-```
-
-```{r testosterone}
-
-hormones %>%
-  filter(hormone == "testosterone") %>%
-  summarise(mean = mean(plasma_conc),
-            median = median(plasma_conc))
-
-p9 <- hormones %>%
-  filter(hormone == "testosterone",
-         study == "characterization") %>%
-  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
-  geom_boxplot() +
-  theme_B3() +
-  labs(y = "Testosterone (ng/mL)", subtitle = element_blank() , x = NULL) +
-  theme(legend.position = "none")  +
-    scale_y_continuous(limits = c(0,8.5)) +
-  scale_fill_manual(values = sexcolors)
-
-df <- hormones %>%
-  filter(hormone == "testosterone",
-         study == "characterization")
-
-p10 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot() +
-  theme_B3() +
-  labs(y = NULL, x = NULL, subtitle = " ") +
-    scale_y_continuous(limits = c(0,8.5)) +
-  theme(legend.position = "none")
-
-p11 <- hormones %>%
-  filter(hormone == "testosterone",
-         treatment %in% c(controlsremoval, levelsremoval )) %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot() +
-  theme_B3() +
-  scale_y_continuous(limits = c(0,8.5)) +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none") +
-  labs( x = NULL, y = "concentration (ng/mL")
-
-p12 <- hormones %>%
-  filter(hormone == "testosterone",
-         treatment %in% c(controlstiming, levelstiming )) %>% 
-ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
-  geom_boxplot() +
-  theme_B3() +
-  scale_y_continuous(limits = c(0,8.5)) +
-  scale_fill_manual(values = colorscharmaip) +
-  theme(legend.position = "none") +
-  labs(y = NULL, x = NULL)
-p12
-
-testosterone <- plot_grid(p9, p10, p11 + labs(y = NULL), p12, nrow = 1, rel_widths =   c(0.175,0.4,0.35,0.25))
-testosterone
-```
-
-```{r sexsteroids2, fig.height=8}
-sexsteroids <- plot_grid(
-       p9 + theme(axis.text.x = element_blank()),  
-       p10 + theme(axis.text.x = element_blank()), 
-       p11 + labs(y = NULL) + theme(axis.text.x = element_blank()), 
-       p12 + theme(axis.text.x = element_blank()),
-        p5 + theme(axis.text.x = element_blank()), 
-       p6 + theme(axis.text.x = element_blank()), 
-       p7 + labs(y = NULL) + theme(axis.text.x = element_blank()), 
-       p8 + theme(axis.text.x = element_blank()), 
-       p1 , 
-       p2 + theme(axis.text.x = element_text(angle = 45, hjust = 1)), 
-       p3 + labs(y = NULL)  + theme(axis.text.x = element_text(angle = 45, hjust = 1)), 
-       p4  + theme(axis.text.x = element_text(angle = 45, hjust = 1)), 
-        nrow = 3, 
-        rel_widths =   c(0.175,0.4,0.35,0.25), align = "hv",
-          label_size = 8, labels = c("A1", "A2", "A3" , "A4",
-                                     "B1", "B2", "B3" , "B4",
-                                     "C1", "C2", "C3" , "C3"))
-sexsteroids
-
-```
-
-## pca of hormones
-
-```{r PCA}
-
-library("factoextra")
-library("FactoMineR")
-
-# char only
-hormoneswidechar <- hormoneswide %>% filter(treatment %in% charlevels)
-df <- hormoneswidechar %>% select(prolactin:progesterone)
-res.pca <- PCA(df)
-
-a1 <- fviz_pca_var(res.pca, col.var = "black", title = "Charactherization only") 
-# Contributions of variables to PC1
-
-b1 <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10) + labs(y = "Dim 1 contibutions (%)",
-                                                                      title = element_blank())
-# Contributions of variables to PC2
-c1 <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10) +
-  labs(y = "Dim 2 contibutions (%)") + theme(title = element_blank())
-
-ones <- plot_grid(a1,b1,c1, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")
-
-# char and manip
-df <- hormoneswide %>% select(prolactin:progesterone)
-res.pca <- PCA(df)
-
-a2 <- fviz_pca_var(res.pca, col.var = "black", title = "Charactherization & manipluation") 
-# Contributions of variables to PC1
-
-b2 <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10) + labs(y = "Dim 1 contibutions (%)",
-                                                                      title = element_blank())
-# Contributions of variables to PC2
-c2 <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10) +
-  labs(y = "Dim 2 contibutions (%)") + theme(title = element_blank())
-
-twos <- plot_grid(a2 + theme(title = element_blank()),b2,c2, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")
-
-plot_grid(ones,twos, nrow = 2)
-
-twos
-```
-
-```{r figure3, fig.height=8, fig.width=7}
-
-plot_grid(twos,sexsteroids, nrow = 2, rel_heights = c(0.25,0.75), align = "hv")
-```
-
-# smooth funciton
-
-```{r summary}
-head(hormones)
-ggplot(hormones, aes(x = as.numeric(treatment), y = plasma_conc, color = hormone)) +
-  geom_smooth()
-
-box <- ggplot(hormones, aes(x = as.numeric(treatment), y = plasma_conc)) +
-  geom_boxplot(aes(fill = treatment)) +
-  #geom_smooth(aes(color = hormone)) +
-  facet_wrap(~hormone, scales = "free_y", nrow = 5) +
-  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16),
-                     labels = alllevels) +
-  theme_B3() +
-  theme(#strip.text = element_blank(),
-        axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "none",
-        panel.background = element_rect(fill = "transparent",colour = NA),
-        plot.background = element_rect(fill = "transparent",colour = NA)) +
-  labs(y = "plasma conc. (ng/mL)", x = "parental stage") +
-  scale_fill_manual(values = colorscharmaip)
-
-line <- ggplot(hormones, aes(x = as.numeric(treatment), y = plasma_conc)) +
-  #geom_boxplot(aes(fill = treatment)) +
-  geom_smooth(aes(color = hormone)) +
-  facet_wrap(~hormone, scales = "free_y", nrow = 5) +
-  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16),
-                     labels = alllevels) +
-  theme_B3() +
-  theme(#strip.text = element_blank(),
-        axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "none",
-        panel.background = element_rect(fill = "transparent",colour = NA),
-        plot.background = element_rect(fill = "transparent",colour = NA)) +
-  labs(y = "plasma conc. (ng/mL)", x = "parental stage")
-
-plot_grid(box, line, nrow = 1)
-```
-
-## stats
-
-```{r}
-
-P3 <- hormones %>% filter(hormone == "progesterone")
-E2 <- hormones %>% filter(hormone == "estradiol")
-T2 <- hormones %>% filter(hormone == "testosterone")
-
-aov1 <- aov(PRLm$logconc ~ PRLm$treatment)
-res1 <- TukeyHSD(aov1)
-res1 <- as.data.frame(res1$`PRLm$treatment`)
-res1$hormone <- "male PRL"
-res1$sig <- ifelse(res1$`p adj` < 0.001, "***",
-                   ifelse(res1$`p adj` < 0.01, "**",
-                           ifelse(res1$`p adj` < 0.05, "*", "NS" )))
-res1$comparison <- row.names(res1)
-res1
-
-aov2 <- aov(PRLf$logconc ~ PRLf$treatment)
-res2 <- TukeyHSD(aov2)
-res2 <- as.data.frame(res2$`PRLf$treatment`)
-res2$hormone <- "female PRL"
-res2$sig <- ifelse(res2$`p adj` < 0.001, "***",
-                   ifelse(res2$`p adj` < 0.01, "**",
-                           ifelse(res2$`p adj` < 0.05, "*","NS"  )))
-res2$comparison <- row.names(res2)
-res2
-
-aov3 <- aov(P3$logconc ~ P3$treatment)
-res3 <- TukeyHSD(aov3)
-res3 <- as.data.frame(res3$`P3$treatment`)
-res3$hormone <- "P3"
-res3$sig <- ifelse(res3$`p adj` < 0.001, "***",
-                   ifelse(res3$`p adj` < 0.01, "**",
-                           ifelse(res3$`p adj` < 0.05, "*", "NS"  )))
-res3$comparison <- row.names(res3)
-res3
-
-aov4 <- aov(E2$logconc ~ E2$treatment)
-res4 <- TukeyHSD(aov4)
-res4 <- as.data.frame(res4$`E2$treatment`)
-res4$hormone <- "E2"
-res4$sig <- ifelse(res4$`p adj` < 0.001, "***",
-                   ifelse(res4$`p adj` < 0.01, "**",
-                           ifelse(res4$`p adj` < 0.05, "*", "NS" )))
-res4$comparison <- row.names(res4)
-res4
-
-aov5 <- aov(T2$logconc ~ T2$treatment)
-res5 <- TukeyHSD(aov5)
-res5 <- as.data.frame(res5$`T2$treatment`)
-res5$hormone <- "T"
-res5$sig <- ifelse(res5$`p adj` < 0.001, "***",
-                   ifelse(res5$`p adj` < 0.01, "**",
-                           ifelse(res5$`p adj` < 0.05, "*", "NS"  )))
-res5$comparison <- row.names(res5)
-res5
-
-resall <- rbind(res3, res5, res4, res2, res1) %>% drop_na()
-resall
-
-table1 <- resall %>%
-  filter(grepl('control', comparison),
-         !grepl('prolong|extend|m.', comparison)) %>%
-   mutate(diff = round(diff, 2)) %>% 
-  mutate(diffsig = paste(diff, sig, sep = ", ")) %>% 
-   select(hormone, comparison, diffsig) %>% 
-  pivot_wider( names_from = hormone, values_from = diffsig)
-table1
-
-table2 <- resall %>%
-  filter(!grepl('control|bldg|lay|n5', comparison),
-         grepl('prolong|extend|m.', comparison)) %>%
-   mutate(diff = round(diff, 2)) %>% 
-  mutate(diffsig = paste(diff, sig, sep = ", ")) %>% 
-   select(hormone, comparison, diffsig) %>% 
-  pivot_wider( names_from = hormone, values_from = diffsig)
-table2
-
 ```

### 2020-02-09 analysis/07_hormones.Rmd
new hormone stuff

@@
                                 characterization = charlevels,
-                                manipulation = maniplevels1)) %>% 
+                                manipulation = maniplevels1),
+                       moltbin = as.integer(moltbin)) %>% 
               dplyr::mutate(bird_id = gsub("[[:punct:]]", "." , id)) %>% 
               dplyr::select(bird_id, treatment, sex, study, 
-                            corticosterone, estradiol, testosterone, progesterone) %>% 
-              pivot_longer(cols = corticosterone:progesterone,
+                            corticosterone, estradiol, testosterone, progesterone, moltbin) %>% 
+              pivot_longer(cols = corticosterone:moltbin,
                            names_to = "hormone", values_to = "plasma_conc",
@@
 
-write.csv(hormones, "../results/hormones.csv", row.names = F)
-
 # make a winder one for correlations
@@
 
-# for facet wrapping
-hormones <- left_join(hormones, birds)
-
+# for faceting
+hormones <- hormones %>% filter(hormone != "moltbin")
 hormones$treatment <- factor(hormones$treatment, levels = alllevels)
-
 hormones$hormone <- factor(hormones$hormone, 
@@
                                       "estradiol", "testosterone"))
+# for stats
 hormones$logconc <- log10(hormones$plasma_conc)
@@
 write.csv(hormoneswide, "../results/07_hormoneswide.csv", row.names = F)
+
+#write.csv(hormones, "../../parentalhormones/data/hormones.csv", row.names = F)
+#write.csv(hormoneswide, "../../parentalhormones/data/hormoneswide.csv", row.names = F)
 ```

### 2020-02-09 analysis/figure-1.Rmd
now with tsne and fig 1

--- 2019-12-22 analysis/02_PCA.Rmd
@@
 ---
-title: "PCA"
+title: "Fig1"
 output: md_document
@@
 
-knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/pca/', cache = T)
-
+knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
 ```
 
-# note. this file is too big for storage in github :(
+# Figure 1. Experimental design and tSNE analysis
 
-## Data: Pseudocounts from Limma
+## wrange data
+
+Note, the input file is too big for storage on github :(
 
@@
 ```
-
-## All data, characterization and manipulations
 
@@
 
+ftsne <-  subsetmaketsne(tissuelevels, charlevels, "female")
+mtsne <-  subsetmaketsne(tissuelevels, charlevels, "male")
+
 ```
 
-```{r tSNE}
-plotcolorfultsnes <- function(tsnedf, whichfactor, whichcolors){
-  p <- ggplot(tsnedf, aes(x = V1, y = V2, color = whichfactor, shape = tissue)) +
-    geom_point(size = 1) +
+## make figure
+
+```{r fig1, fig.width=7}
+
+plottsneelipse <- function(tsnedf, whichfactor, whichcolors){
+  p <- ggplot(tsnedf, aes(x = V1, y = V2)) +
+    geom_point(size = 1, aes(color = whichfactor)) +
     theme_B3() +
@@
           axis.text = element_blank()) +
-    scale_shape_manual(values = myshapes)
-  print(p)
+    stat_ellipse(linetype = 2, aes(color = tissue )) 
+  return(p)
 }
 
-a <- plotcolorfultsnes(chartsne, chartsne$tissue, colorstissue)  
-b <- plotcolorfultsnes(chartsne, chartsne$sex, sexcolors)   
-c <- plotcolorfultsnes(chartsne, chartsne$treatment, colorscharmaip)
+a <- plottsneelipse(chartsne, chartsne$tissue, allcolors) + labs(subtitle = "tissue\n")    
+b <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = " ", subtitle = "tissue * sex\n")    
+c <- plottsneelipse(ftsne, ftsne$treatment, allcolors ) + labs(y = " ", subtitle = "females\ntissue * treatment")
+d <- plottsneelipse(mtsne, mtsne$treatment, allcolors ) + labs(y = " ", subtitle = "males\ntissue * treatment")
 
-abc <- plot_grid(a,b,c, nrow = 1, labels = c("a", "b", "c"), label_size = 12)
-abc
+abcd <- plot_grid(a,b,c,d, nrow = 1, labels = c("b", "c", "d", "e"), label_size = 12 )
 
+expdesign <- png::readPNG("../figures/images/fig1a_fig1a.png")
+expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
+
+fig1 <- plot_grid(expdesign, abcd, nrow = 2, labels = c("a", "b"), label_size = 12, rel_heights = c(0.5,1))
+fig1
 ```
-
-```{r subsetmakepca}
-subsetmakepca <- function(whichtissue, whichtreatment, whichsex){
-
-  colData <- colData %>%
-    dplyr::filter(tissue %in% whichtissue,
-                  treatment %in% whichtreatment,
-                  sex %in% whichsex) 
-  row.names(colData) <- colData$V1
-
-  # save counts that match colData
-  savecols <- as.character(colData$V1) 
-  savecols <- as.vector(savecols)
-
-  countData <- as.data.frame(t(countData))
-  countData <- countData %>% dplyr::select(one_of(savecols)) 
-  countData <- as.data.frame(t(countData))
-
-  mypca <- prcomp(countData)
-
-  mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
-                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
-                  ID = row.names(countData))
-  mypcadf$V1 <- row.names(mypcadf)
-  mypcadf <- left_join(colData, mypcadf)
-  mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
-  return(mypcadf)
-}
-
-charpca <- subsetmakepca(tissuelevels, charlevels, sexlevels)
-hyppca <- subsetmakepca("hypothalamus", charlevels, sexlevels)
-pitpca <- subsetmakepca("pituitary", charlevels, sexlevels)
-gonpca <- subsetmakepca("gonads", charlevels, sexlevels)
-
-```
-
-```{r pca}
-plotcolorfulpcs <- function(mypcadf,  whichfactor, whichcolors){
-  p <- mypcadf %>%
-    ggplot(aes(x = PC1, y = PC2, shape = tissue, color = whichfactor )) +
-    geom_point(size = 1)  +
-    theme_B3() +
-    theme(legend.title = element_blank(),
-         axis.text = element_blank(),
-         legend.position = "none") +
-    labs(x = "PC1", y = "PC2")  +
-    scale_color_manual(values = whichcolors) +
-    scale_shape_manual(values = myshapes)
-  print(p)
-}
-
-d <- plotcolorfulpcs(charpca, charpca$tissue, colorstissue)  
-e <- plotcolorfulpcs(charpca, charpca$sex, sexcolors) 
-f <- plotcolorfulpcs(charpca,charpca$treatment, colorscharmaip)
-
-def <- plot_grid(d,e, f, nrow =  1,  labels = c("d", "e", "f"), label_size = 12)
-def
-```
-
-```{r fviz}
-
-makefvizdf <-  function(whichtissue, whichtreatment, whichsex){
-  colData <- colData %>%
-      dplyr::filter(tissue %in% whichtissue,
-                  treatment %in% whichtreatment,
-                  sex %in% whichsex) 
-  row.names(colData) <- colData$V1
-
-  # save counts that match colData
-  savecols <- as.character(colData$V1) 
-  savecols <- as.vector(savecols)
-
-  countData <- as.data.frame(t(countData))
-  countData <- countData %>% dplyr::select(one_of(savecols)) 
-  countData <- as.data.frame(t(countData))
-
-  mypca <- prcomp(countData)
-  return(mypca)
-}
-
-charfviz <- makefvizdf(tissuelevels, charlevels, sexlevels)
-hypv <- makefvizdf("hypothalamus", charlevels, sexlevels)
-pitfviz <- makefvizdf("pituitary", charlevels, sexlevels)
-gonfviz <- makefvizdf("gonads", charlevels, sexlevels)
-
-g <- fviz_screeplot(charfviz, addlabels = TRUE, ylim = c(0, 50),  ncp = 5, barcolor = "white", barfill = "white") + 
-  labs(title = NULL, y = "PC Variance") + theme_B3() 
-h <- fviz_pca_var(charfviz,  labelsize = 3 , axes.linetype = "blank", 
-                   repel = TRUE ,
-                  select.var= list(contrib = 5)) + 
-      theme_B3() + 
-      labs(x = "PC1 (35.8%)", y = "PC2 (32.5%)", title =  NULL) +
-      theme( axis.text = element_blank())
-
-legend <- png::readPNG("../figures/images/DoveParentsRNAseq_legendchar.png")
-legend <-  grid::rasterGrob(legend, interpolate=TRUE)
-
-gh <- plot_grid(g,h, legend, nrow = 1, rel_widths = c(1, 1, 1), labels = c("h", "i", NULL ), label_size = 12)
-gh
-plot_grid(def, gh, nrow = 2)
-```
-
-```{r PCA-tSNE, fig.height= 7, fig.width=7}
-plot_grid( abc, def, gh, nrow = 3)
-
-```
-
-## tissue specific
-
-```{r tissuespecific, fig.height=7,fig.width=7}
-
-## tsne 
-h4 <- plotcolorfultsnes(hyptsne, hyptsne$sex, sexcolors)  + labs(y = "hypothalamus") 
-h5 <- plotcolorfultsnes(hyptsne, hyptsne$treatment, colorscharmaip)
-
-p4 <- plotcolorfultsnes(pittsne, pittsne$sex, sexcolors)   + labs(y = "pituitary") 
-p5 <- plotcolorfultsnes(pittsne, pittsne$treatment, colorscharmaip)
-
-g4 <- plotcolorfultsnes(gontsne, gontsne$sex, sexcolors)  + labs(y = "gonads")  
-g5 <- plotcolorfultsnes(gontsne, gontsne$treatment, colorscharmaip)
-
-## pca
-
-h1 <- plotcolorfulpcs(hyppca, hyppca$sex, sexcolors) 
-h2 <- plotcolorfulpcs(hyppca, hyppca$treatment, colorscharmaip)
-
-p1 <- plotcolorfulpcs(pitpca, pitpca$sex, sexcolors) 
-p2 <- plotcolorfulpcs(pitpca, pitpca$treatment, colorscharmaip)
-
-g1 <- plotcolorfulpcs(gonpca, gonpca$sex, sexcolors) 
-g2 <- plotcolorfulpcs(gonpca, gonpca$treatment, colorscharmaip)
-
-plot_grid(h1,h2,p1,p2,g1,g2, nrow = 3)
-
-# screeplot for variance
-fviz_screeplot(hypv, addlabels = TRUE,   ncp = 5) + 
-  labs(title = NULL, y = "PC Variance") + theme_B3() 
-fviz_screeplot(pitfviz, addlabels = TRUE,   ncp = 5) + 
-  labs(title = NULL, y = "PC Variance") + theme_B3() 
-fviz_screeplot(gonfviz, addlabels = TRUE,    ncp = 5) + 
-  labs(title = NULL, y = "PC Variance") + theme_B3()
-
-# fviz contributions
-
-plotfriz <- function(frizdf){
-
-     p <- fviz_pca_var(frizdf,  labelsize = 3.5 , axes.linetype = "blank", 
-                   repel = T ,  select.var= list(contrib = 3))  + 
-      labs(title = NULL) + 
-      theme_B3() + 
-      theme(axis.text = element_blank())
-   print(p)
-}
-
-h3 <- plotfriz(hypv) + labs(x = "PC1 (59.8%)", y = "PC2 (37.7%)") 
-p3 <-plotfriz(pitfviz) + labs(x = "PC1 (54.3%)", y = "PC2 (35.7%)") 
-g3 <- plotfriz(gonfviz) + labs(x = "PC1 (82.7%)", y = "PC2 (11.6%)")
-
-## all together now
-
-plot_grid(h4,h5,h2,h3,
-          p4,p5,p2,p3,
-          g4,g5,g2,g3,
-          nrow = 3,
-          labels = "auto", label_size = 10)
-```

### 2020-02-09 R/functions.R

@@ -806,3 +806,85 @@ network_plot.cor_df <- function(rdf,
   
 }
 
+## friz pca with genes and PCs 
+
+makefvizdf <-  function(whichtissue, whichtreatment, whichsex){
+  colData <- colData %>%
+    dplyr::filter(tissue %in% whichtissue,
+                  treatment %in% whichtreatment,
+                  sex %in% whichsex) 
+  row.names(colData) <- colData$V1
+  
+  
+  # save counts that match colData
+  savecols <- as.character(colData$V1) 
+  savecols <- as.vector(savecols) 
+  
+  countData <- as.data.frame(t(countData))
+  countData <- countData %>% dplyr::select(one_of(savecols)) 
+  countData <- as.data.frame(t(countData))
+  
+  mypca <- prcomp(countData)
+  return(mypca)
+}
+
+plotfriz <- function(frizdf){
+  
+  p <- fviz_pca_var(frizdf,  labelsize = 3.5 , axes.linetype = "blank", 
+                    repel = T ,  select.var= list(contrib = 3))  + 
+    labs(title = NULL) + 
+    theme_B3() + 
+    theme(axis.text = element_blank())
+  print(p)
+}
+
+# pca 
+
+subsetmakepca <- function(whichtissue, whichtreatment, whichsex){
+  
+  colData <- colData %>%
+    dplyr::filter(tissue %in% whichtissue,
+                  treatment %in% whichtreatment,
+                  sex %in% whichsex) 
+  row.names(colData) <- colData$V1
+  
+  
+  # save counts that match colData
+  savecols <- as.character(colData$V1) 
+  savecols <- as.vector(savecols) 
+  
+  countData <- as.data.frame(t(countData))
+  countData <- countData %>% dplyr::select(one_of(savecols)) 
+  countData <- as.data.frame(t(countData))
+  
+  mypca <- prcomp(countData)
+  
+  mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
+                        PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
+                        ID = row.names(countData))
+  mypcadf$V1 <- row.names(mypcadf)
+  mypcadf <- left_join(colData, mypcadf)
+  mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
[…]

### 2020-02-09 R/themes.R

@@ -167,6 +167,8 @@ colorsvolcano <-  c(colorscharnew,
                     
                     "NS" = "#bdbdbd")
 
+allcolors <- c(colorscharmaip, sexcolors, colorstissue)
+
 colorsvolcanochar <-  c(colorscharnew,
                     "NS" = "#bdbdbd")
 

### 2020-02-10 analysis/02_DESeq2.Rmd
now with DEseq2 figs and data

---
title: "06_DESeq2"
output: md_document
---

```{r setup, include = FALSE}

library(tidyverse)
library(DESeq2)

library(BiocParallel)
register(MulticoreParam(6))

source("../R/functions.R")  # load custom functions 
source("../R/themes.R")  # load custom themes and color palletes
source("../R/icons.R")  # for images

knitr::opts_chunk$set(fig.path = '../figures/DESeq2/',  message=F, comment=F, warning=F)
```

# Warning: This script takes a very long time to run.

DESeq2 was not designed to run on 300 samples. But, I really like it, so I do it anyways. Some of these commands take like 15 min to run using 6 cores.

## DEseq2 on all chaacterization and manipulations

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

# craete variable that will be critical for subset later on
colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
colData$sextissue <- as.factor(paste(colData$sex, colData$tissue, sep = "_"))
colData$treatment <- factor(colData$treatment, levels = charlevels)
colData$tissue <- factor(colData$tissue, levels = tissuelevel)
levels(colData$treatment)
levels(colData$sex)
levels(colData$sextissue)
levels(colData$tissue)
```

```{r deseq2, eval=T}
for(i in levels(colData$sextissue)){

  newcolData <- subsetcolData2(colData, i)

  # save counts that match colData
  savecols <- as.character(newcolData$V1) 
  savecols <- as.vector(savecols)

  newcountData <- countData %>% dplyr::select(one_of(savecols))

  dds <- DESeqDataSetFromMatrix(countData = newcountData,
                                colData = newcolData,
                                design = ~ treatment )
  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
  print(dds)
  print(dim(dds))
  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis

  vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))

  myfilename = paste0("../results/DEseq2/", i, "_vsd.csv")
  write.csv(vsd, myfilename)

  #return(dds)
  #return(vsd)
  print(head(vsd))

  # save differential gene expression results
  control.bldg <- createDEGdfsave("bldg", "control", i)
  bldg.lay <- createDEGdfsave("lay", "bldg", i)
  lay.inc.d3 <- createDEGdfsave("inc.d3", "lay",  i) 
  inc.d3.inc.d9 <- createDEGdfsave("inc.d9", "inc.d3", i) 
  inc.d9.inc.d17 <- createDEGdfsave("inc.d17", "inc.d9", i)
  inc.d17.hatch <- createDEGdfsave( "hatch", "inc.d17", i) 
  hatch.n5 <- createDEGdfsave( "n5", "hatch",  i) 
  n5.n9 <- createDEGdfsave("n9", "n5",  i) 
}

```

### 2020-02-10 analysis/figure-2.Rmd
now with DEseq2 figs and data

---
title: "Fig2"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
```

## DEGs

```{r fig2, fig.width=7}

DEG_path <- "../results/DEseq2/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
DEG_files

allDEG <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
head(allDEG)

allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
allDEG$direction <- factor(allDEG$direction, levels = charlevels)

allDEG %>%
  group_by(sex, tissue, comparison) %>%
  summarize(n = n())

p1 <- allDEG %>%
  filter(comparison == "control_bldg") %>%
ggplot(aes(x = comparison,  fill = direction)) +
  geom_bar(position = "dodge") +
  facet_grid(tissue~sex) +
  theme_B3() +
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
         axis.text.x = element_blank(),
        legend.position = "none",
        strip.text.y = element_blank())  +
  guides(fill = guide_legend(nrow = 1)) +
  labs(x = " ", y = "Total Differentially Expressed Genes") +
  scale_fill_manual(values = colorscharnew,
                       name = " ",
                       drop = FALSE) +
  scale_color_manual(values = sexcolors)

p2 <- allDEG %>%
  filter(comparison != "control_bldg") %>%
ggplot(aes(x = comparison,  fill = direction)) +
  geom_bar(position = "dodge") +
  facet_grid(tissue~sex) +
  theme_B3() +
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
         axis.text.x = element_blank(),
        legend.position = "none",
        strip.placement = "right")  +
  guides(fill = guide_legend(nrow = 1)) +
  labs(x = "Sequential Parental Care Stages  ", y = NULL) +
  scale_fill_manual(values = colorscharnew,
                       name = " ", drop = FALSE) +
  scale_color_manual(values = sexcolors) +
  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
            position = position_dodge(width = 1),
            size = 2.5, color = "black")

forlegend <- allDEG %>%
ggplot(aes(x = comparison,  fill = direction)) +
  geom_bar(position = "dodge") +
   scale_fill_manual(values = colorscharnew,
                       name = "increased in", drop = FALSE) +
  scale_color_manual(values = sexcolors) +
  guides(color = F) +
  guides(fill= guide_legend(nrow =1)) +
  theme_B3() +
  theme(legend.position = "bottom")

mylegend <- get_legend(forlegend)

barplots <- plot_grid(p1,p2, nrow=1, rel_widths = c(0.9,3))

fig2 <- plot_grid(barplots, mylegend, nrow = 2, rel_heights = c(1,0.1))
fig2
```

### 2020-02-10 analysis/suppl-figure-1.Rmd
suppl fig

---
title: "Suppl.Fig 1"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)

library(factoextra)
source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
```

```{r read}
source("../R/wrangledata.R")
```

## PCA

```{r supplfig1}
subsetmakepca <- function(whichtissue, whichtreatment, whichsex){	
  colData <- colData %>%	
    dplyr::filter(tissue %in% whichtissue,	
                  treatment %in% whichtreatment,	
                  sex %in% whichsex) 	
  row.names(colData) <- colData$V1	
  # save counts that match colData	
  savecols <- as.character(colData$V1) 	
  savecols <- as.vector(savecols)

  countData <- as.data.frame(t(countData))	
  countData <- countData %>% dplyr::select(one_of(savecols)) 	
  countData <- as.data.frame(t(countData))

  mypca <- prcomp(countData)	
  mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 	
                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],	
                  ID = row.names(countData))	
  mypcadf$V1 <- row.names(mypcadf)	
  mypcadf <- left_join(colData, mypcadf)	
  mypcadf <- mypcadf %>% dplyr::select(bird,sex,tissue,treatment,PC1:PC6)	
  return(mypcadf)	
}

plotcolorfulpcs <- function(mypcadf,  whichfactor, whichcolors){	
  p <- mypcadf %>%	
    ggplot(aes(x = PC1, y = PC2, shape = tissue, color = whichfactor )) +	
    geom_point(size = 1)  +	
    theme_B3() +	
    theme(legend.title = element_blank(),	
         axis.text = element_blank(),	
         legend.position = "none") +	
    labs(x = "PC1", y = "PC2")  +	
    scale_color_manual(values = whichcolors) +	
    scale_shape_manual(values = myshapes)	+
    stat_ellipse( )

  return(p)	
}

makefvizdf <-  function(whichtissue, whichtreatment, whichsex){	
  colData <- colData %>%	
      dplyr::filter(tissue %in% whichtissue,	
                  treatment %in% whichtreatment,	
                  sex %in% whichsex) 	
  row.names(colData) <- colData$V1	
  # save counts that match colData	
  savecols <- as.character(colData$V1) 	
  savecols <- as.vector(savecols)

  countData <- as.data.frame(t(countData))	
  countData <- countData %>% dplyr::select(one_of(savecols)) 	
  countData <- as.data.frame(t(countData))	
  mypca <- prcomp(countData)	
  return(mypca)	
}

plotfriz <- function(frizdf){

  p <- fviz_pca_var(frizdf,  labelsize = 3.5 , axes.linetype = "blank", 
                    repel = T , select.var= list(contrib = 3), col.var = "black")  + 
    labs(title = NULL) + 
    theme_B3() + 
    theme(axis.text = element_blank())
  return(p)
}

fpca <- subsetmakepca(tissuelevels, charlevels, "female")	
mpca <- subsetmakepca(tissuelevels, charlevels, "male")	
ffviz <- makefvizdf(tissuelevels, charlevels, "female")	
mfviz<- makefvizdf(tissuelevels, charlevels, "male")

f1 <- plotcolorfulpcs(fpca,fpca$treatment, allcolors) + labs(subtitle = "females")
f2 <- plotfriz(ffviz) + labs(subtitle = " ")

m1 <- plotcolorfulpcs(mpca,mpca$treatment, allcolors) + labs(subtitle = "males")
m2 <- plotfriz(mfviz) + labs(subtitle = " ")

allpcaplots <- plot_grid(f1,m1,f2,m2, nrow = 2, rel_heights = c(0.6,0.4), 
                labels= c("a"," ", "b"), label_size = 12)

forlegend <- plotcolorfulpcs(mpca,mpca$treatment, allcolors) + 
  theme(legend.position = "left",
        legend.direction = "vertical",
        legend.key.size = unit(0.5, 'lines')) +
  guides(color = guide_legend(ncol = 1),
         shape = guide_legend(ncol = 1))
mylegend <- get_legend(forlegend)

supplefig1 <- plot_grid(allpcaplots, mylegend, nrow = 1, rel_widths = c(1,0.2))
supplefig1
```

### 2020-02-10 analysis/figure-1.Rmd
suppl fig

@@
 
-Note, the input file is too big for storage on github :(
-
-```{r read}
-pseudocounts <- read_csv("../results/01_pseudo.counts.csv")
-head(pseudocounts[1:3])
-pseudocounts <- as.data.frame(pseudocounts)
-row.names(pseudocounts) <- pseudocounts$X1
-pseudocounts$X1 <- NULL
-# prep count data for all samples
-countData <- as.data.frame(t(pseudocounts))
-head(countData[1:3])
-```
-
-```{r colData}
-# prep col data for all samples
-colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
-colData$treatment <- factor(colData$treatment, levels = alllevels)
-colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
-colData$tissue <- factor(colData$tissue, levels = tissuelevels)
-row.names(colData) <- colData$V1
-
-# check ready for analysis
-#row.names(countData) == row.names(colData)
-
-head(colData)
+```{r wrangle}
+source("../R/wrangledata.R")
 ```

### 2020-02-10 analysis/figure-3.Rmd
supple fig 3 PRL

---
title: "Fig3"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Plots with Prolactin

```{r}
library(tidyverse)
library(cowplot)
library(magick)
library(png)
library(grid)
library(ggimage)
library(apaTables)

source("../R/themes.R") 
source("../R/functions.R")
source("../R/icons.R")

knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
```

## Circulating levels of prolactin and prolactin expression

```{r data}
prolactin <- read_csv("../results/07_hormones.csv") %>%
    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
    droplevels() 
prolactin$treatment <- factor(prolactin$treatment, levels = alllevels)

vsd_path <- "../results/DEseq2/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

PRLvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts") %>%
  filter(gene == "PRL")  %>%
  dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
  dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
  dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
  dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
  dplyr::select(sex, tissue, treatment, gene, samples, counts)

PRLvsd$treatment <- factor(PRLvsd$treatment, levels = alllevels)

PRLhyp <- PRLvsd %>% filter(tissue == "hypothalamus")
PRLpit <- PRLvsd %>% filter(tissue == "pituitary")
PRLgon <- PRLvsd %>% filter(tissue == "gonad")
```

```{r fig3, fig.width=7}
plotprolactin <- function(df, myy, myylab, mysubtitle ){

  p <-  ggplot(df, aes(x = treatment, y = myy)) +
    geom_boxplot(aes(fill = treatment, color = sex)) +
    theme_B3() +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = sexcolors) +
    labs(y = "prolactin (ng/mL)", x = NULL) +
    theme(legend.position = c(0.85,0.15), legend.direction = "horizontal") + 
  labs(x = "parental stage", subtitle = mysubtitle, y= myylab) +
      guides(fill = guide_legend(nrow = 1)) 
  return(p)
}

a <- plotprolactin(PRLhyp, PRLhyp$counts, "PRL", "hypothalamus") + 
  theme(legend.position = "none", axis.text.x = element_blank(), axis.title.x = element_blank(),
        axis.title.y = element_text(face = "italic"))
b <- plotprolactin(PRLpit, PRLpit$counts, "PRL", "pituitary") + 
  theme(legend.position = "none", axis.text.x = element_blank(), axis.title.x = element_blank(),
        axis.title.y = element_text(face = "italic"))
c <- plotprolactin(prolactin, prolactin$plasma_conc, "prolactin (ng/mL)", "blood") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) 
d <- plotprolactin(PRLgon, PRLgon$counts, "PRL", "gonads") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(face = "italic"))

mylegend <- plotprolactin(PRLvsd, PRLvsd$counts, "PRL", "gonads") +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        legend.key.size = unit(0.5, 'lines')) +
  guides(color = guide_legend(ncol = 1),
         shape = guide_legend(ncol = 1))
mylegend <- get_legend(mylegend)

allPLRplots <- plot_grid(a,b,c,d, labels = "auto", rel_heights = c(0.425,0.575))

fig3 <- plot_grid(allPLRplots, mylegend , nrow = 2, rel_heights = c(1,0.2))
fig3
```

### 2020-02-10 analysis/figure-1.Rmd
attempt to combine all three figs

@@
 source("../R/themes.R")
+source("../R/functions.R")
+source("../R/wrangledata.R")
 
@@
 
-# Figure 1. Experimental design and tSNE analysis
-
-## wrange data
-
-```{r wrangle}
-source("../R/wrangledata.R")
-```
+# Figure 1. Experimental design, tSNE analysis, PCA, and prolactin
 
@@
 
-```{r fig1, fig.width=7}
+```{r }
 
-plottsneelipse <- function(tsnedf, whichfactor, whichcolors){
+plottsneelipse <- function(tsnedf, pointcolor, whichcolors){
   p <- ggplot(tsnedf, aes(x = V1, y = V2)) +
-    geom_point(size = 1, aes(color = whichfactor)) +
+    geom_point(size = 1, aes(color = pointcolor)) +
     theme_B3() +
@@
 b <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = " ", subtitle = "tissue * sex\n")    
-c <- plottsneelipse(ftsne, ftsne$treatment, allcolors ) + labs(y = " ", subtitle = "females\ntissue * treatment")
-d <- plottsneelipse(mtsne, mtsne$treatment, allcolors ) + labs(y = " ", subtitle = "males\ntissue * treatment")
+c <- plottsneelipse(ftsne, ftsne$treatment, allcolors ) + labs(y = " ", subtitle = "treatment * tissue\nfemales")
+d <- plottsneelipse(mtsne, mtsne$treatment, allcolors ) + labs(y = " ", subtitle = "treatment * tissue\nmales")
 
@@
 
-fig1 <- plot_grid(expdesign, abcd, nrow = 2, labels = c("a", "b"), label_size = 12, rel_heights = c(0.5,1))
-fig1
+abcde <- plot_grid(expdesign, abcd, nrow = 2, labels = c("a", "b"), label_size = 12, rel_heights = c(0.5,1))
+abcde
 ```
+
+## pca
+
+```{r supplfig1}
+charpca <- subsetmakepca(tissuelevels, charlevels, sexlevels)	
+charfviz <- makefvizdf(tissuelevels, charlevels, sexlevels)
+
+f <- plotcolorfulpcs(charpca,charpca$treatment, allcolors) + labs(subtitle = " ") +
+  theme(legend.position = c(0.6,0.2), 
+        #legend.direction = "horizontal", 
+        legend.key.size = unit(0.5, 'lines')) + 
+  guides(color = FALSE) +
+  labs(subtitle = "tissue * treatment")   
+g <- plotfriz(charfviz) + labs(subtitle = "  ")
+
+pcaplots <- plot_grid(f,g, labels = c("f", "g"), label_size = 12, align = "hv")
+pcaplots
+```
+
+## PRL
+
+```{r}
+PRLplot <- plotprolactin(PRLpit, PRLpit$counts, "PRL", "treatment * sex, pituitary only") + 
+  theme(legend.position = "none", 
+        axis.text.x = element_text(angle = 45, hjust = 1), 
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(face = "italic"))
+
+fgh <- plot_grid(pcaplots, PRLplot, labels = c(" ", "h"), label_size = 12)
+fgh
+```
+
+## all together now
+
+```{r fig1, fig.width=7, fig.height=6}
+plot_grid(abcde, fgh, nrow = 2, rel_heights = c(2,1.5))
+```

### 2020-02-10 analysis/figure-3.Rmd
attempt to combine all three figs

@@
 
-vsd_path <- "../results/DEseq2/"   # path to the data
-vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
-vsd_pathfiles <- paste0(vsd_path, vsd_files)
-vsd_files
-
-PRLvsd <- vsd_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name")  %>% 
-  dplyr::rename("gene" = "X1") %>% 
-  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
-               names_to = "samples", values_to = "counts") %>%
-  filter(gene == "PRL")  %>%
-  dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
-  dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
-  dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
-                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
-                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
-  dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
-  dplyr::select(sex, tissue, treatment, gene, samples, counts)
-
-PRLvsd$treatment <- factor(PRLvsd$treatment, levels = alllevels)
-
-PRLhyp <- PRLvsd %>% filter(tissue == "hypothalamus")
-PRLpit <- PRLvsd %>% filter(tissue == "pituitary")
-PRLgon <- PRLvsd %>% filter(tissue == "gonad")
 ```
@@
 ```{r fig3, fig.width=7}
-plotprolactin <- function(df, myy, myylab, mysubtitle ){
-
-  p <-  ggplot(df, aes(x = treatment, y = myy)) +
-    geom_boxplot(aes(fill = treatment, color = sex)) +
-    theme_B3() +
-    scale_fill_manual(values = allcolors) +
-    scale_color_manual(values = sexcolors) +
-    labs(y = "prolactin (ng/mL)", x = NULL) +
-    theme(legend.position = c(0.85,0.15), legend.direction = "horizontal") + 
-  labs(x = "parental stage", subtitle = mysubtitle, y= myylab) +
-      guides(fill = guide_legend(nrow = 1)) 
-  return(p)
-}
 

### 2020-02-10 R/functions.R
attempt to combine all three figs

@@ -615,10 +615,7 @@ selectcBRCA1vsds <- function(listofgenes, vsd, colData){
 
 
 
-
-
-
-### modified corrr fucniont
+### modified corrr function
 
 network_plot.cor_df <- function(rdf,
                                 min_cor = .30,
@@ -714,85 +711,84 @@ network_plot.cor_df <- function(rdf,
   
 }
 
-## friz pca with genes and PCs 
+## PCA and FVIZ plots
 
-makefvizdf <-  function(whichtissue, whichtreatment, whichsex){
-  colData <- colData %>%
-    dplyr::filter(tissue %in% whichtissue,
-                  treatment %in% whichtreatment,
-                  sex %in% whichsex) 
-  row.names(colData) <- colData$V1
+subsetmakepca <- function(whichtissue, whichtreatment, whichsex){	
+  colData <- colData %>%	
+    dplyr::filter(tissue %in% whichtissue,	
+                  treatment %in% whichtreatment,	
+                  sex %in% whichsex) 	
+  row.names(colData) <- colData$V1	
+  # save counts that match colData	
+  savecols <- as.character(colData$V1) 	
+  savecols <- as.vector(savecols) 	
   
+  countData <- as.data.frame(t(countData))	
+  countData <- countData %>% dplyr::select(one_of(savecols)) 	
+  countData <- as.data.frame(t(countData))	
   
-  # save counts that match colData
-  savecols <- as.character(colData$V1) 
-  savecols <- as.vector(savecols) 
+  mypca <- prcomp(countData)	
+  mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 	
+                        PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],	
+                        ID = row.names(countData))	
+  mypcadf$V1 <- row.names(mypcadf)	
+  mypcadf <- left_join(colData, mypcadf)	
+  mypcadf <- mypcadf %>% dplyr::select(bird,sex,tissue,treatment,PC1:PC6)	
+  return(mypcadf)	
+}	
+
+
+plotcolorfulpcs <- function(mypcadf,  whichfactor, whichcolors){	
+  p <- mypcadf %>%	
+    ggplot(aes(x = PC1, y = PC2, shape = tissue, color = whichfactor )) +	
+    geom_point(size = 1)  +	
+    theme_B3() +	
[…]

### 2020-02-10 analysis/figure-1.Rmd
better fig 1 and 3

@@
 library(tidyverse)
-library(ggfortify)
-library(factoextra)
 library(cowplot)
 library(Rtsne)
-library(viridis)
-library(forcats)
 
@@
 ```{r tSNE-df}
+
 # prep for tsne
@@
 
-```{r }
+```{r fig1, fig.width=7, fig.height=6}
 
@@
           axis.text = element_blank()) +
-    stat_ellipse(linetype = 2, aes(color = tissue )) 
+    stat_ellipse(linetype = 1, aes(color = tissue )) 
   return(p)
@@
 abcde <- plot_grid(expdesign, abcd, nrow = 2, labels = c("a", "b"), label_size = 12, rel_heights = c(0.5,1))
-abcde
+
+plottsneelipsev2 <- function(tsnedf, pointcolor, whichcolors){
+  p <- ggplot(tsnedf, aes(x = V1, y = V2)) +
+    geom_point(size = 1, aes(color = pointcolor)) +
+    theme_B3() +
+    labs(x = "tSNE 1", y = "tSNE 2") +
+    scale_color_manual(values = whichcolors) +
+    theme(legend.position = "none",
+          axis.text = element_blank()) +
+    stat_ellipse(linetype = 1, aes(color = treatment)) 
+  return(p)
+}
+
+h <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus")  + facet_wrap(~sex)
+i <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary", y = NULL) + facet_wrap(~sex) 
+j <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads", y = NULL)  + facet_wrap(~sex)
+
+hij <- plot_grid(h,i,j, nrow = 1, labels = c("h", "i", "j"), label_size = 12)
+
+fig1 <- plot_grid(abcde, hij, nrow = 2, rel_heights = c(0.6,0.4))
+fig1
 ```
-
-## pca
-
-```{r supplfig1}
-charpca <- subsetmakepca(tissuelevels, charlevels, sexlevels)	
-charfviz <- makefvizdf(tissuelevels, charlevels, sexlevels)
-
-f <- plotcolorfulpcs(charpca,charpca$treatment, allcolors) + labs(subtitle = " ") +
-  theme(legend.position = c(0.6,0.2), 
-        #legend.direction = "horizontal", 
-        legend.key.size = unit(0.5, 'lines')) + 
-  guides(color = FALSE) +
-  labs(subtitle = "tissue * treatment")   
-g <- plotfriz(charfviz) + labs(subtitle = "  ")
-
-pcaplots <- plot_grid(f,g, labels = c("f", "g"), label_size = 12, align = "hv")
-pcaplots
-```
-
-## PRL
-
-```{r}
-PRLplot <- plotprolactin(PRLpit, PRLpit$counts, "PRL", "treatment * sex, pituitary only") + 
-  theme(legend.position = "none", 
-        axis.text.x = element_text(angle = 45, hjust = 1), 
-        axis.title.x = element_blank(),
-        axis.title.y = element_text(face = "italic"))
-
-fgh <- plot_grid(pcaplots, PRLplot, labels = c(" ", "h"), label_size = 12)
-fgh
-```
-
-## all together now
-
-```{r fig1, fig.width=7, fig.height=6}
-plot_grid(abcde, fgh, nrow = 2, rel_heights = c(2,1.5))
-```

### 2020-02-10 analysis/figure-3.Rmd
better fig 1 and 3

@@
 
-# Plots with Prolactin
+# Figure 3
 
-```{r}
+```{r setup}
 library(tidyverse)
 library(cowplot)
-library(magick)
-library(png)
-library(grid)
 library(ggimage)
 library(apaTables)
+library(factoextra)
 
@@
 source("../R/icons.R")
+source("../R/wrangledata.R")
 
@@
 
-## Circulating levels of prolactin and prolactin expression
+## Data: PCA, hormones, PRL expression
 
 ```{r data}
+# pca
+charpca <- subsetmakepca(tissuelevels, charlevels, sexlevels)	
+charfviz <- makefvizdf(tissuelevels, charlevels, sexlevels)
+
+# hormones
 prolactin <- read_csv("../results/07_hormones.csv") %>%
@@
 
+# PRL vsd
+
+vsd_path <- "../results/DEseq2/"   # path to the data
+vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
+vsd_pathfiles <- paste0(vsd_path, vsd_files)
+vsd_files
+
+PRLvsd <- vsd_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name")  %>% 
+  dplyr::rename("gene" = "X1") %>% 
+  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "samples", values_to = "counts") %>%
+  filter(gene == "PRL")  %>%
+  dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
+  dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
+  dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
+  dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
+  dplyr::select(sex, tissue, treatment, gene, samples, counts)
+PRLvsd$treatment <- factor(PRLvsd$treatment, levels = alllevels)
+
+PRLhyp <- PRLvsd %>% filter(tissue == "hypothalamus")
+PRLpit <- PRLvsd %>% filter(tissue == "pituitary")
+PRLgon <- PRLvsd %>% filter(tissue == "gonad")
 ```
 
-```{r fig3, fig.width=7}
+```{r fig3, fig.width=7, fig.height=7}
 
-a <- plotprolactin(PRLhyp, PRLhyp$counts, "PRL", "hypothalamus") + 
+a <- plotcolorfulpcs(charpca,charpca$treatment, allcolors) + labs(subtitle = " ") +
+  theme(legend.position = c(0.6,0.1), 
+        legend.direction = "horizontal", 
+        legend.key.size = unit(0.5, 'lines')) + 
+  guides(color = FALSE) +
+  labs(subtitle = " ")
+
+b <- plotprolactin(PRLhyp, PRLhyp$counts, "PRL", "hypothalamus") + 
+  theme(legend.position = c(0.7,0.9), axis.text.x = element_blank(), 
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(face = "italic"),
+        legend.title = element_blank()) + guides(fill = F)
+
+c <- plotfriz(charfviz) + labs(subtitle = "  ")
+
+d <- plotprolactin(PRLpit, PRLpit$counts, "PRL", "pituitary") + 
   theme(legend.position = "none", axis.text.x = element_blank(), axis.title.x = element_blank(),
         axis.title.y = element_text(face = "italic"))
-b <- plotprolactin(PRLpit, PRLpit$counts, "PRL", "pituitary") + 
-  theme(legend.position = "none", axis.text.x = element_blank(), axis.title.x = element_blank(),
-        axis.title.y = element_text(face = "italic"))
-c <- plotprolactin(prolactin, prolactin$plasma_conc, "prolactin (ng/mL)", "blood") +
+
+e <- plotprolactin(prolactin, prolactin$plasma_conc, "prolactin (ng/mL)", "blood") +
   theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
-d <- plotprolactin(PRLgon, PRLgon$counts, "PRL", "gonads") + 
+
+f <- plotprolactin(PRLgon, PRLgon$counts, "PRL", "gonads") + 
   theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1),
@@
 
-mylegend <- plotprolactin(PRLvsd, PRLvsd$counts, "PRL", "gonads") +
-  theme(legend.position = "bottom",
-        legend.direction = "vertical",
-        legend.key.size = unit(0.5, 'lines')) +
-  guides(color = guide_legend(ncol = 1),
-         shape = guide_legend(ncol = 1))
-mylegend <- get_legend(mylegend)
+plot_grid(a,b,c,d,e,f, labels = "auto", ncol = 2, rel_heights = c(1,1,1.2))
 
-allPLRplots <- plot_grid(a,b,c,d, labels = "auto", rel_heights = c(0.425,0.575))
-
-fig3 <- plot_grid(allPLRplots, mylegend , nrow = 2, rel_heights = c(1,0.2))
-fig3
 ```

### 2020-02-10 analysis/figure-4.Rmd
candidate gene 4.

---
title: "fig4"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)

source("../R/themes.R")

knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
```

```{r fig4, fig.height=7, fig.width=7}
geneids <- read_csv("../metadata/00_geneinfo.csv")

vsd_path <- "../results/DEseq2/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts")

hypgenes <- read_csv("../results/DEseq2/female_hypothalamus_hatch_n5_DEGs.csv") %>%
  arrange(desc(logpadj)) %>% head(., n = 2) %>% pull(gene)
hypgenes

pitgenes <- read_csv("../results/DEseq2/female_pituitary_inc.d9_inc.d17_DEGs.csv") %>%
  arrange(desc(logpadj)) %>% head(., n = 2) %>% pull(gene)
pitgenes

gongenes <- read_csv("../results/DEseq2/female_gonad_lay_inc.d3_DEGs.csv") %>%
  arrange(desc(logpadj)) %>% head(., n = 2) %>% pull(gene)
gongenes

plottopgenes <- function(whichgenes, whichtissue, whichsex, mysubtitle){
  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
    filter(tissue == whichtissue)

  candidates$treatment <- factor(candidates$treatment, levels = alllevels)

  p <- ggplot(candidates,  aes(x = treatment, y = counts, fill = treatment, color = sex)) +
    geom_boxplot() + facet_wrap(~gene, scales = "free_y") + 
    theme_B3() +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
          legend.position = "none",
          axis.title.x = element_blank(),
          strip.text = element_text(face = "italic")) +
    labs(y = mysubtitle) 
  return(p)

}

a <- plottopgenes(hypgenes, "hypothalamus", "female", "hypothalamus") + theme(axis.text.x = element_blank())
b <- plottopgenes(pitgenes, "pituitary", "female", "pituitary") + theme(axis.text.x = element_blank())
c <- plottopgenes(gongenes, "gonad", "female", "gonads")

plot_grid(a,b,c, ncol = 1)

```

### 2020-02-10 R/wrangledata.R

@@ -1,7 +1,7 @@
 # data wrangle for multiple figures
 
-# Note, the input file is too big for storage on github :(
-
+# this is for tsne and pca
+# Note, pseudocounts is too big for storage on github :(
 pseudocounts <- read_csv("../results/01_pseudo.counts.csv")
 head(pseudocounts[1:3])
 pseudocounts <- as.data.frame(pseudocounts)
@@ -11,14 +11,8 @@ pseudocounts$X1 <- NULL
 countData <- as.data.frame(t(pseudocounts))
 head(countData[1:3])
 
+# this is for tsne and pca
 # prep col data for all samples
-colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
-colData$treatment <- factor(colData$treatment, levels = alllevels)
-colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
-colData$tissue <- factor(colData$tissue, levels = tissuelevels)
-row.names(colData) <- colData$V1
-
-# sample data
 colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
 colData$treatment <- factor(colData$treatment, levels = alllevels)
 colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
@@ -27,4 +21,3 @@ row.names(colData) <- colData$V1
 
 # check ready for analysis
 #row.names(countData) == row.names(colData)
-

### 2020-02-11 analysis/figure-4.Rmd
now with top 3 genes

@@
 hypgenes <- read_csv("../results/DEseq2/female_hypothalamus_hatch_n5_DEGs.csv") %>%
-  arrange(desc(logpadj)) %>% head(., n = 2) %>% pull(gene)
+  arrange(desc(logpadj)) %>% head(., n = 3) %>% pull(gene)
 hypgenes
@@
 pitgenes <- read_csv("../results/DEseq2/female_pituitary_inc.d9_inc.d17_DEGs.csv") %>%
-  arrange(desc(logpadj)) %>% head(., n = 2) %>% pull(gene)
+  arrange(desc(logpadj)) %>% head(., n = 3) %>% pull(gene)
 pitgenes
@@
 gongenes <- read_csv("../results/DEseq2/female_gonad_lay_inc.d3_DEGs.csv") %>%
-  arrange(desc(logpadj)) %>% head(., n = 2) %>% pull(gene)
+  arrange(desc(logpadj)) %>% head(., n = 3) %>% pull(gene)
 gongenes

### 2020-02-11 analysis/figure-4.Rmd
now with focused and zoomed out view of fig 4

--- 2020-02-10 analysis/figure-4.Rmd
@@
 
-plottopgenes <- function(whichgenes, whichtissue, whichsex, mysubtitle){
+plottopgenes <- function(whichgenes, whichtissue, whichsex, mysubtitle, whichtstages){
   candidates  <- allvsd %>%
@@
     dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
-    filter(tissue == whichtissue)
+    filter(tissue == whichtissue, sex %in% whichsex)
 
@@
 
-  p <- ggplot(candidates,  aes(x = treatment, y = counts, fill = treatment, color = sex)) +
+  p <- candidates %>%
+    filter(treatment %in% whichtstages) %>%
+    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
     geom_boxplot() + facet_wrap(~gene, scales = "free_y") + 
@@
     scale_color_manual(values = allcolors) +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
+    theme(axis.text.x = element_blank(),
           legend.position = "none",
@@
 
-a <- plottopgenes(hypgenes, "hypothalamus", "female", "hypothalamus") + theme(axis.text.x = element_blank())
-b <- plottopgenes(pitgenes, "pituitary", "female", "pituitary") + theme(axis.text.x = element_blank())
-c <- plottopgenes(gongenes, "gonad", "female", "gonads")
+a <- plottopgenes(hypgenes, "hypothalamus", c("female"), "hypothalamus", c("hatch", "n5")) 
+b <- plottopgenes(pitgenes, "pituitary", c("female"), "pituitary", c("inc.d9", "inc.d17")) 
+c <- plottopgenes(gongenes, "gonad", c("female"), "gonads", c("lay", "inc.d3"))
 
-plot_grid(a,b,c, ncol = 1)
+fig4a <- plot_grid(a,b,c, ncol = 1)
 
+d <- plottopgenes(hypgenes, "hypothalamus", c("female", "male"), NULL, charlevels) 
+e <- plottopgenes(pitgenes, "pituitary", c("female", "male"), NULL, charlevels) 
+f <- plottopgenes(gongenes, "gonad", c("female", "male"), NULL, charlevels) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+fig4b <- plot_grid(d,e,f, ncol = 1)
+
+fig4 <- plot_grid(fig4a, fig4b, rel_widths = c(1,2))
+fig4
 ```

### 2020-02-11 analysis/figure-5.Rmd
adding manipulation deseq data

---
title: "Fig5"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
```

## DEGs

```{r fig2, fig.width=7}

DEG_path <- "../results/DEseq2/manip/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
DEG_files

allDEG <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/manip/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
head(allDEG)

comparisonlevels <- c( "inc.d3_m.inc.d3" ,  "inc.d9_m.inc.d9", 
                        "inc.d17_m.inc.d17", "m.n2_hatch" ,
                        "inc.d9_m.inc.d8", "inc.d17_prolong", "hatch_extend"  )

allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
allDEG$direction <- factor(allDEG$direction, levels = manipulationsandcontrols)

allDEG %>%
  group_by(sex, tissue, comparison) %>%
  summarize(n = n())

allDEG

allDEG %>%
ggplot(aes(x = comparison,  fill = direction)) +
  geom_bar(position = "dodge") +
  facet_grid(tissue~sex) +
  theme_B3() +
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
         axis.text.x = element_blank(),
        legend.position = "none",
        strip.placement = "right")  +
  guides(fill = guide_legend(nrow = 1)) +
  labs(x = "Parental stage-specific manipulations", y = NULL) +
  scale_fill_manual(values = allcolors,
                       name = " ", drop = FALSE) +
  scale_color_manual(values = sexcolors) +
  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
            position = position_dodge(width = 1),
            size = 2.5, color = "black")

```

### 2020-02-11 analysis/02_DESeq2.Rmd
adding manipulation deseq data

@@
 ---
-title: "06_DESeq2"
+title: "02_DESeq2"
 output: md_document
@@
 source("../R/themes.R")  # load custom themes and color palletes
-source("../R/icons.R")  # for images
 
@@
 
-# Warning: This script takes a very long time to run.
+## Warning: This script can take a very long time to run.
 
-DESeq2 was not designed to run on 300 samples. But, I really like it, so I do it anyways. Some of these commands take like 15 min to run using 6 cores.
+DESeq2 was not designed to run on 100+ samples. But, I really like it, so I do it anyways. Some of these commands take like 15 min to run using 6 cores.
 
-## DEseq2 on all chaacterization and manipulations
+## Characterization
 
@@
 ```
+
+## Manipulation
+
+```{r manipulation}
+# import "colData" which contains sample information and "countData" which contains read counts
+countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
+geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+head(geneinfo)
+
+# craete variable that will be critical for subset later on
+colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
+colData$sextissue <- as.factor(paste(colData$sex, colData$tissue, sep = "_"))
+colData$treatment <- factor(colData$treatment, levels = manipulationsandcontrols)
+colData$tissue <- factor(colData$tissue, levels = tissuelevel)
+levels(colData$treatment)
+levels(colData$sex)
+levels(colData$sextissue)
+levels(colData$tissue)
+colData <- colData %>% drop_na()
+
+colData <- colData %>% filter(sex == "male")
+
+createDEGdfsaveManip <- function(up, down, mytissue){
+
+  res <- results(dds, contrast = c("treatment", up, down), independentFiltering = T, alpha = 0.1)
+
+  DEGs <- data.frame(gene = row.names(res),
+                        padj = res$padj, 
+                        logpadj = -log10(res$padj),
+                        lfc = res$log2FoldChange,
+                        sextissue = mytissue)
+  DEGs <- na.omit(DEGs)
+  DEGs <- DEGs %>%
+    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
+                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
+                                                           yes = down, no = "NS"))) %>% 
+    dplyr::arrange(desc(lfc))
+
+  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up))
+
+  # write DEGsframe of only significant genes
+  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
+  print(str(DEGs))
+
+  partialfilename = paste("_", down, "_", up, sep = "")
+  myfilename = paste0("../results/DESeq2/manip/", mytissue, partialfilename, "_DEGs.csv")
+
+  write.csv(DEGs, myfilename, row.names = F)
+  # return DEGs frome with all data, included NS genes
+  #print(head(DEGs))
+}
+
+for(i in levels(colData$sextissue)){
+
+  newcolData <- subsetcolData2(colData, i)
+
+  # save counts that match colData
+  savecols <- as.character(newcolData$V1) 
+  savecols <- as.vector(savecols)
+
+  newcountData <- countData %>% dplyr::select(one_of(savecols))
+
+  dds <- DESeqDataSetFromMatrix(countData = newcountData,
+                                colData = newcolData,
+                                design = ~ treatment )
+  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
+  print(dds)
+  print(dim(dds))
+  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
+
+  vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
+
+  myfilename = paste0("../results/DEseq2/manip/", i, "_vsd.csv")
+  write.csv(vsd, myfilename)
+
+  #return(dds)
+  #return(vsd)
+  #print(head(vsd))
+
+  # save differential gene expression results
+  inc.d3.m.inc.d3 <- createDEGdfsaveManip("m.inc.d3", "inc.d3",  i) 
+  inc.d9.m.inc.d9 <- createDEGdfsaveManip("m.inc.d9", "inc.d9",  i) 
+  inc.d17.m.inc.d17 <- createDEGdfsaveManip("m.inc.d17", "inc.d17",  i) 
+  hatch.m.n2 <- createDEGdfsaveManip("hatch", "m.n2",  i) 
+  inc.d9.m.inc.d8 <- createDEGdfsaveManip("m.inc.d8", "inc.d9",  i) 
+  inc.d17.prolong <- createDEGdfsaveManip("prolong", "inc.d17",  i) 
+  hatch.extend <- createDEGdfsaveManip("extend", "hatch",  i) 
+}
+
+```

### 2020-02-11 analysis/02_DESeq2.Rmd
now with manip figure 5

@@
 
-```{r readfiles}
+```{r readfiles, eval=F, include=F}
 # import "colData" which contains sample information and "countData" which contains read counts
@@
 
-```{r deseq2, eval=T}
+```{r deseq2, eval=F, include=F}
 for(i in levels(colData$sextissue)){
@@
 
+levelsreplace <- c( "m.inc.d8" , "prolong" , "extend")
+levelsremoval <- c( "m.inc.d3" ,    "m.inc.d9" , "m.inc.d17" , "m.n2")
+controlsremovalreplace <- c( "inc.d3" ,    "inc.d9" , "inc.d17" , "hatch")
+
+manipulationsandcontrols <- c(controlsremoval, levelsreplace, levelsremoval)
+
 # craete variable that will be critical for subset later on
@@
 colData <- colData %>% drop_na()
-
-colData <- colData %>% filter(sex == "male")
 

### 2020-02-11 analysis/figure-5.Rmd
now with manip figure 5

@@
 
-```{r fig2, fig.width=7}
+```{r fig5, fig.width=7, fig.height=7}
 
@@
 
-comparisonlevels <- c( "inc.d3_m.inc.d3" ,  "inc.d9_m.inc.d9", 
-                        "inc.d17_m.inc.d17", "m.n2_hatch" ,
-                        "inc.d9_m.inc.d8", "inc.d17_prolong", "hatch_extend"  )
+comparisonlevels <- c( "inc.d9_m.inc.d8", "inc.d17_prolong", "hatch_extend" , 
+                       "inc.d3_m.inc.d3" ,  "inc.d9_m.inc.d9", 
+                        "inc.d17_m.inc.d17", "m.n2_hatch" 
+                        )
+
+levelsreplace <- c( "m.inc.d8" , "prolong" , "extend")
+levelsremoval <- c( "m.inc.d3" ,    "m.inc.d9" , "m.inc.d17" , "m.n2")
+controlsremovalreplace <- c( "inc.d3" ,    "inc.d9" , "inc.d17" , "hatch")
+
+manipulationsandcontrols <- c(controlsremoval, levelsreplace, levelsremoval)
 
@@
 
-allDEG
+expdesign <- png::readPNG("../figures/images/fig5a_fig5a.png")
+expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
 
-allDEG %>%
+replacementplots <- allDEG %>%
+  filter(comparison %in% c("inc.d9_m.inc.d8", "inc.d17_prolong", "hatch_extend" ))  %>%
 ggplot(aes(x = comparison,  fill = direction)) +
@@
         axis.text.x = element_blank(),
+        legend.position = "bottom",
+        strip.text.y =  element_blank())  +
+  guides(fill = guide_legend(nrow = 1)) +
+  labs(x = "Offspring replacement", y = "Total differentially expressed genes") +
+  scale_fill_manual(values = allcolors,
+                       name = "increased in", drop = FALSE) +
+  scale_color_manual(values = sexcolors) +
+  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
+            position = position_dodge(width = 1),
+            size = 2.5, color = "black") +
+  guides(fill =  guide_legend(nrow =2)) +
+  ylim(0,3500)
+
+mylegend <- get_legend(replacementplots)
+
+removalplots <- allDEG %>%
+  filter(comparison %in% c("inc.d3_m.inc.d3" ,  "inc.d9_m.inc.d9", 
+                        "inc.d17_m.inc.d17", "m.n2_hatch" ))  %>%
+ggplot(aes(x = comparison,  fill = direction)) +
+  geom_bar(position = "dodge") +
+  facet_grid(tissue~sex) +
+  theme_B3() +
+  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.text.x = element_blank(),
+         axis.text.y = element_blank(),
         legend.position = "none",
@@
   guides(fill = guide_legend(nrow = 1)) +
-  labs(x = "Parental stage-specific manipulations", y = NULL) +
+  labs(x = "Offspring removal", y = NULL) +
   scale_fill_manual(values = allcolors,
-                       name = " ", drop = FALSE) +
+                       name = "increased in", drop = FALSE) +
   scale_color_manual(values = sexcolors) +
@@
             position = position_dodge(width = 1),
-            size = 2.5, color = "black")
+            size = 2.5, color = "black") +
+  ylim(0,3500)
 
+barplots <- plot_grid( replacementplots + theme(legend.position = "none"),  removalplots,  ncol = 2, rel_widths = c(1,1.2))
+
+fig5 <- plot_grid(expdesign, barplots, mylegend, ncol =1, rel_heights = c(0.25,1,0.1))
+fig5
 ```

### 2020-02-12 analysis/figure-6.Rmd
second manipulatino figure

---
title: "fig6"
output: md_document
---

# Figure 4

```{r setup, include=TRUE}
library(tidyverse)
library(cowplot)

source("../R/themes.R")

knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
```

```{r fig6, fig.height=7, fig.width=7}
geneids <- read_csv("../metadata/00_geneinfo.csv")

vsd_path <- "../results/DEseq2/manip/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1")  %>% 
  pivot_longer(cols = blk.s031.pu.d_female_gonad_prolong:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts")

hypreplace <- read_csv("../results/DEseq2/manip/female_hypothalamus_hatch_extend_DEGs.csv") %>%
  arrange(desc(logpadj)) %>% head(., n = 1) %>% pull(gene)
hypreplace

hypremove <- read_csv("../results/DEseq2/manip/female_hypothalamus_m.n2_hatch_DEGs.csv") %>%
  arrange(desc(logpadj)) %>% head(., n = 1) %>% pull(gene)
hypremove

pitreplace <- read_csv("../results/DEseq2/manip/female_pituitary_hatch_extend_DEGs.csv") %>%
  arrange(desc(logpadj)) %>% head(., n = 1) %>% pull(gene)
pitreplace

pitremove <- read_csv("../results/DEseq2/manip/female_pituitary_m.n2_hatch_DEGs.csv") %>%
  arrange(desc(logpadj)) %>% head(., n = 1) %>% pull(gene)
pitremove

gonreplace <- read_csv("../results/DEseq2/manip/female_gonad_inc.d17_prolong_DEGs.csv") %>%
  arrange(desc(logpadj)) %>% head(., n = 1) %>% pull(gene)
gonreplace

gonremove <- read_csv("../results/DEseq2/manip/female_gonad_inc.d17_m.inc.d17_DEGs.csv") %>%
  arrange(desc(logpadj)) %>% head(., n = 1) %>% pull(gene)
gonremove

plottopgenes <- function(whichgenes, whichtissue, whichsex, mysubtitle, whichtstages){
  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/manip/'), "[", 2)) %>%
    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
    filter(tissue == whichtissue, sex %in% whichsex)

  candidates$treatment <- factor(candidates$treatment, levels = alllevels2)

  p <- candidates %>%
    filter(treatment %in% whichtstages) %>%
    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
    geom_boxplot() + facet_wrap(~gene, scales = "free_y") + 
    theme_B3() +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme(axis.text.x = element_blank(),
          legend.position = "none",
          axis.title.x = element_blank(),
          strip.text = element_text(face = "italic")) +
    labs(y = mysubtitle) 
  return(p)

}

a <- plottopgenes(hypreplace, "hypothalamus", c("female"), "hypothalamus", c("hatch", "extend")) 
b <- plottopgenes(pitreplace, "pituitary", c("female"), "pituitary", c("hatch", "extend")) 
c <- plottopgenes(gonreplace, "gonad", c("female"), "gonads", c("inc.d17", "prolong"))

fig6a <- plot_grid(a,b,c, ncol = 1)

d <- plottopgenes(hypreplace, "hypothalamus", c("female", "male"), NULL, alllevels2) 
e <- plottopgenes(pitreplace, "pituitary", c("female", "male"), NULL, alllevels2) 
f <- plottopgenes(gonreplace, "gonad", c("female", "male"), NULL, alllevels2) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

fig6b <- plot_grid(d,e,f, ncol = 1)

fig6 <- plot_grid(fig6a, fig6b, rel_widths = c(1,2))
fig6
```

### 2020-02-17 analysis/02_DESeq2.Rmd
now with manipuation degs and modules of genes for music

@@
 
-```{r manipulation}
+### Data wrangle
+
+```{r manipulation1, include = F}
 # import "colData" which contains sample information and "countData" which contains read counts
@@
 
-# craete variable that will be critical for subset later on
+colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
+colData$sextissue <- as.factor(paste(colData$sex, colData$tissue, sep = "_"))
+colData$treatment <- factor(colData$treatment, levels = manipulationsandcontrols)
+colData$tissue <- factor(colData$tissue, levels = tissuelevel)
+levels(colData$treatment)
+levels(colData$sex)
+levels(colData$sextissue)
+levels(colData$tissue)
+colData <- colData %>% drop_na()
+```
+### Manipuation: comparisons to respective controls
+
+```{r manipulationcontrol, eval = F}
+
+createDEGdfsaveManip <- function(up, down, mytissue){
+
+  res <- results(dds, contrast = c("treatment", up, down), independentFiltering = T, alpha = 0.1)
+
+  DEGs <- data.frame(gene = row.names(res),
+                        padj = res$padj, 
+                        logpadj = -log10(res$padj),
+                        lfc = res$log2FoldChange,
+                        sextissue = mytissue)
+  DEGs <- na.omit(DEGs)
+  DEGs <- DEGs %>%
+    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
+                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
+                                                           yes = down, no = "NS"))) %>% 
+    dplyr::arrange(desc(lfc))
+
+  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up))
+
+  # write DEGsframe of only significant genes
+  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
+  print(str(DEGs))
+
+  partialfilename = paste("_", down, "_", up, sep = "")
+  myfilename = paste0("../results/DESeq2/manip/", mytissue, partialfilename, "_DEGs.csv")
+
+  write.csv(DEGs, myfilename, row.names = F)
+  # return DEGs frome with all data, included NS genes
+  #print(head(DEGs))
+}
+
+for(i in levels(colData$sextissue)){
+
+  newcolData <- subsetcolData2(colData, i)
+
+  # save counts that match colData
+  savecols <- as.character(newcolData$V1) 
+  savecols <- as.vector(savecols)
+
+  newcountData <- countData %>% dplyr::select(one_of(savecols))
+
+  dds <- DESeqDataSetFromMatrix(countData = newcountData,
+                                colData = newcolData,
+                                design = ~ treatment )
+  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
+  print(dds)
+  print(dim(dds))
+  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
+
+  vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
+
+  myfilename = paste0("../results/DEseq2/manip/", i, "_vsd.csv")
+  write.csv(vsd, myfilename)
+
+  #return(dds)
+  #return(vsd)
+  #print(head(vsd))
+
+  # save differential gene expression results
+  inc.d3.m.inc.d3 <- createDEGdfsaveManip("m.inc.d3", "inc.d3",  i) 
+  inc.d9.m.inc.d9 <- createDEGdfsaveManip("m.inc.d9", "inc.d9",  i) 
+  inc.d17.m.inc.d17 <- createDEGdfsaveManip("m.inc.d17", "inc.d17",  i) 
+  hatch.m.n2 <- createDEGdfsaveManip("m.n2", "hatch",  i) 
+  inc.d9.m.inc.d8 <- createDEGdfsaveManip("m.inc.d8", "inc.d9",  i) 
+  inc.d17.prolong <- createDEGdfsaveManip("prolong", "inc.d17",  i) 
+  hatch.extend <- createDEGdfsaveManip("extend", "hatch",  i) 
+}
+
+```
+
+## Manipulation: comparisons to other manipluations
+
+```{r manipulationwithin}
+# import "colData" which contains sample information and "countData" which contains read counts
+countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
+geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+head(geneinfo)
+
+levelsreplace <- c( "m.inc.d8" , "prolong" , "extend")
+levelsremoval <- c( "m.inc.d3" ,    "m.inc.d9" , "m.inc.d17" , "m.n2")
+controlsremovalreplace <- c( "inc.d3" ,    "inc.d9" , "inc.d17" , "hatch")
+
+manipulationsandcontrols <- c(controlsremoval, levelsreplace, levelsremoval)
+
 colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
@@
   # save differential gene expression results
-  inc.d3.m.inc.d3 <- createDEGdfsaveManip("m.inc.d3", "inc.d3",  i) 
-  inc.d9.m.inc.d9 <- createDEGdfsaveManip("m.inc.d9", "inc.d9",  i) 
-  inc.d17.m.inc.d17 <- createDEGdfsaveManip("m.inc.d17", "inc.d17",  i) 
-  hatch.m.n2 <- createDEGdfsaveManip("hatch", "m.n2",  i) 
-  inc.d9.m.inc.d8 <- createDEGdfsaveManip("m.inc.d8", "inc.d9",  i) 
-  inc.d17.prolong <- createDEGdfsaveManip("prolong", "inc.d17",  i) 
-  hatch.extend <- createDEGdfsaveManip("extend", "hatch",  i) 
+  hatch.m.n2 <- createDEGdfsaveManip("m.n2", "hatch",  i) 
+  prolong.extend <- createDEGdfsaveManip("extend", "prolong",  i) 
+  early.extend <- createDEGdfsaveManip("extend", "m.inc.d8",  i) 
+  early.prolong <- createDEGdfsaveManip("prolong", "m.inc.d8",  i) 
+  m.inc.d9.m.inc.d3 <- createDEGdfsaveManip("m.inc.d9", "m.inc.d3",  i) 
+  m.inc.d17.m.inc.d9 <- createDEGdfsaveManip("m.inc.d17", "m.inc.d9",  i) 
+  m.inc.d17.m.inc.d3 <- createDEGdfsaveManip("m.inc.d17", "m.inc.d3",  i) 
+  m.n2.m.inc.d17 <- createDEGdfsaveManip("m.n2", "m.inc.d17",  i) 
+  m.n2.m.inc.d9 <- createDEGdfsaveManip("m.n2", "m.inc.d9",  i)
+  m.n2.m.inc.d3 <- createDEGdfsaveManip("m.n2", "m.inc.d3",  i)
+  prolong.m.inc.d17 <- createDEGdfsaveManip("prolong", "m.inc.d17",  i) 
+  m.inc.d8.m.inc.d9 <- createDEGdfsaveManip("m.inc.d8", "m.inc.d9",  i) 
+  extend.m.n2 <- createDEGdfsaveManip("extend", "m.n2",  i) 
 }

### 2020-02-17 analysis/22_symphony.Rmd
now with manipuation degs and modules of genes for music

@@
   theme_B3() +
-  theme(axis.ticks = element_blank()) +
+  theme(axis.ticks = element_blank(),
+        axis.text.x = element_text(angle = 10, hjust = 1)) +
   labs(subtitle = "Sinnerman by Nina Simone", y = NULL)
@@
 ```{r soulmusic}
-p1 <- plot_grid(respectsong + theme(legend.position = "none"), feelinggoodsong + theme(legend.position = "none"), rel_widths = c(0.45,0.55))
 
-p3 <- plot_grid(p1, sinnermansong, nrow = 2)
+p3 <- plot_grid(respectsong + theme(legend.position = "bottom"), 
+                sinnermansong + theme(legend.position = "bottom"), 
+                nrow = 1, rel_widths = c(0.45, 0.55))
 p3
@@
 candidategenes <- c("PRL", "PRLR", 
-                 "VIP", "VIPR1", "VIPR2", 
-                 "OXT", "AVP", "AVPR1A", "AVPR1B", 
-                 "GNRH1","GNRHR", "NPVF",
-                 "NR3C1", "NR3C2",
-                 "ESR1", "ESR2", "AR",
-                 "DIO2","LEPR", "DIO3", "DIO1","CYP19A1",
-                 "HSPA14", "HSPA12A",
-                 "PTGES3", "HSD11B2","DRD5", "DRD1", "DRD2","PGE1", "PGF")
+                 #"VIP", "VIPR1", "VIPR2", 
+                 "OXT", "AVP",  "AVPR1B", # "AVPR1A",
+                # "GNRH1","GNRHR", "NPVF",
+                # "NR3C1", "NR3C2",
+                # "ESR1", "ESR2",
+                "AR",
+                # "DIO2","LEPR", "DIO3", "DIO1","CYP19A1",
+                # "HSPA14", "HSPA12A",
+                # "PTGES3", "HSD11B2",
+                 "DRD5", "DRD1", "DRD2",
+                 #"PGE1", "PGF",
+                 "BRCA1", "MYC", "PRKCZ" ,"FOSL2")
 
@@
 
-d4 %>%
-  filter( modulecolor != "turquoise",
-          # modulecolor != "greenyellow",
+turquoisegenes <- d4 %>%
+  filter( modulecolor == "turquoise",
           tissue == "pituitary") %>%
@@
   geom_image(aes(image=image), size = 0.15) +
+  geom_smooth(aes(x = as.numeric(treatment)), se = F, color = "turquoise") +
   facet_wrap(~modulecolor, scales = "free_y", nrow = 2) +
-  labs(subtitle = "WGCNA + canddiate genes in the pituitary", y = "gene expression", x = "parental stage") +
+  labs(subtitle = "WGCNA turquoise module", 
+       y = "gene expression  in the pitutiary", x = "parental stage") +
     facet_wrap(~gene, scales = "free_y", nrow = ) +
     theme_B3() +
-    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
+    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1),
+          strip.text = element_text(face = "italic"))
 
-d4 %>%
-  filter( modulecolor != "turquoise",
-          # modulecolor != "greenyellow",
+redgenes <- d4 %>%
+  filter( modulecolor == "red",
           tissue == "pituitary") %>%
@@
     ggplot(aes(x = treatment, y = m)) +
-  geom_image(aes(image=image), size = 0.1) +
-  labs(subtitle = "WGCNA + canddiate genes in the pituitary", y = "gene expression", x = "parental stage") +
+  geom_image(aes(image=image), size = 0.15) +
+  geom_smooth(aes(x = as.numeric(treatment)), se = F, color = "red") +
+  facet_wrap(~modulecolor, scales = "free_y", nrow = 2) +
+  labs(subtitle = "WGCNA red module", 
+       y = "gene expression  in the pitutiary", x = "parental stage") +
+    facet_wrap(~gene, scales = "free_y", nrow = ) +
     theme_B3() +
-    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_y_log10()
+    theme(legend.position = "none", 
+          axis.text.x = element_text(angle = 45, hjust = 1),
+          strip.text = element_text(face = "italic"))
 
+turquoisegenes
+redgenes
+
+plot_grid(turquoisegenes, redgenes)
+```
+
+```{r}
+
+d4 %>% 
+  filter(gene %in% c("PRL", "PRLR", "ESR2")) %>%
+ggplot( aes(x = treatment, y = m, color = gene)) +
+    geom_smooth(aes(x = as.numeric(treatment)), se = F) +
+    facet_grid(gene~tissue, scales = "free")
+
+head(df)
+head(df2)
 ```

### 2020-02-17 analysis/figure-7.Rmd
now with manipuation degs and modules of genes for music

---
title: "Fig7"
output: md_document
---

# Figure 7

```{r setup, include=TRUE}
library(tidyverse)
library(cowplot)

source("../R/themes.R")

knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
```

```{r fig6, fig.height=7, fig.width=7}
geneids <- read_csv("../metadata/00_geneinfo.csv")

DEG_manip_path <- "../results/DEseq2/manip/"   # path to the data
DEG_manip_files <- dir(DEG_manip_path, pattern = "*_DEGs.csv") 
DEG_manip_pathfiles <- paste0(DEG_manip_path, DEG_manip_files)
DEG_manip_files

manipDEGs <- DEG_manip_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::mutate(comparison = sapply(strsplit(file_name,'./results/DEseq2/manip/'), "[", 2),
                comparison = sapply(strsplit(comparison,'_DEGs.csv'), "[", 1)) %>%
  dplyr::select(comparison, gene, padj,logpadj, lfc, sextissue, direction)
head(manipDEGs)

DEG_char_path <- "../results/DEseq2/"   # path to the data
DEG_char_files <- dir(DEG_char_path, pattern = "*_DEGs.csv") 
DEG_char_pathfiles <- paste0(DEG_char_path, DEG_char_files)
DEG_char_files

charDEGs <- DEG_char_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::mutate(comparison = sapply(strsplit(file_name,'./results/DEseq2/'), "[", 2),
                comparison = sapply(strsplit(comparison,'_DEGs.csv'), "[", 1)) %>%
  dplyr::select(comparison, gene, padj,logpadj, lfc, sextissue, direction)
head(charDEGs)

allDEGs <- rbind(charDEGs, manipDEGs)
head(allDEGs)

nearhatch <- allDEGs %>% 
  dplyr::mutate(comparison = str_c(str_split_fixed(comparison,'\\_', 4)[,c(1,3)],collapse="_")) %>% 
  dplyr::filter(grepl("inc.d17|hatch|prolong|extend|m.inc.d8", comparison)) %>%
  dplyr::filter(!grepl("inc.d9|m.inc.d3", comparison))
head(nearhatch)

unique(nearhatch$comparison)
```

### 2020-02-18 analysis/22_symphony.Rmd
save files

@@
 ```{r}
-
 d4 %>% 

### 2020-02-18 analysis/figure-6.Rmd
save files

@@
 
-# Figure 4
+# Figure 6
 
@@
 
-vsd_path <- "../results/DEseq2/manip/"   # path to the data
-vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
-vsd_pathfiles <- paste0(vsd_path, vsd_files)
-vsd_files
+DEGchar <- read_csv("../results/DEGs-char.csv")
+DEGmanip <- read_csv("../results/DEGs-manip.csv")
+DEGall <- rbind(DEGchar, DEGmanip)
+tail(DEGall)
 
-allvsd <- vsd_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name")  %>% 
-  dplyr::rename("gene" = "X1")  %>% 
-  pivot_longer(cols = blk.s031.pu.d_female_gonad_prolong:y98.o50.x_male_pituitary_inc.d3, 
-               names_to = "samples", values_to = "counts")
+hatchnocontrol <- DEGall %>% filter(grepl("inc.d17", comparison)) %>% filter(!grepl("control", comparison))
+hatchnocontrol %>% group_by(comparison) %>% summarize(n = n())
 
-hypreplace <- read_csv("../results/DEseq2/manip/female_hypothalamus_hatch_extend_DEGs.csv") %>%
-  arrange(desc(logpadj)) %>% head(., n = 1) %>% pull(gene)
-hypreplace
+incd17hatch <- DEGall %>% filter(comparison == "inc.d17_hatch") %>%
+  arrange(gene) %>% distinct(comparison, gene)
+incd17hatch
 
-hypremove <- read_csv("../results/DEseq2/manip/female_hypothalamus_m.n2_hatch_DEGs.csv") %>%
-  arrange(desc(logpadj)) %>% head(., n = 1) %>% pull(gene)
-hypremove
+incd17prolong <- DEGall %>% filter(comparison == "inc.d17_prolong") %>%
+  arrange(gene) %>% distinct(comparison, gene)
+incd17prolong
 
-pitreplace <- read_csv("../results/DEseq2/manip/female_pituitary_hatch_extend_DEGs.csv") %>%
-  arrange(desc(logpadj)) %>% head(., n = 1) %>% pull(gene)
-pitreplace
+bothincs <- full_join(incd17hatch,incd17prolong, by = "gene" )
+bothincs
 
-pitremove <- read_csv("../results/DEseq2/manip/female_pituitary_m.n2_hatch_DEGs.csv") %>%
-  arrange(desc(logpadj)) %>% head(., n = 1) %>% pull(gene)
-pitremove
-
-gonreplace <- read_csv("../results/DEseq2/manip/female_gonad_inc.d17_prolong_DEGs.csv") %>%
-  arrange(desc(logpadj)) %>% head(., n = 1) %>% pull(gene)
-gonreplace
-
-gonremove <- read_csv("../results/DEseq2/manip/female_gonad_inc.d17_m.inc.d17_DEGs.csv") %>%
-  arrange(desc(logpadj)) %>% head(., n = 1) %>% pull(gene)
-gonremove
-
-plottopgenes <- function(whichgenes, whichtissue, whichsex, mysubtitle, whichtstages){
-  candidates  <- allvsd %>%
-    filter(gene %in% whichgenes) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/manip/'), "[", 2)) %>%
-    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
-                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
-                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
-    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
-    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
-    filter(tissue == whichtissue, sex %in% whichsex)
-
-  candidates$treatment <- factor(candidates$treatment, levels = alllevels2)
-
-  p <- candidates %>%
-    filter(treatment %in% whichtstages) %>%
-    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
-    geom_boxplot() + facet_wrap(~gene, scales = "free_y") + 
-    theme_B3() +
-    scale_fill_manual(values = allcolors) +
-    scale_color_manual(values = allcolors) +
-    theme(axis.text.x = element_blank(),
-          legend.position = "none",
-          axis.title.x = element_blank(),
-          strip.text = element_text(face = "italic")) +
-    labs(y = mysubtitle) 
-  return(p)
-
-}
-
-a <- plottopgenes(hypreplace, "hypothalamus", c("female"), "hypothalamus", c("hatch", "extend")) 
-b <- plottopgenes(pitreplace, "pituitary", c("female"), "pituitary", c("hatch", "extend")) 
-c <- plottopgenes(gonreplace, "gonad", c("female"), "gonads", c("inc.d17", "prolong"))
-
-fig6a <- plot_grid(a,b,c, ncol = 1)
-
-d <- plottopgenes(hypreplace, "hypothalamus", c("female", "male"), NULL, alllevels2) 
-e <- plottopgenes(pitreplace, "pituitary", c("female", "male"), NULL, alllevels2) 
-f <- plottopgenes(gonreplace, "gonad", c("female", "male"), NULL, alllevels2) + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-fig6b <- plot_grid(d,e,f, ncol = 1)
-
-fig6 <- plot_grid(fig6a, fig6b, rel_widths = c(1,2))
-fig6
+onlyhatch <- subset(bothincs, is.na(bothincs$comparison.y))
+onlyhatch
 ```

### 2020-02-18 analysis/02_DESeq2.Rmd
now with prolactin hi low

@@
 ```{r setup}
-
 library(tidyverse)
 library(DESeq2)
-
 library(BiocParallel)
 register(MulticoreParam(6))
+library(caret)
 
@@
 
-```{r manipulationwithin}
+```{r manipulationwithin, eval = F}
 # import "colData" which contains sample information and "countData" which contains read counts
@@
 ```
+
+## PRL-driven DESeq2
+
+```{r PRL}
+PRLdata <- read_csv("../results/PRLvsd.csv")
+PRLsamples <- colData %>%
+  mutate(samples = sapply(strsplit(as.character(V1), '.NYNO'), "[", 1))  %>%
+  select(samples) 
+PRLsamples
+
+# join with hi lo, and drop controls
+colDataPRL <- full_join(PRLdata, PRLsamples) %>% 
+  drop_na() %>%
+  select(samples, tissue, sex, treatment, hiloPRL) %>% 
+  filter(!treatment %in% c("control", "bldg"))
+
+# sample 100 only, better for deseq2
+colDataPRL <- colDataPRL[sample(1:nrow(colDataPRL), 100,
+   replace=FALSE),]
+
+colDataPRL <- as.data.frame(colDataPRL)
+row.names(colDataPRL) <- colDataPRL$samples
+
+savecols <- as.character(colDataPRL$samples) 
+savecols <- as.vector(savecols)
+
+newcountData <- countData %>% dplyr::select(one_of(savecols))
+
+colDataPRL %>%
+  group_by(sex, tissue, hiloPRL) %>%
+  summarize(n = n())
+
+dds <- DESeqDataSetFromMatrix(countData = newcountData,
+                                colData = colDataPRL,
+                                design = ~ tissue * sex * hiloPRL )
+dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
+print(dds)
+print(dim(dds))
+dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
+
+vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
+```

### 2020-02-18 analysis/figure-3.Rmd
now with prolactin hi low

@@
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
 
@@
   dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
-  dplyr::select(sex, tissue, treatment, gene, samples, counts)
+  dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
+  drop_na()
 PRLvsd$treatment <- factor(PRLvsd$treatment, levels = alllevels)
@@
 ```
+
+```{r determinePRLhiglow}
+PRLvsd2 <- PRLvsd %>%
+  filter(tissue == "pituitary")
+head(PRLvsd2)
+
+PRLvsd2 %>%
+  group_by(sex) %>%
+  summarize(median = median(counts))
+
+a <- ggplot(PRLvsd2, aes(x = sex, y = counts, fill = sex))  +
+  geom_boxplot() +
+  labs(x = "Sex", y = "Prolactin expression")+
+  theme_B3() + theme(legend.position = "none") +
+  scale_fill_manual(values = sexcolors) +
+  geom_hline(yintercept=18, linetype="dashed", color = "black")
+
+PRLvsd3 <- PRLvsd2 %>%
+  mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
+  drop_na()
+PRLvsd3$hiloPRL <- factor(PRLvsd3$hiloPRL, levels = c("lo", "hi"))
+
+b <- ggplot(PRLvsd3, aes(x = hiloPRL, y = counts, fill = sex))  +
+  geom_boxplot() +
+  labs(x = "Prolactin, binned", y = "Prolactin expression")+
+  theme_B3() + theme(legend.position = "none") +
+  scale_fill_manual(values = sexcolors)
+
+plot_grid(a,b, nrow = 1, rel_widths = c(1,1))
+
+PRLvsd3 %>%
+  group_by(sex, tissue, hiloPRL) %>%
+  summarize(n = n())
+
+write.csv(PRLvsd,"../results/PRLvsd3.csv", row.names = F)
+```

### 2020-02-18 analysis/02_DESeq2.Rmd
working on PRL results

@@
 library(caret) 
+library(cowplot)
 
@@
 PRLdata <- read_csv("../results/PRLvsd.csv")
+
 PRLsamples <- colData %>%
-  mutate(samples = sapply(strsplit(as.character(V1), '.NYNO'), "[", 1))  %>%
+  mutate(samples = V1)  %>%
   select(samples) 
@@
                                 colData = colDataPRL,
-                                design = ~ tissue * sex * hiloPRL )
+                                design = ~sex * hiloPRL )
 dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
@@
 vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
+
+plotDispEsts( dds, ylim = c(1e-6, 1e1) )
+
+ressex <- results( dds, contrast = c("sex", "male", "female") )
+reshiloPRL <- results( dds, contrast = c("hiloPRL", "hi", "lo") )
+hist( ressex$pvalue, breaks=20, col="grey" )
+hist( reshiloPRL$pvalue, breaks=20, col="grey" )
+
+restoDEGs <- function(res, up, down){
+  DEGs <- data.frame(gene = row.names(res),
+                        padj = res$padj, 
+                        logpadj = -log10(res$padj),
+                        lfc = res$log2FoldChange)
+  DEGs <- na.omit(DEGs)
+  DEGs <- DEGs %>%
+    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
+                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
+                                                           yes = down, no = "NS"))) %>% 
+    dplyr::arrange(gene)
+  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up)) 
+  # write DEGsframe of only significant genes
+  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
+  print(str(DEGs))
+  #partialfilename = paste("_", down, "_", up, sep = "")
+  #myfilename = paste0("../results/DESeq2/", mytissue, partialfilename, "_DEGs.csv")
+  #write.csv(DEGs, myfilename, row.names = F)
+  print(head(DEGs))
+  return(DEGs)
+}
+
+sexdegs <- restoDEGs(ressex, "male", "female")
+PRLdegs <- restoDEGs(reshiloPRL, "hi", "lo")
+
+plotbarvolcano <- function(df, mylabels, mycolorname){
+  p1 <- ggplot(df, aes(x = lfc, y = logpadj, color = direction)) +
+    geom_point() +
+    scale_color_manual(values = allcolors, name = mycolorname) +
+    theme_B3() +
+    theme(legend.position = "top") +
+    labs(x = "Log fold change", y = "Log adj. p-value") +
+    xlim(-5,5)
+
+  p2 <- ggplot(df, aes(x = direction,  fill = direction)) +
+    geom_bar() +
+    theme_B3() +
+    scale_fill_manual(values = allcolors) +
+    theme_B3() +
+    theme(legend.position = "none") +
+    labs(x = mycolorname, y = "Total DEGs")
+
+  p <- plot_grid(p1,p2, labels  = mylabels, label_size = 12)
+  plot(p)
+}
+
+top <- plotbarvolcano(sexdegs, c("a","b"), "sex")
+bottom <- plotbarvolcano(PRLdegs, c("c","d"), "prolactin expression")
+
+plot_grid(top,bottom, nrow = 2)
+
+head(PRLdegs)
+head(sexdegs)
+
+sexPRLdesg$direction.x <- as.character(sexPRLdesg$direction.x)
+
+sexPRLdesg <- full_join(sexdegs, PRLdegs, by = "gene") %>%
+  dplyr::select(gene, direction.x, direction.y) %>%
+  tidyr::replace_na(list(direction.x = "no sex effect", direction.y = "no PRL effect"))
+sexPRLdesg
+
+ggplot(sexPRLdesg, aes(x = direction.x, fill = direction.y)) +
+  geom_bar(position = "dodge") +
+  scale_fill_manual(values = allcolors, na.value= "#bdbdbd",
+                    labels= c("low PRL", "hi PRL", "NS"),
+                    name = "PRL DEGs") +
+  scale_x_discrete(breaks = c("female", "male", NA),
+                   labels = c("female", "male", "NS"),
+                   name = "Sex DEGs")
+
+interaction <- sexPRLdesg %>% filter(direction.x %in% c("male", "female") & 
+                                       direction.y %in% c("hi", "lo"))
+interaction
+
+sexonly <- sexPRLdesg %>% filter(direction.x %in% c("male", "female") & 
+                                       !direction.y %in% c("hi", "lo") )
+sexonly
+PRLonly <- sexPRLdesg %>% filter(!direction.x %in% c("male", "female")& 
+                                       direction.y %in% c("hi", "lo"))
+PRLonly
+
+write.csv(sexPRLdesg, "../results/DEG-PRLsex.csv")
 ```

### 2020-02-18 R/themes.R
working on PRL results

@@ -150,6 +150,9 @@ colorstissue <- c("hypothalamus" = "#d95f02",
 
 sexcolors <- c("female" = "#969696", "male" = "#525252")
 
+
+colorsPRL <-  c( "lo"= "#78c679",  "hi"= "#006837")
+
 colorsvolcano <-  c(colorscharnew,
                     
                     "m.inc.d8" = "#cbc9e2", 
@@ -161,7 +164,7 @@ colorsvolcano <-  c(colorscharnew,
                     
                     "NS" = "#bdbdbd")
 
-allcolors <- c(colorscharmaip, sexcolors, colorstissue)
+allcolors <- c(colorscharmaip, sexcolors, colorstissue, colorsPRL)
 
 colorsvolcanochar <-  c(colorscharnew,
                     "NS" = "#bdbdbd")

### 2020-02-18 analysis/02_DESeq2.Rmd
prl degs

@@
 
-```{r readfiles, eval=F, include=F}
+```{r readfiles, eval=T}
 # import "colData" which contains sample information and "countData" which contains read counts
@@
 
-```{r manipulation1, include = F}
+```{r manipulation1, eval = F, include = F}
 # import "colData" which contains sample information and "countData" which contains read counts
@@
 
-plotDispEsts( dds, ylim = c(1e-6, 1e1) )
-
 ressex <- results( dds, contrast = c("sex", "male", "female") )
@@
   DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up)) 
-  # write DEGsframe of only significant genes
   DEGs <- DEGs %>% dplyr::filter(direction != "NS")
-  print(str(DEGs))
-  #partialfilename = paste("_", down, "_", up, sep = "")
-  #myfilename = paste0("../results/DESeq2/", mytissue, partialfilename, "_DEGs.csv")
-  #write.csv(DEGs, myfilename, row.names = F)
   print(head(DEGs))
@@
 top <- plotbarvolcano(sexdegs, c("a","b"), "sex")
-bottom <- plotbarvolcano(PRLdegs, c("c","d"), "prolactin expression")
+bottom <- plotbarvolcano(PRLdegs, c("c","d"), "PRL")
 
-plot_grid(top,bottom, nrow = 2)
+ab <- plot_grid(top,bottom, nrow = 2)
 
-head(PRLdegs)
-head(sexdegs)
+sexPRLdesg <- full_join(sexdegs, PRLdegs, by = "gene")
 
@@
 
-ggplot(sexPRLdesg, aes(x = direction.x, fill = direction.y)) +
+write.csv(sexPRLdesg, "../results/DEG-PRLsex.csv")
+
+sexPRLdesg2 <- sexPRLdesg %>%
+  mutate(higherin = paste(direction.x, direction.y, sep = "_")) %>%
+  mutate(direction = paste(higherin, "PRL", sep = "")) %>%
+  select(gene, direction) %>%
+  group_by(direction) %>%
+  summarize(DEGs = str_c(gene , collapse = " "))
+sexPRLdesg2
+write.csv(sexPRLdesg2, "../results/PRLsexinteraction.csv")
+
+c <- ggplot(sexPRLdesg, aes(x = direction.x, fill = direction.y)) +
   geom_bar(position = "dodge") +
   scale_fill_manual(values = allcolors, na.value= "#bdbdbd",
-                    labels= c("low PRL", "hi PRL", "NS"),
-                    name = "PRL DEGs") +
+                    labels= c("lo", "hi", "NS"),
+                    name = "PRL") +
   scale_x_discrete(breaks = c("female", "male", NA),
                    labels = c("female", "male", "NS"),
-                   name = "Sex DEGs")
+                   name = "Sex")  +
+  labs(y = "Total DEGs") +
+  theme_B3() +
+  theme(legend.position = "top")
 
-interaction <- sexPRLdesg %>% filter(direction.x %in% c("male", "female") & 
-                                       direction.y %in% c("hi", "lo"))
-interaction
+plot_grid(ab, c, nrow = 1, rel_widths = c(2,1.2), labels = c(" ", "d"), label_size = 12)
 
-sexonly <- sexPRLdesg %>% filter(direction.x %in% c("male", "female") & 
-                                       !direction.y %in% c("hi", "lo") )
-sexonly
-PRLonly <- sexPRLdesg %>% filter(!direction.x %in% c("male", "female")& 
-                                       direction.y %in% c("hi", "lo"))
-PRLonly
-
-write.csv(sexPRLdesg, "../results/DEG-PRLsex.csv")
 ```

### 2020-02-19 analysis/02_DESeq2.Rmd
now with prl degs

@@
 
+head(sexPRLdesg)
+
+PRLhi <- sexPRLdesg %>%
+  filter(direction.y == "hi")
+PRLhi
+
+PRLwgcna <- read_csv("../results/08_PRL_associated.csv") %>%
+  dplyr::rename("gene" =  "x") %>%
+  mutate(wgcna = "yes")
+PRLwgcna
+
+PRLwgcnaDEGs <- left_join(PRLwgcna, sexPRLdesg) %>%
+  select(-direction.x) %>%
+  dplyr::rename("PRL" =  "direction.y") 
+PRLwgcnaDEGs %>%
+  group_by(wgcna, PRL) %>%
+  summarize(n = n())
+head(PRLwgcnaDEGs)
+# 90 of the 98 genes in the wgcna module are also DEGs, with increase expression in hi PRL
 ```

### 2020-02-19 analysis/figure-3.Rmd
working on PRL results
better prl figures

@@
 
-c <- plotfriz(charfviz) + labs(subtitle = "  ")
+c <- plotfriz(charfviz) + labs(subtitle = "  ") +
+  ylim(-250000,500000) + xlim(-250000,500000) + 
+  theme(axis.text = element_blank())
 
@@
 
-```{r determinePRLhiglow}
+```{r determinePRLhiglo, fig.height = 6, fig.width=3}
 PRLvsd2 <- PRLvsd %>%
@@
 
-b <- ggplot(PRLvsd3, aes(x = hiloPRL, y = counts, fill = sex))  +
+b <- ggplot(PRLvsd3, aes(x = sex, y = counts, fill = sex, color = hiloPRL))  +
   geom_boxplot() +
-  labs(x = "Prolactin, binned", y = "Prolactin expression")+
+  labs(x = "Prolactin, binned into `lo` and `hi`", y = "Prolactin expression")+
   theme_B3() + theme(legend.position = "none") +
-  scale_fill_manual(values = sexcolors)
+  scale_fill_manual(values = allcolors) +
+  scale_color_manual(values = allcolors)
 
-plot_grid(a,b, nrow = 1, rel_widths = c(1,1))
+plot_grid(a,b, nrow = 2, rel_widths = c(1,1))
 
@@
 
-write.csv(PRLvsd,"../results/PRLvsd3.csv", row.names = F)
+write.csv(PRLvsd3,"../results/PRLvsd.csv", row.names = F)
 ```

### 2020-02-20 analysis/figure-2.Rmd
save files
supple table 1

@@
 library(cowplot)
+library(kableExtra)
 source("../R/themes.R")
@@
 
-```{r fig2, fig.width=7}
+```{r allDEGs}
 
@@
 
-allDEG %>%
+overlappingDEGs <- allDEG %>%
+  filter(tissue == "hypothalamus",
+         comparison != "control_bldg") %>%
+  mutate(group = paste(sex, tissue,comparison, sep = ".")) %>%
+  group_by(gene) %>%
+  summarize(n = n()) %>%
+  arrange(desc(n))
+overlappingDEGs
+
+HYPcandidategenes <- c("OXT", "AVP", "GNRH1",  "AR", "POMC", "AGRP",
+                       "CRH", "AVPR1A", "AVPR1B", "AVPR2",
+                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR")
+
+suppletable1 <- allDEG %>%
+  filter(gene %in% HYPcandidategenes) %>%
   group_by(sex, tissue, comparison) %>%
-  summarize(n = n())
+  summarize(genes = str_c(gene, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = genes ) %>%
+  select(sex, tissue, control_bldg, lay_inc.d3, inc.d3_inc.d9,
+         inc.d9_inc.d17, hatch_n5)  %>%
+  arrange(sex, tissue)
+suppletable1
+
+write.csv(suppletable1, "../results/suppltable-1.csv", row.names = F)
+```
+
+## variance stabilized gene expression  (vsd)
+
+```{r vsd}
+geneids <- read_csv("../metadata/00_geneinfo.csv")
+
+vsd_path <- "../results/DEseq2/"   # path to the data
+vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
+vsd_pathfiles <- paste0(vsd_path, vsd_files)
+vsd_files
+
+allvsd <- vsd_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name")  %>% 
+  dplyr::rename("gene" = "X1") %>% 
+  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "samples", values_to = "counts") 
+```
+
+```{r fig2, fig.width=7, fig.height=7}
 
@@
   filter(comparison == "control_bldg") %>%
+  filter(tissue == "hypothalamus") %>%
 ggplot(aes(x = comparison,  fill = direction)) +
@@
   theme_B3() +
-  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
-         axis.text.x = element_blank(),
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+         #axis.text.x = element_blank(),
         legend.position = "none",
@@
   guides(fill = guide_legend(nrow = 1)) +
-  labs(x = " ", y = "Total Differentially Expressed Genes") +
+  labs(x = " ", y = "Hypothalamic DEGs") +
   scale_fill_manual(values = colorscharnew,
@@
                        drop = FALSE) +
-  scale_color_manual(values = sexcolors)
+  scale_color_manual(values = sexcolors) +
+  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
+            position = position_dodge(width = 1),
+            size = 2.5, color = "black")  +
+  ylim(0,4200)
 
@@
   filter(comparison != "control_bldg") %>%
+    filter(tissue == "hypothalamus") %>%
 ggplot(aes(x = comparison,  fill = direction)) +
   geom_bar(position = "dodge") +
-  facet_grid(tissue~sex) +
+  facet_wrap(~sex) +
   theme_B3() +
-  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
-         axis.text.x = element_blank(),
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+         #axis.text.x = element_blank(),
         legend.position = "none",
@@
             position = position_dodge(width = 1),
-            size = 2.5, color = "black")
+            size = 2.5, color = "black")  +
+  ylim(0,1150)
 
@@
 
-fig2 <- plot_grid(barplots, mylegend, nrow = 2, rel_heights = c(1,0.1))
-fig2
+plottopgenes <- function(whichgenes, whichtissue, whichsex, mysubtitle, whichtstages){
+  candidates  <- allvsd %>%
+    filter(gene %in% whichgenes) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
+    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
+    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
+    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
+    filter(tissue == whichtissue, sex %in% whichsex)  %>%
+    drop_na()
+
+  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
+
+  p <- candidates %>%
+    filter(treatment %in% whichtstages) %>%
+    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
+    geom_boxplot() + 
+    facet_wrap(~gene, scales = "free_y") + 
+    theme_B3() +
+    scale_fill_manual(values = allcolors) +
+    scale_color_manual(values = allcolors) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none",
+          strip.text = element_text(face = "italic")) +
+    labs(y = mysubtitle, x = "Parental stage") 
+  return(p)
+
+}
+
+hypDEGs <- c("OXT", "AVP", "GNRH1",  "AR")
+topgenes <- plottopgenes(hypDEGs, "hypothalamus", c("female", "male"), "Hypothalamic expression", charlevels)
+
+plot_grid(barplots, topgenes, nrow = 2, rel_heights = c(0.4,0.6))
+
 ```
+
+```{r suppletable1}
+kable(suppletable1)
+```
+
+```{r writefiles}
+#write_csv(allDEG, "../results/DEGs-char.csv")
+```

### 2020-02-20 analysis/figure-2.Rmd
suppl table 1

@@
 suppletable1 <- allDEG %>%
-  filter(gene %in% HYPcandidategenes) %>%
+  filter(gene %in% HYPcandidategenes, 
+         tissue == "hypothalamus") %>%
   group_by(sex, tissue, comparison) %>%
@@
   pivot_wider(names_from = comparison, values_from = genes ) %>%
-  select(sex, tissue, control_bldg, lay_inc.d3, inc.d3_inc.d9,
+  select(sex, tissue, control_bldg,  
          inc.d9_inc.d17, hatch_n5)  %>%

### 2020-02-20 analysis/figure-2.Rmd
now with 1 fig showing total degs and a few examples

--- 2020-02-10 analysis/figure-2.Rmd
@@
 library(cowplot)
+library(kableExtra) 
+library(ggtext) # for italic fonts
+library(ggsignif) # fot stats on plots
+
 source("../R/themes.R")
@@
 
-```{r fig2, fig.width=7}
-
+```{r allDEGs}
 DEG_path <- "../results/DEseq2/"   # path to the data
@@
 DEG_pathfiles <- paste0(DEG_path, DEG_files)
-DEG_files
+#DEG_files
 
@@
 
-allDEG %>%
-  group_by(sex, tissue, comparison) %>%
-  summarize(n = n())
+overlappingDEGs <- allDEG %>%
+  filter(tissue == "hypothalamus",
+         comparison != "control_bldg") %>%
+  mutate(group = paste(sex, tissue,comparison, sep = ".")) %>%
+  group_by(gene) %>%
+  summarize(n = n()) %>%
+  arrange(desc(n))
+overlappingDEGs
 
-p1 <- allDEG %>%
-  filter(comparison == "control_bldg") %>%
+candidategenes <- c("OXT", "AVP", "GNRH1",  "AR", "POMC", "AGRP",
+                       "CRH", "AVPR1A", "AVPR1B", "AVPR2",
+                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR",
+                    "SOX9")
+
+nocontrol <- allDEG %>%
+  filter(comparison != "control_bldg")
+
+malegon <- nocontrol %>%
+  filter(sex == "male", tissue == "gonad")
+```
+
+## variance stabilized gene expression  (vsd)
+
+```{r vsd}
+geneids <- read_csv("../metadata/00_geneinfo.csv")
+
+vsd_path <- "../results/DEseq2/"   # path to the data
+vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
+vsd_pathfiles <- paste0(vsd_path, vsd_files)
+vsd_files
+
+allvsd <- vsd_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name")  %>% 
+  dplyr::rename("gene" = "X1") %>% 
+  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "samples", values_to = "counts")
+
+getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
+  candidates  <- allvsd %>%
+    filter(gene %in% whichgenes) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
+    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
+    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
+    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
+    filter(tissue == whichtissue, sex %in% whichsex)  %>%
+    drop_na()
+  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
+  return(candidates)
+}
+
+hypvsd <- getcandidatevsd(candidategenes, "hypothalamus", sexlevels)
+pitvsd <- getcandidatevsd(candidategenes, "pituitary", sexlevels)
+gonvsd <- getcandidatevsd(candidategenes, "gonad", sexlevels)
+head(hypvsd)
+head(pitvsd)
+head(gonvsd)
+```
+
+```{r figurefunctions}
+
+makebargraph <- function(whichtissue, myylab, lowlim){
+  p <- allDEG %>%
+    filter(tissue == whichtissue,
+           comparison != "control_bldg") %>%
   ggplot(aes(x = comparison,  fill = direction)) +
@@
     theme_B3() +
-  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
-         axis.text.x = element_blank(),
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
           legend.position = "none",
@@
     guides(fill = guide_legend(nrow = 1)) +
-  labs(x = " ", y = "Total Differentially Expressed Genes") +
-  scale_fill_manual(values = colorscharnew,
+    labs(x = "Sequential parental care stage comparisons", 
+         y = myylab,
+         subtitle = " ") +
+  scale_fill_manual(values = allcolors,
                        name = " ",
                        drop = FALSE) +
-  scale_color_manual(values = sexcolors)
+  scale_color_manual(values = allcolors)  #+
+ # geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
+ #           position = position_dodge(width = 1),
+ #           size = 2.5, color = "black")  +
+ # ylim(lowlim, higherlim)
+  return(p)
+}
 
-p2 <- allDEG %>%
-  filter(comparison != "control_bldg") %>%
-ggplot(aes(x = comparison,  fill = direction)) +
-  geom_bar(position = "dodge") +
-  facet_grid(tissue~sex) +
+makeboxplots <- function(df, whichgene, myylab, whichsex){
+  p <- df %>%
+    filter(treatment %in% charlevels,
+           gene %in% whichgene,
+           sex %in% whichsex) %>%
+    filter(treatment != "control") %>%
+    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
+    geom_boxplot() + 
+    #geom_point() +
+    facet_wrap(~sex) +
     theme_B3() +
-  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
-         axis.text.x = element_blank(),
-        legend.position = "none",
-        strip.placement = "right")  +
-  guides(fill = guide_legend(nrow = 1)) +
-  labs(x = "Sequential Parental Care Stages  ", y = NULL) +
-  scale_fill_manual(values = colorscharnew,
-                       name = " ", drop = FALSE) +
-  scale_color_manual(values = sexcolors) +
-  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
-            position = position_dodge(width = 1),
-            size = 2.5, color = "black")
+    scale_fill_manual(values = allcolors) +
+    scale_color_manual(values = allcolors) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none") +
+    labs(y = myylab , x = "Parental stage", subtitle = "") +
+    theme(axis.title.y = element_markdown())
+  return(p)
+}
+```
 
-forlegend <- allDEG %>%
-ggplot(aes(x = comparison,  fill = direction)) +
-  geom_bar(position = "dodge") +
-   scale_fill_manual(values = colorscharnew,
-                       name = "increased in", drop = FALSE) +
-  scale_color_manual(values = sexcolors) +
-  guides(color = F) +
-  guides(fill= guide_legend(nrow =1)) +
-  theme_B3() +
-  theme(legend.position = "bottom")
+## HPG
 
-mylegend <- get_legend(forlegend)
+```{r fig2, fig.width=7, fig.height=7}
 
-barplots <- plot_grid(p1,p2, nrow=1, rel_widths = c(0.9,3))
+allDEG %>%
+  filter(gene %in% candidategenes, 
+         tissue == "hypothalamus") %>%
+  group_by(sex, tissue, comparison) %>%
+  summarize(genes = str_c(gene, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = genes ) %>%
+  select(sex, tissue, control_bldg,  
+        inc.d9_inc.d17, hatch_n5)  %>%
+  arrange(sex, tissue)
 
-fig2 <- plot_grid(barplots, mylegend, nrow = 2, rel_heights = c(1,0.1))
-fig2
+write.csv(table1, "../results/table-1.csv", row.names = F)
+
+# hyp
+p1 <- makebargraph("hypothalamus","Hypothalamic DEGs") + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
+p2 <- makeboxplots(hypvsd, "AR","*AR* expression", "male") +
+              geom_signif(comparisons=list(c("inc.d9", "inc.d17")), annotations= "*", 
+              y_position = 8, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
+              theme(axis.text.x = element_blank(), axis.title.x = element_blank())
+
+p3 <- makeboxplots(hypvsd, "DRD1","*DRD1* expression", "female") +
+              geom_signif(comparisons=list(c("hatch", "n5")), annotations= "*", 
+              y_position = 10, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
+              #geom_signif(comparisons=list(c("control", "bldg")), annotations= "*", 
+              #y_position = 10, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) + 
+              theme(axis.text.x = element_blank(), axis.title.x = element_blank())
+
+p23 <- plot_grid(p3,p2, labels = c("b", "c"), label_size = 12)
+
+# pit
+
+p4 <- makebargraph("pituitary","Pituitary DEGs")  +  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), strip.text = element_blank())
+p5 <- makeboxplots(pitvsd, "PRL","*PRL* expression", "male") +  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), strip.text = element_blank())+
+              #geom_signif(comparisons=list(c("control", "bldg")), annotations= "*", 
+              #y_position = 21.5, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
+              geom_signif(comparisons=list(c("inc.d9", "inc.d17")), annotations= "*", 
+              y_position = 21.5, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) 
+p6 <- makeboxplots(pitvsd, "PRL","*PRL* expression", "female")  +  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), strip.text = element_blank()) +
+              geom_signif(comparisons=list(c("inc.d9", "inc.d17")), annotations= "*", 
+              y_position = 21, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
+              #geom_signif(comparisons=list(c("control", "bldg")), annotations= "*", 
+              #y_position = 21, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
+              geom_signif(comparisons=list(c("hatch", "n5")), annotations= "*", 
+              y_position = 21, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 )
+
+p56 <- plot_grid(p6,p5, labels = c("e", "f"), label_size = 12)
+
+# gon
+
+p7 <- makebargraph("gonad","Gonadal DEGs") +  theme(strip.text = element_blank())
+p8 <- makeboxplots(gonvsd, "SOX9","*SOX9* expression", "male")  +  theme(strip.text = element_blank()) +
+              geom_signif(comparisons=list(c("lay", "inc.d3")), annotations= "*", 
+              y_position = 7.5, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 )
+p9 <- makeboxplots(gonvsd, "AVPR1A","*AVPR1A* expression", "female") +  theme(strip.text = element_blank()) +
+              #geom_signif(comparisons=list(c("control", "bldg")), annotations= "*", 
+              #y_position = 9.5, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
+              geom_signif(comparisons=list(c("lay", "inc.d3")), annotations= "*", 
+              y_position = 9.5, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
+              geom_signif(comparisons=list(c("inc.d3", "inc.d9")), annotations= "*", 
+              y_position = 9.5, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 )
+
+p89 <- plot_grid(p9,p8, labels = c("h", "i"), label_size = 12)
+
+a <- plot_grid(p1, p23, nrow = 1, rel_heights = c(0.5,0.5), labels = c("a", " "), label_size = 12)
+b <- plot_grid(p4, p56, nrow = 1, rel_heights = c(0.5,0.5), labels = c("d", " "), label_size = 12)
+c <- plot_grid(p7, p89, nrow = 1, rel_heights = c(0.5,0.5), labels = c("g", " "), label_size = 12)
+plot_grid(a,b,c, nrow = 3, rel_heights = c(1,0.9,1.3))
+
 ```
+
+```{r}
+allDEG %>%
+  group_by(tissue, comparison) %>%
+  summarize(totalDEGs = n()) %>%
+  arrange(tissue, comparison)
+```
+
+```{r writefiles}
+table1 <- allDEG %>%
+  filter(gene %in% candidategenes) %>%
+  group_by(sex, tissue, comparison) %>%
+  summarize(genes = str_c(gene, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = genes ) %>%
+  select(sex, tissue, control_bldg, lay_inc.d3, inc.d3_inc.d9,
+        inc.d9_inc.d17, hatch_n5)  %>%
+  arrange( tissue, sex)
+table1
+
+write_csv(table1, "../results/table1.csv")
+```

### 2020-02-21 analysis/02_DESeq2.Rmd
new fig 2

@@
 register(MulticoreParam(6))
-library(caret) 
+library(caret)  # for sampling data
 library(cowplot)
@@
 
-## Warning: This script can take a very long time to run.
+## Notes
 
-DESeq2 was not designed to run on 100+ samples. But, I really like it, so I do it anyways. Some of these commands take like 15 min to run using 6 cores.
+- DESeq2 was not designed to run on 100+ samples. But, I really like it, so I do it anyways. Some of these commands take like 15 min to run using 6 cores. 
+_ Also, I don't run every chuck every time. When I want new analysese, I add them to the bottom and set `eval = T` for chunks I don't want to rerurn
 
@@
 ```
+
+## Hypothesis testing - external stimuli and/or prolactin
+
+```{r}
+# label each bird with weather or not their pituitary had hi PRL or not
+# then make new column with nest, eggs, or chicks, drop controls
+
+PRLdata <- read_csv("../results/PRLvsd.csv") %>%
+  mutate(V1 = samples) %>%
+  mutate(bird = sapply(strsplit(as.character(samples),'\\_'), "[", 1)) %>%
+  select(bird,hiloPRL)
+
+# join with hi lo, and drop controls
+colDatahypothesis <- full_join(colData, PRLdata, by = "bird") %>% 
+  drop_na() %>%
+  select(V1, tissue, sex, sextissue, treatment, hiloPRL) %>% 
+  filter(!treatment %in% c("control", "bldg")) %>%
+  mutate(external = fct_collapse(treatment, eggs = c("lay", "inc.d3", "inc.d9", "inc.d17"),
+                                            chicks = c("hatch", "n5", "n9")))
+
+colDatahypothesis <- as.data.frame(colDatahypothesis)
+row.names(colDatahypothesis) <- colDatahypothesis$V1
+
+savecols <- as.character(colDatahypothesis$V1) 
+savecols <- as.vector(savecols)
+
+newcountData <- countData %>% dplyr::select(one_of(savecols))
+
+colDatahypothesis %>%
+  group_by(sex, tissue, hiloPRL, external) %>%
+  summarize(n = n())
+
+hypothesesistesting <- function(whichfactor, up, down, mytissue){
+
+  res <- results(dds, contrast = c(whichfactor, up, down), independentFiltering = T, alpha = 0.1)
+
+  DEGs <- data.frame(gene = row.names(res),
+                        padj = res$padj, 
+                        logpadj = -log10(res$padj),
+                        lfc = res$log2FoldChange,
+                        sextissue = mytissue)
+  DEGs <- na.omit(DEGs)
+  DEGs <- DEGs %>%
+    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
+                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
+                                                           yes = down, no = "NS"))) %>% 
+    dplyr::arrange(desc(lfc))
+
+  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up))
+
+  # write DEGsframe of only significant genes
+  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
+  print(str(DEGs))
+
+  partialfilename = paste("_", down, "_", up, sep = "")
+  myfilename = paste0("../results/DESeq2/hypothesis/", mytissue, partialfilename, "_DEGs.csv")
+
+  write.csv(DEGs, myfilename, row.names = F)
+  # return DEGs frome with all data, included NS genes
+  #print(head(DEGs))
+}
+
+for(i in levels(colDatahypothesis$sextissue)){
+
+  newcolData <- subsetcolData2(colDatahypothesis, i)
+
+  # save counts that match colData
+  savecols <- as.character(newcolData$V1) 
+  savecols <- as.vector(savecols)
+
+  newcountData <- countData %>% dplyr::select(one_of(savecols))
+
+  dds <- DESeqDataSetFromMatrix(countData = newcountData,
+                                colData = newcolData,
+                                design = ~ external + hiloPRL )
+  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
+  print(dds)
+  print(dim(dds))
+  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
+
+  vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
+
+  myfilename = paste0("../results/DEseq2/hypothesis/", i, "_vsd.csv")
+  write.csv(vsd, myfilename)
+
+  # save differential gene expression results
+  # sexmalefemale <- hypothesesistesting("sex", "male", "female",  i) 
+  externalschickseggs <- hypothesesistesting("external", "chicks", "eggs",  i) 
+  hiloPRL <- hypothesesistesting("hiloPRL", "hi","lo",  i) 
+}
+```

### 2020-02-21 analysis/figure-2.Rmd
new fig 2

--- 2020-02-20 analysis/figure-2.Rmd
@@
 library(kableExtra) 
+library(ggtext) # for italic fonts
+library(ggsignif) # fot stats on plots
+
 source("../R/themes.R")
@@
 ```{r allDEGs}
-
 DEG_path <- "../results/DEseq2/"   # path to the data
@@
 allDEG$direction <- factor(allDEG$direction, levels = charlevels)
-
-overlappingDEGs <- allDEG %>%
-  filter(tissue == "hypothalamus",
-         comparison != "control_bldg") %>%
-  mutate(group = paste(sex, tissue,comparison, sep = ".")) %>%
-  group_by(gene) %>%
-  summarize(n = n()) %>%
-  arrange(desc(n))
-overlappingDEGs
-
-HYPcandidategenes <- c("OXT", "AVP", "GNRH1",  "AR", "POMC", "AGRP",
-                       "CRH", "AVPR1A", "AVPR1B", "AVPR2",
-                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR")
-
-suppletable1 <- allDEG %>%
-  filter(gene %in% HYPcandidategenes, 
-         tissue == "hypothalamus") %>%
-  group_by(sex, tissue, comparison) %>%
-  summarize(genes = str_c(gene, collapse = " ")) %>%
-  pivot_wider(names_from = comparison, values_from = genes ) %>%
-  select(sex, tissue, control_bldg,  
-         inc.d9_inc.d17, hatch_n5)  %>%
-  arrange(sex, tissue)
-suppletable1
-
-write.csv(suppletable1, "../results/suppltable-1.csv", row.names = F)
 ```
 
-## variance stabilized gene expression  (vsd)
+```{r figurefunctions}
 
-```{r vsd}
-geneids <- read_csv("../metadata/00_geneinfo.csv")
-
-vsd_path <- "../results/DEseq2/"   # path to the data
-vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
-vsd_pathfiles <- paste0(vsd_path, vsd_files)
-vsd_files
-
-allvsd <- vsd_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name")  %>% 
-  dplyr::rename("gene" = "X1") %>% 
-  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
-               names_to = "samples", values_to = "counts") 
-```
-
-```{r fig2, fig.width=7, fig.height=7}
-
-p1 <- allDEG %>%
-  filter(comparison == "control_bldg") %>%
-  filter(tissue == "hypothalamus") %>%
+makebargraph <- function(whichtissue, myylab, lowlim, higherlim){
+  p <- allDEG %>%
+    filter(tissue == whichtissue,
+           comparison != "control_bldg") %>%
   ggplot(aes(x = comparison,  fill = direction)) +
@@
     theme(axis.text.x = element_text(angle = 45, hjust = 1),
-         #axis.text.x = element_blank(),
-        legend.position = "none",
-        strip.text.y = element_blank())  +
+          legend.position = "none")  +
     guides(fill = guide_legend(nrow = 1)) +
-  labs(x = " ", y = "Hypothalamic DEGs") +
-  scale_fill_manual(values = colorscharnew,
+    labs(x = "Sequential parental care stage comparisons", 
+         y = myylab,
+         subtitle = " ") +
+  scale_fill_manual(values = allcolors,
                        name = " ",
                        drop = FALSE) +
-  scale_color_manual(values = sexcolors) +
+  scale_color_manual(values = allcolors) +
   geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
@@
             size = 2.5, color = "black")  +
-  ylim(0,4200)
+  ylim(lowlim, higherlim)
+  return(p)
+}
 
-p2 <- allDEG %>%
-  filter(comparison != "control_bldg") %>%
-    filter(tissue == "hypothalamus") %>%
+```
+
+## Internal versus external
+
+```{r hypothesis}
+DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG2 <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
+head(allDEG)
+
+allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)
+
+allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
+allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
+                                                      "eggs vs. chicks" = "eggs_chicks"))
+allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))
+
+makenewbargraph <- function(whichtissue, myxlab, whichcomparison, lowlim, higherlim){
+  p <- allDEG2 %>%
+    filter(tissue == whichtissue,
+           comparison == whichcomparison) %>%
   ggplot(aes(x = comparison,  fill = direction)) +
     geom_bar(position = "dodge") +
-  facet_wrap(~sex) +
+    facet_grid(tissue~sex) +
     theme_B3() +
     theme(axis.text.x = element_text(angle = 45, hjust = 1),
-         #axis.text.x = element_blank(),
-        legend.position = "none",
-        strip.placement = "right")  +
+          legend.position = "none")  +
     guides(fill = guide_legend(nrow = 1)) +
-  labs(x = "Sequential Parental Care Stages  ", y = NULL) +
-  scale_fill_manual(values = colorscharnew,
-                       name = " ", drop = FALSE) +
-  scale_color_manual(values = sexcolors) +
+    labs(x = myxlab, 
+         y = "Total DEGs") +
   geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
@@
            size = 2.5, color = "black")  + 
-  ylim(0,1150)
-
-forlegend <- allDEG %>%
-ggplot(aes(x = comparison,  fill = direction)) +
-  geom_bar(position = "dodge") +
-   scale_fill_manual(values = colorscharnew,
-                       name = "increased in", drop = FALSE) +
-  scale_color_manual(values = sexcolors) +
-  guides(color = F) +
-  guides(fill= guide_legend(nrow =1)) +
-  theme_B3() +
-  theme(legend.position = "bottom")
-
-mylegend <- get_legend(forlegend)
-
-barplots <- plot_grid(p1,p2, nrow=1, rel_widths = c(0.9,3))
-
-plottopgenes <- function(whichgenes, whichtissue, whichsex, mysubtitle, whichtstages){
-  candidates  <- allvsd %>%
-    filter(gene %in% whichgenes) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
-    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
-                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
-                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
-    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
-    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
-    filter(tissue == whichtissue, sex %in% whichsex)  %>%
-    drop_na()
-
-  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
-
-  p <- candidates %>%
-    filter(treatment %in% whichtstages) %>%
-    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
-    geom_boxplot() + 
-    facet_wrap(~gene, scales = "free_y") + 
-    theme_B3() +
-    scale_fill_manual(values = allcolors) +
-    scale_color_manual(values = allcolors) +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
-          legend.position = "none",
-          strip.text = element_text(face = "italic")) +
-    labs(y = mysubtitle, x = "Parental stage") 
+      ylim(lowlim, higherlim)
   return(p)
-
 }
-
-hypDEGs <- c("OXT", "AVP", "GNRH1",  "AR")
-topgenes <- plottopgenes(hypDEGs, "hypothalamus", c("female", "male"), "Hypothalamic expression", charlevels)
-
-plot_grid(barplots, topgenes, nrow = 2, rel_heights = c(0.4,0.6))
 
@@
 
-```{r suppletable1}
-kable(suppletable1)
+## DEGS
+
+```{r fig2, fig.width=7, fig.height=6}
+# hyp
+p1 <- makebargraph("hypothalamus","DEGs", 0, 1250) + theme(axis.text.x = element_blank(), 
+                                                               axis.title.x = element_blank())
+
+# pit
+p2 <- makebargraph("pituitary","DEGs", 0, 1250)  +  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), 
+                                                              strip.text.x = element_blank())
+
+# gon
+p3 <- makebargraph("gonad","DEGs", 0, 1250) +  theme(strip.text.x = element_blank())
+
+bcd <- plot_grid(p1,p2,p3, nrow = 3, rel_heights = c(1.2,1,1.5), labels = c("b", "c", "d"), label_size = 12)
+
+expdesign <- png::readPNG("../figures/images/DoveParentsRNAseq_DEGS.png")
+expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
+
+plot_grid(expdesign, bcd, labels = c("a", " "), label_size = 12, nrow = 2, rel_heights = c(0.2,1))
+
+b <- makenewbargraph(tissuelevel, "External stimuli", "eggs vs. chicks", 0, 1250)  + theme(strip.text.y = element_blank()) +
+  scale_fill_manual(values = c("#80cdc1", "#018571"))
+c <- makenewbargraph(tissuelevel, "Pituitary PRL expression   ", "lo vs. hi PRL   ", 0, 1250) + labs(y = NULL) + theme(axis.text.y = element_blank()) +
+  scale_fill_manual(values = c("#dfc27d", "#a6611a"))
+
+plot_grid(b, c, rel_widths = c(1,1), nrow = 1, labels = "auto", label_size = 12)
 ```
 
+## total degs
+
+```{r}
+allDEG %>%
+  group_by(tissue, comparison) %>%
+  summarize(totalDEGs = n()) %>%
+  arrange(tissue, comparison)
+```
+
+## candidate genes 
 ```{r writefiles}
-#write_csv(allDEG, "../results/DEGs-char.csv")
+candidategenes <- c("OXT", "AVP", "GNRH1",  "AR", "POMC", "AGRP",
+                       "CRH", "AVPR1A", "AVPR1B", "AVPR2",
+                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9")
+
+table1 <- allDEG %>%
+  filter(gene %in% candidategenes) %>%
+  group_by(sex, tissue, comparison) %>%
+  summarize(genes = str_c(gene, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = genes ) %>%
+  select(sex, tissue, control_bldg, lay_inc.d3, inc.d3_inc.d9,
+        inc.d9_inc.d17, hatch_n5)  %>%
+  arrange( tissue, sex)
+table1
+
+write_csv(table1, "../results/table1.csv")
 ```

### 2020-02-21 analysis/figure-X.Rmd
new fig 2

---
title: "Fig X"
author: "Rayna M Harris"
date: "2/20/2020"
output: html_document
---

# all things box plots

## candidate genes

```{r}
candidategenes <- c("OXT", "AVP", "GNRH1",  "AR", "POMC", "AGRP",
                       "CRH", "AVPR1A", "AVPR1B", "AVPR2",
                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9")

```

## variance stabilized gene expression  (vsd)

```{r vsd}
geneids <- read_csv("../metadata/00_geneinfo.csv")

vsd_path <- "../results/DEseq2/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts")

getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
    filter(tissue == whichtissue, sex %in% whichsex)  %>%
    drop_na()
  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
  return(candidates)
}

hypvsd <- getcandidatevsd(candidategenes, "hypothalamus", sexlevels)
pitvsd <- getcandidatevsd(candidategenes, "pituitary", sexlevels)
gonvsd <- getcandidatevsd(candidategenes, "gonad", sexlevels)
head(hypvsd)
head(pitvsd)
head(gonvsd)
```

## Figs

```{r}
makeboxplots <- function(df, whichgene, myylab, whichsex){
  p <- df %>%
    filter(treatment %in% charlevels,
           gene %in% whichgene,
           sex %in% whichsex) %>%
    filter(treatment != "control") %>%
    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
    geom_boxplot() + 
    #geom_point() +
    facet_wrap(~sex) +
    theme_B3() +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none") +
    labs(y = myylab , x = "Parental stage", subtitle = "") +
    theme(axis.title.y = element_markdown())
  return(p)
}

## hyp 
p2 <- makeboxplots(hypvsd, "AR","*AR* expression", "male") +
              geom_signif(comparisons=list(c("inc.d9", "inc.d17")), annotations= "*", 
              y_position = 8, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
              theme(axis.text.x = element_blank(), axis.title.x = element_blank())

p3 <- makeboxplots(hypvsd, "DRD1","*DRD1* expression", "female") +
              geom_signif(comparisons=list(c("hatch", "n5")), annotations= "*", 
              y_position = 10, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
              #geom_signif(comparisons=list(c("control", "bldg")), annotations= "*", 
              #y_position = 10, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) + 
              theme(axis.text.x = element_blank(), axis.title.x = element_blank())

p23 <- plot_grid(p3,p2, labels = c("b", "c"), label_size = 12)

## pit

p5 <- makeboxplots(pitvsd, "PRL","*PRL* expression", "male") +  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), strip.text = element_blank())+
              #geom_signif(comparisons=list(c("control", "bldg")), annotations= "*", 
              #y_position = 21.5, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
              geom_signif(comparisons=list(c("inc.d9", "inc.d17")), annotations= "*", 
              y_position = 21.5, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) 
p6 <- makeboxplots(pitvsd, "PRL","*PRL* expression", "female")  +  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), strip.text = element_blank()) +
              geom_signif(comparisons=list(c("inc.d9", "inc.d17")), annotations= "*", 
              y_position = 21, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
              #geom_signif(comparisons=list(c("control", "bldg")), annotations= "*", 
              #y_position = 21, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
              geom_signif(comparisons=list(c("hatch", "n5")), annotations= "*", 
              y_position = 21, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 )

p56 <- plot_grid(p6,p5, labels = c("e", "f"), label_size = 12)

## gon

p8 <- makeboxplots(gonvsd, "SOX9","*SOX9* expression", "male")  +  theme(strip.text = element_blank()) +
              geom_signif(comparisons=list(c("lay", "inc.d3")), annotations= "*", 
              y_position = 7.5, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 )
p9 <- makeboxplots(gonvsd, "AVPR1A","*AVPR1A* expression", "female") +  theme(strip.text = element_blank()) +
              #geom_signif(comparisons=list(c("control", "bldg")), annotations= "*", 
              #y_position = 9.5, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
              geom_signif(comparisons=list(c("lay", "inc.d3")), annotations= "*", 
              y_position = 9.5, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 ) +
              geom_signif(comparisons=list(c("inc.d3", "inc.d9")), annotations= "*", 
              y_position = 9.5, tip_length = 0.005, vjust = 0.5, color= "black" , textsize = 3 )

p89 <- plot_grid(p9,p8, labels = c("h", "i"), label_size = 12)

```

### 2020-02-21 analysis/figure-2.Rmd
now with allDEGs file

@@
 allDEG %>%
-  group_by(tissue, comparison) %>%
+  group_by(sex, tissue, comparison) %>%
   summarize(totalDEGs = n()) %>%
@@
 write_csv(table1, "../results/table1.csv")
+
+suppletable1 <- allDEG %>%
+  filter(comparison != "control_bldg") %>%
+  group_by(sex, tissue, comparison) %>%
+  arrange( tissue, sex)
+
+suppletable1 %>%
+  group_by(tissue) %>%
+  summarize(totalDEGs = n())
+
+write_csv(table1, "../results/suppletable1.csv")
+
 ```

### 2020-02-21 analysis/figure-2.Rmd
new fig 2b

@@
 
-b <- makenewbargraph(tissuelevel, "External stimuli", "eggs vs. chicks", 0, 1250)  + theme(strip.text.y = element_blank()) +
-  scale_fill_manual(values = c("#80cdc1", "#018571"))
-c <- makenewbargraph(tissuelevel, "Pituitary PRL expression   ", "lo vs. hi PRL   ", 0, 1250) + labs(y = NULL) + theme(axis.text.y = element_blank()) +
-  scale_fill_manual(values = c("#dfc27d", "#a6611a"))
+```
 
-plot_grid(b, c, rel_widths = c(1,1), nrow = 1, labels = "auto", label_size = 12)
+```{r fig2b, fig.width=5, fig.height=6}
+
+expdesign2 <- png::readPNG("../figures/images/DoveParentsRNAseq_hypothesis.png")
+expdesign2 <- ggdraw() +  draw_image(expdesign2, scale = 1)
+
+b <- makenewbargraph(tissuelevel, NULL, "eggs vs. chicks", 0, 800)  + theme(strip.text.y = element_blank()) +
+  scale_fill_manual(values = allcolors, name = "higher in") +
+  labs(subtitle = " ")
+c <- makenewbargraph(tissuelevel, NULL, "lo vs. hi PRL   ", 0, 800) + labs(y = NULL) + theme(axis.text.y = element_blank()) +
+  scale_fill_manual(values =  allcolors, name = "higher in") +
+  labs(subtitle = " ")
+
+bc <- plot_grid(b, c, rel_widths = c(1,1), nrow = 1, labels = c("b", "c"), label_size = 12)
+
+plot_grid(expdesign2, bc, labels = c("a", " "), label_size = 12, nrow = 2, rel_heights = c(0.4,1))
+
 ```

### 2020-02-21 analysis/figure-2.Rmd
better bar graphs, no facets

--- 2020-02-21 analysis/figure-2.Rmd
@@
 
-makenewbargraph <- function(whichtissue, myxlab, whichcomparison, lowlim, higherlim){
-  p <- allDEG2 %>%
-    filter(tissue == whichtissue,
-           comparison == whichcomparison) %>%
-  ggplot(aes(x = comparison,  fill = direction)) +
-    geom_bar(position = "dodge") +
-    facet_grid(tissue~sex) +
-    theme_B3() +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
-          legend.position = "none")  +
-    guides(fill = guide_legend(nrow = 1)) +
-    labs(x = myxlab, 
-         y = "Total DEGs") +
-  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
-           position = position_dodge(width = 1),
-           size = 2.5, color = "black")  + 
-      ylim(lowlim, higherlim)
-  return(p)
-}
-
 ```
@@
 
-b <- makenewbargraph(tissuelevel, "External stimuli", "eggs vs. chicks", 0, 1250)  + theme(strip.text.y = element_blank()) +
-  scale_fill_manual(values = c("#80cdc1", "#018571"))
-c <- makenewbargraph(tissuelevel, "Pituitary PRL expression   ", "lo vs. hi PRL   ", 0, 1250) + labs(y = NULL) + theme(axis.text.y = element_blank()) +
-  scale_fill_manual(values = c("#dfc27d", "#a6611a"))
+```
 
-plot_grid(b, c, rel_widths = c(1,1), nrow = 1, labels = "auto", label_size = 12)
+```{r fig2b, fig.width=5, fig.height=5}
+
+makenewbargraph <- function(whichtissue, whichsex,  whichcomparison, lowlim, higherlim){
+  p <- allDEG2 %>%
+    filter(tissue == whichtissue,
+           comparison == whichcomparison,
+           sex == whichsex) %>%
+    ggplot(aes(x = comparison,  fill = direction)) +
+    geom_bar(position = "dodge", drop = FALSE) +
+    theme_B3() +
+    theme(legend.position = "none")  +
+    guides(fill = guide_legend(nrow = 1)) +
+    labs( y = whichtissue) +
+  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
+           position = position_dodge(width = 1),
+           size = 2.5, color = "black")  + 
+      ylim(lowlim, higherlim) +
+  scale_fill_manual(values = allcolors, name = "higher in")   + 
+    theme(axis.text.x = element_blank()) 
+  return(p)
+}
+
+b11 <- makenewbargraph("hypothalamus", "female","eggs vs. chicks", 0, 800) +  labs(subtitle = "females", x = NULL, y = "Hyp. DEGS")
+b21 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 800)  + labs(x = NULL, y = "Pit. DEGS")
+b31 <- makenewbargraph("gonad", "female", "eggs vs. chicks", 0, 800) + labs(x = "eggs vs. chicks", y = "Gon. DEGS")   
+b112131 <- plot_grid(b11,b21,b31, nrow = 3, rel_heights = c(1.1,1,1.1))
+
+b12 <- makenewbargraph("hypothalamus", "male",  "eggs vs. chicks", 0, 800) + labs(subtitle = "males", x = NULL)+ 
+  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
+b22 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 800) + labs(x = NULL)+ 
+  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
+b32 <- makenewbargraph("gonad", "male", "eggs vs. chicks", 0, 800) + labs(x = "eggs vs. chicks")   + 
+  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
+b122232 <- plot_grid(b12, b22,b32, nrow = 3, rel_heights = c(1.1,1,1.1))
+
+c11 <- makenewbargraph("hypothalamus", "female", "lo vs. hi PRL   ", 0, 2500) +  labs(subtitle = "females", x = NULL) + 
+  theme(axis.title.y = element_blank())
+c21 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2500)  + labs(x = NULL)+ 
+  theme(axis.title.y = element_blank())
+c31 <- makenewbargraph("gonad", "female", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL")  + 
+  theme(axis.title.y = element_blank()) 
+c112131 <- plot_grid(c11,c21,c31, nrow = 3, rel_heights = c(1.1,1,1.1))
+
+c12 <- makenewbargraph("hypothalamus", "male",  "lo vs. hi PRL   ", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
+  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
+c22 <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = NULL)+ 
+  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
+c32 <- makenewbargraph("gonad", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL")   + 
+  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
+c122232 <- plot_grid(c12, c22,c32, nrow = 3, rel_heights = c(1.1,1,1.1))
+
+hypothesisbars <-  plot_grid(b112131, b122232, c112131, c122232, nrow = 1, rel_widths = c(1.3, 1, 1.1, 1),
+          labels = c("b", " ", "c", " "), label_size = 12)
+
+expdesign2 <- png::readPNG("../figures/images/DoveParentsRNAseq_hypothesis.png")
+expdesign2 <- ggdraw() +  draw_image(expdesign2, scale = 1)
+
+plot_grid(expdesign2, hypothesisbars, rel_heights = c(0.5,1), nrow = 2, labels = c("a", " "), label_size = 12)
+```
+
+```{r GO}
+## shinygo termshttp://bioinformatics.sdstate.edu/go/
+allDEG2 %>% filter(tissue == "pituitary", direction == "chicks") %>% arrange(gene) %>% pull(gene) # nervous systme development, neurogenesis, neuron differentiaion 
+allDEG2 %>% filter(tissue == "pituitary", direction == "eggs") %>% arrange(gene) %>% pull(gene) # miotic cell cycle, cell cycle, cell cycle process
+allDEG2 %>% filter(tissue == "pituitary", direction == "hi") %>% arrange(gene) %>% pull(gene) # this 1000 combined with next 10000
+allDEG2 %>% filter(tissue == "pituitary", direction == "hi") %>% arrange(desc(gene)) %>% pull(gene) # Cellular localization, #Cellular protein localization, Protein targeting to ER
+
+#miotic cell cycle, miotic cell cycle process, cell cycle
+allDEG2 %>% filter(tissue == "pituitary", direction == "lo") %>% arrange(gene)  %>% pull(gene)   # this 1000 combined with next 10000
+allDEG2 %>% filter(tissue == "pituitary", direction == "lo") %>% arrange(desc(gene)) %>% pull(gene) #Anatomical structure morphogenesis, Regulation of developmental process, Regulation of signaling
+
+allDEG2 %>%
+  filter(tissue == "pituitary") %>%
+  select(sex, tissue, comparison, gene) %>%
+  group_by(sex, tissue, gene) %>%
+  summarize(groups = str_c(comparison , collapse = " "))   %>%
+   group_by(sex, tissue, groups) %>%
+  summarize(n = n()) %>%
+  arrange(desc(n))
+
+allDEG2 %>%
+  filter(tissue == "pituitary") %>%
+  select(sex, tissue, comparison, gene) %>%
+  group_by(sex, tissue, gene) %>%
+  summarize(groups = str_c(comparison , collapse = " "))  %>%
+  filter(groups == "eggs vs. chicks") %>%
+  arrange(gene) %>%
+  distinct(gene) %>%
+  pull(gene)  # Nervous system development, Cell development, Anatomical structure morphogenesis
+
+allDEG2 %>%
+  filter(comparison == "eggs vs. chicks") %>%
+  group_by(sex, tissue, comparison, direction) %>%
+  summarize(n = n()) %>%
+  arrange(comparison, direction, tissue, sex)
+
+allDEG2 %>%
+  filter(comparison == "lo vs. hi PRL   ") %>%
+  group_by(sex, tissue, comparison, direction) %>%
+  summarize(n = n()) %>%
+  arrange(comparison, direction, tissue, sex)
+
+```
+```{r}
+allDEG2 %>%
+  filter(tissue == "pituitary") %>%
+  select(sex, tissue, comparison, gene) %>%
+  group_by(sex, tissue, gene) %>%
+  summarize(groups = str_c(comparison , collapse = " "))  %>%
+  filter(groups == "lo vs. hi PRL   ") %>%
+  arrange(gene) %>%
+  distinct(gene) %>%
+  pull(gene)  # Nervous system development, Cell development, Anatomical structure morphogenesis
+
+allDEG2 %>%
+  group_by(sex, tissue, comparison, direction) %>%
+  summarize(n = n())
 ```

### 2020-02-21 analysis/figure-1.Rmd
smaller font

@@
 
-abcd <- plot_grid(a,b,c,d, nrow = 1, labels = c("b", "c", "d", "e"), label_size = 12 )
+abcd <- plot_grid(a,b,c,d, nrow = 1, labels = c("b", "c", "d", "e"), label_size = 8 )
 
@@
 
-abcde <- plot_grid(expdesign, abcd, nrow = 2, labels = c("a", "b"), label_size = 12, rel_heights = c(0.5,1))
+abcde <- plot_grid(expdesign, abcd, nrow = 2, labels = c("a", "b"), label_size = 8, rel_heights = c(0.5,1))
 
@@
 
-hij <- plot_grid(h,i,j, nrow = 1, labels = c("h", "i", "j"), label_size = 12)
+hij <- plot_grid(h,i,j, nrow = 1, labels = c("h", "i", "j"), label_size = 8)
 

### 2020-02-21 analysis/figure-2.Rmd
smaller font

@@
             position = position_dodge(width = 1),
-            size = 2.5, color = "black")  +
+            size = 2, color = "black")  +
   ylim(lowlim, higherlim)
@@
 
-```{r fig2, fig.width=7, fig.height=6}
+```{r fig2, fig.width=6, fig.height=5}
 # hyp
@@
 
-bcd <- plot_grid(p1,p2,p3, nrow = 3, rel_heights = c(1.2,1,1.5), labels = c("b", "c", "d"), label_size = 12)
+bcd <- plot_grid(p1,p2,p3, nrow = 3, rel_heights = c(1.2,1,1.5), labels = c("b", "c", "d"), label_size = 8)
 
@@
 
-plot_grid(expdesign, bcd, labels = c("a", " "), label_size = 12, nrow = 2, rel_heights = c(0.2,1))
+plot_grid(expdesign, bcd, labels = c("a", " "), label_size = 8, nrow = 2, rel_heights = c(0.2,1))
 
@@
            position = position_dodge(width = 1),
-           size = 2.5, color = "black")  + 
+           size = 2, color = "black")  + 
       ylim(lowlim, higherlim) +
@@
 hypothesisbars <-  plot_grid(b112131, b122232, c112131, c122232, nrow = 1, rel_widths = c(1.3, 1, 1.1, 1),
-          labels = c("b", " ", "c", " "), label_size = 12)
+          labels = c("b", " ", "c", " "), label_size = 8)
 
@@
 
-plot_grid(expdesign2, hypothesisbars, rel_heights = c(0.5,1), nrow = 2, labels = c("a", " "), label_size = 12)
+plot_grid(expdesign2, hypothesisbars, rel_heights = c(0.5,1), nrow = 2, labels = c("a", " "), label_size = 8)
 ```

### 2020-02-21 R/themes.R

@@ -1,7 +1,7 @@
 # custom themes
 
 theme_B3 <- function () { 
-  theme_classic(base_size = 11) +
+  theme_classic(base_size = 8) +
     theme(
       panel.grid.major  = element_blank(),  # remove major gridlines
       panel.grid.minor  = element_blank(),  # remove minor gridlines

### 2020-02-21 analysis/figure-2.Rmd
better ylim

@@
 
-b11 <- makenewbargraph("hypothalamus", "female","eggs vs. chicks", 0, 800) +  labs(subtitle = "females", x = NULL, y = "Hyp. DEGS")
-b21 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 800)  + labs(x = NULL, y = "Pit. DEGS")
-b31 <- makenewbargraph("gonad", "female", "eggs vs. chicks", 0, 800) + labs(x = "eggs vs. chicks", y = "Gon. DEGS")   
+b11 <- makenewbargraph("hypothalamus", "female","eggs vs. chicks", 0, 2500) +  labs(subtitle = "females", x = NULL, y = "Hyp. DEGS")
+b21 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2500)  + labs(x = NULL, y = "Pit. DEGS")
+b31 <- makenewbargraph("gonad", "female", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks", y = "Gon. DEGS")   
 b112131 <- plot_grid(b11,b21,b31, nrow = 3, rel_heights = c(1.1,1,1.1))
 
-b12 <- makenewbargraph("hypothalamus", "male",  "eggs vs. chicks", 0, 800) + labs(subtitle = "males", x = NULL)+ 
+b12 <- makenewbargraph("hypothalamus", "male",  "eggs vs. chicks", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
   theme(axis.title.y = element_blank(), axis.text.y = element_blank())
-b22 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 800) + labs(x = NULL)+ 
+b22 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2500) + labs(x = NULL)+ 
   theme(axis.title.y = element_blank(), axis.text.y = element_blank())
-b32 <- makenewbargraph("gonad", "male", "eggs vs. chicks", 0, 800) + labs(x = "eggs vs. chicks")   + 
+b32 <- makenewbargraph("gonad", "male", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks")   + 
   theme(axis.title.y = element_blank(), axis.text.y = element_blank())

### 2020-02-21 analysis/figure-2.Rmd
new candidate genes

@@
 ```{r writefiles}
+geneids <- read_csv("../metadata/00_geneinfo.csv")
+
 candidategenes <- c("OXT", "AVP", "GNRH1",  "AR", "POMC", "AGRP",
                        "CRH", "AVPR1A", "AVPR1B", "AVPR2",
-                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9")
+                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9", 
+                    "ESR1","ESR2", "LBH", "CDK1", "BRCA1",
+                    "PTEN", "CREBBP", "FOS", "JUN", "EGR1",
+                     "BDNF", "GRM2",
+                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16")
 
 table1 <- allDEG %>%
-  filter(gene %in% candidategenes) %>%
+  filter(gene %in% candidategenes,
+         comparison != "group_bldg") %>%
   group_by(sex, tissue, comparison) %>%
@@
   pivot_wider(names_from = comparison, values_from = genes ) %>%
-  select(sex, tissue, control_bldg, lay_inc.d3, inc.d3_inc.d9,
-        inc.d9_inc.d17, hatch_n5)  %>%
+  select(sex, tissue, bldg_lay, lay_inc.d3, inc.d3_inc.d9,
+        inc.d9_inc.d17,inc.d17_hatch, hatch_n5, n5_n9)  %>%
   arrange( tissue, sex)

### 2020-02-21 analysis/figure-2.Rmd
remove GO

@@
 
-```{r GO}
-## shinygo termshttp://bioinformatics.sdstate.edu/go/
-allDEG2 %>% filter(tissue == "pituitary", direction == "chicks") %>% arrange(gene) %>% pull(gene) # nervous systme development, neurogenesis, neuron differentiaion 
-allDEG2 %>% filter(tissue == "pituitary", direction == "eggs") %>% arrange(gene) %>% pull(gene) # miotic cell cycle, cell cycle, cell cycle process
-allDEG2 %>% filter(tissue == "pituitary", direction == "hi") %>% arrange(gene) %>% pull(gene) # this 1000 combined with next 10000
-allDEG2 %>% filter(tissue == "pituitary", direction == "hi") %>% arrange(desc(gene)) %>% pull(gene) # Cellular localization, #Cellular protein localization, Protein targeting to ER
-
-#miotic cell cycle, miotic cell cycle process, cell cycle
-allDEG2 %>% filter(tissue == "pituitary", direction == "lo") %>% arrange(gene)  %>% pull(gene)   # this 1000 combined with next 10000
-allDEG2 %>% filter(tissue == "pituitary", direction == "lo") %>% arrange(desc(gene)) %>% pull(gene) #Anatomical structure morphogenesis, Regulation of developmental process, Regulation of signaling
-
-allDEG2 %>%
-  filter(tissue == "pituitary") %>%
-  select(sex, tissue, comparison, gene) %>%
-  group_by(sex, tissue, gene) %>%
-  summarize(groups = str_c(comparison , collapse = " "))   %>%
-   group_by(sex, tissue, groups) %>%
-  summarize(n = n()) %>%
-  arrange(desc(n))
-
-allDEG2 %>%
-  filter(tissue == "pituitary") %>%
-  select(sex, tissue, comparison, gene) %>%
-  group_by(sex, tissue, gene) %>%
-  summarize(groups = str_c(comparison , collapse = " "))  %>%
-  filter(groups == "eggs vs. chicks") %>%
-  arrange(gene) %>%
-  distinct(gene) %>%
-  pull(gene)  # Nervous system development, Cell development, Anatomical structure morphogenesis
-
-allDEG2 %>%
-  filter(comparison == "eggs vs. chicks") %>%
-  group_by(sex, tissue, comparison, direction) %>%
-  summarize(n = n()) %>%
-  arrange(comparison, direction, tissue, sex)
-
-allDEG2 %>%
-  filter(comparison == "lo vs. hi PRL   ") %>%
-  group_by(sex, tissue, comparison, direction) %>%
-  summarize(n = n()) %>%
-  arrange(comparison, direction, tissue, sex)
-
-```
-```{r}
-allDEG2 %>%
-  filter(tissue == "pituitary") %>%
-  select(sex, tissue, comparison, gene) %>%
-  group_by(sex, tissue, gene) %>%
-  summarize(groups = str_c(comparison , collapse = " "))  %>%
-  filter(groups == "lo vs. hi PRL   ") %>%
-  arrange(gene) %>%
-  distinct(gene) %>%
-  pull(gene)  # Nervous system development, Cell development, Anatomical structure morphogenesis
-
-allDEG2 %>%
-  group_by(sex, tissue, comparison, direction) %>%
-  summarize(n = n())
-```
-
 ## total degs

### 2020-02-21 analysis/figure-2.Rmd
better supple tables

@@
 
-candidategenes <- c("OXT", "AVP", "GNRH1",  "AR", "POMC", "AGRP",
+candidategenes <- c("OXT", "AVP", "GNRH1", "GNRHR", "CGNRH-R",
+                    "AR", "POMC", "AGRP",
                        "CRH", "AVPR1A", "AVPR1B", "AVPR2",
@@
          comparison != "group_bldg") %>%
+    arrange(gene) %>%
+
   group_by(sex, tissue, comparison) %>%
@@
   group_by(sex, tissue, comparison) %>%
-  arrange( tissue, sex)
+  arrange( tissue, sex, direction, gene)
+head(suppletable1)
 
@@
 
-write_csv(table1, "../results/suppletable1.csv")
+write_csv(suppletable1, "../results/suppletable1.csv")
 
+allDEG2 %>%
+  filter(gene %in% candidategenes,
+         comparison != "group_bldg") %>%
+  arrange(gene) %>%
+  group_by(sex, tissue, comparison) %>%
+  summarize(genes = str_c(gene, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = genes ) %>%
+  arrange( tissue, sex)
+
+suppletable2 <- allDEG2 %>%
+  filter(comparison != "control_bldg") %>%
+  group_by(sex, tissue, comparison) %>%
+  arrange( tissue, sex, direction, gene)
+head(suppletable2)
+write_csv(suppletable2, "../results/suppletable1.csv")
 ```

### 2020-02-21 analysis/figure-2.Rmd
more candidate genes, better parental lists

--- 2020-02-21 analysis/figure-2.Rmd
@@
 
-candidategenes <- c("OXT", "AVP", "GNRH1",  "AR", "POMC", "AGRP",
-                       "CRH", "AVPR1A", "AVPR1B", "AVPR2",
+candidategenes <- c("OXT", "AVP", "GNRH1", "GNRHR", "CGNRH-R",
+                    "AR", "POMC", "AGRP", 
+                       "CRH", "AVPR1A", "AVPR1B", "AVPR2","VIP",
                        "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9", 
@@
                     "PTEN", "CREBBP", "FOS", "JUN", "EGR1",
-                     "BDNF", "GRM2",
-                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16")
+                     "BDNF", "GRM2","GRIA1",
+                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16", 
+                    "DIO3", "DIO2", "DIO1")
 
@@
   filter(gene %in% candidategenes,
-         comparison != "group_bldg") %>%
+         comparison != "control_bldg") %>%
+    arrange(gene) %>%
   group_by(sex, tissue, comparison) %>%
@@
 
+table2 <- allDEG2 %>%
+  filter(gene %in% candidategenes) %>%
+  arrange(gene) %>%
+  group_by(sex, tissue, comparison) %>%
+  summarize(genes = str_c(gene, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = genes ) %>%
+  arrange( tissue, sex)
+table2
+
 write_csv(table1, "../results/table1.csv")
+write_csv(table2, "../results/table2.csv")
+```
+
+## supplemental tables, all DEGS
+
+```{r}
 
@@
 
-write_csv(table1, "../results/suppletable1.csv")
+suppletable2 <- allDEG2 %>%
+  filter(comparison != "control_bldg") %>%
+  group_by(sex, tissue, comparison) %>%
+  arrange( tissue, sex, direction, gene)
+head(suppletable2)
+
+write_csv(suppletable1, "../results/suppletable1.csv")
+
+write_csv(suppletable2, "../results/suppletable2.csv")
+
+head(allDEG) 
+```
+
+## female pit, parental/repro only, not PRL
+
+```{r}
+allDEGslim <- allDEG %>% filter(comparison != "control_bldg")
+
+femalepitPRL <- allDEG2 %>% filter(sex == "female", tissue == "pituitary") %>%
+  filter(comparison != "eggs vs. chicks") %>% arrange(gene) %>% pull(gene)
+
+femalepitparental <- allDEGslim %>% 
+  filter(!gene %in% femalepitPRL,
+         sex == "female", tissue == "pituitary") %>%
+  arrange(gene, sex, tissue, comparison, direction) %>%
+  distinct(gene) %>% pull(gene)
+femalepitparental
+
+malepitPRL <- allDEG2 %>% filter(sex == "male", tissue == "pituitary") %>%
+  filter(comparison != "eggs vs. chicks") %>% arrange(gene) %>% pull(gene)
+
+malepitparental <- allDEGslim %>% 
+  filter(!gene %in% malepitPRL,
+         sex == "male", tissue == "pituitary") %>%
+  arrange(gene, sex, tissue, comparison, direction) %>%
+  distinct(gene) %>% pull(gene)
+malepitparental
 

### 2020-02-26 analysis/figure-2.Rmd
now with table 1 in readme

@@
   arrange( tissue, sex)
-table1
+kable(table1)
 
@@
   arrange( tissue, sex)
-table2
+table(table2)
 

### 2020-03-01 analysis/figure-4.Rmd
second manipulatino figure
new output for plotnplay.R : :

@@
 
-```{r setup, include=FALSE}
+# Figure 4
+
+```{r setup, include=TRUE}
 library(tidyverse)
@@
 gongenes
+
+candidategenes <- c("PRL", "PRLR" ,  "AVPR1B" , "MYC" ,  
+                "MAP2KA", "MSH" ,   "PALB2" , "BRCA1", 
+                "OXT", "ESR1", "ESR2", "CISH", 
+                "POU2F1", "RAP1GAP", "CEBPB","ARHGAP32"
+                )
+
+returnvsds <- function(whichgenes){
+
+  df  <- allvsd %>%
+    filter(gene %in% whichgenes) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
+    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
+    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
+    dplyr::select(sex, tissue, treatment, gene, samples, counts)
+
+  df$treatment <- factor(df$treatment, levels = alllevels)
+  print(head(df))
+  return(df)
+}
+
+candidatevsd <- returnvsds(candidategenes)
 
@@
 ```
+
+```{r savefiles}
+write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
+```

### 2020-03-02 analysis/figure-4.Rmd
new fig 4

@@
 
-```{r fig4, fig.height=7, fig.width=7}
+```{r data}
 geneids <- read_csv("../metadata/00_geneinfo.csv")
@@
 
-hypgenes <- read_csv("../results/DEseq2/female_hypothalamus_hatch_n5_DEGs.csv") %>%
-  arrange(desc(logpadj)) %>% head(., n = 2) %>% pull(gene)
-hypgenes
-
-pitgenes <- read_csv("../results/DEseq2/female_pituitary_inc.d9_inc.d17_DEGs.csv") %>%
-  arrange(desc(logpadj)) %>% head(., n = 2) %>% pull(gene)
-pitgenes
-
-gongenes <- read_csv("../results/DEseq2/female_gonad_lay_inc.d3_DEGs.csv") %>%
-  arrange(desc(logpadj)) %>% head(., n = 2) %>% pull(gene)
-gongenes
+hypgenes <- c("POMC", "DRD1")
+pitgenes <- c("PRL", "BRCA1")
+gongenes <- c("AVPR1A", "PRLR")
 
@@
 
+```
+
+```{r fig4, fig.height=7, fig.width=7}
 plottopgenes <- function(whichgenes, whichtissue, whichsex, mysubtitle, whichtstages){
@@
 b <- plottopgenes(pitgenes, "pituitary", c("female"), "pituitary", c("inc.d9", "inc.d17")) 
-c <- plottopgenes(gongenes, "gonad", c("female"), "gonads", c("lay", "inc.d3"))
+c <- plottopgenes(gongenes, "gonad", c("female"), "gonads", c("lay", "inc.d3")) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 

### 2020-03-02 analysis/figure-3.Rmd
new fig 3

@@
 # pca
-charpca <- subsetmakepca(tissuelevels, charlevels, sexlevels)	
-charfviz <- makefvizdf(tissuelevels, charlevels, sexlevels)
+charpca <- subsetmakepca("pituitary", charlevels, sexlevels)	
+charfviz <- makefvizdf("pituitary", charlevels, sexlevels)
 
@@
 a <- plotcolorfulpcs(charpca,charpca$treatment, allcolors) + labs(subtitle = " ") +
-  theme(legend.position = c(0.6,0.1), 
+  theme(legend.position = "none", 
         legend.direction = "horizontal", 
@@
   guides(color = FALSE) +
-  labs(subtitle = " ")
+  labs(subtitle = "pituitary")
 
-b <- plotprolactin(PRLhyp, PRLhyp$counts, "PRL", "hypothalamus") + 
-  theme(legend.position = c(0.7,0.9), axis.text.x = element_blank(), 
-        axis.title.x = element_blank(),
-        axis.title.y = element_text(face = "italic"),
-        legend.title = element_blank()) + guides(fill = F)
-
-c <- plotfriz(charfviz) + labs(subtitle = "  ") +
-  ylim(-250000,500000) + xlim(-250000,500000) + 
+b <- plotfriz(charfviz) + labs(subtitle = "pituitary") +
+  #ylim(-250000,500000) + xlim(-250000,500000) + 
   theme(axis.text = element_blank())
 
-d <- plotprolactin(PRLpit, PRLpit$counts, "PRL", "pituitary") + 
-  theme(legend.position = "none", axis.text.x = element_blank(), axis.title.x = element_blank(),
+c <- plotprolactin(PRLpit, PRLpit$counts, "PRL", "pituitary") + 
+  theme(legend.position = "none", 
+         axis.text.x = element_text(angle = 45, hjust = 1),
         axis.title.y = element_text(face = "italic"))
 
-e <- plotprolactin(prolactin, prolactin$plasma_conc, "prolactin (ng/mL)", "blood") +
+d <- plotprolactin(prolactin, prolactin$plasma_conc, "prolactin (ng/mL)", "blood") +
   theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
 
-f <- plotprolactin(PRLgon, PRLgon$counts, "PRL", "gonads") + 
-  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1),
-        axis.title.y = element_text(face = "italic"))
+abcd <- plot_grid(a,b,c,d, labels = c("b", "c", "d", "e"), ncol = 2, label_size = 8)
 
-plot_grid(a,b,c,d,e,f, labels = "auto", ncol = 2, rel_heights = c(1,1,1.2))
+expdesign <- png::readPNG("../figures/images/parentalstages_fig1a.png")
+expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
+
+abcde <- plot_grid(expdesign, abcd, nrow = 2, labels = c("a", "b"), label_size = 8, rel_heights = c(0.3,1))
+abcde
 

### 2020-03-04 analysis/figure-1.Rmd
new fig 1

@@
 
-a <- plottsneelipse(chartsne, chartsne$tissue, allcolors) + labs(subtitle = "tissue\n")    
-b <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = " ", subtitle = "tissue * sex\n")    
-c <- plottsneelipse(ftsne, ftsne$treatment, allcolors ) + labs(y = " ", subtitle = "treatment * tissue\nfemales")
-d <- plottsneelipse(mtsne, mtsne$treatment, allcolors ) + labs(y = " ", subtitle = "treatment * tissue\nmales")
+a <- plottsneelipse(chartsne, chartsne$tissue, allcolors) + labs(subtitle = "~ tissue")    
+b <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = " ", subtitle = "~ tissue * sex")    
+c <- plottsneelipse(ftsne, ftsne$tissue, allcolors ) + labs(y = " ", subtitle = "female ~ tissue")
+d <- plottsneelipse(mtsne, mtsne$tissue, allcolors ) + labs(y = " ", subtitle = "male * tissue")
 
@@
 
-h <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus")  + facet_wrap(~sex)
-i <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary", y = NULL) + facet_wrap(~sex) 
-j <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads", y = NULL)  + facet_wrap(~sex)
+h <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus ~ treatment")  + facet_wrap(~sex, scales = "free")
+i <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary ~ treatment", y = NULL) + facet_wrap(~sex, scales = "free") + theme(legend.position = )
+j <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads ~ treatment", y = NULL)  + facet_wrap(~sex, scales = "free")
 
@@
 fig1
+
 ```

### 2020-03-04 analysis/figure-3.Rmd
new fig 1

@@
 ```
+
+```{r gon}
+# pca
+charpca <- subsetmakepca("gonads", charlevels, sexlevels)	
+charfviz <- makefvizdf("gonads", charlevels, sexlevels)
+
+a <- plotcolorfulpcs(charpca,charpca$treatment, allcolors) + labs(subtitle = " ") +
+  theme(legend.position = "none", 
+        legend.direction = "horizontal", 
+        legend.key.size = unit(0.5, 'lines')) + 
+  guides(color = FALSE) +
+  labs(subtitle = "gonad")
+
+b <- plotfriz(charfviz) + labs(subtitle = "gonad") +
+  #ylim(-250000,500000) + xlim(-250000,500000) + 
+  theme(axis.text = element_blank())
+
+plot_grid(a,b)
+
+head(charpca)
+
+ggplot(charpca, aes(x = PC1, y = PC2, shape = tissue, color = treatment )) +	
+    geom_point(size = 1)  +	
+  stat_ellipse( ) +
+    theme_B3() +
+  scale_color_manual(values = colorschar)
+
+ggplot(charpca, aes(x = PC1, y = PC3, shape = tissue, color = treatment )) +	
+    geom_point(size = 1)  +	
+  stat_ellipse( ) +
+    theme_B3() +
+  scale_color_manual(values = colorschar) +
+  facet_wrap(~sex)
+
+ggplot(charpca, aes(x = PC1, y = PC3, shape = tissue, color = treatment )) +	
+    geom_point(size = 1)  +	
+  stat_ellipse( ) +
+    theme_B3() +
+  scale_color_manual(values = colorschar) +
+  facet_wrap(~sex)
+```

### 2020-03-04 analysis/figure-4.Rmd
new fig 1

@@
 
+DEGs <- read_csv("../results/DEGs-char.csv")  %>%
+  dplyr::filter(comparison != "control_bldg") %>%
+  dplyr::distinct(gene)  %>% arrange(gene) %>%
+  filter(!grepl("LOC", gene)) #%>% 
+  #pull(gene)
+
+HUGO <- read_table("../metadata/HUGOGeneNomenclature.txt")
+names(HUGO)[1] <- "all"
+HUGO$gene <- sapply(strsplit(HUGO$all,'\t'), "[", 3)
+HUGO$name <- sapply(strsplit(HUGO$all,'\t'), "[", 4)
+HUGO <- HUGO %>%distinct(gene,name)
+
+DEGsHUGO <- inner_join(DEGs, HUGO, by = "gene") %>% pull(gene)
+
+parentalbehavior <- read_table("../results/goterms/parentalbehavior.txt")
+names(parentalbehavior)[1] <- "allGO"
+parentalbehavior$id <- sapply(strsplit(parentalbehavior$allGO,'\t'), "[", 1)
+parentalbehavior$gene <- sapply(strsplit(parentalbehavior$allGO,'\t'), "[", 2)
+parentalbehavior$name <- sapply(strsplit(parentalbehavior$allGO,'\t'), "[", 3)
+parentalbehavior$GO <- sapply(strsplit(parentalbehavior$allGO,'\t'), "[", 6)
+parentalbehavior$allGO <- NULL
+parentalbehaviorgenes <- parentalbehavior %>% 
+  mutate(gene = str_to_upper(gene)) %>%  pull(gene)
+parentalbehaviorgenes
+
 returnvsds <- function(whichgenes){
@@
   df  <- allvsd %>%
-    filter(gene %in% whichgenes) %>%
+    dplyr::filter(!grepl("control", samples)) %>%
+    dplyr::filter(gene %in% whichgenes) %>%
     dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
@@
 candidatevsd <- returnvsds(candidategenes)
-
+#DEGvsd <- returnvsds(DEGsHUGO) too big 1G
+parentalvsd <- returnvsds(parentalbehaviorgenes)
 ```
@@
 write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
+write.csv(parentalvsd, "../../musicalgenes/data/parentalcounts.csv")
+
 ```

### 2020-03-04 R/themes.R

@@ -34,6 +34,8 @@ mytheme <- function(){
 
 charlevels <- c("control", "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9")
 
+charlevelsnocontrol <- c( "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9")
+
 maniplevels <- c( "remove.d03" ,  "remove.d09" ,  "remove.d17" , "remove.d20" ,  "extend" , "prolong", "early")
 
 maniplevels1 <- c( "m.inc.d3" ,  "m.inc.d8" ,  "m.inc.d9" , "m.inc.d17" , "m.n2", "prolong" , "extend")

### 2020-03-06 docs/manuscript/elife-v1.tex
now with manuscritp draft

% !TeX root = RJwrapper.tex
\title{Internal physiology and external cues drive changes in reproductive axis
of parenting pigeons}
\author{by Rayna M Harris \footnote{equal contribtion}, Suzanne H. Austin \footnote{equal contribtion}, Andrew S. Lang, Victoria S. Farrar, April Booth, Tanner Feustel, Matthew D. MacManes, Rebecca M. Calisi \footnote{corresponding author}}

\maketitle

% Any extra LaTeX you need in the preamble

In species that provide parental care, successful rearing of offspring involves a shift from aggressive and sexual behaviors to more caring and nurturing ones.  We hypothesize that these shifts in behavior are related to changes in gene expression that are controlled by internal mechanisms rather than by the external cues.  We found that gene expression in the hypothalamus is fairly insensitive to the natural transition between egg and nestling stages; however, in the pituitary and gonads, gene expression showed much more plasticity over the course of reproduction. We show evidence that natural transitions in reproductive and parental behavior by mediated internal clock controlling gene expression in the pitary and gonads but that the hypothalamus is very sensitive to external cues that signal a distruption in the natural parental care cycle. This research provides a deeper understanding of the interplay of genes, hormones, and parental care and has important implications for understanding molecular processes that underlie behavior transitions.

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Reproduction is a critical period in an organism's life cycle. Parents
must balance their own needs with the energetic costs and
time-constraints required to successfully reproduce and rear their
offspring to independence(1). A complex series of evolutionarily
conserved physiological processes underlie behaviors typically
associated with reproduction in vertebrates, including courtship,
nesting, and parental care. These processes are primarily mediated by
the reproductive axis, which consists of the hypothalamus in the brain,
the pituitary, and the gonads (i.e., the HPG). Our understanding of the
physiological changes that adults navigate as they transition through
reproductive states are primarily limited to a number of circulating
hormones and a small set of well-known genes (cite). We have yet to come
close to conceptualizing the larger symphony of reproductive events that
occur to yield reproductive and parental care behaviors.

Using the model of a biparental vertebrate species, the rock dove
(Columba livia), our team has previously demonstrated the dynamic nature
of gene activity within the reproductive or HPG axis.
\citet{MacManes2017} reported that HPG gene activity varied by sex, with
females exhibiting higher levels of overall gene expression than males.
\citet{Calisi2018} reported that HPG gene activity varied response to
restraint-stress, and Austin et al.~identify genes that responds to both
corticosterone treatment and restraint stress. While each of these
studies advanced our understanding of sex-differences and
transcriptional plasticity in the reproducitve axis, none of these
studies examined activity of the reproducitve axis of bird that are
actively reproducing or caring for offspring. Studying parental care in
both sexes of mammalian species is challenging because only the females
are able to use lactation to feed their young, and the physiological
mechanisms regulating this process are hard to control for. The rock
dove is an excellent system for studying reproduction and parental care
in both sexes because both males and females feed their young an
energy-rich secretion produced from the epithelial lining of their
crops, known as crop milk.

In this study, we investigate the activity of the HPG axis over the
course of reproduction and parental care by characterizing gene
expression as birds transition from non-breeding to nest-building, as
they lay and incubate their clutches, and finally as they become parents
and raise their chicks. Then, we conducted a series of nest
manipulations to ascertain constraints associated with biological timing
versus input from external stimuli. We conducted a comprehensive series
of analyses examine broad patterns of variation as well as the activity
of specific genes in specific tissues. We have also made our data and
scripts available so that others can reuse the data to ask different
questions or reproduce our results. This research provides a deeper
understanding of the interplay of genes, hormones, and parental care and
has important implications for understanding molecular processes that
underlie behavioral transitions. \citep[reviewed in][]{Lea1989} say
this, but other say that \citep[reviewed in][]{Dulac2014}.

\hypertarget{results}{%
\subsection{Results}\label{results}}

To characterize transcriptional activity, we first examined broad
patterns of variation in gene expression in the hypothalamus, pituitary,
and gonads of male and female rock doves across a series of timepoints
that encapsulate important transitions in reproduction and and parental
care (Fig. \ref{figure:fig1}a). t-Distributed Stochastic Neighbor
Embedding (t-SNE) is a nonlinear, dimension reduction technique is used
to visualize high-dimensional data by giving each datapoint a location
in a two-dimensional map (van der Maaten and Hinton 2008). t-SNE is
widely using in single-cell data process for identify related cell
types. Here, we use t-SNE to identify variation associated with tissue,
sex, and treatment (Fig. \ref{figure:fig1}j).

\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions. a) Experimental design. b) t-SNE plot highlighting tissue differences in gene expression. Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple). c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males). t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orange, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
  \label{figure:fig1}
\end{figure}

Our results suggest that tissue is the greatest source of variation in
the dataset, with hypothalamic, pituitary, and gonadal samples all
forming distinct clusters, with only a few pituitary and gonad samples
falling outside the 95\% confidence interval (Fig. \ref{figure:fig1}b).
Sex-differences are most prounounced in the gonads with male and female
samples form two discrete clusters (Fig. \ref{figure:fig1}c). Treatment
accounts for much less of the variation than tissue and sex, and even
after removing sex as a variatble, treatment effects in all three
tissues are hard to visualize simultatenously (Fig.
\ref{figure:fig1}d,e), so we next looked at each tissue and sex
separately. In the hypothalamus, control samples form tight clusters
that stands out from all other timepoints that are largely overlapping
in space (Fig. \ref{figure:fig1}h). The pituitary shows the strongest
treatment effect with one cluster of the ``earlier'' parental timepoints
(nest-building, lay, incubation days 3 and 9), a second cluster of
``later'' parental timpoints (incubation day 17, hatch, and nestling
care days 5 and 9; Fig. \ref{figure:fig1}i). Like the hypothalamus, in
the gonads, the control timepoints appear to be much less variable that
the parental timepoints, but there is no clear separation of gonadal
samples by parental stage (Fig. \ref{figure:fig1}j). While the tSNE
analysis reveals broad patterns across all levels of the experiment,
additional tools are needed it hone in on the specific differences
associated with parental care in each tissue.

Differential gene expression analysis can shed insight into specific
differences between treatment groups and thus deepen our understand of
the transitions between reproduction and parental care behaviors. We
used DESEq2 to identify genes whose expression changed significantly
between sequential timepoints in each tissue and sex separately (Fig.
\ref{figure:fig2}a).

\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth]{../../figures/fig2-1}
  \caption{Differential gene expression differences between sequential parental stages. a) We calculated the differentail gene expression between each consequtive parental stage samples, for a total of 8 pair-wise comparisons for each sex in each tissue. The y-axis of these bar plots displays the total number of differentially expressed genes for each comparsison in the hypothalamus (b), pituitary (c), gonads (d). Bars are collored by the direction of the change in gene expression. Control: control: red or , nest-building (bldg): orange, lay: yellow; incubation day 3 (inc.d3): light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, nestling care day 5 (n5): blue, and n9: light blue.}
  \label{figure:fig2}
\end{figure}

In the hypothalamus only two sequential stage comparisons results in a
sizable change in expression; in females, 1927 genes changes their
expression betwen hatch to nesting care day 5 (with 1076 decreasing and
851 increased their expression), and in the male hypothalamus,87 genes
were diferentially expressed bewteen incubation days 9 and 17 (with 38
decreasing and 49 increasing their expression (Fig.
\ref{figure:fig2}b)). In the pituitary, differential gene expression
differenes are greatest when compariing incubation days 9 and 17, and
this difference is more pronounced in females (Fig. \ref{figure:fig2}c).
The next largest difference in the pituitary is between hatch and n5.
The female putitary shows moderate chagnes from nest-bilding to lay to
incuabation day 3 (Fig. \ref{figure:fig2}c). The male pituitary changes
more that the female pitutiary changes between nestling care days 5 and
9 (Fig. \ref{figure:fig2}c). About 2000 genes also change their
expression in the female pituitary and about 425 in the male pituitary
when comparing incubation days 9 and 17 (Fig. \ref{figure:fig2}). In the
female gonads, 270 genes are decreased while only 56 are increased when
comparing incubation days 9 and 17(Fig. 2). Many more significant
changes in gene expression in the female gonad occur between
nest-building and lay and between lay and clutch completion (incubation
day 3) than in any other comparison (Fig. \ref{figure:fig2}).

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\hypertarget{rnaseq}{%
\subsubsection{RNAseq}\label{rnaseq}}

See \citet{Lang2019} for details RNAseq.

Differential gene expression conducted wtih d Binder
\citep{project_jupyter-proc-scipy-2018, Love2014, Wilke2016}.

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

This file is only a basic article template. For full details of
\emph{The R Journal} style and information on how to prepare your
article for submission, see the
\href{https://journal.r-project.org/share/author-guide.pdf}{Instructions
for Authors}.

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

Thanks to Owen for discussion of data sonfification. Thanks to DIB lab
members for software and methods discussion (Titus, Taylor, Luiz,
Amanda, Alicia, Picaso).

\hypertarget{references}{%
\subsection{References}\label{references}}

\bibliography{RJreferences}

\address{%
Rayna M Harris \footnote{equal contribtion}\\
UC Davis\\
\\
}

\address{%
Suzanne H. Austin \footnote{equal contribtion}\\
UC Davis\\
\\
}

\address{%
Andrew S. Lang\\
UNH\\
\\
}

\address{%
Victoria S. Farrar\\
UC Davis\\
\\
}

\address{%
April Booth\\
UC Davis\\
\\
}

\address{%
Tanner Feustel\\
UC Davis\\
\\
}

\address{%
Matthew D. MacManes\\
UNH\\
\\
}

\address{%
Rebecca M. Calisi \footnote{corresponding author}\\
UC Davis\\
\\
}
\href{mailto:author7@work}{\nolinkurl{author7@work}}

### 2020-03-06 analysis/figure-5.Rmd
save files
now with manuscritp draft

@@
   mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
-  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2))
+
+allDEG <- allDEG %>%
 dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj)
-head(allDEG)
 
-comparisonlevels <- c( "inc.d9_m.inc.d8", "inc.d17_prolong", "hatch_extend" , 
-                       "inc.d3_m.inc.d3" ,  "inc.d9_m.inc.d9", 
-                        "inc.d17_m.inc.d17", "m.n2_hatch" 
-                        )
+allDEG %>% distinct(comparison)
+
+comparisonlevels <-  c(	 "inc.d3_m.inc.d3",			
+                             "inc.d9_m.inc.d9",	
+                             "inc.d17_m.inc.d17",		
+                           "hatch_m.n2",
+
+                           "inc.d9_m.inc.d8",
+
+                         "m.inc.d9_m.n2",
+                         "m.inc.d9_m.inc.d8",
+
+                         "m.inc.d8_extend",
+                         "prolong_extend",
+                           "hatch_extend",
+                         "m.n2_extend",
+
+                         "m.inc.d8_prolong",
+                            "inc.d17_prolong",	
+                           "m.inc.d17_prolong",
+
+                           "m.inc.d3_m.inc.d9",
+                           "m.inc.d3_m.inc.d17",
+                         "m.inc.d3_m.n2",
+                         "m.inc.d9_m.inc.d17")
 
@@
   filter(comparison %in% c("inc.d3_m.inc.d3" ,  "inc.d9_m.inc.d9", 
-                        "inc.d17_m.inc.d17", "m.n2_hatch" ))  %>%
+                        "inc.d17_m.inc.d17", "hatch_m.n2" ))  %>%
 ggplot(aes(x = comparison,  fill = direction)) +
@@
 
-fig5 <- plot_grid(expdesign, barplots, mylegend, ncol =1, rel_heights = c(0.25,1,0.1))
+fig5 <- plot_grid(expdesign, barplots, ncol =1, rel_heights = c(0.5,1))
 fig5
+
 ```
+
+```{r writefiles}
+write_csv(allDEG, "../results/DEGs-manip.csv")
+```
+
+## subset. examine for E, T, hyp, and gonad hypotheses
+
+```{r subset}
+
+subsetDEGs <- function(whichsex, whichtissue, whichcomparison) {
+df <- allDEG %>%
+  filter(sex %in% whichsex,
+         tissue %in% whichtissue,
+         comparison %in% whichcomparison)  %>%
+  group_by(sex, tissue, comparison) #%>%
+  #summarize(DEGs = n()) %>%
+  #arrange(desc(DEGs)) %>%
+  #drop_na()
+print(df)
+}
+
+subsetDEGs(c("male","female"), c("hypothalamus", "pituitary", "gonad"), comparisonlevels)
+fhyphatchext <- subsetDEGs("female", "hypothalamus", "hatch_extend") %>% arrange (gene)
+mhyphatchext <- subsetDEGs("male", "hypothalamus", "hatch_extend") %>% arrange (gene)
+
+ggplot(fhyphatchext, aes(x = lfc, y = logpadj)) + geom_point() + xlim(-5,5) + ylim(0,7)
+ggplot(mhyphatchext, aes(x = lfc, y = logpadj)) + geom_point() + xlim(-5,5) + ylim(0,7)
+
+```

### 2020-03-06 docs/manuscript/elife_v1.Rmd
now with manuscritp draft

--- 2020-03-06 docs/manuscript/elife-v1.tex
@@
-% !TeX root = RJwrapper.tex
-\title{Internal physiology and external cues drive changes in reproductive axis
-of parenting pigeons}
-\author{by Rayna M Harris \footnote{equal contribtion}, Suzanne H.
-Austin \footnote{equal contribtion}, Andrew S.
-Lang, Victoria S.
-Farrar, April Booth, Tanner Feustel, Matthew D.
-MacManes, Rebecca M.
-Calisi \footnote{corresponding author}}
+---
+title: Internal physiology and external cues drive changes in reproductive axis of parenting pigeons
+author:
+  - name: Rayna M Harris ^[equal contribtion] 
+    affiliation: UC Davis
+  - name: Suzanne H. Austin  ^[equal contribtion] 
+    affiliation: UC Davis
+  - name: Andrew S. Lang
+    affiliation: UNH
+  - name: Victoria S. Farrar
+    affiliation: UC Davis
+  - name: April Booth
+    affiliation: UC Davis
+  - name: Tanner Feustel
+    affiliation: UC Davis
+  - name: Matthew D. MacManes
+    affiliation: UNH
+  - name: Rebecca M. Calisi  ^[corresponding author] 
+    affiliation: UC Davis
+    email:  author7@work   
+address:
+  - code: UC Davis
+    address: Department, Street, City, CA, Zip
+  - code: UNH
+    address: Department, Street, City, NH, Zip    
+abstract: |
+output:
+  rticles::rjournal_article:
+    includes:
+      in_header: preamble.tex
 
-\maketitle
+---
 
-% Any extra LaTeX you need in the preamble
+## Introduction
 
-In species that provide parental care, successful rearing of offspring involves a shift from aggressive and sexual behaviors to more caring and nurturing ones.
-We hypothesize that these shifts in behavior are related to changes in gene expression that are controlled by internal mechanisms rather than by the external cues.
-We found that gene expression in the hypothalamus is fairly insensitive to the natural transition between egg and nestling stages; however, in the pituitary and gonads, gene expression showed much more plasticity over the course of reproduction.
-We show evidence that natural transitions in reproductive and parental behavior by mediated internal clock controlling gene expression in the pitary and gonads but that the hypothalamus is very sensitive to external cues that signal a distruption in the natural parental care cycle.
-This research provides a deeper understanding of the interplay of genes, hormones, and parental care and has important implications for understanding molecular processes that underlie behavior transitions.
+Reproduction is a critical period in an organism’s life cycle.
+Parents must balance their own needs with the energetic costs and time-constraints required to successfully reproduce and rear their offspring to independence(1).
+A complex series of evolutionarily conserved physiological processes underlie behaviors typically associated with reproduction in vertebrates, including courtship, nesting, and parental care.
+These processes are primarily mediated by the reproductive axis, which consists of the hypothalamus in the brain, the pituitary, and the gonads (i.e., the HPG).
+Our understanding of the physiological changes that adults navigate as they transition through reproductive states are primarily limited to a number of circulating hormones and a small set of well-known genes (cite).
+We have yet to come close to conceptualizing the larger symphony of reproductive events that occur to yield reproductive and parental care behaviors.
 
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+Using the model of a biparental vertebrate species, the rock dove (Columba livia), our team has previously demonstrated the dynamic nature of gene activity within the reproductive or HPG axis.
+\citet{MacManes2017} reported that HPG gene activity varied by sex, with females exhibiting higher levels of overall gene expression than males.
+\citet{Calisi2018} reported that HPG gene activity varied response to restraint-stress, and Austin et al.
+identify genes that responds to both corticosterone treatment and restraint stress.
+While each of these studies advanced our understanding of sex-differences and transcriptional plasticity in the reproducitve axis, none of these studies examined activity of the reproducitve axis of bird that are actively reproducing or caring for offspring.
+Studying parental care in both sexes of mammalian species is challenging because only the females are able to use lactation to feed their young, and the physiological mechanisms regulating this process are hard to control for.
+The rock dove is an excellent system for studying reproduction and parental care in both sexes because both males and females feed their young an energy-rich secretion produced from the epithelial lining of their crops, known as crop milk.
 
-Reproduction is a critical period in an organism's life cycle. Parents
-must balance their own needs with the energetic costs and
-time-constraints required to successfully reproduce and rear their
-offspring to independence(1). A complex series of evolutionarily
-conserved physiological processes underlie behaviors typically
-associated with reproduction in vertebrates, including courtship,
-nesting, and parental care. These processes are primarily mediated by
-the reproductive axis, which consists of the hypothalamus in the brain,
-the pituitary, and the gonads (i.e., the HPG). Our understanding of the
-physiological changes that adults navigate as they transition through
-reproductive states are primarily limited to a number of circulating
-hormones and a small set of well-known genes (cite). We have yet to come
-close to conceptualizing the larger symphony of reproductive events that
-occur to yield reproductive and parental care behaviors.
+In this study, we investigate the activity of the HPG axis over the course of reproduction and parental care by characterizing gene expression as birds transition from non-breeding to nest-building, as they lay and incubate their clutches, and finally as they become parents and raise their chicks.
+Then, we conducted a series of nest manipulations to ascertain constraints associated with biological timing versus input from external stimuli.
+We conducted a comprehensive series of analyses examine broad patterns of variation as well as the activity of specific genes in specific tissues.
+We have also made our data and scripts available so that others can reuse the data to ask different questions or reproduce our results.
+This research provides a deeper understanding of the interplay of genes, hormones, and parental care and has important implications for understanding molecular processes that underlie behavioral transitions.
+\citep[reviewed in][]{Lea1989} say this, but other say that \citep[reviewed in][]{Dulac2014}.
 
-Using the model of a biparental vertebrate species, the rock dove
-(Columba livia), our team has previously demonstrated the dynamic nature
-of gene activity within the reproductive or HPG axis.
-\citet{MacManes2017} reported that HPG gene activity varied by sex, with
-females exhibiting higher levels of overall gene expression than males.
-\citet{Calisi2018} reported that HPG gene activity varied response to
-restraint-stress, and Austin et al.~identify genes that responds to both
-corticosterone treatment and restraint stress. While each of these
-studies advanced our understanding of sex-differences and
-transcriptional plasticity in the reproducitve axis, none of these
-studies examined activity of the reproducitve axis of bird that are
-actively reproducing or caring for offspring. Studying parental care in
-both sexes of mammalian species is challenging because only the females
-are able to use lactation to feed their young, and the physiological
-mechanisms regulating this process are hard to control for. The rock
-dove is an excellent system for studying reproduction and parental care
-in both sexes because both males and females feed their young an
-energy-rich secretion produced from the epithelial lining of their
-crops, known as crop milk.
+## Results
 
-In this study, we investigate the activity of the HPG axis over the
-course of reproduction and parental care by characterizing gene
-expression as birds transition from non-breeding to nest-building, as
-they lay and incubate their clutches, and finally as they become parents
-and raise their chicks. Then, we conducted a series of nest
-manipulations to ascertain constraints associated with biological timing
-versus input from external stimuli. We conducted a comprehensive series
-of analyses examine broad patterns of variation as well as the activity
-of specific genes in specific tissues. We have also made our data and
-scripts available so that others can reuse the data to ask different
-questions or reproduce our results. This research provides a deeper
-understanding of the interplay of genes, hormones, and parental care and
-has important implications for understanding molecular processes that
-underlie behavioral transitions. \citep[reviewed in][]{Lea1989} say
-this, but other say that \citep[reviewed in][]{Dulac2014}.
-
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-To characterize transcriptional activity, we first examined broad
-patterns of variation in gene expression in the hypothalamus, pituitary,
-and gonads of male and female rock doves across a series of timepoints
-that encapsulate important transitions in reproduction and and parental
-care (Fig. \ref{figure:fig1}a). t-Distributed Stochastic Neighbor
-Embedding (t-SNE) is a nonlinear, dimension reduction technique is used
-to visualize high-dimensional data by giving each datapoint a location
-in a two-dimensional map (van der Maaten and Hinton 2008). t-SNE is
-widely using in single-cell data process for identify related cell
-types. Here, we use t-SNE to identify variation associated with tissue,
-sex, and treatment (Fig. \ref{figure:fig1}j).
+To characterize transcriptional activity, we first examined broad patterns of variation in gene expression in the hypothalamus, pituitary, and gonads of male and female rock doves across a series of timepoints that encapsulate important transitions in reproduction and and parental care (Fig.
+\ref{figure:fig1}a).
+t-Distributed Stochastic Neighbor Embedding (t-SNE) is a nonlinear, dimension reduction technique is used to visualize high-dimensional data by giving each datapoint a location in a two-dimensional map (van der Maaten and Hinton 2008).
+t-SNE is widely using in single-cell data process for identify related cell types.
+Here, we use t-SNE to identify variation associated with tissue, sex, and treatment (Fig.
+\ref{figure:fig1}j).
 
@@
 
-Our results suggest that tissue is the greatest source of variation in
-the dataset, with hypothalamic, pituitary, and gonadal samples all
-forming distinct clusters, with only a few pituitary and gonad samples
-falling outside the 95\% confidence interval (Fig. \ref{figure:fig1}b).
-Sex-differences are most prounounced in the gonads with male and female
-samples form two discrete clusters (Fig. \ref{figure:fig1}c). Treatment
-accounts for much less of the variation than tissue and sex, and even
-after removing sex as a variatble, treatment effects in all three
-tissues are hard to visualize simultatenously (Fig.
-\ref{figure:fig1}d,e), so we next looked at each tissue and sex
-separately. In the hypothalamus, control samples form tight clusters
-that stands out from all other timepoints that are largely overlapping
-in space (Fig. \ref{figure:fig1}h). The pituitary shows the strongest
-treatment effect with one cluster of the ``earlier'' parental timepoints
-(nest-building, lay, incubation days 3 and 9), a second cluster of
-``later'' parental timpoints (incubation day 17, hatch, and nestling
-care days 5 and 9; Fig. \ref{figure:fig1}i). Like the hypothalamus, in
-the gonads, the control timepoints appear to be much less variable that
-the parental timepoints, but there is no clear separation of gonadal
-samples by parental stage (Fig. \ref{figure:fig1}j). While the tSNE
-analysis reveals broad patterns across all levels of the experiment,
-additional tools are needed it hone in on the specific differences
-associated with parental care in each tissue.
+Our results suggest that tissue is the greatest source of variation in the dataset, with hypothalamic, pituitary, and gonadal samples all forming distinct clusters, with only a few pituitary and gonad samples falling outside the 95% confidence interval (Fig.
+\ref{figure:fig1}b).
+Sex-differences are most prounounced in the gonads with male and female samples form two discrete clusters (Fig.
+\ref{figure:fig1}c).
+Treatment accounts for much less of the variation than tissue and sex, and even after removing sex as a variatble, treatment effects in all three tissues are hard to visualize simultatenously (Fig.
+\ref{figure:fig1}d,e), so we next looked at each tissue and sex separately.
+In the hypothalamus, control samples form tight clusters that stands out from all other timepoints that are largely overlapping in space (Fig.
+\ref{figure:fig1}h).
+The pituitary shows the strongest treatment effect with one cluster of the “earlier” parental timepoints (nest-building, lay, incubation days 3 and 9), a second cluster of “later” parental timpoints (incubation day 17, hatch, and nestling care days 5 and 9; Fig.
+\ref{figure:fig1}i).
+Like the hypothalamus, in the gonads, the control timepoints appear to be much less variable that the parental timepoints, but there is no clear separation of gonadal samples by parental stage (Fig.
+\ref{figure:fig1}j).
+While the tSNE analysis reveals broad patterns across all levels of the experiment, additional tools are needed it hone in on the specific differences associated with parental care in each tissue.
 
-Differential gene expression analysis can shed insight into specific
-differences between treatment groups and thus deepen our understand of
-the transitions between reproduction and parental care behaviors. We
-used DESEq2 to identify genes whose expression changed significantly
-between sequential timepoints in each tissue and sex separately (Fig.
+Differential gene expression analysis can shed insight into specific differences between treatment groups and thus deepen our understand of the transitions between reproduction and parental care behaviors.
+We used DESEq2 to identify genes whose expression changed significantly between sequential timepoints in each tissue and sex separately (Fig.
 \ref{figure:fig2}a).
@@
 
-In the hypothalamus only two sequential stage comparisons results in a
-sizable change in expression; in females, 1927 genes changes their
-expression betwen hatch to nesting care day 5 (with 1076 decreasing and
-851 increased their expression), and in the male hypothalamus,87 genes
-were diferentially expressed bewteen incubation days 9 and 17 (with 38
-decreasing and 49 increasing their expression (Fig.
-\ref{figure:fig2}b)). In the pituitary, differential gene expression
-differenes are greatest when compariing incubation days 9 and 17, and
-this difference is more pronounced in females (Fig. \ref{figure:fig2}c).
+In the hypothalamus only two sequential stage comparisons results in a sizable change in expression; in females, 1927 genes changes their expression betwen hatch to nesting care day 5 (with 1076 decreasing and 851 increased their expression), and in the male hypothalamus,87 genes were diferentially expressed bewteen incubation days 9 and 17 (with 38 decreasing and 49 increasing their expression (Fig.
+\ref{figure:fig2}b)).
+In the pituitary, differential gene expression differenes are greatest when compariing incubation days 9 and 17, and this difference is more pronounced in females (Fig.
+\ref{figure:fig2}c).
 The next largest difference in the pituitary is between hatch and n5.
-The female putitary shows moderate chagnes from nest-bilding to lay to
-incuabation day 3 (Fig. \ref{figure:fig2}c). The male pituitary changes
-more that the female pitutiary changes between nestling care days 5 and
-9 (Fig. \ref{figure:fig2}c). About 2000 genes also change their
-expression in the female pituitary and about 425 in the male pituitary
-when comparing incubation days 9 and 17 (Fig. \ref{figure:fig2}). In the
-female gonads, 270 genes are decreased while only 56 are increased when
-comparing incubation days 9 and 17(Fig. 2). Many more significant
-changes in gene expression in the female gonad occur between
-nest-building and lay and between lay and clutch completion (incubation
-day 3) than in any other comparison (Fig. \ref{figure:fig2}).
+The female putitary shows moderate chagnes from nest-bilding to lay to incuabation day 3 (Fig.
+\ref{figure:fig2}c).
+The male pituitary changes more that the female pitutiary changes between nestling care days 5 and 9  (Fig.
+\ref{figure:fig2}c).
+About 2000 genes also change their expression in the female pituitary and about 425 in the male pituitary when comparing incubation days 9 and 17 (Fig.
+\ref{figure:fig2}).
+In the female gonads, 270 genes are decreased while only 56 are increased when comparing incubation days 9 and 17(Fig.
+2).
+Many more significant changes in gene expression in the female gonad occur between nest-building and lay and between lay and clutch completion (incubation day 3) than in any other comparison (Fig.
+\ref{figure:fig2}).
 
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
+## Discussion
 
-\hypertarget{materials-and-methods}{%
-\subsection{Materials and Methods}\label{materials-and-methods}}
+## Materials and Methods
 
-\hypertarget{rnaseq}{%
-\subsubsection{RNAseq}\label{rnaseq}}
+### RNAseq
 
@@
 
-Differential gene expression conducted wtih d Binder
-\citep{project_jupyter-proc-scipy-2018, Love2014, Wilke2016}.
+Differential gene expression conducted wtih d Binder \citep{project_jupyter-proc-scipy-2018, Love2014, Wilke2016}.
 
-\hypertarget{summary}{%
-\subsection{Summary}\label{summary}}
+## Summary
 
-This file is only a basic article template. For full details of
-\emph{The R Journal} style and information on how to prepare your
-article for submission, see the
-\href{https://journal.r-project.org/share/author-guide.pdf}{Instructions
-for Authors}.
+This file is only a basic article template.
+For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
 
-\hypertarget{acknowledgements}{%
-\subsection{Acknowledgements}\label{acknowledgements}}
+## Acknowledgements
 
-Thanks to Owen for discussion of data sonfification. Thanks to DIB lab
-members for software and methods discussion (Titus, Taylor, Luiz,
-Amanda, Alicia, Picaso).
+Thanks to Owen for discussion of data sonfification. Thanks to DIB lab members for software and methods discussion (Titus, Taylor, Luiz, Amanda, Alicia, Picaso).
 
-\hypertarget{references}{%
-\subsection{References}\label{references}}
+## References
 
@@
 
-\address{%
-Rayna M Harris \footnote{equal contribtion}\\
-UC Davis\\
-\\
-}
-
-\address{%
-Suzanne H. Austin \footnote{equal contribtion}\\
-UC Davis\\
-\\
-}
-
-\address{%
-Andrew S. Lang\\
-UNH\\
-\\
-}
-
-\address{%
-Victoria S. Farrar\\
-UC Davis\\
-\\
-}
-
-\address{%
-April Booth\\
-UC Davis\\
-\\
-}
-
-\address{%
-Tanner Feustel\\
-UC Davis\\
-\\
-}
-
-\address{%
-Matthew D. MacManes\\
-UNH\\
-\\
-}
-
-\address{%
-Rebecca M. Calisi \footnote{corresponding author}\\
-UC Davis\\
-\\
-}
-\href{mailto:author7@work}{\nolinkurl{author7@work}}
+<!-- 
+Introductory section which may include references in parentheses
+\citep{RDevelopmentCoreTeam2013}, or cite a reference such as \citet{RDevelopmentCoreTeam2013} in the text.
+-->

### 2020-03-06 docs/manuscript/preamble.tex
now with manuscritp draft

--- 2020-03-06 docs/manuscript/elife-v1.tex
@@
-% !TeX root = RJwrapper.tex
-\title{Internal physiology and external cues drive changes in reproductive axis
-of parenting pigeons}
-\author{by Rayna M Harris \footnote{equal contribtion}, Suzanne H.
-Austin \footnote{equal contribtion}, Andrew S.
-Lang, Victoria S.
-Farrar, April Booth, Tanner Feustel, Matthew D.
-MacManes, Rebecca M.
-Calisi \footnote{corresponding author}}
-
-\maketitle
-
 % Any extra LaTeX you need in the preamble
@@
 This research provides a deeper understanding of the interplay of genes, hormones, and parental care and has important implications for understanding molecular processes that underlie behavior transitions.
-
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
-
-Reproduction is a critical period in an organism's life cycle. Parents
-must balance their own needs with the energetic costs and
-time-constraints required to successfully reproduce and rear their
-offspring to independence(1). A complex series of evolutionarily
-conserved physiological processes underlie behaviors typically
-associated with reproduction in vertebrates, including courtship,
-nesting, and parental care. These processes are primarily mediated by
-the reproductive axis, which consists of the hypothalamus in the brain,
-the pituitary, and the gonads (i.e., the HPG). Our understanding of the
-physiological changes that adults navigate as they transition through
-reproductive states are primarily limited to a number of circulating
-hormones and a small set of well-known genes (cite). We have yet to come
-close to conceptualizing the larger symphony of reproductive events that
-occur to yield reproductive and parental care behaviors.
-
-Using the model of a biparental vertebrate species, the rock dove
-(Columba livia), our team has previously demonstrated the dynamic nature
-of gene activity within the reproductive or HPG axis.
-\citet{MacManes2017} reported that HPG gene activity varied by sex, with
-females exhibiting higher levels of overall gene expression than males.
-\citet{Calisi2018} reported that HPG gene activity varied response to
-restraint-stress, and Austin et al.~identify genes that responds to both
-corticosterone treatment and restraint stress. While each of these
-studies advanced our understanding of sex-differences and
-transcriptional plasticity in the reproducitve axis, none of these
-studies examined activity of the reproducitve axis of bird that are
-actively reproducing or caring for offspring. Studying parental care in
-both sexes of mammalian species is challenging because only the females
-are able to use lactation to feed their young, and the physiological
-mechanisms regulating this process are hard to control for. The rock
-dove is an excellent system for studying reproduction and parental care
-in both sexes because both males and females feed their young an
-energy-rich secretion produced from the epithelial lining of their
-crops, known as crop milk.
-
-In this study, we investigate the activity of the HPG axis over the
-course of reproduction and parental care by characterizing gene
-expression as birds transition from non-breeding to nest-building, as
-they lay and incubate their clutches, and finally as they become parents
-and raise their chicks. Then, we conducted a series of nest
-manipulations to ascertain constraints associated with biological timing
-versus input from external stimuli. We conducted a comprehensive series
-of analyses examine broad patterns of variation as well as the activity
-of specific genes in specific tissues. We have also made our data and
-scripts available so that others can reuse the data to ask different
-questions or reproduce our results. This research provides a deeper
-understanding of the interplay of genes, hormones, and parental care and
-has important implications for understanding molecular processes that
-underlie behavioral transitions. \citep[reviewed in][]{Lea1989} say
-this, but other say that \citep[reviewed in][]{Dulac2014}.
-
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-To characterize transcriptional activity, we first examined broad
-patterns of variation in gene expression in the hypothalamus, pituitary,
-and gonads of male and female rock doves across a series of timepoints
-that encapsulate important transitions in reproduction and and parental
-care (Fig. \ref{figure:fig1}a). t-Distributed Stochastic Neighbor
-Embedding (t-SNE) is a nonlinear, dimension reduction technique is used
-to visualize high-dimensional data by giving each datapoint a location
-in a two-dimensional map (van der Maaten and Hinton 2008). t-SNE is
-widely using in single-cell data process for identify related cell
-types. Here, we use t-SNE to identify variation associated with tissue,
-sex, and treatment (Fig. \ref{figure:fig1}j).
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
-  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.
-a) Experimental design.
-b) t-SNE plot highlighting tissue differences in gene expression.
-Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple).
-c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males).
-t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orange, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
-  \label{figure:fig1}
-\end{figure}
-
-Our results suggest that tissue is the greatest source of variation in
-the dataset, with hypothalamic, pituitary, and gonadal samples all
-forming distinct clusters, with only a few pituitary and gonad samples
-falling outside the 95\% confidence interval (Fig. \ref{figure:fig1}b).
-Sex-differences are most prounounced in the gonads with male and female
-samples form two discrete clusters (Fig. \ref{figure:fig1}c). Treatment
-accounts for much less of the variation than tissue and sex, and even
-after removing sex as a variatble, treatment effects in all three
-tissues are hard to visualize simultatenously (Fig.
-\ref{figure:fig1}d,e), so we next looked at each tissue and sex
-separately. In the hypothalamus, control samples form tight clusters
-that stands out from all other timepoints that are largely overlapping
-in space (Fig. \ref{figure:fig1}h). The pituitary shows the strongest
-treatment effect with one cluster of the ``earlier'' parental timepoints
-(nest-building, lay, incubation days 3 and 9), a second cluster of
-``later'' parental timpoints (incubation day 17, hatch, and nestling
-care days 5 and 9; Fig. \ref{figure:fig1}i). Like the hypothalamus, in
-the gonads, the control timepoints appear to be much less variable that
-the parental timepoints, but there is no clear separation of gonadal
-samples by parental stage (Fig. \ref{figure:fig1}j). While the tSNE
-analysis reveals broad patterns across all levels of the experiment,
-additional tools are needed it hone in on the specific differences
-associated with parental care in each tissue.
-
-Differential gene expression analysis can shed insight into specific
-differences between treatment groups and thus deepen our understand of
-the transitions between reproduction and parental care behaviors. We
-used DESEq2 to identify genes whose expression changed significantly
-between sequential timepoints in each tissue and sex separately (Fig.
-\ref{figure:fig2}a).
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig2-1}
-  \caption{Differential gene expression differences between sequential parental stages.
-a) We calculated the differentail gene expression between each consequtive parental stage samples, for a total of 8 pair-wise comparisons for each sex in each tissue.
-The y-axis of these bar plots displays the total number of differentially expressed genes for each comparsison in the hypothalamus (b), pituitary (c), gonads (d).
-Bars are collored by the direction of the change in gene expression.
-Control: control: red or , nest-building (bldg): orange, lay: yellow; incubation day 3 (inc.d3): light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, nestling care day 5 (n5): blue, and n9: light blue.}
-  \label{figure:fig2}
-\end{figure}
-
-In the hypothalamus only two sequential stage comparisons results in a
-sizable change in expression; in females, 1927 genes changes their
-expression betwen hatch to nesting care day 5 (with 1076 decreasing and
-851 increased their expression), and in the male hypothalamus,87 genes
-were diferentially expressed bewteen incubation days 9 and 17 (with 38
-decreasing and 49 increasing their expression (Fig.
-\ref{figure:fig2}b)). In the pituitary, differential gene expression
-differenes are greatest when compariing incubation days 9 and 17, and
-this difference is more pronounced in females (Fig. \ref{figure:fig2}c).
-The next largest difference in the pituitary is between hatch and n5.
-The female putitary shows moderate chagnes from nest-bilding to lay to
-incuabation day 3 (Fig. \ref{figure:fig2}c). The male pituitary changes
-more that the female pitutiary changes between nestling care days 5 and
-9 (Fig. \ref{figure:fig2}c). About 2000 genes also change their
-expression in the female pituitary and about 425 in the male pituitary
-when comparing incubation days 9 and 17 (Fig. \ref{figure:fig2}). In the
-female gonads, 270 genes are decreased while only 56 are increased when
-comparing incubation days 9 and 17(Fig. 2). Many more significant
-changes in gene expression in the female gonad occur between
-nest-building and lay and between lay and clutch completion (incubation
-day 3) than in any other comparison (Fig. \ref{figure:fig2}).
-
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
-
-\hypertarget{materials-and-methods}{%
-\subsection{Materials and Methods}\label{materials-and-methods}}
-
-\hypertarget{rnaseq}{%
-\subsubsection{RNAseq}\label{rnaseq}}
-
-See \citet{Lang2019} for details RNAseq.
-
-Differential gene expression conducted wtih d Binder
-\citep{project_jupyter-proc-scipy-2018, Love2014, Wilke2016}.
-
-\hypertarget{summary}{%
-\subsection{Summary}\label{summary}}
-
-This file is only a basic article template. For full details of
-\emph{The R Journal} style and information on how to prepare your
-article for submission, see the
-\href{https://journal.r-project.org/share/author-guide.pdf}{Instructions
-for Authors}.
-
-\hypertarget{acknowledgements}{%
-\subsection{Acknowledgements}\label{acknowledgements}}
-
-Thanks to Owen for discussion of data sonfification. Thanks to DIB lab
-members for software and methods discussion (Titus, Taylor, Luiz,
-Amanda, Alicia, Picaso).
-
-\hypertarget{references}{%
-\subsection{References}\label{references}}
-
-\bibliography{RJreferences}
-
-\address{%
-Rayna M Harris \footnote{equal contribtion}\\
-UC Davis\\
-\\
-}
-
-\address{%
-Suzanne H. Austin \footnote{equal contribtion}\\
-UC Davis\\
-\\
-}
-
-\address{%
-Andrew S. Lang\\
-UNH\\
-\\
-}
-
-\address{%
-Victoria S. Farrar\\
-UC Davis\\
-\\
-}
-
-\address{%
-April Booth\\
-UC Davis\\
-\\
-}
-
-\address{%
-Tanner Feustel\\
-UC Davis\\
-\\
-}
-
-\address{%
-Matthew D. MacManes\\
-UNH\\
-\\
-}
-
-\address{%
-Rebecca M. Calisi \footnote{corresponding author}\\
-UC Davis\\
-\\
-}
-\href{mailto:author7@work}{\nolinkurl{author7@work}}

### 2020-03-06 docs/manuscript/elife_v1.Rmd
more pub stuff

@@
       in_header: preamble.tex
-
 ---
@@
 Reproduction is a critical period in an organism’s life cycle.
-Parents must balance their own needs with the energetic costs and time-constraints required to successfully reproduce and rear their offspring to independence(1).
+Parents must balance their own needs with the energetic costs and time-constraints required to successfully reproduce and rear their offspring to independence (cite).
 A complex series of evolutionarily conserved physiological processes underlie behaviors typically associated with reproduction in vertebrates, including courtship, nesting, and parental care.
@@
 
-### RNAseq
+### Animal Care
 
-See \citet{Lang2019} for details RNAseq.
+All animal care and use comply with the Public Health Service Policy on Humane Care and Use of Laboratory Animals and were approved by the University of California, Davis IACUC permit #18895.
+Birds were housed at the University of California, Davis in enclosed aviaries with ~8 sexually reproductive adult pairs per aviary.
+All adults birds are uniquely associated with nests by their unique color band combination.
+Food and water were provided ad libitum, and nests were monitored daily.
 
-Differential gene expression conducted wtih d Binder \citep{project_jupyter-proc-scipy-2018, Love2014, Wilke2016}.
+### Experimental design and tissue processing
 
-## Summary
+To characterize reproduction and parental care, birds were sampled at 8 timepoints across the parental care cycle (Fig.
+1a) as perviously described in Austin et al 2020.
+Briefly, nest-building pairs where at least was individual was seen carrying nesting material or shaping a nest (bldg); the day the 1st egg was laid (lay), clutch completion (also known as incubation day 3 (inc.d3)), mid-incubation (inc.d9), late incubation (inc.d17), the day the first chick hatched (where hatch), nestling care days 5 and 9 (n5 and n9).
+Additionally, a non-breeding (control) timepoint was adding using control birds from previously published studies (2,3).
 
-This file is only a basic article template.
-For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
+To test our hypothesis that gene expression is govern more by internal mechanisms than external cues, we conducted multiple offpsring replacement or removal manipuations…. Add details here.
+
+To control for circadian rhythm confounds, each member of the pair was collected between 0900-1200 (PST).
+We sampled a total of 331 birds (16 treatments with ~10 pairs per treatment).
+Pairs were captured and collected simultaneously within 5 min of entering their cage, were anesthetized using isoflurane until unresponsive (<2 min), and then decapitated.
+Trunk blood, brains, pituitaries, and gonads were collected.
+Processing and analysis of truck blood is described in Austin et al.
+2020.
+
+### RNA sequencing analysis
+
+Processing of brains, pituitaries, and gonads for RNA sequenceing is described in detail in Lang et al.
+2020.
+Briefly, RNA from the hypothalamus, pituitary, and gonads was processed for Illumina sequencing using the NEB Next Ultra Directional RNA Library Prep Kit.
+Reads were pseudomapped (insert Kalisto citation) to the Rock Dove transcriptome v1.1.0 whose transcripts were annotated with genes from Gallus gallus genome v5.
+
+Statistical analyses and modeling were perforeming using R \citep{RDevelopmentCoreTeam2013, Wickham2016}.
+Limma was use to process all samples using the model ~ tissue * sex * treatment (citation).
+t-Distributed Stochastic Neighbor Embedding (t-SNE) analysis was conducted using a perplexity of 10 \citep{VanDerMaaten2008}.
+Principal Component Analysis (PCA) was conducted using the top 500 most varying genes (citation).
+DESeq2 was used to model gene express for sex and tissue combination separately using the model ~treatment \citep{Love2014).
+Weighted correlation network analysis (WGCNA) was used to identify set of genes with simliar patterns of gene expression over the course of reprodution and parental care (citation).
+ANOVAs were used to test whether specific genes were differnetially expressed between between sequential timepoints.
+
+### Data availablility
+
+RNA sequencing data for available through the European Nucleotide Archive (ENA) project ID PRJEB16136 and XXXXX.
+Code and data for reproducing the results described in the manuscript are available at https://github.com/macmanes-lab/DoveParentsRNAseq (upload to zendo and cite).
+Data can be explored in the cloud using Binder \citep{project_jupyter-proc-scipy-2018}.
 

### 2020-03-06 docs/manuscript/elife-v1.tex
more pub stuff

@@
 
-\hypertarget{rnaseq}{%
-\subsubsection{RNAseq}\label{rnaseq}}
+\hypertarget{animal-care}{%
+\subsubsection{Animal Care}\label{animal-care}}
 
-See \citet{Lang2019} for details RNAseq.
+All animal care and use comply with the Public Health Service Policy on
+Humane Care and Use of Laboratory Animals and were approved by the
+University of California, Davis IACUC permit \#18895. Birds were housed
+at the University of California, Davis in enclosed aviaries with
+\textasciitilde{}8 sexually reproductive adult pairs per aviary. All
+adults birds are uniquely associated with nests by their unique color
+band combination. Food and water were provided ad libitum, and nests
+were monitored daily.
 
-Differential gene expression conducted wtih d Binder
-\citep{project_jupyter-proc-scipy-2018, Love2014, Wilke2016}.
+\hypertarget{experimental-design-and-tissue-processing}{%
+\subsubsection{Experimental design and tissue
+processing}\label{experimental-design-and-tissue-processing}}
 
-\hypertarget{summary}{%
-\subsection{Summary}\label{summary}}
+To characterize reproduction and parental care, birds were sampled at 8
+timepoints across the parental care cycle (Fig. 1a) as perviously
+described in Austin et al 2020. Briefly, nest-building pairs where at
+least was individual was seen carrying nesting material or shaping a
+nest (bldg); the day the 1st egg was laid (lay), clutch completion (also
+known as incubation day 3 (inc.d3)), mid-incubation (inc.d9), late
+incubation (inc.d17), the day the first chick hatched (where hatch),
+nestling care days 5 and 9 (n5 and n9). Additionally, a non-breeding
+(control) timepoint was adding using control birds from previously
+published studies (2,3).
 
-This file is only a basic article template. For full details of
-\emph{The R Journal} style and information on how to prepare your
-article for submission, see the
-\href{https://journal.r-project.org/share/author-guide.pdf}{Instructions
-for Authors}.
+To test our hypothesis that gene expression is govern more by internal
+mechanisms than external cues, we conducted multiple offpsring
+replacement or removal manipuations\ldots{}. Add details here.
+
+To control for circadian rhythm confounds, each member of the pair was
+collected between 0900-1200 (PST). We sampled a total of 331 birds (16
+treatments with \textasciitilde{}10 pairs per treatment). Pairs were
+captured and collected simultaneously within 5 min of entering their
+cage, were anesthetized using isoflurane until unresponsive
+(\textless{}2 min), and then decapitated. Trunk blood, brains,
+pituitaries, and gonads were collected. Processing and analysis of truck
+blood is described in Austin et al.~2020.
+
+\hypertarget{rna-sequencing-analysis}{%
+\subsubsection{RNA sequencing analysis}\label{rna-sequencing-analysis}}
+
+Processing of brains, pituitaries, and gonads for RNA sequenceing is
+described in detail in Lang et al.~2020. Briefly, RNA from the
+hypothalamus, pituitary, and gonads was processed for Illumina
+sequencing using the NEB Next Ultra Directional RNA Library Prep Kit.
+Reads were pseudomapped (insert Kalisto citation) to the Rock Dove
+transcriptome v1.1.0 whose transcripts were annotated with genes from
+Gallus gallus genome v5.
+
+Statistical analyses and modeling were perforeming using R
+\citep{RDevelopmentCoreTeam2013, Wickham2016}. Limma was use to process
+all samples using the model \textasciitilde{} tissue * sex * treatment
+(citation). t-Distributed Stochastic Neighbor Embedding (t-SNE) analysis
+was conducted using a perplexity of 10 \citep{VanDerMaaten2008}.
+Principal Component Analysis (PCA) was conducted using the top 500 most
+varying genes (citation). DESeq2 was used to model gene express for sex
+and tissue combination separately using the model
+\textasciitilde{}treatment \textbackslash{}citep\{Love2014). Weighted
+correlation network analysis (WGCNA) was used to identify set of genes
+with simliar patterns of gene expression over the course of reprodution
+and parental care (citation). ANOVAs were used to test whether specific
+genes were differnetially expressed between between sequential
+timepoints.
+
+\hypertarget{data-availablility}{%
+\subsubsection{Data availablility}\label{data-availablility}}
+
+RNA sequencing data for available through the European Nucleotide
+Archive (ENA) project ID PRJEB16136 and XXXXX. Code and data for
+reproducing the results described in the manuscript are available at
+\url{https://github.com/macmanes-lab/DoveParentsRNAseq} (upload to zendo
+and cite). Data can be explored in the cloud using Binder
+\citep{project_jupyter-proc-scipy-2018}.
 

### 2020-03-06 docs/manuscript/elife-v1.tex
done with pub for now

@@
 time-constraints required to successfully reproduce and rear their
-offspring to independence(1). A complex series of evolutionarily
+offspring to independence (cite). A complex series of evolutionarily
 conserved physiological processes underlie behaviors typically
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
+  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.
+a) Experimental design.
+b) t-SNE plot highlighting tissue differences in gene expression.
+Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple).
+c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males).
+t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orangse, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
+  \label{figure:fig1}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig4-1}
+  \caption{Another figure}
+  \label{figure:fig4}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig5-1}
+  \caption{Another another figure}
+  \label{figure:fig5}
+\end{figure}
+
 \hypertarget{discussion}{%
 \subsection{Discussion}\label{discussion}}
+
+Results in context to literature.
+
+Significance.
+
+Limitations.
+
+Data accessibility and reproducibility are key challenges for today's
+researchers. Using Binder and Shiny, we have provides a means for
+scientists, physicians, and the general public to explore the data,
+code, and results from this project at
+\url{https://raynamharris.shinyapps.io/musicalgenes/}. Like Hipposeq
+(Cembrowski et al elife 2016, 2018), this tools allows quick and easy
+data exploration of RNAseq data from mutiple cell types that relate to
+brain function.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/supplfig3}
+  \caption{R shiny, image 1}
+  \label{figure:supplfig3}
+\end{figure}
 

### 2020-03-06 docs/manuscript/elife_v1.Rmd
done with pub for now

@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
+  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.
+a) Experimental design.
+b) t-SNE plot highlighting tissue differences in gene expression.
+Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple).
+c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males).
+t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orangse, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
+  \label{figure:fig1}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig4-1}
+  \caption{Another figure}
+  \label{figure:fig4}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig5-1}
+  \caption{Another another figure}
+  \label{figure:fig5}
+\end{figure}
+
 ## Discussion
+
+Results in context to literature.
+
+Significance.
+
+Limitations.
+
+Data accessibility and reproducibility are key challenges for today's researchers.
+Using Binder and Shiny, we have provides a means for scientists, physicians, and the general public to explore the data, code, and results from this project at <https://raynamharris.shinyapps.io/musicalgenes/>.
+Like Hipposeq (Cembrowski et al elife 2016, 2018), this tools allows quick and easy data exploration of RNAseq data from mutiple cell types that relate to brain function.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/supplfig3}
+  \caption{R shiny, image 1}
+  \label{figure:supplfig3}
+\end{figure}
 

### 2020-03-06 analysis/figure-1.Rmd
new fig 1
now with tsne fig 5

@@
 
-subsetmaketsne <- function(whichtissue, whichtreatment, whichsex){
-
-  colData <- colData %>%
-    dplyr::filter(tissue %in% whichtissue,
-                  treatment %in% whichtreatment,
-                  sex %in% whichsex) 
-  row.names(colData) <- colData$V1
-
-  # save counts that match colData
-  savecols <- as.character(colData$V1) 
-  savecols <- as.vector(savecols)
-
-  countData <- as.data.frame(t(countData))
-  countData <- countData %>% dplyr::select(one_of(savecols)) 
-  countData <- as.data.frame(t(countData))
-
-  euclidist <- dist(countData) # euclidean distances between the rows
-
-  tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=10, theta=0.5, dims=2)
-  tsne_df = as.data.frame(tsne_model$Y)
-
-  # prep for adding columns
-  colData2 <- colData 
-  colData2$V1 <- NULL
-  tsne_df_cols <- cbind(colData2, tsne_df)
-  return(tsne_df_cols)
-}
-
 chartsne <- subsetmaketsne(tissuelevels, charlevels, sexlevels)
@@
 
-plottsneelipse <- function(tsnedf, pointcolor, whichcolors){
-  p <- ggplot(tsnedf, aes(x = V1, y = V2)) +
-    geom_point(size = 1, aes(color = pointcolor)) +
-    theme_B3() +
-    labs(x = "tSNE 1", y = "tSNE 2") +
-    scale_color_manual(values = whichcolors) +
-    theme(legend.position = "none",
-          axis.text = element_blank()) +
-    stat_ellipse(linetype = 1, aes(color = tissue )) 
-  return(p)
-}
-
 a <- plottsneelipse(chartsne, chartsne$tissue, allcolors) + labs(subtitle = "~ tissue")    
@@
 
-plottsneelipsev2 <- function(tsnedf, pointcolor, whichcolors){
-  p <- ggplot(tsnedf, aes(x = V1, y = V2)) +
-    geom_point(size = 1, aes(color = pointcolor)) +
-    theme_B3() +
-    labs(x = "tSNE 1", y = "tSNE 2") +
-    scale_color_manual(values = whichcolors) +
-    theme(legend.position = "none",
-          axis.text = element_blank()) +
-    stat_ellipse(linetype = 1, aes(color = treatment)) 
-  return(p)
-}
-
 h <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus ~ treatment")  + facet_wrap(~sex, scales = "free")
@@
 ```
+
+```{r}
+write.csv(chartsne, "../../musicalgenes/data/tsne.csv")
+```

### 2020-03-06 analysis/figure5-maniptsne.Rmd
now with tsne fig 5

---
title: "Fig5"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(Rtsne)

source("../R/themes.R")
source("../R/functions.R")
source("../R/wrangledata.R")

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
```

# Figure 1. Experimental design, tSNE analysis, PCA, and prolactin

```{r tSNE-df}

# prep for tsne

hyptsne <- subsetmaketsne("hypothalamus", allmaniplevels, sexlevels)
pittsne <- subsetmaketsne("pituitary", allmaniplevels, sexlevels)
gontsne <- subsetmaketsne("gonads", allmaniplevels, sexlevels)

addgroupings <- function(df){

  df <- df %>% mutate(hiloPRL = fct_collapse(treatment, 
                                  lo = c("m.inc.d8", "m.inc.d3", "m.inc.d9", "inc.d3", "inc.d9", "lay"),
                                  hi = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch"  ,   "m.n2", "extend", "n5")),
                extint = fct_collapse(treatment, 
                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
                                  chicks = c( "m.inc.d8", "hatch", "extend", "n5"),
                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17")))
  df$extint <- factor(df$extint, levels = levelsextint)
  return(df)
}

hyptsne <- addgroupings(hyptsne)
pittsne <- addgroupings(pittsne)
gontsne <- addgroupings(gontsne)

```

## make figure

```{r fig5, fig.width=7, fig.height=6}

expdesign2 <- png::readPNG("../figures/images/DoveParentsRNAseq_hypothesis.png")
expdesign2 <- ggdraw() +  draw_image(expdesign2, scale = 1)

b <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus ~ parental stage", x = NULL)  + facet_wrap(~sex, scales = "free")
c <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary ~ parental stage", x = NULL) + facet_wrap(~sex, scales = "free") 
d <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads ~ parental stage")  + facet_wrap(~sex, scales = "free")

bcd <- plot_grid(b,c,d, nrow = 3, labels = c("b", "c", "d"), label_size = 8, rel_heights = c(1,1,1))

e <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + labs(subtitle = "hypothalamus ~ estimated prolactin", x = NULL) + facet_wrap(~sex, scales = "free")  
f <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + labs(subtitle = "pituitary ~ estimated prolactin", x = NULL) + facet_wrap(~sex, scales = "free")   
g <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + labs(subtitle = "gonads ~ estimated prolactin" ) + facet_wrap(~sex, scales = "free")

efg <- plot_grid(e,f,g, nrow = 3, labels = c("h", "i", "j"), label_size = 8, rel_heights = c(1,1,1))

h <- plottsneelipsev2(hyptsne, hyptsne$extint, allcolors ) + labs(subtitle = "hypothalamus ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")  
i <- plottsneelipsev2(pittsne, pittsne$extint, allcolors ) + labs(subtitle = "pituitary ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")   
j <- plottsneelipsev2(gontsne, gontsne$extint, allcolors ) + labs(subtitle = "gonads ~ external enviornment" ) + facet_wrap(~sex, scales = "free")

hij <- plot_grid(h,i,j, nrow = 3, labels = c("e", "f", "g"), label_size = 8, rel_heights = c(1,1,1))

bcdhijefg <- plot_grid(bcd, hij,  efg, nrow = 1)

fig6 <- plot_grid(expdesign2, bcdhijefg, nrow = 2, labels = c("e"), label_size = 8, rel_heights = c(0.4,1))
fig6

```

### 2020-03-06 analysis/figure5-maniptsne.Rmd
now with tsne fig 5

@@
   df <- df %>% mutate(hiloPRL = fct_collapse(treatment, 
-                                  lo = c("m.inc.d8", "m.inc.d3", "m.inc.d9", "inc.d3", "inc.d9", "lay"),
+                                  lo = c("m.inc.d8", "m.inc.d3", "m.inc.d9", "inc.d3", "inc.d9", "lay", "bldg"),
                                   hi = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" ,  "m.n2", "extend", "n5")),
@@
                                   chicks = c( "m.inc.d8", "hatch", "extend", "n5"),
-                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17")))
+                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")))
   df$extint <- factor(df$extint, levels = levelsextint)
@@
 
-expdesign2 <- png::readPNG("../figures/images/DoveParentsRNAseq_hypothesis.png")
+expdesign2 <- png::readPNG("../figures/images/DoveParentsRNAseq_maniptsne.png")
 expdesign2 <- ggdraw() +  draw_image(expdesign2, scale = 1)
@@
 
-fig6 <- plot_grid(expdesign2, bcdhijefg, nrow = 2, labels = c("e"), label_size = 8, rel_heights = c(0.4,1))
+fig6 <- plot_grid(expdesign2, bcdhijefg, nrow = 2, labels = c("e"), label_size = 8, rel_heights = c(0.5,1))
 fig6

### 2020-03-07 analysis/figure5-maniptsne.Rmd
better colors and illustration

@@
 
-bcd <- plot_grid(b,c,d, nrow = 3, labels = c("b", "c", "d"), label_size = 8, rel_heights = c(1,1,1))
+bcd <- plot_grid(b,c,d, nrow = 3, labels = c("b"), label_size = 8, rel_heights = c(1,1,1))
 
-e <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + labs(subtitle = "hypothalamus ~ estimated prolactin", x = NULL) + facet_wrap(~sex, scales = "free")  
-f <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + labs(subtitle = "pituitary ~ estimated prolactin", x = NULL) + facet_wrap(~sex, scales = "free")   
-g <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + labs(subtitle = "gonads ~ estimated prolactin" ) + facet_wrap(~sex, scales = "free")
+e <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + labs(subtitle = "hypothalamus ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")  
+f <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + labs(subtitle = "pituitary ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")   
+g <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + labs(subtitle = "gonads ~ internal physiology" ) + facet_wrap(~sex, scales = "free")
 
-efg <- plot_grid(e,f,g, nrow = 3, labels = c("h", "i", "j"), label_size = 8, rel_heights = c(1,1,1))
+efg <- plot_grid(e,f,g, nrow = 3, labels = c("d"), label_size = 8, rel_heights = c(1,1,1))
 
@@
 
-hij <- plot_grid(h,i,j, nrow = 3, labels = c("e", "f", "g"), label_size = 8, rel_heights = c(1,1,1))
+hij <- plot_grid(h,i,j, nrow = 3, labels = c("c"), label_size = 8, rel_heights = c(1,1,1))
 
@@
 
-fig6 <- plot_grid(expdesign2, bcdhijefg, nrow = 2, labels = c("e"), label_size = 8, rel_heights = c(0.5,1))
+fig6 <- plot_grid(expdesign2, bcdhijefg, nrow = 2, labels = c("a"), label_size = 8, rel_heights = c(0.4,1))
 fig6

### 2020-03-07 analysis/figure-2.Rmd
new fig 1
better fig 2 and 3

@@
 
-```{r figurefunctions}
+```{r fig2, fig.width=7, fig.height=7}
 
@@
 
-```
-
-## Internal versus external
-
-```{r hypothesis}
-DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
-DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
-DEG_pathfiles <- paste0(DEG_path, DEG_files)
-#DEG_files
-
-allDEG2 <- DEG_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name") %>% 
-  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
-  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
-  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
-  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
-  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
-  mutate(comparison = paste(down,up, sep = "_")) %>%
-  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
-  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
-dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
-head(allDEG)
-
-allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)
-
-allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
-allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
-                                                      "eggs vs. chicks" = "eggs_chicks"))
-allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))
-
-```
-
-## DEGS
-
-```{r fig2, fig.width=6, fig.height=5}
 # hyp
@@
                                                                axis.title.x = element_blank())
-
 # pit
@@
                                                               strip.text.x = element_blank())
-
 # gon
 p3 <- makebargraph("gonad","DEGs", 0, 1250) +  theme(strip.text.x = element_blank())
-
 bcd <- plot_grid(p1,p2,p3, nrow = 3, rel_heights = c(1.2,1,1.5), labels = c("b", "c", "d"), label_size = 8)
 
-expdesign <- png::readPNG("../figures/images/DoveParentsRNAseq_DEGS.png")
+expdesign <- png::readPNG("../figures/images/fig_fig2a.png")
 expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
 
-plot_grid(expdesign, bcd, labels = c("a", " "), label_size = 8, nrow = 2, rel_heights = c(0.2,1))
+plot_grid(expdesign, bcd, labels = c("a", " "), label_size = 8, nrow = 2, rel_heights = c(0.25,1))
 
-```
-
-```{r fig2b, fig.width=5, fig.height=5}
-
-makenewbargraph <- function(whichtissue, whichsex,  whichcomparison, lowlim, higherlim){
-  p <- allDEG2 %>%
-    filter(tissue == whichtissue,
-           comparison == whichcomparison,
-           sex == whichsex) %>%
-    ggplot(aes(x = comparison,  fill = direction)) +
-    geom_bar(position = "dodge", drop = FALSE) +
-    theme_B3() +
-    theme(legend.position = "none")  +
-    guides(fill = guide_legend(nrow = 1)) +
-    labs( y = whichtissue) +
-  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
-           position = position_dodge(width = 1),
-           size = 2, color = "black")  + 
-      ylim(lowlim, higherlim) +
-  scale_fill_manual(values = allcolors, name = "higher in")   + 
-    theme(axis.text.x = element_blank()) 
-  return(p)
-}
-
-b11 <- makenewbargraph("hypothalamus", "female","eggs vs. chicks", 0, 2500) +  labs(subtitle = "females", x = NULL, y = "Hyp. DEGS")
-b21 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2500)  + labs(x = NULL, y = "Pit. DEGS")
-b31 <- makenewbargraph("gonad", "female", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks", y = "Gon. DEGS")   
-b112131 <- plot_grid(b11,b21,b31, nrow = 3, rel_heights = c(1.1,1,1.1))
-
-b12 <- makenewbargraph("hypothalamus", "male",  "eggs vs. chicks", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
-  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
-b22 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2500) + labs(x = NULL)+ 
-  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
-b32 <- makenewbargraph("gonad", "male", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks")   + 
-  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
-b122232 <- plot_grid(b12, b22,b32, nrow = 3, rel_heights = c(1.1,1,1.1))
-
-c11 <- makenewbargraph("hypothalamus", "female", "lo vs. hi PRL   ", 0, 2500) +  labs(subtitle = "females", x = NULL) + 
-  theme(axis.title.y = element_blank())
-c21 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2500)  + labs(x = NULL)+ 
-  theme(axis.title.y = element_blank())
-c31 <- makenewbargraph("gonad", "female", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL")  + 
-  theme(axis.title.y = element_blank()) 
-c112131 <- plot_grid(c11,c21,c31, nrow = 3, rel_heights = c(1.1,1,1.1))
-
-c12 <- makenewbargraph("hypothalamus", "male",  "lo vs. hi PRL   ", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
-  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
-c22 <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = NULL)+ 
-  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
-c32 <- makenewbargraph("gonad", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL")   + 
-  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
-c122232 <- plot_grid(c12, c22,c32, nrow = 3, rel_heights = c(1.1,1,1.1))
-
-hypothesisbars <-  plot_grid(b112131, b122232, c112131, c122232, nrow = 1, rel_widths = c(1.3, 1, 1.1, 1),
-          labels = c("b", " ", "c", " "), label_size = 8)
-
-expdesign2 <- png::readPNG("../figures/images/DoveParentsRNAseq_hypothesis.png")
-expdesign2 <- ggdraw() +  draw_image(expdesign2, scale = 1)
-
-plot_grid(expdesign2, hypothesisbars, rel_heights = c(0.5,1), nrow = 2, labels = c("a", " "), label_size = 8)
 ```
@@
 
-table2 <- allDEG2 %>%
-  filter(gene %in% candidategenes) %>%
-  arrange(gene) %>%
-  group_by(sex, tissue, comparison) %>%
-  summarize(genes = str_c(gene, collapse = " ")) %>%
-  pivot_wider(names_from = comparison, values_from = genes ) %>%
-  arrange( tissue, sex)
-table(table2)
-
 write_csv(table1, "../results/table1.csv")
-write_csv(table2, "../results/table2.csv")
 ```
@@
 
-suppletable2 <- allDEG2 %>%
-  filter(comparison != "control_bldg") %>%
-  group_by(sex, tissue, comparison) %>%
-  arrange( tissue, sex, direction, gene)
-head(suppletable2)
-
 write_csv(suppletable1, "../results/suppletable1.csv")
-
-write_csv(suppletable2, "../results/suppletable2.csv")
 
@@
 
-## female pit, parental/repro only, not PRL
+```{r savefiles}
+allDEG %>%
+  drop_na()
 
-```{r}
-allDEGslim <- allDEG %>% filter(comparison != "control_bldg")
-
-femalepitPRL <- allDEG2 %>% filter(sex == "female", tissue == "pituitary") %>%
-  filter(comparison != "eggs vs. chicks") %>% arrange(gene) %>% pull(gene)
-
-femalepitparental <- allDEGslim %>% 
-  filter(!gene %in% femalepitPRL,
-         sex == "female", tissue == "pituitary") %>%
-  arrange(gene, sex, tissue, comparison, direction) %>%
-  distinct(gene) %>% pull(gene)
-femalepitparental
-
-malepitPRL <- allDEG2 %>% filter(sex == "male", tissue == "pituitary") %>%
-  filter(comparison != "eggs vs. chicks") %>% arrange(gene) %>% pull(gene)
-
-malepitparental <- allDEGslim %>% 
-  filter(!gene %in% malepitPRL,
-         sex == "male", tissue == "pituitary") %>%
-  arrange(gene, sex, tissue, comparison, direction) %>%
-  distinct(gene) %>% pull(gene)
-malepitparental
+write.csv(allDEG, "../../musicalgenes/data/allDEG.csv")
 

### 2020-03-07 analysis/figure-3.Rmd
better fig 2 and 3

--- 2020-03-02 analysis/figure-3.Rmd
@@
 source("../R/functions.R")
-source("../R/icons.R")
 source("../R/wrangledata.R")
@@
 
-```{r data}
+```{r PRLfoo}
 # pca
@@
 
+```{r determinePRLhiglo, fig.height = 6, fig.width=3}
+PRLvsd2 <- PRLvsd %>%
+  filter(tissue == "pituitary")
+head(PRLvsd2)
+
+PRLvsd2 %>%
+  group_by(sex) %>%
+  summarize(median = median(counts))
+
+PRLvsd3 <- PRLvsd2 %>%
+  mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
+  drop_na()
+PRLvsd3$hiloPRL <- factor(PRLvsd3$hiloPRL, levels = c("lo", "hi"))
+
+PRLvsd3 %>%
+  group_by(sex, tissue, hiloPRL) %>%
+  summarize(n = n())
+
+```
+
+## Internal versus external
+
+```{r hypothesis}
+DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG2 <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
+head(allDEG2)
+
+allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)
+
+allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
+allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
+                                                      "eggs vs. chicks" = "eggs_chicks"))
+allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))
+
+```
+
 ```{r fig3, fig.width=7, fig.height=7}
+
+expdesign <- png::readPNG("../figures/images/fig_fig3a.png")
+expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
 
@@
 
-abcd <- plot_grid(a,b,c,d, labels = c("b", "c", "d", "e"), ncol = 2, label_size = 8)
+bcde <- plot_grid(a,b,c,d, labels = c("b", "c", "d", "e"), ncol = 2, label_size = 8)
 
-expdesign <- png::readPNG("../figures/images/parentalstages_fig1a.png")
-expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
+makenewbargraph <- function(whichtissue, whichsex,  whichcomparison, lowlim, higherlim){
+  p <- allDEG2 %>%
+    filter(tissue == whichtissue,
+           comparison == whichcomparison,
+           sex == whichsex) %>%
+    ggplot(aes(x = comparison,  fill = direction)) +
+    geom_bar(position = "dodge", drop = FALSE) +
+    theme_B3() +
+    theme(legend.position = "none")  +
+    guides(fill = guide_legend(nrow = 1)) +
+    labs( y = whichtissue) +
+  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
+           position = position_dodge(width = 1),
+           size = 2, color = "black")  + 
+      ylim(lowlim, higherlim) +
+  scale_fill_manual(values = allcolors, name = "higher in")   + 
+    theme(axis.text.x = element_blank()) 
+  return(p)
+}
 
-abcde <- plot_grid(expdesign, abcd, nrow = 2, labels = c("a", "b"), label_size = 8, rel_heights = c(0.3,1))
-abcde
+b11 <- makenewbargraph("hypothalamus", "female","eggs vs. chicks", 0, 2500) +  labs(subtitle = "females", x = NULL, y = " ") + 
+  theme( axis.text = element_blank())
+b21 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2500)  + labs(x = NULL, y = " ") + 
+  theme(axis.text = element_blank())
+b31 <- makenewbargraph("gonad", "female", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks", y = " ")    + 
+  theme(axis.text = element_blank())
+b112131 <- plot_grid(b11,b21,b31, nrow = 3, rel_heights = c(1.1,1,1.1))
+
+b12 <- makenewbargraph("hypothalamus", "male",  "eggs vs. chicks", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
+  theme(axis.title.y = element_blank(), axis.text = element_blank())
+b22 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2500) + labs(x = NULL)+ 
+  theme(axis.title.y = element_blank(), axis.text = element_blank())
+b32 <- makenewbargraph("gonad", "male", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks")   + 
+  theme(axis.title.y = element_blank(), axis.text = element_blank())
+b122232 <- plot_grid(b12, b22,b32, nrow = 3, rel_heights = c(1.1,1,1.1))
+
+c11 <- makenewbargraph("hypothalamus", "female", "lo vs. hi PRL   ", 0, 2500) +  labs(subtitle = "females", x = NULL, y = "hypothalamus DEGs") + 
+  theme( axis.text.x = element_blank()) 
+c21 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2500)  + labs(x = NULL, y = "pituitary DEGs")+ 
+  theme( axis.text.x = element_blank())
+c31 <- makenewbargraph("gonad", "female", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL", y = "gonad DEGs")  + 
+  theme(axis.text.x = element_blank()) 
+c112131 <- plot_grid(c11,c21,c31, nrow = 3, rel_heights = c(1.1,1,1.1))
+
+c12 <- makenewbargraph("hypothalamus", "male",  "lo vs. hi PRL   ", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
+  theme(axis.title.y = element_blank(), axis.text = element_blank())
+c22 <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = NULL)+ 
+  theme(axis.title.y = element_blank(), axis.text = element_blank())
+c32 <- makenewbargraph("gonad", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL")   + 
+  theme(axis.title.y = element_blank(), axis.text = element_blank())
+c122232 <- plot_grid(c12, c22,c32, nrow = 3, rel_heights = c(1.1,1,1.1))
+
+hypothesisbars <-  plot_grid(c112131, c122232, b112131, b122232, nrow = 1, rel_widths = c(1.3, 1, 1.1, 1),
+          labels = c("f", " ", "g", " "), label_size = 8)
+
+dataplots <- plot_grid(bcde, hypothesisbars)
+
+fig3 <- plot_grid(expdesign, dataplots, rel_heights = c(0.25,1), nrow = 2, labels = c(" ", "e"), label_size = 8)
+fig3
 
@@
 
-```{r determinePRLhiglo, fig.height = 6, fig.width=3}
-PRLvsd2 <- PRLvsd %>%
-  filter(tissue == "pituitary")
-head(PRLvsd2)
+```{r savefiles}
+#write.csv(PRLvsd3,"../results/PRLvsd.csv", row.names = F)
 
-PRLvsd2 %>%
-  group_by(sex) %>%
-  summarize(median = median(counts))
-
-a <- ggplot(PRLvsd2, aes(x = sex, y = counts, fill = sex))  +
-  geom_boxplot() +
-  labs(x = "Sex", y = "Prolactin expression")+
-  theme_B3() + theme(legend.position = "none") +
-  scale_fill_manual(values = sexcolors) +
-  geom_hline(yintercept=18, linetype="dashed", color = "black")
-
-PRLvsd3 <- PRLvsd2 %>%
-  mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
-  drop_na()
-PRLvsd3$hiloPRL <- factor(PRLvsd3$hiloPRL, levels = c("lo", "hi"))
-
-b <- ggplot(PRLvsd3, aes(x = sex, y = counts, fill = sex, color = hiloPRL))  +
-  geom_boxplot() +
-  labs(x = "Prolactin, binned into `lo` and `hi`", y = "Prolactin expression")+
-  theme_B3() + theme(legend.position = "none") +
-  scale_fill_manual(values = allcolors) +
-  scale_color_manual(values = allcolors)
-
-plot_grid(a,b, nrow = 2, rel_widths = c(1,1))
-
-PRLvsd3 %>%
-  group_by(sex, tissue, hiloPRL) %>%
-  summarize(n = n())
-
-write.csv(PRLvsd3,"../results/PRLvsd.csv", row.names = F)
 ```

### 2020-03-07 analysis/figure-3.Rmd
better fig 1-4
mew figs

@@
 
-a <- plotcolorfulpcs(charpca,charpca$treatment, allcolors) + labs(subtitle = " ") +
+b <- plotcolorfulpcs(charpca,charpca$treatment, allcolors) + labs(subtitle = " ") +
   theme(legend.position = "none", 
@@
 
-b <- plotfriz(charfviz) + labs(subtitle = "pituitary") +
-  #ylim(-250000,500000) + xlim(-250000,500000) + 
-  theme(axis.text = element_blank())
-
 c <- plotprolactin(PRLpit, PRLpit$counts, "PRL", "pituitary") + 
   theme(legend.position = "none", 
-         axis.text.x = element_text(angle = 45, hjust = 1),
+         axis.text.x = element_blank(),
+        axis.title.x = element_blank(),
         axis.title.y = element_text(face = "italic"))
@@
 
-bcde <- plot_grid(a,b,c,d, labels = c("b", "c", "d", "e"), ncol = 2, label_size = 8)
+bcd <- plot_grid(b,c,d, labels = c("b", "c", "d"), ncol = 1, label_size = 8, rel_heights = c(1,1,1.3))
 
@@
 
-dataplots <- plot_grid(bcde, hypothesisbars)
+dataplots <- plot_grid(bcd, hypothesisbars)
 
-fig3 <- plot_grid(expdesign, dataplots, rel_heights = c(0.25,1), nrow = 2, labels = c(" ", "e"), label_size = 8)
+fig3 <- plot_grid(expdesign, dataplots, rel_heights = c(0.25,1), nrow = 2, labels = c("a"), label_size = 8)
 fig3

### 2020-03-07 analysis/figure-4.Rmd
mew figs

--- 2020-03-02 analysis/figure-4.Rmd
@@
 ---
-title: "fig4"
+title: "Fig5"
 output: md_document
@@
 
-# Figure 4
-
-```{r setup, include=TRUE}
+```{r setup, include=FALSE}
 library(tidyverse)
 library(cowplot)
+library(Rtsne)
 
 source("../R/themes.R")
+source("../R/functions.R")
+source("../R/wrangledata.R")
 
-knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
 ```
 
-```{r data}
-geneids <- read_csv("../metadata/00_geneinfo.csv")
+# Figure 1. Experimental design, tSNE analysis, PCA, and prolactin
 
-vsd_path <- "../results/DEseq2/"   # path to the data
-vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
-vsd_pathfiles <- paste0(vsd_path, vsd_files)
-vsd_files
+```{r tSNE-df}
 
-allvsd <- vsd_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name")  %>% 
-  dplyr::rename("gene" = "X1") %>% 
-  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
-               names_to = "samples", values_to = "counts")
+# prep for tsne
 
-hypgenes <- c("POMC", "DRD1")
-pitgenes <- c("PRL", "BRCA1")
-gongenes <- c("AVPR1A", "PRLR")
+hyptsne <- subsetmaketsne("hypothalamus", allmaniplevels, sexlevels)
+pittsne <- subsetmaketsne("pituitary", allmaniplevels, sexlevels)
+gontsne <- subsetmaketsne("gonads", allmaniplevels, sexlevels)
 
-candidategenes <- c("PRL", "PRLR" ,  "AVPR1B" , "MYC" ,  
-                "MAP2KA", "MSH" ,   "PALB2" , "BRCA1", 
-                "OXT", "ESR1", "ESR2", "CISH", 
-                "POU2F1", "RAP1GAP", "CEBPB","ARHGAP32"
-                )
+addgroupings <- function(df){
 
-returnvsds <- function(whichgenes){
-
-  df  <- allvsd %>%
-    filter(gene %in% whichgenes) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
-    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
-                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
-                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
-    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
-    dplyr::select(sex, tissue, treatment, gene, samples, counts)
-
-  df$treatment <- factor(df$treatment, levels = alllevels)
-  print(head(df))
+  df <- df %>% mutate(hiloPRL = fct_collapse(treatment, 
+                                  lo = c("m.inc.d8", "m.inc.d3", "m.inc.d9", "inc.d3", "inc.d9", "lay", "bldg"),
+                                  hi = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" ,  "m.n2", "extend", "n5")),
+                extint = fct_collapse(treatment, 
+                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
+                                  chicks = c( "m.inc.d8", "hatch", "extend", "n5"),
+                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")))
+  df$extint <- factor(df$extint, levels = levelsextint)
   return(df)
@@
 
-candidatevsd <- returnvsds(candidategenes)
+hyptsne <- addgroupings(hyptsne)
+pittsne <- addgroupings(pittsne)
+gontsne <- addgroupings(gontsne)
 
@@
 
-```{r fig4, fig.height=7, fig.width=7}
-plottopgenes <- function(whichgenes, whichtissue, whichsex, mysubtitle, whichtstages){
-  candidates  <- allvsd %>%
-    filter(gene %in% whichgenes) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
-    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
-                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
-                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
-    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
-    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
-    filter(tissue == whichtissue, sex %in% whichsex)
+## make figure
 
-  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
+```{r fig4, fig.width=7, fig.height=7}
 
-  p <- candidates %>%
-    filter(treatment %in% whichtstages) %>%
-    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
-    geom_boxplot() + facet_wrap(~gene, scales = "free_y") + 
-    theme_B3() +
-    scale_fill_manual(values = allcolors) +
-    scale_color_manual(values = allcolors) +
-    theme(axis.text.x = element_blank(),
-          legend.position = "none",
-          axis.title.x = element_blank(),
-          strip.text = element_text(face = "italic")) +
-    labs(y = mysubtitle) 
-  return(p)
+expdesign2 <- png::readPNG("../figures/images/fig_fig4a.png")
+expdesign2 <- ggdraw() +  draw_image(expdesign2, scale = 1)
 
-}
+b <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus ~ parental stage", x = NULL)  + facet_wrap(~sex, scales = "free")
+c <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary ~ parental stage", x = NULL) + facet_wrap(~sex, scales = "free") 
+d <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads ~ parental stage")  + facet_wrap(~sex, scales = "free")
 
-a <- plottopgenes(hypgenes, "hypothalamus", c("female"), "hypothalamus", c("hatch", "n5")) 
-b <- plottopgenes(pitgenes, "pituitary", c("female"), "pituitary", c("inc.d9", "inc.d17")) 
-c <- plottopgenes(gongenes, "gonad", c("female"), "gonads", c("lay", "inc.d3")) + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+bcd <- plot_grid(b,c,d, nrow = 3, labels = c("b"), label_size = 8, rel_heights = c(1,1,1))
 
-fig4a <- plot_grid(a,b,c, ncol = 1)
+e <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + labs(subtitle = "hypothalamus ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")  
+f <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + labs(subtitle = "pituitary ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")   
+g <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + labs(subtitle = "gonads ~ internal physiology" ) + facet_wrap(~sex, scales = "free")
 
-d <- plottopgenes(hypgenes, "hypothalamus", c("female", "male"), NULL, charlevels) 
-e <- plottopgenes(pitgenes, "pituitary", c("female", "male"), NULL, charlevels) 
-f <- plottopgenes(gongenes, "gonad", c("female", "male"), NULL, charlevels) + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+efg <- plot_grid(e,f,g, nrow = 3, labels = c("d"), label_size = 8, rel_heights = c(1,1,1))
 
-fig4b <- plot_grid(d,e,f, ncol = 1)
+h <- plottsneelipsev2(hyptsne, hyptsne$extint, allcolors ) + labs(subtitle = "hypothalamus ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")  
+i <- plottsneelipsev2(pittsne, pittsne$extint, allcolors ) + labs(subtitle = "pituitary ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")   
+j <- plottsneelipsev2(gontsne, gontsne$extint, allcolors ) + labs(subtitle = "gonads ~ external enviornment" ) + facet_wrap(~sex, scales = "free")
 
-fig4 <- plot_grid(fig4a, fig4b, rel_widths = c(1,2))
-fig4
+hij <- plot_grid(h,i,j, nrow = 3, labels = c("c"), label_size = 8, rel_heights = c(1,1,1))
+
+bcdhijefg <- plot_grid(bcd, hij,  efg, nrow = 1)
+
+fig5 <- plot_grid(expdesign2, bcdhijefg, nrow = 2, labels = c("a"), label_size = 8, rel_heights = c(0.25,1))
+fig5
+
 ```
-
-```{r savefiles}
-write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
-```

### 2020-03-07 docs/manuscript/elife_v1.Rmd
mew figs

@@
 ---
-title: Internal physiology and external cues drive changes in reproductive axis of parenting pigeons
+title: Pituitary prolactin drives transcriptional changes in the reproductive axis of parenting pigeons
 author:
   - name: Rayna M Harris ^[equal contribtion] 
-    affiliation: UC Davis
-  - name: Suzanne H. Austin  ^[equal contribtion] 
-    affiliation: UC Davis
+    affiliation: University of California, Davis
+  - name: Suzanne H. Austin  ^[equal contribtion, currently at the University of Oregon] 
+    affiliation: University of California, Davis
   - name: Andrew S. Lang
-    affiliation: UNH
+    affiliation: University of New Hampshire
   - name: Victoria S. Farrar
-    affiliation: UC Davis
+    affiliation: University of California, Davis
   - name: April Booth
-    affiliation: UC Davis
+    affiliation: University of California, Davis
   - name: Tanner Feustel
-    affiliation: UC Davis
+    affiliation: University of California, Davis
   - name: Matthew D. MacManes
-    affiliation: UNH
+    affiliation: University of New Hampshire
   - name: Rebecca M. Calisi  ^[corresponding author] 
-    affiliation: UC Davis
-    email:  author7@work   
+    affiliation: University of California, Davis
+    email:  rmcalisi@ucdavis.edu  
 address:
-  - code: UC Davis
-    address: Department, Street, City, CA, Zip
-  - code: UNH
-    address: Department, Street, City, NH, Zip    
-abstract: |
+  - code: University of California, Davis
+    address: Department of Neurobiology, Physiology, and Behavior, Davis, CA
+  - code: University of New Hampshire
+    address: Department of Molecular, Cellular and Biomedical Sciences, Durham, NH 
+abstract: In species that provide parental care, successful rearing of offspring involves a shift from aggressive and sexual behaviors to more caring and nurturing ones.
+We hypothesize that these shifts in behavior are related to changes in gene expression that are controlled by internal mechanisms rather than by the external cues.
+We found that gene expression in the hypothalamus is fairly insensitive to the natural transition between egg and nestling stages; however, in the pituitary and gonads, gene expression showed much more plasticity over the course of reproduction.
+We show evidence that natural transitions in reproductive and parental behavior by mediated internal clock controlling gene expression in the pitary and gonads but that the hypothalamus is very sensitive to external cues that signal a distruption in the natural parental care cycle.
+This research provides a deeper understanding of the interplay of genes, hormones, and parental care and has important implications for understanding molecular processes that underlie behavior transitions.
 output:
-  rticles::rjournal_article:
-    includes:
-      in_header: preamble.tex
+  rticles::rjournal_article
 ---
@@
 
+_Characterization of parental care_
+
 To characterize transcriptional activity, we first examined broad patterns of variation in gene expression in the hypothalamus, pituitary, and gonads of male and female rock doves across a series of timepoints that encapsulate important transitions in reproduction and and parental care (Fig.
@@
 \ref{figure:fig1}j).
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
-  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.
-a) Experimental design.
-b) t-SNE plot highlighting tissue differences in gene expression.
-Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple).
-c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males).
-t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orange, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
-  \label{figure:fig1}
-\end{figure}
 
@@
 While the tSNE analysis reveals broad patterns across all levels of the experiment, additional tools are needed it hone in on the specific differences associated with parental care in each tissue.
-
 Differential gene expression analysis can shed insight into specific differences between treatment groups and thus deepen our understand of the transitions between reproduction and parental care behaviors.
@@
 \ref{figure:fig2}a).
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig2-1}
-  \caption{Differential gene expression differences between sequential parental stages.
-a) We calculated the differentail gene expression between each consequtive parental stage samples, for a total of 8 pair-wise comparisons for each sex in each tissue.
-The y-axis of these bar plots displays the total number of differentially expressed genes for each comparsison in the hypothalamus (b), pituitary (c), gonads (d).
-Bars are collored by the direction of the change in gene expression.
-Control: control: red or , nest-building (bldg): orange, lay: yellow; incubation day 3 (inc.d3): light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, nestling care day 5 (n5): blue, and n9: light blue.}
-  \label{figure:fig2}
-\end{figure}
 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
-  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.
-a) Experimental design.
-b) t-SNE plot highlighting tissue differences in gene expression.
-Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple).
-c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males).
-t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orangse, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
-  \label{figure:fig1}
-\end{figure}
+_Internal physiology versus external stimuli hypothesis_
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig4-1}
-  \caption{Another figure}
-  \label{figure:fig4}
-\end{figure}
+We show that samples from the pituitary on 
+                     incubation day 17 cluster more closely with hatch and nestling care 
+                     day 5 samples than with incubation days 3 and 9. This suggest that, 
+                     in the pituitary, internal mechanisms that prepare for chick arrival 
+                     have the greatest effect on gene expression rather than suggesting 
+                     that gene expression responds to changes in the external enviornament.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig5-1}
-  \caption{Another another figure}
-  \label{figure:fig5}
-\end{figure}
+Blah \ref{figure:fig3}, blah, \ref{figure:fig4} blah.
 
-## Discussion
+_Potential implications for cancer research_
 
-Results in context to literature.
+These results confirmed prolactin (PRL) gene expression fluctuates throughout parental care 
+in a manner that is consistent with its role in promoting lactation 
+                and maintaining parental behaviors.
+                Interestingly, we identified about 100 genes whose expression 
+                was correlated with PRL,
+                including (BRCA1, which is associated with breast cancer).
+                Additionally, we identified thousands of genes whose expression 
+                changed over 
+                the course of parental care.
+                Exploring the relationship between PRL and other differentially 
+                expressed genes could provide important insights into the `symphony` of 
+                gene expression 
+                that regulates behavior at the organismal and cellular level.
 
-Significance.
-
-Limitations.
+_Software for data exploration and hypothesis testing_
 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/supplfig3}
-  \caption{R shiny, image 1}
-  \label{figure:supplfig3}
-\end{figure}
+## Discussion
+
+_Results in context to literature. _
+
+blah blah balh
+
+_Significance. _
+
+blah blah balh
+
+_Limitations._
+
+blah blah balh
 
@@
 
-### Animal Care
+_Animal Care_
 
@@
 
-### Experimental design and tissue processing
+_Experimental design and tissue processing_
 
@@
 
-### RNA sequencing analysis
+_RNA sequencing analysis_
 
@@
 
-### Data availablility
+_Data availablility_
 
@@
 
+\newpage
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
+  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.
+a) Experimental design.
+b) t-SNE plot highlighting tissue differences in gene expression.
+Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple).
+c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males).
+t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orange, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
+  \label{figure:fig1}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig2-1}
+  \caption{Differential gene expression differences between sequential parental stages.
+a) We calculated the differentail gene expression between each consequtive parental stage samples, for a total of 8 pair-wise comparisons for each sex in each tissue.
+The y-axis of these bar plots displays the total number of differentially expressed genes for each comparsison in the hypothalamus (b), pituitary (c), gonads (d).
+Bars are collored by the direction of the change in gene expression.
+Control: control: red or , nest-building (bldg): orange, lay: yellow; incubation day 3 (inc.d3): light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, nestling care day 5 (n5): blue, and n9: light blue.}
+  \label{figure:fig2}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig3-1}
+  \caption{Another figure}
+  \label{figure:fig3}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig4-1}
+  \caption{Another another figure}
+  \label{figure:fig4}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/supplfig3}
+  \caption{R shiny, image 1}
+  \label{figure:supplfig3}
+\end{figure}
+
+\newpage
+
+## Author affiliations
+
 <!-- 

### 2020-03-07 docs/manuscript/elife-v1.tex
mew figs

@@
 % !TeX root = RJwrapper.tex
-\title{Internal physiology and external cues drive changes in reproductive axis
-of parenting pigeons}
-\author{by Rayna M Harris \footnote{equal contribtion}, Suzanne H.
-Austin \footnote{equal contribtion}, Andrew S.
-Lang, Victoria S.
-Farrar, April Booth, Tanner Feustel, Matthew D.
-MacManes, Rebecca M.
-Calisi \footnote{corresponding author}}
+\title{Pituitary prolactin drives transcriptional changes in the reproductive
+axis of parenting pigeons}
+\author{by Rayna M Harris \footnote{equal contribtion}, Suzanne H. Austin \footnote{equal contribtion, currently at the
+  University of Oregon}, Andrew S. Lang, Victoria S. Farrar, April Booth, Tanner Feustel, Matthew D. MacManes, Rebecca M. Calisi \footnote{corresponding author}}
 
@@
 
-% Any extra LaTeX you need in the preamble
-
-In species that provide parental care, successful rearing of offspring involves a shift from aggressive and sexual behaviors to more caring and nurturing ones.
-We hypothesize that these shifts in behavior are related to changes in gene expression that are controlled by internal mechanisms rather than by the external cues.
-We found that gene expression in the hypothalamus is fairly insensitive to the natural transition between egg and nestling stages; however, in the pituitary and gonads, gene expression showed much more plasticity over the course of reproduction.
-We show evidence that natural transitions in reproductive and parental behavior by mediated internal clock controlling gene expression in the pitary and gonads but that the hypothalamus is very sensitive to external cues that signal a distruption in the natural parental care cycle.
-This research provides a deeper understanding of the interplay of genes, hormones, and parental care and has important implications for understanding molecular processes that underlie behavior transitions.
+\abstract{%
+In species that provide parental care, successful rearing of offspring
+involves a shift from aggressive and sexual behaviors to more caring and
+nurturing ones. We hypothesize that these shifts in behavior are related
+to changes in gene expression that are controlled by internal mechanisms
+rather than by the external cues. We found that gene expression in the
+hypothalamus is fairly insensitive to the natural transition between egg
+and nestling stages; however, in the pituitary and gonads, gene
+expression showed much more plasticity over the course of reproduction.
+We show evidence that natural transitions in reproductive and parental
+behavior by mediated internal clock controlling gene expression in the
+pitary and gonads but that the hypothalamus is very sensitive to
+external cues that signal a distruption in the natural parental care
+cycle. This research provides a deeper understanding of the interplay of
+genes, hormones, and parental care and has important implications for
+understanding molecular processes that underlie behavior transitions.
+}
 
@@
 
+\emph{Characterization of parental care}
+
 To characterize transcriptional activity, we first examined broad
@@
 sex, and treatment (Fig. \ref{figure:fig1}j).
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
-  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.
-a) Experimental design.
-b) t-SNE plot highlighting tissue differences in gene expression.
-Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple).
-c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males).
-t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orange, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
-  \label{figure:fig1}
-\end{figure}
 
@@
 additional tools are needed it hone in on the specific differences
-associated with parental care in each tissue.
-
-Differential gene expression analysis can shed insight into specific
-differences between treatment groups and thus deepen our understand of
-the transitions between reproduction and parental care behaviors. We
-used DESEq2 to identify genes whose expression changed significantly
-between sequential timepoints in each tissue and sex separately (Fig.
-\ref{figure:fig2}a).
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig2-1}
-  \caption{Differential gene expression differences between sequential parental stages.
-a) We calculated the differentail gene expression between each consequtive parental stage samples, for a total of 8 pair-wise comparisons for each sex in each tissue.
-The y-axis of these bar plots displays the total number of differentially expressed genes for each comparsison in the hypothalamus (b), pituitary (c), gonads (d).
-Bars are collored by the direction of the change in gene expression.
-Control: control: red or , nest-building (bldg): orange, lay: yellow; incubation day 3 (inc.d3): light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, nestling care day 5 (n5): blue, and n9: light blue.}
-  \label{figure:fig2}
-\end{figure}
+associated with parental care in each tissue. Differential gene
+expression analysis can shed insight into specific differences between
+treatment groups and thus deepen our understand of the transitions
+between reproduction and parental care behaviors. We used DESEq2 to
+identify genes whose expression changed significantly between sequential
+timepoints in each tissue and sex separately (Fig. \ref{figure:fig2}a).
 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
-  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.
-a) Experimental design.
-b) t-SNE plot highlighting tissue differences in gene expression.
-Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple).
-c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males).
-t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orangse, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
-  \label{figure:fig1}
-\end{figure}
+\emph{Internal physiology versus external stimuli hypothesis}
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig4-1}
-  \caption{Another figure}
-  \label{figure:fig4}
-\end{figure}
+We show that samples from the pituitary on incubation day 17 cluster
+more closely with hatch and nestling care day 5 samples than with
+incubation days 3 and 9. This suggest that, in the pituitary, internal
+mechanisms that prepare for chick arrival have the greatest effect on
+gene expression rather than suggesting that gene expression responds to
+changes in the external enviornament.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig5-1}
-  \caption{Another another figure}
-  \label{figure:fig5}
-\end{figure}
+Blah \ref{figure:fig3}, blah, \ref{figure:fig4} blah.
 
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
+\emph{Potential implications for cancer research}
 
-Results in context to literature.
+These results confirmed prolactin (PRL) gene expression fluctuates
+throughout parental care in a manner that is consistent with its role in
+promoting lactation and maintaining parental behaviors. Interestingly,
+we identified about 100 genes whose expression was correlated with PRL,
+including (BRCA1, which is associated with breast cancer). Additionally,
+we identified thousands of genes whose expression changed over the
+course of parental care. Exploring the relationship between PRL and
+other differentially expressed genes could provide important insights
+into the \texttt{symphony} of gene expression that regulates behavior at
+the organismal and cellular level.
 
-Significance.
-
-Limitations.
+\emph{Software for data exploration and hypothesis testing}
 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/supplfig3}
-  \caption{R shiny, image 1}
-  \label{figure:supplfig3}
-\end{figure}
+\hypertarget{discussion}{%
+\subsection{Discussion}\label{discussion}}
+
+\emph{Results in context to literature. }
+
+blah blah balh
+
+\emph{Significance. }
+
+blah blah balh
+
+\emph{Limitations.}
+
+blah blah balh
 
@@
 
-\hypertarget{animal-care}{%
-\subsubsection{Animal Care}\label{animal-care}}
+\emph{Animal Care}
 
@@
 
-\hypertarget{experimental-design-and-tissue-processing}{%
-\subsubsection{Experimental design and tissue
-processing}\label{experimental-design-and-tissue-processing}}
+\emph{Experimental design and tissue processing}
 
@@
 
-\hypertarget{rna-sequencing-analysis}{%
-\subsubsection{RNA sequencing analysis}\label{rna-sequencing-analysis}}
+\emph{RNA sequencing analysis}
 
@@
 
-\hypertarget{data-availablility}{%
-\subsubsection{Data availablility}\label{data-availablility}}
+\emph{Data availablility}
 
@@
 
+\newpage
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
+  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.
+a) Experimental design.
+b) t-SNE plot highlighting tissue differences in gene expression.
+Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple).
+c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males).
+t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orange, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
+  \label{figure:fig1}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig2-1}
+  \caption{Differential gene expression differences between sequential parental stages.
+a) We calculated the differentail gene expression between each consequtive parental stage samples, for a total of 8 pair-wise comparisons for each sex in each tissue.
+The y-axis of these bar plots displays the total number of differentially expressed genes for each comparsison in the hypothalamus (b), pituitary (c), gonads (d).
+Bars are collored by the direction of the change in gene expression.
+Control: control: red or , nest-building (bldg): orange, lay: yellow; incubation day 3 (inc.d3): light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, nestling care day 5 (n5): blue, and n9: light blue.}
+  \label{figure:fig2}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig3-1}
+  \caption{Another figure}
+  \label{figure:fig3}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig4-1}
+  \caption{Another another figure}
+  \label{figure:fig4}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/supplfig3}
+  \caption{R shiny, image 1}
+  \label{figure:supplfig3}
+\end{figure}
+
+\newpage
+
+\hypertarget{author-affiliations}{%
+\subsection{Author affiliations}\label{author-affiliations}}
+
 \address{%
 Rayna M Harris \footnote{equal contribtion}\\
-UC Davis\\
+University of California, Davis\\
 \\
@@
 \address{%
-Suzanne H. Austin \footnote{equal contribtion}\\
-UC Davis\\
+Suzanne H. Austin \footnote{equal contribtion, currently at the
+  University of Oregon}\\
+University of California, Davis\\
 \\
@@
 Andrew S. Lang\\
-UNH\\
+University of New Hampshire\\
 \\
@@
 Victoria S. Farrar\\
-UC Davis\\
+University of California, Davis\\
 \\
@@
 April Booth\\
-UC Davis\\
+University of California, Davis\\
 \\
@@
 Tanner Feustel\\
-UC Davis\\
+University of California, Davis\\
 \\
@@
 Matthew D. MacManes\\
-UNH\\
+University of New Hampshire\\
 \\
@@
 Rebecca M. Calisi \footnote{corresponding author}\\
-UC Davis\\
+University of California, Davis\\
 \\
 }
-\href{mailto:author7@work}{\nolinkurl{author7@work}}
+\href{mailto:rmcalisi@ucdavis.edu}{\nolinkurl{rmcalisi@ucdavis.edu}}

### 2020-03-09 analysis/figure-3.Rmd
new figs

@@
 
-b11 <- makenewbargraph("hypothalamus", "female","eggs vs. chicks", 0, 2500) +  labs(subtitle = "females", x = NULL, y = " ") + 
-  theme( axis.text = element_blank())
-b21 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2500)  + labs(x = NULL, y = " ") + 
-  theme(axis.text = element_blank())
-b31 <- makenewbargraph("gonad", "female", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks", y = " ")    + 
-  theme(axis.text = element_blank())
+b11 <- makenewbargraph("hypothalamus", "female","eggs vs. chicks", 0, 2500) +  labs(subtitle = "females", x = NULL, y = "hypothalamus DEGs") + 
+  theme( axis.text.x = element_blank())
+b21 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2500)  + labs(x = NULL, y = "hypothalamus DEGs") + 
+  theme(axis.text.x = element_blank())
+b31 <- makenewbargraph("gonad", "female", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks", y = "hypothalamus DEGs")    + 
+  theme(axis.text.x = element_blank())
 b112131 <- plot_grid(b11,b21,b31, nrow = 3, rel_heights = c(1.1,1,1.1))
@@
 
-c11 <- makenewbargraph("hypothalamus", "female", "lo vs. hi PRL   ", 0, 2500) +  labs(subtitle = "females", x = NULL, y = "hypothalamus DEGs") + 
-  theme( axis.text.x = element_blank()) 
-c21 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2500)  + labs(x = NULL, y = "pituitary DEGs")+ 
-  theme( axis.text.x = element_blank())
-c31 <- makenewbargraph("gonad", "female", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL", y = "gonad DEGs")  + 
-  theme(axis.text.x = element_blank()) 
+c11 <- makenewbargraph("hypothalamus", "female", "lo vs. hi PRL   ", 0, 2500) +  labs(subtitle = "females", x = NULL, y = " ") + 
+  theme( axis.text = element_blank()) 
+c21 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2500)  + labs(x = NULL, y = " ")+ 
+  theme( axis.text = element_blank())
+c31 <- makenewbargraph("gonad", "female", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL", y = " ")  + 
+  theme(axis.text = element_blank()) 
 c112131 <- plot_grid(c11,c21,c31, nrow = 3, rel_heights = c(1.1,1,1.1))
@@
 
-hypothesisbars <-  plot_grid(c112131, c122232, b112131, b122232, nrow = 1, rel_widths = c(1.3, 1, 1.1, 1),
+hypothesisbars <-  plot_grid(b112131, b122232, c112131, c122232, nrow = 1, rel_widths = c(1.3, 1, 1.1, 1),
           labels = c("f", " ", "g", " "), label_size = 8)

### 2020-03-09 docs/manuscript/elife-v1.tex
new figs

@@
 
-\emph{Results in context to literature. }
-
-blah blah balh
-
-\emph{Significance. }
-
-blah blah balh
-
-\emph{Limitations.}
-
-blah blah balh
+This RNA-sequencing project was designed to characterize the
+reproductive axis of bi-parental pigeons across a characteristic
+reproductive cycle and through the parental main parental stages of egg
+incubation and parental care using data- and hypotheses driven
+approaches. We hypothesized that differences in gene expression would be
+driven by either internal physiology or external stimuli. In the
+pituitary, we found that the most explanatory source of variation in
+gene expression was the expression level of prolactin (PRL). Circulating
+levels of prolactin in the blood in this species are known to rise
+around day nine and peak on or the day before hatch. Our results confirm
+this and also identified one hundred genes whose expression is highly
+correlated with prolactin. Today, this variation, which changes more
+within a parental stage (e.g.~egg incubation or nestling care) suggests
+that internal physiology drives changes in gene expression that precede
+changes in external stimuli that lead to behavioral changes (e.g.~crop
+milk or regurgitation). We predicted that either sex steroid peptide
+hormones would play a role in governing the reproductive transitions,
+but we did not anticipate discovering a link between PRL expression to
+genes that are involved with DNA repair and regulation of the cycle.
+Finally, we aimed to make these data and analyses FAIR (findable,
+accessible, interoperable, and reproducible), open and reproducible to
+encourage others to reuse our data to test new hypotheses or explore the
+data with a different perspective or toolset. The scripts that made the
+figures and tables are browseable online thanks to RStudio, GitHub and
+Binder. The data can be visualized and sonified (turned into sound)
+interactively thanks to RStudio, GitHub, and Shiny.
 
@@
 members for software and methods discussion (Titus, Taylor, Luiz,
-Amanda, Alicia, Picaso).
+Amanda, Alicia, Picaso). We thank Pacha for help with database and
+Shiny.
+
+We thank Irene, Erica, Beth Krestoff and Tiffany Chen for assistance in
+managing the aviary and collecting nest stage data. We thank Candice
+Lee, Annie Bond, and Tanner Feustel and others?? for assistance with
+tissue collections. We thank Dr.~Fred Angelier for performing the
+prolactin RIA.This research was supported by NSF IOS 1455960 to R.M.C.
+and M.D.M. Author contributions: RMC designed the experiment. RMC and
+MDM acquired funding. SHA contributed to the experimental design by
+adding several time points. SHA, with the assistance of VF, AB, and RMC,
+conducted the data collection. SHA and a team of trained undergraduate
+students biopsy punched the hypothalami and lateral septum. ASL
+extracted the RNA, created the libraries, and conducted all early
+bioinformatic work. RMH conducted all later analyses of the genomic
+data. RMH and SHA drafted the manuscript and all other authors provided
+edits.
 

### 2020-03-09 docs/manuscript/elife_v1.Rmd
new figs

@@
 
-_Results in context to literature. _
-
-blah blah balh
-
-_Significance. _
-
-blah blah balh
-
-_Limitations._
-
-blah blah balh
+This RNA-sequencing project was designed to characterize the reproductive axis of bi-parental pigeons across a characteristic reproductive cycle and through the parental main parental stages of egg incubation and parental care using data- and hypotheses driven approaches.
+We hypothesized that differences in gene expression would be driven by either internal physiology or external stimuli.
+In the pituitary, we found that the most explanatory source of variation in gene expression was the expression level of prolactin (PRL).
+Circulating levels of prolactin in the blood in this species are known to rise around day nine and peak on or the day before hatch.
+Our results confirm this and also identified one hundred genes whose expression is highly correlated with prolactin.
+Today, this variation, which changes more within a parental stage (e.g.
+egg incubation or nestling care) suggests that internal physiology drives changes in gene expression that precede changes in external stimuli that lead to behavioral changes (e.g.
+crop milk or regurgitation).
+We predicted that either sex steroid peptide hormones would play a role in governing the reproductive transitions, but we did not anticipate discovering a link between PRL expression to genes that are involved with DNA repair and regulation of the cycle.
+Finally, we aimed to make these data and analyses FAIR (findable, accessible, interoperable, and reproducible), open and reproducible to encourage others to reuse our data to test new hypotheses or explore the data with a different perspective or toolset.
+The scripts that made the figures and tables are browseable online thanks to RStudio, GitHub and Binder.
+The data can be visualized and sonified (turned into sound) interactively thanks to RStudio, GitHub, and Shiny.
 
@@
 
-Thanks to Owen for discussion of data sonfification. Thanks to DIB lab members for software and methods discussion (Titus, Taylor, Luiz, Amanda, Alicia, Picaso).
+Thanks to Owen for discussion of data sonfification.
+Thanks to DIB lab members for software and methods discussion (Titus, Taylor, Luiz, Amanda, Alicia, Picaso).
+We thank Pacha for help with database and Shiny.
+
+We thank Irene, Erica, Beth Krestoff and Tiffany Chen for assistance in managing the aviary and collecting nest stage data.
+We thank Candice Lee, Annie Bond, and Tanner Feustel and others??
+for assistance with tissue collections.
+We thank Dr.
+Fred Angelier for performing the prolactin RIA.This research was supported by NSF IOS 1455960 to R.M.C.
+and M.D.M.
+Author contributions: RMC designed the experiment.
+RMC and MDM acquired funding.
+SHA contributed to the experimental design by adding several time points.
+SHA, with the assistance of VF, AB, and RMC, conducted the data collection.
+SHA and a team of trained undergraduate students biopsy punched the hypothalami and lateral septum.
+ASL extracted the RNA, created the libraries, and conducted all early bioinformatic work.
+RMH conducted all later analyses of the genomic data.
+RMH and SHA drafted the manuscript and all other authors provided edits.
 

### 2020-03-09 analysis/figure-X.Rmd
new ePRL figs

@@
 ```{r}
-candidategenes <- c("OXT", "AVP", "GNRH1",  "AR", "POMC", "AGRP",
-                       "CRH", "AVPR1A", "AVPR1B", "AVPR2",
-                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9")
+library(tidyverse)
+source("../R/themes.R")
+
+candidategenes <- c("OXT", "AVP", "GNRH1", "GNRHR", "CGNRH-R",
+                    "AR", "POMC", "AGRP", 
+                       "CRH", "AVPR1A", "AVPR1B", "AVPR2","VIP",
+                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9", 
+                    "ESR1","ESR2", "LBH", "CDK1", "BRCA1",
+                    "PTEN", "CREBBP", "FOS", "JUN", "EGR1",
+                     "BDNF", "GRM2","GRIA1",
+                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16", 
+                    "DIO3", "DIO2", "DIO1")
 
@@
 
-```{r vsd}
+```{r fig-x}
 geneids <- read_csv("../metadata/00_geneinfo.csv")
@@
 head(gonvsd)
+
+candidatevsd <- rbind(hypvsd, pitvsd)
+candidatevsd <- rbind(candidatevsd, gonvsd)
 ```
@@
 
-```{r}
+```{r, fig.width=5, fig.height=5}
 makeboxplots <- function(df, whichgene, myylab, whichsex){
@@
 
+plot_grid(p23, p56, p89, nrow = 3)
+
+candidatevsd$tissue = factor(candidatevsd$tissue, levels = tissuelevel)
+
+candidatevsd %>%
+    filter(treatment %in% charlevels,
+           gene %in% "PRL",
+           sex %in% sexlevels) %>%
+    filter(treatment != "control") %>%
+    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
+    geom_boxplot() + 
+    #geom_point() +
+    facet_wrap(~tissue, scales = "free_y") +
+    theme_B3() +
+    scale_fill_manual(values = allcolors) +
+    scale_color_manual(values = allcolors) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none") +
+    theme(axis.title.y = element_markdown()) +
+  labs(y = "variance-stabilized gene expression", x = "parental stage", 
+       subtitle = "PRL and ePRL expression")
+
+candidatevsd %>%
+    filter(treatment %in% charlevels,
+           gene %in% "VIP",
+           sex %in% sexlevels) %>%
+    filter(treatment != "control") %>%
+    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
+    geom_boxplot() + 
+    #geom_point() +
+    facet_wrap(~tissue, scales = "free_y") +
+    theme_B3() +
+    scale_fill_manual(values = allcolors) +
+    scale_color_manual(values = allcolors) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none") +
+    theme(axis.title.y = element_markdown()) +
+  labs(y = "variance-stabilized gene expression", x = "parental stage", 
+       subtitle = "VIP expression")
+
 ```
+
+```{r writefiles}
+write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
+```

### 2020-03-09 analysis/figure-X.Rmd
extra figs

--- 2020-02-21 analysis/figure-X.Rmd
@@
 date: "2/20/2020"
-output: html_document
+output: md_document
 ---
@@
 ```{r}
-candidategenes <- c("OXT", "AVP", "GNRH1",  "AR", "POMC", "AGRP",
-                       "CRH", "AVPR1A", "AVPR1B", "AVPR2",
-                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9")
+library(tidyverse)
+library(ggtext)
+library(cowplot)
+library(ggpubr)
+
+source("../R/themes.R")
+
+knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
+
+candidategenes <- c("OXT", "AVP", "GNRH1", "GNRHR", "CGNRH-R",
+                    "AR", "POMC", "AGRP", 
+                       "CRH", "AVPR1A", "AVPR1B", "AVPR2","VIP",
+                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9", 
+                    "ESR1","ESR2", "LBH", "CDK1", "BRCA1",
+                    "PTEN", "CREBBP", "FOS", "JUN", "EGR1",
+                     "BDNF", "GRM2","GRIA1",
+                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16", 
+                    "DIO3", "DIO2", "DIO1")
 
@@
 
-```{r vsd}
+```{r }
 geneids <- read_csv("../metadata/00_geneinfo.csv")
@@
 
-```{r}
+```{r fig-x, fig.width=7, fig.height=5}
 makeboxplots <- function(df, whichgene, myylab, whichsex){
@@
 
+plot_grid(p23, p56, p89, nrow = 3)
+
+candidatevsd$tissue = factor(candidatevsd$tissue, levels = tissuelevel)
+
+plotcandidategene <- function(whichgene, mysubtitle){
+  p <- candidatevsd %>%
+    filter(treatment %in% charlevels,
+           gene %in% whichgene,
+           sex %in% sexlevels) %>%
+    filter(treatment != "control") %>%
+    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
+    geom_boxplot() + 
+    #geom_point() +
+    facet_wrap(~tissue, scales = "free_y") +
+    theme_B3() +
+    scale_fill_manual(values = allcolors) +
+    scale_color_manual(values = allcolors) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none") +
+    theme(axis.title.y = element_markdown()) +
+  labs(y = "variance-stabilized gene expression", x = "parental stage", 
+       subtitle = mysubtitle)
+  return(p)
+
+}
+
+a <- plotcandidategene("PRL", "PRL and ePRL expression") + 
+  theme(axis.title.x = element_blank(), axis.text.x = element_blank())
+b <- plotcandidategene("VIP", "VIP expression") + theme(legend.position = "bottom") +
+  guides(fill = FALSE)
+
+plot_grid(a,b, nrow = 2, rel_heights = c(0.4,0.6))
+
+ePRL <- candidatevsd  %>%
+  filter(gene == "PRL") %>%
+  mutate(bird = sapply(strsplit(samples, '\\_'), "[", 1)) %>%
+  select(bird, sex, tissue, treatment, counts)  %>%
+  pivot_wider(names_from = tissue, values_from = counts) 
+ePRL
+
+a <- ggplot(ePRL, aes( y = hypothalamus, x = pituitary)) +
+  geom_point(aes(color = treatment)) +
+  scale_color_manual(values = allcolors) +
+  theme(legend.position = "none") +
+  geom_smooth(method = "lm", aes(color = sex)) +
+  labs(x = "pituitary PRL", y = "hypothalamic ePRL" )
+
+b <- ggplot(ePRL, aes( y = gonad, x = pituitary)) +
+  geom_point(aes(color = treatment)) +
+  scale_color_manual(values = allcolors) +
+  theme(legend.position = "none") +
+  geom_smooth(method = "lm", aes(color = sex))  +
+  labs(x = "pituitary PRL", y = "gonadal ePRL" )
+
+plot_grid(a,b)
+
 ```
+
+```{r writefiles}
+write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
+```

### 2020-03-14 analysis/22_symphony.Rmd
moving to musical genes

@@
 
-turquoisegenes <- d4 %>%
-  filter( modulecolor == "turquoise",
-          tissue == "pituitary") %>%
-  droplevels() %>% 
-    ggplot(aes(x = treatment, y = m)) +
-  geom_image(aes(image=image), size = 0.15) +
-  geom_smooth(aes(x = as.numeric(treatment)), se = F, color = "turquoise") +
-  facet_wrap(~modulecolor, scales = "free_y", nrow = 2) +
-  labs(subtitle = "WGCNA turquoise module", 
-       y = "gene expression  in the pitutiary", x = "parental stage") +
-    facet_wrap(~gene, scales = "free_y", nrow = ) +
-    theme_B3() +
-    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1),
-          strip.text = element_text(face = "italic"))
-
 redgenes <- d4 %>%
@@
     ggplot(aes(x = treatment, y = m)) +
+  geom_errorbar(aes(ymin=m-se, ymax=m+se, color=treatment), width=.1) +
   geom_image(aes(image=image), size = 0.15) +
-  geom_smooth(aes(x = as.numeric(treatment)), se = F, color = "red") +
+  #geom_smooth(aes(x = as.numeric(treatment)), se = F, color = "red") +
   facet_wrap(~modulecolor, scales = "free_y", nrow = 2) +
@@
           axis.text.x = element_text(angle = 45, hjust = 1),
-          strip.text = element_text(face = "italic"))
+          strip.text = element_text(face = "italic")) +
+  scale_color_manual(values = allcolors)
 
-turquoisegenes
 redgenes
 
-plot_grid(turquoisegenes, redgenes)
 ```

### 2020-03-24 docs/manuscript/elife-v1.tex
even better figures
new discussion and methods

@@
 this, but other say that \citep[reviewed in][]{Dulac2014}.
+
+OR
+
+Reproduction is an essential process for all living things, but
+only a subset of vertebrates and invertebrates require parental care for offspring survival.
+Identifying the mechanisms that regulate parental care behavior is like a thing for scientists in many disciplines (ecology, neuroscience, psychology), but it is hard to disentangle the mechanisms that regulate reproduction from those that regulate parental care.
+Or maybe that contrast doesn't work...
+
+Cumbia livia has been used by many scientists to study evolutionary processes and behavior stuff.
+Building on previous research, we sought to identify the transcriptional mechanisms that regulate parental care in the hypothalamic-pituitary-gonadal axis (or the reproductive axis).
+With a highly replicated experimental design, we aimed to distinguish the reproductive process from the behavioral process.
+
+Discuss parental care candidate genes and reproductive candidate genes.
+Cite some literature.
+Oxytocin and arginine vasopressin are known to regulate pair-bonding.
+Oxytocin also promotes cervical contractions.
+These all occur before birth of the offspring and therefore do not count as parental care according to that book I read.
+Oh crap, but they are also involved in pair-bonding with offspring...
+how to dissociate reproduction and parental care.....
+Prolactin is known for promoting lactation, hence it's name.
+In mammals, typically (or always?) only the female lactates to provide nutrition for the offspring thus even in bi-parental species where both parents provide parental care, only the female provides breastfeeding-related care.
+In Cumbia livia, however (maybe other species) both males and females produce "crop milk" thus both males and females can feed and care for their offspring in a way that other species cannot.
+Indeed, prolactin promotes this pseudo-lactation.
+In some species lactation promotes.
 
@@
 
+OR
+
+In the hormone experiment, we replicated the peak in prolactin.
+Then, unexpectedly or can I find this elsewhere, we confirmed that prolactin gene expression in the pituitary is strongly correlated with circulating prolactin.
+Oh, maybe do stats comparing the curve of prolactin hormone and prolactin gene expression.
+could use those same stats that I used for hormone papers or maybe just show correlations...
+
+hypothesis-driven.
+this research is hypothesis-driven...
+we had these hypotheses and came to these conclusions...
+Add GO analysis comparing observed versus expected hits for a given go term.
+\ref{figure:fig5}
+
+A data-driven analysis using PCA identifies prolactin gene expression in the pituitary is the major explainer of variation in our entire dataset.
+This leads us to ask what genes are correlated with prolactin and what genes are differentially expressed when comparing samples that had higher or lower levels of prolactin.
+We found 100 genes that mirrored the prolactin curve.
+GO analysis suggested that these genes regulate the cell cycle.
+Many are associated with breast- and testicular cancers.
+This suggests/confirms?
+that there may be a strong link between prolactin and cancer in the tissues that receive strong inputs from the pituitary (like the mammary glands, the gonads, and the adrenals).
+this research uses a readily available data-driven approach for identifying the most variable genes.
+our data-driven approach supported existing knowledge and shed light on new gene-gene interactions that need future investigation and could potentially yield new insight into unsolved problems in basic biology and biomedicine.
+
+Our software tool provides a unique oportunity for professionals with differnt levels of proficiency in bioinformatics to itneract with the data to advance science and medicien.
+Futuremore, we turn our data into sounds, providing a unique insight into transcripional variation that cannot be captured with the traditional data viauzliations that are the foundation of omics science communication.
+
 \hypertarget{materials-and-methods}{%
@@
 \citep{project_jupyter-proc-scipy-2018}.
+
+\emph{Findable, accessible, interoperable, and reusable}
+
+In addition to making our data and code open by sharing it in open-access repositories (e.g GitHub and that repo were are using that hopefully, Andrew has the link to), we also wanted to make it findable, accessible, and interoperable, and reusable (the tenants of FAIR data) \citep{Wilkinson2016}.
+Our first step in doing this was to create a "binderized" \citep{project_jupyter-proc-scipy-2018} GitHub repository where the R scripts can be executed in an RStudio server on the cloud.
+This means that no local installations or downloads are required, ensuring (or at least increasing the likelihood of) interoperability and reproducibility.
+To make the data more reusable and accessible to those familiar and unfamiliar with R, we created Shiny application (define) that allows users to interactively pick a gene of interest and visualize their expression level in this data set.
+(If Rayna adds a download button, or a new shiny tab then the can get the raw data too.) This saves researcher hours of computing time downloading files and running scripts if they had to start from the links we provided above to the raw data and scripts.
+Finally, we musical representations of the data to our software so that uses who are not familiar with interpreting the complex graphs typically associated with the RNA-seq project can immerse themselves in data exploration.
+We call this software application "Musical Genes".
+We believe our approach is a novel way of communicating scientific information to broad audiences \ref{figure:fig6}.
 
@@
   \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/supplfig3}
-  \caption{R shiny, image 1}
-  \label{figure:supplfig3}
+  \includegraphics[width=1.0\textwidth]{../../figures/images/fig_fig5}
+  \caption{Hypothesis-driven research, data exploration and software development.
+We integrate hypothesis driven experimental testing with data-driven exploration and sofware develop to make new discoveries and advance our understanding reproduction and parental care.}
+  \label{figure:fig5}
 \end{figure}
 
-\newpage
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/supplfig3}
+  \caption{Shiny Musical Genes. Musical Genes is a R-based Shiny application for vizualizing and sonifying gene expression associated with transitions through reproduction and parental care.}
+  \label{figure:fig6}
+\end{figure}
 

### 2020-03-25 analysis/12_GO.Rmd
new GO anlysis, new candidate gnee list

@@
 ```{r}
-candidates <- c("PRL", "PRLR", 
-                 "AVP", "VIP", "AVPR1A", "AVPR1B", "OXT",  "VIPR1", "VIPR2",
-                "GNRH1", "NPY", "LH", "FSH",
-                "CRH", "FSHB", "GH",  "JAK2", "HCRT", "TRH", "TSHB",
+candidates <- c("OXT", "AVP", "GNRH1", "GNRHR", 
+                    "AR", "POMC", "AGRP", 
+                       "CRH", "AVPR1A", "AVPR1B", "AVPR2","VIP",
+                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9", 
+                    "ESR1","ESR2", "LBH", "CDK1", "BRCA1",
+                    "PTEN", "CREBBP", "FOS", "JUN", "EGR1",
+                     "BDNF", "GRM2","GRIA1",
+                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16", 
+                    "DIO3", "DIO2", "DIO1",
+
+                     "VIPR1", "VIPR2", "NPY", "LH", "FSH", "FSHB",
+                    "JAK2", "HCRT", "TRH", "TSHB",
                     "MC3R", "MC4R", "MC5R",  "NR3C1", "NR3C2",
-                "ESR1", "ESR2", "AR", 
-                "STAT5B", "CISH",
-                "GNRHR", "NPVF")
+                    "STAT5B","NPVF")
 
@@
 
-DEG_path <- "../results/"   # path to the data
-DEG_files <- dir(DEG_path, pattern = "03_DEGs.*") # get file names
+DEG_path <- "../results/DEseq2/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs.csv") # get file names
 DEG_pathfiles <- paste0(DEG_path, DEG_files)
@@
   map_df(~read_csv(.x), .id = "file_name") %>%
-  mutate(comparison = sapply(strsplit(as.character(file_name),'../results/03_DEGs.'), "[", 2)) %>% 
-  mutate(comparison = sapply(strsplit(as.character(comparison),'.csv'), "[", 1)) %>% 
+  mutate(comparison = sapply(strsplit(as.character(file_name),'../results/DEseq2/*'), "[", 2)) %>% 
+  mutate(comparison = sapply(strsplit(as.character(comparison),'_DEGs.csv'), "[", 1)) %>% 
   select(gene, comparison)
@@
   pivot_wider(names_from = comparison, values_from = genes) %>% 
-  select(GO, contains("control.bldg"), contains("bldg.lay"), contains("lay.inc.d3"),
-         contains("inc.d3.inc.d9"), contains("inc.d9.inc.d17"), contains("inc.d17.hatch"),
-         contains("hatch.n5"), contains("n5.n9")) %>%
+  select(GO, contains("control_bldg"), contains("bldg_lay"), contains("lay_inc.d3"),
+         contains("inc.d3_inc.d9"), contains("inc.d9_inc.d17"), contains("inc.d17_hatch"),
+         contains("hatch_n5"), contains("n5_n9")) %>%
   filter(GO %in% c("endocrineprocess", "lactation", 
@@
   pivot_wider(names_from = comparison, values_from = genes) %>% 
-  select(GO, contains("control.bldg"), contains("bldg.lay"), contains("lay.inc.d3"),
-         contains("inc.d3.inc.d9"), contains("inc.d9.inc.d17"), contains("inc.d17.hatch"),
-         contains("hatch.n5"), contains("n5.n9"))
+  select(GO, contains("control_bldg"), contains("bldg_lay"), contains("lay_inc.d3"),
+         contains("inc.d3_inc.d9"), contains("inc.d9_inc.d17"), contains("inc.d17_hatch"),
+         contains("hatch_n5"), contains("n5_n9"))
 candidate_DEGs <- as.data.frame(candidate_DEGs)
+
 row.names(candidate_DEGs) <- candidate_DEGs$GO  
@@
 candidate_DEGs <- as.data.frame(candidate_DEGs)
-candidate_DEGs <- tail(candidate_DEGs,10)
 kable(candidate_DEGs)
 
-# Top Go terms: reponse to keytone, response to hormone, repsponse to organonitrogen compound, positive regulation of homeostatic process, maternal beahvior
+bioprocshinyGO <- c("topBiologicalProcessGO",
+                    "Prostate gland growth",
+                    "Cellular response to hormone stimulus",
+                    "Response to endogenous stimulus",
+                    "Mammary gland development",
+                    "Response to hormone",
+                    "Response to endogenous stimulus",
+                    NA,
+                    "Regulation of animal organ morphogenesis",
+                    "Temperature homeostasis",
+                    "Prostate gland growth",
+                    NA,NA,NA,
+                    "Cellular response to hormone stimulus",
+                    NA,
+                    "Regulation of cell communication",
+                    NA,
+                    "Cell-cell signaling",
+                    "Cellular response to hormone stimulus",
+                    "Positive regulation of ossification",
+                    "G2 DNA damage checkpoint"
+                    )
+
+candidate_DEGs <- cbind(candidate_DEGs, bioprocshinyGO)
+kable(candidate_DEGs)
 

### 2020-03-25 analysis/figure-X.Rmd
new GO anlysis, new candidate gnee list

@@
                     "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16", 
-                    "DIO3", "DIO2", "DIO1")
+                    "DIO3", "DIO2", "DIO1",
+
+                     "VIPR1", "VIPR2", "NPY", "LH", "FSH", "FSHB",
+                    "JAK2", "HCRT", "TRH", "TSHB",
+                    "MC3R", "MC4R", "MC5R",  "NR3C1", "NR3C2",
+                    "STAT5B","NPVF")
 

### 2020-03-26 analysis/figure-2.Rmd
new table 1, now with direction and control

@@
 candidategenes <- c("OXT", "AVP", "GNRH1", "GNRHR", "CGNRH-R",
-                    "AR", "POMC", "AGRP", 
-                       "CRH", "AVPR1A", "AVPR1B", "AVPR2","VIP",
-                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9", 
-                    "ESR1","ESR2", "LBH", "CDK1", "BRCA1",
-                    "PTEN", "CREBBP", "FOS", "JUN", "EGR1",
-                     "BDNF", "GRM2","GRIA1",
-                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16", 
-                    "DIO3", "DIO2", "DIO1")
+                    "AR",  "CYP19A1", 
+                     "AVPR1A", "AVPR1B", "AVPR2","VIP",
+                  "DRD1", "DRD2", 
+                  "PRL", "PRLR",  
+                    "ESR1","ESR2", "LBH",
+
+                    "FOS", "JUN", "EGR1", "BDNF"
+                      )
 
 table1 <- allDEG %>%
-  filter(gene %in% candidategenes,
-         comparison != "control_bldg") %>%
-    arrange(gene) %>%
+  mutate(updown = ifelse(lfc > 0, "+", "-"))  %>%
+  mutate(geneupdown = paste(gene, updown, sep = "")) %>%
+  filter(gene %in% candidategenes) %>%
+  arrange(geneupdown) %>%
   group_by(sex, tissue, comparison) %>%
-  summarize(genes = str_c(gene, collapse = " ")) %>%
+  summarize(genes = str_c(geneupdown, collapse = " ")) %>%
   pivot_wider(names_from = comparison, values_from = genes ) %>%
-  select(sex, tissue, bldg_lay, lay_inc.d3, inc.d3_inc.d9,
-        inc.d9_inc.d17,inc.d17_hatch, hatch_n5, n5_n9)  %>%
+  select(sex, tissue, control_bldg, bldg_lay, lay_inc.d3, inc.d3_inc.d9,
+        inc.d9_inc.d17, hatch_n5)  %>%
   arrange( tissue, sex)

### 2020-03-26 analysis/figure-2.Rmd
one less candidate gene

@@
 
-candidategenes <- c("OXT", "AVP", "GNRH1", "GNRHR", "CGNRH-R",
+candidategenesslim <- c("OXT", "AVP", "GNRH1", "GNRHR", 
                     "AR",  "CYP19A1", 

### 2020-03-27 analysis/figure-4.Rmd
reorder figures

@@
 ---
-title: "Fig5"
+title: "Fig4"
 output: md_document
@@
 
-```{r setup, include=FALSE}
+# Figure 3
+
+```{r setup}
 library(tidyverse)
 library(cowplot)
-library(Rtsne)
+library(ggimage)
+library(apaTables)
+library(factoextra)
 
@@
 
-knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
+knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
 ```
 
-# Figure 1. Experimental design, tSNE analysis, PCA, and prolactin
+## Data: PCA, hormones, PRL expression
 
-```{r tSNE-df}
+```{r PRLfoo}
+# pca
+charpca <- subsetmakepca("pituitary", charlevels, sexlevels)	
+charfviz <- makefvizdf("pituitary", charlevels, sexlevels)
 
-# prep for tsne
+# hormones
+prolactin <- read_csv("../results/07_hormones.csv") %>%
+    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
+    droplevels() 
+prolactin$treatment <- factor(prolactin$treatment, levels = alllevels)
 
-hyptsne <- subsetmaketsne("hypothalamus", allmaniplevels, sexlevels)
-pittsne <- subsetmaketsne("pituitary", allmaniplevels, sexlevels)
-gontsne <- subsetmaketsne("gonads", allmaniplevels, sexlevels)
+# PRL vsd
 
-addgroupings <- function(df){
+vsd_path <- "../results/DEseq2/"   # path to the data
+vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
+vsd_pathfiles <- paste0(vsd_path, vsd_files)
+vsd_files
 
-  df <- df %>% mutate(hiloPRL = fct_collapse(treatment, 
-                                  lo = c("m.inc.d8", "m.inc.d3", "m.inc.d9", "inc.d3", "inc.d9", "lay", "bldg"),
-                                  hi = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" ,  "m.n2", "extend", "n5")),
-                extint = fct_collapse(treatment, 
-                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
-                                  chicks = c( "m.inc.d8", "hatch", "extend", "n5"),
-                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")))
-  df$extint <- factor(df$extint, levels = levelsextint)
-  return(df)
-}
+PRLvsd <- vsd_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name")  %>% 
+  dplyr::rename("gene" = "X1") %>% 
+  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "samples", values_to = "counts") %>%
+  filter(gene == "PRL")  %>%
+  dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
+  dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
+  dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
+  dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
+  dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
+  drop_na()
+PRLvsd$treatment <- factor(PRLvsd$treatment, levels = alllevels)
 
-hyptsne <- addgroupings(hyptsne)
-pittsne <- addgroupings(pittsne)
-gontsne <- addgroupings(gontsne)
+PRLhyp <- PRLvsd %>% filter(tissue == "hypothalamus") 
+PRLpit <- PRLvsd %>% filter(tissue == "pituitary")
+PRLgon <- PRLvsd %>% filter(tissue == "gonad")
+```
+
+```{r determinePRLhiglo, fig.height = 6, fig.width=3}
+PRLvsd2 <- PRLvsd %>%
+  filter(tissue == "pituitary")
+head(PRLvsd2)
+
+PRLvsd2 %>%
+  group_by(sex) %>%
+  summarize(median = median(counts))
+
+PRLvsd3 <- PRLvsd2 %>%
+  mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
+  drop_na()
+PRLvsd3$hiloPRL <- factor(PRLvsd3$hiloPRL, levels = c("lo", "hi"))
+
+PRLvsd3 %>%
+  group_by(sex, tissue, hiloPRL) %>%
+  summarize(n = n())
 
@@
 
-## make figure
+## Internal versus external
+
+```{r hypothesis}
+DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG2 <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
+head(allDEG2)
+
+allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)
+
+allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
+allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
+                                                      "eggs vs. chicks" = "eggs_chicks"))
+allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))
+
+```
 
@@
 
-expdesign2 <- png::readPNG("../figures/images/fig_fig4a.png")
-expdesign2 <- ggdraw() +  draw_image(expdesign2, scale = 1)
+expdesign <- png::readPNG("../figures/images/fig_fig3a.png")
+expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
 
-b <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus ~ parental stage", x = NULL)  + facet_wrap(~sex, scales = "free")
-c <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary ~ parental stage", x = NULL) + facet_wrap(~sex, scales = "free") 
-d <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads ~ parental stage")  + facet_wrap(~sex, scales = "free")
+b <- plotcolorfulpcs(charpca,charpca$treatment, allcolors) + labs(subtitle = " ") +
+  theme(legend.position = "none", 
+        legend.direction = "horizontal", 
+        legend.key.size = unit(0.5, 'lines')) + 
+  guides(color = FALSE) +
+  labs(subtitle = "pituitary")
 
-bcd <- plot_grid(b,c,d, nrow = 3, labels = c("b"), label_size = 8, rel_heights = c(1,1,1))
+c <- plotprolactin(PRLpit, PRLpit$counts, "PRL", "pituitary") + 
+  theme(legend.position = "none", 
+         axis.text.x = element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(face = "italic"))
 
-e <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + labs(subtitle = "hypothalamus ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")  
-f <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + labs(subtitle = "pituitary ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")   
-g <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + labs(subtitle = "gonads ~ internal physiology" ) + facet_wrap(~sex, scales = "free")
+d <- plotprolactin(prolactin, prolactin$plasma_conc, "prolactin (ng/mL)", "blood") +
+  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
 
-efg <- plot_grid(e,f,g, nrow = 3, labels = c("d"), label_size = 8, rel_heights = c(1,1,1))
+bcd <- plot_grid(b,c,d, labels = c("b", "c", "d"), ncol = 1, label_size = 8, rel_heights = c(1,1,1.3))
 
-h <- plottsneelipsev2(hyptsne, hyptsne$extint, allcolors ) + labs(subtitle = "hypothalamus ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")  
-i <- plottsneelipsev2(pittsne, pittsne$extint, allcolors ) + labs(subtitle = "pituitary ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")   
-j <- plottsneelipsev2(gontsne, gontsne$extint, allcolors ) + labs(subtitle = "gonads ~ external enviornment" ) + facet_wrap(~sex, scales = "free")
+makenewbargraph <- function(whichtissue, whichsex,  whichcomparison, lowlim, higherlim){
+  p <- allDEG2 %>%
+    filter(tissue == whichtissue,
+           comparison == whichcomparison,
+           sex == whichsex) %>%
+    ggplot(aes(x = comparison,  fill = direction)) +
+    geom_bar(position = "dodge", drop = FALSE) +
+    theme_B3() +
+    theme(legend.position = "none")  +
+    guides(fill = guide_legend(nrow = 1)) +
+    labs( y = whichtissue) +
+  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
+           position = position_dodge(width = 1),
+           size = 2, color = "black")  + 
+      ylim(lowlim, higherlim) +
+  scale_fill_manual(values = allcolors, name = "higher in")   + 
+    theme(axis.text.x = element_blank()) 
+  return(p)
+}
 
-hij <- plot_grid(h,i,j, nrow = 3, labels = c("c"), label_size = 8, rel_heights = c(1,1,1))
+b11 <- makenewbargraph("hypothalamus", "female","eggs vs. chicks", 0, 2500) +  labs(subtitle = "females", x = NULL, y = "hypothalamus DEGs") + 
+  theme( axis.text.x = element_blank())
+b21 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2500)  + labs(x = NULL, y = "hypothalamus DEGs") + 
+  theme(axis.text.x = element_blank())
+b31 <- makenewbargraph("gonad", "female", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks", y = "hypothalamus DEGs")    + 
+  theme(axis.text.x = element_blank())
+b112131 <- plot_grid(b11,b21,b31, nrow = 3, rel_heights = c(1.1,1,1.1))
 
-bcdhijefg <- plot_grid(bcd, hij,  efg, nrow = 1)
+b12 <- makenewbargraph("hypothalamus", "male",  "eggs vs. chicks", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
+  theme(axis.title.y = element_blank(), axis.text = element_blank())
+b22 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2500) + labs(x = NULL)+ 
+  theme(axis.title.y = element_blank(), axis.text = element_blank())
+b32 <- makenewbargraph("gonad", "male", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks")   + 
+  theme(axis.title.y = element_blank(), axis.text = element_blank())
+b122232 <- plot_grid(b12, b22,b32, nrow = 3, rel_heights = c(1.1,1,1.1))
 
-fig5 <- plot_grid(expdesign2, bcdhijefg, nrow = 2, labels = c("a"), label_size = 8, rel_heights = c(0.25,1))
-fig5
+c11 <- makenewbargraph("hypothalamus", "female", "lo vs. hi PRL   ", 0, 2500) +  labs(subtitle = "females", x = NULL, y = " ") + 
+  theme( axis.text = element_blank()) 
+c21 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2500)  + labs(x = NULL, y = " ")+ 
+  theme( axis.text = element_blank())
+c31 <- makenewbargraph("gonad", "female", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL", y = " ")  + 
+  theme(axis.text = element_blank()) 
+c112131 <- plot_grid(c11,c21,c31, nrow = 3, rel_heights = c(1.1,1,1.1))
+
+c12 <- makenewbargraph("hypothalamus", "male",  "lo vs. hi PRL   ", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
+  theme(axis.title.y = element_blank(), axis.text = element_blank())
+c22 <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = NULL)+ 
+  theme(axis.title.y = element_blank(), axis.text = element_blank())
+c32 <- makenewbargraph("gonad", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL")   + 
+  theme(axis.title.y = element_blank(), axis.text = element_blank())
+c122232 <- plot_grid(c12, c22,c32, nrow = 3, rel_heights = c(1.1,1,1.1))
+
+hypothesisbars <-  plot_grid(b112131, b122232, c112131, c122232, nrow = 1, rel_widths = c(1.3, 1, 1.1, 1),
+          labels = c("f", " ", "g", " "), label_size = 8)
+
+dataplots <- plot_grid(bcd, hypothesisbars)
+
+fig4 <- plot_grid(expdesign, dataplots, rel_heights = c(0.25,1), nrow = 2, labels = c("a"), label_size = 8)
+fig4
 
 ```
+
+```{r savefiles}
+#write.csv(PRLvsd3,"../results/PRLvsd.csv", row.names = F)
+
+```

### 2020-03-27 analysis/figure-3.Rmd
reorder figures

--- 2020-03-07 analysis/figure-3.Rmd
@@
 title: "Fig3"
+author: "Rayna M Harris"
+date: "3/26/2020"
 output: md_document
@@
 
-# Figure 3
+# all things box plots
 
-```{r setup}
+## candidate genes
+
+```{r}
 library(tidyverse)
+library(ggtext)
 library(cowplot)
-library(ggimage)
-library(apaTables)
-library(factoextra)
+library(ggpubr)
 
 source("../R/themes.R")
-source("../R/functions.R")
-source("../R/wrangledata.R")
 
-knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
+
+candidategenes <- c("OXT", "AVP", "GNRH1", "GNRHR", "CGNRH-R",
+                    "AR", "POMC", "AGRP", 
+                       "CRH", "AVPR1A", "AVPR1B", "AVPR2","VIP",
+                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9", 
+                    "ESR1","ESR2", "LBH", "CDK1", "BRCA1",
+                    "PTEN", "CREBBP", "FOS", "JUN", "EGR1",
+                     "BDNF", "GRM2","GRIA1",
+                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16", 
+                    "DIO3", "DIO2", "DIO1",
+
+                     "VIPR1", "VIPR2", "NPY", "LH", "FSH", "FSHB",
+                    "JAK2", "HCRT", "TRH", "TSHB",
+                    "MC3R", "MC4R", "MC5R",  "NR3C1", "NR3C2",
+                    "STAT5B","NPVF")
+
+candidategenesslim <- c("OXT", "AVP", "GNRH1", "GNRHR", 
+                    "AR",  "CYP19A1", 
+                     "AVPR1A", "AVPR1B", "AVPR2","VIP",
+                  "DRD1", "DRD2", 
+                  "PRL", "PRLR",  
+                    "ESR1","ESR2", "LBH",
+
+                    "FOS", "JUN", "EGR1", "BDNF"
+                      )
+
 ```
 
-## Data: PCA, hormones, PRL expression
+## variance stabilized gene expression  (vsd)
 
-```{r PRLfoo}
-# pca
-charpca <- subsetmakepca("pituitary", charlevels, sexlevels)	
-charfviz <- makefvizdf("pituitary", charlevels, sexlevels)
-
-# hormones
-prolactin <- read_csv("../results/07_hormones.csv") %>%
-    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
-    droplevels() 
-prolactin$treatment <- factor(prolactin$treatment, levels = alllevels)
-
-# PRL vsd
+```{r }
+geneids <- read_csv("../metadata/00_geneinfo.csv")
 
@@
 
-PRLvsd <- vsd_pathfiles %>%
+allvsd <- vsd_pathfiles %>%
   setNames(nm = .) %>% 
@@
   pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
-               names_to = "samples", values_to = "counts") %>%
-  filter(gene == "PRL")  %>%
+               names_to = "samples", values_to = "counts")
+
+getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
+  candidates  <- allvsd %>%
+    filter(gene %in% whichgenes) %>%
     dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
@@
     dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
+    filter(tissue == whichtissue, sex %in% whichsex)  %>%
     drop_na()
-PRLvsd$treatment <- factor(PRLvsd$treatment, levels = alllevels)
+  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
+  return(candidates)
+}
 
-PRLhyp <- PRLvsd %>% filter(tissue == "hypothalamus") 
-PRLpit <- PRLvsd %>% filter(tissue == "pituitary")
-PRLgon <- PRLvsd %>% filter(tissue == "gonad")
+hypvsd <- getcandidatevsd(candidategenesslim, "hypothalamus", sexlevels)
+pitvsd <- getcandidatevsd(candidategenesslim, "pituitary", sexlevels)
+gonvsd <- getcandidatevsd(candidategenesslim, "gonad", sexlevels)
+head(hypvsd)
+head(pitvsd)
+head(gonvsd)
+
+candidatevsd <- rbind(hypvsd, pitvsd)
+candidatevsd <- rbind(candidatevsd, gonvsd)
+
+unique(candidatevsd$gene)
 ```
 
-```{r determinePRLhiglo, fig.height = 6, fig.width=3}
-PRLvsd2 <- PRLvsd %>%
-  filter(tissue == "pituitary")
-head(PRLvsd2)
+## Figs
 
-PRLvsd2 %>%
-  group_by(sex) %>%
-  summarize(median = median(counts))
+```{r fig3, fig.width=7, fig.height=7}
+makeboxplots <- function(df, whichgene, mysubtitle, whichsex, whichtimepoint){
+  p <- df %>%
+    filter(treatment %in% charlevels,
+           gene %in% whichgene,
+           sex %in% whichsex) %>%
+    filter(treatment %in% whichtimepoint) %>%
+    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
+    geom_boxplot(outlier.shape = NA) + 
+    geom_jitter(size = 0.5, width = 0.1) +
+    facet_wrap(~gene, scales = "free_y", nrow = 1) +
+    theme_B3() +
+    scale_fill_manual(values = allcolors) +
+    scale_color_manual(values = allcolors) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none") +
+    labs(y = "expression" , x = "Parental stages", subtitle = mysubtitle) +
+    theme(strip.text = element_text(face = "italic"))
+  return(p)
+}
 
-PRLvsd3 <- PRLvsd2 %>%
-  mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
-  drop_na()
-PRLvsd3$hiloPRL <- factor(PRLvsd3$hiloPRL, levels = c("lo", "hi"))
+## hyp 
+a <- makeboxplots(hypvsd, c("BDNF","CYP19A1", "DRD1", "EGR1"), 
+                  "Female hypothalamus", "female", c("hatch", "n5"))  +
+  theme(axis.title.x = element_blank())
 
-PRLvsd3 %>%
-  group_by(sex, tissue, hiloPRL) %>%
-  summarize(n = n())
+c <- makeboxplots(pitvsd, c("ESR1","GNRHR"), "Female pituitary", "female", c("bldg", "lay", "inc.d3"))   +
+  theme(axis.title.x = element_blank())
+d <- makeboxplots(pitvsd, c("LBH", "PRL" ), " ", "female", c("inc.d9", "inc.d17", "hatch", "n5"))  +
+  theme(axis.title  = element_blank())
+e <- makeboxplots(pitvsd, c( "AVPR2"), " ", "female", c( "hatch", "n5"))  +
+  theme(axis.title  = element_blank())
+
+cde <- plot_grid(c,d, e, rel_widths = c(2.5,2.5,1), nrow = 1)
+
+g <- makeboxplots(gonvsd, c( "AVPR1A"), "Female gonads", "female", c( "lay", "inc.d3", "inc.d9"))  +
+  theme(axis.title.x = element_blank())
+h <- makeboxplots(gonvsd, c( "EGR1", "FOS", "PRLR"), " ", "female", c( "lay", "inc.d3"))  +
+  theme(axis.title = element_blank())
+
+gh <- plot_grid(g,h,nrow = 1, rel_widths = c(1,2.5))
+
+b <- makeboxplots(hypvsd, c("AR"),"Male hypothalamic expression", "male", 
+                  c( "inc.d9", "inc.d17"))  + labs(caption = " ")
+f <- makeboxplots(pitvsd, c("LBH", "PRL"), "Male pituitary expression", "male", 
+                  c("inc.d9", "inc.d17")) + 
+  theme(axis.title.y  = element_blank(),
+        plot.caption = element_text(face = "italic")) + 
+  labs(caption = "candidate genes: AR, AVP, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, EGR1, ESR1, ESR2, FOS, GNRH1, GNRHR, JUN, LBH, OXT, PRL, PRLR, VIP")
+
+bf <- plot_grid(b,f,nrow = 1, rel_widths = c(1,2))
+
+plot_grid(a, cde, gh, bf, nrow = 4, rel_heights = c(1,1,1,1.2), labels = "auto", label_size = 8)
 
@@
 
-## Internal versus external
-
-```{r hypothesis}
-DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
-DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
-DEG_pathfiles <- paste0(DEG_path, DEG_files)
-#DEG_files
-
-allDEG2 <- DEG_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name") %>% 
-  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
-  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
-  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
-  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
-  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
-  mutate(comparison = paste(down,up, sep = "_")) %>%
-  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
-  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
-dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
-head(allDEG2)
-
-allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)
-
-allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
-allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
-                                                      "eggs vs. chicks" = "eggs_chicks"))
-allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))
-
+```{r writefiles}
+write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
 ```
-
-```{r fig3, fig.width=7, fig.height=7}
-
-expdesign <- png::readPNG("../figures/images/fig_fig3a.png")
-expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
-
-b <- plotcolorfulpcs(charpca,charpca$treatment, allcolors) + labs(subtitle = " ") +
-  theme(legend.position = "none", 
-        legend.direction = "horizontal", 
-        legend.key.size = unit(0.5, 'lines')) + 
-  guides(color = FALSE) +
-  labs(subtitle = "pituitary")
-
-c <- plotprolactin(PRLpit, PRLpit$counts, "PRL", "pituitary") + 
-  theme(legend.position = "none", 
-         axis.text.x = element_blank(),
-        axis.title.x = element_blank(),
-        axis.title.y = element_text(face = "italic"))
-
-d <- plotprolactin(prolactin, prolactin$plasma_conc, "prolactin (ng/mL)", "blood") +
-  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
-
-bcd <- plot_grid(b,c,d, labels = c("b", "c", "d"), ncol = 1, label_size = 8, rel_heights = c(1,1,1.3))
-
-makenewbargraph <- function(whichtissue, whichsex,  whichcomparison, lowlim, higherlim){
-  p <- allDEG2 %>%
-    filter(tissue == whichtissue,
-           comparison == whichcomparison,
-           sex == whichsex) %>%
-    ggplot(aes(x = comparison,  fill = direction)) +
-    geom_bar(position = "dodge", drop = FALSE) +
-    theme_B3() +
-    theme(legend.position = "none")  +
-    guides(fill = guide_legend(nrow = 1)) +
-    labs( y = whichtissue) +
-  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
-           position = position_dodge(width = 1),
-           size = 2, color = "black")  + 
-      ylim(lowlim, higherlim) +
-  scale_fill_manual(values = allcolors, name = "higher in")   + 
-    theme(axis.text.x = element_blank()) 
-  return(p)
-}
-
-b11 <- makenewbargraph("hypothalamus", "female","eggs vs. chicks", 0, 2500) +  labs(subtitle = "females", x = NULL, y = " ") + 
-  theme( axis.text = element_blank())
-b21 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2500)  + labs(x = NULL, y = " ") + 
-  theme(axis.text = element_blank())
-b31 <- makenewbargraph("gonad", "female", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks", y = " ")    + 
-  theme(axis.text = element_blank())
-b112131 <- plot_grid(b11,b21,b31, nrow = 3, rel_heights = c(1.1,1,1.1))
-
-b12 <- makenewbargraph("hypothalamus", "male",  "eggs vs. chicks", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
-  theme(axis.title.y = element_blank(), axis.text = element_blank())
-b22 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2500) + labs(x = NULL)+ 
-  theme(axis.title.y = element_blank(), axis.text = element_blank())
-b32 <- makenewbargraph("gonad", "male", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks")   + 
-  theme(axis.title.y = element_blank(), axis.text = element_blank())
-b122232 <- plot_grid(b12, b22,b32, nrow = 3, rel_heights = c(1.1,1,1.1))
-
-c11 <- makenewbargraph("hypothalamus", "female", "lo vs. hi PRL   ", 0, 2500) +  labs(subtitle = "females", x = NULL, y = "hypothalamus DEGs") + 
-  theme( axis.text.x = element_blank()) 
-c21 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2500)  + labs(x = NULL, y = "pituitary DEGs")+ 
-  theme( axis.text.x = element_blank())
-c31 <- makenewbargraph("gonad", "female", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL", y = "gonad DEGs")  + 
-  theme(axis.text.x = element_blank()) 
-c112131 <- plot_grid(c11,c21,c31, nrow = 3, rel_heights = c(1.1,1,1.1))
-
-c12 <- makenewbargraph("hypothalamus", "male",  "lo vs. hi PRL   ", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
-  theme(axis.title.y = element_blank(), axis.text = element_blank())
-c22 <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = NULL)+ 
-  theme(axis.title.y = element_blank(), axis.text = element_blank())
-c32 <- makenewbargraph("gonad", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL")   + 
-  theme(axis.title.y = element_blank(), axis.text = element_blank())
-c122232 <- plot_grid(c12, c22,c32, nrow = 3, rel_heights = c(1.1,1,1.1))
-
-hypothesisbars <-  plot_grid(c112131, c122232, b112131, b122232, nrow = 1, rel_widths = c(1.3, 1, 1.1, 1),
-          labels = c("f", " ", "g", " "), label_size = 8)
-
-dataplots <- plot_grid(bcd, hypothesisbars)
-
-fig3 <- plot_grid(expdesign, dataplots, rel_heights = c(0.25,1), nrow = 2, labels = c("a"), label_size = 8)
-fig3
-
-```
-
-```{r savefiles}
-#write.csv(PRLvsd3,"../results/PRLvsd.csv", row.names = F)
-
-```

### 2020-03-27 analysis/figure-5.Rmd
reorder figures

--- 2020-02-11 analysis/figure-5.Rmd
@@
 library(cowplot)
+library(Rtsne)
+
 source("../R/themes.R")
+source("../R/functions.R")
+source("../R/wrangledata.R")
 
-knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
+knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
 ```
 
-## DEGs
+# Figure 1. Experimental design, tSNE analysis, PCA, and prolactin
 
-```{r fig2, fig.width=7}
+```{r tSNE-df}
 
-DEG_path <- "../results/DEseq2/manip/"   # path to the data
-DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
-DEG_pathfiles <- paste0(DEG_path, DEG_files)
-DEG_files
+# prep for tsne
 
-allDEG <- DEG_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name") %>% 
-  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/manip/'), "[", 2))  %>% 
-  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
-  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
-  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
-  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
-  mutate(comparison = paste(down,up, sep = "_")) %>%
-  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
-  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
-dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
-head(allDEG)
+hyptsne <- subsetmaketsne("hypothalamus", allmaniplevels, sexlevels)
+pittsne <- subsetmaketsne("pituitary", allmaniplevels, sexlevels)
+gontsne <- subsetmaketsne("gonads", allmaniplevels, sexlevels)
 
-comparisonlevels <- c( "inc.d3_m.inc.d3" ,  "inc.d9_m.inc.d9", 
-                        "inc.d17_m.inc.d17", "m.n2_hatch" ,
-                        "inc.d9_m.inc.d8", "inc.d17_prolong", "hatch_extend"  )
+addgroupings <- function(df){
 
-allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
-allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
-allDEG$direction <- factor(allDEG$direction, levels = manipulationsandcontrols)
+  df <- df %>% mutate(hiloPRL = fct_collapse(treatment, 
+                                  lo = c("m.inc.d8", "m.inc.d3", "m.inc.d9", "inc.d3", "inc.d9", "lay", "bldg"),
+                                  hi = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" ,  "m.n2", "extend", "n5")),
+                extint = fct_collapse(treatment, 
+                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
+                                  chicks = c( "m.inc.d8", "hatch", "extend", "n5"),
+                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")))
+  df$extint <- factor(df$extint, levels = levelsextint)
+  return(df)
+}
 
-allDEG %>%
-  group_by(sex, tissue, comparison) %>%
-  summarize(n = n())
-
-allDEG
-
-allDEG %>%
-ggplot(aes(x = comparison,  fill = direction)) +
-  geom_bar(position = "dodge") +
-  facet_grid(tissue~sex) +
-  theme_B3() +
-  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
-         axis.text.x = element_blank(),
-        legend.position = "none",
-        strip.placement = "right")  +
-  guides(fill = guide_legend(nrow = 1)) +
-  labs(x = "Parental stage-specific manipulations", y = NULL) +
-  scale_fill_manual(values = allcolors,
-                       name = " ", drop = FALSE) +
-  scale_color_manual(values = sexcolors) +
-  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
-            position = position_dodge(width = 1),
-            size = 2.5, color = "black")
+hyptsne <- addgroupings(hyptsne)
+pittsne <- addgroupings(pittsne)
+gontsne <- addgroupings(gontsne)
 
 ```
+
+## make figure
+
+```{r fig5, fig.width=7, fig.height=7}
+
+expdesign2 <- png::readPNG("../figures/images/fig_fig4a.png")
+expdesign2 <- ggdraw() +  draw_image(expdesign2, scale = 1)
+
+b <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus ~ parental stage", x = NULL)  + facet_wrap(~sex, scales = "free")
+c <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary ~ parental stage", x = NULL) + facet_wrap(~sex, scales = "free") 
+d <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads ~ parental stage")  + facet_wrap(~sex, scales = "free")
+
+bcd <- plot_grid(b,c,d, nrow = 3, labels = c("b"), label_size = 8, rel_heights = c(1,1,1))
+
+e <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + labs(subtitle = "hypothalamus ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")  
+f <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + labs(subtitle = "pituitary ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")   
+g <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + labs(subtitle = "gonads ~ internal physiology" ) + facet_wrap(~sex, scales = "free")
+
+efg <- plot_grid(e,f,g, nrow = 3, labels = c("d"), label_size = 8, rel_heights = c(1,1,1))
+
+h <- plottsneelipsev2(hyptsne, hyptsne$extint, allcolors ) + labs(subtitle = "hypothalamus ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")  
+i <- plottsneelipsev2(pittsne, pittsne$extint, allcolors ) + labs(subtitle = "pituitary ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")   
+j <- plottsneelipsev2(gontsne, gontsne$extint, allcolors ) + labs(subtitle = "gonads ~ external enviornment" ) + facet_wrap(~sex, scales = "free")
+
+hij <- plot_grid(h,i,j, nrow = 3, labels = c("c"), label_size = 8, rel_heights = c(1,1,1))
+
+bcdhijefg <- plot_grid(bcd, hij,  efg, nrow = 1)
+
+fig5 <- plot_grid(expdesign2, bcdhijefg, nrow = 2, labels = c("a"), label_size = 8, rel_heights = c(0.25,1))
+fig5
+
+```

### 2020-03-27 analysis/figure-3.Rmd
better caption, better size

@@
                       )
+
+# all candidate genes: AR, AVP, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, EGR1, ESR1, ESR2, FOS, GNRH1, GNRHR, JUN, LBH, OXT, PRL, PRLR, VIP
+
+# DEG candidate genes: AR, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, ESR1, FOS, GNRHR, LBH, PRL, PRLR,
+
+# Candidate genes that were not differentially expressed: AVP, AVPR1B, ESR2,  GNRH1, JUN, OXT, VIP
 
@@
     scale_color_manual(values = allcolors) +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+    theme(#axis.text.x = element_text(angle = 45, hjust = 1),
           legend.position = "none") +
-    labs(y = "expression" , x = "Parental stages", subtitle = mysubtitle) +
+    labs(y = "gene expression" , x = "Sequential stages with candidate DEGs", subtitle = mysubtitle) +
     theme(strip.text = element_text(face = "italic"))
@@
 
+b <- makeboxplots(hypvsd, c("AR"),"Male hypothalamus", "male", 
+                  c( "inc.d9", "inc.d17"))  +
+  theme(axis.title = element_blank())
+
+ab <- plot_grid(a,b, rel_widths = c(4,1),  labels = "auto", label_size = 8)
+
 c <- makeboxplots(pitvsd, c("ESR1","GNRHR"), "Female pituitary", "female", c("bldg", "lay", "inc.d3"))   +
@@
 
-cde <- plot_grid(c,d, e, rel_widths = c(2.5,2.5,1), nrow = 1)
+cde <- plot_grid(c,e,d, rel_widths = c(6,2.25,7), nrow = 1,  labels = c("c"), label_size = 8)
 
 g <- makeboxplots(gonvsd, c( "AVPR1A"), "Female gonads", "female", c( "lay", "inc.d3", "inc.d9"))  +
-  theme(axis.title.x = element_blank())
+  labs(caption = " ", x = " ")
 h <- makeboxplots(gonvsd, c( "EGR1", "FOS", "PRLR"), " ", "female", c( "lay", "inc.d3"))  +
-  theme(axis.title = element_blank())
+  theme(axis.title.y = element_blank()) + labs(caption = " ")
+f <- makeboxplots(pitvsd, c("LBH", "PRL"), "Male pituitary", "male", 
+                  c("inc.d9", "inc.d17")) + 
+  theme(plot.caption = element_text(face = "italic")) + 
+  labs(x = " ",
+       caption = "Candidate genes that were not differentially expressed from bldg to n5: AVP, AVPR1B, ESR2,  GNRH1, JUN, OXT, VIP.")
 
-gh <- plot_grid(g,h,nrow = 1, rel_widths = c(1,2.5))
+ghf <- plot_grid(g,h,f, nrow = 1, rel_widths = c(3,6,4),  labels = c("d", " ", "e"), label_size = 8)
 
-b <- makeboxplots(hypvsd, c("AR"),"Male hypothalamic expression", "male", 
-                  c( "inc.d9", "inc.d17"))  + labs(caption = " ")
-f <- makeboxplots(pitvsd, c("LBH", "PRL"), "Male pituitary expression", "male", 
-                  c("inc.d9", "inc.d17")) + 
-  theme(axis.title.y  = element_blank(),
-        plot.caption = element_text(face = "italic")) + 
-  labs(caption = "candidate genes: AR, AVP, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, EGR1, ESR1, ESR2, FOS, GNRH1, GNRHR, JUN, LBH, OXT, PRL, PRLR, VIP")
-
-bf <- plot_grid(b,f,nrow = 1, rel_widths = c(1,2))
-
-plot_grid(a, cde, gh, bf, nrow = 4, rel_heights = c(1,1,1,1.2), labels = "auto", label_size = 8)
+plot_grid(ab, cde, ghf, nrow = 3, rel_heights = c(1,1,1.25))
 

### 2020-03-28 analysis/figure-3.Rmd
better caption

@@
   theme(plot.caption = element_text(face = "italic")) + 
+  theme(axis.title.y = element_blank()) +
   labs(x = " ",

### 2020-03-28 analysis/12_GO.Rmd
new GO DEG thing

@@
                    "parentalbehavior", "matingbehavior",
-                   "candidategenes")) %>% 
+                   "candidategenes", "cellcycleprocess")) %>% 
   map(~.x) %>%
@@
   map_df(~.x)
+head(GO_DEGs)
 ```

### 2020-03-28 analysis/12_GO.Rmd
GO working now

@@
 ```{r bytissue}
-hyp <- GO_DEGs %>% select(GO, starts_with("hyp"))
-pit <- GO_DEGs %>% select(GO, starts_with("pit"))
-gon <- GO_DEGs %>% select(GO, starts_with("gon"))
+hyp <- GO_DEGs %>% select(GO, starts_with("male_hyp"), starts_with("female_hyp"))
+pit <- GO_DEGs %>% select(GO, starts_with("male_pit"), starts_with("female_pit"))
+gon <- GO_DEGs %>% select(GO, starts_with("male_gon"), starts_with("emale_gon"))
 

### 2020-03-28 analysis/12_GO.Rmd
no control

@@
   pivot_wider(names_from = comparison, values_from = genes) %>% 
-  select(GO, contains("control_bldg"), contains("bldg_lay"), contains("lay_inc.d3"),
+  select(GO, 
+         #contains("control_bldg"), 
+         contains("bldg_lay"), contains("lay_inc.d3"),
          contains("inc.d3_inc.d9"), contains("inc.d9_inc.d17"), contains("inc.d17_hatch"),

### 2020-03-28 analysis/12_GO.Rmd
candidate genes slim

@@
 ```{r}
-candidates <- c("OXT", "AVP", "GNRH1", "GNRHR", 
-                    "AR", "POMC", "AGRP", 
-                       "CRH", "AVPR1A", "AVPR1B", "AVPR2","VIP",
-                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9", 
-                    "ESR1","ESR2", "LBH", "CDK1", "BRCA1",
-                    "PTEN", "CREBBP", "FOS", "JUN", "EGR1",
-                     "BDNF", "GRM2","GRIA1",
-                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16", 
-                    "DIO3", "DIO2", "DIO1",
 
-                     "VIPR1", "VIPR2", "NPY", "LH", "FSH", "FSHB",
-                    "JAK2", "HCRT", "TRH", "TSHB",
-                    "MC3R", "MC4R", "MC5R",  "NR3C1", "NR3C2",
-                    "STAT5B","NPVF")
+candidategenesslim <- c("OXT", "AVP", "GNRH1", "GNRHR", 
+                    "AR",  "CYP19A1", 
+                     "AVPR1A", "AVPR1B", "AVPR2","VIP",
+                  "DRD1", "DRD2", 
+                  "PRL", "PRLR",  
+                    "ESR1","ESR2", "LBH",
+
+                    "FOS", "JUN", "EGR1", "BDNF"
+                      )
+
+candidates <- candidategenesslim
 
@@
 kable(candidate_DEGs) 
-
-bioprocshinyGO <- c("topBiologicalProcessGO",
-                    "Prostate gland growth",
-                    "Cellular response to hormone stimulus",
-                    "Response to endogenous stimulus",
-                    "Mammary gland development",
-                    "Response to hormone",
-                    "Response to endogenous stimulus",
-                    NA,
-                    "Regulation of animal organ morphogenesis",
-                    "Temperature homeostasis",
-                    "Prostate gland growth",
-                    NA,NA,NA,
-                    "Cellular response to hormone stimulus",
-                    NA,
-                    "Regulation of cell communication",
-                    NA,
-                    "Cell-cell signaling",
-                    "Cellular response to hormone stimulus",
-                    "Positive regulation of ossification",
-                    "G2 DNA damage checkpoint"
-                    )
-
-candidate_DEGs <- cbind(candidate_DEGs, bioprocshinyGO)
-kable(candidate_DEGs)
-
 ```

### 2020-03-28 analysis/12_GO.Rmd
new GO term

@@
   filter(GO %in% c("endocrineprocess", "lactation", 
-                   "parentalbehavior", "matingbehavior",
+                   "parentalbehavior", "matingbehavior", "reproduction",
                    "candidategenes", "cellcycleprocess")) %>% 

### 2020-03-28 analysis/12_GO.Rmd
remove stages

@@
 
-## Stage-specific differentially expressed genes
-
-```{r bystage}
-one <- GO_DEGs %>% select(GO, contains("control.bldg"))
-two <- GO_DEGs %>% select(GO, contains("bldg.lay"))
-three <- GO_DEGs %>% select(GO, contains("lay.inc.d3"))
-four <- GO_DEGs %>% select(GO, contains("inc.d3.inc.d9"))
-five <- GO_DEGs %>% select(GO, contains("inc.d9.inc.d17"))
-six <- GO_DEGs %>% select(GO, contains("inc.d17.hatch"))
-seven <- GO_DEGs %>% select(GO, contains("hatch.n5"))
-eight <- GO_DEGs %>% select(GO, contains("n5.n9"))
-
-kable(one) 
-# top GOs of these genes: endocrine porcess, 
-#regulation of blood pressure (by hormones or chemical signals), 
-#system process, regulation of hormone levels
-
-kable(two) 
-# top Gos maternal, parental, reproductive behavior; 
-# reproduction, reproductive process
-
-kable(three) 
-# response to chemical, maternal and parental behaviors, 
-# regulation of RNApolI transcription
-
-kable(four) 
-# insuffiencient genes for GO analysis
-
-kable(five) 
-# (cellular) response to organic substance, endocrine process, 
-# response to stimulus, response to chemical
-
-kable(six)  
-# insuffiencient genes
-
-kable(seven) 
-# endocrine process, system process, hormone transport, 
-# synaptic transmission (dopamineric), secretion by cell
-
-kable(eight) 
-# (positive) regulation of gap junction assembly, 
-# regulation of cardiac conduction, regualation of blood pressure
-
-```
-
 ## Candidate genes

### 2020-03-28 analysis/12_GO.Rmd
?

@@
   map_df(~.x)
-head(GO_DEGs)
+#head(GO_DEGs)
 ```
@@
 ```{r bytissue}
-hyp <- GO_DEGs %>% select(GO, starts_with("male_hyp"), starts_with("female_hyp"))
-pit <- GO_DEGs %>% select(GO, starts_with("male_pit"), starts_with("female_pit"))
-gon <- GO_DEGs %>% select(GO, starts_with("male_gon"), starts_with("emale_gon"))
+fhyp <- GO_DEGs %>% select(GO, starts_with("female_hyp"))
+mhyp <- GO_DEGs %>% select(GO, starts_with("male_hyp"))
 
-kable(hyp)  
-# top GO therms for hyp:
-# endocrine process, regulation of hormone levels,
-# signal release # hormone secretion, # system process
-# hormone transport, cell-cell signaling
+fpit <- GO_DEGs %>% select(GO, starts_with("female_pit"))
+mpit <- GO_DEGs %>% select(GO, starts_with("male_pit"))
 
-kable(pit)
-# endocrine process, mating behavior, system process,
-# reproductive behavior, mating, regulation of blood pressure
-# signal release, regulation of hormone levels
+fgon <- GO_DEGs %>% select(GO, starts_with("female_gon"))
+mgon <- GO_DEGs %>% select(GO, starts_with("male_gon"))
 
-kable(gon)  
-# endocrine process, regulation of blood pressure
-# blood circutation, circulatory system
-```
+kable(fhyp) 
+kable(mhyp)
 
-## Candidate genes
+kable(fpit)
+kable(mpit)
 
-```{r}
-candidate_DEGs <- left_join(candidates, DEGs) %>%
-  arrange(GO, gene) %>%
-  group_by(GO, comparison) %>%
-  summarize(genes = str_c(gene, collapse = ", ")) %>%
-  pivot_wider(names_from = comparison, values_from = genes) %>% 
-  select(GO, contains("control_bldg"), contains("bldg_lay"), contains("lay_inc.d3"),
-         contains("inc.d3_inc.d9"), contains("inc.d9_inc.d17"), contains("inc.d17_hatch"),
-         contains("hatch_n5"), contains("n5_n9"))
-candidate_DEGs <- as.data.frame(candidate_DEGs)
+kable(fgon) 
+kable(mgon)
 
-row.names(candidate_DEGs) <- candidate_DEGs$GO  
-candidate_DEGs <- t(candidate_DEGs)
-candidate_DEGs <- as.data.frame(candidate_DEGs)
-kable(candidate_DEGs) 
 ```
@@
 write.csv(GO_DEGs, "../results/12_GO_DEGs.csv")
-write.csv(hyp, "../results/12_GO_DEGs_hyp.csv")
-write.csv(pit, "../results/12_GO_DEGs_pit.csv")
-write.csv(gon, "../results/12_GO_DEGs_gon.csv")
+#write.csv(hyp, "../results/12_GO_DEGs_hyp.csv")
+#write.csv(pit, "../results/12_GO_DEGs_pit.csv")
+#write.csv(gon, "../results/12_GO_DEGs_gon.csv")
 ```

### 2020-03-28 analysis/figure-6.Rmd
first attempt at candidate gene figure

@@
 ---
-title: "fig6"
+title: "Fig6"
+author: "Rayna M Harris"
+date: "3/26/2020"
 output: md_document
@@
 
-# Figure 6
+## manipulation box plots for candidate genes
 
-```{r setup, include=TRUE}
+```{r}
 library(tidyverse)
+library(ggtext)
 library(cowplot)
+library(ggpubr)
+library(kableExtra)
 
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
+
+candidategenesslim <- c("OXT", "AVP", "GNRH1", "GNRHR", 
+                    "AR",  "CYP19A1", 
+                     "AVPR1A", "AVPR1B", "AVPR2","VIP",
+                  "DRD1", "DRD2", 
+                  "PRL", "PRLR",  
+                    "ESR1","ESR2", "LBH",
+
+                    "FOS", "JUN", "EGR1", "BDNF"
+                      )
+
+candidategenes <- candidategenesslim
+
 ```
 
-```{r fig6, fig.height=7, fig.width=7}
+## variance stabilized gene expression  (vsd)
+
+```{r }
 geneids <- read_csv("../metadata/00_geneinfo.csv")
 
-DEGchar <- read_csv("../results/DEGs-char.csv")
-DEGmanip <- read_csv("../results/DEGs-manip.csv")
-DEGall <- rbind(DEGchar, DEGmanip)
-tail(DEGall)
+vsd_path <- "../results/DEseq2/manip/"   # path to the data
+vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
+vsd_pathfiles <- paste0(vsd_path, vsd_files)
+vsd_files
 
-hatchnocontrol <- DEGall %>% filter(grepl("inc.d17", comparison)) %>% filter(!grepl("control", comparison))
-hatchnocontrol %>% group_by(comparison) %>% summarize(n = n())
+allvsd <- vsd_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name")  %>% 
+  dplyr::rename("gene" = "X1") %>% 
+  pivot_longer(cols = blk.s031.pu.d_female_gonad_prolong:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "samples", values_to = "counts") 
+allvsd <- as.tibble(allvsd)
 
-incd17hatch <- DEGall %>% filter(comparison == "inc.d17_hatch") %>%
-  arrange(gene) %>% distinct(comparison, gene)
-incd17hatch
+getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
+  candidates  <- allvsd %>%
+    filter(gene %in% whichgenes) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/manip/'), "[", 2)) %>%
+    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
+    #dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1))# %>%
+   # dplyr::recode(treatment, "m.hatch" = "m.n2", .default = levels(treatment))
 
-incd17prolong <- DEGall %>% filter(comparison == "inc.d17_prolong") %>%
-  arrange(gene) %>% distinct(comparison, gene)
-incd17prolong
+     dplyr::mutate(treatment = recode(treatment, "m.hatch" = "m.n2"))%>%
 
-bothincs <- full_join(incd17hatch,incd17prolong, by = "gene" )
-bothincs
+    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
+    filter(tissue == whichtissue, sex %in% whichsex)  %>%
+    drop_na()
+  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
+  return(candidates)
+}
 
-onlyhatch <- subset(bothincs, is.na(bothincs$comparison.y))
-onlyhatch
+hypvsd <- getcandidatevsd(candidategenesslim, "hypothalamus", sexlevels)
+pitvsd <- getcandidatevsd(candidategenesslim, "pituitary", sexlevels)
+gonvsd <- getcandidatevsd(candidategenesslim, "gonad", sexlevels)
+head(hypvsd)
+head(pitvsd)
+head(gonvsd)
+
+candidatevsd <- rbind(hypvsd, pitvsd)
+candidatevsd <- rbind(candidatevsd, gonvsd)
+
+head(candidatevsd)
+levels(candidatevsd$treatment)
 ```
+
+## DEGs
+
+```{r allDEGs}
+DEG_path <- "../results/DEseq2/manip/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/manip/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
+head(allDEG)
+
+allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
+allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
+allDEG$direction <- factor(allDEG$direction, levels = alllevels)
+
+head(allDEG)
+
+table1 <- allDEG %>%
+  mutate(updown = ifelse(lfc > 0, "+", "-"))  %>%
+  mutate(geneupdown = paste(gene, updown, sep = "")) %>%
+  filter(gene %in% candidategenes) %>%
+  arrange(geneupdown) %>%
+  group_by(sex, tissue, comparison) %>%
+  summarize(genes = str_c(geneupdown, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = genes ) %>%
+  select(sex, tissue, contains("m."))  %>%
+  arrange( tissue, sex)
+
+(table1$inc.d3_m.inc.d3)
+(table1$inc.d9_m.inc.d9)
+(table1$inc.d17_m.inc.d17)
+(table1$hatch_m.n2)
+```
+
+## Figs
+
+```{r fig6, fig.height=7, fig.width=7}
+makeboxplotsmanip <- function(df, whichgene, mysubtitle, whichsex, whichtimepoint){
+  p <- df %>%
+    filter(treatment %in% alllevels,
+           gene %in% whichgene,
+           sex %in% whichsex) %>%
+    filter(treatment %in% whichtimepoint) %>%
+    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
+    geom_boxplot(outlier.shape = NA) + 
+    geom_jitter(size = 0.5, width = 0.1) +
+    facet_wrap(~gene, scales = "free_y", nrow = 1) +
+    theme_B3() +
+    scale_fill_manual(values = allcolors) +
+    scale_color_manual(values = allcolors) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1),
+          legend.position = "none",
+          plot.caption = element_text(face = "italic")) +
+    labs(y = "gene expression" , x = "Parental Stages and Manipulations", subtitle = mysubtitle) +
+    theme(strip.text = element_text(face = "italic"))
+  return(p)
+}
+
+## hyp 
+a2 <- makeboxplotsmanip(hypvsd, c("DRD1","EGR1","PRLR"), "Female hypothalamus", "female",
+                  c("inc.d3", "m.inc.d3", "inc.d17", "m.inc.d17",  "hatch", "m.n2"))  +
+  theme(axis.title.x = element_blank())
+b2 <- makeboxplotsmanip(hypvsd, c("BDNF","ESR2", "FOS"), " ", "female",
+                  c("inc.d3", "m.inc.d3", "inc.d17", "m.inc.d17"))  +
+  theme(axis.title.x = element_blank())
+
+c <- makeboxplotsmanip(hypvsd, c("AVP","OXT"), " ", "female", 
+                       c("inc.d17", "m.inc.d17"))  +
+  theme(axis.title = element_blank())
+d <- makeboxplotsmanip(hypvsd, c("AVPR1A","CYP19A1","GNRH1"), " ", "female",
+                   c("hatch", "m.n2"))  +
+  theme(axis.title = element_blank())
+e <- makeboxplotsmanip(hypvsd, c("JUN"), " ", "female", 
+                 c("hatch", "m.n2"))  +
+  theme(axis.title = element_blank()) + labs(subtitle = " ")
+
+f1 <- makeboxplotsmanip(hypvsd, c("DRD1", "EGR1", "PRLR"), "Male hypothalamus", "male",
+                  c( "inc.d17", "m.inc.d17", "hatch", "m.n2"))  +
+  theme(axis.title.x = element_blank())
+
+f2 <- makeboxplotsmanip(hypvsd, c( "PRL"), " ", "male",
+                  c( "inc.d17", "m.inc.d17", "hatch", "m.n2"))  +
+  theme(axis.title = element_blank())
+
+h <- makeboxplotsmanip(hypvsd, c("AR", "AVP", "BDNF", "ESR1", "LBH"), " ", "male",
+                  c( "inc.d17", "m.inc.d17"))  + labs(caption = " ")
+
+i <- makeboxplotsmanip(hypvsd, c("GNRH1"), " ", "male",
+                  c("hatch", "m.n2"))  +
+  theme(axis.title = element_blank()) +
+  labs(x = " ",
+       caption = "Candidate genes that were not differentially expressed following removal: AVPR1B, AVPR2, ESR1, GNRHR, VIP")
+
+ab2 <- plot_grid(a2,c, rel_widths = c(18,5), labels = "auto", label_size = 8)
+cde <- plot_grid(b2,d,e, rel_widths = c(12,6,2), nrow =  1, align = "h",  
+                 labels = c("c", "d", " "), label_size = 8)
+f12 <- plot_grid(f1,f2, rel_widths = c(3,1), labels = c("e",  " "), label_size = 8)
+hi <- plot_grid(h,i, ncol = 2, rel_widths = c(10,2),labels = c("f",  "g"), label_size = 8, align = "h")
+
+plot_grid(ab2, cde, f12,hi, nrow = 4, rel_heights = c(1,1,1,1.25))
+
+```
+
+```{r writefiles}
+write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
+```

### 2020-03-29 analysis/figure-6.Rmd
now with better fig 3 and 6

@@
 gonvsd <- getcandidatevsd(candidategenesslim, "gonad", sexlevels)
-head(hypvsd)
-head(pitvsd)
-head(gonvsd)
 
@@
 
-## Figs
-
-```{r fig6, fig.height=7, fig.width=7}
+```{r boxplotfunction}
 makeboxplotsmanip <- function(df, whichgene, mysubtitle, whichsex, whichtimepoint){
@@
     labs(y = "gene expression" , x = "Parental Stages and Manipulations", subtitle = mysubtitle) +
-    theme(strip.text = element_text(face = "italic"))
+    theme(strip.text = element_text(face = "italic", size = 5)) + 
+    scale_y_continuous(labels = scales::number_format(accuracy = 1))
   return(p)
 }
+```
 
-## hyp 
-a2 <- makeboxplotsmanip(hypvsd, c("DRD1","EGR1","PRLR"), "Female hypothalamus", "female",
+## Figs
+
+```{r fig6, fig.height=8, fig.width=10}
+
+## hypothatlamus
+a <- makeboxplotsmanip(hypvsd, c("DRD1","EGR1","PRLR"), "Female hypothalamus", "female",
                   c("inc.d3", "m.inc.d3", "inc.d17", "m.inc.d17",  "hatch", "m.n2"))  +
   theme(axis.title.x = element_blank())
-b2 <- makeboxplotsmanip(hypvsd, c("BDNF","ESR2", "FOS"), " ", "female",
+b <- makeboxplotsmanip(hypvsd, c("BDNF","ESR2", "FOS"), " ", "female",
                   c("inc.d3", "m.inc.d3", "inc.d17", "m.inc.d17"))  +
-  theme(axis.title.x = element_blank())
+  theme(axis.title = element_blank())
 
@@
 
-f1 <- makeboxplotsmanip(hypvsd, c("DRD1", "EGR1", "PRLR"), "Male hypothalamus", "male",
+f <- makeboxplotsmanip(hypvsd, c("DRD1", "EGR1", "PRLR"), "Male hypothalamus", "male",
                   c( "inc.d17", "m.inc.d17", "hatch", "m.n2"))  +
@@
 
-f2 <- makeboxplotsmanip(hypvsd, c( "PRL"), " ", "male",
+g <- makeboxplotsmanip(hypvsd, c( "PRL"), " ", "male",
                   c( "inc.d17", "m.inc.d17", "hatch", "m.n2"))  +
@@
 h <- makeboxplotsmanip(hypvsd, c("AR", "AVP", "BDNF", "ESR1", "LBH"), " ", "male",
-                  c( "inc.d17", "m.inc.d17"))  + labs(caption = " ")
+                  c( "inc.d17", "m.inc.d17")) +
+  theme(axis.title = element_blank())
 
@@
                   c("hatch", "m.n2"))  +
-  theme(axis.title = element_blank()) +
-  labs(x = " ",
-       caption = "Candidate genes that were not differentially expressed following removal: AVPR1B, AVPR2, ESR1, GNRHR, VIP")
+  theme(axis.title = element_blank())
 
-ab2 <- plot_grid(a2,c, rel_widths = c(18,5), labels = "auto", label_size = 8)
-cde <- plot_grid(b2,d,e, rel_widths = c(12,6,2), nrow =  1, align = "h",  
-                 labels = c("c", "d", " "), label_size = 8)
-f12 <- plot_grid(f1,f2, rel_widths = c(3,1), labels = c("e",  " "), label_size = 8)
-hi <- plot_grid(h,i, ncol = 2, rel_widths = c(10,2),labels = c("f",  "g"), label_size = 8, align = "h")
+j <- makeboxplotsmanip(pitvsd, c("AR", "PRL", "LBH"), "Female pituitary", "female",
+                  c( "inc.d17", "m.inc.d17", "hatch", "m.n2"))  + theme(axis.title.x = element_blank())
 
-plot_grid(ab2, cde, f12,hi, nrow = 4, rel_heights = c(1,1,1,1.25))
+k <- makeboxplotsmanip(pitvsd, c("DRD1"), " ", "female",
+                  c( "inc.d3", "m.inc.d3"))  + theme(axis.title = element_blank())
 
+l <- makeboxplotsmanip(pitvsd, c("BDNF", "PRLR"), " ", "female",
+                  c( "inc.d17", "m.inc.d17"))  +  theme(axis.title = element_blank())
+
+m <- makeboxplotsmanip(pitvsd, c("AVPR1B", "AVPR2", "EGR1", "GNRHR", "JUN"), " ", "female",
+                  c("hatch", "m.n2"))  + theme(axis.title = element_blank())
+
+n <- makeboxplotsmanip(pitvsd, c("LBH"), "Male pituitary", "male",
+                  c("inc.d17", "m.inc.d17", "hatch", "m.n2"))  
+o <- makeboxplotsmanip(pitvsd, c("DRD1", "ESR2"), " ", "male",
+                  c("inc.d17", "m.inc.d17"))  + theme(axis.title  = element_blank())
+p <- makeboxplotsmanip(pitvsd, c("AVPR1A", "PRLR"), " ", "male",
+                  c("hatch", "m.n2"))  + theme(axis.title  = element_blank())
+
+q <- makeboxplotsmanip(gonvsd, c("LBH"), "Male gonads ", "male",
+                  c("inc.d17", "m.inc.d17"))  + theme(axis.title  = element_blank())
+r <- makeboxplotsmanip(gonvsd, c("JUN"), "Female gonads ", "female",
+                  c("inc.d17", "m.inc.d17"))  + theme(axis.title  = element_blank())
+
+abcd <- plot_grid(a,b,c,d,e, nrow = 1, rel_widths = c(16,10,4,6,2), align = "h", labels = c("a", "b", "c", "d"), label_size = 8)
+efgh <- plot_grid(f,g,h,i, nrow = 1, rel_widths = c(10,3,10,2), align = "h", labels = c("e", " ", "f", "g"), label_size = 8)
+jklm <- plot_grid(j,k,l,m, nrow = 1, rel_widths = c(12,2,4,10), align = "h",  labels = c("h", "i", "j", "k"), label_size = 8)
+nopqr <- plot_grid(n,o,p,q,r, nrow = 1, rel_widths = c(4,4,4,2,2), align = "h",  labels = c("l", "m", "n", "o", "p"), label_size = 8)
+
+plot_grid(abcd, efgh, jklm,nopqr, ncol = 1)
 ```

### 2020-03-29 analysis/figure-3.Rmd
now with better fig 3 and 6

--- 2020-03-27 analysis/figure-3.Rmd
@@
   theme(plot.caption = element_text(face = "italic")) + 
+  theme(axis.title.y = element_blank()) +
   labs(x = " ",
-       caption = "Candidate genes that were not differentially expressed from bldg to n5: AVP, AVPR1B, ESR2,  GNRH1, JUN, OXT, VIP.")
+       caption = "Candidate genes that were not differentially expressed from bldg to n5: AVP, AVPR1B, ESR2,  GNRH1, JUN, and OXT.")
 

### 2020-04-06 analysis/16_PRL.Rmd
tables with prl counts

@@
 source("../R/functions.R")
-source("../R/icons.R")
 
@@
 ```{r data}
-hormones <- read_csv("../results/hormones.csv")
+hormones <- read_csv("../results/07_hormones.csv")
 hormones$treatment <- factor(hormones$treatment, levels = alllevels)
@@
 
-p1 +
-  annotation_custom(control, ymin = -20, ymax = 0, xmin = -7.8) +
-  annotation_custom(bldg, ymin = -20, ymax = 0, xmin = -5.8) +
-  annotation_custom(lay, ymin = -20, ymax = 0, xmin = -3.8) +
-  annotation_custom(inc, ymin = -20, ymax = 0, xmin = -1.8) +
-  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 0.6) +
-  annotation_custom(inc, ymin = -20, ymax = 0, xmin = 2.4) +
-  annotation_custom(hatch, ymin = -20, ymax = 0, xmin = 4.4) +
-  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 6.4) +
-  annotation_custom(nestling, ymin = -20, ymax = 0, xmin = 8.4) + 
-  ylim(c(-10,120))
-
 ```
 
-## Prolactin (_PRL_) and _BRCA1_ expression in the pituitary
+## Prolactin expression
 
 ```{r candidatevsds}
-colData.pit <-  read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
-colData.pit <- colData.pit %>% filter(tissue == "pituitary")
+prl <- read_csv("../results/candidatecounts.csv") %>%
+  filter(gene ==  "PRL")  %>% 
+  mutate(bird_id = sapply(strsplit(samples, '_'), "[", 1)) %>%
+  select(-X1, -samples) %>% 
+  select(bird_id, everything())
 
-vsd.pit <- read.csv("../results/06_pitallvsd.csv", header = T, row.names = 1)
+prl
 
-savecols <- as.character(colData.pit$V1) 
-savecols <- as.vector(savecols) 
-vsd.pit <- vsd.pit %>% dplyr::select(one_of(savecols)) 
-vsd.pit$genes <- row.names(vsd.pit)
+hormonesslim <- hormones %>%
+  select(bird_id, hormone, plasma_conc, logconc)
+head(hormonesslim)
 
-candidategenes <- c("PRL", "BRCA1", "PRLR")
-candidates.pit <- vsd.pit %>% dplyr::filter(genes %in% candidategenes)
-candidates.pit <- as.data.frame(candidates.pit)
-row.names(candidates.pit) <- candidates.pit$genes
-candidates.pit$genes <- NULL
-candidates.pit <- as.data.frame(t(candidates.pit))
-candidates.pit$V1 <- row.names(candidates.pit) 
-candidates.pit <- left_join(colData.pit, candidates.pit) 
-candidates.pit$treatment <- factor(candidates.pit$treatment, levels = alllevels)
-candidates.pit$study2 <- ifelse(candidates.pit$treatment %in% charlevels, "charcterization",
-                              ifelse(candidates.pit$treatment %in% levelstiming, "timing",
-                                     ifelse(candidates.pit$treatment %in% levelsremoval, "removal", NA)))
-tail(candidates.pit)
-
+hormonesandPRL <- full_join(prl, hormonesslim, by = "bird_id")
+hormonesandPRL
 ```
@@
 p2
-p2+
-  annotation_custom(control, ymin = 13.5, ymax = 15, xmin = -7.5) +
-  annotation_custom(bldg, ymin = 13.5, ymax = 15, xmin = -5.5) +
-  annotation_custom(lay, ymin = 13.5, ymax = 15, xmin = -3.5) +
-  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = -1.5) +
-  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = 0.5) +
-  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = 2.5) +
-  annotation_custom(hatch, ymin = 13.5, ymax = 15, xmin = 4.5) +
-  annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 6.5) +
-  annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 8.5)
 
-```
-
-```{r hormonesandgenes}
-plot_grid(p1 + theme(legend.position ="none",
-                     axis.title.x = element_blank(),
-                     axis.text.x = element_blank()),
-          p2 + theme(legend.position = c(0.85,0.3)) +
-            labs(subtitle = NULL) +
-  annotation_custom(control, ymin = 13.5, ymax = 15.5, xmin = -7.5) +
-  annotation_custom(bldg, ymin = 13.5, ymax = 15.5, xmin = -5.5) +
-  annotation_custom(lay, ymin = 13.5, ymax = 15.5, xmin = -3.5) +
-  annotation_custom(inc, ymin = 13.5, ymax = 15.5, xmin = -1.5) +
-  annotation_custom(inc, ymin = 13.5, ymax = 15.5, xmin = 0.5) +
-  annotation_custom(inc, ymin = 13.5, ymax = 15.5, xmin = 2.5) +
-  annotation_custom(hatch, ymin = 13.5, ymax = 15.5, xmin = 4.5) +
-  annotation_custom(nestling, ymin = 13.5, ymax = 15.5, xmin = 6.5) +
-  annotation_custom(nestling, ymin = 13.5, ymax = 15.5, xmin = 8.5), 
-          nrow = 2,rel_heights = c(0.4,0.6)) 
 ```
@@
   labs(x = "increasing time >>", y = "BRCA1", subtitle = "Pituitary") +
-  annotation_custom(control, ymin = 6.65, ymax = 7, xmin = -7.8) +
-  annotation_custom(bldg, ymin = 6.65, ymax = 7, xmin = -5.8) +
-  annotation_custom(lay, ymin = 6.65, ymax = 7, xmin = -3.8) +
-  annotation_custom(inc, ymin = 6.65, ymax = 7, xmin = -1.8) +
-  annotation_custom(inc, ymin = 6.65, ymax = 7, xmin = 0.6) +
-  annotation_custom(inc, ymin = 6.65, ymax = 7, xmin = 2.4) +
-  annotation_custom(hatch, ymin = 6.65, ymax = 7, xmin = 4.4) +
-  annotation_custom(nestling, ymin = 6.65, ymax = 7, xmin = 6.4) +
-  annotation_custom(nestling, ymin = 6.65, ymax = 7, xmin = 8.4) +
+
   ylim(6.8,8.5) +
@@
   theme(axis.title.y  = element_text(face = "italic"),
-        axis.text.x = element_blank())  +
-  annotation_custom(control, ymin = 13.5, ymax = 15, xmin = -7.5) +
-  annotation_custom(bldg, ymin = 13.5, ymax = 15, xmin = -5.5) +
-  annotation_custom(lay, ymin = 13.5, ymax = 15, xmin = -3.5) +
-  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = -1.5) +
-  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = 0.5) +
-  annotation_custom(inc, ymin = 13.5, ymax = 15, xmin = 2.5) +
-  annotation_custom(hatch, ymin = 13.5, ymax = 15, xmin = 4.5) +
-  annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 6.5) +
-  annotation_custom(nestling, ymin = 13.5, ymax = 15, xmin = 8.5)
+        axis.text.x = element_blank())
 
@@
 
+write.csv(prl, "../results/16_PRL.csv", row.names = F)
+write.csv(hormonesandPRL, "../results/16_hormonesandPRL.csv", row.names = F)
+
 ```

### 2020-04-11 analysis/16_PRL.Rmd
differnt  analysis with log counts

@@
 ```
+
+```{r include=F, eval=F}
+# data from VF
+library(tidyverse)
+counts <- read_csv("../results/2020-04-10_Suzanne_gene_counts_VF.csv") %>%
+  mutate(logcounts = log(counts)) %>%
+  select(bird_id, sex, tissue, stage, timept, geneid, counts, logcounts) %>%
+  drop_na()
+head(counts)
+
+counts %>%
+  filter(geneid == "prl") %>%
+  ggplot(aes(x = as.factor(timept), y = logcounts)) +
+  geom_boxplot(aes(color = sex)) +
+  facet_wrap(~ tissue)
+```

### 2020-04-13 analysis/figure-3.Rmd
tables with prl counts
new DEG table

--- 2020-03-27 analysis/figure-3.Rmd
@@
                     "ESR1","ESR2", "LBH",  
+                    "FOS", "JUN", "EGR1", "BDNF")
 
-                    "FOS", "JUN", "EGR1", "BDNF"
-                      )
+# candidategenesslim: AR, AVP, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, EGR1, ESR1, ESR2, FOS, GNRH1, GNRHR, JUN, LBH, OXT, PRL, PRLR, VIP
+# DEG candidategenesslim: AR, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, ESR1, FOS, GNRHR, LBH, PRL, PRLR, 
+# non-DEG candidategenesslim: AVP, AVPR1B, ESR2,  GNRH1, JUN, OXT, VIP
 
-# all candidate genes: AR, AVP, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, EGR1, ESR1, ESR2, FOS, GNRH1, GNRHR, JUN, LBH, OXT, PRL, PRLR, VIP
-
-# DEG candidate genes: AR, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, ESR1, FOS, GNRHR, LBH, PRL, PRLR,
-
-# Candidate genes that were not differentially expressed: AVP, AVPR1B, ESR2,  GNRH1, JUN, OXT, VIP
-
+hormonepapergenes <- c("GNRH1", "CGNRH-R", "GNRHR",  
+                       "FSHB", "FSHR", "LHCGR", # no LSH
+                       "VIP", "VIPR1",
+                       "ESR1", "AR", "CYP19A1", "HSD17B1", # receptors synthesis
+                       "PGR" , "HSD3B2", 
+                       "PRL", "PRLR")
 ```
@@
 ```{r }
-geneids <- read_csv("../metadata/00_geneinfo.csv")
+geneids <- read_csv("../metadata/00_geneinfo.csv") %>% select(-X1)
 
@@
 
-hypvsd <- getcandidatevsd(candidategenesslim, "hypothalamus", sexlevels)
-pitvsd <- getcandidatevsd(candidategenesslim, "pituitary", sexlevels)
-gonvsd <- getcandidatevsd(candidategenesslim, "gonad", sexlevels)
+hypvsd <- getcandidatevsd(hormonepapergenes, "hypothalamus", sexlevels)
+pitvsd <- getcandidatevsd(hormonepapergenes, "pituitary", sexlevels)
+gonvsd <- getcandidatevsd(hormonepapergenes, "gonad", sexlevels)
 head(hypvsd)
-head(pitvsd)
-head(gonvsd)
+tail(gonvsd)
 
@@
 
-```{r fig3, fig.width=7, fig.height=7}
-makeboxplots <- function(df, whichgene, mysubtitle, whichsex, whichtimepoint){
+```{r DEGs}
+# summary DEG results from DESeq2
+candidateDEGS <- read_csv("../../musicalgenes/data/allDEG.csv") %>%
+  select(-X1) %>%
+  filter(gene %in% hormonepapergenes)
+head(candidateDEGS)
+
+# get genes to plot
+candidateDEGS %>% filter(tissue == "hypothalamus", sex == "female", comparison != "control_bldg") 
+candidateDEGS %>% filter(tissue == "hypothalamus", sex == "male", comparison != "control_bldg") 
+candidateDEGS %>% filter(tissue == "pituitary", sex == "male", comparison != "control_bldg") 
+candidateDEGS %>% filter(tissue == "pituitary", sex == "female", comparison != "control_bldg") 
+candidateDEGS %>% filter(tissue == "gonad", sex == "female", comparison != "control_bldg")
+
+# summarize all differences
+candidateDEGtable <- candidateDEGS %>%
+  mutate(posneg = ifelse(lfc >= 0, "+", "-")) %>%
+  mutate(geneposneg = paste(gene, posneg, sep = "")) %>%
+  select(sex, tissue, comparison, geneposneg) %>%
+  group_by(sex, tissue, comparison) %>%
+   arrange(geneposneg) %>%
+  summarize(genes = str_c(geneposneg, collapse = " ")) %>% 
+  pivot_wider(names_from = comparison, values_from = genes)  %>% 
+  ungroup(tissue) %>%
+  mutate(tissue = factor(tissue, levels =  tissuelevel))  %>% 
+  arrange(sex, tissue)  %>% 
+  select(sex, tissue, bldg_lay, lay_inc.d3, inc.d9_inc.d17, hatch_n5, control_bldg)
+candidateDEGtable
+
+newboxplot <- function(df, mygenes, whichsex, whichstage){
+
   p <- df %>%
-    filter(treatment %in% charlevels,
-           gene %in% whichgene,
-           sex %in% whichsex) %>%
-    filter(treatment %in% whichtimepoint) %>%
+    filter(gene %in% mygenes, 
+           sex %in% whichsex,
+           treatment %in% whichstage) %>%
+    droplevels() %>%
+    mutate(tissue = factor(tissue, levels = tissuelevel)) %>%
     ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
-    geom_boxplot(outlier.shape = NA) + 
+    geom_boxplot() +
     geom_jitter(size = 0.5, width = 0.1) +
-    facet_wrap(~gene, scales = "free_y", nrow = 1) +
+    facet_grid(tissue~gene, scales = "free_y") +
     theme_B3() +
+    scale_color_manual(values = allcolors) +
     scale_fill_manual(values = allcolors) +
-    scale_color_manual(values = allcolors) +
-    theme(#axis.text.x = element_text(angle = 45, hjust = 1),
-          legend.position = "none") +
-    labs(y = "gene expression" , x = "Sequential stages with candidate DEGs", subtitle = mysubtitle) +
-    theme(strip.text = element_text(face = "italic"))
+    theme(legend.position = "none",
+          axis.text.x = element_text(angle = 45, hjust = 1),
+          strip.text.x = element_text(face = "italic"),
+          axis.title.x = element_blank())  +
+    labs(x = "Parental stage",
+         y = "variance stabilized expression")
   return(p)
@@
 
-## hyp 
-a <- makeboxplots(hypvsd, c("BDNF","CYP19A1", "DRD1", "EGR1"), 
-                  "Female hypothalamus", "female", c("hatch", "n5"))  +
-  theme(axis.title.x = element_blank())
+a1 <- newboxplot(hypvsd, c("PGR",  "CYP19A1", "LHCGR" ), "female", c( "hatch", "n5")) + 
+  theme(strip.text.y = element_blank()) + labs(subtitle = "Females")
+a2 <- newboxplot(hypvsd, c("AR"), "male", c( "hatch", "n5")) + 
+  theme(axis.title.y = element_blank()) + labs(subtitle = "Males")
 
-b <- makeboxplots(hypvsd, c("AR"),"Male hypothalamus", "male", 
-                  c( "inc.d9", "inc.d17"))  +
-  theme(axis.title = element_blank())
+b1 <- newboxplot(gonvsd, c("HSD3B2"), "female", c( "bldg", "lay"))  +
+  theme(strip.text.y = element_blank(),
+        axis.title.y = element_blank()) + labs(subtitle = "Females")
+b2 <- newboxplot(gonvsd, c("VIPR1",  "PRLR" ), "female", c( "lay", "inc.d3"))  + 
+  theme(axis.title.y = element_blank()) + labs(subtitle = " ")
 
-ab <- plot_grid(a,b, rel_widths = c(4,1),  labels = "auto", label_size = 8)
+ab <- plot_grid(a1,a2,b1,b2, rel_widths = c(6,2.5,2,4), labels = c("a", " ", "b", " "), label_size = 8, nrow = 1)
 
-c <- makeboxplots(pitvsd, c("ESR1","GNRHR"), "Female pituitary", "female", c("bldg", "lay", "inc.d3"))   +
-  theme(axis.title.x = element_blank())
-d <- makeboxplots(pitvsd, c("LBH", "PRL" ), " ", "female", c("inc.d9", "inc.d17", "hatch", "n5"))  +
-  theme(axis.title  = element_blank())
-e <- makeboxplots(pitvsd, c( "AVPR2"), " ", "female", c( "hatch", "n5"))  +
-  theme(axis.title  = element_blank())
+c1 <- newboxplot(pitvsd, c( "ESR1", "GNRHR"), "female", c("bldg", "lay",  "inc.d3")) 
+c2 <- newboxplot(pitvsd, c( "PRL"), "female", c("inc.d9", "inc.d17", "hatch", "n5")) 
+c3 <- newboxplot(pitvsd, c("VIPR1",  "PRL" ), "male", c( "inc.d9", "inc.d17"))
 
-cde <- plot_grid(c,e,d, rel_widths = c(6,2.25,7), nrow = 1,  labels = c("c"), label_size = 8)
+c <- plot_grid(c1 + theme(strip.text.y = element_blank()) + labs(subtitle = "Females"),
+          c2 + theme(strip.text.y = element_blank(),
+                    axis.title.y = element_blank()) + labs(subtitle = " ") ,
+          c3 + theme(axis.title.y = element_blank()) + labs(subtitle = "Males"), 
+          nrow = 1, rel_widths = c(6,4,4),
+          labels = c("c"), label_size = 8)
 
-g <- makeboxplots(gonvsd, c( "AVPR1A"), "Female gonads", "female", c( "lay", "inc.d3", "inc.d9"))  +
-  labs(caption = " ", x = " ")
-h <- makeboxplots(gonvsd, c( "EGR1", "FOS", "PRLR"), " ", "female", c( "lay", "inc.d3"))  +
-  theme(axis.title.y = element_blank()) + labs(caption = " ")
-f <- makeboxplots(pitvsd, c("LBH", "PRL"), "Male pituitary", "male", 
-                  c("inc.d9", "inc.d17")) + 
-  theme(plot.caption = element_text(face = "italic")) + 
-  labs(x = " ",
-       caption = "Candidate genes that were not differentially expressed from bldg to n5: AVP, AVPR1B, ESR2,  GNRH1, JUN, OXT, VIP.")
-
-ghf <- plot_grid(g,h,f, nrow = 1, rel_widths = c(3,6,4),  labels = c("d", " ", "e"), label_size = 8)
-
-plot_grid(ab, cde, ghf, nrow = 3, rel_heights = c(1,1,1.25))
+plot_grid(ab,c, nrow = 2)
 
@@
 ```{r writefiles}
-write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
+#write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
+#write.csv(candidatevsd, "../results/candidatecounts.csv")
+write.csv(candidateDEGtable, "../results/candidateDEGtable.csv")
+
 ```

### 2020-04-13 analysis/figure-3.Rmd
candidate hormone degs and corrrs

@@
 
-```{r DEGs}
+```{r fig3, fig.width = 7}
 # summary DEG results from DESeq2
@@
 
+# correlations
+library(corrr)
+
+plotcorrs <- function(whichtissue){
+
+  x <- candidatevsd %>%
+    filter(tissue == whichtissue)  %>%
+    pivot_wider(names_from = gene, values_from = counts) %>%
+    select(-sex, - tissue, - treatment, -samples) %>%
+    correlate()
+  print(fashion(x))
+
+  p <- rplot(x, colors = c("#0571b0", "#92c5de", "#f7f7f7","#f7f7f7","#f7f7f7", "#f4a582", "#ca0020")) +
+    theme_B3() +
+    theme(legend.position = "bottom",
+          axis.text.x = element_text(angle = 45, hjust = 1),
+          axis.text = element_text(face = "italic")) +
+    labs(subtitle = whichtissue)
+  return(p)
+}
+
+a <- plotcorrs("hypothalamus") 
+b <- plotcorrs("pituitary") + theme(axis.text.y = element_blank())
+c <- plotcorrs("gonad") + theme(axis.text.y = element_blank())
+
+plot_grid(a,b,c, nrow = 1, rel_widths = c(1.2,1,1))
+
 ```

### 2020-04-13 analysis/figure-3.Rmd
now with hormoens

@@
 
+hormones <- read_csv("../results/07_hormoneswide.csv") %>%
+  select(-study, -treatment, -sex)
+head(hormones)
+
+head(candidatevsd)
+
+vsdhormones <- candidatevsd %>%
+  pivot_wider(names_from = gene, values_from = counts) %>%
+  mutate(bird_id = sapply(strsplit(samples,"\\_"), "[", 1)) %>%
+  left_join(.,hormones, by =  "bird_id")
+vsdhormones
+
 plotcorrs <- function(whichtissue){
 
-  x <- candidatevsd %>%
+  x <- vsdhormones  %>%
     filter(tissue == whichtissue) %>%
-    pivot_wider(names_from = gene, values_from = counts) %>%
-    select(-sex, - tissue, - treatment, -samples) %>%
+    select(-sex, - tissue, - treatment, -samples, -bird_id, -moltbin) %>%
     correlate()

### 2020-04-13 analysis/figure-3.Rmd
more correlations

@@
 
-plotcorrs <- function(whichtissue){
+vsdhormoneshyp <- vsdhormones %>% filter(tissue == "hypothalamus" )
+vsdhormonespit <- vsdhormones %>% filter(tissue == "pituitary" )
+vsdhormonesgon <- vsdhormones %>% filter(tissue == "gonad" )
+
+plotcorrs <- function(whichtissue, whichsex){
 
   x <- vsdhormones  %>%
-    filter(tissue == whichtissue) %>%
+    filter(tissue == whichtissue, 
+           sex == whichsex) %>%
     select(-sex, - tissue, - treatment, -samples, -bird_id, -moltbin) %>%
@@
 
-  p <- rplot(x, colors = c("#0571b0", "#92c5de", "#f7f7f7","#f7f7f7","#f7f7f7", "#f4a582", "#ca0020")) +
+  p <-rplot(x, colors = c("#0571b0", "#92c5de", "#f7f7f7","#f7f7f7","#f7f7f7", "#f4a582", "#ca0020")) +
     theme_B3() +
@@
 
-a <- plotcorrs("hypothalamus") 
-b <- plotcorrs("pituitary") + theme(axis.text.y = element_blank())
-c <- plotcorrs("gonad") + theme(axis.text.y = element_blank())
+a <- plotcorrs("hypothalamus", "female") + theme(legend.position = "none")
+b <- plotcorrs("pituitary", "female") + theme(axis.text.y = element_blank()) + theme(legend.position = "none")
+c <- plotcorrs("gonad", "female") + theme(axis.text.y = element_blank()) + theme(legend.position = "none")
 
-plot_grid(a,b,c, nrow = 1, rel_widths = c(1.2,1,1))
+d <- plotcorrs("hypothalamus", "male") 
+e <- plotcorrs("pituitary", "male") + theme(axis.text.y = element_blank())
+f <- plotcorrs("gonad", "male") + theme(axis.text.y = element_blank())
 
+plot_grid(a,b,c, d,e,f, nrow = 2, rel_widths = c(1.2,1,1), rel_heights = c(1,1.2))
+
+plotspecificcorrs <- function(df, myx, myy, myxlab, myylab){
+  p <- df %>%
+    mutate(tissue = factor(tissue, levels = tissuelevel)) %>%
+    ggplot(aes(x = myx, y = myy)) +
+    geom_point(aes(color = treatment)) +
+    geom_smooth(method = "lm", aes(color = sex)) +
+    facet_wrap(~sex, scales = "free", nrow = 2) +
+    scale_color_manual(values = allcolors) +
+    labs(x = myxlab, y = myylab) +
+    theme_B3() +
+    theme(legend.position = "none",
+          axis.title = element_text(face = "italic"))
+  return(p)
+}
+
+a <- plotspecificcorrs(vsdhormoneshyp, vsdhormoneshyp$FSHB, vsdhormoneshyp$PRL, "FSHB", "PRL") + labs(subtitle = "Hypothalamus")
+b <- plotspecificcorrs(vsdhormonespit, vsdhormonespit$prolactin, vsdhormonespit$PRL, "circulating prolactin", "PRL") +
+  theme(axis.title.x = element_text(face = "plain")) + labs(subtitle = "Pituitary")
+c <- plotspecificcorrs(vsdhormonesgon, vsdhormonesgon$ESR1, vsdhormonesgon$PGR, "ESR1", "PGR")  + labs(subtitle = "Gonads")
+
+plot_grid(a,b,c,nrow = 1, labels = "auto", label_size = 8)
 ```

### 2020-04-15 analysis/figure-6.Rmd
file for victoria

@@
 ```{r writefiles}
-write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
+#write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
+
+# for victoria
+prlmanip <- candidatevsd %>%
+  filter(gene == "PRL", 
+         treatment %in% c("extend", "m.inc.d8"))
+write.csv(prlmanip, "../results/prlmanip.csv")
 ```

### 2020-04-16 analysis/12_GO.Rmd
curleychampagne genes

@@
 ```{r write}
-write.csv(GO_DEGs, "../results/12_GO_DEGs.csv")
+#write.csv(GO_DEGs, "../results/12_GO_DEGs.csv")
 #write.csv(hyp, "../results/12_GO_DEGs_hyp.csv")

### 2020-04-17 analysis/figure-2.Rmd
now with fig and 2 togetehr

@@
 
+## this file no longer works!
+
+The code here is incomplete. Saving it to use for creating tables.
+
 ```{r setup, include=FALSE}
-library(tidyverse)
-library(cowplot)
-library(kableExtra) 
-library(ggtext) # for italic fonts
-library(ggsignif) # fot stats on plots
 
@@
 
-## DEGs
-
-```{r allDEGs}
-DEG_path <- "../results/DEseq2/"   # path to the data
-DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
-DEG_pathfiles <- paste0(DEG_path, DEG_files)
-#DEG_files
-
-allDEG <- DEG_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name") %>% 
-  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/'), "[", 2))  %>% 
-  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
-  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
-  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
-  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
-  mutate(comparison = paste(down,up, sep = "_")) %>%
-  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
-  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
-dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
-head(allDEG)
-
-allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
-allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
-allDEG$direction <- factor(allDEG$direction, levels = charlevels)
-```
-
 ```{r fig2, fig.width=7, fig.height=7}
-
-makebargraph <- function(whichtissue, myylab, lowlim, higherlim){
-  p <- allDEG %>%
-    filter(tissue == whichtissue,
-           comparison != "control_bldg") %>%
-  ggplot(aes(x = comparison,  fill = direction)) +
-    geom_bar(position = "dodge") +
-    facet_grid(tissue~sex) +
-    theme_B3() +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
-          legend.position = "none")  +
-    guides(fill = guide_legend(nrow = 1)) +
-    labs(x = "Sequential parental care stage comparisons", 
-         y = myylab,
-         subtitle = " ") +
-  scale_fill_manual(values = allcolors,
-                       name = " ",
-                       drop = FALSE) +
-  scale_color_manual(values = allcolors) +
-  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
-            position = position_dodge(width = 1),
-            size = 2, color = "black")  +
-  ylim(lowlim, higherlim)
-  return(p)
-}
-
-# hyp
-p1 <- makebargraph("hypothalamus","DEGs", 0, 1250) + theme(axis.text.x = element_blank(), 
-                                                               axis.title.x = element_blank())
-# pit
-p2 <- makebargraph("pituitary","DEGs", 0, 1250)  +  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), 
-                                                              strip.text.x = element_blank())
-# gon
-p3 <- makebargraph("gonad","DEGs", 0, 1250) +  theme(strip.text.x = element_blank())
-bcd <- plot_grid(p1,p2,p3, nrow = 3, rel_heights = c(1.2,1,1.5), labels = c("b", "c", "d"), label_size = 8)
-
-expdesign <- png::readPNG("../figures/images/fig_fig2a.png")
-expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
-
-plot_grid(expdesign, bcd, labels = c("a", " "), label_size = 8, nrow = 2, rel_heights = c(0.25,1))
 
@@
 
-write.csv(allDEG, "../../musicalgenes/data/allDEG.csv")
+#write.csv(allDEG, "../../musicalgenes/data/allDEG.csv")
 

### 2020-04-17 analysis/figure-1.Rmd
better fig 1-4
now with fig and 2 togetehr

@@
 
+library(kableExtra) 
+library(ggtext) # for italic fonts
+
 source("../R/themes.R")
@@
 
-# Figure 1. Experimental design, tSNE analysis, PCA, and prolactin
+#  Experimental design, tSNE analysis, and bar chars
+
+## tsne
 
@@
 
+## DEGs
+
+```{r allDEGs}
+DEG_path <- "../results/DEseq2/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
+head(allDEG)
+
+allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
+allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
+allDEG$direction <- factor(allDEG$direction, levels = charlevels)
+```
+
 ## make figure
 
-```{r fig1, fig.width=7, fig.height=6}
+```{r fig1, fig.width=7.25, fig.height=7.25}
 
-a <- plottsneelipse(chartsne, chartsne$tissue, allcolors) + labs(subtitle = "~ tissue")    
+a <- png::readPNG("../figures/images/fig_fig1a.png")
+a <- ggdraw() +  draw_image(a, scale = 1)
+
 b <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = " ", subtitle = "~ tissue * sex")    
-c <- plottsneelipse(ftsne, ftsne$tissue, allcolors ) + labs(y = " ", subtitle = "female ~ tissue")
-d <- plottsneelipse(mtsne, mtsne$tissue, allcolors ) + labs(y = " ", subtitle = "male * tissue")
+c <- plottsneelipse(ftsne, ftsne$tissue, allcolors ) + labs(y = " ", subtitle = "~ tissue, females only")
+d <- plottsneelipse(mtsne, mtsne$tissue, allcolors ) + labs(y = " ", subtitle = "~ tissue, males only")
 
-abcd <- plot_grid(a,b,c,d, nrow = 1, labels = c("b", "c", "d", "e"), label_size = 8 )
+bcd <- plot_grid(b,c,d, nrow = 1, labels = c("B", "C", "D"), label_size = 8 )
+abcd <- plot_grid(a, bcd, nrow = 2, labels = c("A"), label_size = 8, rel_heights = c(1,1))
 
-expdesign <- png::readPNG("../figures/images/fig1a_fig1a.png")
-expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
+e <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
+  labs( x = NULL)  + 
+  facet_wrap(~sex, scales = "free")
 
-abcde <- plot_grid(expdesign, abcd, nrow = 2, labels = c("a", "b"), label_size = 8, rel_heights = c(0.5,1))
+g <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
+  labs(  x = NULL) + 
+  facet_wrap(~sex, scales = "free") +
+  theme(strip.text = element_blank())
 
-h <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus ~ treatment")  + facet_wrap(~sex, scales = "free")
-i <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary ~ treatment", y = NULL) + facet_wrap(~sex, scales = "free") + theme(legend.position = )
-j <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads ~ treatment", y = NULL)  + facet_wrap(~sex, scales = "free")
+i <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
+   facet_wrap(~sex, scales = "free") +
+  theme(strip.text = element_blank()) +
+  labs(x = "tSNE1 \n \n \n \n")
 
-hij <- plot_grid(h,i,j, nrow = 1, labels = c("h", "i", "j"), label_size = 8)
+egi <- plot_grid(e,g,i, nrow = 3, rel_heights = c(1,1,1.4),
+                 labels = c("E", "G", "I"), label_size = 8)
 
-fig1 <- plot_grid(abcde, hij, nrow = 2, rel_heights = c(0.6,0.4))
+# hyp
+f <- makebargraph("hypothalamus","DEGs", 0, 1250) + 
+  theme(axis.text.x = element_blank(), 
+        axis.title.x = element_blank())  
+# pit
+h <- makebargraph("pituitary","DEGs", 0, 1250)  +  
+  theme(axis.text.x = element_blank(), 
+        axis.title.x = element_blank(), 
+        strip.text.x = element_blank())  
+# gon
+j <- makebargraph("gonad","DEGs", 0, 1250) +  
+  theme(strip.text.x = element_blank()) +
+  scale_x_discrete(labels = comparisonlabels)
+
+fhj <- plot_grid(f,h,j, nrow = 3, rel_heights = c(1,1,1.4),
+                 labels = c("F", "H", "J"), label_size = 8)
+
+efghij <- plot_grid(egi, fhj, nrow = 1, align = "h", rel_widths = c(1.5,2))
+
+fig1 <- plot_grid(abcd, efghij, nrow = 2, rel_heights = c(1,1.5))
 fig1
-
 ```
@@
 ```{r}
-write.csv(chartsne, "../../musicalgenes/data/tsne.csv")
+#write.csv(chartsne, "../../musicalgenes/data/tsne.csv")
 ```

### 2020-04-17 analysis/figure-1.Rmd
now all of fig2 script in fig 1 script

@@
 
+# supple table 1 of all but control-bldg DEGs
+
 ```{r}
+suppletable1 <- allDEG %>%
+  filter(comparison != "control_bldg") %>%
+  group_by(sex, tissue, comparison) %>%
+  arrange( tissue, sex, direction, gene)
+head(suppletable1)
+
+write_csv(suppletable1, "../results/suppletable1.csv")
+
+# save file for musical genes https://raynamharris.shinyapps.io/musicalgenes/
+#write.csv(allDEG, "../../musicalgenes/data/allDEG.csv")
+
+# tsne files too big to save
 #write.csv(chartsne, "../../musicalgenes/data/tsne.csv")

### 2020-04-17 analysis/figure-2.Rmd
candidate gene now figure 2

@@
 ---
-title: "Fig2"
+title: "Fig3"
+author: "Rayna M Harris"
+date: "3/26/2020"
 output: md_document
@@
 
-## this file no longer works!
+# all things box plots
 
-The code here is incomplete. Saving it to use for creating tables.
+## candidate genes
 
-```{r setup, include=FALSE}
+```{r}
+library(tidyverse)
+library(ggtext)
+library(cowplot)
+library(ggpubr)
 
@@
 knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
-```
 
-```{r fig2, fig.width=7, fig.height=7}
+candidategenes <- c("OXT", "AVP", "GNRH1", "GNRHR", "CGNRH-R",
+                    "AR", "POMC", "AGRP", 
+                       "CRH", "AVPR1A", "AVPR1B", "AVPR2","VIP",
+                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9", 
+                    "ESR1","ESR2", "LBH", "CDK1", "BRCA1",
+                    "PTEN", "CREBBP", "FOS", "JUN", "EGR1",
+                     "BDNF", "GRM2","GRIA1",
+                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16", 
+                    "DIO3", "DIO2", "DIO1",
 
-```
-
-## total degs
-
-```{r}
-allDEG %>%
-  group_by(sex, tissue, comparison) %>%
-  summarize(totalDEGs = n()) %>%
-  arrange(tissue, comparison)
-```
-
-## candidate genes 
-```{r writefiles}
-geneids <- read_csv("../metadata/00_geneinfo.csv")
+                     "VIPR1", "VIPR2", "NPY", "LH", "FSH", "FSHB",
+                    "JAK2", "HCRT", "TRH", "TSHB",
+                    "MC3R", "MC4R", "MC5R",  "NR3C1", "NR3C2",
+                    "STAT5B","NPVF")
 
@@
                     "ESR1","ESR2", "LBH",  
+                    "FOS", "JUN", "EGR1", "BDNF")
 
-                    "FOS", "JUN", "EGR1", "BDNF"
+# candidategenesslim: AR, AVP, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, EGR1, ESR1, ESR2, FOS, GNRH1, GNRHR, JUN, LBH, OXT, PRL, PRLR, VIP
+# DEG candidategenesslim: AR, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, ESR1, FOS, GNRHR, LBH, PRL, PRLR, 
+# non-DEG candidategenesslim: AVP, AVPR1B, ESR2,  GNRH1, JUN, OXT, VIP
+
+hormonepapergenes <- c("GNRH1", "CGNRH-R", "GNRHR",  
+                       "FSHB", "FSHR", "LHCGR", # no LSH
+                       "VIP", "VIPR1",
+                       "ESR1", "AR", "CYP19A1", "HSD17B1", # receptors synthesis
+                       "PGR" , "HSD3B2", 
+                       "PRL", "PRLR")
+
+curleychampagnegenes <- c("AVPR1A", "ADRA2A",    
+                         "COMT", "CRH", "CRHBP", "CRHR1", "CRHR2", 
+                          "DRD1", "DRD4","ESR1", "ESR2", #ERa ERbeta
+                          "FOS", "HTR2C", "MEST", "NR3C1", #GR
+                          "OPRM1", "PGR", "PRL", "PRLR",  "SLC6A4" #5HTT
                          )
 
-table1 <- allDEG %>%
-  mutate(updown = ifelse(lfc > 0, "+", "-"))  %>%
-  mutate(geneupdown = paste(gene, updown, sep = "")) %>%
-  filter(gene %in% candidategenes) %>%
-  arrange(geneupdown) %>%
-  group_by(sex, tissue, comparison) %>%
-  summarize(genes = str_c(geneupdown, collapse = " ")) %>%
-  pivot_wider(names_from = comparison, values_from = genes ) %>%
-  select(sex, tissue, control_bldg, bldg_lay, lay_inc.d3, inc.d3_inc.d9,
-        inc.d9_inc.d17, hatch_n5)  %>%
-  arrange( tissue, sex)
-kable(table1)
+parentalcareGO <- c("AVP", "AVPR1A", "BRINP1", "CREBRF", "DBH", "DRD1",  
+                    "GNAQ", "HUBB", "KALRN", "MBD2","NPAS1", "NPAS3",
+                     "NR3C1", "OPRK1", "OXT", "OXTR", "PRL", "PTEN")
 
-write_csv(table1, "../results/table1.csv")
+# can't find NOS1 or OXTR or FOX1B or HTR5A
 ```
 
-## supplemental tables, all DEGS
+## variance stabilized gene expression  (vsd)
 
-```{r}
+```{r }
+geneids <- read_csv("../metadata/00_geneinfo.csv") %>% select(-X1)
 
-suppletable1 <- allDEG %>%
-  filter(comparison != "control_bldg") %>%
-  group_by(sex, tissue, comparison) %>%
-  arrange( tissue, sex, direction, gene)
-head(suppletable1)
+vsd_path <- "../results/DEseq2/"   # path to the data
+vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
+vsd_pathfiles <- paste0(vsd_path, vsd_files)
+vsd_files
 
-suppletable1 %>%
-  group_by(tissue) %>%
-  summarize(totalDEGs = n())
+allvsd <- vsd_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name")  %>% 
+  dplyr::rename("gene" = "X1") %>% 
+  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "samples", values_to = "counts")
 
-write_csv(suppletable1, "../results/suppletable1.csv")
+getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
+  candidates  <- allvsd %>%
+    filter(gene %in% whichgenes) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
+    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
+    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
+    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
+    filter(tissue == whichtissue, sex %in% whichsex)  %>%
+    drop_na()
+  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
+  return(candidates)
+}
 
-head(allDEG) 
+hypvsd <- getcandidatevsd(curleychampagnegenes, "hypothalamus", sexlevels)
+pitvsd <- getcandidatevsd(curleychampagnegenes, "pituitary", sexlevels)
+gonvsd <- getcandidatevsd(curleychampagnegenes, "gonad", sexlevels)
+head(hypvsd)
+tail(gonvsd)
+
+candidatevsd <- rbind(hypvsd, pitvsd)
+candidatevsd <- rbind(candidatevsd, gonvsd)
+
+unique(candidatevsd$gene)
 ```
 
-```{r savefiles}
-allDEG %>%
-  drop_na()
+## Figs
 
-#write.csv(allDEG, "../../musicalgenes/data/allDEG.csv")
+```{r }
+# summary DEG results from DESeq2
+candidateDEGS <- read_csv("../../musicalgenes/data/allDEG.csv") %>%
+  select(-X1) %>%
+  filter(gene %in% curleychampagnegenes)
+head(candidateDEGS)
+
+# get genes to plot
+candidateDEGS %>% filter(tissue == "hypothalamus", sex == "female", comparison != "control_bldg") 
+candidateDEGS %>% filter(tissue == "hypothalamus", sex == "male", comparison != "control_bldg") 
+candidateDEGS %>% filter(tissue == "pituitary", sex == "male", comparison != "control_bldg") 
+candidateDEGS %>% filter(tissue == "pituitary", sex == "female", comparison != "control_bldg") 
+candidateDEGS %>% filter(tissue == "gonad", sex == "female", comparison != "control_bldg")
+
+# summarize all differences
+candidateDEGtable <- candidateDEGS %>%
+  mutate(posneg = ifelse(lfc >= 0, "+", "-")) %>%
+  mutate(geneposneg = paste(gene, posneg, sep = "")) %>%
+  select(sex, tissue, comparison, geneposneg) %>%
+  group_by(sex, tissue, comparison) %>%
+   arrange(geneposneg) %>%
+  summarize(genes = str_c(geneposneg, collapse = " ")) %>% 
+  pivot_wider(names_from = comparison, values_from = genes)  %>% 
+  ungroup(tissue) %>%
+  mutate(tissue = factor(tissue, levels =  tissuelevel))  %>% 
+  arrange(sex, tissue)  %>% 
+  select(sex, tissue, bldg_lay, lay_inc.d3, inc.d3_inc.d9,  
+         inc.d9_inc.d17,  hatch_n5, control_bldg)
+candidateDEGtable
+```
+
+```{r fig3, fig.width = 7}
+
+newboxplot <- function(df, mygenes, whichsex, whichstage){
+
+  p <- df %>%
+    filter(gene %in% mygenes, 
+           sex %in% whichsex,
+           treatment %in% whichstage) %>%
+    droplevels() %>%
+    mutate(tissue = factor(tissue, levels = tissuelevel)) %>%
+    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
+    geom_boxplot() +
+    geom_jitter(size = 0.5, width = 0.1) +
+    facet_grid(tissue~gene, scales = "free_y") +
+    theme_B3() +
+    scale_color_manual(values = allcolors) +
+    scale_fill_manual(values = allcolors) +
+    theme(legend.position = "none",
+          axis.text.x = element_text(angle = 45, hjust = 1),
+          strip.text.x = element_text(face = "italic"),
+          axis.title.x = element_blank())  +
+    labs(x = "Parental stage",
+         y = "variance stabilized expression")
+  return(p)
+}
+
+a1 <- newboxplot(hypvsd, c("PGR",  "CYP19A1", "LHCGR" ), "female", c( "hatch", "n5")) + 
+  theme(strip.text.y = element_blank()) + labs(subtitle = "Females")
+a2 <- newboxplot(hypvsd, c("AR"), "male", c( "hatch", "n5")) + 
+  theme(axis.title.y = element_blank()) + labs(subtitle = "Males")
+
+b1 <- newboxplot(gonvsd, c("HSD3B2"), "female", c( "bldg", "lay"))  +
+  theme(strip.text.y = element_blank(),
+        axis.title.y = element_blank()) + labs(subtitle = "Females")
+b2 <- newboxplot(gonvsd, c("VIPR1",  "PRLR" ), "female", c( "lay", "inc.d3"))  + 
+  theme(axis.title.y = element_blank()) + labs(subtitle = " ")
+
+ab <- plot_grid(a1,a2,b1,b2, rel_widths = c(6,2.5,2,4), labels = c("a", " ", "b", " "), label_size = 8, nrow = 1)
+
+c1 <- newboxplot(pitvsd, c( "ESR1", "GNRHR"), "female", c("bldg", "lay",  "inc.d3")) 
+c2 <- newboxplot(pitvsd, c( "PRL"), "female", c("inc.d9", "inc.d17", "hatch", "n5")) 
+c3 <- newboxplot(pitvsd, c("VIPR1",  "PRL" ), "male", c( "inc.d9", "inc.d17"))
+
+c <- plot_grid(c1 + theme(strip.text.y = element_blank()) + labs(subtitle = "Females"),
+          c2 + theme(strip.text.y = element_blank(),
+                    axis.title.y = element_blank()) + labs(subtitle = " ") ,
+          c3 + theme(axis.title.y = element_blank()) + labs(subtitle = "Males"), 
+          nrow = 1, rel_widths = c(6,4,4),
+          labels = c("c"), label_size = 8)
+
+plot_grid(ab,c, nrow = 2)
+
+# correlations
+library(corrr)
+
+hormones <- read_csv("../results/07_hormoneswide.csv") %>%
+  select(-study, -treatment, -sex)
+head(hormones)
+
+head(candidatevsd)
+
+vsdhormones <- candidatevsd %>%
+  pivot_wider(names_from = gene, values_from = counts) %>%
+  mutate(bird_id = sapply(strsplit(samples,"\\_"), "[", 1)) %>%
+  left_join(.,hormones, by =  "bird_id")
+vsdhormones
+
+vsdhormoneshyp <- vsdhormones %>% filter(tissue == "hypothalamus" )
+vsdhormonespit <- vsdhormones %>% filter(tissue == "pituitary" )
+vsdhormonesgon <- vsdhormones %>% filter(tissue == "gonad" )
+
+plotcorrs <- function(whichtissue, whichsex){
+
+  x <- vsdhormones  %>%
+    filter(tissue == whichtissue, 
+           sex == whichsex) %>%
+    select(-sex, - tissue, - treatment, -samples, -bird_id, -moltbin) %>%
+    correlate()
+  print(fashion(x))
+
+  p <-rplot(x, colors = c("#0571b0", "#92c5de", "#f7f7f7","#f7f7f7","#f7f7f7", "#f4a582", "#ca0020")) +
+    theme_B3() +
+    theme(legend.position = "bottom",
+          axis.text.x = element_text(angle = 45, hjust = 1),
+          axis.text = element_text(face = "italic")) +
+    labs(subtitle = whichtissue)
+  return(p)
+}
+
+a <- plotcorrs("hypothalamus", "female") + theme(legend.position = "none")
+b <- plotcorrs("pituitary", "female") + theme(axis.text.y = element_blank()) + theme(legend.position = "none")
+c <- plotcorrs("gonad", "female") + theme(axis.text.y = element_blank()) + theme(legend.position = "none")
+
+d <- plotcorrs("hypothalamus", "male") 
+e <- plotcorrs("pituitary", "male") + theme(axis.text.y = element_blank())
+f <- plotcorrs("gonad", "male") + theme(axis.text.y = element_blank())
+
+plot_grid(a,b,c, d,e,f, nrow = 2, rel_widths = c(1.2,1,1), rel_heights = c(1,1.2),
+          labels = c("females", " ", " ",
+                     "males", " ", ""),
+          label_size = 8)
+
+plotspecificcorrs <- function(df, myx, myy, myxlab, myylab){
+  p <- df %>%
+    mutate(tissue = factor(tissue, levels = tissuelevel)) %>%
+    ggplot(aes(x = myx, y = myy)) +
+    geom_point(aes(color = treatment)) +
+    geom_smooth(method = "lm", aes(color = sex)) +
+    facet_wrap(~sex, scales = "free", nrow = 2) +
+    scale_color_manual(values = allcolors) +
+    labs(x = myxlab, y = myylab) +
+    theme_B3() +
+    theme(legend.position = "none",
+          axis.title = element_text(face = "italic"))
+  return(p)
+}
+
+a <- plotspecificcorrs(vsdhormoneshyp, vsdhormoneshyp$FSHB, vsdhormoneshyp$PRL, "FSHB", "PRL") + labs(subtitle = "Hypothalamus")
+b <- plotspecificcorrs(vsdhormonespit, vsdhormonespit$prolactin, vsdhormonespit$PRL, "circulating prolactin", "PRL") +
+  theme(axis.title.x = element_text(face = "plain")) + labs(subtitle = "Pituitary")
+c <- plotspecificcorrs(vsdhormonesgon, vsdhormonesgon$ESR1, vsdhormonesgon$PGR, "ESR1", "PGR")  + labs(subtitle = "Gonads")
+
+plot_grid(a,b,c,nrow = 1, labels = "auto", label_size = 8)
+```
+
+```{r writefiles}
+#write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
+#write.csv(candidatevsd, "../results/candidatecounts.csv")
+write.csv(candidateDEGtable, "../results/candidateDEGtable.csv")
 

### 2020-04-17 analysis/figure-2.Rmd
remove old candidate gene lists

@@
 knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
-
-candidategenes <- c("OXT", "AVP", "GNRH1", "GNRHR", "CGNRH-R",
-                    "AR", "POMC", "AGRP", 
-                       "CRH", "AVPR1A", "AVPR1B", "AVPR2","VIP",
-                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9", 
-                    "ESR1","ESR2", "LBH", "CDK1", "BRCA1",
-                    "PTEN", "CREBBP", "FOS", "JUN", "EGR1",
-                     "BDNF", "GRM2","GRIA1",
-                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16", 
-                    "DIO3", "DIO2", "DIO1",
-
-                     "VIPR1", "VIPR2", "NPY", "LH", "FSH", "FSHB",
-                    "JAK2", "HCRT", "TRH", "TSHB",
-                    "MC3R", "MC4R", "MC5R",  "NR3C1", "NR3C2",
-                    "STAT5B","NPVF")
-
-candidategenesslim <- c("OXT", "AVP", "GNRH1", "GNRHR", 
-                    "AR",  "CYP19A1", 
-                    "AVPR1A", "AVPR1B", "AVPR2","VIP",
-                    "DRD1", "DRD2", 
-                    "PRL", "PRLR",  
-                    "ESR1","ESR2", "LBH",  
-                    "FOS", "JUN", "EGR1", "BDNF")
-
-# candidategenesslim: AR, AVP, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, EGR1, ESR1, ESR2, FOS, GNRH1, GNRHR, JUN, LBH, OXT, PRL, PRLR, VIP
-# DEG candidategenesslim: AR, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, ESR1, FOS, GNRHR, LBH, PRL, PRLR, 
-# non-DEG candidategenesslim: AVP, AVPR1B, ESR2,  GNRH1, JUN, OXT, VIP
 

### 2020-04-17 analysis/figure-2.Rmd
new candidate gene/GO analysis

@@
 ---
-title: "Fig3"
+title: "Fig2: Candidate genes"
 author: "Rayna M Harris"
@@
 
-# all things box plots
-
-## candidate genes
+# Candidate gene analysis
 
@@
 library(ggpubr)
+library(knitr)
+library(kableExtra)
 
@@
 knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
+```
 
-hormonepapergenes <- c("GNRH1", "CGNRH-R", "GNRHR",  
-                       "FSHB", "FSHR", "LHCGR", # no LSH
-                       "VIP", "VIPR1",
-                       "ESR1", "AR", "CYP19A1", "HSD17B1", # receptors synthesis
-                       "PGR" , "HSD3B2", 
-                       "PRL", "PRLR")
+## Candidate Genes (from literature and relevant GO terms)
 
+```{r GO}
+
+# all genes in this parental care dataset
+
+geneids <- read_csv("../metadata/00_geneinfo.csv") %>%
+  select(-X1) %>%
+  rename("NCBI" = "entrezid",
+         "gene" = "Name")
+
+# From Ch 17 Evolution of parental care
 curleychampagnegenes <- c("AVPR1A", "ADRA2A",    
@@
                           "FOS", "HTR2C", "MEST", "NR3C1", #GR
-                          "OPRM1", "PGR", "PRL", "PRLR",  "SLC6A4" #5HTT
-                         )
+                          "OPRM1", "PGR", "PRL", "PRLR",  "SLC6A4") #5HTT
+curleychampagnegenes <- as.data.frame(curleychampagnegenes) %>%
+  mutate(GO = "parentalcare", gene = curleychampagnegenes)  %>%
+  select(GO, gene)
+curleychampagnegenes
 
-parentalcareGO <- c("AVP", "AVPR1A", "BRINP1", "CREBRF", "DBH", "DRD1",  
-                    "GNAQ", "HUBB", "KALRN", "MBD2","NPAS1", "NPAS3",
-                     "NR3C1", "OPRK1", "OXT", "OXTR", "PRL", "PTEN")
+# Candidate GO terms 
+GO_path <- "../metadata/goterms/"   # path to the data
+GO_files <- dir(GO_path, pattern = "*.txt") # get file names
+GO_pathfiles <- paste0(GO_path, GO_files)
 
-# can't find NOS1 or OXTR or FOX1B or HTR5A
+GOgenesLong <- GO_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_table(.x, col_types = cols(), col_names = FALSE), .id = "file_name") %>% 
+  mutate(GO = sapply(strsplit(as.character(file_name),'../metadata/goterms/'), "[", 2)) %>% 
+  mutate(GO = sapply(strsplit(as.character(GO),'.txt'), "[", 1)) %>% 
+  mutate(gene = sapply(strsplit(as.character(X1), "[\\\\]|[^[:print:]]" ), "[", 2)) %>% 
+  select(GO, gene)  %>%
+  filter(gene != "Symbol") %>%
+  distinct(GO,gene)  %>%
+  mutate(gene = toupper(gene)) %>%
+  rbind(., curleychampagnegenes) %>%
+  arrange(gene) %>%
+  left_join(., geneids, by = "gene") %>%
+  drop_na() 
+GOgenesLong
+
+GOgenesWide <- GOgenesLong %>% 
+  pivot_wider(
+    names_from = GO,
+    values_from = gene) %>%
+  mutate(numGOs = 4 - rowSums(is.na(.))) %>%
+  arrange(desc(numGOs)) %>%
+  select(-numGOs) %>%
+  left_join(geneids, by = c("geneid","NCBI"))
+GOgenesWide
+
+# genes I can't find in dataset 
+# NOS1 or OXTR or FOX1B or HTR5A
+```
+
+## Candidate DEGs
+
+```{r DEGs}
+
+candidategenes <- GOgenesLong %>% distinct(gene) %>% pull(gene)
+
+# summary DEG results from DESeq2
+candidateDEGS <- read_csv("../results/suppletable1.csv") %>%
+  filter(gene %in% candidategenes) %>%
+  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
+         sex = recode(sex, "female" = "F", "male" = "M" ),
+         tissue = recode(tissue, 
+                         "hypothalamus" = "H",
+                         "pituitary" = "P", "gonad" = "G")) %>%
+  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
+  select(gene, res, comparison)  %>%
+  group_by(gene,  comparison) %>%
+  summarize(res = str_c(res, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = res) %>%
+  select(gene, bldg_lay, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
+candidateDEGS
+
+table1 <- left_join(candidateDEGS, GOgenesWide) %>%
+  select(gene, bldg_lay:n5_n9, parentalcare, parentalbehavior, NCBI, ) %>%
+  mutate(parentalcare = if_else(is.na(parentalcare), " ", "X"),
+         parentalbehavior = if_else(is.na(parentalbehavior), " ", "X")) %>%
+  rename("Literature" = "parentalcare", "GO" =  "parentalbehavior")
+table1[is.na(table1)] <- " "
+table1
+
+kable(table1)
 ```
@@
 
-```{r }
-geneids <- read_csv("../metadata/00_geneinfo.csv") %>% select(-X1)
+```{r eval = F}
 
@@
 
-hypvsd <- getcandidatevsd(curleychampagnegenes, "hypothalamus", sexlevels)
-pitvsd <- getcandidatevsd(curleychampagnegenes, "pituitary", sexlevels)
-gonvsd <- getcandidatevsd(curleychampagnegenes, "gonad", sexlevels)
-head(hypvsd)
-tail(gonvsd)
+head(allvsd)
+candidategenes <- GOgenesLong %>% distinct(gene) %>% pull(gene)
 
+hypvsd <- getcandidatevsd(candidategenes, "hypothalamus", sexlevels)
+pitvsd <- getcandidatevsd(candidategenes, "pituitary", sexlevels)
+gonvsd <- getcandidatevsd(candidategenes, "gonad", sexlevels)
 candidatevsd <- rbind(hypvsd, pitvsd)
 candidatevsd <- rbind(candidatevsd, gonvsd)
-
-unique(candidatevsd$gene)
+head(candidatevsd)
 ```
 
-## Figs
-
-```{r }
-# summary DEG results from DESeq2
-candidateDEGS <- read_csv("../../musicalgenes/data/allDEG.csv") %>%
-  select(-X1) %>%
-  filter(gene %in% curleychampagnegenes)
-head(candidateDEGS)
-
-# get genes to plot
-candidateDEGS %>% filter(tissue == "hypothalamus", sex == "female", comparison != "control_bldg") 
-candidateDEGS %>% filter(tissue == "hypothalamus", sex == "male", comparison != "control_bldg") 
-candidateDEGS %>% filter(tissue == "pituitary", sex == "male", comparison != "control_bldg") 
-candidateDEGS %>% filter(tissue == "pituitary", sex == "female", comparison != "control_bldg") 
-candidateDEGS %>% filter(tissue == "gonad", sex == "female", comparison != "control_bldg")
-
-# summarize all differences
-candidateDEGtable <- candidateDEGS %>%
-  mutate(posneg = ifelse(lfc >= 0, "+", "-")) %>%
-  mutate(geneposneg = paste(gene, posneg, sep = "")) %>%
-  select(sex, tissue, comparison, geneposneg) %>%
-  group_by(sex, tissue, comparison) %>%
-   arrange(geneposneg) %>%
-  summarize(genes = str_c(geneposneg, collapse = " ")) %>% 
-  pivot_wider(names_from = comparison, values_from = genes)  %>% 
-  ungroup(tissue) %>%
-  mutate(tissue = factor(tissue, levels =  tissuelevel))  %>% 
-  arrange(sex, tissue)  %>% 
-  select(sex, tissue, bldg_lay, lay_inc.d3, inc.d3_inc.d9,  
-         inc.d9_inc.d17,  hatch_n5, control_bldg)
-candidateDEGtable
-```
-
-```{r fig3, fig.width = 7}
+```{r fig2, fig.width = 7, eval = F}
 
@@
 #write.csv(candidatevsd, "../results/candidatecounts.csv")
-write.csv(candidateDEGtable, "../results/candidateDEGtable.csv")
-
+write.csv(table1, "../results/table1.csv")
 ```

### 2020-04-17 analysis/figure-2.Rmd
now with candidate genes as fig 2

@@
 table1 <- left_join(candidateDEGS, GOgenesWide) %>%
-  select(gene, bldg_lay:n5_n9, parentalcare, parentalbehavior, NCBI, ) %>%
+  select(gene, bldg_lay:n5_n9, parentalcare, parentalbehavior, NCBI) %>%
   mutate(parentalcare = if_else(is.na(parentalcare), " ", "X"),
@@
 
-```{r eval = F}
+```{r vsds}
 
@@
 
-head(allvsd)
 candidategenes <- GOgenesLong %>% distinct(gene) %>% pull(gene)
@@
 
-```{r fig2, fig.width = 7, eval = F}
+```{r fig2, fig.width = 7.25, fig.height = 7.25}
 
-newboxplot <- function(df, mygenes, whichsex, whichstage){
+# significant candidate genes
+hypDEGs <- c("CRHBP", "CRHR2", "DRD1", "GNAQ", 
+             "HTR2C", "OPRK1", "PGR")
+pitDEGs <- c("ESR1", "MEST", "PRL", "PTEN")
+gonadDEGs <- c("AVPR1A", "CREBRF", "FOS", "GNAQ", 
+               "NR3C1", "OPRM1", "PRLR")
 
-  p <- df %>%
-    filter(gene %in% mygenes, 
-           sex %in% whichsex,
-           treatment %in% whichstage) %>%
-    droplevels() %>%
-    mutate(tissue = factor(tissue, levels = tissuelevel)) %>%
-    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
-    geom_boxplot() +
-    geom_jitter(size = 0.5, width = 0.1) +
-    facet_grid(tissue~gene, scales = "free_y") +
+makelineplots <- function(whichtissue, whichgenes){
+
+  p <- candidatevsd %>%
+    filter(tissue == whichtissue,
+          gene %in% whichgenes) %>%
+    mutate(treatment = factor(treatment, levels = charlevels)) %>%
+    mutate(treatmentNum = as.numeric(treatment)) %>%
+    ggplot(aes(x = treatmentNum, y = counts)) +
+    geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
+    #geom_smooth(aes(color = sex)) +
+    geom_jitter(size = 0.5, aes(color = sex)) +
+    facet_grid(gene~sex, scales = "free") +
+    scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9),
+                       labels = charlevels) +
+    scale_fill_manual(values = allcolors) +
+    scale_color_manual(values = allcolors) +
     theme_B3() + 
-    scale_color_manual(values = allcolors) +
-    scale_fill_manual(values = allcolors) +
     theme(legend.position = "none",
-          axis.text.x = element_text(angle = 45, hjust = 1),
-          strip.text.x = element_text(face = "italic"),
-          axis.title.x = element_blank())  +
-    labs(x = "Parental stage",
-         y = "variance stabilized expression")
+          strip.text.y = element_text(face = "italic"),
+          axis.text.x = element_text(angle = 45, hjust = 1)) +
+    labs(subtitle = whichtissue, 
+         y = "gene expression (variance stabilized)",
+         x = "Parental Care Stage") 
   return(p)
@@
 
-a1 <- newboxplot(hypvsd, c("PGR",  "CYP19A1", "LHCGR" ), "female", c( "hatch", "n5")) + 
-  theme(strip.text.y = element_blank()) + labs(subtitle = "Females")
-a2 <- newboxplot(hypvsd, c("AR"), "male", c( "hatch", "n5")) + 
-  theme(axis.title.y = element_blank()) + labs(subtitle = "Males")
+a <- makelineplots("hypothalamus", hypDEGs)
+b <- makelineplots("pituitary", pitDEGs) + labs(y = NULL)
+c <- makelineplots("gonad", gonadDEGs) + labs(y = NULL)
 
-b1 <- newboxplot(gonvsd, c("HSD3B2"), "female", c( "bldg", "lay"))  +
-  theme(strip.text.y = element_blank(),
-        axis.title.y = element_blank()) + labs(subtitle = "Females")
-b2 <- newboxplot(gonvsd, c("VIPR1",  "PRLR" ), "female", c( "lay", "inc.d3"))  + 
-  theme(axis.title.y = element_blank()) + labs(subtitle = " ")
-
-ab <- plot_grid(a1,a2,b1,b2, rel_widths = c(6,2.5,2,4), labels = c("a", " ", "b", " "), label_size = 8, nrow = 1)
-
-c1 <- newboxplot(pitvsd, c( "ESR1", "GNRHR"), "female", c("bldg", "lay",  "inc.d3")) 
-c2 <- newboxplot(pitvsd, c( "PRL"), "female", c("inc.d9", "inc.d17", "hatch", "n5")) 
-c3 <- newboxplot(pitvsd, c("VIPR1",  "PRL" ), "male", c( "inc.d9", "inc.d17"))
-
-c <- plot_grid(c1 + theme(strip.text.y = element_blank()) + labs(subtitle = "Females"),
-          c2 + theme(strip.text.y = element_blank(),
-                    axis.title.y = element_blank()) + labs(subtitle = " ") ,
-          c3 + theme(axis.title.y = element_blank()) + labs(subtitle = "Males"), 
-          nrow = 1, rel_widths = c(6,4,4),
-          labels = c("c"), label_size = 8)
-
-plot_grid(ab,c, nrow = 2)
-
-# correlations
-library(corrr)
-
-hormones <- read_csv("../results/07_hormoneswide.csv") %>%
-  select(-study, -treatment, -sex)
-head(hormones)
-
-head(candidatevsd)
-
-vsdhormones <- candidatevsd %>%
-  pivot_wider(names_from = gene, values_from = counts) %>%
-  mutate(bird_id = sapply(strsplit(samples,"\\_"), "[", 1)) %>%
-  left_join(.,hormones, by =  "bird_id")
-vsdhormones
-
-vsdhormoneshyp <- vsdhormones %>% filter(tissue == "hypothalamus" )
-vsdhormonespit <- vsdhormones %>% filter(tissue == "pituitary" )
-vsdhormonesgon <- vsdhormones %>% filter(tissue == "gonad" )
-
-plotcorrs <- function(whichtissue, whichsex){
-
-  x <- vsdhormones  %>%
-    filter(tissue == whichtissue, 
-           sex == whichsex) %>%
-    select(-sex, - tissue, - treatment, -samples, -bird_id, -moltbin) %>%
-    correlate()
-  print(fashion(x))
-
-  p <-rplot(x, colors = c("#0571b0", "#92c5de", "#f7f7f7","#f7f7f7","#f7f7f7", "#f4a582", "#ca0020")) +
-    theme_B3() +
-    theme(legend.position = "bottom",
-          axis.text.x = element_text(angle = 45, hjust = 1),
-          axis.text = element_text(face = "italic")) +
-    labs(subtitle = whichtissue)
-  return(p)
-}
-
-a <- plotcorrs("hypothalamus", "female") + theme(legend.position = "none")
-b <- plotcorrs("pituitary", "female") + theme(axis.text.y = element_blank()) + theme(legend.position = "none")
-c <- plotcorrs("gonad", "female") + theme(axis.text.y = element_blank()) + theme(legend.position = "none")
-
-d <- plotcorrs("hypothalamus", "male") 
-e <- plotcorrs("pituitary", "male") + theme(axis.text.y = element_blank())
-f <- plotcorrs("gonad", "male") + theme(axis.text.y = element_blank())
-
-plot_grid(a,b,c, d,e,f, nrow = 2, rel_widths = c(1.2,1,1), rel_heights = c(1,1.2),
-          labels = c("females", " ", " ",
-                     "males", " ", ""),
-          label_size = 8)
-
-plotspecificcorrs <- function(df, myx, myy, myxlab, myylab){
-  p <- df %>%
-    mutate(tissue = factor(tissue, levels = tissuelevel)) %>%
-    ggplot(aes(x = myx, y = myy)) +
-    geom_point(aes(color = treatment)) +
-    geom_smooth(method = "lm", aes(color = sex)) +
-    facet_wrap(~sex, scales = "free", nrow = 2) +
-    scale_color_manual(values = allcolors) +
-    labs(x = myxlab, y = myylab) +
-    theme_B3() +
-    theme(legend.position = "none",
-          axis.title = element_text(face = "italic"))
-  return(p)
-}
-
-a <- plotspecificcorrs(vsdhormoneshyp, vsdhormoneshyp$FSHB, vsdhormoneshyp$PRL, "FSHB", "PRL") + labs(subtitle = "Hypothalamus")
-b <- plotspecificcorrs(vsdhormonespit, vsdhormonespit$prolactin, vsdhormonespit$PRL, "circulating prolactin", "PRL") +
-  theme(axis.title.x = element_text(face = "plain")) + labs(subtitle = "Pituitary")
-c <- plotspecificcorrs(vsdhormonesgon, vsdhormonesgon$ESR1, vsdhormonesgon$PGR, "ESR1", "PGR")  + labs(subtitle = "Gonads")
-
-plot_grid(a,b,c,nrow = 1, labels = "auto", label_size = 8)
+plot_grid(a,b,c, nrow = 1, labels = "AUTO", label_size = 8)
 ```

### 2020-04-17 analysis/figure-2.Rmd
smaller points

@@
     #geom_smooth(aes(color = sex)) +
-    geom_jitter(size = 0.5, aes(color = sex)) +
+    geom_jitter(size = 0.25, aes(color = sex)) +
     facet_grid(gene~sex, scales = "free") +

### 2020-04-17 analysis/figure-2.Rmd
wide figure

--- 2020-04-17 analysis/figure-2.Rmd
@@
 
-makelineplots <- function(whichtissue, whichgenes){
+candidateboxplot <- function(whichtissue, whichgenes, whichsex){
 
   p <- candidatevsd %>%
-    filter(tissue == whichtissue,
-          gene %in% whichgenes) %>%
+    filter(tissue %in% whichtissue,
+          gene %in% whichgenes,
+          sex %in% whichsex) %>%
     mutate(treatment = factor(treatment, levels = charlevels)) %>%
@@
     #geom_smooth(aes(color = sex)) +
-    geom_jitter(size = 0.5, aes(color = sex)) +
-    facet_grid(gene~sex, scales = "free") +
+    geom_jitter(size = 0.25, aes(color = sex)) +
+    facet_wrap(~gene, scales = "free_y", nrow = 1) +
     scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9),
@@
     theme(legend.position = "none",
-          strip.text.y = element_text(face = "italic"),
-          axis.text.x = element_text(angle = 45, hjust = 1)) +
-    labs(subtitle = whichtissue, 
-         y = "gene expression (variance stabilized)",
-         x = "Parental Care Stage") 
+          strip.text = element_text(face = "italic"),
+          axis.text.x = element_blank()) +
+    labs(subtitle = whichsex,
+         y = "gene expression",
+         x = NULL) 
   return(p)
@@
 
-a <- makelineplots("hypothalamus", hypDEGs)
-b <- makelineplots("pituitary", pitDEGs) + labs(y = NULL)
-c <- makelineplots("gonad", gonadDEGs) + labs(y = NULL)
+a1 <- candidateboxplot("hypothalamus", hypDEGs, "female") + labs(x = NULL, title = "hypothalamus") 
+a2 <- candidateboxplot("hypothalamus", hypDEGs, "male") + labs(x = NULL) +
+  theme(strip.text = element_blank())
 
-plot_grid(a,b,c, nrow = 1, labels = "AUTO", label_size = 8)
+b1 <- candidateboxplot("pituitary", pitDEGs, "female") + labs(x = NULL, title = "pituitary")
+b2 <- candidateboxplot("pituitary", pitDEGs, "male") + labs(x = NULL) +
+  theme(strip.text = element_blank())
+
+c1 <- candidateboxplot("gonad", gonadDEGs, "female") + labs(x = NULL, title = "gonad")
+c2 <- candidateboxplot("gonad", gonadDEGs, "male") + 
+  theme(strip.text = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1))
+
+plot_grid(a1,a2, b1,b2, c1,c2, nrow = 6, labels = c("A", " ","B", " ", "C", " "), label_size = 8,
+          rel_heights = c(1, 0.75, 1, 0.75, 1, 1))
 ```

### 2020-04-18 analysis/figure-1.Rmd
now with full table

@@
 suppletable1 <- allDEG %>%
-  filter(comparison != "control_bldg") %>%
+  #filter(comparison != "control_bldg") %>%
   group_by(sex, tissue, comparison) %>%

### 2020-04-18 analysis/figure-2.Rmd
now with full table

@@
   rename("Literature" = "parentalcare", "GO" =  "parentalbehavior")
-table1[is.na(table1)] <- " "
-table1
+table1$numDEGs <- rowSums(is.na(table1)) # count NAs to know how many are NS
+table1 <- table1 %>% 
+  mutate(sig = ifelse(numDEGs == 6, "NS", "DEG")) %>% 
+  arrange(sig, gene)  %>%  select(-sig, -numDEGs)
+table1[is.na(table1)] <- " " # replace NA with blank space so it's pretty
+#table1
 

### 2020-04-18 analysis/figure-2.Rmd
first pass at new WGCNA

@@
 write.csv(table1, "../results/table1.csv")
+write.csv(candidategenes, "../results/candidategenes.csv")
 ```

### 2020-04-18 analysis/08_WGCNA_1.R

@@ -1,11 +1,4 @@
-
-
-#=====================================================================================
-#
-#  Code chunk 1
-#
-#=====================================================================================
-
+# WGCNA pipeline
 
 # Load the packages
 library(tidyverse)
@@ -14,93 +7,61 @@ library(dplyr)
 library(magrittr)
 library(forcats)
 
-
-# read the count data
-countData <- read.csv("../results/00_countData_characterization.csv", row.names = 1)
-head(countData[1:2])
+source("../R/functions.R")
 
 # read the sample meta data or column data
-colData <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, stringsAsFactors = T)
-
-
-## rename rownames and colnames for better vizualiation
-colData <- colData %>%
-  mutate(sex = fct_recode(sex,
-                          "F" = "female",
-                          "M" = "male"),
+colData <- read_csv("../metadata/00_colData_characterization.csv") %>%
+  mutate(sex = factor(sex),
+         tissue = factor(tissue),
+         treatment = factor(treatment)) %>%
+  mutate(sex = fct_recode(sex, "F" = "female", "M" = "male"),
          tissue = fct_recode(tissue,
-                             "pit" = "pituitary",
-                             "hyp" = "hypothalamus",
-                             "gon" = "gonad"),
-         hypothesis = fct_recode(treatment,
-                                 "non.parental" = "control",
-                                 "non.parental" = "bldg",
-                                 "incubation" = "lay",
-                                 "incubation" = "inc.d3",
-                                 "incubation" = "inc.d9",
-                                 "incubation" = "inc.d17",
-                                 "hatchling.care" = "hatch",
-                                 "hatchling.care" = "n5",
-                                 "hatchling.care" = "n9"))
-
-# wgcna needs numeric identifiers
-colData$ID <- as.numeric(colData$bird)
-
-# create a short sample desctiption, make it the row and columns names of colData and countData
-colData$sample <- paste(colData$treatment, colData$sex, colData$tissue,  colData$ID, sep = ".")
-  
-# set row and count names to be the same
+                             "P" = "pituitary",
[…]

### 2020-04-18 R/functions.R
first pass at new WGCNA

@@ -23,6 +23,23 @@ subsetcolData2 <- function(colData, eachgroup){
 }
 
 
+subsetcolData3 <- function(colData, eachgroup){
+  
+  colData <- colData %>%
+    dplyr::filter(sextissue %in% eachgroup) %>%
+    droplevels()
+  colData <- as.data.frame(colData)
+  row.names(colData) <- colData$sample
+  return(colData)
+}
+
+
+subsetcountData3 <- function(df){
+  savecols <- as.character(df$sample) 
+  savecols <- as.vector(savecols)  
+  df2 <- countData %>% dplyr::select(one_of(savecols)) 
+  return(df2)
+}
 
 
 ############ numDEGs #########

### 2020-04-18 analysis/figure-1.Rmd
now with better image placement

@@
 
-b <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = " ", subtitle = "~ tissue * sex")    
-c <- plottsneelipse(ftsne, ftsne$tissue, allcolors ) + labs(y = " ", subtitle = "~ tissue, females only")
-d <- plottsneelipse(mtsne, mtsne$tissue, allcolors ) + labs(y = " ", subtitle = "~ tissue, males only")
+b <- png::readPNG("../figures/images/fig_fig1b.png")
+b <- ggdraw() +  draw_image(b, scale = 1)
 
-bcd <- plot_grid(b,c,d, nrow = 1, labels = c("B", "C", "D"), label_size = 8 )
-abcd <- plot_grid(a, bcd, nrow = 2, labels = c("A"), label_size = 8, rel_heights = c(1,1))
+c <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = " ", subtitle = "~ tissue * sex")    
+d <- plottsneelipse(ftsne, ftsne$tissue, allcolors ) + labs(y = " ", subtitle = "~ tissue, females only")
+e <- plottsneelipse(mtsne, mtsne$tissue, allcolors ) + labs(y = " ", subtitle = "~ tissue, males only")
+
+bcde <- plot_grid(b,c,d,e, nrow = 1, labels = c("B", "C", "D", "E"), label_size = 8 )
+abcd <- plot_grid(a, bcde, nrow = 2, labels = c("A"), label_size = 8, rel_heights = c(1,1))
 
@@
 egi <- plot_grid(e,g,i, nrow = 3, rel_heights = c(1,1,1.4),
-                 labels = c("E", "G", "I"), label_size = 8)
+                 labels = c("F", "H", "J"), label_size = 8)
 
@@
 fhj <- plot_grid(f,h,j, nrow = 3, rel_heights = c(1,1,1.4),
-                 labels = c("F", "H", "J"), label_size = 8)
+                 labels = c("G", "I", "K"), label_size = 8)
 

### 2020-04-21 analysis/figure-1.Rmd
new fig 1

--- 2020-04-18 analysis/figure-1.Rmd
@@
 # prep for tsne
-
 chartsne <- subsetmaketsne(tissuelevels, charlevels, sexlevels)
@@
 mtsne <-  subsetmaketsne(tissuelevels, charlevels, "male")
-
 ```
@@
 ```{r fig1, fig.width=7.25, fig.height=7.25}
-
-a <- png::readPNG("../figures/images/fig_fig1a.png")
+a <- png::readPNG("../figures/images/fig_fig1b.png")
 a <- ggdraw() +  draw_image(a, scale = 1)
@@
 
-bcd <- plot_grid(b,c,d, nrow = 1, labels = c("B", "C", "D"), label_size = 8 )
-abcd <- plot_grid(a, bcd, nrow = 2, labels = c("A"), label_size = 8, rel_heights = c(1,1))
+abcd <- plot_grid(a,b,c,d, nrow = 1, labels = c("A", "B", "C", "D"), label_size = 8 )
 
-e <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
-  labs( x = NULL)  + 
+e <- png::readPNG("../figures/images/fig_fig1a.png")
+e <- ggdraw() +  draw_image(e, scale = 1)
+
+f <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
+  labs(x = NULL, subtitle = "hypothalamus")  + 
   facet_wrap(~sex, scales = "free")
 
-g <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
-  labs(  x = NULL) + 
+h <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
+  labs(x = NULL, subtitle = "pituitary") + 
   facet_wrap(~sex, scales = "free") +
@@
 
-i <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
+j <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
    facet_wrap(~sex, scales = "free") +
   theme(strip.text = element_blank()) +
-  labs(x = "tSNE1 \n \n \n \n")
-
-egi <- plot_grid(e,g,i, nrow = 3, rel_heights = c(1,1,1.4),
-                 labels = c("E", "G", "I"), label_size = 8)
-
-# hyp
-f <- makebargraph("hypothalamus","DEGs", 0, 1250) + 
-  theme(axis.text.x = element_blank(), 
-        axis.title.x = element_blank())  
-# pit
-h <- makebargraph("pituitary","DEGs", 0, 1250)  +  
-  theme(axis.text.x = element_blank(), 
-        axis.title.x = element_blank(), 
-        strip.text.x = element_blank())  
-# gon
-j <- makebargraph("gonad","DEGs", 0, 1250) +  
-  theme(strip.text.x = element_blank()) +
-  scale_x_discrete(labels = comparisonlabels)
+  labs(x = "tSNE1 \n \n \n \n", subtitle = "gonads")
 
@@
 
-efghij <- plot_grid(egi, fhj, nrow = 1, align = "h", rel_widths = c(1.5,2))
+# hyp
+g <- makebargraph("hypothalamus","DEGs", 0, 4300) + 
+  theme(axis.text.x = element_blank(), 
+        axis.title.x = element_blank())  +
+  labs(subtitle = "hypothalamus")
+# pit
+i <- makebargraph("pituitary","DEGs", 0, 4300)  +  
+  theme(axis.text.x = element_blank(), 
+        axis.title.x = element_blank(), 
+        strip.text.x = element_blank())  +
+  labs(subtitle = "pituitary")
+# gon
+k <- makebargraph("gonad","DEGs", 0, 4300) +  
+  theme(strip.text.x = element_blank()) +
+  scale_x_discrete(labels = comparisonlabels)+
+  labs(subtitle = "gonads")
 
-fig1 <- plot_grid(abcd, efghij, nrow = 2, rel_heights = c(1,1.5))
+gik <- plot_grid(g,i,k, nrow = 3, rel_heights = c(1,1,1.4),
+                 labels = c("G", "I", "K"), label_size = 8)
+
+fhjgik <- plot_grid(fhj, gik, nrow = 1, align = "h", rel_widths = c(1.5,2))
+
+fig1 <- plot_grid(abcd, e, fhjgik, nrow = 3, rel_heights = c(0.7,0.5,1.5),
+                  labels = c(" ","E"), label_size = 8 )
 fig1

### 2020-04-21 docs/manuscript/elife-v1.tex
now with all figs
new fig 1

--- 2020-03-09 docs/manuscript/elife-v1.tex
@@
 % !TeX root = RJwrapper.tex
-\title{Pituitary prolactin drives transcriptional changes in the reproductive
-axis of parenting pigeons}
+\title{Transcriptional changes in the reproductive axis of parenting pigeons}
 \author{by Rayna M Harris \footnote{equal contribtion}, Suzanne H. Austin \footnote{equal contribtion, currently at the
@@
 
-Blah \ref{figure:fig3}, blah, \ref{figure:fig4} blah.
+Blah, blah, \ref{figure:fig4} blah.
 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig3-1}
-  \caption{Another figure}
-  \label{figure:fig3}
-\end{figure}
-
-\newpage
-
 \begin{figure}[ht]
@@
   \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/supplfig3}
+  \includegraphics[width=1.0\textwidth]{../../figures/fig5-1}
+  \caption{Hypothesis-driven data exploration and software development.}
+  \label{figure:fig5}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/supplfig1}
   \caption{R shiny, image 1}

### 2020-04-21 docs/manuscript/elife_v1.Rmd
even better figures
now with all figs
new fig 1

@@
 ---
-title: Pituitary prolactin drives transcriptional changes in the reproductive axis of parenting pigeons
+title: Transcriptional changes in the reproductive axis of parenting pigeons
 author:

### 2020-04-21 R/functions.R

@@ -911,15 +911,13 @@ plottsneelipsev2 <- function(tsnedf, pointcolor, whichcolors){
 
 makebargraph <- function(whichtissue, myylab, lowlim, higherlim){
   p <- allDEG %>%
-    filter(tissue == whichtissue,
-           comparison != "control_bldg") %>%
+    filter(tissue == whichtissue) %>%
     ggplot(aes(x = comparison,  fill = direction)) +
     geom_bar(position = "dodge") +
-    facet_grid(tissue~sex) +
+    facet_wrap(~sex) +
     theme_B3() +
     theme(axis.text.x = element_text(angle = 45, hjust = 1),
           legend.position = "none")  +
-    guides(fill = guide_legend(nrow = 1)) +
     labs(x = NULL, 
          y = myylab,
          subtitle = " ") +
@@ -929,7 +927,7 @@ makebargraph <- function(whichtissue, myylab, lowlim, higherlim){
     scale_color_manual(values = allcolors) +
     geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
               position = position_dodge(width = 1),
-              size = 2, color = "black")  +
+              size = 1.5, color = "black")  +
     ylim(lowlim, higherlim)
   return(p)
 }

### 2020-04-21 analysis/figure-2.Rmd
now with candidate genes in wangle data, finally

@@
 ```{r GO}
-
-# all genes in this parental care dataset
-
-geneids <- read_csv("../metadata/00_geneinfo.csv") %>%
-  select(-X1) %>%
-  rename("NCBI" = "entrezid",
-         "gene" = "Name")
-
-# From Ch 17 Evolution of parental care
-curleychampagnegenes <- c("AVPR1A", "ADRA2A",    
-                         "COMT", "CRH", "CRHBP", "CRHR1", "CRHR2", 
-                          "DRD1", "DRD4","ESR1", "ESR2", #ERa ERbeta
-                          "FOS", "HTR2C", "MEST", "NR3C1", #GR
-                          "OPRM1", "PGR", "PRL", "PRLR",  "SLC6A4") #5HTT
-curleychampagnegenes <- as.data.frame(curleychampagnegenes) %>%
-  mutate(GO = "parentalcare", gene = curleychampagnegenes)  %>%
-  select(GO, gene)
-curleychampagnegenes
-
-# Candidate GO terms 
-GO_path <- "../metadata/goterms/"   # path to the data
-GO_files <- dir(GO_path, pattern = "*.txt") # get file names
-GO_pathfiles <- paste0(GO_path, GO_files)
-
-GOgenesLong <- GO_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_table(.x, col_types = cols(), col_names = FALSE), .id = "file_name") %>% 
-  mutate(GO = sapply(strsplit(as.character(file_name),'../metadata/goterms/'), "[", 2)) %>% 
-  mutate(GO = sapply(strsplit(as.character(GO),'.txt'), "[", 1)) %>% 
-  mutate(gene = sapply(strsplit(as.character(X1), "[\\\\]|[^[:print:]]" ), "[", 2)) %>% 
-  select(GO, gene)  %>%
-  filter(gene != "Symbol") %>%
-  distinct(GO,gene)  %>%
-  mutate(gene = toupper(gene)) %>%
-  rbind(., curleychampagnegenes) %>%
-  arrange(gene) %>%
-  left_join(., geneids, by = "gene") %>%
-  drop_na() 
-GOgenesLong
-
-GOgenesWide <- GOgenesLong %>% 
-  pivot_wider(
-    names_from = GO,
-    values_from = gene) %>%
-  mutate(numGOs = 4 - rowSums(is.na(.))) %>%
-  arrange(desc(numGOs)) %>%
-  select(-numGOs) %>%
-  left_join(geneids, by = c("geneid","NCBI"))
-GOgenesWide
-
-# genes I can't find in dataset 
-# NOS1 or OXTR or FOX1B or HTR5A
+source("../R/wrangledata.R")
 ```
@@
 ```{r DEGs}
-
-candidategenes <- GOgenesLong %>% distinct(gene) %>% pull(gene)
-
 # summary DEG results from DESeq2

### 2020-04-21 analysis/figure-6.Rmd
now with candidate genes in wangle data, finally

--- 2020-03-28 analysis/figure-6.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
+```
 
-candidategenesslim <- c("OXT", "AVP", "GNRH1", "GNRHR", 
-                    "AR",  "CYP19A1", 
-                     "AVPR1A", "AVPR1B", "AVPR2","VIP",
-                  "DRD1", "DRD2", 
-                  "PRL", "PRLR",  
-                    "ESR1","ESR2", "LBH",
+## Candidate Genes (from literature and relevant GO terms)
 
-                    "FOS", "JUN", "EGR1", "BDNF"
-                      )
-
-candidategenes <- candidategenesslim
-
+```{r GO}
+source("../R/wrangledata.R")
 ```
@@
 
-```{r }
-geneids <- read_csv("../metadata/00_geneinfo.csv")
-
+```{r vsd}
 vsd_path <- "../results/DEseq2/manip/"   # path to the data
@@
 
-hypvsd <- getcandidatevsd(candidategenesslim, "hypothalamus", sexlevels)
-pitvsd <- getcandidatevsd(candidategenesslim, "pituitary", sexlevels)
-gonvsd <- getcandidatevsd(candidategenesslim, "gonad", sexlevels)
-head(hypvsd)
-head(pitvsd)
-head(gonvsd)
+hypvsd <- getcandidatevsd(candidategenes, "hypothalamus", sexlevels)
+pitvsd <- getcandidatevsd(candidategenes, "pituitary", sexlevels)
+gonvsd <- getcandidatevsd(candidategenes, "gonad", sexlevels)
 
@@
 head(candidatevsd)
+
 levels(candidatevsd$treatment)
@@
 
-table1 <- allDEG %>%
-  mutate(updown = ifelse(lfc > 0, "+", "-"))  %>%
-  mutate(geneupdown = paste(gene, updown, sep = "")) %>%
+candidateDEGS <- allDEG %>%
   filter(gene %in% candidategenes) %>%
-  arrange(geneupdown) %>%
-  group_by(sex, tissue, comparison) %>%
-  summarize(genes = str_c(geneupdown, collapse = " ")) %>%
-  pivot_wider(names_from = comparison, values_from = genes ) %>%
-  select(sex, tissue, contains("m."))  %>%
-  arrange( tissue, sex)
+  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
+         sex = recode(sex, "female" = "F", "male" = "M" ),
+         tissue = recode(tissue, 
+                         "hypothalamus" = "H",
+                         "pituitary" = "P", "gonad" = "G")) %>%
+  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
+  select(gene, res, comparison)  %>%
+  group_by(gene,  comparison) %>%
+  summarize(res = str_c(res, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = res) %>%
+  select(gene, contains("m.")) 
+candidateDEGS
 
-(table1$inc.d3_m.inc.d3)
-(table1$inc.d9_m.inc.d9)
-(table1$inc.d17_m.inc.d17)
-(table1$hatch_m.n2)
-```
-
-## Figs
-
-```{r fig6, fig.height=7, fig.width=7}
-makeboxplotsmanip <- function(df, whichgene, mysubtitle, whichsex, whichtimepoint){
-  p <- df %>%
-    filter(treatment %in% alllevels,
-           gene %in% whichgene,
-           sex %in% whichsex) %>%
-    filter(treatment %in% whichtimepoint) %>%
-    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
-    geom_boxplot(outlier.shape = NA) + 
-    geom_jitter(size = 0.5, width = 0.1) +
-    facet_wrap(~gene, scales = "free_y", nrow = 1) +
-    theme_B3() +
-    scale_fill_manual(values = allcolors) +
-    scale_color_manual(values = allcolors) +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1),
-          legend.position = "none",
-          plot.caption = element_text(face = "italic")) +
-    labs(y = "gene expression" , x = "Parental Stages and Manipulations", subtitle = mysubtitle) +
-    theme(strip.text = element_text(face = "italic"))
-  return(p)
-}
-
-## hyp 
-a2 <- makeboxplotsmanip(hypvsd, c("DRD1","EGR1","PRLR"), "Female hypothalamus", "female",
-                  c("inc.d3", "m.inc.d3", "inc.d17", "m.inc.d17",  "hatch", "m.n2"))  +
-  theme(axis.title.x = element_blank())
-b2 <- makeboxplotsmanip(hypvsd, c("BDNF","ESR2", "FOS"), " ", "female",
-                  c("inc.d3", "m.inc.d3", "inc.d17", "m.inc.d17"))  +
-  theme(axis.title.x = element_blank())
-
-c <- makeboxplotsmanip(hypvsd, c("AVP","OXT"), " ", "female", 
-                       c("inc.d17", "m.inc.d17"))  +
-  theme(axis.title = element_blank())
-d <- makeboxplotsmanip(hypvsd, c("AVPR1A","CYP19A1","GNRH1"), " ", "female",
-                   c("hatch", "m.n2"))  +
-  theme(axis.title = element_blank())
-e <- makeboxplotsmanip(hypvsd, c("JUN"), " ", "female", 
-                 c("hatch", "m.n2"))  +
-  theme(axis.title = element_blank()) + labs(subtitle = " ")
-
-f1 <- makeboxplotsmanip(hypvsd, c("DRD1", "EGR1", "PRLR"), "Male hypothalamus", "male",
-                  c( "inc.d17", "m.inc.d17", "hatch", "m.n2"))  +
-  theme(axis.title.x = element_blank())
-
-f2 <- makeboxplotsmanip(hypvsd, c( "PRL"), " ", "male",
-                  c( "inc.d17", "m.inc.d17", "hatch", "m.n2"))  +
-  theme(axis.title = element_blank())
-
-h <- makeboxplotsmanip(hypvsd, c("AR", "AVP", "BDNF", "ESR1", "LBH"), " ", "male",
-                  c( "inc.d17", "m.inc.d17"))  + labs(caption = " ")
-
-i <- makeboxplotsmanip(hypvsd, c("GNRH1"), " ", "male",
-                  c("hatch", "m.n2"))  +
-  theme(axis.title = element_blank()) +
-  labs(x = " ",
-       caption = "Candidate genes that were not differentially expressed following removal: AVPR1B, AVPR2, ESR1, GNRHR, VIP")
-
-ab2 <- plot_grid(a2,c, rel_widths = c(18,5), labels = "auto", label_size = 8)
-cde <- plot_grid(b2,d,e, rel_widths = c(12,6,2), nrow =  1, align = "h",  
-                 labels = c("c", "d", " "), label_size = 8)
-f12 <- plot_grid(f1,f2, rel_widths = c(3,1), labels = c("e",  " "), label_size = 8)
-hi <- plot_grid(h,i, ncol = 2, rel_widths = c(10,2),labels = c("f",  "g"), label_size = 8, align = "h")
-
-plot_grid(ab2, cde, f12,hi, nrow = 4, rel_heights = c(1,1,1,1.25))
-
+table2 <- left_join(candidateDEGS, GOgenesWide) %>%
+  select(gene, m.inc.d17_prolong:m.inc.d9_m.inc.d17, parentalcare, parentalbehavior, NCBI) %>%
+  mutate(parentalcare = if_else(is.na(parentalcare), " ", "X"),
+         parentalbehavior = if_else(is.na(parentalbehavior), " ", "X")) %>%
+  rename("Literature" = "parentalcare", "GO" =  "parentalbehavior")
+table2$numDEGs <- rowSums(is.na(table2)) # count NAs to know how many are NS
+table2 <- table2 %>% 
+  mutate(sig = ifelse(numDEGs == 6, "NS", "DEG")) %>% 
+  arrange(sig, gene)  %>%  select(-sig, -numDEGs)
+table2[is.na(table2)] <- " " # replace NA with blank space so it's pretty
+table2
 ```
@@
 ```{r writefiles}
-write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
+write.csv(table2, "../results/table2.csv")
+
+# for victoria
+prlmanip <- candidatevsd %>%
+  filter(gene == "PRL", 
+         treatment %in% c("extend", "m.inc.d8"))
+write.csv(prlmanip, "../results/prlmanip.csv")
 ```

### 2020-04-21 analysis/figure-6.Rmd
working on manip figure

@@
 
+head(allvsd)
+
 getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
@@
                 treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
-    #dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1))# %>%
-   # dplyr::recode(treatment, "m.hatch" = "m.n2", .default = levels(treatment))
+    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
+     dplyr::mutate(treatment = recode(treatment, "m.hatch" = "m.n2",
+                                                  "extend.hatch" = "m.n2",
+                                      "inc.prolong" = "prolong")) %>%
 
-     dplyr::mutate(treatment = recode(treatment, "m.hatch" = "m.n2"))%>%
+    dplyr::mutate(day = fct_collapse(treatment,
+                                     "day 3" = c("inc.d3", "m.inc.d3" ),
+                                     "day 9" = c("inc.d9", "m.inc.d9", "m.inc.d8"  ),
+                                     "day 17" = c("inc.d17", "m.inc.d17", "prolong"  ),
+                                     "hatch" = c("hatch", "m.n2", "extend"  )
+                                     )) %>%
 
-    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
+    dplyr::mutate(manip = fct_collapse(treatment,
+                                       "reference" = charlevels,
+                                       "removal" = c("m.inc.d3", "m.inc.d9", "m.inc.d17","m.n2"),
+                                       "replacement" = c("m.inc.d8", "prolong", "extend"))) %>%
+
+    dplyr::select(sex, tissue, treatment, day,  manip, gene, samples, counts) %>%
     filter(tissue == whichtissue, sex %in% whichsex)  %>%
@@
 gonvsd <- getcandidatevsd(candidategenes, "gonad", sexlevels)
-
 candidatevsd <- rbind(hypvsd, pitvsd)
 candidatevsd <- rbind(candidatevsd, gonvsd)
-
 head(candidatevsd)
-
-levels(candidatevsd$treatment)
 ```
@@
 allDEG$direction <- factor(allDEG$direction, levels = alllevels)
-
 head(allDEG)
@@
 
+```{r fig6}
+candidatevsd %>%
+  filter(gene == "PRL") %>%
+  ggplot(aes(y = counts, x = day, color = manip)) +
+  geom_boxplot() +
+  facet_wrap(tissue~sex, scales = "free") +
+  theme(legend.position = "bottom") 
+```
+
 ```{r writefiles}

### 2020-04-22 analysis/figure-6.Rmd
first attempt new fig 6" "

@@
 head(allvsd)
+```
 
+```{r candidate vsd}
+# for plotting gene expression over time
 getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
@@
     dplyr::mutate(day = fct_collapse(treatment,
-                                     "day 3" = c("inc.d3", "m.inc.d3" ),
-                                     "day 9" = c("inc.d9", "m.inc.d9", "m.inc.d8"  ),
-                                     "day 17" = c("inc.d17", "m.inc.d17", "prolong"  ),
-                                     "hatch" = c("hatch", "m.n2", "extend"  )
+                                     "1" = c("inc.d3", "m.inc.d3" ),
+                                     "2" = c("inc.d9", "m.inc.d9", "m.inc.d8"  ),
+                                     "2" = c("inc.d17", "m.inc.d17", "prolong"  ),
+                                     "4" = c("hatch", "m.n2", "extend"  )
                                      )) %>%
-
+    dplyr::mutate(day = as.numeric(day)) %>%
     dplyr::mutate(manip = fct_collapse(treatment,
@@
                                        "replacement" = c("m.inc.d8", "prolong", "extend"))) %>%
-
+    dplyr::mutate(manip = factor(manip, levels = c("reference", "removal", "replacement"))) %>%
     dplyr::select(sex, tissue, treatment, day,  manip, gene, samples, counts) %>%
@@
 table2
+
+table2 %>% select(gene,inc.d3_m.inc.d3, inc.d9_m.inc.d9, inc.d17_m.inc.d17, hatch_m.n2)
 ```
 
-```{r fig6}
-candidatevsd %>%
+```{r fig6, fig.height=7.25, fig.width=10}
+
+plotcandidatemanip <- function(whichtissue, whichsex, whichgenes){
+
+  p <- candidatevsd %>%
+  filter(tissue == whichtissue, sex == whichsex) %>%
+  filter(gene %in% whichgenes) %>%
+  filter(treatment %in% c(levelsremoval, controlsremoval)) %>%
+  mutate(treatment = factor(treatment, levels = alllevels3)) %>%
+  ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
+  geom_boxplot() +
+  facet_wrap(~gene, scales = "free_y", nrow = 2) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        strip.text = element_text(face = "italic")) +
+  scale_color_manual(values = allcolors) +
+  scale_fill_manual(values = allcolors)
+
+  return(p)
+}
+
+p1 <- plotcandidatemanip("hypothalamus", "female", c("AVP", "BRINP1", "COMT", "CREBRF", "CRH", "CRHBP", "CRHR1", "CRHR2", "OPRK1")) +
+  labs(subtitle = "hypothalamus") +
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) 
+p2 <- plotcandidatemanip("pituitary", "female", c("CREBRF", "CRHR1", "DRD1", "DRD4", "GNAQ")) +
+  labs(subtitle = "pituitary")
+p3 <- plotcandidatemanip("gonad", "female", c("BRINP1", "OPRK1"))  +
+  labs(subtitle = "gonads")
+
+p23 <- plot_grid(p2,p3, nrow = 1, rel_widths = c(3,1))
+
+plot_grid(p1,p23, nrow = 2, rel_heights = c(1,1.2))
+
+## hypothesis
+```
+
+```{r, eval = F}
+
+## not working
+meanse <- candidatevsd %>%
+  dplyr::group_by(sex, tissue, treatment, day, manip, gene) %>%
+  dplyr::summarise(m = mean(counts), 
+                   se = sd(counts)/sqrt(length(counts))) %>%
+  dplyr::mutate(m = round(m,0))
+head(meanse)
+
+meanse %>%
   filter(gene == "PRL") %>%
-  ggplot(aes(y = counts, x = day, color = manip)) +
-  geom_boxplot() +
-  facet_wrap(tissue~sex, scales = "free") +
-  theme(legend.position = "bottom") 
+  ggplot(aes(x = day, y = m, color = manip)) +
+  geom_errorbar(aes(ymin=m-se, ymax=m+se), width=.1) +
+    geom_point(size = 1) +
+  facet_wrap(sex~tissue, scales = "free")
 ```

### 2020-04-23 R/themes.R
bigger font

@@ -1,7 +1,7 @@
 # custom themes
 
 theme_B3 <- function () { 
-  theme_classic(base_size = 6,
+  theme_classic(base_size = 7,
                 base_family = 'Helvetica') +
     theme(
       panel.grid.major  = element_blank(),  # remove major gridlines

### 2020-04-23 analysis/figure-1.Rmd
bigger font

@@
 
-b <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = " ", subtitle = "~ tissue * sex")    
+b <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = "tSNE 2 ", subtitle = "~ tissue * sex")    
 c <- plottsneelipse(ftsne, ftsne$tissue, allcolors ) + labs(y = " ", subtitle = "~ tissue, females only")

### 2020-04-23 analysis/figure-2.Rmd
new fig 2

@@
 a1 <- candidateboxplot("hypothalamus", hypDEGs, "female") + labs(x = NULL, title = "hypothalamus")
-a2 <- candidateboxplot("hypothalamus", hypDEGs, "male") + labs(x = NULL) +
-  theme(strip.text = element_blank())
 
 b1 <- candidateboxplot("pituitary", pitDEGs, "female") + labs(x = NULL, title = "pituitary")
-b2 <- candidateboxplot("pituitary", pitDEGs, "male") + labs(x = NULL) +
-  theme(strip.text = element_blank())
 
-c1 <- candidateboxplot("gonad", gonadDEGs, "female") + labs(x = NULL, title = "gonad")
-c2 <- candidateboxplot("gonad", gonadDEGs, "male") + 
-  theme(strip.text = element_blank(),
-        axis.text.x = element_text(angle = 45, hjust = 1))
+b2 <- candidateboxplot("pituitary", "PRL", "male") + labs(x = NULL, y = NULL,  title = "pituitary")
 
-plot_grid(a1,a2, b1,b2, c1,c2, nrow = 6, labels = c("A", " ","B", " ", "C", " "), label_size = 8,
-          rel_heights = c(1, 0.75, 1, 0.75, 1, 1))
+b12 <- plot_grid(b1,b2,nrow = 1, labels = c("B", "C"),
+          label_size = 8, rel_widths = c(4,1))
+
+c1 <- candidateboxplot("gonad", gonadDEGs, "female") + 
+  labs(x = NULL, title = "gonad")
+
+plot_grid(a1, b12, c1, nrow = 3, labels = c("A", " ","D"), 
+          label_size = 8)
 ```

### 2020-04-25 analysis/figure-2.Rmd
now with fig 4

@@
 source("../R/themes.R")
+source("../R/wrangledata.R")
 
 knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
-```
-
-## Candidate Genes (from literature and relevant GO terms)
-
-```{r GO}
-source("../R/wrangledata.R")
 ```
@@
 
-## variance stabilized gene expression  (vsd)
-
-```{r vsds}
-
-vsd_path <- "../results/DEseq2/"   # path to the data
-vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
-vsd_pathfiles <- paste0(vsd_path, vsd_files)
-vsd_files
-
-allvsd <- vsd_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name")  %>% 
-  dplyr::rename("gene" = "X1") %>% 
-  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
-               names_to = "samples", values_to = "counts")
-
-getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
-  candidates  <- allvsd %>%
-    filter(gene %in% whichgenes) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
-    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
-                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
-                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
-    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
-    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
-    filter(tissue == whichtissue, sex %in% whichsex)  %>%
-    drop_na()
-  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
-  return(candidates)
-}
-
-candidategenes <- GOgenesLong %>% distinct(gene) %>% pull(gene)
-
-hypvsd <- getcandidatevsd(candidategenes, "hypothalamus", sexlevels)
-pitvsd <- getcandidatevsd(candidategenes, "pituitary", sexlevels)
-gonvsd <- getcandidatevsd(candidategenes, "gonad", sexlevels)
-candidatevsd <- rbind(hypvsd, pitvsd)
-candidatevsd <- rbind(candidatevsd, gonvsd)
-head(candidatevsd)
-```
+## Candidate variance stabilized gene expression (vsds)
 
 ```{r fig2, fig.width = 7.25, fig.height = 7.25}
+
+# `candidatevsd` loaded with source(../R/wrangleata.R)
 

### 2020-04-25 analysis/figure-3.Rmd
fig tweeks
now with fig 4

--- 2020-04-13 analysis/figure-3.Rmd
@@
 ---
-title: "Fig3"
+title: "Fig3: Correlations in gene expression"
 author: "Rayna M Harris"
@@
 
-# all things box plots
+# Correlations in gene expression
 
-## candidate genes
-
-```{r}
+```{r setup}
 library(tidyverse)
-library(ggtext)
 library(cowplot)
+library(corrr)
 library(ggpubr)
@@
 source("../R/themes.R")
+source("../R/functions.R")
+source("../R/wrangledata.R")
 
 knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
-
-candidategenes <- c("OXT", "AVP", "GNRH1", "GNRHR", "CGNRH-R",
-                    "AR", "POMC", "AGRP", 
-                       "CRH", "AVPR1A", "AVPR1B", "AVPR2","VIP",
-                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9", 
-                    "ESR1","ESR2", "LBH", "CDK1", "BRCA1",
-                    "PTEN", "CREBBP", "FOS", "JUN", "EGR1",
-                     "BDNF", "GRM2","GRIA1",
-                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16", 
-                    "DIO3", "DIO2", "DIO1",
-
-                     "VIPR1", "VIPR2", "NPY", "LH", "FSH", "FSHB",
-                    "JAK2", "HCRT", "TRH", "TSHB",
-                    "MC3R", "MC4R", "MC5R",  "NR3C1", "NR3C2",
-                    "STAT5B","NPVF")
-
-candidategenesslim <- c("OXT", "AVP", "GNRH1", "GNRHR", 
-                    "AR",  "CYP19A1", 
-                    "AVPR1A", "AVPR1B", "AVPR2","VIP",
-                    "DRD1", "DRD2", 
-                    "PRL", "PRLR",  
-                    "ESR1","ESR2", "LBH",  
-                    "FOS", "JUN", "EGR1", "BDNF")
-
-# candidategenesslim: AR, AVP, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, EGR1, ESR1, ESR2, FOS, GNRH1, GNRHR, JUN, LBH, OXT, PRL, PRLR, VIP
-# DEG candidategenesslim: AR, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, ESR1, FOS, GNRHR, LBH, PRL, PRLR, 
-# non-DEG candidategenesslim: AVP, AVPR1B, ESR2,  GNRH1, JUN, OXT, VIP
-
-hormonepapergenes <- c("GNRH1", "CGNRH-R", "GNRHR",  
-                       "FSHB", "FSHR", "LHCGR", # no LSH
-                       "VIP", "VIPR1",
-                       "ESR1", "AR", "CYP19A1", "HSD17B1", # receptors synthesis
-                       "PGR" , "HSD3B2", 
-                       "PRL", "PRLR")
 ```
 
-## variance stabilized gene expression  (vsd)
+## data wrangle
 
-```{r }
-geneids <- read_csv("../metadata/00_geneinfo.csv") %>% select(-X1)
+```{r data}
 
-vsd_path <- "../results/DEseq2/"   # path to the data
-vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
-vsd_pathfiles <- paste0(vsd_path, vsd_files)
-vsd_files
+# load `candidatevsd` with `source("../R/wrangledata.R")`
+head(candidatevsd)
 
-allvsd <- vsd_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name")  %>% 
-  dplyr::rename("gene" = "X1") %>% 
-  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
-               names_to = "samples", values_to = "counts")
+curleychampagnegenes <- curleychampagnegenes %>% pull(gene)
 
-getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
-  candidates  <- allvsd %>%
-    filter(gene %in% whichgenes) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
-    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
-                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
-                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
-    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
-    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
-    filter(tissue == whichtissue, sex %in% whichsex)  %>%
-    drop_na()
-  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
-  return(candidates)
+# widen for correlations
+
+candidatevsdwide <- candidatevsd  %>%
+    pivot_wider(names_from = gene, values_from = counts) 
+head(candidatevsdwide)
+
+makecorrdf <- function(whichsex, whichtissue, whichgenes){
+
+  corrrdf <- candidatevsd %>%
+    filter(sex == whichsex, tissue == whichtissue,
+           gene %in% whichgenes) %>%
+    pivot_wider(names_from = gene, values_from = counts) %>%
+    select(-sex,-tissue, -treatment, -samples) %>%
+    correlate() %>%
+    rearrange()
+  print(head(corrrdf))
+  return(corrrdf)
 }
 
-hypvsd <- getcandidatevsd(hormonepapergenes, "hypothalamus", sexlevels)
-pitvsd <- getcandidatevsd(hormonepapergenes, "pituitary", sexlevels)
-gonvsd <- getcandidatevsd(hormonepapergenes, "gonad", sexlevels)
-head(hypvsd)
-tail(gonvsd)
+c1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
+c2 <- makecorrdf("female", "pituitary", curleychampagnegenes)  
+c3 <- makecorrdf("female", "gonad", curleychampagnegenes)
 
-candidatevsd <- rbind(hypvsd, pitvsd)
-candidatevsd <- rbind(candidatevsd, gonvsd)
+c4 <- makecorrdf("male", "hypothalamus", curleychampagnegenes)  
+c5 <- makecorrdf("male", "pituitary", curleychampagnegenes)  
+c6 <- makecorrdf("male", "gonad", curleychampagnegenes)
 
-unique(candidatevsd$gene)
+subsetcandidatevsdwide <- function(whichsex, whichtissue){
+  df <- candidatevsdwide %>%
+    filter(sex == whichsex, tissue == whichtissue,
+           treatment != "control") 
+  return(df)
+}
+
+FH <- subsetcandidatevsdwide("female", "hypothalamus")
+FP <- subsetcandidatevsdwide("female", "pituitary")
+FG <- subsetcandidatevsdwide("female", "gonad")
+MH <- subsetcandidatevsdwide("male", "hypothalamus")
+MP <- subsetcandidatevsdwide("male", "pituitary")
+MG <- subsetcandidatevsdwide("male", "gonad") 
 ```
 
-## Figs
+### figure
 
-```{r fig3, fig.width = 7}
-# summary DEG results from DESeq2
-candidateDEGS <- read_csv("../../musicalgenes/data/allDEG.csv") %>%
-  select(-X1) %>%
-  filter(gene %in% hormonepapergenes)
-head(candidateDEGS)
+```{r fig3, fig.width = 7.25, fig.height = 7.25}
 
-# get genes to plot
-candidateDEGS %>% filter(tissue == "hypothalamus", sex == "female", comparison != "control_bldg") 
-candidateDEGS %>% filter(tissue == "hypothalamus", sex == "male", comparison != "control_bldg") 
-candidateDEGS %>% filter(tissue == "pituitary", sex == "male", comparison != "control_bldg") 
-candidateDEGS %>% filter(tissue == "pituitary", sex == "female", comparison != "control_bldg") 
-candidateDEGS %>% filter(tissue == "gonad", sex == "female", comparison != "control_bldg")
+plotcorrplot <- function(df, subtitle){  
+  p <- df %>%
+    rplot2() +
+    theme_B3() +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1),
+          axis.text = element_text(face = "italic"))
 
-# summarize all differences
-candidateDEGtable <- candidateDEGS %>%
-  mutate(posneg = ifelse(lfc >= 0, "+", "-")) %>%
-  mutate(geneposneg = paste(gene, posneg, sep = "")) %>%
-  select(sex, tissue, comparison, geneposneg) %>%
-  group_by(sex, tissue, comparison) %>%
-   arrange(geneposneg) %>%
-  summarize(genes = str_c(geneposneg, collapse = " ")) %>% 
-  pivot_wider(names_from = comparison, values_from = genes)  %>% 
-  ungroup(tissue) %>%
-  mutate(tissue = factor(tissue, levels =  tissuelevel))  %>% 
-  arrange(sex, tissue)  %>% 
-  select(sex, tissue, bldg_lay, lay_inc.d3, inc.d9_inc.d17, hatch_n5, control_bldg)
-candidateDEGtable
-
-newboxplot <- function(df, mygenes, whichsex, whichstage){
-
-  p <- df %>%
-    filter(gene %in% mygenes, 
-           sex %in% whichsex,
-           treatment %in% whichstage) %>%
-    droplevels() %>%
-    mutate(tissue = factor(tissue, levels = tissuelevel)) %>%
-    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
-    geom_boxplot() +
-    geom_jitter(size = 0.5, width = 0.1) +
-    facet_grid(tissue~gene, scales = "free_y") +
-    theme_B3() +
-    scale_color_manual(values = allcolors) +
-    scale_fill_manual(values = allcolors) +
-    theme(legend.position = "none",
-          axis.text.x = element_text(angle = 45, hjust = 1),
-          strip.text.x = element_text(face = "italic"),
-          axis.title.x = element_blank())  +
-    labs(x = "Parental stage",
-         y = "variance stabilized expression")
   return(p)
@@
 
-a1 <- newboxplot(hypvsd, c("PGR",  "CYP19A1", "LHCGR" ), "female", c( "hatch", "n5")) + 
-  theme(strip.text.y = element_blank()) + labs(subtitle = "Females")
-a2 <- newboxplot(hypvsd, c("AR"), "male", c( "hatch", "n5")) + 
-  theme(axis.title.y = element_blank()) + labs(subtitle = "Males")
+p1 <- plotcorrplot(c1, "FH") + theme(legend.position = "none", axis.text.x = element_blank()) + labs(subtitle = "Hypothalamus")
+p2 <- plotcorrplot(c2, "FP") + theme(legend.position = "none", axis.text.x = element_blank()) + 
+  labs(subtitle = "Pituitary") 
+p3 <- plotcorrplot(c3, "FG") + theme(legend.position = "none") + labs(subtitle = "Gonad")
 
-b1 <- newboxplot(gonvsd, c("HSD3B2"), "female", c( "bldg", "lay"))  +
-  theme(strip.text.y = element_blank(),
-        axis.title.y = element_blank()) + labs(subtitle = "Females")
-b2 <- newboxplot(gonvsd, c("VIPR1",  "PRLR" ), "female", c( "lay", "inc.d3"))  + 
-  theme(axis.title.y = element_blank()) + labs(subtitle = " ")
+p4 <- plotcorrplot(c4, "MH") + theme(legend.position = "none", axis.text.y = element_blank(), axis.text.x = element_blank()) + labs(subtitle = " ")
+p5 <- plotcorrplot(c5, "MP") + theme(legend.position = "none", axis.text.y = element_blank(), axis.text.x = element_blank())  +
+  labs(subtitle = " ") 
+p6 <- plotcorrplot(c6, " ") + theme(legend.position = "none", axis.text.y = element_blank()) + labs(subtitle = " ")
 
-ab <- plot_grid(a1,a2,b1,b2, rel_widths = c(6,2.5,2,4), labels = c("a", " ", "b", " "), label_size = 8, nrow = 1)
+a <- plot_grid(p1 + labs(title = "Female"),
+               p4 + labs(title = "Male"),
+               p2, p5, p3, p6, nrow = 3, rel_widths = c(1.2,1),
+               rel_heights = c(1,1,1.2))
 
-c1 <- newboxplot(pitvsd, c( "ESR1", "GNRHR"), "female", c("bldg", "lay",  "inc.d3")) 
-c2 <- newboxplot(pitvsd, c( "PRL"), "female", c("inc.d9", "inc.d17", "hatch", "n5")) 
-c3 <- newboxplot(pitvsd, c("VIPR1",  "PRL" ), "male", c( "inc.d9", "inc.d17"))
-
-c <- plot_grid(c1 + theme(strip.text.y = element_blank()) + labs(subtitle = "Females"),
-          c2 + theme(strip.text.y = element_blank(),
-                    axis.title.y = element_blank()) + labs(subtitle = " ") ,
-          c3 + theme(axis.title.y = element_blank()) + labs(subtitle = "Males"), 
-          nrow = 1, rel_widths = c(6,4,4),
-          labels = c("c"), label_size = 8)
-
-plot_grid(ab,c, nrow = 2)
-
-# correlations
-library(corrr)
-
-plotcorrs <- function(whichtissue){
-
-  x <- candidatevsd %>%
-    filter(tissue == whichtissue)  %>%
-    pivot_wider(names_from = gene, values_from = counts) %>%
-    select(-sex, - tissue, - treatment, -samples) %>%
-    correlate()
-  print(fashion(x))
-
-  p <- rplot(x, colors = c("#0571b0", "#92c5de", "#f7f7f7","#f7f7f7","#f7f7f7", "#f4a582", "#ca0020")) +
+scattercorrelations <- function(df, gene1, gene2, mylinecolor, myxlab, myylab){
+  p <- ggplot(df, aes(x = gene1, y = gene2)) +
+    labs(subtitle = " ", x = myxlab, y = myylab) + 
+    scale_color_manual(values = allcolors) +
+    geom_smooth(method = "lm",  color = mylinecolor) +
+    geom_point(aes(color = treatment))  +
     theme_B3() +  
-    theme(legend.position = "bottom",
-          axis.text.x = element_text(angle = 45, hjust = 1),
-          axis.text = element_text(face = "italic")) +
-    labs(subtitle = whichtissue)
+    theme(axis.title = element_text(face = "italic"), 
+            legend.position = "none") +
+    stat_cor( size = 2)
   return(p)
@@
 
-a <- plotcorrs("hypothalamus") 
-b <- plotcorrs("pituitary") + theme(axis.text.y = element_blank())
-c <- plotcorrs("gonad") + theme(axis.text.y = element_blank())
+# "female" = "#969696", "male" = "#525252"
 
-plot_grid(a,b,c, nrow = 1, rel_widths = c(1.2,1,1))
+p7 <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "Hypothalamus")
+p8 <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = "Pituitary")
+p9 <- scattercorrelations(FG, FG$ESR2, FG$CRHBP, "#969696", "ESR2", "CRHBP")   + labs(subtitle = "Gonad")
+
+p10 <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C") + theme(axis.title.y = element_blank())
+p11 <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1") + theme(axis.title.y = element_blank())
+p12 <- scattercorrelations(MG, MG$ESR2, MG$CRHBP, "#525252", "ESR2", "CRHBP")  + theme(axis.title.y = element_blank())
+
+b <- plot_grid(p7 + labs(title = "Female"), 
+               p10 + labs(title = "Male"), 
+               p8, p11, p9, p12, nrow = 3, rel_heights = c(1,1,1.2))
+
+ab <- plot_grid(a,b, rel_widths = c(1.5,1), labels = c("A", "B"), label_size = 8)
+
+p13 <- plotcorrplot(c1, "FH") + theme(legend.position = "bottom")
+p14 <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + 
+  theme(legend.position = "bottom",
+        legend.title = element_blank())
+
+legend1 <- p13 %>% get_legend()
+legend2 <- p14 %>% get_legend()
+
+legends <- plot_grid(legend1, legend2, rel_widths = c(1.5,1))
+
+plot_grid(ab, legends, nrow = 2, rel_heights = c(1,0.1))
 
 ```
-
-```{r writefiles}
-#write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
-#write.csv(candidatevsd, "../results/candidatecounts.csv")
-write.csv(candidateDEGtable, "../results/candidateDEGtable.csv")
-
-```

### 2020-04-26 analysis/figure-4.Rmd
new fig 4

--- 2020-03-07 analysis/figure-4.Rmd
@@
 
-```{r setup, include=FALSE}
+# Figure 4: All things prolactin
+
+```{r setup}
 library(tidyverse)
 library(cowplot)
-library(Rtsne)
+library(ggimage)
+library(apaTables)
+library(factoextra)
 
@@
 
-knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
+knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
 ```
 
-# Figure 1. Experimental design, tSNE analysis, PCA, and prolactin
+## PCA data
 
-```{r tSNE-df}
+```{r PCA}
+# pca
+pca1f <- subsetmakepca("pituitary", charlevelsnocontrol, "female")	
+pca1m <- subsetmakepca("pituitary", charlevelsnocontrol, "male")
 
-# prep for tsne
-
-hyptsne <- subsetmaketsne("hypothalamus", allmaniplevels, sexlevels)
-pittsne <- subsetmaketsne("pituitary", allmaniplevels, sexlevels)
-gontsne <- subsetmaketsne("gonads", allmaniplevels, sexlevels)
-
-addgroupings <- function(df){
-
-  df <- df %>% mutate(hiloPRL = fct_collapse(treatment, 
-                                  lo = c("m.inc.d8", "m.inc.d3", "m.inc.d9", "inc.d3", "inc.d9", "lay", "bldg"),
-                                  hi = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" ,  "m.n2", "extend", "n5")),
-                extint = fct_collapse(treatment, 
-                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
-                                  chicks = c( "m.inc.d8", "hatch", "extend", "n5"),
-                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")))
-  df$extint <- factor(df$extint, levels = levelsextint)
-  return(df)
-}
-
-hyptsne <- addgroupings(hyptsne)
-pittsne <- addgroupings(pittsne)
-gontsne <- addgroupings(gontsne)
+pca2f <- makefvizdf("pituitary", charlevelsnocontrol, "female")	
+pca2m <- makefvizdf("pituitary", charlevelsnocontrol, "male")
 
@@
 
-## make figure
+## PRL hormone
 
-```{r fig4, fig.width=7, fig.height=7}
+```{r hormones}
 
-expdesign2 <- png::readPNG("../figures/images/fig_fig4a.png")
-expdesign2 <- ggdraw() +  draw_image(expdesign2, scale = 1)
+# hormones
+prolactin <- read_csv("../results/07_hormones.csv") %>%
+    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
+    droplevels() 
+prolactin$treatment <- factor(prolactin$treatment, levels = alllevels)
+```
 
-b <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus ~ parental stage", x = NULL)  + facet_wrap(~sex, scales = "free")
-c <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary ~ parental stage", x = NULL) + facet_wrap(~sex, scales = "free") 
-d <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads ~ parental stage")  + facet_wrap(~sex, scales = "free")
+# subset PRL vsd
 
-bcd <- plot_grid(b,c,d, nrow = 3, labels = c("b"), label_size = 8, rel_heights = c(1,1,1))
+```{r}
+# `candidatevsd` loaded with `source("../R/wrangledata.R")`
 
-e <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + labs(subtitle = "hypothalamus ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")  
-f <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + labs(subtitle = "pituitary ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")   
-g <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + labs(subtitle = "gonads ~ internal physiology" ) + facet_wrap(~sex, scales = "free")
+PRLpit <- candidatevsd %>% filter(tissue == "pituitary", gene == "PRL")
+PRLpitF <- candidatevsd %>% filter(tissue == "pituitary", gene == "PRL", 
+                                   sex == "female" , treatment != "control")
+PRLpitM <- candidatevsd %>% filter(tissue == "pituitary", gene == "PRL", 
+                                   sex == "male", treatment != "control")
+```
 
-efg <- plot_grid(e,f,g, nrow = 3, labels = c("d"), label_size = 8, rel_heights = c(1,1,1))
+## Internal versus external hyotheses
 
-h <- plottsneelipsev2(hyptsne, hyptsne$extint, allcolors ) + labs(subtitle = "hypothalamus ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")  
-i <- plottsneelipsev2(pittsne, pittsne$extint, allcolors ) + labs(subtitle = "pituitary ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")   
-j <- plottsneelipsev2(gontsne, gontsne$extint, allcolors ) + labs(subtitle = "gonads ~ external enviornment" ) + facet_wrap(~sex, scales = "free")
+```{r hypothesis}
 
-hij <- plot_grid(h,i,j, nrow = 3, labels = c("c"), label_size = 8, rel_heights = c(1,1,1))
+PRLpit %>%
+  group_by(sex) %>%
+  summarize(median = median(counts))
 
-bcdhijefg <- plot_grid(bcd, hij,  efg, nrow = 1)
+PRLvsd3 <- PRLpit %>%
+  mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
+  drop_na()
+PRLvsd3$hiloPRL <- factor(PRLvsd3$hiloPRL, levels = c("lo", "hi"))
 
-fig5 <- plot_grid(expdesign2, bcdhijefg, nrow = 2, labels = c("a"), label_size = 8, rel_heights = c(0.25,1))
-fig5
+PRLvsd3 %>%
+  group_by(sex, tissue, hiloPRL) %>%
+  summarize(n = n())
 
+DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG2 <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
+head(allDEG2)
+
+allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)
+
+allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
+allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
+                                                      "eggs vs. chicks" = "eggs_chicks"))
+allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))
+
+PRLDEGs <- allDEG2 %>%
+  filter(tissue == "pituitary", comparison == "lo vs. hi PRL   ", direction == "hi") %>%
+  arrange(desc(comparison))
+PRLDEGs
 ```
+
+```{r fig4, fig.width=7.25, fig.height=7.25}
+
+expdesign <- png::readPNG("../figures/images/fig_fig4a.png")
+expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
+
+a <-plotcolorfulpcs(pca1f,pca1f$treatment, allcolors) + labs(subtitle = "Female pituitary") 
+b <- plotfriz(pca2f) + labs(subtitle = " ") 
+c <- plotprolactin(PRLpitF, PRLpitF$counts, "PRL", " ") + 
+  theme(legend.position = "none", 
+        axis.title.y = element_text(face = "italic"))  
+d <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2100)   
+e <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2100) + theme(axis.title.y = element_blank())
+
+f <- plotcolorfulpcs(pca1m, pca1m$treatment, allcolors) + labs(subtitle = "Male pituitary")
+g <- plotfriz(pca2m) + labs(subtitle = " ") 
+h <- plotprolactin(PRLpitM, PRLpitM$counts, "PRL", " ") + theme(legend.position = "none", 
+        axis.title.y = element_text(face = "italic"))  
+i <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2100)  
+j <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2100) + theme(axis.title.y = element_blank())
+
+d2m <- plot_grid(a,b,c,d,e,f,g,h,i,j, nrow = 2, rel_widths = c(1,1,2,0.8, 0.75), 
+                 labels = c("D", "E", "F", "G", "H", "I", "J", "K", "L", "M"), label_size = 8)
+
+fig4 <- plot_grid(expdesign,d2m, ncol = 1, rel_heights = c(0.25,1))
+fig4
+```
+
+```{r savefiles}
+#write.csv(PRLvsd3,"../results/PRLvsd.csv", row.names = F)
+```

### 2020-04-26 R/functions.R
now with WGCNA results incorporated into fig 4

@@ -795,8 +795,8 @@ plotfriz <- function(frizdf){
   p <- fviz_pca_var(frizdf,  
                     axes.linetype = "blank", 
                     repel = T , 
-                    select.var= list(contrib = 3),
-                    labelsize = 3)  + 
+                    select.var= list(contrib = 5),
+                    labelsize = 2)  + 
     labs(title = NULL) + 
     theme_B3() +
     theme(axis.text = element_blank(),
@@ -1012,23 +1012,38 @@ rplot2 <- function(rdf,
                         size = "size", alpha = "size",
                         label = "label")) +
     plot_
-  
-  #   # plot
-  #   ggplot(aes_string(x = "x", y = "y", color = "r",
-  #                                       size = "size", alpha = "size",
-  #                                       label = "label")) +
-  #   geom_point(shape = shape) +
-  #   scale_colour_gradientn(limits = c(-1, 1), colors = colours) +
-  #   labs(x = "", y ="") +
-  #   theme_classic()
-  # 
-  # if (print_cor) {
-  #   p <- p + geom_text(color = "black", size = 3, show.legend = FALSE)
-  # }
-  # 
-  # if (!legend) {
-  #   p <- p + theme(legend.position = "none")
-  # }
-  # 
-  # p
 }
+
+
+##### 3 funcitons  for correlation plots
+
+makecorrdf <- function(whichsex, whichtissue, whichgenes){
+  
+  corrrdf <- candidatevsd %>%
+    filter(sex == whichsex, tissue == whichtissue,
+           gene %in% whichgenes) %>%
+    pivot_wider(names_from = gene, values_from = counts) %>%
+    select(-sex,-tissue, -treatment, -samples) %>%
+    correlate() %>%
+    rearrange()
+  print(head(corrrdf))
+  return(corrrdf)
+}
+
+subsetcandidatevsdwide <- function(whichsex, whichtissue){
+  df <- candidatevsdwide %>%
+    filter(sex == whichsex, tissue == whichtissue,
+           treatment != "control") 
+  return(df)
+}
+
+plotcorrplot <- function(df, subtitle){  
+  p <- df %>%
+    rplot2() +
+    theme_B3() +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1),
+          axis.text = element_text(face = "italic"))
+  
+  return(p)
+}
+

### 2020-04-26 analysis/figure-3.Rmd
now with WGCNA results incorporated into fig 4

@@
 
-candidatevsdwide <- candidatevsd  %>%
-    pivot_wider(names_from = gene, values_from = counts) 
-head(candidatevsdwide)
-
-makecorrdf <- function(whichsex, whichtissue, whichgenes){
-
-  corrrdf <- candidatevsd %>%
-    filter(sex == whichsex, tissue == whichtissue,
-           gene %in% whichgenes) %>%
-    pivot_wider(names_from = gene, values_from = counts) %>%
-    select(-sex,-tissue, -treatment, -samples) %>%
-    correlate() %>%
-    rearrange()
-  print(head(corrrdf))
-  return(corrrdf)
-}
-
 c1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
@@
 
-subsetcandidatevsdwide <- function(whichsex, whichtissue){
-  df <- candidatevsdwide %>%
-    filter(sex == whichsex, tissue == whichtissue,
-           treatment != "control") 
-  return(df)
-}
+candidatevsdwide <- candidatevsd  %>%
+    pivot_wider(names_from = gene, values_from = counts) 
+head(candidatevsdwide)
 
@@
 ```{r fig3, fig.width = 7.25, fig.height = 7.25}
-
-plotcorrplot <- function(df, subtitle){  
-  p <- df %>%
-    rplot2() +
-    theme_B3() +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1),
-          axis.text = element_text(face = "italic"))
-
-  return(p)
-}
 

### 2020-04-26 analysis/figure-4.Rmd
now with WGCNA results incorporated into fig 4

@@
 library(factoextra)
+library(corrr)
+library(ggpubr)
 
@@
 
-## PRL hormone
-
-```{r hormones}
-
-# hormones
-prolactin <- read_csv("../results/07_hormones.csv") %>%
-    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
-    droplevels() 
-prolactin$treatment <- factor(prolactin$treatment, levels = alllevels)
-```
-
 # subset PRL vsd
@@
                                    sex == "male", treatment != "control")
-```
-
-## Internal versus external hyotheses
-
-```{r hypothesis}
 
@@
   summarize(n = n())
+```
+
+## Internal versus external hyotheses
+
+```{r hypothesis}
 
@@
 
-expdesign <- png::readPNG("../figures/images/fig_fig4a.png")
-expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
-
 a <-plotcolorfulpcs(pca1f,pca1f$treatment, allcolors) + labs(subtitle = "Female pituitary") 
@@
 
-fig4 <- plot_grid(expdesign,d2m, ncol = 1, rel_heights = c(0.25,1))
+## resutls from WGCNA prl module
+
+PRLgenes <- read_csv("../results/PRLmodule.csv") %>% pull(x)
+
+candidatevsd <- getcandidatevsd(PRLgenes, "pituitary", sexlevels)
+
+df <- candidatevsd %>%
+    pivot_wider(names_from = gene, values_from = counts) %>%
+    select(-sex,-tissue, -treatment, -samples) %>%
+    correlate() %>%
+  focus(PRL)  %>%
+  arrange(rowname)
+df
+
+k <- df %>%
+  filter(PRL > 0.65) %>% 
+  mutate(PRLrounded = round(PRL, 2))%>% 
+  ggplot(aes(x = reorder(rowname, PRL), y = PRL)) +
+  geom_bar(stat = "identity") +
+  theme_B3() +
+  theme(axis.text.y = element_text(face = "italic")) +
+  labs(x = " ", y = "Correlation with PRL", subtitle = "Female and male pituitary") +
+
+  coord_flip() +
+  scale_y_continuous(expand = c(0, 0))
+
+l <- candidatevsd %>%
+  filter(gene == "LAPTM4B", treatment != "control") %>%
+  ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
+  geom_boxplot() +
+  scale_fill_manual(values = allcolors) +
+  scale_color_manual(values = allcolors)  +
+  theme_B3()  +
+  theme(legend.position = "none", axis.title.y = element_text(face = "italic")) +
+  labs(x = "Parental stage", y = "LAPTM4B", subtitle = " " )
+
+m <- candidatevsd %>%
+  filter(gene %in% c("LAPTM4B", "PRL"),  treatment != "control") %>%
+  pivot_wider(names_from = gene, values_from = counts) %>%
+  ggplot(aes(x = PRL, y = LAPTM4B)) +
+  geom_point(aes(color = treatment)) +
+  geom_smooth(aes(color = sex), method = "lm") +
+  scale_fill_manual(values = allcolors) +
+  scale_color_manual(values = allcolors)  +
+  theme_B3() +
+  theme(axis.title = element_text(face = "italic"),
+        legend.position = "none") +
+  stat_cor( size = 2) + labs(subtitle = " ")
+
+klm <- plot_grid(k,l,m, rel_widths = c(1,1,0.75), nrow = 1, 
+                 labels = c("N", "O", "P"), label_size = 8)
+
+expdesign <- png::readPNG("../figures/images/fig_fig4a.png")
+expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
+
+fig4 <- plot_grid(expdesign,d2m, klm, ncol = 1, rel_heights = c(0.25, 0.75, 0.4))
 fig4

### 2020-04-26 analysis/figure-2.Rmd
new fig 2

@@
 
-```{r setup}
+```{r setup, message=F}
 library(tidyverse)
@@
 library(kableExtra)
+library(corrr)
 
 source("../R/themes.R")
+source("../R/functions.R")
 source("../R/wrangledata.R")
@@
 
-table1 <- left_join(candidateDEGS, GOgenesWide) %>%
+## table 1 summary candidate genes
+table1 <- left_join(candidateDEGS, parentalcaregenes) %>%
   select(gene, bldg_lay:n5_n9, parentalcare, parentalbehavior, NCBI) %>%
@@
 table1[is.na(table1)] <- " " # replace NA with blank space so it's pretty
-#table1
+table1
 
@@
 
-## Candidate variance stabilized gene expression (vsds)
+## Candidate Correlations
 
-```{r fig2, fig.width = 7.25, fig.height = 7.25}
+```{r}
+# load `candidatevsd` with `source("../R/wrangledata.R")`
+head(candidatevsd)
 
-# `candidatevsd` loaded with source(../R/wrangleata.R)
+curleychampagnegenes <- curleychampagnegenes %>% pull(gene)
 
-# significant candidate genes
-hypDEGs <- c("CRHBP", "CRHR2", "DRD1", "GNAQ", 
-             "HTR2C", "OPRK1", "PGR")
-pitDEGs <- c("ESR1", "MEST", "PRL", "PTEN")
-gonadDEGs <- c("AVPR1A", "CREBRF", "FOS", "GNAQ", 
-               "NR3C1", "OPRM1", "PRLR")
+hyp1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
+pit1 <- makecorrdf("female", "pituitary", curleychampagnegenes)  
+gon1 <- makecorrdf("female", "gonad", curleychampagnegenes)
 
+hyp2 <- makecorrdf("male", "hypothalamus", curleychampagnegenes)  
+pit2 <- makecorrdf("male", "pituitary", curleychampagnegenes)  
+gon2 <- makecorrdf("male", "gonad", curleychampagnegenes)
+
+hyp3 <- makecorrdf(sexlevels, "hypothalamus", curleychampagnegenes)  
+pit3 <- makecorrdf(sexlevels, "pituitary", curleychampagnegenes)  
+gon3 <- makecorrdf(sexlevels, "gonad", curleychampagnegenes)
+
+candidatevsdwide <- candidatevsd  %>%
+    pivot_wider(names_from = gene, values_from = counts) 
+FH <- subsetcandidatevsdwide("female", "hypothalamus")
+FP <- subsetcandidatevsdwide("female", "pituitary")
+FG <- subsetcandidatevsdwide("female", "gonad")
+MH <- subsetcandidatevsdwide("male", "hypothalamus")
+MP <- subsetcandidatevsdwide("male", "pituitary")
+MG <- subsetcandidatevsdwide("male", "gonad") 
+```
+
+### Candidate gene functions
+
+```{r}
 candidateboxplot <- function(whichtissue, whichgenes, whichsex){
@@
     geom_jitter(size = 0.25, aes(color = sex)) +
-    facet_wrap(~gene, scales = "free_y", nrow = 1) +
+    facet_wrap(~sex, scales = "free_y", nrow = 2) +
     scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9),
@@
     theme(legend.position = "none",
-          strip.text = element_text(face = "italic"),
+          axis.title.y = element_text(face = "italic"),
           axis.text.x = element_blank()) +
-    labs(subtitle = whichsex,
-         y = "gene expression",
+    labs(y = whichgenes,
          x = NULL) 
@@
 
-a1 <- candidateboxplot("hypothalamus", hypDEGs, "female") + labs(x = NULL, title = "hypothalamus")
+scattercorrelations <- function(df, gene1, gene2, mylinecolor, myxlab, myylab){
+  p <- ggplot(df, aes(x = gene1, y = gene2)) +
+    labs(subtitle = " ", x = myxlab, y = myylab) + 
+    scale_color_manual(values = allcolors) +
+    geom_smooth(method = "lm",  color = mylinecolor) +
+    geom_point(aes(color = treatment))  +
+    theme_B3() +  
+    theme(axis.title = element_text(face = "italic"), 
+            legend.position = "none") +
+    stat_cor( size = 2)
+  return(p)
+}
 
-b1 <- candidateboxplot("pituitary", pitDEGs, "female") + labs(x = NULL, title = "pituitary")
+```
 
-b2 <- candidateboxplot("pituitary", "PRL", "male") + labs(x = NULL, y = NULL,  title = "pituitary")
+```{r fig2, fig.width = 7.25, fig.height = 7.25}
 
-b12 <- plot_grid(b1,b2,nrow = 1, labels = c("B", "C"),
-          label_size = 8, rel_widths = c(4,1))
+## plot sig candidate genes
+a <- candidateboxplot("hypothalamus", "DRD1", sexlevels) + labs(x = NULL, title = "Hypothalamus")
 
-c1 <- candidateboxplot("gonad", gonadDEGs, "female") + 
-  labs(x = NULL, title = "gonad")
+g <- candidateboxplot("pituitary", "PRL", sexlevels) + labs(x = NULL, title = "Pituitary")
 
-plot_grid(a1, b12, c1, nrow = 3, labels = c("A", " ","D"), 
-          label_size = 8)
+m <- candidateboxplot("gonad", "CREBRF", sexlevels) + labs(x = NULL, title = "Gonad") + 
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+c <- plotcorrplot(hyp3, "females and males") + theme(legend.position = "none")  + labs(title = "")
+i <- plotcorrplot(pit3, "females and males") + theme(legend.position = "none") + labs(title = "")
+o <- plotcorrplot(gon3, "females and males") + theme(legend.position = "bottom")  + labs(title = "")
+
+e <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female")+ labs(title = "")
+k <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = "female")+ labs(title = "")
+q <- scattercorrelations(FG, FG$ESR2, FG$CRHBP, "#969696", "ESR2", "CRHBP")   + labs(subtitle = "female")+ labs(title = "")
+
+f <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C") + theme(axis.title.y = element_blank()) + labs(subtitle = "male") + labs(title = "")
+l <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1") + theme(axis.title.y = element_blank()) + labs(subtitle = "male") + labs(title = "")
+r <- scattercorrelations(MG, MG$ESR2, MG$CRHBP, "#525252", "ESR2", "CRHBP")  + theme(axis.title.y = element_blank()) + labs(subtitle = "male") + labs(title = "")
+
+plot_grid(a,c,e,f,g,i,k,l,m,o,q,r, nrow = 3, rel_widths = c(1,1.2,0.8,0.75), rel_heights = c(1,1,1.2),
+          labels = "AUTO", label_size =  8)
 ```

### 2020-04-29 analysis/figure-2.Rmd
now with control in table 1

@@
   pivot_wider(names_from = comparison, values_from = res) %>%
-  select(gene, bldg_lay, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
+  select(gene, control_bldg, bldg_lay, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
 candidateDEGS
@@
 table1 <- left_join(candidateDEGS, parentalcaregenes) %>%
-  select(gene, bldg_lay:n5_n9, parentalcare, parentalbehavior, NCBI) %>%
+  select(gene, control_bldg:n5_n9, parentalcare, parentalbehavior, NCBI) %>%
   mutate(parentalcare = if_else(is.na(parentalcare), " ", "X"),
@@
 table1[is.na(table1)] <- " " # replace NA with blank space so it's pretty
-table1
+head(table1)
 

### 2020-04-29 analysis/figure-2.Rmd
better table

@@
 table1$numDEGs <- rowSums(is.na(table1)) # count NAs to know how many are NS
+
 table1 <- table1 %>% 
   mutate(sig = ifelse(numDEGs == 6, "NS", "DEG")) %>% 
-  arrange(sig, gene)  %>%  select(-sig, -numDEGs)
+  arrange(gene)  %>%  select(-sig, -numDEGs)
 table1[is.na(table1)] <- " " # replace NA with blank space so it's pretty

### 2020-05-03 analysis/figure-2.Rmd
?

@@
 
+###  tissue specific expression
+
+```{r}
+head(allvsd)
+
+tissuespecificity <- allvsd %>%
+  mutate(tissue = sapply(strsplit(as.character(samples),'\\_'), "[", 3)) %>% 
+  group_by(X1) %>%
+  summarize(tissues = str_c(tissue, collapse = ", ")) %>%
+  filter(tissues != "gon, hyp, pit") %>%
+  group_by(tissues) %>%
+  summarize(genes = str_c(X1, collapse = ", ")) %>% 
+  mutate(n = str_count(genes, ",") + 1) %>% 
+  arrange(desc(n))
+tissuespecificity
+
+```
+
 ```{r writefiles}

### 2020-05-03 analysis/22_symphony.Rmd
transcriptional symphony

@@
 ---
-title: "music"
+title: "Transcriptoinal symphony"
 output: md_document
@@
 source("../R/themes.R")
+
+theme_music <- function () { 
+  theme_classic(base_size = 12,
+                base_family = 'Helvetica') +
+    theme(
+      panel.grid.major  = element_blank(),  # remove major gridlines
+      panel.grid.minor  = element_blank()  # remove minor gridlines
+    )
+}
+
 knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/favegenes/")
@@
 
-## R Markdown
+## Music example
 
-```{r respect}
-song <- c("What", "you", "want,", "ba-", "by", "I", "got.")
-stage <- c( "bldg", "lay", "inc.d9", "inc.d17", "hatch", "n5", "n9")
-time <- c(1,2,3,4,5,6,7)
-position <- c(1,2,3,4,5,6,7)
-gene1 <- c(11,9,7,8,8,8,6)
-gene2 <- c(7, 9,5,6,8,8,4)
-gene3 <- c(4, 9,2,4,8,8,1)
-
-respect <- as.data.frame(t(rbind(position, gene1,gene2,gene3)))
-respect$song <- song
-respect$stage <- stage
-respect$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
-                        size=7, replace = TRUE)
-respect <- respect %>%
-  pivot_longer(gene1:gene3, names_to = "music", values_to = "genes")  %>% 
-  mutate(songstage = paste(song,stage, sep = "\n\n"))
-
-songstageslim <- unique(respect$songstage)
-
-respectsong <- ggplot(respect, aes(x = position, y = genes)) +
-  geom_line(aes(color = music)) +
-  geom_image(aes(image=image), size = 0.15) +
-  scale_x_continuous(breaks = position,
-                     labels = songstageslim,
-                     name = element_blank()) +
-  scale_y_continuous(breaks = c(3, 5, 7, 9),
-                     labels = c( "F",   "A",   "C",  "E"),
-                     limits = c(-1,12)) +
-  geom_hline(yintercept=c(2.15,4.15,6.15,8.15,10.15))   +
-  theme_B3() +
-  theme(axis.ticks = element_blank()) +
-  labs(subtitle = "RESPECT by Otis Redding, Jr.", y = NULL)
-respectsong
-```
-
-```{r feelinggood}
-song <- c("Birds", "fly-", "ing", "high,", "you", "know", "how", "I", "feel.")
-stage <- charlevels
-position <- c(1,2,3,4,5,6,7,8,9)
-gene1 <- c(-2,0,1,2,2,1,1,-2,0)
-gene2 <- gene1 - 1
-gene1 <- gene2
-gene3 <- gene2
-
-feelinggood <- as.data.frame(t(rbind(position, gene1,gene2,gene3)))
-feelinggood$song <- song
-feelinggood$stage <- stage
-feelinggood$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
-                        size=9, replace = TRUE)
-feelinggood <- feelinggood %>%
-  pivot_longer(gene1:gene3, names_to = "music", values_to = "genes")   %>%
-  mutate(songstage = paste(song,stage, sep = "\n\n"))
-feelinggood
-
-songstageslim <- unique(feelinggood$songstage)
-
-feelinggoodsong <- ggplot(feelinggood, aes(x = position, y = genes)) +
-  geom_line(aes(color = music)) +
-  geom_image(aes(image=image), size = 0.15) +
-  scale_x_continuous(breaks = position,
-                    labels = songstageslim,
-                     name = element_blank()) +
-  scale_y_continuous(breaks = c(3, 5, 7, 9),
-                     labels = c( "F",   "A",   "C",  "E"),
-                     limits = c(-6,12)) +
-  geom_hline(yintercept=c(2.15,4.15,6.15,8.15,10.15))   +
-  theme_B3() +
-  theme(axis.ticks = element_blank()) +
-  labs(subtitle = "Feeling Good by L. Bricusse & A. Newley", y = NULL)
-feelinggoodsong
-
-```
-
-```{r sinnerman}
-song <- c("Oh,", "sin-", "ner", "man,,", "where", "_", "he", "gon'", "run", "_", "to?")
+```{r transcriptionalsymphony}
+song <- c("Oh,", "sin-", "ner", "man,", "where", "_", "he", "gon'", "run", "_", "to?")
 stage <- c( "bldg", "lay", "inc.d3", "m.inc.d3" ,
@@
 position <- c(1,2,3,4,5,6,7,8,9,10,11)
-gene1 <- c(0,0,0,0,2,2,1,0,0,0,0)
+gene1 <- c(4,4,4,4,6,6,5,4,4,4,4)
 gene2 <- gene1 + 2
@@
 sinnerman <- sinnerman %>%
-  pivot_longer(gene1:gene3, names_to = "music", values_to = "genes")   %>%
-  mutate(songstage = paste(song,stage, sep = "\n\n"))
-sinnerman
-
-songstageslim <- unique(sinnerman$songstage)
+  pivot_longer(gene1:gene3, names_to = "music", values_to = "genes")
 
 sinnermansong <- ggplot(sinnerman, aes(x = position, y = genes)) +
-  geom_line(aes(color = music)) +
-  geom_image(aes(image=image), size = 0.1) +
+  #geom_line(aes(color = music)) +
+  geom_image(aes(image=image), size = 0.15) +
   scale_x_continuous(breaks = position,
-                    labels = songstageslim,
+                    labels = song,
                      name = element_blank()) +
@@
   geom_hline(yintercept=c(2.15,4.15,6.15,8.15,10.15))   +
-  theme_B3() +
-  theme(axis.ticks = element_blank(),
-        axis.text.x = element_text(angle = 10, hjust = 1)) +
+  theme_music() +
+  theme(axis.ticks = element_blank()) +
   labs(subtitle = "Sinnerman by Nina Simone", y = NULL)
@@
 ```
-```{r soulmusic}
-
-p3 <- plot_grid(respectsong + theme(legend.position = "bottom"), 
-                sinnermansong + theme(legend.position = "bottom"), 
-                nrow = 1, rel_widths = c(0.45, 0.55))
-p3
-```
-
-```{r symphonymodules}
-modules <- read_csv("../results/08_genes_modules.csv") %>%
-  rename(modulecolor = `net$colors`)
-
-candidategenes <- c("PRL", "PRLR", 
-                 #"VIP", "VIPR1", "VIPR2", 
-                 "OXT", "AVP",  "AVPR1B", # "AVPR1A",
-                # "GNRH1","GNRHR", "NPVF",
-                # "NR3C1", "NR3C2",
-                # "ESR1", "ESR2",
-                "AR",
-                # "DIO2","LEPR", "DIO3", "DIO1","CYP19A1",
-                # "HSPA14", "HSPA12A",
-                # "PTGES3", "HSD11B2",
-                 "DRD5", "DRD1", "DRD2",
-                 #"PGE1", "PGF",
-                 "BRCA1", "MYC", "PRKCZ" ,"FOSL2")
-
-modulescandidates <- modules %>% filter(gene %in% candidategenes)  %>%
-  group_by(modulecolor)  %>%
-  summarize(gene = str_c(gene, collapse = ", "))
-
-modulescandidates
-
-datapath <- "../results/"   # path to the data
-datafiles <- dir(datapath, pattern = "*allvsd.csv") # get file names
-datapathfile <- paste0(datapath, datafiles)
-
-df <- datapathfile %>%
-  setNames(nm = .) %>%
-  map_df(~read_csv(.x, col_types = cols(), col_names = T), .id = "filename") %>% 
-  mutate(tissue = sapply(strsplit(as.character(filename),'../results/06_'), "[", 2)) %>% 
-  mutate(tissue = sapply(strsplit(as.character(tissue),'allvsd.csv'), "[", 1))  %>% 
-  select(tissue, X1, everything())
-
-df2 <-  df  %>%
-  filter(X1 %in% c(candidategenes)) %>%
-  pivot_longer(cols = L.Blu13_male_gonad_control.NYNO:y98.o50.x_male_pituitary_inc.d3, 
-               names_to = "sample", values_to = "vsd") %>%
-   mutate(sex = sapply(strsplit(as.character(sample),'_'), "[", 2)) %>%
-   mutate(treatment = sapply(strsplit(as.character(sample),'_'), "[", 4))  %>%
-   mutate(treatment = sapply(strsplit(as.character(treatment),'.NYNO'), "[", 1)) %>%
-  mutate(bird = sapply(strsplit(as.character(sample),'_'), "[", 1)) %>%
-  filter(treatment %in% charlevels) %>%
-  select(bird, sex, treatment, tissue, X1, vsd) %>%
-  mutate(tissue = fct_recode(tissue, "hypothalamus" = "hyp",
-                    "pituitary" = "pit",
-                    "gonads" = "gon"
-                    )) %>%
-  rename(gene = X1) %>%
-  drop_na() %>%
-  droplevels()
-head(df2)
-
-df2$treatment <- factor(df2$treatment, levels = alllevels)
-df2$tissue <- factor(df2$tissue, levels = tissuelevels)
-df2$gene <- factor(df2$gene)
-
-df3 <- df2 %>% 
-  mutate(treatment = fct_relevel(treatment, charlevels)) %>% 
-  group_by(treatment, tissue, gene)  %>% 
-  summarize(m = mean(vsd, na.rm = T), se = sd(vsd,  na.rm = T)/sqrt(length(vsd))) %>%
-  mutate(image = "../figures/images/DoveParentsRNAseq_note.png") 
-head(df3)
-
-d4 <- left_join(df3, modules, by = "gene")
-
-for (i in levels(d4$tissue)) {
-  p <-  d4 %>%
-    filter(tissue == i) %>%
-    ggplot(aes(x = treatment, y = m)) +
-    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
-    geom_point(size = 1, aes(color = gene)) +
-    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
-    scale_alpha_manual(values = c(0.5, 1)) +
-    labs(subtitle = i, y = "average expression", x = "parental stage") +
-    facet_wrap(~modulecolor, nrow = 2) +
-    theme_B3() +
-    theme(legend.position = "bottom")
- print(p)
-}
-
-d4 %>%
-  filter(modulecolor != "turquoise",
-         tissue == "pituitary") %>%
-  droplevels() %>% 
-    ggplot(aes(x = treatment, y = m)) +
-    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
-    geom_point(size = 1, aes(color = gene)) +
-    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
-    scale_alpha_manual(values = c(0.5, 1)) +
-    labs(subtitle = "WGCNA + canddiate genes in the pituitary", y = "average expression", x = "parental stage") +
-    facet_wrap(~modulecolor, scales = "free_y", nrow = 2) +
-    theme_B3() +
-    theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
-    guides(color = guide_legend(nrow = 2, byrow = T))
-
-redgenes <- d4 %>%
-  filter( modulecolor == "red",
-          tissue == "pituitary") %>%
-  droplevels() %>% 
-    ggplot(aes(x = treatment, y = m)) +
-  geom_errorbar(aes(ymin=m-se, ymax=m+se, color=treatment), width=.1) +
-  geom_image(aes(image=image), size = 0.15) +
-  #geom_smooth(aes(x = as.numeric(treatment)), se = F, color = "red") +
-  facet_wrap(~modulecolor, scales = "free_y", nrow = 2) +
-  labs(subtitle = "WGCNA red module", 
-       y = "gene expression  in the pitutiary", x = "parental stage") +
-    facet_wrap(~gene, scales = "free_y", nrow = ) +
-    theme_B3() +
-    theme(legend.position = "none", 
-          axis.text.x = element_text(angle = 45, hjust = 1),
-          strip.text = element_text(face = "italic")) +
-  scale_color_manual(values = allcolors)
-
-redgenes
-
-```
-
-```{r}
-d4 %>% 
-  filter(gene %in% c("PRL", "PRLR", "ESR2")) %>%
-ggplot( aes(x = treatment, y = m, color = gene)) +
-    geom_smooth(aes(x = as.numeric(treatment)), se = F) +
-    facet_grid(gene~tissue, scales = "free")
-
-head(df)
-head(df2)
-```
-
-## outline
-
-- data-driven
-- hypothesis-drive
-- sonically-driven
-- visually-driven
-- statistcally-driven
-
-Data and hypothesis driven research for the visually, sonically, and statistcially inclined biologists

### 2020-05-03 analysis/figure-2.Rmd
new fig 2

--- 2020-04-29 analysis/figure-2.Rmd
@@
 library(corrr)
+library(ggsignif)
 
@@
 MG <- subsetcandidatevsdwide("male", "gonad")
+
 ```
@@
           gene %in% whichgenes,
-          sex %in% whichsex) %>%
+          sex %in% whichsex,
+          treatment != "control") %>%
     mutate(treatment = factor(treatment, levels = charlevels)) %>%
-    mutate(treatmentNum = as.numeric(treatment)) %>%
-    ggplot(aes(x = treatmentNum, y = counts)) +
+    ggplot(aes(x = treatment, y = counts)) +
     geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
-    #geom_smooth(aes(color = sex)) +
     geom_jitter(size = 0.25, aes(color = sex)) +
-    facet_wrap(~sex, scales = "free_y", nrow = 2) +
-    scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9),
-                       labels = charlevels) +
+    facet_wrap(~gene, scales = "free_y", nrow = 2) +
     scale_fill_manual(values = allcolors) +
@@
     theme(legend.position = "none",
-          axis.title.y = element_text(face = "italic"),
+          strip.text.y = element_text(face = "italic"),
+          strip.text.x = element_text(face = "italic"),
           axis.text.x = element_blank()) +
-    labs(y = whichgenes,
-         x = NULL) 
+    labs(y = paste(whichsex, "gene expression", sep = " "),
+         x = NULL) +
+    geom_signif(comparisons = list(c( "bldg", "lay"),
+                                 c( "lay", "inc.d3"),
+                                 c("inc.d3", "inc.d9"),
+                                 c( "inc.d9", "inc.d17"),
+                                 c( "inc.d17", "hatch"),
+                                 c("hatch", "n5"),
+                                 c( "n5", "n9")),  
+                map_signif_level=TRUE,
+                textsize = 1.5, family = 'Helvetica',
+                vjust = 1.5, size = 0)
+
   return(p)
@@
     scale_color_manual(values = allcolors) +
-    geom_smooth(method = "lm",  color = mylinecolor) +
+    geom_smooth(method = "glm",  color = mylinecolor) +
     geom_point(aes(color = treatment))  +
@@
 ## plot sig candidate genes
-a <- candidateboxplot("hypothalamus", "DRD1", sexlevels) + labs(x = NULL, title = "Hypothalamus")
+a <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female") + labs(title = "Hypothalamus")
+e <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = "female")+ labs(title = "Pituitary")
+i <- scattercorrelations(FG, FG$ESR1, FG$PGR, "#969696", "ESR1", "PGR")   + labs(subtitle = "female")+ labs(title = "Gonad")
 
-g <- candidateboxplot("pituitary", "PRL", sexlevels) + labs(x = NULL, title = "Pituitary")
+b <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C")  + labs(subtitle = "male") + labs(title = "")
+f <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = "male") + labs(title = "")
+j <- scattercorrelations(MG, MG$ESR1, MG$PGR, "#525252", "ESR1", "PGR")  + labs(subtitle = "male") + labs(title = "")
 
-m <- candidateboxplot("gonad", "CREBRF", sexlevels) + labs(x = NULL, title = "Gonad") + 
+c <- candidateboxplot("hypothalamus", c("DRD1", "HTR2C"), "female") + labs(x = NULL, title = "") 
+g <- candidateboxplot("pituitary", c( "AVPR1A", "CRHR1" ), "female") + labs(x = NULL, title = "")
+k <- candidateboxplot("gonad", c("ESR1", "PGR"), "female") + labs(x = NULL, title = "") + 
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
 
-c <- plotcorrplot(hyp3, "females and males") + theme(legend.position = "none")  + labs(title = "")
-i <- plotcorrplot(pit3, "females and males") + theme(legend.position = "none") + labs(title = "")
-o <- plotcorrplot(gon3, "females and males") + theme(legend.position = "bottom")  + labs(title = "")
+d <- candidateboxplot("hypothalamus",  c("DRD1", "HTR2C"), "male") + labs(x = NULL, title = " ") 
+h <- candidateboxplot("pituitary", c("AVPR1A", "CRHR1"), "male") + labs(x = NULL, title = " ")
+l <- candidateboxplot("gonad", c("ESR1", "PGR"), "male") + labs(x = NULL, title = " ") + 
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
 
-e <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female")+ labs(title = "")
-k <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = "female")+ labs(title = "")
-q <- scattercorrelations(FG, FG$ESR2, FG$CRHBP, "#969696", "ESR2", "CRHBP")   + labs(subtitle = "female")+ labs(title = "")
+plot_grid(a,b,c,d,e,f,g,h,i,j,k,l, nrow = 3, 
+          rel_widths = c(0.8,0.8,1,1), rel_heights = c(1,1,1.2),
+          labels = c("A", " ", "B", "",
+                     "C", " ", "D", "",
+                     "E", " ", "F", ""), label_size =  8)
 
-f <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C") + theme(axis.title.y = element_blank()) + labs(subtitle = "male") + labs(title = "")
-l <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1") + theme(axis.title.y = element_blank()) + labs(subtitle = "male") + labs(title = "")
-r <- scattercorrelations(MG, MG$ESR2, MG$CRHBP, "#525252", "ESR2", "CRHBP")  + theme(axis.title.y = element_blank()) + labs(subtitle = "male") + labs(title = "")
+```
 
-plot_grid(a,c,e,f,g,i,k,l,m,o,q,r, nrow = 3, rel_widths = c(1,1.2,0.8,0.75), rel_heights = c(1,1,1.2),
-          labels = "AUTO", label_size =  8)
+## suplement correlations
+
+```{r supplfig2, fig.width = 7.25, fig.height = 7.25}
+
+a <- plotcorrplot(hyp1, "females") + theme(legend.position = "none", axis.text.x = element_blank())  + labs(title = "")
+c <- plotcorrplot(pit1, "females") + theme(legend.position = "none", axis.text.x = element_blank()) + labs(title = "")
+e <- plotcorrplot(gon1, "females") + theme(legend.position = "bottom")  + labs(title = "")
+
+b <- plotcorrplot(hyp2, "males") + theme(legend.position = "none", axis.text.x = element_blank())  + labs(title = "")
+d <- plotcorrplot(pit2, "males") + theme(legend.position = "none", axis.text.x = element_blank()) + labs(title = "")
+f <- plotcorrplot(gon2, "males") + theme(legend.position = "bottom")  + labs(title = "")
+
+plot_grid(a,b,c,d,e,f, nrow = 3, rel_heights = c(1,1,1.5))
+
 ```

### 2020-05-03 R/themes.R
new fig 2

@@ -4,6 +4,7 @@ theme_B3 <- function () {
   theme_classic(base_size = 7,
                 base_family = 'Helvetica') +
     theme(
+      strip.background  = element_blank(),
       panel.grid.major  = element_blank(),  # remove major gridlines
       panel.grid.minor  = element_blank()  # remove minor gridlines
     )

### 2020-05-03 analysis/figure-2.Rmd
now all have labels

@@
           strip.text.x = element_text(face = "italic"),
-          axis.text.x = element_blank()) +
+          axis.text.x = element_text(angle = 45, hjust = 1)) +
     labs(y = paste(whichsex, "gene expression", sep = " "),
@@
 g <- candidateboxplot("pituitary", c( "AVPR1A", "CRHR1" ), "female") + labs(x = NULL, title = "")
-k <- candidateboxplot("gonad", c("ESR1", "PGR"), "female") + labs(x = NULL, title = "") + 
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+k <- candidateboxplot("gonad", c("ESR1", "PGR"), "female") + labs(x = NULL, title = "")
 
@@
 h <- candidateboxplot("pituitary", c("AVPR1A", "CRHR1"), "male") + labs(x = NULL, title = " ")
-l <- candidateboxplot("gonad", c("ESR1", "PGR"), "male") + labs(x = NULL, title = " ") + 
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+l <- candidateboxplot("gonad", c("ESR1", "PGR"), "male") + labs(x = NULL, title = " ")
 

### 2020-05-03 analysis/figure-2.Rmd
now with music notes

@@
 library(ggsignif)
+library(magick)
+library(scales)
+library(ggimage)
 
@@
 
-plot_grid(a,b,c,d,e,f,g,h,i,j,k,l, nrow = 3, 
+atof <- plot_grid(a,b,c,d,e,f,g,h,i,j,k,l, nrow = 3, 
           rel_widths = c(0.8,0.8,1,1), rel_heights = c(1,1,1.2),
@@
 
+musicplot <- function(whichgene, whichtissue){
+
+    treble <- magick::image_read("../figures/images/fig_treble.png")
+    base <- magick::image_read("../figures/images/fig_base.png")
+
+    p <- candidatevsd %>%
+      group_by(treatment, tissue, gene, sex)  %>% 
+      summarize(median = median(counts, na.rm = T), 
+                se = sd(counts,  na.rm = T)/sqrt(length(counts))) %>%
+      dplyr::mutate(scaled = rescale(median, to = c(0, 7))) %>%
+      dplyr::mutate(image = "../figures/images/musicnote.png")   %>%
+      filter(
+        gene %in% whichgene,
+        tissue %in% whichtissue      ) %>% 
+      collect() %>%
+      drop_na() %>%
+      filter(treatment != "control") %>%
+      ggplot( aes(x = treatment, y = median)) +
+      geom_errorbar(aes(ymin = median - se, 
+                        ymax = median + se, color = "white"),  width=0) +
+      geom_image(aes(image=image), size = 0.1)+
+      theme_void(base_size = 7) +
+      theme(legend.position = "none",
+            title = element_text(face = "italic"),
+            strip.text = element_text(color = "white")) +
+      scale_color_manual(values = allcolors) +
+      labs(y = " ", subtitle = " " ) +
+      facet_wrap(~sex, nrow = 2, scales = "free_y")
+      return(p)
+}
+
+scales <- png::readPNG("../figures/images/fig_music.png")
+scales <- ggdraw() +  draw_image(scales, scale = 1)
+
+p1 <- musicplot(c("DRD1", "HTR2C"), "hypothalamus") + labs(subtitle = " D1DR & HTR2C", title = " " )
+p2 <- musicplot(c("AVPR1A", "CRHR1"), "pituitary") + labs(subtitle = " AVPR1A & CRHR1 ", title = " " )
+p3 <- musicplot(c("ESR1", "PGR"), "gonad") + labs(subtitle = "  ESR1 & PGR", title = " ")
+
+p123 <- plot_grid(p1, NULL,  p2, NULL,p3, NULL,ncol = 1, rel_heights = c(1,0.2,1,0.2,1.2,0.2))
+
+plot_grid(atof, scales, p123, rel_widths = c(1.5,0.1, 0.3), ncol = 3)
 ```

### 2020-05-06 analysis/figure-2.Rmd
now with better fig 2, start of fig 3

--- 2020-05-03 analysis/figure-2.Rmd
@@
           gene %in% whichgenes,
-          sex %in% whichsex,
-          treatment != "control") %>%
+          sex %in% whichsex) %>%
     mutate(treatment = factor(treatment, levels = charlevels)) %>%
@@
     geom_jitter(size = 0.25, aes(color = sex)) +
-    facet_wrap(~gene, scales = "free_y", nrow = 2) +
+    facet_wrap(~sex, scales = "free_y", nrow = 1) +
     scale_fill_manual(values = allcolors) +
@@
     theme(legend.position = "none",
-          strip.text.y = element_text(face = "italic"),
-          strip.text.x = element_text(face = "italic"),
-          axis.text.x = element_text(angle = 45, hjust = 1)) +
-    labs(y = paste(whichsex, "gene expression", sep = " "),
+          axis.title.y = element_text(face = "italic"),
+          axis.text.x = element_blank(),
+          axis.ticks = element_blank()) +
+    labs(y = whichgenes,
          x = NULL) +
-    geom_signif(comparisons = list(c( "bldg", "lay"),
+    geom_signif(comparisons = list(c( "control", "bldg"),
+                                 c( "bldg", "lay"),
                                  c( "lay", "inc.d3"),
@@
 }
+
 ```
@@
 
-## plot sig candidate genes
-a <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female") + labs(title = "Hypothalamus")
-e <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = "female")+ labs(title = "Pituitary")
-i <- scattercorrelations(FG, FG$ESR1, FG$PGR, "#969696", "ESR1", "PGR")   + labs(subtitle = "female")+ labs(title = "Gonad")
+header <- png::readPNG("../figures/images/fig_fig2a.png")
+header <- ggdraw() +  draw_image(header, scale = 1)
 
-b <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C")  + labs(subtitle = "male") + labs(title = "")
-f <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = "male") + labs(title = "")
-j <- scattercorrelations(MG, MG$ESR1, MG$PGR, "#525252", "ESR1", "PGR")  + labs(subtitle = "male") + labs(title = "")
+b <- plotcorrplot(hyp3, "females and males") + theme(legend.position = "none")  + labs(title = "Hypothalamus")
+g <- plotcorrplot(pit3, NULL) + theme(legend.position = "none") + labs(title = "Pituitary")
+l <- plotcorrplot(gon3, NULL) + theme(legend.position = "none")  + labs(title = "Gonad")
 
-c <- candidateboxplot("hypothalamus", c("DRD1", "HTR2C"), "female") + labs(x = NULL, title = "") 
-g <- candidateboxplot("pituitary", c( "AVPR1A", "CRHR1" ), "female") + labs(x = NULL, title = "")
-k <- candidateboxplot("gonad", c("ESR1", "PGR"), "female") + labs(x = NULL, title = "")
+c <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female") + labs(title = "")
+h <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = NULL)+ labs(title = "")
+m <- scattercorrelations(FG, FG$ESR1, FG$PGR, "#969696", "ESR1", "PGR")   + labs(subtitle = NULL)+ labs(title = "")
 
-d <- candidateboxplot("hypothalamus",  c("DRD1", "HTR2C"), "male") + labs(x = NULL, title = " ") 
-h <- candidateboxplot("pituitary", c("AVPR1A", "CRHR1"), "male") + labs(x = NULL, title = " ")
-l <- candidateboxplot("gonad", c("ESR1", "PGR"), "male") + labs(x = NULL, title = " ")
+d <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C")  + labs(subtitle = "male") + labs(title = "")
+i <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = NULL) + labs(title = "")
+n <- scattercorrelations(MG, MG$ESR1, MG$PGR, "#525252", "ESR1", "PGR")  + labs(subtitle = NULL) + labs(title = "")
 
-plot_grid(a,b,c,d,e,f,g,h,i,j,k,l, nrow = 3, 
-          rel_widths = c(0.8,0.8,1,1), rel_heights = c(1,1,1.2),
-          labels = c("A", " ", "B", "",
-                     "C", " ", "D", "",
-                     "E", " ", "F", ""), label_size =  8)
+d1 <- candidateboxplot("hypothalamus", c("DRD1"), sexlevels) + labs(x = NULL , title = " ")  + theme(axis.line.x = element_blank()) 
+d2 <- candidateboxplot("hypothalamus", c("HTR2C"), sexlevels) + labs(x = "Parental stage" ) + theme(strip.text = element_blank())
+
+g1 <- candidateboxplot("pituitary", c("AVPR1A"), sexlevels) + labs(x = NULL, title = " " )  + theme( strip.text = element_blank(), axis.line.x = element_blank())
+g2 <- candidateboxplot("pituitary", c("CRHR1"), sexlevels) + labs(x = "Parental stage"  )  + theme(strip.text = element_blank())
+
+j1 <- candidateboxplot("gonad", c("ESR1"), sexlevels) + labs(x = NULL , title = " ") + theme(strip.text = element_blank(), axis.line.x = element_blank())
+j2 <- candidateboxplot("gonad", c("PGR"), sexlevels) + labs(x = "Parental stage"  ) + theme(strip.text = element_blank())
+
+btn <- plot_grid(b,c,d, g,h,i,l,m,n, ncol = 3, labels = c("B", "C", " ","E", "F", " ", "H", "I"), label_size = 8, rel_widths = c(1.4,1,1), rel_heights = c(1.1,1,1))
+dgj <- plot_grid(d1,d2,g1,g2,j1,j2, ncol = 1, labels = c("D", " ", "G", " " , "J"), label_size = 8, rel_heights = c(1.3,1,1.2,1,1.2,1))
+btoj <- plot_grid(btn,dgj, ncol = 2 , rel_widths = c(1.75,1))
+
+plot_grid(header, btoj,  rel_heights = c(0.2,1), ncol = 1)
 
@@
 
-## suplement correlations
+```{r fig3, fig.height=3, fig.width=7.25}
 
-```{r supplfig2, fig.width = 7.25, fig.height = 7.25}
+musicplot <- function(whichgene, whichtissue){
 
-a <- plotcorrplot(hyp1, "females") + theme(legend.position = "none", axis.text.x = element_blank())  + labs(title = "")
-c <- plotcorrplot(pit1, "females") + theme(legend.position = "none", axis.text.x = element_blank()) + labs(title = "")
-e <- plotcorrplot(gon1, "females") + theme(legend.position = "bottom")  + labs(title = "")
+    p <- candidatevsd %>%
+      group_by(treatment, tissue, gene, sex)  %>% 
+      summarize(median = median(counts, na.rm = T), 
+                se = sd(counts,  na.rm = T)/sqrt(length(counts))) %>%
+      dplyr::mutate(scaled = rescale(median, to = c(0, 7))) %>%
+      dplyr::mutate(image = "../figures/images/musicnote.png")   %>%
+      filter(
+        gene %in% whichgene,
+        tissue %in% whichtissue      ) %>% 
+      collect() %>%
+      drop_na() %>%
+      ggplot( aes(x = treatment, y = median)) +
+      geom_errorbar(aes(ymin = median - se, 
+                        ymax = median + se, color = "white"),  width=0) +
+      geom_image(aes(image=image), size = 0.1)+
+      theme_B3() +
+      theme(legend.position = "none",
+            title = element_text(face = "italic"),
+            strip.text = element_text(color = "white"),
+            axis.text.x = element_text(angle = 45, hjust = 1),
+            axis.text.y = element_blank(),
+            axis.line = element_blank(), axis.ticks = element_blank()) +
+      scale_color_manual(values = allcolors) +
+      labs(y = " ",x = NULL ) +
+      facet_wrap(~sex, nrow = 2, scales = "free_y")
+      return(p)
+}
 
-b <- plotcorrplot(hyp2, "males") + theme(legend.position = "none", axis.text.x = element_blank())  + labs(title = "")
-d <- plotcorrplot(pit2, "males") + theme(legend.position = "none", axis.text.x = element_blank()) + labs(title = "")
-f <- plotcorrplot(gon2, "males") + theme(legend.position = "bottom")  + labs(title = "")
+music <- png::readPNG("../figures/images/fig_music.png")
+music <- ggdraw() +  draw_image(music, scale = 1)
 
-plot_grid(a,b,c,d,e,f, nrow = 3, rel_heights = c(1,1,1.5))
+musicalgenes <- png::readPNG("../figures/images/fig_musicalgenes.png")
+musicalgenes <- ggdraw() +  draw_image(musicalgenes, scale = 1)
+
+p1 <- musicplot(c("AVP", "OXT"), "hypothalamus") + labs(subtitle = "Hypothalamic AVP & OXT" ) + theme(axis.text.x = element_blank())
+p2 <- musicplot(c("AVP", "OXT"), "pituitary") + labs(subtitle = "Pituitary AVP & OXT " ) + theme(axis.text.x = element_blank())
+p3 <- musicplot(c("AVP", "OXT"), "gonad") + labs(subtitle = "Gonadal  AVP & OXT"  )
+
+p123 <- plot_grid(music, p1, music, p3, ncol = 2, rel_widths = c(0.125,1))
+
+plot_grid(musicalgenes, p123, nrow = 1, rel_widths = c(1.5,1))
 

### 2020-05-06 analysis/02_DESeq2.Rmd
new fig 2b
now with total sex differenes

@@
 
-```{r PRL}
+```{r PRL, eval = F}
 PRLdata <- read_csv("../results/PRLvsd.csv")
@@
 plot_grid(ab, c, nrow = 1, rel_widths = c(2,1.2), labels = c(" ", "d"), label_size = 12)
-
-head(sexPRLdesg)
-
-PRLhi <- sexPRLdesg %>%
-  filter(direction.y == "hi")
-PRLhi
-
-PRLwgcna <- read_csv("../results/08_PRL_associated.csv") %>%
-  dplyr::rename("gene" =  "x") %>%
-  mutate(wgcna = "yes")
-PRLwgcna
-
-PRLwgcnaDEGs <- left_join(PRLwgcna, sexPRLdesg) %>%
-  select(-direction.x) %>%
-  dplyr::rename("PRL" =  "direction.y") 
-PRLwgcnaDEGs %>%
-  group_by(wgcna, PRL) %>%
-  summarize(n = n())
-head(PRLwgcnaDEGs)
-# 90 of the 98 genes in the wgcna module are also DEGs, with increase expression in hi PRL
 ```
@@
 
-```{r}
+```{r, eval = F}
 # label each bird with weather or not their pituitary had hi PRL or not
@@
   group_by(sex, tissue, hiloPRL, external) %>%
+  summarize(n = n())
+
+colDatahypothesis %>%
+  filter(tissue == "pituitary")  %>%
+  group_by(sex, tissue, hiloPRL, treatment) %>%
   summarize(n = n())
@@
 ```
+
+## tissue and sex differences
+
+```{r}
+
+for(i in "gonad"){
+
+  newcolData <- subsetcolData4(colData, i)
+  print(head(newcolData))
+  # save counts that match colData
+  savecols <- as.character(newcolData$V1) 
+  savecols <- as.vector(savecols)
+
+  newcountData <- countData %>% dplyr::select(one_of(savecols))
+
+  dds <- DESeqDataSetFromMatrix(countData = newcountData,
+                                colData = newcolData,
+                                design = ~ sex )
+
+  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
+  print(dds)
+  print(dim(dds))
+  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
+
+  vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
+
+  myfilename = paste0("../results/DEseq2/sex/", i, "_vsd.csv")
+  write.csv(vsd, myfilename)
+
+  print(head(vsd))
+
+  # save differential gene expression results
+  sex <- createDEGdfsavesex("male", "female", i)
+}
+
+```

### 2020-05-06 analysis/figure-1.Rmd
rm whitespace
now with total sex differenes

--- 2020-04-21 analysis/figure-1.Rmd
@@
 library(Rtsne)
-
 library(kableExtra) 
@@
 allDEG$direction <- factor(allDEG$direction, levels = charlevels)
+
+hypsex <- read_csv("../results/DEseq2/sex/hypothalamus_female_male_DEGs.csv")
+pitsex <- read_csv("../results/DEseq2/sex/pituitary_female_male_DEGs.csv")
+gonsex <- read_csv("../results/DEseq2/sex/gonad_female_male_DEGs.csv")
+
 ```
@@
 ```{r fig1, fig.width=7.25, fig.height=7.25}
-a <- png::readPNG("../figures/images/fig_fig1b.png")
+
+a <- png::readPNG("../figures/images/fig_fig1a.png")
 a <- ggdraw() +  draw_image(a, scale = 1)
 
-b <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = " ", subtitle = "~ tissue * sex")    
-c <- plottsneelipse(ftsne, ftsne$tissue, allcolors ) + labs(y = " ", subtitle = "~ tissue, females only")
-d <- plottsneelipse(mtsne, mtsne$tissue, allcolors ) + labs(y = " ", subtitle = "~ tissue, males only")
+b <- plottsneelipse(chartsne, chartsne$tissue, allcolors)   + labs(y = "tSNE 2 ", subtitle = " ")  
+c <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = "tSNE 2 ", subtitle = " ")
 
-abcd <- plot_grid(a,b,c,d, nrow = 1, labels = c("A", "B", "C", "D"), label_size = 8 )
+d1 <- sexbarplots(hypsex, 0, 7000) + labs(y = "DEGS", subtitle = "hypothalamus" )
+d2 <- sexbarplots(pitsex, 0, 7000) + labs(subtitle = "pituitary" ) + 
+  theme(axis.text.y = element_blank(), axis.line.y = element_blank(), axis.ticks = element_blank())
+d3 <- sexbarplots(gonsex, 0, 7000) + labs(subtitle = "gonads" ) +
+  theme(axis.text.y = element_blank(), axis.line.y = element_blank(), axis.ticks = element_blank())
 
-e <- png::readPNG("../figures/images/fig_fig1a.png")
-e <- ggdraw() +  draw_image(e, scale = 1)
-
-f <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
+e <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
   labs(x = NULL, subtitle = "hypothalamus")  + 
@@
 
-h <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
+g <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
   labs(x = NULL, subtitle = "pituitary") + 
@@
 
-j <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
+i <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
    facet_wrap(~sex, scales = "free") +
@@
 
-fhj <- plot_grid(f,h,j, nrow = 3, rel_heights = c(1,1,1.4),
-                 labels = c("F", "H", "J"), label_size = 8)
-
 # hyp
-g <- makebargraph("hypothalamus","DEGs", 0, 4300) + 
+f <- makebargraph("hypothalamus","DEGs", 0, 4300) + 
   theme(axis.text.x = element_blank(), 
@@
 # pit
-i <- makebargraph("pituitary","DEGs", 0, 4300)  +  
+h <- makebargraph("pituitary","DEGs", 0, 4300)  +  
   theme(axis.text.x = element_blank(), 
@@
 # gon
-k <- makebargraph("gonad","DEGs", 0, 4300) +  
+j <- makebargraph("gonad","DEGs", 0, 4300) +  
   theme(strip.text.x = element_blank()) +
@@
 
-gik <- plot_grid(g,i,k, nrow = 3, rel_heights = c(1,1,1.4),
-                 labels = c("G", "I", "K"), label_size = 8)
+bcd <- plot_grid(b,c,d1,d2,d3, nrow = 1, labels = c("B", "C", "D"), 
+                   label_size = 8, rel_widths = c(1.2,1.2,1.2,1,1))
 
-fhjgik <- plot_grid(fhj, gik, nrow = 1, align = "h", rel_widths = c(1.5,2))
+etoj <- plot_grid(e,f,g,h,i,j, ncol = 2, rel_heights = c(1,1,1.4), rel_widths = c(2.4,3.2),
+                  labels = c("E", "F", "G", "H", "I", "J"), label_size = 8)
 
-fig1 <- plot_grid(abcd, e, fhjgik, nrow = 3, rel_heights = c(0.7,0.5,1.5),
-                  labels = c(" ","E"), label_size = 8 )
+fig1 <- plot_grid(a, bcd, etoj, nrow = 3, rel_heights = c(0.8,0.7,2),
+                  labels = c("A"), label_size = 8)
 fig1

### 2020-05-06 analysis/figure-1.Rmd
bigger font

@@
 
-d1 <- sexbarplots(hypsex, 0, 7000) + labs(y = "DEGS", subtitle = "hypothalamus" )
-d2 <- sexbarplots(pitsex, 0, 7000) + labs(subtitle = "pituitary" ) + 
+d1 <- sexbarplots(hypsex, 0, 7100) + labs(y = "DEGS", subtitle = "hypothalamus" )
+d2 <- sexbarplots(pitsex, 0, 7100) + labs(subtitle = "pituitary" ) + 
   theme(axis.text.y = element_blank(), axis.line.y = element_blank(), axis.ticks = element_blank())
-d3 <- sexbarplots(gonsex, 0, 7000) + labs(subtitle = "gonads" ) +
+d3 <- sexbarplots(gonsex, 0, 7100) + labs(subtitle = "gonads" ) +
   theme(axis.text.y = element_blank(), axis.line.y = element_blank(), axis.ticks = element_blank())

### 2020-05-06 analysis/02_DESeq2.Rmd
now with tissue specific degs

@@
 
-```{r}
+```{r eval = F}
 
-for(i in "gonad"){
+## first or sex
+
+for(i in levels(colData$tissue)){
 
@@
 
+## then for tissue
+
+dds <- DESeqDataSetFromMatrix(countData = countData,
+                                colData = colData,
+                                design = ~ tissue )
+dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
+print(dds)
+print(dim(dds))
+dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
+
+vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
+write.csv(vsd, "../results/DEseq2/tissue/tissuevsds.csv")
+print(head(vsd))
+
+# save differential gene expression results
+HP <- createDEGdfsavestissue("pituitary", "hypothalamus")
+HG <- createDEGdfsavestissue("gonad", "hypothalamus")
+PG <- createDEGdfsavestissue("gonad", "pituitary")
+
 ```

### 2020-05-06 analysis/figure-1.Rmd
better fig 1

@@
 
+hyppit <- read_csv("../results/DEseq2/tissue/hypothalamus_pituitary_DEGs.csv") 
+hypgon <- read_csv("../results/DEseq2/tissue/hypothalamus_gonad_DEGs.csv")
+pitgon <- read_csv("../results/DEseq2/tissue/pituitary_gonad_DEGs.csv")
+
+hyppit$direction <- factor(hyppit$direction, levels = c("hypothalamus", "pituitary"))
+hypgon$direction <- factor(hypgon$direction, levels = c("hypothalamus",  "gonad"))
+pitgon$direction <- factor(pitgon$direction, levels = c( "pituitary", "gonad"))
 ```
@@
 
-b <- plottsneelipse(chartsne, chartsne$tissue, allcolors)   + labs(y = "tSNE 2 ", subtitle = " ")  
-c <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = "tSNE 2 ", subtitle = " ")
+b1 <- plottsneelipse(chartsne, chartsne$tissue, allcolors)   + labs(y = "tSNE 2 ", subtitle = " ")  
+b2 <- sexbarplots(hyppit, 0, 8100) + labs(subtitle = " ", y = "DEGs w/ + LFC") +
+  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
+  scale_x_discrete(labels=c("hypothalamus" = "hyp", "pituitary" = "pit" ))
+b3 <- sexbarplots(hypgon, 0, 8100) + labs(subtitle = " " ) +
+  theme(axis.text.y = element_blank(), axis.line.y = element_blank(), 
+        axis.ticks = element_blank(), axis.title.y = element_blank()) +
+    scale_x_discrete(labels=c("hypothalamus" = "hyp",  "gonad" = "gon"))
+b4 <- sexbarplots(pitgon, 0, 8100) + labs(subtitle = " " ) +
+  theme(axis.text.y = element_blank(), axis.line.y = element_blank(),  
+        axis.ticks = element_blank(), axis.title.y = element_blank()) +
+    scale_x_discrete(labels=c(  "pituitary" = "pit", "gonad" = "gon"))
 
-d1 <- sexbarplots(hypsex, 0, 7100) + labs(y = "DEGS", subtitle = "hypothalamus" )
-d2 <- sexbarplots(pitsex, 0, 7100) + labs(subtitle = "pituitary" ) + 
-  theme(axis.text.y = element_blank(), axis.line.y = element_blank(), axis.ticks = element_blank())
-d3 <- sexbarplots(gonsex, 0, 7100) + labs(subtitle = "gonads" ) +
-  theme(axis.text.y = element_blank(), axis.line.y = element_blank(), axis.ticks = element_blank())
+c1 <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = "tSNE 2 ", subtitle = " ")    
+c2 <- sexbarplots(hypsex, 0, 8100) + labs(y = "DEGs w/ + LFC", subtitle = "hypothalamus" ) +
+  theme(axis.text.y = element_blank(), axis.ticks = element_blank())
+c3 <- sexbarplots(pitsex, 0, 8100) + labs(subtitle = "pituitary" ) + 
+  theme(axis.text.y = element_blank(), axis.line.y = element_blank(),  
+        axis.ticks = element_blank(), axis.title.y = element_blank())
+c4 <- sexbarplots(gonsex, 0, 8100) + labs(subtitle = "gonads" ) +
+  theme(axis.text.y = element_blank(), axis.line.y = element_blank(),  
+        axis.ticks = element_blank(), axis.title.y = element_blank())
 
-e <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
+d1 <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
   labs(x = NULL, subtitle = "hypothalamus")  + 
@@
 
-g <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
+d3 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
   labs(x = NULL, subtitle = "pituitary") + 
@@
 
-i <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
+d5 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
    facet_wrap(~sex, scales = "free") +
   theme(strip.text = element_blank()) +
-  labs(x = "tSNE1 \n \n \n \n", subtitle = "gonads")
+  labs(x = "tSNE1 \n \n DEGs = differentially expressed genes \n + LFC = positive log fold change", 
+       subtitle = "gonads")
 
 # hyp
-f <- makebargraph("hypothalamus","DEGs", 0, 4300) + 
-  theme(axis.text.x = element_blank(), 
+d2 <- makebargraph("hypothalamus","DEGs w/ + LFC", 0, 4300) + 
+  theme(axis.ticks = element_blank(),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
         axis.title.x = element_blank())  +
-  labs(subtitle = "hypothalamus")
+  labs(subtitle = " ")
 # pit
-h <- makebargraph("pituitary","DEGs", 0, 4300)  +  
-  theme(axis.text.x = element_blank(), 
+d4 <- makebargraph("pituitary","DEGs w/ + LFC", 0, 4300)  +  
+  theme(axis.ticks = element_blank(),
+        axis.text.x = element_blank(), 
+        axis.text.y = element_blank(),
         axis.title.x = element_blank(), 
         strip.text.x = element_blank())  +
-  labs(subtitle = "pituitary")
+  labs(subtitle = " ")
 # gon
-j <- makebargraph("gonad","DEGs", 0, 4300) +  
-  theme(strip.text.x = element_blank()) +
+d6 <- makebargraph("gonad","DEGs w/ + LFC", 0, 4300) +  
+  theme(axis.ticks = element_blank(),
+        axis.text.y = element_blank(),
+        strip.text.x = element_blank()) +
   scale_x_discrete(labels = comparisonlabels)+
-  labs(subtitle = "gonads")
+  labs(subtitle = " ")
 
-bcd <- plot_grid(b,c,d1,d2,d3, nrow = 1, labels = c("B", "C", "D"), 
-                   label_size = 8, rel_widths = c(1.2,1.2,1.2,1,1))
+bc <- plot_grid(b1,b2,b3,b4, c1,c2,c3,c4, nrow = 1, rel_widths = c(1.5,1.1,0.9,0.9,1.5,1.1,0.9,0.9),
+                labels = c("B", "", "", "", "C"), label_size = 8)
 
-etoj <- plot_grid(e,f,g,h,i,j, ncol = 2, rel_heights = c(1,1,1.4), rel_widths = c(2.4,3.2),
-                  labels = c("E", "F", "G", "H", "I", "J"), label_size = 8)
+d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 2, rel_heights = c(1,0.9,1.3), rel_widths = c(1,2),
+                  labels = c("D"), label_size = 8)
 
-fig1 <- plot_grid(a, bcd, etoj, nrow = 3, rel_heights = c(0.8,0.7,2),
+fig1 <- plot_grid(a,bc, d, nrow = 3, rel_heights = c(0.7,0.7,2),
                   labels = c("A"), label_size = 8)

### 2020-05-06 analysis/figure-2.Rmd
new fig 2

@@
     geom_jitter(size = 0.25, aes(color = sex)) +
-    facet_wrap(~sex, scales = "free_y", nrow = 1) +
+    facet_wrap(~sex,  nrow = 1) +
     scale_fill_manual(values = allcolors) +
@@
 
-header <- png::readPNG("../figures/images/fig_fig2a.png")
-header <- ggdraw() +  draw_image(header, scale = 1)
+a <- png::readPNG("../figures/images/fig_fig2a.png")
+a <- ggdraw() +  draw_image(a, scale = 1)
 
-b <- plotcorrplot(hyp3, "females and males") + theme(legend.position = "none")  + labs(title = "Hypothalamus")
-g <- plotcorrplot(pit3, NULL) + theme(legend.position = "none") + labs(title = "Pituitary")
-l <- plotcorrplot(gon3, NULL) + theme(legend.position = "none")  + labs(title = "Gonad")
+b1 <- plotcorrplot(hyp3, "females and males") + theme(legend.position = "none")  + labs(y = "Hypothalamus", title = "Candidate gene correlations")
+b2 <- plotcorrplot(pit3, NULL) + theme(legend.position = "none") + labs(y = "Pituitary")
+b3 <- plotcorrplot(gon3, NULL) + theme(legend.position = "none")  + labs(y = "Gonad")
 
-c <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female") + labs(title = "")
-h <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = NULL)+ labs(title = "")
-m <- scattercorrelations(FG, FG$ESR1, FG$PGR, "#969696", "ESR1", "PGR")   + labs(subtitle = NULL)+ labs(title = "")
+c1 <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female", title =  "Top correlation")  
+c3 <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = NULL) 
+c5 <- scattercorrelations(FG, FG$ESR1, FG$PGR, "#969696", "ESR1", "PGR")   + labs(subtitle = NULL)
 
-d <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C")  + labs(subtitle = "male") + labs(title = "")
-i <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = NULL) + labs(title = "")
-n <- scattercorrelations(MG, MG$ESR1, MG$PGR, "#525252", "ESR1", "PGR")  + labs(subtitle = NULL) + labs(title = "")
+c2 <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C")  + labs(subtitle = "male", title = "", y = NULL)  
+c4 <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = NULL, y = NULL)  
+c6 <- scattercorrelations(MG, MG$ESR1, MG$PGR, "#525252", "ESR1", "PGR")  + labs(subtitle = NULL, y = NULL)
 
-d1 <- candidateboxplot("hypothalamus", c("DRD1"), sexlevels) + labs(x = NULL , title = " ")  + theme(axis.line.x = element_blank()) 
-d2 <- candidateboxplot("hypothalamus", c("HTR2C"), sexlevels) + labs(x = "Parental stage" ) + theme(strip.text = element_blank())
+d1 <- candidateboxplot("hypothalamus", c("DRD1"), sexlevels) + labs(x = NULL, title = "Temporal expression pattern" )  
+d2 <- candidateboxplot("hypothalamus", c("HTR2C"), sexlevels) + labs(x = NULL ) + theme(strip.text = element_blank())
+d3 <- candidateboxplot("pituitary", c("AVPR1A"), sexlevels) + labs(x = NULL )  + theme( strip.text = element_blank())
+d4 <- candidateboxplot("pituitary", c("CRHR1"), sexlevels) + labs(x = NULL  )  + theme(strip.text = element_blank())
+d5 <- candidateboxplot("gonad", c("ESR1"), sexlevels) + labs(x = NULL) + theme(strip.text = element_blank())
+d6 <- candidateboxplot("gonad", c("PGR"), sexlevels) + labs(x = NULL ) + theme(strip.text = element_blank())
 
-g1 <- candidateboxplot("pituitary", c("AVPR1A"), sexlevels) + labs(x = NULL, title = " " )  + theme( strip.text = element_blank(), axis.line.x = element_blank())
-g2 <- candidateboxplot("pituitary", c("CRHR1"), sexlevels) + labs(x = "Parental stage"  )  + theme(strip.text = element_blank())
+b <- plot_grid(b1,b2,b3, ncol = 1, rel_heights = c(1.1,1,1))
+c <- plot_grid(c1,c2,c3,c4,c5,c6, ncol = 2, rel_heights = c(1.1,1,1), rel_widths = c(1.1,1))
+d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.1,1.1, 1,1, 1,1))
 
-j1 <- candidateboxplot("gonad", c("ESR1"), sexlevels) + labs(x = NULL , title = " ") + theme(strip.text = element_blank(), axis.line.x = element_blank())
-j2 <- candidateboxplot("gonad", c("PGR"), sexlevels) + labs(x = "Parental stage"  ) + theme(strip.text = element_blank())
+bcd <- plot_grid(b,c,d, ncol = 3, rel_widths = c(1,1,1), labels = c("B", "C", "D"), label_size = 8)
 
-btn <- plot_grid(b,c,d, g,h,i,l,m,n, ncol = 3, labels = c("B", "C", " ","E", "F", " ", "H", "I"), label_size = 8, rel_widths = c(1.4,1,1), rel_heights = c(1.1,1,1))
-dgj <- plot_grid(d1,d2,g1,g2,j1,j2, ncol = 1, labels = c("D", " ", "G", " " , "J"), label_size = 8, rel_heights = c(1.3,1,1.2,1,1.2,1))
-btoj <- plot_grid(btn,dgj, ncol = 2 , rel_widths = c(1.75,1))
-
-plot_grid(header, btoj,  rel_heights = c(0.2,1), ncol = 1)
-
+fig2 <- plot_grid(a,bcd, ncol = 1, labels = c("A"), label_size = 8, rel_heights = c(0.1,1))
+fig2
 ```

### 2020-05-06 R/themes.R

@@ -1,7 +1,7 @@
 # custom themes
 
 theme_B3 <- function () { 
-  theme_classic(base_size = 8,
+  theme_classic(base_size = 7,
                 base_family = 'Helvetica') +
     theme(
       strip.background  = element_blank(),

### 2020-05-06 analysis/figure-2.Rmd
better PRL fig, now fig 3 not 4

@@
 c <- plot_grid(c1,c2,c3,c4,c5,c6, ncol = 2, rel_heights = c(1.1,1,1), rel_widths = c(1.1,1))
-d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.1,1.1, 1,1, 1,1))
+d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))
 

### 2020-05-06 analysis/figure-3.Rmd
better PRL fig, now fig 3 not 4

@@
 ---
-title: "Fig3: Correlations in gene expression"
-author: "Rayna M Harris"
-date: "3/26/2020"
+title: "Fig3"
 output: md_document
@@
 
-# Correlations in gene expression
+# Figure 4: All things prolactin
 
@@
 library(cowplot)
+library(ggimage)
+library(apaTables)
+library(factoextra)
 library(corrr)
 library(ggpubr)
+library(ggrepel)
+library(ggsignif)
 
@@
 
-knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
+knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
 ```
 
-## data wrangle
+## PCA data
 
-```{r data}
+```{r PCA}
+# pca
+pca1f <- subsetmakepca("pituitary", charlevels, "female")	
+pca2f <- makefvizdf("pituitary", charlevels, "female")
 
-# load `candidatevsd` with `source("../R/wrangledata.R")`
-head(candidatevsd)
+pca1m <- subsetmakepca("pituitary", charlevels, "male")	
+pca2m <- makefvizdf("pituitary", charlevels, "male")
 
-curleychampagnegenes <- curleychampagnegenes %>% pull(gene)
-
-# widen for correlations
-
-c1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
-c2 <- makecorrdf("female", "pituitary", curleychampagnegenes)  
-c3 <- makecorrdf("female", "gonad", curleychampagnegenes)
-
-c4 <- makecorrdf("male", "hypothalamus", curleychampagnegenes)  
-c5 <- makecorrdf("male", "pituitary", curleychampagnegenes)  
-c6 <- makecorrdf("male", "gonad", curleychampagnegenes)
-
-candidatevsdwide <- candidatevsd  %>%
-    pivot_wider(names_from = gene, values_from = counts) 
-head(candidatevsdwide)
-
-FH <- subsetcandidatevsdwide("female", "hypothalamus")
-FP <- subsetcandidatevsdwide("female", "pituitary")
-FG <- subsetcandidatevsdwide("female", "gonad")
-MH <- subsetcandidatevsdwide("male", "hypothalamus")
-MP <- subsetcandidatevsdwide("male", "pituitary")
-MG <- subsetcandidatevsdwide("male", "gonad") 
 ```
 
-### figure
+# subset PRL vsd
 
-```{r fig3, fig.width = 7.25, fig.height = 7.25}
+```{r}
+# `candidatevsd` loaded with `source("../R/wrangledata.R")`
 
-p1 <- plotcorrplot(c1, "FH") + theme(legend.position = "none", axis.text.x = element_blank()) + labs(subtitle = "Hypothalamus")
-p2 <- plotcorrplot(c2, "FP") + theme(legend.position = "none", axis.text.x = element_blank()) + 
-  labs(subtitle = "Pituitary") 
-p3 <- plotcorrplot(c3, "FG") + theme(legend.position = "none") + labs(subtitle = "Gonad")
+PRLpit <- candidatevsd %>% filter(tissue == "pituitary", gene == "PRL") %>%
+  mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
+  drop_na()
+PRLpit$hiloPRL <- factor(PRLpit$hiloPRL, levels = c("lo", "hi"))
 
-p4 <- plotcorrplot(c4, "MH") + theme(legend.position = "none", axis.text.y = element_blank(), axis.text.x = element_blank()) + labs(subtitle = " ")
-p5 <- plotcorrplot(c5, "MP") + theme(legend.position = "none", axis.text.y = element_blank(), axis.text.x = element_blank())  +
-  labs(subtitle = " ") 
-p6 <- plotcorrplot(c6, " ") + theme(legend.position = "none", axis.text.y = element_blank()) + labs(subtitle = " ")
+PRLpit %>%
+  group_by(sex) %>%
+  summarize(median = median(counts))
 
-a <- plot_grid(p1 + labs(title = "Female"),
-               p4 + labs(title = "Male"),
-               p2, p5, p3, p6, nrow = 3, rel_widths = c(1.2,1),
-               rel_heights = c(1,1,1.2))
+PRLpit %>%
+  group_by(sex, tissue, hiloPRL) %>%
+  summarize(n = n())
 
-scattercorrelations <- function(df, gene1, gene2, mylinecolor, myxlab, myylab){
-  p <- ggplot(df, aes(x = gene1, y = gene2)) +
-    labs(subtitle = " ", x = myxlab, y = myylab) + 
-    scale_color_manual(values = allcolors) +
-    geom_smooth(method = "lm",  color = mylinecolor) +
-    geom_point(aes(color = treatment))  +
-    theme_B3() +  
-    theme(axis.title = element_text(face = "italic"), 
-            legend.position = "none") +
-    stat_cor( size = 2)
-  return(p)
-}
-
-# "female" = "#969696", "male" = "#525252"
-
-p7 <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "Hypothalamus")
-p8 <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = "Pituitary")
-p9 <- scattercorrelations(FG, FG$ESR2, FG$CRHBP, "#969696", "ESR2", "CRHBP")   + labs(subtitle = "Gonad")
-
-p10 <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C") + theme(axis.title.y = element_blank())
-p11 <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1") + theme(axis.title.y = element_blank())
-p12 <- scattercorrelations(MG, MG$ESR2, MG$CRHBP, "#525252", "ESR2", "CRHBP")  + theme(axis.title.y = element_blank())
-
-b <- plot_grid(p7 + labs(title = "Female"), 
-               p10 + labs(title = "Male"), 
-               p8, p11, p9, p12, nrow = 3, rel_heights = c(1,1,1.2))
-
-ab <- plot_grid(a,b, rel_widths = c(1.5,1), labels = c("A", "B"), label_size = 8)
-
-p13 <- plotcorrplot(c1, "FH") + theme(legend.position = "bottom")
-p14 <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + 
-  theme(legend.position = "bottom",
-        legend.title = element_blank())
-
-legend1 <- p13 %>% get_legend()
-legend2 <- p14 %>% get_legend()
-
-legends <- plot_grid(legend1, legend2, rel_widths = c(1.5,1))
-
-plot_grid(ab, legends, nrow = 2, rel_heights = c(1,0.1))
+PRLpitF <- PRLpit %>% filter( sex == "female" )
+PRLpitM <- PRLpit %>% filter( sex == "male")
 
 ```
+
+## Internal versus external hyotheses
+
+```{r hypothesis}
+
+DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG2 <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
+head(allDEG2)
+
+allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)
+
+allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
+allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
+                                                      "eggs vs. chicks" = "eggs_chicks"))
+allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))
+
+PRLDEGs <- allDEG2 %>%
+  filter(tissue == "pituitary", comparison == "lo vs. hi PRL   ",
+         direction == "hi") %>%
+  arrange(desc(comparison))
+PRLDEGs
+
+## genes WGCNA prl module
+
+PRLgenes <- read_csv("../results/PRLmodule.csv") %>% pull(x)
+
+candidatevsd <- getcandidatevsd(PRLgenes, "pituitary", sexlevels)
+
+df <- candidatevsd %>%
+    pivot_wider(names_from = gene, values_from = counts) %>%
+    select(-sex,-tissue, -treatment, -samples) %>%
+    correlate() %>%
+  focus(PRL)  %>%
+  arrange(rowname)
+df
+
+```
+
+```{r fig3, fig.width=7.25, fig.height=7.25}
+
+a1 <- plotpc12(pca1f, pca2f, pca1f$treatment, allcolors, "Female pituitary") 
+a2 <- plotpc12(pca1m, pca2m, pca1m$treatment, allcolors, "Male pituitary")
+a <- plot_grid(a1,a2, labels = c("A"),label_size = 8)
+
+bcd1 <- png::readPNG("../figures/images/fig_fig3a.png")
+bcd1 <- ggdraw() +  draw_image(bcd1, scale = 1)
+
+b2 <- plotprolactin(PRLpitF, PRLpitF$counts, "PRL", "Female gene expression") 
+c2 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2200)   + theme(axis.text.x = element_blank() )
+d2 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2200) + theme(axis.text.x = element_blank())   
+b3 <- plotprolactin(PRLpitM, PRLpitM$counts, "PRL", "Male gene expression") + theme(axis.text.x = element_text())
+c3 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2200)  
+d3 <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2200)
+
+bcd23 <- plot_grid(b2,c2,d2,b3,c3,d3, nrow = 2, rel_widths = c(2,1,1), rel_heights = c(1,1.1))
+
+## top correlations
+e <- df %>%
+  arrange(desc(PRL)) %>% 
+  head(10)  %>%
+  mutate(PRLrounded = round(PRL, 2))%>% 
+  ggplot(aes(x = reorder(rowname, PRL), y = PRL)) +
+  geom_bar(stat = "identity") +
+  theme_B3() +
+  theme(axis.text.y = element_text(face = "italic")) +
+  labs(x = " ", y = "Correlation with PRL", subtitle = "Top 10 co-regulated genes") +
+  coord_flip() +
+  scale_y_continuous(expand = c(0, 0))
+
+# most sig DEGs
+
+f <-  PRLDEGs %>%
+  arrange(desc(lfc)) %>%
+  filter(sex == "female") %>% head(10) %>% 
+  ggplot(aes(x = reorder(gene, lfc), y = lfc)) + 
+  geom_bar(stat = "identity", fill = "#969696") +
+  theme_B3() +
+  theme(axis.text.y = element_text(face = "italic")) +
+  labs(x = " ", y = "Log-fold change, lo vs. hi PRL", subtitle = "Top 10 female DEGs") +
+  coord_flip() +
+  scale_y_continuous(expand = c(0, 0))
+
+g <-  PRLDEGs %>%
+  arrange(desc(lfc)) %>%
+  filter(sex == "male") %>% head(10) %>% 
+  ggplot(aes(x = reorder(gene, lfc), y = lfc)) + 
+  geom_bar(stat = "identity", fill = "#525252") +
+  theme_B3() +
+  theme(axis.text.y = element_text(face = "italic")) +
+  labs(x = " ", y = "Log-fold change, lo vs. hi PRL", subtitle = "Top 10 male DEGs") +
+  coord_flip() +
+  scale_y_continuous(expand = c(0, 0))
+
+efg <- plot_grid(e,f,g, nrow = 1, labels = c("E", "F", "G"), label_size = 8)
+
+plot_grid(a, bcd1, bcd23, efg, nrow = 4, rel_heights = c(1,1,2,1.2))
+
+```
+
+```{r savefiles}
+write.csv(PRLpit,"../results/PRLvsd.csv", row.names = F)
+
+```

### 2020-05-07 analysis/figure-3.Rmd
better fig 4

@@
 
+envDEGs <- allDEG2 %>%
+  filter(tissue == "pituitary", comparison == "eggs vs. chicks",
+         direction == "chicks") %>%
+  arrange(desc(comparison))
+envDEGs
+
 ## genes WGCNA prl module
@@
 
-```
-
-```{r fig3, fig.width=7.25, fig.height=7.25}
-
-a1 <- plotpc12(pca1f, pca2f, pca1f$treatment, allcolors, "Female pituitary") 
-a2 <- plotpc12(pca1m, pca2m, pca1m$treatment, allcolors, "Male pituitary")
-a <- plot_grid(a1,a2, labels = c("A"),label_size = 8)
-
-bcd1 <- png::readPNG("../figures/images/fig_fig3a.png")
-bcd1 <- ggdraw() +  draw_image(bcd1, scale = 1)
-
-b2 <- plotprolactin(PRLpitF, PRLpitF$counts, "PRL", "Female gene expression") 
-c2 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2200)   + theme(axis.text.x = element_blank() )
-d2 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2200) + theme(axis.text.x = element_blank())   
-b3 <- plotprolactin(PRLpitM, PRLpitM$counts, "PRL", "Male gene expression") + theme(axis.text.x = element_text())
-c3 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2200)  
-d3 <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2200)
-
-bcd23 <- plot_grid(b2,c2,d2,b3,c3,d3, nrow = 2, rel_widths = c(2,1,1), rel_heights = c(1,1.1))
-
 ## top correlations
-e <- df %>%
+c <- df %>%
   arrange(desc(PRL)) %>% 
@@
 
+```
+
+```{r fig3, fig.width=7.25, fig.height=7.25}
+
+a1 <- plotpc12(pca1f, pca2f, pca1f$treatment, allcolors, "Female pituitary gene expression", " ")   
+a2 <- plotprolactin(PRLpitF, PRLpitF$counts, "PRL", " ") + theme(axis.text.x = element_text())
+
+a3 <- plotpc12(pca1m, pca2m, pca1m$treatment, allcolors, "Male pituitary gene expression", NULL)
+a4 <- plotprolactin(PRLpitM, PRLpitM$counts, "PRL", " ") + theme(axis.text.x = element_text())
+
+a <- plot_grid(a1,a2,a3,a4, labels = c("A"), label_size = 8, rel_heights = c(1.1,1))
+
+b1 <- png::readPNG("../figures/images/fig_fig3b.png")
+b1 <- ggdraw() +  draw_image(b1, scale = 1)
+
+b2 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2200) + labs(subtitle = "female")  + labs(title = " ")  
+b3 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2200)  + labs(subtitle = "male")  + labs(title = " ") +
+  theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), 
+        axis.text.y = element_blank(), axis.line.y = element_blank())
+
+b4 <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2200) 
+b5 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2200)   + 
+  theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), 
+        axis.text.y = element_blank(), axis.line.y = element_blank())
+
+b25 <- plot_grid(b2,b3,b4,b5, nrow = 2, rel_widths = c(1.2,1), rel_heights = c(1.2,1))
+
 # most sig DEGs
 
-f <-  PRLDEGs %>%
-  arrange(desc(lfc)) %>%
-  filter(sex == "female") %>% head(10) %>% 
-  ggplot(aes(x = reorder(gene, lfc), y = lfc)) + 
-  geom_bar(stat = "identity", fill = "#969696") +
-  theme_B3() +
-  theme(axis.text.y = element_text(face = "italic")) +
-  labs(x = " ", y = "Log-fold change, lo vs. hi PRL", subtitle = "Top 10 female DEGs") +
-  coord_flip() +
-  scale_y_continuous(expand = c(0, 0))
+c1 <-  plottopDEGs(envDEGs, "female", "#3A80B9", "LFC eggs vs. chicks", "Top 10 female DEGs") + labs(title = " ")
+c2 <-  plottopDEGs(PRLDEGs, "male", "#3A80B9", "LFC eggs vs. chicks", "Top 10 male DEGs") + labs(title = " ")
+c3 <-  plottopDEGs(PRLDEGs, "female", "#19757A", "LFC lo vs. hi PRL", NULL)
+c4 <-  plottopDEGs(PRLDEGs, "male", "#19757A", "LFC lo vs. hi PRL", NULL)
 
-g <-  PRLDEGs %>%
-  arrange(desc(lfc)) %>%
-  filter(sex == "male") %>% head(10) %>% 
-  ggplot(aes(x = reorder(gene, lfc), y = lfc)) + 
-  geom_bar(stat = "identity", fill = "#525252") +
-  theme_B3() +
-  theme(axis.text.y = element_text(face = "italic")) +
-  labs(x = " ", y = "Log-fold change, lo vs. hi PRL", subtitle = "Top 10 male DEGs") +
-  coord_flip() +
-  scale_y_continuous(expand = c(0, 0))
+b <- plot_grid(b1,b25,nrow = 1, rel_widths = c(1.2,1))
 
-efg <- plot_grid(e,f,g, nrow = 1, labels = c("E", "F", "G"), label_size = 8)
+c <- plot_grid(c1,c2,c3,c4, align = "v" , rel_heights = c(1.2,1))
 
-plot_grid(a, bcd1, bcd23, efg, nrow = 4, rel_heights = c(1,1,2,1.2))
+bc <- plot_grid(b,c, labels = c("B","C"), label_size = 8, rel_widths = c(1.2,1))
 
+fig3 <- plot_grid(a, bc, nrow = 2, rel_heights = c(1,1))
+fig3
+
+#pdf(file="../figures/fig3.pdf", width=7.25, height=7.25)
+#plot(fig3)
+#dev.off()
 ```

### 2020-05-07 analysis/figure-2.Rmd
now with correaltions WGCNA and DESEq in Fig 3

@@
 
-### Candidate gene functions
-
-```{r}
-candidateboxplot <- function(whichtissue, whichgenes, whichsex){
-
-  p <- candidatevsd %>%
-    filter(tissue %in% whichtissue,
-          gene %in% whichgenes,
-          sex %in% whichsex) %>%
-    mutate(treatment = factor(treatment, levels = charlevels)) %>%
-    ggplot(aes(x = treatment, y = counts)) +
-    geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
-    geom_jitter(size = 0.25, aes(color = sex)) +
-    facet_wrap(~sex,  nrow = 1) +
-    scale_fill_manual(values = allcolors) +
-    scale_color_manual(values = allcolors) +
-    theme_B3() + 
-    theme(legend.position = "none",
-          axis.title.y = element_text(face = "italic"),
-          axis.text.x = element_blank(),
-          axis.ticks = element_blank()) +
-    labs(y = whichgenes,
-         x = NULL) +
-    geom_signif(comparisons = list(c( "control", "bldg"),
-                                 c( "bldg", "lay"),
-                                 c( "lay", "inc.d3"),
-                                 c("inc.d3", "inc.d9"),
-                                 c( "inc.d9", "inc.d17"),
-                                 c( "inc.d17", "hatch"),
-                                 c("hatch", "n5"),
-                                 c( "n5", "n9")),  
-                map_signif_level=TRUE,
-                textsize = 1.5, family = 'Helvetica',
-                vjust = 1.5, size = 0)
-
-  return(p)
-}
-
-scattercorrelations <- function(df, gene1, gene2, mylinecolor, myxlab, myylab){
-  p <- ggplot(df, aes(x = gene1, y = gene2)) +
-    labs(subtitle = " ", x = myxlab, y = myylab) + 
-    scale_color_manual(values = allcolors) +
-    geom_smooth(method = "glm",  color = mylinecolor) +
-    geom_point(aes(color = treatment))  +
-    theme_B3() +  
-    theme(axis.title = element_text(face = "italic"), 
-            legend.position = "none") +
-    stat_cor( size = 2)
-  return(p)
-}
-
-```
+## Figure
 

### 2020-05-07 analysis/figure-3.Rmd
now with correaltions WGCNA and DESEq in Fig 3

@@
 
+pca1 <- subsetmakepca("pituitary", charlevels, sexlevels)	
+pca2 <- makefvizdf("pituitary", charlevels, sexlevels)
+
 ```
 
-# subset PRL vsd
+# Hi Lo PRL and WGCNA PRL
 
@@
 PRLpitM <- PRLpit %>% filter( sex == "male")
+
+```
+
+### WGCNA prolactin module
+
+```{r}
+## genes WGCNA prl module
+
+WGCNAgenes <- read_csv("../results/PRLmodule.csv") %>% pull(x)
+
+WGCNAvsd <- getcandidatevsd(WGCNAgenes, "pituitary", sexlevels)
 
@@
          direction == "hi") %>%
-  arrange(desc(comparison))
+  arrange(desc(logpadj))
 PRLDEGs
@@
          direction == "chicks") %>%
-  arrange(desc(comparison))
+  arrange(desc(logpadj))
 envDEGs
 
-## genes WGCNA prl module
+hilogenes <- PRLDEGs %>% head(20) %>% distinct(gene) %>% pull(gene)
+eggchickgenes <- envDEGs %>% head(20) %>% distinct(gene) %>% pull(gene)
 
-PRLgenes <- read_csv("../results/PRLmodule.csv") %>% pull(x)
+hypothesisDEGs <- c(hilogenes, eggchickgenes, WGCNAgenes) %>% unique()
 
-candidatevsd <- getcandidatevsd(PRLgenes, "pituitary", sexlevels)
+candidatevsd <- getcandidatevsd(hypothesisDEGs, "pituitary", sexlevels)
 
-df <- candidatevsd %>%
-    pivot_wider(names_from = gene, values_from = counts) %>%
-    select(-sex,-tissue, -treatment, -samples) %>%
-    correlate() %>%
-  focus(PRL)  %>%
-  arrange(rowname)
-df
+#candidateboxplot("pituitary", c("PNOC"), "female") + labs(x = NULL )  + theme( strip.text = element_blank())
+#candidateboxplot("pituitary", c("KPNA2"), "male") + labs(x = NULL )  + theme( strip.text = element_blank())
 
-## top correlations
-c <- df %>%
-  arrange(desc(PRL)) %>% 
-  head(10)  %>%
-  mutate(PRLrounded = round(PRL, 2))%>% 
-  ggplot(aes(x = reorder(rowname, PRL), y = PRL)) +
-  geom_bar(stat = "identity") +
-  theme_B3() +
-  theme(axis.text.y = element_text(face = "italic")) +
-  labs(x = " ", y = "Correlation with PRL", subtitle = "Top 10 co-regulated genes") +
-  coord_flip() +
-  scale_y_continuous(expand = c(0, 0))
-
+pithilo <- makecorrdf("female", "pituitary", hilogenes)
+piteggchick <- makecorrdf("female", "pituitary", eggchickgenes)
+pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[1:20])
 ```
@@
 
-a1 <- plotpc12(pca1f, pca2f, pca1f$treatment, allcolors, "Female pituitary gene expression", " ")   
-a2 <- plotprolactin(PRLpitF, PRLpitF$counts, "PRL", " ") + theme(axis.text.x = element_text())
+a1 <- plotpc12(pca1, pca2, pca1$treatment, allcolors, "Pituitary gene expression", " ")   
+a2 <- plotprolactin(PRLpit, PRLpit$counts, "PRL", " ") + theme(axis.text.x = element_text())
 
-a3 <- plotpc12(pca1m, pca2m, pca1m$treatment, allcolors, "Male pituitary gene expression", NULL)
-a4 <- plotprolactin(PRLpitM, PRLpitM$counts, "PRL", " ") + theme(axis.text.x = element_text())
-
-a <- plot_grid(a1,a2,a3,a4, labels = c("A"), label_size = 8, rel_heights = c(1.1,1))
+a <- plot_grid(a1,a2, labels = c("A"), label_size = 8)
 
@@
 
-fig3 <- plot_grid(a, bc, nrow = 2, rel_heights = c(1,1))
+d1 <-plotcorrplot(piteggchick, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 eggs vs. chicks DEGs")
+d2 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 lo vs. PRL DEGs")
+d3 <-plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes from WGCNA module with PRL")
+
+d <- plot_grid(d1,d2,d3, nrow = 1, align = "h", labels = c("D"), label_size = 8)
+
+fig3 <- plot_grid(a, bc, d, nrow = 3, rel_heights = c(1.2,2,1.8))
 fig3

### 2020-05-07 analysis/figure-3.Rmd
better fig 3, now with 3 correlation matreces

@@
 a1 <- plotpc12(pca1, pca2, pca1$treatment, allcolors, "Pituitary gene expression", " ")   
-a2 <- plotprolactin(PRLpit, PRLpit$counts, "PRL", " ") + theme(axis.text.x = element_text())
+a2 <- plotprolactin(PRLpit, PRLpit$counts, "PRL", " ") + theme(axis.text.x = element_text())  +
+  theme(legend.position = c(0.8,0.2)) + guides(fill = FALSE)
 
@@
 
-# most sig DEGs
+b6 <-  plottopDEGs(envDEGs, "female", "#3A80B9", "LFC eggs vs. chicks", "Top 10 female DEGs") + labs(title = " ")
+b7 <-  plottopDEGs(PRLDEGs, "male", "#3A80B9", "LFC eggs vs. chicks", "Top 10 male DEGs") + labs(title = " ")
+b8 <-  plottopDEGs(PRLDEGs, "female", "#19757A", "LFC lo vs. hi PRL", NULL)
+b9 <-  plottopDEGs(PRLDEGs, "male", "#19757A", "LFC lo vs. hi PRL", NULL)
 
-c1 <-  plottopDEGs(envDEGs, "female", "#3A80B9", "LFC eggs vs. chicks", "Top 10 female DEGs") + labs(title = " ")
-c2 <-  plottopDEGs(PRLDEGs, "male", "#3A80B9", "LFC eggs vs. chicks", "Top 10 male DEGs") + labs(title = " ")
-c3 <-  plottopDEGs(PRLDEGs, "female", "#19757A", "LFC lo vs. hi PRL", NULL)
-c4 <-  plottopDEGs(PRLDEGs, "male", "#19757A", "LFC lo vs. hi PRL", NULL)
+b15 <- plot_grid(b1,b25,nrow = 1, rel_widths = c(1.2,1))
+b69 <- plot_grid(b6,b7,b8,b9, align = "v" , rel_heights = c(1.2,1))
+b <- plot_grid(b15,b69, labels = c("B"), label_size = 8, rel_widths = c(1.2,1))
 
-b <- plot_grid(b1,b25,nrow = 1, rel_widths = c(1.2,1))
+c1 <- plotcorrplot(piteggchick, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 eggs vs. chicks DEGs")
+c2 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 lo vs. PRL DEGs") 
+d <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes from WGCNA module with PRL") + theme(legend.position = "right")
 
-c <- plot_grid(c1,c2,c3,c4, align = "v" , rel_heights = c(1.2,1))
+cd <- plot_grid(c1,c2,d, nrow = 1, align = "h", labels = c("C", "", "D"), label_size = 8, rel_widths = c(1,1,1.3))
 
-bc <- plot_grid(b,c, labels = c("B","C"), label_size = 8, rel_widths = c(1.2,1))
-
-d1 <-plotcorrplot(piteggchick, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 eggs vs. chicks DEGs")
-d2 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 lo vs. PRL DEGs")
-d3 <-plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes from WGCNA module with PRL")
-
-d <- plot_grid(d1,d2,d3, nrow = 1, align = "h", labels = c("D"), label_size = 8)
-
-fig3 <- plot_grid(a, bc, d, nrow = 3, rel_heights = c(1.2,2,1.8))
+fig3 <- plot_grid(a, b, cd, nrow = 3, rel_heights = c(1.2,2,1.8))
 fig3
 
-#pdf(file="../figures/fig3.pdf", width=7.25, height=7.25)
-#plot(fig3)
-#dev.off()
+pdf(file="../figures/fig3-1.pdf", width=7.25, height=7.25)
+plot(fig3)
+dev.off()
 ```

### 2020-05-07 analysis/figure-4.Rmd
working on new fig 4

--- 2020-04-26 analysis/figure-4.Rmd
@@
 ---
-title: "Fig4"
+title: "Fig5"
 output: md_document
@@
 
-# Figure 4: All things prolactin
-
-```{r setup}
+```{r setup, include=FALSE}
 library(tidyverse)
 library(cowplot)
-library(ggimage)
-library(apaTables)
-library(factoextra)
+library(Rtsne)
+library(ggsignif)
 
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
 ```
 
-## PCA data
+# Figure 1. Experimental design, tSNE analysis, PCA, and prolactin
 
-```{r PCA}
-# pca
-pca1f <- subsetmakepca("pituitary", charlevelsnocontrol, "female")	
-pca1m <- subsetmakepca("pituitary", charlevelsnocontrol, "male")
+```{r tSNE-df}
 
-pca2f <- makefvizdf("pituitary", charlevelsnocontrol, "female")	
-pca2m <- makefvizdf("pituitary", charlevelsnocontrol, "male")
+# prep for tsne
+
+hyptsne <- subsetmaketsne("hypothalamus", allmaniplevels, sexlevels)
+pittsne <- subsetmaketsne("pituitary", allmaniplevels, sexlevels)
+gontsne <- subsetmaketsne("gonads", allmaniplevels, sexlevels)
+
+addgroupings <- function(df){
+
+  df <- df %>% mutate(hiloPRL = fct_collapse(treatment, 
+                                  lo = c("m.inc.d8", "m.inc.d3", "m.inc.d9", "inc.d3", "inc.d9", "lay", "bldg"),
+                                  hi = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" ,  "m.n2", "extend", "n5")),
+                extint = fct_collapse(treatment, 
+                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
+                                  chicks = c( "m.inc.d8", "hatch", "extend", "n5"),
+                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")))
+  df$extint <- factor(df$extint, levels = levelsextint)
+  return(df)
+}
+
+hyptsne <- addgroupings(hyptsne)
+pittsne <- addgroupings(pittsne)
+gontsne <- addgroupings(gontsne)
 
@@
 
-## PRL hormone
+## variance stabilized gene expression  (vsd)
 
-```{r hormones}
+```{r vsd}
+vsd_path <- "../results/DEseq2/manip/"   # path to the data
+vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
+vsd_pathfiles <- paste0(vsd_path, vsd_files)
+vsd_files
 
-# hormones
-prolactin <- read_csv("../results/07_hormones.csv") %>%
-    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
-    droplevels() 
-prolactin$treatment <- factor(prolactin$treatment, levels = alllevels)
+allvsd <- vsd_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name")  %>% 
+  dplyr::rename("gene" = "X1") %>% 
+  pivot_longer(cols = blk.s031.pu.d_female_gonad_prolong:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "samples", values_to = "counts") 
+allvsd <- as_tibble(allvsd)
+
+head(allvsd)
+
+# for plotting gene expression over time
+getcandidatevsdmanip <- function(whichgenes, whichtissue, whichsex){
+  candidates  <- allvsd %>%
+    filter(gene %in% whichgenes) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/manip/'), "[", 2)) %>%
+    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
+    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
+     dplyr::mutate(treatment = recode(treatment, "m.hatch" = "m.n2",
+                                                  "extend.hatch" = "m.n2",
+                                      "inc.prolong" = "prolong")) %>%
+
+    dplyr::mutate(day = fct_collapse(treatment,
+                                     "1" = c("inc.d3", "m.inc.d3" ),
+                                     "2" = c("inc.d9", "m.inc.d9", "m.inc.d8"  ),
+                                     "2" = c("inc.d17", "m.inc.d17", "prolong"  ),
+                                     "4" = c("hatch", "m.n2", "extend"  )
+                                     )) %>%
+    dplyr::mutate(day = as.numeric(day)) %>%
+    dplyr::mutate(manip = fct_collapse(treatment,
+                                       "reference" = charlevels,
+                                       "removal" = c("m.inc.d3", "m.inc.d9", "m.inc.d17","m.n2"),
+                                       "replacement" = c("m.inc.d8", "prolong", "extend"))) %>%
+    dplyr::mutate(manip = factor(manip, levels = c("reference", "removal", "replacement"))) %>%
+    dplyr::select(sex, tissue, treatment, day,  manip, gene, samples, counts) %>%
+    filter(tissue == whichtissue, sex %in% whichsex)  %>%
+    drop_na()
+  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
+  return(candidates)
+}
+
+hypvsd <- getcandidatevsdmanip(candidategenes, "hypothalamus", sexlevels)
+pitvsd <- getcandidatevsdmanip(candidategenes, "pituitary", sexlevels)
+gonvsd <- getcandidatevsdmanip(candidategenes, "gonad", sexlevels)
+candidatevsd <- rbind(hypvsd, pitvsd)
+candidatevsd <- rbind(candidatevsd, gonvsd)
+head(candidatevsd)
+
+hypvsd2 <- getcandidatevsdmanip(hypothesisDEGs, "hypothalamus", sexlevels)
+pitvsd2 <- getcandidatevsdmanip(hypothesisDEGs, "pituitary", sexlevels)
+gonvsd2 <- getcandidatevsdmanip(hypothesisDEGs, "gonad", sexlevels)
+candidatevsd2 <- rbind(hypvsd2, pitvsd2)
+candidatevsd2 <- rbind(candidatevsd2, gonvsd2)
+head(candidatevsd2) 
 ```
 
-# subset PRL vsd
+## DEGs
 
-```{r}
-# `candidatevsd` loaded with `source("../R/wrangledata.R")`
-
-PRLpit <- candidatevsd %>% filter(tissue == "pituitary", gene == "PRL")
-PRLpitF <- candidatevsd %>% filter(tissue == "pituitary", gene == "PRL", 
-                                   sex == "female" , treatment != "control")
-PRLpitM <- candidatevsd %>% filter(tissue == "pituitary", gene == "PRL", 
-                                   sex == "male", treatment != "control")
-```
-
-## Internal versus external hyotheses
-
-```{r hypothesis}
-
-PRLpit %>%
-  group_by(sex) %>%
-  summarize(median = median(counts))
-
-PRLvsd3 <- PRLpit %>%
-  mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
-  drop_na()
-PRLvsd3$hiloPRL <- factor(PRLvsd3$hiloPRL, levels = c("lo", "hi"))
-
-PRLvsd3 %>%
-  group_by(sex, tissue, hiloPRL) %>%
-  summarize(n = n())
-
-DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
+```{r allDEGs}
+DEG_path <- "../results/DEseq2/manip/"   # path to the data
 DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
@@
 
-allDEG2 <- DEG_pathfiles %>%
+allDEG <- DEG_pathfiles %>%
   setNames(nm = .) %>% 
   map_df(~read_csv(.x), .id = "file_name") %>% 
-  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/manip/'), "[", 2))  %>% 
   mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
@@
 dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
-head(allDEG2)
+head(allDEG)
 
-allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)
+allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
+allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
+allDEG$direction <- factor(allDEG$direction, levels = alllevels)
+head(allDEG)
 
-allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
-allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
-                                                      "eggs vs. chicks" = "eggs_chicks"))
-allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))
+candidateDEGS <- allDEG %>%
+  filter(gene %in% candidategenes) %>%
+  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
+         sex = recode(sex, "female" = "F", "male" = "M" ),
+         tissue = recode(tissue, 
+                         "hypothalamus" = "H",
+                         "pituitary" = "P", "gonad" = "G")) %>%
+  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
+  select(gene, res, comparison)  %>%
+  group_by(gene,  comparison) %>%
+  summarize(res = str_c(res, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = res) %>%
+  select(gene, contains("m.")) 
+candidateDEGS
 
-PRLDEGs <- allDEG2 %>%
-  filter(tissue == "pituitary", comparison == "lo vs. hi PRL   ", direction == "hi") %>%
-  arrange(desc(comparison))
-PRLDEGs
+hypothesisDEGStable <- allDEG %>%
+  filter(gene %in% hypothesisDEGs) %>%
+  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
+         sex = recode(sex, "female" = "F", "male" = "M" ),
+         tissue = recode(tissue, 
+                         "hypothalamus" = "H",
+                         "pituitary" = "P", "gonad" = "G")) %>%
+  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
+  select(gene, res, comparison)  %>%
+  group_by(gene,  comparison) %>%
+  summarize(res = str_c(res, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = res) %>%
+  select(gene, contains("m.")) 
+hypothesisDEGStable
 ```
 
-```{r fig4, fig.width=7.25, fig.height=7.25}
+## make figure
 
-expdesign <- png::readPNG("../figures/images/fig_fig4a.png")
-expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)
+```{r fig5, fig.width=7, fig.height=7}
 
-a <-plotcolorfulpcs(pca1f,pca1f$treatment, allcolors) + labs(subtitle = "Female pituitary") 
-b <- plotfriz(pca2f) + labs(subtitle = " ") 
-c <- plotprolactin(PRLpitF, PRLpitF$counts, "PRL", " ") + 
-  theme(legend.position = "none", 
-        axis.title.y = element_text(face = "italic"))  
-d <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2100)   
-e <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2100) + theme(axis.title.y = element_blank())
+a <- png::readPNG("../figures/images/fig_fig4a.png")
+a <- ggdraw() +  draw_image(a, scale = 1)
 
-f <- plotcolorfulpcs(pca1m, pca1m$treatment, allcolors) + labs(subtitle = "Male pituitary")
-g <- plotfriz(pca2m) + labs(subtitle = " ") 
-h <- plotprolactin(PRLpitM, PRLpitM$counts, "PRL", " ") + theme(legend.position = "none", 
-        axis.title.y = element_text(face = "italic"))  
-i <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2100)  
-j <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2100) + theme(axis.title.y = element_blank())
+b <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary ~ parental stage", x = NULL) + facet_wrap(~sex, scales = "free")
 
-d2m <- plot_grid(a,b,c,d,e,f,g,h,i,j, nrow = 2, rel_widths = c(1,1,2,0.8, 0.75), 
-                 labels = c("D", "E", "F", "G", "H", "I", "J", "K", "L", "M"), label_size = 8)
+c <- plottsneelipsev2(pittsne, pittsne$extint, allcolors ) + labs(subtitle = "pituitary ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")
 
-fig4 <- plot_grid(expdesign,d2m, ncol = 1, rel_heights = c(0.25,1))
-fig4
+d <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + labs(subtitle = "pituitary ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")
+
+bcd <- plot_grid(b,c,d, nrow = 1)
+
+b <- plotcandidatemanipquad(candidatevsd, "hypothalamus", c("DRD1")) 
+c <- plotcandidatemanipquad(candidatevsd2,"pituitary", c("PRL")) 
+d <- plotcandidatemanipquad(candidatevsd,"gonad", c("PGR"))
+
+plot_grid(a, b,c,d , ncol = 1, rel_heights = c(1,1,1,1))
+
 ```
 
-```{r savefiles}
-#write.csv(PRLvsd3,"../results/PRLvsd.csv", row.names = F)
+## supplementary figs
+
+```{r}
+
+b <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus ~ parental stage", x = NULL)  + facet_wrap(~sex, scales = "free")
+
+d <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads ~ parental stage")  + facet_wrap(~sex, scales = "free")
+
+e <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + labs(subtitle = "hypothalamus ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")
+
+g <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + labs(subtitle = "gonads ~ internal physiology" ) + facet_wrap(~sex, scales = "free")
+
+efg <- plot_grid(e,f,g, nrow = 3, labels = c("d"), label_size = 8, rel_heights = c(1,1,1))
+
+h <- plottsneelipsev2(hyptsne, hyptsne$extint, allcolors ) + labs(subtitle = "hypothalamus ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")
+
+j <- plottsneelipsev2(gontsne, gontsne$extint, allcolors ) + labs(subtitle = "gonads ~ external enviornment" ) + facet_wrap(~sex, scales = "free")
+
 ```

### 2020-05-08 analysis/figure-3.Rmd
working on new fig 4
now with temp fig 4

@@
 
-DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
-DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
-DEG_pathfiles <- paste0(DEG_path, DEG_files)
-#DEG_files
+#deg wrangling moved to datawrangling.R
 
-allDEG2 <- DEG_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name") %>% 
-  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
-  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
-  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
-  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
-  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
-  mutate(comparison = paste(down,up, sep = "_")) %>%
-  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
-  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
-dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
-head(allDEG2)
-
-allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)
-
-allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
-allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
-                                                      "eggs vs. chicks" = "eggs_chicks"))
-allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))
-
-PRLDEGs <- allDEG2 %>%
-  filter(tissue == "pituitary", comparison == "lo vs. hi PRL   ",
-         direction == "hi") %>%
-  arrange(desc(logpadj))
-PRLDEGs
-
-envDEGs <- allDEG2 %>%
-  filter(tissue == "pituitary", comparison == "eggs vs. chicks",
-         direction == "chicks") %>%
-  arrange(desc(logpadj))
-envDEGs
-
-hilogenes <- PRLDEGs %>% head(20) %>% distinct(gene) %>% pull(gene)
-eggchickgenes <- envDEGs %>% head(20) %>% distinct(gene) %>% pull(gene)
-
-hypothesisDEGs <- c(hilogenes, eggchickgenes, WGCNAgenes) %>% unique()
+hypothesisDEGs <- c(hypothesisDEGs, WGCNAgenes) %>% unique()
 

### 2020-05-08 analysis/figure-4.Rmd
now with temp fig 4

@@
 
-```{r fig5, fig.width=7, fig.height=7}
+```{r fig4, fig.width=7.25, fig.height=7.25}
 
@@
 
-b <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary ~ parental stage", x = NULL) + facet_wrap(~sex, scales = "free")
+b1 <- plotcandidatemanipquad(candidatevsd, "hypothalamus", c("DRD1")) 
+b2 <- plotcandidatemanipquad(candidatevsd2,"pituitary", c("PRL")) 
+b3 <- plotcandidatemanipquad(candidatevsd,"gonad", c("PGR"))
 
-c <- plottsneelipsev2(pittsne, pittsne$extint, allcolors ) + labs(subtitle = "pituitary ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")
+b <- plot_grid(b1,b2,b3, ncol = 1)
 
-d <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + labs(subtitle = "pituitary ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")
+c1 <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
+  labs(subtitle = "Hypothalamus ", x = NULL)  
+c2 <- plottsneelipsev2(hyptsne, hyptsne$extint, allcolors ) + 
+  labs(subtitle = " ", x = NULL) + 
+  theme(axis.text.y = element_blank(), axis.title.y = element_blank())
+c3 <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + 
+  labs(subtitle = " ", x = NULL) + 
+  theme(axis.text.y = element_blank(), axis.title.y = element_blank()) 
+c4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
+  labs(subtitle = "Pituitary ", x = NULL) 
+c5 <- plottsneelipsev2(pittsne, pittsne$extint, allcolors ) + 
+  labs(subtitle = " ", x = NULL) + 
+  theme(axis.text.y = element_blank(), axis.title.y = element_blank()) 
+c6 <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + 
+  labs(subtitle = " ", x = NULL) + 
+  theme(axis.text.y = element_blank(), axis.title.y = element_blank())
+c7 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
+  labs(subtitle = "Ggonads ")  
+c8 <- plottsneelipsev2(gontsne, gontsne$extint, allcolors ) + 
+  labs(subtitle = " " )    +
+  theme(axis.text.y = element_blank(), axis.title.y = element_blank())
+c9 <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + 
+  labs(subtitle = " " ) + 
+  theme(axis.text.y = element_blank(), axis.title.y = element_blank())
 
-bcd <- plot_grid(b,c,d, nrow = 1)
+c <- plot_grid(c1,c2,c3,c4,c5,c6,c7,c8,c9, nrow = 3)
 
-b <- plotcandidatemanipquad(candidatevsd, "hypothalamus", c("DRD1")) 
-c <- plotcandidatemanipquad(candidatevsd2,"pituitary", c("PRL")) 
-d <- plotcandidatemanipquad(candidatevsd,"gonad", c("PGR"))
+bc <- plot_grid(b, c, nrow = 1, labels = c("", "C"), label_size = 8)
 
-plot_grid(a, b,c,d , ncol = 1, rel_heights = c(1,1,1,1))
+plot_grid(a, bc, ncol = 1, rel_heights  = c(0.25,1),
+          labels = "AUTO", label_size = 8)
 
 ```
-
-## supplementary figs
-
-```{r}
-
-b <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus ~ parental stage", x = NULL)  + facet_wrap(~sex, scales = "free")
-
-d <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads ~ parental stage")  + facet_wrap(~sex, scales = "free")
-
-e <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + labs(subtitle = "hypothalamus ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")
-
-g <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + labs(subtitle = "gonads ~ internal physiology" ) + facet_wrap(~sex, scales = "free")
-
-efg <- plot_grid(e,f,g, nrow = 3, labels = c("d"), label_size = 8, rel_heights = c(1,1,1))
-
-h <- plottsneelipsev2(hyptsne, hyptsne$extint, allcolors ) + labs(subtitle = "hypothalamus ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")
-
-j <- plottsneelipsev2(gontsne, gontsne$extint, allcolors ) + labs(subtitle = "gonads ~ external enviornment" ) + facet_wrap(~sex, scales = "free")
-
-```

### 2020-05-08 R/functions.R

@@ -1000,15 +1000,14 @@ plottsneelipse <- function(tsnedf, pointcolor, whichcolors){
 
 plottsneelipsev2 <- function(tsnedf, pointcolor, whichcolors){
   p <- ggplot(tsnedf, aes(x = V1, y = V2)) +
-    stat_ellipse(linetype = 1, aes(color = pointcolor)) +
+   # stat_ellipse(linetype = 1, aes(color = pointcolor)) +
     geom_point(size = 0.5, aes(color = pointcolor)) +
     theme_B3() +
     labs(x = "tSNE 1", y = "tSNE 2",
          subtitle = " ") +
     scale_color_manual(values = whichcolors) +
     theme(legend.position = "none",
-          axis.text = element_blank(), axis.ticks = element_blank()) +
-    stat_ellipse(linetype = 1, aes(color = pointcolor)) 
+          axis.text = element_blank(), axis.ticks = element_blank()) 
   return(p)
 }
 
@@ -1246,11 +1245,12 @@ plotcandidatemanipquad <- function(df, whichtissue, whichgenes){
     geom_jitter(size = 0.25, aes(color = sex)) +
     theme_B3() +
     theme(legend.position = "none",
-          axis.text.x = element_text(angle = 45, hjust = 1),
+          axis.text.x = element_blank(),
+          axis.title.x = element_blank(),
           strip.text = element_text(face = "italic")) +
     scale_color_manual(values = allcolors) +
     scale_fill_manual(values = allcolors)  +
-    labs(title = whichtissue, subtitle = whichsex ,y = whichgenes)
+    labs(subtitle = paste(whichsex, whichtissue, sep = " "), y = whichgenes)
   
   return(p)
   
@@ -1290,7 +1290,7 @@ plotcandidatemanipquad <- function(df, whichtissue, whichgenes){
                 textsize = 3, family = 'Helvetica',
                 vjust = 1.5, size =0)
   
-  p <- plot_grid(p1,p2,p3,p4, nrow = 1, rel_widths = c(1,1))
+  p <- plot_grid(p1,p3,p2,p4, nrow = 2, rel_widths = c(1,1.2))
   
   
   return(p)

### 2020-05-08 R/wrangledata.R

@@ -100,3 +100,50 @@ pitvsd <- getcandidatevsd(candidategenes, "pituitary", sexlevels)
 gonvsd <- getcandidatevsd(candidategenes, "gonad", sexlevels)
 candidatevsd <- rbind(hypvsd, pitvsd)
 candidatevsd <- rbind(candidatevsd, gonvsd)
+
+
+
+### manip 
+
+
+DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG2 <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+  dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
+head(allDEG2)
+
+allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)
+
+allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
+allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
+                                                      "eggs vs. chicks" = "eggs_chicks"))
+allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))
+
+
+PRLDEGs <- allDEG2 %>%
[…]

### 2020-05-09 analysis/02_DESeq2.Rmd
better fig 1
fig 4 updates

@@
 
+str(HP)
 ```
+
+## FOR DEGs of all TREATMENT comparsions, char y manip
+
+```{r}
+countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
+geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
+colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
+
+colData$sextissue <- as.factor(paste(colData$sex, colData$tissue, sep = "_"))
+colData$treatment <- factor(colData$treatment, levels = alllevels)
+colData$tissue <- factor(colData$tissue, levels = tissuelevel)
+levels(colData$treatment)
+levels(colData$sex)
+levels(colData$sextissue)
+levels(colData$tissue)
+colData <- colData %>% drop_na()
+# 987 samples, 8 variables
+
+createDEGdftreatment <- function(up, down, mytissue){
+
+  res <- results(dds, contrast = c("treatment", up, down), 
+                 independentFiltering = T, alpha = 0.1)
+
+  DEGs <- data.frame(gene = row.names(res),
+                        padj = res$padj, 
+                        logpadj = -log10(res$padj),
+                        lfc = res$log2FoldChange,
+                        sextissue = mytissue)
+  DEGs <- na.omit(DEGs)
+  DEGs <- DEGs %>%
+    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
+                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
+                                                           yes = down, no = "NS"))) %>% 
+    dplyr::arrange(desc(lfc))
+
+  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up))
+
+  # write DEGsframe of only significant genes
+  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
+  print(str(DEGs))
+
+  partialfilename = paste("_", down, "_", up, sep = "")
+  myfilename = paste0("../results/DESeq2/treatment/", mytissue, partialfilename, "_DEGs.csv")
+
+  write.csv(DEGs, myfilename, row.names = F)
+  # return DEGs frome with all data, included NS genes
+  #print(head(DEGs))
+}
+
+for(i in levels(colData$sextissue)){
+
+  newcolData <- subsetcolData2(colData, i)
+
+  # save counts that match colData
+  savecols <- as.character(newcolData$V1) 
+  savecols <- as.vector(savecols)
+
+  newcountData <- countData %>% dplyr::select(one_of(savecols))
+
+  dds <- DESeqDataSetFromMatrix(countData = newcountData,
+                                colData = newcolData,
+                                design = ~ treatment )
+  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
+  print(dds)
+  print(dim(dds))
+  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
+
+  vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
+
+  myfilename = paste0("../results/DEseq2/treatment/", i, "_vsd.csv")
+  write.csv(vsd, myfilename)
+
+  #return(dds)
+  #return(vsd)
+  #print(head(vsd))
+
+  # save differential gene expression results
+
+  control.bldg <- createDEGdftreatment("bldg", "control", i)
+  bldg.lay <- createDEGdftreatment("lay", "bldg", i)
+  lay.inc.d3 <- createDEGdftreatment("inc.d3", "lay",  i) 
+  inc.d3.inc.d9 <- createDEGdftreatment("inc.d9", "inc.d3", i) 
+  inc.d9.inc.d17 <- createDEGdftreatment("inc.d17", "inc.d9", i)
+  inc.d17.hatch <- createDEGdftreatment( "hatch", "inc.d17", i) 
+  hatch.n5 <- createDEGdftreatment( "n5", "hatch",  i) 
+  n5.n9 <- createDEGdftreatment("n9", "n5",  i)
+
+  control.lay <- createDEGdftreatment("lay", "control", i)
+  control.inc.d3 <- createDEGdftreatment("inc.d3", "control",  i) 
+  control.inc.d9 <- createDEGdftreatment("inc.d9", "control", i) 
+  control.inc.d17 <- createDEGdftreatment("inc.d17", "control", i)
+  control.hatch <- createDEGdftreatment( "hatch", "control", i) 
+  control.n5 <- createDEGdftreatment( "n5", "control",  i) 
+  control.n9 <- createDEGdftreatment("n9", "control",  i)
+
+  control.extend <- createDEGdftreatment("extend", "control",  i) 
+  control.prolong <- createDEGdftreatment("prolong", "control",  i) 
+  control.early <- createDEGdftreatment("m.inc.d8", "control",  i)
+
+  control.m.inc.d3 <- createDEGdftreatment("m.inc.d3", "control",  i) 
+  control.m.inc.d9 <- createDEGdftreatment("m.inc.d9", "control",  i) 
+  control.m.inc.d17 <- createDEGdftreatment("m.inc.d17", "control",  i)
+  control.m.n2 <- createDEGdftreatment("m.n2", "control",  i)
+
+  bldg.inc.d3 <- createDEGdftreatment("inc.d3", "bldg",  i) 
+  bldg.inc.d9 <- createDEGdftreatment("inc.d9", "bldg", i) 
+  bldg.inc.d17 <- createDEGdftreatment("inc.d17", "bldg", i)
+  bldg.hatch <- createDEGdftreatment( "hatch", "bldg", i) 
+  bldg.n5 <- createDEGdftreatment( "n5", "bldg",  i) 
+  bldg.n9 <- createDEGdftreatment("n9", "bldg",  i)
+
+  bldg.extend <- createDEGdftreatment("extend", "bldg",  i) 
+  bldg.prolong <- createDEGdftreatment("prolong", "bldg",  i) 
+  bldg.early <- createDEGdftreatment("m.inc.d8", "bldg",  i)
+
+  bldg.m.inc.d3 <- createDEGdftreatment("m.inc.d3", "bldg",  i) 
+  bldg.m.inc.d9 <- createDEGdftreatment("m.inc.d9", "bldg",  i) 
+  bldg.m.inc.d17 <- createDEGdftreatment("m.inc.d17", "bldg",  i)
+  bldg.m.n2 <- createDEGdftreatment("m.n2", "bldg",  i)
+
+  hatch.m.n2 <- createDEGdftreatment("m.n2", "hatch",  i) 
+  hatch.prolong <- createDEGdftreatment("prolong", "hatch",  i) 
+  hatch.m.inc.d8 <- createDEGdftreatment("prolong", "hatch",  i)
+
+  inc.d17.m.n2 <- createDEGdftreatment("m.n2", "inc.d17",  i) 
+  inc.d17.prolong <- createDEGdftreatment("prolong", "inc.d17",  i)
+
+  prolong.extend <- createDEGdftreatment("extend", "prolong",  i) 
+  early.extend <- createDEGdftreatment("extend", "m.inc.d8",  i) 
+  early.prolong <- createDEGdftreatment("prolong", "m.inc.d8",  i)
+
+  m.inc.d9.m.inc.d3 <- createDEGdftreatment("m.inc.d9", "m.inc.d3",  i) 
+  m.inc.d17.m.inc.d9 <- createDEGdftreatment("m.inc.d17", "m.inc.d9",  i) 
+  m.inc.d17.m.inc.d3 <- createDEGdftreatment("m.inc.d17", "m.inc.d3",  i) 
+  m.n2.m.inc.d17 <- createDEGdftreatment("m.n2", "m.inc.d17",  i) 
+  m.n2.m.inc.d9 <- createDEGdftreatment("m.n2", "m.inc.d9",  i)
+  m.n2.m.inc.d3 <- createDEGdftreatment("m.n2", "m.inc.d3",  i)
+
+  m.inc.d3.inc.d3 <- createDEGdftreatment("m.inc.d3", "inc.d3",  i) 
+  m.inc.d9.inc.d9 <- createDEGdftreatment("m.inc.d9", "inc.d9",  i) 
+  m.inc.d17.inc.d17 <- createDEGdftreatment("m.inc.d17", "inc.d17",  i) 
+  m.n2.hatch <- createDEGdftreatment("m.n2", "hatch",  i)
+
+}
+
+```

### 2020-05-09 analysis/figure-1.Rmd
working on incorporating new desseq results

@@
 source("../R/functions.R")
-source("../R/wrangledata.R")
+#source("../R/wrangledata.R")
 
@@
 ```{r allDEGs}
-DEG_path <- "../results/DEseq2/"   # path to the data
+DEG_path <- "../results/DEseq2/treatment/"   # path to the data
 DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
@@
   map_df(~read_csv(.x), .id = "file_name") %>% 
-  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
   mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
@@
 
+# select only characterization plots
 allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
-allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
+allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevelschar)
 allDEG$direction <- factor(allDEG$direction, levels = charlevels)
 
+allDEG <- allDEG %>% drop_na()
+
+## other DEG stuff
 hypsex <- read_csv("../results/DEseq2/sex/hypothalamus_female_male_DEGs.csv")
@@
 pitgon$direction <- factor(pitgon$direction, levels = c( "pituitary", "gonad"))
+
+str(allDEG)
+unique(allDEG$comparison)
 ```
@@
 b1 <- plottsneelipse(chartsne, chartsne$tissue, allcolors)  + labs(y = "tSNE 2 ", subtitle = " ")
+
 b2 <- sexbarplots(hyppit, 0, 8100) + labs(subtitle = " ", y = "DEGs w/ + LFC") +
@@
 
+theme_B3 <- function () { 
+  theme_classic(base_size = 20,
+                base_family = 'Helvetica') +
+    theme(
+      strip.background  = element_blank(),
+      panel.grid.major  = element_blank(),  # remove major gridlines
+      panel.grid.minor  = element_blank()  # remove minor gridlines
+    )
+}
 # hyp
@@
   labs(subtitle = " ")
+
+d6
 

### 2020-05-09 analysis/figure-4.Rmd
fig 4 updates
small tests working. like these new plots

@@
 b1 <- plotcandidatemanipquad(candidatevsd, "hypothalamus", c("DRD1")) 
-b2 <- plotcandidatemanipquad(candidatevsd2,"pituitary", c("PRL")) 
+b2 <- plotcandidatemanipquad(candidatevsd,"pituitary", c("PRL")) 
 b3 <- plotcandidatemanipquad(candidatevsd,"gonad", c("PGR"))
@@
 
-plot_grid(a, bc, ncol = 1, rel_heights  = c(0.25,1),
+fig4 <- plot_grid(a, bc, ncol = 1, rel_heights  = c(0.25,1),
           labels = "AUTO", label_size = 8)
+fig4
+
+pdf(file="../figures/fig4-1.pdf", width=7.25, height=7.25)
+plot(fig4)
+dev.off()
+
+candidatevsd <- candidatevsd %>%
+  dplyr::mutate(treatment = fct_collapse(treatment,
+                                         "m.n2" = c("m.n2",  "m.hatch" ),
+                                         "prolong" = c("prolong", "inc.prolong") ,
+                                         "extend" = c("extend", "extend.hatch")
+                                         ))  %>%
+  dplyr::mutate(day = fct_collapse(treatment,
+                                    "0" = c("control", "bldg"),
+                                    "1" = c("lay"),
+                                     "3" = c("inc.d3", "m.inc.d3" ),
+                                     "9" = c("inc.d9", "m.inc.d9", "m.inc.d8"  ),
+                                     "17" = c("inc.d17", "m.inc.d17", "prolong"  ),
+                                     "18" = c("hatch", "m.n2" ),
+                                     "22" = c("n5",  "extend" ),
+                                     "26" = c("n9"  )
+                                     )) %>%
+   mutate(treatment = factor(treatment, levels = alllevels)) %>%
+  dplyr::mutate(day = factor(day, levels = c("0", "1", "3", "9", "17", "18", "22", "26")))
+
+candidatevsd %>%
+    filter(tissue == "hypothalamus", sex == "female") %>%
+    filter(gene %in% "DRD1") %>%
+    drop_na()  %>%
+    ggplot(aes(y =  counts, x = day, fill = treatment, color = sex)) +
+    geom_boxplot(outlier.shape = NA) +
+    geom_jitter(size = 0.25, aes(color = sex)) +
+    theme_B3() +
+    theme(legend.position = "bottom",
+          #axis.text.x = element_blank(),
+          #axis.title.x = element_blank(),
+          strip.text = element_text(face = "italic")) +
+    scale_color_manual(values = allcolors) +
+    scale_fill_manual(values = allcolors)  +
+    labs(subtitle = paste("female", "hypothalamus", sep = " "), y = "DRD1")
+
+candidatevsd %>%
+    filter(tissue == "pituitary", sex == "female") %>%
+    filter(gene %in% "PRL") %>%
+    drop_na()  %>%
+    ggplot(aes(y =  counts, x = day, fill = treatment, color = sex)) +
+    geom_boxplot(outlier.shape = NA) +
+    geom_jitter(size = 0.25, aes(color = sex)) +
+    theme_B3() +
+    theme(legend.position = "bottom",
+          #axis.text.x = element_blank(),
+          #axis.title.x = element_blank(),
+          strip.text = element_text(face = "italic")) +
+    scale_color_manual(values = allcolors) +
+    scale_fill_manual(values = allcolors)  +
+    labs(subtitle = paste("female", "pituitary", sep = " "), y = "PRL")
 

### 2020-05-11 analysis/03_datawrangling.Rmd
moving wrangle data files

---
title: "03_datawrangle"
output: md_document
---

```{r}
library(tidyverse)
library(readr)

source("../R/themes.R")
```

# data wrangle for multiple figures

## this is for tsne and pca

Note, pseudocounts is too big for storage on github :(

```{r}
pseudocounts <- read_csv("../results/01_pseudo.counts.csv")
head(pseudocounts[1:3])
pseudocounts <- as.data.frame(pseudocounts)
row.names(pseudocounts) <- pseudocounts$X1
pseudocounts$X1 <- NULL
# prep count data for all samples
pseudocounts <- as.data.frame(t(pseudocounts))
head(pseudocounts[1:3])

pseudocountsLong <- pseudocounts %>%
  mutate(samples = row.names(.)) %>%
  pivot_longer(-samples, names_to = "gene", values_to = "pseudocounts")
#pseudocountsLong$samples <- row.names(pseudocountsLong)
head(pseudocountsLong)
```

### prep col data for all samples

```{r}
colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
colData$treatment <- factor(colData$treatment, levels = alllevels)
colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
colData$tissue <- factor(colData$tissue, levels = tissuelevels)
row.names(colData) <- colData$V1

# check ready for analysis
# row.names(countData) == row.names(colData)
head(row.names(pseudocounts) == row.names(colData))
```

### candidate gene anlayses

```{r}
# all genes in this parental care dataset 
geneids <- read_csv("../metadata/00_geneinfo.csv") %>%
  select(-X1) %>%
  rename("NCBI" = "entrezid",
         "gene" = "Name")

# genes from Ch 17 Evolution of parental care
curleychampagnegenes <- c("AVP", "AVPR1A", "ADRA2A",    
                          "COMT", "CRH", "CRHBP", "CRHR1", "CRHR2", 
                          "DRD1", "DRD4","ESR1", "ESR2", #ERa ERbeta
                          "FOS", "HTR2C", "MEST", "NR3C1", #GR
                          "OPRM1", "OXT" , "PGR", "PRL", "PRLR",  "SLC6A4") #5HTT
curleychampagnegenes
curleychampagnegenes <- as.data.frame(curleychampagnegenes) %>%
  mutate(GO = "parentalcare", gene = curleychampagnegenes)  %>%
  select(GO, gene)

# candidate gens from parental care GO terms 
GO_path <- "../metadata/goterms/"   # path to the data
GO_files <- dir(GO_path, pattern = "*.txt") # get file names
GO_pathfiles <- paste0(GO_path, GO_files)

GOgenesLong <- GO_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_table(.x, col_types = cols(), col_names = FALSE), .id = "file_name") %>% 
  mutate(GO = sapply(strsplit(as.character(file_name),'../metadata/goterms/'), "[", 2)) %>% 
  mutate(GO = sapply(strsplit(as.character(GO),'.txt'), "[", 1)) %>% 
  mutate(gene = sapply(strsplit(as.character(X1), "[\\\\]|[^[:print:]]" ), "[", 2)) %>% 
  select(GO, gene)  %>%
  filter(gene != "Symbol") %>%
  distinct(GO,gene)  %>%
  mutate(gene = toupper(gene)) %>%
  rbind(., curleychampagnegenes) %>%
  arrange(gene) %>%
  left_join(., geneids, by = "gene") %>%
  arrange(gene) %>%
  drop_na() 
head(GOgenesLong)

parentalcaregenes <- GOgenesLong %>% 
  pivot_wider(
    names_from = GO,
    values_from = gene) %>%
  mutate(numGOs = 4 - rowSums(is.na(.))) %>%
  arrange(desc(numGOs)) %>%
  select(-numGOs) %>%
  left_join(geneids, by = c("geneid","NCBI"))
head(parentalcaregenes)

candidategenes <- parentalcaregenes %>% distinct(gene) %>% pull(gene)

# note: I can't find these genes in dataset: NOS1 or OXTR or FOX1B or HTR5A

```

### variance stabilized gene expression  (vsd)

```{r}
vsd_path <- "../results/DEseq2/treatment/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

## before pivoting, check names of df with
## head(names(allvsd)) and tail(names(allvsd))

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts") 
allvsd %>% select(-file_name) %>% head()
```

### vsd for all and candidate genes

```{r}
candidategenes <- GOgenesLong %>% distinct(gene) %>% pull(gene)

getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/treatment/'), "[", 2)) %>%
    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                  tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                  treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
    filter(tissue == whichtissue, sex %in% whichsex)  %>%
    drop_na()
  #candidates$treatment <- factor(candidates$treatment, levels = alllevels)
  return(candidates)
}

hypvsd <- getcandidatevsd(candidategenes, "hypothalamus", sexlevels)
pitvsd <- getcandidatevsd(candidategenes, "pituitary", sexlevels)
gonvsd <- getcandidatevsd(candidategenes, "gonad", sexlevels)
candidatevsd <- rbind(hypvsd, pitvsd)
candidatevsd <- rbind(candidatevsd, gonvsd)

head(candidatevsd)
```

### 2020-05-12 analysis/00_datawrangling.Rmd
slimming data prep

--- 2019-12-19 analysis/00_datawrangling.Rmd
@@
 # https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html
+
 library(tidyverse)
+
+source("../R/themes.R")
 ```
@@
 
-## countData and geneinfo
+## import Kallisto transcript data, make gene info file
 
@@
 kallistodata <- read.table("../results/kallistocounts.txt", 
-                        sep = ",", row.names = NULL)
+                        sep = ",", row.names = NULL) %>%
+  rename("NCBI" = "entrezid",
+         "gene" = "Name")
 head(kallistodata)
@@
 geneinfo <- kallistodata %>%
-  select(Name, geneid, entrezid)
+  select(gene, geneid, NCBI) 
 head(geneinfo)
+```
 
-# check for genes that have multple transcripts expressed
+## # check for genes that have multple transcripts expressed
+
+```{r}
 isoforms <- kallistodata %>%
-  group_by(Name) %>%
+  group_by(gene) %>%
   summarize(n = n()) %>%
-  arrange(desc(n), Name) %>%
-  filter(n > 1)
+  filter(n > 1) %>%
+  group_by(n) %>% 
+  summarize(genes = str_c(gene, collapse = ", ")) %>%
+  arrange(desc(n)) %>%
+  rename("n(isoforms)" = "n") %>% 
+  mutate("n(counts)" = str_count(genes, ",") + 1)
 head(isoforms)
 
-# for example, this gene has 
-geneinfo %>% filter(Name == "GRIN1")
+# for example, these gene have 2 and 3 isoforms
+geneinfo %>% filter(gene == "GRIN1")
+geneinfo %>% filter(gene == "CACNA1C")
+```
 
+## group data by gene
+
+```{r}
+head(kallistodata)
 # aggregate transcript counts to gene counts
 countData <- kallistodata %>% 
-  select(-row.names, -geneid, -entrezid) %>% 
-  rename(gene = Name) %>%
+  select(-row.names, -geneid, -NCBI) %>% 
   pivot_longer(-gene, names_to = "samples", values_to = "counts") %>%  
@@
   filter(gene != "")
-head(countData)
-
 countData <- as.data.frame(countData)
@@
 ```{r colData}
-colData <- read.table(file.path( "../metadata/kallistosamples.txt"), header = F, stringsAsFactors = F)
+colData <- read.table(file.path( "../metadata/kallistosamples.txt"),
+                      header = F, stringsAsFactors = F) %>%
+  # use strsplit to cut the filename into meaningful columns
+  mutate(bird = sapply(strsplit(V1,'\\_'), "[", 1),
+         sex = sapply(strsplit(V1,'\\_'), "[", 2),
+         tissue = sapply(strsplit(V1,'\\_'), "[", 3),
+         temp = sapply(strsplit(V1,'\\_'), "[", 4)) %>%
+  mutate(treatmenttemp = sapply(strsplit(temp,'\\.'), "[", 1),
+         NYNO = sapply(strsplit(temp,'\\.'), "[", 2)) %>%
 
-# use strsplit to cut the filename into meaningful columns
-colData$bird <- sapply(strsplit(as.character(colData$V1),'\\_'), "[", 1)
-colData$sex <- sapply(strsplit(as.character(colData$V1),'\\_'), "[", 2)
-colData$tissue <- sapply(strsplit(as.character(colData$V1),'\\_'), "[", 3)
-
-# to create columns with variable names in the filename 
-colData$temp <- sapply(strsplit(as.character(colData$V1),'\\_'), "[", 4)
-colData$treatmentTemp <- sapply(strsplit(as.character(colData$temp),'\\.'), "[", 1)
-colData$NYNO <- sapply(strsplit(as.character(colData$temp),'\\.'), "[", 2)
-
-# relabel incorrectly named variables 
-colData$treatment <- ifelse(grepl("extend-hatch", colData$treatmentTemp), "extend",
-                            ifelse(grepl("inc-prolong", colData$treatmentTemp), "prolong",
-                                   ifelse(grepl("m.hatch", colData$treatmentTemp), "m.n2",
-                                   colData$treatmentTemp)))
-
-# drop no-longer-needed cols
-colData <- colData %>% dplyr::select(-c(temp, treatmentTemp, NYNO))
-
-# replace dashes with periods and make rowname, (to make deseq happy)
-colData$bird <- gsub("-", ".", colData$bird)
-colData$treatment <- gsub("-", ".", colData$treatment)
-colData$V1 <- gsub("-", ".", colData$V1)
-row.names(colData) <- colData$V1
-
-# make variables factors and specify level
+  # rename variables
+  mutate(treatment = ifelse(grepl("extend-hatch", treatmenttemp), "extend",
+                            ifelse(grepl("inc-prolong", treatmenttemp), "prolong",
+                                   ifelse(grepl("m.hatch", treatmenttemp), "m.n2",
+                                          ifelse(grepl("m.inc.d8", treatmenttemp), "early",
+                                   treatmenttemp))))) %>%
+   select(-temp, -NYNO, -treatmenttemp ) %>%
+  # replace dashes with periods (to make deseq happy)
+  mutate(bird = gsub("-", ".", bird),
+         treatment = gsub("-", ".", treatment),
+         V1 = gsub("-", ".", V1))
+# specify levels that will be factors
 cols <- c("sex", "treatment", "tissue")
 colData[cols] <- lapply(colData[cols], factor) 
-colData$tissue <- factor(colData$tissue, 
-                         levels = c("hypothalamus", "pituitary", "gonad"))
+colData <- colData %>%
+  mutate(tissue = factor(tissue, levels = c("hypothalamus", "pituitary", "gonad"))) %>% 
+  mutate(tissue = fct_recode(tissue, "gonads" = "gonad")) %>%
+  mutate(group = paste(sex, tissue, treatment, sep = "."),
+         study = ifelse(grepl("m.|extend|prolong", treatment), 
+                        "manipulation", "charcterization")) %>%
+  mutate(study = factor(study, levels = c("charcterization", "manipulation")))
+colData <- as.data.frame(colData)
+row.names(colData) <- colData$V1
+head(colData)
+```
 
-# create a combinatorial variable
-colData$group <- paste(colData$sex, colData$tissue, colData$treatment, sep=".")
+## check that rownames and colnames match for DESeq
 
-# add variable to distinguish characterization study and manipulation
-colData$study <- ifelse(grepl("m.|extend|prolong", colData$treatment), "manipulation", "charcterization")
-
-colData$study <- factor(colData$study, 
-                         levels = c("charcterization", "manipulation"))
-
-## view colData
+```{r}
 str(colData)
-head(colData, 5)
-
-# check that rownames and colnames match for DESeq
 ncol(countData) == nrow(colData)
@@
 colData %>% select(sex, tissue, treatment, study)  %>%  summary()
-
+table(colData$sex, colData$treatment, colData$tissue)
 ```
 
-## save bird and sample and count data
+## save bird data
 
 ```{r}
-samples <- colData
-
 birds <- colData %>%
@@
 birds
-
-```
-
-## subset for manipulation study
-
-```{r manip}
-# create a new dfs for just the manipulation study
-colData_manipluation <- colData %>%
-  dplyr::filter(grepl('m.|extend|prolong', treatment)) %>%
-  droplevels()
-row.names(colData_manipluation) <- colData_manipluation$V1
-
-# print sample sizes
-colData_manipluation %>% select(sex,treatment, tissue)  %>%  summary()
-
-# create list with filenames and select countData columns matching the file names
-savecols <- as.character(colData_manipluation$V1) 
-savecols <- as.vector(savecols) 
-countData_manipluation <- countData %>% dplyr::select(one_of(savecols))
-
-# confirm that row and col lenghts are equal
-ncol(countData_manipluation) == nrow(colData_manipluation)
-```
-
-## subset for characterization study
-
-```{r characterization}
-# filter out data related to the manipulation study
-colData_characterization <- colData %>%
-  dplyr::filter(!grepl('m.|extend|prolong', treatment)) %>%
-  droplevels()
-row.names(colData_characterization) <- colData_characterization$V1
-
-# print sample sizes
-colData_characterization %>% select(sex,treatment, tissue)  %>%  summary()
-
-savecols <- as.character(colData_characterization$V1) 
-savecols <- as.vector(savecols) 
-countData_characterization <- countData %>% dplyr::select(one_of(savecols))
-
-# check that row and col lenghts are equal
-ncol(countData_characterization) == nrow(colData_characterization)
 ```
@@
 write.csv(countData, "../results/00_counts.csv")
-write.csv(samples, "../metadata/00_samples.csv")
+write.csv(isoforms, "../results/00_geneswithisoforms.csv", row.names = T)
+
+write.csv(colData, "../metadata/00_colData.csv")
 write.csv(birds, "../metadata/00_birds.csv")
-
-write.csv(colData_manipluation, "../metadata/00_colData_manipluation.csv", row.names = TRUE) 
-write.csv(countData_manipluation, "../results/00_countData_manipluation.csv", row.names = TRUE) 
-write.csv(colData_characterization, "../metadata/00_colData_characterization.csv", row.names = TRUE) 
-write.csv(countData_characterization, "../results/00_countData_characterization.csv", row.names = TRUE)
-
 write.csv(geneinfo, "../metadata/00_geneinfo.csv", row.names = TRUE)
 
-write.csv(isoforms, "../results/00_geneswithisoformss.csv", row.names = T)
 ```

### 2020-05-12 analysis/01_limma_all.Rmd
working on large limma files

@@
 
-knitr::opts_chunk$set(fig.path = '../figures/01_limma/',cache=TRUE)
+knitr::opts_chunk$set(fig.path = '../figures/01_limma/')
 ```
@@
 # import "colData" which contains sample information and "countData" which contains read counts
-colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
+colData <- read.csv("../metadata/00_colData.csv", header = T, row.names = 1)
 countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
-head(colData)
+
+#countData  <- head(countData, 1000) # suset for quick analysis
+
 geneinfo <- row.names(countData)
@@
 ```{r edgeR}
+head(colData$group)
 # create a large DGEList with 3 elements
@@
 
-write.csv(parentalobject$pseudo.counts, "../results/01_pseudo.counts.csv")
+write.csv(parentalobject$pseudo.counts, "../results/01_limma.csv")
 ```

### 2020-05-12 analysis/03_datawrangling.Rmd
slimming data prep
working on large limma files

@@
 
-Note, pseudocounts is too big for storage on github :(
+Note, limmacounts is too big for storage on github :(
 
 ```{r}
-pseudocounts <- read_csv("../results/01_pseudo.counts.csv")
-head(pseudocounts[1:3])
-pseudocounts <- as.data.frame(pseudocounts)
-row.names(pseudocounts) <- pseudocounts$X1
-pseudocounts$X1 <- NULL
+limmacounts <- read_csv("../results/01_limma.csv")
+head(limmacounts[1:3])
+limmacounts <- as.data.frame(limmacounts)
+row.names(limmacounts) <- limmacounts$X1
+limmacounts$X1 <- NULL
 # prep count data for all samples
-pseudocounts <- as.data.frame(t(pseudocounts))
-head(pseudocounts[1:3])
+limmacounts <- as.data.frame(t(limmacounts))
+head(limmacounts[1:3])
 
-pseudocountsLong <- pseudocounts %>%
+limmacountsLong <- limmacounts %>%
   mutate(samples = row.names(.)) %>%
-  pivot_longer(-samples, names_to = "gene", values_to = "pseudocounts")
-#pseudocountsLong$samples <- row.names(pseudocountsLong)
-head(pseudocountsLong)
+  pivot_longer(-samples, names_to = "gene", values_to = "limmacounts")
+#limmacountsLong$samples <- row.names(limmacountsLong)
+head(limmacountsLong)
 ```
@@
 ```{r}
-colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
-colData$treatment <- factor(colData$treatment, levels = alllevels)
-colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
-colData$tissue <- factor(colData$tissue, levels = tissuelevels)
+colData <- read.csv("../metadata/00_colData.csv", header = T, row.names = 1) %>%
+  mutate(treatment = factor(treatment, levels = alllevels),
+         tissue = factor(tissue, levels = tissuelevels))
+colData <- as.data.frame(colData)
 row.names(colData) <- colData$V1
@@
 # row.names(countData) == row.names(colData)
-head(row.names(pseudocounts) == row.names(colData))
+head(row.names(limmacounts) == row.names(colData))
 ```
@@
 # all genes in this parental care dataset 
-geneids <- read_csv("../metadata/00_geneinfo.csv") %>%
-  select(-X1) %>%
-  rename("NCBI" = "entrezid",
-         "gene" = "Name")
+geneids <- read_csv("../metadata/00_geneinfo.csv") %>% select(-X1)
 

### 2020-05-12 analysis/03_datawrangling.Rmd
removing old themes and functions

@@
 source("../R/themes.R")
-```
-
-# data wrangle for multiple figures
-
-## this is for tsne and pca
-
-Note, limmacounts is too big for storage on github :(
-
-```{r}
-limmacounts <- read_csv("../results/01_limma.csv")
-head(limmacounts[1:3])
-limmacounts <- as.data.frame(limmacounts)
-row.names(limmacounts) <- limmacounts$X1
-limmacounts$X1 <- NULL
-# prep count data for all samples
-limmacounts <- as.data.frame(t(limmacounts))
-head(limmacounts[1:3])
-
-limmacountsLong <- limmacounts %>%
-  mutate(samples = row.names(.)) %>%
-  pivot_longer(-samples, names_to = "gene", values_to = "limmacounts")
-#limmacountsLong$samples <- row.names(limmacountsLong)
-head(limmacountsLong)
-```
-
-### prep col data for all samples
-
-```{r}
-colData <- read.csv("../metadata/00_colData.csv", header = T, row.names = 1) %>%
-  mutate(treatment = factor(treatment, levels = alllevels),
-         tissue = factor(tissue, levels = tissuelevels))
-colData <- as.data.frame(colData)
-row.names(colData) <- colData$V1
-
-# check ready for analysis
-# row.names(countData) == row.names(colData)
-head(row.names(limmacounts) == row.names(colData))
 ```

### 2020-05-12 analysis/figure-1.Rmd
small tests working. like these new plots
removing old themes and functions

@@
 source("../R/functions.R")
-#source("../R/wrangledata.R")
 
@@
 #  Experimental design, tSNE analysis, and bar chars
+
+## import limma counts and sample info
+
+```{r data}
+countData <- read_csv("../results/01_limma.csv") %>%
+  column_to_rownames(var = "X1")
+countData <- as.data.frame(t(countData))
+head(countData[1:3])
+
+colData <- read_csv("../metadata/00_colData.csv") %>%
+  mutate(treatment = factor(treatment, levels = alllevels),
+         tissue = factor(tissue, levels = tissuelevels)) %>% 
+  column_to_rownames(var = "X1") 
+head(colData)
+# check ready for analysis
+# row.names(countData) == row.names(colData)
+head(row.names(countData) == row.names(colData))
+```
 
@@
 
-theme_B3 <- function () { 
-  theme_classic(base_size = 20,
-                base_family = 'Helvetica') +
-    theme(
-      strip.background  = element_blank(),
-      panel.grid.major  = element_blank(),  # remove major gridlines
-      panel.grid.minor  = element_blank()  # remove minor gridlines
-    )
-}
 # hyp
@@
         strip.text.x = element_blank()) +
-  scale_x_discrete(labels = comparisonlabels)+
+  scale_x_discrete(labels = comparisonlabelschar)+
   labs(subtitle = " ")
-
-d6
 

### 2020-05-12 analysis/figure-3.Rmd
fig3 working, back to datawrangling.R script

@@
 source("../R/functions.R")
-source("../R/wrangledata.R")
+
+source("../R/datawrangling.R")
 
 knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
+```
+
+## import limma counts and sample info
+
+```{r data}
+countData <- read_csv("../results/01_limma.csv") %>%
+  column_to_rownames(var = "X1")
+countData <- as.data.frame(t(countData))
+head(countData[1:3])
+
+colData <- read_csv("../metadata/00_colData.csv") %>%
+  mutate(treatment = factor(treatment, levels = alllevels),
+         tissue = factor(tissue, levels = tissuelevels)) %>% 
+  column_to_rownames(var = "X1") 
+head(colData)
+# check ready for analysis
+# row.names(countData) == row.names(colData)
+head(row.names(countData) == row.names(colData))
 ```

### 2020-05-12 R/datawrangling.R

@@ -1,21 +1,10 @@
----
-title: "03_datawrangle"
-output: md_document
----
-
-```{r}
 library(tidyverse)
 library(readr)
 
 source("../R/themes.R")
-```
-
-
-
 
 ### candidate gene anlayses
 
-```{r}
 # all genes in this parental care dataset 
 geneids <- read_csv("../metadata/00_geneinfo.csv") %>% select(-X1) 
 
@@ -66,12 +55,10 @@ candidategenes <- parentalcaregenes %>% distinct(gene) %>% pull(gene)
 
 # note: I can't find these genes in dataset: NOS1 or OXTR or FOX1B or HTR5A
 
-```
-
 
 ### variance stabilized gene expression  (vsd) 
 
-```{r}
+
 vsd_path <- "../results/DEseq2/treatment/"   # path to the data
 vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
 vsd_pathfiles <- paste0(vsd_path, vsd_files)
@@ -87,16 +74,10 @@ allvsd <- vsd_pathfiles %>%
   pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
                names_to = "samples", values_to = "counts") 
 allvsd %>% select(-file_name) %>% head()
-```
-
-
-
 
 
 ### vsd for all and candidate genes
 
-
-```{r}
 candidategenes <- GOgenesLong %>% distinct(gene) %>% pull(gene)
 
 getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
@@ -123,6 +104,50 @@ candidatevsd <- rbind(hypvsd, pitvsd)
 candidatevsd <- rbind(candidatevsd, gonvsd)
 
 head(candidatevsd)
-```
 
 
+### manip 	
+
+
+
+DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data	
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names	
+DEG_pathfiles <- paste0(DEG_path, DEG_files)	
+#DEG_files	
+
+allDEG2 <- DEG_pathfiles %>%	
+  setNames(nm = .) %>% 	
+  map_df(~read_csv(.x), .id = "file_name") %>% 	
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 	
[…]

### 2020-05-12 analysis/figure-3.Rmd
fig 3 working

@@
 
-PRLpit <- candidatevsd %>% filter(tissue == "pituitary", gene == "PRL") %>%
+PRLpit <- candidatevsd %>% 
+  filter(tissue == "pituitary", 
+         gene == "PRL",
+         treatment %in% charlevels) %>%
+  mutate(treatment = factor(treatment, levels = charlevels))  %>%
   mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
@@
 a1 <- plotpc12(pca1, pca2, pca1$treatment, allcolors, "Pituitary gene expression", " ")   
-a2 <- plotprolactin(PRLpit, PRLpit$counts, "PRL", " ") + theme(axis.text.x = element_text())  +
-  theme(legend.position = c(0.8,0.2)) + guides(fill = FALSE)
+a2 <- plotprolactin(PRLpit, PRLpit$counts, "PRL", " ") + 
+  theme(axis.text.x = element_text())  +
+  theme(legend.position = c(0.8,0.2)) + 
+  guides(fill = FALSE)
 

### 2020-05-12 analysis/figure-3.Rmd
attempting to save and read DEGs and vsd files

@@
 ```{r PCA}
-# pca
+# pca relies on count data from limma
+
 pca1f <- subsetmakepca("pituitary", charlevels, "female")	
@@
 pca2 <- makefvizdf("pituitary", charlevels, sexlevels)	
-
 ```
@@
 ```{r}
-# `candidatevsd` loaded with `source("../R/wrangledata.R")`
+candidatevsd <- read_csv("../results/03_candidatevsd.csv")
+head(candidatevsd)
 
@@
 
+hypothesisDEGs <- read_csv("../results/03_hypothesisDEGs.csv")
 hypothesisDEGs <- c(hypothesisDEGs, WGCNAgenes) %>% unique()

### 2020-05-12 analysis/03_DESeq2part2.Rmd
attempting to save and read DEGs and vsd files

--- 2020-05-12 analysis/03_datawrangling.Rmd
@@
 ---
-title: "03_datawrangle"
+title: "03_DESeq2part2"
 output: md_document
@@
 head(candidatevsd)
+
 ```
+
+## All DEGs (but currently only char DEGs)
+
+```{r}
+DEG_path <- "../results/DEseq2/treatment/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
+head(allDEG)
+
+# select only characterization plots
+allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
+allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevelschar)
+allDEG$direction <- factor(allDEG$direction, levels = charlevels)
+
+allDEG <- allDEG %>% drop_na()
+```
+
+### hypothesis testing
+
+```{r}
+
+DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data	
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names	
+DEG_pathfiles <- paste0(DEG_path, DEG_files)	
+#DEG_files
+
+allDEG2 <- DEG_pathfiles %>%	
+  setNames(nm = .) %>% 	
+  map_df(~read_csv(.x), .id = "file_name") %>% 	
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 	
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 	
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%	
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%	
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%	
+  mutate(comparison = paste(down,up, sep = "_")) %>%	
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%	
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%	
+  dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 	
+head(allDEG2)
+
+allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)
+
+allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))	
+allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",	
+                                                      "eggs vs. chicks" = "eggs_chicks"))	
+allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))
+
+PRLDEGs <- allDEG2 %>%	
+  filter(tissue == "pituitary", comparison == "lo vs. hi PRL   ",	
+         direction == "hi") %>%	
+  arrange(desc(logpadj))
+
+envDEGs <- allDEG2 %>%	
+  filter(tissue == "pituitary", comparison == "eggs vs. chicks",	
+         direction == "chicks") %>%	
+  arrange(desc(logpadj))
+
+hilogenes <- PRLDEGs %>% head(20) %>% distinct(gene) %>% pull(gene)	
+eggchickgenes <- envDEGs %>% head(20) %>% distinct(gene) %>% pull(gene)
+
+hypothesisDEGs <- c(hilogenes, eggchickgenes) %>% unique()
+```
+
+## save files
+
+```{r write}
+
+write.csv(allDEG, "../results/03_hypothesisDEGs.csv")
+write.csv(allDEG, "../results/03_allDEGs.csv")
+write.csv(allvsd, "../results/03_allvsd.csv")
+write.csv(candidatevsd, "../results/03_candidatevsd.csv")
+```

### 2020-05-12 analysis/figure-1.Rmd
added code to print pdf next time
attempting to save and read DEGs and vsd files

@@
 
+# uses count data from limma
+
 # prep for tsne
@@
 
-## DEGs
+## Treatment specific DEGs
 
-```{r allDEGs}
-DEG_path <- "../results/DEseq2/treatment/"   # path to the data
-DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
-DEG_pathfiles <- paste0(DEG_path, DEG_files)
-#DEG_files
+```{r}
+allDEG <- read_csv("../results/03_allDEG.csv")
+str(allDEG)
+unique(allDEG$comparison)
+```
 
-allDEG <- DEG_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name") %>% 
-  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
-  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
-  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
-  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
-  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
-  mutate(comparison = paste(down,up, sep = "_")) %>%
-  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
-  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
-dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
-head(allDEG)
+## Sex and Tissue -related DEGs
 
-# select only characterization plots
-allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
-allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevelschar)
-allDEG$direction <- factor(allDEG$direction, levels = charlevels)
-
-allDEG <- allDEG %>% drop_na()
+```{r DEGs}
 
@@
 
-str(allDEG)
-unique(allDEG$comparison)
 ```
@@
 fig1
+
+pdf(file="../figures/fig1-1.pdf", width=7.25, height=7.25)
+plot(fig1)
+dev.off()
 ```

### 2020-05-12 analysis/figure-2.Rmd
added code to print pdf next time
attempting to save and read DEGs and vsd files

@@
 source("../R/functions.R")
-source("../R/wrangledata.R")
 
@@
 # load `candidatevsd` with `source("../R/wrangledata.R")`
+candidatevsd <- read_csv("../results/03_candidatevsd.csv")
 head(candidatevsd)
@@
 fig2
+
+pdf(file="../figures/fig2-1.pdf", width=7.25, height=7.25)
+plot(fig2)
+dev.off()
 ```

### 2020-05-12 analysis/03_DESeq2part2.Rmd
now with 03 files saved

--- 2020-05-12 analysis/03_datawrangling.Rmd
@@
 source("../R/themes.R")
+
+knitr::opts_chunk$set(cache = TRUE, cache.lazy = FALSE)
 ```
@@
 
-```{r}
+```{r candidategenes}
 # all genes in this parental care dataset 
@@
 GOgenesLong <- GO_pathfiles %>%
-  setNames(nm = .) %>% 
+  stats::setNames(nm = .) %>% 
   map_df(~read_table(.x, col_types = cols(), col_names = FALSE), .id = "file_name") %>% 
@@
 
-```{r}
+```{r allvsd}
 vsd_path <- "../results/DEseq2/treatment/"   # path to the data
@@
 allvsd <- vsd_pathfiles %>%
-  setNames(nm = .) %>% 
+  stats::setNames(nm = .) %>% 
   map_df(~read_csv(.x), .id = "file_name")  %>% 
@@
 
-```{r}
+```{r candidatevsd}
 candidategenes <- GOgenesLong %>% distinct(gene) %>% pull(gene)
@@
 head(candidatevsd)
+
 ```
+
+## All DEGs (but currently only char DEGs)
+
+```{r allDEG}
+DEG_path <- "../results/DEseq2/treatment/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG <- DEG_pathfiles %>%
+  stats::setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
+head(allDEG)
+
+# select only characterization plots
+allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
+allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevelschar)
+allDEG$direction <- factor(allDEG$direction, levels = charlevels)
+
+allDEG <- allDEG %>% drop_na()
+```
+
+### hypothesis testing
+
+```{r hypothesisDEGS}
+
+DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data	
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names	
+DEG_pathfiles <- paste0(DEG_path, DEG_files)	
+#DEG_files
+
+allDEG2 <- DEG_pathfiles %>%	
+  stats::setNames(nm = .) %>% 	
+  map_df(~read_csv(.x), .id = "file_name") %>% 	
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 	
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 	
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%	
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%	
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%	
+  mutate(comparison = paste(down,up, sep = "_")) %>%	
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%	
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%	
+  dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 	
+head(allDEG2)
+
+allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)
+
+allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))	
+allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",	
+                                                      "eggs vs. chicks" = "eggs_chicks"))	
+allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))
+
+PRLDEGs <- allDEG2 %>%	
+  filter(tissue == "pituitary", comparison == "lo vs. hi PRL   ",	
+         direction == "hi") %>%	
+  arrange(desc(logpadj))
+
+envDEGs <- allDEG2 %>%	
+  filter(tissue == "pituitary", comparison == "eggs vs. chicks",	
+         direction == "chicks") %>%	
+  arrange(desc(logpadj))
+
+hilogenes <- PRLDEGs %>% head(20) %>% distinct(gene) %>% pull(gene)	
+eggchickgenes <- envDEGs %>% head(20) %>% distinct(gene) %>% pull(gene)
+
+hypothesisDEGs <- c(hilogenes, eggchickgenes) %>% unique()
+```
+
+## save files
+
+```{r write}
+write.csv(hypothesisDEGs, "../results/03_hypothesisDEGs.csv")
+write.csv(allDEG, "../results/03_allDEGs.csv")
+write.csv(candidatevsd, "../results/03_candidatevsd.csv")
+
+#library(crunch)
+#write.csv.gz(allvsd, "../results/03_allvsd.csv.gz", na = "")
+```

### 2020-05-12 analysis/03_DESeq2part2.Rmd

--- 2020-05-12 analysis/03_DESeq2part2.Rmd
@@
 
-```{r}
+```{r allDEG}
 DEG_path <- "../results/DEseq2/treatment/"   # path to the data
@@
 
-```{r}
-
+```{r hypothesisDEGS}
 DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data	
@@
 ```{r write}
-
-write.csv(allDEG, "../results/03_hypothesisDEGs.csv")
+write.csv(hypothesisDEGs, "../results/03_hypothesisDEGs.csv")
 write.csv(allDEG, "../results/03_allDEGs.csv")
-write.csv(allvsd, "../results/03_allvsd.csv")
 write.csv(candidatevsd, "../results/03_candidatevsd.csv")

### 2020-05-13 analysis/02_DESeq2.Rmd
not working...

@@
 register(MulticoreParam(6))
-library(caret)  # for sampling data
 library(cowplot)
@@
 _ Also, I don't run every chuck every time. When I want new analysese, I add them to the bottom and set `eval = T` for chunks I don't want to rerurn
-
-## Characterization
-
-```{r readfiles, eval=T}
-# import "colData" which contains sample information and "countData" which contains read counts
-countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
-geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
-head(geneinfo)
-
-# craete variable that will be critical for subset later on
-colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
-colData$sextissue <- as.factor(paste(colData$sex, colData$tissue, sep = "_"))
-colData$treatment <- factor(colData$treatment, levels = charlevels)
-colData$tissue <- factor(colData$tissue, levels = tissuelevel)
-levels(colData$treatment)
-levels(colData$sex)
-levels(colData$sextissue)
-levels(colData$tissue)
-```
-
-```{r deseq2, eval=F, include=F}
-for(i in levels(colData$sextissue)){
-
-  newcolData <- subsetcolData2(colData, i)
-
-  # save counts that match colData
-  savecols <- as.character(newcolData$V1) 
-  savecols <- as.vector(savecols)
-
-  newcountData <- countData %>% dplyr::select(one_of(savecols))
-
-  dds <- DESeqDataSetFromMatrix(countData = newcountData,
-                                colData = newcolData,
-                                design = ~ treatment )
-  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
-  print(dds)
-  print(dim(dds))
-  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
-
-  vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
-
-  myfilename = paste0("../results/DEseq2/", i, "_vsd.csv")
-  write.csv(vsd, myfilename)
-
-  #return(dds)
-  #return(vsd)
-  print(head(vsd))
-
-  # save differential gene expression results
-  control.bldg <- createDEGdfsave("bldg", "control", i)
-  bldg.lay <- createDEGdfsave("lay", "bldg", i)
-  lay.inc.d3 <- createDEGdfsave("inc.d3", "lay",  i) 
-  inc.d3.inc.d9 <- createDEGdfsave("inc.d9", "inc.d3", i) 
-  inc.d9.inc.d17 <- createDEGdfsave("inc.d17", "inc.d9", i)
-  inc.d17.hatch <- createDEGdfsave( "hatch", "inc.d17", i) 
-  hatch.n5 <- createDEGdfsave( "n5", "hatch",  i) 
-  n5.n9 <- createDEGdfsave("n9", "n5",  i) 
-}
-
-```
-
-## Manipulation
-
-### Data wrangle
-
-```{r manipulation1, eval = F, include = F}
-# import "colData" which contains sample information and "countData" which contains read counts
-countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
-geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
-head(geneinfo)
-
-levelsreplace <- c( "m.inc.d8" , "prolong" , "extend")
-levelsremoval <- c( "m.inc.d3" ,    "m.inc.d9" , "m.inc.d17" , "m.n2")
-controlsremovalreplace <- c( "inc.d3" ,    "inc.d9" , "inc.d17" , "hatch")
-
-manipulationsandcontrols <- c(controlsremoval, levelsreplace, levelsremoval)
-
-colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
-colData$sextissue <- as.factor(paste(colData$sex, colData$tissue, sep = "_"))
-colData$treatment <- factor(colData$treatment, levels = manipulationsandcontrols)
-colData$tissue <- factor(colData$tissue, levels = tissuelevel)
-levels(colData$treatment)
-levels(colData$sex)
-levels(colData$sextissue)
-levels(colData$tissue)
-colData <- colData %>% drop_na()
-```
-### Manipuation: comparisons to respective controls
-
-```{r manipulationcontrol, eval = F}
-
-createDEGdfsaveManip <- function(up, down, mytissue){
-
-  res <- results(dds, contrast = c("treatment", up, down), independentFiltering = T, alpha = 0.1)
-
-  DEGs <- data.frame(gene = row.names(res),
-                        padj = res$padj, 
-                        logpadj = -log10(res$padj),
-                        lfc = res$log2FoldChange,
-                        sextissue = mytissue)
-  DEGs <- na.omit(DEGs)
-  DEGs <- DEGs %>%
-    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
-                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
-                                                           yes = down, no = "NS"))) %>% 
-    dplyr::arrange(desc(lfc))
-
-  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up))
-
-  # write DEGsframe of only significant genes
-  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
-  print(str(DEGs))
-
-  partialfilename = paste("_", down, "_", up, sep = "")
-  myfilename = paste0("../results/DESeq2/manip/", mytissue, partialfilename, "_DEGs.csv")
-
-  write.csv(DEGs, myfilename, row.names = F)
-  # return DEGs frome with all data, included NS genes
-  #print(head(DEGs))
-}
-
-for(i in levels(colData$sextissue)){
-
-  newcolData <- subsetcolData2(colData, i)
-
-  # save counts that match colData
-  savecols <- as.character(newcolData$V1) 
-  savecols <- as.vector(savecols)
-
-  newcountData <- countData %>% dplyr::select(one_of(savecols))
-
-  dds <- DESeqDataSetFromMatrix(countData = newcountData,
-                                colData = newcolData,
-                                design = ~ treatment )
-  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
-  print(dds)
-  print(dim(dds))
-  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
-
-  vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
-
-  myfilename = paste0("../results/DEseq2/manip/", i, "_vsd.csv")
-  write.csv(vsd, myfilename)
-
-  #return(dds)
-  #return(vsd)
-  #print(head(vsd))
-
-  # save differential gene expression results
-  inc.d3.m.inc.d3 <- createDEGdfsaveManip("m.inc.d3", "inc.d3",  i) 
-  inc.d9.m.inc.d9 <- createDEGdfsaveManip("m.inc.d9", "inc.d9",  i) 
-  inc.d17.m.inc.d17 <- createDEGdfsaveManip("m.inc.d17", "inc.d17",  i) 
-  hatch.m.n2 <- createDEGdfsaveManip("m.n2", "hatch",  i) 
-  inc.d9.m.inc.d8 <- createDEGdfsaveManip("m.inc.d8", "inc.d9",  i) 
-  inc.d17.prolong <- createDEGdfsaveManip("prolong", "inc.d17",  i) 
-  hatch.extend <- createDEGdfsaveManip("extend", "hatch",  i) 
-}
-
-```
-
-## Manipulation: comparisons to other manipluations
-
-```{r manipulationwithin, eval = F}
-# import "colData" which contains sample information and "countData" which contains read counts
-countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
-geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
-head(geneinfo)
-
-levelsreplace <- c( "m.inc.d8" , "prolong" , "extend")
-levelsremoval <- c( "m.inc.d3" ,    "m.inc.d9" , "m.inc.d17" , "m.n2")
-controlsremovalreplace <- c( "inc.d3" ,    "inc.d9" , "inc.d17" , "hatch")
-
-manipulationsandcontrols <- c(controlsremoval, levelsreplace, levelsremoval)
-
-colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
-colData$sextissue <- as.factor(paste(colData$sex, colData$tissue, sep = "_"))
-colData$treatment <- factor(colData$treatment, levels = manipulationsandcontrols)
-colData$tissue <- factor(colData$tissue, levels = tissuelevel)
-levels(colData$treatment)
-levels(colData$sex)
-levels(colData$sextissue)
-levels(colData$tissue)
-colData <- colData %>% drop_na()
-
-createDEGdfsaveManip <- function(up, down, mytissue){
-
-  res <- results(dds, contrast = c("treatment", up, down), independentFiltering = T, alpha = 0.1)
-
-  DEGs <- data.frame(gene = row.names(res),
-                        padj = res$padj, 
-                        logpadj = -log10(res$padj),
-                        lfc = res$log2FoldChange,
-                        sextissue = mytissue)
-  DEGs <- na.omit(DEGs)
-  DEGs <- DEGs %>%
-    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
-                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
-                                                           yes = down, no = "NS"))) %>% 
-    dplyr::arrange(desc(lfc))
-
-  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up))
-
-  # write DEGsframe of only significant genes
-  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
-  print(str(DEGs))
-
-  partialfilename = paste("_", down, "_", up, sep = "")
-  myfilename = paste0("../results/DESeq2/manip/", mytissue, partialfilename, "_DEGs.csv")
-
-  write.csv(DEGs, myfilename, row.names = F)
-  # return DEGs frome with all data, included NS genes
-  #print(head(DEGs))
-}
-
-for(i in levels(colData$sextissue)){
-
-  newcolData <- subsetcolData2(colData, i)
-
-  # save counts that match colData
-  savecols <- as.character(newcolData$V1) 
-  savecols <- as.vector(savecols)
-
-  newcountData <- countData %>% dplyr::select(one_of(savecols))
-
-  dds <- DESeqDataSetFromMatrix(countData = newcountData,
-                                colData = newcolData,
-                                design = ~ treatment )
-  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
-  print(dds)
-  print(dim(dds))
-  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
-
-  vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
-
-  myfilename = paste0("../results/DEseq2/manip/", i, "_vsd.csv")
-  write.csv(vsd, myfilename)
-
-  #return(dds)
-  #return(vsd)
-  #print(head(vsd))
-
-  # save differential gene expression results
-  hatch.m.n2 <- createDEGdfsaveManip("m.n2", "hatch",  i) 
-  prolong.extend <- createDEGdfsaveManip("extend", "prolong",  i) 
-  early.extend <- createDEGdfsaveManip("extend", "m.inc.d8",  i) 
-  early.prolong <- createDEGdfsaveManip("prolong", "m.inc.d8",  i) 
-  m.inc.d9.m.inc.d3 <- createDEGdfsaveManip("m.inc.d9", "m.inc.d3",  i) 
-  m.inc.d17.m.inc.d9 <- createDEGdfsaveManip("m.inc.d17", "m.inc.d9",  i) 
-  m.inc.d17.m.inc.d3 <- createDEGdfsaveManip("m.inc.d17", "m.inc.d3",  i) 
-  m.n2.m.inc.d17 <- createDEGdfsaveManip("m.n2", "m.inc.d17",  i) 
-  m.n2.m.inc.d9 <- createDEGdfsaveManip("m.n2", "m.inc.d9",  i)
-  m.n2.m.inc.d3 <- createDEGdfsaveManip("m.n2", "m.inc.d3",  i)
-  prolong.m.inc.d17 <- createDEGdfsaveManip("prolong", "m.inc.d17",  i) 
-  m.inc.d8.m.inc.d9 <- createDEGdfsaveManip("m.inc.d8", "m.inc.d9",  i) 
-  extend.m.n2 <- createDEGdfsaveManip("extend", "m.n2",  i) 
-}
-
-```
 

### 2020-05-13 analysis/03_DESeq2part2.Rmd
not working...

@@
 
+## WGCNA
+
+```{r WGCNA}
+## genes WGCNA prl module
+
+WGCNAgenes <- read_csv("../results/04_PRLmodule.csv") %>% pull(x)
+WGCNAvsd <- getcandidatevsd(WGCNAgenes, "pituitary", sexlevels) 
+```
+
 ## save files
@@
 write.csv(candidatevsd, "../results/03_candidatevsd.csv")
+write.csv(WGCNAvsd, "../results/03_WGCNAvsd.csv")
+
 ```

### 2020-05-13 analysis/figure-3.Rmd
not working...

@@
 source("../R/functions.R")
-
-source("../R/datawrangling.R")
 
@@
 
-```{r}
+```{r PRLpit}
 candidatevsd <- read_csv("../results/03_candidatevsd.csv")
@@
 
-```{r}
+```{r WGCNA}
 ## genes WGCNA prl module
 
-WGCNAgenes <- read_csv("../results/PRLmodule.csv") %>% pull(x)
-
-WGCNAvsd <- getcandidatevsd(WGCNAgenes, "pituitary", sexlevels)
-
+WGCNAgenes <- read_csv("../results/04_PRLmodule.csv") %>% pull(x)
+WGCNAvsd <- read_csv("../results/03_WGCNAvsd.csv")
 ```
@@
 
-candidatevsd <- getcandidatevsd(hypothesisDEGs, "pituitary", sexlevels)
+candidatevsd <-read_csv("../results/03_candidatevsd.csv")
 

### 2020-05-13 analysis/03_datawrangling.Rmd
less stuff if 03_DESeq2

@@
 head(candidatevsd)
+
 ```
+
+## All DEGs (but currently only char DEGs)
+
+```{r allDEG}
+DEG_path <- "../results/DEseq2/treatment/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
+head(allDEG)
+
+# select only characterization plots
+allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
+allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevelschar)
+allDEG$direction <- factor(allDEG$direction, levels = charlevels)
+
+allDEG <- allDEG %>% drop_na()
+```
+
+## save files
+
+```{r write}
+write.csv(allDEG, "../results/03_allDEGs.csv")
+write.csv(candidatevsd, "../results/03_candidatevsd.csv")
+```

### 2020-05-13 analysis/02_DESeq2.Rmd
new hypothesis DEGs

@@
 
-## Notes
+# DESeq2
 
 - DESeq2 was not designed to run on 100+ samples. But, I really like it, so I do it anyways. Some of these commands take like 15 min to run using 6 cores. 
-_ Also, I don't run every chuck every time. When I want new analysese, I add them to the bottom and set `eval = T` for chunks I don't want to rerurn
+_ Also, I don't run every chuck every time. When I want new analysese, I add them to the bottom and set `eval = F` for chunks I don't want to rerurn
 
-## PRL-driven DESeq2
+```{r}
+countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
+geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
 
-```{r PRL, eval = F}
-PRLdata <- read_csv("../results/PRLvsd.csv")
-
-PRLsamples <- colData %>%
-  mutate(samples = V1)  %>%
-  select(samples) 
-PRLsamples
-
-# join with hi lo, and drop controls
-colDataPRL <- full_join(PRLdata, PRLsamples) %>% 
-  drop_na() %>%
-  select(samples, tissue, sex, treatment, hiloPRL) %>% 
-  filter(!treatment %in% c("control", "bldg"))
-
-# sample 100 only, better for deseq2
-colDataPRL <- colDataPRL[sample(1:nrow(colDataPRL), 100,
-   replace=FALSE),]
-
-colDataPRL <- as.data.frame(colDataPRL)
-row.names(colDataPRL) <- colDataPRL$samples
-
-savecols <- as.character(colDataPRL$samples) 
-savecols <- as.vector(savecols)
-
-newcountData <- countData %>% dplyr::select(one_of(savecols))
-
-colDataPRL %>%
-  group_by(sex, tissue, hiloPRL) %>%
-  summarize(n = n())
-
-dds <- DESeqDataSetFromMatrix(countData = newcountData,
-                                colData = colDataPRL,
-                                design = ~sex * hiloPRL )
-dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
-print(dds)
-print(dim(dds))
-dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
-
-vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
-
-ressex <- results( dds, contrast = c("sex", "male", "female") )
-reshiloPRL <- results( dds, contrast = c("hiloPRL", "hi", "lo") )
-hist( ressex$pvalue, breaks=20, col="grey" )
-hist( reshiloPRL$pvalue, breaks=20, col="grey" )
-
-restoDEGs <- function(res, up, down){
-  DEGs <- data.frame(gene = row.names(res),
-                        padj = res$padj, 
-                        logpadj = -log10(res$padj),
-                        lfc = res$log2FoldChange)
-  DEGs <- na.omit(DEGs)
-  DEGs <- DEGs %>%
-    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
-                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
-                                                           yes = down, no = "NS"))) %>% 
-    dplyr::arrange(gene)
-  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up)) 
-  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
-  print(head(DEGs))
-  return(DEGs)
-}
-
-sexdegs <- restoDEGs(ressex, "male", "female")
-PRLdegs <- restoDEGs(reshiloPRL, "hi", "lo")
-
-plotbarvolcano <- function(df, mylabels, mycolorname){
-  p1 <- ggplot(df, aes(x = lfc, y = logpadj, color = direction)) +
-    geom_point() +
-    scale_color_manual(values = allcolors, name = mycolorname) +
-    theme_B3() +
-    theme(legend.position = "top") +
-    labs(x = "Log fold change", y = "Log adj. p-value") +
-    xlim(-5,5)
-
-  p2 <- ggplot(df, aes(x = direction,  fill = direction)) +
-    geom_bar() +
-    theme_B3() +
-    scale_fill_manual(values = allcolors) +
-    theme_B3() +
-    theme(legend.position = "none") +
-    labs(x = mycolorname, y = "Total DEGs")
-
-  p <- plot_grid(p1,p2, labels  = mylabels, label_size = 12)
-  plot(p)
-}
-
-top <- plotbarvolcano(sexdegs, c("a","b"), "sex")
-bottom <- plotbarvolcano(PRLdegs, c("c","d"), "PRL")
-
-ab <- plot_grid(top,bottom, nrow = 2)
-
-sexPRLdesg <- full_join(sexdegs, PRLdegs, by = "gene")
-
-sexPRLdesg$direction.x <- as.character(sexPRLdesg$direction.x)
-
-sexPRLdesg <- full_join(sexdegs, PRLdegs, by = "gene") %>%
-  dplyr::select(gene, direction.x, direction.y) %>%
-  tidyr::replace_na(list(direction.x = "no sex effect", direction.y = "no PRL effect"))
-sexPRLdesg
-
-write.csv(sexPRLdesg, "../results/DEG-PRLsex.csv")
-
-sexPRLdesg2 <- sexPRLdesg %>%
-  mutate(higherin = paste(direction.x, direction.y, sep = "_")) %>%
-  mutate(direction = paste(higherin, "PRL", sep = "")) %>%
-  select(gene, direction) %>%
-  group_by(direction) %>%
-  summarize(DEGs = str_c(gene , collapse = " "))
-sexPRLdesg2
-write.csv(sexPRLdesg2, "../results/PRLsexinteraction.csv")
-
-c <- ggplot(sexPRLdesg, aes(x = direction.x, fill = direction.y)) +
-  geom_bar(position = "dodge") +
-  scale_fill_manual(values = allcolors, na.value= "#bdbdbd",
-                    labels= c("lo", "hi", "NS"),
-                    name = "PRL") +
-  scale_x_discrete(breaks = c("female", "male", NA),
-                   labels = c("female", "male", "NS"),
-                   name = "Sex")  +
-  labs(y = "Total DEGs") +
-  theme_B3() +
-  theme(legend.position = "top")
-
-plot_grid(ab, c, nrow = 1, rel_widths = c(2,1.2), labels = c(" ", "d"), label_size = 12)
-```
-
-## Hypothesis testing - external stimuli and/or prolactin
-
-```{r, eval = F}
-# label each bird with weather or not their pituitary had hi PRL or not
-# then make new column with nest, eggs, or chicks, drop controls
-
-PRLdata <- read_csv("../results/PRLvsd.csv") %>%
-  mutate(V1 = samples) %>%
-  mutate(bird = sapply(strsplit(as.character(samples),'\\_'), "[", 1)) %>%
-  select(bird,hiloPRL)
-
-# join with hi lo, and drop controls
-colDatahypothesis <- full_join(colData, PRLdata, by = "bird") %>% 
-  drop_na() %>%
-  select(V1, tissue, sex, sextissue, treatment, hiloPRL) %>% 
-  filter(!treatment %in% c("control", "bldg")) %>%
-  mutate(external = fct_collapse(treatment, eggs = c("lay", "inc.d3", "inc.d9", "inc.d17"),
-                                            chicks = c("hatch", "n5", "n9")))
-
-colDatahypothesis <- as.data.frame(colDatahypothesis)
-row.names(colDatahypothesis) <- colDatahypothesis$V1
-
-savecols <- as.character(colDatahypothesis$V1) 
-savecols <- as.vector(savecols)
-
-newcountData <- countData %>% dplyr::select(one_of(savecols))
-
-colDatahypothesis %>%
-  group_by(sex, tissue, hiloPRL, external) %>%
-  summarize(n = n())
-
-colDatahypothesis %>%
-  filter(tissue == "pituitary")  %>%
-  group_by(sex, tissue, hiloPRL, treatment) %>%
-  summarize(n = n())
-
-hypothesesistesting <- function(whichfactor, up, down, mytissue){
-
-  res <- results(dds, contrast = c(whichfactor, up, down), independentFiltering = T, alpha = 0.1)
-
-  DEGs <- data.frame(gene = row.names(res),
-                        padj = res$padj, 
-                        logpadj = -log10(res$padj),
-                        lfc = res$log2FoldChange,
-                        sextissue = mytissue)
-  DEGs <- na.omit(DEGs)
-  DEGs <- DEGs %>%
-    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
-                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
-                                                           yes = down, no = "NS"))) %>% 
-    dplyr::arrange(desc(lfc))
-
-  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up))
-
-  # write DEGsframe of only significant genes
-  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
-  print(str(DEGs))
-
-  partialfilename = paste("_", down, "_", up, sep = "")
-  myfilename = paste0("../results/DESeq2/hypothesis/", mytissue, partialfilename, "_DEGs.csv")
-
-  write.csv(DEGs, myfilename, row.names = F)
-  # return DEGs frome with all data, included NS genes
-  #print(head(DEGs))
-}
-
-for(i in levels(colDatahypothesis$sextissue)){
-
-  newcolData <- subsetcolData2(colDatahypothesis, i)
-
-  # save counts that match colData
-  savecols <- as.character(newcolData$V1) 
-  savecols <- as.vector(savecols)
-
-  newcountData <- countData %>% dplyr::select(one_of(savecols))
-
-  dds <- DESeqDataSetFromMatrix(countData = newcountData,
-                                colData = newcolData,
-                                design = ~ external + hiloPRL )
-  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
-  print(dds)
-  print(dim(dds))
-  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis
-
-  vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
-
-  myfilename = paste0("../results/DEseq2/hypothesis/", i, "_vsd.csv")
-  write.csv(vsd, myfilename)
-
-  # save differential gene expression results
-  # sexmalefemale <- hypothesesistesting("sex", "male", "female",  i) 
-  externalschickseggs <- hypothesesistesting("external", "chicks", "eggs",  i) 
-  hiloPRL <- hypothesesistesting("hiloPRL", "hi","lo",  i) 
-}
+colData <- read.csv("../metadata/00_colData.csv", header = T, row.names = 1) %>%
+  mutate(tissue = factor(tissue, levels = tissuelevels),
+         treatment <- factor(treatment, levels = alllevels)) %>%
+  mutate(sextissue = as.factor(paste(sex, tissue, sep = "_"))) %>%
+  column_to_rownames(var = "V1")
+colData <- colData %>% drop_na()
+# 987 samples, 8 variables
 ```
@@
 
-```{r eval = F}
+```{r sextissue, eval = F}
 
@@
 
-## FOR DEGs of all TREATMENT comparsions, char y manip
+## FOR DEGs of all TREATMENT comparisons, char y manip
 
-```{r}
-countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
-geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
-colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
-
-colData$sextissue <- as.factor(paste(colData$sex, colData$tissue, sep = "_"))
-colData$treatment <- factor(colData$treatment, levels = alllevels)
-colData$tissue <- factor(colData$tissue, levels = tissuelevel)
-levels(colData$treatment)
-levels(colData$sex)
-levels(colData$sextissue)
-levels(colData$tissue)
-colData <- colData %>% drop_na()
-# 987 samples, 8 variables
+```{r treatment, eval = F}
 

### 2020-05-13 analysis/04_DESeq2.Rmd
new hypothesis DEGs

---
title: "04_DESeq2"
author: "Rayna M Harris"
date: "5/12/2020"
output: md_document
---

```{r}
library(tidyverse)
library(DESeq2)

library(BiocParallel)
register(MulticoreParam(6))

source("../R/themes.R")
source("../R/functions.R")
```

### Hypothesis related data wrangle

- PRL WGCNA
- hi v lo PRL
- extral environment

```{r data}
countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)

## sample info
colData <- read_csv("../metadata/00_colData.csv") %>%
  mutate(tissue = factor(tissue, levels = tissuelevels),
         treatment = factor(treatment, levels = alllevels)) %>%
  mutate(sextissue = as.factor(paste(sex,tissue, sep = "_"))) %>%
  column_to_rownames(var = "X1") %>%
  mutate("samples" = V1) 
head(colData)

## calcualte hi lo PRL, add to col Data
PRLvsd <- read_csv("../results/03_candidatevsd.csv") %>%
  filter(gene == "PRL", tissue == "pituitary") 
PRLvsd %>%
  group_by(sex) %>%
  summarize(avecounts = mean(counts))
PRLvsd <- PRLvsd %>%
  mutate(hiloPRL = ifelse(counts >= 18.5, "hi", "lo")) %>%
  mutate(bird = sapply(strsplit(samples, '\\_'), "[", 1)) %>%
  select(bird, hiloPRL) 
colData <- left_join(colData, PRLvsd, by = "bird")   
unique(colData$hiloPRL)

## add external hypothesis

colData <- colData %>%
  mutate(external = fct_collapse(treatment,
                                 eggs = c("lay", "inc.d3", "inc.d9", "inc.d17", "prolong"),
                                 chicks = c("hatch", "n5", "n9", "extend" , "m.inc.d8" ),
                                 nest = c("bldg", "m.inc.d3",  "m.inc.d9",  "m.inc.d17",  "m.n2"),
                                 control = c("control"))) %>%
  drop_na()

```

## Hypothesis

```{r hypothesis}

createDEGdfhypothesis <- function(variable, up, down, mytissue){

  res <- results(dds, contrast = c(variable, up, down), 
                 independentFiltering = T, alpha = 0.1)

  DEGs <- data.frame(gene = row.names(res),
                        padj = res$padj, 
                        logpadj = -log10(res$padj),
                        lfc = res$log2FoldChange,
                        sextissue = mytissue)
  DEGs <- na.omit(DEGs)
  DEGs <- DEGs %>%
    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
                                                           yes = down, no = "NS"))) %>% 
    dplyr::arrange(desc(lfc))

  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up))

  # write DEGsframe of only significant genes
  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
  print(str(DEGs))

  partialfilename = paste("_", down, "_", up, sep = "")
  myfilename = paste0("../results/DESeq2/hypothesis/", mytissue, partialfilename, "_DEGs.csv")

  write.csv(DEGs, myfilename, row.names = F)
  # return DEGs frome with all data, included NS genes
  #print(head(DEGs))
}

for(i in levels(colData$sextissue)){

  newcolData <- subsetcolData2(colData, i)

  # save counts that match colData
  savecols <- as.character(newcolData$V1) 
  savecols <- as.vector(savecols)

  newcountData <- countData %>% dplyr::select(one_of(savecols))

  dds <- DESeqDataSetFromMatrix(countData = newcountData,
                                colData = newcolData,
                                design = ~ hiloPRL + external )
  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
  print(dds)
  print(dim(dds))
  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis

  vsd <- as.data.frame(assay(varianceStabilizingTransformation(dds, blind=FALSE)))

  myfilename = paste0("../results/DEseq2/hypothesis/", i, "_vsd.csv")
  write.csv(vsd, myfilename)

  createDEGdfhypothesis("hiloPRL", "hi", "lo", i)
  createDEGdfhypothesis("external", "chicks", "eggs", i)
  createDEGdfhypothesis("external", "chicks", "nest", i)
  createDEGdfhypothesis("external", "eggs", "nest", i)

}
```

### 2020-05-13 analysis/figure-3.Rmd
new hypothesis DEGs

@@
 
-# Hi Lo PRL and WGCNA PRL
-
-```{r PRLpit}
-candidatevsd <- read_csv("../results/03_candidatevsd.csv")
-head(candidatevsd)
-
-PRLpit <- candidatevsd %>% 
-  filter(tissue == "pituitary", 
-         gene == "PRL",
-         treatment %in% charlevels) %>%
-  mutate(treatment = factor(treatment, levels = charlevels))  %>%
-  mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
-  drop_na()
-PRLpit$hiloPRL <- factor(PRLpit$hiloPRL, levels = c("lo", "hi"))
-
-PRLpit %>%
-  group_by(sex) %>%
-  summarize(median = median(counts))
-
-PRLpit %>%
-  group_by(sex, tissue, hiloPRL) %>%
-  summarize(n = n())
-
-PRLpitF <- PRLpit %>% filter( sex == "female" )
-PRLpitM <- PRLpit %>% filter( sex == "male")
-
-```
-
-### WGCNA prolactin module
-
-```{r WGCNA}
-## genes WGCNA prl module
-
-WGCNAgenes <- read_csv("../results/04_PRLmodule.csv") %>% pull(x)
-WGCNAvsd <- read_csv("../results/03_WGCNAvsd.csv")
-```
-
-## Internal versus external hyotheses
-
-```{r hypothesis}
-
-#deg wrangling moved to datawrangling.R
-
-hypothesisDEGs <- read_csv("../results/03_hypothesisDEGs.csv")
-hypothesisDEGs <- c(hypothesisDEGs, WGCNAgenes) %>% unique()
-
-candidatevsd <-read_csv("../results/03_candidatevsd.csv")
-
-#candidateboxplot("pituitary", c("PNOC"), "female") + labs(x = NULL )  + theme( strip.text = element_blank())
-#candidateboxplot("pituitary", c("KPNA2"), "male") + labs(x = NULL )  + theme( strip.text = element_blank())
-
-pithilo <- makecorrdf("female", "pituitary", hilogenes)
-piteggchick <- makecorrdf("female", "pituitary", eggchickgenes)
-pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[1:20])
-```
-
 ```{r fig3, fig.width=7.25, fig.height=7.25}

### 2020-05-13 R/themes.R

@@ -21,10 +21,10 @@ charlevelsnocontrol <- c( "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch",
 
 # manipulation
 maniplevels <- c( "rm inc.d3" ,  "rm inc.d9" ,  "rm inc.d17" , "rm inc.d20" ,  "extend" , "prolong", "early")
-maniplevels1 <- c( "m.inc.d3" ,  "m.inc.d8" ,  "m.inc.d9" , "m.inc.d17" , "m.n2", "prolong" , "extend")
+maniplevels1 <- c( "m.inc.d3" ,  "early" ,  "m.inc.d9" , "m.inc.d17" , "m.n2", "prolong" , "extend")
 
 # mix
-levelstiming <- c( "m.inc.d8" , "prolong" , "extend")
+levelstiming <- c( "early" , "prolong" , "extend")
 levelsremoval <- c( "m.inc.d3" ,    "m.inc.d9" , "m.inc.d17" , "m.n2")
 controlstiming <- c( "inc.d9" , "hatch" , "inc.d17")
 controlsremoval <- c( "inc.d3" ,    "inc.d9" , "inc.d17" , "hatch")
@@ -32,7 +32,7 @@ allmaniplevels <- c(maniplevels1, controlsremoval, "n5", "lay", "bldg")
 
 # mix of characterization and manipulation, ordered by time
 alllevels <- c("control", "bldg", "lay", "inc.d3", "m.inc.d3" ,  
-               "inc.d9", "m.inc.d9" , "m.inc.d8" ,
+               "inc.d9", "m.inc.d9" , "early" ,
                "inc.d17",  "m.inc.d17","prolong" ,
                "hatch",  "m.n2",   "extend",
                "n5", "n9")
@@ -40,7 +40,7 @@ alllevels <- c("control", "bldg", "lay", "inc.d3", "m.inc.d3" ,
 # characteriation first, then manip by time
 alllevels2 <- c("control", "bldg", "lay", "inc.d3", "inc.d9", 
                 "inc.d17", "hatch", "n5", "n9",
-                "m.inc.d3" ,  "m.inc.d8" ,  "m.inc.d9" , 
+                "m.inc.d3" ,  "early" ,  "m.inc.d9" , 
                 "m.inc.d17" , "m.n2", "prolong" , "extend")
 
 # tissue
@@ -81,7 +81,7 @@ colorsmanip <- c("m.inc.d3" = "#CDCDCD",
[…]

### 2020-05-13 analysis/04_DESeq2.Rmd
need to rerun

@@
 
-### Hypothesis related data wrangle
+### Hypothesis-related data wrangle
 
@@
                                  eggs = c("lay", "inc.d3", "inc.d9", "inc.d17", "prolong"),
-                                 chicks = c("hatch", "n5", "n9", "extend" , "m.inc.d8" ),
+                                 chicks = c("hatch", "n5", "n9", "extend" , "early" ),
                                  nest = c("bldg", "m.inc.d3",  "m.inc.d9",  "m.inc.d17",  "m.n2"),

### 2020-05-13 analysis/03_datawrangling.Rmd
rename
now with WGCNA as step 5, wrangle as 6

@@
 ---
-title: "03_DESeq2part2"
+title: "03_wrangle"
 output: md_document
@@
 curleychampagnegenes <- as.data.frame(curleychampagnegenes) %>%
-  mutate(GO = "parentalcare", gene = curleychampagnegenes)  %>%
+  mutate(GO = "literature", gene = curleychampagnegenes)  %>%
   select(GO, gene)
@@
     names_from = GO,
-    values_from = gene) %>%
-  mutate(numGOs = 4 - rowSums(is.na(.))) %>%
-  arrange(desc(numGOs)) %>%
-  select(-numGOs) %>%
-  left_join(geneids, by = c("geneid","NCBI"))
+    values_from = gene,
+    values_fill = list(gene = "")) %>%
+  left_join(geneids, by = c("geneid","NCBI")) %>%
+  dplyr::rename("GO"= "parentalbehavior" )
 head(parentalcaregenes)
@@
 candidategenes <- parentalcaregenes %>% distinct(gene) %>% pull(gene)
-
 # note: I can't find these genes in dataset: NOS1 or OXTR or FOX1B or HTR5A
-
 ```
@@
 write.csv(candidatevsd, "../results/03_candidatevsd.csv")
+write.csv(parentalcaregenes, "../metadata/03_parentalcaregenes.csv")
 ```

### 2020-05-13 analysis/06_datawrangling.Rmd
now with WGCNA as step 5, wrangle as 6

---
title: "05_wrangle"
output: md_document
---

```{r setup}
library(tidyverse)
library(readr)

source("../R/themes.R")
```

## Wrangle data for hyotheses testing

```{r }
# col data with internal (hiloPRL) and external  hypothesese
colData <- read_csv("../metadata/04_colData.csv")
tail(colData)

# candidate genes from GO and literature
candidategenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% pull(gene)
candidategenes

## genes WGCNA prl module
WGCNAgenes <- read_csv("../results/05_PRLmodule.csv") %>% pull(x)
WGCNAgenes
```

## candidate genes

```{r}

```

```{R}
pithilo <- makecorrdf("female", "pituitary", hilogenes)
piteggchick <- makecorrdf("female", "pituitary", eggchickgenes)
pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[1:20])

write.csv(WGCNAvsd, "../results/03_WGCNAvsd.csv")
write.csv(hypothesisDEGs, "../results/03_hypothesisDEGs.csv")

```

### 2020-05-13 analysis/old/20_forrechelle.Rmd
moving old scripts like moved old figures

--- 2020-01-22 analysis/20_forrechelle.Rmd
@@
 
-candidategenes <- c("PRLR", "PRL", "CISH", "JAK2", "SOCS3", "GH" )
+candidategenes <- c("PRLR", "PRL", "CISH", "JAK2", "SOCS3", "GH", "FSHB", "CGA" )
+
 datadrivengenes <- c("COX1", "PRL", "CHGA", "ATP2B4", "LOC107055658")
@@
                  "NR3C1", "NR3C2",
-                 "ESR1", "ESR2", "AR")
+                 "ESR1", "ESR2", "AR",
+                 "FSH", "CGA")
 
 timepoints <- charlevels
-
 ```
@@
 genes <- as.data.frame(df$X1)
+
 ```
@@
 
-```{r candidates}
+```{r candidates1}
 
@@
 
+```
+
+```{r candidates2}
+
 df3 <- df2 %>% 
@@
 }
+
+```
+
+```{r candidates3}
 

### 2020-05-13 analysis/old/22_symphony.Rmd
moving old scripts like moved old figures

--- 2020-05-03 analysis/22_symphony.Rmd
@@
 library(cowplot)
+library(ggsignif)
 
 source("../R/themes.R")
+source("../R/functions.R")
+source("../R/wrangledata.R")
 
@@
 ```
+
+```{r, fig.height=6}
+head(candidatevsd)
+
+candidatevsd %>%
+  filter(gene == "PRL")  %>%
+  ggplot(aes(x=treatment, y=counts)) + 
+  geom_boxplot() +
+  theme_B3() +
+  geom_signif(comparisons = list(c("control", "bldg"),
+                                 c( "bldg", "lay"),
+                                 c( "lay", "inc.d3"),
+                                 c("inc.d3", "inc.d9"),
+                                 c( "inc.d9", "inc.d17"),
+                                 c( "inc.d17", "hatch"),
+                                 c("hatch", "n5"),
+                                 c( "n5", "n9")),  map_signif_level=TRUE)  +
+  facet_wrap(tissue~sex, scales = "free_y")
+```

### 2020-05-13 analysis/06_datawrangling.Rmd
now with files saved for futuer analysis

@@
 library(readr)
+library(corrr)
 
 source("../R/themes.R")
+source("../R/functions.R")
 ```
@@
 
-```{r }
+```{r data}
 # col data with internal (hiloPRL) and external  hypothesese
-colData <- read_csv("../metadata/04_colData.csv")
+colData <- read_csv("../metadata/04_colData.csv") %>%
+  column_to_rownames(var = "V1")
 tail(colData)
+```
+
+### variance stabilized gene expression  (vsd)
+
+```{r allvsd}
+vsd_path <- "../results/DEseq2/hypothesis/"   # path to the data
+vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
+vsd_pathfiles <- paste0(vsd_path, vsd_files)
+vsd_files
+
+## before pivoting, check names of df with
+## head(names(allvsd)) and tail(names(allvsd))
+
+allvsd <- vsd_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name")  %>% 
+  dplyr::rename("gene" = "X1") %>% 
+  pivot_longer(cols = blk.s061.pu.y_female_gonad_inc.d9:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "samples", values_to = "counts") 
+allvsd %>% select(-file_name) %>% head()
+
+head(allvsd)
+```
+
+## All DEGs (but currently only char DEGs)
+
+```{r allDEG}
+DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_DEGs.csv'), "[", 1))  %>% 
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+  select(DEG, sex, tissue, direction, everything())   %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+  dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj)  %>%
+  mutate(tissue = factor(tissue)) %>% 
+  drop_na()
+head(allDEG)
+```
+
+## gene lists
+
+```{r}
 
 # candidate genes from GO and literature
-candidategenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% pull(gene)
-candidategenes
+parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% pull(gene)
+parentalcaregenes
 
@@
 WGCNAgenes
-```
 
-## candidate genes
+unique(allDEG$comparison)
 
-```{r}
+## hilogenes
+
+hilogenes <- allDEG %>%
+  filter(comparison == "lo_hi")  %>%
+  drop_na() %>%
+  arrange(padj) %>% 
+  top_n(20) %>% pull(gene)
+
+eggchickgenes <- allDEG %>%
+  filter(comparison != "lo_hi")  %>%
+  drop_na() %>%
+  arrange(padj)  %>% 
+  top_n(20) %>% pull(gene)
 
@@
 
+### vsd for all and candidate, hypothesis, and data-driven genes
+
+```{r candidatevsd}
+candidategenes <- c(parentalcaregenes, WGCNAgenes, hilogenes, eggchickgenes)
+
+getcandidatevsd2 <- function(whichgenes, whichtissue, whichsex){
+  candidates  <- allvsd %>%
+    filter(gene %in% whichgenes) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/hypothesis/'), "[", 2)) %>%
+    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                  tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                  treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
+    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
+    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
+    filter(tissue == whichtissue, sex %in% whichsex)  %>%
+    drop_na()
+  #candidates$treatment <- factor(candidates$treatment, levels = alllevels)
+  return(candidates)
+}
+
+hypvsd <- getcandidatevsd2(candidategenes, "hypothalamus", sexlevels)
+pitvsd <- getcandidatevsd2(candidategenes, "pituitary", sexlevels)
+gonvsd <- getcandidatevsd2(candidategenes, "gonad", sexlevels)
+candidatevsd <- rbind(hypvsd, pitvsd)
+candidatevsd <- rbind(candidatevsd, gonvsd)
+
+head(candidatevsd)
+```
+
 ```{R}
-pithilo <- makecorrdf("female", "pituitary", hilogenes)
-piteggchick <- makecorrdf("female", "pituitary", eggchickgenes)
-pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[1:20])
-
-write.csv(WGCNAvsd, "../results/03_WGCNAvsd.csv")
-write.csv(hypothesisDEGs, "../results/03_hypothesisDEGs.csv")
-
+write.csv(candidatevsd, "../results/06_candidatevsd.csv", row.names = F)
+write.csv(allDEG, "../results/06_allDEG.csv", row.names = F)
 ```

### 2020-05-13 analysis/figure-3.Rmd
now with files saved for futuer analysis

@@
 
-colData <- read_csv("../metadata/00_colData.csv") %>%
+colData <- read_csv("../metadata/04_colData.csv") %>%
   mutate(treatment = factor(treatment, levels = alllevels),
@@
 pca2 <- makefvizdf("pituitary", charlevels, sexlevels)	
+```
+
+## candidate genes, hypotheses genes, and data-driven genes
+
+```{r}
+
+candidatevsd <- read_csv("../results/06_candidatevsd.csv")
+
+pithilo <- makecorrdf("female", "pituitary", hilogenes)
+piteggchick <- makecorrdf("female", "pituitary", eggchickgenes)
+pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[1:20])
+parentalcare <- makecorrdf("female", "pituitary", parentalcaregenes[1:20])
+
 ```

### 2020-05-13 R/themes.R

@@ -43,6 +43,8 @@ alllevels2 <- c("control", "bldg", "lay", "inc.d3", "inc.d9",
                 "m.inc.d3" ,  "early" ,  "m.inc.d9" , 
                 "m.inc.d17" , "m.n2", "prolong" , "extend")
 
+hypothesislevels <- c(  "lo" , "hi" ,  "nest",  "eggs", "chicks" )
+
 # tissue
 tissuelevels <- c("hypothalamus", "pituitary", "gonads")
 tissuelevel <- c("hypothalamus", "pituitary", "gonad")

### 2020-05-13 analysis/figure-3.Rmd
now working for fig 3... but includes manip data...

--- 2020-05-13 analysis/figure-3.Rmd
@@
 library(ggsignif)
+library(forcats)
 
@@
 
-colData <- read_csv("../metadata/00_colData.csv") %>%
+colData <- read_csv("../metadata/04_colData.csv") %>%
   mutate(treatment = factor(treatment, levels = alllevels),
-         tissue = factor(tissue, levels = tissuelevels)) %>% 
-  column_to_rownames(var = "X1") 
+         tissue = factor(tissue, levels = tissuelevels),
+         temprowname = V1) %>% 
+  column_to_rownames(var = "temprowname") 
 head(colData)
@@
 
-pca1f <- subsetmakepca("pituitary", charlevels, "female")	
-pca2f <- makefvizdf("pituitary", charlevels, "female")
+pca1f <- subsetmakepca("pituitary", alllevels, "female")	
+pca2f <- makefvizdf("pituitary", alllevels, "female")
 
-pca1m <- subsetmakepca("pituitary", charlevels, "male")	
-pca2m <- makefvizdf("pituitary", charlevels, "male")
+pca1m <- subsetmakepca("pituitary", alllevels, "male")	
+pca2m <- makefvizdf("pituitary", alllevels, "male")
 
-pca1 <- subsetmakepca("pituitary", charlevels, sexlevels)	
-pca2 <- makefvizdf("pituitary", charlevels, sexlevels)	
+pca1 <- subsetmakepca("pituitary", alllevels, sexlevels)	
+pca2 <- makefvizdf("pituitary", alllevels, sexlevels)	
+```
+
+## candidate genes, hypotheses genes, and data-driven genes
+
+```{r}
+# candidate counts for boxplots
+candidatevsd <- read_csv("../results/06_candidatevsd.csv") %>%
+  mutate(treatment = factor(treatment)) %>%
+  mutate(tissue = factor(tissue, levels = tissuelevel),
+        treatment = factor(treatment, levels = alllevels)) %>%
+  drop_na()
+
+## prlpit
+PRLpit <- candidatevsd %>% 
+  filter(gene == "PRL", tissue == "pituitary")
+
+# list of pvalue and DEGs
+allDEG <- read_csv("../results/06_allDEG.csv") 
+allDEG2 <- allDEG
+
+# candidate genes from GO and literature
+parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% pull(gene)
+
+## genes WGCNA prl module
+WGCNAgenes <- read_csv("../results/05_PRLmodule.csv") %>% pull(x)
+
+# DEGs infor for PRL and external
+hilodf <- allDEG %>%
+  filter(comparison == "lo_hi")  %>%
+  arrange(padj)
+
+eggchickdf <- allDEG %>%
+  filter(comparison != "lo_hi")  %>%
+  arrange(padj)
+
+eggchickgenes <- eggchickdf %>%  top_n(20) %>% pull(gene)
+hilogenes <- hilodf %>%  top_n(20) %>% pull(gene)
+
+## for correlations
+pithilo <- makecorrdf("female", "pituitary", hilogenes)
+piteggchick <- makecorrdf("female", "pituitary", eggchickgenes)
+pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[1:20])
+parentalcare <- makecorrdf("female", "pituitary", parentalcaregenes[1:20])
 ```
@@
 a2 <- plotprolactin(PRLpit, PRLpit$counts, "PRL", " ") + 
-  theme(axis.text.x = element_text())  +
+  theme(axis.text.x = element_text(angle = 25))  +
   theme(legend.position = c(0.8,0.2)) + 
   guides(fill = FALSE) 
-
-a <- plot_grid(a1,a2, labels = c("A"), label_size = 8)
+a <- plot_grid(a1,a2, labels = c("A"), label_size = 8, rel_widths = c(1,1.1))
 
@@
 b1 <- ggdraw() +  draw_image(b1, scale = 1)
-
-b2 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2200) + labs(subtitle = "female")  + labs(title = " ")  
-b3 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2200)  + labs(subtitle = "male")  + labs(title = " ") +
+b2 <- makenewbargraph("pituitary", "female","eggs_chicks", 0, 400) + labs(subtitle = "female")  + labs(title = " ")  
+b3 <- makenewbargraph("pituitary", "male", "eggs_chicks", 0, 400)  + labs(subtitle = "male")  + labs(title = " ") +
   theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), 
         axis.text.y = element_blank(), axis.line.y = element_blank())
-
-b4 <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2200) 
-b5 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2200)   + 
+b4 <- makenewbargraph("pituitary", "male", "lo_hi", 0, 4000) 
+b5 <- makenewbargraph("pituitary", "female", "lo_hi", 0, 4000)   + 
   theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), 
         axis.text.y = element_blank(), axis.line.y = element_blank())
-
 b25 <- plot_grid(b2,b3,b4,b5, nrow = 2, rel_widths = c(1.2,1), rel_heights = c(1.2,1))
-
-b6 <-  plottopDEGs(envDEGs, "female", "#3A80B9", "LFC eggs vs. chicks", "Top 10 female DEGs") + labs(title = " ")
-b7 <-  plottopDEGs(PRLDEGs, "male", "#3A80B9", "LFC eggs vs. chicks", "Top 10 male DEGs") + labs(title = " ")
-b8 <-  plottopDEGs(PRLDEGs, "female", "#19757A", "LFC lo vs. hi PRL", NULL)
-b9 <-  plottopDEGs(PRLDEGs, "male", "#19757A", "LFC lo vs. hi PRL", NULL)
-
+b6 <-  plottopDEGs(eggchickdf, "female", "#3A80B9", "LFC eggs vs. chicks", "Top 10 female DEGs") + labs(title = " ")
+b7 <-  plottopDEGs(hilodf, "male", "#3A80B9", "LFC eggs vs. chicks", "Top 10 male DEGs") + labs(title = " ")
+b8 <-  plottopDEGs(hilodf, "female", "#19757A", "LFC lo vs. hi PRL", NULL)
+b9 <-  plottopDEGs(hilodf, "male", "#19757A", "LFC lo vs. hi PRL", NULL)
 b15 <- plot_grid(b1,b25,nrow = 1, rel_widths = c(1.2,1))
@@
 d <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes from WGCNA module with PRL") + theme(legend.position = "right")
-
 cd <- plot_grid(c1,c2,d, nrow = 1, align = "h", labels = c("C", "", "D"), label_size = 8, rel_widths = c(1,1,1.3))

### 2020-05-13 R/themes.R

@@ -40,8 +40,10 @@ alllevels <- c("control", "bldg", "lay", "inc.d3", "m.inc.d3" ,
 # characteriation first, then manip by time
 alllevels2 <- c("control", "bldg", "lay", "inc.d3", "inc.d9", 
                 "inc.d17", "hatch", "n5", "n9",
-                "m.inc.d3" ,  "early" ,  "m.inc.d9" , 
-                "m.inc.d17" , "m.n2", "prolong" , "extend")
+                "m.inc.d3" ,  
+                "early" ,  "m.inc.d9" , 
+                "prolong" , "m.inc.d17" , 
+                "extend", "m.n2")
 
 hypothesislevels <- c(  "lo" , "hi" ,  "nest",  "eggs", "chicks" )
 

### 2020-05-15 analysis/figure-1.Rmd
fig 1 still working yea
changed axes

@@
 ```{r}
-allDEG <- read_csv("../results/03_allDEG.csv")
+allDEG <- read_csv("../results/03_allDEGs.csv") %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelschar))
 str(allDEG)
@@
 # hyp
-d2 <- makebargraph("hypothalamus","DEGs w/ + LFC", 0, 4300) + 
+d2 <- makebargraph("hypothalamus","DEGs w/ + LFC", 0, 4500) + 
   theme(axis.ticks = element_blank(),
@@
 # pit
-d4 <- makebargraph("pituitary","DEGs w/ + LFC", 0, 4300)  +  
+d4 <- makebargraph("pituitary","DEGs w/ + LFC", 0, 4500)  +  
   theme(axis.ticks = element_blank(),
@@
 # gon
-d6 <- makebargraph("gonad","DEGs w/ + LFC", 0, 4300) +  
+d6 <- makebargraph("gonad","DEGs w/ + LFC", 0, 4500) +  
   theme(axis.ticks = element_blank(),

### 2020-05-15 R/functions.R

@@ -948,6 +948,8 @@ makebargraph <- function(whichtissue, myylab, lowlim, higherlim){
 }
 
 
+
+
 sexbarplots <- function(df, lowlim, higherlim){
    
   DEGs <- df %>% dplyr::filter(direction != "NS")

### 2020-05-15 analysis/figure-1.Rmd
rm control from bar graph, need to make suppl files

@@
 b1 <- plottsneelipse(chartsne, chartsne$tissue, allcolors)  + labs(y = "tSNE 2 ", subtitle = " ")  
-
 b2 <- sexbarplots(hyppit, 0, 8100) + labs(subtitle = " ", y = "DEGs w/ + LFC") +
@@
     scale_x_discrete(labels=c(  "pituitary" = "pit", "gonad" = "gon"))
-
 c1 <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = "tSNE 2 ", subtitle = " ")    
@@
         axis.ticks = element_blank(), axis.title.y = element_blank())
+bc <- plot_grid(b1,b2,b3,b4, c1,c2,c3,c4, nrow = 1, rel_widths = c(1.5,1.1,0.9,0.9,1.5,1.1,0.9,0.9),
+                labels = c("B", "", "", "", "C"), label_size = 8)
 
@@
   facet_wrap(~sex, scales = "free")
-
 d3 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
@@
   theme(strip.text = element_blank())
-
 d5 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
@@
        subtitle = "gonads")
-
-# hyp
-d2 <- makebargraph("hypothalamus","DEGs w/ + LFC", 0, 4500) + 
-  theme(axis.ticks = element_blank(),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        axis.title.x = element_blank())  +
-  labs(subtitle = " ")
-# pit
-d4 <- makebargraph("pituitary","DEGs w/ + LFC", 0, 4500)  +  
-  theme(axis.ticks = element_blank(),
-        axis.text.x = element_blank(), 
-        axis.text.y = element_blank(),
+d2 <- makebargraph("hypothalamus","DEGs w/ + LFC", 0, 1800) + 
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank())  
+d4 <- makebargraph("pituitary","DEGs w/ + LFC", 0, 1800)  +  
+  theme(axis.text.x = element_blank(), 
         axis.title.x = element_blank(), 
-        strip.text.x = element_blank())  +
-  labs(subtitle = " ")
-# gon
-d6 <- makebargraph("gonad","DEGs w/ + LFC", 0, 4500) +  
-  theme(axis.ticks = element_blank(),
-        axis.text.y = element_blank(),
-        strip.text.x = element_blank()) +
+        strip.text.x = element_blank())   
+d6 <- makebargraph("gonad","DEGs w/ + LFC", 0, 1800) +  
+  theme(strip.text.x = element_blank()) +
   scale_x_discrete(labels = comparisonlabelschar)+
   labs(subtitle = " ")
-
-bc <- plot_grid(b1,b2,b3,b4, c1,c2,c3,c4, nrow = 1, rel_widths = c(1.5,1.1,0.9,0.9,1.5,1.1,0.9,0.9),
-                labels = c("B", "", "", "", "C"), label_size = 8)
-
 d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 2, rel_heights = c(1,0.9,1.3), rel_widths = c(1,2),

### 2020-05-15 analysis/03_datawrangling.Rmd
now with files saved for futuer analysis
now with suppl table and figs

@@
 
-candidategenes <- parentalcaregenes %>% distinct(gene) %>% pull(gene)
 # note: I can't find these genes in dataset: NOS1 or OXTR or FOX1B or HTR5A
@@
 
-## All DEGs (but currently only char DEGs)
+## All DEGs
 
@@
 
-# select only characterization plots
-allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
-allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevelschar)
-allDEG$direction <- factor(allDEG$direction, levels = charlevels)
-
-allDEG <- allDEG %>% drop_na()
+unique(allDEG$comparison)
 ```
@@
 ```{r write}
-write.csv(allDEG, "../results/03_allDEGs.csv")
+write.csv(allDEG, "../results/03_allDEG.csv", row.names = F)
 write.csv(candidatevsd, "../results/03_candidatevsd.csv")

### 2020-05-15 analysis/figure-1.Rmd
now with suppl table and figs

@@
 
-```{r}
-allDEG <- read_csv("../results/03_allDEGs.csv") %>%
+```{r suppltables}
+allDEG <- read_csv("../results/03_allDEG.csv") %>%
+  mutate(tissue = factor(tissue, levels = tissuelevel),
+         direction = factor(direction, levels = alllevels))
+
+# for fig 1
+DEGchar <- allDEG %>% 
+  filter(comparison %in% comparisonlevelschar) %>%
   mutate(comparison = factor(comparison, levels = comparisonlevelschar))
-str(allDEG)
-unique(allDEG$comparison)
+
+# for supple fig 1
+DEGcontrol <- allDEG %>% 
+  filter(grepl("control", comparison),
+         !grepl("m.|early|extend|prolong", comparison))  %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol))
+
+DEGbldg <- allDEG %>% 
+  filter(grepl("bldg", comparison),
+         !grepl("m.|early|extend|prolong", comparison))  %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))
+
+suppltable1 <- DEGchar %>%
+  arrange( tissue, sex, direction, gene)
+head(suppltable1)
+write_csv(suppltable1, "../results/suppltable1.csv")
+
+suppltable2 <- rbind(DEGcontrol, DEGbldg)  %>%
+  arrange( tissue, sex, direction, gene)
+head(suppltable2)
+write_csv(suppltable2, "../results/suppltable1.csv")
+
 ```
@@
 
-```{r DEGs}
+```{r DEGs, message=F}
 
@@
 
-## make figure
+## Figure 1
 
@@
        subtitle = "gonads")
-d2 <- makebargraph("hypothalamus","DEGs w/ + LFC", 0, 1800) + 
+d2 <- makebargraph(DEGchar, "hypothalamus","DEGs w/ + LFC", 0, 1800) + 
   theme(axis.text.x = element_blank(),
         axis.title.x = element_blank())  
-d4 <- makebargraph("pituitary","DEGs w/ + LFC", 0, 1800)  +  
+d4 <- makebargraph(DEGchar, "pituitary","DEGs w/ + LFC", 0, 1800)  +  
   theme(axis.text.x = element_blank(), 
@@
         strip.text.x = element_blank())   
-d6 <- makebargraph("gonad","DEGs w/ + LFC", 0, 1800) +  
+d6 <- makebargraph(DEGchar, "gonad","DEGs w/ + LFC", 0, 1800) +  
   theme(strip.text.x = element_blank()) +
-  scale_x_discrete(labels = comparisonlabelschar)+
-  labs(subtitle = " ")
+  scale_x_discrete(labels = comparisonlabelschar)
 d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 2, rel_heights = c(1,0.9,1.3), rel_widths = c(1,2),
@@
 
-# supple table 1 of all but control-bldg DEGs
+## suppl fig 1 controls versus bldg
 
-```{r}
-suppletable1 <- allDEG %>%
-  #filter(comparison != "control_bldg") %>%
-  group_by(sex, tissue, comparison) %>%
-  arrange( tissue, sex, direction, gene)
-head(suppletable1)
+```{r supplfig-1, fig.width=5, fig.height=5}
+s1a <- makebargraphsuppl(DEGcontrol, "hypothalamus","DEGs w/ + LFC", 0, 5000) + 
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank()) + 
+  labs(subtitle = "hypothalamus") 
+s1c <- makebargraphsuppl(DEGcontrol, "pituitary","DEGs w/ + LFC", 0, 5000)  +  
+  theme(axis.text.x = element_blank(), 
+        axis.title.x = element_blank(), 
+        strip.text.x = element_blank())    + 
+  labs(subtitle = "pituitary") 
+s1e <- makebargraphsuppl(DEGcontrol, "gonad","DEGs w/ + LFC", 0, 5000) +  
+  theme(strip.text.x = element_blank()) +
+  scale_x_discrete(labels = comparisonlevelscontrol) +
+  labs(subtitle = "gonads")
 
-write_csv(suppletable1, "../results/suppletable1.csv")
+s1b <- makebargraphsuppl(DEGbldg, "hypothalamus", NULL, 0, 1000) + 
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank())  
+s1d <- makebargraphsuppl(DEGbldg, "pituitary",NULL, 0, 1000)  +  
+  theme(axis.text.x = element_blank(), 
+        axis.title.x = element_blank(), 
+        strip.text.x = element_blank())   
+s1f <- makebargraphsuppl(DEGbldg, "gonad",NULL, 0, 1000) +  
+  theme(strip.text.x = element_blank()) +
+  scale_x_discrete(labels = comparisonlevelsbldg)
 
+s1 <- plot_grid(s1a,s1b,s1c,s1d,s1e,s1f, ncol = 2, rel_heights = c(1,0.9,1.3))
+
+plot_grid(a, s1, ncol = 1, rel_heights = c(0.25,1))
+```
+
+```{r eval = F}
 # save file for musical genes https://raynamharris.shinyapps.io/musicalgenes/

### 2020-05-15 analysis/figure-1.Rmd
now with better suppl fig and large file tracked

@@
          !grepl("m.|early|extend|prolong", comparison))  %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))
+  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))  %>%
+  drop_na()
 
@@
 ```{r supplfig-1, fig.width=5, fig.height=5}
+
+sa <- png::readPNG("../figures/images/fig_supplfig1.png")
+sa <- ggdraw() +  draw_image(sa, scale = 1)
+
 s1a <- makebargraphsuppl(DEGcontrol, "hypothalamus","DEGs w/ + LFC", 0, 5000) + 
@@
 
-plot_grid(a, s1, ncol = 1, rel_heights = c(0.25,1))
+supplfig1 <- plot_grid(sa, s1, ncol = 1, rel_heights = c(0.75,2))
+supplfig1
+
+pdf(file="../figures/supplfig-1-1.pdf", width=5, height=5)
+plot(supplfig1)
+dev.off()
 ```

### 2020-05-15 analysis/figure-1.Rmd
better suppl fig

@@
 
-sa <- png::readPNG("../figures/images/fig_supplfig1.png")
+sa <- png::readPNG("../figures/images/fig_supplfig-1.png")
 sa <- ggdraw() +  draw_image(sa, scale = 1)
@@
 
-s1 <- plot_grid(s1a,s1b,s1c,s1d,s1e,s1f, ncol = 2, rel_heights = c(1,0.9,1.3))
+s1 <- plot_grid(s1a,s1b,s1c,s1d,s1e,s1f, ncol = 2, rel_heights = c(1,0.9,1.3),
+                labels = c("B", "C"), label_size = 8)
 
-supplfig1 <- plot_grid(sa, s1, ncol = 1, rel_heights = c(0.75,2))
+supplfig1 <- plot_grid(sa, s1, ncol = 1, rel_heights = c(0.75,2), labels = c("A"), label_size = 8)
 supplfig1

### 2020-05-15 analysis/figure-1.Rmd
now with all three types of comparisons

@@
 
-```{r supplfig-1, fig.width=5, fig.height=5}
-
-sa <- png::readPNG("../figures/images/fig_supplfig-1.png")
-sa <- ggdraw() +  draw_image(sa, scale = 1)
+```{r supplfig-1, fig.width=7.25, fig.height=5}
 
@@
 
-s1 <- plot_grid(s1a,s1b,s1c,s1d,s1e,s1f, ncol = 2, rel_heights = c(1,0.9,1.3),
-                labels = c("B", "C"), label_size = 8)
+s1g <- makebargraphsuppl(DEGchar, "hypothalamus", NULL, 0, 2000) + 
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank())  
+s1h <- makebargraphsuppl(DEGchar, "pituitary",NULL, 0, 2000)  +  
+  theme(axis.text.x = element_blank(), 
+        axis.title.x = element_blank(), 
+        strip.text.x = element_blank())   
+s1i <- makebargraphsuppl(DEGchar, "gonad",NULL, 0, 2000) +  
+  theme(strip.text.x = element_blank()) +
+  scale_x_discrete(labels = comparisonlevelschar)
 
-supplfig1 <- plot_grid(sa, s1, ncol = 1, rel_heights = c(0.75,2), labels = c("A"), label_size = 8)
+s1 <- plot_grid(s1a, s1b, s1g, 
+          s1c, s1d, s1h, 
+          s1e, s1f, s1i, nrow = 3,
+          rel_heights = c(1,0.9,1.3),
+          labels = c("A", "B", "C"), label_size = 8, rel_widths = c(1.1,1,1))
+
+sa <- png::readPNG("../figures/images/fig_supplfig-1.png")
+sa <- ggdraw() +  draw_image(sa, scale = 1)
+
+supplfig1 <- plot_grid(sa, s1, ncol = 1, rel_heights = c(0.7,2))
 supplfig1
 
-pdf(file="../figures/supplfig-1-1.pdf", width=5, height=5)
+pdf(file="../figures/supplfig-1-1.pdf", width=7.25, height=5)
 plot(supplfig1)
 dev.off()
+
 ```

### 2020-05-18 docs/manuscript/elife-v1.tex
now torking

--- 2020-03-09 docs/manuscript/elife-v1.tex
@@
 
-\emph{Characterization of parental care}
+\emph{Data-driven characterization of reproductive and parental stage
+groups.}
 
-To characterize transcriptional activity, we first examined broad
-patterns of variation in gene expression in the hypothalamus, pituitary,
-and gonads of male and female rock doves across a series of timepoints
-that encapsulate important transitions in reproduction and and parental
-care (Fig. \ref{figure:fig1}a). t-Distributed Stochastic Neighbor
-Embedding (t-SNE) is a nonlinear, dimension reduction technique is used
-to visualize high-dimensional data by giving each datapoint a location
-in a two-dimensional map (van der Maaten and Hinton 2008). t-SNE is
-widely using in single-cell data process for identify related cell
-types. Here, we use t-SNE to identify variation associated with tissue,
-sex, and treatment (Fig. \ref{figure:fig1}j).
+We first examined broad patterns of variation in gene expression in the
+hypothalamus, pituitary, and gonads of male and female rock doves across
+a series of timepoints that encapsulate important transitions in
+reproduction and and parental care (Fig. 1A). t-Distributed Stochastic
+Neighbor Embedding (t-SNE) is a nonlinear, dimension reduction technique
+is used to visualize high-dimensional data (van der Maaten and Hinton
+2008) and is widely used in in RNA sequencing to identify related and
+distinct cell types (refs). Differential expespression profiling is
+useful for pair-wise comparisons of differentially expressed genes
+(DEGs) with positive log fold change (+LFC) in one group compared to
+another. Combined, the provide stronger support for similarities and
+differences between cell-types and treatment groups in multi-dimential
+datasets. First, the data supports that the hypothalamus, pituitary, and
+gonads have distinct gene expression profiles as evidenced by little
+overlap in t-SNE space and an average of 13,000 DEGs between each tissue
+(Fig. 1B). Second, the data supports that sex-differences are much more
+promient in the gonads than in the brain (13,000 versus 2,000 DEGs)
+(Fig. 1C). Because of these widespread tissue and sex-differences,
+further ananlyses were conducted separately for each tissue and sex
+(Fig. 1D, Suppl. Fig 1). In all the tissues, the control samples are
+quiet different from all the other timepoints samples, and only a
+fraction of the transcriptome is altered during the course of parental
+care at the timepoints sampled. The most notable differences male and
+female hpothalamic control samples are very different from on the rest
+(\textasciitilde{}7000 DEGs, and only the two nestling care stages (n5
+and n9)) show differential expression of more than 1,000 genes. In the
+pituitary, late inubcation and hatch are quite simlilar to each other
+and distinct from the earlier incubation stages stages, and temporal
+changes in gene expression appears to be less variable in males than
+females. In the female gonads there appear to be more changes in gene
+expression early one, between nest building and early incuation, but the
+male gonads show very little variation associated with parental stage.
 
-Our results suggest that tissue is the greatest source of variation in
-the dataset, with hypothalamic, pituitary, and gonadal samples all
-forming distinct clusters, with only a few pituitary and gonad samples
-falling outside the 95\% confidence interval (Fig. \ref{figure:fig1}b).
-Sex-differences are most prounounced in the gonads with male and female
-samples form two discrete clusters (Fig. \ref{figure:fig1}c). Treatment
-accounts for much less of the variation than tissue and sex, and even
-after removing sex as a variatble, treatment effects in all three
-tissues are hard to visualize simultatenously (Fig.
-\ref{figure:fig1}d,e), so we next looked at each tissue and sex
-separately. In the hypothalamus, control samples form tight clusters
-that stands out from all other timepoints that are largely overlapping
-in space (Fig. \ref{figure:fig1}h). The pituitary shows the strongest
-treatment effect with one cluster of the ``earlier'' parental timepoints
-(nest-building, lay, incubation days 3 and 9), a second cluster of
-``later'' parental timpoints (incubation day 17, hatch, and nestling
-care days 5 and 9; Fig. \ref{figure:fig1}i). Like the hypothalamus, in
-the gonads, the control timepoints appear to be much less variable that
-the parental timepoints, but there is no clear separation of gonadal
-samples by parental stage (Fig. \ref{figure:fig1}j). While the tSNE
-analysis reveals broad patterns across all levels of the experiment,
-additional tools are needed it hone in on the specific differences
-associated with parental care in each tissue. Differential gene
-expression analysis can shed insight into specific differences between
-treatment groups and thus deepen our understand of the transitions
-between reproduction and parental care behaviors. We used DESEq2 to
-identify genes whose expression changed significantly between sequential
-timepoints in each tissue and sex separately (Fig. \ref{figure:fig2}a).
+\emph{Hypothesis-driven analysis of candidate parental-care genes.}
 
-In the hypothalamus only two sequential stage comparisons results in a
-sizable change in expression; in females, 1927 genes changes their
-expression betwen hatch to nesting care day 5 (with 1076 decreasing and
-851 increased their expression), and in the male hypothalamus,87 genes
-were diferentially expressed bewteen incubation days 9 and 17 (with 38
-decreasing and 49 increasing their expression (Fig.
-\ref{figure:fig2}b)). In the pituitary, differential gene expression
-differenes are greatest when compariing incubation days 9 and 17, and
-this difference is more pronounced in females (Fig. \ref{figure:fig2}c).
-The next largest difference in the pituitary is between hatch and n5.
-The female putitary shows moderate chagnes from nest-bilding to lay to
-incuabation day 3 (Fig. \ref{figure:fig2}c). The male pituitary changes
-more that the female pitutiary changes between nestling care days 5 and
-9 (Fig. \ref{figure:fig2}c). About 2000 genes also change their
-expression in the female pituitary and about 425 in the male pituitary
-when comparing incubation days 9 and 17 (Fig. \ref{figure:fig2}). In the
-female gonads, 270 genes are decreased while only 56 are increased when
-comparing incubation days 9 and 17(Fig. 2). Many more significant
-changes in gene expression in the female gonad occur between
-nest-building and lay and between lay and clutch completion (incubation
-day 3) than in any other comparison (Fig. \ref{figure:fig2}).
-
-\emph{Internal physiology versus external stimuli hypothesis}
-
-We show that samples from the pituitary on incubation day 17 cluster
-more closely with hatch and nestling care day 5 samples than with
-incubation days 3 and 9. This suggest that, in the pituitary, internal
-mechanisms that prepare for chick arrival have the greatest effect on
-gene expression rather than suggesting that gene expression responds to
-changes in the external enviornament.
-
-Blah \ref{figure:fig3}, blah, \ref{figure:fig4} blah.
-
-\emph{Potential implications for cancer research}
-
-These results confirmed prolactin (PRL) gene expression fluctuates
-throughout parental care in a manner that is consistent with its role in
-promoting lactation and maintaining parental behaviors. Interestingly,
-we identified about 100 genes whose expression was correlated with PRL,
-including (BRCA1, which is associated with breast cancer). Additionally,
-we identified thousands of genes whose expression changed over the
-course of parental care. Exploring the relationship between PRL and
-other differentially expressed genes could provide important insights
-into the \texttt{symphony} of gene expression that regulates behavior at
-the organismal and cellular level.
+\begin{itemize}
+\tightlist
+\item
+  Champagne Curley Genes (Fig 2)
+\item
+  GO genes (Fig 2 Suppl Fig 1)
+\item
+  Cancer genes (Fig 2 Suppl Fig 2)
+\item
+  Shiny genes (Fig 2 Suppl Fig 3)
+\end{itemize}
 
@@
 
+\textbackslash{}begin\{figure\}{[}h{]} \centering
+\includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
+\textbackslash{}caption\{Broad patterns of variation in hypothalamic,
+pituitary, and gonadal gene expression in male and female rock doves
+across reproductive and parental transitions.** A) Experimental design.
+B) The t-SNE plot shows all 636 samples as points in space colored by
+tissue with ellipses drawn around show 95\% confidence interval, and the
+bar plots show the total number of differentially expressioned genes
+(DEGs) whose expression was significantly higher in one tissue versus
+the other. C) This t-SNE plot shows the same data points colored by sex
+while the ellipses delinate the source tissue; bar plots show the DEGs
+between the sexes for each tissue. D) These t-SNE and bar plots provide
+insight into the temporal differences in gene expression within each
+tissue for each sex. Here, the bar plots highlight the number of DEGs
+between each sequential time point. For comparisons to controls, see
+fupplemental figure 1.\} \label{figure:fig1}
+\textbackslash{}end\{figure\}
+
 \begin{figure}[h]
   \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
-  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.
-a) Experimental design.
-b) t-SNE plot highlighting tissue differences in gene expression.
-Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple).
-c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males).
-t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orange, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
-  \label{figure:fig1}
+  \includegraphics[width=1.0\textwidth]{../../figures/supplfig-1-1}
+  \caption{Summary of the total number of differentially expressed genes (DEGs) from pair-wise comparison using DESeq2.** Bar plots showing the total number of differentially expressed genes (p < 0.1) and colored by the direction of increased expression (or positive log fold change (+LFC))  are shown for three different analysis: relative to the control group (A), the nest-building (bldg) group (B), and the previous sampling timepoint (C).}
+  \label{figure:supplfig1}
 \end{figure}
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering
   \includegraphics[width=1.0\textwidth]{../../figures/fig3-1}
-  \caption{Another figure}
+  \caption{Another another figure}
   \label{figure:fig3}
@@
 \newpage
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/supplfig3}
-  \caption{R shiny, image 1}
-  \label{figure:supplfig3}
-\end{figure}
 

### 2020-05-18 docs/manuscript/elife_v1.Rmd
now torking

@@
 
-_Characterization of parental care_
+_Data-driven characterization of reproductive and parental stage groups._
 
-To characterize transcriptional activity, we first examined broad patterns of variation in gene expression in the hypothalamus, pituitary, and gonads of male and female rock doves across a series of timepoints that encapsulate important transitions in reproduction and and parental care (Fig.
-\ref{figure:fig1}a).
-t-Distributed Stochastic Neighbor Embedding (t-SNE) is a nonlinear, dimension reduction technique is used to visualize high-dimensional data by giving each datapoint a location in a two-dimensional map (van der Maaten and Hinton 2008).
-t-SNE is widely using in single-cell data process for identify related cell types.
-Here, we use t-SNE to identify variation associated with tissue, sex, and treatment (Fig.
-\ref{figure:fig1}j).
+We first examined broad patterns of variation in gene expression in the hypothalamus, pituitary, and gonads of male and female rock doves across a series of timepoints that encapsulate important transitions in reproduction and and parental care (Fig.
+1A).
+t-Distributed Stochastic Neighbor Embedding (t-SNE) is a nonlinear, dimension reduction technique is used to visualize high-dimensional data (van der Maaten and Hinton 2008) and is widely used in in RNA sequencing to identify related and distinct cell types (refs).
+Differential expespression profiling is useful for pair-wise comparisons of differentially expressed genes (DEGs) with positive log fold change (+LFC) in one group compared to another.
+Combined, the provide stronger support for similarities and differences between cell-types and treatment groups in multi-dimential datasets.
+First, the data supports that the hypothalamus, pituitary, and gonads have distinct gene expression profiles as evidenced by little overlap in t-SNE space and an average of 13,000 DEGs between each tissue (Fig.
+1B).
+Second, the data supports that sex-differences are much more promient in the gonads than in the brain (13,000 versus 2,000 DEGs) (Fig.
+1C).
+Because of these widespread tissue and sex-differences, further ananlyses were conducted separately for each tissue and sex (Fig.
+1D, Suppl.
+Fig 1).
+In all the tissues, the control samples are quiet different from all the other timepoints samples, and only a fraction of the transcriptome is altered during the course of parental care at the timepoints sampled.
+The most notable differences male and female hpothalamic control samples are very different from on the rest (~7000 DEGs, and only the two nestling care stages (n5 and n9)) show differential expression of more than 1,000 genes.
+In the pituitary, late inubcation and hatch are quite simlilar to each other and distinct from the earlier incubation stages stages, and temporal changes in gene expression appears to be less variable in males than females.
+In the female gonads there appear to be more changes in gene expression early one, between nest building and early incuation, but the male gonads show very little variation associated with parental stage.
 
-Our results suggest that tissue is the greatest source of variation in the dataset, with hypothalamic, pituitary, and gonadal samples all forming distinct clusters, with only a few pituitary and gonad samples falling outside the 95% confidence interval (Fig.
-\ref{figure:fig1}b).
-Sex-differences are most prounounced in the gonads with male and female samples form two discrete clusters (Fig.
-\ref{figure:fig1}c).
-Treatment accounts for much less of the variation than tissue and sex, and even after removing sex as a variatble, treatment effects in all three tissues are hard to visualize simultatenously (Fig.
-\ref{figure:fig1}d,e), so we next looked at each tissue and sex separately.
-In the hypothalamus, control samples form tight clusters that stands out from all other timepoints that are largely overlapping in space (Fig.
-\ref{figure:fig1}h).
-The pituitary shows the strongest treatment effect with one cluster of the “earlier” parental timepoints (nest-building, lay, incubation days 3 and 9), a second cluster of “later” parental timpoints (incubation day 17, hatch, and nestling care days 5 and 9; Fig.
-\ref{figure:fig1}i).
-Like the hypothalamus, in the gonads, the control timepoints appear to be much less variable that the parental timepoints, but there is no clear separation of gonadal samples by parental stage (Fig.
-\ref{figure:fig1}j).
-While the tSNE analysis reveals broad patterns across all levels of the experiment, additional tools are needed it hone in on the specific differences associated with parental care in each tissue.
-Differential gene expression analysis can shed insight into specific differences between treatment groups and thus deepen our understand of the transitions between reproduction and parental care behaviors.
-We used DESEq2 to identify genes whose expression changed significantly between sequential timepoints in each tissue and sex separately (Fig.
-\ref{figure:fig2}a).
+_Hypothesis-driven analysis of candidate parental-care genes._
 
-In the hypothalamus only two sequential stage comparisons results in a sizable change in expression; in females, 1927 genes changes their expression betwen hatch to nesting care day 5 (with 1076 decreasing and 851 increased their expression), and in the male hypothalamus,87 genes were diferentially expressed bewteen incubation days 9 and 17 (with 38 decreasing and 49 increasing their expression (Fig.
-\ref{figure:fig2}b)).
-In the pituitary, differential gene expression differenes are greatest when compariing incubation days 9 and 17, and this difference is more pronounced in females (Fig.
-\ref{figure:fig2}c).
-The next largest difference in the pituitary is between hatch and n5.
-The female putitary shows moderate chagnes from nest-bilding to lay to incuabation day 3 (Fig.
-\ref{figure:fig2}c).
-The male pituitary changes more that the female pitutiary changes between nestling care days 5 and 9  (Fig.
-\ref{figure:fig2}c).
-About 2000 genes also change their expression in the female pituitary and about 425 in the male pituitary when comparing incubation days 9 and 17 (Fig.
-\ref{figure:fig2}).
-In the female gonads, 270 genes are decreased while only 56 are increased when comparing incubation days 9 and 17(Fig.
-2).
-Many more significant changes in gene expression in the female gonad occur between nest-building and lay and between lay and clutch completion (incubation day 3) than in any other comparison (Fig.
-\ref{figure:fig2}).
-
-_Internal physiology versus external stimuli hypothesis_
-
-We show that samples from the pituitary on 
-                     incubation day 17 cluster more closely with hatch and nestling care 
-                     day 5 samples than with incubation days 3 and 9. This suggest that, 
-                     in the pituitary, internal mechanisms that prepare for chick arrival 
-                     have the greatest effect on gene expression rather than suggesting 
-                     that gene expression responds to changes in the external enviornament.
-
-Blah, blah, \ref{figure:fig4} blah.
-
-_Potential implications for cancer research_
-
-These results confirmed prolactin (PRL) gene expression fluctuates throughout parental care 
-in a manner that is consistent with its role in promoting lactation 
-                and maintaining parental behaviors.
-                Interestingly, we identified about 100 genes whose expression 
-                was correlated with PRL,
-                including (BRCA1, which is associated with breast cancer).
-                Additionally, we identified thousands of genes whose expression 
-                changed over 
-                the course of parental care.
-                Exploring the relationship between PRL and other differentially 
-                expressed genes could provide important insights into the `symphony` of 
-                gene expression 
-                that regulates behavior at the organismal and cellular level.
+- Champagne Curley Genes (Fig 2)
+- GO genes (Fig 2 Suppl Fig 1)
+- Cancer genes (Fig 2 Suppl Fig 2)
+- Shiny genes (Fig 2 Suppl Fig 3)
 
@@
   \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
-  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.
-a) Experimental design.
-b) t-SNE plot highlighting tissue differences in gene expression.
-Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple).
-c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males).
-t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orange, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
+  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.** A) Experimental design.
+B) The t-SNE plot shows all 636 samples as points in space colored by tissue with ellipses drawn around show 95% confidence interval, and the bar plots show the total number of differentially expressioned genes (DEGs) whose expression was significantly higher in one tissue versus the other.
+C) This t-SNE plot shows the same data points colored by sex while the ellipses delinate the source tissue; bar plots show the DEGs between the sexes for each tissue.
+D) These t-SNE and bar plots provide insight into the temporal differences in gene expression within each tissue for each sex.
+Here, the bar plots highlight the number of DEGs between each sequential time point.
+For comparisons to controls, see fupplemental figure 1.}
   \label{figure:fig1}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/supplfig-1-1}
+  \caption{Summary of the total number of differentially expressed genes (DEGs) from pair-wise comparison using DESeq2.** Bar plots showing the total number of differentially expressed genes (p < 0.1) and colored by the direction of increased expression (or positive log fold change (+LFC))  are shown for three different analysis: relative to the control group (A), the nest-building (bldg) group (B), and the previous sampling timepoint (C).}
+  \label{figure:supplfig1}
 \end{figure}
@@
   \centering
+  \includegraphics[width=1.0\textwidth]{../../figures/fig3-1}
+  \caption{Another another figure}
+  \label{figure:fig3}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht]
+  \centering
   \includegraphics[width=1.0\textwidth]{../../figures/fig4-1}
@@
 \newpage
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig5-1}
-  \caption{Hypothesis-driven data exploration and software development.}
-  \label{figure:fig5}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/supplfig1}
-  \caption{R shiny, image 1}
-  \label{figure:supplfig3}
-\end{figure}
 

### 2020-05-18 analysis/00_datawrangling.Rmd
with lots of the clutter removed from the output

@@
 
-```{r setup}
+```{r setup, message=F}
 # https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html
@@
 source("../R/themes.R")
+
+knitr::opts_chunk$set(message=F, warning=F)
 ```
@@
          "gene" = "Name")
-head(kallistodata)
+#head(kallistodata)
 
@@
 ```{r}
-head(kallistodata)
+#head(kallistodata)
 # aggregate transcript counts to gene counts
@@
 row.names(colData) <- colData$V1
-head(colData)
+str(colData)
+
+## check that rownames and colnames match for DESeq
+ncol(countData) == nrow(colData)
 ```
 
-## check that rownames and colnames match for DESeq
+## summarize data
 
 ```{r}
-str(colData)
-ncol(countData) == nrow(colData)
 
@@
 table(colData$sex, colData$treatment, colData$tissue)
+
+length(unique(colData$bird))
+table(colData$sex, colData$tissue)
 ```
@@
   distinct()
-birds
+head(birds)
 ```

### 2020-05-19 analysis/figure-1.Rmd
better fig1 and suppl fig 1

--- 2020-05-15 analysis/figure-1.Rmd
@@
 
-sa <- png::readPNG("../figures/images/fig_supplfig-1.png")
-sa <- ggdraw() +  draw_image(sa, scale = 1)
-
 s1a <- makebargraphsuppl(DEGcontrol, "hypothalamus","DEGs w/ + LFC", 0, 5000) + 
@@
   theme(strip.text.x = element_blank()) +
-  scale_x_discrete(labels = comparisonlevelscontrol) +
+  scale_x_discrete(labels = comparisonlabelscontrol) +
   labs(subtitle = "gonads")
@@
   theme(strip.text.x = element_blank()) +
-  scale_x_discrete(labels = comparisonlevelsbldg)
+  scale_x_discrete(labels = comparisonlabelsbldg)
 
-s1 <- plot_grid(s1a,s1b,s1c,s1d,s1e,s1f, ncol = 2, rel_heights = c(1,0.9,1.3),
-                labels = c("B", "C"), label_size = 8)
+s1 <- plot_grid(s1a, s1b, 
+          s1c, s1d, 
+          s1e, s1f, nrow = 3,
+          rel_heights = c(1,0.9,1.3),
+          labels = c("A", "B"), label_size = 8, rel_widths = c(1.1,1))
 
-supplfig1 <- plot_grid(sa, s1, ncol = 1, rel_heights = c(0.75,2), labels = c("A"), label_size = 8)
+sa <- png::readPNG("../figures/images/fig_supplfig-1.png")
+sa <- ggdraw() +  draw_image(sa, scale = 1)
+
+supplfig1 <- plot_grid(s1, sa, ncol = 1, rel_heights = c(2,0.7))
 supplfig1

### 2020-05-19 R/functions.R

@@ -913,7 +913,7 @@ plottsneelipse <- function(tsnedf, pointcolor, whichcolors){
 plottsneelipsev2 <- function(tsnedf, pointcolor, whichcolors){
   p <- ggplot(tsnedf, aes(x = V1, y = V2)) +
    # stat_ellipse(linetype = 1, aes(color = pointcolor)) +
-    geom_point(size = 0.5, aes(color = pointcolor)) +
+    geom_point(size = 1, aes(color = pointcolor)) +
     theme_B3() +
     labs(x = "tSNE 1", y = "tSNE 2",
          subtitle = " ") +

### 2020-05-20 analysis/figure-2.Rmd
missing candidate gene list

@@
 ```{r DEGs}
+
 # summary DEG results from DESeq2
-candidateDEGS <- read_csv("../results/suppletable1.csv") %>%
+candidateDEGS <- read_csv("../results/suppltable1.csv") %>%
   filter(gene %in% candidategenes) %>%
@@
 
-bcd <- plot_grid(b,c,d, ncol = 3, rel_widths = c(1,1,1), labels = c("B", "C", "D"), label_size = 8)
+bcd <- plot_grid(b,c,d, ncol = 3, rel_widths = c(1,1,1), labels = c("A", "B", "C"), label_size = 8)
 
-fig2 <- plot_grid(a,bcd, ncol = 1, labels = c("A"), label_size = 8, rel_heights = c(0.1,1))
+fig2 <- plot_grid(bcd, a, ncol = 1, rel_heights = c(1,0.1))
 fig2

### 2020-05-20 analysis/figure-1.Rmd
music removed, correlations moved to suppl, fig2 needs more

@@
 head(suppltable2)
-write_csv(suppltable2, "../results/suppltable1.csv")
+write_csv(suppltable2, "../results/suppltable2.csv")
 

### 2020-05-20 analysis/figure-2.Rmd
music removed, correlations moved to suppl, fig2 needs more

@@
 
+## Candidate genes
+
+```{r candidategenes}
+parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% select(-X1)
+head(parentalcaregenes)
+
+curleychampagnegenes <- parentalcaregenes %>% distinct(literature) %>% drop_na() %>% pull(literature)
+GOgenes <- parentalcaregenes %>% distinct(GO)  %>% drop_na() %>% pull(GO)
+candidategenes <- parentalcaregenes %>% pull(gene) 
+candidategenes
+```
+
 ## Candidate DEGs
@@
   pivot_wider(names_from = comparison, values_from = res) %>%
-  select(gene, control_bldg, bldg_lay, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
+  select(gene, bldg_lay, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
 candidateDEGS
@@
 table1 <- left_join(candidateDEGS, parentalcaregenes) %>%
-  select(gene, control_bldg:n5_n9, parentalcare, parentalbehavior, NCBI) %>%
-  mutate(parentalcare = if_else(is.na(parentalcare), " ", "X"),
-         parentalbehavior = if_else(is.na(parentalbehavior), " ", "X")) %>%
-  rename("Literature" = "parentalcare", "GO" =  "parentalbehavior")
-table1$numDEGs <- rowSums(is.na(table1)) # count NAs to know how many are NS
-
-table1 <- table1 %>% 
-  mutate(sig = ifelse(numDEGs == 6, "NS", "DEG")) %>% 
-  arrange(gene)  %>%  select(-sig, -numDEGs)
+  select(gene, bldg_lay:n5_n9, literature, GO, NCBI) %>%
+  mutate(literature = if_else(is.na(literature), " ", "X"),
+         GO = if_else(is.na(GO), " ", "X"))
 table1[is.na(table1)] <- " " # replace NA with blank space so it's pretty
@@
 
-kable(table1)
 ```
 
-## Candidate Correlations
+## Candidate VSDs
 
-```{r}
+```{r candidatevsds}
 # load `candidatevsd` with `source("../R/wrangledata.R")`
-candidatevsd <- read_csv("../results/03_candidatevsd.csv")
-head(candidatevsd)
-
-curleychampagnegenes <- curleychampagnegenes %>% pull(gene)
-
-hyp1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
-pit1 <- makecorrdf("female", "pituitary", curleychampagnegenes)  
-gon1 <- makecorrdf("female", "gonad", curleychampagnegenes)
-
-hyp2 <- makecorrdf("male", "hypothalamus", curleychampagnegenes)  
-pit2 <- makecorrdf("male", "pituitary", curleychampagnegenes)  
-gon2 <- makecorrdf("male", "gonad", curleychampagnegenes)
-
-hyp3 <- makecorrdf(sexlevels, "hypothalamus", curleychampagnegenes)  
-pit3 <- makecorrdf(sexlevels, "pituitary", curleychampagnegenes)  
-gon3 <- makecorrdf(sexlevels, "gonad", curleychampagnegenes)
+candidatevsd <- read_csv("../results/03_candidatevsd.csv") %>% 
+  select(-X1) %>%
+  filter(treatment %in% charlevels)
+tail(candidatevsd)
 
@@
 MG <- subsetcandidatevsdwide("male", "gonad") 
+```
 
+## Candidate Correlations - SUppl fig 1
+
+```{r correlations}
+hyp1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
+pit1 <- makecorrdf("female", "pituitary", curleychampagnegenes)  
+gon1 <- makecorrdf("female", "gonad", curleychampagnegenes)
+
+hyp2 <- makecorrdf("male", "hypothalamus", curleychampagnegenes)  
+pit2 <- makecorrdf("male", "pituitary", curleychampagnegenes)  
+gon2 <- makecorrdf("male", "gonad", curleychampagnegenes)
+
+hyp3 <- makecorrdf("female", "hypothalamus", GOgenes)  
+pit3 <- makecorrdf("female", "pituitary", GOgenes)  
+gon3 <- makecorrdf("female", "gonad", GOgenes)
+
+hyp4 <- makecorrdf("male", "hypothalamus", GOgenes)  
+pit4 <- makecorrdf("male", "pituitary", GOgenes)  
+gon4 <- makecorrdf("male", "gonad", GOgenes)  
+```
+
+```{r supplfig-2, fig.width=7.25, fig.height=7.25}
+
+b1 <- plotcorrplot(hyp1, "females") + labs(y = "Hypothalamus", title = "Parental Care Literature")  + 
+  scale_x_discrete(position = "top")   + theme(axis.text.x = element_text(vjust = -0.25))
+b2 <- plotcorrplot(pit1, NULL)  + labs(y = "Pituitary") + 
+  theme(axis.text.x = element_blank())
+b3 <- plotcorrplot(gon1, NULL)  + labs(y = "Gonad") + theme(axis.text.x = element_text(vjust = 0.25))
+
+b4 <- plotcorrplot(hyp2, "males") + labs( title =  " ") + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = -0.25)) +
+  scale_x_discrete(position = "top") 
+b5 <- plotcorrplot(pit2, NULL)    + 
+  theme(axis.text.x = element_blank(), axis.text.y = element_blank())
+b6 <- plotcorrplot(gon2, NULL)  + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = 0.25))
+
+b7 <- plotcorrplot(hyp3, "females") + labs( title = "Parental Care GO") + 
+   scale_x_discrete(position = "top") + theme(axis.text.x = element_text(vjust = -0.25))
+b8 <- plotcorrplot(pit3, NULL)   + 
+  theme(axis.text.x = element_blank())
+b9 <- plotcorrplot(gon3, NULL)  + theme(axis.text.x = element_text(vjust = 0.25))
+
+b10 <- plotcorrplot(hyp4, "males") + labs( title =  " ") + 
+  theme(axis.text.y = element_blank(),
+        axis.text.x = element_text(vjust = -0.25)) + scale_x_discrete(position = "top") 
+b11 <- plotcorrplot(pit4, NULL)    + 
+  theme(axis.text.x = element_blank() , axis.text.y = element_blank())
+b12 <- plotcorrplot(gon4, NULL)   + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = 0.25))
+
+forlegend <- plotcorrplot(gon4, NULL) + theme(legend.position = "bottom") 
+mylegend <- get_legend(forlegend)
+
+allcorplots <- plot_grid(b1,b4,b7,b10,
+          b2,b5,b8,b11,
+          b3,b6,b9,b12, rel_heights = c(1.4,1,1.2),
+          labels = c("A", " ", "B"), label_size = 8, rel_widths = c(1.4,1.2,1.2,1))
+
+supplfig2 <- plot_grid(allcorplots, mylegend, ncol = 1, rel_heights = c(1,0.1))
+supplfig2
+
+pdf(file="../figures/supplfig-2.pdf", width=7.25, height=7.25)
+plot(supplfig2)
+dev.off()
 ```
@@
 
-b1 <- plotcorrplot(hyp3, "females and males") + theme(legend.position = "none")  + labs(y = "Hypothalamus", title = "Candidate gene correlations")
-b2 <- plotcorrplot(pit3, NULL) + theme(legend.position = "none") + labs(y = "Pituitary")
-b3 <- plotcorrplot(gon3, NULL) + theme(legend.position = "none")  + labs(y = "Gonad")
-
-c1 <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female", title =  "Top correlation")  
+c1 <- scattercorrelations(FH, FH$DRD4, FH$HTR2C, "#969696", "DRD4", "HTR2C") + labs(subtitle = "female", title =  "Top correlation")  
 c3 <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = NULL) 
@@
 
-c2 <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C")  + labs(subtitle = "male", title = "", y = NULL)  
+c2 <- scattercorrelations(MH, MH$DRD4, MH$HTR2C, "#525252", "DRD4", "HTR2C")  + labs(subtitle = "male", title = "", y = NULL)  
 c4 <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = NULL, y = NULL)  
@@
 
-d1 <- candidateboxplot("hypothalamus", c("DRD1"), sexlevels) + labs(x = NULL, title = "Temporal expression pattern" )  
+d1 <- candidateboxplot("hypothalamus", c("DRD4"), sexlevels) + labs(x = NULL, title = "Temporal expression pattern" )  
 d2 <- candidateboxplot("hypothalamus", c("HTR2C"), sexlevels) + labs(x = NULL ) + theme(strip.text = element_blank())
@@
 
-b <- plot_grid(b1,b2,b3, ncol = 1, rel_heights = c(1.1,1,1))
+#b <- plot_grid(b1,b2,b3, ncol = 1, rel_heights = c(1.1,1,1))
 c <- plot_grid(c1,c2,c3,c4,c5,c6, ncol = 2, rel_heights = c(1.1,1,1), rel_widths = c(1.1,1))
@@
 
-bcd <- plot_grid(b,c,d, ncol = 3, rel_widths = c(1,1,1), labels = c("A", "B", "C"), label_size = 8)
+bcd <- plot_grid(c,d, ncol = 2, rel_widths = c(1,1), labels = c("A", "B"), label_size = 8)
 
@@
 
-```{r fig3, fig.height=3, fig.width=7.25}
-
-musicplot <- function(whichgene, whichtissue){
-
-    p <- candidatevsd %>%
-      group_by(treatment, tissue, gene, sex)  %>% 
-      summarize(median = median(counts, na.rm = T), 
-                se = sd(counts,  na.rm = T)/sqrt(length(counts))) %>%
-      dplyr::mutate(scaled = rescale(median, to = c(0, 7))) %>%
-      dplyr::mutate(image = "../figures/images/musicnote.png")   %>%
-      filter(
-        gene %in% whichgene,
-        tissue %in% whichtissue      ) %>% 
-      collect() %>%
-      drop_na() %>%
-      ggplot( aes(x = treatment, y = median)) +
-      geom_errorbar(aes(ymin = median - se, 
-                        ymax = median + se, color = "white"),  width=0) +
-      geom_image(aes(image=image), size = 0.1)+
-      theme_B3() +
-      theme(legend.position = "none",
-            title = element_text(face = "italic"),
-            strip.text = element_text(color = "white"),
-            axis.text.x = element_text(angle = 45, hjust = 1),
-            axis.text.y = element_blank(),
-            axis.line = element_blank(), axis.ticks = element_blank()) +
-      scale_color_manual(values = allcolors) +
-      labs(y = " ",x = NULL ) +
-      facet_wrap(~sex, nrow = 2, scales = "free_y")
-      return(p)
-}
-
-music <- png::readPNG("../figures/images/fig_music.png")
-music <- ggdraw() +  draw_image(music, scale = 1)
-
-musicalgenes <- png::readPNG("../figures/images/fig_musicalgenes.png")
-musicalgenes <- ggdraw() +  draw_image(musicalgenes, scale = 1)
-
-p1 <- musicplot(c("AVP", "OXT"), "hypothalamus") + labs(subtitle = "Hypothalamic AVP & OXT" ) + theme(axis.text.x = element_blank())
-p2 <- musicplot(c("AVP", "OXT"), "pituitary") + labs(subtitle = "Pituitary AVP & OXT " ) + theme(axis.text.x = element_blank())
-p3 <- musicplot(c("AVP", "OXT"), "gonad") + labs(subtitle = "Gonadal  AVP & OXT"  )
-
-p123 <- plot_grid(music, p1, music, p3, ncol = 2, rel_widths = c(0.125,1))
-
-plot_grid(musicalgenes, p123, nrow = 1, rel_widths = c(1.5,1))
-
-```
-
 ```{r writefiles}
@@
 write.csv(table1, "../results/table1.csv")
-write.csv(candidategenes, "../results/candidategenes.csv")
 ```

### 2020-05-20 analysis/old/musicplot.R

@@ -1,0 +1,46 @@
+
+musicplot <- function(whichgene, whichtissue){
+  
+  p <- candidatevsd %>%
+    group_by(treatment, tissue, gene, sex)  %>% 
+    summarize(median = median(counts, na.rm = T), 
+              se = sd(counts,  na.rm = T)/sqrt(length(counts))) %>%
+    dplyr::mutate(scaled = rescale(median, to = c(0, 7))) %>%
+    dplyr::mutate(image = "../figures/images/musicnote.png")   %>%
+    filter(
+      gene %in% whichgene,
+      tissue %in% whichtissue      ) %>% 
+    collect() %>%
+    drop_na() %>%
+    ggplot( aes(x = treatment, y = median)) +
+    geom_errorbar(aes(ymin = median - se, 
+                      ymax = median + se, color = "white"),  width=0) +
+    geom_image(aes(image=image), size = 0.1)+
+    theme_B3() +
+    theme(legend.position = "none",
+          title = element_text(face = "italic"),
+          strip.text = element_text(color = "white"),
+          axis.text.x = element_text(angle = 45, hjust = 1),
+          axis.text.y = element_blank(),
+          axis.line = element_blank(), axis.ticks = element_blank()) +
+    scale_color_manual(values = allcolors) +
+    labs(y = " ",x = NULL ) +
+    facet_wrap(~sex, nrow = 2, scales = "free_y")
+  return(p)
+}
+
+music <- png::readPNG("../figures/images/fig_music.png")
+music <- ggdraw() +  draw_image(music, scale = 1)
+
+musicalgenes <- png::readPNG("../figures/images/fig_musicalgenes.png")
+musicalgenes <- ggdraw() +  draw_image(musicalgenes, scale = 1)
+
+
+p1 <- musicplot(c("AVP", "OXT"), "hypothalamus") + labs(subtitle = "Hypothalamic AVP & OXT" ) + theme(axis.text.x = element_blank())
+p2 <- musicplot(c("AVP", "OXT"), "pituitary") + labs(subtitle = "Pituitary AVP & OXT " ) + theme(axis.text.x = element_blank())
+p3 <- musicplot(c("AVP", "OXT"), "gonad") + labs(subtitle = "Gonadal  AVP & OXT"  )
+
+p123 <- plot_grid(music, p1, music, p3, ncol = 2, rel_widths = c(0.125,1))
+
+plot_grid(musicalgenes, p123, nrow = 1, rel_widths = c(1.5,1))
+

### 2020-05-20 analysis/figure-2.Rmd
updated fig 2 and suppl fig 2

@@
          GO = if_else(is.na(GO), " ", "X"))
-table1[is.na(table1)] <- " " # replace NA with blank space so it's pretty
 head(table1)
-
 ```
@@
   select(-X1) %>%
-  filter(treatment %in% charlevels)
+  filter(treatment %in% charlevelsnocontrol) %>%
+  drop_na()
 tail(candidatevsd)
@@
 ```{r correlations}
+
+makecorrdf <- function(whichsex, whichtissue, whichgenes){
+  corrrdf <- candidatevsd %>%
+    filter(sex == whichsex, tissue == whichtissue,
+           gene %in% whichgenes) %>%
+    drop_na() %>%
+    pivot_wider(names_from = gene, values_from = counts) %>%
+    select(-sex, -tissue, -treatment, -samples) %>%
+    correlate() 
+  # print(head(corrrdf))
+
+  corrrdflong <- corrrdf %>%
+    pivot_longer(-rowname, names_to = "gene2", values_to = "corr") %>%
+    rename("gene1" = "rowname") %>%
+    arrange(desc(corr)) %>%
+    mutate(tissue = whichtissue,
+           sex = whichsex) %>%
+    select(tissue, sex, gene1, gene2, corr)
+  print("The top two correlations in this tissue and sex are:")
+  print(corrrdflong[c(1,3),])
+
+  return(corrrdf)
+}
+
 hyp1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
@@
 
-forlegend <- plotcorrplot(gon4, NULL) + theme(legend.position = "bottom") 
+forlegend <- plotcorrplot(gon4, NULL) + theme(legend.position = "bottom") + labs(color = "Correlation")
 mylegend <- get_legend(forlegend)
@@
 
-a <- png::readPNG("../figures/images/fig_fig2a.png")
-a <- ggdraw() +  draw_image(a, scale = 1)
+candidateboxplot <- function(whichtissue, whichgenes, whichsex){
 
-c1 <- scattercorrelations(FH, FH$DRD4, FH$HTR2C, "#969696", "DRD4", "HTR2C") + labs(subtitle = "female", title =  "Top correlation")  
-c3 <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = NULL) 
-c5 <- scattercorrelations(FG, FG$ESR1, FG$PGR, "#969696", "ESR1", "PGR")   + labs(subtitle = NULL)
+  p <- candidatevsd %>%
+    filter(tissue %in% whichtissue,
+           gene %in% whichgenes,
+           sex %in% whichsex) %>%
+    mutate(treatment = factor(treatment, levels = charlevels)) %>%
+    ggplot(aes(x = treatment, y = counts)) +
+    geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
+    geom_jitter(size = 0.25, aes(color = sex)) +
+    #facet_wrap(~sex,  nrow = 1) +
+    scale_fill_manual(values = allcolors) +
+    scale_color_manual(values = allcolors) +
+    theme_B3() + 
+    theme(legend.position = "none",
+          axis.title.y = element_text(face = "italic"),
+          axis.text.x = element_blank(),
+          axis.ticks = element_blank(),
+          axis.line.x = element_blank()) +
+    labs(y = whichgenes,
+         x = NULL) +
+    geom_signif(comparisons = list(c( "control", "bldg"),
+                                   c( "bldg", "lay"),
+                                   c( "lay", "inc.d3"),
+                                   c("inc.d3", "inc.d9"),
+                                   c( "inc.d9", "inc.d17"),
+                                   c( "inc.d17", "hatch"),
+                                   c("hatch", "n5"),
+                                   c( "n5", "n9")),  
+                map_signif_level=TRUE,
+                textsize = 1.5, family = 'Helvetica',
+                vjust = 1.5, size = 0)
 
-c2 <- scattercorrelations(MH, MH$DRD4, MH$HTR2C, "#525252", "DRD4", "HTR2C")  + labs(subtitle = "male", title = "", y = NULL)  
-c4 <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = NULL, y = NULL)  
-c6 <- scattercorrelations(MG, MG$ESR1, MG$PGR, "#525252", "ESR1", "PGR")  + labs(subtitle = NULL, y = NULL)
+  return(p)
+}
 
-d1 <- candidateboxplot("hypothalamus", c("DRD4"), sexlevels) + labs(x = NULL, title = "Temporal expression pattern" )  
-d2 <- candidateboxplot("hypothalamus", c("HTR2C"), sexlevels) + labs(x = NULL ) + theme(strip.text = element_blank())
-d3 <- candidateboxplot("pituitary", c("AVPR1A"), sexlevels) + labs(x = NULL )  + theme( strip.text = element_blank())
-d4 <- candidateboxplot("pituitary", c("CRHR1"), sexlevels) + labs(x = NULL  )  + theme(strip.text = element_blank())
-d5 <- candidateboxplot("gonad", c("ESR1"), sexlevels) + labs(x = NULL) + theme(strip.text = element_blank())
-d6 <- candidateboxplot("gonad", c("PGR"), sexlevels) + labs(x = NULL ) + theme(strip.text = element_blank())
+c1 <- scattercorrelations(FH, FH$CRHR2, "CRHR2", FH$HTR2C, "HTR2C", "#969696" ) +  labs( title =  "Hypothlamic expression", subtitle = "females")  
+c3 <- scattercorrelations(FP, FP$AVPR1A, "AVPR1A", FP$CRHR1,  "CRHR1", "#969696")  + labs(title = "Pituitary expression", subtitle = "females") 
+c5 <- scattercorrelations(FG, FG$CRHBP, "CRHBP",  FG$ESR2,  "ESR2", "#969696")   + labs(title = "Gonadal expression", subtitle = "females")
+
+c2 <- scattercorrelations(MH, MH$DRD1, "DRD1", MH$HTR2C, "HTR2C", "#525252")  + labs(title = " ", subtitle = "males" )  
+c4 <- scattercorrelations(MP, MP$AVPR1A,  "AVPR1A",  MP$CRHR1, "CRHR1", "#525252")  + labs(title = " ", subtitle = "males" ) 
+c6 <- scattercorrelations(MG, MG$PGR, "PGR", MG$PRLR, "PRLR", "#525252")   + labs(title = " ", subtitle = "males" )
+
+d1 <- candidateboxplot("hypothalamus", c("CRHR2"), "female") + labs(x = NULL, title = " ", subtitle = " " )  
+d2 <- candidateboxplot("hypothalamus", c("HTR2C"), "female") + labs(x = NULL )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
+d3 <- candidateboxplot("pituitary", c("AVPR1A"), "female") + labs(x = NULL , subtitle = " " ) 
+d4 <- candidateboxplot("pituitary", c("CRHR1"), "female") + labs(x = NULL  )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
+d5 <- candidateboxplot("gonad", c("CRHBP"), "female") + labs(x = NULL, subtitle = " " )
+d6 <- candidateboxplot("gonad", c("ESR2"), "female") + labs(x = NULL ) + theme(axis.text.x = element_text(angle = 45, vjust = 1))
+
+e1 <- candidateboxplot("hypothalamus", c("DRD1"), "male") + labs(x = NULL, title = " " , subtitle = " " )  
+e2 <- candidateboxplot("hypothalamus", c("HTR2C"), "male") + labs(x = NULL )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
+e3 <- candidateboxplot("pituitary", c("AVPR1A"), "male") + labs(x = NULL , subtitle = " " ) 
+e4 <- candidateboxplot("pituitary", c("CRHR1"), "male") + labs(x = NULL  )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
+e5 <- candidateboxplot("gonad", c("PGR"), "male") + labs(x = NULL, subtitle = " " ) 
+e6 <- candidateboxplot("gonad", c("PRLR"), "male") + labs(x = NULL )   + theme(axis.text.x = element_text(angle = 45, vjust = 1))
 
 #b <- plot_grid(b1,b2,b3, ncol = 1, rel_heights = c(1.1,1,1))
-c <- plot_grid(c1,c2,c3,c4,c5,c6, ncol = 2, rel_heights = c(1.1,1,1), rel_widths = c(1.1,1))
+c123 <- plot_grid(c1,c3,c5, ncol = 1, rel_heights = c(1.1,1,1), labels = c("A", "B", "C"), label_size = 8)
+c456 <- plot_grid(c2,c4,c6, ncol = 1, rel_heights = c(1.1,1,1))
 d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))
+e <- plot_grid(e1,e2,d3,e4,e5,e6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))
 
-bcd <- plot_grid(c,d, ncol = 2, rel_widths = c(1,1), labels = c("A", "B"), label_size = 8)
-
-fig2 <- plot_grid(bcd, a, ncol = 1, rel_heights = c(1,0.1))
+fig2 <- plot_grid(c123,d,c456,e, nrow  = 1)
 fig2
 
-pdf(file="../figures/fig2-1.pdf", width=7.25, height=7.25)
-plot(fig2)
-dev.off()
+#pdf(file="../figures/fig2-1.pdf", width=7.25, height=7.25)
+#plot(fig2)
+#dev.off()
 ```

### 2020-05-20 analysis/figure-2.Rmd
even better fig 2

@@
          x = NULL) +
-    geom_signif(comparisons = list(c( "control", "bldg"),
-                                   c( "bldg", "lay"),
+    geom_signif(comparisons = list( c( "bldg", "lay"),
                                    c( "lay", "inc.d3"),
@@
                 textsize = 1.5, family = 'Helvetica',
-                vjust = 1.5, size = 0)
+                vjust = 1.5, size = 0.2)
 
@@
 
-c1 <- scattercorrelations(FH, FH$CRHR2, "CRHR2", FH$HTR2C, "HTR2C", "#969696" ) +  labs( title =  "Hypothlamic expression", subtitle = "females")  
-c3 <- scattercorrelations(FP, FP$AVPR1A, "AVPR1A", FP$CRHR1,  "CRHR1", "#969696")  + labs(title = "Pituitary expression", subtitle = "females") 
-c5 <- scattercorrelations(FG, FG$CRHBP, "CRHBP",  FG$ESR2,  "ESR2", "#969696")   + labs(title = "Gonadal expression", subtitle = "females")
+c1 <- scattercorrelations(FH, FH$CRHR2, "CRHR2", FH$HTR2C, "HTR2C", "#969696" ) +  labs(title = " ", subtitle = " " )  
+c3 <- scattercorrelations(FP, FP$AVPR1A, "AVPR1A", FP$CRHR1,  "CRHR1", "#969696")  + labs(title = " ", subtitle = " ") 
+c5 <- scattercorrelations(FG, FG$CRHBP, "CRHBP",  FG$ESR2,  "ESR2", "#969696")   + labs(title = " ", subtitle = " ")
 
-c2 <- scattercorrelations(MH, MH$DRD1, "DRD1", MH$HTR2C, "HTR2C", "#525252")  + labs(title = " ", subtitle = "males" )  
-c4 <- scattercorrelations(MP, MP$AVPR1A,  "AVPR1A",  MP$CRHR1, "CRHR1", "#525252")  + labs(title = " ", subtitle = "males" ) 
-c6 <- scattercorrelations(MG, MG$PGR, "PGR", MG$PRLR, "PRLR", "#525252")   + labs(title = " ", subtitle = "males" )
+c2 <- scattercorrelations(MH, MH$DRD1, "DRD1", MH$HTR2C, "HTR2C", "#525252")  + labs(title = " ", subtitle = " " )  
+c4 <- scattercorrelations(MP, MP$AVPR1A,  "AVPR1A",  MP$CRHR1, "CRHR1", "#525252")  + labs(title = " ", subtitle = " " ) 
+c6 <- scattercorrelations(MG, MG$PGR, "PGR", MG$PRLR, "PRLR", "#525252")   + labs(title = " ", subtitle = " " )
 
-d1 <- candidateboxplot("hypothalamus", c("CRHR2"), "female") + labs(x = NULL, title = " ", subtitle = " " )  
+d1 <- candidateboxplot("hypothalamus", c("CRHR2"), "female") + labs(x = NULL, title =  "Hypothlamic expression", subtitle = "females" )  
 d2 <- candidateboxplot("hypothalamus", c("HTR2C"), "female") + labs(x = NULL )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
-d3 <- candidateboxplot("pituitary", c("AVPR1A"), "female") + labs(x = NULL , subtitle = " " ) 
+d3 <- candidateboxplot("pituitary", c("AVPR1A"), "female") + labs(x = NULL , title = "Pituitary expression", subtitle = "females") 
 d4 <- candidateboxplot("pituitary", c("CRHR1"), "female") + labs(x = NULL  )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
-d5 <- candidateboxplot("gonad", c("CRHBP"), "female") + labs(x = NULL, subtitle = " " )
+d5 <- candidateboxplot("gonad", c("CRHBP"), "female") + labs(x = NULL, title = "Gonadal expression", subtitle = "females")
 d6 <- candidateboxplot("gonad", c("ESR2"), "female") + labs(x = NULL ) + theme(axis.text.x = element_text(angle = 45, vjust = 1))
 
-e1 <- candidateboxplot("hypothalamus", c("DRD1"), "male") + labs(x = NULL, title = " " , subtitle = " " )  
+e1 <- candidateboxplot("hypothalamus", c("DRD1"), "male") + labs(x = NULL, title = " " , subtitle = "males" )  
 e2 <- candidateboxplot("hypothalamus", c("HTR2C"), "male") + labs(x = NULL )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
-e3 <- candidateboxplot("pituitary", c("AVPR1A"), "male") + labs(x = NULL , subtitle = " " ) 
+e3 <- candidateboxplot("pituitary", c("AVPR1A"), "male") + labs(x = NULL , subtitle = "males" ) 
 e4 <- candidateboxplot("pituitary", c("CRHR1"), "male") + labs(x = NULL  )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
-e5 <- candidateboxplot("gonad", c("PGR"), "male") + labs(x = NULL, subtitle = " " ) 
+e5 <- candidateboxplot("gonad", c("PGR"), "male") + labs(x = NULL, subtitle = "males" ) 
 e6 <- candidateboxplot("gonad", c("PRLR"), "male") + labs(x = NULL )   + theme(axis.text.x = element_text(angle = 45, vjust = 1))
@@
 #b <- plot_grid(b1,b2,b3, ncol = 1, rel_heights = c(1.1,1,1))
-c123 <- plot_grid(c1,c3,c5, ncol = 1, rel_heights = c(1.1,1,1), labels = c("A", "B", "C"), label_size = 8)
+c123 <- plot_grid(c1,c3,c5, ncol = 1, rel_heights = c(1.1,1,1))
 c456 <- plot_grid(c2,c4,c6, ncol = 1, rel_heights = c(1.1,1,1))
-d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))
-e <- plot_grid(e1,e2,d3,e4,e5,e6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))
+d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1), labels = c("A", " ",  "B", " ", "C"), label_size = 8)
+e <- plot_grid(e1,e2,e3,e4,e5,e6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))
 
-fig2 <- plot_grid(c123,d,c456,e, nrow  = 1)
+fig2 <- plot_grid(d, c123,e,c456,nrow  = 1)
 fig2
 
-#pdf(file="../figures/fig2-1.pdf", width=7.25, height=7.25)
-#plot(fig2)
-#dev.off()
+pdf(file="../figures/fig2-1.pdf", width=7.25, height=7.25)
+plot(fig2)
+dev.off()
 ```

### 2020-05-21 analysis/figure-1.Rmd
better figs

@@
 bc <- plot_grid(b1,b2,b3,b4, c1,c2,c3,c4, nrow = 1, rel_widths = c(1.5,1.1,0.9,0.9,1.5,1.1,0.9,0.9),
-                labels = c("B", "", "", "", "C"), label_size = 8)
+                labels = c("A", "", "", "", "B"), label_size = 8)
 
@@
   theme(strip.text.x = element_blank()) +
-  scale_x_discrete(labels = comparisonlabelschar)
-d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 2, rel_heights = c(1,0.9,1.3), rel_widths = c(1,2),
-                  labels = c("D"), label_size = 8)
+  scale_x_discrete(labels = comparisonlabelschar) +
+  labs(x = "Sequential comparisons of differential gene expression")
 
-fig1 <- plot_grid(a,bc, d, nrow = 3, rel_heights = c(0.7,0.7,2),
-                  labels = c("A"), label_size = 8)
+d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 2, rel_heights = c(1,0.9,1.4), rel_widths = c(1,2),
+                  labels = c("C"), label_size = 8)
+
+fig1 <- plot_grid(bc, a, d, nrow = 3, rel_heights = c(0.7,0.7,2))
 fig1
@@
 
-```{r eval = F}
+## Save files
+
+```{r write}
 # save file for musical genes https://raynamharris.shinyapps.io/musicalgenes/
@@
 #write.csv(chartsne, "../../musicalgenes/data/tsne.csv")
+
 ```

### 2020-05-21 analysis/figure-2.Rmd
better fig 3

@@
 ```{r correlations}
-
-makecorrdf <- function(whichsex, whichtissue, whichgenes){
-  corrrdf <- candidatevsd %>%
-    filter(sex == whichsex, tissue == whichtissue,
-           gene %in% whichgenes) %>%
-    drop_na() %>%
-    pivot_wider(names_from = gene, values_from = counts) %>%
-    select(-sex, -tissue, -treatment, -samples) %>%
-    correlate() 
-  # print(head(corrrdf))
-
-  corrrdflong <- corrrdf %>%
-    pivot_longer(-rowname, names_to = "gene2", values_to = "corr") %>%
-    rename("gene1" = "rowname") %>%
-    arrange(desc(corr)) %>%
-    mutate(tissue = whichtissue,
-           sex = whichsex) %>%
-    select(tissue, sex, gene1, gene2, corr)
-  print("The top two correlations in this tissue and sex are:")
-  print(corrrdflong[c(1,3),])
-
-  return(corrrdf)
-}
 

### 2020-05-21 analysis/figure-3.Rmd
better fig 3

@@
 
-pca1f <- subsetmakepca("pituitary", alllevels, "female")	
-pca2f <- makefvizdf("pituitary", alllevels, "female")
+pca1f <- subsetmakepca("pituitary", charlevelsnocontrol, "female")	
+pca2f <- makefvizdf("pituitary", charlevelsnocontrol, "female")
 
-pca1m <- subsetmakepca("pituitary", alllevels, "male")	
-pca2m <- makefvizdf("pituitary", alllevels, "male")
+pca1m <- subsetmakepca("pituitary", charlevelsnocontrol, "male")	
+pca2m <- makefvizdf("pituitary", charlevelsnocontrol, "male")
 
-pca1 <- subsetmakepca("pituitary", alllevels, sexlevels)	
-pca2 <- makefvizdf("pituitary", alllevels, sexlevels)	
+pca1 <- subsetmakepca("pituitary", charlevelsnocontrol, sexlevels)	
+pca2 <- makefvizdf("pituitary", charlevelsnocontrol, sexlevels)	
 ```
@@
   mutate(tissue = factor(tissue, levels = tissuelevel),
-        treatment = factor(treatment, levels = alllevels)) %>%
+        treatment = factor(treatment, levels = charlevelsnocontrol)) %>%
   drop_na()
@@
 # list of pvalue and DEGs
-allDEG <- read_csv("../results/06_allDEG.csv") 
-allDEG2 <- allDEG
+allDEG <- read_csv("../results/06_allDEG.csv")  %>%
+  mutate(direction = factor(direction, levels = hypothesislevels))
 
 # candidate genes from GO and literature
-parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% pull(gene)
+parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% 
+  distinct(literature)  %>% drop_na() %>% pull(literature)
 
@@
 eggchickdf <- allDEG %>%
-  filter(comparison != "lo_hi")  %>%
+  filter(!comparison %in% c("lo_hi", "nest_eggs", "nest_chicks"))  %>%
   arrange(padj)
@@
 pithilo <- makecorrdf("female", "pituitary", hilogenes)
-piteggchick <- makecorrdf("female", "pituitary", eggchickgenes)
 pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[1:20])
-parentalcare <- makecorrdf("female", "pituitary", parentalcaregenes[1:20])
+
+## for scatter correlation
+
+candidatevsdwide <- candidatevsd  %>%
+    pivot_wider(names_from = gene, values_from = counts) 
+pitwide <- subsetcandidatevsdwide(sexlevels, "pituitary")
+
 ```
@@
 
-a1 <- plotpc12(pca1, pca2, pca1$treatment, allcolors, "Pituitary gene expression", " ")   
-a2 <- plotprolactin(PRLpit, PRLpit$counts, "PRL", " ") + 
+a1 <- plotpc12(pca1, pca2, pca1$treatment, allcolors, "Pituitary gene expression", NULL)   
+a2 <- plotprolactin(PRLpit, PRLpit$counts, "PRL", NULL) + 
   theme(axis.text.x = element_text(angle = 25))  +
-  theme(legend.position = c(0.8,0.2)) + 
-  guides(fill = FALSE) 
-a <- plot_grid(a1,a2, labels = c("A"), label_size = 8, rel_widths = c(1,1.1))
+  facet_wrap(~sex)
+a <- plot_grid(a1,a2, labels = c("A", "B"), label_size = 8, rel_widths = c(1,1))
 
@@
 b1 <- ggdraw() +  draw_image(b1, scale = 1)
-b2 <- makenewbargraph("pituitary", "female","eggs_chicks", 0, 400) + labs(subtitle = "female")  + labs(title = " ")  
-b3 <- makenewbargraph("pituitary", "male", "eggs_chicks", 0, 400)  + labs(subtitle = "male")  + labs(title = " ") +
+b2 <- makenewbargraph("pituitary", "female","eggs_chicks", 0, 4000) + labs(subtitle = "female")  + labs(title = " ") +
+  scale_x_discrete(labels = "eggs vs. chicks")
+b3 <- makenewbargraph("pituitary", "male", "eggs_chicks", 0, 4000)  + labs(subtitle = "male")  + labs(title = " ") +
   theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), 
-        axis.text.y = element_blank(), axis.line.y = element_blank())
-b4 <- makenewbargraph("pituitary", "male", "lo_hi", 0, 4000) 
+        axis.text.y = element_blank(), axis.line.y = element_blank()) +
+  scale_x_discrete(labels = "eggs vs. chicks")
+b4 <- makenewbargraph("pituitary", "male", "lo_hi", 0, 4000)  +
+  scale_x_discrete(labels = "lo vs. hi PRL")
 b5 <- makenewbargraph("pituitary", "female", "lo_hi", 0, 4000)   + 
   theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), 
-        axis.text.y = element_blank(), axis.line.y = element_blank())
+        axis.text.y = element_blank(), axis.line.y = element_blank()) +
+  scale_x_discrete(labels = "lo vs. hi PRL")
 b25 <- plot_grid(b2,b3,b4,b5, nrow = 2, rel_widths = c(1.2,1), rel_heights = c(1.2,1))
-b6 <-  plottopDEGs(eggchickdf, "female", "#3A80B9", "LFC eggs vs. chicks", "Top 10 female DEGs") + labs(title = " ")
-b7 <-  plottopDEGs(hilodf, "male", "#3A80B9", "LFC eggs vs. chicks", "Top 10 male DEGs") + labs(title = " ")
-b8 <-  plottopDEGs(hilodf, "female", "#19757A", "LFC lo vs. hi PRL", NULL)
-b9 <-  plottopDEGs(hilodf, "male", "#19757A", "LFC lo vs. hi PRL", NULL)
+
+b6 <-  plottopDEGs(eggchickdf, "female",  "LFC eggs vs. chicks", "Top 10 female DEGs") + labs(title = " ")
+b7 <-  plottopDEGs(eggchickdf, "male",  "LFC eggs vs. chicks", "Top 10 male DEGs") + labs(title = " ")
+b8 <-  plottopDEGs(hilodf, "female",  "LFC lo vs. hi PRL", NULL)
+b9 <-  plottopDEGs(hilodf, "male",  "LFC lo vs. hi PRL", NULL)
 b15 <- plot_grid(b1,b25,nrow = 1, rel_widths = c(1.2,1))
 b69 <- plot_grid(b6,b7,b8,b9, align = "v" , rel_heights = c(1.2,1))
-b <- plot_grid(b15,b69, labels = c("B"), label_size = 8, rel_widths = c(1.2,1))
+b <- plot_grid(b15,b69, labels = c("C"), label_size = 8, rel_widths = c(1.2,1))
 
-c1 <- plotcorrplot(piteggchick, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 eggs vs. chicks DEGs")
-c2 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 lo vs. PRL DEGs") 
-d <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes from WGCNA module with PRL") + theme(legend.position = "right")
-cd <- plot_grid(c1,c2,d, nrow = 1, align = "h", labels = c("C", "", "D"), label_size = 8, rel_widths = c(1,1,1.3))
+c1 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 lo vs. hi PRL DEGs") 
+c2 <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes from WGCNA module with PRL") + theme(legend.position = "right")
+c3 <- candidateboxplot("pituitary", c("KPNA2"), sexlevels)
+c4 <- candidateboxplot("pituitary", c("CDK1"), sexlevels) + 
+  theme(axis.text.x = element_text(angle = 45, hjust =1), strip.text = element_blank())  + labs(x = " ")
+c34 <- plot_grid(c3,c4, nrow = 2, rel_heights = c(1,1.2))
 
-fig3 <- plot_grid(a, b, cd, nrow = 3, rel_heights = c(1.2,2,1.8))
+c <- plot_grid(c1,c2,c34, nrow = 1, labels = c("D", "", "E"), label_size = 8, rel_widths = c(1.1,1.4,1))
+
+fig3 <- plot_grid(a, b, c, nrow = 3, rel_heights = c(1.2,2,1.8))
 fig3

### 2020-05-23 analysis/03_datawrangling.Rmd
better fig1 and suppl fig 1
fig 3 from thursday

@@
 # note: I can't find these genes in dataset: NOS1 or OXTR or FOX1B or HTR5A
+
+GOgeneonly <- parentalcaregenes %>%
+  filter(GO != "") %>%
+  pull(GO)
+GOgeneonly
 ```
@@
 allvsd %>% select(-file_name) %>% head()
+
 ```

### 2020-05-23 analysis/06_datawrangling.Rmd
now working for fig 3... but includes manip data...
fig 3 from thursday

@@
 ---
-title: "05_wrangle"
+title: "06_wrangle"
 output: md_document
@@
 head(allvsd)
+
+library(stringr)
+allvsd %>%
+  filter(str_detect(samples, "Merc"))
+
 ```
@@
 
-```{r}
+```{r genelists}
 
@@
 write.csv(allDEG, "../results/06_allDEG.csv", row.names = F)
+
+candidategenes
 ```

### 2020-05-23 analysis/figure-3.Rmd
now with box plots
fig 3 from thursday

@@
 write.csv(PRLpit,"../results/PRLvsd.csv", row.names = F)
+```
+
+## PRL and PRLR in three tissues
+
+```{r}
+p1 <- candidateboxplot("hypothalamus", c("PRL"), sexlevels) + labs(subtitle = "hypothalamus", title = "PRL")
+p2 <- candidateboxplot("pituitary", c("PRL"), sexlevels) + labs(subtitle = "pituitary")
+p3 <- candidateboxplot("gonad", c("PRL"), sexlevels) + labs(subtitle = "gonads") + theme(axis.text.x = element_text(angle = 45))
+
+p4 <- candidateboxplot("hypothalamus", c("PRLR"), sexlevels) + labs(subtitle = "blood", title = " ") 
+p5 <- candidateboxplot("pituitary", c("PRLR"), sexlevels) + labs(subtitle = "cropmilk")
+p6 <- candidateboxplot("gonad", c("PRLR"), sexlevels) + theme(axis.text.x = element_text(angle = 45))+ labs(subtitle = "spleen")
+
+plot_grid(p1,p4,p2,p5,p3,p6, ncol = 2, rel_heights = c(1.2,1,1.2))
 

### 2020-05-23 analysis/06_datawrangling.Rmd
small chagnes, nothing run'
now with control and bldg, hopefully
gonad bldg still missing

@@
 
-```{r allvsd}
+```{r allvsd, message = F, warning  = F}
 vsd_path <- "../results/DEseq2/hypothesis/"   # path to the data
@@
 head(allvsd)
-
-library(stringr)
-allvsd %>%
-  filter(str_detect(samples, "Merc"))
-
 ```
@@
 
-```{R}
+```{r write files}
 write.csv(candidatevsd, "../results/06_candidatevsd.csv", row.names = F)
@@
 
-candidategenes
 ```

### 2020-05-23 analysis/figure-3.Rmd
small chagnes, nothing run'
gonad bldg still missing

--- 2020-05-21 analysis/figure-3.Rmd
@@
 
-# Figure 4: All things prolactin
+# Figure 3: All things prolactin
 
@@
   mutate(tissue = factor(tissue, levels = tissuelevel),
-        treatment = factor(treatment, levels = charlevelsnocontrol)) %>%
+        treatment = factor(treatment, levels = alllevels)) %>%
   drop_na()
+head(candidatevsd)
 
@@
 write.csv(PRLpit,"../results/PRLvsd.csv", row.names = F)
+```
+
+## PRL and PRLR in three tissues
+
+```{r eval = F}
+p1 <- candidateboxplot("hypothalamus", c("PRL"), sexlevels) + labs(subtitle = "hypothalamus", title = "PRL")
+p2 <- candidateboxplot("pituitary", c("PRL"), sexlevels) + labs(subtitle = "pituitary")
+p3 <- candidateboxplot("gonad", c("PRL"), sexlevels) + labs(subtitle = "gonads") + theme(axis.text.x = element_text(angle = 45))
+
+p4 <- candidateboxplot("hypothalamus", c("PRLR"), sexlevels) + labs(subtitle = " ", title = " ") 
+p5 <- candidateboxplot("pituitary", c("PRLR"), sexlevels) + labs(subtitle = " ")
+p6 <- candidateboxplot("gonad", c("PRLR"), sexlevels) + theme(axis.text.x = element_text(angle = 45))+ labs(subtitle = " ")
+
+plot_grid(p1,p4,p2,p5,p3,p6, ncol = 2, rel_heights = c(1.2,1,1.2))
 

### 2020-05-23 R/functions.R

@@ -1068,17 +1068,6 @@ makecorrdf <- function(whichsex, whichtissue, whichgenes){
     select(-sex, -tissue, -treatment, -samples) %>%
     correlate() 
   # print(head(corrrdf))
-  
-  corrrdflong <- corrrdf %>%
-    pivot_longer(-rowname, names_to = "gene2", values_to = "corr") %>%
-    rename("gene1" = "rowname") %>%
-    arrange(desc(corr)) %>%
-    mutate(tissue = whichtissue,
-           sex = whichsex) %>%
-    select(tissue, sex, gene1, gene2, corr)
-  print("The top two correlations in this tissue and sex are:")
-  print(corrrdflong[c(1,3),]) 
-  
   return(corrrdf)
 }
 

### 2020-05-25 analysis/06_datawrangling.Rmd
new genes lists, but candidate vsd is empty :(

@@
 library(corrr)
+library(forcats)
 
@@
 source("../R/functions.R")
+source("../R/genelists.R")
 ```
@@
 ```{r allvsd, message = F, warning  = F}
-vsd_path <- "../results/DEseq2/hypothesis/"   # path to the data
+vsd_path <- "../results/DEseq2/treatment/"   # path to the data
 vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
@@
   dplyr::rename("gene" = "X1") %>% 
-  pivot_longer(cols = blk.s061.pu.y_female_gonad_inc.d9:y98.o50.x_male_pituitary_inc.d3, 
+  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
                names_to = "samples", values_to = "counts")
+
 allvsd %>%  select(-file_name) %>% head()
 
-head(allvsd)
+colData %>% filter(tissue == "gonads", treatment == "bldg")
+
+allvsd %>% filter(samples == "blk11.x_female_gonad_bldg")
 ```
@@
 
-# candidate genes from GO and literature
-parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% pull(gene)
-parentalcaregenes
-
-## genes WGCNA prl module
-WGCNAgenes <- read_csv("../results/05_PRLmodule.csv") %>% pull(x)
-WGCNAgenes
-
-unique(allDEG$comparison)
-
-## hilogenes
+## "internal clock" PRL genes
 
@@
 
+## external environment gens 
 eggchickgenes <- allDEG %>%
@@
 ```{r candidatevsd}
-candidategenes <- c(parentalcaregenes, WGCNAgenes, hilogenes, eggchickgenes)
+# all list of candidate gens accumlated here
+
+candidategenes <- c(parentalcaregenes, WGCNAgenes, hilogenes, eggchickgenes, 
+                    suszynskaagenes, shaidgenes)
 
@@
     dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/hypothesis/'), "[", 2)) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/treatment/'), "[", 2)) %>%
     dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
@@
 gonvsd <- getcandidatevsd2(candidategenes, "gonad", sexlevels)
+
 candidatevsd <- rbind(hypvsd, pitvsd)
 candidatevsd <- rbind(candidatevsd, gonvsd) 
-
-head(candidatevsd)
+candidatevsd <- candidatevsd %>%
+  mutate(treatment = factor(treatment)) %>%
+  mutate(treatment = fct_recode(treatment, "early" = "m.inc.d8" )) %>%
+  mutate(treatment = factor(treatment, alllevels))
+tail(candidatevsd)
 ```
 
-```{r write files}
+```{r writefiles}
 write.csv(candidatevsd, "../results/06_candidatevsd.csv", row.names = F)
 write.csv(allDEG, "../results/06_allDEG.csv", row.names = F)
-
 ```

### 2020-05-25 analysis/figure-3.Rmd
new genes lists, but candidate vsd is empty :(

--- 2020-05-21 analysis/figure-3.Rmd
@@
 source("../R/functions.R")
+source("../R/genelists.R")
 
@@
 
-pca1f <- subsetmakepca("pituitary", charlevelsnocontrol, "female")	
-pca2f <- makefvizdf("pituitary", charlevelsnocontrol, "female")
+pca1f <- subsetmakepca("pituitary", charlevels, "female")	
+pca2f <- makefvizdf("pituitary", charlevels, "female")
 
-pca1m <- subsetmakepca("pituitary", charlevelsnocontrol, "male")	
-pca2m <- makefvizdf("pituitary", charlevelsnocontrol, "male")
+pca1m <- subsetmakepca("pituitary", charlevels, "male")	
+pca2m <- makefvizdf("pituitary", charlevels, "male")
 
-pca1 <- subsetmakepca("pituitary", charlevelsnocontrol, sexlevels)	
-pca2 <- makefvizdf("pituitary", charlevelsnocontrol, sexlevels)	
+pca1 <- subsetmakepca("pituitary", charlevels, sexlevels)	
+pca2 <- makefvizdf("pituitary", charlevels, sexlevels)	
 ```
@@
   mutate(treatment = factor(treatment)) %>%
-  mutate(tissue = factor(tissue, levels = tissuelevel),
-        treatment = factor(treatment, levels = charlevelsnocontrol)) %>%
+  mutate(tissue = factor(tissue, levels = charlevels),
+        treatment = factor(treatment, levels = charlevels)) %>%
   drop_na()
+head(candidatevsd)
 
@@
   arrange(padj) 
+hilogenes <- hilodf %>%  top_n(20) %>% pull(gene)
 
@@
   arrange(padj)  
+eggchickgenes <- eggchickdf %>%  top_n(20) %>% pull(gene)
 
-eggchickgenes <- eggchickdf %>%  top_n(20) %>% pull(gene)
-hilogenes <- hilodf %>%  top_n(20) %>% pull(gene)
-
-## for correlations
+## data-driven PRL genes
 pithilo <- makecorrdf("female", "pituitary", hilogenes)
@@
 
-## for scatter correlation
+## https://www.gynecologiconcology-online.net/article/S0090-8258(19)30069-1/fulltext
+cancergenes <- makecorrdf("female", "pituitary", c(shaidgenes,suszynskaagenes))
 
+## for scatter correlation (not currently usesed)
 candidatevsdwide <- candidatevsd  %>%
@@
 c1 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 lo vs. hi PRL DEGs") 
-c2 <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes from WGCNA module with PRL") + theme(legend.position = "right")
-c3 <- candidateboxplot("pituitary", c("KPNA2"), sexlevels)
-c4 <- candidateboxplot("pituitary", c("CDK1"), sexlevels) + 
-  theme(axis.text.x = element_text(angle = 45, hjust =1), strip.text = element_blank())  + labs(x = " ")
-c34 <- plot_grid(c3,c4, nrow = 2, rel_heights = c(1,1.2))
-
-c <- plot_grid(c1,c2,c34, nrow = 1, labels = c("D", "", "E"), label_size = 8, rel_widths = c(1.1,1.4,1))
+c2 <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes from WGCNA module with PRL") 
+c3 <- plotcorrplot(cancergenes, NULL) + theme(legend.position = "right") + labs(subtitle = "Breast, ovary, piuitary cancer genes") 
+c <- plot_grid(c1,c2,c3, nrow = 1, labels = c("D", "E", "F"), label_size = 8, rel_widths = c(1,1,1.4))
+c
 
@@
 
-pdf(file="../figures/fig3-1.pdf", width=7.25, height=7.25)
-plot(fig3)
-dev.off()
 ```
@@
 ```{r savefiles}
-write.csv(PRLpit,"../results/PRLvsd.csv", row.names = F)
+#write.csv(PRLpit,"../results/PRLvsd.csv", row.names = F)
 
+#pdf(file="../figures/fig3-1.pdf", width=7.25, height=7.25)
+#plot(fig3)
+#dev.off()
 ```
+
+```{r forvictoria, eval = F}
+## PRL and PRLR in three tissues
+
+candidatevsd <- candidatevsd %>%
+  filter(treatment %in%  c("early", "extend", charlevels))
+
+p1 <- candidateboxplot("hypothalamus", c("PRL"), sexlevels) + labs(subtitle = "hypothalamus", title = "PRL")
+p2 <- candidateboxplot("pituitary", c("PRL"), sexlevels) + labs(subtitle = "pituitary")
+p3 <- candidateboxplot("gonad", c("PRL"), sexlevels) + labs(subtitle = "gonads") + theme(axis.text.x = element_text(angle = 45))
+
+p4 <- candidateboxplot("hypothalamus", c("PRLR"), sexlevels) + labs(subtitle = " ", title = " ") 
+p5 <- candidateboxplot("pituitary", c("PRLR"), sexlevels) + labs(subtitle = " ")
+p6 <- candidateboxplot("gonad", c("PRLR"), sexlevels) + theme(axis.text.x = element_text(angle = 45))+ labs(subtitle = " ")
+
+plot_grid(p1,p4,p2,p5,p3,p6, ncol = 2, rel_heights = c(1.2,1,1.2))
+```

### 2020-05-25 analysis/figure-3.Rmd
fig 3 working now

@@
 colData <- read_csv("../metadata/04_colData.csv") %>%
-  mutate(treatment = factor(treatment, levels = alllevels),
-         tissue = factor(tissue, levels = tissuelevels),
+  filter(treatment %in% charlevels) %>%
+  mutate(tissue = factor(tissue, levels = tissuelevels),
+         treatment = factor(treatment, levels = charlevels),
          temprowname = V1) %>% 
@@
 # pca relies on count data from limma
-
 pca1f <- subsetmakepca("pituitary", charlevels, "female")	
@@
 ```{r}
+# all genes lists in "./R/genelists.R"
+
 # candidate counts for boxplots
 candidatevsd <- read_csv("../results/06_candidatevsd.csv") %>%
+  filter(treatment %in% charlevels) %>%
   mutate(treatment = factor(treatment)) %>%
-  mutate(tissue = factor(tissue, levels = charlevels),
+  mutate(tissue = factor(tissue, levels = tissuelevel),
         treatment = factor(treatment, levels = charlevels)) %>%
@@
   mutate(direction = factor(direction, levels = hypothesislevels))
-
-# candidate genes from GO and literature
-parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% 
-  distinct(literature)  %>% drop_na() %>% pull(literature)
-
-## genes WGCNA prl module
-WGCNAgenes <- read_csv("../results/05_PRLmodule.csv") %>% pull(x)
 
@@
 pithilo <- makecorrdf("female", "pituitary", hilogenes)
-pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[1:20])
+pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[32:52])
 
 ## https://www.gynecologiconcology-online.net/article/S0090-8258(19)30069-1/fulltext
-cancergenes <- makecorrdf("female", "pituitary", c(shaidgenes,suszynskaagenes))
+cancergenes <- makecorrdf("female", "pituitary", c(shaidgenes,suszynskaagenes, "PRL"))
 
@@
 
-c1 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 lo vs. hi PRL DEGs") 
-c2 <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes from WGCNA module with PRL") 
-c3 <- plotcorrplot(cancergenes, NULL) + theme(legend.position = "right") + labs(subtitle = "Breast, ovary, piuitary cancer genes") 
-c <- plot_grid(c1,c2,c3, nrow = 1, labels = c("D", "E", "F"), label_size = 8, rel_widths = c(1,1,1.4))
-c
+c1 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 DEGs, low vs. high PRL") 
+c2 <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes in WGCNA module with PRL") 
+c3 <- plotcorrplot(cancergenes, NULL) + theme(legend.position = "right") + labs(subtitle = "Breast, ovary, & piuitary cancer genes") 
+c <- plot_grid(c1,c2,c3, nrow = 1, labels = c("D", "E", "F"), label_size = 8, rel_widths = c(1,1,1.3))
 
@@
 fig3
-
 ```
@@
 ```{r savefiles}
-#write.csv(PRLpit,"../results/PRLvsd.csv", row.names = F)
+write.csv(PRLpit,"../results/PRLvsd.csv", row.names = F)
 
-#pdf(file="../figures/fig3-1.pdf", width=7.25, height=7.25)
-#plot(fig3)
-#dev.off()
+pdf(file="../figures/fig3-1.pdf", width=7.25, height=7.25)
+plot(fig3)
+dev.off()
 ```

### 2020-05-25 R/genelists.R

@@ -2,17 +2,15 @@
 
 ## candidate genes from GO and literature
 parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% pull(gene)
-parentalcaregenes
 
 ## genes WGCNA prl module
 WGCNAgenes <- read_csv("../results/05_PRLmodule.csv") %>% pull(x)
-WGCNAgenes
 
 ## candidate genes from breast and ovarian cancers 
 ## https://www.gynecologiconcology-online.net/article/S0090-8258(19)30069-1/fulltext
-suszynskaagenes <- c("BRCA1","BRCA2", "CDKN2A", "PTEN", "PALB2", "TP53", "CDH1", "ATM", "BARD1", "MESH6",
-                     "MSH2", "BRIP1", "NBN", "FANCC", "FANCM", "RAD51C","RAD51D")
+suszynskaagenes <- c("BRCA1","BRCA2", "CDKN2A", "PTEN", "PALB2", "TP53", "CDH1", "ATM",
+                     "BARD1", "MESH6","MSH2", "BRIP1", "NBN", "FANCC", "FANCM", 
+                     "RAD51C","RAD51D")
 
-shaidgenes <- c("GNAS", "USB8", "PIK3CA", "GPR101","RAS",
-                "MEN1", "AIP", "DICER1", "PRKAR1A", "PRKACA", 
-                "SDH", "GPR101")
+shaidgenes <- c("GNAS", "USB8", "PIK3CA", "GPR101","RAS","MEN1", "AIP", "DICER1", 
+                "PRKAR1A", "PRKACA","SDH", "GPR101")

### 2020-05-26 analysis/figure-3.Rmd
better fig 3

@@
 
+## number of genes total vs plotted
+
+eggchickdf %>% filter(tissue == "pituitary") %>% nrow()
+hilodf %>% filter(tissue == "pituitary") %>% nrow()
+length(WGCNAgenes)
 ```
@@
 c3 <- plotcorrplot(cancergenes, NULL) + theme(legend.position = "right") + labs(subtitle = "Breast, ovary, & piuitary cancer genes") 
-c <- plot_grid(c1,c2,c3, nrow = 1, labels = c("D", "E", "F"), label_size = 8, rel_widths = c(1,1,1.3))
+c <- plot_grid(c1,c2,c3, nrow = 1, labels = c("D"), label_size = 8, rel_widths = c(1,1,1.3))
 

### 2020-05-26 analysis/figure-4.Rmd
progress on fig 4

@@
 ---
-title: "Fig5"
+title: "Fig 4 and Supple figs"
 output: md_document
@@
 source("../R/functions.R")
-#source("../R/wrangledata.R")
+source("../R/genelists.R")
 
@@
 # Figure 1. Experimental design, tSNE analysis, PCA, and prolactin
+
+## import limma counts and sample info
+
+```{r data}
+countData <- read_csv("../results/01_limma.csv") %>%
+  column_to_rownames(var = "X1")
+countData <- as.data.frame(t(countData))
+head(countData[1:3])
+
+colData <- read_csv("../metadata/00_colData.csv") %>%
+  mutate(treatment = factor(treatment, levels = alllevels),
+         tissue = factor(tissue, levels = tissuelevels)) %>% 
+  column_to_rownames(var = "X1") 
+head(colData)
+# check ready for analysis
+# row.names(countData) == row.names(colData)
+head(row.names(countData) == row.names(colData))
+```
+
+## tsne
 
@@
 
-  df <- df %>% mutate(hiloPRL = fct_collapse(treatment, 
-                                  lo = c("m.inc.d8", "m.inc.d3", "m.inc.d9", "inc.d3", "inc.d9", "lay", "bldg"),
-                                  hi = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" ,  "m.n2", "extend", "n5")),
+  df <- df %>% mutate(earlylate = fct_collapse(treatment, 
+                                  early = c("early", "m.inc.d3", "m.inc.d9", "inc.d3", 
+                                         "inc.d9", "lay", "bldg"),
+                                  late = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" , 
+                                         "m.n2", "extend", "n5")),
                 extint = fct_collapse(treatment, 
                                   eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
-                                  chicks = c( "m.inc.d8", "hatch", "extend", "n5"),
+                                  chicks = c( "early", "hatch", "extend", "n5"),
                                   loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")))
@@
 ```{r vsd}
-vsd_path <- "../results/DEseq2/manip/"   # path to the data
+vsd_path <- "../results/DEseq2/hypothesis/"   # path to the data
 vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
@@
 vsd_files
+
+## before pivoting, check names of df with
+## head(names(allvsd)) and tail(names(allvsd))
 
@@
   dplyr::rename("gene" = "X1") %>% 
-  pivot_longer(cols = blk.s031.pu.d_female_gonad_prolong:y98.o50.x_male_pituitary_inc.d3, 
+  pivot_longer(cols = blk.s061.pu.y_female_gonad_inc.d9:y98.o50.x_male_pituitary_inc.d3, 
                names_to = "samples", values_to = "counts") 
@@
 
+allvsd$samples
+```
+
+```{r candidatevsd}
+
 # for plotting gene expression over time
-getcandidatevsdmanip <- function(whichgenes, whichtissue, whichsex){
+getcandidatevsdmanip <- function(whichgenes, whichtissue){
+
   candidates  <- allvsd %>%
     filter(gene %in% whichgenes) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/manip/'), "[", 2)) %>%
-    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1),
+                  sextissue = sapply(strsplit(sextissue, '../results/DEseq2/hypothesis/'), "[", 2),
+                  sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                 tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
-                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
-    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
-     dplyr::mutate(treatment = recode(treatment, "m.hatch" = "m.n2",
-                                                  "extend.hatch" = "m.n2",
-                                      "inc.prolong" = "prolong")) %>%
-
-    dplyr::mutate(day = fct_collapse(treatment,
-                                     "1" = c("inc.d3", "m.inc.d3" ),
-                                     "2" = c("inc.d9", "m.inc.d9", "m.inc.d8"  ),
-                                     "2" = c("inc.d17", "m.inc.d17", "prolong"  ),
-                                     "4" = c("hatch", "m.n2", "extend"  )
-                                     )) %>%
-    dplyr::mutate(day = as.numeric(day)) %>%
-    dplyr::mutate(manip = fct_collapse(treatment,
-                                       "reference" = charlevels,
-                                       "removal" = c("m.inc.d3", "m.inc.d9", "m.inc.d17","m.n2"),
-                                       "replacement" = c("m.inc.d8", "prolong", "extend"))) %>%
-    dplyr::mutate(manip = factor(manip, levels = c("reference", "removal", "replacement"))) %>%
-    dplyr::select(sex, tissue, treatment, day,  manip, gene, samples, counts) %>%
-    filter(tissue == whichtissue, sex %in% whichsex)  %>%
-    drop_na()
-  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
+                treatment = sapply(strsplit(samples, '\\_'), "[", 4),
+                treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1),
+                treatment = recode(treatment, "m.hatch" = "m.n2", "extend.hatch" = "m.n2",
+                                                   "inc.prolong" = "prolong", "m.inc.d8" = "early"),
+                treatment =  factor(candidates$treatment, levels = alllevels),
+                  earlylate = fct_collapse(treatment, 
+                                  early = c("early", "m.inc.d3", "m.inc.d9", "inc.d3", 
+                                         "inc.d9", "lay", "bldg"),
+                                  late = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" , 
+                                         "m.n2", "extend", "n5")),
+                extint = fct_collapse(treatment, 
+                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
+                                  chicks = c( "early", "hatch", "extend", "n5"),
+                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")),
+                manip = factor(manip, levels = c("reference", "removal", "replacement")))
   return(candidates)
 } 
+  %>%
+    dplyr::select(gene, counts, sex, tissue, treatment, day, manip, earlylate, extint, samples)  %>%
+    dplyr::filter(tissue == whichtissue) 
+  return(candidates)
 
-hypvsd <- getcandidatevsdmanip(candidategenes, "hypothalamus", sexlevels)
-pitvsd <- getcandidatevsdmanip(candidategenes, "pituitary", sexlevels)
-gonvsd <- getcandidatevsdmanip(candidategenes, "gonad", sexlevels)
+hypvsd <- getcandidatevsdmanip(c("PRL", "DRD1"), "hypothalamus")
+pitvsd <- getcandidatevsdmanip(candidategenes, "pituitary")
+gonvsd <- getcandidatevsdmanip(candidategenes, "gonad")
 candidatevsd <- rbind(hypvsd, pitvsd)
@@
 
-hypvsd2 <- getcandidatevsdmanip(hypothesisDEGs, "hypothalamus", sexlevels)
-pitvsd2 <- getcandidatevsdmanip(hypothesisDEGs, "pituitary", sexlevels)
-gonvsd2 <- getcandidatevsdmanip(hypothesisDEGs, "gonad", sexlevels)
-candidatevsd2 <- rbind(hypvsd2, pitvsd2)
-candidatevsd2 <- rbind(candidatevsd2, gonvsd2)
-head(candidatevsd2) 
+```
+
+```{r timingfigure, fig.height= 7.25in}
+
+plotcandidatetiming <- function(df, whichgene, whichtissue){
+
+  dfslim <- df %>%
+    dplyr::filter(gene  == whichgene, treatment %in% timing) 
+  print(head(dfslim))
+
+  p <- dfslim %>%
+    ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
+    geom_boxplot(outlier.shape = NA) +
+    geom_jitter(size = 0.25, aes(color = sex)) +
+    theme_B3() +
+    theme(legend.position = "none",
+          #axis.text.x = element_blank(),
+          axis.title.x = element_blank(),
+          plot.subtitle = element_text(face = "italic")) +
+    scale_color_manual(values = allcolors) +
+    scale_fill_manual(values = allcolors)  +
+    labs(title = paste(whichtissue, sep = " "), subtitle = whichgene, y = "gene expression") +
+    facet_wrap(~sex)  +
+    geom_signif(comparisons = list(c( "inc.d9", "early"),
+                                   c( "inc.d17", "prolong"),
+                                   c( "hatch", "prolong"),
+                                   c( "hatch", "extend")),
+                map_signif_level=TRUE,
+                textsize = 3, family = 'Helvetica',
+                vjust = 1.5, size = 0)
+  return(p)
+
+}
+
+a <- plotcandidatetiming(hypvsd, "DRD1", "hypothalamus")
+b <- plotcandidatetiming(pitvsd, "PRL", "pituitary")
+c <- plotcandidatetiming(gonvsd, "CRHBP", "gonads")
+
+plot_grid(a,b,c, nrow = 3)
+
 ```
@@
   theme(axis.text.y = element_blank(), axis.title.y = element_blank())
-c3 <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + 
+c3 <- plottsneelipsev2(hyptsne, hyptsne$earlylate, allcolors ) + 
   labs(subtitle = " ", x = NULL) + 
@@
   theme(axis.text.y = element_blank(), axis.title.y = element_blank()) 
-c6 <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + 
+c6 <- plottsneelipsev2(pittsne, pittsne$earlylate, allcolors ) + 
   labs(subtitle = " ", x = NULL) + 
@@
   theme(axis.text.y = element_blank(), axis.title.y = element_blank())
-c9 <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + 
+c9 <- plottsneelipsev2(gontsne, gontsne$earlylate, allcolors ) + 
   labs(subtitle = " " ) + 
@@
                                      "3" = c("inc.d3", "m.inc.d3" ),
-                                     "9" = c("inc.d9", "m.inc.d9", "m.inc.d8"  ),
+                                     "9" = c("inc.d9", "m.inc.d9", "early"  ),
                                      "17" = c("inc.d17", "m.inc.d17", "prolong"  ),

### 2020-05-26 analysis/figure-4.Rmd
now with screenshot of musical genes

@@
 head(allvsd)
-
-allvsd$samples
 ```
@@
                                                    "inc.prolong" = "prolong", "m.inc.d8" = "early"),
-                treatment =  factor(candidates$treatment, levels = alllevels),
+                treatment =  factor(treatment, levels = alllevels),
                 earlylate = fct_collapse(treatment, 
@@
                                   chicks = c( "early", "hatch", "extend", "n5"),
-                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")),
-                manip = factor(manip, levels = c("reference", "removal", "replacement")))
+                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg"))) %>%
+    dplyr::select(gene, counts, sex, tissue, treatment, earlylate, extint, samples)  %>%
+    dplyr::filter(tissue == whichtissue) 
   return(candidates)
 }
-  %>%
-    dplyr::select(gene, counts, sex, tissue, treatment, day, manip, earlylate, extint, samples)  %>%
-    dplyr::filter(tissue == whichtissue) 
-  return(candidates)
 
-hypvsd <- getcandidatevsdmanip(c("PRL", "DRD1"), "hypothalamus")
+hypvsd <- getcandidatevsdmanip(candidategenes, "hypothalamus")
 pitvsd <- getcandidatevsdmanip(candidategenes, "pituitary")

### 2020-05-26 analysis/figure-4.Rmd
working on removal, gonads have missing data

@@
 ---
-title: "Fig 4 and Supple figs"
+title: "Removal only"
 output: md_document
@@
 
-knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
+knitr::opts_chunk$set(echo = TRUE, message = F, warning = F,
+                      fig.path = '../figures/', fig.width = 7.25, fig.height = 7.25)
 ```
 
-# Figure 1. Experimental design, tSNE analysis, PCA, and prolactin
+# Figure 4. What is the effect of removal
 
@@
 # prep for tsne
-
-hyptsne <- subsetmaketsne("hypothalamus", allmaniplevels, sexlevels)
-pittsne <- subsetmaketsne("pituitary", allmaniplevels, sexlevels)
-gontsne <- subsetmaketsne("gonads", allmaniplevels, sexlevels)
+hyptsne <- subsetmaketsne("hypothalamus", removallevels, sexlevels)
+pittsne <- subsetmaketsne("pituitary", removallevels, sexlevels)
+gontsne <- subsetmaketsne("gonads", removallevels, sexlevels)
 
@@
                                   early = c("early", "m.inc.d3", "m.inc.d9", "inc.d3", 
-                                         "inc.d9", "lay", "bldg"),
+                                         "inc.d9",  "bldg"),
                                   late = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" , 
-                                         "m.n2", "extend", "n5")),
+                                         "m.n2", "extend", "n5"),
+                                  reference = "control"),
                 extint = fct_collapse(treatment, 
-                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
-                                  chicks = c( "early", "hatch", "extend", "n5"),
-                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")))
-  df$extint <- factor(df$extint, levels = levelsextint)
+                                  nest = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg"),
+                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong"),
+                                  chicks = c("early", "hatch", "extend", "n5"),
+                                  reference = "control")) 
   return(df)
@@
 gontsne <- addgroupings(gontsne)
+head(gontsne)
 
+t1 <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) +  labs(subtitle = "Hypothalamus ") +
+  facet_wrap(~sex) + theme(axis.title.x = element_blank())
+t2 <- plottsneelipsev2(hyptsne, hyptsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
+  facet_wrap(~sex) + theme(axis.title = element_blank())
+t3 <- plottsneelipsev2(hyptsne, hyptsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
+  facet_wrap(~sex) + theme(axis.title = element_blank())
+t123 <- plot_grid(t1, t2, t3, nrow = 1, rel_widths = c(1.1,1,1))
+
+t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(subtitle = "Pituitary ") +
+  facet_wrap(~sex) + theme(axis.title.x = element_blank(), strip.text = element_blank())
+t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
+  facet_wrap(~sex) + theme(axis.title = element_blank(), strip.text = element_blank())
+t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
+  facet_wrap(~sex) + theme(axis.title = element_blank(), strip.text = element_blank())
+t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
+
+t7 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors) +  labs(subtitle = "Gonads ") +
+  facet_wrap(~sex) + theme(strip.text = element_blank())
+t8 <- plottsneelipsev2(gontsne, gontsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank(), strip.text = element_blank())
+t9 <- plottsneelipsev2(gontsne, gontsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank(), strip.text = element_blank())
+t789 <- plot_grid(t7, t8, t9, nrow = 1)
+
+tsneplots <- plot_grid(t123, t456,t789, ncol = 1, rel_widths = c(1.1,1,1))
+tsneplots
 ```
@@
                                   early = c("early", "m.inc.d3", "m.inc.d9", "inc.d3", 
-                                         "inc.d9", "lay", "bldg"),
+                                         "inc.d9",  "bldg"),
                                   late = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" , 
-                                         "m.n2", "extend", "n5")),
+                                         "m.n2", "extend", "n5"),
+                                  reference = "control"),
                 extint = fct_collapse(treatment, 
-                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
-                                  chicks = c( "early", "hatch", "extend", "n5"),
-                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg"))) %>%
-    dplyr::select(gene, counts, sex, tissue, treatment, earlylate, extint, samples)  %>%
-    dplyr::filter(tissue == whichtissue) 
+                                  nest = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg"),
+                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong"),
+                                  chicks = c("early", "hatch", "extend", "n5"),
+                                  reference = "control"))  %>%
+    dplyr::select(gene, counts, sex, tissue, treatment, earlylate, extint, samples)  
   return(candidates)
@@
 
-hypvsd <- getcandidatevsdmanip(candidategenes, "hypothalamus")
-pitvsd <- getcandidatevsdmanip(candidategenes, "pituitary")
-gonvsd <- getcandidatevsdmanip(candidategenes, "gonad")
-candidatevsd <- rbind(hypvsd, pitvsd)
-candidatevsd <- rbind(candidatevsd, gonvsd)
-head(candidatevsd)
+candidatevsd <- getcandidatevsdmanip(candidategenes)
+hypvsd <- candidatevsd %>% filter(tissue %in% "hypothalamus") %>% drop_na()
+pitvsd <- candidatevsd %>% filter(tissue %in% "pituitary")%>% drop_na()
+gonvsd <- candidatevsd %>% filter(tissue %in% "gonad")%>% drop_na()
 
@@
 
-plotcandidatetiming <- function(df, whichgene, whichtissue){
+plotcandidateremoval <- function(df, whichgene, whichtissue){
 
-  dfslim <- df %>%
-    dplyr::filter(gene  == whichgene, treatment %in% timing) 
-  print(head(dfslim))
-
-  p <- dfslim %>%
+  p <- df %>%
+    dplyr::filter(gene  == whichgene, treatment %in% removallevels)  %>%
     ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
@@
     theme(legend.position = "none",
-          #axis.text.x = element_blank(),
           axis.title.x = element_blank(),
@@
     labs(title = paste(whichtissue, sep = " "), subtitle = whichgene, y = "gene expression") +
-    facet_wrap(~sex)  +
-    geom_signif(comparisons = list(c( "inc.d9", "early"),
-                                   c( "inc.d17", "prolong"),
-                                   c( "hatch", "prolong"),
-                                   c( "hatch", "extend")),
+    #facet_wrap(~sex)  +
+    geom_signif(comparisons = list(c( "inc.d3", "m.inc.d3"),
+                                   c( "inc.d9", "m.inc.d9"),
+                                   c( "inc.d17", "m.inc.d17"),
+                                   c( "hatch", "m.n2")),
                 map_signif_level=TRUE,
                 textsize = 3, family = 'Helvetica',
-                vjust = 1.5, size = 0)
+                vjust = 1.5, size = 0.5)
   return(p)
@@
 
-a <- plotcandidatetiming(hypvsd, "DRD1", "hypothalamus")
-b <- plotcandidatetiming(pitvsd, "PRL", "pituitary")
-c <- plotcandidatetiming(gonvsd, "CRHBP", "gonads")
+a <- plotcandidateremoval(hypvsd, "DRD1", "hypothalamus") + theme(axis.title.x = element_blank())
+b <- plotcandidateremoval(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank())
+c <- plotcandidateremoval(gonvsd, "CRHBP", "gonads")
 

### 2020-05-26 analysis/figure-4.Rmd
better with vsd from treatment

@@
 
-# Figure 4. What is the effect of removal
+# Figure 4. What is the effect of removal?
 
@@
 
-hyptsne <- addgroupings(hyptsne)
 pittsne <- addgroupings(pittsne)
-gontsne <- addgroupings(gontsne)
-head(gontsne)
 
-t1 <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) +  labs(subtitle = "Hypothalamus ") +
-  facet_wrap(~sex) + theme(axis.title.x = element_blank())
-t2 <- plottsneelipsev2(hyptsne, hyptsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
-  facet_wrap(~sex) + theme(axis.title = element_blank())
-t3 <- plottsneelipsev2(hyptsne, hyptsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
-  facet_wrap(~sex) + theme(axis.title = element_blank())
-t123 <- plot_grid(t1, t2, t3, nrow = 1, rel_widths = c(1.1,1,1))
+t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "Pituitary ", subtitle = "parental stage") +
+  facet_wrap(~sex) 
+t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(title = "  ", subtitle = "external enviornment")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal timing")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
+t456
 
-t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(subtitle = "Pituitary ") +
-  facet_wrap(~sex) + theme(axis.title.x = element_blank(), strip.text = element_blank())
-t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
-  facet_wrap(~sex) + theme(axis.title = element_blank(), strip.text = element_blank())
-t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
-  facet_wrap(~sex) + theme(axis.title = element_blank(), strip.text = element_blank())
-t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
-
-t7 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors) +  labs(subtitle = "Gonads ") +
-  facet_wrap(~sex) + theme(strip.text = element_blank())
-t8 <- plottsneelipsev2(gontsne, gontsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
-  facet_wrap(~sex) + theme(axis.title.y = element_blank(), strip.text = element_blank())
-t9 <- plottsneelipsev2(gontsne, gontsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
-  facet_wrap(~sex) + theme(axis.title.y = element_blank(), strip.text = element_blank())
-t789 <- plot_grid(t7, t8, t9, nrow = 1)
-
-tsneplots <- plot_grid(t123, t456,t789, ncol = 1, rel_widths = c(1.1,1,1))
-tsneplots
 ```
@@
 ```{r vsd}
-vsd_path <- "../results/DEseq2/hypothesis/"   # path to the data
+vsd_path <- "../results/DEseq2/treatment/"   # path to the data
 vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
@@
 
-## before pivoting, check names of df with
-## head(names(allvsd)) and tail(names(allvsd))
-
 allvsd <- vsd_pathfiles %>%
@@
   map_df(~read_csv(.x), .id = "file_name")  %>% 
-  dplyr::rename("gene" = "X1") %>% 
-  pivot_longer(cols = blk.s061.pu.y_female_gonad_inc.d9:y98.o50.x_male_pituitary_inc.d3, 
+  dplyr::rename("gene" = "X1") 
+## before pivoting, check names of df with
+## head(names(allvsd)) and tail(names(allvsd))
+allvsd <- allvsd %>% 
+  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
                names_to = "samples", values_to = "counts") 
 allvsd <- as_tibble(allvsd)
+head(allvsd)
+tail(allvsd)
 
-head(allvsd)
 ```
@@
     dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1),
-                  sextissue = sapply(strsplit(sextissue, '../results/DEseq2/hypothesis/'), "[", 2),
+                  sextissue = sapply(strsplit(sextissue, '../results/DEseq2/treatment/'), "[", 2),
                   sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
@@
 candidatevsd <- getcandidatevsdmanip(candidategenes)
-hypvsd <- candidatevsd %>% filter(tissue %in% "hypothalamus") %>% drop_na()
-pitvsd <- candidatevsd %>% filter(tissue %in% "pituitary")%>% drop_na()
-gonvsd <- candidatevsd %>% filter(tissue %in% "gonad")%>% drop_na()
+hypvsd <- candidatevsd %>% filter(tissue %in% "hypothalamus")  
+pitvsd <- candidatevsd %>% filter(tissue %in% "pituitary") 
+gonvsd <- candidatevsd %>% filter(tissue %in% "gonad")
 
+# not all genes appear to be expressed in the gonads... these are
+filter(gonvsd, treatment == "control") %>% head()
 ```
@@
   p <- df %>%
-    dplyr::filter(gene  == whichgene, treatment %in% removallevels)  %>%
+    dplyr::filter(gene  == whichgene, 
+                  treatment %in% removallevels)  %>%
     ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
@@
     labs(title = paste(whichtissue, sep = " "), subtitle = whichgene, y = "gene expression") +
-    #facet_wrap(~sex)  +
     geom_signif(comparisons = list(c( "inc.d3", "m.inc.d3"),
@@
 
-a <- plotcandidateremoval(hypvsd, "DRD1", "hypothalamus") + theme(axis.title.x = element_blank())
-b <- plotcandidateremoval(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank())
-c <- plotcandidateremoval(gonvsd, "CRHBP", "gonads")
+a <- plotcandidateremoval(hypvsd, "HTR2C", "hypothalamus") + theme(axis.title.x = element_blank())
+b <- plotcandidateremoval(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank(), strip.text = element_blank())
+c <- plotcandidateremoval(gonvsd, "ESR2", "gonads") + theme(strip.text = element_blank())
+boxplots1 <- plot_grid(a,b,c, nrow = 3)
+boxplots1
 
-plot_grid(a,b,c, nrow = 3)
+plotcandidatetiming <- function(df, whichgene, whichtissue){
 
+  p <- df %>%
+    dplyr::filter(gene  == whichgene, 
+                  treatment %in% timinglevels,
+                  !treatment %in%  c("control", "bldg", "inc.d3"))  %>%
+    ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
+    geom_boxplot(outlier.shape = NA) +
+    geom_jitter(size = 0.25, aes(color = sex)) +
+    theme_B3() +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          plot.subtitle = element_text(face = "italic")) +
+    scale_color_manual(values = allcolors) +
+    scale_fill_manual(values = allcolors)  +
+    labs(title = paste(whichtissue, sep = " "), subtitle = whichgene, y = "gene expression") +
+    geom_signif(comparisons = list(c("inc.d9", "early"), 
+                                  c( "inc.d17", "prolong"), c( "prolong", "hatch"),
+                                  c( "hatch", "extend"), c( "n5", "extend")),
+                map_signif_level=TRUE,
+                textsize = 3, family = 'Helvetica',
+                vjust = 1.5, size = 0.5) 
+  return(p)
+
+}
+
+e <- plotcandidatetiming(hypvsd, "HTR2C", "hypothalamus") + theme(axis.title.x = element_blank())
+f <- plotcandidatetiming(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank(), strip.text = element_blank())
+g <- plotcandidatetiming(gonvsd, "ESR2", "gonads") + theme(strip.text = element_blank())
+boxplots2 <- plot_grid(e,f,g, nrow = 3)
+boxplots2
 ```
@@
 ```{r allDEGs}
-DEG_path <- "../results/DEseq2/manip/"   # path to the data
+DEG_path <- "../results/DEseq2/treatment/"   # path to the data
 DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
@@
   map_df(~read_csv(.x), .id = "file_name") %>% 
-  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/manip/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
   mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
@@
   mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
-dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
-head(allDEG)
-
-allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
-allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
-allDEG$direction <- factor(allDEG$direction, levels = alllevels)
+  dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) # %>%
+ # mutate(tissue = factor(tissue, levels = tissuelevel),
+  #       comparison = factor(comparison , levels = comparisonlevelsremoval),
+  #       direction = factor(direction, levels = alllevels))
 head(allDEG)
@@
                          "hypothalamus" = "H",
-                         "pituitary" = "P", "gonad" = "G")) %>%
+                         "pituitary" = "P", 
+                         "gonad" = "G", "gonads" = "G")) %>%
   mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
@@
   pivot_wider(names_from = comparison, values_from = res) %>%
-  select(gene, contains("m.")) 
+  select(gene, inc.d3_m.inc.d3, inc.d9_m.inc.d9, inc.d17_m.inc.d17, hatch_m.n2,
+          inc.d17_prolong, hatch_prolong)
 candidateDEGS
 
-hypothesisDEGStable <- allDEG %>%
-  filter(gene %in% hypothesisDEGs) %>%
-  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
-         sex = recode(sex, "female" = "F", "male" = "M" ),
-         tissue = recode(tissue, 
-                         "hypothalamus" = "H",
-                         "pituitary" = "P", "gonad" = "G")) %>%
-  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
-  select(gene, res, comparison)  %>%
-  group_by(gene,  comparison) %>%
-  summarize(res = str_c(res, collapse = " ")) %>%
-  pivot_wider(names_from = comparison, values_from = res) %>%
-  select(gene, contains("m.")) 
-hypothesisDEGStable
 ```
@@
 ```
+
+## supplementary
+
+```{r}
+
+hyptsne <- addgroupings(hyptsne)
+gontsne <- addgroupings(gontsne)
+
+t1 <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) +  labs(subtitle = "Hypothalamus ") +
+  facet_wrap(~sex) + theme(axis.title.x = element_blank())
+t2 <- plottsneelipsev2(hyptsne, hyptsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
+  facet_wrap(~sex) + theme(axis.title = element_blank())
+t3 <- plottsneelipsev2(hyptsne, hyptsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
+  facet_wrap(~sex) + theme(axis.title = element_blank())
+t123 <- plot_grid(t1, t2, t3, nrow = 1, rel_widths = c(1.1,1,1))
+
+t7 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors) +  labs(subtitle = "Gonads ") +
+  facet_wrap(~sex) + theme(strip.text = element_blank())
+t8 <- plottsneelipsev2(gontsne, gontsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank(), strip.text = element_blank())
+t9 <- plottsneelipsev2(gontsne, gontsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank(), strip.text = element_blank())
+t789 <- plot_grid(t7, t8, t9, nrow = 1)
+
+```

### 2020-05-26 analysis/figure-4.Rmd
working but not ready

@@
 library(ggsignif)
+library(corrr)
 
@@
 candidateDEGS <- allDEG %>%
-  filter(gene %in% candidategenes) %>%
+  filter(gene %in% parentalcaregenes) %>%
   mutate(posneg = ifelse(lfc >= 0, "+", "-"),
@@
 
-b1 <- plotcandidatemanipquad(candidatevsd, "hypothalamus", c("DRD1")) 
-b2 <- plotcandidatemanipquad(candidatevsd,"pituitary", c("PRL")) 
-b3 <- plotcandidatemanipquad(candidatevsd,"gonad", c("PGR"))
+boxplots <- plot_grid(boxplots1, boxplots2, nrow = 1, rel_widths = c(1.5,1))
+plot_grid(boxplots, t456, a, ncol = 1, rel_heights = c(3,1,1))
 
-b <- plot_grid(b1,b2,b3, ncol = 1)
+#pdf(file="../figures/fig4-1.pdf", width=7.25, height=7.25)
+#plot(fig4)
+#dev.off()
 
-c1 <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
-  labs(subtitle = "Hypothalamus ", x = NULL)  
-c2 <- plottsneelipsev2(hyptsne, hyptsne$extint, allcolors ) + 
-  labs(subtitle = " ", x = NULL) + 
-  theme(axis.text.y = element_blank(), axis.title.y = element_blank())
-c3 <- plottsneelipsev2(hyptsne, hyptsne$earlylate, allcolors ) + 
-  labs(subtitle = " ", x = NULL) + 
-  theme(axis.text.y = element_blank(), axis.title.y = element_blank()) 
-c4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
-  labs(subtitle = "Pituitary ", x = NULL) 
-c5 <- plottsneelipsev2(pittsne, pittsne$extint, allcolors ) + 
-  labs(subtitle = " ", x = NULL) + 
-  theme(axis.text.y = element_blank(), axis.title.y = element_blank()) 
-c6 <- plottsneelipsev2(pittsne, pittsne$earlylate, allcolors ) + 
-  labs(subtitle = " ", x = NULL) + 
-  theme(axis.text.y = element_blank(), axis.title.y = element_blank())
-c7 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
-  labs(subtitle = "Ggonads ")  
-c8 <- plottsneelipsev2(gontsne, gontsne$extint, allcolors ) + 
-  labs(subtitle = " " )    +
-  theme(axis.text.y = element_blank(), axis.title.y = element_blank())
-c9 <- plottsneelipsev2(gontsne, gontsne$earlylate, allcolors ) + 
-  labs(subtitle = " " ) + 
-  theme(axis.text.y = element_blank(), axis.title.y = element_blank())
+makecorrdf <- function(whichsex, whichtissue, whichgenes){
+  corrrdf <- candidatevsd %>% 
+    filter(sex == whichsex, tissue == whichtissue,
+           gene %in% whichgenes) %>%
+    drop_na() %>%
+    pivot_wider(names_from = gene, values_from = counts) %>%
+    select(-sex, -tissue, -treatment, -samples, -earlylate, -extint) %>%
+    correlate() 
+  print(head(corrrdf))
+  return(corrrdf)
+}
 
-c <- plot_grid(c1,c2,c3,c4,c5,c6,c7,c8,c9, nrow = 3)
+candidatevsd <- candidatevsd %>% filter(treatment %in% maniplevels)
 
-bc <- plot_grid(b, c, nrow = 1, labels = c("", "C"), label_size = 8)
+hyp1 <- makecorrdf(sexlevels, "hypothalamus", curleychampagnegenes)  
+pit1 <- makecorrdf(sexlevels, "pituitary", curleychampagnegenes)  
+gon1 <- makecorrdf(sexlevels, "gonad", curleychampagnegenes)
 
-fig4 <- plot_grid(a, bc, ncol = 1, rel_heights  = c(0.25,1),
-          labels = "AUTO", label_size = 8)
-fig4
+hyp2 <- makecorrdf(sexlevels, "hypothalamus", GOgenes)  
+pit2 <- makecorrdf(sexlevels, "pituitary", GOgenes)  
+gon2 <- makecorrdf(sexlevels, "gonad", GOgenes)
 
-pdf(file="../figures/fig4-1.pdf", width=7.25, height=7.25)
-plot(fig4)
-dev.off()
+hyp3 <- makecorrdf(sexlevels, "hypothalamus", cancergenes)  
+pit3 <- makecorrdf(sexlevels, "pituitary", GOgecancergenesnes)  
+gon3 <- makecorrdf(sexlevels, "gonad", cancergenes)
 
-candidatevsd <- candidatevsd %>%
-  dplyr::mutate(treatment = fct_collapse(treatment,
-                                         "m.n2" = c("m.n2",  "m.hatch" ),
-                                         "prolong" = c("prolong", "inc.prolong") ,
-                                         "extend" = c("extend", "extend.hatch")
-                                         ))  %>%
-  dplyr::mutate(day = fct_collapse(treatment,
-                                    "0" = c("control", "bldg"),
-                                    "1" = c("lay"),
-                                     "3" = c("inc.d3", "m.inc.d3" ),
-                                     "9" = c("inc.d9", "m.inc.d9", "early"  ),
-                                     "17" = c("inc.d17", "m.inc.d17", "prolong"  ),
-                                     "18" = c("hatch", "m.n2" ),
-                                     "22" = c("n5",  "extend" ),
-                                     "26" = c("n9"  )
-                                     )) %>%
-   mutate(treatment = factor(treatment, levels = alllevels)) %>%
-  dplyr::mutate(day = factor(day, levels = c("0", "1", "3", "9", "17", "18", "22", "26")))
+b1 <- plotcorrplot(hyp1, "Literature") + labs(title = "hyp")
+b2 <- plotcorrplot(pit1, NULL)  + labs(title = "pit")
+b3 <- plotcorrplot(gon1, NULL)  + labs(title = "gon")
 
-candidatevsd %>%
-    filter(tissue == "hypothalamus", sex == "female") %>%
-    filter(gene %in% "DRD1") %>%
-    drop_na()  %>%
-    ggplot(aes(y =  counts, x = day, fill = treatment, color = sex)) +
-    geom_boxplot(outlier.shape = NA) +
-    geom_jitter(size = 0.25, aes(color = sex)) +
-    theme_B3() +
-    theme(legend.position = "bottom",
-          #axis.text.x = element_blank(),
-          #axis.title.x = element_blank(),
-          strip.text = element_text(face = "italic")) +
-    scale_color_manual(values = allcolors) +
-    scale_fill_manual(values = allcolors)  +
-    labs(subtitle = paste("female", "hypothalamus", sep = " "), y = "DRD1")
+b4 <- plotcorrplot(hyp2, "GO")  + labs(title = " ")
+b5 <- plotcorrplot(pit2, NULL)   + labs(title = " ")
+b6 <- plotcorrplot(gon2, NULL)  + labs(title = " ")
 
-candidatevsd %>%
-    filter(tissue == "pituitary", sex == "female") %>%
-    filter(gene %in% "PRL") %>%
-    drop_na()  %>%
-    ggplot(aes(y =  counts, x = day, fill = treatment, color = sex)) +
-    geom_boxplot(outlier.shape = NA) +
-    geom_jitter(size = 0.25, aes(color = sex)) +
-    theme_B3() +
-    theme(legend.position = "bottom",
-          #axis.text.x = element_blank(),
-          #axis.title.x = element_blank(),
-          strip.text = element_text(face = "italic")) +
-    scale_color_manual(values = allcolors) +
-    scale_fill_manual(values = allcolors)  +
-    labs(subtitle = paste("female", "pituitary", sep = " "), y = "PRL")
+b7 <- plotcorrplot(hyp3, "Caner") + labs(title = " ")
+b8 <- plotcorrplot(pit3, NULL)   + labs(title = " ")
+b9 <- plotcorrplot(gon3, NULL)  + labs(title = " ")
 
+forlegend <- plotcorrplot(gon3, NULL) + theme(legend.position = "bottom") + labs(color = "Correlation")
+mylegend <- get_legend(forlegend)
+
+allcorplots <- plot_grid(b1,b4,b7,
+          b2,b5,b8,
+          b3,b6,b9)
+
+supplfigX <- plot_grid(allcorplots, mylegend, ncol = 1, rel_heights = c(1,0.1))
+supplfigX
 ```
@@
 t789 <- plot_grid(t7, t8, t9, nrow = 1)
+```
 
+## write files
+
+```{r write}
+#write.csv(candidateDEGS, "../results/table2.csv", row.names = F)
 ```

### 2020-05-26 R/genelists.R

@@ -1,7 +1,10 @@
 # genelists
 
 ## candidate genes from GO and literature
-parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% pull(gene)
+LitGOdf <- read_csv("../metadata/03_parentalcaregenes.csv") 
+curleychampagnegenes <- LitGOdf %>% filter(literature != "NA" ) %>% pull(gene)
+GOgenes <- LitGOdf %>% filter(GO != "NA") %>% pull(gene)
+parentalcaregenes <- LitGOdf %>% pull(gene)
 
 ## genes WGCNA prl module
 WGCNAgenes <- read_csv("../results/05_PRLmodule.csv") %>% pull(x)
@@ -15,4 +18,6 @@ suszynskaagenes <- c("BRCA1","BRCA2", "CDKN2A", "PTEN", "PALB2", "TP53", "CDH1",
 shaidgenes <- c("GNAS", "USB8", "PIK3CA", "GPR101","RAS","MEN1", "AIP", "DICER1", 
                 "PRKAR1A", "PRKACA","SDH", "GPR101")
 
-candidategenes <- c(parentalcaregenes)
+cancergenes <- c(suszynskaagenes, shaidgenes)
+
+candidategenes <- c(parentalcaregenes, WGCNAgenes, cancergenes)

### 2020-05-26 analysis/figure-4.Rmd
removing correlations fig 4

@@
 
-makecorrdf <- function(whichsex, whichtissue, whichgenes){
-  corrrdf <- candidatevsd %>% 
-    filter(sex == whichsex, tissue == whichtissue,
-           gene %in% whichgenes) %>%
-    drop_na() %>%
-    pivot_wider(names_from = gene, values_from = counts) %>%
-    select(-sex, -tissue, -treatment, -samples, -earlylate, -extint) %>%
-    correlate() 
-  print(head(corrrdf))
-  return(corrrdf)
-}
-
-candidatevsd <- candidatevsd %>% filter(treatment %in% maniplevels)
-
-hyp1 <- makecorrdf(sexlevels, "hypothalamus", curleychampagnegenes)  
-pit1 <- makecorrdf(sexlevels, "pituitary", curleychampagnegenes)  
-gon1 <- makecorrdf(sexlevels, "gonad", curleychampagnegenes)
-
-hyp2 <- makecorrdf(sexlevels, "hypothalamus", GOgenes)  
-pit2 <- makecorrdf(sexlevels, "pituitary", GOgenes)  
-gon2 <- makecorrdf(sexlevels, "gonad", GOgenes)
-
-hyp3 <- makecorrdf(sexlevels, "hypothalamus", cancergenes)  
-pit3 <- makecorrdf(sexlevels, "pituitary", GOgecancergenesnes)  
-gon3 <- makecorrdf(sexlevels, "gonad", cancergenes)
-
-b1 <- plotcorrplot(hyp1, "Literature") + labs(title = "hyp")
-b2 <- plotcorrplot(pit1, NULL)  + labs(title = "pit")
-b3 <- plotcorrplot(gon1, NULL)  + labs(title = "gon")
-
-b4 <- plotcorrplot(hyp2, "GO")  + labs(title = " ")
-b5 <- plotcorrplot(pit2, NULL)   + labs(title = " ")
-b6 <- plotcorrplot(gon2, NULL)  + labs(title = " ")
-
-b7 <- plotcorrplot(hyp3, "Caner") + labs(title = " ")
-b8 <- plotcorrplot(pit3, NULL)   + labs(title = " ")
-b9 <- plotcorrplot(gon3, NULL)  + labs(title = " ")
-
-forlegend <- plotcorrplot(gon3, NULL) + theme(legend.position = "bottom") + labs(color = "Correlation")
-mylegend <- get_legend(forlegend)
-
-allcorplots <- plot_grid(b1,b4,b7,
-          b2,b5,b8,
-          b3,b6,b9)
-
-supplfigX <- plot_grid(allcorplots, mylegend, ncol = 1, rel_heights = c(1,0.1))
-supplfigX
 ```

### 2020-05-26 analysis/figure-4.Rmd
done with fig 4

@@
 
-t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "Pituitary ", subtitle = "parental stage") +
+t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "pituitary ", subtitle = "parental stage and manipulations") +
   facet_wrap(~sex) 
@@
   facet_wrap(~sex) + theme(axis.title.y = element_blank())
-t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal timing")   +
+t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal clock")   +
   facet_wrap(~sex) + theme(axis.title.y = element_blank())
 t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
-t456
 
@@
 
-```{r timingfigure, fig.height= 7.25in}
+```{r boxplots}
 
@@
 
-a <- plotcandidateremoval(hypvsd, "HTR2C", "hypothalamus") + theme(axis.title.x = element_blank())
-b <- plotcandidateremoval(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank(), strip.text = element_blank())
+a <- plotcandidateremoval(hypvsd, "HTR2C", "hypothalamus") + theme(axis.title.x = element_blank(), axis.text.x = element_blank())
+b <- plotcandidateremoval(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank(), strip.text = element_blank(), axis.text.x = element_blank())
 c <- plotcandidateremoval(gonvsd, "ESR2", "gonads") + theme(strip.text = element_blank())
 boxplots1 <- plot_grid(a,b,c, nrow = 3)
-boxplots1
 
@@
 
-e <- plotcandidatetiming(hypvsd, "HTR2C", "hypothalamus") + theme(axis.title.x = element_blank())
-f <- plotcandidatetiming(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank(), strip.text = element_blank())
+e <- plotcandidatetiming(hypvsd, "HTR2C", "hypothalamus") + theme(axis.title.x = element_blank(), axis.text.x = element_blank()) 
+f <- plotcandidatetiming(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank(), strip.text = element_blank(), axis.text.x = element_blank())
 g <- plotcandidatetiming(gonvsd, "ESR2", "gonads") + theme(strip.text = element_blank())
 boxplots2 <- plot_grid(e,f,g, nrow = 3)
-boxplots2
 ```
@@
 
-a <- png::readPNG("../figures/images/fig_fig4a.png")
-a <- ggdraw() +  draw_image(a, scale = 1)
+design <- png::readPNG("../figures/images/fig_fig4a.png")
+design <- ggdraw() +  draw_image(design, scale = 1)
 
-boxplots <- plot_grid(boxplots1, boxplots2, nrow = 1, rel_widths = c(1.5,1))
-plot_grid(boxplots, t456, a, ncol = 1, rel_heights = c(3,1,1))
+boxplots <- plot_grid(boxplots1, boxplots2, nrow = 1, rel_widths = c(1.5,1),
+                      labels = "AUTO", label_size = 8)
+fig4 <- plot_grid(boxplots, design, t456, ncol = 1, rel_heights = c(2.75,1,1),
+                      labels = c(" ", " " , "C"), label_size = 8)
+fig4
 
-#pdf(file="../figures/fig4-1.pdf", width=7.25, height=7.25)
-#plot(fig4)
-#dev.off()
+pdf(file="../figures/fig4-1.pdf", width=7.25, height=7.25)
+plot(fig4)
+dev.off()
 
-```
-
-## supplementary
-
-```{r}
-
-hyptsne <- addgroupings(hyptsne)
-gontsne <- addgroupings(gontsne)
-
-t1 <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) +  labs(subtitle = "Hypothalamus ") +
-  facet_wrap(~sex) + theme(axis.title.x = element_blank())
-t2 <- plottsneelipsev2(hyptsne, hyptsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
-  facet_wrap(~sex) + theme(axis.title = element_blank())
-t3 <- plottsneelipsev2(hyptsne, hyptsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
-  facet_wrap(~sex) + theme(axis.title = element_blank())
-t123 <- plot_grid(t1, t2, t3, nrow = 1, rel_widths = c(1.1,1,1))
-
-t7 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors) +  labs(subtitle = "Gonads ") +
-  facet_wrap(~sex) + theme(strip.text = element_blank())
-t8 <- plottsneelipsev2(gontsne, gontsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
-  facet_wrap(~sex) + theme(axis.title.y = element_blank(), strip.text = element_blank())
-t9 <- plottsneelipsev2(gontsne, gontsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
-  facet_wrap(~sex) + theme(axis.title.y = element_blank(), strip.text = element_blank())
-t789 <- plot_grid(t7, t8, t9, nrow = 1)
 ```
@@
 ```{r write}
-#write.csv(candidateDEGS, "../results/table2.csv", row.names = F)
+write.csv(candidateDEGS, "../results/table2.csv", row.names = F)
 ```

### 2020-05-26 analysis/02_DESeq2.Rmd
more manip comparisons

@@
 
-```{r}
+```{r coldata}
 countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
@@
          treatment <- factor(treatment, levels = alllevels)) %>%
-  mutate(sextissue = as.factor(paste(sex, tissue, sep = "_"))) %>%
-  column_to_rownames(var = "V1")
+  mutate(sextissue = as.factor(paste(sex, tissue, sep = "_"))) 
 colData <- colData %>% drop_na()
 # 987 samples, 8 variables
+
+ncol(countData) == nrow(colData)
+
 ```
@@
 
-```{r treatment, eval = F}
+```{r treatment, eval = T}
 
@@
 
-  control.bldg <- createDEGdftreatment("bldg", "control", i)
-  bldg.lay <- createDEGdftreatment("lay", "bldg", i)
-  lay.inc.d3 <- createDEGdftreatment("inc.d3", "lay",  i) 
-  inc.d3.inc.d9 <- createDEGdftreatment("inc.d9", "inc.d3", i) 
-  inc.d9.inc.d17 <- createDEGdftreatment("inc.d17", "inc.d9", i)
-  inc.d17.hatch <- createDEGdftreatment( "hatch", "inc.d17", i) 
-  hatch.n5 <- createDEGdftreatment( "n5", "hatch",  i) 
-  n5.n9 <- createDEGdftreatment("n9", "n5",  i)
+  inc.d9.early <- createDEGdftreatment("inc.d9", "early",  i) 
+  earyl.hatch <- createDEGdftreatment("early", "hatch",  i)
 
-  control.lay <- createDEGdftreatment("lay", "control", i)
-  control.inc.d3 <- createDEGdftreatment("inc.d3", "control",  i) 
-  control.inc.d9 <- createDEGdftreatment("inc.d9", "control", i) 
-  control.inc.d17 <- createDEGdftreatment("inc.d17", "control", i)
-  control.hatch <- createDEGdftreatment( "hatch", "control", i) 
-  control.n5 <- createDEGdftreatment( "n5", "control",  i) 
-  control.n9 <- createDEGdftreatment("n9", "control",  i)
+  inc.d17.prolong <- createDEGdftreatment("inc.d17", "prolong",  i) 
+  prolong.hatch <- createDEGdftreatment("prolong", "hatch",  i)
 
-  control.extend <- createDEGdftreatment("extend", "control",  i) 
-  control.prolong <- createDEGdftreatment("prolong", "control",  i) 
-  control.early <- createDEGdftreatment("m.inc.d8", "control",  i)
+  hatch.extend <- createDEGdftreatment("hatch", "extend",  i) 
+  extend.n5 <- createDEGdftreatment("extend", "n5",  i)
 
-  control.m.inc.d3 <- createDEGdftreatment("m.inc.d3", "control",  i) 
-  control.m.inc.d9 <- createDEGdftreatment("m.inc.d9", "control",  i) 
-  control.m.inc.d17 <- createDEGdftreatment("m.inc.d17", "control",  i)
-  control.m.n2 <- createDEGdftreatment("m.n2", "control",  i)
-
-  bldg.inc.d3 <- createDEGdftreatment("inc.d3", "bldg",  i) 
-  bldg.inc.d9 <- createDEGdftreatment("inc.d9", "bldg", i) 
-  bldg.inc.d17 <- createDEGdftreatment("inc.d17", "bldg", i)
-  bldg.hatch <- createDEGdftreatment( "hatch", "bldg", i) 
-  bldg.n5 <- createDEGdftreatment( "n5", "bldg",  i) 
-  bldg.n9 <- createDEGdftreatment("n9", "bldg",  i)
-
-  bldg.extend <- createDEGdftreatment("extend", "bldg",  i) 
-  bldg.prolong <- createDEGdftreatment("prolong", "bldg",  i) 
-  bldg.early <- createDEGdftreatment("m.inc.d8", "bldg",  i)
-
-  bldg.m.inc.d3 <- createDEGdftreatment("m.inc.d3", "bldg",  i) 
-  bldg.m.inc.d9 <- createDEGdftreatment("m.inc.d9", "bldg",  i) 
-  bldg.m.inc.d17 <- createDEGdftreatment("m.inc.d17", "bldg",  i)
-  bldg.m.n2 <- createDEGdftreatment("m.n2", "bldg",  i)
-
-  hatch.m.n2 <- createDEGdftreatment("m.n2", "hatch",  i) 
-  hatch.prolong <- createDEGdftreatment("prolong", "hatch",  i) 
-  hatch.m.inc.d8 <- createDEGdftreatment("prolong", "hatch",  i)
-
-  inc.d17.m.n2 <- createDEGdftreatment("m.n2", "inc.d17",  i) 
-  inc.d17.prolong <- createDEGdftreatment("prolong", "inc.d17",  i)
-
-  prolong.extend <- createDEGdftreatment("extend", "prolong",  i) 
-  early.extend <- createDEGdftreatment("extend", "m.inc.d8",  i) 
-  early.prolong <- createDEGdftreatment("prolong", "m.inc.d8",  i)
-
-  m.inc.d9.m.inc.d3 <- createDEGdftreatment("m.inc.d9", "m.inc.d3",  i) 
-  m.inc.d17.m.inc.d9 <- createDEGdftreatment("m.inc.d17", "m.inc.d9",  i) 
-  m.inc.d17.m.inc.d3 <- createDEGdftreatment("m.inc.d17", "m.inc.d3",  i) 
-  m.n2.m.inc.d17 <- createDEGdftreatment("m.n2", "m.inc.d17",  i) 
-  m.n2.m.inc.d9 <- createDEGdftreatment("m.n2", "m.inc.d9",  i)
-  m.n2.m.inc.d3 <- createDEGdftreatment("m.n2", "m.inc.d3",  i)
-
-  m.inc.d3.inc.d3 <- createDEGdftreatment("m.inc.d3", "inc.d3",  i) 
-  m.inc.d9.inc.d9 <- createDEGdftreatment("m.inc.d9", "inc.d9",  i) 
-  m.inc.d17.inc.d17 <- createDEGdftreatment("m.inc.d17", "inc.d17",  i) 
-  m.n2.hatch <- createDEGdftreatment("m.n2", "hatch",  i)
+  ## previously run samples in comparisons.R
 

### 2020-05-26 R/comparisons.R

@@ -1,0 +1,60 @@
+control.bldg <- createDEGdftreatment("bldg", "control", i)
+bldg.lay <- createDEGdftreatment("lay", "bldg", i)
+lay.inc.d3 <- createDEGdftreatment("inc.d3", "lay",  i) 
+inc.d3.inc.d9 <- createDEGdftreatment("inc.d9", "inc.d3", i) 
+inc.d9.inc.d17 <- createDEGdftreatment("inc.d17", "inc.d9", i)
+inc.d17.hatch <- createDEGdftreatment( "hatch", "inc.d17", i) 
+hatch.n5 <- createDEGdftreatment( "n5", "hatch",  i) 
+n5.n9 <- createDEGdftreatment("n9", "n5",  i) 
+
+control.lay <- createDEGdftreatment("lay", "control", i)
+control.inc.d3 <- createDEGdftreatment("inc.d3", "control",  i) 
+control.inc.d9 <- createDEGdftreatment("inc.d9", "control", i) 
+control.inc.d17 <- createDEGdftreatment("inc.d17", "control", i)
+control.hatch <- createDEGdftreatment( "hatch", "control", i) 
+control.n5 <- createDEGdftreatment( "n5", "control",  i) 
+control.n9 <- createDEGdftreatment("n9", "control",  i) 
+
+control.extend <- createDEGdftreatment("extend", "control",  i) 
+control.prolong <- createDEGdftreatment("prolong", "control",  i) 
+
+control.m.inc.d3 <- createDEGdftreatment("m.inc.d3", "control",  i) 
+control.m.inc.d9 <- createDEGdftreatment("m.inc.d9", "control",  i) 
+control.m.inc.d17 <- createDEGdftreatment("m.inc.d17", "control",  i)
+control.m.n2 <- createDEGdftreatment("m.n2", "control",  i) 
+
+bldg.inc.d3 <- createDEGdftreatment("inc.d3", "bldg",  i) 
+bldg.inc.d9 <- createDEGdftreatment("inc.d9", "bldg", i) 
+bldg.inc.d17 <- createDEGdftreatment("inc.d17", "bldg", i)
+bldg.hatch <- createDEGdftreatment( "hatch", "bldg", i) 
[…]

### 2020-05-26 R/functions.R

@@ -23,6 +23,8 @@ subsetcolData2 <- function(colData, eachgroup){
 }
 
 
+
+
 subsetcolData3 <- function(colData, eachgroup){
   
   colData <- colData %>%

### 2020-05-27 analysis/figure-4.Rmd
now with fig 4 supplement

@@
   select(gene, inc.d3_m.inc.d3, inc.d9_m.inc.d9, inc.d17_m.inc.d17, hatch_m.n2,
-          inc.d17_prolong, hatch_prolong)
-candidateDEGS
-
+         inc.d17_prolong, hatch_prolong,extend_hatch, n5_extend, 
+         early_inc.d9 ,hatch_early)
+head(candidateDEGS[1:5])
 ```
@@
 
+## suppl fig 1
+
+```{r supplfig-4}
+
+comparisonlevelsmanip <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2",
+         "early_inc.d9" ,"hatch_early", "inc.d17_prolong", "hatch_prolong","extend_hatch", "n5_extend")
+
+# for fig 1
+DEGmanip <- allDEG %>% 
+  filter(comparison %in% comparisonlevelsmanip) %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelsmanip))
+
+DEGmanip %>% 
+  group_by(comparison, sex, tissue) %>%
+  summarize(totalDEGs = n())  %>%
+  arrange(desc(totalDEGs))
+
+s1a <- makebargraphsuppl(DEGmanip, "hypothalamus","DEGs w/ + LFC", 0, 3200) + 
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank()) + 
+  labs(subtitle = "hypothalamus") + geom_vline(xintercept = 4.5, linetype = "dashed", color = "grey")
+
+s1b <- makebargraphsuppl(DEGmanip, "pituitary","DEGs w/ + LFC", 0, 3200) + 
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), strip.text = element_blank()) + 
+  labs(subtitle = "pituitary")  + geom_vline(xintercept = 4.5, linetype = "dashed", color = "grey")
+
+s1c <- makebargraphsuppl(DEGmanip, "gonad","DEGs w/ + LFC", 0, 3200) + 
+  theme(strip.text = element_blank()) + 
+  labs(subtitle = "gonad", x = "Comparison of parental stage and manipulation groups") + geom_vline(xintercept = 4.5, linetype = "dashed", color = "grey") + 
+  scale_x_discrete(breaks= comparisonlevelsmanip,
+                      labels= comparisonlevelsmanip,
+                   drop=FALSE)
+
+supplfig4 <- plot_grid(s1a, s1b, s1c, ncol = 1, rel_heights = c(1,1,1.5))
+supplfig4
+
+```
+
 ## write files
@@
 ```{r write}
+
+pdf(file="../figures/supplfig-4-1.pdf", width=5, height=5)
+plot(supplfig4)
+dev.off()
+
 write.csv(candidateDEGS, "../results/table2.csv", row.names = F)

### 2020-05-27 analysis/figure-1.Rmd
better fig 1
better fig 1 supplements

@@
   drop_na()
-
-suppltable1 <- DEGchar %>%
-  arrange( tissue, sex, direction, gene)
-head(suppltable1)
-write_csv(suppltable1, "../results/suppltable1.csv")
-
-suppltable2 <- rbind(DEGcontrol, DEGbldg)  %>%
-  arrange( tissue, sex, direction, gene)
-head(suppltable2)
-write_csv(suppltable2, "../results/suppltable2.csv")
 
@@
 
-pdf(file="../figures/supplfig-1-1.pdf", width=5, height=5)
-plot(supplfig1)
-dev.off()
+# for volcano plots
+hyphatch <- DEGchar %>% filter(tissue == "hypothalamus", comparison == "hatch_n5") 
+pitinc <- DEGchar %>% filter(tissue == "pituitary", comparison == "inc.d9_inc.d17") 
+goninc <- DEGchar %>% filter(tissue == "gonad", comparison == "lay_inc.d3")
+
+s2a <- plot.volcano("hypothalamus", sexlevels,  "hatch_n5") + facet_wrap(~sex) + 
+  labs(subtitle = "hypothalamus", title = "Volcano plots") 
+s2c <- plot.volcano("pituitary", sexlevels,  "inc.d9_inc.d17") + 
+  facet_wrap(~sex) + 
+  labs(subtitle = "pitutiary") + 
+  theme(legend.title = element_blank(), strip.text = element_blank())
+s2e <- plot.volcano("gonad", sexlevels,  "lay_inc.d3") + facet_wrap(~sex) + 
+  labs( x = "Log-fold change (LFC)", subtitle = "gonads") + 
+  theme(legend.title = element_blank(), strip.text = element_blank())
+
+s2b <- makebargraph(hyphatch, "hypothalamus","Total DEGs with +LFC", 0, 1800)  + 
+  labs(subtitle = " ", title = "Bar graphs") + scale_x_discrete(labels = "hatch\nvs n5")
+s2d <- makebargraph(pitinc, "pituitary","Total DEGs with +LFC", 0, 1800) + 
+  labs(subtitle = " ") + scale_x_discrete(labels = "inc.d9 vs\ninc.d17") + 
+  theme(strip.text = element_blank())
+s2f <- makebargraph(goninc, "gonad","Total DEGs with +LFC", 0, 1800) + 
+  labs(subtitle = " ", x = "stages compared") + scale_x_discrete(labels = "lay vs\ninc.d3") + 
+  theme(strip.text = element_blank())
+supplfig2 <- plot_grid(s2a, s2b, s2c, s2d, s2e, s2f,
+          rel_widths = c(2,1), rel_heights = c(1.1,1,1.1), ncol = 2,
+          labels = c("A", "B"), label_size = 8)
+supplfig2
 ```
@@
 
+suppltable1 <- DEGchar %>%
+  arrange( tissue, sex, direction, gene)
+head(suppltable1)
+write_csv(suppltable1, "../results/suppltable1.csv")
+
+suppltable2 <- rbind(DEGcontrol, DEGbldg)  %>%
+  arrange( tissue, sex, direction, gene)
+head(suppltable2)
+write_csv(suppltable2, "../results/suppltable2.csv")
+
+pdf(file="../figures/supplfig-1-1.pdf", width=5, height=5)
+plot(supplfig1)
+dev.off()
+
+pdf(file="../figures/supplfig-1-2.pdf", height=5, width=5)
+plot(supplfig2)
+dev.off()
 ```

### 2020-05-29 analysis/figure-4.Rmd
better suppl files

@@
                 map_signif_level=TRUE,
-                textsize = 3, family = 'Helvetica',
+                textsize = 1.5, family = 'Helvetica',
                 vjust = 1.5, size = 0.5) 

### 2020-05-29 analysis/figure-4.Rmd
better exp desig fig 4

--- 2020-05-27 analysis/figure-4.Rmd
@@
     ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
-    geom_boxplot(outlier.shape = NA) +
+    geom_boxplot(outlier.shape = NA, lwd=0.5) +
     geom_jitter(size = 0.25, aes(color = sex)) +
@@
 
-a <- plotcandidateremoval(hypvsd, "HTR2C", "hypothalamus") + theme(axis.title.x = element_blank(), axis.text.x = element_blank())
-b <- plotcandidateremoval(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank(), strip.text = element_blank(), axis.text.x = element_blank())
+a <- plotcandidateremoval(hypvsd, "HTR2C", "hypothalamus") + theme(axis.title.x = element_blank())
+b <- plotcandidateremoval(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank(), strip.text = element_blank())
 c <- plotcandidateremoval(gonvsd, "ESR2", "gonads") + theme(strip.text = element_blank())
@@
 
-e <- plotcandidatetiming(hypvsd, "HTR2C", "hypothalamus") + theme(axis.title.x = element_blank(), axis.text.x = element_blank()) 
-f <- plotcandidatetiming(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank(), strip.text = element_blank(), axis.text.x = element_blank())
-g <- plotcandidatetiming(gonvsd, "ESR2", "gonads") + theme(strip.text = element_blank())
+e <- plotcandidatetiming(hypvsd, "HTR2C", " ") + theme(axis.title.x = element_blank()) + labs(y = NULL, subtitle = " ", title = " ") 
+f <- plotcandidatetiming(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank(), strip.text = element_blank()) + labs(y = NULL, subtitle = " ", title = " ") 
+g <- plotcandidatetiming(gonvsd, "ESR2", "gonads") + theme(strip.text = element_blank()) + labs(y = NULL, subtitle = " ", title = " ") 
 boxplots2 <- plot_grid(e,f,g, nrow = 3)
+boxplots2
 ```
@@
                       labels = "AUTO", label_size = 8)
-fig4 <- plot_grid(boxplots, design, t456, ncol = 1, rel_heights = c(2.75,1,1),
+fig4 <- plot_grid(boxplots, design, t456, ncol = 1, rel_heights = c(2.75,1.3,1),
                       labels = c(" ", " " , "C"), label_size = 8)

### 2020-05-29 analysis/figure-4.Rmd
better suppl file

@@
 
-pdf(file="../figures/fig4-1.pdf", width=7.25, height=7.25)
-plot(fig4)
-dev.off()
-
 ```
@@
 
-supplfig4 <- plot_grid(s1a, s1b, s1c, ncol = 1, rel_heights = c(1,1,1.5))
+contrasts <- png::readPNG("../figures/images/fig_fig4sup1.png")
+contrasts <- ggdraw() +  draw_image(contrasts, scale = 1)
+
+supplfig4 <- plot_grid(contrasts, s1a, s1b, s1c, ncol = 1, rel_heights = c(0.5, 1,1,1.5))
 supplfig4
@@
 
+pdf(file="../figures/fig4-1.pdf", width=7.25, height=7.25)
+plot(fig4)
+dev.off()
+
 pdf(file="../figures/supplfig-4-1.pdf", width=5, height=5)

### 2020-05-29 analysis/figure-4.Rmd
fig 3 with new infographic

@@
 boxplots2 <- plot_grid(e,f,g, nrow = 3)
-boxplots2
 ```

### 2020-05-31 analysis/figure-2.Rmd
now colors, include control for fig 2

@@
 title: "Fig2: Candidate genes"
-author: "Rayna M Harris"
-date: "3/26/2020"
 output: md_document
@@
   select(-X1) %>%
-  filter(treatment %in% charlevelsnocontrol) %>%
+  filter(treatment %in% charlevels) %>%
   drop_na()
-tail(candidatevsd)
+head(candidatevsd)
 
@@
 
-pdf(file="../figures/supplfig-2.pdf", width=7.25, height=7.25)
-plot(supplfig2)
-dev.off()
 ```
@@
          x = NULL) +
-    geom_signif(comparisons = list( c( "bldg", "lay"),
+    geom_signif(comparisons = list( c( "control", "bldg"),
+                                    c( "bldg", "lay"),
                                    c( "lay", "inc.d3"),
@@
 
-c1 <- scattercorrelations(FH, FH$CRHR2, "CRHR2", FH$HTR2C, "HTR2C", "#969696" ) +  labs(title = " ", subtitle = " " )  
+c1 <- scattercorrelations(FH, FH$DRD1, "DRD1", FH$HTR2C, "HTR2C", "#969696" ) +  labs(title = " ", subtitle = " " )  
 c3 <- scattercorrelations(FP, FP$AVPR1A, "AVPR1A", FP$CRHR1,  "CRHR1", "#969696")  + labs(title = " ", subtitle = " ") 
-c5 <- scattercorrelations(FG, FG$CRHBP, "CRHBP",  FG$ESR2,  "ESR2", "#969696")   + labs(title = " ", subtitle = " ")
+c5 <- scattercorrelations(FG, FG$PGR, "PGR",  FG$ESR1,  "ESR1", "#969696")   + labs(title = " ", subtitle = " ")
 
@@
 c4 <- scattercorrelations(MP, MP$AVPR1A,  "AVPR1A",  MP$CRHR1, "CRHR1", "#525252")  + labs(title = " ", subtitle = " " ) 
-c6 <- scattercorrelations(MG, MG$PGR, "PGR", MG$PRLR, "PRLR", "#525252")   + labs(title = " ", subtitle = " " )
+c6 <- scattercorrelations(MG, MG$PGR, "PGR", MG$ESR1, "ESR1", "#525252")   + labs(title = " ", subtitle = " " )
 
-d1 <- candidateboxplot("hypothalamus", c("CRHR2"), "female") + labs(x = NULL, title =  "Hypothlamic expression", subtitle = "females" )  
+d1 <- candidateboxplot("hypothalamus", c("DRD1"), "female") + labs(x = NULL, title =  "Hypothlamic expression", subtitle = "females" )  
 d2 <- candidateboxplot("hypothalamus", c("HTR2C"), "female") + labs(x = NULL )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
@@
 d4 <- candidateboxplot("pituitary", c("CRHR1"), "female") + labs(x = NULL  )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
-d5 <- candidateboxplot("gonad", c("CRHBP"), "female") + labs(x = NULL, title = "Gonadal expression", subtitle = "females")
-d6 <- candidateboxplot("gonad", c("ESR2"), "female") + labs(x = NULL ) + theme(axis.text.x = element_text(angle = 45, vjust = 1))
+d5 <- candidateboxplot("gonad", c("PGR"), "female") + labs(x = NULL, title = "Gonadal expression", subtitle = "females")
+d6 <- candidateboxplot("gonad", c("ESR1"), "female") + labs(x = NULL ) + theme(axis.text.x = element_text(angle = 45, vjust = 1))
 
 e1 <- candidateboxplot("hypothalamus", c("DRD1"), "male") + labs(x = NULL, title = " " , subtitle = "males" )  
-e2 <- candidateboxplot("hypothalamus", c("HTR2C"), "male") + labs(x = NULL )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
-e3 <- candidateboxplot("pituitary", c("AVPR1A"), "male") + labs(x = NULL , subtitle = "males" ) 
-e4 <- candidateboxplot("pituitary", c("CRHR1"), "male") + labs(x = NULL  )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
-e5 <- candidateboxplot("gonad", c("PGR"), "male") + labs(x = NULL, subtitle = "males" ) 
-e6 <- candidateboxplot("gonad", c("PRLR"), "male") + labs(x = NULL )   + theme(axis.text.x = element_text(angle = 45, vjust = 1))
+e2 <- candidateboxplot("hypothalamus", c("HTR2C"), "male") + labs(x = NULL, title = " " )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
+e3 <- candidateboxplot("pituitary", c("AVPR1A"), "male") + labs(x = NULL , subtitle = "males", title = " " ) 
+e4 <- candidateboxplot("pituitary", c("CRHR1"), "male") + labs(x = NULL, title = " "  )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
+e5 <- candidateboxplot("gonad", c("PGR"), "male") + labs(x = NULL, subtitle = "males", title = " " ) 
+e6 <- candidateboxplot("gonad", c("ESR1"), "male") + labs(x = NULL , title = " ")   + theme(axis.text.x = element_text(angle = 45, vjust = 1))
 
@@
 
-pdf(file="../figures/fig2-1.pdf", width=7.25, height=7.25)
-plot(fig2)
-dev.off()
 ```
@@
 write.csv(table1, "../results/table1.csv")
+
+pdf(file="../figures/fig2-1.pdf", width=7.25, height=7.25)
+plot(fig2)
+dev.off()
+
+pdf(file="../figures/supplfig-2.pdf", width=7.25, height=7.25)
+plot(supplfig2)
+dev.off()
 ```

### 2020-06-02 analysis/figure-3.Rmd
huh

@@
 c2 <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes in WGCNA module with PRL") 
-c3 <- plotcorrplot(cancergenes, NULL) + theme(legend.position = "right") + labs(subtitle = "Breast, ovary, & piuitary cancer genes") 
+c3 <- plotcorrplot(cancergenes, NULL) + theme(legend.position = "right") + labs(subtitle = "Breast, ovary, & piuitary cancer genes", color = "Correlation")
 c <- plot_grid(c1,c2,c3, nrow = 1, labels = c("D"), label_size = 8, rel_widths = c(1,1,1.3))

### 2020-07-31 analysis/03_datawrangling.Rmd
new candidate gene table

@@
                           "OPRM1", "OXT" , "PGR", "PRL", "PRLR",  "SLC6A4") #5HTT
-curleychampagnegenes
-curleychampagnegenes <- as.data.frame(curleychampagnegenes) %>%
-  mutate(GO = "literature", gene = curleychampagnegenes)  %>%
+
+literaturegenes <- c(curleychampagnegenes, 
+                     "VIP", "GNIH", "GAL", "TH", "THRB",
+                     "GNRHR", "GNRH1", "CGNRH-R", "FSHB", "FSHR")
+literaturegenes
+
+literaturegenes <- as.data.frame(literaturegenes) %>%
+  mutate(GO = "literature", gene = literaturegenes)  %>%
   select(GO, gene)
@@
   mutate(gene = toupper(gene)) %>%
-  rbind(., curleychampagnegenes) %>%
+  rbind(., literaturegenes) %>%
   arrange(gene) %>%
@@
 
-# note: I can't find these genes in dataset: NOS1 or OXTR or FOX1B or HTR5A
-
-GOgeneonly <- parentalcaregenes %>%
-  filter(GO != "") %>%
-  pull(GO)
-GOgeneonly
 ```
@@
 unique(allDEG$comparison)
+
+tempDEGs <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2))
+head(tempDEGs)
+
+wideDEGs <- tempDEGs %>%
+  mutate(lfc = round(lfc, 2),
+         padj =  round(padj,2)) %>%
+  mutate(sextissue = paste(sex, tissue, sep = "_"),
+         lfcpadj = paste(lfc, padj, sep = ", "))  %>%
+  dplyr::select(sextissue,comparison, gene, lfcpadj) %>%
+  pivot_wider(names_from = comparison, values_from = lfcpadj)
+
+wideDEGscandidates <- wideDEGs %>%
+  filter(gene %in% GOgenesLong$gene)
+wideDEGscandidates
+
+wideDEGsBldg <- wideDEGscandidates %>%
+  select(sextissue,gene, contains("bldg"))
+wideDEGsBldg
+
+wideDEGsControl <- wideDEGscandidates %>%
+  select(sextissue,gene, contains("control"))
+wideDEGsControl
+
+# sequential in 2 or more comparisons
+wideDEGsSequential <- wideDEGscandidates %>%
+  select(sextissue,gene, "control_bldg", "bldg_lay", "lay_inc.d3", "inc.d3_inc.d9",
+         "inc.d9_inc.d17", "inc.d17_hatch", "hatch_n5", "n5_n9") %>%
+  mutate(totalNA = rowSums(is.na(.))) %>%
+  filter(totalNA < 8) %>%
+  arrange(gene, sextissue) %>%
+  select(-totalNA) %>% 
+  replace(., is.na(.), "")
+wideDEGsSequential
+
 ```
@@
 write.csv(parentalcaregenes, "../metadata/03_parentalcaregenes.csv")
+write.csv(wideDEGsSequential, "../results/03_wideDEGsSequential.csv")
 ```

### 2020-07-31 analysis/03_datawrangling.Rmd
better file names

@@
 write.csv(candidatevsd, "../results/03_candidatevsd.csv")
-write.csv(parentalcaregenes, "../metadata/03_parentalcaregenes.csv")
-write.csv(wideDEGsSequential, "../results/03_wideDEGsSequential.csv")
+write.csv(wideDEGsSequential, "../results/table1.csv", row.names = F)
 ```

### 2020-07-31 analysis/03_datawrangling.Rmd
new candidate genes, show external

@@
   filter(totalNA < 8) %>%
-  arrange(gene, sextissue) %>%
+  arrange(totalNA, gene, sextissue) %>%
+  #arrange( gene, sextissue) %>%
   select(-totalNA) %>% 

### 2020-07-31 analysis/figure-2.Rmd
new candidate genes, show external

@@
 parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% select(-X1)
-head(parentalcaregenes)
-
-curleychampagnegenes <- parentalcaregenes %>% distinct(literature) %>% drop_na() %>% pull(literature)
-GOgenes <- parentalcaregenes %>% distinct(GO)  %>% drop_na() %>% pull(GO)
 candidategenes <- parentalcaregenes %>% pull(gene) 
 candidategenes
+
+GOgenes <- parentalcaregenes %>% distinct(GO)  %>% drop_na() %>% pull(GO)
+GOgenes
+
+literaturegenes <- parentalcaregenes %>% 
+  distinct(literature) %>% 
+  drop_na() %>% 
+  pull(literature)
+literaturegenes
+
+litNotGO <-  literaturegenes[!literaturegenes %in% GOgenes]
+litNotGOknown <- litNotGO[!litNotGO %in% c("ADRA2A", "MEST", "SLC6A4")]
+litNotGOknown
 ```
@@
   pivot_wider(names_from = comparison, values_from = res) %>%
-  select(gene, bldg_lay, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
+  select(gene, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
 candidateDEGS
@@
 table1 <- left_join(candidateDEGS, parentalcaregenes) %>%
-  select(gene, bldg_lay:n5_n9, literature, GO, NCBI) %>%
+  select(gene, lay_inc.d3:n5_n9, literature, GO, NCBI) %>%
   mutate(literature = if_else(is.na(literature), " ", "X"),
          GO = if_else(is.na(GO), " ", "X"))
-head(table1)
+(table1)
 ```
@@
   filter(treatment %in% charlevels) %>%
+  mutate(treatment = factor(treatment, levels = charlevels)) %>%
   drop_na()
@@
 ```{r correlations}
+hyp1 <- makecorrdf("female", "hypothalamus", litNotGOknown)  
+pit1 <- makecorrdf("female", "pituitary", litNotGOknown)  
+gon1 <- makecorrdf("female", "gonad", litNotGOknown)
 
-hyp1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
-pit1 <- makecorrdf("female", "pituitary", curleychampagnegenes)  
-gon1 <- makecorrdf("female", "gonad", curleychampagnegenes)
-
-hyp2 <- makecorrdf("male", "hypothalamus", curleychampagnegenes)  
-pit2 <- makecorrdf("male", "pituitary", curleychampagnegenes)  
-gon2 <- makecorrdf("male", "gonad", curleychampagnegenes)
+hyp2 <- makecorrdf("male", "hypothalamus", litNotGOknown)  
+pit2 <- makecorrdf("male", "pituitary", litNotGOknown)  
+gon2 <- makecorrdf("male", "gonad", litNotGOknown)
 
@@
 
-candidateboxplot <- function(whichtissue, whichgenes, whichsex){
-
-  p <- candidatevsd %>%
-    filter(tissue %in% whichtissue,
-           gene %in% whichgenes,
-           sex %in% whichsex) %>%
-    mutate(treatment = factor(treatment, levels = charlevels)) %>%
-    ggplot(aes(x = treatment, y = counts)) +
-    geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
-    geom_jitter(size = 0.25, aes(color = sex)) +
-    #facet_wrap(~sex,  nrow = 1) +
-    scale_fill_manual(values = allcolors) +
-    scale_color_manual(values = allcolors) +
-    theme_B3() + 
-    theme(legend.position = "none",
-          axis.title.y = element_text(face = "italic"),
-          axis.text.x = element_blank(),
-          axis.ticks = element_blank(),
-          axis.line.x = element_blank()) +
-    labs(y = whichgenes,
-         x = NULL) +
-    geom_signif(comparisons = list( c( "control", "bldg"),
-                                    c( "bldg", "lay"),
-                                   c( "lay", "inc.d3"),
-                                   c("inc.d3", "inc.d9"),
-                                   c( "inc.d9", "inc.d17"),
-                                   c( "inc.d17", "hatch"),
-                                   c("hatch", "n5"),
-                                   c( "n5", "n9")),  
-                map_signif_level=TRUE,
-                textsize = 1.5, family = 'Helvetica',
-                vjust = 1.5, size = 0.2)
-
-  return(p)
-}
-
 c1 <- scattercorrelations(FH, FH$DRD1, "DRD1", FH$HTR2C, "HTR2C", "#969696" ) +  labs(title = " ", subtitle = " " )  
@@
 
+```{r newHPGfig}
+
+a <- candidateboxplot("gonad", c("COMT"), "female") + 
+  labs(title =  "Females", 
+       subtitle = "Hypothlamic expression" )
+
+b <- externalboxplots("gonad", c("COMT"), "female") + 
+  labs(title =  "Females", 
+       subtitle = "Hypothlamic expression" )
+
+c <- candidateboxplot("pituitary", c("PRL"), "female") + 
+  labs(subtitle = "Pituitary expression" )
+
+d <- externalboxplots("pituitary", c("PRL"), "female") + 
+   labs(subtitle = "Pituitary expression" )
+
+e <- candidateboxplot("gonad", c("AVPR1A"), "female") + 
+  labs( subtitle = "Gonad expression" ) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+f <- externalboxplots("pituitary", c("AVPR1A"), "female") + 
+   labs( subtitle = "Gonad expression" ) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+females <- plot_grid(a,b,c,d,e,f, ncol = 2, rel_widths = c(2,1))
+
+a2 <- candidateboxplot("gonad", c("COMT"), "male") + 
+  labs(title =  "Males", 
+       subtitle = "Hypothlamic expression" )
+
+b2 <- externalboxplots("gonad", c("COMT"), "male") + 
+  labs(title =  "Males", 
+       subtitle = "Hypothlamic expression" )
+
+c2 <- candidateboxplot("pituitary", c("PRL"), "male") + 
+  labs(subtitle = "Pituitary expression" )
+
+d2 <- externalboxplots("pituitary", c("PRL"), "male") + 
+   labs(subtitle = "Pituitary expression" )
+
+e2 <- candidateboxplot("gonad", c("AVPR1A"), "male") + 
+  labs( subtitle = "Gonad expression" ) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+f2 <- externalboxplots("pituitary", c("AVPR1A"), "male") + 
+   labs( subtitle = "Gonad expression" )  + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+males <- plot_grid(a2,b2,c2,d2,e2,f2, ncol = 2, rel_widths = c(2,1))
+
+newfig2 <- plot_grid(females, males)
+newfig2
+```
+
 ```{r writefiles}
@@
 #write.csv(candidatevsd, "../results/candidatecounts.csv")
-write.csv(table1, "../results/table1.csv")
+write.csv(table1, "../results/table1-v2.csv")
 
@@
 dev.off()
+
+pdf(file="../figures/fig2-2.pdf", width=7.25, height=7.25)
+plot(newfig2)
+dev.off()
 ```

### 2020-07-31 R/functions.R

@@ -1224,3 +1224,78 @@ plotcandidatemanipquad <- function(df, whichtissue, whichgenes){
 }
 
 
+## candidate gene box plot 
+
+candidateboxplot <- function(whichtissue, whichgenes, whichsex){
+  
+  p <- candidatevsd %>%
+    filter(tissue %in% whichtissue,
+           gene %in% whichgenes,
+           sex %in% whichsex) %>%
+    mutate(treatment = factor(treatment, levels = charlevels)) %>%
+    ggplot(aes(x = treatment, y = counts)) +
+    geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
+    geom_jitter(size = 0.25, aes(color = sex)) +
+    #facet_wrap(~sex,  nrow = 1) +
+    scale_fill_manual(values = allcolors) +
+    scale_color_manual(values = allcolors) +
+    theme_B3() + 
+    theme(legend.position = "none",
+          axis.title.y = element_text(face = "italic"),
+          axis.text.x = element_blank(),
+          axis.ticks = element_blank(),
+          axis.line.x = element_blank()) +
+    labs(y = whichgenes,
+         x = NULL) +
+    geom_signif(comparisons = list( c( "control", "bldg"),
+                                    c( "bldg", "lay"),
+                                    c( "lay", "inc.d3"),
+                                    c("inc.d3", "inc.d9"),
+                                    c( "inc.d9", "inc.d17"),
+                                    c( "inc.d17", "hatch"),
+                                    c("hatch", "n5"),
+                                    c( "n5", "n9")),  
+                map_signif_level=TRUE,
+                textsize = 1.5, family = 'Helvetica',
+                vjust = 1.5, size = 0.2) 
+  
+  return(p)
+}
+
+
+## candidate gene box plot 
+
+externalboxplots <- function(whichtissue, whichgenes, whichsex){
+  
+  p <- candidatevsd %>%
+    filter(tissue %in% whichtissue,
+           gene %in% whichgenes,
+           sex %in% whichsex) %>%
+    mutate(treatment = factor(treatment, levels = charlevels)) %>%
+    mutate(external = fct_collapse(treatment, 
+                                   "none" = c("control", "bldg"),
+                                   "eggs" = c("lay" , "inc.d3", 
+                                              "inc.d9", "inc.d17"),
+                                   "chicks" = c("hatch", "n5", "n9"))) %>%
[…]

### 2020-07-31 R/themes.R

@@ -113,7 +113,8 @@ colorsnewgrouping <-  c("control" = "#F8766D",
                  "chickcare"= "#FF61C3",
                  "manip" = "#959595")
 
-colorhypothesis <- c("nest" = "#8B4513",
+colorhypothesis <- c("none" = "#cc5500",
+                     "nest" = "#8B4513",
                      "eggs" = "#7FA163",
                      "chicks" = "#3A80B9",
                      "lo" = "#B5C4B2",
@@ -123,7 +124,8 @@ colorhypothesis <- c("nest" = "#8B4513",
 
 colorstissue <- c("hypothalamus" = "#d95f02",
                   "pituitary" = "#1b9e77",
-                  "gonads" =  "#7570b3")
+                  "gonads" =  "#7570b3",
+                  "gonad" =  "#7570b3")
 
 sexcolors <- c("female" = "#969696", "male" = "#525252")
 
@@ -139,7 +141,8 @@ colorsvolcano <-  c(colorschar,
                     "NS" = "#bdbdbd")
 
  
-allcolors <- c(colorscharmaip, sexcolors, colorstissue, colorhypothesis,  "gonad" =  "#7570b3")
+allcolors <- c(colorscharmaip, sexcolors, 
+               colorstissue, colorhypothesis)
  
 colorsvolcanochar <-  c(colorschar,
                     "NS" = "#bdbdbd")

### 2020-07-31 analysis/figure-2.Rmd
better new fig 2

@@
 
-```
+a <- candidateboxplot("hypothalamus", c("COMT"), "female") + 
+  labs(title =  "Sequential", 
+       subtitle = "Female hypothalamus" )
 
-```{r newHPGfig}
-
-a <- candidateboxplot("gonad", c("COMT"), "female") + 
-  labs(title =  "Females", 
-       subtitle = "Hypothlamic expression" )
-
-b <- externalboxplots("gonad", c("COMT"), "female") + 
-  labs(title =  "Females", 
-       subtitle = "Hypothlamic expression" )
+b <- externalboxplots("hypothalamus", c("COMT"), "female") + 
+  labs(title =  "External", 
+       subtitle = " " )
 
 c <- candidateboxplot("pituitary", c("PRL"), "female") + 
-  labs(subtitle = "Pituitary expression" )
+  labs(subtitle = "Female pituitary" )
 
 d <- externalboxplots("pituitary", c("PRL"), "female") + 
-   labs(subtitle = "Pituitary expression" )
+   labs(subtitle = " " )
 
 e <- candidateboxplot("gonad", c("AVPR1A"), "female") + 
-  labs( subtitle = "Gonad expression" ) + 
+  labs( subtitle = "Female gonads" ) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
@@
 f <- externalboxplots("pituitary", c("AVPR1A"), "female") + 
-   labs( subtitle = "Gonad expression" ) + 
+   labs( subtitle = " " ) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
@@
 
-a2 <- candidateboxplot("gonad", c("COMT"), "male") + 
-  labs(title =  "Males", 
-       subtitle = "Hypothlamic expression" )
+a2 <- candidateboxplot("hypothalamus", c("COMT"), "male") + 
+  labs(title =  "Sequential", 
+       subtitle = "Male hypothalamus" )
 
-b2 <- externalboxplots("gonad", c("COMT"), "male") + 
-  labs(title =  "Males", 
-       subtitle = "Hypothlamic expression" )
+b2 <- externalboxplots("hypothalamus", c("COMT"), "male") + 
+  labs(title =  "External", 
+       subtitle = "  " )
 
 c2 <- candidateboxplot("pituitary", c("PRL"), "male") + 
-  labs(subtitle = "Pituitary expression" )
+  labs(subtitle = "Male pituitary" )
 
 d2 <- externalboxplots("pituitary", c("PRL"), "male") + 
-   labs(subtitle = "Pituitary expression" )
+   labs(subtitle = " " )
 
 e2 <- candidateboxplot("gonad", c("AVPR1A"), "male") + 
-  labs( subtitle = "Gonad expression" ) + 
+  labs( subtitle = "Male gonads" ) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
@@
 f2 <- externalboxplots("pituitary", c("AVPR1A"), "male") + 
-   labs( subtitle = "Gonad expression" )  + 
+   labs( subtitle = " " )  + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
@@
 
-newfig2 <- plot_grid(females, males)
+newfig2 <- plot_grid(females, males, labels = c("A", "B"),
+                     label_size  = 8)
 newfig2

### 2020-08-01 analysis/figure-2.Rmd
broken candidate gene plot

@@
 
-d1 <- candidateboxplot("hypothalamus", c("DRD1"), "female") + labs(x = NULL, title =  "Hypothlamic expression", subtitle = "females" )  
-d2 <- candidateboxplot("hypothalamus", c("HTR2C"), "female") + labs(x = NULL )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
-d3 <- candidateboxplot("pituitary", c("AVPR1A"), "female") + labs(x = NULL , title = "Pituitary expression", subtitle = "females") 
-d4 <- candidateboxplot("pituitary", c("CRHR1"), "female") + labs(x = NULL  )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
-d5 <- candidateboxplot("gonad", c("PGR"), "female") + labs(x = NULL, title = "Gonadal expression", subtitle = "females")
-d6 <- candidateboxplot("gonad", c("ESR1"), "female") + labs(x = NULL ) + theme(axis.text.x = element_text(angle = 45, vjust = 1))
+plot3candidates <- function(hypgene, pitgene, gongene){
 
-e1 <- candidateboxplot("hypothalamus", c("DRD1"), "male") + labs(x = NULL, title = " " , subtitle = "males" )  
-e2 <- candidateboxplot("hypothalamus", c("HTR2C"), "male") + labs(x = NULL, title = " " )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
-e3 <- candidateboxplot("pituitary", c("AVPR1A"), "male") + labs(x = NULL , subtitle = "males", title = " " ) 
-e4 <- candidateboxplot("pituitary", c("CRHR1"), "male") + labs(x = NULL, title = " "  )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
-e5 <- candidateboxplot("gonad", c("PGR"), "male") + labs(x = NULL, subtitle = "males", title = " " ) 
-e6 <- candidateboxplot("gonad", c("ESR1"), "male") + labs(x = NULL , title = " ")   + theme(axis.text.x = element_text(angle = 45, vjust = 1))
+  df <- candidatevsd %>%
+    filter(gene %in% c(hypgene, pitgene, gongene)) %>%
+    mutate(treatment = factor(treatment, levels = charlevels)) %>%
+    mutate(external = fct_collapse(treatment, 
+                                   "none" = c("control", "bldg"),
+                                   "eggs" = c("lay" , "inc.d3", 
+                                              "inc.d9", "inc.d17"),
+                                   "chicks" = c("hatch", "n5", "n9")))
 
-#b <- plot_grid(b1,b2,b3, ncol = 1, rel_heights = c(1.1,1,1))
-c123 <- plot_grid(c1,c3,c5, ncol = 1, rel_heights = c(1.1,1,1))
-c456 <- plot_grid(c2,c4,c6, ncol = 1, rel_heights = c(1.1,1,1))
-d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1), labels = c("A", " ",  "B", " ", "C"), label_size = 8)
-e <- plot_grid(e1,e2,e3,e4,e5,e6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))
+  a1 <- newcandidateboxplot("hypothalamus", hypgene, "female", "female hypothalamus", "sequential") 
+  a2 <- externalboxplots("hypothalamus", hypgene, "female", " ", "external") 
+  a3 <- newcandidateboxplot("hypothalamus", hypgene, "male", "male hypothalamus", "sequential")
+  a4 <- externalboxplots( "hypothalamus", hypgene, "female", " ", "external")
 
-fig2 <- plot_grid(d, c123,e,c456,nrow  = 1)
-fig2
+  hypplots <- plot_grid(a1,a2, a3, a4, rel_widths = c(2,1,2,1), nrow = 1)
 
-a <- candidateboxplot("hypothalamus", c("COMT"), "female") + 
-  labs(title =  "Sequential", 
-       subtitle = "Female hypothalamus" )
+  b1 <- newcandidateboxplot("pituitary", pitgene, "female", "female pituitary", "sequential")
+  b2 <- externalboxplots("pituitary", pitgene, "female", " ", "external") 
+  b3 <- newcandidateboxplot("pituitary", pitgene, "male", "male pituitary", "sequential")
+  b4 <- externalboxplots( "pituitary", pitgene, "female", " ", "external")
 
-b <- externalboxplots("hypothalamus", c("COMT"), "female") + 
-  labs(title =  "External", 
-       subtitle = " " )
+  pitplots <- plot_grid(b1,b2, b3, b4, rel_widths = c(2,1,2,1), nrow = 1)
 
-c <- candidateboxplot("pituitary", c("PRL"), "female") + 
-  labs(subtitle = "Female pituitary" )
-
-d <- externalboxplots("pituitary", c("PRL"), "female") + 
-   labs(subtitle = " " )
-
-e <- candidateboxplot("gonad", c("AVPR1A"), "female") + 
-  labs( subtitle = "Female gonads" ) + 
+  c1 <- newcandidateboxplot("gonad", gongene, "female", "female gonad", "sequential") +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  c2 <- externalboxplots("gonad", gongene, "female", " ", "external") + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  c3 <- newcandidateboxplot("gonad", gongene, "male", "male gonad", "sequential") + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  c4 <- externalboxplots( "gonad", gongene, "female", " ", "external") + 
     theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
-f <- externalboxplots("pituitary", c("AVPR1A"), "female") + 
-   labs( subtitle = " " ) + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  gonplots <- plot_grid(c1,c2, c3, c4, rel_widths = c(2,1,2,1), nrow = 1)
 
-females <- plot_grid(a,b,c,d,e,f, ncol = 2, rel_widths = c(2,1))
+  fig2alt <- plot_grid(hypplots, pitplots, gonplots, nrow = 3,
+                       labels = "AUTO", label_size = 8,
+                       rel_heights = c(1,1,1.2))
+  print(fig2alt)
 
-a2 <- candidateboxplot("hypothalamus", c("COMT"), "male") + 
-  labs(title =  "Sequential", 
-       subtitle = "Male hypothalamus" )
+}
 
-b2 <- externalboxplots("hypothalamus", c("COMT"), "male") + 
-  labs(title =  "External", 
-       subtitle = "  " )
-
-c2 <- candidateboxplot("pituitary", c("PRL"), "male") + 
-  labs(subtitle = "Male pituitary" )
-
-d2 <- externalboxplots("pituitary", c("PRL"), "male") + 
-   labs(subtitle = " " )
-
-e2 <- candidateboxplot("gonad", c("AVPR1A"), "male") + 
-  labs( subtitle = "Male gonads" ) + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-f2 <- externalboxplots("pituitary", c("AVPR1A"), "male") + 
-   labs( subtitle = " " )  + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-males <- plot_grid(a2,b2,c2,d2,e2,f2, ncol = 2, rel_widths = c(2,1))
-
-newfig2 <- plot_grid(females, males, labels = c("A", "B"),
-                     label_size  = 8)
-newfig2
+plot3candidates("COMT","PRL", "AVPR1A")
+plot3candidates("PRL","PRL", "PRL")
+plot3candidates("PRL","FSH", "ESR1")
 ```

### 2020-08-01 analysis/figure-2.Rmd
still broken

@@
   mutate(treatment = factor(treatment, levels = charlevels)) %>%
-  drop_na()
+  drop_na() %>%
+    mutate(external = fct_collapse(treatment, 
+                                   "none" = c("control", "bldg"),
+                                   "eggs" = c("lay" , "inc.d3", 
+                                              "inc.d9", "inc.d17"),
+                                   "chicks" = c("hatch", "n5", "n9")))
 head(candidatevsd)
@@
 candidatevsdwide <- candidatevsd  %>%
+  select(-external) %>%
   pivot_wider(names_from = gene, values_from = counts) 
@@
 c6 <- scattercorrelations(MG, MG$PGR, "PGR", MG$ESR1, "ESR1", "#525252")   + labs(title = " ", subtitle = " " ) 
+```
+
+```{r newfig}
 
 plot3candidates <- function(hypgene, pitgene, gongene){
-
-  df <- candidatevsd %>%
-    filter(gene %in% c(hypgene, pitgene, gongene)) %>%
-    mutate(treatment = factor(treatment, levels = charlevels)) %>%
-    mutate(external = fct_collapse(treatment, 
-                                   "none" = c("control", "bldg"),
-                                   "eggs" = c("lay" , "inc.d3", 
-                                              "inc.d9", "inc.d17"),
-                                   "chicks" = c("hatch", "n5", "n9")))
 
@@
 
-plot3candidates("COMT","PRL", "AVPR1A")
+plot3candidates("COMT","PRL","AVPR1A")
 plot3candidates("PRL","PRL", "PRL")

### 2020-08-06 analysis/boneyard.Rmd
new fig 1, now with a boneyard

--- 2020-05-27 analysis/figure-1.Rmd
@@
----
-title: "Fig1"
-output: md_document
----
-
-```{r setup, include=FALSE}
-library(tidyverse)
-library(cowplot)
-library(Rtsne)
-library(kableExtra) 
-library(ggtext) # for italic fonts
-
-source("../R/themes.R")
-source("../R/functions.R")
-
-knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
-```
-
-#  Experimental design, tSNE analysis, and bar chars
-
-## import limma counts and sample info
-
-```{r data}
-countData <- read_csv("../results/01_limma.csv") %>%
-  column_to_rownames(var = "X1")
-countData <- as.data.frame(t(countData))
-head(countData[1:3])
-
-colData <- read_csv("../metadata/00_colData.csv") %>%
-  mutate(treatment = factor(treatment, levels = alllevels),
-         tissue = factor(tissue, levels = tissuelevels)) %>% 
-  column_to_rownames(var = "X1") 
-head(colData)
-# check ready for analysis
-# row.names(countData) == row.names(colData)
-head(row.names(countData) == row.names(colData))
-```
-
-## tsne
-
-```{r tSNE-df}
-
-# uses count data from limma
-
-# prep for tsne
-chartsne <- subsetmaketsne(tissuelevels, charlevels, sexlevels)
-hyptsne <- subsetmaketsne("hypothalamus", charlevels, sexlevels)
-pittsne <- subsetmaketsne("pituitary", charlevels, sexlevels)
-gontsne <- subsetmaketsne("gonads", charlevels, sexlevels)
-
-ftsne <-  subsetmaketsne(tissuelevels, charlevels, "female")
-mtsne <-  subsetmaketsne(tissuelevels, charlevels, "male")
-```
-
 ## Treatment specific DEGs
@@
 
-# for fig 1
+# for figure
 DEGchar <- allDEG %>% 
@@
 
-# for supple fig 1
+# for suppl figures
 DEGcontrol <- allDEG %>% 
@@
   drop_na()
-
 ```
@@
 
-## Figure 1
+```{r barplots}
 
-```{r fig1, fig.width=7.25, fig.height=7.25}
-
-a <- png::readPNG("../figures/images/fig_fig1a.png")
-a <- ggdraw() +  draw_image(a, scale = 1)
-
-b1 <- plottsneelipse(chartsne, chartsne$tissue, allcolors)  + labs(y = "tSNE 2 ", subtitle = " ")  
 b2 <- sexbarplots(hyppit, 0, 8100) + labs(subtitle = " ", y = "DEGs w/ + LFC") +
@@
                 labels = c("A", "", "", "", "B"), label_size = 8)
-
-d1 <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
-  labs(x = NULL, subtitle = "hypothalamus")  + 
-  facet_wrap(~sex, scales = "free")
-d3 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
-  labs(x = NULL, subtitle = "pituitary") + 
-  facet_wrap(~sex, scales = "free") +
-  theme(strip.text = element_blank())
-d5 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
-   facet_wrap(~sex, scales = "free") +
-  theme(strip.text = element_blank()) +
-  labs(x = "tSNE1 \n \n DEGs = differentially expressed genes \n + LFC = positive log fold change", 
-       subtitle = "gonads")
-d2 <- makebargraph(DEGchar, "hypothalamus","DEGs w/ + LFC", 0, 1800) + 
-  theme(axis.text.x = element_blank(),
-        axis.title.x = element_blank())  
-d4 <- makebargraph(DEGchar, "pituitary","DEGs w/ + LFC", 0, 1800)  +  
-  theme(axis.text.x = element_blank(), 
-        axis.title.x = element_blank(), 
-        strip.text.x = element_blank())   
-d6 <- makebargraph(DEGchar, "gonad","DEGs w/ + LFC", 0, 1800) +  
-  theme(strip.text.x = element_blank()) +
-  scale_x_discrete(labels = comparisonlabelschar) +
-  labs(x = "Sequential comparisons of differential gene expression")
-
-d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 2, rel_heights = c(1,0.9,1.4), rel_widths = c(1,2),
-                  labels = c("C"), label_size = 8)
-
-fig1 <- plot_grid(bc, a, d, nrow = 3, rel_heights = c(0.7,0.7,2))
-fig1
-
-pdf(file="../figures/fig1-1.pdf", width=7.25, height=7.25)
-plot(fig1)
-dev.off()
 ```
@@
 
-## Save files
-
-```{r write}
-# save file for musical genes https://raynamharris.shinyapps.io/musicalgenes/
-#write.csv(allDEG, "../../musicalgenes/data/allDEG.csv")
-
-# tsne files too big to save
-#write.csv(chartsne, "../../musicalgenes/data/tsne.csv")
-
+```{r}
 suppltable1 <- DEGchar %>%
@@
 
-pdf(file="../figures/supplfig-1-1.pdf", width=5, height=5)
-plot(supplfig1)
-dev.off()
-
-pdf(file="../figures/supplfig-1-2.pdf", height=5, width=5)
-plot(supplfig2)
-dev.off()
 ```

### 2020-08-06 analysis/figure-1.Rmd
new fig 1, now with a boneyard

--- 2020-05-21 analysis/figure-1.Rmd
@@
 head(colData)
+
 # check ready for analysis
@@
 
-```{r tSNE-df}
-
-# uses count data from limma
-
-# prep for tsne
-chartsne <- subsetmaketsne(tissuelevels, charlevels, sexlevels)
-hyptsne <- subsetmaketsne("hypothalamus", charlevels, sexlevels)
-pittsne <- subsetmaketsne("pituitary", charlevels, sexlevels)
-gontsne <- subsetmaketsne("gonads", charlevels, sexlevels)
-
-ftsne <-  subsetmaketsne(tissuelevels, charlevels, "female")
-mtsne <-  subsetmaketsne(tissuelevels, charlevels, "male")
-```
-
-## Treatment specific DEGs
-
-```{r suppltables}
-allDEG <- read_csv("../results/03_allDEG.csv") %>%
-  mutate(tissue = factor(tissue, levels = tissuelevel),
-         direction = factor(direction, levels = alllevels))
-
-# for fig 1
-DEGchar <- allDEG %>% 
-  filter(comparison %in% comparisonlevelschar) %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelschar))
-
-# for supple fig 1
-DEGcontrol <- allDEG %>% 
-  filter(grepl("control", comparison),
-         !grepl("m.|early|extend|prolong", comparison))  %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol))
-
-DEGbldg <- allDEG %>% 
-  filter(grepl("bldg", comparison),
-         !grepl("m.|early|extend|prolong", comparison))  %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))  %>%
-  drop_na()
-
-suppltable1 <- DEGchar %>%
-  arrange( tissue, sex, direction, gene)
-head(suppltable1)
-write_csv(suppltable1, "../results/suppltable1.csv")
-
-suppltable2 <- rbind(DEGcontrol, DEGbldg)  %>%
-  arrange( tissue, sex, direction, gene)
-head(suppltable2)
-write_csv(suppltable2, "../results/suppltable2.csv")
-
-```
-
-## Sex and Tissue -related DEGs
-
-```{r DEGs, message=F}
-
-## other DEG stuff
-hypsex <- read_csv("../results/DEseq2/sex/hypothalamus_female_male_DEGs.csv")
-pitsex <- read_csv("../results/DEseq2/sex/pituitary_female_male_DEGs.csv")
-gonsex <- read_csv("../results/DEseq2/sex/gonad_female_male_DEGs.csv")
-
-hyppit <- read_csv("../results/DEseq2/tissue/hypothalamus_pituitary_DEGs.csv") 
-hypgon <- read_csv("../results/DEseq2/tissue/hypothalamus_gonad_DEGs.csv")
-pitgon <- read_csv("../results/DEseq2/tissue/pituitary_gonad_DEGs.csv")
-
-hyppit$direction <- factor(hyppit$direction, levels = c("hypothalamus", "pituitary"))
-hypgon$direction <- factor(hypgon$direction, levels = c("hypothalamus",  "gonad"))
-pitgon$direction <- factor(pitgon$direction, levels = c( "pituitary", "gonad"))
-
+```{r tSNE-df, fig.width=7, fig.height=5.25}
+# prep for tsne useing count data from limma and the custom `subsetmaketsne` function
+hyptsne <- subsetmaketsne("hypothalamus", alllevels, sexlevels)
+pittsne <- subsetmaketsne("pituitary", alllevels, sexlevels)
+gontsne <- subsetmaketsne("gonads", alllevels, sexlevels)
 ```
@@
 
-```{r fig1, fig.width=7.25, fig.height=7.25}
+```{r fig1, fig.width=7, fig.height=5.5}
 
@@
 
-b1 <- plottsneelipse(chartsne, chartsne$tissue, allcolors)  + labs(y = "tSNE 2 ", subtitle = " ")  
-b2 <- sexbarplots(hyppit, 0, 8100) + labs(subtitle = " ", y = "DEGs w/ + LFC") +
-  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
-  scale_x_discrete(labels=c("hypothalamus" = "hyp", "pituitary" = "pit" ))
-b3 <- sexbarplots(hypgon, 0, 8100) + labs(subtitle = " " ) +
-  theme(axis.text.y = element_blank(), axis.line.y = element_blank(), 
-        axis.ticks = element_blank(), axis.title.y = element_blank()) +
-    scale_x_discrete(labels=c("hypothalamus" = "hyp",  "gonad" = "gon"))
-b4 <- sexbarplots(pitgon, 0, 8100) + labs(subtitle = " " ) +
-  theme(axis.text.y = element_blank(), axis.line.y = element_blank(),  
-        axis.ticks = element_blank(), axis.title.y = element_blank()) +
-    scale_x_discrete(labels=c(  "pituitary" = "pit", "gonad" = "gon"))
-c1 <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = "tSNE 2 ", subtitle = " ")    
-c2 <- sexbarplots(hypsex, 0, 8100) + labs(y = "DEGs w/ + LFC", subtitle = "hypothalamus" ) +
-  theme(axis.text.y = element_blank(), axis.ticks = element_blank())
-c3 <- sexbarplots(pitsex, 0, 8100) + labs(subtitle = "pituitary" ) + 
-  theme(axis.text.y = element_blank(), axis.line.y = element_blank(),  
-        axis.ticks = element_blank(), axis.title.y = element_blank())
-c4 <- sexbarplots(gonsex, 0, 8100) + labs(subtitle = "gonads" ) +
-  theme(axis.text.y = element_blank(), axis.line.y = element_blank(),  
-        axis.ticks = element_blank(), axis.title.y = element_blank())
-bc <- plot_grid(b1,b2,b3,b4, c1,c2,c3,c4, nrow = 1, rel_widths = c(1.5,1.1,0.9,0.9,1.5,1.1,0.9,0.9),
-                labels = c("A", "", "", "", "B"), label_size = 8)
+e1 <- plottsneelipsev3(hyptsne, hyptsne$treatment, allcolors) + 
+  labs(subtitle = "hypothalamus") 
+e2 <- plottsneelipsev3(pittsne, pittsne$treatment, allcolors ) + 
+  labs(subtitle = "pituitary") 
+  theme(strip.text = element_blank())
+e3 <- plottsneelipsev3(gontsne, gontsne$treatment, allcolors ) + 
+  labs(subtitle = "gonads", caption = "Perplexity = 50")
 
-d1 <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + 
-  labs(x = NULL, subtitle = "hypothalamus")  + 
-  facet_wrap(~sex, scales = "free")
-d3 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + 
-  labs(x = NULL, subtitle = "pituitary") + 
-  facet_wrap(~sex, scales = "free") +
-  theme(strip.text = element_blank())
-d5 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + 
-   facet_wrap(~sex, scales = "free") +
-  theme(strip.text = element_blank()) +
-  labs(x = "tSNE1 \n \n DEGs = differentially expressed genes \n + LFC = positive log fold change", 
-       subtitle = "gonads")
-d2 <- makebargraph(DEGchar, "hypothalamus","DEGs w/ + LFC", 0, 1800) + 
-  theme(axis.text.x = element_blank(),
-        axis.title.x = element_blank())  
-d4 <- makebargraph(DEGchar, "pituitary","DEGs w/ + LFC", 0, 1800)  +  
-  theme(axis.text.x = element_blank(), 
-        axis.title.x = element_blank(), 
-        strip.text.x = element_blank())   
-d6 <- makebargraph(DEGchar, "gonad","DEGs w/ + LFC", 0, 1800) +  
-  theme(strip.text.x = element_blank()) +
-  scale_x_discrete(labels = comparisonlabelschar) +
-  labs(x = "Sequential comparisons of differential gene expression")
+e <- plot_grid(e1,e2,e3, ncol = 3, labels = c("E"), label_size = 8)
 
-d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 2, rel_heights = c(1,0.9,1.4), rel_widths = c(1,2),
-                  labels = c("C"), label_size = 8)
-
-fig1 <- plot_grid(bc, a, d, nrow = 3, rel_heights = c(0.7,0.7,2))
+fig1 <- plot_grid(a, e, nrow = 2, rel_heights = c(2.5,2))
 fig1
-
-pdf(file="../figures/fig1-1.pdf", width=7.25, height=7.25)
-plot(fig1)
-dev.off()
-```
-
-## suppl fig 1 controls versus bldg
-
-```{r supplfig-1, fig.width=5, fig.height=5}
-
-s1a <- makebargraphsuppl(DEGcontrol, "hypothalamus","DEGs w/ + LFC", 0, 5000) + 
-  theme(axis.text.x = element_blank(),
-        axis.title.x = element_blank()) + 
-  labs(subtitle = "hypothalamus") 
-s1c <- makebargraphsuppl(DEGcontrol, "pituitary","DEGs w/ + LFC", 0, 5000)  +  
-  theme(axis.text.x = element_blank(), 
-        axis.title.x = element_blank(), 
-        strip.text.x = element_blank())    + 
-  labs(subtitle = "pituitary") 
-s1e <- makebargraphsuppl(DEGcontrol, "gonad","DEGs w/ + LFC", 0, 5000) +  
-  theme(strip.text.x = element_blank()) +
-  scale_x_discrete(labels = comparisonlabelscontrol) +
-  labs(subtitle = "gonads")
-
-s1b <- makebargraphsuppl(DEGbldg, "hypothalamus", NULL, 0, 1000) + 
-  theme(axis.text.x = element_blank(),
-        axis.title.x = element_blank())  
-s1d <- makebargraphsuppl(DEGbldg, "pituitary",NULL, 0, 1000)  +  
-  theme(axis.text.x = element_blank(), 
-        axis.title.x = element_blank(), 
-        strip.text.x = element_blank())   
-s1f <- makebargraphsuppl(DEGbldg, "gonad",NULL, 0, 1000) +  
-  theme(strip.text.x = element_blank()) +
-  scale_x_discrete(labels = comparisonlabelsbldg)
-
-s1 <- plot_grid(s1a, s1b, 
-          s1c, s1d, 
-          s1e, s1f, nrow = 3,
-          rel_heights = c(1,0.9,1.3),
-          labels = c("A", "B"), label_size = 8, rel_widths = c(1.1,1))
-
-sa <- png::readPNG("../figures/images/fig_supplfig-1.png")
-sa <- ggdraw() +  draw_image(sa, scale = 1)
-
-supplfig1 <- plot_grid(s1, sa, ncol = 1, rel_heights = c(2,0.7))
-supplfig1
-
-pdf(file="../figures/supplfig-1-1.pdf", width=5, height=5)
-plot(supplfig1)
-dev.off()
 ```
@@
 
+pdf(file="../figures/fig1-1.pdf", width=7.25, height=7.25)
+plot(fig1)
+dev.off()
 ```

### 2020-08-06 R/functions.R

@@ -813,7 +813,7 @@ plotprolactin <- function(df, myy, myylab, mysubtitle ){
   return(p)
 }
 
-## tsne  figure 1 and 5
+## subsetmaketsne 
 
 subsetmaketsne <- function(whichtissue, whichtreatment, whichsex){
   
@@ -833,7 +833,8 @@ subsetmaketsne <- function(whichtissue, whichtreatment, whichsex){
   
   euclidist <- dist(countData) # euclidean distances between the rows
   
-  tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=10, theta=0.5, dims=2)
+  tsne_model <- Rtsne(euclidist, check_duplicates=FALSE,
+                      perplexity = 50)
   tsne_df = as.data.frame(tsne_model$Y) 
   
   # prep for adding columns
@@ -870,6 +871,18 @@ plottsneelipsev2 <- function(tsnedf, pointcolor, whichcolors){
   return(p)
 }
 
+plottsneelipsev3 <- function(tsnedf, pointcolor, whichcolors){
+  p <- ggplot(tsnedf, aes(x = V1, y = V2)) +
+    # stat_ellipse(linetype = 1, aes(color = pointcolor)) +
+    geom_point(size = 1.5, aes(color = pointcolor, shape = sex)) +
+    theme_B3() +
+    labs(x = "tSNE 1", y = "tSNE 2", 
+         title = " ", caption = " ") +
+    scale_color_manual(values = whichcolors) +
+    theme(legend.position = "none",
+          axis.text = element_blank(), axis.ticks = element_blank()) 
+  return(p)
+}
 
 makebargraph <- function(df, whichtissue, myylab, lowlim, higherlim){
   p <- df %>%

### 2020-08-06 analysis/figure-1.Rmd
better fig 1

@@
 
-```{r tSNE-df, fig.width=7, fig.height=5.25}
+```{r tSNE-df}
 # prep for tsne useing count data from limma and the custom `subsetmaketsne` function
@@
 
-```{r fig1, fig.width=7, fig.height=5.5}
+```{r fig1, fig.width=7, fig.height=4.5}
 
@@
 e3 <- plottsneelipsev3(gontsne, gontsne$treatment, allcolors ) + 
-  labs(subtitle = "gonads", caption = "Perplexity = 50")
+  labs(subtitle = "gonads")
 
-e <- plot_grid(e1,e2,e3, ncol = 3, labels = c("E"), label_size = 8)
+e <- plot_grid(e1,e2,e3, ncol = 3, labels = c("D"), label_size = 8)
 
-fig1 <- plot_grid(a, e, nrow = 2, rel_heights = c(2.5,2))
+fig1 <- plot_grid(a, e, nrow = 2, rel_heights = c(2.4,1.6))
 fig1
@@
 ```{r write}
-# save file for musical genes https://raynamharris.shinyapps.io/musicalgenes/
-#write.csv(allDEG, "../../musicalgenes/data/allDEG.csv")
-
-# tsne files too big to save
-#write.csv(chartsne, "../../musicalgenes/data/tsne.csv")
-
-pdf(file="../figures/fig1-1.pdf", width=7.25, height=7.25)
+pdf(file="../figures/fig1-1.pdf", width=7, height=4.5)
 plot(fig1)

### 2020-08-06 analysis/00_datawrangling.Rmd
revisiting kallisto

@@
 kallistodata <- read.table("../results/kallistocounts.txt", 
-                        sep = ",", row.names = NULL) %>%
-  rename("NCBI" = "entrezid",
+                        sep = ",", row.names = NULL) 
+kallistodata <- kallistodata %>%
+  dplyr::rename("NCBI" = "entrezid",
          "gene" = "Name")
-#head(kallistodata)
+head(kallistodata)
 

### 2020-08-06 analysis/figure-1.Rmd
higher res?
remove unneded packages

@@
 library(Rtsne)
-library(kableExtra) 
-library(ggtext) # for italic fonts
 
@@
 dev.off()
+
+png("../figures/fig1-1.png", width = 7, height = 4.5, 
+    units = 'in', res = 300)
+plot(fig1) # Make plot
+dev.off()
 ```

### 2020-08-06 analysis/boneyard.Rmd
not sure i like it but same 3 figs for hyp pit gon

@@
 ## other DEG stuff
-hypsex <- read_csv("../results/DEseq2/sex/hypothalamus_female_male_DEGs.csv")
+
 pitsex <- read_csv("../results/DEseq2/sex/pituitary_female_male_DEGs.csv")
@@
     scale_x_discrete(labels=c(  "pituitary" = "pit", "gonad" = "gon"))
-c1 <- plottsneelipse(chartsne, chartsne$sex, allcolors)   + labs(y = "tSNE 2 ", subtitle = " ")    
-c2 <- sexbarplots(hypsex, 0, 8100) + labs(y = "DEGs w/ + LFC", subtitle = "hypothalamus" ) +
-  theme(axis.text.y = element_blank(), axis.ticks = element_blank())
+
 c3 <- sexbarplots(pitsex, 0, 8100) + labs(subtitle = "pituitary" ) + 
@@
 
-s1a <- makebargraphsuppl(DEGcontrol, "hypothalamus","DEGs w/ + LFC", 0, 5000) + 
-  theme(axis.text.x = element_blank(),
-        axis.title.x = element_blank()) + 
-  labs(subtitle = "hypothalamus") 
 s1c <- makebargraphsuppl(DEGcontrol, "pituitary","DEGs w/ + LFC", 0, 5000)  +  
@@
 
-s1b <- makebargraphsuppl(DEGbldg, "hypothalamus", NULL, 0, 1000) + 
-  theme(axis.text.x = element_blank(),
-        axis.title.x = element_blank())  
 s1d <- makebargraphsuppl(DEGbldg, "pituitary",NULL, 0, 1000)  +  
@@
 # for volcano plots
-hyphatch <- DEGchar %>% filter(tissue == "hypothalamus", comparison == "hatch_n5") 
+
 pitinc <- DEGchar %>% filter(tissue == "pituitary", comparison == "inc.d9_inc.d17") 

### 2020-08-06 analysis/figure-1.Rmd
now with session info
not sure i like it but same 3 figs for hyp pit gon

@@
 
-```{r setup, include=FALSE}
+```{r setup, include=T}
 library(tidyverse)
@@
 dev.off()
+
+sessionInfo()
 ```

### 2020-08-06 analysis/gonad.Rmd
not sure i like it but same 3 figs for hyp pit gon

---
title: "Gonad"
output: md_document
---

# Gonad

```{r setup, include=T}
library(tidyverse)
library(cowplot)
library(ggsignif)

source("../R/themes.R")
source("../R/functions.R")

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
```

## Treatment specific DEGs in the Hypothalamsu

```{r suppltables}
allDEG <- read_csv("../results/03_allDEG.csv") %>%
  mutate(tissue = factor(tissue, levels = tissuelevel),
         direction = factor(direction, levels = alllevels),
         label = gsub("_","\nvs. ", comparison)) %>%
  filter(tissue == "gonad")
allDEG

# for suppl figures
DEGcontrol <- allDEG %>% 
  filter(grepl("control", comparison),
         !grepl("m.|early|extend|prolong", comparison))  %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol))

DEGbldg <- allDEG %>% 
  filter(grepl("bldg", comparison),
         !grepl("m.|early|extend|prolong", comparison))  %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))  %>%
  drop_na()

DEGchar <- allDEG %>% 
  filter(comparison %in% comparisonlevelschar,
         !grepl("control|bldg", comparison)) %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelschar))

candidatevsd <- read_csv("../results/03_candidatevsd.csv") %>% 
  select(-X1) %>%
  filter(treatment %in% charlevels) %>%
  mutate(treatment = factor(treatment, levels = charlevels)) %>%
  drop_na() %>%
    mutate(external = fct_collapse(treatment, 
                                   "none" = c("control", "bldg"),
                                   "eggs" = c("lay" , "inc.d3", 
                                              "inc.d9", "inc.d17"),
                                   "chicks" = c("hatch", "n5", "n9")))
```

```{r gonad, fig.width=7, fig.height=7}

a <- plot.volcano("gonad", sexlevels,  "control_bldg") + 
  facet_wrap(~sex) +  labs(title = "Gonad") 
b <- makebargraph(DEGcontrol, "gonad","DEGs w/ + LFC", 0, 4200, comparisonlabelscontrol) 
  labs(title = " ")

c <- makebargraph(DEGbldg, "gonad", "DEGs w/ + LFC", 0, 2200, comparisonlabelsbldg) 
d <- makebargraph(DEGchar, "gonad","DEGs w/ + LFC", 0, 1500, comparisonlabelscharnobldg)

e <- candidateboxplot("gonad",  c("ESR1"), sexlevels)
f <- candidateboxplot("gonad",  c("PGR"), sexlevels)

ab <- plot_grid(a,b,rel_widths = c(1,3), labels = c("A", "B"), label_size = 8)
cd <- plot_grid(c,d,rel_widths = c(1,1), labels = c("C", "D"), label_size = 8)
ef <- plot_grid(e,f,rel_widths = c(1,1), labels = c("E", "F"), label_size = 8)

goncharfig <- plot_grid(ab,cd,ef, ncol = 1)
goncharfig
```

## Save files

```{r write}
pdf(file="../figures/pitcharfig.pdf", width=7, height=7)
plot(goncharfig)
dev.off()

png("../figures/pitcharfig.png", width = 7, height = 7, 
    units = 'in', res = 300)
plot(goncharfig) # Make plot
dev.off()

sessionInfo()
```

### 2020-08-06 analysis/hypothalamus.Rmd
not sure i like it but same 3 figs for hyp pit gon

--- 2020-08-06 analysis/gonad.Rmd
@@
 ---
-title: "Gonad"
+title: "Hypothalamus"
 output: md_document
@@
 
-# Gonad
+# Hypothalamus
 
@@
          label = gsub("_","\nvs. ", comparison)) %>%
-  filter(tissue == "gonad")
+  filter(tissue == "hypothalamus")
 allDEG
@@
                                    "chicks" = c("hatch", "n5", "n9")))
+head(candidatevsd)
 ```
 
-```{r gonad, fig.width=7, fig.height=7}
+```{r hypothalamus, fig.width=7, fig.height=7}
 
-a <- plot.volcano("gonad", sexlevels,  "control_bldg") + 
-  facet_wrap(~sex) +  labs(title = "Gonad") 
-b <- makebargraph(DEGcontrol, "gonad","DEGs w/ + LFC", 0, 4200, comparisonlabelscontrol) 
+a <- plot.volcano("hypothalamus", sexlevels,  "control_bldg") + 
+  facet_wrap(~sex) +  labs(title = "Hypothalamus") 
+b <- makebargraph(DEGcontrol, "hypothalamus","DEGs w/ + LFC", 0, 4800, comparisonlabelscontrol) 
   labs(title = " ")
 
-c <- makebargraph(DEGbldg, "gonad", "DEGs w/ + LFC", 0, 2200, comparisonlabelsbldg) 
-d <- makebargraph(DEGchar, "gonad","DEGs w/ + LFC", 0, 1500, comparisonlabelscharnobldg)
+c <- makebargraph(DEGbldg, "hypothalamus", "DEGs w/ + LFC", 0, 480, comparisonlabelsbldg) 
+d <- makebargraph(DEGchar, "hypothalamus","DEGs w/ + LFC", 0, 1700, comparisonlabelscharnobldg)
 
-e <- candidateboxplot("gonad",  c("ESR1"), sexlevels)
-f <- candidateboxplot("gonad",  c("PGR"), sexlevels)
+e <- candidateboxplot("hypothalamus",  c("DRD1"), sexlevels)
+f <- candidateboxplot("hypothalamus",  c("COMT"), sexlevels)
 
@@
 
-goncharfig <- plot_grid(ab,cd,ef, ncol = 1)
-goncharfig
+hypcharfig <- plot_grid(ab,cd,ef, ncol = 1)
+hypcharfig
 ```
@@
 ```{r write}
-pdf(file="../figures/pitcharfig.pdf", width=7, height=7)
-plot(goncharfig)
+pdf(file="../figures/hypcharfig.pdf", width=7, height=7)
+plot(hypcharfig)
 dev.off()
 
-png("../figures/pitcharfig.png", width = 7, height = 7, 
+png("../figures/hypcharfig.png", width = 7, height = 7, 
     units = 'in', res = 300)
-plot(goncharfig) # Make plot
+plot(hypcharfig) # Make plot
 dev.off()

### 2020-08-06 analysis/pituitary.Rmd
not sure i like it but same 3 figs for hyp pit gon

--- 2020-08-06 analysis/gonad.Rmd
@@
 ---
-title: "Gonad"
+title: "Pituitary"
 output: md_document
@@
 
-# Gonad
+# Pituitary
 
@@
          label = gsub("_","\nvs. ", comparison)) %>%
-  filter(tissue == "gonad")
+  filter(tissue == "pituitary")
 allDEG
@@
 
-```{r gonad, fig.width=7, fig.height=7}
+```{r pituitary, fig.width=7, fig.height=7}
 
-a <- plot.volcano("gonad", sexlevels,  "control_bldg") + 
-  facet_wrap(~sex) +  labs(title = "Gonad") 
-b <- makebargraph(DEGcontrol, "gonad","DEGs w/ + LFC", 0, 4200, comparisonlabelscontrol) 
+a <- plot.volcano("pituitary", sexlevels,  "control_bldg") + 
+  facet_wrap(~sex) +  labs(title = "Pituitary") 
+b <- makebargraph(DEGcontrol, "pituitary","DEGs w/ + LFC", 0, 4200, comparisonlabelscontrol) 
   labs(title = " ")
 
-c <- makebargraph(DEGbldg, "gonad", "DEGs w/ + LFC", 0, 2200, comparisonlabelsbldg) 
-d <- makebargraph(DEGchar, "gonad","DEGs w/ + LFC", 0, 1500, comparisonlabelscharnobldg)
+c <- makebargraph(DEGbldg, "pituitary", "DEGs w/ + LFC", 0, 2200, comparisonlabelsbldg) 
+d <- makebargraph(DEGchar, "pituitary","DEGs w/ + LFC", 0, 1500, comparisonlabelscharnobldg)
 
-e <- candidateboxplot("gonad",  c("ESR1"), sexlevels)
-f <- candidateboxplot("gonad",  c("PGR"), sexlevels)
+e <- candidateboxplot("pituitary",  c("PRL"), sexlevels)
+f <- candidateboxplot("pituitary",  c("FSHB"), sexlevels)
 
@@
 
-goncharfig <- plot_grid(ab,cd,ef, ncol = 1)
-goncharfig
+pitcharfig <- plot_grid(ab,cd,ef, ncol = 1)
+pitcharfig
 ```
@@
 pdf(file="../figures/pitcharfig.pdf", width=7, height=7)
-plot(goncharfig)
+plot(pitcharfig)
 dev.off()
@@
     units = 'in', res = 300)
-plot(goncharfig) # Make plot
+plot(pitcharfig) # Make plot
 dev.off()

### 2020-08-06 R/functions.R

@@ -421,11 +421,32 @@ plot.volcano <- function(whichtissue, whichsex,  whichcomparison){
                alpha = 0.75, na.rm = T) + 
     theme_B3() +
     scale_color_manual(values = allcolors, 
-                       name = "stage with\nincreased\nexpression\n(+LFC)") +
-    xlim(-10,5) +
-    labs(y = "-log10(adj. p-value)", x = NULL)
+                       name = "increased expression in:") +
+    labs(y = "-log10(adj. p-value)", 
+         x = "Log-fold change (LFC)") +
+    theme(legend.position = "bottom",
+          legend.box = "vertical", 
+          legend.direction = "vertical",
+          legend.key.height=unit(0, "cm"),      
+          plot.margin = unit(c(1,0.5,0,0.5), "lines")) +
+    guides(color = guide_legend(nrow = 1))
   return(volcano)
 }
+
+plot.volcano.sex <- function(df){
+  
+  volcano <- df %>%
+    ggplot(aes(x = lfc, y = logpadj)) + 
+    geom_point(aes(color = direction), size = 1, 
+               alpha = 0.75, na.rm = T) + 
+    theme_B3() +
+    scale_color_manual(values = allcolors, 
+                       name = NULL) +
+    labs(y = "-log10(adj. p-value)", x = "Log-fold change (LFC)") +
+    theme(legend.position = "top") 
+  return(volcano)
+}
+
 
 
 # made for beta testing
@@ -883,7 +904,7 @@ plottsneelipsev3 <- function(tsnedf, pointcolor, whichcolors){
   return(p)
 }
 
-makebargraph <- function(df, whichtissue, myylab, lowlim, higherlim){
+makebargraph <- function(df, whichtissue, myylab, lowlim, higherlim, mylabels){
   p <- df %>%
     filter(tissue == whichtissue) %>%
     ggplot(aes(x = comparison,  fill = direction)) +
@@ -891,12 +912,8 @@ makebargraph <- function(df, whichtissue, myylab, lowlim, higherlim){
     facet_wrap(~sex) +
     theme_B3() +
     theme(legend.position = "none",
-          axis.text.x = element_text(angle = 45, hjust = 1),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank())  +
-    labs(x = NULL, 
-         y = myylab,
-         subtitle = " ") +
+          axis.text.x = element_text(angle = 45, hjust = 1))  +
+    labs(x = NULL, y = myylab) +
[…]

### 2020-08-07 analysis/boneyard.Rmd
now with 1 total DEGs figure

@@
 ```
+
+```{r}
+candidatevsd <- read_csv("../results/03_candidatevsd.csv") %>% 
+  select(-X1) %>%
+  filter(treatment %in% charlevels) %>%
+  mutate(treatment = factor(treatment, levels = charlevels)) %>%
+  drop_na() %>%
+    mutate(external = fct_collapse(treatment, 
+                                   "none" = c("control", "bldg"),
+                                   "eggs" = c("lay" , "inc.d3", 
+                                              "inc.d9", "inc.d17"),
+                                   "chicks" = c("hatch", "n5", "n9")))
+head(candidatevsd)
+```

### 2020-08-07 analysis/figure-2-new.Rmd
now with 1 total DEGs figure

---
title: "Fig 2"
output: md_document
---

# Fig 2

```{r setup, include=T}
library(tidyverse)
library(cowplot)
library(ggsignif)

source("../R/themes.R")
source("../R/functions.R")

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
```

## Treatment specific DEGs

```{r suppltables}
allDEG <- read_csv("../results/03_allDEG.csv") %>%
  mutate(tissue = factor(tissue, levels = tissuelevel),
         direction = factor(direction, levels = alllevels),
         label = gsub("_","\nvs. ", comparison)) 
head(allDEG)

# for suppl figures
DEGcontrol <- allDEG %>% 
  filter(grepl("control", comparison),
         !grepl("m.|early|extend|prolong", comparison))  %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol))

DEGbldg <- allDEG %>% 
  filter(grepl("bldg", comparison),
         !grepl("m.|early|extend|prolong", comparison))  %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))  %>%
  drop_na()

DEGchar <- allDEG %>% 
  filter(comparison %in% comparisonlevelschar,
         !grepl("control|bldg", comparison)) %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelschar))
```

```{r fig2-new, fig.width=7, fig.height=7}

a <- plot.volcano("hypothalamus", sexlevels,  "control_bldg") + 
  facet_wrap(~sex) + labs(subtitle = " ")

b <- makebargraph(DEGcontrol, "hypothalamus","DEGs w/ + LFC", 0, 4800, comparisonlabelscontrol) +
  labs(x = "Control versus all other reproductive and parental stages")

c <- makebargraph(DEGbldg, "hypothalamus", "DEGs w/ + LFC", 0, 480, comparisonlabelsbldg) +
  labs(x = "Nest-building versus all other parental stages")
d <- makebargraph(DEGchar, "hypothalamus","DEGs w/ + LFC", 0, 1700, comparisonlabelscharnobldg) +
  labs(x = "Comparison of sequential parental stages")

e <- makebargraph(DEGbldg, "pituitary", "DEGs w/ + LFC", 0, 2200, comparisonlabelsbldg)

f <- makebargraph(DEGchar, "pituitary","DEGs w/ + LFC", 0, 1500, comparisonlabelscharnobldg)

g <- makebargraph(DEGbldg, "gonad", "DEGs w/ + LFC", 0, 2800, comparisonlabelsbldg) 
h <- makebargraph(DEGchar, "gonad","DEGs w/ + LFC", 0, 1200, comparisonlabelscharnobldg)

ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
                labels = c("A Hypothalamus", " " ), label_size = 8, hjust = 0)
cd <- plot_grid(c,d,rel_widths = c(1,1))
ef <- plot_grid(e,f,rel_widths = c(1,1),
                labels = c("B Pitutary"), label_size = 8 ,hjust = 0)
gh <- plot_grid(g,h,rel_widths = c(1,1), 
                labels = c("C Gonads"), label_size = 8, hjust = 0)

fig <- plot_grid(ab,cd,ef, gh, ncol = 1)
fig
```

## Save files

```{r write}
pdf(file="../figures/fig2-new-1.pdf", width=7, height=7)
plot(fig)
dev.off()

png("../figures/fig2-new-1.png", width = 7, height = 7, 
    units = 'in', res = 300)
plot(fig) # Make plot
dev.off()

sessionInfo()
```

### 2020-08-07 analysis/figure-2-new.Rmd
fig 2 margin legend foo

@@
 
-a <- plot.volcano("hypothalamus", sexlevels,  "control_bldg") + 
-  facet_wrap(~sex) + labs(subtitle = " ")
-
 b <- makebargraph(DEGcontrol, "hypothalamus","DEGs w/ + LFC", 0, 4800, comparisonlabelscontrol) +
-  labs(x = "Control versus all other reproductive and parental stages")
+  labs(subtitle = "Control versus all other reproductive and parental stages")
 
 c <- makebargraph(DEGbldg, "hypothalamus", "DEGs w/ + LFC", 0, 480, comparisonlabelsbldg) +
-  labs(x = "Nest-building versus all other parental stages")
+  labs(subtitle = "Nest-building versus all other parental stages")
 d <- makebargraph(DEGchar, "hypothalamus","DEGs w/ + LFC", 0, 1700, comparisonlabelscharnobldg) +
-  labs(x = "Comparison of sequential parental stages")
+  labs(subtitle = "Comparison of sequential parental stages")
 
@@
 h <- makebargraph(DEGchar, "gonad","DEGs w/ + LFC", 0, 1200, comparisonlabelscharnobldg) 
-
-ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
-                labels = c("A Hypothalamus", " " ), label_size = 8, hjust = 0)
 cd <- plot_grid(c,d,rel_widths = c(1,1))
@@
                 labels = c("C Gonads"), label_size = 8, hjust = 0)
+
+a <- plot.volcano("hypothalamus", sexlevels,  "control_bldg") + 
+  facet_wrap(~sex) + labs(subtitle = " ")
+ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
+                labels = c("A Hypothalamus"), label_size = 8, hjust = 0)
 

### 2020-08-07 R/functions.R

@@ -427,8 +427,8 @@ plot.volcano <- function(whichtissue, whichsex,  whichcomparison){
     theme(legend.position = "top",
           legend.direction = "vertical",
           legend.key.height = unit(0, "cm"),      
-          plot.margin = unit(c(0,0,0,0), "lines"),
-          legend.spacing.y = unit(0, "cm")) +
+          legend.spacing.y = unit(0, "cm"),
+          legend.margin=margin(0,0, -5, 0)) +
     guides(color = guide_legend(nrow = 1))
   return(volcano)
 }

### 2020-08-07 analysis/figure-2-new.Rmd
small edits

--- 2020-08-07 analysis/figure-2-new.Rmd
@@
 
-```{r fig2-new, fig.width=7, fig.height=7}
-
+```{r fig2-new, fig.width=7.25, fig.height=7.25}
 a <- plot.volcano("hypothalamus", sexlevels,  "control_bldg") + 
   facet_wrap(~sex) + labs(subtitle = " ")
+b <- makebargraph(DEGcontrol, "hypothalamus","DEGs w/ + LFC", 0, 5000, comparisonlabelscontrol) +
+  labs(title = " ",
+       subtitle = "Control versus all other reproductive and parental stages")
 
-b <- makebargraph(DEGcontrol, "hypothalamus","DEGs w/ + LFC", 0, 4800, comparisonlabelscontrol) +
-  labs(x = "Control versus all other reproductive and parental stages")
+c <- makebargraph(DEGbldg, "hypothalamus", "DEGs w/ + LFC", 0, 490, comparisonlabelsbldg) +
+  labs(subtitle = "Nest-building versus all other parental stages")
+d <- makebargraph(DEGchar, "hypothalamus", NULL, 0, 1800, comparisonlabelscharnobldg) +
+  labs(subtitle = "Comparison of sequential parental stages")
 
-c <- makebargraph(DEGbldg, "hypothalamus", "DEGs w/ + LFC", 0, 480, comparisonlabelsbldg) +
-  labs(x = "Nest-building versus all other parental stages")
-d <- makebargraph(DEGchar, "hypothalamus","DEGs w/ + LFC", 0, 1700, comparisonlabelscharnobldg) +
-  labs(x = "Comparison of sequential parental stages")
+e <- makebargraph(DEGbldg, "pituitary", "DEGs w/ + LFC", 0, 2300, comparisonlabelsbldg)
 
-e <- makebargraph(DEGbldg, "pituitary", "DEGs w/ + LFC", 0, 2200, comparisonlabelsbldg)
+f <- makebargraph(DEGchar, "pituitary",NULL, 0, 1600, comparisonlabelscharnobldg)
 
-f <- makebargraph(DEGchar, "pituitary","DEGs w/ + LFC", 0, 1500, comparisonlabelscharnobldg)
-
-g <- makebargraph(DEGbldg, "gonad", "DEGs w/ + LFC", 0, 2800, comparisonlabelsbldg) 
-h <- makebargraph(DEGchar, "gonad","DEGs w/ + LFC", 0, 1200, comparisonlabelscharnobldg)
+g <- makebargraph(DEGbldg, "gonad", "DEGs w/ + LFC", 0, 2900, comparisonlabelsbldg) 
+h <- makebargraph(DEGchar, "gonad", NULL, 0, 1200, comparisonlabelscharnobldg)
 
 ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
-                labels = c("A Hypothalamus", " " ), label_size = 8, hjust = 0)
-cd <- plot_grid(c,d,rel_widths = c(1,1))
-ef <- plot_grid(e,f,rel_widths = c(1,1),
+                labels = c("A Hypothalamus"), label_size = 8, hjust = 0)
+cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h")
+ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h",
                 labels = c("B Pitutary"), label_size = 8 ,hjust = 0)
-gh <- plot_grid(g,h,rel_widths = c(1,1), 
+gh <- plot_grid(g,h,rel_widths = c(1.1,1), align = "h",
                 labels = c("C Gonads"), label_size = 8, hjust = 0)
@@
     units = 'in', res = 300)
-plot(fig) # Make plot
+plot(fig) 
 dev.off()

### 2020-08-07 analysis/03_datawrangling.Rmd
pca and some candidate now suppl fig 1

@@
 
-literaturegenes <- c(curleychampagnegenes, 
+datadrivengenes <- c("COX1", "COX2", "CYTB", "ATP2B4", "LOC107055658", 
+                     "CHGA", "TF", "OVALY", "ANXA5")
+
+literaturegenes <- c(curleychampagnegenes, datadrivengenes,
                      "VIP", "GNIH", "GAL", "TH", "THRB",

### 2020-08-07 analysis/boneyard.Rmd
pca and some candidate now suppl fig 1

--- 2020-08-06 analysis/boneyard.Rmd
@@
 ```
+
+```{r}
+## prlpit
+PRLpit <- candidatevsd %>% 
+  filter(gene == "PRL", tissue == "pituitary")
+
+# list of pvalue and DEGs
+allDEG <- read_csv("../results/06_allDEG.csv")  %>%
+  mutate(direction = factor(direction, levels = hypothesislevels))
+
+# DEGs infor for PRL and external
+hilodf <- allDEG %>%
+  filter(comparison == "lo_hi")  %>%
+  arrange(padj) 
+hilogenes <- hilodf %>%  top_n(20) %>% pull(gene)
+
+eggchickdf <- allDEG %>%
+  filter(!comparison %in% c("lo_hi", "nest_eggs", "nest_chicks"))  %>%
+  arrange(padj)  
+eggchickgenes <- eggchickdf %>%  top_n(20) %>% pull(gene)
+
+## data-driven PRL genes
+pithilo <- makecorrdf("female", "pituitary", hilogenes)
+pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[32:52])
+
+## https://www.gynecologiconcology-online.net/article/S0090-8258(19)30069-1/fulltext
+cancergenes <- makecorrdf("female", "pituitary", c(shaidgenes,suszynskaagenes, "PRL"))
+
+## for scatter correlation (not currently usesed)
+candidatevsdwide <- candidatevsd  %>%
+    pivot_wider(names_from = gene, values_from = counts) 
+pitwide <- subsetcandidatevsdwide(sexlevels, "pituitary")
+
+## number of genes total vs plotted
+
+eggchickdf %>% filter(tissue == "pituitary") %>% nrow()
+hilodf %>% filter(tissue == "pituitary") %>% nrow()
+length(WGCNAgenes)
+
+b1 <- png::readPNG("../figures/images/fig_fig3b.png")
+b1 <- ggdraw() +  draw_image(b1, scale = 1)
+
+b2 <- makenewbargraph("pituitary", "female","eggs_chicks", 0, 4000) + labs(subtitle = "female")  + labs(title = " ") +
+  scale_x_discrete(labels = "eggs vs. chicks")
+b3 <- makenewbargraph("pituitary", "male", "eggs_chicks", 0, 4000)  + labs(subtitle = "male")  + labs(title = " ") +
+  theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), 
+        axis.text.y = element_blank(), axis.line.y = element_blank()) +
+  scale_x_discrete(labels = "eggs vs. chicks")
+b4 <- makenewbargraph("pituitary", "male", "lo_hi", 0, 4000)  +
+  scale_x_discrete(labels = "lo vs. hi PRL")
+b5 <- makenewbargraph("pituitary", "female", "lo_hi", 0, 4000)   + 
+  theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), 
+        axis.text.y = element_blank(), axis.line.y = element_blank()) +
+  scale_x_discrete(labels = "lo vs. hi PRL")
+b25 <- plot_grid(b2,b3,b4,b5, nrow = 2, rel_widths = c(1.2,1), rel_heights = c(1.2,1))
+
+b6 <-  plottopDEGs(eggchickdf, "female",  "LFC eggs vs. chicks", "Top 10 female DEGs") + labs(title = " ")
+b7 <-  plottopDEGs(eggchickdf, "male",  "LFC eggs vs. chicks", "Top 10 male DEGs") + labs(title = " ")
+b8 <-  plottopDEGs(hilodf, "female",  "LFC lo vs. hi PRL", NULL)
+b9 <-  plottopDEGs(hilodf, "male",  "LFC lo vs. hi PRL", NULL)
+b15 <- plot_grid(b1,b25,nrow = 1, rel_widths = c(1.2,1))
+b69 <- plot_grid(b6,b7,b8,b9, align = "v" , rel_heights = c(1.2,1))
+b <- plot_grid(b15,b69, labels = c("C"), label_size = 8, rel_widths = c(1.2,1))
+
+c1 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 DEGs, low vs. high PRL") 
+c2 <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes in WGCNA module with PRL") 
+c3 <- plotcorrplot(cancergenes, NULL) + theme(legend.position = "right") + labs(subtitle = "Breast, ovary, & piuitary cancer genes", color = "Correlation")
+c <- plot_grid(c1,c2,c3, nrow = 1, labels = c("D"), label_size = 8, rel_widths = c(1,1,1.3))
+
+fig3 <- plot_grid(a, b, c, nrow = 3, rel_heights = c(1.2,2,1.8))
+fig3
+```

### 2020-08-07 analysis/figure-1.Rmd
pca and some candidate now suppl fig 1

@@
 
+## Supplemental PCA data
+
+```{r}
+# pca relies on count data from limma
+pca_h1 <- subsetmakepca("hypothalamus", alllevels, sexlevels)	
+pca_h2 <- makefvizdf("hypothalamus", alllevels, sexlevels)
+
+pca_p1 <- subsetmakepca("pituitary", alllevels, sexlevels)	
+pca_p2 <- makefvizdf("pituitary", alllevels, sexlevels)
+
+pca_g1 <- subsetmakepca("gonads", alllevels, sexlevels)	
+pca_g2 <- makefvizdf("gonads", alllevels, sexlevels)	
+```
+
+## Supplemental Data-driven genes vsds for box plots
+
+```{r candidates}
+# genes lists in "./analysis/03_datawrangling.Rmd" and old lists in "./R/genelists.R"
+
+candidatevsd <- read_csv("../results/03_candidatevsd.csv") %>% 
+  select(-X1) %>%
+  filter(treatment %in% alllevels) %>%
+  mutate(treatment = factor(treatment, levels = charlevels)) %>%
+  drop_na() 
+head(candidatevsd)
+```
+
+## Fig 3
+
+```{r supplfig, fig.height=7.25, fig.width=7.25}
+
+a <- plotpc12(pca_h1, pca_h2, pca_h1$treatment, allcolors, "Hypothalamus", NULL) 
+c <- plotpc12(pca_p1, pca_p2, pca_p1$treatment, allcolors, "Pituitary", NULL) 
+e <- plotpc12(pca_g1, pca_g2, pca_g1$treatment, allcolors, "Gonads", NULL)
+
+b1 <- fviz_screeplot(pca_h2, addlabels = F, barfill = "grey", barcolor = "grey") +
+  theme_B3() + labs(title = " ")
+d1 <- fviz_screeplot(pca_p2, addlabels = F, barfill = "grey", barcolor = "grey") +
+  theme_B3() + labs(title = " ")
+f1 <- fviz_screeplot(pca_g2, addlabels = F, barfill = "grey", barcolor = "grey") +
+  theme_B3() + labs(title = " ")
+
+b2 <- candidateboxplot("hypothalamus", c("COX1"), sexlevels)
+d2 <- candidateboxplot("pituitary", c("PRL"), sexlevels)
+f2 <- candidateboxplot("gonad", c("OVALY"), sexlevels)
+
+supplfig <- plot_grid(a,b1,b2, c,d1,d2,e,f1,f2, nrow = 3, rel_widths = c(2,1,2))
+```
+
 ## Save files
@@
     units = 'in', res = 300)
-plot(fig1) # Make plot
+plot(fig1) 
+dev.off()
+
+pdf(file="../figures/supplfig-1.pdf", width=7.5, height=7.5)
+plot(supplfig)
+dev.off()
+
+png("../figures/supplfig-1.png", width = 7.5, height = 7.5, 
+    units = 'in', res = 300)
+plot(supplfig)  
 dev.off()

### 2020-08-07 analysis/figure-1.Rmd
new suppl fig 1

@@
 library(Rtsne)
+library(factoextra)
+library(ggsignif)
 
@@
 
-```{r}
+```{r pca}
 # pca relies on count data from limma
@@
 
-```{r supplfig, fig.height=7.25, fig.width=7.25}
+```{r supplfig-1, fig.height=7.25, fig.width=7.25}
 
@@
 
-pdf(file="../figures/supplfig-1.pdf", width=7.5, height=7.5)
+pdf(file="../figures/supplfig-1-1.pdf", width=7.5, height=7.5)
 plot(supplfig)
@@
 
-png("../figures/supplfig-1.png", width = 7.5, height = 7.5, 
+png("../figures/supplfig-1-1.png", width = 7.5, height = 7.5, 
     units = 'in', res = 300)

### 2020-08-07 analysis/figure-3.Rmd
working on new fig 3 with all manip files, fig 4 is blank

--- 2020-05-25 analysis/figure-3.Rmd
@@
 ---
-title: "Fig3"
+title: "Replacement only"
 output: md_document
@@
 
-# Figure 3: All things prolactin
-
-```{r setup}
+```{r setup, include=FALSE}
 library(tidyverse)
 library(cowplot)
-library(ggimage)
-library(apaTables)
-library(factoextra)
+library(Rtsne)
+library(ggsignif)
 library(corrr)
-library(ggpubr)
-library(ggrepel)
-library(ggsignif)
-library(forcats)
 
@@
 
-knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE, message = F, warning = F,
+                      fig.path = '../figures/', fig.width = 7.25, fig.height = 7.25)
 ```
+
+# Figure 3 and 4. What are the effects of removal  and replacement?
 
@@
 
-colData <- read_csv("../metadata/04_colData.csv") %>%
-  filter(treatment %in% charlevels) %>%
-  mutate(tissue = factor(tissue, levels = tissuelevels),
-         treatment = factor(treatment, levels = charlevels),
-         temprowname = V1) %>% 
-  column_to_rownames(var = "temprowname") 
+colData <- read_csv("../metadata/00_colData.csv") %>%
+  mutate(treatment = factor(treatment, levels = alllevels),
+         tissue = factor(tissue, levels = tissuelevels)) %>% 
+  column_to_rownames(var = "X1") 
 head(colData)
@@
 
-## PCA data
+## tsne
 
-```{r PCA}
-# pca relies on count data from limma
-pca1f <- subsetmakepca("pituitary", charlevels, "female")	
-pca2f <- makefvizdf("pituitary", charlevels, "female")
+```{r tSNE}
 
-pca1m <- subsetmakepca("pituitary", charlevels, "male")	
-pca2m <- makefvizdf("pituitary", charlevels, "male")
+# removal
+hyptsne <- subsetmaketsne("hypothalamus", removallevels, sexlevels)
+pittsne <- subsetmaketsne("pituitary", removallevels, sexlevels)
+gontsne <- subsetmaketsne("gonads", removallevels, sexlevels)
 
-pca1 <- subsetmakepca("pituitary", charlevels, sexlevels)	
-pca2 <- makefvizdf("pituitary", charlevels, sexlevels)	
-```
+pittsne <- addgroupings(pittsne)
 
-## candidate genes, hypotheses genes, and data-driven genes
+t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "pituitary ", subtitle = "parental stage and manipulations") +
+  facet_wrap(~sex) 
+t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(title = "  ", subtitle = "external enviornment")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal clock")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
+t456
 
-```{r}
-# all genes lists in "./R/genelists.R"
+# replacement
+hyptsne <- subsetmaketsne("hypothalamus", timinglevels, sexlevels)
+pittsne <- subsetmaketsne("pituitary", timinglevels, sexlevels)
+gontsne <- subsetmaketsne("gonads", timinglevels, sexlevels)
 
-# candidate counts for boxplots
-candidatevsd <- read_csv("../results/06_candidatevsd.csv") %>%
-  filter(treatment %in% charlevels) %>%
-  mutate(treatment = factor(treatment)) %>%
-  mutate(tissue = factor(tissue, levels = tissuelevel),
-        treatment = factor(treatment, levels = charlevels)) %>%
-  drop_na()
-head(candidatevsd)
+pittsne <- addgroupings(pittsne)
 
-## prlpit
-PRLpit <- candidatevsd %>% 
-  filter(gene == "PRL", tissue == "pituitary")
-
-# list of pvalue and DEGs
-allDEG <- read_csv("../results/06_allDEG.csv")  %>%
-  mutate(direction = factor(direction, levels = hypothesislevels))
-
-# DEGs infor for PRL and external
-hilodf <- allDEG %>%
-  filter(comparison == "lo_hi")  %>%
-  arrange(padj) 
-hilogenes <- hilodf %>%  top_n(20) %>% pull(gene)
-
-eggchickdf <- allDEG %>%
-  filter(!comparison %in% c("lo_hi", "nest_eggs", "nest_chicks"))  %>%
-  arrange(padj)  
-eggchickgenes <- eggchickdf %>%  top_n(20) %>% pull(gene)
-
-## data-driven PRL genes
-pithilo <- makecorrdf("female", "pituitary", hilogenes)
-pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[32:52])
-
-## https://www.gynecologiconcology-online.net/article/S0090-8258(19)30069-1/fulltext
-cancergenes <- makecorrdf("female", "pituitary", c(shaidgenes,suszynskaagenes, "PRL"))
-
-## for scatter correlation (not currently usesed)
-candidatevsdwide <- candidatevsd  %>%
-    pivot_wider(names_from = gene, values_from = counts) 
-pitwide <- subsetcandidatevsdwide(sexlevels, "pituitary")
+t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "pituitary ", subtitle = "parental stage and manipulations") +
+  facet_wrap(~sex) 
+t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(title = "  ", subtitle = "external enviornment")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal clock")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
+t456
 
@@
 
-```{r fig3, fig.width=7.25, fig.height=7.25}
+## variance stabilized gene expression  (vsd)
 
-a1 <- plotpc12(pca1, pca2, pca1$treatment, allcolors, "Pituitary gene expression", NULL)   
-a2 <- plotprolactin(PRLpit, PRLpit$counts, "PRL", NULL) + 
-  theme(axis.text.x = element_text(angle = 25))  +
-  facet_wrap(~sex)
-a <- plot_grid(a1,a2, labels = c("A", "B"), label_size = 8, rel_widths = c(1,1))
+```{r vsd}
+vsd_path <- "../results/DEseq2/treatment/"   # path to the data
+vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
+vsd_pathfiles <- paste0(vsd_path, vsd_files)
+vsd_files
 
-b1 <- png::readPNG("../figures/images/fig_fig3b.png")
-b1 <- ggdraw() +  draw_image(b1, scale = 1)
-b2 <- makenewbargraph("pituitary", "female","eggs_chicks", 0, 4000) + labs(subtitle = "female")  + labs(title = " ") +
-  scale_x_discrete(labels = "eggs vs. chicks")
-b3 <- makenewbargraph("pituitary", "male", "eggs_chicks", 0, 4000)  + labs(subtitle = "male")  + labs(title = " ") +
-  theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), 
-        axis.text.y = element_blank(), axis.line.y = element_blank()) +
-  scale_x_discrete(labels = "eggs vs. chicks")
-b4 <- makenewbargraph("pituitary", "male", "lo_hi", 0, 4000)  +
-  scale_x_discrete(labels = "lo vs. hi PRL")
-b5 <- makenewbargraph("pituitary", "female", "lo_hi", 0, 4000)   + 
-  theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), 
-        axis.text.y = element_blank(), axis.line.y = element_blank()) +
-  scale_x_discrete(labels = "lo vs. hi PRL")
-b25 <- plot_grid(b2,b3,b4,b5, nrow = 2, rel_widths = c(1.2,1), rel_heights = c(1.2,1))
+allvsd <- vsd_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name")  %>% 
+  dplyr::rename("gene" = "X1") 
+## before pivoting, check names of df with
+## head(names(allvsd)) and tail(names(allvsd))
+allvsd <- allvsd %>% 
+  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "samples", values_to = "counts") 
+allvsd <- as_tibble(allvsd)
+head(allvsd[2:4])
+tail(allvsd[2:4])
 
-b6 <-  plottopDEGs(eggchickdf, "female",  "LFC eggs vs. chicks", "Top 10 female DEGs") + labs(title = " ")
-b7 <-  plottopDEGs(eggchickdf, "male",  "LFC eggs vs. chicks", "Top 10 male DEGs") + labs(title = " ")
-b8 <-  plottopDEGs(hilodf, "female",  "LFC lo vs. hi PRL", NULL)
-b9 <-  plottopDEGs(hilodf, "male",  "LFC lo vs. hi PRL", NULL)
-b15 <- plot_grid(b1,b25,nrow = 1, rel_widths = c(1.2,1))
-b69 <- plot_grid(b6,b7,b8,b9, align = "v" , rel_heights = c(1.2,1))
-b <- plot_grid(b15,b69, labels = c("C"), label_size = 8, rel_widths = c(1.2,1))
-
-c1 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 DEGs, low vs. high PRL") 
-c2 <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes in WGCNA module with PRL") 
-c3 <- plotcorrplot(cancergenes, NULL) + theme(legend.position = "right") + labs(subtitle = "Breast, ovary, & piuitary cancer genes") 
-c <- plot_grid(c1,c2,c3, nrow = 1, labels = c("D", "E", "F"), label_size = 8, rel_widths = c(1,1,1.3))
-
-fig3 <- plot_grid(a, b, c, nrow = 3, rel_heights = c(1.2,2,1.8))
-fig3
 ```
 
-```{r savefiles}
-write.csv(PRLpit,"../results/PRLvsd.csv", row.names = F)
+```{r candidatevsd}
 
-pdf(file="../figures/fig3-1.pdf", width=7.25, height=7.25)
-plot(fig3)
-dev.off()
+# for plotting gene expression over time
+getcandidatevsdmanip <- function(whichgenes, whichtissue){
+
+  candidates  <- allvsd %>%
+    filter(gene %in% whichgenes) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1),
+                  sextissue = sapply(strsplit(sextissue, '../results/DEseq2/treatment/'), "[", 2),
+                  sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                treatment = sapply(strsplit(samples, '\\_'), "[", 4),
+                treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1),
+                treatment = recode(treatment, "m.hatch" = "m.n2", "extend.hatch" = "m.n2",
+                                                   "inc.prolong" = "prolong", "m.inc.d8" = "early"),
+                treatment =  factor(treatment, levels = alllevels),
+                earlylate = fct_collapse(treatment, 
+                                  early = c("early", "m.inc.d3", "m.inc.d9", "inc.d3", 
+                                         "inc.d9",  "bldg"),
+                                  late = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" , 
+                                         "m.n2", "extend", "n5"),
+                                  reference = "control"),
+                extint = fct_collapse(treatment, 
+                                  nest = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg"),
+                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong"),
+                                  chicks = c("early", "hatch", "extend", "n5"),
+                                  reference = "control"))  %>%
+    dplyr::select(gene, counts, sex, tissue, treatment, earlylate, extint, samples)  
+  return(candidates)
+}
+
+candidatevsd <- getcandidatevsdmanip(candidategenes)
+
+hypvsd <- candidatevsd %>% filter(tissue %in% "hypothalamus")  
+pitvsd <- candidatevsd %>% filter(tissue %in% "pituitary") 
+gonvsd <- candidatevsd %>% filter(tissue %in% "gonad")
+
+hypvsdf <- candidatevsd %>% filter(tissue %in% "hypothalamus", sex == "female")  
+pitvsdf <- candidatevsd %>% filter(tissue %in% "pituitary", sex == "female") 
+gonvsdf <- candidatevsd %>% filter(tissue %in% "gonad", sex == "female")
+
+hypvsdm <- candidatevsd %>% filter(tissue %in% "hypothalamus", sex == "male")  
+pitvsdm <- candidatevsd %>% filter(tissue %in% "pituitary", sex == "male") 
+gonvsdm <- candidatevsd %>% filter(tissue %in% "gonad", sex == "male")
+
+# not all genes appear to be expressed in the gonads... these are
+filter(gonvsd, treatment == "control") %>% head()
 ```
 
-```{r forvictoria, eval = F}
-## PRL and PRLR in three tissues
+## candidate gene box plots
 
-candidatevsd <- candidatevsd %>%
-  filter(treatment %in%  c("early", "extend", charlevels))
+```{r}
+plotcandidatemanp <- function(df, whichgene, whichtissue){
 
-p1 <- candidateboxplot("hypothalamus", c("PRL"), sexlevels) + labs(subtitle = "hypothalamus", title = "PRL")
-p2 <- candidateboxplot("pituitary", c("PRL"), sexlevels) + labs(subtitle = "pituitary")
-p3 <- candidateboxplot("gonad", c("PRL"), sexlevels) + labs(subtitle = "gonads") + theme(axis.text.x = element_text(angle = 45))
+  p <- df %>%
+    dplyr::filter(gene  == whichgene)  %>%
+    ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
+    geom_boxplot(outlier.shape = NA, lwd=0.5) +
+    facet_wrap(~sex) +
+    geom_jitter(size = 0.25, aes(color = sex)) +
+    theme_B3() +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          axis.text.x = element_text(angle = 45, hjust = 1),
+          plot.subtitle = element_text(face = "italic")) +
+    scale_color_manual(values = allcolors) +
+    scale_fill_manual(values = allcolors)  +
+    labs(title = paste(whichtissue, sep = " "), subtitle = whichgene, y = "gene expression") +
+    geom_signif(comparisons = list(c("inc.d9", "early"), 
+                                  c( "inc.d17", "prolong"), 
+                                  c( "prolong", "hatch")),
+                map_signif_level=TRUE,
+                textsize = 2, family = 'Helvetica',
+                vjust = 0, size = 0.5, step_increase = 0.075) +
+  geom_signif(comparisons = list(c( "inc.d3", "m.inc.d3"),
+                                   c( "inc.d9", "m.inc.d9"),
+                                   c( "inc.d17", "m.inc.d17"),
+                                   c( "hatch", "m.n2")),
+                map_signif_level=TRUE,
+                textsize = 2, family = 'Helvetica',
+                vjust = 0,  size = 0.5, step_increase = 0.075) +
+    geom_signif(comparisons = list(c( "hatch", "extend"), 
+                                  c( "n5", "extend")),
+                map_signif_level=TRUE,
+                textsize = 2, family = 'Helvetica',
+                vjust = 0, size = 0.5, step_increase = 0.075)
 
-p4 <- candidateboxplot("hypothalamus", c("PRLR"), sexlevels) + labs(subtitle = " ", title = " ") 
-p5 <- candidateboxplot("pituitary", c("PRLR"), sexlevels) + labs(subtitle = " ")
-p6 <- candidateboxplot("gonad", c("PRLR"), sexlevels) + theme(axis.text.x = element_text(angle = 45))+ labs(subtitle = " ")
+  return(p)
 
-plot_grid(p1,p4,p2,p5,p3,p6, ncol = 2, rel_heights = c(1.2,1,1.2))
+}
+
+a <- plotcandidatemanp(hypvsdf, "HTR2C", "hypothalamus") 
+c <- plotcandidatemanp(pitvsdf, "PRL", "pituitary")
+e <- plotcandidatemanp(gonvsdf, "ESR2", "gonads") 
+b <- plotcandidatemanp(hypvsdm, "HTR2C", "hypothalamus") 
+d <- plotcandidatemanp(pitvsdm, "PRL", "pituitary")
+f <- plotcandidatemanp(gonvsdm, "ESR2", "gonads") 
+boxplots3 <- plot_grid(a,b,c,d,e,f, nrow = 3, 
+                       labels = c("A", "", "B", " ", "C"),
+                       hjust = 0, label_size = 8)
+boxplots3
+
 ```
+
+## DEGs
+
+```{r allDEGs}
+DEG_path <- "../results/DEseq2/treatment/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+  dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) # %>%
+ # mutate(tissue = factor(tissue, levels = tissuelevel),
+  #       comparison = factor(comparison , levels = comparisonlevelsremoval),
+  #       direction = factor(direction, levels = alllevels))
+head(allDEG)
+
+candidateDEGS <- allDEG %>%
+  filter(gene %in% parentalcaregenes) %>%
+  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
+         sex = recode(sex, "female" = "F", "male" = "M" ),
+         tissue = recode(tissue, 
+                         "hypothalamus" = "H",
+                         "pituitary" = "P", 
+                         "gonad" = "G", "gonads" = "G")) %>%
+  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
+  select(gene, res, comparison)  %>%
+  group_by(gene,  comparison) %>%
+  summarize(res = str_c(res, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = res) %>%
+  select(gene, inc.d3_m.inc.d3, inc.d9_m.inc.d9, inc.d17_m.inc.d17, hatch_m.n2,
+         inc.d17_prolong, hatch_prolong,extend_hatch, n5_extend, 
+         early_inc.d9 ,hatch_early)
+head(candidateDEGS[1:5])
+```
+
+## make figure
+
+```{r fig4, fig.width=7.25, fig.height=7.25}
+
+design <- png::readPNG("../figures/images/fig_fig4a.png")
+design <- ggdraw() +  draw_image(design, scale = 1)
+
+boxplots <- plot_grid(boxplots1, boxplots2, nrow = 1, rel_widths = c(1.5,1),
+                      labels = "AUTO", label_size = 8)
+fig4 <- plot_grid(boxplots, design, t456, ncol = 1, rel_heights = c(2.75,1.3,1),
+                      labels = c(" ", " " , "C"), label_size = 8)
+fig4
+
+```
+
+## suppl fig 1
+
+```{r supplfig-4}
+
+comparisonlevelsmanip <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2",
+         "early_inc.d9" ,"hatch_early", "inc.d17_prolong", "hatch_prolong","extend_hatch", "n5_extend")
+
+# for fig 1
+DEGmanip <- allDEG %>% 
+  filter(comparison %in% comparisonlevelsmanip) %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelsmanip))
+
+DEGmanip %>% 
+  group_by(comparison, sex, tissue) %>%
+  summarize(totalDEGs = n())  %>%
+  arrange(desc(totalDEGs))
+
+s1a <- makebargraph(DEGmanip, "hypothalamus","DEGs w/ + LFC", 0, 3200, DEGmanip$comparison) + 
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank()) + 
+  labs(subtitle = "hypothalamus") + geom_vline(xintercept = 4.5, linetype = "dashed", color = "grey")
+
+s1b <- makebargraph(DEGmanip, "pituitary","DEGs w/ + LFC", 0, 3200, DEGmanip$comparison) + 
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), strip.text = element_blank()) + 
+  labs(subtitle = "pituitary")  + geom_vline(xintercept = 4.5, linetype = "dashed", color = "grey")
+
+s1c <- makebargraph(DEGmanip, "gonad","DEGs w/ + LFC", 0, 3200, DEGmanip$comparison) + 
+  theme(strip.text = element_blank()) + 
+  labs(subtitle = "gonad", x = "Comparison of parental stage and manipulation groups") + geom_vline(xintercept = 4.5, linetype = "dashed", color = "grey") + 
+  scale_x_discrete(breaks= comparisonlevelsmanip,
+                      labels= comparisonlevelsmanip,
+                   drop=FALSE)
+
+contrasts <- png::readPNG("../figures/images/fig_fig4sup1.png")
+contrasts <- ggdraw() +  draw_image(contrasts, scale = 1)
+
+supplfig4 <- plot_grid(contrasts, s1a, s1b, s1c, ncol = 1, rel_heights = c(0.5, 1,1,1.5))
+supplfig4
+
+```
+
+## write files
+
+```{r write}
+
+pdf(file="../figures/fig4-1.pdf", width=7.25, height=7.25)
+plot(fig4)
+dev.off()
+
+pdf(file="../figures/supplfig-4-1.pdf", width=5, height=5)
+plot(supplfig4)
+dev.off()
+
+write.csv(candidateDEGS, "../results/table2.csv", row.names = F)
+```

### 2020-08-07 analysis/boneyard.Rmd
rm facet_wrap

@@
 ```
+
+## tsne
+
+```{r tSNE}
+
+# removal
+hyptsne <- subsetmaketsne("hypothalamus", removallevels, sexlevels)
+pittsne <- subsetmaketsne("pituitary", removallevels, sexlevels)
+gontsne <- subsetmaketsne("gonads", removallevels, sexlevels)
+
+pittsne <- addgroupings(pittsne)
+
+t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "pituitary ", subtitle = "parental stage and manipulations") +
+  facet_wrap(~sex) 
+t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(title = "  ", subtitle = "external enviornment")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal clock")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
+t456
+
+# replacement
+hyptsne <- subsetmaketsne("hypothalamus", timinglevels, sexlevels)
+pittsne <- subsetmaketsne("pituitary", timinglevels, sexlevels)
+gontsne <- subsetmaketsne("gonads", timinglevels, sexlevels)
+
+pittsne <- addgroupings(pittsne)
+
+t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "pituitary ", subtitle = "parental stage and manipulations") +
+  facet_wrap(~sex) 
+t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(title = "  ", subtitle = "external enviornment")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal clock")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
+t456
+
+```
+
+## DEGs
+
+```{r allDEGs}
+DEG_path <- "../results/DEseq2/treatment/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+  dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) # %>%
+ # mutate(tissue = factor(tissue, levels = tissuelevel),
+  #       comparison = factor(comparison , levels = comparisonlevelsremoval),
+  #       direction = factor(direction, levels = alllevels))
+head(allDEG)
+
+candidateDEGS <- allDEG %>%
+  filter(gene %in% parentalcaregenes) %>%
+  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
+         sex = recode(sex, "female" = "F", "male" = "M" ),
+         tissue = recode(tissue, 
+                         "hypothalamus" = "H",
+                         "pituitary" = "P", 
+                         "gonad" = "G", "gonads" = "G")) %>%
+  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
+  select(gene, res, comparison)  %>%
+  group_by(gene,  comparison) %>%
+  summarize(res = str_c(res, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = res) %>%
+  select(gene, inc.d3_m.inc.d3, inc.d9_m.inc.d9, inc.d17_m.inc.d17, hatch_m.n2,
+         inc.d17_prolong, hatch_prolong,extend_hatch, n5_extend, 
+         early_inc.d9 ,hatch_early)
+head(candidateDEGS[1:5])
+```
+
+## suppl fig 1
+
+```{r supplfig-4}
+
+comparisonlevelsmanip <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2",
+         "early_inc.d9" ,"hatch_early", "inc.d17_prolong", "hatch_prolong","extend_hatch", "n5_extend")
+
+# for fig 1
+DEGmanip <- allDEG %>% 
+  filter(comparison %in% comparisonlevelsmanip) %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelsmanip))
+
+DEGmanip %>% 
+  group_by(comparison, sex, tissue) %>%
+  summarize(totalDEGs = n())  %>%
+  arrange(desc(totalDEGs))
+
+s1a <- makebargraph(DEGmanip, "hypothalamus","DEGs w/ + LFC", 0, 3200, DEGmanip$comparison) + 
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank()) + 
+  labs(subtitle = "hypothalamus") + geom_vline(xintercept = 4.5, linetype = "dashed", color = "grey")
+
+s1b <- makebargraph(DEGmanip, "pituitary","DEGs w/ + LFC", 0, 3200, DEGmanip$comparison) + 
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), strip.text = element_blank()) + 
+  labs(subtitle = "pituitary")  + geom_vline(xintercept = 4.5, linetype = "dashed", color = "grey")
+
+s1c <- makebargraph(DEGmanip, "gonad","DEGs w/ + LFC", 0, 3200, DEGmanip$comparison) + 
+  theme(strip.text = element_blank()) + 
+  labs(subtitle = "gonad", x = "Comparison of parental stage and manipulation groups") + geom_vline(xintercept = 4.5, linetype = "dashed", color = "grey") + 
+  scale_x_discrete(breaks= comparisonlevelsmanip,
+                      labels= comparisonlevelsmanip,
+                   drop=FALSE)
+
+contrasts <- png::readPNG("../figures/images/fig_fig4sup1.png")
+contrasts <- ggdraw() +  draw_image(contrasts, scale = 1)
+
+supplfig3 <- plot_grid(contrasts, s1a, s1b, s1c, ncol = 1, rel_heights = c(0.5, 1,1,1.5))
+supplfig3
+
+pdf(file="../figures/supplfig-3-1.pdf", width=5, height=5)
+plot(supplfig3)
+dev.off()
+
+write.csv(candidateDEGS, "../results/table2.csv", row.names = F)
+```

### 2020-08-07 analysis/figure-3.Rmd
rm facet_wrap

--- 2020-05-26 analysis/figure-3.Rmd
@@
 ---
-title: "Fig3"
+title: "Removal and Replacement only"
 output: md_document
@@
 
-## PCA data
+## variance stabilized gene expression  (vsd)
 
-```{r PCA}
-# pca relies on count data from limma
-pca1f <- subsetmakepca("pituitary", charlevels, "female")	
-pca2f <- makefvizdf("pituitary", charlevels, "female")
+```{r vsd}
+vsd_path <- "../results/DEseq2/treatment/"   # path to the data
+vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
+vsd_pathfiles <- paste0(vsd_path, vsd_files)
+vsd_files
 
-pca1m <- subsetmakepca("pituitary", charlevels, "male")	
-pca2m <- makefvizdf("pituitary", charlevels, "male")
+allvsd <- vsd_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name")  %>% 
+  dplyr::rename("gene" = "X1") 
+## before pivoting, check names of df with
+## head(names(allvsd)) and tail(names(allvsd))
+allvsd <- allvsd %>% 
+  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "samples", values_to = "counts") 
+allvsd <- as_tibble(allvsd)
+head(allvsd[2:4])
+tail(allvsd[2:4])
 
-pca1 <- subsetmakepca("pituitary", charlevels, sexlevels)	
-pca2 <- makefvizdf("pituitary", charlevels, sexlevels)	
 ```
 
-## candidate genes, hypotheses genes, and data-driven genes
+```{r candidatevsd}
 
-```{r}
-# all genes lists in "./R/genelists.R"
+# for plotting gene expression over time
+getcandidatevsdmanip <- function(whichgenes, whichtissue){
 
-# candidate counts for boxplots
-candidatevsd <- read_csv("../results/06_candidatevsd.csv") %>%
-  filter(treatment %in% charlevels) %>%
-  mutate(treatment = factor(treatment)) %>%
-  mutate(tissue = factor(tissue, levels = tissuelevel),
-        treatment = factor(treatment, levels = charlevels)) %>%
-  drop_na()
-head(candidatevsd)
+  candidates  <- allvsd %>%
+    filter(gene %in% whichgenes) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1),
+                  sextissue = sapply(strsplit(sextissue, '../results/DEseq2/treatment/'), "[", 2),
+                  sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                treatment = sapply(strsplit(samples, '\\_'), "[", 4),
+                treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1),
+                treatment = recode(treatment, "m.hatch" = "m.n2", "extend.hatch" = "m.n2",
+                                                   "inc.prolong" = "prolong", "m.inc.d8" = "early"),
+                treatment =  factor(treatment, levels = alllevels),
+                earlylate = fct_collapse(treatment, 
+                                  early = c("early", "m.inc.d3", "m.inc.d9",
+                                           "lay", "inc.d3", "inc.d9" ),
 
-## prlpit
-PRLpit <- candidatevsd %>% 
-  filter(gene == "PRL", tissue == "pituitary")
+                                  late = c("inc.d17", "m.inc.d17", "prolong" ,  "hatch" , 
+                                         "m.n2", "extend", "n5", "n9"),
+                                  reference = c("control", "bldg")),
+                extint = fct_collapse(treatment, 
+                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17"),
+                                  eggs = c("lay","inc.d3", "inc.d9", "inc.d17", "prolong"),
+                                  chicks = c("early", "hatch", "extend", "n5", "n9"),
+                                  reference = c("control", "bldg")))  %>%
+    dplyr::select(gene, counts, sex, tissue, treatment, earlylate, extint, samples)  
+  return(candidates)
+}
 
-# list of pvalue and DEGs
-allDEG <- read_csv("../results/06_allDEG.csv")  %>%
-  mutate(direction = factor(direction, levels = hypothesislevels))
+candidatevsd <- getcandidatevsdmanip(candidategenes)
 
-# DEGs infor for PRL and external
-hilodf <- allDEG %>%
-  filter(comparison == "lo_hi")  %>%
-  arrange(padj) 
-hilogenes <- hilodf %>%  top_n(20) %>% pull(gene)
+hypvsd <- candidatevsd %>% filter(tissue %in% "hypothalamus")  
+pitvsd <- candidatevsd %>% filter(tissue %in% "pituitary") 
+gonvsd <- candidatevsd %>% filter(tissue %in% "gonad")
 
-eggchickdf <- allDEG %>%
-  filter(!comparison %in% c("lo_hi", "nest_eggs", "nest_chicks"))  %>%
-  arrange(padj)  
-eggchickgenes <- eggchickdf %>%  top_n(20) %>% pull(gene)
+hypvsdf <- candidatevsd %>% filter(tissue %in% "hypothalamus", sex == "female")  %>% drop_na()
+pitvsdf <- candidatevsd %>% filter(tissue %in% "pituitary", sex == "female") %>% drop_na() 
+gonvsdf <- candidatevsd %>% filter(tissue %in% "gonad", sex == "female")  %>% drop_na()
 
-## data-driven PRL genes
-pithilo <- makecorrdf("female", "pituitary", hilogenes)
-pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[32:52])
+hypvsdm <- candidatevsd %>% filter(tissue %in% "hypothalamus", sex == "male")  %>% drop_na()
+pitvsdm <- candidatevsd %>% filter(tissue %in% "pituitary", sex == "male") %>% drop_na()
+gonvsdm <- candidatevsd %>% filter(tissue %in% "gonad", sex == "male") %>% drop_na()
 
-## https://www.gynecologiconcology-online.net/article/S0090-8258(19)30069-1/fulltext
-cancergenes <- makecorrdf("female", "pituitary", c(shaidgenes,suszynskaagenes, "PRL"))
-
-## for scatter correlation (not currently usesed)
-candidatevsdwide <- candidatevsd  %>%
-    pivot_wider(names_from = gene, values_from = counts) 
-pitwide <- subsetcandidatevsdwide(sexlevels, "pituitary")
-
-## number of genes total vs plotted
-
-eggchickdf %>% filter(tissue == "pituitary") %>% nrow()
-hilodf %>% filter(tissue == "pituitary") %>% nrow()
-length(WGCNAgenes)
+# not all genes appear to be expressed in all tissues. added drop_na()
+filter(gonvsd, treatment == "control") %>% head()
 ```
 
-```{r fig3, fig.width=7.25, fig.height=7.25}
+## candidate gene box plots
 
-a1 <- plotpc12(pca1, pca2, pca1$treatment, allcolors, "Pituitary gene expression", NULL)   
-a2 <- plotprolactin(PRLpit, PRLpit$counts, "PRL", NULL) + 
-  theme(axis.text.x = element_text(angle = 25))  +
-  facet_wrap(~sex)
-a <- plot_grid(a1,a2, labels = c("A", "B"), label_size = 8, rel_widths = c(1,1))
+```{r fig3, fig.height=7.25, fig.width=7.25}
+plotcandidatemanp <- function(df, whichgene, whichtissue){
 
-b1 <- png::readPNG("../figures/images/fig_fig3b.png")
-b1 <- ggdraw() +  draw_image(b1, scale = 1)
-b2 <- makenewbargraph("pituitary", "female","eggs_chicks", 0, 4000) + labs(subtitle = "female")  + labs(title = " ") +
-  scale_x_discrete(labels = "eggs vs. chicks")
-b3 <- makenewbargraph("pituitary", "male", "eggs_chicks", 0, 4000)  + labs(subtitle = "male")  + labs(title = " ") +
-  theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), 
-        axis.text.y = element_blank(), axis.line.y = element_blank()) +
-  scale_x_discrete(labels = "eggs vs. chicks")
-b4 <- makenewbargraph("pituitary", "male", "lo_hi", 0, 4000)  +
-  scale_x_discrete(labels = "lo vs. hi PRL")
-b5 <- makenewbargraph("pituitary", "female", "lo_hi", 0, 4000)   + 
-  theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), 
-        axis.text.y = element_blank(), axis.line.y = element_blank()) +
-  scale_x_discrete(labels = "lo vs. hi PRL")
-b25 <- plot_grid(b2,b3,b4,b5, nrow = 2, rel_widths = c(1.2,1), rel_heights = c(1.2,1))
+  p <- df %>%
+    dplyr::filter(gene  == whichgene)  %>%
+    ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
+    geom_boxplot(outlier.shape = NA, lwd=0.5) +
+    #facet_wrap(~sex) +
+    geom_jitter(size = 0.25, aes(color = sex)) +
+    theme_B3() +
+    theme(legend.position = "none",
 
-b6 <-  plottopDEGs(eggchickdf, "female",  "LFC eggs vs. chicks", "Top 10 female DEGs") + labs(title = " ")
-b7 <-  plottopDEGs(eggchickdf, "male",  "LFC eggs vs. chicks", "Top 10 male DEGs") + labs(title = " ")
-b8 <-  plottopDEGs(hilodf, "female",  "LFC lo vs. hi PRL", NULL)
-b9 <-  plottopDEGs(hilodf, "male",  "LFC lo vs. hi PRL", NULL)
-b15 <- plot_grid(b1,b25,nrow = 1, rel_widths = c(1.2,1))
-b69 <- plot_grid(b6,b7,b8,b9, align = "v" , rel_heights = c(1.2,1))
-b <- plot_grid(b15,b69, labels = c("C"), label_size = 8, rel_widths = c(1.2,1))
+          axis.text.x = element_text(angle = 45, hjust = 1),
+          plot.subtitle = element_text(face = "italic")) +
+    scale_color_manual(values = allcolors) +
+    scale_fill_manual(values = allcolors)  +
+    labs(title = whichtissue, 
+         subtitle = whichgene, 
+         y = "gene expression", x = "Semi-sequential stages") +
+    geom_signif(comparisons = list(c("inc.d9", "early"), 
+                                  c( "inc.d17", "prolong"), 
+                                  c( "prolong", "hatch")),
+                map_signif_level=TRUE,
+                textsize = 2, family = 'Helvetica',
+                vjust = 0, size = 0.5, step_increase = 0.075) +
+  geom_signif(comparisons = list(c( "inc.d3", "m.inc.d3"),
+                                   c( "inc.d9", "m.inc.d9"),
+                                   c( "inc.d17", "m.inc.d17"),
+                                   c( "hatch", "m.n2")),
+                map_signif_level=TRUE,
+                textsize = 2, family = 'Helvetica',
+                vjust = 0,  size = 0.5, step_increase = 0.075) +
+    geom_signif(comparisons = list(c( "hatch", "extend"), 
+                                  c( "n5", "extend")),
+                map_signif_level=TRUE,
+                textsize = 2, family = 'Helvetica',
+                vjust = 0, size = 0.5, step_increase = 0.075)
 
-c1 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 DEGs, low vs. high PRL") 
-c2 <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes in WGCNA module with PRL") 
-c3 <- plotcorrplot(cancergenes, NULL) + theme(legend.position = "right") + labs(subtitle = "Breast, ovary, & piuitary cancer genes") 
-c <- plot_grid(c1,c2,c3, nrow = 1, labels = c("D"), label_size = 8, rel_widths = c(1,1,1.3))
+  return(p)
 
-fig3 <- plot_grid(a, b, c, nrow = 3, rel_heights = c(1.2,2,1.8))
+}
+
+boxplotextint <- function(df, whichgene, whichtissue){
+
+  p1 <- df %>%
+    dplyr::filter(gene  == whichgene)  %>%
+    mutate(earlylate = factor(earlylate, levels = levelhypothesis)) %>%
+    ggplot(aes(y =  counts, x = earlylate, fill = earlylate, color = sex)) +
+    geom_boxplot(outlier.shape = NA, lwd=0.5) +
+    #facet_wrap(~sex) +
+    geom_jitter(size = 0.25, aes(color = treatment)) +
+    theme_B3() +
+  scale_color_manual(values = allcolors) +
+    scale_fill_manual(values = allcolors)  +
+  geom_signif(comparisons = list(c( "reference", "early"), 
+                                  c( "early", "late"),
+                                 c( "reference", "late")),
+                map_signif_level=TRUE,
+                textsize = 2, family = 'Helvetica',
+                vjust = 0, size = 0.5, step_increase = 0.075) +
+  labs(title = paste(whichtissue, sep = " "), subtitle = " ", y = NULL) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1))
+
+p2 <- df %>%
+    dplyr::filter(gene  == whichgene)  %>%
+      mutate(extint = factor(extint, levels = levelhypothesis)) %>%
+    ggplot(aes(y =  counts, x = extint, fill = extint, color = sex)) +
+    geom_boxplot(outlier.shape = NA, lwd=0.5) +
+    #facet_wrap(~sex) +
+    geom_jitter(size = 0.25, aes(color = treatment)) +
+    theme_B3() +
+  scale_color_manual(values = allcolors) +
+    scale_fill_manual(values = allcolors)  +
+  geom_signif(comparisons = list(c( "reference", "eggs"), 
+                                  c( "eggs", "chicks"),
+                                 c( "chicks", "loss"),
+                                 c( "reference", "chicks"),
+                                 c( "eggs", "loss"),
+                                 c( "reference", "loss")),
+                map_signif_level=TRUE,
+                textsize = 2, family = 'Helvetica',
+                vjust = 0, size = 0.5, step_increase = 0.075) +
+  labs(title = " ", subtitle = " " , y = NULL) +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1))
+
+p <- plot_grid(p1,p2)
+return(p)
+}
+
+a <- plotcandidatemanp(hypvsdf, "HTR2C", "female hypothalamus") 
+b <- boxplotextint(hypvsdf, "HTR2C", " ") 
+c <- plotcandidatemanp(hypvsdm, "HTR2C", "male hypothalamus") 
+d <- boxplotextint(hypvsdm, "HTR2C", " ")
+
+e <- plotcandidatemanp(pitvsdf, "PRL", "female pituitary")
+f <- boxplotextint(pitvsdf, "PRL", " ")
+
+g <- plotcandidatemanp(pitvsdm, "PRL", "male pituitary")
+h <- boxplotextint(pitvsdm, "PRL", " ")
+
+i <- plotcandidatemanp(gonvsdf, "ESR2", "female gonads") 
+j <- boxplotextint(gonvsdf, "ESR2", " ")
+
+k <- plotcandidatemanp(gonvsdm, "ESR2", "male gonads") 
+l <- boxplotextint(gonvsdm, "ESR2", " ")
+
+fig3 <- plot_grid(a,b,c,d,e,f, 
+                       g,h,i,j,k,l, rel_widths = c(1.75,1,1.75,1),
+                       labels = c("A", "", "", "", 
+                                  "B", "", "", "", 
+                                  "C", "", "", ""),
+                       label_size = 8, hjust = 0)
 fig3
@@
 
-```{r savefiles}
-write.csv(PRLpit,"../results/PRLvsd.csv", row.names = F)
+## write files
+
+```{r write}
 
@@
 dev.off()
+
+write.csv(candidatevsd, "../results/fig3-suppltable.csv")
+
 ```
-
-```{r forvictoria, eval = F}
-## PRL and PRLR in three tissues
-
-candidatevsd <- candidatevsd %>%
-  filter(treatment %in%  c("early", "extend", charlevels))
-
-p1 <- candidateboxplot("hypothalamus", c("PRL"), sexlevels) + labs(subtitle = "hypothalamus", title = "PRL")
-p2 <- candidateboxplot("pituitary", c("PRL"), sexlevels) + labs(subtitle = "pituitary")
-p3 <- candidateboxplot("gonad", c("PRL"), sexlevels) + labs(subtitle = "gonads") + theme(axis.text.x = element_text(angle = 45))
-
-p4 <- candidateboxplot("hypothalamus", c("PRLR"), sexlevels) + labs(subtitle = " ", title = " ") 
-p5 <- candidateboxplot("pituitary", c("PRLR"), sexlevels) + labs(subtitle = " ")
-p6 <- candidateboxplot("gonad", c("PRLR"), sexlevels) + theme(axis.text.x = element_text(angle = 45))+ labs(subtitle = " ")
-
-plot_grid(p1,p4,p2,p5,p3,p6, ncol = 2, rel_heights = c(1.2,1,1.2))
-```

### 2020-08-07 R/themes.R

@@ -1,5 +1,8 @@
 # custom themes
 
+## hex colors chosen using https://colorbrewer2.org/ 
+## hex colors blended using from https://meyerweb.com/eric/tools/color-blend/#F8766D:D39200::hex 
+
 theme_B3 <- function () { 
   theme_classic(base_size = 7,
                 base_family = 'Helvetica') +
@@ -118,13 +121,16 @@ colorsnewgrouping <-  c("control" = "#F8766D",
                  "manip" = "#959595")
 
 colorhypothesis <- c("none" = "#cc5500",
-                     "nest" = "#8B4513",
+                     "loss" = "#8B4513",
                      "eggs" = "#7FA163",
                      "chicks" = "#3A80B9",
                      "lo" = "#B5C4B2",
                      "hi" = "#19757A",
                      "early" = "#B5C4B2",
-                     "late" = "#19757A")
+                     "late" = "#19757A",
+                     "reference" = "#F8766D" )
+
+levelhypothesis <- c("reference", "eggs", "chicks", "loss", "early", "late")
 
 colorstissue <- c("hypothalamus" = "#d95f02",
                   "pituitary" = "#1b9e77",

### 2020-08-07 analysis/boneyard.Rmd
now with suppl fig 3 and fig 3

--- 2020-08-07 analysis/boneyard.Rmd
@@
 ```
+
+## tsne
+
+```{r tSNE}
+
+# removal
+hyptsne <- subsetmaketsne("hypothalamus", removallevels, sexlevels)
+pittsne <- subsetmaketsne("pituitary", removallevels, sexlevels)
+gontsne <- subsetmaketsne("gonads", removallevels, sexlevels)
+
+pittsne <- addgroupings(pittsne)
+
+t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "pituitary ", subtitle = "parental stage and manipulations") +
+  facet_wrap(~sex) 
+t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(title = "  ", subtitle = "external enviornment")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal clock")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
+t456
+
+# replacement
+hyptsne <- subsetmaketsne("hypothalamus", timinglevels, sexlevels)
+pittsne <- subsetmaketsne("pituitary", timinglevels, sexlevels)
+gontsne <- subsetmaketsne("gonads", timinglevels, sexlevels)
+
+pittsne <- addgroupings(pittsne)
+
+t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "pituitary ", subtitle = "parental stage and manipulations") +
+  facet_wrap(~sex) 
+t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(title = "  ", subtitle = "external enviornment")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal clock")   +
+  facet_wrap(~sex) + theme(axis.title.y = element_blank())
+t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
+t456
+
+```

### 2020-08-07 analysis/figure-3.Rmd
now with suppl fig 3 and fig 3

@@
 
-```{r fig3, fig.height=7.25, fig.width=7.25}
+```{r }
 plotcandidatemanp <- function(df, whichgene, whichtissue){
@@
     theme(legend.position = "none",
-
           axis.text.x = element_text(angle = 45, hjust = 1),
@@
     scale_fill_manual(values = allcolors)  +
-    labs(title = whichtissue, 
-         subtitle = whichgene, 
+    labs( subtitle = whichgene, 
          y = "gene expression", x = "Semi-sequential stages") +
@@
                 vjust = 0, size = 0.5, step_increase = 0.075) +
-  labs(title = paste(whichtissue, sep = " "), subtitle = " ", y = NULL) +
+  labs(subtitle = " ", y = NULL) +
   theme(legend.position = "none",
@@
                 vjust = 0, size = 0.5, step_increase = 0.075) +
-  labs(title = " ", subtitle = " " , y = NULL) +
+  labs(subtitle = " " , y = NULL) +
   theme(legend.position = "none",
@@
 
-p <- plot_grid(p1,p2)
+p <- plot_grid(p1,p2, rel_widths = c(3.5,4))
 return(p)
@@
 
+```
+
+## DEGs
+
+```{r allDEGs}
+DEG_path <- "../results/DEseq2/treatment/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+  dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) # %>%
+ # mutate(tissue = factor(tissue, levels = tissuelevel),
+  #       comparison = factor(comparison , levels = comparisonlevelsremoval),
+  #       direction = factor(direction, levels = alllevels))
+head(allDEG)
+
+candidateDEGS <- allDEG %>%
+  filter(gene %in% parentalcaregenes) %>%
+  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
+         sex = recode(sex, "female" = "F", "male" = "M" ),
+         tissue = recode(tissue, 
+                         "hypothalamus" = "H",
+                         "pituitary" = "P", 
+                         "gonad" = "G", "gonads" = "G")) %>%
+  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
+  select(gene, res, comparison)  %>%
+  group_by(gene,  comparison) %>%
+  summarize(res = str_c(res, collapse = " ")) %>%
+  pivot_wider(names_from = comparison, values_from = res) %>%
+  select(gene, inc.d3_m.inc.d3, inc.d9_m.inc.d9, inc.d17_m.inc.d17, hatch_m.n2,
+         inc.d17_prolong, hatch_prolong,extend_hatch, n5_extend, 
+         early_inc.d9 ,hatch_early) %>%
+  arrange(gene)
+head(candidateDEGS[1:5])
+
+DEGmanip <- allDEG %>% 
+  filter(comparison %in% comparisonlevelsmanip) %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelsmanip))
+
+DEGmanipf <- DEGmanip %>% filter(sex == "female")
+DEGmanipm <- DEGmanip %>% filter(sex == "male")
+
+DEGmanip %>% 
+  group_by(comparison, sex, tissue) %>%
+  summarize(totalDEGs = n())  %>%
+  arrange(desc(totalDEGs))
+```
+
+## fig
+
+```{r fig3, fig.height=7.25, fig.width=7.25}
 a <- plotcandidatemanp(hypvsdf, "HTR2C", "female hypothalamus") 
 b <- boxplotextint(hypvsdf, "HTR2C", " ")
-c <- plotcandidatemanp(hypvsdm, "HTR2C", "male hypothalamus") 
-d <- boxplotextint(hypvsdm, "HTR2C", " ")
 
-e <- plotcandidatemanp(pitvsdf, "PRL", "female pituitary")
-f <- boxplotextint(pitvsdf, "PRL", " ")
+d <- plotcandidatemanp(pitvsdf, "PRL", "female pituitary")
+e <- boxplotextint(pitvsdf, "PRL", " ")
 
-g <- plotcandidatemanp(pitvsdm, "PRL", "male pituitary")
-h <- boxplotextint(pitvsdm, "PRL", " ")
+g <- plotcandidatemanp(gonvsdf, "ESR2", "female gonads") 
+h <- boxplotextint(gonvsdf, "ESR2", " ")
 
-i <- plotcandidatemanp(gonvsdf, "ESR2", "female gonads") 
-j <- boxplotextint(gonvsdf, "ESR2", " ")
+c <- makebargraph(DEGmanipf, "hypothalamus","DEGs w/ + LFC", 0, 3200, DEGmanipf$comparison) + 
+  labs(title = "female hypothalamus", x = "Comparison of manipulated group to parental controls") + 
+  theme(strip.text = element_blank()) + 
+  scale_x_discrete(breaks= comparisonlevelsmanip,
+                      labels= comparisonlabelssmanip,
+                   drop=FALSE)
 
-k <- plotcandidatemanp(gonvsdm, "ESR2", "male gonads") 
-l <- boxplotextint(gonvsdm, "ESR2", " ")
+f <- makebargraph(DEGmanipf, "pituitary","DEGs w/ + LFC", 0, 3200, DEGmanipf$comparison) + 
+  labs(title = "female pituitary",x = "Comparison of manipulated group to parental controls")  + 
+  theme(strip.text = element_blank()) + 
+  scale_x_discrete(breaks= comparisonlevelsmanip,
+                      labels= comparisonlabelssmanip,
+                   drop=FALSE)
 
-fig3 <- plot_grid(a,b,c,d,e,f, 
-                       g,h,i,j,k,l, rel_widths = c(1.75,1,1.75,1),
-                       labels = c("A", "", "", "", 
-                                  "B", "", "", "", 
-                                  "C", "", "", ""),
-                       label_size = 8, hjust = 0)
+i <- makebargraph(DEGmanipf, "gonad","DEGs w/ + LFC", 0, 3200, DEGmanipf$comparison) + 
+  theme(strip.text = element_blank()) + 
+  labs(title = "female hypothalamus",x = "Comparison of manipulated group to parental controls") + 
+  scale_x_discrete(breaks= comparisonlevelsmanip,
+                      labels= comparisonlabelssmanip,
+                   drop=FALSE)
+fig3 <- plot_grid(c, a,b,f,d,e,i,g,h, ncol = 3, rel_widths = c(2,1.8,1), 
+                  labels = c("A", "", "",
+                             "B", "", "",
+                             "C", "", ""), label_size = 8, hjust = 0)
 fig3
+
 ```
 
+## suppl fig
+
+```{r supplfig-3}
+
+a <- plotcandidatemanp(hypvsdm, "HTR2C", "male hypothalamus") 
+b <- boxplotextint(hypvsdm, "HTR2C", " ")
+
+d <- plotcandidatemanp(pitvsdm, "PRL", "male pituitary")
+e <- boxplotextint(pitvsdm, "PRL", " ")
+
+g <- plotcandidatemanp(gonvsdm, "ESR2", "male gonads") 
+h <- boxplotextint(gonvsdm, "ESR2", " ")
+
+c <- makebargraph(DEGmanipm, "hypothalamus","DEGs w/ + LFC", 0, 3200, DEGmanipm$comparison) + 
+  labs(subtitle = "male hypothalamus", x = "Comparison of manipulated group to parental controls") 
+  scale_x_discrete(breaks= comparisonlevelsmanip,
+                      labels= comparisonlevelsmanip,
+                   drop=FALSE)
+
+f <- makebargraph(DEGmanipm, "pituitary","DEGs w/ + LFC", 0, 3200, DEGmanipm$comparison) + 
+  labs(subtitle = "male pituitary", x = "Comparison of manipulated group to parental controls") 
+  scale_x_discrete(breaks= comparisonlevelsmanip,
+                      labels= comparisonlevelsmanip,
+                   drop=FALSE)
+
+i <- makebargraph(DEGmanipm, "gonad","DEGs w/ + LFC", 0, 3200, DEGmanipm$comparison) + 
+  theme(strip.text = element_blank()) + 
+  labs(subtitle = "male gonad", x = "Comparison of manipulated group to parental controls")  + 
+  scale_x_discrete(breaks= comparisonlevelsmanip,
+                      labels= comparisonlevelsmanip,
+                   drop=FALSE)
+
+supplfig3 <- plot_grid(c, a,b,f,d,e,i,g,h, ncol = 3, rel_widths = c(2,1.8,1), 
+                  labels = c("A", "", "",
+                             "B", "", "",
+                             "C", "", ""), label_size = 8, hjust = 0)
+supplfig3
+
+```
 ## write files
@@
 
+pdf(file="../figures/supplfig-3-1.pdf", width=5, height=5)
+plot(supplfig3)
+dev.off()
+
+write.csv(candidateDEGS, "../results/table2.csv", row.names = F)
+
 write.csv(candidatevsd, "../results/fig3-suppltable.csv")

### 2020-08-07 analysis/figure-3.Rmd
typo fix

@@
   theme(strip.text = element_blank()) + 
-  labs(title = "female hypothalamus",x = "Comparison of manipulated group to parental controls") + 
+  labs(title = "female gonads",x = "Comparison of manipulated group to parental controls") + 
   scale_x_discrete(breaks= comparisonlevelsmanip,

### 2020-08-19 analysis/boneyard.Rmd
now with summary of DEGs characteriztion

@@
 ```
+
+## Candidate Correlations - SUppl fig 1
+
+```{r correlations}
+hyp1 <- makecorrdf("female", "hypothalamus", litNotGOknown)  
+pit1 <- makecorrdf("female", "pituitary", litNotGOknown)  
+gon1 <- makecorrdf("female", "gonad", litNotGOknown)
+
+hyp2 <- makecorrdf("male", "hypothalamus", litNotGOknown)  
+pit2 <- makecorrdf("male", "pituitary", litNotGOknown)  
+gon2 <- makecorrdf("male", "gonad", litNotGOknown)
+
+hyp3 <- makecorrdf("female", "hypothalamus", GOgenes)  
+pit3 <- makecorrdf("female", "pituitary", GOgenes)  
+gon3 <- makecorrdf("female", "gonad", GOgenes)
+
+hyp4 <- makecorrdf("male", "hypothalamus", GOgenes)  
+pit4 <- makecorrdf("male", "pituitary", GOgenes)  
+gon4 <- makecorrdf("male", "gonad", GOgenes)  
+```
+
+```{r supplfig-2, fig.width=7.25, fig.height=7.25}
+
+b1 <- plotcorrplot(hyp1, "females") + labs(y = "Hypothalamus", title = "Parental Care Literature")  + 
+  scale_x_discrete(position = "top")   + theme(axis.text.x = element_text(vjust = -0.25))
+b2 <- plotcorrplot(pit1, NULL)  + labs(y = "Pituitary") + 
+  theme(axis.text.x = element_blank())
+b3 <- plotcorrplot(gon1, NULL)  + labs(y = "Gonad") + theme(axis.text.x = element_text(vjust = 0.25))
+
+b4 <- plotcorrplot(hyp2, "males") + labs( title =  " ") + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = -0.25)) +
+  scale_x_discrete(position = "top") 
+b5 <- plotcorrplot(pit2, NULL)    + 
+  theme(axis.text.x = element_blank(), axis.text.y = element_blank())
+b6 <- plotcorrplot(gon2, NULL)  + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = 0.25))
+
+b7 <- plotcorrplot(hyp3, "females") + labs( title = "Parental Care GO") + 
+   scale_x_discrete(position = "top") + theme(axis.text.x = element_text(vjust = -0.25))
+b8 <- plotcorrplot(pit3, NULL)   + 
+  theme(axis.text.x = element_blank())
+b9 <- plotcorrplot(gon3, NULL)  + theme(axis.text.x = element_text(vjust = 0.25))
+
+b10 <- plotcorrplot(hyp4, "males") + labs( title =  " ") + 
+  theme(axis.text.y = element_blank(),
+        axis.text.x = element_text(vjust = -0.25)) + scale_x_discrete(position = "top") 
+b11 <- plotcorrplot(pit4, NULL)    + 
+  theme(axis.text.x = element_blank() , axis.text.y = element_blank())
+b12 <- plotcorrplot(gon4, NULL)   + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = 0.25))
+
+forlegend <- plotcorrplot(gon4, NULL) + theme(legend.position = "bottom") + labs(color = "Correlation")
+mylegend <- get_legend(forlegend)
+
+allcorplots <- plot_grid(b1,b4,b7,b10,
+          b2,b5,b8,b11,
+          b3,b6,b9,b12, rel_heights = c(1.4,1,1.2),
+          labels = c("A", " ", "B"), label_size = 8, rel_widths = c(1.4,1.2,1.2,1))
+
+supplfig2 <- plot_grid(allcorplots, mylegend, ncol = 1, rel_heights = c(1,0.1))
+supplfig2
+
+```
+
+## Figure
+
+```{r fig2, fig.width = 7.25, fig.height = 7.25}
+
+c1 <- scattercorrelations(FH, FH$DRD1, "DRD1", FH$HTR2C, "HTR2C", "#969696" ) +  labs(title = " ", subtitle = " " )  
+c3 <- scattercorrelations(FP, FP$AVPR1A, "AVPR1A", FP$CRHR1,  "CRHR1", "#969696")  + labs(title = " ", subtitle = " ") 
+c5 <- scattercorrelations(FG, FG$PGR, "PGR",  FG$ESR1,  "ESR1", "#969696")   + labs(title = " ", subtitle = " ")
+
+c2 <- scattercorrelations(MH, MH$DRD1, "DRD1", MH$HTR2C, "HTR2C", "#525252")  + labs(title = " ", subtitle = " " )  
+c4 <- scattercorrelations(MP, MP$AVPR1A,  "AVPR1A",  MP$CRHR1, "CRHR1", "#525252")  + labs(title = " ", subtitle = " " ) 
+c6 <- scattercorrelations(MG, MG$PGR, "PGR", MG$ESR1, "ESR1", "#525252")   + labs(title = " ", subtitle = " " ) 
+```
+
+```{r newfig}
+
+plot3candidates <- function(hypgene, pitgene, gongene){
+
+  a1 <- newcandidateboxplot("hypothalamus", hypgene, "female", "female hypothalamus", "sequential") 
+  a2 <- externalboxplots("hypothalamus", hypgene, "female", " ", "external") 
+  a3 <- newcandidateboxplot("hypothalamus", hypgene, "male", "male hypothalamus", "sequential")
+  a4 <- externalboxplots( "hypothalamus", hypgene, "female", " ", "external")
+
+  hypplots <- plot_grid(a1,a2, a3, a4, rel_widths = c(2,1,2,1), nrow = 1)
+
+  b1 <- newcandidateboxplot("pituitary", pitgene, "female", "female pituitary", "sequential")
+  b2 <- externalboxplots("pituitary", pitgene, "female", " ", "external") 
+  b3 <- newcandidateboxplot("pituitary", pitgene, "male", "male pituitary", "sequential")
+  b4 <- externalboxplots( "pituitary", pitgene, "female", " ", "external")
+
+  pitplots <- plot_grid(b1,b2, b3, b4, rel_widths = c(2,1,2,1), nrow = 1)
+
+  c1 <- newcandidateboxplot("gonad", gongene, "female", "female gonad", "sequential") +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  c2 <- externalboxplots("gonad", gongene, "female", " ", "external") + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  c3 <- newcandidateboxplot("gonad", gongene, "male", "male gonad", "sequential") + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  c4 <- externalboxplots( "gonad", gongene, "female", " ", "external") + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+  gonplots <- plot_grid(c1,c2, c3, c4, rel_widths = c(2,1,2,1), nrow = 1)
+
+  fig2alt <- plot_grid(hypplots, pitplots, gonplots, nrow = 3,
+                       labels = "AUTO", label_size = 8,
+                       rel_heights = c(1,1,1.2))
+  print(fig2alt)
+
+}
+
+plot3candidates("COMT","PRL","AVPR1A")
+plot3candidates("PRL","PRL", "PRL")
+plot3candidates("PRL","FSH", "ESR1")
+```

### 2020-08-19 analysis/figure-2.Rmd
now with summary of DEGs characteriztion

@@
 ---
-title: "Fig2: Candidate genes"
+title: "Table 1. Candidate genes and differentially expressed genes"
 output: md_document
@@
 
-# Candidate gene analysis
+# Table 1. Candidate genes and differentially expressed genes
 
@@
 
-## Candidate DEGs
+## Differentially expressed genes (DEGs)
 
-```{r DEGs}
-
+```{r}
 # summary DEG results from DESeq2
-candidateDEGS <- read_csv("../results/suppltable1.csv") %>%
-  filter(gene %in% candidategenes) %>%
+allDEGS <- read_csv("../results/suppltable1.csv") %>%
    mutate(posneg = ifelse(lfc >= 0, "+", "-"),
@@
                          "hypothalamus" = "H",
-                         "pituitary" = "P", "gonad" = "G")) %>%
-  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
-  select(gene, res, comparison)  %>%
-  group_by(gene,  comparison) %>%
-  summarize(res = str_c(res, collapse = " ")) %>%
-  pivot_wider(names_from = comparison, values_from = res) %>%
-  select(gene, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
-candidateDEGS
+                         "pituitary" = "P", "gonad" = "G"),
+         comparison = recode(comparison, 
+                         "bldg_lay" = "Bldg2Lay",
+                         "lay_inc.d3" = "Lay2Inc3", 
+                         "inc.d3_inc.d9" = "Inc3toInc9",
+                         "inc.d9_inc.d17" = "Inc9toInc17",
+                         "inc.d17_hatch" = "In17toHatch",
+                         "hatch_n5" = "Hatch2N5",
+                         "n5_n9" = "N5toN9"),
+         group = paste(sex, tissue, sep = "")) %>%
+  mutate(res = paste("(", posneg, ")", sep = "")) %>%
+  mutate(compres = paste(comparison, res, sep = "")) %>%
+  select(gene, group, compres) 
+allDEGS
 
-## table 1 summary candidate genes
-table1 <- left_join(candidateDEGS, parentalcaregenes) %>%
-  select(gene, lay_inc.d3:n5_n9, literature, GO, NCBI) %>%
-  mutate(literature = if_else(is.na(literature), " ", "X"),
-         GO = if_else(is.na(GO), " ", "X"))
-(table1)
-```
+DEGsbytissue <- function(whichtissueshorthand, whichsex){
 
-## Candidate VSDs
+  df <- allDEGS %>%
+  filter(grepl(whichtissueshorthand, group),
+         grepl(whichsex,group)) %>%
+  group_by(group,gene) %>%
+  summarize(results = str_c(compres, collapse = "; ")) %>%
+  mutate(n = (str_count(results, pattern = ";"))+1) %>%
+  arrange(desc(n))  %>%
+  filter(n>1)
+  #group_by(comparisons) %>%
+  #summarize(genes = str_c(gene, collapse = "; "))
 
-```{r candidatevsds}
-# load `candidatevsd` with `source("../R/wrangledata.R")`
-candidatevsd <- read_csv("../results/03_candidatevsd.csv") %>% 
-  select(-X1) %>%
-  filter(treatment %in% charlevels) %>%
-  mutate(treatment = factor(treatment, levels = charlevels)) %>%
-  drop_na() %>%
-    mutate(external = fct_collapse(treatment, 
-                                   "none" = c("control", "bldg"),
-                                   "eggs" = c("lay" , "inc.d3", 
-                                              "inc.d9", "inc.d17"),
-                                   "chicks" = c("hatch", "n5", "n9")))
-head(candidatevsd)
+  print(head(df))
+  return(df)
 
-candidatevsdwide <- candidatevsd  %>%
-  select(-external) %>%
-  pivot_wider(names_from = gene, values_from = counts) 
-FH <- subsetcandidatevsdwide("female", "hypothalamus")
-FP <- subsetcandidatevsdwide("female", "pituitary")
-FG <- subsetcandidatevsdwide("female", "gonad")
-MH <- subsetcandidatevsdwide("male", "hypothalamus")
-MP <- subsetcandidatevsdwide("male", "pituitary")
-MG <- subsetcandidatevsdwide("male", "gonad") 
-```
+}
 
-## Candidate Correlations - SUppl fig 1
+hmDEGs <- DEGsbytissue("H", "M")
+pmDEGs <- DEGsbytissue("P", "M")
+gmDEGs <- DEGsbytissue("G", "M")
 
-```{r correlations}
-hyp1 <- makecorrdf("female", "hypothalamus", litNotGOknown)  
-pit1 <- makecorrdf("female", "pituitary", litNotGOknown)  
-gon1 <- makecorrdf("female", "gonad", litNotGOknown)
+hfDEGs <- DEGsbytissue("H", "F")
+pfDEGs <- DEGsbytissue("P", "F")
+gfDEGs <- DEGsbytissue("G", "F")
 
-hyp2 <- makecorrdf("male", "hypothalamus", litNotGOknown)  
-pit2 <- makecorrdf("male", "pituitary", litNotGOknown)  
-gon2 <- makecorrdf("male", "gonad", litNotGOknown)
+maleDEGs <- rbind(hmDEGs, pmDEGs) %>%
+  rbind(., gmDEGs)
+maleDEGs
 
-hyp3 <- makecorrdf("female", "hypothalamus", GOgenes)  
-pit3 <- makecorrdf("female", "pituitary", GOgenes)  
-gon3 <- makecorrdf("female", "gonad", GOgenes)
+femaleDEGs <- rbind(hfDEGs, pfDEGs) %>%
+  rbind(., gfDEGs)
+femaleDEGs
 
-hyp4 <- makecorrdf("male", "hypothalamus", GOgenes)  
-pit4 <- makecorrdf("male", "pituitary", GOgenes)  
-gon4 <- makecorrdf("male", "gonad", GOgenes)  
-```
-
-```{r supplfig-2, fig.width=7.25, fig.height=7.25}
-
-b1 <- plotcorrplot(hyp1, "females") + labs(y = "Hypothalamus", title = "Parental Care Literature")  + 
-  scale_x_discrete(position = "top")   + theme(axis.text.x = element_text(vjust = -0.25))
-b2 <- plotcorrplot(pit1, NULL)  + labs(y = "Pituitary") + 
-  theme(axis.text.x = element_blank())
-b3 <- plotcorrplot(gon1, NULL)  + labs(y = "Gonad") + theme(axis.text.x = element_text(vjust = 0.25))
-
-b4 <- plotcorrplot(hyp2, "males") + labs( title =  " ") + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = -0.25)) +
-  scale_x_discrete(position = "top") 
-b5 <- plotcorrplot(pit2, NULL)    + 
-  theme(axis.text.x = element_blank(), axis.text.y = element_blank())
-b6 <- plotcorrplot(gon2, NULL)  + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = 0.25))
-
-b7 <- plotcorrplot(hyp3, "females") + labs( title = "Parental Care GO") + 
-   scale_x_discrete(position = "top") + theme(axis.text.x = element_text(vjust = -0.25))
-b8 <- plotcorrplot(pit3, NULL)   + 
-  theme(axis.text.x = element_blank())
-b9 <- plotcorrplot(gon3, NULL)  + theme(axis.text.x = element_text(vjust = 0.25))
-
-b10 <- plotcorrplot(hyp4, "males") + labs( title =  " ") + 
-  theme(axis.text.y = element_blank(),
-        axis.text.x = element_text(vjust = -0.25)) + scale_x_discrete(position = "top") 
-b11 <- plotcorrplot(pit4, NULL)    + 
-  theme(axis.text.x = element_blank() , axis.text.y = element_blank())
-b12 <- plotcorrplot(gon4, NULL)   + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = 0.25))
-
-forlegend <- plotcorrplot(gon4, NULL) + theme(legend.position = "bottom") + labs(color = "Correlation")
-mylegend <- get_legend(forlegend)
-
-allcorplots <- plot_grid(b1,b4,b7,b10,
-          b2,b5,b8,b11,
-          b3,b6,b9,b12, rel_heights = c(1.4,1,1.2),
-          labels = c("A", " ", "B"), label_size = 8, rel_widths = c(1.4,1.2,1.2,1))
-
-supplfig2 <- plot_grid(allcorplots, mylegend, ncol = 1, rel_heights = c(1,0.1))
-supplfig2
+summarizedDEGs <- rbind(femaleDEGs,maleDEGs) %>%
+  group_by(group, results) %>%
+  summarize(genes = str_c(gene, collapse = "; ")) 
+summarizedDEGs
 
@@
 
-## Figure
+```{r writefiles}
+write.csv(summarizedDEGs,"../results/summarizedDEGsv2.csv")
 
-```{r fig2, fig.width = 7.25, fig.height = 7.25}
-
-c1 <- scattercorrelations(FH, FH$DRD1, "DRD1", FH$HTR2C, "HTR2C", "#969696" ) +  labs(title = " ", subtitle = " " )  
-c3 <- scattercorrelations(FP, FP$AVPR1A, "AVPR1A", FP$CRHR1,  "CRHR1", "#969696")  + labs(title = " ", subtitle = " ") 
-c5 <- scattercorrelations(FG, FG$PGR, "PGR",  FG$ESR1,  "ESR1", "#969696")   + labs(title = " ", subtitle = " ")
-
-c2 <- scattercorrelations(MH, MH$DRD1, "DRD1", MH$HTR2C, "HTR2C", "#525252")  + labs(title = " ", subtitle = " " )  
-c4 <- scattercorrelations(MP, MP$AVPR1A,  "AVPR1A",  MP$CRHR1, "CRHR1", "#525252")  + labs(title = " ", subtitle = " " ) 
-c6 <- scattercorrelations(MG, MG$PGR, "PGR", MG$ESR1, "ESR1", "#525252")   + labs(title = " ", subtitle = " " ) 
 ```
-
-```{r newfig}
-
-plot3candidates <- function(hypgene, pitgene, gongene){
-
-  a1 <- newcandidateboxplot("hypothalamus", hypgene, "female", "female hypothalamus", "sequential") 
-  a2 <- externalboxplots("hypothalamus", hypgene, "female", " ", "external") 
-  a3 <- newcandidateboxplot("hypothalamus", hypgene, "male", "male hypothalamus", "sequential")
-  a4 <- externalboxplots( "hypothalamus", hypgene, "female", " ", "external")
-
-  hypplots <- plot_grid(a1,a2, a3, a4, rel_widths = c(2,1,2,1), nrow = 1)
-
-  b1 <- newcandidateboxplot("pituitary", pitgene, "female", "female pituitary", "sequential")
-  b2 <- externalboxplots("pituitary", pitgene, "female", " ", "external") 
-  b3 <- newcandidateboxplot("pituitary", pitgene, "male", "male pituitary", "sequential")
-  b4 <- externalboxplots( "pituitary", pitgene, "female", " ", "external")
-
-  pitplots <- plot_grid(b1,b2, b3, b4, rel_widths = c(2,1,2,1), nrow = 1)
-
-  c1 <- newcandidateboxplot("gonad", gongene, "female", "female gonad", "sequential") +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1))
-  c2 <- externalboxplots("gonad", gongene, "female", " ", "external") + 
-    theme(axis.text.x = element_text(angle = 45, hjust = 1))
-  c3 <- newcandidateboxplot("gonad", gongene, "male", "male gonad", "sequential") + 
-    theme(axis.text.x = element_text(angle = 45, hjust = 1))
-  c4 <- externalboxplots( "gonad", gongene, "female", " ", "external") + 
-    theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-  gonplots <- plot_grid(c1,c2, c3, c4, rel_widths = c(2,1,2,1), nrow = 1)
-
-  fig2alt <- plot_grid(hypplots, pitplots, gonplots, nrow = 3,
-                       labels = "AUTO", label_size = 8,
-                       rel_heights = c(1,1,1.2))
-  print(fig2alt)
-
-}
-
-plot3candidates("COMT","PRL","AVPR1A")
-plot3candidates("PRL","PRL", "PRL")
-plot3candidates("PRL","FSH", "ESR1")
-```
-
-```{r writefiles}
-#write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
-#write.csv(candidatevsd, "../results/candidatecounts.csv")
-write.csv(table1, "../results/table1-v2.csv")
-
-pdf(file="../figures/fig2-1.pdf", width=7.25, height=7.25)
-plot(fig2)
-dev.off()
-
-pdf(file="../figures/supplfig-2.pdf", width=7.25, height=7.25)
-plot(supplfig2)
-dev.off()
-
-pdf(file="../figures/fig2-2.pdf", width=7.25, height=7.25)
-plot(newfig2)
-dev.off()
-```

### 2020-08-20 analysis/03_datawrangling.Rmd
shared manipluation DEGs

@@
 
+### manipuluation
+
+```{r}
+head(allDEG)
+
+makemanipulationDEGtables <- function(whichcomparisons){
+  df <- allDEG %>%
+    filter(comparison %in% whichcomparisons) %>%
+     mutate(posneg = ifelse(lfc >= 0, "+", "-"),
+         sex = recode(sex, "female" = "F", "male" = "M" ),
+         tissue = recode(tissue, 
+                         "hypothalamus" = "H",
+                         "pituitary" = "P", "gonad" = "G"),
+         group = paste(sex, tissue, sep = "")) %>%
+  mutate(res = paste("(", posneg, ")", sep = "")) %>%
+  mutate(compres = paste(comparison, res, sep = "")) %>%
+  select(gene, group, compres) %>%
+  group_by(group,gene) %>%
+  summarize(results = str_c(compres, collapse = "; ")) %>%
+  mutate(n = (str_count(results, pattern = ";"))+1) %>%
+  arrange(desc(n))  %>%
+  filter(n>1) %>%
+  group_by(group,results) %>%
+  summarize(genes = str_c(gene, collapse = "; "))  %>%
+  arrange(group, results)
+  print(head(df))
+return(df) 
+}
+
+allDEG %>%
+  group_by(comparison) %>%
+  summarize(n = n())
+
+removalcomps <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2")
+early <- c("early_inc.d9", "hatch_early")
+late <- c("hatch_prolong", "inc.d17_prolong")
+
+removalDEGs <- makemanipulationDEGtables(removalcomps)
+hatchDEGs <- makemanipulationDEGtables(early)
+prolongDEGs <- makemanipulationDEGtables(late)
+
+write.csv(removalDEGs, "../results/removalDEGs.csv")
+write.csv(hatchDEGs, "../results/hatchDEGs.csv")
+write.csv(prolongDEGs, "../results/prolongDEGs.csv")
+
+```
+
 ## save files

### 2020-08-21 analysis/03_datawrangling.Rmd
now with fewer DEGs

--- 2020-08-07 analysis/03_datawrangling.Rmd
@@
 
+### manipuluation
+
+```{r}
+head(allDEG)
+
+filteredDEGs <- allDEG %>%
+  filter(lfc > 1.1 | lfc < -1.1 )
+dim(allDEG)
+dim(filteredDEGs)
+
+makemanipulationDEGtables <- function(whichcomparisons){
+  df <- filteredDEGs %>%
+    filter(comparison %in% whichcomparisons) %>%
+     mutate(posneg = ifelse(lfc >= 0, "+", "-"),
+         sex = recode(sex, "female" = "F", "male" = "M" ),
+         tissue = recode(tissue, 
+                         "hypothalamus" = "H",
+                         "pituitary" = "P", "gonad" = "G"),
+         group = paste(sex, tissue, sep = "")) %>%
+  mutate(res = paste("(", posneg, ")", sep = "")) %>%
+  mutate(compres = paste(comparison, res, sep = "")) %>%
+  select(gene, group, compres) %>%
+  group_by(group,gene) %>%
+  summarize(results = str_c(compres, collapse = "; ")) %>%
+  mutate(n = (str_count(results, pattern = ";"))+1) %>%
+  arrange(desc(n))  %>%
+  filter(n>1) %>%
+  group_by(group,results) %>%
+  summarize(genes = str_c(gene, collapse = "; "))  %>%
+  #arrange(group, results) %>%
+  mutate(n = (str_count(genes, pattern = ";"))+1) %>%
+  arrange(desc(n))
+  print(head(df))
+return(df) 
+}
+
+removalcomps <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2")
+early <- c("early_inc.d9", "hatch_early")
+late <- c("hatch_prolong", "inc.d17_prolong")
+
+removalDEGs <- makemanipulationDEGtables(removalcomps)
+hatchDEGs <- makemanipulationDEGtables(early)
+prolongDEGs <- makemanipulationDEGtables(late)
+
+manipDEGs <- rbind(removalDEGs, hatchDEGs) %>%
+  rbind(., prolongDEGs) %>%
+  arrange(desc(n))
+head(manipDEGs)
+```
+
 ## save files
@@
 write.csv(wideDEGsSequential, "../results/table1.csv", row.names = F)
+
+write.csv(removalDEGs, "../results/removalDEGs.csv")
+write.csv(hatchDEGs, "../results/hatchDEGs.csv")
+write.csv(prolongDEGs, "../results/prolongDEGs.csv")
+
+write.csv(manipDEGs, "../results/manipDEGs.csv")
+
 ```

### 2020-08-21 analysis/figure-2-new.Rmd
now with new bar plots

@@
          direction = factor(direction, levels = alllevels),
-         label = gsub("_","\nvs. ", comparison)) 
+         label = gsub("_","\nvs. ", comparison)) %>%
+  filter(lfc > 1.1 | lfc < -1.1 )
 head(allDEG)
@@
          !grepl("m.|early|extend|prolong", comparison))  %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol))
+  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol)) %>%
+  group_by(sex, tissue, comparison, direction, label) %>%
+    summarise(n = n()) %>%
+    mutate(n = ifelse(direction == "control", n*-1, n*1 ))
 
@@
   filter(grepl("bldg", comparison),
-         !grepl("m.|early|extend|prolong", comparison))  %>%
+         !grepl("m.|early|extend|prolong|control", comparison))  %>%
+  drop_na() %>%
   mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))  %>%
-  drop_na()
+  group_by(sex, tissue, comparison, direction, label) %>%
+    summarise(n = n()) %>%
+    mutate(n = ifelse(direction == "bldg", n*-1, n*1 ))
 
@@
          !grepl("control|bldg", comparison)) %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelschar))
+  mutate(comparison = factor(comparison, levels = comparisonlevelschar)) %>%
+  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
+  group_by(sex, tissue, comparison, direction, label, updown) %>%
+  summarise(n = n()) %>%
+  mutate(n = n*updown ) %>%
+  select(-updown)
+
 ```
@@
 a <- plot.volcano("hypothalamus", sexlevels,  "control_bldg") + 
-  facet_wrap(~sex) + labs(subtitle = " ")
-b <- makebargraph(DEGcontrol, "hypothalamus","DEGs w/ + LFC", 0, 5000, comparisonlabelscontrol) +
-  labs(title = " ",
-       subtitle = "Control versus all other reproductive and parental stages")
+  facet_wrap(~sex)
 
-c <- makebargraph(DEGbldg, "hypothalamus", "DEGs w/ + LFC", 0, 490, comparisonlabelsbldg) +
-  labs(subtitle = "Nest-building versus all other parental stages")
-d <- makebargraph(DEGchar, "hypothalamus", NULL, 0, 1800, comparisonlabelscharnobldg) +
-  labs(subtitle = "Comparison of sequential parental stages")
+b <- makebargraphv3(DEGcontrol, "hypothalamus","No. of DEGs\n (-) decreased, incrased (+)", comparisonlabelscontrol) +
+  labs(title = "Control versus all other reproductive and parental stages",
+       subtitle = "hypothalamus") +
+  geom_rect(mapping=aes(xmin=0.5, xmax=1.5, ymin=-1000, ymax = 350, fill = F), color="black", alpha=0.5) +
+  annotate("text", x = 1, y = -1075, label = "B", size = 2.5)
 
-e <- makebargraph(DEGbldg, "pituitary", "DEGs w/ + LFC", 0, 2300, comparisonlabelsbldg)
+c <- makebargraphv3(DEGbldg, "hypothalamus", "No. of DEGs\n (-) decreased, incrased (+)", comparisonlabelsbldg) +
+  labs(title = "Nest-building versus all other parental stages",
+       subtitle = "hypothalamus")
+d <- makebargraphv3(DEGchar, "hypothalamus", NULL,  comparisonlabelscharnobldg) +
+  labs(title = "Comparison of sequential parental stages",
+       subtitle = "hypothalamus")
 
-f <- makebargraph(DEGchar, "pituitary",NULL, 0, 1600, comparisonlabelscharnobldg)
+e <- makebargraphv3(DEGbldg, "pituitary", "No. of DEGs\n (-) decreased, incrased (+)", comparisonlabelsbldg) +
+  labs(subtitle = "pituitary")
+f <- makebargraphv3(DEGchar, "pituitary", NULL,comparisonlabelscharnobldg) +
+  labs(subtitle = "pituitary")
 
-g <- makebargraph(DEGbldg, "gonad", "DEGs w/ + LFC", 0, 2900, comparisonlabelsbldg) 
-h <- makebargraph(DEGchar, "gonad", NULL, 0, 1200, comparisonlabelscharnobldg)
+g <- makebargraphv3(DEGbldg, "gonad", "No. of DEGs\n (-) decreased, incrased (+)", comparisonlabelsbldg) +
+  labs(subtitle = "gonads")
+h <- makebargraphv3(DEGchar, "gonad", NULL, comparisonlabelscharnobldg) +
+  labs(subtitle = "gonads")
 
-ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
-                labels = c("A Hypothalamus"), label_size = 8, hjust = 0)
-cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h")
-ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h",
-                labels = c("B Pitutary"), label_size = 8 ,hjust = 0)
-gh <- plot_grid(g,h,rel_widths = c(1.1,1), align = "h",
-                labels = c("C Gonads"), label_size = 8, hjust = 0)
+ab <- plot_grid(b,a,rel_widths = c(2.5,1), 
+                labels = c("A", "B"), label_size = 8, hjust = 0)
+cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h",
+                labels = c("C", "D"), label_size = 8, hjust = 0)
+ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h")
+gh <- plot_grid(g,h,rel_widths = c(1.1,1), align = "h")
 
@@
 fig
+
 ```

### 2020-08-21 R/functions.R

@@ -929,6 +929,29 @@ makebargraph <- function(df, whichtissue, myylab, lowlim, higherlim, mylabels){
 
 
 
+makebargraphv3 <- function(df, whichtissue, myylab, mylabels){
+  
+  p <- df %>%
+    filter(tissue == whichtissue) %>%
+    ggplot(aes(x = comparison, y = n, fill = direction)) +
+    geom_bar(stat="identity") +
+    facet_wrap(~sex) +
+    theme_B3() +
+    theme(legend.position = "none")  +
+    scale_fill_manual(values = allcolors,
+                      name = " ",
+                      drop = FALSE) +
+    scale_color_manual(values = allcolors) +
+    geom_text(stat='identity', aes(label= n), vjust =-0.5, 
+              position = position_dodge(width = 1),
+              size = 1.75, color = "black") +
+    labs(x = NULL, y = myylab)  +
+    scale_x_discrete(labels = mylabels)
+  return(p)
+}
+
+
+
 sexbarplots <- function(df, lowlim, higherlim){
    
   DEGs <- df %>% dplyr::filter(direction != "NS")

### 2020-08-21 analysis/figure-2-new.Rmd
better labels

@@
 
-b <- makebargraphv3(DEGcontrol, "hypothalamus","No. of DEGs\n (-) decreased, incrased (+)", comparisonlabelscontrol) +
+b <- makebargraphv3(DEGcontrol, "hypothalamus","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
   labs(title = "Control versus all other reproductive and parental stages",
@@
 
-c <- makebargraphv3(DEGbldg, "hypothalamus", "No. of DEGs\n (-) decreased, incrased (+)", comparisonlabelsbldg) +
+c <- makebargraphv3(DEGbldg, "hypothalamus", "No. of DEGs\n (-)decreased  increased (+)", comparisonlabelsbldg) +
   labs(title = "Nest-building versus all other parental stages",
@@
 
-e <- makebargraphv3(DEGbldg, "pituitary", "No. of DEGs\n (-) decreased, incrased (+)", comparisonlabelsbldg) +
+e <- makebargraphv3(DEGbldg, "pituitary", "No. of DEGs\n (-)decreased  increased (+)", comparisonlabelsbldg) +
   labs(subtitle = "pituitary")
@@
 
-g <- makebargraphv3(DEGbldg, "gonad", "No. of DEGs\n (-) decreased, incrased (+)", comparisonlabelsbldg) +
+g <- makebargraphv3(DEGbldg, "gonad", "No. of DEGs\n (-)decreased  increased (+)", comparisonlabelsbldg) +
   labs(subtitle = "gonads")

### 2020-08-21 analysis/boneyard.Rmd
now appropriately names, some DEG anlaysese moved to boneyard

@@
 ```
+
+# Table 1. Candidate genes and differentially expressed genes
+
+## Genes of interest
+
+```{r candidategenes}
+parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% select(-X1)
+candidategenes <- parentalcaregenes %>% pull(gene) 
+candidategenes
+
+GOgenes <- parentalcaregenes %>% distinct(GO)  %>% drop_na() %>% pull(GO)
+GOgenes
+
+literaturegenes <- parentalcaregenes %>% 
+  distinct(literature) %>% 
+  drop_na() %>% 
+  pull(literature)
+literaturegenes
+
+litNotGO <-  literaturegenes[!literaturegenes %in% GOgenes]
+litNotGOknown <- litNotGO[!litNotGO %in% c("ADRA2A", "MEST", "SLC6A4")]
+litNotGOknown
+```
+
+## Differentially expressed genes (DEGs)
+
+```{r}
+# summary DEG results from DESeq2
+allDEGS <- read_csv("../results/suppltable1.csv") %>%
+   mutate(posneg = ifelse(lfc >= 0, "+", "-"),
+         sex = recode(sex, "female" = "F", "male" = "M" ),
+         tissue = recode(tissue, 
+                         "hypothalamus" = "H",
+                         "pituitary" = "P", "gonad" = "G"),
+         comparison = recode(comparison, 
+                         "bldg_lay" = "Bldg2Lay",
+                         "lay_inc.d3" = "Lay2Inc3", 
+                         "inc.d3_inc.d9" = "Inc3toInc9",
+                         "inc.d9_inc.d17" = "Inc9toInc17",
+                         "inc.d17_hatch" = "In17toHatch",
+                         "hatch_n5" = "Hatch2N5",
+                         "n5_n9" = "N5toN9"),
+         group = paste(sex, tissue, sep = "")) %>%
+  mutate(res = paste("(", posneg, ")", sep = "")) %>%
+  mutate(compres = paste(comparison, res, sep = "")) %>%
+  select(gene, group, compres) 
+allDEGS
+
+DEGsbytissue <- function(whichtissueshorthand, whichsex){
+
+  df <- allDEGS %>%
+  filter(grepl(whichtissueshorthand, group),
+         grepl(whichsex,group)) %>%
+  group_by(group,gene) %>%
+  summarize(results = str_c(compres, collapse = "; ")) %>%
+  mutate(n = (str_count(results, pattern = ";"))+1) %>%
+  arrange(desc(n))  %>%
+  filter(n>1)
+  #group_by(comparisons) %>%
+  #summarize(genes = str_c(gene, collapse = "; "))
+
+  print(head(df))
+  return(df)
+
+}
+
+hmDEGs <- DEGsbytissue("H", "M")
+pmDEGs <- DEGsbytissue("P", "M")
+gmDEGs <- DEGsbytissue("G", "M")
+
+hfDEGs <- DEGsbytissue("H", "F")
+pfDEGs <- DEGsbytissue("P", "F")
+gfDEGs <- DEGsbytissue("G", "F")
+
+maleDEGs <- rbind(hmDEGs, pmDEGs) %>%
+  rbind(., gmDEGs)
+maleDEGs
+
+femaleDEGs <- rbind(hfDEGs, pfDEGs) %>%
+  rbind(., gfDEGs)
+femaleDEGs
+
+summarizedDEGs <- rbind(femaleDEGs,maleDEGs) %>%
+  group_by(group, results) %>%
+  summarize(genes = str_c(gene, collapse = "; ")) 
+summarizedDEGs
+
+```
+
+```{r writefiles}
+write.csv(summarizedDEGs,"../results/summarizedDEGsv2.csv")
+
+```

### 2020-08-21 analysis/figure-2-new.Rmd
now appropriately names, some DEG anlaysese moved to boneyard

@@
 ```{r write}
-pdf(file="../figures/fig2-new-1.pdf", width=7, height=7)
+pdf(file="../figures/fig2-1.pdf", width=7, height=7)
 plot(fig)
@@
 
-png("../figures/fig2-new-1.png", width = 7, height = 7, 
+png("../figures/fig2-1.png", width = 7, height = 7, 
     units = 'in', res = 300)

### 2020-08-21 analysis/boneyard.Rmd
figs 2-4 made with 1 script

@@
 ```
+
+```{r}
+e <- makebargraphv3(DEGbldg, "pituitary", "No. of DEGs\n (-)decreased  increased (+)", comparisonlabelsbldg) +
+  labs(subtitle = "pituitary")
+f <- makebargraphv3(DEGchar, "pituitary", NULL,comparisonlabelscharnobldg) +
+  labs(subtitle = "pituitary")
+
+g <- makebargraphv3(DEGbldg, "gonad", "No. of DEGs\n (-)decreased  increased (+)", comparisonlabelsbldg) +
+  labs(subtitle = "gonads")
+h <- makebargraphv3(DEGchar, "gonad", NULL, comparisonlabelscharnobldg) +
+  labs(subtitle = "gonads")
+
+ab <- plot_grid(b,a,rel_widths = c(2.5,1), 
+                labels = c("A", "B"), label_size = 8, hjust = 0)
+cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h",
+                labels = c("C", "D"), label_size = 8, hjust = 0)
+ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h")
+gh <- plot_grid(g,h,rel_widths = c(1.1,1), align = "h")
+
+fig <- plot_grid(ab,cd,ef, gh, ncol = 1)
+fig
+```

### 2020-08-21 analysis/figure-2.Rmd
shared manipluation DEGs
figs 2-4 made with 1 script

@@
 ---
-title: "Table 1. Candidate genes and differentially expressed genes"
+title: "Fig 2"
 output: md_document
@@
 
-# Table 1. Candidate genes and differentially expressed genes
+# Fig 2
 
-```{r setup, message=F}
+```{r setup, include=T}
 library(tidyverse)
-library(ggtext)
 library(cowplot)
-library(ggpubr)
-library(knitr)
-library(kableExtra)
-library(corrr)
 library(ggsignif)
-library(magick)
-library(scales)
-library(ggimage)
 
@@
 source("../R/functions.R")
+source("../R/genelists.R")
 
-knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
+knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
 ```
 
-## Candidate genes
+## Treatment specific DEGs
 
-```{r candidategenes}
-parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% select(-X1)
-candidategenes <- parentalcaregenes %>% pull(gene) 
-candidategenes
+```{r suppltables}
+allDEG <- read_csv("../results/03_allDEG.csv") %>%
+  mutate(tissue = factor(tissue, levels = tissuelevel),
+         direction = factor(direction, levels = alllevels),
+         label = gsub("_","\nvs. ", comparison)) %>%
+  filter(lfc > 1.1 | lfc < -1.1 )
+head(allDEG)
 
-GOgenes <- parentalcaregenes %>% distinct(GO)  %>% drop_na() %>% pull(GO)
-GOgenes
+# for suppl figures
+DEGcontrol <- allDEG %>% 
+  filter(grepl("control", comparison),
+         !grepl("m.|early|extend|prolong", comparison))  %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol)) %>%
+  group_by(sex, tissue, comparison, direction, label) %>%
+    summarise(n = n()) %>%
+    mutate(n = ifelse(direction == "control", n*-1, n*1 ))
 
-literaturegenes <- parentalcaregenes %>% 
-  distinct(literature) %>% 
+DEGbldg <- allDEG %>% 
+  filter(grepl("bldg", comparison),
+         !grepl("m.|early|extend|prolong|control", comparison))  %>%
   drop_na() %>%
-  pull(literature)
-literaturegenes
+  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))  %>%
+  group_by(sex, tissue, comparison, direction, label) %>%
+    summarise(n = n()) %>%
+    mutate(n = ifelse(direction == "bldg", n*-1, n*1 ))
 
-litNotGO <-  literaturegenes[!literaturegenes %in% GOgenes]
-litNotGOknown <- litNotGO[!litNotGO %in% c("ADRA2A", "MEST", "SLC6A4")]
-litNotGOknown
+DEGchar <- allDEG %>% 
+  filter(comparison %in% comparisonlevelschar,
+         !grepl("control|bldg", comparison)) %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelschar)) %>%
+  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
+  group_by(sex, tissue, comparison, direction, label, updown) %>%
+  summarise(n = n()) %>%
+  mutate(n = n*updown ) %>%
+  select(-updown)
+
+DEGremove <- allDEG %>% 
+  filter(comparison %in% comparisonlevelsremoval) %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelsremoval)) %>%
+  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
+  group_by(sex, tissue, comparison, direction, label, updown) %>%
+  summarise(n = n()) %>%
+  mutate(n = n*updown ) %>%
+  select(-updown)
+DEGremove
+
+DEGreplace <- allDEG %>% 
+  filter(comparison %in% comparisonlevelsreplace) %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelsreplace)) %>%
+  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
+  group_by(sex, tissue, comparison, direction, label, updown) %>%
+  summarise(n = n()) %>%
+  mutate(n = n*updown ) %>%
+  select(-updown)
+DEGreplace
 ```
 
-## Differentially expressed genes (DEGs)
+## Variance stabilized data
 
-```{r}
-# summary DEG results from DESeq2
-allDEGS <- read_csv("../results/suppltable1.csv") %>%
-   mutate(posneg = ifelse(lfc >= 0, "+", "-"),
-         sex = recode(sex, "female" = "F", "male" = "M" ),
-         tissue = recode(tissue, 
-                         "hypothalamus" = "H",
-                         "pituitary" = "P", "gonad" = "G"),
-         comparison = recode(comparison, 
-                         "bldg_lay" = "Bldg2Lay",
-                         "lay_inc.d3" = "Lay2Inc3", 
-                         "inc.d3_inc.d9" = "Inc3toInc9",
-                         "inc.d9_inc.d17" = "Inc9toInc17",
-                         "inc.d17_hatch" = "In17toHatch",
-                         "hatch_n5" = "Hatch2N5",
-                         "n5_n9" = "N5toN9"),
-         group = paste(sex, tissue, sep = "")) %>%
-  mutate(res = paste("(", posneg, ")", sep = "")) %>%
-  mutate(compres = paste(comparison, res, sep = "")) %>%
-  select(gene, group, compres) 
-allDEGS
+```{r data}
+countData <- read_csv("../results/01_limma.csv") %>%
+  column_to_rownames(var = "X1")
+countData <- as.data.frame(t(countData))
+head(countData[1:3])
 
-DEGsbytissue <- function(whichtissueshorthand, whichsex){
+colData <- read_csv("../metadata/00_colData.csv") %>%
+  mutate(treatment = factor(treatment, levels = alllevels),
+         tissue = factor(tissue, levels = tissuelevels)) %>% 
+  column_to_rownames(var = "X1") 
+head(colData)
+# check ready for analysis
+# row.names(countData) == row.names(colData)
+head(row.names(countData) == row.names(colData))
 
-  df <- allDEGS %>%
-  filter(grepl(whichtissueshorthand, group),
-         grepl(whichsex,group)) %>%
-  group_by(group,gene) %>%
-  summarize(results = str_c(compres, collapse = "; ")) %>%
-  mutate(n = (str_count(results, pattern = ";"))+1) %>%
-  arrange(desc(n))  %>%
-  filter(n>1)
-  #group_by(comparisons) %>%
-  #summarize(genes = str_c(gene, collapse = "; "))
+vsd_path <- "../results/DEseq2/treatment/"   # path to the data
+vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
+vsd_pathfiles <- paste0(vsd_path, vsd_files)
+vsd_files
 
-  print(head(df))
-  return(df)
+allvsd <- vsd_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name")  %>% 
+  dplyr::rename("gene" = "X1") 
+## before pivoting, check names of df with
+## head(names(allvsd)) and tail(names(allvsd))
+allvsd <- allvsd %>% 
+  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "samples", values_to = "counts")  %>% 
+    drop_na() 
+allvsd <- as_tibble(allvsd)
+head(allvsd[2:4])
+tail(allvsd[2:4])
 
+# for plotting gene expression over time
+getcandidatevsdmanip <- function(whichgenes, whichtissue){
+
+  candidates  <- allvsd %>%
+    filter(gene %in% whichgenes) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1),
+                  sextissue = sapply(strsplit(sextissue, '../results/DEseq2/treatment/'), "[", 2),
+                  sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                sex = factor(sex),
+                  tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                treatment = sapply(strsplit(samples, '\\_'), "[", 4),
+                treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1),
+                treatment = recode(treatment, "m.hatch" = "m.n2", "extend.hatch" = "m.n2",
+                                                   "inc.prolong" = "prolong", "m.inc.d8" = "early"),
+                treatment =  factor(treatment, levels = alllevels),
+                earlylate = fct_collapse(treatment, 
+                                  early = c("early", "m.inc.d3", "m.inc.d9",
+                                           "lay", "inc.d3", "inc.d9" ),
+
+                                  late = c("inc.d17", "m.inc.d17", "prolong" ,  "hatch" , 
+                                         "m.n2", "extend", "n5", "n9"),
+                                  reference = c("control", "bldg")),
+                extint = fct_collapse(treatment, 
+                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17"),
+                                  eggs = c("lay","inc.d3", "inc.d9", "inc.d17", "prolong"),
+                                  chicks = c("early", "hatch", "extend", "n5", "n9"),
+                                  reference = c("control", "bldg")))  %>%
+    dplyr::select(gene, counts, sex, tissue, treatment, earlylate, extint, samples) 
+  return(candidates)
 }
 
-hmDEGs <- DEGsbytissue("H", "M")
-pmDEGs <- DEGsbytissue("P", "M")
-gmDEGs <- DEGsbytissue("G", "M")
+candidatevsd <- getcandidatevsdmanip(candidategenes)
 
-hfDEGs <- DEGsbytissue("H", "F")
-pfDEGs <- DEGsbytissue("P", "F")
-gfDEGs <- DEGsbytissue("G", "F")
+hypvsd <- candidatevsd %>% filter(tissue %in% "hypothalamus")  
+pitvsd <- candidatevsd %>% filter(tissue %in% "pituitary") 
+gonvsd <- candidatevsd %>% filter(tissue %in% "gonad")
 
-maleDEGs <- rbind(hmDEGs, pmDEGs) %>%
-  rbind(., gmDEGs)
-maleDEGs
+hypvsdf <- candidatevsd %>% filter(tissue %in% "hypothalamus", sex == "female")  
+pitvsdf <- candidatevsd %>% filter(tissue %in% "pituitary", sex == "female")
+gonvsdf <- candidatevsd %>% filter(tissue %in% "gonad", sex == "female")  
+hypvsdm <- candidatevsd %>% filter(tissue %in% "hypothalamus", sex == "male")  
+pitvsdm <- candidatevsd %>% filter(tissue %in% "pituitary", sex == "male") 
+gonvsdm <- candidatevsd %>% filter(tissue %in% "gonad", sex == "male")
 
-femaleDEGs <- rbind(hfDEGs, pfDEGs) %>%
-  rbind(., gfDEGs)
-femaleDEGs
-
-summarizedDEGs <- rbind(femaleDEGs,maleDEGs) %>%
-  group_by(group, results) %>%
-  summarize(genes = str_c(gene, collapse = "; ")) 
-summarizedDEGs
-
+summary(hypvsd)
+summary(pitvsd)
 ```
 
-```{r writefiles}
-write.csv(summarizedDEGs,"../results/summarizedDEGsv2.csv")
+```{r fig2, fig.width=7.5, fig.height=7.25}
+a <- plot.volcano("hypothalamus", sexlevels,  "control_bldg") + 
+  facet_wrap(~sex)
 
+b <- makebargraphv3(DEGcontrol, "hypothalamus","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
+  labs(x = "Control versus all other reproductive and parental stages",
+       title = "Hypothalamus") +
+  geom_rect(mapping=aes(xmin=0.5, xmax=1.5, ymin=-1000, ymax = 350, fill = F), color="black", alpha=0.5) +
+  annotate("text", x = 1, y = -1075, label = "D", size = 2.5)
+
+c <- makebargraphv3(DEGbldg, "hypothalamus", "No. of DEGs\n (-)decreased  increased (+)", comparisonlabelsbldg) +
+  labs(x = "Nest-building versus all other parental stages")
+d <- makebargraphv3(DEGchar, "hypothalamus", NULL,  comparisonlabelscharnobldg) +
+  labs(x = "Comparison of sequential parental stages")
+
+e <- makebargraphv3(DEGremove, "hypothalamus", "No. of DEGs\n (-)decreased  increased (+)", comparisonlevelsremoval) +
+  labs(x = "Offspring removal versus temporal control")
+f <- makebargraphv3(DEGreplace, "hypothalamus", NULL,  comparisonlevelsreplace) +
+  labs(x = "Offspring removal versus temporal or external control")
+
+g <- plotcandidatechar(hypvsdf, "AVP") + labs(subtitle = "Hypothalamus")
+h <- plotcandidatechar(hypvsdm, "AVP") + labs(y = NULL, x = "") + labs(subtitle = " ")
+i <- plotremoval(hypvsdf, "AVP")+ labs(y = NULL) + labs(subtitle = " ")
+j <- plotremoval(hypvsdm, "AVP") + labs(y = NULL, x = "")+ labs(subtitle = " ")
+k <- plotreplacement(hypvsdf, "AVP") + labs(y = NULL)+ labs(subtitle = " ")
+l <- plotreplacement(hypvsdm, "AVP") + labs(y = NULL, x = "")+ labs(subtitle = " ")
+
+ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
+                labels = c("D", "E"), label_size = 8, hjust = 0)
+cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h",
+                labels = c("F", "G"), label_size = 8, hjust = 0)
+ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h",
+                labels = c("H", "I"), label_size = 8, hjust = 0)
+
+#h <- boxplotextint(hypvsdf, "HTR2C", " ")
+
+ghi <- plot_grid(g,h,i, j,k,l,nrow = 1,
+                labels = c("A", "", "B", "", "C"), label_size = 8, hjust = 0,
+                rel_widths = c(9,9,8,8,6,6))
+
+fig2 <- plot_grid(ghi,ab,cd,ef,  ncol = 1)
+fig2
 ```
+
+```{r fig3, fig.width=7.5, fig.height=7.25}
+a <- plot.volcano("pituitary", sexlevels,  "control_bldg") + 
+  facet_wrap(~sex)
+
+b <- makebargraphv3(DEGcontrol, "pituitary","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
+  labs(x = "Control versus all other reproductive and parental stages",
+       title = "Pituitary")
+
+c <- makebargraphv3(DEGbldg, "pituitary", "No. of DEGs\n (-)decreased  increased (+)", comparisonlabelsbldg) +
+  labs(x = "Nest-building versus all other parental stages")
+d <- makebargraphv3(DEGchar, "pituitary", NULL,  comparisonlabelscharnobldg) +
+  labs(x = "Comparison of sequential parental stages")
+
+e <- makebargraphv3(DEGremove, "pituitary", "No. of DEGs\n (-)decreased  increased (+)", comparisonlevelsremoval) +
+  labs(x = "Offspring removal versus temporal control")
+f <- makebargraphv3(DEGreplace, "pituitary", NULL,  comparisonlevelsreplace) +
+  labs(x = "Offspring removal versus temporal or external control")
+
+g <- plotcandidatechar(pitvsdf, "PRL") + labs(subtitle = "Pituitary")
+h <- plotcandidatechar(pitvsdm, "PRL") + labs(y = NULL, x = "") + labs(subtitle = " ")
+i <- plotremoval(pitvsdf, "PRL")+ labs(y = NULL) + labs(subtitle = " ")
+j <- plotremoval(pitvsdm, "PRL") + labs(y = NULL, x = "")+ labs(subtitle = " ")
+k <- plotreplacement(pitvsdf, "PRL") + labs(y = NULL)+ labs(subtitle = " ")
+l <- plotreplacement(pitvsdm, "PRL") + labs(y = NULL, x = "")+ labs(subtitle = " ")
+
+ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
+                labels = c("D", "E"), label_size = 8, hjust = 0)
+cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h",
+                labels = c("F", "G"), label_size = 8, hjust = 0)
+ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h",
+                labels = c("H", "I"), label_size = 8, hjust = 0)
+
+ghi <- plot_grid(g,h,i, j,k,l,nrow = 1,
+                labels = c("A", "", "B", "", "C"), label_size = 8, hjust = 0,
+                rel_widths = c(9,9,8,8,6,6))
+
+fig3 <- plot_grid(ghi,ab,cd,ef,  ncol = 1)
+fig3
+```
+
+```{r fig4, fig.width=7.5, fig.height=7.25}
+a <- plot.volcano("gonad", sexlevels,  "control_bldg") + 
+  facet_wrap(~sex)
+
+b <- makebargraphv3(DEGcontrol, "gonad","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
+  labs(x = "Control versus all other reproductive and parental stages",
+       title = "Pituitary")
+
+c <- makebargraphv3(DEGbldg, "gonad", "No. of DEGs\n (-)decreased  increased (+)", comparisonlabelsbldg) +
+  labs(x = "Nest-building versus all other parental stages")
+d <- makebargraphv3(DEGchar, "gonad", NULL,  comparisonlabelscharnobldg) +
+  labs(x = "Comparison of sequential parental stages")
+
+e <- makebargraphv3(DEGremove, "gonad", "No. of DEGs\n (-)decreased  increased (+)", comparisonlevelsremoval) +
+  labs(x = "Offspring removal versus temporal control")
+f <- makebargraphv3(DEGreplace, "gonad", NULL,  comparisonlevelsreplace) +
+  labs(x = "Offspring removal versus temporal or external control")
+
+g <- plotcandidatechar(pitvsdf, "ESR1") + labs(subtitle = "Gonads")
+h <- plotcandidatechar(pitvsdm, "ESR1") + labs(y = NULL, x = "") + labs(subtitle = " ")
+i <- plotremoval(pitvsdf, "ESR1")+ labs(y = NULL) + labs(subtitle = " ")
+j <- plotremoval(pitvsdm, "ESR1") + labs(y = NULL, x = "")+ labs(subtitle = " ")
+k <- plotreplacement(pitvsdf, "ESR1") + labs(y = NULL)+ labs(subtitle = " ")
+l <- plotreplacement(pitvsdm, "ESR1") + labs(y = NULL, x = "")+ labs(subtitle = " ")
+
+ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
+                labels = c("D", "E"), label_size = 8, hjust = 0)
+cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h",
+                labels = c("F", "G"), label_size = 8, hjust = 0)
+ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h",
+                labels = c("H", "I"), label_size = 8, hjust = 0)
+
+ghi <- plot_grid(g,h,i, j,k,l,nrow = 1,
+                labels = c("A", "", "B", "", "C"), label_size = 8, hjust = 0,
+                rel_widths = c(9,9,8,8,6,6))
+
+fig4 <- plot_grid(ghi,ab,cd,ef,  ncol = 1)
+fig4
+```
+
+## Save files
+
+```{r write}
+pdf(file="../figures/fig2-1.pdf", width=7, height=7)
+plot(fig2)
+dev.off()
+
+png("../figures/fig2-1.png", width = 7, height = 7, 
+    units = 'in', res = 300)
+plot(fig2) 
+dev.off()
+
+pdf(file="../figures/fig3-1.pdf", width=7, height=7)
+plot(fig3)
+dev.off()
+
+png("../figures/fig3-1.png", width = 7, height = 7, 
+    units = 'in', res = 300)
+plot(fig3) 
+dev.off()
+
+pdf(file="../figures/fig4-1.pdf", width=7, height=7)
+plot(fig4)
+dev.off()
+
+png("../figures/fig4-1.png", width = 7, height = 7, 
+    units = 'in', res = 300)
+plot(fig4) 
+dev.off()
+
+sessionInfo()
+```

### 2020-08-21 analysis/figure-2.Rmd
typos

@@
 b <- makebargraphv3(DEGcontrol, "hypothalamus","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
-  labs(x = "Control versus all other reproductive and parental stages",
-       title = "Hypothalamus") +
+  labs(x = "Control versus all other reproductive and parental stages") +
   geom_rect(mapping=aes(xmin=0.5, xmax=1.5, ymin=-1000, ymax = 350, fill = F), color="black", alpha=0.5) +
@@
 b <- makebargraphv3(DEGcontrol, "pituitary","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
-  labs(x = "Control versus all other reproductive and parental stages",
-       title = "Pituitary")
+  labs(x = "Control versus all other reproductive and parental stages")
 
@@
 b <- makebargraphv3(DEGcontrol, "gonad","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
-  labs(x = "Control versus all other reproductive and parental stages",
-       title = "Pituitary")
+  labs(x = "Control versus all other reproductive and parental stages")
 

### 2020-08-24 analysis/figure-2.Rmd
now with htr2c for hyp

@@
 
-g <- plotcandidatechar(hypvsdf, "AVP") + labs(subtitle = "Hypothalamus")
-h <- plotcandidatechar(hypvsdm, "AVP") + labs(y = NULL, x = "") + labs(subtitle = " ")
-i <- plotremoval(hypvsdf, "AVP")+ labs(y = NULL) + labs(subtitle = " ")
-j <- plotremoval(hypvsdm, "AVP") + labs(y = NULL, x = "")+ labs(subtitle = " ")
-k <- plotreplacement(hypvsdf, "AVP") + labs(y = NULL)+ labs(subtitle = " ")
-l <- plotreplacement(hypvsdm, "AVP") + labs(y = NULL, x = "")+ labs(subtitle = " ")
+g <- plotcandidatechar(hypvsdf, "HTR2C") + labs(subtitle = "Hypothalamus")
+h <- plotcandidatechar(hypvsdm, "HTR2C") + labs(y = NULL, x = "") + labs(subtitle = " ")
+i <- plotremoval(hypvsdf, "HTR2C")+ labs(y = NULL) + labs(subtitle = " ")
+j <- plotremoval(hypvsdm, "HTR2C") + labs(y = NULL, x = "")+ labs(subtitle = " ")
+k <- plotreplacement(hypvsdf, "HTR2C") + labs(y = NULL)+ labs(subtitle = " ")
+l <- plotreplacement(hypvsdm, "HTR2C") + labs(y = NULL, x = "")+ labs(subtitle = " ")
 
@@
                 labels = c("H", "I"), label_size = 8, hjust = 0)
-
-#h <- boxplotextint(hypvsdf, "HTR2C", " ")
 

### 2020-08-24 analysis/figure-2.Rmd
working on upsets

@@
 dev.off()
+```
 
+```{r refs}
 sessionInfo()
 ```
+
+```{r}
+head(allDEG)
+
+library(UpSetR)
+head(allDEG)
+
+MG <- allDEG %>%
+  filter(sex == "male", tissue == "hypothalamus", 
+         comparison %in% comparisonlevelsreplace) %>%
+  select(sex, tissue, comparison, gene) %>%
+  mutate(seen = 1) %>%
+  pivot_wider(names_from = comparison, values_from = seen)
+MG
+
+upset(MG) 
+```

### 2020-08-26 analysis/03_datawrangling.Rmd
now with removal venn

@@
 library(readr)
+library(VennDiagram)
 
 source("../R/themes.R")
+
+knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = '../figures/')
 ```
@@
 
-```{r allDEG}
+```{r allDEG, results=F}
 DEG_path <- "../results/DEseq2/treatment/"   # path to the data
@@
 ```
+
+### overlapp
+
+```{r}
+removalcomps <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2")
+early <- c("early_inc.d9", "hatch_early")
+late <- c("hatch_prolong", "inc.d17_prolong")
+
+filteredDEGs %>%
+  filter(comparison %in% removalcomps)  %>%
+  select(sex:gene) %>%
+  pivot_wider(names_from = comparison, values_from = direction)
+
+hatch_m.n2DEGs <- filteredDEGs %>%
+  filter(comparison == "hatch_m.n2") %>%
+  drop_na() %>%
+  pull(gene)
+
+inc.d17_m.inc.d17DEGs <- filteredDEGs %>%
+  filter(comparison == "inc.d17_m.inc.d17") %>%
+  drop_na() %>%
+  pull(gene)
+
+inc.d3_m.inc.d3DEGs <- filteredDEGs %>%
+  filter(comparison == "inc.d3_m.inc.d3") %>%
+  drop_na() %>%
+  pull(gene)
+
+inc.d9_m.inc.d9DEGs <- filteredDEGs %>%
+  filter(comparison == "inc.d9_m.inc.d9") %>%
+  drop_na() %>%
+  pull(gene)
+
+# Generate 3 sets of 200 words
+# Chart
+
+venn.diagram(
+  x = list(hatch_m.n2DEGs, inc.d17_m.inc.d17DEGs, 
+           inc.d3_m.inc.d3DEGs, inc.d9_m.inc.d9DEGs),
+  category.names = c("Inc 3" , "Hatch" , "Inc 9", " Inc 17"),
+  filename = '../figures/venn.png',
+  output=FALSE,
+  col=c("#CDCDCD", '#262625', '#959595','#959595'),
+  fill = c("#CDCDCD", "#262625", "#959595", "#959595"),
+  main = "No. of genes that respond to offspring removal",
+  sub = "9 shared genes: \nPHGDH, NRG2, LHX9, SSTR3, VTN, SARM1, ALK, LPAR1, and APOH" )
+
+intersect(hatch_m.n2DEGs, inc.d17_m.inc.d17DEGs) %>%
+  intersect(.,inc.d3_m.inc.d3DEGs) %>%
+  intersect(.,inc.d9_m.inc.d9DEGs)
+```

### 2020-08-26 analysis/03_datawrangling.Rmd
new female stuff

@@
 ```{r write}
-write.csv(allDEG, "../results/03_allDEG.csv", row.names = F)
-write.csv(candidatevsd, "../results/03_candidatevsd.csv")
-write.csv(wideDEGsSequential, "../results/table1.csv", row.names = F)
+#write.csv(allDEG, "../results/03_allDEG.csv", row.names = F)
+#write.csv(candidatevsd, "../results/03_candidatevsd.csv")
+#write.csv(wideDEGsSequential, "../results/table1.csv", row.names = F)
 
-write.csv(removalDEGs, "../results/removalDEGs.csv")
-write.csv(hatchDEGs, "../results/hatchDEGs.csv")
-write.csv(prolongDEGs, "../results/prolongDEGs.csv")
+#write.csv(removalDEGs, "../results/removalDEGs.csv")
+#write.csv(hatchDEGs, "../results/hatchDEGs.csv")
+#write.csv(prolongDEGs, "../results/prolongDEGs.csv")
 
-write.csv(manipDEGs, "../results/manipDEGs.csv")
+#write.csv(manipDEGs, "../results/manipDEGs.csv")
 
@@
 
-### overlapp
+### overlap
 
@@
 late <- c("hatch_prolong", "inc.d17_prolong")
+
+filteredDEGs <- allDEG %>%
+  filter(lfc > 1.1 | lfc < -1.1 ) %>%
+  filter(tissue == "gonad")
 
@@
   category.names = c("Inc 3" , "Hatch" , "Inc 9", " Inc 17"),
-  filename = '../figures/venn.png',
+  filename = '../figures/venn-gonad.png',
   output=FALSE,
+  print.mode = "raw",
   col=c("#CDCDCD", '#262625', '#959595','#959595'),
   fill = c("#CDCDCD", "#262625", "#959595", "#959595"),
-  main = "No. of genes that respond to offspring removal",
-  sub = "9 shared genes: \nPHGDH, NRG2, LHX9, SSTR3, VTN, SARM1, ALK, LPAR1, and APOH" )
+  main = "Gonads",
+  sub = "zero genes respond to all four offspring removal manipulations"
+  #main = "Pituitary",
+  #sub = "Six genes respond to all four offspring removal manipulations\n SSTR3, VTN, SARM1, ALK, LPAR1, and PHGDH"
+  #main = "Hypothalamus",
+  #sub = "Zero genes respond to all four offspring removal manipulations"
+  #main = "No. of genes that respond to offspring removal",
+  #sub = "9 shared genes: \nPHGDH, NRG2, LHX9, SSTR3, VTN, SARM1, ALK, LPAR1, and APOH"
+  )
 

### 2020-08-26 analysis/03_datawrangling.Rmd
venn diagrams

@@
 
-### overlap
+### removal overlap
 
 ```{r}
-removalcomps <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2")
-early <- c("early_inc.d9", "hatch_early")
-late <- c("hatch_prolong", "inc.d17_prolong")
 
@@
   filter(lfc > 1.1 | lfc < -1.1 ) %>%
-  filter(tissue == "gonad")
+  filter(tissue == "pituitary")
 
-filteredDEGs %>%
-  filter(comparison %in% removalcomps)  %>%
-  select(sex:gene) %>%
-  pivot_wider(names_from = comparison, values_from = direction)
+removalcomps <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2")
 
@@
 
-# Generate 3 sets of 200 words
-# Chart
+intersect(hatch_m.n2DEGs, inc.d17_m.inc.d17DEGs) %>%
+  intersect(.,inc.d3_m.inc.d3DEGs) %>%
+  intersect(.,inc.d9_m.inc.d9DEGs)
 
 venn.diagram(
-  x = list(hatch_m.n2DEGs, inc.d17_m.inc.d17DEGs, 
-           inc.d3_m.inc.d3DEGs, inc.d9_m.inc.d9DEGs),
+  x = list(inc.d3_m.inc.d3DEGs,hatch_m.n2DEGs,
+            inc.d9_m.inc.d9DEGs, inc.d17_m.inc.d17DEGs),
   category.names = c("Inc 3" , "Hatch" , "Inc 9", " Inc 17"),
-  filename = '../figures/venn-gonad.png',
+  filename = '../figures/venn-pit.png',
   output=FALSE,
@@
   fill = c("#CDCDCD", "#262625", "#959595", "#959595"),
-  main = "Gonads",
-  sub = "zero genes respond to all four offspring removal manipulations"
-  #main = "Pituitary",
-  #sub = "Six genes respond to all four offspring removal manipulations\n SSTR3, VTN, SARM1, ALK, LPAR1, and PHGDH"
-  #main = "Hypothalamus",
-  #sub = "Zero genes respond to all four offspring removal manipulations"
-  #main = "No. of genes that respond to offspring removal",
-  #sub = "9 shared genes: \nPHGDH, NRG2, LHX9, SSTR3, VTN, SARM1, ALK, LPAR1, and APOH"
+  main = "Pituitary - Offspring Removal",
+  sub = "Differentially expressed genes relative to internal controls"
   )
 
-intersect(hatch_m.n2DEGs, inc.d17_m.inc.d17DEGs) %>%
-  intersect(.,inc.d3_m.inc.d3DEGs) %>%
-  intersect(.,inc.d9_m.inc.d9DEGs)
 ```
+
+## chicks / nesting care
+
+```{r}
+
+filteredDEGs <- allDEG %>%
+  filter(lfc > 1.1 | lfc < -1.1 ) %>%
+  filter(tissue == "hypothalamus")
+
+chicks <- c("bldg_hatch", "bldg_n5", "bldg_n9",
+            "bldg_extend", "bldg_early")
+
+bldg_hatchDEGs <- filteredDEGs %>%
+  filter(comparison == "bldg_hatch") %>%
+  drop_na() %>%
+  pull(gene)
+
+bldg_n5DEGs <- filteredDEGs %>%
+  filter(comparison == "bldg_n5") %>%
+  drop_na() %>%
+  pull(gene)
+
+bldg_n9DEGs <- filteredDEGs %>%
+  filter(comparison == "bldg_n9") %>%
+  drop_na() %>%
+  pull(gene)
+
+bldg_extendDEGs <- filteredDEGs %>%
+  filter(comparison == "bldg_extend") %>%
+  drop_na() %>%
+  pull(gene)
+
+bldg_earlyDEGs <- filteredDEGs %>%
+  filter(comparison == "bldg_early") %>%
+  drop_na() %>%
+  pull(gene)
+
+# chicks
+venn.diagram(
+  x = list(bldg_n5DEGs, bldg_hatchDEGs, 
+           bldg_earlyDEGs,bldg_extendDEGs,  bldg_n9DEGs),
+  category.names = c( "N5" ,"Hatch" , "Early", " Extend", "N9"),
+  filename = '../figures/venn-chicks-hyp.png',
+  output=FALSE,
+  print.mode = "raw",
+  col=c('#3182bd', "#6baed6",  '#cbc9e2', '#6a51a3', '#08519c'),
+  fill = c("#3182bd", "#6baed6", "#cbc9e2", '#6a51a3', "#08519c" ),
+  main = "Hypothalamus - Nestling Care",
+  sub = "Differentially expressesed genes relative to nest building controls"
+  )
+
+```
+
+## eggs / incubation
+
+```{r}
+
+filteredDEGs <- allDEG %>%
+  filter(lfc > 1.1 | lfc < -1.1 ) %>%
+  filter(tissue == "hypothalamus")
+
+eggs <- c("bldg_lay", 
+         "bldg_inc.d3", "bldg_inc.d9", "bldg_inc.d17",
+         "bldg_prolong")
+
+bldg_layDEGs <- filteredDEGs %>%
+  filter(comparison == "bldg_lay") %>%
+  drop_na() %>%
+  pull(gene)
+
+bldg_inc.d3DEGs <- filteredDEGs %>%
+  filter(comparison == "bldg_inc.d3") %>%
+  drop_na() %>%
+  pull(gene)
+
+bldg_inc.d9DEGs <- filteredDEGs %>%
+  filter(comparison == "bldg_inc.d9") %>%
+  drop_na() %>%
+  pull(gene)
+
+bldg_inc.d17DEGs <- filteredDEGs %>%
+  filter(comparison == "bldg_inc.d17") %>%
+  drop_na() %>%
+  pull(gene)
+
+bldg_prolongDEGs <- filteredDEGs %>%
+  filter(comparison == "bldg_prolong") %>%
+  drop_na() %>%
+  pull(gene)
+
+# chicks
+venn.diagram(
+  x = list(bldg_inc.d3DEGs, bldg_layDEGs,  bldg_prolongDEGs,
+           bldg_inc.d17DEGs, bldg_inc.d9DEGs),
+  category.names = c("Inc3" , "Lay" , "Prolong", " Inc17", "Inc9"),
+  filename = '../figures/venn-eggs-hyp.png',
+  output=FALSE,
+  print.mode = "raw",
+  col=c( '#78c679', "#fed98e",  '#9e9ac8','#006837', '#31a354'),
+  fill = c("#78c679", "#fed98e",  "#9e9ac8",'#006837', "#31a354"),
+  main = "Hypothalamus - Incubation",
+  sub = "Differentially expressesed genes relative to nest building controls"
+  )
+
+```

### 2020-08-26 analysis/03_datawrangling.Rmd
better fig size

@@
   print.mode = "raw",
+  imagetype = "png",
   col=c("#CDCDCD", '#262625', '#959595','#959595'),
@@
   print.mode = "raw",
+  imagetype = "png",
   col=c('#3182bd', "#6baed6",  '#cbc9e2', '#6a51a3', '#08519c'),
@@
   print.mode = "raw",
+  imagetype = "png",
+  height = 1500,
+	width = 1500,
   col=c( '#78c679', "#fed98e",  '#9e9ac8','#006837', '#31a354'),
@@
 ```
+```{r}
+library(cowplot)
+eggshyp <- png::readPNG("../figures/venn-eggs-hyp.png")
+eggshyp <- ggdraw() +  draw_image(eggshyp, scale = 1)
+
+plot_grid(eggshyp,eggshyp, eggshyp, nrow = 1)
+```

### 2020-08-26 analysis/03_datawrangling.Rmd
now with venns in main fig

@@
   filter(lfc > 1.1 | lfc < -1.1 ) %>%
-  filter(tissue == "hypothalamus")
+  filter(tissue == "pituitary")
 
@@
   category.names = c( "N5" ,"Hatch" , "Early", " Extend", "N9"),
-  filename = '../figures/venn-chicks-hyp.png',
+  filename = '../figures/venn-chicks-pit.png',
   output=FALSE,
@@
   imagetype = "png",
+  height = 1500,
+	width = 1750,
+  resolution = 500,
+  cex = 0.75,
+  cat.cex = 0.75,
   col=c('#3182bd', "#6baed6",  '#cbc9e2', '#6a51a3', '#08519c'),
-  fill = c("#3182bd", "#6baed6", "#cbc9e2", '#6a51a3', "#08519c" ),
-  main = "Hypothalamus - Nestling Care",
-  sub = "Differentially expressesed genes relative to nest building controls"
+  fill = c("#3182bd", "#6baed6", "#cbc9e2", '#6a51a3', "#08519c" )
   )
@@
   category.names = c("Inc3" , "Lay" , "Prolong", " Inc17", "Inc9"),
-  filename = '../figures/venn-eggs-hyp.png',
+  filename = '../figures/venn-eggs-pit.png',
   output=FALSE,
@@
   height = 1500,
-	width = 1500,
+	width = 1750,
+  resolution = 500,
+  cex = 0.75,
+  cat.cex = 0.75,
   col=c( '#78c679', "#fed98e",  '#9e9ac8','#006837', '#31a354'),
-  fill = c("#78c679", "#fed98e",  "#9e9ac8",'#006837', "#31a354"),
-  main = "Hypothalamus - Incubation",
-  sub = "Differentially expressesed genes relative to nest building controls"
+  fill = c("#78c679", "#fed98e",  "#9e9ac8",'#006837', "#31a354")
   )
@@
 ```
-```{r}
-library(cowplot)
-eggshyp <- png::readPNG("../figures/venn-eggs-hyp.png")
-eggshyp <- ggdraw() +  draw_image(eggshyp, scale = 1)
-
-plot_grid(eggshyp,eggshyp, eggshyp, nrow = 1)
-```

### 2020-08-26 analysis/figure-2.Rmd
now with venns in main fig

--- 2020-08-24 analysis/figure-2.Rmd
@@
 ```{r fig2, fig.width=7.5, fig.height=7.25}
+g <- plotcandidatechar(hypvsdf, "HTR2C") + labs(subtitle = "Hypothalamus")
+h <- plotcandidatechar(hypvsdm, "HTR2C") + labs(y = NULL, x = "") + labs(subtitle = " ")
+i <- plotremoval(hypvsdf, "HTR2C")+ labs(y = NULL) + labs(subtitle = " ")
+j <- plotremoval(hypvsdm, "HTR2C") + labs(y = NULL, x = "")+ labs(subtitle = " ")
+k <- plotreplacement(hypvsdf, "HTR2C") + labs(y = NULL)+ labs(subtitle = " ")
+l <- plotreplacement(hypvsdm, "HTR2C") + labs(y = NULL, x = "")+ labs(subtitle = " ")
+
 a <- plot.volcano("hypothalamus", sexlevels,  "control_bldg") + 
@@
 
-g <- plotcandidatechar(hypvsdf, "HTR2C") + labs(subtitle = "Hypothalamus")
-h <- plotcandidatechar(hypvsdm, "HTR2C") + labs(y = NULL, x = "") + labs(subtitle = " ")
-i <- plotremoval(hypvsdf, "HTR2C")+ labs(y = NULL) + labs(subtitle = " ")
-j <- plotremoval(hypvsdm, "HTR2C") + labs(y = NULL, x = "")+ labs(subtitle = " ")
-k <- plotreplacement(hypvsdf, "HTR2C") + labs(y = NULL)+ labs(subtitle = " ")
-l <- plotreplacement(hypvsdm, "HTR2C") + labs(y = NULL, x = "")+ labs(subtitle = " ")
+m <- png::readPNG("../figures/venn-eggs-hyp.png")
+m <- ggdraw() +  draw_image(m, scale = 1)
+
+n <- png::readPNG("../figures/venn-chicks-hyp.png")
+n <- ggdraw() +  draw_image(n, scale = 1)
+
+ghi <- plot_grid(g,h,k,l,nrow = 1,
+                labels = c("A", "", "B"), label_size = 8, hjust = 0,
+                rel_widths = c(9,9,6,6))
 
 ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
-                labels = c("D", "E"), label_size = 8, hjust = 0)
+                labels = c("C", "D"), label_size = 8, hjust = 0)
+
 cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h",
-                labels = c("F", "G"), label_size = 8, hjust = 0)
-ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h",
-                labels = c("H", "I"), label_size = 8, hjust = 0)
+                labels = c("E", "F"), label_size = 8, hjust = 0)
 
-ghi <- plot_grid(g,h,i, j,k,l,nrow = 1,
-                labels = c("A", "", "B", "", "C"), label_size = 8, hjust = 0,
-                rel_widths = c(9,9,8,8,6,6))
+ef <- plot_grid(f,m,n, rel_widths = c(2,1,1), nrow =1,
+                labels = c("G", "H", "I"), label_size = 8, hjust = 0)
 
@@
 fig2
+
 ```
@@
   labs(x = "Offspring removal versus temporal control")
-f <- makebargraphv3(DEGreplace, "pituitary", NULL,  comparisonlevelsreplace) +
+f <- makebargraphv3(DEGreplace, "pituitary", NULL,  comparisonlablelssreplace) +
   labs(x = "Offspring removal versus temporal or external control")
@@
 
+m <- png::readPNG("../figures/venn-eggs-pit.png")
+m <- ggdraw() +  draw_image(m, scale = 1)
+
+n <- png::readPNG("../figures/venn-chicks-pit.png")
+n <- ggdraw() +  draw_image(n, scale = 1)
+
+ghi <- plot_grid(g,h,k,l,nrow = 1,
+                labels = c("A", "", "B"), label_size = 8, hjust = 0,
+                rel_widths = c(9,9,6,6))
+
 ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
-                labels = c("D", "E"), label_size = 8, hjust = 0)
+                labels = c("C", "D"), label_size = 8, hjust = 0)
+
 cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h",
-                labels = c("F", "G"), label_size = 8, hjust = 0)
-ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h",
-                labels = c("H", "I"), label_size = 8, hjust = 0)
+                labels = c("E", "F"), label_size = 8, hjust = 0)
 
-ghi <- plot_grid(g,h,i, j,k,l,nrow = 1,
-                labels = c("A", "", "B", "", "C"), label_size = 8, hjust = 0,
-                rel_widths = c(9,9,8,8,6,6))
+ef <- plot_grid(f,m,n, rel_widths = c(2,1,1), nrow =1,
+                labels = c("G", "H", "I"), label_size = 8, hjust = 0)
 
@@
   labs(x = "Offspring removal versus temporal control")
-f <- makebargraphv3(DEGreplace, "gonad", NULL,  comparisonlevelsreplace) +
+f <- makebargraphv3(DEGreplace, "gonad", NULL,  comparisonlablelssreplace) +
   labs(x = "Offspring removal versus temporal or external control")
@@
 
+m <- png::readPNG("../figures/venn-eggs-gon.png")
+m <- ggdraw() +  draw_image(m, scale = 1)
+
+n <- png::readPNG("../figures/venn-chicks-gon.png")
+n <- ggdraw() +  draw_image(n, scale = 1)
+
+ghi <- plot_grid(g,h,k,l,nrow = 1,
+                labels = c("A", "", "B"), label_size = 8, hjust = 0,
+                rel_widths = c(9,9,6,6))
+
 ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
-                labels = c("D", "E"), label_size = 8, hjust = 0)
+                labels = c("C", "D"), label_size = 8, hjust = 0)
+
 cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h",
-                labels = c("F", "G"), label_size = 8, hjust = 0)
-ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h",
-                labels = c("H", "I"), label_size = 8, hjust = 0)
+                labels = c("E", "F"), label_size = 8, hjust = 0)
 
-ghi <- plot_grid(g,h,i, j,k,l,nrow = 1,
-                labels = c("A", "", "B", "", "C"), label_size = 8, hjust = 0,
-                rel_widths = c(9,9,8,8,6,6))
+ef <- plot_grid(f,m,n, rel_widths = c(2,1,1), nrow =1,
+                labels = c("G", "H", "I"), label_size = 8, hjust = 0)
 
@@
 dev.off()
-
-sessionInfo()
 ```

### 2020-08-26 R/themes.R

@@ -90,6 +90,11 @@ comparisonlevelsremoval <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.in
 comparisonlevelsreplace <- c("hatch_early", "inc.d17_prolong",  "hatch_prolong", 
                              "extend_hatch", "n5_extend")
 
+comparisonlablelssreplace <- c("hatch\nearly", "inc.d17\nprolong",  "hatch\nprolong", 
+                             "extend\nhatch", "n5\nextend")
+
+
+
 comparisonlevelsmanip <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9",  "early_inc.d9" ,
                            "hatch_early",  "inc.d17_m.inc.d17", 
                            "hatch_prolong", "hatch_m.n2",

### 2020-08-26 analysis/figure-2.Rmd
not working, trying to add more comparisons with labels

@@
 DEGreplace
+
+DEGcontrolreplace <- allDEG %>% 
+  filter(comparison %in% comparisonlevelscontrolreplace)  %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelscontrolreplace)) %>%
+  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
+  group_by(sex, tissue, comparison, direction, label, updown) %>%
+  summarise(n = n()) %>%
+  mutate(n = n*updown ) %>%
+  select(-updown)
+
 ```
@@
 
-b <- makebargraphv3(DEGcontrol, "hypothalamus","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
+b <- makebargraphv3(DEGcontrolreplace, "hypothalamus","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrolreaplce) +
   labs(x = "Control versus all other reproductive and parental stages") +
@@
 
-ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
+ab <- plot_grid(a,b,rel_widths = c(1,3), 
                 labels = c("C", "D"), label_size = 8, hjust = 0)
@@
 
-b <- makebargraphv3(DEGcontrol, "pituitary","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
+b <- makebargraphv3(DEGcontrolreplace, "pituitary","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrolreaplce) +
   labs(x = "Control versus all other reproductive and parental stages")
@@
 
-b <- makebargraphv3(DEGcontrol, "gonad","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
+b <- makebargraphv3(DEGcontrolreplace, "gonad","No. of DEGs\n(-) decreased  increased (+)", comparisonlevelscontrolreplace, comparisonlevelscontrolreplace) +
   labs(x = "Control versus all other reproductive and parental stages")
 
-c <- makebargraphv3(DEGbldg, "gonad", "No. of DEGs\n (-)decreased  increased (+)", comparisonlabelsbldg) +
+c <- makebargraphv3(DEGbldg, "gonad", "No. of DEGs\n (-)decreased  increased (+)", comparisonlevelsbldg, comparisonlabelsbldg) +
   labs(x = "Nest-building versus all other parental stages")
-d <- makebargraphv3(DEGchar, "gonad", NULL,  comparisonlabelscharnobldg) +
+d <- makebargraphv3(DEGchar, "gonad", NULL,  comparisonlevelscharnobldg, comparisonlabelscharnobldg) +
   labs(x = "Comparison of sequential parental stages")
 
-e <- makebargraphv3(DEGremove, "gonad", "No. of DEGs\n (-)decreased  increased (+)", comparisonlevelsremoval) +
+e <- makebargraphv3(DEGremove, "gonad", "No. of DEGs\n (-)decreased  increased (+)", comparisonlevelsremoval, comparisonlevelsremoval) +
   labs(x = "Offspring removal versus temporal control")
-f <- makebargraphv3(DEGreplace, "gonad", NULL,  comparisonlablelssreplace) +
+f <- makebargraphv3(DEGreplace, "gonad", NULL,  comparisonlevelsreplace, comparisonlablelssreplace) +
   labs(x = "Offspring removal versus temporal or external control")

### 2020-08-27 analysis/03_datawrangling.Rmd
now with sequential comparisons

@@
   dplyr::rename("gene" = "X1") %>% 
-  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
+  #pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
+               #names_to = "samples", values_to = "counts") 
+  pivot_longer(cols = L.G118_female_hypothalamus_control.NYNO:y98.g54_female_hypothalamus_m.hatch, 
                names_to = "samples", values_to = "counts") 
@@
 
+tail((allvsd))
 ```
@@
 wideDEGsSequential <- wideDEGscandidates %>%
-  select(sextissue,gene, "control_bldg", "bldg_lay", "lay_inc.d3", "inc.d3_inc.d9",
-         "inc.d9_inc.d17", "inc.d17_hatch", "hatch_n5", "n5_n9") %>%
+  select(sextissue,gene, "control_bldg", "bldg_lay" #, 
+         #"lay_inc.d3", "inc.d3_inc.d9",
+         #"inc.d9_inc.d17", "inc.d17_hatch", "hatch_n5", "n5_n9"
+         ) %>%
   mutate(totalNA = rowSums(is.na(.))) %>%
@@
   arrange(totalNA, gene, sextissue) %>%
-  #arrange( gene, sextissue) %>%
+  arrange( gene, sextissue) %>%
   select(-totalNA) %>% 
-  replace(., is.na(.), "")
+  replace(., is.na(.), "NS")
 wideDEGsSequential
@@
 filteredDEGs <- allDEG %>%
-  filter(lfc > 1.1 | lfc < -1.1 )
+  filter(lfc > 0.14 | lfc < -0.14 )
 dim(allDEG)
@@
 filteredDEGs <- allDEG %>%
-  filter(lfc > 1.1 | lfc < -1.1 ) %>%
-  filter(tissue == "pituitary")
+  filter(lfc > 0.14 | lfc < -0.14 ) %>%
+  filter(tissue == "hypothalamus")
 
@@
   category.names = c("Inc 3" , "Hatch" , "Inc 9", " Inc 17"),
-  filename = '../figures/venn-pit.png',
+  filename = '../figures/venn-hyp.png',
   output=FALSE,
@@
   fill = c("#CDCDCD", "#262625", "#959595", "#959595"),
-  main = "Pituitary - Offspring Removal",
+  main = "Hypothalamus - Offspring Removal",
   sub = "Differentially expressed genes relative to internal controls"
@@
   filter(lfc > 1.1 | lfc < -1.1 ) %>%
-  filter(tissue == "pituitary")
+  filter(tissue == "hypothalamus")
 

### 2020-08-27 R/functions.R

@@ -126,7 +126,7 @@ createDEGdftreatmentvcontrols <- function(whichdds, whichgroup, downs, ups){
           # write DEGsframe of only significant genes
           DEGs <- DEGs %>% dplyr::filter(direction != "NS")
           
-          top6 <- DEGs %>% head(.) %>% pull(gene) %>% droplevels()
+          top6 <- DEGs %>% head(.) %>% pull(gene) 
           print(top6)
           
           # return DEGs frome with all data, included NS genes

### 2020-08-28 analysis/03_datawrangling.Rmd
working on better tables

@@
 head(parentalcaregenes)
-
 ```
@@
                #names_to = "samples", values_to = "counts") 
-  pivot_longer(cols = L.G118_female_hypothalamus_control.NYNO:y98.g54_female_hypothalamus_m.hatch, 
+  pivot_longer(cols = L.Blu13_male_hypothalamus_control.NYNO:y98.o50.x_male_hypothalamus_inc.d3, 
                names_to = "samples", values_to = "counts") 
@@
 
-tail((allvsd))
+tail(names(allvsd))
 ```
@@
 
-wideDEGs <- tempDEGs %>%
-  mutate(lfc = round(lfc, 2),
-         padj =  round(padj,2)) %>%
-  mutate(sextissue = paste(sex, tissue, sep = "_"),
-         lfcpadj = paste(lfc, padj, sep = ", "))  %>%
-  dplyr::select(sextissue,comparison, gene, lfcpadj) %>%
-  pivot_wider(names_from = comparison, values_from = lfcpadj)
-
-wideDEGscandidates <- wideDEGs %>%
-  filter(gene %in% GOgenesLong$gene)
+# sequential in 2 or more comparisons
+wideDEGscandidates <- tempDEGs %>%
+  filter(gene %in% candidategenes) %>%
+  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
+         sex = recode(sex, "female" = "F", "male" = "M" ),
+         tissue = recode(tissue, 
+                         "hypothalamus" = "H",
+                         "pituitary" = "P", "gonad" = "G"),
+         group = paste(sex, tissue, sep = "")) %>%
+  mutate(res = paste("(", posneg, ")", sep = "")) %>%
+  mutate(compres = paste(group, res, sep = "")) %>%
+  group_by(gene, comparison) %>%
+  summarize(results = str_c(compres, collapse = "; ")) %>%
+  pivot_wider(names_from = comparison, values_from = results) 
 wideDEGscandidates
 
-wideDEGsBldg <- wideDEGscandidates %>%
-  select(sextissue,gene, contains("bldg"))
-wideDEGsBldg
-
-wideDEGsControl <- wideDEGscandidates %>%
-  select(sextissue,gene, contains("control"))
-wideDEGsControl
-
-# sequential in 2 or more comparisons
-wideDEGsSequential <- wideDEGscandidates %>%
-  select(sextissue,gene, "control_bldg", "bldg_lay" #, 
-         #"lay_inc.d3", "inc.d3_inc.d9",
-         #"inc.d9_inc.d17", "inc.d17_hatch", "hatch_n5", "n5_n9"
-         ) %>%
-  mutate(totalNA = rowSums(is.na(.))) %>%
-  filter(totalNA < 8) %>%
-  arrange(totalNA, gene, sextissue) %>%
-  arrange( gene, sextissue) %>%
-  select(-totalNA) %>% 
-  replace(., is.na(.), "NS")
-wideDEGsSequential
-
+unique(wideDEGscandidates$compres)
 ```
@@
 removalcomps <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2")
-early <- c("early_inc.d9", "hatch_early")
-late <- c("hatch_prolong", "inc.d17_prolong")
+earlys <- c("inc.d9_early", "inc.d17_early", "hatch_early", "n5_early")
+prolongs <- c("inc.d9_prolong","inc.d17_prolong", "hatch_prolong",  "n5_prolong")
+extends <- c("inc.d9_extend","inc.d17_extend", "hatch_extend",  "n5_extend")
 
 removalDEGs <- makemanipulationDEGtables(removalcomps)
-hatchDEGs <- makemanipulationDEGtables(early)
-prolongDEGs <- makemanipulationDEGtables(late)
+earlyDEGs <- makemanipulationDEGtables(earlys)
+prolongDEGs <- makemanipulationDEGtables(prolongs)
+extendDEGs <- makemanipulationDEGtables(extends)
 
@@
 ```{r write}
-#write.csv(allDEG, "../results/03_allDEG.csv", row.names = F)
-#write.csv(candidatevsd, "../results/03_candidatevsd.csv")
-#write.csv(wideDEGsSequential, "../results/table1.csv", row.names = F)
-
-#write.csv(removalDEGs, "../results/removalDEGs.csv")
-#write.csv(hatchDEGs, "../results/hatchDEGs.csv")
-#write.csv(prolongDEGs, "../results/prolongDEGs.csv")
-
-#write.csv(manipDEGs, "../results/manipDEGs.csv")
-
+write.csv(allDEG, "../results/03_allDEG.csv", row.names = F)
+write.csv(candidatevsd, "../results/03_candidatevsd.csv")
+write.csv(wideDEGsSequential, "../results/table1.csv", row.names = F)
+write.csv(manipDEGs, "../results/manipDEGs.csv")
 ```

### 2020-08-28 analysis/03_datawrangling.Rmd
new DEGs

@@
   dplyr::rename("gene" = "X1") %>% 
-  #pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
-               #names_to = "samples", values_to = "counts") 
-  pivot_longer(cols = L.Blu13_male_hypothalamus_control.NYNO:y98.o50.x_male_hypothalamus_inc.d3, 
+  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
                names_to = "samples", values_to = "counts") 
@@
 
-manipDEGs <- rbind(removalDEGs, hatchDEGs) %>%
+manipDEGs <- rbind(removalDEGs, earlyDEGs) %>%
   rbind(., prolongDEGs) %>%
+  rbind(., extendDEGs) %>%
   arrange(desc(n))
@@
 write.csv(candidatevsd, "../results/03_candidatevsd.csv")
-write.csv(wideDEGsSequential, "../results/table1.csv", row.names = F)
+#write.csv(wideDEGsSequential, "../results/table1.csv", row.names = F)
 write.csv(manipDEGs, "../results/manipDEGs.csv")
+write.csv(wideDEGscandidates, "../results/03_wideDEGscandidates.csv", row.names = F)
+
 ```
@@
 
-```{r}
+```{r venn-rm}
 
@@
 
-```{r}
+```{r venn-chicks}
 
@@
 
-```{r}
+```{r venn-eggs}
 

### 2020-08-28 analysis/figure-2.Rmd
reorg, move data crunk out of figure 2

@@
 knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
-```
-
-## Treatment specific DEGs
-
-```{r suppltables}
-allDEG <- read_csv("../results/03_allDEG.csv") %>%
-  mutate(tissue = factor(tissue, levels = tissuelevel),
-         direction = factor(direction, levels = alllevels),
-         label = gsub("_","\nvs. ", comparison)) %>%
-  filter(lfc > 1.1 | lfc < -1.1 )
-head(allDEG)
-
-# for suppl figures
-DEGcontrol <- allDEG %>% 
-  filter(grepl("control", comparison),
-         !grepl("m.|early|extend|prolong", comparison))  %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol)) %>%
-  group_by(sex, tissue, comparison, direction, label) %>%
-    summarise(n = n()) %>%
-    mutate(n = ifelse(direction == "control", n*-1, n*1 ))
-
-DEGbldg <- allDEG %>% 
-  filter(grepl("bldg", comparison),
-         !grepl("m.|early|extend|prolong|control", comparison))  %>%
-  drop_na() %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))  %>%
-  group_by(sex, tissue, comparison, direction, label) %>%
-    summarise(n = n()) %>%
-    mutate(n = ifelse(direction == "bldg", n*-1, n*1 ))
-
-DEGchar <- allDEG %>% 
-  filter(comparison %in% comparisonlevelschar,
-         !grepl("control|bldg", comparison)) %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelschar)) %>%
-  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
-  group_by(sex, tissue, comparison, direction, label, updown) %>%
-  summarise(n = n()) %>%
-  mutate(n = n*updown ) %>%
-  select(-updown)
-
-DEGremove <- allDEG %>% 
-  filter(comparison %in% comparisonlevelsremoval) %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelsremoval)) %>%
-  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
-  group_by(sex, tissue, comparison, direction, label, updown) %>%
-  summarise(n = n()) %>%
-  mutate(n = n*updown ) %>%
-  select(-updown)
-DEGremove
-
-DEGreplace <- allDEG %>% 
-  filter(comparison %in% comparisonlevelsreplace) %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelsreplace)) %>%
-  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
-  group_by(sex, tissue, comparison, direction, label, updown) %>%
-  summarise(n = n()) %>%
-  mutate(n = n*updown ) %>%
-  select(-updown)
-DEGreplace
-
-DEGcontrolreplace <- allDEG %>% 
-  filter(comparison %in% comparisonlevelscontrolreplace)  %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelscontrolreplace)) %>%
-  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
-  group_by(sex, tissue, comparison, direction, label, updown) %>%
-  summarise(n = n()) %>%
-  mutate(n = n*updown ) %>%
-  select(-updown)
-
-```
-
-## Variance stabilized data
-
-```{r data}
-countData <- read_csv("../results/01_limma.csv") %>%
-  column_to_rownames(var = "X1")
-countData <- as.data.frame(t(countData))
-head(countData[1:3])
-
-colData <- read_csv("../metadata/00_colData.csv") %>%
-  mutate(treatment = factor(treatment, levels = alllevels),
-         tissue = factor(tissue, levels = tissuelevels)) %>% 
-  column_to_rownames(var = "X1") 
-head(colData)
-# check ready for analysis
-# row.names(countData) == row.names(colData)
-head(row.names(countData) == row.names(colData))
-
-vsd_path <- "../results/DEseq2/treatment/"   # path to the data
-vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
-vsd_pathfiles <- paste0(vsd_path, vsd_files)
-vsd_files
-
-allvsd <- vsd_pathfiles %>%
-  setNames(nm = .) %>% 
-  map_df(~read_csv(.x), .id = "file_name")  %>% 
-  dplyr::rename("gene" = "X1") 
-## before pivoting, check names of df with
-## head(names(allvsd)) and tail(names(allvsd))
-allvsd <- allvsd %>% 
-  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
-               names_to = "samples", values_to = "counts")  %>% 
-    drop_na() 
-allvsd <- as_tibble(allvsd)
-head(allvsd[2:4])
-tail(allvsd[2:4])
-
-# for plotting gene expression over time
-getcandidatevsdmanip <- function(whichgenes, whichtissue){
-
-  candidates  <- allvsd %>%
-    filter(gene %in% whichgenes) %>%
-    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1),
-                  sextissue = sapply(strsplit(sextissue, '../results/DEseq2/treatment/'), "[", 2),
-                  sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
-                sex = factor(sex),
-                  tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
-                treatment = sapply(strsplit(samples, '\\_'), "[", 4),
-                treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1),
-                treatment = recode(treatment, "m.hatch" = "m.n2", "extend.hatch" = "m.n2",
-                                                   "inc.prolong" = "prolong", "m.inc.d8" = "early"),
-                treatment =  factor(treatment, levels = alllevels),
-                earlylate = fct_collapse(treatment, 
-                                  early = c("early", "m.inc.d3", "m.inc.d9",
-                                           "lay", "inc.d3", "inc.d9" ),
-
-                                  late = c("inc.d17", "m.inc.d17", "prolong" ,  "hatch" , 
-                                         "m.n2", "extend", "n5", "n9"),
-                                  reference = c("control", "bldg")),
-                extint = fct_collapse(treatment, 
-                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17"),
-                                  eggs = c("lay","inc.d3", "inc.d9", "inc.d17", "prolong"),
-                                  chicks = c("early", "hatch", "extend", "n5", "n9"),
-                                  reference = c("control", "bldg")))  %>%
-    dplyr::select(gene, counts, sex, tissue, treatment, earlylate, extint, samples) 
-  return(candidates)
-}
-
-candidatevsd <- getcandidatevsdmanip(candidategenes)
-
-hypvsd <- candidatevsd %>% filter(tissue %in% "hypothalamus")  
-pitvsd <- candidatevsd %>% filter(tissue %in% "pituitary") 
-gonvsd <- candidatevsd %>% filter(tissue %in% "gonad")
-
-hypvsdf <- candidatevsd %>% filter(tissue %in% "hypothalamus", sex == "female")  
-pitvsdf <- candidatevsd %>% filter(tissue %in% "pituitary", sex == "female")
-gonvsdf <- candidatevsd %>% filter(tissue %in% "gonad", sex == "female")  
-hypvsdm <- candidatevsd %>% filter(tissue %in% "hypothalamus", sex == "male")  
-pitvsdm <- candidatevsd %>% filter(tissue %in% "pituitary", sex == "male") 
-gonvsdm <- candidatevsd %>% filter(tissue %in% "gonad", sex == "male")
-
-summary(hypvsd)
-summary(pitvsd)
 ```

### 2020-08-28 analysis/03_DEGs.R

@@ -1,0 +1,184 @@
+library(tidyverse)
+library(readr)
+library(VennDiagram)
+
+source("../R/themes.R")
+source("../R/genelists.R")
+
+## All DEGs 
+
+DEG_path <- "../results/DEseq2/treatment/"   # path to the data
+DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
+DEG_pathfiles <- paste0(DEG_path, DEG_files)
+#DEG_files
+
+allDEG <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
+  mutate(comparison = paste(down,up, sep = "_")) %>%
+  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
+  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
+  dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
+head(allDEG)
+
+unique(allDEG$comparison)
+
+tempDEGs <- DEG_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name") %>% 
+  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
+  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
+  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
+  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
+  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
[…]

### 2020-08-28 analysis/03_vsd.R

@@ -1,0 +1,78 @@
+library(tidyverse)
+library(readr)
+library(VennDiagram)
+
+source("../R/themes.R")
+source("../R/genelists.R")
+
+### variance stabilized gene expression  (vsd) 
+
+
+vsd_path <- "../results/DEseq2/treatment/"   # path to the data
+vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
+vsd_pathfiles <- paste0(vsd_path, vsd_files)
+vsd_files
+
+allvsd <- vsd_pathfiles %>%
+  setNames(nm = .) %>% 
+  map_df(~read_csv(.x), .id = "file_name")  %>% 
+  dplyr::rename("gene" = "X1") 
+
+## before pivoting, check names of df with
+## head(names(allvsd)) and tail(names(allvsd))
+
+allvsd <- allvsd %>% 
+  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
+               names_to = "samples", values_to = "counts")  %>% 
+  drop_na() 
+allvsd <- as_tibble(allvsd)
+head(allvsd[2:4])
+tail(allvsd[2:4])
+
+# for plotting gene expression over time
+getcandidatevsdmanip <- function(whichgenes, whichtissue){
+  
+  candidates  <- allvsd %>%
+    filter(gene %in% whichgenes) %>%
+    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1),
+                  sextissue = sapply(strsplit(sextissue, '../results/DEseq2/treatment/'), "[", 2),
+                  sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
+                  sex = factor(sex),
+                  tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
+                  treatment = sapply(strsplit(samples, '\\_'), "[", 4),
+                  treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1),
+                  treatment = recode(treatment, "m.hatch" = "m.n2", "extend.hatch" = "m.n2",
+                                     "inc.prolong" = "prolong", "m.inc.d8" = "early"),
+                  treatment =  factor(treatment, levels = alllevels),
+                  earlylate = fct_collapse(treatment, 
+                                           early = c("early", "m.inc.d3", "m.inc.d9",
+                                                     "lay", "inc.d3", "inc.d9" ),
+                                           
[…]

### 2020-08-28 analysis/venn.R

@@ -1,0 +1,169 @@
+
+
+### removal overlap
+
+```{r venn-rm}
+
+filteredDEGs <- allDEG %>%
+  filter(lfc > 0.14 | lfc < -0.14 ) %>%
+  filter(tissue == "hypothalamus")
+
+removalcomps <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2")
+
+hatch_m.n2DEGs <- filteredDEGs %>%
+  filter(comparison == "hatch_m.n2") %>%
+  drop_na() %>%
+  pull(gene)
+
+inc.d17_m.inc.d17DEGs <- filteredDEGs %>%
+  filter(comparison == "inc.d17_m.inc.d17") %>%
+  drop_na() %>%
+  pull(gene)
+
+inc.d3_m.inc.d3DEGs <- filteredDEGs %>%
+  filter(comparison == "inc.d3_m.inc.d3") %>%
+  drop_na() %>%
+  pull(gene)
+
+inc.d9_m.inc.d9DEGs <- filteredDEGs %>%
+  filter(comparison == "inc.d9_m.inc.d9") %>%
+  drop_na() %>%
+  pull(gene)
+
+
+intersect(hatch_m.n2DEGs, inc.d17_m.inc.d17DEGs) %>%
+  intersect(.,inc.d3_m.inc.d3DEGs) %>%
+  intersect(.,inc.d9_m.inc.d9DEGs)
+
+venn.diagram(
+  x = list(inc.d3_m.inc.d3DEGs,hatch_m.n2DEGs,
+           inc.d9_m.inc.d9DEGs, inc.d17_m.inc.d17DEGs),
+  category.names = c("Inc 3" , "Hatch" , "Inc 9", " Inc 17"),
+  filename = '../figures/venn-hyp.png',
+  output=FALSE,
+  print.mode = "raw",
+  imagetype = "png",
+  col=c("#CDCDCD", '#262625', '#959595','#959595'),
+  fill = c("#CDCDCD", "#262625", "#959595", "#959595"),
+  main = "Hypothalamus - Offspring Removal",
+  sub = "Differentially expressed genes relative to internal controls"
+)
+
+
+```
+
+## chicks / nesting care
+
+
+```{r venn-chicks}
+
+filteredDEGs <- allDEG %>%
+  filter(lfc > 1.1 | lfc < -1.1 ) %>%
+  filter(tissue == "hypothalamus")
[…]

### 2020-08-28 analysis/boneyard.Rmd
snake make working now

@@
 ```
+
+```{r}
+# from DEGs
+
+# for suppl figures
+DEGcontrol <- allDEG %>% 
+  filter(grepl("control", comparison),
+         !grepl("m.|early|extend|prolong", comparison))  %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol)) %>%
+  group_by(sex, tissue, comparison, direction, label) %>%
+  summarise(n = n()) %>%
+  mutate(n = ifelse(direction == "control", n*-1, n*1 ))
+
+DEGbldg <- allDEG %>% 
+  filter(grepl("bldg", comparison),
+         !grepl("m.|early|extend|prolong|control", comparison))  %>%
+  drop_na() %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))  %>%
+  group_by(sex, tissue, comparison, direction, label) %>%
+  summarise(n = n()) %>%
+  mutate(n = ifelse(direction == "bldg", n*-1, n*1 ))
+
+DEGchar <- allDEG %>% 
+  filter(comparison %in% comparisonlevelschar,
+         !grepl("control|bldg", comparison)) %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelschar)) %>%
+  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
+  group_by(sex, tissue, comparison, direction, label, updown) %>%
+  summarise(n = n()) %>%
+  mutate(n = n*updown ) %>%
+  select(-updown)
+
+DEGremove <- allDEG %>% 
+  filter(comparison %in% comparisonlevelsremoval) %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelsremoval)) %>%
+  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
+  group_by(sex, tissue, comparison, direction, label, updown) %>%
+  summarise(n = n()) %>%
+  mutate(n = n*updown ) %>%
+  select(-updown)
+DEGremove
+
+DEGreplace <- allDEG %>% 
+  filter(comparison %in% comparisonlevelsreplace) %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelsreplace)) %>%
+  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
+  group_by(sex, tissue, comparison, direction, label, updown) %>%
+  summarise(n = n()) %>%
+  mutate(n = n*updown ) %>%
+  select(-updown)
+DEGreplace
+
+DEGcontrolreplace <- allDEG %>% 
+  filter(comparison %in% comparisonlevelscontrolreplace)  %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelscontrolreplace)) %>%
+  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
+  group_by(sex, tissue, comparison, direction, label, updown) %>%
+  summarise(n = n()) %>%
+  mutate(n = n*updown ) %>%
+  select(-updown)
+```

### 2020-08-28 analysis/03_DEGs.R

@@ -1,9 +1,8 @@
 library(tidyverse)
 library(readr)
-library(VennDiagram)
 
-source("../R/themes.R")
-source("../R/genelists.R")
+source("R/themes.R")
+source("R/genelists.R")
 
 ## All DEGs 
 
@@ -114,71 +113,6 @@ manipDEGs <- rbind(removalDEGs, earlyDEGs) %>%
 head(manipDEGs)
 
 
-# for suppl figures
-DEGcontrol <- allDEG %>% 
-  filter(grepl("control", comparison),
-         !grepl("m.|early|extend|prolong", comparison))  %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol)) %>%
-  group_by(sex, tissue, comparison, direction, label) %>%
-  summarise(n = n()) %>%
-  mutate(n = ifelse(direction == "control", n*-1, n*1 ))
-
-
-DEGbldg <- allDEG %>% 
-  filter(grepl("bldg", comparison),
-         !grepl("m.|early|extend|prolong|control", comparison))  %>%
-  drop_na() %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))  %>%
-  group_by(sex, tissue, comparison, direction, label) %>%
-  summarise(n = n()) %>%
-  mutate(n = ifelse(direction == "bldg", n*-1, n*1 ))
-
-DEGchar <- allDEG %>% 
-  filter(comparison %in% comparisonlevelschar,
-         !grepl("control|bldg", comparison)) %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelschar)) %>%
-  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
-  group_by(sex, tissue, comparison, direction, label, updown) %>%
-  summarise(n = n()) %>%
-  mutate(n = n*updown ) %>%
-  select(-updown)
-
-DEGremove <- allDEG %>% 
-  filter(comparison %in% comparisonlevelsremoval) %>%
-  mutate(comparison = factor(comparison, levels = comparisonlevelsremoval)) %>%
-  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
-  group_by(sex, tissue, comparison, direction, label, updown) %>%
-  summarise(n = n()) %>%
-  mutate(n = n*updown ) %>%
-  select(-updown)
-DEGremove
-
-DEGreplace <- allDEG %>% 
[…]

### 2020-08-28 analysis/boneyard.Rmd
move unrelated things

--- 2020-08-21 analysis/boneyard.Rmd
@@
 bc <- plot_grid(b1,b2,b3,b4, c1,c2,c3,c4, nrow = 1, rel_widths = c(1.5,1.1,0.9,0.9,1.5,1.1,0.9,0.9),
-                labels = c("A", "", "", "", "B"), label_size = 8)
+                labels = c("A", "", "", "", "B")_size = 8)
 ```
@@
           rel_heights = c(1,0.9,1.3),
-          labels = c("A", "B"), label_size = 8, rel_widths = c(1.1,1))
+          labels = c("A", "B")_size = 8, rel_widths = c(1.1,1))
 
@@
           rel_widths = c(2,1), rel_heights = c(1.1,1,1.1), ncol = 2,
-          labels = c("A", "B"), label_size = 8)
+          labels = c("A", "B")_size = 8)
 supplfig2
@@
 b69 <- plot_grid(b6,b7,b8,b9, align = "v" , rel_heights = c(1.2,1))
-b <- plot_grid(b15,b69, labels = c("C"), label_size = 8, rel_widths = c(1.2,1))
+b <- plot_grid(b15,b69s = c("C")_size = 8, rel_widths = c(1.2,1))
 
@@
 c3 <- plotcorrplot(cancergenes, NULL) + theme(legend.position = "right") + labs(subtitle = "Breast, ovary, & piuitary cancer genes", color = "Correlation")
-c <- plot_grid(c1,c2,c3, nrow = 1, labels = c("D"), label_size = 8, rel_widths = c(1,1,1.3))
+c <- plot_grid(c1,c2,c3, nrow = 1s = c("D")_size = 8, rel_widths = c(1,1,1.3))
 
@@
           b3,b6,b9,b12, rel_heights = c(1.4,1,1.2),
-          labels = c("A", " ", "B"), label_size = 8, rel_widths = c(1.4,1.2,1.2,1))
+          labels = c("A", " ", "B")_size = 8, rel_widths = c(1.4,1.2,1.2,1))
 
@@
   fig2alt <- plot_grid(hypplots, pitplots, gonplots, nrow = 3,
-                       labels = "AUTO", label_size = 8,
+                       labels = "AUTO"_size = 8,
                        rel_heights = c(1,1,1.2))
@@
 ab <- plot_grid(b,a,rel_widths = c(2.5,1), 
-                labels = c("A", "B"), label_size = 8, hjust = 0)
+                labels = c("A", "B")_size = 8, hjust = 0)
 cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h",
-                labels = c("C", "D"), label_size = 8, hjust = 0)
+                labels = c("C", "D")_size = 8, hjust = 0)
 ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h")

### 2020-08-28 analysis/venn.R

@@ -166,4 +166,12 @@ venn.diagram(
   fill = c("#78c679", "#fed98e",  "#9e9ac8",'#006837', "#31a354")
 )
 
+
+m <- png::readPNG("figures/venn-eggs-hyp.png")
+m <- ggdraw() +  draw_image(m, scale = 1)
+
+n <- png::readPNG("figures/venn-chicks-hyp.png")
+n <- ggdraw() +  draw_image(n, scale = 1)
+
+
 ```

### 2020-08-28 analysis/boneyard/04_DESeq2.Rmd
cleaner repo

--- 2020-05-13 analysis/04_DESeq2.Rmd
@@
 
-```{r hypothesis}
+```{r hypothesis,  message = F, warning = F}
 
@@
   DEGs <- DEGs %>% dplyr::filter(direction != "NS")
-  print(str(DEGs))
+  #print(str(DEGs))
 
@@
 
-  createDEGdfhypothesis("hiloPRL", "hi", "lo", i)
-  createDEGdfhypothesis("external", "chicks", "eggs", i)
-  createDEGdfhypothesis("external", "chicks", "nest", i)
-  createDEGdfhypothesis("external", "eggs", "nest", i)
+  #createDEGdfhypothesis("hiloPRL", "hi", "lo", i)
+  #createDEGdfhypothesis("external", "chicks", "eggs", i)
+  #createDEGdfhypothesis("external", "chicks", "nest", i)
+  #createDEGdfhypothesis("external", "eggs", "nest", i)
 
 }
+
 ```
+
+```{r}
+write.csv(colData, "../metadata/04_colData.csv", row.names = F)
+```

### 2020-08-28 analysis/01_limma.R

@@ -1,55 +1,0 @@
-# Limma
-
-# DESeq2 is _not_ recommended for experiments with more than 100 samples ([see Mike Love's post](https://mikelove.wordpress.com/2016/09/28/deseq2-or-edger/)), so I decided to try the limma package. I followed [this tutorial](https://www.bioconductor.org/packages/devel/workflows/vignettes/RNAseq123/inst/doc/limmaWorkflow.html).
-
-library(limma)
-library(Glimma)
-library(edgeR)
-
-# This script follows 00_datawrangling.R
-
-# import "colData" which contains sample information and "countData" which contains read counts
-colData <- read.csv("metadata/00_colData.csv", header = T, row.names = 1)
-countData <- read.csv("results/00_counts.csv", header = T, row.names = 1)
-head(colData)
-head(countData[1:3])
-
-print("subset for quick run")
-countData  <- head(countData, 1000) # suset for quick analysis
-
-geneinfo <- row.names(countData)
-
-# Then, I followed the steps from <https://github.com/macmanes-lab/RockDove/blob/master/parental_care/parental_analysis.Rmd>.
-
-head(colData$group)
-print("create a large DGEList with 3 elements")
-parentalobject <- DGEList(counts=countData, genes=geneinfo, group=colData$group)
-
-print("transform raw counts to countspermillion")
-cpms <- cpm(parentalobject)
-
-print("calculate number of lowly lowly expressed genes and remove them")
-table(rowSums(parentalobject$counts==0)==10)
-keep_genes <- rowSums(cpms >= 1) >= 10
-dge <- parentalobject[keep_genes, ]
-
-print("specific the design")
-parentaldesign <- model.matrix(~ colData$group )
-colnames(parentaldesign) <- levels(colData$group)
-
-print("The TMM normalization")
-parentalobject <- calcNormFactors(parentalobject)
-parentalobject <- estimateCommonDisp(parentalobject)
-parentalobject <- estimateTagwiseDisp(parentalobject)
-parentalobject <- estimateDisp(parentalobject, parentaldesign)
-parentalobject <- estimateGLMCommonDisp(parentalobject, parentaldesign, verbose=TRUE)
-parentalobject <- estimateGLMTrendedDisp(parentalobject, parentaldesign)
-parentalobject <- estimateGLMTagwiseDisp(parentalobject, parentaldesign)
-
-print("find and print data")
-names(parentalobject)
[…]

### 2020-08-28 analysis/06_fig3.R

@@ -1,0 +1,132 @@
+# Fig 3
+
+library(tidyverse)
+library(cowplot)
+library(ggsignif)
+
+source("R/themes.R")
+source("R/functions.R")
+source("R/genelists.R")
+
+
+pitvsdf <- wranglevsds("results/03_pitvsdf.csv")
+pitvsdm <- wranglevsds("results/03_pitvsdm.csv")
+
+print("done with vsds")
+
+allDEG <- read_csv("results/03_allDEG.csv")
+
+DEGcontrol <- allDEG %>% 
+  filter(grepl("control", comparison),
+         !grepl("m.|early|extend|prolong", comparison))  %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol)) %>%
+  group_by(sex, tissue, comparison, direction) %>%
+  summarise(n = n()) %>%
+  mutate(n = ifelse(direction == "control", n*-1, n*1 ))
+
+
+DEGbldg <- allDEG %>% 
+  filter(grepl("bldg", comparison),
+         !grepl("m.|early|extend|prolong|control", comparison))  %>%
+  drop_na() %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))  %>%
+  group_by(sex, tissue, comparison, direction) %>%
+  summarise(n = n()) %>%
+  mutate(n = ifelse(direction == "bldg", n*-1, n*1 ))
+
+DEGchar <- allDEG %>% 
+  filter(comparison %in% comparisonlevelschar,
+         !grepl("control|bldg", comparison)) %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelschar)) %>%
+  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
+  group_by(sex, tissue, comparison, direction, updown) %>%
+  summarise(n = n()) %>%
+  mutate(n = n*updown ) %>%
+  select(-updown)
+
+DEGremove <- allDEG %>% 
+  filter(comparison %in% comparisonlevelsremoval) %>%
+  mutate(comparison = factor(comparison, levels = comparisonlevelsremoval)) %>%
+  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
+  group_by(sex, tissue, comparison, direction, updown) %>%
+  summarise(n = n()) %>%
+  mutate(n = n*updown ) %>%
+  select(-updown)
+DEGremove
+
[…]

### 2020-09-02 analysis/boneyard/boneyard.Rmd
ready for new fig 23

--- 2020-08-28 analysis/boneyard.Rmd
@@
+```{r}
+comparisonlevelscontrol <- c("control_bldg", "control_lay",  "control_inc.d3",  
+                          "control_inc.d9",   "control_inc.d17",  
+                          "control_hatch", "control_n5", "control_n9")
+
+comparisonlevelsbldg <- c( "bldg_lay",  "bldg_inc.d3",  
+                             "bldg_inc.d9", "bldg_inc.d17",  
+                             "bldg_hatch", "bldg_n5", "bldg_n9")
+
+comparisonlabelscharnobldg <- c("lay\ninc3", 
+                          "inc3\ninc9", "inc.d9\ninc17", 
+                          "inc17\nhatch" , "hatch\nn5", "n5\nn9")
+
+comparisonlabelscontrol <- c("bldg\ncontrol" ,"lay\ncontrol", "inc3\ncontrol", 
+                                "inc9\ncontrol", "inc17\ncontrol", 
+                                "hatch\ncontrol" , "n5\ncontrol", "n9\ncontrol")
+
+comparisonlabelsbldg <- c("lay\nbldg", "inc3\nbldg", 
+                                "inc9\nbldg", "inc17\nbldg", 
+                                "hatch\nbldg" , "n5\nbldg", "n9\nbldg")
+
+comparisonlevelsremoval <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", 
+                             "inc.d17_m.inc.d17", "hatch_m.n2")
+
+comparisonlablelsremoval <- c("m.inc.d3", "m.inc.d9", 
+                              "m.inc.d17", "m.n2")
+
+comparisonlevelsreplace <- c("hatch_early", "inc.d17_prolong",  "hatch_prolong", 
+                             "extend_hatch", "n5_extend")
+
+comparisonlablelssreplace <- c("hatch\nearly", "inc.d17\nprolong",  "hatch\nprolong", 
+                             "extend\nhatch", "n5\nextend")
+
+```
+
 ## Treatment specific DEGs

### 2020-09-02 analysis/05_figs23.R

@@ -20,63 +20,68 @@ allDEG <- read_csv("results/03_allDEG.csv")
 
 ## figure 2 candidate genes
 
-makefig2 <- function(tissue, dff, dfm, candidategene, comp1, comp2, label1){
+makefig2 <- function(label1, label2, tissue, dff, dfm, 
+                     candidategene, comp1, comp2, 
+                     treatment1, treatment2, treatment3, treatment4){
   
   fsubtitle = paste("Female", tissue, sep = " ")
   msubtitle = paste("Male", tissue, sep = " ")
 
-  a <- plotcandidatechar(dff, candidategene) + labs(subtitle = fsubtitle) +
+  a <- plotcandidatechar(dff, candidategene, treatment1, treatment2) + labs(subtitle = fsubtitle) +
     theme(axis.title.x = element_blank(), axis.text.x = element_blank())
-  b <- plotremoval(dff, candidategene) + labs(y = NULL) + labs(subtitle = " ")+
+  b <- plotcandidatemanip(dff, candidategene, treatment3, treatment4) + labs(y = NULL) + labs(subtitle = " ") +
     theme(axis.title.x = element_blank(), axis.text.x = element_blank())
-  c <- plotreplacement(dff, candidategene) + labs(y = NULL) + labs(subtitle = " ")+
-    theme(axis.title.x = element_blank(), axis.text.x = element_blank())
-  d <- plot.volcano(tissue, "female",  comp1) + labs(subtitle = " ") +
+  
+  c <- plot.volcano(tissue, "female",  comp1) + labs(subtitle = " ") +
     theme(axis.title.x = element_blank(),
           legend.position = "none")
-  d2 <- plot.volcano(tissue, "female",  comp2) + labs(subtitle = " ")  +
+  d <- plot.volcano(tissue, "female",  comp2) + labs(subtitle = " ")  +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           legend.position = "none")
   
-  abc <- plot_grid(a,b,c, nrow = 1)
-  abcd <- plot_grid(abc,d,d2, nrow = 1, rel_widths = c(3,1,0.9))
+  abcd <- plot_grid(a,b,c,d,nrow = 1, rel_widths = c(9,12,7.35,7),
+                    labels = c(label1, "", label2), label_size = 8)
   
 
-  h <- plotcandidatechar(dfm, candidategene) + labs(subtitle = msubtitle)
[…]

### 2020-09-07 manuscript/elife-v1.tex
fixed bug in figures plus ms reorg

--- 2020-05-18 docs/manuscript/elife-v1.tex
@@
 \title{Transcriptional changes in the reproductive axis of parenting pigeons}
-\author{by Rayna M Harris \footnote{equal contribtion}, Suzanne H. Austin \footnote{equal contribtion, currently at the
-  University of Oregon}, Andrew S. Lang, Victoria S. Farrar, April Booth, Tanner Feustel, Matthew D. MacManes, Rebecca M. Calisi \footnote{corresponding author}}
+\author{by Suzanne H. Austin \footnote{equal contribtion, currently at the
+  University of Oregon}, Rayna M Harris \footnote{equal contribtion}, Andrew S.
+Lang, Victoria S.
+Farrar, April Booth, Tanner Feustel, Matthew D.
+MacManes, Rebecca M.
+Calisi \footnote{corresponding author}}
 
@@
 
-\textbackslash{}begin\{figure\}{[}h{]} \centering
-\includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
-\textbackslash{}caption\{Broad patterns of variation in hypothalamic,
-pituitary, and gonadal gene expression in male and female rock doves
-across reproductive and parental transitions.** A) Experimental design.
-B) The t-SNE plot shows all 636 samples as points in space colored by
-tissue with ellipses drawn around show 95\% confidence interval, and the
-bar plots show the total number of differentially expressioned genes
-(DEGs) whose expression was significantly higher in one tissue versus
-the other. C) This t-SNE plot shows the same data points colored by sex
-while the ellipses delinate the source tissue; bar plots show the DEGs
-between the sexes for each tissue. D) These t-SNE and bar plots provide
-insight into the temporal differences in gene expression within each
-tissue for each sex. Here, the bar plots highlight the number of DEGs
-between each sequential time point. For comparisons to controls, see
-fupplemental figure 1.\} \label{figure:fig1}
-\textbackslash{}end\{figure\}
+\hypertarget{figures}{%
+\subsection{Figures}\label{figures}}
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/supplfig-1-1}
-  \caption{Summary of the total number of differentially expressed genes (DEGs) from pair-wise comparison using DESeq2.** Bar plots showing the total number of differentially expressed genes (p < 0.1) and colored by the direction of increased expression (or positive log fold change (+LFC))  are shown for three different analysis: relative to the control group (A), the nest-building (bldg) group (B), and the previous sampling timepoint (C).}
-  \label{figure:supplfig1}
-\end{figure}
+\includegraphics[width=0.95\linewidth]{fig1-1}
 
-\newpage
+\textbf{Figure 1. Experimental design.} A) We investigated nine
+timepoints that spanned the majority of reproductive efforts in this
+species. These time-points consisted of a ``control'' or non-parenting
+state (from MacManes et al 2016 and Calisi et al.~2017), nest-building,
+clutch initiation or the onset of incubation, clutch completion and
+early incubation, mid-incubation, late incubation, initiation of
+nestling care, early nestling care, and mid-nestling care. To teste
+whether external or internal factors were regulating gene activity, we
+also conducted a series of offspring removal (B) and replacement (C)
+manipulations to test whether transcriptional changes were dependent
+upon offspring presence or were regulated by an internal clock. (D) We
+used t-Distributed Stochastic Neighbor Embedding (t-SNE) to reduce the
+dimensionality of the transcriptomes from the hypothalamus, pituitary,
+and gonads of male and female rock. Circles and triangles represent
+female and male samples, respectively, and points are colored by
+treatment.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig2-1}
-  \caption{Differential gene expression differences between sequential parental stages.
-a) We calculated the differentail gene expression between each consequtive parental stage samples, for a total of 8 pair-wise comparisons for each sex in each tissue.
-The y-axis of these bar plots displays the total number of differentially expressed genes for each comparsison in the hypothalamus (b), pituitary (c), gonads (d).
-Bars are collored by the direction of the change in gene expression.
-Control: control: red or , nest-building (bldg): orange, lay: yellow; incubation day 3 (inc.d3): light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, nestling care day 5 (n5): blue, and n9: light blue.}
-  \label{figure:fig2}
-\end{figure}
+\includegraphics[width=0.95\linewidth]{fig2-1}
 
-\newpage
+\textbf{Figure 2. Hypothesis-driven and data-driven approaches to
+identifying changes in gene expression.} Box-and-whisker plots show
+changes in gene expression of serotonin receptor (HTR2C) in the
+hypothalamus (A), prolactin (PRL) in the pituitary (B), and estrogen
+receptor 1 (ESR1) in the gonads.Bar and statistics are shown for one
+contrast of interest between treatment groups. Boxes are colored by
+treatment. Volcano plots show the log-fold change (LFC) and -log10(FDR)
+for all genes that are significantly different differentially between
+the aforementioned treatment groups. Boxes are colored by the treatment
+were gene expression was higher.
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig3-1}
-  \caption{Another another figure}
-  \label{figure:fig3}
-\end{figure}
+\includegraphics[width=0.95\linewidth]{fig3-1}
 
-\newpage
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig4-1}
-  \caption{Another another figure}
-  \label{figure:fig4}
-\end{figure}
-
-\newpage
+\textbf{Figure 3. Summary of differentially expressed genes.} Bar plots
+show the total number of significantly differentially expressed genes
+for all two-way comparisons in the hypothalamus (A), pituitary (B),
+gonads (C) . Contrasts are made relative to the non-breeding controls,
+the nest-building group, the previous stage, or the internal or external
+control group. Bars are colored by the treatment where gene expression
+was higher. A negative number of differentially expressed genes means
+that many genes had increase expression in the reference group.
 
@@
 \address{%
-Rayna M Harris \footnote{equal contribtion}\\
+Suzanne H. Austin \footnote{equal contribtion, currently at the
+  University of Oregon}\\
 University of California, Davis\\
@@
 }
+\href{mailto:suzannehaustin@gmail.com}{\nolinkurl{suzannehaustin@gmail.com}}
 
 \address{%
-Suzanne H. Austin \footnote{equal contribtion, currently at the
-  University of Oregon}\\
+Rayna M Harris \footnote{equal contribtion}\\
 University of California, Davis\\

### 2020-09-07 manuscript/elife_v1.Rmd
fixed bug in figures plus ms reorg

--- 2020-05-18 docs/manuscript/elife_v1.Rmd
@@
 author:
+  - name: Suzanne H. Austin  ^[equal contribtion, currently at the University of Oregon] 
+    affiliation: University of California, Davis
+    email:  suzannehaustin@gmail.com
   - name: Rayna M Harris ^[equal contribtion] 
-    affiliation: University of California, Davis
-  - name: Suzanne H. Austin  ^[equal contribtion, currently at the University of Oregon] 
     affiliation: University of California, Davis
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
-  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions.** A) Experimental design.
-B) The t-SNE plot shows all 636 samples as points in space colored by tissue with ellipses drawn around show 95% confidence interval, and the bar plots show the total number of differentially expressioned genes (DEGs) whose expression was significantly higher in one tissue versus the other.
-C) This t-SNE plot shows the same data points colored by sex while the ellipses delinate the source tissue; bar plots show the DEGs between the sexes for each tissue.
-D) These t-SNE and bar plots provide insight into the temporal differences in gene expression within each tissue for each sex.
-Here, the bar plots highlight the number of DEGs between each sequential time point.
-For comparisons to controls, see fupplemental figure 1.}
-  \label{figure:fig1}
-\end{figure}
+## Figures
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/supplfig-1-1}
-  \caption{Summary of the total number of differentially expressed genes (DEGs) from pair-wise comparison using DESeq2.** Bar plots showing the total number of differentially expressed genes (p < 0.1) and colored by the direction of increased expression (or positive log fold change (+LFC))  are shown for three different analysis: relative to the control group (A), the nest-building (bldg) group (B), and the previous sampling timepoint (C).}
-  \label{figure:supplfig1}
-\end{figure}
+\includegraphics[width=0.95\linewidth]{fig1-1}
 
-\newpage
+**Figure 1.
+Experimental design.** A) We investigated nine timepoints that spanned the majority of reproductive efforts in this species.
+These time-points consisted of a "control" or non-parenting state (from MacManes et al 2016 and Calisi et al.
+2017), nest-building, clutch initiation or the onset of incubation, clutch completion and early incubation, mid-incubation, late incubation, initiation of nestling care, early nestling care, and mid-nestling care.
+To teste whether external or internal factors were regulating gene activity, we also conducted a series of offspring removal (B) and replacement (C) manipulations to test whether transcriptional changes were dependent upon offspring presence or were regulated by an internal clock.
+(D) We used t-Distributed Stochastic Neighbor Embedding (t-SNE) to reduce the dimensionality of the transcriptomes from the hypothalamus, pituitary, and gonads of male and female rock.
+Circles and triangles represent female and male samples, respectively, and points are colored by treatment.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig2-1}
-  \caption{Differential gene expression differences between sequential parental stages.
-a) We calculated the differentail gene expression between each consequtive parental stage samples, for a total of 8 pair-wise comparisons for each sex in each tissue.
-The y-axis of these bar plots displays the total number of differentially expressed genes for each comparsison in the hypothalamus (b), pituitary (c), gonads (d).
-Bars are collored by the direction of the change in gene expression.
-Control: control: red or , nest-building (bldg): orange, lay: yellow; incubation day 3 (inc.d3): light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, nestling care day 5 (n5): blue, and n9: light blue.}
-  \label{figure:fig2}
-\end{figure}
+\includegraphics[width=0.95\linewidth]{fig2-1}
 
-\newpage
+**Figure 2.
+Hypothesis-driven and data-driven approaches to identifying changes in gene expression.** Box-and-whisker plots show changes in gene expression of serotonin receptor (HTR2C) in the hypothalamus (A), prolactin (PRL) in the pituitary (B), and estrogen receptor 1 (ESR1) in the gonads.Bar and statistics are shown for one contrast of interest between treatment groups.
+Boxes are colored by treatment.
+Volcano plots show the log-fold change (LFC) and -log10(FDR) for all genes that are significantly different differentially between the aforementioned treatment groups.
+Boxes are colored by the treatment were gene expression was higher.
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig3-1}
-  \caption{Another another figure}
-  \label{figure:fig3}
-\end{figure}
+\includegraphics[width=0.95\linewidth]{fig3-1}
 
-\newpage
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=1.0\textwidth]{../../figures/fig4-1}
-  \caption{Another another figure}
-  \label{figure:fig4}
-\end{figure}
-
-\newpage
+**Figure 3.
+Summary of differentially expressed genes.** Bar plots show the total number of significantly differentially expressed genes for all two-way comparisons in the hypothalamus (A), pituitary (B), gonads (C) .
+Contrasts are made relative to the non-breeding controls, the nest-building group, the previous stage, or the internal or external control group.
+Bars are colored by the treatment where gene expression was higher.
+A negative number of differentially expressed genes means that many genes had increase expression in the reference group.
 

### 2020-09-07 manuscript/elife-v1.tex
now with a table and some results

@@
 % !TeX root = RJwrapper.tex
-\title{Transcriptional changes in the reproductive axis of parenting pigeons}
+\title{The reproductive and parental care transcriptome of the rock dove}
 \author{by Suzanne H. Austin \footnote{equal contribtion, currently at the
@@
 \abstract{%
-In species that provide parental care, successful rearing of offspring
-involves a shift from aggressive and sexual behaviors to more caring and
-nurturing ones. We hypothesize that these shifts in behavior are related
-to changes in gene expression that are controlled by internal mechanisms
-rather than by the external cues. We found that gene expression in the
-hypothalamus is fairly insensitive to the natural transition between egg
-and nestling stages; however, in the pituitary and gonads, gene
-expression showed much more plasticity over the course of reproduction.
-We show evidence that natural transitions in reproductive and parental
-behavior by mediated internal clock controlling gene expression in the
-pitary and gonads but that the hypothalamus is very sensitive to
-external cues that signal a distruption in the natural parental care
-cycle. This research provides a deeper understanding of the interplay of
-genes, hormones, and parental care and has important implications for
-understanding molecular processes that underlie behavior transitions.
+In animals that exhibit a parental care reproductive strategy,
+successful rearing of offspring typically involves a shift from sexual
+behaviors to more caring and nurturing ones. Decades of work have
+highlighted the role of particular hormones in the maintenance of
+parental care behavior, the majority of which are produced by a
+biological system essential for reproduction, the
+hypothalamic-pituitary-gonadal (HPG) axis. However, we know very little
+about when and how the HPG axis transitions into parental care behaviors
+in any vertebrate, and this knowledge is fundamentally important to our
+understanding of the mechanisms mediating parental care. First, we
+characterized the transcriptome of the HPG axis in both mothers and
+fathers over 9 points of the parental care stage using a socially
+monogamous and bi-parental model, the rock dove (Columba livia). Then,
+we experimentally manipulated offspring presence at 7 different time
+points using classic offspring replacement and removal manipulations to
+uncover cause and effect of these transcriptional changes. We offer the
+most complete characterization and understanding of changes in neural,
+pituitary and gonadal transcription and translation ever reported in any
+vertebrate over the course of parental care.
 }
@@
 
-Reproduction is a critical period in an organism's life cycle. Parents
-must balance their own needs with the energetic costs and
-time-constraints required to successfully reproduce and rear their
-offspring to independence (cite). A complex series of evolutionarily
-conserved physiological processes underlie behaviors typically
-associated with reproduction in vertebrates, including courtship,
-nesting, and parental care. These processes are primarily mediated by
-the reproductive axis, which consists of the hypothalamus in the brain,
-the pituitary, and the gonads (i.e., the HPG). Our understanding of the
-physiological changes that adults navigate as they transition through
-reproductive states are primarily limited to a number of circulating
-hormones and a small set of well-known genes (cite). We have yet to come
-close to conceptualizing the larger symphony of reproductive events that
-occur to yield reproductive and parental care behaviors.
+Parental care is essential to many organisms' reproductive strategy and
+predominates avian and mammalian groups (Clutton-Brock, 1991; Royle et
+al., 2012). Depending on the species, reproductive and parental
+behaviors may include territory establishment, courtship and mating,
+nest-building, nest defense, incubation, and the brooding and
+provisioning of young. The goal of these behaviors, either directly or
+indirectly, is the successful production of offspring resulting in a
+direct fitness benefit (Reviewed in Santos, E.S.A, 2012). Underlying
+these behaviors are the anatomical and physiological changes that
+facilitate the transition into and across reproductive and parenting
+stages. These changes have fascinated biologists and inspired a
+prodigious body of research. Much of this research has focused on
+understanding reproduction and parental care from an endocrine and
+neuroendocrine perspective (reviewed in Ondrasek 2016, Kelly 2020,
+Wingfield, Silver, Buntin). Reproduction and parental care is largely
+regulated by a set of three tissues: The Hypothalamus in the brain, the
+Pituitary, and the Gonads (testes and ovaries). Together, these tissues
+are often referred to as the `reproductive' axis or the HPG. To activate
+reproduction, the HPG initiates an endocrine cascade starting with the
+production and release of gonadotropin-releasing hormone (GnRH) in the
+hypothalamus which, in turn, stimulates the anterior pituitary to
+produce luteinizing hormone (LH) and follicle-stimulating hormone (FSH).
+LH and FSH travel to the gonads via the bloodstream where they induce
+gonadal growth and steroidogenesis of estradiol, testosterone and
+progesterone). These sex steroids have well established roles in mating
+behavior, nest-building, territory defense, and egg-laying in birds
+(reviewed in Buntin). Another key hormone, prolactin, is produced in the
+pituitary and is associated with the induction and maintenance of
+behaviors such as incubation in birds and lactation (in doves and
+mammals) (birds: reviewed in Angelier, Austin et al.~in prep, mammals:
+Austin and Word). Other hormones like vasopressin and oxytocin are
+related to affiliative behaviors (Yoshihara C, 2018) while dopamine,
+serotonin aggressive behaviors or avoidance (Rilling, JK, 2013; Adams
+DB, 2006). While these and many other hormones have been reported to
+play integral roles in various reproductive and parental care behaviors,
+we have yet to come close to conceptualizing the larger physiological
+and genomic coordination of events that yield such behaviors.
 
-Using the model of a biparental vertebrate species, the rock dove
-(Columba livia), our team has previously demonstrated the dynamic nature
-of gene activity within the reproductive or HPG axis.
-\citet{MacManes2017} reported that HPG gene activity varied by sex, with
-females exhibiting higher levels of overall gene expression than males.
-\citet{Calisi2018} reported that HPG gene activity varied response to
-restraint-stress, and Austin et al.~identify genes that responds to both
-corticosterone treatment and restraint stress. While each of these
-studies advanced our understanding of sex-differences and
-transcriptional plasticity in the reproducitve axis, none of these
-studies examined activity of the reproducitve axis of bird that are
-actively reproducing or caring for offspring. Studying parental care in
-both sexes of mammalian species is challenging because only the females
-are able to use lactation to feed their young, and the physiological
-mechanisms regulating this process are hard to control for. The rock
-dove is an excellent system for studying reproduction and parental care
-in both sexes because both males and females feed their young an
-energy-rich secretion produced from the epithelial lining of their
-crops, known as crop milk.
+To begin to address this larger goal, we investigated genomic changes in
+the reproductive axis using the model of the rock dove (Columba livia).
+This species is socially and genetically monogamous and has bi-parental
+incubation and offspring care thereby making investigations of sexual
+variation across reproduction possible. Doves and pigeons have been used
+extensively to study the physiological mechanisms of reproduction and
+parental care, including the proximate and ultimate mechanisms that
+regulate associated behaviors (cite a bunch of new and old papers or
+books). Our team has produced a series of papers with the goal of
+illuminating how transcriptomic activity in the reproductive axis varied
+by sex (MacManes et al.~2016), in response to a stressor (Calisi et
+al.~2017, Lang et al.~2020), and following a hormonal manipulation
+(Austin et al.~in prep). By using a genomic approach that spans the
+reproductive timeline, we have the ability to elucidate a
+transcriptional symphony that coordinates the anatomical, physiological
+and behavioral changes associated with transitioning from non-breeding
+to parenting in a vertebrate species. The aims of this research were
+three-fold: Our first aim was to produce a high quality and reproducible
+transcriptional characterization of the HPG over reproduction and
+parental care in male and female rock doves. We were specifically
+interested in understanding how gene activity in the HPG changed as
+individuals transitioned from a sexually mature, non-reproductive state
+into reproduction and parenting. To accomplish this aim, we investigated
+nine time points that spanned the majority of reproductive efforts in
+this species. These time points consisted of 1) a control or
+non-parenting state (from MacManes et al 2016 and Calisi et al.~2017),
+2) nest-building, 3) clutch initiation and onset of incubation, 4)
+clutch completion and early incubation, 5) mid-incubation, 6) late
+incubation, 7) initiation of nestling care, 8) early nestling care, and
+9) mid-nestling care. Because the majority of time points included in
+this study involved some form of direct offspring care, for simplicity,
+we will discuss these as parental care rather than differentiating
+between reproductive and parenting behaviors. However, it should be
+understood that parental care remains only one component of a larger set
+of reproductive behaviors. Our second aim focused on characterizing the
+major patterns of transcriptional changes between the tissues, sexes,
+and parental stages. We specifically investigate how gene expression in
+the HPG changed across various transition points. For example, from
+non-breeding to mating and nest-building behaviors, from nest-building
+to egg-laying to incubation, from incubation to nestling care, and from
+highly dependent nestlings (brooding and crop milk production) to older,
+thermoregulatory independent nestlings that are fed a regurgitant of
+seeds. Our third aim was to identify the drivers underlying changes in
+transcription. Specifically, we tested whether external or internal
+factors were regulating gene activity. We conducted a series of
+offspring removal and replacement manipulations to test whether
+transcriptional changes were dependent upon offspring presence or were
+regulated by an internal clock.
 
-In this study, we investigate the activity of the HPG axis over the
-course of reproduction and parental care by characterizing gene
-expression as birds transition from non-breeding to nest-building, as
-they lay and incubate their clutches, and finally as they become parents
-and raise their chicks. Then, we conducted a series of nest
-manipulations to ascertain constraints associated with biological timing
-versus input from external stimuli. We conducted a comprehensive series
-of analyses examine broad patterns of variation as well as the activity
-of specific genes in specific tissues. We have also made our data and
-scripts available so that others can reuse the data to ask different
-questions or reproduce our results. This research provides a deeper
-understanding of the interplay of genes, hormones, and parental care and
-has important implications for understanding molecular processes that
-underlie behavioral transitions. \citep[reviewed in][]{Lea1989} say
-this, but other say that \citep[reviewed in][]{Dulac2014}.
+Additionally, because these data are extensive, we took the opportunity
+to provide them in an easily accessible form that can be used for
+independent data exploration by the reader. Given the homology of genes
+and gene networks shared between rock doves and other vertebrate
+species, the findings from these RNA-seq experiments could be of
+interest to scientists and clinicians studying diverse systems,
+particularly in non-model organisms like our study species. Overall,
+this study provides a deeper understanding of the interplay of genes
+across parenting and has important implications for understanding
+molecular processes that underlie behavioral transitions.
 
@@
 
-\emph{Data-driven characterization of reproductive and parental stage
-groups.}
+\hypertarget{the-rock-dove-reproduction-and-parental-care-transcriptome}{%
+\subsubsection{The rock dove reproduction and parental care
+transcriptome}\label{the-rock-dove-reproduction-and-parental-care-transcriptome}}
 
-We first examined broad patterns of variation in gene expression in the
-hypothalamus, pituitary, and gonads of male and female rock doves across
-a series of timepoints that encapsulate important transitions in
-reproduction and and parental care (Fig. 1A). t-Distributed Stochastic
-Neighbor Embedding (t-SNE) is a nonlinear, dimension reduction technique
-is used to visualize high-dimensional data (van der Maaten and Hinton
-2008) and is widely used in in RNA sequencing to identify related and
-distinct cell types (refs). Differential expespression profiling is
-useful for pair-wise comparisons of differentially expressed genes
-(DEGs) with positive log fold change (+LFC) in one group compared to
-another. Combined, the provide stronger support for similarities and
-differences between cell-types and treatment groups in multi-dimential
-datasets. First, the data supports that the hypothalamus, pituitary, and
-gonads have distinct gene expression profiles as evidenced by little
-overlap in t-SNE space and an average of 13,000 DEGs between each tissue
-(Fig. 1B). Second, the data supports that sex-differences are much more
-promient in the gonads than in the brain (13,000 versus 2,000 DEGs)
-(Fig. 1C). Because of these widespread tissue and sex-differences,
-further ananlyses were conducted separately for each tissue and sex
-(Fig. 1D, Suppl. Fig 1). In all the tissues, the control samples are
-quiet different from all the other timepoints samples, and only a
-fraction of the transcriptome is altered during the course of parental
-care at the timepoints sampled. The most notable differences male and
-female hpothalamic control samples are very different from on the rest
-(\textasciitilde{}7000 DEGs, and only the two nestling care stages (n5
-and n9)) show differential expression of more than 1,000 genes. In the
-pituitary, late inubcation and hatch are quite simlilar to each other
-and distinct from the earlier incubation stages stages, and temporal
-changes in gene expression appears to be less variable in males than
-females. In the female gonads there appear to be more changes in gene
-expression early one, between nest building and early incuation, but the
-male gonads show very little variation associated with parental stage.
+We created the rock dove transcriptome by sequenced 987 RNA samples from
+the hypothalamus, pituitary and gonads from the female and male of 166
+rock dove pairs across (Fig. 1). These samples come from nine timepoints
+that encapsulate important transitions in reproduction (Fig. 1A), four
+offspring removal manipulations (Fig. 1B), and three offspring
+replacement manipulations (Fig. 1C) to determine the influence of
+internal mechanisms and external stimuli on gene expression. Using
+T-distributed Stochastic Neighbor Embedding (tSNE) we show broad
+patterns of variation due to treatment or sex in each tissue (Fig. 1D).
+In the hypothalamus, tSNE dimension visually separates most of the
+samples that were subject to an offspring manipulation from those from
+naturally occurring parental stages. In the pituitary however, samples
+cluster more by wether they were before incubation day 9 or after
+incubation day 17. Neither the hypothalamus or pituitary show strong sex
+differences at this level, but the gonads samples from very distinct,
+sex-specific clusters.
 
-\emph{Hypothesis-driven analysis of candidate parental-care genes.}
-
-\begin{itemize}
-\tightlist
-\item
-  Champagne Curley Genes (Fig 2)
-\item
-  GO genes (Fig 2 Suppl Fig 1)
-\item
-  Cancer genes (Fig 2 Suppl Fig 2)
-\item
-  Shiny genes (Fig 2 Suppl Fig 3)
-\end{itemize}
-
-\emph{Software for data exploration and hypothesis testing}
-
-Data accessibility and reproducibility are key challenges for today's
-researchers. Using Binder and Shiny, we have provides a means for
-scientists, physicians, and the general public to explore the data,
-code, and results from this project at
-\url{https://raynamharris.shinyapps.io/musicalgenes/}. Like Hipposeq
-(Cembrowski et al elife 2016, 2018), this tools allows quick and easy
-data exploration of RNAseq data from mutiple cell types that relate to
-brain function.
+\begin{Schunk}
+\begin{figure}
+\includegraphics{elife_v1_files/figure-latex/fig1-1} \caption[Experimental design]{Experimental design.
+A) We investigated nine timepoints that spanned the majority of reproductive efforts in this species.
+These timepoints consisted of a "control" or non-parenting state (from MacManes et al 2016 and Calisi et al.
+2017), nest-building, clutch initiation or the onset of incubation, clutch completion and early incubation, mid-incubation, late incubation, initiation of nestling care, early nestling care, and mid-nestling care.
+To teste whether external or internal factors were regulating gene activity, we also conducted a series of offspring removal (B) and replacement (C) manipulations to test whether transcriptional changes were dependent upon offspring presence or were regulated by an internal clock.
+(D) We used t-Distributed Stochastic Neighbor Embedding (t-SNE) to reduce the dimensionality of the transcriptomes from the hypothalamus, pituitary, and gonads of male and female rock.
+Circles and triangles represent female and male samples, respectively, and points are colored by treatment.}\label{fig:fig1}
+\end{figure}
+\end{Schunk}
 
@@
 \subsection{Discussion}\label{discussion}}
-
-This RNA-sequencing project was designed to characterize the
-reproductive axis of bi-parental pigeons across a characteristic
-reproductive cycle and through the parental main parental stages of egg
-incubation and parental care using data- and hypotheses driven
-approaches. We hypothesized that differences in gene expression would be
-driven by either internal physiology or external stimuli. In the
-pituitary, we found that the most explanatory source of variation in
-gene expression was the expression level of prolactin (PRL). Circulating
-levels of prolactin in the blood in this species are known to rise
-around day nine and peak on or the day before hatch. Our results confirm
-this and also identified one hundred genes whose expression is highly
-correlated with prolactin. Today, this variation, which changes more
-within a parental stage (e.g.~egg incubation or nestling care) suggests
-that internal physiology drives changes in gene expression that precede
-changes in external stimuli that lead to behavioral changes (e.g.~crop
-milk or regurgitation). We predicted that either sex steroid peptide
-hormones would play a role in governing the reproductive transitions,
-but we did not anticipate discovering a link between PRL expression to
-genes that are involved with DNA repair and regulation of the cycle.
-Finally, we aimed to make these data and analyses FAIR (findable,
-accessible, interoperable, and reproducible), open and reproducible to
-encourage others to reuse our data to test new hypotheses or explore the
-data with a different perspective or toolset. The scripts that made the
-figures and tables are browseable online thanks to RStudio, GitHub and
-Binder. The data can be visualized and sonified (turned into sound)
-interactively thanks to RStudio, GitHub, and Shiny.
 
@@
 
+\newpage
+
+\hypertarget{tables}{%
+\subsection{Tables}\label{tables}}
+
+\begin{Schunk}
+\begin{table}
+
+\caption{\label{tab:table1}Changes in candidate gene expression.
+Significant chagns in gene expression are shown for X candidate genes.
+Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G).
+Comparisons are denoted in the column names.
+The first group is the reference group.}
+\centering
+\begin{tabular}[t]{l|l|l|l|l|l|l|l}
+\hline
+gene & bldg-lay & lay-inc3 & inc3-inc9 & inc9-inc17 & inc17-hatch & hatch-n5 & n5-n9\\
+\hline
+AR &  &  &  &  &  &  & \\
+\hline
+AVPR1B &  &  &  &  &  &  & \\
+\hline
+AVPR2 &  &  &  &  &  & FP(+) & \\
+\hline
+CRH &  &  &  &  &  &  & \\
+\hline
+CRHR1 &  &  &  &  &  &  & \\
+\hline
+DIO1 &  &  &  &  &  &  & \\
+\hline
+DIO2 &  & FG(-) &  &  &  &  & \\
+\hline
+DIO3 & FP(-) &  &  &  &  &  & \\
+\hline
+ESR2 &  &  &  &  &  &  & \\
+\hline
+FSHB &  &  &  &  &  &  & \\
+\hline
+FSHR &  &  &  &  &  &  & \\
+\hline
+GABRQ &  &  &  &  &  &  & \\
+\hline
+GALR1 &  &  &  &  &  &  & \\
+\hline
+GNRH1 &  &  &  &  &  &  & \\
+\hline
+GNRHR & FP(-) &  &  &  &  &  & \\
+\hline
+JAK2 &  &  &  &  &  &  & \\
+\hline
+LHCGR &  &  &  &  &  &  & \\
+\hline
+NPFFR1 &  &  &  &  &  &  & \\
+\hline
+NPVF &  &  &  &  &  &  & \\
+\hline
+NPY &  &  &  &  &  &  & \\
+\hline
+NR3C2 &  &  &  &  &  &  & \\
+\hline
+POMC &  &  &  &  &  &  & \\
+\hline
+PRL &  &  &  & FP(+); MP(+) &  &  & \\
+\hline
+PRLH & FP(-) & FP(+) &  & FP(+) &  &  & \\
+\hline
+PRLHR &  &  &  &  &  &  & \\
+\hline
+SERPINA4 &  &  &  &  &  &  & \\
+\hline
+VIPR1 &  &  &  &  &  &  & \\
+\hline
+\end{tabular}
+\end{table}
+
+\end{Schunk}
+
+\hypertarget{figures}{%
+\subsection{Figures}\label{figures}}
+
+\begin{Schunk}
+\begin{figure}
+\includegraphics{elife_v1_files/figure-latex/fig2-1} \caption[Hypothesis-driven and data-driven approaches to identifying changes in gene expression.Box-and-whisker plots show changes in gene expression of serotonin receptor (HTR2C) in the hypothalamus (A), prolactin (PRL) in the pituitary (B), and estrogen receptor 1 (ESR1) in the gonads.Bar and statistics are shown for one contrast of interest between treatment groups]{Hypothesis-driven and data-driven approaches to identifying changes in gene expression.Box-and-whisker plots show changes in gene expression of serotonin receptor (HTR2C) in the hypothalamus (A), prolactin (PRL) in the pituitary (B), and estrogen receptor 1 (ESR1) in the gonads.Bar and statistics are shown for one contrast of interest between treatment groups.
+Boxes are colored by treatment.
+Volcano plots show the log-fold change (LFC) and -log10(FDR) for all genes that are significantly different differentially between the aforementioned treatment groups.
+Boxes are colored by the treatment were gene expression was higher.}\label{fig:fig2}
+\end{figure}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{figure}
+\includegraphics{elife_v1_files/figure-latex/fig3-1} \caption[Summary of differentially expressed genes]{Summary of differentially expressed genes.
+Bar plots show the total number of significantly differentially expressed genes for all two-way comparisons in the hypothalamus (A), pituitary (B), gonads (C) .
+Contrasts are made relative to the non-breeding controls, the nest-building group, the previous stage, or the internal or external control group.
+Bars are colored by the treatment where gene expression was higher.
+A negative number of differentially expressed genes means that many genes had increase expression in the reference group.}\label{fig:fig3}
+\end{figure}
+\end{Schunk}
+
+\newpage
+
+\hypertarget{references}{%
+\subsection{References}\label{references}}
+
+\bibliography{RJreferences}
+
 \hypertarget{acknowledgements}{%
@@
 Amanda, Alicia, Picaso). We thank Pacha for help with database and
-Shiny.
-
-We thank Irene, Erica, Beth Krestoff and Tiffany Chen for assistance in
-managing the aviary and collecting nest stage data. We thank Candice
-Lee, Annie Bond, and Tanner Feustel and others?? for assistance with
-tissue collections. We thank Dr.~Fred Angelier for performing the
-prolactin RIA.This research was supported by NSF IOS 1455960 to R.M.C.
-and M.D.M. Author contributions: RMC designed the experiment. RMC and
-MDM acquired funding. SHA contributed to the experimental design by
-adding several time points. SHA, with the assistance of VF, AB, and RMC,
-conducted the data collection. SHA and a team of trained undergraduate
-students biopsy punched the hypothalami and lateral septum. ASL
-extracted the RNA, created the libraries, and conducted all early
-bioinformatic work. RMH conducted all later analyses of the genomic
-data. RMH and SHA drafted the manuscript and all other authors provided
-edits.
-
-\hypertarget{references}{%
-\subsection{References}\label{references}}
-
-\bibliography{RJreferences}
-
-\newpage
-
-\hypertarget{figures}{%
-\subsection{Figures}\label{figures}}
-
-\includegraphics[width=0.95\linewidth]{fig1-1}
-
-\textbf{Figure 1. Experimental design.} A) We investigated nine
-timepoints that spanned the majority of reproductive efforts in this
-species. These time-points consisted of a ``control'' or non-parenting
-state (from MacManes et al 2016 and Calisi et al.~2017), nest-building,
-clutch initiation or the onset of incubation, clutch completion and
-early incubation, mid-incubation, late incubation, initiation of
-nestling care, early nestling care, and mid-nestling care. To teste
-whether external or internal factors were regulating gene activity, we
-also conducted a series of offspring removal (B) and replacement (C)
-manipulations to test whether transcriptional changes were dependent
-upon offspring presence or were regulated by an internal clock. (D) We
-used t-Distributed Stochastic Neighbor Embedding (t-SNE) to reduce the
-dimensionality of the transcriptomes from the hypothalamus, pituitary,
-and gonads of male and female rock. Circles and triangles represent
-female and male samples, respectively, and points are colored by
-treatment.
-
-\includegraphics[width=0.95\linewidth]{fig2-1}
-
-\textbf{Figure 2. Hypothesis-driven and data-driven approaches to
-identifying changes in gene expression.} Box-and-whisker plots show
-changes in gene expression of serotonin receptor (HTR2C) in the
-hypothalamus (A), prolactin (PRL) in the pituitary (B), and estrogen
-receptor 1 (ESR1) in the gonads.Bar and statistics are shown for one
-contrast of interest between treatment groups. Boxes are colored by
-treatment. Volcano plots show the log-fold change (LFC) and -log10(FDR)
-for all genes that are significantly different differentially between
-the aforementioned treatment groups. Boxes are colored by the treatment
-were gene expression was higher.
-
-\includegraphics[width=0.95\linewidth]{fig3-1}
-
-\textbf{Figure 3. Summary of differentially expressed genes.} Bar plots
-show the total number of significantly differentially expressed genes
-for all two-way comparisons in the hypothalamus (A), pituitary (B),
-gonads (C) . Contrasts are made relative to the non-breeding controls,
-the nest-building group, the previous stage, or the internal or external
-control group. Bars are colored by the treatment where gene expression
-was higher. A negative number of differentially expressed genes means
-that many genes had increase expression in the reference group.
-
-\newpage
+Shiny. We thank Irene, Erica, Beth Krestoff and Tiffany Chen for
+assistance in managing the aviary and collecting nest stage data. We
+thank Candice Lee, Annie Bond, and Tanner Feustel and others?? for
+assistance with tissue collections. We thank Dr.~Fred Angelier for
+performing the prolactin RIA.This research was supported by NSF IOS
+1455960 to R.M.C. and M.D.M. Author contributions: RMC designed the
+experiment. RMC and MDM acquired funding. SHA contributed to the
+experimental design by adding several time points. SHA, with the
+assistance of VF, AB, and RMC, conducted the data collection. SHA and a
+team of trained undergraduate students biopsy punched the hypothalami
+and lateral septum. ASL extracted the RNA, created the libraries, and
+conducted all early bioinformatic work. RMH conducted all later analyses
+of the genomic data. RMH and SHA drafted the manuscript and all other
+authors provided edits.
 

### 2020-09-07 manuscript/elife_v1.Rmd
now with a table and some results

@@
 ---
-title: Transcriptional changes in the reproductive axis of parenting pigeons
+title:  The reproductive and parental care transcriptome of the rock dove
 author:
@@
     address: Department of Molecular, Cellular and Biomedical Sciences, Durham, NH 
-abstract: In species that provide parental care, successful rearing of offspring involves a shift from aggressive and sexual behaviors to more caring and nurturing ones.
-We hypothesize that these shifts in behavior are related to changes in gene expression that are controlled by internal mechanisms rather than by the external cues.
-We found that gene expression in the hypothalamus is fairly insensitive to the natural transition between egg and nestling stages; however, in the pituitary and gonads, gene expression showed much more plasticity over the course of reproduction.
-We show evidence that natural transitions in reproductive and parental behavior by mediated internal clock controlling gene expression in the pitary and gonads but that the hypothalamus is very sensitive to external cues that signal a distruption in the natural parental care cycle.
-This research provides a deeper understanding of the interplay of genes, hormones, and parental care and has important implications for understanding molecular processes that underlie behavior transitions.
+abstract: In animals that exhibit a parental care reproductive strategy, successful rearing of offspring typically involves a shift from sexual behaviors to more caring and nurturing ones.
+Decades of work have highlighted the role of particular hormones in the maintenance of parental care behavior, the majority of which are produced by a biological system essential for reproduction, the hypothalamic-pituitary-gonadal (HPG) axis.
+However, we know very little about when and how the HPG axis transitions into parental care behaviors in any vertebrate, and this knowledge is fundamentally important to our understanding of the mechanisms mediating parental care.
+First, we characterized the transcriptome of the HPG axis in both mothers and fathers over 9 points of the parental care stage using a socially monogamous and bi-parental model, the rock dove (Columba livia).
+Then, we experimentally manipulated offspring presence at 7 different time points using classic offspring replacement and removal manipulations to uncover cause and effect of these transcriptional changes.
+We offer the most complete characterization and understanding of changes in neural, pituitary and gonadal transcription and translation ever reported in any vertebrate over the course of parental care.
+
 output:
@@
 
+```{r setup, include = F}
+library(tidyverse)
+library(kableExtra)
+library(cowplot)
+```
+
 ## Introduction
 
-Reproduction is a critical period in an organism’s life cycle.
-Parents must balance their own needs with the energetic costs and time-constraints required to successfully reproduce and rear their offspring to independence (cite).
-A complex series of evolutionarily conserved physiological processes underlie behaviors typically associated with reproduction in vertebrates, including courtship, nesting, and parental care.
-These processes are primarily mediated by the reproductive axis, which consists of the hypothalamus in the brain, the pituitary, and the gonads (i.e., the HPG).
-Our understanding of the physiological changes that adults navigate as they transition through reproductive states are primarily limited to a number of circulating hormones and a small set of well-known genes (cite).
-We have yet to come close to conceptualizing the larger symphony of reproductive events that occur to yield reproductive and parental care behaviors.
+Parental care is essential to many organisms’ reproductive strategy and predominates avian and mammalian groups (Clutton-Brock, 1991; Royle et al., 2012).
+Depending on the species, reproductive and parental behaviors may include territory establishment, courtship and mating, nest-building, nest defense, incubation, and the brooding and provisioning of young.
+The goal of these behaviors, either directly or indirectly, is the successful production of offspring resulting in a direct fitness benefit  (Reviewed in Santos, E.S.A, 2012).
+Underlying these behaviors are the anatomical and physiological changes that facilitate the transition into and across reproductive and parenting stages.
+These changes have fascinated biologists and inspired a prodigious body of research.
+Much of this research has focused on understanding reproduction and parental care from an endocrine and neuroendocrine perspective (reviewed in Ondrasek 2016, Kelly 2020, Wingfield, Silver, Buntin).
+Reproduction and parental care is largely regulated by a set of three tissues: The Hypothalamus in the brain, the Pituitary, and the Gonads (testes and ovaries).
+Together, these tissues are often referred to as the ‘reproductive’ axis or the HPG.
+To activate reproduction, the HPG initiates an endocrine cascade starting with the production and release of gonadotropin-releasing hormone (GnRH) in the hypothalamus which, in turn,  stimulates the anterior pituitary to produce luteinizing hormone (LH) and follicle-stimulating hormone (FSH).
+LH and FSH travel to the gonads via the bloodstream where they induce gonadal growth and steroidogenesis of estradiol, testosterone and progesterone).
+These sex steroids have well established roles in mating behavior, nest-building, territory defense, and egg-laying in birds (reviewed in Buntin).
+Another key hormone, prolactin, is produced in the pituitary and is associated with the induction and maintenance of behaviors such as  incubation in birds and lactation (in doves and mammals) (birds: reviewed in Angelier, Austin et al.
+in prep, mammals: Austin and Word).
+Other hormones like vasopressin and oxytocin are related to affiliative behaviors (Yoshihara C, 2018) while dopamine, serotonin aggressive behaviors or avoidance (Rilling, JK, 2013; Adams DB, 2006).
+While these and many other hormones have been reported to play integral roles in various reproductive and parental care behaviors, we have yet to come close to conceptualizing the larger physiological and genomic coordination of events that yield such behaviors.
 
-Using the model of a biparental vertebrate species, the rock dove (Columba livia), our team has previously demonstrated the dynamic nature of gene activity within the reproductive or HPG axis.
-\citet{MacManes2017} reported that HPG gene activity varied by sex, with females exhibiting higher levels of overall gene expression than males.
-\citet{Calisi2018} reported that HPG gene activity varied response to restraint-stress, and Austin et al.
-identify genes that responds to both corticosterone treatment and restraint stress.
-While each of these studies advanced our understanding of sex-differences and transcriptional plasticity in the reproducitve axis, none of these studies examined activity of the reproducitve axis of bird that are actively reproducing or caring for offspring.
-Studying parental care in both sexes of mammalian species is challenging because only the females are able to use lactation to feed their young, and the physiological mechanisms regulating this process are hard to control for.
-The rock dove is an excellent system for studying reproduction and parental care in both sexes because both males and females feed their young an energy-rich secretion produced from the epithelial lining of their crops, known as crop milk.
+To begin to address this larger goal, we investigated genomic changes in the reproductive axis using the model of the rock dove (Columba livia).
+This species is socially and genetically monogamous and has bi-parental incubation and offspring care thereby making investigations of sexual variation across reproduction possible.
+Doves and pigeons have been used extensively to study the physiological mechanisms of reproduction and parental care, including the proximate and ultimate mechanisms that regulate associated behaviors (cite a bunch of new and old papers or books).
+Our team has produced a series of papers with the goal of illuminating how transcriptomic activity in the reproductive axis varied by sex (MacManes et al.
+2016), in response to a stressor (Calisi et al.
+2017, Lang et al.
+2020), and following a hormonal manipulation (Austin et al.
+in prep).
+By using a genomic approach that spans the reproductive timeline, we have the ability to elucidate a transcriptional symphony that coordinates the anatomical, physiological and behavioral changes associated with transitioning from non-breeding to parenting in a vertebrate species.
+The aims of this research were three-fold: Our first aim was to produce a high quality and reproducible transcriptional characterization of the HPG over reproduction and parental care in male and female rock doves.
+We were specifically interested in understanding how gene activity in the HPG changed as individuals transitioned from a sexually mature, non-reproductive state into reproduction and parenting.
+To accomplish this aim, we investigated nine time points that spanned the majority of reproductive efforts in this species.
+These time points consisted of 1) a control or non-parenting state (from MacManes et al 2016 and Calisi et al.
+2017), 2) nest-building, 3) clutch initiation and onset of incubation, 4) clutch completion and early incubation, 5) mid-incubation, 6) late incubation, 7) initiation of nestling care, 8) early nestling care, and 9) mid-nestling care.
+Because the majority of time points included in this study involved some form of direct offspring care, for simplicity, we will discuss these as parental care rather than differentiating between reproductive and parenting behaviors.
+However, it should be understood that parental care remains only one component of a larger set of reproductive behaviors.
+Our second aim focused on characterizing the major patterns of transcriptional changes between the tissues, sexes, and parental stages.
+We specifically investigate how gene expression in the HPG changed across various transition points.
+For example, from non-breeding to mating and nest-building behaviors, from nest-building to egg-laying to incubation, from incubation to nestling care, and from highly dependent nestlings (brooding and crop milk production) to older, thermoregulatory independent nestlings that are fed a regurgitant of seeds.
+Our third aim was to identify the drivers underlying changes in transcription.
+Specifically, we tested whether external or internal factors were regulating gene activity.
+We conducted a series of offspring removal and replacement manipulations to test whether transcriptional changes were dependent upon offspring presence or were regulated by an internal clock.
 
-In this study, we investigate the activity of the HPG axis over the course of reproduction and parental care by characterizing gene expression as birds transition from non-breeding to nest-building, as they lay and incubate their clutches, and finally as they become parents and raise their chicks.
-Then, we conducted a series of nest manipulations to ascertain constraints associated with biological timing versus input from external stimuli.
-We conducted a comprehensive series of analyses examine broad patterns of variation as well as the activity of specific genes in specific tissues.
-We have also made our data and scripts available so that others can reuse the data to ask different questions or reproduce our results.
-This research provides a deeper understanding of the interplay of genes, hormones, and parental care and has important implications for understanding molecular processes that underlie behavioral transitions.
-\citep[reviewed in][]{Lea1989} say this, but other say that \citep[reviewed in][]{Dulac2014}.
+Additionally, because these data are extensive, we took the opportunity to provide them in an easily accessible form that can be used for independent data exploration by the reader.
+Given the homology of genes and gene networks shared between rock doves and other vertebrate species, the findings from these RNA-seq experiments could be of interest to scientists and clinicians studying diverse systems, particularly in non-model organisms like our study species.
+Overall, this study provides a deeper understanding of the interplay of genes across parenting and has important implications for understanding molecular processes that underlie behavioral transitions.
 
@@
 
-_Data-driven characterization of reproductive and parental stage groups._
+### The rock dove reproduction and parental care transcriptome
 
-We first examined broad patterns of variation in gene expression in the hypothalamus, pituitary, and gonads of male and female rock doves across a series of timepoints that encapsulate important transitions in reproduction and and parental care (Fig.
-1A).
-t-Distributed Stochastic Neighbor Embedding (t-SNE) is a nonlinear, dimension reduction technique is used to visualize high-dimensional data (van der Maaten and Hinton 2008) and is widely used in in RNA sequencing to identify related and distinct cell types (refs).
-Differential expespression profiling is useful for pair-wise comparisons of differentially expressed genes (DEGs) with positive log fold change (+LFC) in one group compared to another.
-Combined, the provide stronger support for similarities and differences between cell-types and treatment groups in multi-dimential datasets.
-First, the data supports that the hypothalamus, pituitary, and gonads have distinct gene expression profiles as evidenced by little overlap in t-SNE space and an average of 13,000 DEGs between each tissue (Fig.
-1B).
-Second, the data supports that sex-differences are much more promient in the gonads than in the brain (13,000 versus 2,000 DEGs) (Fig.
-1C).
-Because of these widespread tissue and sex-differences, further ananlyses were conducted separately for each tissue and sex (Fig.
-1D, Suppl.
-Fig 1).
-In all the tissues, the control samples are quiet different from all the other timepoints samples, and only a fraction of the transcriptome is altered during the course of parental care at the timepoints sampled.
-The most notable differences male and female hpothalamic control samples are very different from on the rest (~7000 DEGs, and only the two nestling care stages (n5 and n9)) show differential expression of more than 1,000 genes.
-In the pituitary, late inubcation and hatch are quite simlilar to each other and distinct from the earlier incubation stages stages, and temporal changes in gene expression appears to be less variable in males than females.
-In the female gonads there appear to be more changes in gene expression early one, between nest building and early incuation, but the male gonads show very little variation associated with parental stage.
+We created the rock dove transcriptome by sequenced 987 RNA samples from the hypothalamus, pituitary and gonads from the female and male of 166 rock dove pairs across (Fig.
+1).
+These samples come from nine timepoints that encapsulate important transitions in reproduction (Fig.
+1A), four offspring removal manipulations (Fig.
+1B), and three offspring replacement manipulations (Fig.
+1C) to determine the influence of internal mechanisms and external stimuli on gene expression.
+Using T-distributed Stochastic Neighbor Embedding (tSNE) we show broad patterns of variation due to treatment or sex in each tissue (Fig.
+1D).
+In the hypothalamus, tSNE dimension visually separates most of the samples that were subject to an offspring manipulation from those from naturally occurring parental stages.
+In the pituitary however, samples cluster more by wether they were before incubation day 9 or after incubation day 17.
+Neither the hypothalamus or pituitary show strong sex differences at this level, but the gonads samples from very distinct, sex-specific clusters.
 
-_Hypothesis-driven analysis of candidate parental-care genes._
-
-- Champagne Curley Genes (Fig 2)
-- GO genes (Fig 2 Suppl Fig 1)
-- Cancer genes (Fig 2 Suppl Fig 2)
-- Shiny genes (Fig 2 Suppl Fig 3)
-
-_Software for data exploration and hypothesis testing_
-
-Data accessibility and reproducibility are key challenges for today's researchers.
-Using Binder and Shiny, we have provides a means for scientists, physicians, and the general public to explore the data, code, and results from this project at <https://raynamharris.shinyapps.io/musicalgenes/>.
-Like Hipposeq (Cembrowski et al elife 2016, 2018), this tools allows quick and easy data exploration of RNAseq data from mutiple cell types that relate to brain function.
+```{r fig1, echo = F, message = F, fig.cap = 'Experimental design.
+A) We investigated nine timepoints that spanned the majority of reproductive efforts in this species.
+These timepoints consisted of a "control" or non-parenting state (from MacManes et al 2016 and Calisi et al.
+2017), nest-building, clutch initiation or the onset of incubation, clutch completion and early incubation, mid-incubation, late incubation, initiation of nestling care, early nestling care, and mid-nestling care.
+To teste whether external or internal factors were regulating gene activity, we also conducted a series of offspring removal (B) and replacement (C) manipulations to test whether transcriptional changes were dependent upon offspring presence or were regulated by an internal clock.
+(D) We used t-Distributed Stochastic Neighbor Embedding (t-SNE) to reduce the dimensionality of the transcriptomes from the hypothalamus, pituitary, and gonads of male and female rock.
+Circles and triangles represent female and male samples, respectively, and points are colored by treatment.'}
+a <- png::readPNG("../figures/fig1-1.png")
+a <- ggdraw() +  draw_image(a, scale = 0.95)
+a
+```
 
 ## Discussion
-
-This RNA-sequencing project was designed to characterize the reproductive axis of bi-parental pigeons across a characteristic reproductive cycle and through the parental main parental stages of egg incubation and parental care using data- and hypotheses driven approaches.
-We hypothesized that differences in gene expression would be driven by either internal physiology or external stimuli.
-In the pituitary, we found that the most explanatory source of variation in gene expression was the expression level of prolactin (PRL).
-Circulating levels of prolactin in the blood in this species are known to rise around day nine and peak on or the day before hatch.
-Our results confirm this and also identified one hundred genes whose expression is highly correlated with prolactin.
-Today, this variation, which changes more within a parental stage (e.g.
-egg incubation or nestling care) suggests that internal physiology drives changes in gene expression that precede changes in external stimuli that lead to behavioral changes (e.g.
-crop milk or regurgitation).
-We predicted that either sex steroid peptide hormones would play a role in governing the reproductive transitions, but we did not anticipate discovering a link between PRL expression to genes that are involved with DNA repair and regulation of the cycle.
-Finally, we aimed to make these data and analyses FAIR (findable, accessible, interoperable, and reproducible), open and reproducible to encourage others to reuse our data to test new hypotheses or explore the data with a different perspective or toolset.
-The scripts that made the figures and tables are browseable online thanks to RStudio, GitHub and Binder.
-The data can be visualized and sonified (turned into sound) interactively thanks to RStudio, GitHub, and Shiny.
 
@@
 Statistical analyses and modeling were perforeming using R \citep{RDevelopmentCoreTeam2013, Wickham2016}.
-Limma was use to process all samples using the model ~ tissue * sex * treatment (citation).
+Limma was use to process all samples using the model ~ tissue \* sex \* treatment (citation).
 t-Distributed Stochastic Neighbor Embedding (t-SNE) analysis was conducted using a perplexity of 10 \citep{VanDerMaaten2008}.
@@
 
+\newpage
+
+## Tables
+
+```{r table1, echo =FALSE, message = F }
+table1 <- read_csv("../results/table1.csv")
+table1[is.na(table1)] <- " "
+kable(table1,
+      col.names = c('gene', 'bldg-lay',  'lay-inc3', 'inc3-inc9', 'inc9-inc17', 
+                    'inc17-hatch', 'hatch-n5', 'n5-n9'),
+      caption = "Changes in candidate gene expression.
+Significant chagns in gene expression are shown for X candidate genes.
+Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G).
+Comparisons are denoted in the column names.
+The first group is the reference group.")
+```
+
+## Figures
+
+```{r fig2, echo = F, message = F, fig.cap = 'Hypothesis-driven and data-driven approaches to identifying changes in gene expression.Box-and-whisker plots show changes in gene expression of serotonin receptor (HTR2C) in the hypothalamus (A), prolactin (PRL) in the pituitary (B), and estrogen receptor 1 (ESR1) in the gonads.Bar and statistics are shown for one contrast of interest between treatment groups.
+Boxes are colored by treatment.
+Volcano plots show the log-fold change (LFC) and -log10(FDR) for all genes that are significantly different differentially between the aforementioned treatment groups.
+Boxes are colored by the treatment were gene expression was higher.'}
+a <- png::readPNG("../figures/fig2-1.png")
+a <- ggdraw() +  draw_image(a, scale = 0.95)
+a
+```
+
+```{r fig3, echo = F, message = F, fig.cap = 'Summary of differentially expressed genes.
+Bar plots show the total number of significantly differentially expressed genes for all two-way comparisons in the hypothalamus (A), pituitary (B), gonads (C) .
+Contrasts are made relative to the non-breeding controls, the nest-building group, the previous stage, or the internal or external control group.
+Bars are colored by the treatment where gene expression was higher.
+A negative number of differentially expressed genes means that many genes had increase expression in the reference group.'}
+a <- png::readPNG("../figures/fig2-1.png")
+a <- ggdraw() +  draw_image(a, scale = 0.95)
+a
+```
+
+\newpage
+
+## References
+
+\bibliography{RJreferences}
+
 ## Acknowledgements
@@
 We thank Pacha for help with database and Shiny.
-
 We thank Irene, Erica, Beth Krestoff and Tiffany Chen for assistance in managing the aviary and collecting nest stage data.
@@
 
-## References
-
-\bibliography{RJreferences}
-
-\newpage
-
-## Figures
-
-\includegraphics[width=0.95\linewidth]{fig1-1}
-
-**Figure 1.
-Experimental design.** A) We investigated nine timepoints that spanned the majority of reproductive efforts in this species.
-These time-points consisted of a "control" or non-parenting state (from MacManes et al 2016 and Calisi et al.
-2017), nest-building, clutch initiation or the onset of incubation, clutch completion and early incubation, mid-incubation, late incubation, initiation of nestling care, early nestling care, and mid-nestling care.
-To teste whether external or internal factors were regulating gene activity, we also conducted a series of offspring removal (B) and replacement (C) manipulations to test whether transcriptional changes were dependent upon offspring presence or were regulated by an internal clock.
-(D) We used t-Distributed Stochastic Neighbor Embedding (t-SNE) to reduce the dimensionality of the transcriptomes from the hypothalamus, pituitary, and gonads of male and female rock.
-Circles and triangles represent female and male samples, respectively, and points are colored by treatment.
-
-\includegraphics[width=0.95\linewidth]{fig2-1}
-
-**Figure 2.
-Hypothesis-driven and data-driven approaches to identifying changes in gene expression.** Box-and-whisker plots show changes in gene expression of serotonin receptor (HTR2C) in the hypothalamus (A), prolactin (PRL) in the pituitary (B), and estrogen receptor 1 (ESR1) in the gonads.Bar and statistics are shown for one contrast of interest between treatment groups.
-Boxes are colored by treatment.
-Volcano plots show the log-fold change (LFC) and -log10(FDR) for all genes that are significantly different differentially between the aforementioned treatment groups.
-Boxes are colored by the treatment were gene expression was higher.
-
-\includegraphics[width=0.95\linewidth]{fig3-1}
-
-**Figure 3.
-Summary of differentially expressed genes.** Bar plots show the total number of significantly differentially expressed genes for all two-way comparisons in the hypothalamus (A), pituitary (B), gonads (C) .
-Contrasts are made relative to the non-breeding controls, the nest-building group, the previous stage, or the internal or external control group.
-Bars are colored by the treatment where gene expression was higher.
-A negative number of differentially expressed genes means that many genes had increase expression in the reference group.
-
-\newpage
-
 ## Author affiliations

### 2020-09-07 analysis/05_figs23.R

@@ -66,8 +66,8 @@ cd <- makefig2("B", "E",  "pituitary", pitf, pitm, "PRL",
                "inc.d9_inc.d17", "hatch_early", 
                "inc.d9", "inc.d17",  "hatch", "early")
 ef <- makefig2("C", "F", "gonads", gonf, gonm, "ESR2",
-               "control_lay", "lay_inc.d3",
-               "control", "lay", "inc.d3", "m.inc.d3")
+               "control_lay", "inc.d17_prolong",
+               "control", "lay", "inc.d17", "prolong")
 
 fig2 <- plot_grid(ab,cd,ef, nrow = 3)
 

### 2020-09-07 analysis/06_tables.R

@@ -1,0 +1,21 @@
+source("R/themes.R")
+library(tidyverse)
+
+cDEGs <- read_csv("results/03_wideDEGscandidates.csv") %>%
+  select(-X1)
+
+
+mycomparsisons <- c(levelssequential)
+
+# helper function to add column of NAs if no DEGs exist
+fncols <- function(data, cname) {
+  add <-cname[!cname%in%names(data)]
+  if(length(add)!=0) data[add] <- NA
+  data
+}
+
+table1 <- fncols(cDEGs, mycomparsisons) %>%
+  select(gene,levelssequential) 
+table1
+
+write_csv(table1, "results/table1.csv")

### 2020-09-08 manuscript/elife-v1.tex
now with 3 awesome tables

@@
 replacement manipulations (Fig. 1C) to determine the influence of
-internal mechanisms and external stimuli on gene expression. Using
-T-distributed Stochastic Neighbor Embedding (tSNE) we show broad
+internal mechanisms and external stimuli on gene expression.
+
+Using T-distributed Stochastic Neighbor Embedding (tSNE) we show broad
 patterns of variation due to treatment or sex in each tissue (Fig. 1D).
@@
 sex-specific clusters.
-
-\begin{Schunk}
-\begin{figure}
-\includegraphics{elife_v1_files/figure-latex/fig1-1} \caption[Experimental design]{Experimental design.
-A) We investigated nine timepoints that spanned the majority of reproductive efforts in this species.
-These timepoints consisted of a "control" or non-parenting state (from MacManes et al 2016 and Calisi et al.
-2017), nest-building, clutch initiation or the onset of incubation, clutch completion and early incubation, mid-incubation, late incubation, initiation of nestling care, early nestling care, and mid-nestling care.
-To teste whether external or internal factors were regulating gene activity, we also conducted a series of offspring removal (B) and replacement (C) manipulations to test whether transcriptional changes were dependent upon offspring presence or were regulated by an internal clock.
-(D) We used t-Distributed Stochastic Neighbor Embedding (t-SNE) to reduce the dimensionality of the transcriptomes from the hypothalamus, pituitary, and gonads of male and female rock.
-Circles and triangles represent female and male samples, respectively, and points are colored by treatment.}\label{fig:fig1}
-\end{figure}
-\end{Schunk}
 
@@
 
-\caption{\label{tab:table1}Changes in candidate gene expression.
-Significant chagns in gene expression are shown for X candidate genes.
+\caption{\label{tab:table1}Changes in candidate gene expression over the course of parental care in rock doves.
+Significant changes in gene expression are shown for X candidate genes.
 Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G).
@@
 \hline
-AR &  &  &  &  &  &  & \\
-\hline
-AVPR1B &  &  &  &  &  &  & \\
-\hline
 AVPR2 &  &  &  &  &  & FP(+) & \\
-\hline
-CRH &  &  &  &  &  &  & \\
-\hline
-CRHR1 &  &  &  &  &  &  & \\
-\hline
-DIO1 &  &  &  &  &  &  & \\
 \hline
@@
 \hline
-ESR2 &  &  &  &  &  &  & \\
-\hline
-FSHB &  &  &  &  &  &  & \\
-\hline
-FSHR &  &  &  &  &  &  & \\
-\hline
-GABRQ &  &  &  &  &  &  & \\
-\hline
-GALR1 &  &  &  &  &  &  & \\
-\hline
-GNRH1 &  &  &  &  &  &  & \\
-\hline
 GNRHR & FP(-) &  &  &  &  &  & \\
-\hline
-JAK2 &  &  &  &  &  &  & \\
-\hline
-LHCGR &  &  &  &  &  &  & \\
-\hline
-NPFFR1 &  &  &  &  &  &  & \\
-\hline
-NPVF &  &  &  &  &  &  & \\
-\hline
-NPY &  &  &  &  &  &  & \\
-\hline
-NR3C2 &  &  &  &  &  &  & \\
-\hline
-POMC &  &  &  &  &  &  & \\
 \hline
@@
 PRLH & FP(-) & FP(+) &  & FP(+) &  &  & \\
-\hline
-PRLHR &  &  &  &  &  &  & \\
-\hline
-SERPINA4 &  &  &  &  &  &  & \\
-\hline
-VIPR1 &  &  &  &  &  &  & \\
 \hline
@@
 
+\begin{Schunk}
+\begin{table}
+
+\caption{\label{tab:table2}Changes in candidate gene expression in response to removal.
+Significant changess in gene expression are shown for X candidate genes.
+Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G).
+Comparisons are denoted in the column names.
+The first group is the reference group.}
+\centering
+\begin{tabular}[t]{l|l|l|l|l}
+\hline
+gene & inc.d3\_m.inc.d3 & inc.d9\_m.inc.d9 & inc.d17\_m.inc.d17 & hatch\_m.n2\\
+\hline
+AVPR1B &  &  &  & FP(+)\\
+\hline
+AVPR2 &  &  &  & FP(+)\\
+\hline
+DIO3 &  &  & FP(+); MH(-) & FH(-); FP(+)\\
+\hline
+ESR2 &  &  & FH(-); MH(-) & \\
+\hline
+FSHB &  &  & FP(-) & FH(-); FP(-)\\
+\hline
+GABRQ &  &  & FH(+) & FH(+)\\
+\hline
+JAK2 &  &  & FP(+) & FP(+)\\
+\hline
+LHCGR &  &  & FH(-) & FH(-); MH(-)\\
+\hline
+NPVF & FH(-) &  & FH(-) & FH(-); MH(-)\\
+\hline
+NPY & FP(-) &  &  & MP(+)\\
+\hline
+PRL &  &  & FP(-) & MP(-)\\
+\hline
+PRLH &  &  &  & MH(+)\\
+\hline
+SERPINA4 &  &  & FP(+) & FP(+)\\
+\hline
+\end{tabular}
+\end{table}
+
+\end{Schunk}
+
+\begin{Schunk}
+\begin{table}
+
+\caption{\label{tab:table3}Changes in candidate gene expression in response to replacement.
+Significant changess in gene expression are shown for X candidate genes.
+Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G).
+Comparisons are denoted in the column names.
+The first group is the reference group.}
+\centering
+\begin{tabular}[t]{l|l|l|l|l|l}
+\hline
+gene & inc.d9\_early & inc.d17\_prolong & hatch\_early & hatch\_prolong & hatch\_extend\\
+\hline
+AR &  &  & FP(+) &  & \\
+\hline
+AVPR2 &  &  & FP(+) &  & \\
+\hline
+CRHR1 &  & MG(+) &  &  & \\
+\hline
+ESR2 &  & MG(+) &  &  & \\
+\hline
+GABRQ & MP(+) & FH(+) & MH(+); MP(+) & FH(+); MG(+) & FH(+)\\
+\hline
+GNRH1 &  &  & FH(+) &  & \\
+\hline
+LHCGR &  & FH(-) & MH(-) & FH(-) & FH(-); MH(-)\\
+\hline
+NPVF & MH(-) & FH(-) & FH(-); MH(-) & FH(-) & FH(-)\\
+\hline
+NPY &  & MP(-) & MP(+) &  & \\
+\hline
+POMC &  &  & FH(-) & FH(-) & \\
+\hline
+PRL &  &  & FP(-); MP(-) &  & \\
+\hline
+PRLHR &  &  & MP(+) &  & \\
+\hline
+SERPINA4 &  & MG(+) & MH(+) &  & \\
+\hline
+\end{tabular}
+\end{table}
+
+\end{Schunk}
+
+levelsmanip
+
 \hypertarget{figures}{%
 \subsection{Figures}\label{figures}}
+
+\begin{Schunk}
+\begin{figure}
+\includegraphics{elife_v1_files/figure-latex/fig1-1} \caption[Experimental design]{Experimental design.
+A) We investigated nine timepoints that spanned the majority of reproductive efforts in this species.
+These timepoints consisted of a "control" or non-parenting state (from MacManes et al 2016 and Calisi et al.
+2017), nest-building, clutch initiation or the onset of incubation, clutch completion and early incubation, mid-incubation, late incubation, initiation of nestling care, early nestling care, and mid-nestling care.
+To teste whether external or internal factors were regulating gene activity, we also conducted a series of offspring removal (B) and replacement (C) manipulations to test whether transcriptional changes were dependent upon offspring presence or were regulated by an internal clock.
+(D) We used t-Distributed Stochastic Neighbor Embedding (t-SNE) to reduce the dimensionality of the transcriptomes from the hypothalamus, pituitary, and gonads of male and female rock.
+Circles and triangles represent female and male samples, respectively, and points are colored by treatment.}\label{fig:fig1}
+\end{figure}
+\end{Schunk}
 

### 2020-09-08 manuscript/elife_v1.Rmd
now with 3 awesome tables

@@
 library(cowplot)
+
+source("../R/themes.R")
 ```
@@
 1C) to determine the influence of internal mechanisms and external stimuli on gene expression.
+
 Using T-distributed Stochastic Neighbor Embedding (tSNE) we show broad patterns of variation due to treatment or sex in each tissue (Fig.
@@
 Neither the hypothalamus or pituitary show strong sex differences at this level, but the gonads samples from very distinct, sex-specific clusters.
-
-```{r fig1, echo = F, message = F, fig.cap = 'Experimental design.
-A) We investigated nine timepoints that spanned the majority of reproductive efforts in this species.
-These timepoints consisted of a "control" or non-parenting state (from MacManes et al 2016 and Calisi et al.
-2017), nest-building, clutch initiation or the onset of incubation, clutch completion and early incubation, mid-incubation, late incubation, initiation of nestling care, early nestling care, and mid-nestling care.
-To teste whether external or internal factors were regulating gene activity, we also conducted a series of offspring removal (B) and replacement (C) manipulations to test whether transcriptional changes were dependent upon offspring presence or were regulated by an internal clock.
-(D) We used t-Distributed Stochastic Neighbor Embedding (t-SNE) to reduce the dimensionality of the transcriptomes from the hypothalamus, pituitary, and gonads of male and female rock.
-Circles and triangles represent female and male samples, respectively, and points are colored by treatment.'}
-a <- png::readPNG("../figures/fig1-1.png")
-a <- ggdraw() +  draw_image(a, scale = 0.95)
-a
-```
 
@@
 
+```{r cDEGs, include = F}
+cDEGs <- read_csv("../results/03_wideDEGscandidates.csv") %>%
+  select(-X1)
+
+# helper function to add column of NAs if no DEGs exist
+fncols <- function(data, cname) {
+  add <-cname[!cname%in%names(data)]
+  if(length(add)!=0) data[add] <- NA
+  data
+}
+
+maketables123 <- function(df, whichlevels){
+  table <- fncols(cDEGs, whichlevels) %>% 
+    select(gene, whichlevels)  %>%
+    mutate(allres = rowSums(is.na(.)))
+
+  numcomparisons <- ncol(table) -2
+
+  table <- table %>%
+    filter(allres < numcomparisons) %>%
+    select(-allres)
+  table[is.na(table)] <- " "
+  print(table)
+  return(table)
+
+}
+
+table1 <- maketables123(cDEGs, levelssequential) 
+table2 <- maketables123(cDEGs, levelsrm) 
+table3 <- maketables123(cDEGs, levelsreplace)
+
+write_csv(table1, "../results/table1.csv")
+write_csv(table2, "../results/table2.csv")
+write_csv(table3, "../results/table3.csv")
+
+```
+
 _Data availablility_
@@
 ```{r table1, echo =FALSE, message = F }
-table1 <- read_csv("../results/table1.csv")
-table1[is.na(table1)] <- " "
 kable(table1,
@@
                     'inc17-hatch', 'hatch-n5', 'n5-n9'),
-      caption = "Changes in candidate gene expression.
-Significant chagns in gene expression are shown for X candidate genes.
+      caption = "Changes in candidate gene expression over the course of parental care in rock doves.
+Significant changes in gene expression are shown for X candidate genes.
 Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G).
@@
 
+```{r table2, echo =FALSE, message = F }
+kable(table2,
+      caption = "Changes in candidate gene expression in response to removal.
+Significant changess in gene expression are shown for X candidate genes.
+Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G).
+Comparisons are denoted in the column names.
+The first group is the reference group.")
+```
+
+```{r table3, echo =FALSE, message = F }
+kable(table3,
+      caption = "Changes in candidate gene expression in response to replacement.
+Significant changess in gene expression are shown for X candidate genes.
+Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G).
+Comparisons are denoted in the column names.
+The first group is the reference group.")
+```
+
+levelsmanip
+
 ## Figures
+
+```{r fig1, echo = F, message = F, fig.cap = 'Experimental design.
+A) We investigated nine timepoints that spanned the majority of reproductive efforts in this species.
+These timepoints consisted of a "control" or non-parenting state (from MacManes et al 2016 and Calisi et al.
+2017), nest-building, clutch initiation or the onset of incubation, clutch completion and early incubation, mid-incubation, late incubation, initiation of nestling care, early nestling care, and mid-nestling care.
+To teste whether external or internal factors were regulating gene activity, we also conducted a series of offspring removal (B) and replacement (C) manipulations to test whether transcriptional changes were dependent upon offspring presence or were regulated by an internal clock.
+(D) We used t-Distributed Stochastic Neighbor Embedding (t-SNE) to reduce the dimensionality of the transcriptomes from the hypothalamus, pituitary, and gonads of male and female rock.
+Circles and triangles represent female and male samples, respectively, and points are colored by treatment.'}
+a <- png::readPNG("../figures/fig1-1.png")
+a <- ggdraw() +  draw_image(a, scale = 0.95)
+a
+```
 

### 2020-09-09 manuscript/elife_v1.Rmd
rm old files
vsd working, degs not working

--- 2020-09-07 manuscript/elife_v1.Rmd
@@
 Much of this research has focused on understanding reproduction and parental care from an endocrine and neuroendocrine perspective (reviewed in Ondrasek 2016, Kelly 2020, Wingfield, Silver, Buntin).
+
 Reproduction and parental care is largely regulated by a set of three tissues: The Hypothalamus in the brain, the Pituitary, and the Gonads (testes and ovaries).
@@
 
+```{r table2, echo =FALSE, message = F }
+table2 <- read_csv("../results/table2.csv")
+table2[is.na(table2)] <- " "
+kable(table2,
+      caption = "Changes in candidate gene expression in response to removal.
+Significant changess in gene expression are shown for X candidate genes.
+Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G).
+Comparisons are denoted in the column names.
+The first group is the reference group.")
+```
+
+```{r table3, echo =FALSE, message = F }
+table3 <- read_csv("../results/table3.csv")
+table3[is.na(table3)] <- " "
+kable(table3,
+      caption = "Changes in candidate gene expression in response to replacement.
+Significant changess in gene expression are shown for X candidate genes.
+Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G).
+Comparisons are denoted in the column names.
+The first group is the reference group.")
+```
+
+levelsmanip
+
 ## Figures
+
+```{r fig1, echo = F, message = F, fig.cap = 'Experimental design.
+A) We investigated nine timepoints that spanned the majority of reproductive efforts in this species.
+These timepoints consisted of a "control" or non-parenting state (from MacManes et al 2016 and Calisi et al.
+2017), nest-building, clutch initiation or the onset of incubation, clutch completion and early incubation, mid-incubation, late incubation, initiation of nestling care, early nestling care, and mid-nestling care.
+To teste whether external or internal factors were regulating gene activity, we also conducted a series of offspring removal (B) and replacement (C) manipulations to test whether transcriptional changes were dependent upon offspring presence or were regulated by an internal clock.
+(D) We used t-Distributed Stochastic Neighbor Embedding (t-SNE) to reduce the dimensionality of the transcriptomes from the hypothalamus, pituitary, and gonads of male and female rock.
+Circles and triangles represent female and male samples, respectively, and points are colored by treatment.'}
+a <- png::readPNG("../figures/fig1-1.png")
+a <- ggdraw() +  draw_image(a, scale = 0.95)
+a
+```
 
